### 2024-02-05 paper-quollr.Rmd
Added initial plan set up

---
title: "quollr: An R Package for Visalizing 2D Models in High Dimensional Space"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  - name: Jayani P.G. Lakshika
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email:  \email{jayani.piyadigamage@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)

```

```{r load-libraries}
library(quollr)
```

# Introduction

# Implementation

## Installation

The package can be installed from CRAN using

> ```install.packages("quollr")```

and from GitHub using

> ```remotes::install_github("JayaniLakshika/quollr")```

to install the development version.

## Web site

More documentation of the package can be found at the web site (need to add the link).

## Data sets

## Functions

## Tests

# Examples

# Conclusion

# Acknowledgements

This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template. The source code for reproducing this paper can be found at: [https://github.com/JayaniLakshika/paper-quollr](https://github.com/JayaniLakshika/paper-quollr).

### 2024-02-05 paper-quollr.tex
Added initial plan set up

% !TeX root = RJwrapper.tex
\title{quollr: An R Package for Visalizing 2D Models in High Dimensional Space}

\author{by Jayani P.G. Lakshika, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}

\maketitle

\abstract{%
An abstract of less than 150 words.
}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{implementation}{%
\section{Implementation}\label{implementation}}

\hypertarget{installation}{%
\subsection{Installation}\label{installation}}

The package can be installed from CRAN using

\begin{quote}
\texttt{install.packages("quollr")}
\end{quote}

and from GitHub using

\begin{quote}
\texttt{remotes::install\_github("JayaniLakshika/quollr")}
\end{quote}

to install the development version.

\hypertarget{web-site}{%
\subsection{Web site}\label{web-site}}

More documentation of the package can be found at the web site (need to add the link).

\hypertarget{data-sets}{%
\subsection{Data sets}\label{data-sets}}

\hypertarget{functions}{%
\subsection{Functions}\label{functions}}

\hypertarget{tests}{%
\subsection{Tests}\label{tests}}

\hypertarget{examples}{%
\section{Examples}\label{examples}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This article is created using \CRANpkg{knitr} (Xie 2015) and \CRANpkg{rmarkdown} (Xie, Allaire, and Grolemund 2018) in R with the \texttt{rjtools::rjournal\_article} template. The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.

\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{{R} Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.

\end{CSLReferences}

\address{%
Jayani P.G. Lakshika\\
Monash University\\%
Department of Econometrics and Business Statistics, VIC 3800 Australia\\
%
\url{https://jayanilakshika.netlify.app/}\\%
\textit{ORCiD: \href{https://orcid.org/0000-0002-6265-6481}{0000-0002-6265-6481}}\\%
\email{jayani.piyadigamage@monash.edu}%
}

\address{%
Dianne Cook\\
Monash University\\%
Department of Econometrics and Business Statistics, VIC 3800 Australia\\
%
\url{http://www.dicook.org/}\\%
\textit{ORCiD: \href{https://orcid.org/0000-0002-3813-7155}{0000-0002-3813-7155}}\\%
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
}

\address{%
Paul Harrison\\
Monash University\\%
MGBP, BDInstitute, VIC 3800 Australia\\
%
%
\textit{ORCiD: \href{https://orcid.org/0000-0002-3980-268X}{0000-0002-3980-268X}}\\%
\href{mailto:paul.harrison@monash.edu}{\nolinkurl{paul.harrison@monash.edu}}%
}

\address{%
Michael Lydeamore\\
Monash University\\%
Department of Econometrics and Business Statistics, VIC 3800 Australia\\
%
%
\textit{ORCiD: \href{https://orcid.org/0000-0001-6515-827X}{0000-0001-6515-827X}}\\%
\href{mailto:michael.lydeamore@monash.edu}{\nolinkurl{michael.lydeamore@monash.edu}}%
}

\address{%
Thiyanga S. Talagala\\
University of Sri Jayewardenepura\\%
Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka\\
%
\url{https://thiyanga.netlify.app/}\\%
\textit{ORCiD: \href{https://orcid.org/0000-0002-0656-9789}{0000-0002-0656-9789}}\\%
\href{mailto:ttalagala@sjp.ac.lk}{\nolinkurl{ttalagala@sjp.ac.lk}}%
}

### 2024-02-13 paper-quollr.Rmd
Added main workflow for the paper

@@
                       message = FALSE)
-
+source("quollr_code.R", local = TRUE)
 ```
@@
 
+### Compute hexagonal bin configurations
+
+### Generate full hex grid
+
+### Construct the 2D model with different options
+
+### Construct the high-D model with different options
+
+### Generate the triangular mesh
+
+### Compute parameter defaults
+
+#### Shift the hexagonal grid origin
+
+#### Benchmark value to remove the low-density hexagons
+
+#### Benchmark value to remove the long edges
+
+### Model function
+
+### Predict 2D embeddings
+
+### Compute residuals
+
+### Visualizations
+
+#### geom_trimesh
+
+#### coloured_long_edges
+
+#### remove long edges
+
+#### show_langevitour
+
 ## Tests

### 2024-02-13 paper-quollr.tex
Added main workflow for the paper

@@
 \subsection{Functions}\label{functions}}
+
+\hypertarget{compute-hexagonal-bin-configurations}{%
+\subsubsection{Compute hexagonal bin configurations}\label{compute-hexagonal-bin-configurations}}
+
+\hypertarget{generate-full-hex-grid}{%
+\subsubsection{Generate full hex grid}\label{generate-full-hex-grid}}
+
+\hypertarget{construct-the-2d-model-with-different-options}{%
+\subsubsection{Construct the 2D model with different options}\label{construct-the-2d-model-with-different-options}}
+
+\hypertarget{construct-the-high-d-model-with-different-options}{%
+\subsubsection{Construct the high-D model with different options}\label{construct-the-high-d-model-with-different-options}}
+
+\hypertarget{generate-the-triangular-mesh}{%
+\subsubsection{Generate the triangular mesh}\label{generate-the-triangular-mesh}}
+
+\hypertarget{compute-parameter-defaults}{%
+\subsubsection{Compute parameter defaults}\label{compute-parameter-defaults}}
+
+\hypertarget{shift-the-hexagonal-grid-origin}{%
+\paragraph{Shift the hexagonal grid origin}\label{shift-the-hexagonal-grid-origin}}
+
+\hypertarget{benchmark-value-to-remove-the-low-density-hexagons}{%
+\paragraph{Benchmark value to remove the low-density hexagons}\label{benchmark-value-to-remove-the-low-density-hexagons}}
+
+\hypertarget{benchmark-value-to-remove-the-long-edges}{%
+\paragraph{Benchmark value to remove the long edges}\label{benchmark-value-to-remove-the-long-edges}}
+
+\hypertarget{model-function}{%
+\subsubsection{Model function}\label{model-function}}
+
+\hypertarget{predict-2d-embeddings}{%
+\subsubsection{Predict 2D embeddings}\label{predict-2d-embeddings}}
+
+\hypertarget{compute-residuals}{%
+\subsubsection{Compute residuals}\label{compute-residuals}}
+
+\hypertarget{visualizations}{%
+\subsubsection{Visualizations}\label{visualizations}}
+
+\hypertarget{geom_trimesh}{%
+\paragraph{geom\_trimesh}\label{geom_trimesh}}
+
+\hypertarget{coloured_long_edges}{%
+\paragraph{coloured\_long\_edges}\label{coloured_long_edges}}
+
+\hypertarget{remove-long-edges}{%
+\paragraph{remove long edges}\label{remove-long-edges}}
+
+\hypertarget{show_langevitour}{%
+\paragraph{show\_langevitour}\label{show_langevitour}}
 

### 2024-02-13 quollr_code.R

@@ -1,0 +1,1026 @@
+calculate_effective_x_bins <- function(.data, x = UMAP1, cell_area = 1){
+
+  if (any(is.na(.data |> dplyr::pull({{ x }})))) {
+    stop("NAs present")
+  }
+
+  if (any(is.infinite(.data |> dplyr::pull({{ x }})))) {
+    stop("Inf present")
+  }
+
+  if ((cell_area <= 0) || (is.infinite(cell_area))) {
+    stop("Invalid cell area value")
+
+  }
+
+  ## To compute the diameter of the hexagon
+  cell_diameter <- sqrt(2 * cell_area / sqrt(3))
+
+  ## To compute the range along x-axis
+  xwidth <- diff(range(.data |>
+                         dplyr::pull({{ x }})))
+
+  num_bins <- ceiling(xwidth/cell_diameter)
+  #num_bins <- floor(xwidth/cell_diameter + 1.5001)
+  num_bins
+
+}
+
+calculate_effective_shape_value <- function(.data, x = UMAP1, y = UMAP2){
+
+  if (any(is.na(.data |> dplyr::pull({{ x }}))) || any(is.na(.data |> dplyr::pull({{ y }})))) {
+    stop("NAs present")
+  }
+
+  if (any(is.infinite(.data |> dplyr::pull({{ x }}))) || any(is.infinite(.data |> dplyr::pull({{ y }})))) {
+    stop("Inf present")
+  }
+
+  if ((length(.data |> dplyr::pull({{ x }})) == 1) || (length(.data |> dplyr::pull({{ y }})) == 1)) {
+    stop("Presence one observation only")
+
+  }
+
+  ## To compute the range along x-axis
+  xwidth <- diff(range(.data |> dplyr::pull({{ x }})))
+  ## To compute the range along y-axis
+  yheight <- diff(range(.data |> dplyr::pull({{ y }})))
+
+
+  shape <- yheight/xwidth
+  shape
+}
+
+extract_hexbin_centroids <- function(nldr_df, num_bins, shape_val = 1, x = UMAP1, y = UMAP2) {
+
+  ## To create the hexbin object
+  hb_data <- hexbin::hexbin(x = nldr_df |> dplyr::pull({{ x }}),
+                            y = nldr_df |> dplyr::pull({{ y }}),
+                            xbins = num_bins, IDs = TRUE,
+                            shape = shape_val)
+
+  ## To create the hexbin centroid info dataset
[…]

### 2024-02-15 paper-quollr.Rmd
Main workflow
Develop the hexagonal binning code from scratch, write the main steps, write the shift fuction from the scratch as well

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, 
-                      warning = FALSE, 
+knitr::opts_chunk$set(warning = FALSE, 
                       message = FALSE)
@@
 ```{r load-libraries}
-library(quollr)
+#library(quollr)
+library(readr)
+library(ggplot2)
+library(dplyr)
+library(ggbeeswarm)
+```
+
+```{r}
+#| echo: false
+
+training_data <- read_rds(file = "data/s_curve_noise_training.rds")
+s_curve_noise_umap <- read_rds(file = "data/s_curve_noise_umap.rds")
 ```
@@
 
-# Implementation
+# Methodology
 
-## Installation
+## Usage
 
-The package can be installed from CRAN using
-
-> ```install.packages("quollr")```
-
-and from GitHub using
-
-> ```remotes::install_github("JayaniLakshika/quollr")```
-
-to install the development version.
-
-## Web site
-
-More documentation of the package can be found at the web site (need to add the link).
-
-## Data sets
-
-## Functions
+- dependancies
+- basic example
 
@@
 
+```{r}
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", cell_area = 1)
+num_bins_x
+```
+
+```{r}
+
+shape_val <- calculate_effective_shape_value(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2")
+shape_val
+```
+
+```{r}
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP2", y = "UMAP2", shape_val = 1.833091)
+num_bins_y
+```
+
 ### Generate full hex grid
+
+Generating full hexagonal grid contains main three steps:
+
+1. Generate all the hexagonal bin centroids
+
+Steps:
+
+- First compute hex grid bound values along the x and y axis and generate the all the points wthin the hex box
+
+```{r}
+cell_area <- 1
+cell_diameter <- sqrt(2 * cell_area / sqrt(3))
+
+hex_size <- cell_diameter/2
+
+buffer_size <- hex_size/2
+
+x_bounds <- seq(min(s_curve_noise_umap[["UMAP1"]]) - buffer_size,
+                  max(s_curve_noise_umap[["UMAP1"]]) + buffer_size, length.out = num_bins_x)
+
+y_bounds <- seq(min(s_curve_noise_umap[["UMAP2"]]) - buffer_size,
+                max(s_curve_noise_umap[["UMAP2"]]) + buffer_size, length.out = num_bins_y)
+
+box_points <- expand.grid(x = x_bounds, y = y_bounds)
+
+ggplot() +
+  geom_point(data = box_points, aes(x = x, y = y), color = "red")
+
+```
+
+- Second for each x-value, find which y values are in the even row
+
+```{r}
+ box_points <- box_points |>
+    dplyr::arrange(x) |>
+    dplyr::group_by(x) |>
+    dplyr::group_modify(~ generate_even_y(.x)) |>
+    tibble::as_tibble()
+
+ggplot() +
+  geom_point(data = box_points,
+             aes(x = x, y = y, colour = as.factor(is_even)))
+```
+
+- Then, shift the x values of the even rows
+
+```{r}
+## Shift for even values in x-axis
+x_shift <- unique(box_points$x)[2] - unique(box_points$x)[1]
+
+box_points$x <- box_points$x + x_shift/2 * ifelse(box_points$is_even == 1, 1, 0)
+
+ggplot() +
+  geom_point(data = box_points, aes(x = x, y = y), color = "red")
+
+```
+
+```{r}
+all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
+                                                 x = "UMAP1", y = "UMAP2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y, 
+                                                 buffer_size = NA, hex_size = NA)
+
+glimpse(all_centroids_df)
+```
+
+2. Generate hexagonal coordinates
+
+Steps:
+- Compute horizontal width of the hexagon
+
+- Compute vertical width of the hexagon and multiply by a factor for overlapping ($sqrt(3) / 2 * 1.15$)
+
+- Obtain hexagon polygon coordinates
+
+- Obtain the number of hexagons in the full grid
+
+- Generate the coordinates for the hexagons
+
+```{r}
+hex_grid <- gen_hex_coordinates(all_centroids_df)
+glimpse(hex_grid)
+```
+
+```{r}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
+```
+
+3. Map hexagonal IDs
+
+Steps:
+
+- Filter the data set with specific y value
+
+- Order the x values for a specific y value
+
+- Repeat the process for all unique y values
+
+```{r}
+full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
+
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
+```
+
+4. Map polygon IDs
+
+Steps:
+
+- Filter specific hexagon
+
+- Filter specific polygon
+
+- Check the selected hexagonal centroid exists within the polygon
+
+- if so assign that id to centroid, if not check until find the polygon which contains the centroid
+
+```{r}
+full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
+```
+
+4. Assign data into hexagons
+
+- Compute distances between nldr coordinates and hex bin centroids
+
+- Find the hexagonal centroid that have the minimum distance
+
+```{r}
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
+```
+
+5. Compute standardized counts
+
+- Compute number of data points within each hexagon
+
+- Compute standardise count by dividing the counts by the maximum
+
+```{r}
+df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+```
+
+6. Extract full grid info
+
+- Assign standardize counts for hex bins
+
+- Join with the hexagonal coordinates
+
+```{r}
+hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+```
+
+```{r}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
+```
+
+```{r}
+ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+
+```
+
+#### Buffer size
+
+When generating hexagonal bins in R, a buffer is often included to ensure that the data points are evenly distributed within the bins and to prevent edge effects. The buffer helps in two main ways:
+
+1.
+**Preventing Edge Effects**: Without a buffer, the outermost data points might fall near the boundary of the hexagonal grid, leading to incomplete bins or uneven distribution of data.
+By adding a buffer, you create a margin around the outer edges of the grid, ensuring that all data points are fully enclosed within the bins.
+
+2.
+**Ensuring Even Distribution**: The buffer allows for a smoother transition between adjacent bins.
+This helps in cases where data points are not perfectly aligned with the grid lines, ensuring that each data point is assigned to the nearest bin without bias towards any specific direction.
+
+Overall, including a buffer when generating hexagonal bins helps to produce more accurate and robust binning results, particularly when dealing with real-world data that may have irregular distributions or boundary effects.
 
@@
 
+```{r}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-ID) |>
+  dplyr::bind_cols(s_curve_noise_umap_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
+```
+
 ### Generate the triangular mesh
+
+```{r}
+df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
+  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
+  dplyr::distinct() |>
+  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
+
+df_bin_centroids
+```
+
+```{r}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
+```
 
@@
 
+1. Assign shift along the x and y axis (limited the amount should less than the cell_diameter)
+
+2. Generate bounds with shift origin
+
+```{r}
+all_centroids_df_shift <- extract_coord_of_shifted_hex_grid(nldr_df = s_curve_noise_umap, 
+                                                 x = "UMAP1", y = "UMAP2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y,
+                                                 shift_x = 0.2690002, shift_y = 0.271183,
+                                                 buffer_size = NA, hex_size = NA)
+
+glimpse(all_centroids_df_shift)
+```
+
+```{r}
+hex_grid <- gen_hex_coordinates(all_centroids_df_shift)
+glimpse(hex_grid)
+```
+
+```{r}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = all_centroids_df_shift, aes(x = x, y = y), color = "red")
+```
+
+```{r}
+full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df_shift)
+
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
+```
+
+```{r}
+full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
+```
+
+```{r}
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
+```
+
+```{r}
+df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+```
+
+```{r}
+hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+```
+
+```{r}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
+```
+
+```{r}
+ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+
+```
+```{r}
+df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
+  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
+  dplyr::distinct() |>
+  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
+
+df_bin_centroids
+```
+
+```{r}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
+```
+
+```{r, eval=FALSE}
+bin_centroids_shift <- ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
+  geom_point(color = "#bdbdbd") +
+  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "#feb24c") +
+  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+hex_grid_shift <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
+  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
+               fill = NA, color = "#bdbdbd") +
+  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+## Before shift
+before_shift_plot <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  theme_void() +
+  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+## After shift
+after_shift_plot <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
 #### Benchmark value to remove the low-density hexagons
@@
 #### Benchmark value to remove the long edges
+
+```{r}
+## Compute 2D distances
+distance <- cal_2d_dist(.data = tr_from_to_df)
+
+## To plot the distribution of distance
+plot_dist <- function(distance_df){
+  distance_df$group <- "1"
+  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
+    geom_quasirandom()+
+    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
+  return(dist_plot)
+}
+
+plot_dist(distance)
+
+benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+```
 
@@
 
+```{r}
+trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
+  geom_point(size = 0.1) +
+  geom_trimesh() +
+  coord_equal()
+
+trimesh
+```
+
 #### coloured_long_edges
+
+```{r}
+trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
+                                triangular_object = tr1_object, distance_col = distance)
+
+trimesh_gr
+```
 
@@
 
+```{r}
+trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
+                                     triangular_object = tr1_object, distance_col = distance)
+trimesh_removed
+```
+
 #### show_langevitour
+
+```{r}
+tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
+                          distance = distance, distance_col = distance)
+tour1
+```
 

### 2024-02-15 paper-quollr.tex
Develop the hexagonal binning code from scratch, write the main steps, write the shift fuction from the scratch as well

@@
 
+\begin{verbatim}
+#library(quollr)
+library(readr)
+library(ggplot2)
+library(dplyr)
+library(ggbeeswarm)
+\end{verbatim}
+
 \hypertarget{introduction}{%
@@
 
-\hypertarget{implementation}{%
-\section{Implementation}\label{implementation}}
+\hypertarget{methodology}{%
+\section{Methodology}\label{methodology}}
 
-\hypertarget{installation}{%
-\subsection{Installation}\label{installation}}
+\hypertarget{usage}{%
+\subsection{Usage}\label{usage}}
 
-The package can be installed from CRAN using
-
-\begin{quote}
-\texttt{install.packages("quollr")}
-\end{quote}
-
-and from GitHub using
-
-\begin{quote}
-\texttt{remotes::install\_github("JayaniLakshika/quollr")}
-\end{quote}
-
-to install the development version.
-
-\hypertarget{web-site}{%
-\subsection{Web site}\label{web-site}}
-
-More documentation of the package can be found at the web site (need to add the link).
-
-\hypertarget{data-sets}{%
-\subsection{Data sets}\label{data-sets}}
-
-\hypertarget{functions}{%
-\subsection{Functions}\label{functions}}
+\begin{itemize}
+\tightlist
+\item
+  dependancies
+\item
+  basic example
+\end{itemize}
 
@@
 
+\begin{verbatim}
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", cell_area = 1)
+num_bins_x
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 6
+\end{verbatim}
+
+\begin{verbatim}
+shape_val <- calculate_effective_shape_value(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2")
+shape_val
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 2.019414
+\end{verbatim}
+
+\begin{verbatim}
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP2", y = "UMAP2", shape_val = 1.833091)
+num_bins_y
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 12
+\end{verbatim}
+
 \hypertarget{generate-full-hex-grid}{%
 \subsubsection{Generate full hex grid}\label{generate-full-hex-grid}}
+
+Generating full hexagonal grid contains main three steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Generate all the hexagonal bin centroids
+\end{enumerate}
+
+Steps:
+
+\begin{itemize}
+\tightlist
+\item
+  First compute hex grid bound values along the x and y axis and generate the all the points wthin the hex box
+\end{itemize}
+
+\begin{verbatim}
+cell_area <- 1
+cell_diameter <- sqrt(2 * cell_area / sqrt(3))
+
+hex_size <- cell_diameter/2
+
+buffer_size <- hex_size/2
+
+x_bounds <- seq(min(s_curve_noise_umap[["UMAP1"]]) - buffer_size,
+                  max(s_curve_noise_umap[["UMAP1"]]) + buffer_size, length.out = num_bins_x)
+
+y_bounds <- seq(min(s_curve_noise_umap[["UMAP2"]]) - buffer_size,
+                max(s_curve_noise_umap[["UMAP2"]]) + buffer_size, length.out = num_bins_y)
+
+box_points <- expand.grid(x = x_bounds, y = y_bounds)
+
+ggplot() +
+  geom_point(data = box_points, aes(x = x, y = y), color = "red")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+\begin{itemize}
+\tightlist
+\item
+  Second for each x-value, find which y values are in the even row
+\end{itemize}
+
+\begin{verbatim}
+ box_points <- box_points |>
+    dplyr::arrange(x) |>
+    dplyr::group_by(x) |>
+    dplyr::group_modify(~ generate_even_y(.x)) |>
+    tibble::as_tibble()
+
+ggplot() +
+  geom_point(data = box_points,
+             aes(x = x, y = y, colour = as.factor(is_even)))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+\begin{itemize}
+\tightlist
+\item
+  Then, shift the x values of the even rows
+\end{itemize}
+
+\begin{verbatim}
+## Shift for even values in x-axis
+x_shift <- unique(box_points$x)[2] - unique(box_points$x)[1]
+
+box_points$x <- box_points$x + x_shift/2 * ifelse(box_points$is_even == 1, 1, 0)
+
+ggplot() +
+  geom_point(data = box_points, aes(x = x, y = y), color = "red")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+\begin{verbatim}
+all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
+                                                 x = "UMAP1", y = "UMAP2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y, 
+                                                 buffer_size = NA, hex_size = NA)
+
+glimpse(all_centroids_df)
+\end{verbatim}
+
+\begin{verbatim}
+#> Rows: 72
+#> Columns: 2
+#> $ x <dbl> -3.5390002, -2.9126765, -3.5390002, -2.9126765, -3.5390002, -2.91267~
+#> $ y <dbl> -6.0111830, -4.9111506, -3.8111182, -2.7110858, -1.6110534, -0.51102~
+\end{verbatim}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Generate hexagonal coordinates
+\end{enumerate}
+
+Steps:
+- Compute horizontal width of the hexagon
+
+\begin{itemize}
+\item
+  Compute vertical width of the hexagon and multiply by a factor for overlapping (\(sqrt(3) / 2 * 1.15\))
+\item
+  Obtain hexagon polygon coordinates
+\item
+  Obtain the number of hexagons in the full grid
+\item
+  Generate the coordinates for the hexagons
+\end{itemize}
+
+\begin{verbatim}
+hex_grid <- gen_hex_coordinates(all_centroids_df)
+glimpse(hex_grid)
+\end{verbatim}
+
+\begin{verbatim}
+#> Rows: 432
+#> Columns: 3
+#> $ x  <dbl> -2.912676, -2.912676, -3.539000, -4.165324, -4.165324, -3.539000, -~
+#> $ y  <dbl> -5.645998, -6.376368, -6.741553, -6.376368, -5.645998, -5.280813, -~
+#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
+\end{verbatim}
+
+\begin{verbatim}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Map hexagonal IDs
+\end{enumerate}
+
+Steps:
+
+\begin{itemize}
+\item
+  Filter the data set with specific y value
+\item
+  Order the x values for a specific y value
+\item
+  Repeat the process for all unique y values
+\end{itemize}
+
+\begin{verbatim}
+full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
+
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Map polygon IDs
+\end{enumerate}
+
+Steps:
+
+\begin{itemize}
+\item
+  Filter specific hexagon
+\item
+  Filter specific polygon
+\item
+  Check the selected hexagonal centroid exists within the polygon
+\item
+  if so assign that id to centroid, if not check until find the polygon which contains the centroid
+\end{itemize}
+
+\begin{verbatim}
+full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
+\end{verbatim}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Assign data into hexagons
+\end{enumerate}
+
+\begin{itemize}
+\item
+  Compute distances between nldr coordinates and hex bin centroids
+\item
+  Find the hexagonal centroid that have the minimum distance
+\end{itemize}
+
+\begin{verbatim}
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
+\end{verbatim}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  Compute standardized counts
+\end{enumerate}
+
+\begin{itemize}
+\item
+  Compute number of data points within each hexagon
+\item
+  Compute standardise count by dividing the counts by the maximum
+\end{itemize}
+
+\begin{verbatim}
+df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+\end{verbatim}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{5}
+\tightlist
+\item
+  Extract full grid info
+\end{enumerate}
+
+\begin{itemize}
+\item
+  Assign standardize counts for hex bins
+\item
+  Join with the hexagonal coordinates
+\end{itemize}
+
+\begin{verbatim}
+hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+\end{verbatim}
+
+\begin{verbatim}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}
+
+\begin{verbatim}
+ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}
+
+\hypertarget{buffer-size}{%
+\paragraph{Buffer size}\label{buffer-size}}
+
+When generating hexagonal bins in R, a buffer is often included to ensure that the data points are evenly distributed within the bins and to prevent edge effects. The buffer helps in two main ways:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Preventing Edge Effects}: Without a buffer, the outermost data points might fall near the boundary of the hexagonal grid, leading to incomplete bins or uneven distribution of data.
+By adding a buffer, you create a margin around the outer edges of the grid, ensuring that all data points are fully enclosed within the bins.
+\item
+  \textbf{Ensuring Even Distribution}: The buffer allows for a smoother transition between adjacent bins.
+This helps in cases where data points are not perfectly aligned with the grid lines, ensuring that each data point is assigned to the nearest bin without bias towards any specific direction.
+\end{enumerate}
+
+Overall, including a buffer when generating hexagonal bins helps to produce more accurate and robust binning results, particularly when dealing with real-world data that may have irregular distributions or boundary effects.
 
@@
 
+\begin{verbatim}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-ID) |>
+  dplyr::bind_cols(s_curve_noise_umap_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
+\end{verbatim}
+
 \hypertarget{generate-the-triangular-mesh}{%
 \subsubsection{Generate the triangular mesh}\label{generate-the-triangular-mesh}}
+
+\begin{verbatim}
+df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
+  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
+  dplyr::distinct() |>
+  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
+
+df_bin_centroids
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 20 x 4
+#>         x      y hexID std_counts
+#>     <dbl>  <dbl> <int>      <dbl>
+#>  1 -2.91  -4.91      7      0.5  
+#>  2 -3.54  -3.81     13      0.5  
+#>  3 -2.29  -6.01      2      0.5  
+#>  4 -1.66  -4.91      8      0.75 
+#>  5 -2.29  -3.81     14      0.5  
+#>  6 -1.03  -3.81     15      0.375
+#>  7 -0.407 -2.71     21      0.125
+#>  8 -1.03  -1.61     27      0.125
+#>  9  0.219 -1.61     28      1    
+#> 10  0.845 -0.511    34      0.625
+#> 11  0.845  1.69     46      0.125
+#> 12  0.845  3.89     58      0.625
+#> 13  0.219  4.99     64      0.125
+#> 14  0.845  6.09     70      0.375
+#> 15  1.47   0.589    41      1    
+#> 16  1.47   2.79     53      0.125
+#> 17  2.10   3.89     59      0.625
+#> 18  1.47   4.99     65      0.5  
+#> 19  2.10   6.09     71      0.375
+#> 20  2.72   4.99     66      0.5
+\end{verbatim}
+
+\begin{verbatim}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
+\end{verbatim}
 
@@
 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Assign shift along the x and y axis (limited the amount should less than the cell\_diameter)
+\item
+  Generate bounds with shift origin
+\end{enumerate}
+
+\begin{verbatim}
+all_centroids_df_shift <- extract_coord_of_shifted_hex_grid(nldr_df = s_curve_noise_umap, 
+                                                 x = "UMAP1", y = "UMAP2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y,
+                                                 shift_x = 0.2690002, shift_y = 0.271183,
+                                                 buffer_size = NA, hex_size = NA)
+
+glimpse(all_centroids_df_shift)
+\end{verbatim}
+
+\begin{verbatim}
+#> Rows: 72
+#> Columns: 2
+#> $ x <dbl> -3.2700000, -2.6436763, -3.2700000, -2.6436763, -3.2700000, -2.64367~
+#> $ y <dbl> -5.7400000, -4.6399676, -3.5399352, -2.4399028, -1.3398704, -0.23983~
+\end{verbatim}
+
+\begin{verbatim}
+hex_grid <- gen_hex_coordinates(all_centroids_df_shift)
+glimpse(hex_grid)
+\end{verbatim}
+
+\begin{verbatim}
+#> Rows: 432
+#> Columns: 3
+#> $ x  <dbl> -2.643676, -2.643676, -3.270000, -3.896324, -3.896324, -3.270000, -~
+#> $ y  <dbl> -5.3748152, -6.1051848, -6.4703696, -6.1051848, -5.3748152, -5.0096~
+#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
+\end{verbatim}
+
+\begin{verbatim}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = all_centroids_df_shift, aes(x = x, y = y), color = "red")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-23-1.pdf}
+
+\begin{verbatim}
+full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df_shift)
+
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-24-1.pdf}
+
+\begin{verbatim}
+full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
+\end{verbatim}
+
+\begin{verbatim}
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
+\end{verbatim}
+
+\begin{verbatim}
+df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+\end{verbatim}
+
+\begin{verbatim}
+hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+\end{verbatim}
+
+\begin{verbatim}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-29-1.pdf}
+
+\begin{verbatim}
+ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-30-1.pdf}
+
+\begin{verbatim}
+df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
+  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
+  dplyr::distinct() |>
+  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
+
+df_bin_centroids
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 21 x 4
+#>         x     y hexID std_counts
+#>     <dbl> <dbl> <int>      <dbl>
+#>  1 -3.27  -5.74     1      0.286
+#>  2 -2.64  -4.64     7      0.571
+#>  3 -3.27  -3.54    13      0.714
+#>  4 -2.02  -5.74     2      0.857
+#>  5 -1.39  -4.64     8      0.429
+#>  6 -2.02  -3.54    14      0.286
+#>  7 -0.765 -3.54    15      0.429
+#>  8 -0.138 -2.44    21      0.143
+#>  9 -0.765 -1.34    27      0.286
+#> 10  0.488 -1.34    28      1    
+#> # i 11 more rows
+\end{verbatim}
+
+\begin{verbatim}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
+\end{verbatim}
+
+\begin{verbatim}
+bin_centroids_shift <- ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
+  geom_point(color = "#bdbdbd") +
+  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "#feb24c") +
+  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+hex_grid_shift <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
+  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
+               fill = NA, color = "#bdbdbd") +
+  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+## Before shift
+before_shift_plot <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  theme_void() +
+  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+## After shift
+after_shift_plot <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
+\end{verbatim}
+
 \hypertarget{benchmark-value-to-remove-the-low-density-hexagons}{%
@@
 \paragraph{Benchmark value to remove the long edges}\label{benchmark-value-to-remove-the-long-edges}}
+
+\begin{verbatim}
+## Compute 2D distances
+distance <- cal_2d_dist(.data = tr_from_to_df)
+
+## To plot the distribution of distance
+plot_dist <- function(distance_df){
+  distance_df$group <- "1"
+  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
+    geom_quasirandom()+
+    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
+  return(dist_plot)
+}
+
+plot_dist(distance)
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-34-1.pdf}
+
+\begin{verbatim}
+benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+\end{verbatim}
 
@@
 
+\begin{verbatim}
+trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
+  geom_point(size = 0.1) +
+  geom_trimesh() +
+  coord_equal()
+
+trimesh
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-35-1.pdf}
+
 \hypertarget{coloured_long_edges}{%
 \paragraph{coloured\_long\_edges}\label{coloured_long_edges}}
+
+\begin{verbatim}
+trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
+                                triangular_object = tr1_object, distance_col = distance)
+
+trimesh_gr
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-36-1.pdf}
 
@@
 
+\begin{verbatim}
+trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
+                                     triangular_object = tr1_object, distance_col = distance)
+trimesh_removed
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-37-1.pdf}
+
 \hypertarget{show_langevitour}{%
 \paragraph{show\_langevitour}\label{show_langevitour}}
+
+\begin{verbatim}
+tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
+                          distance = distance, distance_col = distance)
+tour1
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-38-1.pdf}
 

### 2024-02-15 quollr_code.R

@@ -1,6 +1,6 @@
-calculate_effective_x_bins <- function(.data, x = UMAP1, cell_area = 1){
+calculate_effective_x_bins <- function(.data, x = "UMAP1", cell_area = 1){
 
-  if (any(is.na(.data |> dplyr::pull({{ x }})))) {
+  if (anyNA(.data[[rlang::as_string(rlang::ensym(x))]])) {
     stop("NAs present")
   }
 
@@ -20,8 +20,28 @@ calculate_effective_x_bins <- function(.data, x = UMAP1, cell_area = 1){
   xwidth <- diff(range(.data |>
                          dplyr::pull({{ x }})))
 
-  num_bins <- ceiling(xwidth/cell_diameter)
-  #num_bins <- floor(xwidth/cell_diameter + 1.5001)
+  #num_bins <- ceiling(xwidth/cell_diameter)
+  num_bins <- floor(xwidth/cell_diameter + 1.5001)
+  num_bins
+
+}
+
+calculate_effective_y_bins <- function(.data, x = "UMAP2", y = "UMAP2", shape_val = NA){
+
+  if (anyNA(.data[[rlang::as_string(rlang::ensym(y))]])) {
+    stop("NAs present")
+  }
+
+  if (any(is.infinite(.data |> dplyr::pull({{ y }})))) {
+    stop("Inf present")
+  }
+
+  if (is.na(shape_val)) {
+    shape_val <- calculate_effective_shape_value(.data = .data, x = x, y = y)
+
+  }
+
+  num_bins <- 2 * floor(((num_bins_x - 1) * shape_val)/sqrt(3) + 1.5001)
   num_bins
 
 }
@@ -50,6 +70,194 @@ calculate_effective_shape_value <- function(.data, x = UMAP1, y = UMAP2){
   shape <- yheight/xwidth
   shape
 }
+
+
+# Function to generate even y-values for a given x-value
+generate_even_y <- function(data) {
+  data <- data |>
+    dplyr::arrange(y) |>
+    dplyr::mutate(is_even = ifelse(seq_along(y) %% 2 == 0, 1, 0))
+
+  return(data)
+}
+
+generate_full_grid_centroids <- function(nldr_df, x = "UMAP1", y = "UMAP2",
+                                         num_bins_x, num_bins_y, buffer_size = NA, hex_size = NA){
+
+  ## If number of bins along the x-axis is not given
+  if (is.na(num_bins_x)) {
+    ## compute the number of bins along the x-axis
[…]

### 2024-02-15 paper-quollr.Rmd
Added updated benchmark function

@@
 - dependancies
+
+```{r, eval=FALSE}
+library(tools)
+package_dependencies("quollr")
+```
+
 - basic example
@@
 
+If shift_x happen to the positive direction of x it should input as a positive value, if not other way
+If shift_y happen to the positive direction of y it should input as a positive value, if not other way
+
 1. Assign shift along the x and y axis (limited the amount should less than the cell_diameter)
@@
 tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
-                          distance = distance, distance_col = distance)
+                          distance = distance, distance_col = "distance")
 tour1

### 2024-02-15 paper-quollr.tex
Added updated benchmark function

@@
   dependancies
+\end{itemize}
+
+\begin{verbatim}
+library(tools)
+package_dependencies("quollr")
+\end{verbatim}
+
+\begin{itemize}
+\tightlist
 \item
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-10-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-18-1.pdf}
 
@@
 
+If shift\_x happen to the positive direction of x it should input as a positive value, if not other way
+If shift\_y happen to the positive direction of y it should input as a positive value, if not other way
+
 \begin{enumerate}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-23-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-24-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-25-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-29-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-30-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-30-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-31-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-34-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-35-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-35-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-36-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-36-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-37-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-37-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-38-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-38-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-39-1.pdf}
 

### 2024-02-15 paper-quollr.R

@@ -22,6 +22,11 @@ training_data <- read_rds(file = "data/s_curve_noise_training.rds")
 s_curve_noise_umap <- read_rds(file = "data/s_curve_noise_umap.rds")
 
 
+## ----eval=FALSE---------------------------------------------------------------
+#> library(tools)
+#> package_dependencies("quollr")
+
+
 ## -----------------------------------------------------------------------------
 num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", cell_area = 1)
 num_bins_x
@@ -344,6 +349,6 @@ trimesh_removed
 
 ## -----------------------------------------------------------------------------
 tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
-                          distance = distance, distance_col = distance)
+                          distance = distance, distance_col = "distance")
 tour1
 

### 2024-02-16 paper-quollr.Rmd
Add application MedLEA data

@@
 source("quollr_code.R", local = TRUE)
+source("nldr_code.R", local = TRUE)
 ```
@@
 library(ggbeeswarm)
+library(Rtsne)
+library(umap)
+library(phateR)
+library(reticulate)
+library(rsample)
+
+set.seed(20230531)
+
+use_python("~/miniforge3/envs/pcamp_env/bin/python")
+use_condaenv("pcamp_env")
+
+reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
+reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
+
 ```
@@
 ```{r}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP2", y = "UMAP2", shape_val = 1.833091)
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2", shape_val = 1.833091, num_bins_x = num_bins_x)
 num_bins_y
@@
 benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+benchmark <- 3
 ```
@@
 trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
-                                triangular_object = tr1_object, distance_col = distance)
+                                triangular_object = tr1_object, distance_col = "distance")
 
@@
 trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
-                                     triangular_object = tr1_object, distance_col = distance)
+                                     triangular_object = tr1_object, distance_col = "distance")
 trimesh_removed
@@
 #### show_langevitour
+
+```{r}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-ID) |>
+  dplyr::bind_cols(s_curve_noise_umap_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
+```
 
@@
 
-# Examples
+# Application
+
+```{r}
+medlea_df <- read_csv("data/medlea_dataset.csv")
+names(medlea_df)[2:(NCOL(medlea_df) - 1)] <- paste0("x", 1:(NCOL(medlea_df) - 2))
+
+medlea_df <- medlea_df |> ## Since only contains zeros
+  select(-x10)
+
+#medlea_df[,2:(NCOL(medlea_df) - 1)] <- scale(medlea_df[,2:(NCOL(medlea_df) - 1)])
+
+calculate_pca <- function(feature_dataset, num_pcs){
+  pcaY_cal <- prcomp(feature_dataset, center = TRUE, scale = TRUE)
+  PCAresults <- data.frame(pcaY_cal$x[, 1:num_pcs])
+  summary_pca <- summary(pcaY_cal)
+  var_explained_df <- data.frame(PC= paste0("PC",1:50),
+                               var_explained=(pcaY_cal$sdev[1:50])^2/sum((pcaY_cal$sdev[1:50])^2))
+  return(list(prcomp_out = pcaY_cal,pca_components = PCAresults, summary = summary_pca, var_explained_pca  = var_explained_df))
+}
+features <- medlea_df[,2:(NCOL(medlea_df) - 1)]
+pca_ref_calc <- calculate_pca(features, 8) 
+pca_ref_calc$summary
+
+var_explained_df <- pca_ref_calc$var_explained_pca
+data_pca <- pca_ref_calc$pca_components |>
+  mutate(ID = 1:NROW(pca_ref_calc$pca_components),
+         shape_label = medlea_df$Shape_label)
+
+var_explained_df |>
+  ggplot(aes(x = PC,y = var_explained, group = 1))+
+  geom_point(size=1)+
+  geom_line()+
+  labs(title="Scree plot: PCA on scaled data") +
+  scale_x_discrete(limits = paste0(rep("PC", 50), 1:50)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+data_split <- initial_split(data_pca)
+training_data <- training(data_split) |>
+  arrange(ID)
+test_data <- testing(data_split) |>
+  arrange(ID)
+```
+
+```{r}
+UMAP_fit <- umap(training_data |> dplyr::select(-c(ID, shape_label)), n_neighbors = 37, n_components =  2)
+
+UMAP_data <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
+
+UMAP_data <- UMAP_data |>
+  mutate(ID = training_data$id)
+
+UMAP_data_with_label <- UMAP_data |>
+  mutate(shape_label = training_data$shape_label)
+```
+
+```{r}
+UMAP_data_with_label |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+```
+
+```{r}
+tSNE_data <- Fit_tSNE(training_data |> dplyr::select(-c(ID, shape_label)), opt_perplexity = calculate_effective_perplexity(training_data |> dplyr::select(-c(ID, shape_label))), with_seed = 20240110)
+
+tSNE_data <- tSNE_data |>
+  select(-ID) |>
+  mutate(ID = training_data$ID)
+
+tSNE_data_with_label <- tSNE_data |>
+  mutate(shape_label = training_data$shape_label)
+
+tSNE_data_with_label |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+```
+
+```{r}
+PHATE_data <- Fit_PHATE(training_data |> dplyr::select(-c(ID, shape_label)), knn = 5, with_seed = 20240110)
+PHATE_data <- PHATE_data |>
+  select(PHATE1, PHATE2)
+PHATE_data <- PHATE_data |>
+  mutate(ID = training_data$ID)
+
+PHATE_data_with_label <- PHATE_data |>
+  mutate(shape_label = training_data$shape_label)
+
+PHATE_data_with_label |>
+    ggplot(aes(x = PHATE1,
+               y = PHATE2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+```
+
+```{r}
+tem_dir <- tempdir()
+
+Fit_TriMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
+
+path <- file.path(tem_dir, "df_2_without_class.csv")
+path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
+
+Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
+
+TriMAP_data <- read_csv(path2)
+TriMAP_data <- TriMAP_data |>
+  mutate(ID = training_data$ID)
+
+TriMAP_data_with_label <- TriMAP_data |>
+  mutate(shape_label = training_data$shape_label)
+
+TriMAP_data_with_label |>
+    ggplot(aes(x = TriMAP1,
+               y = TriMAP2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+
+```
+
+```{r}
+tem_dir <- tempdir()
+
+Fit_PacMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
+
+path <- file.path(tem_dir, "df_2_without_class.csv")
+path2 <- file.path(tem_dir, "dataset_3_PaCMAP_values.csv")
+
+Fit_PaCMAP(as.integer(2), as.integer(10), "random", 0.9, as.integer(2), path, path2)
+
+PaCMAP_data <- read_csv(path2)
+PaCMAP_data <- PaCMAP_data |>
+  mutate(ID = training_data$ID)
+
+PaCMAP_data_with_label <- PaCMAP_data |>
+  mutate(shape_label = training_data$shape_label)
+
+PaCMAP_data_with_label |>
+    ggplot(aes(x = PaCMAP1,
+               y = PaCMAP2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+```
+
+```{r}
+num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", cell_area = 1)
+num_bins_x <- 13
+```
+
+```{r}
+shape_val <- calculate_effective_shape_value(.data = tSNE_data, x = "tSNE1", y = "tSNE2")
+shape_val
+```
+
+```{r}
+num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, x = "tSNE1", y = "tSNE2", shape_val = shape_val, num_bins_x = num_bins_x)
+num_bins_y
+```
+
+```{r}
+all_centroids_df <- generate_full_grid_centroids(nldr_df = tSNE_data, 
+                                                 x = "tSNE1", y = "tSNE2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y, 
+                                                 buffer_size = NA, hex_size = NA)
+
+hex_grid <- gen_hex_coordinates(all_centroids_df)
+
+full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
+
+full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
+
+tSNE_data_with_id <- assign_data(tSNE_data, full_grid_with_hexbin_id)
+
+df_with_std_counts <- compute_std_counts(nldr_df = tSNE_data_with_id)
+
+hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+
+ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+```
+
+```{r}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = tSNE_data, aes(x = tSNE1, y = tSNE2), color = "blue")
+```
+
+```{r}
+df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
+  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
+  dplyr::distinct() |>
+  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
+
+df_bin_centroids
+```
+
+```{r}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
+```
+
+```{r}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-c(ID, shape_label)) |>
+  dplyr::bind_cols(tSNE_data_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "PC") ## Need to pass ID column name
+```
+
+```{r}
+## Compute 2D distances
+distance <- cal_2d_dist(.data = tr_from_to_df)
+
+plot_dist(distance)
+
+benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+```
+
+```{r}
+trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
+  geom_point(size = 0.1) +
+  geom_trimesh() +
+  coord_equal()
+
+trimesh
+```
+
+```{r}
+trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
+                                triangular_object = tr1_object, distance_col = "distance")
+
+trimesh_gr
+```
+
+```{r}
+trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
+                                     triangular_object = tr1_object, distance_col = "distance")
+trimesh_removed
+```
+
+```{r}
+tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
+                          distance = distance, distance_col = "distance", column_start_text = "PC")
+tour1
+```
 

### 2024-02-16 paper-quollr.tex
Add application MedLEA data

@@
 library(ggbeeswarm)
+library(Rtsne)
+library(umap)
+library(phateR)
+library(reticulate)
+library(rsample)
+
+set.seed(20230531)
+
+use_python("~/miniforge3/envs/pcamp_env/bin/python")
+use_condaenv("pcamp_env")
+
+reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
+reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
 \end{verbatim}
@@
 \begin{verbatim}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP2", y = "UMAP2", shape_val = 1.833091)
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2", shape_val = 1.833091, num_bins_x = num_bins_x)
 num_bins_y
@@
 
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
+\hypertarget{application}{%
+\section{Application}\label{application}}
+
+\begin{verbatim}
+medlea_df <- read_csv("data/medlea_dataset.csv")
+names(medlea_df)[2:(NCOL(medlea_df) - 1)] <- paste0("x", 1:(NCOL(medlea_df) - 2))
+
+medlea_df <- medlea_df |> ## Since only contains zeros
+  select(-x10)
+
+#medlea_df[,2:(NCOL(medlea_df) - 1)] <- scale(medlea_df[,2:(NCOL(medlea_df) - 1)])
+
+calculate_pca <- function(feature_dataset, num_pcs){
+  pcaY_cal <- prcomp(feature_dataset, center = TRUE, scale = TRUE)
+  PCAresults <- data.frame(pcaY_cal$x[, 1:num_pcs])
+  summary_pca <- summary(pcaY_cal)
+  var_explained_df <- data.frame(PC= paste0("PC",1:50),
+                               var_explained=(pcaY_cal$sdev[1:50])^2/sum((pcaY_cal$sdev[1:50])^2))
+  return(list(prcomp_out = pcaY_cal,pca_components = PCAresults, summary = summary_pca, var_explained_pca  = var_explained_df))
+}
+features <- medlea_df[,2:(NCOL(medlea_df) - 1)]
+pca_ref_calc <- calculate_pca(features, 8) 
+pca_ref_calc$summary
+\end{verbatim}
+
+\begin{verbatim}
+#> Importance of components:
+#>                           PC1    PC2    PC3     PC4     PC5     PC6     PC7
+#> Standard deviation     3.1691 3.0609 2.7226 1.87967 1.71219 1.34192 1.27525
+#> Proportion of Variance 0.1969 0.1837 0.1453 0.06928 0.05748 0.03531 0.03189
+#> Cumulative Proportion  0.1969 0.3806 0.5260 0.59526 0.65274 0.68805 0.71993
+#>                            PC8     PC9    PC10    PC11    PC12    PC13   PC14
+#> Standard deviation     1.16992 1.13465 1.06628 1.03279 0.97899 0.96264 0.9528
+#> Proportion of Variance 0.02684 0.02524 0.02229 0.02091 0.01879 0.01817 0.0178
+#> Cumulative Proportion  0.74677 0.77202 0.79431 0.81522 0.83402 0.85219 0.8700
+#>                          PC15   PC16    PC17    PC18    PC19    PC20    PC21
+#> Standard deviation     0.9116 0.9090 0.79750 0.76725 0.72414 0.65310 0.61052
+#> Proportion of Variance 0.0163 0.0162 0.01247 0.01154 0.01028 0.00836 0.00731
+#> Cumulative Proportion  0.8863 0.9025 0.91496 0.92650 0.93678 0.94514 0.95245
+#>                          PC22    PC23    PC24    PC25    PC26   PC27    PC28
+#> Standard deviation     0.6019 0.55399 0.52293 0.46638 0.41959 0.3976 0.34697
+#> Proportion of Variance 0.0071 0.00602 0.00536 0.00426 0.00345 0.0031 0.00236
+#> Cumulative Proportion  0.9596 0.96557 0.97093 0.97520 0.97865 0.9818 0.98411
+#>                           PC29    PC30    PC31    PC32    PC33    PC34    PC35
+#> Standard deviation     0.33415 0.30618 0.29237 0.28458 0.26033 0.25420 0.22792
+#> Proportion of Variance 0.00219 0.00184 0.00168 0.00159 0.00133 0.00127 0.00102
+#> Cumulative Proportion  0.98630 0.98814 0.98982 0.99140 0.99273 0.99400 0.99502
+#>                           PC36    PC37    PC38   PC39    PC40    PC41    PC42
+#> Standard deviation     0.21644 0.20437 0.19127 0.1744 0.15586 0.15252 0.12519
+#> Proportion of Variance 0.00092 0.00082 0.00072 0.0006 0.00048 0.00046 0.00031
+#> Cumulative Proportion  0.99594 0.99676 0.99747 0.9981 0.99855 0.99900 0.99931
+#>                           PC43    PC44    PC45    PC46    PC47    PC48    PC49
+#> Standard deviation     0.10485 0.08598 0.08008 0.06491 0.04841 0.04094 0.03791
+#> Proportion of Variance 0.00022 0.00014 0.00013 0.00008 0.00005 0.00003 0.00003
+#> Cumulative Proportion  0.99952 0.99967 0.99980 0.99988 0.99992 0.99996 0.99999
+#>                           PC50    PC51
+#> Standard deviation     0.02347 0.01421
+#> Proportion of Variance 0.00001 0.00000
+#> Cumulative Proportion  1.00000 1.00000
+\end{verbatim}
+
+\begin{verbatim}
+var_explained_df <- pca_ref_calc$var_explained_pca
+data_pca <- pca_ref_calc$pca_components |>
+  mutate(ID = 1:NROW(pca_ref_calc$pca_components),
+         shape_label = medlea_df$Shape_label)
+
+var_explained_df |>
+  ggplot(aes(x = PC,y = var_explained, group = 1))+
+  geom_point(size=1)+
+  geom_line()+
+  labs(title="Scree plot: PCA on scaled data") +
+  scale_x_discrete(limits = paste0(rep("PC", 50), 1:50)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-40-1.pdf}
+
+\begin{verbatim}
+data_split <- initial_split(data_pca)
+training_data <- training(data_split) |>
+  arrange(ID)
+test_data <- testing(data_split) |>
+  arrange(ID)
+\end{verbatim}
+
+\begin{verbatim}
+UMAP_fit <- umap(training_data |> dplyr::select(-c(ID, shape_label)), n_neighbors = 37, n_components =  2)
+
+UMAP_data <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
+
+UMAP_data <- UMAP_data |>
+  mutate(ID = training_data$id)
+
+UMAP_data_with_label <- UMAP_data |>
+  mutate(shape_label = training_data$shape_label)
+\end{verbatim}
+
+\begin{verbatim}
+UMAP_data_with_label |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-42-1.pdf}
+
+\begin{verbatim}
+tSNE_data <- Fit_tSNE(training_data |> dplyr::select(-c(ID, shape_label)), opt_perplexity = calculate_effective_perplexity(training_data |> dplyr::select(-c(ID, shape_label))), with_seed = 20240110)
+
+tSNE_data <- tSNE_data |>
+  select(-ID) |>
+  mutate(ID = training_data$ID)
+
+tSNE_data_with_label <- tSNE_data |>
+  mutate(shape_label = training_data$shape_label)
+
+tSNE_data_with_label |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-43-1.pdf}
+
+\begin{verbatim}
+PHATE_data <- Fit_PHATE(training_data |> dplyr::select(-c(ID, shape_label)), knn = 5, with_seed = 20240110)
+\end{verbatim}
+
+\begin{verbatim}
+#> Calculating PHATE...
+#>   Running PHATE on 824 observations and 8 variables.
+#>   Calculating graph and diffusion operator...
+#>     Calculating KNN search...
+#>     Calculating affinities...
+#>   Calculated graph and diffusion operator in 0.01 seconds.
+#>   Calculating optimal t...
+#>     Automatically selected t = 22
+#>   Calculated optimal t in 0.39 seconds.
+#>   Calculating diffusion potential...
+#>   Calculated diffusion potential in 0.32 seconds.
+#>   Calculating metric MDS...
+#>   Calculated metric MDS in 8.83 seconds.
+#> Calculated PHATE in 9.55 seconds.
+\end{verbatim}
+
+\begin{verbatim}
+PHATE_data <- PHATE_data |>
+  select(PHATE1, PHATE2)
+PHATE_data <- PHATE_data |>
+  mutate(ID = training_data$ID)
+
+PHATE_data_with_label <- PHATE_data |>
+  mutate(shape_label = training_data$shape_label)
+
+PHATE_data_with_label |>
+    ggplot(aes(x = PHATE1,
+               y = PHATE2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-44-1.pdf}
+
+\begin{verbatim}
+tem_dir <- tempdir()
+
+Fit_TriMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
+
+path <- file.path(tem_dir, "df_2_without_class.csv")
+path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
+
+Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
+
+TriMAP_data <- read_csv(path2)
+TriMAP_data <- TriMAP_data |>
+  mutate(ID = training_data$ID)
+
+TriMAP_data_with_label <- TriMAP_data |>
+  mutate(shape_label = training_data$shape_label)
+
+TriMAP_data_with_label |>
+    ggplot(aes(x = TriMAP1,
+               y = TriMAP2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-45-1.pdf}
+
+\begin{verbatim}
+tem_dir <- tempdir()
+
+Fit_PacMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
+
+path <- file.path(tem_dir, "df_2_without_class.csv")
+path2 <- file.path(tem_dir, "dataset_3_PaCMAP_values.csv")
+
+Fit_PaCMAP(as.integer(2), as.integer(10), "random", 0.9, as.integer(2), path, path2)
+
+PaCMAP_data <- read_csv(path2)
+PaCMAP_data <- PaCMAP_data |>
+  mutate(ID = training_data$ID)
+
+PaCMAP_data_with_label <- PaCMAP_data |>
+  mutate(shape_label = training_data$shape_label)
+
+PaCMAP_data_with_label |>
+    ggplot(aes(x = PaCMAP1,
+               y = PaCMAP2, color = shape_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}
+
+\begin{verbatim}
+num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", cell_area = 1)
+num_bins_x <- 13
+\end{verbatim}
+
+\begin{verbatim}
+shape_val <- calculate_effective_shape_value(.data = tSNE_data, x = "tSNE1", y = "tSNE2")
+shape_val
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 1.13882
+\end{verbatim}
+
+\begin{verbatim}
+num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, x = "tSNE1", y = "tSNE2", shape_val = 0.8417289, num_bins_x = num_bins_x)
+num_bins_y
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 14
+\end{verbatim}
+
+\begin{verbatim}
+all_centroids_df <- generate_full_grid_centroids(nldr_df = tSNE_data, 
+                                                 x = "tSNE1", y = "tSNE2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y, 
+                                                 buffer_size = NA, hex_size = NA)
+
+hex_grid <- gen_hex_coordinates(all_centroids_df)
+
+full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
+
+full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
+
+tSNE_data_with_id <- assign_data(tSNE_data, full_grid_with_hexbin_id)
+
+df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+
+hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+
+ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-50-1.pdf}
+
+\begin{verbatim}
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = tSNE_data, aes(x = tSNE1, y = tSNE2), color = "blue")
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-51-1.pdf}
+
+\begin{verbatim}
+df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
+  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
+  dplyr::distinct() |>
+  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
+
+df_bin_centroids
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 21 x 4
+#>        x     y hexID std_counts
+#>    <dbl> <dbl> <int>      <dbl>
+#>  1 -27.7 -34.8     1      0.286
+#>  2 -25.4 -30.1    14      0.286
+#>  3 -27.7 -25.4    27      0.286
+#>  4 -25.4 -20.7    40      0.286
+#>  5 -27.7 -16.0    53      0.429
+#>  6 -25.4 -11.3    66      0.143
+#>  7 -23.2 -34.8     2      0.857
+#>  8 -20.9 -30.1    15      0.429
+#>  9 -23.2 -25.4    28      1    
+#> 10 -20.9 -20.7    41      0.714
+#> # i 11 more rows
+\end{verbatim}
+
+\begin{verbatim}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
+\end{verbatim}
+
+\begin{verbatim}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-c(ID, shape_label)) |>
+  dplyr::bind_cols(tSNE_data_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "PC") ## Need to pass ID column name
+\end{verbatim}
+
+\begin{verbatim}
+## Compute 2D distances
+distance <- cal_2d_dist(.data = tr_from_to_df)
+
+plot_dist(distance)
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-55-1.pdf}
+
+\begin{verbatim}
+benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+\end{verbatim}
+
+\begin{verbatim}
+trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
+  geom_point(size = 0.1) +
+  geom_trimesh() +
+  coord_equal()
+
+trimesh
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-56-1.pdf}
+
+\begin{verbatim}
+trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
+                                triangular_object = tr1_object, distance_col = distance)
+
+trimesh_gr
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-57-1.pdf}
+
+\begin{verbatim}
+trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
+                                     triangular_object = tr1_object, distance_col = distance)
+trimesh_removed
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-58-1.pdf}
+
+\begin{verbatim}
+tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
+                          distance = distance, distance_col = "distance", column_start_text = "PC")
+tour1
+\end{verbatim}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-59-1.pdf}
 

### 2024-02-16 scripts/function_scripts/functions_tri_hex_TriMAP.R

@@ -1,0 +1,349 @@
+Fit_TriMAP_data <- function(df_2_without_class, tem_dir){
+  write.csv(df_2_without_class, file.path(tem_dir, "df_2_without_class.csv"), row.names = FALSE,
+            quote = TRUE)
+}
+
+
+### Plot tSNE in 2D
+
+plot_TriMAP_2D <- function(tSNE_df){
+  tSNE_df_plot <- tSNE_df %>%
+    ggplot(aes(x = TriMAP1,
+               y = TriMAP2))+
+    geom_point() +
+    coord_equal()
+  return(tSNE_df_plot)
+}
+
+## To find optimal bin value
+
+find_opt_bin_val <- function (x)
+{
+  #h <- stats::IQR(x)
+  #if (h == 0)
+  #  h <- stats::mad(x, constant = 2)
+  #if (h > 0)
+  #  ceiling(diff(range(x))/(2 * h * length(x)^(-1/3)))
+  #else 1L
+  h <- 2 * IQR(x) / length(x)^(1/3) # bin width
+  return(h)
+}
+
+
+## To create hexbin object
+create_hex_bin <- function(tSNE_df, num_bins, shape){
+  hb <- hexbin(tSNE_df$TriMAP1, tSNE_df$TriMAP2, num_bins, IDs = TRUE, shape = shape)
+  return(hb)
+}
+
+### To Draw a hex bins with the points
+draw_hex_bin <- function(tSNE_df, num_bins){
+  hex_bin_plot <- ggplot(tSNE_df, aes(TriMAP1, TriMAP2)) +
+    geom_hex(bins = num_bins) + geom_point()+
+    scale_fill_viridis_c()
+
+  return(hex_bin_plot)
+}
+
+## To create the dataframe which contain coordinate values with the hexbin ID
+create_hexbin_ID_for_all <- function(df_2, tSNE_df, num_bins, shape){
+  ### Merge tSNE dataset in 2D with original dataset which contains 4D coordinates
+  df_2 <- df_2 %>%
+    mutate(ID=row_number())
+
+  df_new <- df_2 %>% inner_join(tSNE_df, by = "ID")
+
+  ### Fit hexbins and store hexbin IDs
+  hb <- create_hex_bin(tSNE_df, num_bins, shape)
+
+  ### Add hexbin Ids as a column to the original dataset
+  df_new$hb_id <- hb@cID
+
+  ## To change column names to lower case
+  names(df_new) <- tolower(names(df_new))
+
[…]

### 2024-02-16 scripts/function_scripts/functions_tri_hex_tsne_without_pc.R

@@ -1,0 +1,377 @@
+### Fit tSNE without scaling and create the dataframe
+
+Fit_tSNE <- function(df_2, opt_perplexity, num_dim){
+  tSNE_fit <- df_2 %>%
+    select(where(is.numeric)) %>%
+    Rtsne(perplexity = opt_perplexity, pca = FALSE, pca_center = FALSE, normalize = FALSE, dims = num_dim)
+    #Rtsne(perplexity = opt_perplexity)
+
+  tSNE_df <- tSNE_fit$Y %>%
+    as.data.frame()  %>%
+    mutate(ID=row_number())
+
+  names(tSNE_df)[1:(ncol(tSNE_df)-1)] <- paste0(rep("tSNE",(ncol(tSNE_df)-1)), 1:(ncol(tSNE_df)-1))
+  return(tSNE_df)
+}
+
+### Plot tSNE in 2D
+
+plot_tSNE_2D <- function(tSNE_df){
+  tSNE_df_plot <- tSNE_df %>%
+    ggplot(aes(x = tSNE1,
+               y = tSNE2))+
+    geom_point() +
+    coord_equal()
+  return(tSNE_df_plot)
+}
+
+## To find optimal bin value
+
+find_opt_bin_val <- function (x)
+{
+  #h <- stats::IQR(x)
+  #if (h == 0)
+  #  h <- stats::mad(x, constant = 2)
+  #if (h > 0)
+  #  ceiling(diff(range(x))/(2 * h * length(x)^(-1/3)))
+  #else 1L
+  h <- 2 * IQR(x) / length(x)^(1/3) # bin width
+  return(h)
+}
+
+find_avg_val <- function(tSNE_df){
+  bw1 <- find_opt_bin_val(tSNE_df$tSNE1)
+  bw2 <- find_opt_bin_val(tSNE_df$tSNE2)
+
+  #tSNE1 <- data.frame(tSNE_val = tSNE_df$tSNE1)
+  #tSNE2 <- data.frame(tSNE_val = tSNE_df$tSNE2)
+
+  #tSNE1$type <- "tSNE1"
+  #tSNE2$type <- "tSNE2"
+
+  #cmb_tSNE <- rbind(tSNE1, tSNE2)
+
+  #ggplot(cmb_tSNE, aes(tSNE_val, fill = type)) + geom_histogram(alpha = 0.2)
+
+  #avg <- round((bw1 + bw2)/2,0)
+  sum <- round(bw1 + bw2,0)
+  return(sum)
[…]

### 2024-02-17 paper-quollr.Rmd
Add hex coord generation from scratch

@@
 - basic example
+<!--
+### Appropriate hex size value to have regular hexagons
+
+```{r}
+
+min_value_x <- min(s_curve_noise_umap$UMAP1)
+max_value_x <- max(s_curve_noise_umap$UMAP1)
+min_value_y <- min(s_curve_noise_umap$UMAP2)
+max_value_y <- max(s_curve_noise_umap$UMAP2)
+
+# Calculate the maximum span along x and y axes
+x_span <- max_value_x - min_value_x
+y_span <- max_value_y - min_value_y
+
+hex_size <- min(x_span, y_span)/5.5
+
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", hex_size = hex_size)
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, y = "UMAP2", hex_size = hex_size)
+
+all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
+                                                 x = "UMAP1", y = "UMAP2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y, 
+                                                 buffer_size = NA, hex_size = hex_size)
+hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = hex_size)
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
+```
+-->
 
@@
 ```{r}
-num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", cell_area = 1)
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", hex_size = NA)
 num_bins_x
@@
 ```{r}
-
-shape_val <- calculate_effective_shape_value(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2")
-shape_val
-```
-
-```{r}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2", shape_val = 1.833091, num_bins_x = num_bins_x)
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, y = "UMAP2", hex_size = NA)
 num_bins_y
@@
 cell_area <- 1
-cell_diameter <- sqrt(2 * cell_area / sqrt(3))
 
-hex_size <- cell_diameter/2
+hex_size <- sqrt(2 * cell_area / sqrt(3))
 
@@
 ```{r}
-hex_grid <- gen_hex_coordinates(all_centroids_df)
+hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = NA)
 glimpse(hex_grid)
@@
 
+```{r}
+## As an option first quantile considered as a default
+benchmark_to_rm_lwd_hex <- quantile(df_bin_centroids$std_counts)[2] + 0.01
+
+## To identify low density hexagons
+df_bin_centroids_low <- df_bin_centroids |>
+  dplyr::filter(std_counts <= benchmark_to_rm_lwd_hex)
+
+## To identify low-density hexagons needed to remove by investigating neighbouring mean density
+identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_all = df_bin_centroids, num_bins_x = num_bins_x,
+                     df_bin_centroids_low = df_bin_centroids_low)
+```
+
 #### Benchmark value to remove the long edges
@@
 ## Tests
+
+All functions have tests written and implemented using the \CRANpkg{testthat} [@testthat] in R.
 
@@
 ```{r}
-num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", cell_area = 1)
-num_bins_x <- 13
+num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", hex_size = NA)
 ```
@@
 ```{r}
-shape_val <- calculate_effective_shape_value(.data = tSNE_data, x = "tSNE1", y = "tSNE2")
-shape_val
-```
-
-```{r}
-num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, x = "tSNE1", y = "tSNE2", shape_val = shape_val, num_bins_x = num_bins_x)
+num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, y = "tSNE2", hex_size = NA)
 num_bins_y

### 2024-02-17 paper-quollr.tex
Add hex coord generation from scratch

@@
 \begin{verbatim}
-num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", cell_area = 1)
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", hex_size = NA)
 num_bins_x
@@
 \begin{verbatim}
-#> [1] 6
-\end{verbatim}
-
-\begin{verbatim}
-shape_val <- calculate_effective_shape_value(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2")
-shape_val
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 2.019414
-\end{verbatim}
-
-\begin{verbatim}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2", shape_val = 1.833091, num_bins_x = num_bins_x)
+#> [1] 4
+\end{verbatim}
+
+\begin{verbatim}
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, y = "UMAP2", hex_size = NA)
 num_bins_y
@@
 \begin{verbatim}
-#> [1] 12
+#> [1] 8
 \end{verbatim}
@@
 cell_area <- 1
-cell_diameter <- sqrt(2 * cell_area / sqrt(3))
-
-hex_size <- cell_diameter/2
+
+hex_size <- sqrt(2 * cell_area / sqrt(3))
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
 
@@
 \begin{verbatim}
-#> Rows: 72
+#> Rows: 32
 #> Columns: 2
-#> $ x <dbl> -3.5390002, -2.9126765, -3.5390002, -2.9126765, -3.5390002, -2.91267~
-#> $ y <dbl> -6.0111830, -4.9111506, -3.8111182, -2.7110858, -1.6110534, -0.51102~
+#> $ x <dbl> -3.8076427, -2.6742223, -3.8076427, -2.6742223, -3.8076427, -2.67422~
+#> $ y <dbl> -6.2798254, -4.4744481, -2.6690708, -0.8636935, 0.9416838, 2.7470611~
 \end{verbatim}
@@
 \begin{verbatim}
-hex_grid <- gen_hex_coordinates(all_centroids_df)
+hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = NA)
 glimpse(hex_grid)
@@
 \begin{verbatim}
-#> Rows: 432
+#> Rows: 192
 #> Columns: 3
-#> $ x  <dbl> -2.912676, -2.912676, -3.539000, -4.165324, -4.165324, -3.539000, -~
-#> $ y  <dbl> -5.645998, -6.376368, -6.741553, -6.376368, -5.645998, -5.280813, -~
+#> $ x  <dbl> -2.674222, -2.674222, -3.807643, -4.941063, -4.941063, -3.807643, -~
+#> $ y  <dbl> -5.6804828, -6.8791681, -7.4785108, -6.8791681, -5.6804828, -5.0811~
 #> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-10-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-18-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}
 
@@
 \begin{verbatim}
-#> # A tibble: 20 x 4
 #>             x          y hexID std_counts
-#>     <dbl>  <dbl> <int>      <dbl>
-#>  1 -2.91  -4.91      7      0.5  
-#>  2 -3.54  -3.81     13      0.5  
-#>  3 -2.29  -6.01      2      0.5  
-#>  4 -1.66  -4.91      8      0.75 
-#>  5 -2.29  -3.81     14      0.5  
-#>  6 -1.03  -3.81     15      0.375
-#>  7 -0.407 -2.71     21      0.125
-#>  8 -1.03  -1.61     27      0.125
-#>  9  0.219 -1.61     28      1    
-#> 10  0.845 -0.511    34      0.625
-#> 11  0.845  1.69     46      0.125
-#> 12  0.845  3.89     58      0.625
-#> 13  0.219  4.99     64      0.125
-#> 14  0.845  6.09     70      0.375
-#> 15  1.47   0.589    41      1    
-#> 16  1.47   2.79     53      0.125
-#> 17  2.10   3.89     59      0.625
-#> 18  1.47   4.99     65      0.5  
-#> 19  2.10   6.09     71      0.375
-#> 20  2.72   4.99     66      0.5
+#> 1  -2.6742223 -4.4744481     5     1.0000
+#> 2  -1.5408019 -6.2798254     2     0.3125
+#> 3  -0.4073814 -4.4744481     6     0.0625
+#> 4  -1.5408019 -2.6690708    10     0.2500
+#> 5  -0.4073814 -0.8636935    14     0.5000
+#> 6   0.7260390 -2.6690708    11     0.1250
+#> 7   1.8594594 -0.8636935    15     0.1875
+#> 8   0.7260390  0.9416838    19     0.6250
+#> 9   1.8594594  2.7470611    23     0.2500
+#> 10  0.7260390  4.5524384    27     0.5625
+#> 11  1.8594594  6.3578158    31     0.3750
+#> 12  2.9928798  4.5524384    28     0.4375
 \end{verbatim}
@@
 \begin{verbatim}
-#> Rows: 72
+#> Rows: 32
 #> Columns: 2
-#> $ x <dbl> -3.2700000, -2.6436763, -3.2700000, -2.6436763, -3.2700000, -2.64367~
-#> $ y <dbl> -5.7400000, -4.6399676, -3.5399352, -2.4399028, -1.3398704, -0.23983~
+#> $ x <dbl> -3.5386425, -2.4052221, -3.5386425, -2.4052221, -3.5386425, -2.40522~
+#> $ y <dbl> -6.0086424, -4.2032651, -2.3978878, -0.5925105, 1.2128668, 3.0182441~
 \end{verbatim}
@@
 \begin{verbatim}
-#> Rows: 432
+#> Rows: 192
 #> Columns: 3
-#> $ x  <dbl> -2.643676, -2.643676, -3.270000, -3.896324, -3.896324, -3.270000, -~
-#> $ y  <dbl> -5.3748152, -6.1051848, -6.4703696, -6.1051848, -5.3748152, -5.0096~
+#> $ x  <dbl> -2.405222, -2.405222, -3.538643, -4.672063, -4.672063, -3.538643, -~
+#> $ y  <dbl> -5.409299776, -6.607985117, -7.207327787, -6.607985117, -5.40929977~
 #> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-23-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-25-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-24-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-30-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-29-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-31-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-30-1.pdf}
 
@@
 \begin{verbatim}
-#> # A tibble: 21 x 4
 #>             x          y hexID std_counts
-#>     <dbl> <dbl> <int>      <dbl>
-#>  1 -3.27  -5.74     1      0.286
-#>  2 -2.64  -4.64     7      0.571
-#>  3 -3.27  -3.54    13      0.714
-#>  4 -2.02  -5.74     2      0.857
-#>  5 -1.39  -4.64     8      0.429
-#>  6 -2.02  -3.54    14      0.286
-#>  7 -0.765 -3.54    15      0.429
-#>  8 -0.138 -2.44    21      0.143
-#>  9 -0.765 -1.34    27      0.286
-#> 10  0.488 -1.34    28      1    
-#> # i 11 more rows
+#> 1  -3.5386425 -6.0086424     1 0.21428571
+#> 2  -2.4052221 -4.2032651     5 1.00000000
+#> 3  -1.2718017 -6.0086424     2 0.42857143
+#> 4  -0.1383812 -4.2032651     6 0.07142857
+#> 5  -1.2718017 -2.3978878    10 0.21428571
+#> 6  -0.1383812 -0.5925105    14 0.50000000
+#> 7   0.9950392 -2.3978878    11 0.28571429
+#> 8   2.1284596 -0.5925105    15 0.14285714
+#> 9   0.9950392  1.2128668    19 0.64285714
+#> 10  2.1284596  3.0182441    23 0.35714286
+#> 11  0.9950392  4.8236214    27 1.00000000
+#> 12  2.1284596  6.6289988    31 0.07142857
+#> 13  3.2618800  4.8236214    28 0.42857143
 \end{verbatim}
@@
 
+\begin{verbatim}
+## As an option first quantile considered as a default
+benchmark_to_rm_lwd_hex <- quantile(df_bin_centroids$std_counts)[2] + 0.01
+
+## To identify low density hexagons
+df_bin_centroids_low <- df_bin_centroids |>
+  dplyr::filter(std_counts <= benchmark_to_rm_lwd_hex)
+
+## To identify low-density hexagons needed to remove by investigating neighbouring mean density
+identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_all = df_bin_centroids, num_bins_x = num_bins_x,
+                     df_bin_centroids_low = df_bin_centroids_low)
+\end{verbatim}
+
 \hypertarget{benchmark-value-to-remove-the-long-edges}{%
@@
 benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+benchmark <- 3
 \end{verbatim}
@@
 \paragraph{show\_langevitour}\label{show_langevitour}}
+
+\begin{verbatim}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-ID) |>
+  dplyr::bind_cols(s_curve_noise_umap_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
+\end{verbatim}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-39-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-40-1.pdf}
 
@@
 \subsection{Tests}\label{tests}}
+
+All functions have tests written and implemented using the \CRANpkg{testthat} (Wickham 2011) in R.
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-40-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-41-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-42-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-43-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-43-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-44-1.pdf}
 
@@
 #>     Automatically selected t = 22
-#>   Calculated optimal t in 0.39 seconds.
+#>   Calculated optimal t in 0.49 seconds.
 #>   Calculating diffusion potential...
-#>   Calculated diffusion potential in 0.32 seconds.
+#>   Calculated diffusion potential in 0.38 seconds.
 #>   Calculating metric MDS...
-#>   Calculated metric MDS in 8.83 seconds.
-#> Calculated PHATE in 9.55 seconds.
+#>   Calculated metric MDS in 7.83 seconds.
+#> Calculated PHATE in 8.72 seconds.
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-44-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-45-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-45-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}
-
-\begin{verbatim}
-num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", cell_area = 1)
-num_bins_x <- 13
-\end{verbatim}
-
-\begin{verbatim}
-shape_val <- calculate_effective_shape_value(.data = tSNE_data, x = "tSNE1", y = "tSNE2")
-shape_val
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 1.13882
-\end{verbatim}
-
-\begin{verbatim}
-num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, x = "tSNE1", y = "tSNE2", shape_val = 0.8417289, num_bins_x = num_bins_x)
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-47-1.pdf}
+
+\begin{verbatim}
+num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", hex_size = NA)
+\end{verbatim}
+
+\begin{verbatim}
+num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, y = "tSNE2", hex_size = NA)
 num_bins_y
@@
 \begin{verbatim}
-#> [1] 14
+#> [1] 38
 \end{verbatim}
@@
 
-df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+df_with_std_counts <- compute_std_counts(nldr_df = tSNE_data_with_id)
 
@@
 \begin{verbatim}
-#> # A tibble: 21 x 4
 #>               x             y hexID std_counts
-#>    <dbl> <dbl> <int>      <dbl>
-#>  1 -27.7 -34.8     1      0.286
-#>  2 -25.4 -30.1    14      0.286
-#>  3 -27.7 -25.4    27      0.286
-#>  4 -25.4 -20.7    40      0.286
-#>  5 -27.7 -16.0    53      0.429
-#>  6 -25.4 -11.3    66      0.143
-#>  7 -23.2 -34.8     2      0.857
-#>  8 -20.9 -30.1    15      0.429
-#>  9 -23.2 -25.4    28      1    
-#> 10 -20.9 -20.7    41      0.714
-#> # i 11 more rows
+#> 1   -26.9783746  -3.336399929   552     0.6875
+#> 2   -25.0331153  -3.336399929   553     0.3750
+#> 3   -26.0057449  -1.663988875   582     0.2500
+#> 4   -23.0878560  -6.681222036   496     0.5625
+#> 5   -24.0604856  -5.008810982   525     0.3125
+#> 6   -23.0878560  -3.336399929   554     0.3125
+#> 7   -24.0604856  -1.663988875   583     0.1875
+#> 8   -23.0878560   0.008422179   612     0.0625
+#> 9   -22.1152263  -8.353633090   468     0.0625
+#> 10  -22.1152263  -5.008810982   526     0.0625
+#> 11  -19.1973374 -13.370866252   382     0.0625
+#> 12  -19.1973374  16.732532718   904     0.1250
+#> 13  -19.1973374  20.077354826   962     0.1250
+#> 14  -17.2520781 -13.370866252   383     0.1875
+#> 15  -18.2247077 -11.698455198   412     0.2500
+#> 16  -17.2520781  16.732532718   905     0.1250
+#> 17  -18.2247077  18.404943772   934     0.6875
+#> 18  -17.2520781  20.077354826   963     0.1875
+#> 19  -16.2794484 -18.388099413   297     0.6250
+#> 20  -15.3068188 -16.715688359   326     0.2500
+#> 21  -16.2794484 -15.043277306   355     0.3750
+#> 22  -14.3341891 -18.388099413   298     0.0625
+#> 23  -13.3615595 -16.715688359   327     0.1875
+#> 24  -14.3341891 -15.043277306   356     0.0625
+#> 25  -14.3341891  -5.008810982   530     0.0625
+#> 26  -13.3615595  -3.336399929   559     0.3125
+#> 27  -13.3615595  20.077354826   965     0.2500
+#> 28  -14.3341891  21.749765879   994     0.0625
+#> 29  -13.3615595  23.422176933  1023     0.5000
+#> 30  -12.3889298  -8.353633090   473     0.1875
+#> 31  -12.3889298  -5.008810982   531     0.0625
+#> 32  -11.4163002  -3.336399929   560     0.1250
+#> 33  -12.3889298  -1.663988875   589     0.1250
+#> 34  -12.3889298  18.404943772   937     0.0625
+#> 35  -11.4163002  20.077354826   966     0.5625
+#> 36  -12.3889298  21.749765879   995     0.1875
+#> 37  -11.4163002  23.422176933  1024     0.1875
+#> 38  -12.3889298  25.094587987  1053     0.3125
+#> 39  -11.4163002  26.766999041  1082     0.0625
+#> 40   -9.4710409 -13.370866252   387     0.1875
+#> 41  -10.4436705 -11.698455198   416     0.2500
+#> 42   -9.4710409 -10.026044144   445     0.2500
+#> 43  -10.4436705  -8.353633090   474     0.2500
+#> 44   -9.4710409  -6.681222036   503     0.0625
+#> 45   -9.4710409   6.698066395   735     0.1875
+#> 46   -9.4710409  10.042888502   793     0.0625
+#> 47   -9.4710409  16.732532718   909     0.0625
+#> 48  -10.4436705  18.404943772   938     0.3125
+#> 49   -9.4710409  20.077354826   967     0.2500
+#> 50  -10.4436705  21.749765879   996     0.2500
+#> 51   -9.4710409  23.422176933  1025     0.1875
+#> 52  -10.4436705  25.094587987  1054     0.1875
+#> 53   -9.4710409  26.766999041  1083     0.1875
+#> 54   -7.5257816 -20.060510467   272     0.3750
+#> 55   -8.4984112 -18.388099413   301     0.0625
+#> 56   -7.5257816 -16.715688359   330     0.1250
+#> 57   -8.4984112 -15.043277306   359     0.3125
+#> 58   -7.5257816 -13.370866252   388     0.2500
+#> 59   -8.4984112 -11.698455198   417     0.1875
+#> 60   -7.5257816 -10.026044144   446     0.4375
+#> 61   -8.4984112  -8.353633090   475     0.3750
+#> 62   -7.5257816  -6.681222036   504     0.3125
+#> 63   -8.4984112  -5.008810982   533     0.0625
+#> 64   -7.5257816  -3.336399929   562     0.1250
+#> 65   -7.5257816   0.008422179   620     0.1875
+#> 66   -7.5257816   3.353244287   678     0.1250
+#> 67   -8.4984112   5.025655341   707     0.1250
+#> 68   -8.4984112   8.370477449   765     0.3750
+#> 69   -7.5257816  10.042888502   794     0.0625
+#> 70   -8.4984112  18.404943772   939     0.1250
+#> 71   -7.5257816  20.077354826   968     0.1250
+#> 72   -8.4984112  21.749765879   997     0.1875
+#> 73   -7.5257816  23.422176933  1026     0.1250
+#> 74   -6.5531519 -18.388099413   302     0.1250
+#> 75   -5.5805223 -16.715688359   331     0.1250
+#> 76   -6.5531519 -15.043277306   360     0.3125
+#> 77   -5.5805223 -13.370866252   389     0.3125
+#> 78   -6.5531519 -11.698455198   418     0.0625
+#> 79   -5.5805223 -10.026044144   447     0.1875
+#> 80   -6.5531519  -8.353633090   476     0.3750
+#> 81   -5.5805223  -6.681222036   505     0.1250
+#> 82   -6.5531519  -5.008810982   534     0.1875
+#> 83   -6.5531519  -1.663988875   592     0.0625
+#> 84   -5.5805223   0.008422179   621     0.0625
+#> 85   -6.5531519   5.025655341   708     0.1875
+#> 86   -5.5805223   6.698066395   737     0.2500
+#> 87   -6.5531519   8.370477449   766     0.0625
+#> 88   -5.5805223  10.042888502   795     0.3125
+#> 89   -5.5805223  13.387710610   853     0.0625
+#> 90   -5.5805223  16.732532718   911     0.1250
+#> 91   -6.5531519  18.404943772   940     0.1250
+#> 92   -3.6352630 -16.715688359   332     0.1875
+#> 93   -4.6078926 -15.043277306   361     0.1875
+#> 94   -3.6352630 -13.370866252   390     0.1250
+#> 95   -4.6078926 -11.698455198   419     0.1875
+#> 96   -3.6352630 -10.026044144   448     0.1875
+#> 97   -4.6078926  -8.353633090   477     0.3125
+#> 98   -3.6352630  -6.681222036   506     0.3125
+#> 99   -4.6078926  -5.008810982   535     0.1250
+#> 100  -3.6352630   0.008422179   622     0.2500
+#> 101  -3.6352630   6.698066395   738     0.2500
+#> 102  -4.6078926   8.370477449   767     0.1250
+#> 103  -3.6352630  10.042888502   796     0.1875
+#> 104  -4.6078926  11.715299556   825     0.0625
+#> 105  -3.6352630  13.387710610   854     0.2500
+#> 106  -4.6078926  15.060121664   883     0.1250
+#> 107  -1.6900037 -33.439798898    43     0.0625
+#> 108  -2.6626333 -18.388099413   304     0.0625
+#> 109  -1.6900037 -16.715688359   333     0.1250
+#> 110  -2.6626333 -15.043277306   362     0.2500
+#> 111  -1.6900037 -13.370866252   391     0.3125
+#> 112  -2.6626333 -11.698455198   420     0.1250
+#> 113  -1.6900037 -10.026044144   449     0.2500
+#> 114  -2.6626333  -8.353633090   478     0.2500
+#> 115  -1.6900037  -6.681222036   507     0.3750
+#> 116  -2.6626333  -5.008810982   536     0.1250
+#> 117  -2.6626333  -1.663988875   594     0.4375
+#> 118  -2.6626333   5.025655341   710     0.1250
+#> 119  -1.6900037  10.042888502   797     0.0625
+#> 120  -2.6626333  11.715299556   826     0.4375
+#> 121  -1.6900037  13.387710610   855     0.1250
+#> 122  -2.6626333  15.060121664   884     0.0625
+#> 123   0.2552556 -33.439798898    44     1.0000
+#> 124   0.2552556 -13.370866252   392     0.1875
+#> 125  -0.7173740 -11.698455198   421     0.1250
+#> 126   0.2552556 -10.026044144   450     0.1250
+#> 127  -0.7173740  -8.353633090   479     0.0625
+#> 128  -0.7173740   1.680833233   653     0.0625
+#> 129   0.2552556   3.353244287   682     0.5000
+#> 130  -0.7173740   5.025655341   711     0.3125
+#> 131  -0.7173740   8.370477449   769     0.1875
+#> 132   0.2552556  10.042888502   798     0.2500
+#> 133  -0.7173740  11.715299556   827     0.1250
+#> 134  -0.7173740  15.060121664   885     0.1875
+#> 135   0.2552556  16.732532718   914     0.1875
+#> 136   1.2278853 -35.112209952    16     0.0625
+#> 137   2.2005149 -33.439798898    45     0.3750
+#> 138   1.2278853 -31.767387844    74     0.0625
+#> 139   2.2005149 -30.094976790   103     0.6250
+#> 140   1.2278853 -28.422565737   132     0.0625
+#> 141   2.2005149 -26.750154683   161     0.1250
+#> 142   1.2278853 -25.077743629   190     0.0625
+#> 143   2.2005149 -23.405332575   219     0.1250
+#> 144   1.2278853 -21.732921521   248     0.0625
+#> 145   1.2278853 -18.388099413   306     0.0625
+#> 146   2.2005149 -16.715688359   335     0.0625
+#> 147   1.2278853 -15.043277306   364     0.3125
+#> 148   1.2278853   1.680833233   654     0.3750
+#> 149   2.2005149   3.353244287   683     0.1875
+#> 150   1.2278853   5.025655341   712     0.2500
+#> 151   2.2005149   6.698066395   741     0.1250
+#> 152   1.2278853   8.370477449   770     0.0625
+#> 153   2.2005149  10.042888502   799     0.1875
+#> 154   1.2278853  11.715299556   828     0.3750
+#> 155   2.2005149  13.387710610   857     0.1875
+#> 156   1.2278853  15.060121664   886     0.3750
+#> 157   2.2005149  16.732532718   915     0.1875
+#> 158   1.2278853  18.404943772   944     0.0625
+#> 159   3.1731446 -31.767387844    75     0.0625
+#> 160   4.1457742 -30.094976790   104     0.2500
+#> 161   3.1731446 -28.422565737   133     0.3750
+#> 162   3.1731446 -25.077743629   191     0.6875
+#> 163   4.1457742 -23.405332575   220     0.0625
+#> 164   3.1731446  -8.353633090   481     0.1250
+#> 165   3.1731446   1.680833233   655     0.1875
+#> 166   3.1731446   5.025655341   713     0.0625
+#> 167   3.1731446   8.370477449   771     0.1250
+#> 168   4.1457742  10.042888502   800     0.1250
+#> 169   3.1731446  11.715299556   829     0.1875
+#> 170   4.1457742  13.387710610   858     0.1250
+#> 171   3.1731446  15.060121664   887     0.1250
+#> 172   5.1184039 -25.077743629   192     0.1250
+#> 173   6.0910335   0.008422179   627     0.1250
+#> 174   6.0910335   3.353244287   685     0.0625
+#> 175   6.0910335   6.698066395   743     0.1250
+#> 176   6.0910335  10.042888502   801     0.1250
+#> 177   6.0910335  13.387710610   859     0.1250
+#> 178   8.0362928   0.008422179   628     0.0625
+#> 179   7.0636632   1.680833233   657     0.1250
+#> 180   8.0362928   3.353244287   686     0.3125
+#> 181   7.0636632   5.025655341   715     0.2500
+#> 182   8.0362928   6.698066395   744     0.1250
+#> 183   7.0636632   8.370477449   773     0.0625
+#> 184   8.0362928  10.042888502   802     0.3125
+#> 185   7.0636632  11.715299556   831     0.0625
+#> 186   8.0362928  13.387710610   860     0.0625
+#> 187   7.0636632  15.060121664   889     0.2500
+#> 188   8.0362928  16.732532718   918     0.1875
+#> 189   7.0636632  18.404943772   947     0.0625
+#> 190   9.9815521   0.008422179   629     0.0625
+#> 191   9.0089225   1.680833233   658     0.1250
+#> 192   9.9815521   3.353244287   687     0.1250
+#> 193   9.0089225   5.025655341   716     0.1250
+#> 194   9.9815521   6.698066395   745     0.1250
+#> 195   9.0089225   8.370477449   774     0.1875
+#> 196   9.9815521  10.042888502   803     0.1875
+#> 197   9.0089225  11.715299556   832     0.2500
+#> 198   9.9815521  13.387710610   861     0.0625
+#> 199   9.0089225  15.060121664   890     0.2500
+#> 200   9.9815521  16.732532718   919     0.0625
+#> 201  11.9268114   0.008422179   630     0.1250
+#> 202  10.9541818   1.680833233   659     0.3125
+#> 203  11.9268114   3.353244287   688     0.2500
+#> 204  11.9268114   6.698066395   746     0.1875
+#> 205  11.9268114  10.042888502   804     0.0625
+#> 206  10.9541818  11.715299556   833     0.1875
+#> 207  11.9268114  13.387710610   862     0.1250
+#> 208  10.9541818  15.060121664   891     0.1875
+#> 209  13.8720707  -3.336399929   573     0.2500
+#> 210  12.8994411  -1.663988875   602     0.1250
+#> 211  13.8720707   0.008422179   631     0.0625
+#> 212  12.8994411   1.680833233   660     0.1250
+#> 213  13.8720707   3.353244287   689     0.1250
+#> 214  12.8994411   5.025655341   718     0.2500
+#> 215  12.8994411   8.370477449   776     0.0625
+#> 216  13.8720707  10.042888502   805     0.2500
+#> 217  12.8994411  11.715299556   834     0.3750
+#> 218  13.8720707  13.387710610   863     0.1250
+#> 219  12.8994411  15.060121664   892     0.3750
+#> 220  14.8447004  -5.008810982   545     0.4375
+#> 221  15.8173300  -3.336399929   574     0.1250
+#> 222  15.8173300   0.008422179   632     0.3750
+#> 223  14.8447004   5.025655341   719     0.1875
+#> 224  14.8447004   8.370477449   777     0.0625
+#> 225  15.8173300  10.042888502   806     0.1250
+#> 226  14.8447004  11.715299556   835     0.2500
+#> 227  14.8447004  15.060121664   893     0.0625
+#> 228  16.7899597  -5.008810982   546     0.0625
+#> 229  17.7625893  -3.336399929   575     0.0625
+#> 230  16.7899597  -1.663988875   604     0.1875
+#> 231  17.7625893   0.008422179   633     0.2500
+#> 232  17.7625893   3.353244287   691     0.1250
+#> 233  16.7899597   5.025655341   720     0.0625
+#> 234  17.7625893   6.698066395   749     0.0625
+#> 235  17.7625893  10.042888502   807     0.0625
+#> 236  16.7899597  11.715299556   836     0.0625
+#> 237  19.7078486  -3.336399929   576     0.1250
+#> 238  18.7352190  -1.663988875   605     0.3750
+#> 239  19.7078486   0.008422179   634     0.2500
+#> 240  18.7352190   1.680833233   663     0.4375
+#> 241  19.7078486   3.353244287   692     0.3125
+#> 242  18.7352190   5.025655341   721     0.0625
+#> 243  19.7078486   6.698066395   750     0.2500
+#> 244  18.7352190   8.370477449   779     0.1875
+#> 245  19.7078486  10.042888502   808     0.1875
+#> 246  20.6804783  -1.663988875   606     0.3125
+#> 247  21.6531079   0.008422179   635     0.0625
+#> 248  20.6804783   1.680833233   664     0.1875
+#> 249  21.6531079   3.353244287   693     0.2500
+#> 250  21.6531079   6.698066395   751     0.2500
+#> 251  20.6804783   8.370477449   780     0.1875
+#> 252  21.6531079  10.042888502   809     0.5000
+#> 253  20.6804783  11.715299556   838     0.0625
+#> 254  22.6257376  -1.663988875   607     0.1250
+#> 255  23.5983672   0.008422179   636     0.1875
+#> 256  22.6257376   1.680833233   665     0.0625
+#> 257  23.5983672   3.353244287   694     0.1875
+#> 258  23.5983672   6.698066395   752     0.1875
+#> 259  22.6257376   8.370477449   781     0.1250
+#> 260  24.5709969   1.680833233   666     0.3750
+#> 261  24.5709969   5.025655341   724     0.1875
+#> 262  25.5436265   6.698066395   753     0.0625
+#> 263  24.5709969   8.370477449   782     0.2500
+#> 264  26.5162562   1.680833233   667     0.1875
 \end{verbatim}
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-testthat}{}}%
+Wickham, Hadley. 2011. {``Testthat: Get Started with Testing.''} \emph{The R Journal} 3: 5--10. \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%

### 2024-02-17 scripts/quollr_code_old.R

@@ -1,0 +1,1428 @@
+calculate_effective_x_bins <- function(.data, x = "UMAP1", cell_area = 1){
+
+  if (anyNA(.data[[rlang::as_string(rlang::ensym(x))]])) {
+    stop("NAs present")
+  }
+
+  if (any(is.infinite(.data |> dplyr::pull({{ x }})))) {
+    stop("Inf present")
+  }
+
+  if ((cell_area <= 0) || (is.infinite(cell_area))) {
+    stop("Invalid cell area value")
+
+  }
+
+  ## To compute the radius of the hexagon outer circle
+  cell_radius <- 2 * sqrt(2 * cell_area / sqrt(3))
+
+  ## To compute the range along x-axis
+  xwidth <- diff(range(.data |>
+                         dplyr::pull({{ x }})))
+
+  num_bins <- ceiling(xwidth/cell_diameter)
+  #num_bins <- floor(xwidth/cell_diameter + 1.5001)
+  num_bins
+
+}
+
+calculate_effective_y_bins <- function(.data, x = "UMAP2", y = "UMAP2", shape_val = NA, num_bins_x){
+
+  if (anyNA(.data[[rlang::as_string(rlang::ensym(y))]])) {
+    stop("NAs present")
+  }
+
+  if (any(is.infinite(.data |> dplyr::pull({{ y }})))) {
+    stop("Inf present")
+  }
+
+  if (is.na(shape_val)) {
+    shape_val <- calculate_effective_shape_value(.data = .data, x = x, y = y)
+
+  }
+
+  num_bins <- 2 * floor(((num_bins_x - 1) * shape_val)/sqrt(3) + 1.5001)
+  num_bins
+
+}
+
+calculate_effective_shape_value <- function(.data, x = UMAP1, y = UMAP2){
+
+  if (any(is.na(.data |> dplyr::pull({{ x }}))) || any(is.na(.data |> dplyr::pull({{ y }})))) {
+    stop("NAs present")
+  }
+
+  if (any(is.infinite(.data |> dplyr::pull({{ x }}))) || any(is.infinite(.data |> dplyr::pull({{ y }})))) {
+    stop("Inf present")
+  }
+
+  if ((length(.data |> dplyr::pull({{ x }})) == 1) || (length(.data |> dplyr::pull({{ y }})) == 1)) {
+    stop("Presence one observation only")
+
+  }
+
+  ## To compute the range along x-axis
[…]

### 2024-02-18 paper-quollr.tex
Updated version with scratch codes

@@
   basic example
+  --\textgreater{}
 \end{itemize}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-41-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-42-1.pdf}
 
@@
 #>     Automatically selected t = 22
-#>   Calculated optimal t in 0.49 seconds.
+#>   Calculated optimal t in 0.26 seconds.
 #>   Calculating diffusion potential...
-#>   Calculated diffusion potential in 0.38 seconds.
+#>   Calculated diffusion potential in 0.36 seconds.
 #>   Calculating metric MDS...
-#>   Calculated metric MDS in 7.83 seconds.
-#> Calculated PHATE in 8.72 seconds.
+#>   Calculated metric MDS in 8.01 seconds.
+#> Calculated PHATE in 8.64 seconds.
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-47-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-47-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-48-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-50-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-51-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-51-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-52-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-55-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-56-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-56-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-57-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-57-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-58-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-58-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-59-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-59-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-60-1.pdf}
 

### 2024-02-18 paper-quollr.R

@@ -43,6 +43,32 @@ s_curve_noise_umap <- read_rds(file = "data/s_curve_noise_umap.rds")
 
 
 ## -----------------------------------------------------------------------------
+
+min_value_x <- min(s_curve_noise_umap$UMAP1)
+max_value_x <- max(s_curve_noise_umap$UMAP1)
+min_value_y <- min(s_curve_noise_umap$UMAP2)
+max_value_y <- max(s_curve_noise_umap$UMAP2)
+
+# Calculate the maximum span along x and y axes
+x_span <- max_value_x - min_value_x
+y_span <- max_value_y - min_value_y
+
+hex_size <- min(x_span, y_span)/5.5
+
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", hex_size = hex_size)
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, y = "UMAP2", hex_size = hex_size)
+
+all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
+                                                 x = "UMAP1", y = "UMAP2", 
+                                                 num_bins_x = num_bins_x, 
+                                                 num_bins_y = num_bins_y, 
+                                                 buffer_size = NA, hex_size = hex_size)
+hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = hex_size)
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
+
+
+## -----------------------------------------------------------------------------
 num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", hex_size = NA)
 num_bins_x
 

### 2024-02-19 paper-quollr.Rmd
Update with speed up codes

@@
 #library(quollr)
+library(knitr)
+library(kableExtra)
 library(readr)
@@
 training_data <- read_rds(file = "data/s_curve_noise_training.rds")
+test_data <- read_rds(file = "data/s_curve_noise_test.rds")
+
+s_curve_noise <- read_rds(file = "data/s_curve_noise.rds")
 s_curve_noise_umap <- read_rds(file = "data/s_curve_noise_umap.rds")
@@
 
-- dependancies
+- dependencies
 
@@
 - basic example
+
+### Datasets
+
+The `quollr` package comes with several data sets that load with the package.
+These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.
+
+```{r, echo=FALSE}
+datasets_tb <- tibble(dt = c("s_curve_noise",
+                             "s_curve_noise_training", 
+                             "s_curve_noise_test", 
+                             "s_curve_noise_umap"), 
+                      text = c("Simulated 3D S-curve data with additional four noise dimensions.",
+                               "Training data derived from S-curve data.", 
+                               "Test data derived from S-curve data.", 
+                               "UMAP 2D embedding data of S-curve data (n_neighbors: 15, min_dist: 0.1)."))
+```
+
+```{r datasets-tb-html, eval=is_html_output(), echo=FALSE}
+datasets_tb |> 
+  kable(caption = "quollr datasets", col.names = c("data", "explanation")) 
+```
+
+```{r datasets-tb-pdf, eval=is_latex_output(), echo=FALSE}
+datasets_tb |>
+  kable(caption = "quollr datasets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 <!--
@@
 
+Hexagonal binning is a powerful technique for visualizing the density of data points in a 2-d space.
+Unlike traditional rectangular bins, hexagonal bins offer several advantages, including a more uniform representation of density and reduced visual bias.
+However, to effectively do the hexagonal binning, it's essential to compute the appropriate configurations based on the characteristics of the dataset.
+
+Before computing hexagonal bin configurations, we need to determine the range of data along the x and y axes to establish the boundary for hexagonal binning.
+Additionally, choosing an appropriate hexagon size (the radius of the outer circle) is essential.
+By default, the `calculate_effective_x_bins()` and `calculate_effective_y_bins()` functions use a hexagon size of 1.07457.
+However, users can adjust the hexagon size to fit the data range and achieve regular hexagons without overlapping.
+
+The hexagon size directly affects the number of bins generated.
+A higher hexagonal size will result in fewer bins, while a lower hexagonal size will lead to more bins.
+Therefore, there's always a trade-off depending on the dataset used.
+Users should consider their specific data characteristics when selecting the hexagon size.
+
 ```{r}
-num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", hex_size = NA)
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, 
+                                         x = "UMAP1", hex_size = NA)
 num_bins_x
@@
 ```{r}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, y = "UMAP2", hex_size = NA)
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, 
+                                         y = "UMAP2", hex_size = NA)
 num_bins_y
@@
 
+```{r}
+df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
+  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
+  dplyr::distinct() |>
+  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
+
+df_bin_centroids
+```
+
 ### Construct the high-D model with different options
@@
 ```{r}
-df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
-  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
-  dplyr::distinct() |>
-  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
-
-df_bin_centroids
-```
-
-```{r}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
 tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
@@
 ```
+
 ```{r}
@@
 ```{r}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
 tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
@@
 ## Compute 2D distances
-distance <- cal_2d_dist(.data = tr_from_to_df)
+distance <- cal_2d_dist(tr_from_to_df_coord = tr_from_to_df)
 
@@
 
-benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
-benchmark <- 3
+benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
+benchmark
+benchmark <- 5
 ```
@@
 
-### Predict 2D embeddings
-
-### Compute residuals
-
-### Visualizations
-
-#### geom_trimesh
+`fit_high_d_model()` function is used to generate the 2D model and the high-D model.
 
 ```{r}
-trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
-  geom_point(size = 0.1) +
-  geom_trimesh() +
-  coord_equal()
-
-trimesh
+fit_high_d_model(training_data = training_data, nldr_df_with_id = s_curve_noise_umap, x = "UMAP1",
+                             y = "UMAP1", num_bins_x = NA, num_bins_y = NA,
+                             hex_size = NA, buffer_size = NA,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
 ```
-
-#### coloured_long_edges
-
-```{r}
-trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
-                                triangular_object = tr1_object, distance_col = "distance")
-
-trimesh_gr
-```
-
-#### remove long edges
-
-```{r}
-trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
-                                     triangular_object = tr1_object, distance_col = "distance")
-trimesh_removed
-```
-
-#### show_langevitour
 
@@
 
+### Predict 2D embeddings
+
+To predict the 2D embeddings for a new data point using the trained high-D model, we follow a series of steps outlined below:
+
+1.
+**Compute high-dimensional Euclidean distance**: Calculate the Euclidean distance between each new data point and the lift model's coordinates in the high-dimensional space.
+This distance metric helps identify the nearest lift model coordinates to each new data point.
+
+2.
+**Find the nearest lift model coordinates**: Determine the lift model's high-dimensional coordinates that are closest to each new data point based on the computed distances.
+This step helps establish a correspondence between the new data points and the lift model's representation in the high-dimensional space.
+
+3.
+**Map the hexagonal bin ID**: Assign each new data point to a hexagonal bin based on its nearest lift model coordinates.
+This mapping ensures that each data point is associated with a specific region in the hexagonal grid representation of the high-dimensional space.
+
+4.
+**Map the hexagonal bin centroid coordinates in 2D**: Transform the hexagonal bin centroid coordinates from the high-dimensional space to the 2D embedding space using the trained lift model.
+This mapping provides the 2D embeddings for the new data points, allowing them to be visualized and analyzed in a lower-dimensional space.
+
+The `predict_2d_embeddings()` function is used to predict 2D embeddings for a new data point.
+The inputs for the function are test data, bin centroid coordinates in 2D, lifting model coordinates in high-D, and the type of the NLDR technique.
+The output contains the predicted 2D embeddings along with the predicted hexagonal id and the ID of the test data.
+
 ```{r}
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
-                          distance = distance, distance_col = "distance")
+pred_df_test <- predict_2d_embeddings(test_data = training_data,
+df_bin_centroids = df_bin_centroids, df_bin = df_bin, type_NLDR = "UMAP")
+
+glimpse(pred_df_test)
+```
+
+### Making summaries
+
+There are two important summaries that should be made when constructing the model on a dataset using a specific NLDR technique, the Mean Square Error (MSE) and Akaike Information Criterion (AIC) which measure prediction accuracy.
+The function that perform the summaries called `generate_eval_df()`.
+The output of this function is a list which contains MSE and AIC.
+The code below uses `generate_eval_df()` to find the MSE and the AIC values for the model constructed on UMAP 2D embeddings generated from the `s_curve_noise` training data.
+
+```{r}
+generate_summary(test_data = training_data, prediction_df = pred_df_test,
+                 df_bin = df_bin, col_start = "x")
+```
+
+### Visualizations
+
+We use static visualizations to understand the model constructed in 2D. On the other hand, dynamic visualization is used to see how the model looks in high-D space.
+
+#### Static visualizations
+
+Static visualizations main involves two types of results. One is the triangulation and the other is the long edge removal. Both types of visualizations provide ggplot objects.
+
+- Triangulation result: To visualize the results of triangulation, we input a dataset containing hexagonal bin centroid coordinates where 2D embedding data exists.
+`geom_trimesh()` is used to visualize this result.
+
+- Long edge removal: The long edge removal process involves identifying and removing long edges from the triangular mesh.
+Table  shows the main arguments of the functions.
+We offer two functions for visualizing this process:
+
+- `colour_long_edges()`: This function colors the long edges within the triangular mesh by red.
+
+- `remove_long_edges()`: After identifying long edges, this function draws the triangular mesh without the long edges.
+
+#### Dynamic visaulizations
+
+The `show_langevitour()` function enables dynamic visualization of the 2D model alongside the high-dimensional (high-D) data in its original space.
+This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data.
+The main arguments are shown in Table
+
+```{r}
+trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
+  geom_point(size = 0.1) +
+  geom_trimesh() +
+  coord_equal()
+
+trimesh
+```
+
+```{r}
+trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
+
+trimesh_gr
+```
+
+```{r}
+trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
+trimesh_removed
+```
+
+```{r}
+tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, 
+                          benchmark_value = benchmark,
+                          distance = distance, distance_col = "distance", 
+                          use_default_benchmark_val = FALSE, 
+                          column_start_text = "x")
 tour1
@@
 num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", hex_size = NA)
+num_bins_x
 ```
@@
   geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
@@
 ```{r}
-df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
-  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
-  dplyr::distinct() |>
-  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
-
-df_bin_centroids
+df_bin_centroids <- extract_hexbin_centroids(hex_full_count_df)
 ```
@@
 
-benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
 ```
@@
 ```{r}
-trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
-                                triangular_object = tr1_object, distance_col = "distance")
+trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
 
@@
 ```{r}
-trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
-                                     triangular_object = tr1_object, distance_col = "distance")
+trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
 trimesh_removed

### 2024-02-19 paper-quollr.tex
Update with speed up codes

--- 2024-02-16 paper-quollr.tex
@@
 \item
-  dependancies
+  dependencies
 \end{itemize}
@@
 
+\hypertarget{datasets}{%
+\subsubsection{Datasets}\label{datasets}}
+
+The \texttt{quollr} package comes with several data sets that load with the package. These are described in Table \ref{tab:datasets-tb-pdf}.
+
+\begin{table}
+
+\caption{\label{tab:datasets-tb-pdf}quollr datasets}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+data & explanation\\
+\midrule
+s\_curve\_noise & Simulated 3D S-curve data with additional four noise dimensions.\\
+s\_curve\_noise\_training & Training data derived from S-curve data.\\
+s\_curve\_noise\_test & Test data derived from S-curve data.\\
+s\_curve\_noise\_umap & UMAP 2D embedding data of S-curve data (n\_neighbors: 15, min\_dist: 0.1).\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+--\textgreater{}
+
 \hypertarget{compute-hexagonal-bin-configurations}{%
@@
 
-\begin{verbatim}
-num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", cell_area = 1)
+Hexagonal binning is a powerful technique for visualizing the density of data points in a 2-d space.
+Unlike traditional rectangular bins, hexagonal bins offer several advantages, including a more uniform representation of density and reduced visual bias.
+However, to effectively do the hexagonal binning, it's essential to compute the appropriate configurations based on the characteristics of the dataset.
+
+Before computing hexagonal bin configurations, we need to determine the range of data along the x and y axes to establish the boundary for hexagonal binning.
+Additionally, choosing an appropriate hexagon size (the radius of the outer circle) is essential.
+By default, the \texttt{calculate\_effective\_x\_bins()} and \texttt{calculate\_effective\_y\_bins()} functions use a hexagon size of 1.07457.
+However, users can adjust the hexagon size to fit the data range and achieve regular hexagons without overlapping.
+
+The hexagon size directly affects the number of bins generated.
+A higher hexagonal size will result in fewer bins, while a lower hexagonal size will lead to more bins.
+Therefore, there's always a trade-off depending on the dataset used.
+Users should consider their specific data characteristics when selecting the hexagon size.
+
+\begin{verbatim}
+num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, 
+                                         x = "UMAP1", hex_size = NA)
 num_bins_x
@@
 \begin{verbatim}
-#> [1] 6
-\end{verbatim}
-
-\begin{verbatim}
-shape_val <- calculate_effective_shape_value(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2")
-shape_val
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 2.019414
-\end{verbatim}
-
-\begin{verbatim}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2", shape_val = 1.833091, num_bins_x = num_bins_x)
+#> [1] 4
+\end{verbatim}
+
+\begin{verbatim}
+num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, 
+                                         y = "UMAP2", hex_size = NA)
 num_bins_y
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-18-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-19-1.pdf}
 
@@
 \subsubsection{Construct the 2D model with different options}\label{construct-the-2d-model-with-different-options}}
-
-\hypertarget{construct-the-high-d-model-with-different-options}{%
-\subsubsection{Construct the high-D model with different options}\label{construct-the-high-d-model-with-different-options}}
-
-\begin{verbatim}
-## To generate a data set with high-D and 2D training data
-df_all <- training_data |> dplyr::select(-ID) |>
-  dplyr::bind_cols(s_curve_noise_umap_with_id)
-
-## To generate averaged high-D data
-
-df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
-\end{verbatim}
-
-\hypertarget{generate-the-triangular-mesh}{%
-\subsubsection{Generate the triangular mesh}\label{generate-the-triangular-mesh}}
 
@@
 \begin{verbatim}
-#> # A tibble: 20 x 4
 #>             x          y hexID std_counts
-#>     <dbl>  <dbl> <int>      <dbl>
-#>  1 -2.91  -4.91      7      0.5  
-#>  2 -3.54  -3.81     13      0.5  
-#>  3 -2.29  -6.01      2      0.5  
-#>  4 -1.66  -4.91      8      0.75 
-#>  5 -2.29  -3.81     14      0.5  
-#>  6 -1.03  -3.81     15      0.375
-#>  7 -0.407 -2.71     21      0.125
-#>  8 -1.03  -1.61     27      0.125
-#>  9  0.219 -1.61     28      1    
-#> 10  0.845 -0.511    34      0.625
-#> 11  0.845  1.69     46      0.125
-#> 12  0.845  3.89     58      0.625
-#> 13  0.219  4.99     64      0.125
-#> 14  0.845  6.09     70      0.375
-#> 15  1.47   0.589    41      1    
-#> 16  1.47   2.79     53      0.125
-#> 17  2.10   3.89     59      0.625
-#> 18  1.47   4.99     65      0.5  
-#> 19  2.10   6.09     71      0.375
-#> 20  2.72   4.99     66      0.5
-\end{verbatim}
-
-\begin{verbatim}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+#> 1  -2.6742223 -4.4744481     5     1.0000
+#> 2  -1.5408019 -6.2798254     2     0.3125
+#> 3  -0.4073814 -4.4744481     6     0.0625
+#> 4  -1.5408019 -2.6690708    10     0.2500
+#> 5  -0.4073814 -0.8636935    14     0.5000
+#> 6   0.7260390 -2.6690708    11     0.1250
+#> 7   1.8594594 -0.8636935    15     0.1875
+#> 8   0.7260390  0.9416838    19     0.6250
+#> 9   1.8594594  2.7470611    23     0.2500
+#> 10  0.7260390  4.5524384    27     0.5625
+#> 11  1.8594594  6.3578158    31     0.3750
+#> 12  2.9928798  4.5524384    28     0.4375
+\end{verbatim}
+
+\hypertarget{construct-the-high-d-model-with-different-options}{%
+\subsubsection{Construct the high-D model with different options}\label{construct-the-high-d-model-with-different-options}}
+
+\begin{verbatim}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-ID) |>
+  dplyr::bind_cols(s_curve_noise_umap_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
+\end{verbatim}
+
+\hypertarget{generate-the-triangular-mesh}{%
+\subsubsection{Generate the triangular mesh}\label{generate-the-triangular-mesh}}
+
+\begin{verbatim}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
 tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
@@
 \begin{verbatim}
-#> Rows: 72
+#> Rows: 32
 #> Columns: 2
-#> $ x <dbl> -3.2700000, -2.6436763, -3.2700000, -2.6436763, -3.2700000, -2.64367~
-#> $ y <dbl> -5.7400000, -4.6399676, -3.5399352, -2.4399028, -1.3398704, -0.23983~
+#> $ x <dbl> -3.8076427, -2.6742223, -3.8076427, -2.6742223, -3.8076427, -2.67422~
+#> $ y <dbl> -6.2798254, -4.4744481, -2.6690708, -0.8636935, 0.9416838, 2.7470611~
 \end{verbatim}
@@
 \begin{verbatim}
-#> Rows: 432
+#> Rows: 192
 #> Columns: 3
-#> $ x  <dbl> -2.643676, -2.643676, -3.270000, -3.896324, -3.896324, -3.270000, -~
-#> $ y  <dbl> -5.3748152, -6.1051848, -6.4703696, -6.1051848, -5.3748152, -5.0096~
+#> $ x  <dbl> -2.674222, -2.674222, -3.807643, -4.941063, -4.941063, -3.807643, -~
+#> $ y  <dbl> -5.6804828, -6.8791681, -7.4785108, -6.8791681, -5.6804828, -5.0811~
 #> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-25-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-26-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-31-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-32-1.pdf}
 
@@
 \begin{verbatim}
-#> # A tibble: 21 x 4
 #>             x          y hexID std_counts
-#>     <dbl> <dbl> <int>      <dbl>
-#>  1 -3.27  -5.74     1      0.286
-#>  2 -2.64  -4.64     7      0.571
-#>  3 -3.27  -3.54    13      0.714
-#>  4 -2.02  -5.74     2      0.857
-#>  5 -1.39  -4.64     8      0.429
-#>  6 -2.02  -3.54    14      0.286
-#>  7 -0.765 -3.54    15      0.429
-#>  8 -0.138 -2.44    21      0.143
-#>  9 -0.765 -1.34    27      0.286
-#> 10  0.488 -1.34    28      1    
-#> # i 11 more rows
-\end{verbatim}
-
-\begin{verbatim}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+#> 1  -2.6742223 -4.4744481     5     1.0000
+#> 2  -1.5408019 -6.2798254     2     0.3125
+#> 3  -0.4073814 -4.4744481     6     0.0625
+#> 4  -1.5408019 -2.6690708    10     0.2500
+#> 5  -0.4073814 -0.8636935    14     0.5000
+#> 6   0.7260390 -2.6690708    11     0.1250
+#> 7   1.8594594 -0.8636935    15     0.1875
+#> 8   0.7260390  0.9416838    19     0.6250
+#> 9   1.8594594  2.7470611    23     0.2500
+#> 10  0.7260390  4.5524384    27     0.5625
+#> 11  1.8594594  6.3578158    31     0.3750
+#> 12  2.9928798  4.5524384    28     0.4375
+\end{verbatim}
+
+\begin{verbatim}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
 tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-35-1.pdf}
-
-\begin{verbatim}
-benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-37-1.pdf}
+
+\begin{verbatim}
+benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
+benchmark
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 2.603
+\end{verbatim}
+
+\begin{verbatim}
+benchmark <- 5
 \end{verbatim}
@@
 
+\texttt{fit\_high\_d\_model()} function is used to generate the 2D model and the high-D model.
+
+\begin{verbatim}
+fit_high_d_model(training_data = training_data, nldr_df_with_id = s_curve_noise_umap, x = "UMAP1",
+                             y = "UMAP1", num_bins_x = NA, num_bins_y = NA,
+                             hex_size = NA, buffer_size = NA,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+\end{verbatim}
+
+\begin{verbatim}
+#> $df_bin
+#> # A tibble: 6 x 8
+#>   hb_id     x1    x2     x3        x4        x5        x6        x7
+#>   <int>  <dbl> <dbl>  <dbl>     <dbl>     <dbl>     <dbl>     <dbl>
+#> 1     1 -0.470 0.473 -1.75   0.00121   0.00244  -0.0314   -0.000525
+#> 2     2  0.170 1.23  -1.76   0.00610  -0.00313   0.00286  -0.00123 
+#> 3     6  0.760 1.19  -0.407 -0.000154  0.00342   0.0219    0.000410
+#> 4    11 -0.452 1.17   0.294  0.00414  -0.000295  0.00334   0.00207 
+#> 5    14 -0.613 1.69   1.79   0.0119   -0.00418  -0.000519 -0.00251 
+#> 6    15  0.366 0.999  1.70  -0.00115   0.00375  -0.00747  -0.000931
+#> 
+#> $df_bin_centroids
+#>            x         y hexID std_counts
+#> 1 -3.8076427 -3.807643     1 0.41666667
+#> 2 -1.5408019 -3.807643     2 0.62500000
+#> 3 -0.4073814 -1.540802     6 0.41666667
+#> 4 -0.4073814  2.992880    14 0.08333333
+#> 5  0.7260390  0.726039    11 0.58333333
+#> 6  1.8594594  2.992880    15 1.00000000
+\end{verbatim}
+
+\begin{verbatim}
+## To generate a data set with high-D and 2D training data
+df_all <- training_data |> dplyr::select(-ID) |>
+  dplyr::bind_cols(s_curve_noise_umap_with_id)
+
+## To generate averaged high-D data
+
+df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
+\end{verbatim}
+
 \hypertarget{predict-2d-embeddings}{%
@@
 
-\hypertarget{compute-residuals}{%
-\subsubsection{Compute residuals}\label{compute-residuals}}
+To predict the 2D embeddings for a new data point using the trained high-D model, we follow a series of steps outlined below:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Compute high-dimensional Euclidean distance}: Calculate the Euclidean distance between each new data point and the lift model's coordinates in the high-dimensional space.
+This distance metric helps identify the nearest lift model coordinates to each new data point.
+\item
+  \textbf{Find the nearest lift model coordinates}: Determine the lift model's high-dimensional coordinates that are closest to each new data point based on the computed distances.
+This step helps establish a correspondence between the new data points and the lift model's representation in the high-dimensional space.
+\item
+  \textbf{Map the hexagonal bin ID}: Assign each new data point to a hexagonal bin based on its nearest lift model coordinates.
+This mapping ensures that each data point is associated with a specific region in the hexagonal grid representation of the high-dimensional space.
+\item
+  \textbf{Map the hexagonal bin centroid coordinates in 2D}: Transform the hexagonal bin centroid coordinates from the high-dimensional space to the 2D embedding space using the trained lift model.
+This mapping provides the 2D embeddings for the new data points, allowing them to be visualized and analyzed in a lower-dimensional space.
+\end{enumerate}
+
+The \texttt{predict\_2d\_embeddings()} function is used to predict 2D embeddings for a new data point.
+The inputs for the function are test data, bin centroid coordinates in 2D, lifting model coordinates in high-D, and the type of the NLDR technique.
+The output contains the predicted 2D embeddings along with the predicted hexagonal id and the ID of the test data.
+
+\begin{verbatim}
+pred_df_test <- predict_2d_embeddings(test_data = training_data,
+df_bin_centroids = df_bin_centroids, df_bin = df_bin, type_NLDR = "UMAP")
+
+glimpse(pred_df_test)
+\end{verbatim}
+
+\begin{verbatim}
+#> Rows: 75
+#> Columns: 4
+#> $ pred_UMAP_1 <dbl> -2.6742223, 1.8594594, 0.7260390, -1.5408019, -1.5408019, ~
+#> $ pred_UMAP_2 <dbl> -4.4744481, -0.8636935, 0.9416838, -6.2798254, -2.6690708,~
+#> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
+#> $ pred_hb_id  <int> 5, 15, 19, 2, 10, 31, 15, 11, 15, 28, 11, 5, 28, 5, 11, 23~
+\end{verbatim}
+
+\hypertarget{making-summaries}{%
+\subsubsection{Making summaries}\label{making-summaries}}
+
+There are two important summaries that should be made when constructing the model on a dataset using a specific NLDR technique, the Mean Square Error (MSE) and Akaike Information Criterion (AIC) which measure prediction accuracy.
+The function that perform the summaries called \texttt{generate\_eval\_df()}.
+The output of this function is a list which contains MSE and AIC.
+The code below uses \texttt{generate\_eval\_df()} to find the MSE and the AIC values for the model constructed on UMAP 2D embeddings generated from the \texttt{s\_curve\_noise} training data.
+
+\begin{verbatim}
+generate_summary(test_data = training_data, prediction_df = pred_df_test,
+                 df_bin = df_bin, col_start = "x")
+\end{verbatim}
+
+\begin{verbatim}
+#> $mse
+#> [1] 0.2983091
+#> 
+#> $aic
+#> [1] -467.0531
+\end{verbatim}
 
@@
 
-\hypertarget{geom_trimesh}{%
-\paragraph{geom\_trimesh}\label{geom_trimesh}}
+We use static visualizations to understand the model constructed in 2D. On the other hand, dynamic visualization is used to see how the model looks in high-D space.
+
+\hypertarget{static-visualizations}{%
+\paragraph{Static visualizations}\label{static-visualizations}}
+
+Static visualizations main involves two types of results. One is the triangulation and the other is the long edge removal. Both types of visualizations provide ggplot objects.
+
+\begin{itemize}
+\item
+  Triangulation result: To visualize the results of triangulation, we input a dataset containing hexagonal bin centroid coordinates where 2D embedding data exists.
+\texttt{geom\_trimesh()} is used to visualize this result.
+\item
+  Long edge removal: The long edge removal process involves identifying and removing long edges from the triangular mesh.
+Table shows the main arguments of the functions.
+We offer two functions for visualizing this process:
+\item
+  \texttt{colour\_long\_edges()}: This function colors the long edges within the triangular mesh by red.
+\item
+  \texttt{remove\_long\_edges()}: After identifying long edges, this function draws the triangular mesh without the long edges.
+\end{itemize}
+
+\hypertarget{dynamic-visaulizations}{%
+\paragraph{Dynamic visaulizations}\label{dynamic-visaulizations}}
+
+The \texttt{show\_langevitour()} function enables dynamic visualization of the 2D model alongside the high-dimensional (high-D) data in its original space.
+This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data.
+The main arguments are shown in Table
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-36-1.pdf}
-
-\hypertarget{coloured_long_edges}{%
-\paragraph{coloured\_long\_edges}\label{coloured_long_edges}}
-
-\begin{verbatim}
-trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
-                                triangular_object = tr1_object, distance_col = distance)
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-42-1.pdf}
+
+\begin{verbatim}
+trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-37-1.pdf}
-
-\hypertarget{remove-long-edges}{%
-\paragraph{remove long edges}\label{remove-long-edges}}
-
-\begin{verbatim}
-trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
-                                     triangular_object = tr1_object, distance_col = distance)
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-43-1.pdf}
+
+\begin{verbatim}
+trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
 trimesh_removed
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-38-1.pdf}
-
-\hypertarget{show_langevitour}{%
-\paragraph{show\_langevitour}\label{show_langevitour}}
-
-\begin{verbatim}
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
-                          distance = distance, distance_col = "distance")
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-44-1.pdf}
+
+\begin{verbatim}
+tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, 
+                          benchmark_value = benchmark,
+                          distance = distance, distance_col = "distance", 
+                          use_default_benchmark_val = FALSE, 
+                          column_start_text = "x")
 tour1
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-39-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-45-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-40-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-42-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-48-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-43-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-49-1.pdf}
 
@@
 #>     Automatically selected t = 22
-#>   Calculated optimal t in 0.39 seconds.
+#>   Calculated optimal t in 2.20 seconds.
 #>   Calculating diffusion potential...
-#>   Calculated diffusion potential in 0.32 seconds.
+#>   Calculated diffusion potential in 0.44 seconds.
 #>   Calculating metric MDS...
-#>   Calculated metric MDS in 8.83 seconds.
-#> Calculated PHATE in 9.55 seconds.
+#>   Calculated metric MDS in 20.54 seconds.
+#> Calculated PHATE in 23.19 seconds.
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-44-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-50-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-45-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-51-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}
-
-\begin{verbatim}
-num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", cell_area = 1)
-num_bins_x <- 13
-\end{verbatim}
-
-\begin{verbatim}
-shape_val <- calculate_effective_shape_value(.data = tSNE_data, x = "tSNE1", y = "tSNE2")
-shape_val
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 1.13882
-\end{verbatim}
-
-\begin{verbatim}
-num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, x = "tSNE1", y = "tSNE2", shape_val = 0.8417289, num_bins_x = num_bins_x)
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-52-1.pdf}
+
+\begin{verbatim}
+num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", hex_size = NA)
+num_bins_x
+\end{verbatim}
+
+\begin{verbatim}
+#> [1] 29
+\end{verbatim}
+
+\begin{verbatim}
+num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, y = "tSNE2", hex_size = NA)
 num_bins_y
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-50-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-55-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-51-1.pdf}
-
-\begin{verbatim}
-df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
-  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
-  dplyr::distinct() |>
-  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
-
-df_bin_centroids
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 21 x 4
-#>        x     y hexID std_counts
-#>    <dbl> <dbl> <int>      <dbl>
-#>  1 -27.7 -34.8     1      0.286
-#>  2 -25.4 -30.1    14      0.286
-#>  3 -27.7 -25.4    27      0.286
-#>  4 -25.4 -20.7    40      0.286
-#>  5 -27.7 -16.0    53      0.429
-#>  6 -25.4 -11.3    66      0.143
-#>  7 -23.2 -34.8     2      0.857
-#>  8 -20.9 -30.1    15      0.429
-#>  9 -23.2 -25.4    28      1    
-#> 10 -20.9 -20.7    41      0.714
-#> # i 11 more rows
-\end{verbatim}
-
-\begin{verbatim}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-56-1.pdf}
+
+\begin{verbatim}
+df_bin_centroids <- extract_hexbin_centroids(hex_full_count_df)
+\end{verbatim}
+
+\begin{verbatim}
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
 tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-55-1.pdf}
-
-\begin{verbatim}
-benchmark <- find_benchmark_value(.data = distance, distance_col = "distance")
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-60-1.pdf}
+
+\begin{verbatim}
+benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-56-1.pdf}
-
-\begin{verbatim}
-trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark,
-                                triangular_object = tr1_object, distance_col = distance)
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-61-1.pdf}
+
+\begin{verbatim}
+trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-57-1.pdf}
-
-\begin{verbatim}
-trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark,
-                                     triangular_object = tr1_object, distance_col = distance)
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-62-1.pdf}
+
+\begin{verbatim}
+trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
+                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
 trimesh_removed
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-58-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-63-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-59-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-64-1.pdf}
 

### 2024-02-20 quollr_code.R
Added shift hex grid origin to main function

@@ -177,7 +177,8 @@ generate_full_grid_centroids <- function(nldr_df, x = "UMAP1", y = "UMAP2",
 }
 
 extract_coord_of_shifted_hex_grid <- function(nldr_df, x = "UMAP1", y = "UMAP2",
-                                         num_bins_x, num_bins_y, shift_x = 0, shift_y = 0, buffer_size = NA, hex_size = NA){
+                                         num_bins_x, num_bins_y, shift_x = 0, shift_y = 0,
+                                         buffer_size = NA, hex_size = NA){
 
   ## hex size is not provided
   if (is.na(hex_size)) {
@@ -966,7 +967,10 @@ fit_high_d_model <- function(training_data, nldr_df_with_id, x = "UMAP1",
                              is_bin_centroid = TRUE,
                              is_rm_lwd_hex = FALSE,
                              benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x") {
+                             is_avg_high_d = TRUE, column_start_text = "x",
+                             is_shift_origin = FALSE,
+                             shift_x = NA, shift_y = NA
+                             ) {
 
   ## hex size is not provided
   if (is.na(hex_size)) {
@@ -1009,13 +1013,26 @@ fit_high_d_model <- function(training_data, nldr_df_with_id, x = "UMAP1",
   }
 
   ### Generate the full grid
+  ## If shift happen or not
+  if (isTRUE(is_shift_origin)) {
 
-  all_centroids_df <- generate_full_grid_centroids(nldr_df = nldr_df_with_id,
-                                                   x = x, y = y,
-                                                   num_bins_x = num_bins_x,
-                                                   num_bins_y = num_bins_y,
-                                                   buffer_size = buffer_size, hex_size = hex_size)
+    all_centroids_df <- extract_coord_of_shifted_hex_grid(nldr_df = nldr_df_with_id,
+                                                          x = x, y = y,
+                                                          num_bins_x = num_bins_x,
+                                                          num_bins_y = num_bins_y,
+                                                          shift_x = shift_x, shift_y = shift_y,
+                                                          buffer_size = buffer_size, hex_size = hex_size)
 
+  } else {
+
+    all_centroids_df <- generate_full_grid_centroids(nldr_df = nldr_df_with_id,
+                                                     x = x, y = y,
+                                                     num_bins_x = num_bins_x,
+                                                     num_bins_y = num_bins_y,
[…]

### 2024-02-21 paper-quollr.Rmd
Added shift hex grid origin to main function
Attempt 1 to have regular hexagons, still need to find out which hex_size and buffer_size need to initialize, needed to restrict the user inputs for hex_size and buffer_size because too large hex_size and buffer_size are not compatibel with data range

@@
 
+```{r}
+library(scales)
+s_curve_noise_umap$UMAP1 <- rescale(s_curve_noise_umap$UMAP1)
+
+y_min <- -sqrt(3)/2
+y_max <- sqrt(3)/2
+
+s_curve_noise_umap$UMAP2 <- ((s_curve_noise_umap$UMAP2 - min(s_curve_noise_umap$UMAP2))/(max(s_curve_noise_umap$UMAP2) - min(s_curve_noise_umap$UMAP2))) * (y_max - y_min)
+
+# Define the desired size of the hexagons
+hex_size <- 0.2
+
+# Compute the horizontal spacing (dx) and vertical spacing (dy) between hexagon centers
+dx <- sqrt(3) * hex_size
+dy <- 1.5 * hex_size
+
+x_length <- 1
+y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))
+
+# Compute the number of bins along the x-axis and y-axis
+# num_bins_x <- ceiling(x_length / dx)
+# num_bins_y <- ceiling(y_length_scaled / dy)
+```
+
+```{r}
+# s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)] <- scale(s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)])
+```
+
 # Introduction
@@
 
-<!--
-### Appropriate hex size value to have regular hexagons
-
-```{r}
-
-min_value_x <- min(s_curve_noise_umap$UMAP1)
-max_value_x <- max(s_curve_noise_umap$UMAP1)
-min_value_y <- min(s_curve_noise_umap$UMAP2)
-max_value_y <- max(s_curve_noise_umap$UMAP2)
-
-# Calculate the maximum span along x and y axes
-x_span <- max_value_x - min_value_x
-y_span <- max_value_y - min_value_y
-
-hex_size <- min(x_span, y_span)/5.5
-
-num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap, x = "UMAP1", hex_size = hex_size)
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap, y = "UMAP2", hex_size = hex_size)
-
-all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
-                                                 x = "UMAP1", y = "UMAP2", 
-                                                 num_bins_x = num_bins_x, 
-                                                 num_bins_y = num_bins_y, 
-                                                 buffer_size = NA, hex_size = hex_size)
-hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = hex_size)
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
-```
--->
-
 ### Compute hexagonal bin configurations
@@
 num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap,
-                                         x = "UMAP1", hex_size = NA)
+                                         x = "UMAP1", hex_size = 0.2)
 num_bins_x
@@
 num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap,
-                                         y = "UMAP2", hex_size = NA)
+                                         y = "UMAP2", hex_size = 0.2)
 num_bins_y
@@
 
-- First compute hex grid bound values along the x and y axis and generate the all the points wthin the hex box
-
-```{r}
-cell_area <- 1
-
-hex_size <- sqrt(2 * cell_area / sqrt(3))
-
-buffer_size <- hex_size/2
-
-x_bounds <- seq(min(s_curve_noise_umap[["UMAP1"]]) - buffer_size,
-                  max(s_curve_noise_umap[["UMAP1"]]) + buffer_size, length.out = num_bins_x)
-
-y_bounds <- seq(min(s_curve_noise_umap[["UMAP2"]]) - buffer_size,
-                max(s_curve_noise_umap[["UMAP2"]]) + buffer_size, length.out = num_bins_y)
-
-box_points <- expand.grid(x = x_bounds, y = y_bounds)
-
-ggplot() +
-  geom_point(data = box_points, aes(x = x, y = y), color = "red")
-
-```
+- First compute hex grid bound values along the x and y axis and generate the all the points within the hex box
 
@@
 
-```{r}
- box_points <- box_points |>
-    dplyr::arrange(x) |>
-    dplyr::group_by(x) |>
-    dplyr::group_modify(~ generate_even_y(.x)) |>
-    tibble::as_tibble()
-
-ggplot() +
-  geom_point(data = box_points,
-             aes(x = x, y = y, colour = as.factor(is_even)))
-```
-
 - Then, shift the x values of the even rows
-
-```{r}
-## Shift for even values in x-axis
-x_shift <- unique(box_points$x)[2] - unique(box_points$x)[1]
-
-box_points$x <- box_points$x + x_shift/2 * ifelse(box_points$is_even == 1, 1, 0)
-
-ggplot() +
-  geom_point(data = box_points, aes(x = x, y = y), color = "red")
-
-```
 
@@
                                                  num_bins_y = num_bins_y, 
-                                                 buffer_size = NA, hex_size = NA)
+                                                 buffer_x = NA, buffer_y = NA, 
+                                                 hex_size = 0.2)
 
@@
 ```{r}
-hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = NA)
+hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = 0.2)
 glimpse(hex_grid)
@@
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
+  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red") +
+  coord_fixed()
 ```
@@
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
+  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID)) +
+  coord_fixed()
 ```
@@
 
-4. Assign data into hexagons
+5. Assign data into hexagons
 
@@
 
-5. Compute standardized counts
+6. Compute standardized counts
 
@@
 
-6. Extract full grid info
+7. Extract full grid info
 
@@
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
+  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue") +
+  coord_fixed()
 ```
@@
   geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+  coord_fixed()
 
@@
 
-```{r}
-all_centroids_df_shift <- extract_coord_of_shifted_hex_grid(nldr_df = s_curve_noise_umap, 
-                                                 x = "UMAP1", y = "UMAP2", 
-                                                 num_bins_x = num_bins_x, 
-                                                 num_bins_y = num_bins_y,
-                                                 shift_x = 0.2690002, shift_y = 0.271183,
-                                                 buffer_size = NA, hex_size = NA)
-
-glimpse(all_centroids_df_shift)
-```
-
-```{r}
-hex_grid <- gen_hex_coordinates(all_centroids_df_shift)
-glimpse(hex_grid)
-```
-
-```{r}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = all_centroids_df_shift, aes(x = x, y = y), color = "red")
-```
-
-```{r}
-full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df_shift)
-
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
-```
-
-```{r}
-full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-```
-
-```{r}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
-```
-
-```{r}
-df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
-```
-
-```{r}
-hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
-```
-
-```{r}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
-```
-
-```{r}
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
-
-```
-
-```{r}
-df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
-  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
-  dplyr::distinct() |>
-  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
-
-df_bin_centroids
-```
-
-```{r}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-```
-
-```{r, eval=FALSE}
-bin_centroids_shift <- ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
-  geom_point(color = "#bdbdbd") +
-  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "#feb24c") +
-  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-hex_grid_shift <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
-  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
-  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
-               fill = NA, color = "#bdbdbd") +
-  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-## Before shift
-before_shift_plot <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_void() +
-  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-## After shift
-after_shift_plot <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
 #### Benchmark value to remove the low-density hexagons
-
-```{r}
-## As an option first quantile considered as a default
-benchmark_to_rm_lwd_hex <- quantile(df_bin_centroids$std_counts)[2] + 0.01
-
-## To identify low density hexagons
-df_bin_centroids_low <- df_bin_centroids |>
-  dplyr::filter(std_counts <= benchmark_to_rm_lwd_hex)
-
-## To identify low-density hexagons needed to remove by investigating neighbouring mean density
-identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_all = df_bin_centroids, num_bins_x = num_bins_x,
-                     df_bin_centroids_low = df_bin_centroids_low)
-```
 
@@
 benchmark
-benchmark <- 5
 ```
@@
 `fit_high_d_model()` function is used to generate the 2D model and the high-D model.
-
-```{r}
-fit_high_d_model(training_data = training_data, nldr_df_with_id = s_curve_noise_umap, x = "UMAP1",
-                             y = "UMAP1", num_bins_x = NA, num_bins_y = NA,
-                             hex_size = NA, buffer_size = NA,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-```
 
@@
 
+### Find non-empty bins
+
 ## Tests
@@
 # Application
-
-```{r}
-medlea_df <- read_csv("data/medlea_dataset.csv")
-names(medlea_df)[2:(NCOL(medlea_df) - 1)] <- paste0("x", 1:(NCOL(medlea_df) - 2))
-
-medlea_df <- medlea_df |> ## Since only contains zeros
-  select(-x10)
-
-#medlea_df[,2:(NCOL(medlea_df) - 1)] <- scale(medlea_df[,2:(NCOL(medlea_df) - 1)])
-
-calculate_pca <- function(feature_dataset, num_pcs){
-  pcaY_cal <- prcomp(feature_dataset, center = TRUE, scale = TRUE)
-  PCAresults <- data.frame(pcaY_cal$x[, 1:num_pcs])
-  summary_pca <- summary(pcaY_cal)
-  var_explained_df <- data.frame(PC= paste0("PC",1:50),
-                               var_explained=(pcaY_cal$sdev[1:50])^2/sum((pcaY_cal$sdev[1:50])^2))
-  return(list(prcomp_out = pcaY_cal,pca_components = PCAresults, summary = summary_pca, var_explained_pca  = var_explained_df))
-}
-features <- medlea_df[,2:(NCOL(medlea_df) - 1)]
-pca_ref_calc <- calculate_pca(features, 8) 
-pca_ref_calc$summary
-
-var_explained_df <- pca_ref_calc$var_explained_pca
-data_pca <- pca_ref_calc$pca_components |>
-  mutate(ID = 1:NROW(pca_ref_calc$pca_components),
-         shape_label = medlea_df$Shape_label)
-
-var_explained_df |>
-  ggplot(aes(x = PC,y = var_explained, group = 1))+
-  geom_point(size=1)+
-  geom_line()+
-  labs(title="Scree plot: PCA on scaled data") +
-  scale_x_discrete(limits = paste0(rep("PC", 50), 1:50)) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-data_split <- initial_split(data_pca)
-training_data <- training(data_split) |>
-  arrange(ID)
-test_data <- testing(data_split) |>
-  arrange(ID)
-```
-
-```{r}
-UMAP_fit <- umap(training_data |> dplyr::select(-c(ID, shape_label)), n_neighbors = 37, n_components =  2)
-
-UMAP_data <- UMAP_fit$layout |>
-  as.data.frame()
-names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
-
-UMAP_data <- UMAP_data |>
-  mutate(ID = training_data$id)
-
-UMAP_data_with_label <- UMAP_data |>
-  mutate(shape_label = training_data$shape_label)
-```
-
-```{r}
-UMAP_data_with_label |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-```
-
-```{r}
-tSNE_data <- Fit_tSNE(training_data |> dplyr::select(-c(ID, shape_label)), opt_perplexity = calculate_effective_perplexity(training_data |> dplyr::select(-c(ID, shape_label))), with_seed = 20240110)
-
-tSNE_data <- tSNE_data |>
-  select(-ID) |>
-  mutate(ID = training_data$ID)
-
-tSNE_data_with_label <- tSNE_data |>
-  mutate(shape_label = training_data$shape_label)
-
-tSNE_data_with_label |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-```
-
-```{r}
-PHATE_data <- Fit_PHATE(training_data |> dplyr::select(-c(ID, shape_label)), knn = 5, with_seed = 20240110)
-PHATE_data <- PHATE_data |>
-  select(PHATE1, PHATE2)
-PHATE_data <- PHATE_data |>
-  mutate(ID = training_data$ID)
-
-PHATE_data_with_label <- PHATE_data |>
-  mutate(shape_label = training_data$shape_label)
-
-PHATE_data_with_label |>
-    ggplot(aes(x = PHATE1,
-               y = PHATE2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-```
-
-```{r}
-tem_dir <- tempdir()
-
-Fit_TriMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
-
-path <- file.path(tem_dir, "df_2_without_class.csv")
-path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
-
-Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
-
-TriMAP_data <- read_csv(path2)
-TriMAP_data <- TriMAP_data |>
-  mutate(ID = training_data$ID)
-
-TriMAP_data_with_label <- TriMAP_data |>
-  mutate(shape_label = training_data$shape_label)
-
-TriMAP_data_with_label |>
-    ggplot(aes(x = TriMAP1,
-               y = TriMAP2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-
-```
-
-```{r}
-tem_dir <- tempdir()
-
-Fit_PacMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
-
-path <- file.path(tem_dir, "df_2_without_class.csv")
-path2 <- file.path(tem_dir, "dataset_3_PaCMAP_values.csv")
-
-Fit_PaCMAP(as.integer(2), as.integer(10), "random", 0.9, as.integer(2), path, path2)
-
-PaCMAP_data <- read_csv(path2)
-PaCMAP_data <- PaCMAP_data |>
-  mutate(ID = training_data$ID)
-
-PaCMAP_data_with_label <- PaCMAP_data |>
-  mutate(shape_label = training_data$shape_label)
-
-PaCMAP_data_with_label |>
-    ggplot(aes(x = PaCMAP1,
-               y = PaCMAP2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-```
-
-```{r}
-num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", hex_size = NA)
-num_bins_x
-```
-
-```{r}
-num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, y = "tSNE2", hex_size = NA)
-num_bins_y
-```
-
-```{r}
-all_centroids_df <- generate_full_grid_centroids(nldr_df = tSNE_data, 
-                                                 x = "tSNE1", y = "tSNE2", 
-                                                 num_bins_x = num_bins_x, 
-                                                 num_bins_y = num_bins_y, 
-                                                 buffer_size = NA, hex_size = NA)
-
-hex_grid <- gen_hex_coordinates(all_centroids_df)
-
-full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
-
-full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-
-tSNE_data_with_id <- assign_data(tSNE_data, full_grid_with_hexbin_id)
-
-df_with_std_counts <- compute_std_counts(nldr_df = tSNE_data_with_id)
-
-hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
-
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
-```
-
-```{r}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = tSNE_data, aes(x = tSNE1, y = tSNE2), color = "blue")
-```
-
-```{r}
-df_bin_centroids <- extract_hexbin_centroids(hex_full_count_df)
-```
-
-```{r}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-```
-
-```{r}
-## To generate a data set with high-D and 2D training data
-df_all <- training_data |> dplyr::select(-c(ID, shape_label)) |>
-  dplyr::bind_cols(tSNE_data_with_id)
-
-## To generate averaged high-D data
-
-df_bin <- avg_highD_data(.data = df_all, column_start_text = "PC") ## Need to pass ID column name
-```
-
-```{r}
-## Compute 2D distances
-distance <- cal_2d_dist(tr_from_to_df_coord = tr_from_to_df)
-
-plot_dist(distance)
-
-benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
-```
-
-```{r}
-trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
-  geom_point(size = 0.1) +
-  geom_trimesh() +
-  coord_equal()
-
-trimesh
-```
-
-```{r}
-trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
-
-trimesh_gr
-```
-
-```{r}
-trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
-trimesh_removed
-```
-
-```{r}
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
-                          distance = distance, distance_col = "distance", column_start_text = "PC")
-tour1
-```
 

### 2024-02-21 paper-quollr.tex
Added shift hex grid origin to main function
Attempt 1 to have regular hexagons, still need to find out which hex_size and buffer_size need to initialize, needed to restrict the user inputs for hex_size and buffer_size because too large hex_size and buffer_size are not compatibel with data range

@@
 reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
+\end{verbatim}
+
+\begin{verbatim}
+library(scales)
+s_curve_noise_umap$UMAP1 <- rescale(s_curve_noise_umap$UMAP1)
+
+y_min <- -sqrt(3)/2
+y_max <- sqrt(3)/2
+
+s_curve_noise_umap$UMAP2 <- ((s_curve_noise_umap$UMAP2 - min(s_curve_noise_umap$UMAP2))/(max(s_curve_noise_umap$UMAP2) - min(s_curve_noise_umap$UMAP2))) * (y_max - y_min)
+
+# Define the desired size of the hexagons
+hex_size <- 0.2
+
+# Compute the horizontal spacing (dx) and vertical spacing (dy) between hexagon centers
+dx <- sqrt(3) * hex_size
+dy <- 1.5 * hex_size
+
+x_length <- 1
+y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))
+
+# Compute the number of bins along the x-axis and y-axis
+# num_bins_x <- ceiling(x_length / dx)
+# num_bins_y <- ceiling(y_length_scaled / dy)
+\end{verbatim}
+
+\begin{verbatim}
+# s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)] <- scale(s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)])
 \end{verbatim}
@@
 
---\textgreater{}
-
 \hypertarget{compute-hexagonal-bin-configurations}{%
@@
 \begin{verbatim}
-#> [1] 4
+#> [1] 3
 \end{verbatim}
@@
 \begin{verbatim}
-#> [1] 8
+#> [1] 6
 \end{verbatim}
@@
 \begin{itemize}
-\tightlist
 \item
-  First compute hex grid bound values along the x and y axis and generate the all the points wthin the hex box
-\end{itemize}
-
-\begin{verbatim}
-cell_area <- 1
-
-hex_size <- sqrt(2 * cell_area / sqrt(3))
-
-buffer_size <- hex_size/2
-
-x_bounds <- seq(min(s_curve_noise_umap[["UMAP1"]]) - buffer_size,
-                  max(s_curve_noise_umap[["UMAP1"]]) + buffer_size, length.out = num_bins_x)
-
-y_bounds <- seq(min(s_curve_noise_umap[["UMAP2"]]) - buffer_size,
-                max(s_curve_noise_umap[["UMAP2"]]) + buffer_size, length.out = num_bins_y)
-
-box_points <- expand.grid(x = x_bounds, y = y_bounds)
-
-ggplot() +
-  geom_point(data = box_points, aes(x = x, y = y), color = "red")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
-
-\begin{itemize}
-\tightlist
+  First compute hex grid bound values along the x and y axis and generate the all the points within the hex box
 \item
   Second for each x-value, find which y values are in the even row
-\end{itemize}
-
-\begin{verbatim}
- box_points <- box_points |>
-    dplyr::arrange(x) |>
-    dplyr::group_by(x) |>
-    dplyr::group_modify(~ generate_even_y(.x)) |>
-    tibble::as_tibble()
-
-ggplot() +
-  geom_point(data = box_points,
-             aes(x = x, y = y, colour = as.factor(is_even)))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf}
-
-\begin{itemize}
-\tightlist
 \item
@@
 \end{itemize}
-
-\begin{verbatim}
-## Shift for even values in x-axis
-x_shift <- unique(box_points$x)[2] - unique(box_points$x)[1]
-
-box_points$x <- box_points$x + x_shift/2 * ifelse(box_points$is_even == 1, 1, 0)
-
-ggplot() +
-  geom_point(data = box_points, aes(x = x, y = y), color = "red")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 \begin{verbatim}
-#> Rows: 32
+#> Rows: 18
 #> Columns: 2
-#> $ x <dbl> -3.8076427, -2.6742223, -3.8076427, -2.6742223, -3.8076427, -2.67422~
-#> $ y <dbl> -6.2798254, -4.4744481, -2.6690708, -0.8636935, 0.9416838, 2.7470611~
+#> $ x <dbl> 0.1464102, 0.4928203, 0.8392305, 0.3196152, 0.6660254, 1.0124356, 0.~
+#> $ y <dbl> 0.1845299, 0.1845299, 0.1845299, 0.4845299, 0.4845299, 0.4845299, 0.~
 \end{verbatim}
@@
 \begin{verbatim}
-hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = NA)
+hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = 0.2)
 glimpse(hex_grid)
@@
 \begin{verbatim}
-#> Rows: 192
+#> Rows: 108
 #> Columns: 3
-#> $ x  <dbl> -2.674222, -2.674222, -3.807643, -4.941063, -4.941063, -3.807643, -~
-#> $ y  <dbl> -5.6804828, -6.8791681, -7.4785108, -6.8791681, -5.6804828, -5.0811~
+#> $ x  <dbl> 0.31961524, 0.31961524, 0.14641016, -0.02679492, -0.02679492, 0.146~
+#> $ y  <dbl> 0.28412287, 0.08493702, -0.01465590, 0.08493702, 0.28412287, 0.3837~
 #> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
@@
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red")
+  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red") +
+  coord_fixed()
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-10-1.pdf}
 
@@
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
+  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID)) +
+  coord_fixed()
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
+\setcounter{enumi}{4}
 \tightlist
@@
 \def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{4}
+\setcounter{enumi}{5}
 \tightlist
@@
 \def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{5}
+\setcounter{enumi}{6}
 \tightlist
@@
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
+  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue") +
+  coord_fixed()
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-18-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}
 
@@
   geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+  coord_fixed()
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-19-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}
 
@@
 #>            x         y hexID std_counts
-#> 1  -2.6742223 -4.4744481     5     1.0000
-#> 2  -1.5408019 -6.2798254     2     0.3125
-#> 3  -0.4073814 -4.4744481     6     0.0625
-#> 4  -1.5408019 -2.6690708    10     0.2500
-#> 5  -0.4073814 -0.8636935    14     0.5000
-#> 6   0.7260390 -2.6690708    11     0.1250
-#> 7   1.8594594 -0.8636935    15     0.1875
-#> 8   0.7260390  0.9416838    19     0.6250
-#> 9   1.8594594  2.7470611    23     0.2500
-#> 10  0.7260390  4.5524384    27     0.5625
-#> 11  1.8594594  6.3578158    31     0.3750
-#> 12  2.9928798  4.5524384    28     0.4375
+#> 1  0.1464102 0.1845299     1 1.00000000
+#> 2  0.4928203 0.1845299     2 0.04761905
+#> 3  0.3196152 0.4845299     4 0.19047619
+#> 4  0.6660254 0.4845299     5 0.23809524
+#> 5  0.4928203 0.7845299     8 0.19047619
+#> 6  0.8392305 0.7845299     9 0.33333333
+#> 7  0.6660254 1.0845299    11 0.28571429
+#> 8  1.0124356 1.0845299    12 0.04761905
+#> 9  0.8392305 1.3845299    15 0.38095238
+#> 10 0.6660254 1.6845299    17 0.47619048
+#> 11 1.0124356 1.6845299    18 0.38095238
 \end{verbatim}
@@
 
-\begin{verbatim}
-all_centroids_df_shift <- extract_coord_of_shifted_hex_grid(nldr_df = s_curve_noise_umap, 
-                                                 x = "UMAP1", y = "UMAP2", 
-                                                 num_bins_x = num_bins_x, 
-                                                 num_bins_y = num_bins_y,
-                                                 shift_x = 0.2690002, shift_y = 0.271183,
-                                                 buffer_size = NA, hex_size = NA)
-
-glimpse(all_centroids_df_shift)
-\end{verbatim}
-
-\begin{verbatim}
-#> Rows: 32
-#> Columns: 2
-#> $ x <dbl> -3.8076427, -2.6742223, -3.8076427, -2.6742223, -3.8076427, -2.67422~
-#> $ y <dbl> -6.2798254, -4.4744481, -2.6690708, -0.8636935, 0.9416838, 2.7470611~
-\end{verbatim}
-
-\begin{verbatim}
-hex_grid <- gen_hex_coordinates(all_centroids_df_shift)
-glimpse(hex_grid)
-\end{verbatim}
-
-\begin{verbatim}
-#> Rows: 192
-#> Columns: 3
-#> $ x  <dbl> -2.674222, -2.674222, -3.807643, -4.941063, -4.941063, -3.807643, -~
-#> $ y  <dbl> -5.6804828, -6.8791681, -7.4785108, -6.8791681, -5.6804828, -5.0811~
-#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
-\end{verbatim}
-
-\begin{verbatim}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = all_centroids_df_shift, aes(x = x, y = y), color = "red")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-25-1.pdf}
-
-\begin{verbatim}
-full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df_shift)
-
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-26-1.pdf}
-
-\begin{verbatim}
-full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-\end{verbatim}
-
-\begin{verbatim}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
-\end{verbatim}
-
-\begin{verbatim}
-df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
-\end{verbatim}
-
-\begin{verbatim}
-hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
-\end{verbatim}
-
-\begin{verbatim}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-31-1.pdf}
-
-\begin{verbatim}
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-32-1.pdf}
-
-\begin{verbatim}
-df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
-  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
-  dplyr::distinct() |>
-  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
-
-df_bin_centroids
-\end{verbatim}
-
-\begin{verbatim}
-#>             x          y hexID std_counts
-#> 1  -2.6742223 -4.4744481     5     1.0000
-#> 2  -1.5408019 -6.2798254     2     0.3125
-#> 3  -0.4073814 -4.4744481     6     0.0625
-#> 4  -1.5408019 -2.6690708    10     0.2500
-#> 5  -0.4073814 -0.8636935    14     0.5000
-#> 6   0.7260390 -2.6690708    11     0.1250
-#> 7   1.8594594 -0.8636935    15     0.1875
-#> 8   0.7260390  0.9416838    19     0.6250
-#> 9   1.8594594  2.7470611    23     0.2500
-#> 10  0.7260390  4.5524384    27     0.5625
-#> 11  1.8594594  6.3578158    31     0.3750
-#> 12  2.9928798  4.5524384    28     0.4375
-\end{verbatim}
-
-\begin{verbatim}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-\end{verbatim}
-
-\begin{verbatim}
-bin_centroids_shift <- ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
-  geom_point(color = "#bdbdbd") +
-  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "#feb24c") +
-  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-hex_grid_shift <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
-  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
-  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
-               fill = NA, color = "#bdbdbd") +
-  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-## Before shift
-before_shift_plot <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_void() +
-  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-## After shift
-after_shift_plot <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
-\end{verbatim}
-
 \hypertarget{benchmark-value-to-remove-the-low-density-hexagons}{%
 \paragraph{Benchmark value to remove the low-density hexagons}\label{benchmark-value-to-remove-the-low-density-hexagons}}
-
-\begin{verbatim}
-## As an option first quantile considered as a default
-benchmark_to_rm_lwd_hex <- quantile(df_bin_centroids$std_counts)[2] + 0.01
-
-## To identify low density hexagons
-df_bin_centroids_low <- df_bin_centroids |>
-  dplyr::filter(std_counts <= benchmark_to_rm_lwd_hex)
-
-## To identify low-density hexagons needed to remove by investigating neighbouring mean density
-identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_all = df_bin_centroids, num_bins_x = num_bins_x,
-                     df_bin_centroids_low = df_bin_centroids_low)
-\end{verbatim}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-37-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-21-1.pdf}
 
@@
 \begin{verbatim}
-#> [1] 2.603
-\end{verbatim}
-
-\begin{verbatim}
-benchmark <- 5
+#> [1] 0.6
 \end{verbatim}
@@
 \texttt{fit\_high\_d\_model()} function is used to generate the 2D model and the high-D model.
-
-\begin{verbatim}
-fit_high_d_model(training_data = training_data, nldr_df_with_id = s_curve_noise_umap, x = "UMAP1",
-                             y = "UMAP1", num_bins_x = NA, num_bins_y = NA,
-                             hex_size = NA, buffer_size = NA,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-\end{verbatim}
-
-\begin{verbatim}
-#> $df_bin
-#> # A tibble: 6 x 8
-#>   hb_id     x1    x2     x3        x4        x5        x6        x7
-#>   <int>  <dbl> <dbl>  <dbl>     <dbl>     <dbl>     <dbl>     <dbl>
-#> 1     1 -0.470 0.473 -1.75   0.00121   0.00244  -0.0314   -0.000525
-#> 2     2  0.170 1.23  -1.76   0.00610  -0.00313   0.00286  -0.00123 
-#> 3     6  0.760 1.19  -0.407 -0.000154  0.00342   0.0219    0.000410
-#> 4    11 -0.452 1.17   0.294  0.00414  -0.000295  0.00334   0.00207 
-#> 5    14 -0.613 1.69   1.79   0.0119   -0.00418  -0.000519 -0.00251 
-#> 6    15  0.366 0.999  1.70  -0.00115   0.00375  -0.00747  -0.000931
-#> 
-#> $df_bin_centroids
-#>            x         y hexID std_counts
-#> 1 -3.8076427 -3.807643     1 0.41666667
-#> 2 -1.5408019 -3.807643     2 0.62500000
-#> 3 -0.4073814 -1.540802     6 0.41666667
-#> 4 -0.4073814  2.992880    14 0.08333333
-#> 5  0.7260390  0.726039    11 0.58333333
-#> 6  1.8594594  2.992880    15 1.00000000
-\end{verbatim}
 
@@
 #> Columns: 4
-#> $ pred_UMAP_1 <dbl> -2.6742223, 1.8594594, 0.7260390, -1.5408019, -1.5408019, ~
-#> $ pred_UMAP_2 <dbl> -4.4744481, -0.8636935, 0.9416838, -6.2798254, -2.6690708,~
+#> $ pred_UMAP_1 <dbl> 0.1464102, 0.8392305, 1.0124356, 0.1464102, 0.3196152, 0.6~
+#> $ pred_UMAP_2 <dbl> 0.1845299, 0.7845299, 1.0845299, 0.1845299, 0.4845299, 1.6~
 #> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
-#> $ pred_hb_id  <int> 5, 15, 19, 2, 10, 31, 15, 11, 15, 28, 11, 5, 28, 5, 11, 23~
+#> $ pred_hb_id  <int> 1, 9, 12, 1, 4, 17, 9, 5, 9, 18, 5, 1, 18, 1, 5, 15, 2, 15~
 \end{verbatim}
@@
 #> $mse
-#> [1] 0.2983091
+#> [1] 0.372034
 #> 
 #> $aic
-#> [1] -467.0531
+#> [1] -365.1042
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-42-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-25-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-43-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-26-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-44-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-27-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-45-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-28-1.pdf}
+
+\hypertarget{find-non-empty-bins}{%
+\subsubsection{Find non-empty bins}\label{find-non-empty-bins}}
 
@@
 \section{Application}\label{application}}
-
-\begin{verbatim}
-medlea_df <- read_csv("data/medlea_dataset.csv")
-names(medlea_df)[2:(NCOL(medlea_df) - 1)] <- paste0("x", 1:(NCOL(medlea_df) - 2))
-
-medlea_df <- medlea_df |> ## Since only contains zeros
-  select(-x10)
-
-#medlea_df[,2:(NCOL(medlea_df) - 1)] <- scale(medlea_df[,2:(NCOL(medlea_df) - 1)])
-
-calculate_pca <- function(feature_dataset, num_pcs){
-  pcaY_cal <- prcomp(feature_dataset, center = TRUE, scale = TRUE)
-  PCAresults <- data.frame(pcaY_cal$x[, 1:num_pcs])
-  summary_pca <- summary(pcaY_cal)
-  var_explained_df <- data.frame(PC= paste0("PC",1:50),
-                               var_explained=(pcaY_cal$sdev[1:50])^2/sum((pcaY_cal$sdev[1:50])^2))
-  return(list(prcomp_out = pcaY_cal,pca_components = PCAresults, summary = summary_pca, var_explained_pca  = var_explained_df))
-}
-features <- medlea_df[,2:(NCOL(medlea_df) - 1)]
-pca_ref_calc <- calculate_pca(features, 8) 
-pca_ref_calc$summary
-\end{verbatim}
-
-\begin{verbatim}
-#> Importance of components:
-#>                           PC1    PC2    PC3     PC4     PC5     PC6     PC7
-#> Standard deviation     3.1691 3.0609 2.7226 1.87967 1.71219 1.34192 1.27525
-#> Proportion of Variance 0.1969 0.1837 0.1453 0.06928 0.05748 0.03531 0.03189
-#> Cumulative Proportion  0.1969 0.3806 0.5260 0.59526 0.65274 0.68805 0.71993
-#>                            PC8     PC9    PC10    PC11    PC12    PC13   PC14
-#> Standard deviation     1.16992 1.13465 1.06628 1.03279 0.97899 0.96264 0.9528
-#> Proportion of Variance 0.02684 0.02524 0.02229 0.02091 0.01879 0.01817 0.0178
-#> Cumulative Proportion  0.74677 0.77202 0.79431 0.81522 0.83402 0.85219 0.8700
-#>                          PC15   PC16    PC17    PC18    PC19    PC20    PC21
-#> Standard deviation     0.9116 0.9090 0.79750 0.76725 0.72414 0.65310 0.61052
-#> Proportion of Variance 0.0163 0.0162 0.01247 0.01154 0.01028 0.00836 0.00731
-#> Cumulative Proportion  0.8863 0.9025 0.91496 0.92650 0.93678 0.94514 0.95245
-#>                          PC22    PC23    PC24    PC25    PC26   PC27    PC28
-#> Standard deviation     0.6019 0.55399 0.52293 0.46638 0.41959 0.3976 0.34697
-#> Proportion of Variance 0.0071 0.00602 0.00536 0.00426 0.00345 0.0031 0.00236
-#> Cumulative Proportion  0.9596 0.96557 0.97093 0.97520 0.97865 0.9818 0.98411
-#>                           PC29    PC30    PC31    PC32    PC33    PC34    PC35
-#> Standard deviation     0.33415 0.30618 0.29237 0.28458 0.26033 0.25420 0.22792
-#> Proportion of Variance 0.00219 0.00184 0.00168 0.00159 0.00133 0.00127 0.00102
-#> Cumulative Proportion  0.98630 0.98814 0.98982 0.99140 0.99273 0.99400 0.99502
-#>                           PC36    PC37    PC38   PC39    PC40    PC41    PC42
-#> Standard deviation     0.21644 0.20437 0.19127 0.1744 0.15586 0.15252 0.12519
-#> Proportion of Variance 0.00092 0.00082 0.00072 0.0006 0.00048 0.00046 0.00031
-#> Cumulative Proportion  0.99594 0.99676 0.99747 0.9981 0.99855 0.99900 0.99931
-#>                           PC43    PC44    PC45    PC46    PC47    PC48    PC49
-#> Standard deviation     0.10485 0.08598 0.08008 0.06491 0.04841 0.04094 0.03791
-#> Proportion of Variance 0.00022 0.00014 0.00013 0.00008 0.00005 0.00003 0.00003
-#> Cumulative Proportion  0.99952 0.99967 0.99980 0.99988 0.99992 0.99996 0.99999
-#>                           PC50    PC51
-#> Standard deviation     0.02347 0.01421
-#> Proportion of Variance 0.00001 0.00000
-#> Cumulative Proportion  1.00000 1.00000
-\end{verbatim}
-
-\begin{verbatim}
-var_explained_df <- pca_ref_calc$var_explained_pca
-data_pca <- pca_ref_calc$pca_components |>
-  mutate(ID = 1:NROW(pca_ref_calc$pca_components),
-         shape_label = medlea_df$Shape_label)
-
-var_explained_df |>
-  ggplot(aes(x = PC,y = var_explained, group = 1))+
-  geom_point(size=1)+
-  geom_line()+
-  labs(title="Scree plot: PCA on scaled data") +
-  scale_x_discrete(limits = paste0(rep("PC", 50), 1:50)) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-46-1.pdf}
-
-\begin{verbatim}
-data_split <- initial_split(data_pca)
-training_data <- training(data_split) |>
-  arrange(ID)
-test_data <- testing(data_split) |>
-  arrange(ID)
-\end{verbatim}
-
-\begin{verbatim}
-UMAP_fit <- umap(training_data |> dplyr::select(-c(ID, shape_label)), n_neighbors = 37, n_components =  2)
-
-UMAP_data <- UMAP_fit$layout |>
-  as.data.frame()
-names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
-
-UMAP_data <- UMAP_data |>
-  mutate(ID = training_data$id)
-
-UMAP_data_with_label <- UMAP_data |>
-  mutate(shape_label = training_data$shape_label)
-\end{verbatim}
-
-\begin{verbatim}
-UMAP_data_with_label |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-48-1.pdf}
-
-\begin{verbatim}
-tSNE_data <- Fit_tSNE(training_data |> dplyr::select(-c(ID, shape_label)), opt_perplexity = calculate_effective_perplexity(training_data |> dplyr::select(-c(ID, shape_label))), with_seed = 20240110)
-
-tSNE_data <- tSNE_data |>
-  select(-ID) |>
-  mutate(ID = training_data$ID)
-
-tSNE_data_with_label <- tSNE_data |>
-  mutate(shape_label = training_data$shape_label)
-
-tSNE_data_with_label |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-49-1.pdf}
-
-\begin{verbatim}
-PHATE_data <- Fit_PHATE(training_data |> dplyr::select(-c(ID, shape_label)), knn = 5, with_seed = 20240110)
-\end{verbatim}
-
-\begin{verbatim}
-#> Calculating PHATE...
-#>   Running PHATE on 824 observations and 8 variables.
-#>   Calculating graph and diffusion operator...
-#>     Calculating KNN search...
-#>     Calculating affinities...
-#>   Calculated graph and diffusion operator in 0.01 seconds.
-#>   Calculating optimal t...
-#>     Automatically selected t = 22
-#>   Calculated optimal t in 2.20 seconds.
-#>   Calculating diffusion potential...
-#>   Calculated diffusion potential in 0.44 seconds.
-#>   Calculating metric MDS...
-#>   Calculated metric MDS in 20.54 seconds.
-#> Calculated PHATE in 23.19 seconds.
-\end{verbatim}
-
-\begin{verbatim}
-PHATE_data <- PHATE_data |>
-  select(PHATE1, PHATE2)
-PHATE_data <- PHATE_data |>
-  mutate(ID = training_data$ID)
-
-PHATE_data_with_label <- PHATE_data |>
-  mutate(shape_label = training_data$shape_label)
-
-PHATE_data_with_label |>
-    ggplot(aes(x = PHATE1,
-               y = PHATE2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-50-1.pdf}
-
-\begin{verbatim}
-tem_dir <- tempdir()
-
-Fit_TriMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
-
-path <- file.path(tem_dir, "df_2_without_class.csv")
-path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
-
-Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
-
-TriMAP_data <- read_csv(path2)
-TriMAP_data <- TriMAP_data |>
-  mutate(ID = training_data$ID)
-
-TriMAP_data_with_label <- TriMAP_data |>
-  mutate(shape_label = training_data$shape_label)
-
-TriMAP_data_with_label |>
-    ggplot(aes(x = TriMAP1,
-               y = TriMAP2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-51-1.pdf}
-
-\begin{verbatim}
-tem_dir <- tempdir()
-
-Fit_PacMAP_data(training_data |> dplyr::select(-c(ID, shape_label)), tem_dir)
-
-path <- file.path(tem_dir, "df_2_without_class.csv")
-path2 <- file.path(tem_dir, "dataset_3_PaCMAP_values.csv")
-
-Fit_PaCMAP(as.integer(2), as.integer(10), "random", 0.9, as.integer(2), path, path2)
-
-PaCMAP_data <- read_csv(path2)
-PaCMAP_data <- PaCMAP_data |>
-  mutate(ID = training_data$ID)
-
-PaCMAP_data_with_label <- PaCMAP_data |>
-  mutate(shape_label = training_data$shape_label)
-
-PaCMAP_data_with_label |>
-    ggplot(aes(x = PaCMAP1,
-               y = PaCMAP2, color = shape_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd"))
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-52-1.pdf}
-
-\begin{verbatim}
-num_bins_x <- calculate_effective_x_bins(.data = tSNE_data, x = "tSNE1", hex_size = NA)
-num_bins_x
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 29
-\end{verbatim}
-
-\begin{verbatim}
-num_bins_y <- calculate_effective_y_bins(.data = tSNE_data, y = "tSNE2", hex_size = NA)
-num_bins_y
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 38
-\end{verbatim}
-
-\begin{verbatim}
-all_centroids_df <- generate_full_grid_centroids(nldr_df = tSNE_data, 
-                                                 x = "tSNE1", y = "tSNE2", 
-                                                 num_bins_x = num_bins_x, 
-                                                 num_bins_y = num_bins_y, 
-                                                 buffer_size = NA, hex_size = NA)
-
-hex_grid <- gen_hex_coordinates(all_centroids_df)
-
-full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
-
-full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-
-tSNE_data_with_id <- assign_data(tSNE_data, full_grid_with_hexbin_id)
-
-df_with_std_counts <- compute_std_counts(nldr_df = tSNE_data_with_id)
-
-hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
-
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-55-1.pdf}
-
-\begin{verbatim}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = tSNE_data, aes(x = tSNE1, y = tSNE2), color = "blue")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-56-1.pdf}
-
-\begin{verbatim}
-df_bin_centroids <- extract_hexbin_centroids(hex_full_count_df)
-\end{verbatim}
-
-\begin{verbatim}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-\end{verbatim}
-
-\begin{verbatim}
-## To generate a data set with high-D and 2D training data
-df_all <- training_data |> dplyr::select(-c(ID, shape_label)) |>
-  dplyr::bind_cols(tSNE_data_with_id)
-
-## To generate averaged high-D data
-
-df_bin <- avg_highD_data(.data = df_all, column_start_text = "PC") ## Need to pass ID column name
-\end{verbatim}
-
-\begin{verbatim}
-## Compute 2D distances
-distance <- cal_2d_dist(tr_from_to_df_coord = tr_from_to_df)
-
-plot_dist(distance)
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-60-1.pdf}
-
-\begin{verbatim}
-benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
-\end{verbatim}
-
-\begin{verbatim}
-trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
-  geom_point(size = 0.1) +
-  geom_trimesh() +
-  coord_equal()
-
-trimesh
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-61-1.pdf}
-
-\begin{verbatim}
-trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
-
-trimesh_gr
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-62-1.pdf}
-
-\begin{verbatim}
-trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
-trimesh_removed
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-63-1.pdf}
-
-\begin{verbatim}
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark,
-                          distance = distance, distance_col = "distance", column_start_text = "PC")
-tour1
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-64-1.pdf}
 

### 2024-02-22 paper-quollr.Rmd
Add the hexagonal binning version 2 which initialize the parameters x_start, y_start, buffer_x, buffer_y, hex_size. Now need to find limitations for each parameters and recheck the default for hex_size

@@
                                                  num_bins_y = num_bins_y, 
+                                                 x_start = NA, y_start = NA, 
                                                  buffer_x = NA, buffer_y = NA, 
@@
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red") +
+  geom_point(aes(x = c_x, y = c_y), color = "red") +
   coord_fixed()
@@
 
-3. Map hexagonal IDs
-
-Steps:
-
-- Filter the data set with specific y value
-
-- Order the x values for a specific y value
-
-- Repeat the process for all unique y values
-
 ```{r}
-full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
+full_grid_centroids_with_hexbin_id <- hex_grid |>
+  dplyr::select("c_x", "c_y", "hexID") |>
+    dplyr::distinct()
 
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
   coord_fixed()
-```
-
-4. Map polygon IDs
-
-Steps:
-
-- Filter specific hexagon
-
-- Filter specific polygon
-
-- Check the selected hexagonal centroid exists within the polygon
-
-- if so assign that id to centroid, if not check until find the polygon which contains the centroid
-
-```{r}
-full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
 ```
@@
 ```{r}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_centroids_with_hexbin_id)
 ```
@@
 ```{r}
-hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+hex_full_count_df <- generate_full_grid_info(hex_grid, df_with_std_counts)
 ```
@@
 ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_polygon(color = "black", aes(group = id, fill = std_counts)) +
   geom_text(aes(x = c_x, y = c_y, label = hexID)) +

### 2024-02-22 paper-quollr.tex
Add the hexagonal binning version 2 which initialize the parameters x_start, y_start, buffer_x, buffer_y, hex_size. Now need to find limitations for each parameters and recheck the default for hex_size

@@
 \begin{verbatim}
-#> [1] 3
+#> [1] 4
 \end{verbatim}
@@
 \begin{verbatim}
-#> [1] 6
+#> [1] 7
 \end{verbatim}
@@
 \begin{verbatim}
-#> Rows: 18
+#> Rows: 28
 #> Columns: 2
-#> $ x <dbl> 0.1464102, 0.4928203, 0.8392305, 0.3196152, 0.6660254, 1.0124356, 0.~
-#> $ y <dbl> 0.1845299, 0.1845299, 0.1845299, 0.4845299, 0.4845299, 0.4845299, 0.~
+#> $ x <dbl> 0.0000000, 0.3464102, 0.6928203, 1.0392305, 0.1732051, 0.5196152, 0.~
+#> $ y <dbl> 0.0, 0.0, 0.0, 0.0, 0.3, 0.3, 0.3, 0.3, 0.6, 0.6, 0.6, 0.6, 0.9, 0.9~
 \end{verbatim}
@@
 \begin{verbatim}
-#> Rows: 108
-#> Columns: 3
-#> $ x  <dbl> 0.31961524, 0.31961524, 0.14641016, -0.02679492, -0.02679492, 0.146~
-#> $ y  <dbl> 0.28412287, 0.08493702, -0.01465590, 0.08493702, 0.28412287, 0.3837~
-#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
+#> Rows: 168
+#> Columns: 6
+#> $ c_x   <dbl> 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.0000000~
+#> $ c_y   <dbl> 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0,~
+#> $ x     <dbl> 0.1732051, 0.1732051, 0.0000000, -0.1732051, -0.1732051, 0.00000~
+#> $ y     <dbl> 0.09959292, -0.09959292, -0.19918584, -0.09959292, 0.09959292, 0~
+#> $ id    <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4~
+#> $ hexID <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4~
 \end{verbatim}
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  Map hexagonal IDs
-\end{enumerate}
-
-Steps:
-
-\begin{itemize}
-\item
-  Filter the data set with specific y value
-\item
-  Order the x values for a specific y value
-\item
-  Repeat the process for all unique y values
-\end{itemize}
-
 \begin{verbatim}
-full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)
+full_grid_centroids_with_hexbin_id <- hex_grid |>
+  dplyr::select("c_x", "c_y", "hexID") |>
+    dplyr::distinct()
 
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
   coord_fixed()
@@
 \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  Map polygon IDs
-\end{enumerate}
-
-Steps:
-
-\begin{itemize}
-\item
-  Filter specific hexagon
-\item
-  Filter specific polygon
-\item
-  Check the selected hexagonal centroid exists within the polygon
-\item
-  if so assign that id to centroid, if not check until find the polygon which contains the centroid
-\end{itemize}
-
-\begin{verbatim}
-full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-\end{verbatim}
 
@@
 \begin{verbatim}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_centroids_with_hexbin_id)
 \end{verbatim}
@@
 \begin{verbatim}
-hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
+hex_full_count_df <- generate_full_grid_info(hex_grid, df_with_std_counts)
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-15-1.pdf}
 
@@
 #>            x   y hexID std_counts
-#> 1  0.1464102 0.1845299     1 1.00000000
-#> 2  0.4928203 0.1845299     2 0.04761905
-#> 3  0.3196152 0.4845299     4 0.19047619
-#> 4  0.6660254 0.4845299     5 0.23809524
-#> 5  0.4928203 0.7845299     8 0.19047619
-#> 6  0.8392305 0.7845299     9 0.33333333
-#> 7  0.6660254 1.0845299    11 0.28571429
-#> 8  1.0124356 1.0845299    12 0.04761905
-#> 9  0.8392305 1.3845299    15 0.38095238
-#> 10 0.6660254 1.6845299    17 0.47619048
-#> 11 1.0124356 1.6845299    18 0.38095238
+#> 1  0.0000000 0.0     1 0.40000000
+#> 2  0.3464102 0.0     2 0.26666667
+#> 3  0.1732051 0.3     5 0.86666667
+#> 4  0.5196152 0.3     6 0.13333333
+#> 5  0.3464102 0.6    10 0.20000000
+#> 6  0.6928203 0.6    11 0.53333333
+#> 7  0.8660254 0.9    15 0.80000000
+#> 8  0.6928203 1.2    19 0.06666667
+#> 9  0.5196152 1.5    22 0.06666667
+#> 10 0.8660254 1.5    23 1.00000000
+#> 11 0.6928203 1.8    27 0.46666667
+#> 12 1.0392305 1.8    28 0.20000000
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-21-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-20-1.pdf}
 
@@
 #> Columns: 4
-#> $ pred_UMAP_1 <dbl> 0.1464102, 0.8392305, 1.0124356, 0.1464102, 0.3196152, 0.6~
-#> $ pred_UMAP_2 <dbl> 0.1845299, 0.7845299, 1.0845299, 0.1845299, 0.4845299, 1.6~
+#> $ pred_UMAP_1 <dbl> 0.1732051, 0.8660254, 0.8660254, 0.0000000, 0.5196152, 0.6~
+#> $ pred_UMAP_2 <dbl> 0.3, 0.9, 0.9, 0.0, 0.3, 1.8, 0.9, 0.6, 0.9, 1.8, 0.6, 0.3~
 #> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
-#> $ pred_hb_id  <int> 1, 9, 12, 1, 4, 17, 9, 5, 9, 18, 5, 1, 18, 1, 5, 15, 2, 15~
+#> $ pred_hb_id  <int> 5, 15, 15, 1, 6, 27, 15, 11, 15, 28, 11, 5, 23, 5, 11, 22,~
 \end{verbatim}
@@
 #> $mse
-#> [1] 0.372034
+#> [1] 0.2612147
 #> 
 #> $aic
-#> [1] -365.1042
+#> [1] -536.7666
 \end{verbatim}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-26-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-25-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-27-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-26-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-28-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-27-1.pdf}
 

### 2024-02-22 quollr_code.R

@@ -1,4 +1,4 @@
-calculate_effective_x_bins <- function(.data, x = "UMAP1", hex_size = NA){
+calculate_effective_x_bins <- function(.data, x = "UMAP1", hex_size = NA, buffer_x = NA){
 
   if (anyNA(.data[[rlang::as_string(rlang::ensym(x))]])) {
     stop("NAs present")
@@ -23,18 +23,28 @@ calculate_effective_x_bins <- function(.data, x = "UMAP1", hex_size = NA){
     }
   }
 
+  if (is.na(buffer_x)) {
+    #buffer_x <- sqrt(3) * hex_size/2
+    # buffer_x <- min(min(.data[[rlang::as_string(rlang::sym(x))]]) - sqrt(3) * hex_size/2,
+    #                 max(.data[[rlang::as_string(rlang::sym(x))]]) + sqrt(3) * hex_size/2,
+    #                 s_curve_noise_umap[which.min(s_curve_noise_umap$UMAP2), "UMAP1"] |> dplyr::pull("UMAP1") - (sqrt(3) * hex_size/2),
+    #                 s_curve_noise_umap[which.max(s_curve_noise_umap$UMAP2), "UMAP1"] |> dplyr::pull("UMAP1") +
+    #                   (sqrt(3) * hex_size/2))
+    buffer_x <- sqrt(3) * hex_size/2
+  }
+
   ## To compute the range along x-axis
   xwidth <- diff(range(.data |>
-                         dplyr::pull({{ x }})))
+                         dplyr::pull({{ x }})))  + buffer_x
 
-  horizontal_spacing <- sqrt(3) * hex_size
+  horizontal_spacing <- (sqrt(3) * hex_size)
 
   num_bins <- ceiling(xwidth/horizontal_spacing)
   num_bins
 
 }
 
-calculate_effective_y_bins <- function(.data, y = "UMAP2", hex_size = NA){
+calculate_effective_y_bins <- function(.data, y = "UMAP2", hex_size = NA, buffer_y = NA){
 
   if (anyNA(.data[[rlang::as_string(rlang::ensym(y))]])) {
     stop("NAs present")
@@ -57,12 +67,24 @@ calculate_effective_y_bins <- function(.data, y = "UMAP2", hex_size = NA){
     }
   }
 
+  if (is.na(buffer_y)) {
+
+    #buffer_y <- 1.5 * hex_size/2
+    # buffer_y <- min(min(.data[[rlang::as_string(rlang::sym(y))]]) - 1.5 * hex_size/2,
[…]

### 2024-02-22 paper-quollr.Rmd
Changed the buffer size defaults and add restriction that could only have values less than vector or horizontal spacing

@@
                                                  x_start = NA, y_start = NA, 
-                                                 buffer_x = NA, buffer_y = NA, 
                                                  hex_size = 0.2)

### 2024-02-22 quollr_code.R

@@ -11,10 +11,11 @@ calculate_effective_x_bins <- function(.data, x = "UMAP1", hex_size = NA, buffer
 
 
   if (is.na(hex_size)) {
-    cell_area <- 1
-
-    ## To compute the radius of the hexagon outer circle
-    hex_size <- sqrt(2 * cell_area / (3 *sqrt(3)))
+    # cell_area <- 1
+    #
+    # ## To compute the radius of the hexagon outer circle
+    # hex_size <- sqrt(2 * cell_area / (3 *sqrt(3)))
+    hex_size <- 0.2
 
   } else {
     if ((hex_size <= 0) || (is.infinite(hex_size))) {
@@ -30,14 +31,25 @@ calculate_effective_x_bins <- function(.data, x = "UMAP1", hex_size = NA, buffer
     #                 s_curve_noise_umap[which.min(s_curve_noise_umap$UMAP2), "UMAP1"] |> dplyr::pull("UMAP1") - (sqrt(3) * hex_size/2),
     #                 s_curve_noise_umap[which.max(s_curve_noise_umap$UMAP2), "UMAP1"] |> dplyr::pull("UMAP1") +
     #                   (sqrt(3) * hex_size/2))
-    buffer_x <- sqrt(3) * hex_size/2
+    buffer_x <- sqrt(3) * hex_size
+    message(paste0("Buffer along the x-axis set to ", buffer_x, "."))
+  } else {
+
+    ## Buffer size is exceeds
+    if (buffer_x > (sqrt(3) * hex_size)) {
+      stop(paste0("Buffer along the y-axis exceeds than ", sqrt(3) * hex_size, ".
+                  Need to assign a value less than ", sqrt(3) * hex_size, "."))
+
+    }
+
+
   }
 
   ## To compute the range along x-axis
   xwidth <- diff(range(.data |>
                          dplyr::pull({{ x }})))  + buffer_x
 
-  horizontal_spacing <- (sqrt(3) * hex_size)
+  horizontal_spacing <- sqrt(3) * hex_size
 
   num_bins <- ceiling(xwidth/horizontal_spacing)
   num_bins
@@ -55,10 +67,11 @@ calculate_effective_y_bins <- function(.data, y = "UMAP2", hex_size = NA, buffer
   }
 
   if (is.na(hex_size)) {
-    cell_area <- 1
-
-    ## To compute the radius of the hexagon outer circle
[…]

### 2024-02-22 paper-quollr.Rmd
Change benchmark value to preserve S-curve structure
Since one point was observed at the border in the hexgrid...understood that x_start and y_start shouldn't be always the min values along the axes, need to reduce by some factor which could be able to fit all the bins inside the grid. Therefore reduced by the half of vector/horizontal spacing. Why half? because we don't want to over estimate the x and y start, if a point which is at boudary or the outside the boundary means that it should drag to the grid by the spacing aspects.

@@
 
-Steps:
-
-- First compute hex grid bound values along the x and y axis and generate the all the points within the hex box
-
-- Second for each x-value, find which y values are in the even row
-
-- Then, shift the x values of the even rows
-
 ```{r}
@@
 2. Generate hexagonal coordinates
-
-Steps:
-- Compute horizontal width of the hexagon
-
-- Compute vertical width of the hexagon and multiply by a factor for overlapping ($sqrt(3) / 2 * 1.15$)
-
-- Obtain hexagon polygon coordinates
-
-- Obtain the number of hexagons in the full grid
-
-- Generate the coordinates for the hexagons
 
@@
 benchmark
+
+benchmark <- 0.75
 ```

### 2024-02-22 paper-quollr.tex
Change benchmark value to preserve S-curve structure
Since one point was observed at the border in the hexgrid...understood that x_start and y_start shouldn't be always the min values along the axes, need to reduce by some factor which could be able to fit all the bins inside the grid. Therefore reduced by the half of vector/horizontal spacing. Why half? because we don't want to over estimate the x and y start, if a point which is at boudary or the outside the boundary means that it should drag to the grid by the spacing aspects.

@@
 
-Steps:
-
-\begin{itemize}
-\item
-  First compute hex grid bound values along the x and y axis and generate the all the points within the hex box
-\item
-  Second for each x-value, find which y values are in the even row
-\item
-  Then, shift the x values of the even rows
-\end{itemize}
-
 \begin{verbatim}
@@
 #> Columns: 2
-#> $ x <dbl> 0.0000000, 0.3464102, 0.6928203, 1.0392305, 0.1732051, 0.5196152, 0.~
-#> $ y <dbl> 0.0, 0.0, 0.0, 0.0, 0.3, 0.3, 0.3, 0.3, 0.6, 0.6, 0.6, 0.6, 0.9, 0.9~
+#> $ x <dbl> -0.1732051, 0.1732051, 0.5196152, 0.8660254, 0.0000000, 0.3464102, 0~
+#> $ y <dbl> -0.15, -0.15, -0.15, -0.15, 0.15, 0.15, 0.15, 0.15, 0.45, 0.45, 0.45~
 \end{verbatim}
@@
 
-Steps:
-- Compute horizontal width of the hexagon
-
-\begin{itemize}
-\item
-  Compute vertical width of the hexagon and multiply by a factor for overlapping (\(sqrt(3) / 2 * 1.15\))
-\item
-  Obtain hexagon polygon coordinates
-\item
-  Obtain the number of hexagons in the full grid
-\item
-  Generate the coordinates for the hexagons
-\end{itemize}
-
 \begin{verbatim}
@@
 #> Columns: 6
-#> $ c_x   <dbl> 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.0000000~
-#> $ c_y   <dbl> 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0,~
-#> $ x     <dbl> 0.1732051, 0.1732051, 0.0000000, -0.1732051, -0.1732051, 0.00000~
-#> $ y     <dbl> 0.09959292, -0.09959292, -0.19918584, -0.09959292, 0.09959292, 0~
+#> $ c_x   <dbl> -0.1732051, -0.1732051, -0.1732051, -0.1732051, -0.1732051, -0.1~
+#> $ c_y   <dbl> -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -~
+#> $ x     <dbl> 0.0000000, 0.0000000, -0.1732051, -0.3464102, -0.3464102, -0.173~
+#> $ y     <dbl> -0.05040708, -0.24959292, -0.34918584, -0.24959292, -0.05040708,~
 #> $ id    <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4~
@@
 #>            x     y hexID std_counts
-#> 1  0.0000000 0.0     1 0.40000000
-#> 2  0.3464102 0.0     2 0.26666667
-#> 3  0.1732051 0.3     5 0.86666667
-#> 4  0.5196152 0.3     6 0.13333333
-#> 5  0.3464102 0.6    10 0.20000000
-#> 6  0.6928203 0.6    11 0.53333333
-#> 7  0.8660254 0.9    15 0.80000000
-#> 8  0.6928203 1.2    19 0.06666667
-#> 9  0.5196152 1.5    22 0.06666667
-#> 10 0.8660254 1.5    23 1.00000000
-#> 11 0.6928203 1.8    27 0.46666667
-#> 12 1.0392305 1.8    28 0.20000000
+#> 1  0.1732051 -0.15     2  0.2352941
+#> 2  0.0000000  0.15     5  0.5294118
+#> 3  0.3464102  0.15     6  0.4117647
+#> 4  0.1732051  0.45    10  0.1764706
+#> 5  0.5196152  0.45    11  0.3529412
+#> 6  0.6928203  0.75    15  0.7058824
+#> 7  0.8660254  1.05    20  0.4705882
+#> 8  0.6928203  1.35    23  0.2941176
+#> 9  1.0392305  1.35    24  0.2352941
+#> 10 0.8660254  1.65    28  1.0000000
 \end{verbatim}
@@
 
+\begin{verbatim}
+benchmark <- 0.75
+\end{verbatim}
+
 \hypertarget{model-function}{%
@@
 #> Columns: 4
-#> $ pred_UMAP_1 <dbl> 0.1732051, 0.8660254, 0.8660254, 0.0000000, 0.5196152, 0.6~
-#> $ pred_UMAP_2 <dbl> 0.3, 0.9, 0.9, 0.0, 0.3, 1.8, 0.9, 0.6, 0.9, 1.8, 0.6, 0.3~
+#> $ pred_UMAP_1 <dbl> 0.1732051, 0.6928203, 0.8660254, 0.1732051, 0.1732051, 0.8~
+#> $ pred_UMAP_2 <dbl> 0.45, 0.75, 1.05, -0.15, 0.45, 1.65, 0.75, 0.75, 0.75, 1.6~
 #> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
-#> $ pred_hb_id  <int> 5, 15, 15, 1, 6, 27, 15, 11, 15, 28, 11, 5, 23, 5, 11, 22,~
+#> $ pred_hb_id  <int> 10, 15, 20, 2, 10, 28, 15, 15, 15, 28, 15, 10, 24, 5, 15, ~
 \end{verbatim}
@@
 #> $mse
-#> [1] 0.2612147
+#> [1] 0.3229922
 #> 
 #> $aic
-#> [1] -536.7666
+#> [1] -453.3167
 \end{verbatim}

### 2024-02-22 quollr_code.R

@@ -198,8 +198,7 @@ generate_full_grid_centroids <- function(nldr_df, x = "UMAP1", y = "UMAP2",
   if (is.na(x_start)) {
 
     # Define starting point
-    x_start <- min(nldr_df[[rlang::as_string(rlang::sym(x))]])
-    message(paste0("x_start set to ", x_start, "."))
+    x_start <- min(nldr_df[[rlang::as_string(rlang::sym(x))]]) - (sqrt(3) * hex_size/2)
 
   } else {
     max_x_start <- min(nldr_df[[rlang::as_string(rlang::sym(x))]]) + (sqrt(3) * hex_size)
@@ -215,7 +214,8 @@ generate_full_grid_centroids <- function(nldr_df, x = "UMAP1", y = "UMAP2",
 
   if (is.na(y_start)) {
     # Define starting point
-    y_start <- min(nldr_df[[rlang::as_string(rlang::sym(y))]])
+    y_start <- min(nldr_df[[rlang::as_string(rlang::sym(y))]]) - (1.5 * hex_size/2)
+
     message(paste0("y_start set to ", y_start, "."))
 
 

### 2024-02-22 paper-quollr.Rmd
Simplify the fuctions to do small tasks

@@
 
-use_python("~/miniforge3/envs/pcamp_env/bin/python")
-use_condaenv("pcamp_env")
-
-reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
-reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
+# use_python("~/miniforge3/envs/pcamp_env/bin/python")
+# use_condaenv("pcamp_env")
+# 
+# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
+# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
 
@@
 
+# y_min <- 0
+# y_max <- sqrt(3)/2
+
 s_curve_noise_umap$UMAP2 <- ((s_curve_noise_umap$UMAP2 - min(s_curve_noise_umap$UMAP2))/(max(s_curve_noise_umap$UMAP2) - min(s_curve_noise_umap$UMAP2))) * (y_max - y_min)
@@
 # Compute the horizontal spacing (dx) and vertical spacing (dy) between hexagon centers
-dx <- sqrt(3) * hex_size
-dy <- 1.5 * hex_size
+# dx <- sqrt(3) * hex_size
+# dy <- 1.5 * hex_size
 
-x_length <- 1
-y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))
+# x_length <- 1
+# y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))
 
@@
 ```{r}
-full_grid_centroids_with_hexbin_id <- hex_grid |>
-  dplyr::select("c_x", "c_y", "hexID") |>
-    dplyr::distinct()
-
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
@@
 ```{r}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_centroids_with_hexbin_id)
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, hex_grid)
 ```
@@
 ```{r}
-df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+df_with_std_counts <- compute_std_counts(nldr_df_with_hex_id = s_curve_noise_umap_with_id)
 ```
@@
 ```{r}
-hex_full_count_df <- generate_full_grid_info(hex_grid, df_with_std_counts)
+hex_full_count_df <- generate_full_grid_info(nldr_df = s_curve_noise_umap, hex_grid)
 ```

### 2024-02-22 paper-quollr.tex
Simplify the fuctions to do small tasks

@@
 
-use_python("~/miniforge3/envs/pcamp_env/bin/python")
-use_condaenv("pcamp_env")
-
-reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
-reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
+# use_python("~/miniforge3/envs/pcamp_env/bin/python")
+# use_condaenv("pcamp_env")
+# 
+# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
+# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
 \end{verbatim}
@@
 \begin{verbatim}
-full_grid_centroids_with_hexbin_id <- hex_grid |>
-  dplyr::select("c_x", "c_y", "hexID") |>
-    dplyr::distinct()
-
 ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
@@
 \begin{verbatim}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_centroids_with_hexbin_id)
+s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, hex_grid)
 \end{verbatim}
@@
 \begin{verbatim}
-df_with_std_counts <- compute_std_counts(nldr_df = s_curve_noise_umap_with_id)
+df_with_std_counts <- compute_std_counts(nldr_df_with_hex_id = s_curve_noise_umap_with_id)
 \end{verbatim}
@@
 \begin{verbatim}
-hex_full_count_df <- generate_full_grid_info(hex_grid, df_with_std_counts)
+hex_full_count_df <- generate_full_grid_info(nldr_df = s_curve_noise_umap, hex_grid)
 \end{verbatim}

### 2024-02-22 paper-quollr.tex
Change buffer_x and buffer_y defaults to have one more bin along x and y (need to experiment with some other data sets to set up the default)

@@
 \begin{verbatim}
-#> [1] 4
+#> [1] 5
 \end{verbatim}
@@
 \begin{verbatim}
-#> [1] 7
+#> [1] 8
 \end{verbatim}
@@
 \begin{verbatim}
-#> Rows: 28
+#> Rows: 40
 #> Columns: 2
-#> $ x <dbl> -0.1732051, 0.1732051, 0.5196152, 0.8660254, 0.0000000, 0.3464102, 0~
-#> $ y <dbl> -0.15, -0.15, -0.15, -0.15, 0.15, 0.15, 0.15, 0.15, 0.45, 0.45, 0.45~
+#> $ x <dbl> -0.1732051, 0.1732051, 0.5196152, 0.8660254, 1.2124356, 0.0000000, 0~
+#> $ y <dbl> -0.15, -0.15, -0.15, -0.15, -0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.4~
 \end{verbatim}
@@
 \begin{verbatim}
-#> Rows: 168
+#> Rows: 240
 #> Columns: 6
@@
 #> 1  0.1732051 -0.15     2  0.2352941
-#> 2  0.0000000  0.15     5  0.5294118
-#> 3  0.3464102  0.15     6  0.4117647
-#> 4  0.1732051  0.45    10  0.1764706
-#> 5  0.5196152  0.45    11  0.3529412
-#> 6  0.6928203  0.75    15  0.7058824
-#> 7  0.8660254  1.05    20  0.4705882
-#> 8  0.6928203  1.35    23  0.2941176
-#> 9  1.0392305  1.35    24  0.2352941
-#> 10 0.8660254  1.65    28  1.0000000
+#> 2  0.0000000  0.15     6  0.5294118
+#> 3  0.3464102  0.15     7  0.4117647
+#> 4  0.1732051  0.45    12  0.1764706
+#> 5  0.5196152  0.45    13  0.3529412
+#> 6  0.6928203  0.75    18  0.7058824
+#> 7  0.8660254  1.05    24  0.4705882
+#> 8  0.6928203  1.35    28  0.2941176
+#> 9  1.0392305  1.35    29  0.2352941
+#> 10 0.8660254  1.65    34  1.0000000
 \end{verbatim}
@@
 #> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
-#> $ pred_hb_id  <int> 10, 15, 20, 2, 10, 28, 15, 15, 15, 28, 15, 10, 24, 5, 15, ~
+#> $ pred_hb_id  <int> 12, 18, 24, 2, 12, 34, 18, 18, 18, 34, 18, 12, 29, 6, 18, ~
 \end{verbatim}

### 2024-02-26 paper-quollr.Rmd
Add the hexagonal binning logic with images

@@
 ```
+## Preprocessing
 
-### Compute hexagonal bin configurations
+In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
+To achieve this, we implement two key scaling steps.
+Firstly, we scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval.
+Secondly, we scale the second 2D embedding component to range between $-\frac{\sqrt{3}}{2}$ and $\frac{\sqrt{3}}{2}$.
+This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.
+
+## Compute hexagonal bin configurations
 
@@
 Additionally, choosing an appropriate hexagon size (the radius of the outer circle) is essential.
-By default, the `calculate_effective_x_bins()` and `calculate_effective_y_bins()` functions use a hexagon size of 1.07457.
+By default, the `calculate_effective_x_bins()` and `calculate_effective_y_bins()` functions use a hexagon size of $0.2$.
 However, users can adjust the hexagon size to fit the data range and achieve regular hexagons without overlapping.
@@
 
+```{r hex-size-img, out.width = "100%", out.height = "100%", fig.cap = "Hexagonal size parameter", fig.alt="Hexagonal size parameter", echo=FALSE}
+knitr::include_graphics("figures/hex_size_img.png")
+```
+
+```{r spacing-img, out.width = "100%", out.height = "100%", fig.cap = "Horizontal and vertical spacing parameters", fig.alt="Horizontal and vertical spacing parameters", echo=FALSE}
+knitr::include_graphics("figures/spacing_img.png")
+```
+
+### Number of bins along the x-axis ($b_1$)
+
+In the process of computing hexagonal bin configurations, the first step involves determining the number of bins along the x-axis.
+This begins with the initialization of parameters such as the hexagonal size ($s$) and the buffer along the x-axis.
+Subsequently, the range of the first 2D embedding component ($r_1$) is calculated to ensure comprehensive coverage of the data.
+This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
+The default buffer amount along the x-axis is $\sqrt{3} * s * 1.5$.
+Next, the horizontal spacing ($h$) between hexagons is computed based on the hexagon size.
+Finally, the number of bins along the x-axis is computed based on the adjusted range ($r_1 + buffer_{x}$) and the horizontal spacing ($h$).
+
+$$
+h = \sqrt{3} * s
+$${#eq-equation3}
+
+$$
+ b_1 = \frac{r_1 + buffer_{x}}{h}
+$${#eq-equation4}
+
 ```{r}
@@
 ```
+
+### Number of bins along the y-axis ($b_2$)
+
+In the process of computing hexagonal bin configurations, the second step involves determining the number of bins along the y-axis.
+This begins with the initialization of parameters such as the hexagonal size and the buffer along the y-axis.
+Subsequently, the range of the second 2D embedding component ($r_2$) is calculated to ensure comprehensive coverage of the data.
+This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
+The default buffer amount along the y-axis is $1.5 * s * 1.5$.
+Next, the vertical spacing ($v$) between hexagons is computed based on the hexagon size.
+Finally, the number of bins along the y-axis is computed based on the adjusted range ($r_2 + buffer_{y}$) and the vertical spacing ($v$).
+
+$$
+v = 1.5 * s
+$${#eq-equation5}
+
+$$
+ b_2 = \frac{r_2 + buffer_{y}}{v}
+$${#eq-equation6}
 
@@
 
-Generating full hexagonal grid contains main three steps:
+Generating full hexagonal grid contains main four steps:
 
 1. Generate all the hexagonal bin centroids
+
+The default $x_{start}$ and $y_{start}$ can be defined as:
+
+$$
+x_{start} = min(emb_1) - \frac{\sqrt{3} * s}{2}
+$${#eq-equation7}
+
+$$
+y_{start} = min(emb_2) - \frac{1.5 * s}{2}
+$${#eq-equation8}
+
+```{r gen-fullcentroids, out.width = "100%", out.height = "100%", fig.cap = "Workflow of generating hexagonal bin centroids", fig.alt="Workflow of generating hexagonal bin centroids", echo=FALSE}
+knitr::include_graphics("figures/workflow_all_centroids.png")
+```
 
@@
 2. Generate hexagonal coordinates
+
+$$
+dx = \frac{x_{(2)} - x_{(1)}}{2}
+$${#eq-equation9}
+
+$$
+dy = \frac{2*(y_{(2)} - y_{(1)}) * 1.15}{\sqrt{3}}
+$${#eq-equation10}
+
+```{r hex-coordinates, out.width = "100%", out.height = "100%", fig.cap = "Hexagon coordinates", fig.alt="Hexagon coordinates", echo=FALSE}
+knitr::include_graphics("figures/hex_coordinates_img.png")
+```
+
+```{r gen-allcoordinates, out.width = "100%", out.height = "100%", fig.cap = "Workflow of generating hexagon coordinates", fig.alt="Workflow of generating hexagon coordinates", echo=FALSE}
+knitr::include_graphics("figures/workflow_all_coordinates.png")
+```
 
@@
 
-5. Assign data into hexagons
+3. Assign data into hexagons
 
@@
 
-6. Compute standardized counts
+4. Compute standardized counts within each hexagon
 
@@
 ```
-
-7. Extract full grid info
-
-- Assign standardize counts for hex bins
-
-- Join with the hexagonal coordinates
 
@@
 ```
-
+<!--
 #### Buffer size
@@
 Overall, including a buffer when generating hexagonal bins helps to produce more accurate and robust binning results, particularly when dealing with real-world data that may have irregular distributions or boundary effects.
-
+-->
 ### Construct the 2D model with different options
@@
 ### Compute parameter defaults
-
-#### Shift the hexagonal grid origin
-
-If shift_x happen to the positive direction of x it should input as a positive value, if not other way
-If shift_y happen to the positive direction of y it should input as a positive value, if not other way
-
-1. Assign shift along the x and y axis (limited the amount should less than the cell_diameter)
-
-2. Generate bounds with shift origin
 

### 2024-02-26 paper-quollr.tex
Add the hexagonal binning logic with images

@@
 
+\hypertarget{preprocessing}{%
+\subsection{Preprocessing}\label{preprocessing}}
+
+In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
+To achieve this, we implement two key scaling steps.
+Firstly, we scale the first 2D embedding component to range between \(0\) and \(1\), ensuring that all data points fall within this normalized interval.
+Secondly, we scale the second 2D embedding component to range between \(-\frac{\sqrt{3}}{2}\) and \(\frac{\sqrt{3}}{2}\).
+This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.
+
 \hypertarget{compute-hexagonal-bin-configurations}{%
-\subsubsection{Compute hexagonal bin configurations}\label{compute-hexagonal-bin-configurations}}
+\subsection{Compute hexagonal bin configurations}\label{compute-hexagonal-bin-configurations}}
 
@@
 Additionally, choosing an appropriate hexagon size (the radius of the outer circle) is essential.
-By default, the \texttt{calculate\_effective\_x\_bins()} and \texttt{calculate\_effective\_y\_bins()} functions use a hexagon size of 1.07457.
+By default, the \texttt{calculate\_effective\_x\_bins()} and \texttt{calculate\_effective\_y\_bins()} functions use a hexagon size of \(0.2\).
 However, users can adjust the hexagon size to fit the data range and achieve regular hexagons without overlapping.
@@
 Users should consider their specific data characteristics when selecting the hexagon size.
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=1\textheight]{figures/hex_size_img} \caption{Hexagonal size parameter}\label{fig:hex-size-img}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=1\textheight]{figures/spacing_img} \caption{Horizontal and vertical spacing parameters}\label{fig:spacing-img}
+\end{figure}
+
+\hypertarget{number-of-bins-along-the-x-axis-b_1}{%
+\subsubsection{\texorpdfstring{Number of bins along the x-axis (\(b_1\))}{Number of bins along the x-axis (b\_1)}}\label{number-of-bins-along-the-x-axis-b_1}}
+
+In the process of computing hexagonal bin configurations, the first step involves determining the number of bins along the x-axis.
+This begins with the initialization of parameters such as the hexagonal size (\(s\)) and the buffer along the x-axis.
+Subsequently, the range of the first 2D embedding component (\(r_1\)) is calculated to ensure comprehensive coverage of the data.
+This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
+The default buffer amount along the x-axis is \(\sqrt{3} * s * 1.5\).
+Next, the horizontal spacing (\(h\)) between hexagons is computed based on the hexagon size.
+Finally, the number of bins along the x-axis is computed based on the adjusted range (\(r_1 + buffer_{x}\)) and the horizontal spacing (\(h\)).
+
+\[
+h = \sqrt{3} * s
+\]\{\#eq-equation3\}
+
+\[
+ b_1 = \frac{r_1 + buffer_{x}}{h}
+\]\{\#eq-equation4\}
 
@@
 \end{verbatim}
+
+\hypertarget{number-of-bins-along-the-y-axis-b_2}{%
+\subsubsection{\texorpdfstring{Number of bins along the y-axis (\(b_2\))}{Number of bins along the y-axis (b\_2)}}\label{number-of-bins-along-the-y-axis-b_2}}
+
+In the process of computing hexagonal bin configurations, the second step involves determining the number of bins along the y-axis.
+This begins with the initialization of parameters such as the hexagonal size and the buffer along the y-axis.
+Subsequently, the range of the second 2D embedding component (\(r_2\)) is calculated to ensure comprehensive coverage of the data.
+This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
+The default buffer amount along the y-axis is \(1.5 * s * 1.5\).
+Next, the vertical spacing (\(v\)) between hexagons is computed based on the hexagon size.
+Finally, the number of bins along the y-axis is computed based on the adjusted range (\(r_2 + buffer_{y}\)) and the vertical spacing (\(v\)).
+
+\[
+v = 1.5 * s
+\]\{\#eq-equation5\}
+
+\[
+ b_2 = \frac{r_2 + buffer_{y}}{v}
+\]\{\#eq-equation6\}
 
@@
 
-Generating full hexagonal grid contains main three steps:
+Generating full hexagonal grid contains main four steps:
 
@@
 \end{enumerate}
+
+The default \(x_{start}\) and \(y_{start}\) can be defined as:
+
+\[
+x_{start} = min(emb_1) - \frac{\sqrt{3} * s}{2}
+\]\{\#eq-equation7\}
+
+\[
+y_{start} = min(emb_2) - \frac{1.5 * s}{2}
+\]\{\#eq-equation8\}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=1\textheight]{figures/workflow_all_centroids} \caption{Workflow of generating hexagonal bin centroids}\label{fig:gen-fullcentroids}
+\end{figure}
 
@@
 \end{enumerate}
+
+\[
+dx = \frac{x_{(2)} - x_{(1)}}{2}
+\]\{\#eq-equation9\}
+
+\[
+dy = \frac{2*(y_{(2)} - y_{(1)}) * 1.15}{\sqrt{3}}
+\]\{\#eq-equation10\}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=1\textheight]{figures/hex_coordinates_img} \caption{Hexagon coordinates}\label{fig:hex-coordinates}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=1\textheight]{figures/workflow_all_coordinates} \caption{Workflow of generating hexagon coordinates}\label{fig:gen-allcoordinates}
+\end{figure}
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{4}
+\setcounter{enumi}{2}
 \tightlist
@@
 \def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{5}
+\setcounter{enumi}{3}
 \tightlist
 \item
-  Compute standardized counts
+  Compute standardized counts within each hexagon
 \end{enumerate}
@@
 \end{verbatim}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{6}
-\tightlist
-\item
-  Extract full grid info
-\end{enumerate}
-
-\begin{itemize}
-\item
-  Assign standardize counts for hex bins
-\item
-  Join with the hexagonal coordinates
-\end{itemize}
 
@@
 \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}
-
-\hypertarget{buffer-size}{%
-\paragraph{Buffer size}\label{buffer-size}}
-
-When generating hexagonal bins in R, a buffer is often included to ensure that the data points are evenly distributed within the bins and to prevent edge effects. The buffer helps in two main ways:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \textbf{Preventing Edge Effects}: Without a buffer, the outermost data points might fall near the boundary of the hexagonal grid, leading to incomplete bins or uneven distribution of data.
-By adding a buffer, you create a margin around the outer edges of the grid, ensuring that all data points are fully enclosed within the bins.
-\item
-  \textbf{Ensuring Even Distribution}: The buffer allows for a smoother transition between adjacent bins.
-This helps in cases where data points are not perfectly aligned with the grid lines, ensuring that each data point is assigned to the nearest bin without bias towards any specific direction.
-\end{enumerate}
-
-Overall, including a buffer when generating hexagonal bins helps to produce more accurate and robust binning results, particularly when dealing with real-world data that may have irregular distributions or boundary effects.
-
-\hypertarget{construct-the-2d-model-with-different-options}{%
-\subsubsection{Construct the 2D model with different options}\label{construct-the-2d-model-with-different-options}}
+\#\#\# Construct the 2D model with different options
 
@@
 \subsubsection{Compute parameter defaults}\label{compute-parameter-defaults}}
-
-\hypertarget{shift-the-hexagonal-grid-origin}{%
-\paragraph{Shift the hexagonal grid origin}\label{shift-the-hexagonal-grid-origin}}
-
-If shift\_x happen to the positive direction of x it should input as a positive value, if not other way
-If shift\_y happen to the positive direction of y it should input as a positive value, if not other way
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Assign shift along the x and y axis (limited the amount should less than the cell\_diameter)
-\item
-  Generate bounds with shift origin
-\end{enumerate}
 

### 2024-02-26 paper-quollr.R
Add buffer parameter image

@@ -109,6 +109,10 @@ knitr::include_graphics("figures/hex_size_img.png")
 knitr::include_graphics("figures/spacing_img.png")
 
 
+## ----buffer-img, out.width = "100%", out.height = "100%", fig.cap = "Buffer parameters", fig.alt="Buffer parameters", echo=FALSE----
+knitr::include_graphics("figures/buffer_img.png")
+
+
 ## -----------------------------------------------------------------------------
 num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap,
                                          x = "UMAP1", hex_size = 0.2)

### 2024-03-04 paper-quollr.Rmd
Add buffer parameter image
Added weigted mean cal

--- 2024-02-22 paper-quollr.Rmd
@@
 bibliography: RJreferences.bib
-
 ---
@@
                       message = FALSE)
-source("quollr_code.R", local = TRUE)
-source("nldr_code.R", local = TRUE)
+
 ```
@@
 ```{r load-libraries}
-#library(quollr)
+library(quollr)
+library(tibble)
 library(knitr)
 library(kableExtra)
-library(readr)
 library(ggplot2)
-library(dplyr)
-library(ggbeeswarm)
-library(Rtsne)
-library(umap)
-library(phateR)
-library(reticulate)
-library(rsample)
 
 set.seed(20230531)
-
-# use_python("~/miniforge3/envs/pcamp_env/bin/python")
-# use_condaenv("pcamp_env")
-# 
-# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
-# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
-
-```
-
-```{r}
-#| echo: false
-
-training_data <- read_rds(file = "data/s_curve_noise_training.rds")
-test_data <- read_rds(file = "data/s_curve_noise_test.rds")
-
-s_curve_noise <- read_rds(file = "data/s_curve_noise.rds")
-s_curve_noise_umap <- read_rds(file = "data/s_curve_noise_umap.rds")
-```
-
-```{r}
-library(scales)
-s_curve_noise_umap$UMAP1 <- rescale(s_curve_noise_umap$UMAP1)
-
-y_min <- -sqrt(3)/2
-y_max <- sqrt(3)/2
-
-# y_min <- 0
-# y_max <- sqrt(3)/2
-
-s_curve_noise_umap$UMAP2 <- ((s_curve_noise_umap$UMAP2 - min(s_curve_noise_umap$UMAP2))/(max(s_curve_noise_umap$UMAP2) - min(s_curve_noise_umap$UMAP2))) * (y_max - y_min)
-
-# Define the desired size of the hexagons
-hex_size <- 0.2
-
-# Compute the horizontal spacing (dx) and vertical spacing (dy) between hexagon centers
-# dx <- sqrt(3) * hex_size
-# dy <- 1.5 * hex_size
-
-# x_length <- 1
-# y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))
-
-# Compute the number of bins along the x-axis and y-axis
-# num_bins_x <- ceiling(x_length / dx)
-# num_bins_y <- ceiling(y_length_scaled / dy)
-```
-
-```{r}
-# s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)] <- scale(s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)])
 ```
@@
 -   dependencies
-
-```{r, eval=FALSE}
-library(tools)
-package_dependencies("quollr")
-```
 
@@
                              "s_curve_noise_test", 
-                             "s_curve_noise_umap"), 
+                             "s_curve_noise_umap",
+                             "s_curve_noise_umap_predict",
+                             "s_curve_noise_umap_scaled"), 
                       text = c("Simulated 3D S-curve data with additional four noise dimensions.",
@@
                                "Test data derived from S-curve data.", 
-                               "UMAP 2D embedding data of S-curve data (n_neighbors: 15, min_dist: 0.1)."))
+                               "UMAP 2D embedding data of S-curve data (n_neighbors: 15, min_dist: 0.1).",
+                               "Predicted UMAP 2D embedding data of S-curve data",
+                               "Scaled UMAP 2D embedding data of S-curve data"))
 ```
@@
 
-### Compute hexagonal bin configurations
+## Preprocessing
 
-Hexagonal binning is a powerful technique for visualizing the density of data points in a 2-d space.
-Unlike traditional rectangular bins, hexagonal bins offer several advantages, including a more uniform representation of density and reduced visual bias.
-However, to effectively do the hexagonal binning, it's essential to compute the appropriate configurations based on the characteristics of the dataset.
+In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
+To achieve this, we implement two key scaling steps.
+Firstly, we scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval.
+Secondly, we scale the second 2D embedding component to range between $0$ and $y_{max}$.
+This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.
 
-Before computing hexagonal bin configurations, we need to determine the range of data along the x and y axes to establish the boundary for hexagonal binning.
-Additionally, choosing an appropriate hexagon size (the radius of the outer circle) is essential.
-By default, the `calculate_effective_x_bins()` and `calculate_effective_y_bins()` functions use a hexagon size of 1.07457.
-However, users can adjust the hexagon size to fit the data range and achieve regular hexagons without overlapping.
-
-The hexagon size directly affects the number of bins generated.
-A higher hexagonal size will result in fewer bins, while a lower hexagonal size will lead to more bins.
-Therefore, there's always a trade-off depending on the dataset used.
-Users should consider their specific data characteristics when selecting the hexagon size.
-
-```{r}
-num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap,
-                                         x = "UMAP1", hex_size = 0.2)
-num_bins_x 
+```{=tex}
+\begin{equation}\label{eq:aspect_ratio}
+ar = \frac{r_2}{r_1} \tag{1}
+\end{equation}
+```
+```{=tex}
+\begin{equation}\label{eq:y_max}
+y_{max} = ceiling(\frac{ar}{hr}) * hr \tag{2}
+\end{equation}
 ```
@@
 ```{r}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap,
-                                         y = "UMAP2", hex_size = 0.2)
-num_bins_y 
+scaled_data <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", 
+                y = "UMAP2", hex_ratio = NA)
+glimpse(scaled_data)
 ```
 
-### Generate full hex grid
+## Construct the 2D model
 
-Generating full hexagonal grid contains main three steps:
+### Compute hexagonal bin configurations
 
-1. Generate all the hexagonal bin centroids
+## Construct the high-D model
+
+## Model function
+
+The `fit_highd_model()` function is used to generate both the 2D and high-D models.
 
 ```{r}
-all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
-                                                 x = "UMAP1", y = "UMAP2", 
-                                                 num_bins_x = num_bins_x, 
-                                                 num_bins_y = num_bins_y, 
-                                                 x_start = NA, y_start = NA, 
-                                                 hex_size = 0.2)
-
-glimpse(all_centroids_df)
+fit_highd_model(training_data = s_curve_noise_training, x = "UMAP1", y = "UMAP2",
+nldr_df_with_id = s_curve_noise_umap_scaled, col_start_2d = "UMAP", col_start_highd = "x")
 ```
 
-2. Generate hexagonal coordinates
-
-```{r}
-hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = 0.2)
-glimpse(hex_grid)
-```
-
-```{r}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(aes(x = c_x, y = c_y), color = "red") +
-  coord_fixed()
-```
-
-```{r}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  coord_fixed()
-```
-
-5. Assign data into hexagons
-
-- Compute distances between nldr coordinates and hex bin centroids
-
-- Find the hexagonal centroid that have the minimum distance
-
-```{r}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, hex_grid)
-```
-
-6. Compute standardized counts
-
-- Compute number of data points within each hexagon
-
-- Compute standardise count by dividing the counts by the maximum
-
-```{r}
-df_with_std_counts <- compute_std_counts(nldr_df_with_hex_id = s_curve_noise_umap_with_id)
-```
-
-7. Extract full grid info
-
-- Assign standardize counts for hex bins
-
-- Join with the hexagonal coordinates
-
-```{r}
-hex_full_count_df <- generate_full_grid_info(nldr_df = s_curve_noise_umap, hex_grid)
-```
-
-```{r}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue") +
-  coord_fixed()
-```
-
-```{r}
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-  coord_fixed()
-
-```
-
-#### Buffer size
-
-When generating hexagonal bins in R, a buffer is often included to ensure that the data points are evenly distributed within the bins and to prevent edge effects. The buffer helps in two main ways:
-
-1.
-**Preventing Edge Effects**: Without a buffer, the outermost data points might fall near the boundary of the hexagonal grid, leading to incomplete bins or uneven distribution of data.
-By adding a buffer, you create a margin around the outer edges of the grid, ensuring that all data points are fully enclosed within the bins.
-
-2.
-**Ensuring Even Distribution**: The buffer allows for a smoother transition between adjacent bins.
-This helps in cases where data points are not perfectly aligned with the grid lines, ensuring that each data point is assigned to the nearest bin without bias towards any specific direction.
-
-Overall, including a buffer when generating hexagonal bins helps to produce more accurate and robust binning results, particularly when dealing with real-world data that may have irregular distributions or boundary effects.
-
-### Construct the 2D model with different options
-
-```{r}
-df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
-  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
-  dplyr::distinct() |>
-  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
-
-df_bin_centroids
-```
-
-### Construct the high-D model with different options
-
-```{r}
-## To generate a data set with high-D and 2D training data
-df_all <- training_data |> dplyr::select(-ID) |>
-  dplyr::bind_cols(s_curve_noise_umap_with_id)
-
-## To generate averaged high-D data
-
-df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
-```
-
-### Generate the triangular mesh
-
-```{r}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-```
-
-### Compute parameter defaults
-
-#### Shift the hexagonal grid origin
-
-If shift_x happen to the positive direction of x it should input as a positive value, if not other way
-If shift_y happen to the positive direction of y it should input as a positive value, if not other way
-
-1. Assign shift along the x and y axis (limited the amount should less than the cell_diameter)
-
-2. Generate bounds with shift origin
-
-#### Benchmark value to remove the low-density hexagons
-
-#### Benchmark value to remove the long edges
-
-```{r}
-## Compute 2D distances
-distance <- cal_2d_dist(tr_from_to_df_coord = tr_from_to_df)
-
-## To plot the distribution of distance
-plot_dist <- function(distance_df){
-  distance_df$group <- "1"
-  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
-    geom_quasirandom()+
-    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
-  return(dist_plot)
-}
-
-plot_dist(distance)
-
-benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
-benchmark
-
-benchmark <- 0.75
-```
-
-### Model function
-
-`fit_high_d_model()` function is used to generate the 2D model and the high-D model.
-
-```{r}
-## To generate a data set with high-D and 2D training data
-df_all <- training_data |> dplyr::select(-ID) |>
-  dplyr::bind_cols(s_curve_noise_umap_with_id)
-
-## To generate averaged high-D data
-
-df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
-```
+## Model summaries
 
@@
 
-To predict the 2D embeddings for a new data point using the trained high-D model, we follow a series of steps outlined below:
+There are some of NLDR techniques that don't give any functions for predictions. In that sense our methodology facilitates to compute 2D embedding with our workflow.
 
-1.
-**Compute high-dimensional Euclidean distance**: Calculate the Euclidean distance between each new data point and the lift model's coordinates in the high-dimensional space.
-This distance metric helps identify the nearest lift model coordinates to each new data point.
+### Goodness of fit statistics
 
-2.
-**Find the nearest lift model coordinates**: Determine the lift model's high-dimensional coordinates that are closest to each new data point based on the computed distances.
-This step helps establish a correspondence between the new data points and the lift model's representation in the high-dimensional space.
+There are two Goodness of fit statistics were produced. MSE and AIC which interpret the model accuracy.
 
-3.
-**Map the hexagonal bin ID**: Assign each new data point to a hexagonal bin based on its nearest lift model coordinates.
-This mapping ensures that each data point is associated with a specific region in the hexagonal grid representation of the high-dimensional space.
-
-4.
-**Map the hexagonal bin centroid coordinates in 2D**: Transform the hexagonal bin centroid coordinates from the high-dimensional space to the 2D embedding space using the trained lift model.
-This mapping provides the 2D embeddings for the new data points, allowing them to be visualized and analyzed in a lower-dimensional space.
-
-The `predict_2d_embeddings()` function is used to predict 2D embeddings for a new data point.
-The inputs for the function are test data, bin centroid coordinates in 2D, lifting model coordinates in high-D, and the type of the NLDR technique.
-The output contains the predicted 2D embeddings along with the predicted hexagonal id and the ID of the test data.
-
-```{r}
-pred_df_test <- predict_2d_embeddings(test_data = training_data,
-df_bin_centroids = df_bin_centroids, df_bin = df_bin, type_NLDR = "UMAP")
-
-glimpse(pred_df_test)
-```
-
-### Making summaries
-
-There are two important summaries that should be made when constructing the model on a dataset using a specific NLDR technique, the Mean Square Error (MSE) and Akaike Information Criterion (AIC) which measure prediction accuracy.
-The function that perform the summaries called `generate_eval_df()`.
-The output of this function is a list which contains MSE and AIC.
-The code below uses `generate_eval_df()` to find the MSE and the AIC values for the model constructed on UMAP 2D embeddings generated from the `s_curve_noise` training data.
-
-```{r}
-generate_summary(test_data = training_data, prediction_df = pred_df_test,
-                 df_bin = df_bin, col_start = "x")
-```
-
-### Visualizations
+## Visualizations
 
@@
 
+`r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:lgvis-tb-html)', '\\@ref(tab:lgvis-tb-pdf)'))`
+
+```{r, echo=FALSE}
+lg_vis_tb <- tibble(dt = c(".data", 
+                           "benchmark_value", 
+                           "triangular_object", 
+                           "distance_col"), 
+                    text = c("The data frame containing the edge information.",
+                             "The threshold value to determine long edges.", 
+                             "The triangular object containing the mesh information.", 
+                             "The column name in `.data` representing the distances."))
+```
+
+```{r lgvis-tb-html, eval=is_html_output(), echo=FALSE}
+lg_vis_tb |> 
+  kable(caption = "The main arguments for `vis_lg_mesh()` and `vis_rmlg_mesh()`", col.names = c("argument", "explanation")) 
+```
+
+```{r lgvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
+lg_vis_tb |>
+  kable(caption = "The main arguments for `vis\\_lg\\_mesh()` and `vis\\_rmlg\\_mesh()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
 #### Dynamic visaulizations
@@
 This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data.
-The main arguments are shown in Table
+The main arguments are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:dyvis-tb-html)', '\\@ref(tab:dyvis-tb-pdf)'))`.
 
-```{r}
-trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
-  geom_point(size = 0.1) +
-  geom_trimesh() +
-  coord_equal()
+```{r, echo=FALSE}
+dyn_vis_tb <- tibble(dt = c("df", 
+                            "df_b", 
+                            "df_b_with_center_data", 
+                            "benchmark_value",
+                            "distance_df",
+                            "distance_col",
+                            "use_default_benchmark_val",
+                            "column_start_text"), 
+                     text = c("A data frame containing the high-dimensional data.",
+                              "A data frame containing the high-dimensional coordinates of bin centroids/means.",
+                              "The dataset with hexbin centroids/ means.", 
+                              "The benchmark value used to remove long edges (optional).", 
+                              "The distance dataframe.",
+                              "The name of the distance column.",
+                              "Logical, indicating whether to use default benchmark value  to remove long edges(default is FALSE).",
+                              "The text that begin the column name of the high-dimensional data."))
+```
 
-trimesh
+```{r dyvis-tb-html, eval=is_html_output(), echo=FALSE}
+dyn_vis_tb |> 
+  kable(caption = "The main arguments for `show_langevitour()`", col.names = c("argument", "explanation")) 
+```
+
+```{r dyvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
+dyn_vis_tb |>
+  kable(caption = "The main arguments for `show\\_langevitour()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
+
+## Distance with bin means
+
+```{r, echo=FALSE}
+compute_weights <- function(nldr_df_with_hex_id) {
+
+  ## To get the 2D embeddings average of each bin
+  bin_val_hexagons <- nldr_df_with_hex_id |>
+    dplyr::group_by(hb_id) |>
+    dplyr::summarise(dplyr::across(tidyselect::everything(), mean))
+
+  ## Rename columns of averaged 2D embeddings
+  names(bin_val_hexagons) <- c("hb_id", paste0("avg_", tolower(names(nldr_df_with_hex_id)[1:2])))
+
+  ## To calculate distances from average point
+  nldr_with_avg_all <- dplyr::inner_join(bin_val_hexagons , nldr_df_with_hex_id,
+                                         by = c("hb_id" = "hb_id"))
+
+  col_names <- c(names(bin_val_hexagons), names(nldr_df_with_hex_id)[1:2], "distance")
+
+  ## Initialize the vectors to store data
+  hexids <- integer(0)
+  emb1_vec <- numeric(0)
+  emb2_vec <- numeric(0)
+  weight_vec <- numeric(0)
+
+  for(hb_id in unique(nldr_with_avg_all$hb_id)){
+
+    ## These are the weights for weighted mean
+    weighted_mean_df <- nldr_with_avg_all |>
+      dplyr::filter(hb_id == hb_id) |>
+      cal_2d_dist(start_x = col_names[2], start_y = col_names[3], end_x = col_names[4],
+                  end_y = col_names[5], select_vars = col_names)
+
+    hexids <- c(hexids, weighted_mean_df$hb_id)
+    emb1_vec <- c(emb1_vec, weighted_mean_df[[rlang::as_string(rlang::sym(col_names[4]))]])
+    emb2_vec <- c(emb2_vec, weighted_mean_df[[rlang::as_string(rlang::sym(col_names[5]))]])
+    weight_vec <- c(weight_vec, 1/ (weighted_mean_df$distance + 0.05))
+
+  }
+
+  weight_list <- list(hb_id = hexids, emb1_vec = emb1_vec, emb2_vec = emb2_vec,
+                      weight_vec = weight_vec)
+  names(weight_list) <- c("hb_id", names(nldr_df_with_hex_id)[1:2], "weights")
+
+  return(weight_list)
+
+}
+```
+
+```{r, echo=FALSE}
+weighted_highD_data <- function(training_data, nldr_df_with_hex_id,
+                                column_start_text = "x") {
+  ## Remove ID column from training data
+  training_data <- training_data |>
+    dplyr::select(-ID)
+
+  ## Join training data with 2D embeddings
+  df_all <- dplyr::bind_cols(training_data, nldr_df_with_hex_id)
+
+  ## To obtain weights corresponding to 2D embeddings
+  weight_df <- as.data.frame(do.call(cbind,
+                                     compute_weights(nldr_df_with_hex_id = nldr_df_with_hex_id)))
+
+  ## Initialize the columns that need to use for joining
+  joined_col_names <- names(nldr_df_with_hex_id)[1:2]
+
+  ## To join the training data, 2D embeddings and weights
+  weighted_mean_all <- dplyr::inner_join(df_all, weight_df,
+                                         by = c("hb_id" = "hb_id",
+                                                stats::setNames(joined_col_names,
+                                                                joined_col_names)))
+  ## List to store weighted means
+  weighted_mean_df_list <- list()
+
+  for (j in 1:NCOL(training_data)) {
+
+    ## To compute weighted mean across all high-D coordinates
+    weighted_mean_df_list[[j]] <- weighted_mean_all |>
+      dplyr::select(hb_id, names(training_data)[j], weights) |>
+      dplyr::group_by(hb_id) |>
+      dplyr::summarise(dplyr::across(names(training_data)[j], ~ weighted.mean(., weights)))
+
+  }
+
+  ## To combine the elements given in the list
+  weighted_mean <- Reduce(function(dtf1,dtf2) dplyr::full_join(dtf1,dtf2,by="hb_id"),
+                          weighted_mean_df_list)
+
+  ## Column names start with x
+  weighted_mean <- weighted_mean |>
+    dplyr::select(hb_id, tidyselect::starts_with(column_start_text))
+
+  return(weighted_mean)
+}
+```
+
+```{r, echo=FALSE}
+
+extract_hexbin_mean <- function(nldr_df_with_hex_id, counts_df) {
+
+  ## To arrange the hexagon IDs
+  counts_df <- counts_df |>
+    dplyr::arrange(hb_id)
+
+  ## To compute hexagonal bin means
+  hex_mean_df <- nldr_df_with_hex_id |>
+    dplyr::group_by(hb_id) |>
+    dplyr::summarise(dplyr::across(tidyselect::everything(), mean)) |>
+    dplyr::arrange(hb_id) |>
+    dplyr::filter(hb_id %in% counts_df$hb_id) |>
+    dplyr::mutate(std_counts = counts_df$std_counts)
+
+  ## Rename columns
+  names(hex_mean_df) <- c("hexID", "c_x", "c_y", "std_counts")
+
+  return(hex_mean_df)
+}
 ```
@@
 ```{r}
-trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
+bin_list <- calc_bins(data = s_curve_noise_umap_scaled, 
+                      x = "UMAP1", y = "UMAP2", 
+                      hex_size = NA, buffer_x = NA, 
+                      buffer_y = NA)
+num_bins_x <- bin_list$num_x
+num_bins_y <- bin_list$num_y
 
-trimesh_gr
+hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, 
+                      x = "UMAP1", y = "UMAP2", 
+                      num_bins_x = num_bins_x, num_bins_y = num_bins_y, 
+                      x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, 
+                      hex_size = NA, col_start = "UMAP")
+
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+nldr_df_with_hex_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))
+
+## To obtain bin centroids
+df_bin_centroids <- extract_hexbin_mean(nldr_df_with_hex_id = nldr_df_with_hex_id,
+                                             counts_df = counts_df)
+
+df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), nldr_df_with_hex_id)
+
+ggplot() + 
+  geom_trimesh(data = df_bin_centroids, mapping = aes(x = c_x, y = c_y)) +
+  coord_fixed()
+
+tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
+tr_from_to_df <- gen_edges(tri_object = tr1_object)
+distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
+                           start_x = "x_from", start_y = "y_from", 
+                           end_x = "x_to", end_y = "y_to", 
+                           select_vars = c("from", "to", "distance"))
+
+## averaged high-D data
+df_bin <- weighted_highD_data(training_data = s_curve_noise_training, nldr_df_with_hex_id = nldr_df_with_hex_id, column_start_text = "x")
+
+vis_lg_mesh(distance_edges = distance_df, benchmark_value = 1,
+tr_coord_df = tr_from_to_df, distance_col = "distance")
+
+vis_rmlg_mesh(distance_edges = distance_df, benchmark_value = 1,
+tr_coord_df = tr_from_to_df, distance_col = "distance")
+
+show_langevitour(df = df_all, df_b = df_bin, 
+                 df_b_with_center_data = df_bin_centroids, 
+                 benchmark_value = 1, distance = distance_df, 
+                 distance_col = "distance", 
+                 use_default_benchmark_val = FALSE, col_start = "x")
 ```
-
-```{r}
-trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
-trimesh_removed
-```
-
-```{r}
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, 
-                          benchmark_value = benchmark,
-                          distance = distance, distance_col = "distance", 
-                          use_default_benchmark_val = FALSE, 
-                          column_start_text = "x")
-tour1
-```
-
-### Find non-empty bins
 
@@
 This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template.
-The source code for reproducing this paper can be found at: [https://github.com/JayaniLakshika/paper-quollr](https://github.com/JayaniLakshika/paper-quollr).
+The source code for reproducing this paper can be found at: <https://github.com/JayaniLakshika/paper-quollr>.

### 2024-03-04 paper-quollr.tex
Add buffer parameter image
Added weigted mean cal

--- 2024-02-22 paper-quollr.tex
@@
 \begin{verbatim}
-#library(quollr)
+library(quollr)
+library(tibble)
 library(knitr)
 library(kableExtra)
-library(readr)
 library(ggplot2)
-library(dplyr)
-library(ggbeeswarm)
-library(Rtsne)
-library(umap)
-library(phateR)
-library(reticulate)
-library(rsample)
 
 set.seed(20230531)
-
-# use_python("~/miniforge3/envs/pcamp_env/bin/python")
-# use_condaenv("pcamp_env")
-# 
-# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
-# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))
-\end{verbatim}
-
-\begin{verbatim}
-library(scales)
-s_curve_noise_umap$UMAP1 <- rescale(s_curve_noise_umap$UMAP1)
-
-y_min <- -sqrt(3)/2
-y_max <- sqrt(3)/2
-
-# y_min <- 0
-# y_max <- sqrt(3)/2
-
-s_curve_noise_umap$UMAP2 <- ((s_curve_noise_umap$UMAP2 - min(s_curve_noise_umap$UMAP2))/(max(s_curve_noise_umap$UMAP2) - min(s_curve_noise_umap$UMAP2))) * (y_max - y_min)
-
-# Define the desired size of the hexagons
-hex_size <- 0.2
-
-# Compute the horizontal spacing (dx) and vertical spacing (dy) between hexagon centers
-# dx <- sqrt(3) * hex_size
-# dy <- 1.5 * hex_size
-
-# x_length <- 1
-# y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))
-
-# Compute the number of bins along the x-axis and y-axis
-# num_bins_x <- ceiling(x_length / dx)
-# num_bins_y <- ceiling(y_length_scaled / dy)
-\end{verbatim}
-
-\begin{verbatim}
-# s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)] <- scale(s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)])
 \end{verbatim}
@@
 \begin{itemize}
-\tightlist
 \item
   dependencies
-\end{itemize}
-
-\begin{verbatim}
-library(tools)
-package_dependencies("quollr")
-\end{verbatim}
-
-\begin{itemize}
-\tightlist
 \item
@@
 s\_curve\_noise\_umap & UMAP 2D embedding data of S-curve data (n\_neighbors: 15, min\_dist: 0.1).\\
+s\_curve\_noise\_umap\_predict & Predicted UMAP 2D embedding data of S-curve data\\
+\addlinespace
+s\_curve\_noise\_umap\_scaled & Scaled UMAP 2D embedding data of S-curve data\\
 \bottomrule
@@
 
+\hypertarget{preprocessing}{%
+\subsection{Preprocessing}\label{preprocessing}}
+
+In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
+To achieve this, we implement two key scaling steps.
+Firstly, we scale the first 2D embedding component to range between \(0\) and \(1\), ensuring that all data points fall within this normalized interval.
+Secondly, we scale the second 2D embedding component to range between \(0\) and \(y_{max}\).
+This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.
+
+\begin{equation}\label{eq:aspect_ratio}
+ar = \frac{r_2}{r_1} \tag{1}
+\end{equation}
+
+\begin{equation}\label{eq:y_max}
+y_{max} = ceiling(\frac{ar}{hr}) * hr \tag{2}
+\end{equation}
+
+\begin{verbatim}
+scaled_data <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", 
+                y = "UMAP2", hex_ratio = NA)
+glimpse(scaled_data)
+\end{verbatim}
+
+\begin{verbatim}
+#> List of 2
+#>  $ scaled_UMAP1: num [1:75] 0.0804 0.7386 0.8399 0.1672 0.2629 ...
+#>  $ scaled_UMAP2: num [1:75] 0.366 1.1464 1.2392 0.0494 0.4556 ...
+\end{verbatim}
+
+\hypertarget{construct-the-2d-model}{%
+\subsection{Construct the 2D model}\label{construct-the-2d-model}}
+
 \hypertarget{compute-hexagonal-bin-configurations}{%
@@
 
-Hexagonal binning is a powerful technique for visualizing the density of data points in a 2-d space.
-Unlike traditional rectangular bins, hexagonal bins offer several advantages, including a more uniform representation of density and reduced visual bias.
-However, to effectively do the hexagonal binning, it's essential to compute the appropriate configurations based on the characteristics of the dataset.
+\hypertarget{construct-the-high-d-model}{%
+\subsection{Construct the high-D model}\label{construct-the-high-d-model}}
 
-Before computing hexagonal bin configurations, we need to determine the range of data along the x and y axes to establish the boundary for hexagonal binning.
-Additionally, choosing an appropriate hexagon size (the radius of the outer circle) is essential.
-By default, the \texttt{calculate\_effective\_x\_bins()} and \texttt{calculate\_effective\_y\_bins()} functions use a hexagon size of 1.07457.
-However, users can adjust the hexagon size to fit the data range and achieve regular hexagons without overlapping.
+\hypertarget{model-function}{%
+\subsection{Model function}\label{model-function}}
 
-The hexagon size directly affects the number of bins generated.
-A higher hexagonal size will result in fewer bins, while a lower hexagonal size will lead to more bins.
-Therefore, there's always a trade-off depending on the dataset used.
-Users should consider their specific data characteristics when selecting the hexagon size.
+The \texttt{fit\_highd\_model()} function is used to generate both the 2D and high-D models.
 
 \begin{verbatim}
-num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap,
-                                         x = "UMAP1", hex_size = 0.2)
-num_bins_x 
+fit_highd_model(training_data = s_curve_noise_training, x = "UMAP1", y = "UMAP2",
+nldr_df_with_id = s_curve_noise_umap_scaled, col_start_2d = "UMAP", col_start_highd = "x")
 \end{verbatim}
@@
 \begin{verbatim}
-#> [1] 5
+#> $df_bin
+#> # A tibble: 10 x 8
+#>    hb_id       x1     x2     x3        x4        x5       x6        x7
+#>    <dbl>    <dbl>  <dbl>  <dbl>     <dbl>     <dbl>    <dbl>     <dbl>
+#>  1     2 -0.637   1.74   -1.76   0.00953  -0.00143  -0.0117  -0.00152 
+#>  2     6 -0.498   0.524  -1.73  -0.000237  0.00234  -0.0297  -0.000242
+#>  3     7  0.294   1.40   -1.88   0.00890  -0.00803  -0.0123  -0.00120 
+#>  4    12  0.309   0.0421 -1.83   0.00656   0.00823   0.00489 -0.00389 
+#>  5    13  0.868   0.747  -0.781 -0.00408   0.000857  0.0248   0.00170 
+#>  6    18  0.357   1.27   -0.169  0.00607   0.00124   0.0152   0.00204 
+#>  7    24 -0.792   1.25    0.514 -0.000777  0.000464  0.00602  0.000371
+#>  8    28 -0.597   1.19    1.77   0.000240 -0.00417  -0.0185  -0.000786
+#>  9    29 -0.00544 0.211   1.92   0.00116   0.00266   0.00949 -0.00209 
+#> 10    34  0.622   1.21    1.64  -0.000560  0.00540  -0.00741 -0.000886
+#> 
+#> $df_bin_centroids
+#>    hexID       c_x   c_y std_counts
+#> 1      2 0.1732051 -0.15  0.2352941
+#> 2      6 0.0000000  0.15  0.5294118
+#> 3      7 0.3464102  0.15  0.4117647
+#> 4     12 0.1732051  0.45  0.1764706
+#> 5     13 0.5196152  0.45  0.3529412
+#> 6     18 0.6928203  0.75  0.7058824
+#> 7     24 0.8660254  1.05  0.4705882
+#> 8     28 0.6928203  1.35  0.2941176
+#> 9     29 1.0392305  1.35  0.2352941
+#> 10    34 0.8660254  1.65  1.0000000
 \end{verbatim}
 
-\begin{verbatim}
-num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap,
-                                         y = "UMAP2", hex_size = 0.2)
-num_bins_y 
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 8
-\end{verbatim}
-
-\hypertarget{generate-full-hex-grid}{%
-\subsubsection{Generate full hex grid}\label{generate-full-hex-grid}}
-
-Generating full hexagonal grid contains main three steps:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Generate all the hexagonal bin centroids
-\end{enumerate}
-
-\begin{verbatim}
-all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
-                                                 x = "UMAP1", y = "UMAP2", 
-                                                 num_bins_x = num_bins_x, 
-                                                 num_bins_y = num_bins_y, 
-                                                 x_start = NA, y_start = NA, 
-                                                 hex_size = 0.2)
-
-glimpse(all_centroids_df)
-\end{verbatim}
-
-\begin{verbatim}
-#> Rows: 40
-#> Columns: 2
-#> $ x <dbl> -0.1732051, 0.1732051, 0.5196152, 0.8660254, 1.2124356, 0.0000000, 0~
-#> $ y <dbl> -0.15, -0.15, -0.15, -0.15, -0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.4~
-\end{verbatim}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  Generate hexagonal coordinates
-\end{enumerate}
-
-\begin{verbatim}
-hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = 0.2)
-glimpse(hex_grid)
-\end{verbatim}
-
-\begin{verbatim}
-#> Rows: 240
-#> Columns: 6
-#> $ c_x   <dbl> -0.1732051, -0.1732051, -0.1732051, -0.1732051, -0.1732051, -0.1~
-#> $ c_y   <dbl> -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -0.15, -~
-#> $ x     <dbl> 0.0000000, 0.0000000, -0.1732051, -0.3464102, -0.3464102, -0.173~
-#> $ y     <dbl> -0.05040708, -0.24959292, -0.34918584, -0.24959292, -0.05040708,~
-#> $ id    <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4~
-#> $ hexID <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4~
-\end{verbatim}
-
-\begin{verbatim}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(aes(x = c_x, y = c_y), color = "red") +
-  coord_fixed()
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-10-1.pdf}
-
-\begin{verbatim}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  coord_fixed()
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{4}
-\tightlist
-\item
-  Assign data into hexagons
-\end{enumerate}
-
-\begin{itemize}
-\item
-  Compute distances between nldr coordinates and hex bin centroids
-\item
-  Find the hexagonal centroid that have the minimum distance
-\end{itemize}
-
-\begin{verbatim}
-s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, hex_grid)
-\end{verbatim}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{5}
-\tightlist
-\item
-  Compute standardized counts
-\end{enumerate}
-
-\begin{itemize}
-\item
-  Compute number of data points within each hexagon
-\item
-  Compute standardise count by dividing the counts by the maximum
-\end{itemize}
-
-\begin{verbatim}
-df_with_std_counts <- compute_std_counts(nldr_df_with_hex_id = s_curve_noise_umap_with_id)
-\end{verbatim}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{6}
-\tightlist
-\item
-  Extract full grid info
-\end{enumerate}
-
-\begin{itemize}
-\item
-  Assign standardize counts for hex bins
-\item
-  Join with the hexagonal coordinates
-\end{itemize}
-
-\begin{verbatim}
-hex_full_count_df <- generate_full_grid_info(nldr_df = s_curve_noise_umap, hex_grid)
-\end{verbatim}
-
-\begin{verbatim}
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue") +
-  coord_fixed()
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-15-1.pdf}
-
-\begin{verbatim}
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-  coord_fixed()
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-16-1.pdf}
-
-\hypertarget{buffer-size}{%
-\paragraph{Buffer size}\label{buffer-size}}
-
-When generating hexagonal bins in R, a buffer is often included to ensure that the data points are evenly distributed within the bins and to prevent edge effects. The buffer helps in two main ways:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \textbf{Preventing Edge Effects}: Without a buffer, the outermost data points might fall near the boundary of the hexagonal grid, leading to incomplete bins or uneven distribution of data.
-By adding a buffer, you create a margin around the outer edges of the grid, ensuring that all data points are fully enclosed within the bins.
-\item
-  \textbf{Ensuring Even Distribution}: The buffer allows for a smoother transition between adjacent bins.
-This helps in cases where data points are not perfectly aligned with the grid lines, ensuring that each data point is assigned to the nearest bin without bias towards any specific direction.
-\end{enumerate}
-
-Overall, including a buffer when generating hexagonal bins helps to produce more accurate and robust binning results, particularly when dealing with real-world data that may have irregular distributions or boundary effects.
-
-\hypertarget{construct-the-2d-model-with-different-options}{%
-\subsubsection{Construct the 2D model with different options}\label{construct-the-2d-model-with-different-options}}
-
-\begin{verbatim}
-df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
-  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
-  dplyr::distinct() |>
-  dplyr::rename(c("x" = "c_x", "y" = "c_y"))
-
-df_bin_centroids
-\end{verbatim}
-
-\begin{verbatim}
-#>            x     y hexID std_counts
-#> 1  0.1732051 -0.15     2  0.2352941
-#> 2  0.0000000  0.15     6  0.5294118
-#> 3  0.3464102  0.15     7  0.4117647
-#> 4  0.1732051  0.45    12  0.1764706
-#> 5  0.5196152  0.45    13  0.3529412
-#> 6  0.6928203  0.75    18  0.7058824
-#> 7  0.8660254  1.05    24  0.4705882
-#> 8  0.6928203  1.35    28  0.2941176
-#> 9  1.0392305  1.35    29  0.2352941
-#> 10 0.8660254  1.65    34  1.0000000
-\end{verbatim}
-
-\hypertarget{construct-the-high-d-model-with-different-options}{%
-\subsubsection{Construct the high-D model with different options}\label{construct-the-high-d-model-with-different-options}}
-
-\begin{verbatim}
-## To generate a data set with high-D and 2D training data
-df_all <- training_data |> dplyr::select(-ID) |>
-  dplyr::bind_cols(s_curve_noise_umap_with_id)
-
-## To generate averaged high-D data
-
-df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
-\end{verbatim}
-
-\hypertarget{generate-the-triangular-mesh}{%
-\subsubsection{Generate the triangular mesh}\label{generate-the-triangular-mesh}}
-
-\begin{verbatim}
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-\end{verbatim}
-
-\hypertarget{compute-parameter-defaults}{%
-\subsubsection{Compute parameter defaults}\label{compute-parameter-defaults}}
-
-\hypertarget{shift-the-hexagonal-grid-origin}{%
-\paragraph{Shift the hexagonal grid origin}\label{shift-the-hexagonal-grid-origin}}
-
-If shift\_x happen to the positive direction of x it should input as a positive value, if not other way
-If shift\_y happen to the positive direction of y it should input as a positive value, if not other way
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Assign shift along the x and y axis (limited the amount should less than the cell\_diameter)
-\item
-  Generate bounds with shift origin
-\end{enumerate}
-
-\hypertarget{benchmark-value-to-remove-the-low-density-hexagons}{%
-\paragraph{Benchmark value to remove the low-density hexagons}\label{benchmark-value-to-remove-the-low-density-hexagons}}
-
-\hypertarget{benchmark-value-to-remove-the-long-edges}{%
-\paragraph{Benchmark value to remove the long edges}\label{benchmark-value-to-remove-the-long-edges}}
-
-\begin{verbatim}
-## Compute 2D distances
-distance <- cal_2d_dist(tr_from_to_df_coord = tr_from_to_df)
-
-## To plot the distribution of distance
-plot_dist <- function(distance_df){
-  distance_df$group <- "1"
-  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
-    geom_quasirandom()+
-    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
-  return(dist_plot)
-}
-
-plot_dist(distance)
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-20-1.pdf}
-
-\begin{verbatim}
-benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
-benchmark
-\end{verbatim}
-
-\begin{verbatim}
-#> [1] 0.6
-\end{verbatim}
-
-\begin{verbatim}
-benchmark <- 0.75
-\end{verbatim}
-
-\hypertarget{model-function}{%
-\subsubsection{Model function}\label{model-function}}
-
-\texttt{fit\_high\_d\_model()} function is used to generate the 2D model and the high-D model.
-
-\begin{verbatim}
-## To generate a data set with high-D and 2D training data
-df_all <- training_data |> dplyr::select(-ID) |>
-  dplyr::bind_cols(s_curve_noise_umap_with_id)
-
-## To generate averaged high-D data
-
-df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
-\end{verbatim}
+\hypertarget{model-summaries}{%
+\subsection{Model summaries}\label{model-summaries}}
 
@@
 
-To predict the 2D embeddings for a new data point using the trained high-D model, we follow a series of steps outlined below:
+There are some of NLDR techniques that don't give any functions for predictions. In that sense our methodology facilitates to compute 2D embedding with our workflow.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \textbf{Compute high-dimensional Euclidean distance}: Calculate the Euclidean distance between each new data point and the lift model's coordinates in the high-dimensional space.
-This distance metric helps identify the nearest lift model coordinates to each new data point.
-\item
-  \textbf{Find the nearest lift model coordinates}: Determine the lift model's high-dimensional coordinates that are closest to each new data point based on the computed distances.
-This step helps establish a correspondence between the new data points and the lift model's representation in the high-dimensional space.
-\item
-  \textbf{Map the hexagonal bin ID}: Assign each new data point to a hexagonal bin based on its nearest lift model coordinates.
-This mapping ensures that each data point is associated with a specific region in the hexagonal grid representation of the high-dimensional space.
-\item
-  \textbf{Map the hexagonal bin centroid coordinates in 2D}: Transform the hexagonal bin centroid coordinates from the high-dimensional space to the 2D embedding space using the trained lift model.
-This mapping provides the 2D embeddings for the new data points, allowing them to be visualized and analyzed in a lower-dimensional space.
-\end{enumerate}
+\hypertarget{goodness-of-fit-statistics}{%
+\subsubsection{Goodness of fit statistics}\label{goodness-of-fit-statistics}}
 
-The \texttt{predict\_2d\_embeddings()} function is used to predict 2D embeddings for a new data point.
-The inputs for the function are test data, bin centroid coordinates in 2D, lifting model coordinates in high-D, and the type of the NLDR technique.
-The output contains the predicted 2D embeddings along with the predicted hexagonal id and the ID of the test data.
-
-\begin{verbatim}
-pred_df_test <- predict_2d_embeddings(test_data = training_data,
-df_bin_centroids = df_bin_centroids, df_bin = df_bin, type_NLDR = "UMAP")
-
-glimpse(pred_df_test)
-\end{verbatim}
-
-\begin{verbatim}
-#> Rows: 75
-#> Columns: 4
-#> $ pred_UMAP_1 <dbl> 0.1732051, 0.6928203, 0.8660254, 0.1732051, 0.1732051, 0.8~
-#> $ pred_UMAP_2 <dbl> 0.45, 0.75, 1.05, -0.15, 0.45, 1.65, 0.75, 0.75, 0.75, 1.6~
-#> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
-#> $ pred_hb_id  <int> 12, 18, 24, 2, 12, 34, 18, 18, 18, 34, 18, 12, 29, 6, 18, ~
-\end{verbatim}
-
-\hypertarget{making-summaries}{%
-\subsubsection{Making summaries}\label{making-summaries}}
-
-There are two important summaries that should be made when constructing the model on a dataset using a specific NLDR technique, the Mean Square Error (MSE) and Akaike Information Criterion (AIC) which measure prediction accuracy.
-The function that perform the summaries called \texttt{generate\_eval\_df()}.
-The output of this function is a list which contains MSE and AIC.
-The code below uses \texttt{generate\_eval\_df()} to find the MSE and the AIC values for the model constructed on UMAP 2D embeddings generated from the \texttt{s\_curve\_noise} training data.
-
-\begin{verbatim}
-generate_summary(test_data = training_data, prediction_df = pred_df_test,
-                 df_bin = df_bin, col_start = "x")
-\end{verbatim}
-
-\begin{verbatim}
-#> $mse
-#> [1] 0.3229922
-#> 
-#> $aic
-#> [1] -453.3167
-\end{verbatim}
+There are two Goodness of fit statistics were produced. MSE and AIC which interpret the model accuracy.
 
 \hypertarget{visualizations}{%
-\subsubsection{Visualizations}\label{visualizations}}
+\subsection{Visualizations}\label{visualizations}}
 
@@
 
+\ref{tab:lgvis-tb-pdf}
+
+\begin{table}
+
+\caption{\label{tab:lgvis-tb-pdf}The main arguments for `vis\_lg\_mesh()` and `vis\_rmlg\_mesh()`}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+argument & explanation\\
+\midrule
+.data & The data frame containing the edge information.\\
+benchmark\_value & The threshold value to determine long edges.\\
+triangular\_object & The triangular object containing the mesh information.\\
+distance\_col & The column name in `.data` representing the distances.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \hypertarget{dynamic-visaulizations}{%
@@
 This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data.
-The main arguments are shown in Table
+The main arguments are shown in Table \ref{tab:dyvis-tb-pdf}.
+
+\begin{table}
+
+\caption{\label{tab:dyvis-tb-pdf}The main arguments for `show\_langevitour()`}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+argument & explanation\\
+\midrule
+df & A data frame containing the high-dimensional data.\\
+df\_b & A data frame containing the high-dimensional coordinates of bin centroids/means.\\
+df\_b\_with\_center\_data & The dataset with hexbin centroids/ means.\\
+benchmark\_value & The benchmark value used to remove long edges (optional).\\
+distance\_df & The distance dataframe.\\
+\addlinespace
+distance\_col & The name of the distance column.\\
+use\_default\_benchmark\_val & Logical, indicating whether to use default benchmark value  to remove long edges(default is FALSE).\\
+column\_start\_text & The text that begin the column name of the high-dimensional data.\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{distance-with-bin-means}{%
+\subsection{Distance with bin means}\label{distance-with-bin-means}}
 
 \begin{verbatim}
-trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
-  geom_point(size = 0.1) +
-  geom_trimesh() +
-  coord_equal()
+bin_list <- calc_bins(data = s_curve_noise_umap_scaled, 
+                      x = "UMAP1", y = "UMAP2", 
+                      hex_size = NA, buffer_x = NA, 
+                      buffer_y = NA)
+num_bins_x <- bin_list$num_x
+num_bins_y <- bin_list$num_y
 
-trimesh
+hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, 
+                      x = "UMAP1", y = "UMAP2", 
+                      num_bins_x = num_bins_x, num_bins_y = num_bins_y, 
+                      x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, 
+                      hex_size = NA, col_start = "UMAP")
+
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+nldr_df_with_hex_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))
+
+## To obtain bin centroids
+df_bin_centroids <- extract_hexbin_mean(nldr_df_with_hex_id = nldr_df_with_hex_id,
+                                             counts_df = counts_df)
+
+df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), nldr_df_with_hex_id)
+
+ggplot() + 
+  geom_trimesh(data = df_bin_centroids, mapping = aes(x = c_x, y = c_y)) +
+  coord_fixed()
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf}
 
 \begin{verbatim}
-trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
+tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
+tr_from_to_df <- gen_edges(tri_object = tr1_object)
+distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
+                           start_x = "x_from", start_y = "y_from", 
+                           end_x = "x_to", end_y = "y_to", 
+                           select_vars = c("from", "to", "distance"))
 
-trimesh_gr
+## averaged high-D data
+df_bin <- weighted_highD_data(training_data = s_curve_noise_training, nldr_df_with_hex_id = nldr_df_with_hex_id, column_start_text = "x")
+
+vis_lg_mesh(distance_edges = distance_df, benchmark_value = 1,
+tr_coord_df = tr_from_to_df, distance_col = "distance")
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-25-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-2.pdf}
 
 \begin{verbatim}
-trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
-                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
-trimesh_removed
+vis_rmlg_mesh(distance_edges = distance_df, benchmark_value = 1,
+tr_coord_df = tr_from_to_df, distance_col = "distance")
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-26-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-3.pdf}
 
 \begin{verbatim}
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, 
-                          benchmark_value = benchmark,
-                          distance = distance, distance_col = "distance", 
-                          use_default_benchmark_val = FALSE, 
-                          column_start_text = "x")
-tour1
+show_langevitour(df = df_all, df_b = df_bin, 
+                 df_b_with_center_data = df_bin_centroids, 
+                 benchmark_value = 1, distance = distance_df, 
+                 distance_col = "distance", 
+                 use_default_benchmark_val = FALSE, col_start = "x")
 \end{verbatim}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-27-1.pdf}
-
-\hypertarget{find-non-empty-bins}{%
-\subsubsection{Find non-empty bins}\label{find-non-empty-bins}}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-4.pdf}
 

### 2024-03-30 paper-quollr.Rmd
Added main structure of the quollr paper

@@
 ---
-title: "quollr: An R Package for Visalizing 2D Models in High Dimensional Space"
+title: "quollr: An R Package for Visualizing 2D Models from Nonlinear Dimension Reductions in High Dimensional Space"
 abstract: >
@@
 
+<!--intro with S-curve with 5 methods-->
+
 # Methodology
 
+<!--explain algorithm steps in summary-->
+
+# Implementation
+
+The package can be installed from CRAN:
+
+```install.packages("quollr")```
+
+The development version can be installed from GitHub:
+
+```devtools::install_github("JayaniLakshika/quollr")```
+
+The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `s_curve_noise_training` and `s_curve_noise_umap`.
+
+<!-- add about main function to gen 2D and highD models-->
+## Constructing the 2D model
+
+<!--Discuss the model can be generated with bin centroids or bin means-->
+
+### Preprocessing
+
+<!--Discuss scaling the NLDR data-->
+
+### Binning data
+
+### Obtain bin centroids/ means
+
+### Triangulate bin centroids/ means
+
+<!-- add about to do hex binning at once-->
+
+## Lifting the model into high dimensions
+
+## Tuning model parameters
+
+## Compute MSE
+
+## Visualizations
+
+### 2D model visualization
+
+<!--2D model vis: triangular mesh-->
+
+### p-D model visualization
+
+<!--highD model vis: langevitour with the model-->
+
+<!--
 ## Usage
@@
 
--   basic example
+-   basic example-->
 
-### Datasets
-
-The `quollr` package comes with several data sets that load with the package.
-These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.
-
-```{r, echo=FALSE}
-datasets_tb <- tibble(dt = c("s_curve_noise",
-                             "s_curve_noise_training", 
-                             "s_curve_noise_test", 
-                             "s_curve_noise_umap",
-                             "s_curve_noise_umap_predict",
-                             "s_curve_noise_umap_scaled"), 
-                      text = c("Simulated 3D S-curve data with additional four noise dimensions.",
-                               "Training data derived from S-curve data.", 
-                               "Test data derived from S-curve data.", 
-                               "UMAP 2D embedding data of S-curve data (n_neighbors: 15, min_dist: 0.1).",
-                               "Predicted UMAP 2D embedding data of S-curve data",
-                               "Scaled UMAP 2D embedding data of S-curve data"))
-```
-
-```{r datasets-tb-html, eval=is_html_output(), echo=FALSE}
-datasets_tb |> 
-  kable(caption = "quollr datasets", col.names = c("data", "explanation")) 
-```
-
-```{r datasets-tb-pdf, eval=is_latex_output(), echo=FALSE}
-datasets_tb |>
-  kable(caption = "quollr datasets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-## Preprocessing
-
-In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
-To achieve this, we implement two key scaling steps.
-Firstly, we scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval.
-Secondly, we scale the second 2D embedding component to range between $0$ and $y_{max}$.
-This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.
-
-```{=tex}
-\begin{equation}\label{eq:aspect_ratio}
-ar = \frac{r_2}{r_1} \tag{1}
-\end{equation}
-```
-```{=tex}
-\begin{equation}\label{eq:y_max}
-y_{max} = ceiling(\frac{ar}{hr}) * hr \tag{2}
-\end{equation}
-```
-
-```{r}
-scaled_data <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", 
-                y = "UMAP2", hex_ratio = NA)
-glimpse(scaled_data)
-```
-
-## Construct the 2D model
-
-### Compute hexagonal bin configurations
-
-## Construct the high-D model
-
-## Model function
-
-The `fit_highd_model()` function is used to generate both the 2D and high-D models.
-
-```{r}
-fit_highd_model(training_data = s_curve_noise_training, x = "UMAP1", y = "UMAP2",
-nldr_df_with_id = s_curve_noise_umap_scaled, col_start_2d = "UMAP", col_start_highd = "x")
-```
-
-## Model summaries
-
-### Predict 2D embeddings
-
-There are some of NLDR techniques that don't give any functions for predictions. In that sense our methodology facilitates to compute 2D embedding with our workflow.
-
-### Goodness of fit statistics
-
-There are two Goodness of fit statistics were produced. MSE and AIC which interpret the model accuracy.
-
-## Visualizations
-
-We use static visualizations to understand the model constructed in 2D. On the other hand, dynamic visualization is used to see how the model looks in high-D space.
-
-#### Static visualizations
-
-Static visualizations main involves two types of results. One is the triangulation and the other is the long edge removal. Both types of visualizations provide ggplot objects.
-
--   Triangulation result: To visualize the results of triangulation, we input a dataset containing hexagonal bin centroid coordinates where 2D embedding data exists.
-`geom_trimesh()` is used to visualize this result.
-
--   Long edge removal: The long edge removal process involves identifying and removing long edges from the triangular mesh.
-Table shows the main arguments of the functions.
-We offer two functions for visualizing this process:
-
--   `colour_long_edges()`: This function colors the long edges within the triangular mesh by red.
-
--   `remove_long_edges()`: After identifying long edges, this function draws the triangular mesh without the long edges.
-
-`r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:lgvis-tb-html)', '\\@ref(tab:lgvis-tb-pdf)'))`
-
-```{r, echo=FALSE}
-lg_vis_tb <- tibble(dt = c(".data", 
-                           "benchmark_value", 
-                           "triangular_object", 
-                           "distance_col"), 
-                    text = c("The data frame containing the edge information.",
-                             "The threshold value to determine long edges.", 
-                             "The triangular object containing the mesh information.", 
-                             "The column name in `.data` representing the distances."))
-```
-
-```{r lgvis-tb-html, eval=is_html_output(), echo=FALSE}
-lg_vis_tb |> 
-  kable(caption = "The main arguments for `vis_lg_mesh()` and `vis_rmlg_mesh()`", col.names = c("argument", "explanation")) 
-```
-
-```{r lgvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
-lg_vis_tb |>
-  kable(caption = "The main arguments for `vis\\_lg\\_mesh()` and `vis\\_rmlg\\_mesh()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-#### Dynamic visaulizations
-
-The `show_langevitour()` function enables dynamic visualization of the 2D model alongside the high-dimensional (high-D) data in its original space.
-This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data.
-The main arguments are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:dyvis-tb-html)', '\\@ref(tab:dyvis-tb-pdf)'))`.
-
-```{r, echo=FALSE}
-dyn_vis_tb <- tibble(dt = c("df", 
-                            "df_b", 
-                            "df_b_with_center_data", 
-                            "benchmark_value",
-                            "distance_df",
-                            "distance_col",
-                            "use_default_benchmark_val",
-                            "column_start_text"), 
-                     text = c("A data frame containing the high-dimensional data.",
-                              "A data frame containing the high-dimensional coordinates of bin centroids/means.",
-                              "The dataset with hexbin centroids/ means.", 
-                              "The benchmark value used to remove long edges (optional).", 
-                              "The distance dataframe.",
-                              "The name of the distance column.",
-                              "Logical, indicating whether to use default benchmark value  to remove long edges(default is FALSE).",
-                              "The text that begin the column name of the high-dimensional data."))
-```
-
-```{r dyvis-tb-html, eval=is_html_output(), echo=FALSE}
-dyn_vis_tb |> 
-  kable(caption = "The main arguments for `show_langevitour()`", col.names = c("argument", "explanation")) 
-```
-
-```{r dyvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
-dyn_vis_tb |>
-  kable(caption = "The main arguments for `show\\_langevitour()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
-```
-
-## Distance with bin means
-
-```{r, echo=FALSE}
-compute_weights <- function(nldr_df_with_hex_id) {
-
-  ## To get the 2D embeddings average of each bin
-  bin_val_hexagons <- nldr_df_with_hex_id |>
-    dplyr::group_by(hb_id) |>
-    dplyr::summarise(dplyr::across(tidyselect::everything(), mean))
-
-  ## Rename columns of averaged 2D embeddings
-  names(bin_val_hexagons) <- c("hb_id", paste0("avg_", tolower(names(nldr_df_with_hex_id)[1:2])))
-
-  ## To calculate distances from average point
-  nldr_with_avg_all <- dplyr::inner_join(bin_val_hexagons , nldr_df_with_hex_id,
-                                         by = c("hb_id" = "hb_id"))
-
-  col_names <- c(names(bin_val_hexagons), names(nldr_df_with_hex_id)[1:2], "distance")
-
-  ## Initialize the vectors to store data
-  hexids <- integer(0)
-  emb1_vec <- numeric(0)
-  emb2_vec <- numeric(0)
-  weight_vec <- numeric(0)
-
-  for(hb_id in unique(nldr_with_avg_all$hb_id)){
-
-    ## These are the weights for weighted mean
-    weighted_mean_df <- nldr_with_avg_all |>
-      dplyr::filter(hb_id == hb_id) |>
-      cal_2d_dist(start_x = col_names[2], start_y = col_names[3], end_x = col_names[4],
-                  end_y = col_names[5], select_vars = col_names)
-
-    hexids <- c(hexids, weighted_mean_df$hb_id)
-    emb1_vec <- c(emb1_vec, weighted_mean_df[[rlang::as_string(rlang::sym(col_names[4]))]])
-    emb2_vec <- c(emb2_vec, weighted_mean_df[[rlang::as_string(rlang::sym(col_names[5]))]])
-    weight_vec <- c(weight_vec, 1/ (weighted_mean_df$distance + 0.05))
-
-  }
-
-  weight_list <- list(hb_id = hexids, emb1_vec = emb1_vec, emb2_vec = emb2_vec,
-                      weight_vec = weight_vec)
-  names(weight_list) <- c("hb_id", names(nldr_df_with_hex_id)[1:2], "weights")
-
-  return(weight_list)
-
-}
-```
-
-```{r, echo=FALSE}
-weighted_highD_data <- function(training_data, nldr_df_with_hex_id,
-                                column_start_text = "x") {
-  ## Remove ID column from training data
-  training_data <- training_data |>
-    dplyr::select(-ID)
-
-  ## Join training data with 2D embeddings
-  df_all <- dplyr::bind_cols(training_data, nldr_df_with_hex_id)
-
-  ## To obtain weights corresponding to 2D embeddings
-  weight_df <- as.data.frame(do.call(cbind,
-                                     compute_weights(nldr_df_with_hex_id = nldr_df_with_hex_id)))
-
-  ## Initialize the columns that need to use for joining
-  joined_col_names <- names(nldr_df_with_hex_id)[1:2]
-
-  ## To join the training data, 2D embeddings and weights
-  weighted_mean_all <- dplyr::inner_join(df_all, weight_df,
-                                         by = c("hb_id" = "hb_id",
-                                                stats::setNames(joined_col_names,
-                                                                joined_col_names)))
-  ## List to store weighted means
-  weighted_mean_df_list <- list()
-
-  for (j in 1:NCOL(training_data)) {
-
-    ## To compute weighted mean across all high-D coordinates
-    weighted_mean_df_list[[j]] <- weighted_mean_all |>
-      dplyr::select(hb_id, names(training_data)[j], weights) |>
-      dplyr::group_by(hb_id) |>
-      dplyr::summarise(dplyr::across(names(training_data)[j], ~ weighted.mean(., weights)))
-
-  }
-
-  ## To combine the elements given in the list
-  weighted_mean <- Reduce(function(dtf1,dtf2) dplyr::full_join(dtf1,dtf2,by="hb_id"),
-                          weighted_mean_df_list)
-
-  ## Column names start with x
-  weighted_mean <- weighted_mean |>
-    dplyr::select(hb_id, tidyselect::starts_with(column_start_text))
-
-  return(weighted_mean)
-}
-```
-
-```{r, echo=FALSE}
-
-extract_hexbin_mean <- function(nldr_df_with_hex_id, counts_df) {
-
-  ## To arrange the hexagon IDs
-  counts_df <- counts_df |>
-    dplyr::arrange(hb_id)
-
-  ## To compute hexagonal bin means
-  hex_mean_df <- nldr_df_with_hex_id |>
-    dplyr::group_by(hb_id) |>
-    dplyr::summarise(dplyr::across(tidyselect::everything(), mean)) |>
-    dplyr::arrange(hb_id) |>
-    dplyr::filter(hb_id %in% counts_df$hb_id) |>
-    dplyr::mutate(std_counts = counts_df$std_counts)
-
-  ## Rename columns
-  names(hex_mean_df) <- c("hexID", "c_x", "c_y", "std_counts")
-
-  return(hex_mean_df)
-}
-```
-
-```{r}
-bin_list <- calc_bins(data = s_curve_noise_umap_scaled, 
-                      x = "UMAP1", y = "UMAP2", 
-                      hex_size = NA, buffer_x = NA, 
-                      buffer_y = NA)
-num_bins_x <- bin_list$num_x
-num_bins_y <- bin_list$num_y
-
-hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, 
-                      x = "UMAP1", y = "UMAP2", 
-                      num_bins_x = num_bins_x, num_bins_y = num_bins_y, 
-                      x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, 
-                      hex_size = NA, col_start = "UMAP")
-
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
-counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
-nldr_df_with_hex_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))
-
-## To obtain bin centroids
-df_bin_centroids <- extract_hexbin_mean(nldr_df_with_hex_id = nldr_df_with_hex_id,
-                                             counts_df = counts_df)
-
-df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), nldr_df_with_hex_id)
-
-ggplot() + 
-  geom_trimesh(data = df_bin_centroids, mapping = aes(x = c_x, y = c_y)) +
-  coord_fixed()
-
-tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
-tr_from_to_df <- gen_edges(tri_object = tr1_object)
-distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
-                           start_x = "x_from", start_y = "y_from", 
-                           end_x = "x_to", end_y = "y_to", 
-                           select_vars = c("from", "to", "distance"))
-
-## averaged high-D data
-df_bin <- weighted_highD_data(training_data = s_curve_noise_training, nldr_df_with_hex_id = nldr_df_with_hex_id, column_start_text = "x")
-
-vis_lg_mesh(distance_edges = distance_df, benchmark_value = 1,
-tr_coord_df = tr_from_to_df, distance_col = "distance")
-
-vis_rmlg_mesh(distance_edges = distance_df, benchmark_value = 1,
-tr_coord_df = tr_from_to_df, distance_col = "distance")
-
-show_langevitour(df = df_all, df_b = df_bin, 
-                 df_b_with_center_data = df_bin_centroids, 
-                 benchmark_value = 1, distance = distance_df, 
-                 distance_col = "distance", 
-                 use_default_benchmark_val = FALSE, col_start = "x")
-```
+## Prediction
 
@@
 
+<!--select a dataset from cardinalR-->
+
 # Conclusion

### 2024-03-30 paper-quollr.tex
Added main structure of the quollr paper

@@
 % !TeX root = RJwrapper.tex
-\title{quollr: An R Package for Visalizing 2D Models in High Dimensional Space}
+\title{quollr: An R Package for Visualizing 2D Models from Nonlinear Dimension Reductions in High Dimensional Space}
 
@@
 
-\hypertarget{usage}{%
-\subsection{Usage}\label{usage}}
+\hypertarget{implementation}{%
+\section{Implementation}\label{implementation}}
 
-\begin{itemize}
-\item
-  dependencies
-\item
-  basic example
-\end{itemize}
+The package can be installed from CRAN:
 
-\hypertarget{datasets}{%
-\subsubsection{Datasets}\label{datasets}}
+\texttt{install.packages("quollr")}
 
-The \texttt{quollr} package comes with several data sets that load with the package. These are described in Table \ref{tab:datasets-tb-pdf}.
+The development version can be installed from GitHub:
 
-\begin{table}
+\texttt{devtools::install\_github("JayaniLakshika/quollr")}
 
-\caption{\label{tab:datasets-tb-pdf}quollr datasets}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-data & explanation\\
-\midrule
-s\_curve\_noise & Simulated 3D S-curve data with additional four noise dimensions.\\
-s\_curve\_noise\_training & Training data derived from S-curve data.\\
-s\_curve\_noise\_test & Test data derived from S-curve data.\\
-s\_curve\_noise\_umap & UMAP 2D embedding data of S-curve data (n\_neighbors: 15, min\_dist: 0.1).\\
-s\_curve\_noise\_umap\_predict & Predicted UMAP 2D embedding data of S-curve data\\
-\addlinespace
-s\_curve\_noise\_umap\_scaled & Scaled UMAP 2D embedding data of S-curve data\\
-\bottomrule
-\end{tabular}
-\end{table}
+The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded.
+We also want to load the built-in data sets \texttt{s\_curve\_noise\_training} and \texttt{s\_curve\_noise\_umap}.
+
+\hypertarget{constructing-the-2d-model}{%
+\subsection{Constructing the 2D model}\label{constructing-the-2d-model}}
 
 \hypertarget{preprocessing}{%
-\subsection{Preprocessing}\label{preprocessing}}
+\subsubsection{Preprocessing}\label{preprocessing}}
 
-In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
-To achieve this, we implement two key scaling steps.
-Firstly, we scale the first 2D embedding component to range between \(0\) and \(1\), ensuring that all data points fall within this normalized interval.
-Secondly, we scale the second 2D embedding component to range between \(0\) and \(y_{max}\).
-This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.
+\hypertarget{binning-data}{%
+\subsubsection{Binning data}\label{binning-data}}
 
-\begin{equation}\label{eq:aspect_ratio}
-ar = \frac{r_2}{r_1} \tag{1}
-\end{equation}
+\hypertarget{obtain-bin-centroids-means}{%
+\subsubsection{Obtain bin centroids/ means}\label{obtain-bin-centroids-means}}
 
-\begin{equation}\label{eq:y_max}
-y_{max} = ceiling(\frac{ar}{hr}) * hr \tag{2}
-\end{equation}
+\hypertarget{triangulate-bin-centroids-means}{%
+\subsubsection{Triangulate bin centroids/ means}\label{triangulate-bin-centroids-means}}
 
-\begin{verbatim}
-scaled_data <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", 
-                y = "UMAP2", hex_ratio = NA)
-glimpse(scaled_data)
-\end{verbatim}
+\hypertarget{lifting-the-model-into-high-dimensions}{%
+\subsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}}
 
-\begin{verbatim}
-#> List of 2
-#>  $ scaled_UMAP1: num [1:75] 0.0804 0.7386 0.8399 0.1672 0.2629 ...
-#>  $ scaled_UMAP2: num [1:75] 0.366 1.1464 1.2392 0.0494 0.4556 ...
-\end{verbatim}
+\hypertarget{tuning-model-parameters}{%
+\subsection{Tuning model parameters}\label{tuning-model-parameters}}
 
-\hypertarget{construct-the-2d-model}{%
-\subsection{Construct the 2D model}\label{construct-the-2d-model}}
-
-\hypertarget{compute-hexagonal-bin-configurations}{%
-\subsubsection{Compute hexagonal bin configurations}\label{compute-hexagonal-bin-configurations}}
-
-\hypertarget{construct-the-high-d-model}{%
-\subsection{Construct the high-D model}\label{construct-the-high-d-model}}
-
-\hypertarget{model-function}{%
-\subsection{Model function}\label{model-function}}
-
-The \texttt{fit\_highd\_model()} function is used to generate both the 2D and high-D models.
-
-\begin{verbatim}
-fit_highd_model(training_data = s_curve_noise_training, x = "UMAP1", y = "UMAP2",
-nldr_df_with_id = s_curve_noise_umap_scaled, col_start_2d = "UMAP", col_start_highd = "x")
-\end{verbatim}
-
-\begin{verbatim}
-#> $df_bin
-#> # A tibble: 10 x 8
-#>    hb_id       x1     x2     x3        x4        x5       x6        x7
-#>    <dbl>    <dbl>  <dbl>  <dbl>     <dbl>     <dbl>    <dbl>     <dbl>
-#>  1     2 -0.637   1.74   -1.76   0.00953  -0.00143  -0.0117  -0.00152 
-#>  2     6 -0.498   0.524  -1.73  -0.000237  0.00234  -0.0297  -0.000242
-#>  3     7  0.294   1.40   -1.88   0.00890  -0.00803  -0.0123  -0.00120 
-#>  4    12  0.309   0.0421 -1.83   0.00656   0.00823   0.00489 -0.00389 
-#>  5    13  0.868   0.747  -0.781 -0.00408   0.000857  0.0248   0.00170 
-#>  6    18  0.357   1.27   -0.169  0.00607   0.00124   0.0152   0.00204 
-#>  7    24 -0.792   1.25    0.514 -0.000777  0.000464  0.00602  0.000371
-#>  8    28 -0.597   1.19    1.77   0.000240 -0.00417  -0.0185  -0.000786
-#>  9    29 -0.00544 0.211   1.92   0.00116   0.00266   0.00949 -0.00209 
-#> 10    34  0.622   1.21    1.64  -0.000560  0.00540  -0.00741 -0.000886
-#> 
-#> $df_bin_centroids
-#>    hexID       c_x   c_y std_counts
-#> 1      2 0.1732051 -0.15  0.2352941
-#> 2      6 0.0000000  0.15  0.5294118
-#> 3      7 0.3464102  0.15  0.4117647
-#> 4     12 0.1732051  0.45  0.1764706
-#> 5     13 0.5196152  0.45  0.3529412
-#> 6     18 0.6928203  0.75  0.7058824
-#> 7     24 0.8660254  1.05  0.4705882
-#> 8     28 0.6928203  1.35  0.2941176
-#> 9     29 1.0392305  1.35  0.2352941
-#> 10    34 0.8660254  1.65  1.0000000
-\end{verbatim}
-
-\hypertarget{model-summaries}{%
-\subsection{Model summaries}\label{model-summaries}}
-
-\hypertarget{predict-2d-embeddings}{%
-\subsubsection{Predict 2D embeddings}\label{predict-2d-embeddings}}
-
-There are some of NLDR techniques that don't give any functions for predictions. In that sense our methodology facilitates to compute 2D embedding with our workflow.
-
-\hypertarget{goodness-of-fit-statistics}{%
-\subsubsection{Goodness of fit statistics}\label{goodness-of-fit-statistics}}
-
-There are two Goodness of fit statistics were produced. MSE and AIC which interpret the model accuracy.
+\hypertarget{compute-mse}{%
+\subsection{Compute MSE}\label{compute-mse}}
 
@@
 
-We use static visualizations to understand the model constructed in 2D. On the other hand, dynamic visualization is used to see how the model looks in high-D space.
+\hypertarget{d-model-visualization}{%
+\subsubsection{2D model visualization}\label{d-model-visualization}}
 
-\hypertarget{static-visualizations}{%
-\paragraph{Static visualizations}\label{static-visualizations}}
+\hypertarget{p-d-model-visualization}{%
+\subsubsection{p-D model visualization}\label{p-d-model-visualization}}
 
-Static visualizations main involves two types of results. One is the triangulation and the other is the long edge removal. Both types of visualizations provide ggplot objects.
-
-\begin{itemize}
-\item
-  Triangulation result: To visualize the results of triangulation, we input a dataset containing hexagonal bin centroid coordinates where 2D embedding data exists.
-\texttt{geom\_trimesh()} is used to visualize this result.
-\item
-  Long edge removal: The long edge removal process involves identifying and removing long edges from the triangular mesh.
-Table shows the main arguments of the functions.
-We offer two functions for visualizing this process:
-\item
-  \texttt{colour\_long\_edges()}: This function colors the long edges within the triangular mesh by red.
-\item
-  \texttt{remove\_long\_edges()}: After identifying long edges, this function draws the triangular mesh without the long edges.
-\end{itemize}
-
-\ref{tab:lgvis-tb-pdf}
-
-\begin{table}
-
-\caption{\label{tab:lgvis-tb-pdf}The main arguments for `vis\_lg\_mesh()` and `vis\_rmlg\_mesh()`}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-argument & explanation\\
-\midrule
-.data & The data frame containing the edge information.\\
-benchmark\_value & The threshold value to determine long edges.\\
-triangular\_object & The triangular object containing the mesh information.\\
-distance\_col & The column name in `.data` representing the distances.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{dynamic-visaulizations}{%
-\paragraph{Dynamic visaulizations}\label{dynamic-visaulizations}}
-
-The \texttt{show\_langevitour()} function enables dynamic visualization of the 2D model alongside the high-dimensional (high-D) data in its original space.
-This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data.
-The main arguments are shown in Table \ref{tab:dyvis-tb-pdf}.
-
-\begin{table}
-
-\caption{\label{tab:dyvis-tb-pdf}The main arguments for `show\_langevitour()`}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-argument & explanation\\
-\midrule
-df & A data frame containing the high-dimensional data.\\
-df\_b & A data frame containing the high-dimensional coordinates of bin centroids/means.\\
-df\_b\_with\_center\_data & The dataset with hexbin centroids/ means.\\
-benchmark\_value & The benchmark value used to remove long edges (optional).\\
-distance\_df & The distance dataframe.\\
-\addlinespace
-distance\_col & The name of the distance column.\\
-use\_default\_benchmark\_val & Logical, indicating whether to use default benchmark value  to remove long edges(default is FALSE).\\
-column\_start\_text & The text that begin the column name of the high-dimensional data.\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{distance-with-bin-means}{%
-\subsection{Distance with bin means}\label{distance-with-bin-means}}
-
-\begin{verbatim}
-bin_list <- calc_bins(data = s_curve_noise_umap_scaled, 
-                      x = "UMAP1", y = "UMAP2", 
-                      hex_size = NA, buffer_x = NA, 
-                      buffer_y = NA)
-num_bins_x <- bin_list$num_x
-num_bins_y <- bin_list$num_y
-
-hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, 
-                      x = "UMAP1", y = "UMAP2", 
-                      num_bins_x = num_bins_x, num_bins_y = num_bins_y, 
-                      x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, 
-                      hex_size = NA, col_start = "UMAP")
-
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
-counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
-nldr_df_with_hex_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))
-
-## To obtain bin centroids
-df_bin_centroids <- extract_hexbin_mean(nldr_df_with_hex_id = nldr_df_with_hex_id,
-                                             counts_df = counts_df)
-
-df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), nldr_df_with_hex_id)
-
-ggplot() + 
-  geom_trimesh(data = df_bin_centroids, mapping = aes(x = c_x, y = c_y)) +
-  coord_fixed()
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf}
-
-\begin{verbatim}
-tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
-tr_from_to_df <- gen_edges(tri_object = tr1_object)
-distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
-                           start_x = "x_from", start_y = "y_from", 
-                           end_x = "x_to", end_y = "y_to", 
-                           select_vars = c("from", "to", "distance"))
-
-## averaged high-D data
-df_bin <- weighted_highD_data(training_data = s_curve_noise_training, nldr_df_with_hex_id = nldr_df_with_hex_id, column_start_text = "x")
-
-vis_lg_mesh(distance_edges = distance_df, benchmark_value = 1,
-tr_coord_df = tr_from_to_df, distance_col = "distance")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-2.pdf}
-
-\begin{verbatim}
-vis_rmlg_mesh(distance_edges = distance_df, benchmark_value = 1,
-tr_coord_df = tr_from_to_df, distance_col = "distance")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-3.pdf}
-
-\begin{verbatim}
-show_langevitour(df = df_all, df_b = df_bin, 
-                 df_b_with_center_data = df_bin_centroids, 
-                 benchmark_value = 1, distance = distance_df, 
-                 distance_col = "distance", 
-                 use_default_benchmark_val = FALSE, col_start = "x")
-\end{verbatim}
-
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-4.pdf}
+\hypertarget{prediction}{%
+\subsection{Prediction}\label{prediction}}
 

### 2024-03-30 paper_old.qmd
Added main structure of the quollr paper

---
title: "quollr: An R Package for Visalizing 2D Models in High Dimensional Space"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  - name: Jayani P.G. Lakshika
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email:  \email{jayani.piyadigamage@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)

```

```{r load-libraries}
library(quollr)
library(tibble)
library(knitr)
library(kableExtra)
library(ggplot2)

set.seed(20230531)
```

# Introduction

# Methodology

## Usage

-   dependencies

-   basic example

### Datasets

The `quollr` package comes with several data sets that load with the package. These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.

```{r, echo=FALSE}
datasets_tb <- tibble(dt = c("s_curve_noise",
                             "s_curve_noise_training", 
                             "s_curve_noise_test", 
                             "s_curve_noise_umap",
                             "s_curve_noise_umap_predict",
                             "s_curve_noise_umap_scaled"), 
                      text = c("Simulated 3D S-curve data with additional four noise dimensions.",
                               "Training data derived from S-curve data.", 
                               "Test data derived from S-curve data.", 
                               "UMAP 2D embedding data of S-curve data (n_neighbors: 15, min_dist: 0.1).",
                               "Predicted UMAP 2D embedding data of S-curve data",
                               "Scaled UMAP 2D embedding data of S-curve data"))
```

```{r datasets-tb-html, eval=is_html_output(), echo=FALSE}
datasets_tb |> 
  kable(caption = "quollr datasets", col.names = c("data", "explanation")) 
```

```{r datasets-tb-pdf, eval=is_latex_output(), echo=FALSE}
datasets_tb |>
  kable(caption = "quollr datasets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

## Preprocessing

In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization. To achieve this, we implement two key scaling steps. Firstly, we scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval. Secondly, we scale the second 2D embedding component to range between $0$ and $y_{max}$. This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.

```{=tex}
\begin{equation}\label{eq:aspect_ratio}
ar = \frac{r_2}{r_1} \tag{1}
\end{equation}
```
```{=tex}
\begin{equation}\label{eq:y_max}
y_{max} = ceiling(\frac{ar}{hr}) * hr \tag{2}
\end{equation}
```
```{r}
scaled_data <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", 
                y = "UMAP2", hex_ratio = NA)
glimpse(scaled_data)
```

## Construct the 2D model

### Compute hexagonal bin configurations

## Construct the high-D model

## Model function

The `fit_highd_model()` function is used to generate both the 2D and high-D models.

```{r}
fit_highd_model(training_data = s_curve_noise_training, x = "UMAP1", y = "UMAP2",
nldr_df_with_id = s_curve_noise_umap_scaled, col_start_2d = "UMAP", col_start_highd = "x")
```

This function returns a list containing the estimated p-value as well as other useful statistics, including the empirical local correlations measured in the two time periods. See Table 4 for details concerning other arguments that may be passed to this function.

## Model summaries

### Predict 2D embeddings

There are some of NLDR techniques that don't give any functions for predictions. In that sense our methodology facilitates to compute 2D embedding with our workflow.

### Goodness of fit statistics

There are two Goodness of fit statistics were produced. MSE and AIC which interpret the model accuracy.

## Visualizations

We use static visualizations to understand the model constructed in 2D. On the other hand, dynamic visualization is used to see how the model looks in high-D space.

#### Static visualizations

Static visualizations main involves two types of results. One is the triangulation and the other is the long edge removal. Both types of visualizations provide ggplot objects.

-   Triangulation result: To visualize the results of triangulation, we input a dataset containing hexagonal bin centroid coordinates where 2D embedding data exists. `geom_trimesh()` is used to visualize this result.

-   Long edge removal: The long edge removal process involves identifying and removing long edges from the triangular mesh. Table shows the main arguments of the functions. We offer two functions for visualizing this process:

-   `colour_long_edges()`: This function colors the long edges within the triangular mesh by red.

-   `remove_long_edges()`: After identifying long edges, this function draws the triangular mesh without the long edges.

`r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:lgvis-tb-html)', '\\@ref(tab:lgvis-tb-pdf)'))`

```{r, echo=FALSE}
lg_vis_tb <- tibble(dt = c(".data", 
                           "benchmark_value", 
                           "triangular_object", 
                           "distance_col"), 
                    text = c("The data frame containing the edge information.",
                             "The threshold value to determine long edges.", 
                             "The triangular object containing the mesh information.", 
                             "The column name in `.data` representing the distances."))
```

```{r lgvis-tb-html, eval=is_html_output(), echo=FALSE}
lg_vis_tb |> 
  kable(caption = "The main arguments for `vis_lg_mesh()` and `vis_rmlg_mesh()`", col.names = c("argument", "explanation")) 
```

```{r lgvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
lg_vis_tb |>
  kable(caption = "The main arguments for `vis\\_lg\\_mesh()` and `vis\\_rmlg\\_mesh()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

#### Dynamic visaulizations

The `show_langevitour()` function enables dynamic visualization of the 2D model alongside the high-dimensional (high-D) data in its original space. This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data. The main arguments are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:dyvis-tb-html)', '\\@ref(tab:dyvis-tb-pdf)'))`.

```{r, echo=FALSE}
dyn_vis_tb <- tibble(dt = c("df", 
                            "df_b", 
                            "df_b_with_center_data", 
                            "benchmark_value",
                            "distance_df",
                            "distance_col",
                            "use_default_benchmark_val",
                            "column_start_text"), 
                     text = c("A data frame containing the high-dimensional data.",
                              "A data frame containing the high-dimensional coordinates of bin centroids/means.",
                              "The dataset with hexbin centroids/ means.", 
                              "The benchmark value used to remove long edges (optional).", 
                              "The distance dataframe.",
                              "The name of the distance column.",
                              "Logical, indicating whether to use default benchmark value  to remove long edges(default is FALSE).",
                              "The text that begin the column name of the high-dimensional data."))
```

```{r dyvis-tb-html, eval=is_html_output(), echo=FALSE}
dyn_vis_tb |> 
  kable(caption = "The main arguments for `show_langevitour()`", col.names = c("argument", "explanation")) 
```

```{r dyvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
dyn_vis_tb |>
  kable(caption = "The main arguments for `show\\_langevitour()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

## Distance with bin means

```{r, echo=FALSE}

extract_hexbin_mean <- function(nldr_df_with_hex_id, counts_df) {

  ## To arrange the hexagon IDs
  counts_df <- counts_df |>
    dplyr::arrange(hb_id)

  ## To compute hexagonal bin means
  hex_mean_df <- nldr_df_with_hex_id |>
    dplyr::group_by(hb_id) |>
    dplyr::summarise(dplyr::across(tidyselect::everything(), mean)) |>
    dplyr::arrange(hb_id) |>
    dplyr::filter(hb_id %in% counts_df$hb_id) |>
    dplyr::mutate(std_counts = counts_df$std_counts)

  ## Rename columns
  names(hex_mean_df) <- c("hexID", "c_x", "c_y", "std_counts")

  return(hex_mean_df)
}
```

```{r}
bin_list <- calc_bins(data = s_curve_noise_umap_scaled, 
                      x = "UMAP1", y = "UMAP2", 
                      hex_size = NA, buffer_x = NA, 
                      buffer_y = NA)
num_bins_x <- bin_list$num_x
num_bins_y <- bin_list$num_y

hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, 
                      x = "UMAP1", y = "UMAP2", 
                      num_bins_x = num_bins_x, num_bins_y = num_bins_y, 
                      x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, 
                      hex_size = NA, col_start = "UMAP")

all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
nldr_df_with_hex_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))

## To obtain bin centroids
df_bin_centroids <- extract_hexbin_mean(nldr_df_with_hex_id = nldr_df_with_hex_id,
                                             counts_df = counts_df)

df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), nldr_df_with_hex_id)

ggplot() + 
  geom_trimesh(data = df_bin_centroids, mapping = aes(x = c_x, y = c_y)) +
  coord_fixed()

tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df <- gen_edges(tri_object = tr1_object)
distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
                           start_x = "x_from", start_y = "y_from", 
                           end_x = "x_to", end_y = "y_to", 
                           select_vars = c("from", "to", "distance"))

## averaged high-D data
df_bin <- avg_highd_data(data = df_all, col_start = "x")

vis_lg_mesh(distance_edges = distance_df, benchmark_value = 1,
tr_coord_df = tr_from_to_df, distance_col = "distance")

vis_rmlg_mesh(distance_edges = distance_df, benchmark_value = 1,
tr_coord_df = tr_from_to_df, distance_col = "distance")

show_langevitour(df = df_all, df_b = df_bin, 
                 df_b_with_center_data = df_bin_centroids, 
                 benchmark_value = 1, distance = distance_df, 
                 distance_col = "distance", 
                 use_default_benchmark_val = FALSE, col_start = "x")
```

## Tests

All functions have tests written and implemented using the \CRANpkg{testthat} [@testthat] in R.

# Application

# Conclusion

# Acknowledgements

This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template. The source code for reproducing this paper can be found at: <https://github.com/JayaniLakshika/paper-quollr>.

### 2024-03-30 quollr_code.R

@@ -1,1361 +1,0 @@
-calculate_effective_x_bins <- function(.data, x = "UMAP1", hex_size = NA, buffer_x = NA){
-
-  if (anyNA(.data[[rlang::as_string(rlang::ensym(x))]])) {
-    stop("NAs present")
-  }
-
-  if (any(is.infinite(.data |> dplyr::pull({{ x }})))) {
-    stop("Inf present")
-  }
-
-
-
-  if (is.na(hex_size)) {
-    # cell_area <- 1
-    #
-    # ## To compute the radius of the hexagon outer circle
-    # hex_size <- sqrt(2 * cell_area / (3 *sqrt(3)))
-    hex_size <- 0.2
-
-  } else {
-    if ((hex_size <= 0) || (is.infinite(hex_size))) {
-      stop("Invalid hex size value")
-
-    }
-  }
-
-  if (is.na(buffer_x)) {
-    #buffer_x <- sqrt(3) * hex_size/2
-    # buffer_x <- min(min(.data[[rlang::as_string(rlang::sym(x))]]) - sqrt(3) * hex_size/2,
-    #                 max(.data[[rlang::as_string(rlang::sym(x))]]) + sqrt(3) * hex_size/2,
-    #                 s_curve_noise_umap[which.min(s_curve_noise_umap$UMAP2), "UMAP1"] |> dplyr::pull("UMAP1") - (sqrt(3) * hex_size/2),
-    #                 s_curve_noise_umap[which.max(s_curve_noise_umap$UMAP2), "UMAP1"] |> dplyr::pull("UMAP1") +
-    #                   (sqrt(3) * hex_size/2))
-    buffer_x <- sqrt(3) * hex_size * 1.5
-    message(paste0("Buffer along the x-axis set to ", buffer_x, "."))
-  } else {
-
-    ## Buffer size is exceeds
-    if (buffer_x > (sqrt(3) * hex_size)) {
-      stop(paste0("Buffer along the y-axis exceeds than ", sqrt(3) * hex_size, ".
-                  Need to assign a value less than ", sqrt(3) * hex_size, "."))
-
-    }
-
-
-  }
-
-  ## To compute the range along x-axis
-  xwidth <- diff(range(.data |>
-                         dplyr::pull({{ x }})))  + buffer_x
-
-  horizontal_spacing <- sqrt(3) * hex_size
-
-  num_bins <- ceiling(xwidth/horizontal_spacing)
-  num_bins
-
-}
-
[…]

### 2024-03-30 paper-quollr.Rmd
Added main texts for the some steps

@@
 
+<!-- 20 pages-->
+
 # Introduction
@@
 ### Preprocessing
+<!--Discuss scaling the NLDR data-->
 
-<!--Discuss scaling the NLDR data-->
+The function `gen_scaled_data()` is used to prepare the NLDR data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
+
+<!-- add plot before and after scaling NLDR data-->
 
@@
 
+<!-- add about main function for hexagonal bining-->
+
 ### Obtain bin centroids/ means
+
+To obtain hexagonal bin centroids, `extract_hexbin_centroids()`is used. The function `extract_hexbin_mean()` is used to obtain average within each hexagon.
 
@@
 
-<!-- add about to do hex binning at once-->
+The `tri_bin_centroids()` is used to do triangulate bin centroids/ bin means. Then, `gen_edges()` compute the edges to obtain the triangular mesh which is the 2D model.
 
@@
 
+The $p$-D model is generated by `avg_highd_data()` by passing the $p$-D data.
+
 ## Tuning model parameters
 
-## Compute MSE
+Mainly there are two model parameter need to consider: benchmark value to remove the low-density hexagons, and benchmark value to remove the long edges.
+`find_low_dens_hex()` is used to find the hexagons which contains less number of points by considering the density of their neighboring points as well.
+In here, user can first decide which are the low-density hexagons and pass them to this function to check whether these removal of low-density hexagons can affect the model fit by looking at the the neighbors.
+On the other hand, `find_lg_benchmark()` function is used to compute the threshold for removing long edges.
+
+## Compute residuals and Mean Square Error (MSE)
+
+As a Goodness of fit statistics for the model, `gen_summary()` is used to compute residuals and MSE. For the fitted model, how residuals and MSE is changing is shown in Figure 2.
+
+<!--add a plot with total bins Vs MSE, non-empty bins Vs Error-->
 
@@
 
+The package provides four basic visualizations which includes three visualizations regarding 2D model (static vis) and one is regrading $p$-D model (dynamic vis).
+Each visualization can be done their own functions which describes in this section.
+
+<!--add trimesh, colored trimesh, removed edge trimesh, screenshots from langevitour with the model and data-->
+
 ### 2D model visualization
 
-<!--2D model vis: triangular mesh-->
+<!--2D model vis: triangular mesh, colored trimesh, removed edge trimesh-->
+To visualize the 2D model, mainly three functions are used.
+As shown in Figure 1, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`.
+After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges.
+As shown in Figure 1, the small and long edges are colored by black and red respectively.
+Following this, `vis_rmlg_mesh()` is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure 1).
+In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in 2D.
 
@@
 <!--highD model vis: langevitour with the model-->
+
+The $p$-D model overlaid on data is visualized by the function `show_langevitour()`. This is a useful dynamic visualization result to validate visually whether the model fits the data well or not.
 
@@
 
+`predict_emb()` function is used to predict 2D embedding for a new data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
+
 ## Tests
@@
 <!--select a dataset from cardinalR-->
+<!-- diff shape clusters-->
+<!-- talk with best fit-->
 

### 2024-03-30 paper-quollr.tex
Added main texts for the some steps

@@
 
+The function \texttt{gen\_scaled\_data()} is used to prepare the NLDR data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
+
 \hypertarget{binning-data}{%
@@
 
+To obtain hexagonal bin centroids, \texttt{extract\_hexbin\_centroids()}is used. The function \texttt{extract\_hexbin\_mean()} is used to obtain average within each hexagon.
+
 \hypertarget{triangulate-bin-centroids-means}{%
 \subsubsection{Triangulate bin centroids/ means}\label{triangulate-bin-centroids-means}}
+
+The \texttt{tri\_bin\_centroids()} is used to do triangulate bin centroids/ bin means. Then, \texttt{gen\_edges()} compute the edges to obtain the triangular mesh which is the 2D model.
 
@@
 
+The \(p\)-D model is generated by \texttt{avg\_highd\_data()} by passing the \(p\)-D data.
+
 \hypertarget{tuning-model-parameters}{%
@@
 
-\hypertarget{compute-mse}{%
-\subsection{Compute MSE}\label{compute-mse}}
+Mainly there are two model parameter need to consider: benchmark value to remove the low-density hexagons, and benchmark value to remove the long edges.
+\texttt{find\_low\_dens\_hex()} is used to find the hexagons which contains less number of points by considering the density of their neighboring points as well.
+In here, user can first decide which are the low-density hexagons and pass them to this function to check whether these removal of low-density hexagons can affect the model fit by looking at the the neighbors.
+On the other hand, \texttt{find\_lg\_benchmark()} function is used to compute the threshold for removing long edges.
+
+\hypertarget{compute-residuals-and-mean-square-error-mse}{%
+\subsection{Compute residuals and Mean Square Error (MSE)}\label{compute-residuals-and-mean-square-error-mse}}
+
+As a Goodness of fit statistics for the model, \texttt{gen\_summary()} is used to compute residuals and MSE. For the fitted model, how residuals and MSE is changing is shown in Figure 2.
 
@@
 
+The package provides four basic visualizations which includes three visualizations regarding 2D model (static vis) and one is regrading \(p\)-D model (dynamic vis).
+Each visualization can be done their own functions which describes in this section.
+
 \hypertarget{d-model-visualization}{%
 \subsubsection{2D model visualization}\label{d-model-visualization}}
+
+To visualize the 2D model, mainly three functions are used.
+As shown in Figure 1, \texttt{geom\_trimesh()} to visualize the triangular mesh by adding a new layer to \texttt{ggplot()}.
+After identifying benchmark value to remove long edge, \texttt{vis\_lg\_mesh()} is used to visualize the triangular mesh by coloring the small and long edges.
+As shown in Figure 1, the small and long edges are colored by black and red respectively.
+Following this, \texttt{vis\_rmlg\_mesh()} is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure 1).
+In \texttt{vis\_lg\_mesh()} and \texttt{vis\_rmlg\_mesh()}, \texttt{benchmark\_value} argument controls the edge removal in 2D.
 
@@
 
+The \(p\)-D model overlaid on data is visualized by the function \texttt{show\_langevitour()}.
+This is a useful dynamic visualization result to validate visually whether the model fits the data well or not.
+
 \hypertarget{prediction}{%
 \subsection{Prediction}\label{prediction}}
+
+\texttt{predict\_emb()} function is used to predict 2D embedding for a new data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
 

### 2024-03-30 paper-quollr.Rmd
Add the steps to write about hexagonal binning

@@
 
+The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
+When constructing the 2D model, the user can choose either to fir the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
+
 <!-- add about main function to gen 2D and highD models-->
+<!--Discuss the model can be generated with bin centroids or bin means-->
+
 ## Constructing the 2D model
-
-<!--Discuss the model can be generated with bin centroids or bin means-->
 
@@
 
+The mains steps for the hexagonal binning algorithm can be executed by the main function `hex_binning()`, or can be run separately for more flexibility.
 <!-- add about main function for hexagonal bining-->
+
+**Step 1: Compute hexagonal grid configurations**
+
+**Step 2: Generate the hexagonal bin centroids**
+
+**Step 3: Create the hexagonal grid**
+
+**Step 4: Assign NLDR data to hexagons**
 

### 2024-03-30 paper-quollr.tex
Add the steps to write about hexagonal binning

@@
 
+The mains steps for the algorithm can be executed by the main function \texttt{fit\_highd\_model()}, or can be run separately for more flexibility.
+When constructing the 2D model, the user can choose either to fir the 2D model with hexagonal bin centroids or bin means using \texttt{is\_bin\_centroid} argument.
+
 \hypertarget{constructing-the-2d-model}{%
@@
 \subsubsection{Binning data}\label{binning-data}}
+
+The mains steps for the hexagonal binning algorithm can be executed by the main function \texttt{hex\_binning()}, or can be run separately for more flexibility.
+
+\textbf{Step 1: Compute hexagonal grid configurations}
+
+\textbf{Step 2: Generate the hexagonal bin centroids}
+
+\textbf{Step 3: Create the hexagonal grid}
+
+\textbf{Step 4: Assign NLDR data to hexagons}
 

### 2024-03-31 paper-quollr.Rmd
Add dataset for application

@@
 knitr::opts_chunk$set(warning = FALSE, 
-                      message = FALSE)
+                      message = FALSE,
+                      echo = FALSE)
 
@@
 library(ggplot2)
+library(dplyr)
 
-set.seed(20230531)
+set.seed(20240110)
+```
+
+```{r plot-theme}
+theme_set(theme_linedraw() +
+   theme(
+     aspect.ratio = 1,
+     plot.background = element_rect(fill = 'transparent', colour = NA),
+     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+     panel.background = element_rect(fill = 'transparent', 
+                                     colour = NA),
+     panel.grid.major = element_blank(), 
+     panel.grid.minor = element_blank(), 
+     axis.title.x = element_blank(), axis.title.y = element_blank(),
+     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+     legend.background = element_rect(fill = 'transparent', 
+                                      colour = NA),
+     legend.key = element_rect(fill = 'transparent', 
+                               colour = NA),
+     legend.position = "none", 
+     legend.title = element_text(size=5), 
+     legend.text = element_text(size=4),
+     legend.key.height = unit(0.25, 'cm'),
+     legend.key.width = unit(0.25, 'cm')
+   )
+
+)
 ```
@@
 
+```{r, echo=FALSE}
+clusters_different_shapes_diff_num_points <- function(sample_size = 400, with_seed = NULL, cluster_size_vec = c(50, 50, 50, 50, 100, 100), num_gussian_clusters = 4, num_non_gaussian_clusters = 2,
+                                      cluster_sd_gau = 0.05, cluster_sd_non_gau = 0.1, num_dims = 7, a = 2, b = 4) {
+
+  # To check the seed is not assigned
+  if (!is.null(with_seed)) {
+    set.seed(with_seed)
+  }
+
+  num_clusters <- num_gussian_clusters + num_non_gaussian_clusters
+
+  ## Generate Gaussian clusters
+
+  # Create a vector of possible values (0 and 1)
+  values <- c(0, 1)
+
+  # Create an expanded grid with 0's and 1's
+  mean_val_grid <- tidyr::expand_grid(!!!setNames(rep(list(values), num_dims),
+                                                  paste0("mean_dim", 1:num_dims)))
+
+  # To select combinations for assigned number of clusters
+
+  mean_val_grid_gau <- mean_val_grid |>
+    dplyr::slice_sample(n = num_gussian_clusters)
+
+  mean_val_grid_non_gau <- mean_val_grid |>
+    dplyr::slice_sample(n = num_non_gaussian_clusters)
+
+  # To generate empty tibble
+  column_names <- paste0(rep("x", num_dims), 1:num_dims)
+  df <- tibble(!!!setNames(rep(list(NULL), length(column_names)), column_names))
+
+  for (i in 1:num_gussian_clusters) {
+
+    # To filter the mean values for specific cluster
+    mean_val_for_cluster <- mean_val_grid_gau |>
+      dplyr::filter(dplyr::row_number() == i) |>
+      unlist(use.names = FALSE)
+
+    # Initialize an empty list to store the vectors with column
+    # values
+    dim_val_list <- list()
+
+    for (j in 1:num_dims) {
+
+      dim_val_list[[column_names[j]]] <- rnorm(cluster_size_vec[i], mean = mean_val_for_cluster[j],
+                                               sd = cluster_sd_gau)
+
+    }
+    # To generate a tibble for a cluster
+    df_gau_cluster <- tibble::as_tibble(dim_val_list)
+
+    df <- dplyr::bind_rows(df, df_gau_cluster)
+
+  }
+
+  for (i in 1:num_non_gaussian_clusters) {
+
+    phi <- runif(cluster_size_vec[(num_clusters - i)], max = 2*pi)
+    rho <- sqrt(runif(cluster_size_vec[(num_clusters - i)]))
+
+    # To filter the mean values for specific cluster
+    presence_of_elipse_cluster <- mean_val_grid_non_gau |>
+      dplyr::filter(dplyr::row_number() == i) |>
+      unlist(use.names = FALSE)
+
+    # Initialize an empty list to store the vectors with column
+    # values
+    dim_val_list_n <- list()
+
+    for (j in 1:num_dims) {
+      if(presence_of_elipse_cluster[j] == 1){
+        dim_val_list_n[[column_names[j]]] <- sqrt(a)*rho*cos(phi) + b
+        ## Surface of poolar coordinate
+      } else {
+        dim_val_list_n[[column_names[j]]] <- rnorm(cluster_size_vec[(num_clusters - i)], mean = 0,
+                                                   sd = cluster_sd_non_gau)
+
+      }
+
+    }
+    # To generate a tibble for a cluster
+    df_non_gau_cluster <- tibble::as_tibble(dim_val_list_n)
+
+    df <- dplyr::bind_rows(df, df_non_gau_cluster)
+
+  }
+
+  df
+
+}
+```
+
+```{r}
+clust_df <- clusters_different_shapes_diff_num_points(sample_size = 1500, with_seed = NULL, cluster_size_vec = c(250, 150, 150, 150, 350, 450), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.2, cluster_sd_non_gau = 0.3, num_dims = 7, a = 2, b = 4)
+
+langevitour::langevitour(clust_df)
+
+clust_df <- clust_df |>
+  dplyr::mutate(ID = row_number())
+```
+
+```{r}
+library(Rtsne)
+
+tSNE_fit <- clust_df |>
+  dplyr::select(-ID) |>
+  dplyr::select(where(is.numeric)) |>
+  Rtsne::Rtsne(perplexity = 39, pca = FALSE, pca_center = FALSE, normalize = FALSE, dims = 2)
+
+tSNE_df <- tSNE_fit$Y |> 
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::mutate(ID = clust_df$ID)
+
+names(tSNE_df) <- c("tSNE1", "tSNE2", "ID")
+
+tSNE_df_plot <- tSNE_df |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.5, colour="#e41a1c", size = 0.5) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
+        axis.text = element_text(size = 5),
+        axis.title = element_text(size = 7))
+
+tSNE_df_plot
+```
+
+<!-- MSE plot to find which is the effective number of bins to use-->
+```{r}
+tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_df, 
+                                    x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tSNE_df))    
+## tSNE
+hex_size_vec <- seq(0.02, 2, by = 0.01)
+vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+mse_df_gau <- dplyr::bind_rows(vec)[0, ]
+mse_df_gau <- mse_df_gau |>
+  dplyr::mutate_if(is.character, as.numeric)
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = tsne_gau_scaled, 
+            x = "tSNE1", y = "tSNE2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = clust_df, 
+                                   nldr_df_with_id = tsne_gau_scaled, 
+                                   x = "tSNE1", y = "tSNE2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "tSNE", 
+                                   col_start_highd = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = clust_df, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = clust_df, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "x")
+
+  mse_df_gau <- mse_df_gau |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    aic = eval_list$aic,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i])
+
+}
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_gau |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+### Group split by duplicated bins
+duplicate_df_list <- mse_df_gau |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df_gau |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df_clust <- dplyr::bind_rows(not_dupli_df, duplicate_df) 
+```
+
+```{r}
+mse_clust_plot <- ggplot(mse_df_clust, aes(x = num_bins,
+                                       y = log(mse)
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 588, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+mse_clust_plot
+```
+
+<!-- model with bin centroids-->
+```{r}
+
+## Decide by looking at MSE plot
+num_bins_x_tsne_gau <- 21
+num_bins_y_tsne_gau <- 28
+hex_size_tsne_gau <- 0.03
+## non-empty:198
+hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
+                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "tSNE")
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
+
+model_object_tsne_gau <- fit_highd_model( training_data = clust_df, 
+                                 nldr_df_with_id = tsne_gau_scaled, 
+                                 x = "tSNE1", y = "tSNE2", 
+                                 num_bins_x = num_bins_x_tsne_gau, 
+                                 num_bins_y = num_bins_y_tsne_gau, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_tsne_gau,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "tSNE", 
+                                 col_start_highd = "x")
+df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
+df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+## Triangulate bin centroids
+tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
+tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+## Compute 2D distances
+distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+# distance_plot <- plot_dist(distance_tsne_gau) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+## To find the benchmark value
+benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value =
+benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+trimesh_removed_tsne_gau
+
+df_all_tsne_gau <- dplyr::bind_cols(clust_df |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
+
+### Define type column
+df <- df_all_tsne_gau |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_tsne_gau |>
+  dplyr::filter(hb_id %in% df_bin_centroids_tsne_gau$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_tsne_gau$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+df_exe <- dplyr::bind_rows(df_b, df)
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_tsne_gau |>
+  dplyr::filter(distance < benchmark_tsne_gau)
+## Since erase brushing is considerd.
+langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = distance_df_small_edges$from,
+                         lineTo = distance_df_small_edges$to,
+                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.7, NROW(df))),
+                         levelColors = c("#6a3d9a", "#33a02c"))
+
+```
+
+<!-- model with bin means-->
+```{r}
+## Decide by looking at MSE plot
+num_bins_x_tsne_gau <- 21
+num_bins_y_tsne_gau <- 28
+hex_size_tsne_gau <- 0.03
+## non-empty:198
+hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
+                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start
+= "tSNE")
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" =
+"hb_id"))
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
+
+model_object_tsne_gau <- fit_highd_model( training_data = clust_df, 
+                                 nldr_df_with_id = tsne_gau_scaled, 
+                                 x = "tSNE1", y = "tSNE2", 
+                                 num_bins_x = num_bins_x_tsne_gau, 
+                                 num_bins_y = num_bins_y_tsne_gau, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_tsne_gau,
+                                 is_bin_centroid = FALSE,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "tSNE", 
+                                 col_start_highd = "x")
+df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
+df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+## Triangulate bin centroids
+tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
+tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+## Compute 2D distances
+distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+# distance_plot <- plot_dist(distance_tsne_gau) +
+  # ylab(expression(d^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+        # axis.title = element_text(size = 12))
+
+# distance_plot
+## To find the benchmark value
+benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col =
+"distance")
+
+benchmark_tsne_gau <- 0.17
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, 
+                                          benchmark_value = benchmark_tsne_gau, 
+                                          tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size =
+3)
+trimesh_removed_tsne_gau
+df_all_tsne_gau <- dplyr::bind_cols(clust_df |> dplyr::select(-ID),
+tsne_data_with_hb_id_gau)
+### Define type column
+df <- df_all_tsne_gau |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+df_b <- df_bin_tsne_gau |>
+  dplyr::filter(hb_id %in% df_bin_centroids_tsne_gau$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_tsne_gau$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+df_exe <- dplyr::bind_rows(df_b, df)
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_tsne_gau |>
+  dplyr::filter(distance < benchmark_tsne_gau)
+## Since erase brushing is considerd.
+langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = distance_df_small_edges$from,
+                         lineTo = distance_df_small_edges$to,
+                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)),
+rep(0.7, NROW(df))),
+                         levelColors = c("#6a3d9a", "#33a02c"))
+```
+
 # Conclusion

### 2024-03-31 paper-quollr.tex
Add dataset for application

@@
 }
-
-\begin{verbatim}
-library(quollr)
-library(tibble)
-library(knitr)
-library(kableExtra)
-library(ggplot2)
-
-set.seed(20230531)
-\end{verbatim}
 
@@
 
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-2.pdf}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-2.pdf}
+
 \hypertarget{conclusion}{%

### 2024-03-31 paper-quollr.Rmd
Add about main functions

@@
 library(dplyr)
+library(ggbeeswarm)
 
@@
 <!--Discuss the model can be generated with bin centroids or bin means-->
+
+```{r}
+
+## Decide by looking at MSE plot
+num_bins_x_s_curve <- 6
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.13
+hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
+                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+## remove low-density hexagons
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+hex_full_count_df_loop2 <- hex_grid_with_counts |>
+  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high"))
+hex_full_count_df_filtered <- df_bin_centroids |>
+  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high")) |>
+  dplyr::filter(type == "low")
+p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
+p2_n_rm <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = hex_full_count_df_filtered, color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X",
+size = 0.5)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
+
+model_object <- fit_highd_model( training_data = s_curve_noise_training, 
+                                 nldr_df_with_id = s_curve_noise_umap, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_s_curve, 
+                                 num_bins_y = num_bins_y_s_curve, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_s_curve,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+df_bin_centroids_s_curve <- model_object$df_bin_centroids
+df_bin_s_curve <- model_object$df_bin
+cell_count_plot <- ggplot(df_bin_centroids_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y =
+std_counts)) +
+  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
+  geom_hline(yintercept = 0.2, colour = "#de2d26") +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 7),
+        axis.text.x = element_text(angle = 90))
+benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_s_curve$std_counts, 
+                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
+df_bin_centroids_low <- df_bin_centroids_s_curve |>
+  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
+identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_s_curve, 
+                                      num_bins_x = num_bins_x_s_curve, 
+                                      df_bin_centroids_low = df_bin_centroids_low)
+df_bin_centroids_s_curve <- df_bin_centroids_s_curve |>
+  dplyr::filter(!(hexID %in% identify_rm_bins))
+pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = df_bin_centroids_s_curve, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_equal() +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+rm_centroid_df <- df_bin_centroids_low |> 
+  dplyr::filter(hexID %in% identify_rm_bins)
+p2_n_rm_correct <- ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = hex_full_count_df_filtered[6, ], color = "red", aes(x = c_x, y = c_y, group = hexID, label =
+"X", size = 0.5)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
+## Triangulate bin centroids
+tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
+tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
+## Compute 2D distances
+distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+# 
+# distance_plot
+## To find the benchmark value
+benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
+#distance_plot <- plot_dist(distance_s_curve) +
+#   geom_hline(yintercept = benchmark_s_curve, linetype="solid",
+#                color = "red", size=0.8, alpha = 0.5) +
+#  ylab(expression(d^{(2)})) +
+#  theme(axis.text = element_text(size = 5),
+#        axis.title = element_text(size = 12))
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve,
+tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_s_curve_umap <- trimesh_s_curve_umap +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "")
+trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value =
+benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_s_curve <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), umap_data_with_hb_id)
+
+show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
+```
 

### 2024-03-31 paper-quollr.tex
Add about main functions

@@
 When constructing the 2D model, the user can choose either to fir the 2D model with hexagonal bin centroids or bin means using \texttt{is\_bin\_centroid} argument.
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-1-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-2-1.pdf}
-
 \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-3-1.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-4-1.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-2.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
 
 \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-2.pdf}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-2.pdf}
 

### 2024-03-31 paper-quollr.Rmd
Added best fit for S-curve data in the package

--- 2024-03-31 paper-quollr.Rmd
@@
 
+```{r}
+
+## Decide by looking at MSE plot
+num_bins_x_s_curve <- 8
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.1
+hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
+                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+## remove low-density hexagons
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+p2_n <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
+
+model_object <- fit_highd_model( training_data = s_curve_noise_training, 
+                                 nldr_df_with_id = s_curve_noise_umap_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_s_curve, 
+                                 num_bins_y = num_bins_y_s_curve, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_s_curve,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+df_bin_centroids_s_curve <- model_object$df_bin_centroids
+df_bin_s_curve <- model_object$df_bin
+# cell_count_plot <- ggplot(df_bin_centroids_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y =
+# std_counts)) +
+#   geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
+#   geom_hline(yintercept = 0.2, colour = "#de2d26") +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 7),
+#         axis.text.x = element_text(angle = 90))
+# benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_s_curve$std_counts, 
+#                 probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
+# df_bin_centroids_low <- df_bin_centroids_s_curve |>
+#   dplyr::filter(std_counts <= benchmark_value_rm_lwd)
+# identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_s_curve, 
+#                                       num_bins_x = num_bins_x_s_curve, 
+#                                       df_bin_centroids_low = df_bin_centroids_low)
+# df_bin_centroids_s_curve <- df_bin_centroids_s_curve |>
+#   dplyr::filter(!(hexID %in% identify_rm_bins))
+pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = df_bin_centroids_s_curve, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_equal() +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+## Triangulate bin centroids
+tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
+tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
+## Compute 2D distances
+distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+# 
+# distance_plot
+## To find the benchmark value
+benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
+benchmark_s_curve <- 0.5
+#distance_plot <- plot_dist(distance_s_curve) +
+#   geom_hline(yintercept = benchmark_s_curve, linetype="solid",
+#                color = "red", size=0.8, alpha = 0.5) +
+#  ylab(expression(d^{(2)})) +
+#  theme(axis.text = element_text(size = 5),
+#        axis.title = element_text(size = 12))
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve,
+tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_s_curve_umap <- trimesh_s_curve_umap +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "")
+trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value =
+benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_s_curve <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), umap_data_with_hb_id)
+
+show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
+```
+
 ## Constructing the 2D model
@@
 **Step 4: Assign NLDR data to hexagons**
+
+```{r}
+
+```
 

### 2024-03-31 paper-quollr.tex
Added best fit for S-curve data in the package

@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-3-1.pdf}
-
 \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-4-1.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-2.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
 
 \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-2.pdf}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-2.pdf}
 

### 2024-03-31 paper-quollr.R

@@ -50,10 +50,10 @@ theme_set(theme_linedraw() +
 ## -----------------------------------------------------------------------------
 
 ## Decide by looking at MSE plot
-num_bins_x_s_curve <- 6
+num_bins_x_s_curve <- 8
 num_bins_y_s_curve <- 14
-hex_size_s_curve <- 0.13
-hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap, x = "UMAP1", 
+hex_size_s_curve <- 0.1
+hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
                       y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
                       num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
                       buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
@@ -67,12 +67,8 @@ df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df,
                                              counts_df = counts_df)
 ## remove low-density hexagons
 hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-hex_full_count_df_loop2 <- hex_grid_with_counts |>
-  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high"))
-hex_full_count_df_filtered <- df_bin_centroids |>
-  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high")) |>
-  dplyr::filter(type == "low")
-p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+
+p2_n <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
   #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
@@ -88,24 +84,7 @@ vjust = -0.5),
         legend.text = element_text(size=6)) +
   guides(fill = guide_colourbar(title = "Standardized count")) #+
[…]

### 2024-04-01 paper-quollr.Rmd
Add prediction plots for training and test data

--- 2024-03-31 paper-quollr.Rmd
@@
 library(dplyr)
+library(ggbeeswarm)
+library(patchwork)
 
@@
 
+```{r}
+
+## Decide by looking at MSE plot
+num_bins_x_s_curve <- 8
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.1
+hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
+                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+## remove low-density hexagons
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+centroid_plot <- ggplot(data = all_centroids_df, aes(x = c_x, y = c_y)) +
+  geom_point(color = "#33a02c") +
+  coord_fixed() +
+  theme_void() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6))
+
+full_grid <- ggplot(data = hex_grid, aes(x = x, y = y)) +
+  geom_polygon(fill = "#ffffff", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_fixed() +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6))
+
+std_pts_grid <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2), color = "black", alpha = 1) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_fixed() +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6))
+
+umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
+
+model_object <- fit_highd_model( training_data = s_curve_noise_training, 
+                                 nldr_df_with_id = s_curve_noise_umap_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_s_curve, 
+                                 num_bins_y = num_bins_y_s_curve, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_s_curve,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+df_bin_centroids_s_curve <- model_object$df_bin_centroids
+df_bin_s_curve <- model_object$df_bin
+
+## Triangulate bin centroids
+tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
+tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
+## Compute 2D distances
+distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
+benchmark_s_curve <- 0.5
+
+trimesh_s_curve <- ggplot(df_bin_centroids_s_curve, aes(x = c_x, y = c_y)) +
+  geom_trimesh() +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "")
+
+trimesh_colored_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve,
+tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+
+trimesh_colored_s_curve_umap <- trimesh_colored_s_curve_umap +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "")
+
+trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value =
+benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_s_curve <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), umap_data_with_hb_id)
+
+# show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
+
+### Define type column
+df <- df_all_s_curve |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_s_curve |>
+  dplyr::filter(hb_id %in% df_bin_centroids_s_curve$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_s_curve$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+df_exe <- dplyr::bind_rows(df_b, df)
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_s_curve |>
+  dplyr::filter(distance < benchmark_s_curve)
+## Since erase brushing is considerd.
+tour <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = distance_df_small_edges$from,
+                         lineTo = distance_df_small_edges$to,
+                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(1, NROW(df))),
+                         levelColors = c("#6a3d9a", "#33a02c"))
+```
+
 ## Constructing the 2D model
@@
 <!-- add about main function for hexagonal bining-->
+
+```{r, out.width="100%", fig.cap = "Illustration of key steps of the hexagonal bining algorithm: (a) first, all hexagonal bin centroids are created; (b) second, full hexagon grid is created; and (c) finally, assign the data to hexagons.
+The centroids are used to generate hexagonal coordinates.
+Once the full hexagonal grid is generated, by computing the 2D euclidean distance from each NLDR points to hexagonal bin centroids, the nearest NLDR points are allocated to the nearest hexagoanl bin."}
+
+centroid_plot + full_grid + std_pts_grid +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 3) &
+  theme(plot.tag = element_text(size = 8))
+```
 
@@
 
+## Prediction
+
+`predict_emb()` function is used to predict 2D embedding for a new $p$-D data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
+
+```{r}
+## Prediction for training data
+pred_emb_list_training <- predict_emb(test_data = s_curve_noise_training, 
+                             df_bin_centroids = df_bin_centroids_s_curve, 
+                             df_bin = df_bin_s_curve, type_NLDR = "UMAP")
+pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list_training))
+
+pred_points_training <- ggplot(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2)) +
+  geom_point() +
+  geom_point(data = pred_df_training, aes(x = pred_UMAP_1, y = pred_UMAP_2), colour = "#de2d26") +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+## Prediction for test data
+pred_emb_list_test <- predict_emb(test_data = s_curve_noise_test, 
+                             df_bin_centroids = df_bin_centroids_s_curve, 
+                             df_bin = df_bin_s_curve, type_NLDR = "UMAP")
+pred_df_test <- as.data.frame(do.call(cbind, pred_emb_list_test))
+
+pred_points_test <- ggplot(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2)) +
+  geom_point() +
+  geom_point(data = pred_df_test, aes(x = pred_UMAP_1, y = pred_UMAP_2), colour = "#de2d26") +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r, out.width="100%", fig.cap = "Predicted 2D embedding overlaid with UMAP embedding of S-curve training data: (a) prediction of 2D embeddings for S-curve training data, and (b) prediction of 2D embeddings for S-curve test data.
+Once the high-D model is generated, first, for the new data points the nearest high-D mappings of hexagonal bin centroids find by computing Euclidean distance.
+Then, mapping back the 2D hexagonal bin centroids gives the prediction for the new data points."}
+
+pred_points_training + pred_points_test +
+  plot_layout(ncol = 2) &
+  theme(plot.tag = element_text(size = 8))
+```
+
 ## Compute residuals and Mean Square Error (MSE)
@@
 
+```{r, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges.
+The edges which has the length greater than $0.5$ are assigned as long edges.
+To obtain smooth representation in 2D space, the long edges are removed."}
+
+trimesh_s_curve + trimesh_colored_s_curve_umap + trimesh_removed_s_curve_umap +
+  plot_layout(ncol = 3) &
+  theme(plot.tag = element_text(size = 8))
+```
+
 ### p-D model visualization
@@
 The $p$-D model overlaid on data is visualized by the function `show_langevitour()`. This is a useful dynamic visualization result to validate visually whether the model fits the data well or not.
+
+<!--need to add screenshots of the frames and add the link of the video-->
+```{r}
+tour
+```
 
@@
 -   basic example-->
-
-## Prediction
-
-`predict_emb()` function is used to predict 2D embedding for a new data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
 

### 2024-04-01 paper-quollr.tex
Add prediction plots for training and test data

--- 2024-03-31 paper-quollr.tex
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-1-1.pdf}
-
 \hypertarget{constructing-the-2d-model}{%
@@
 The mains steps for the hexagonal binning algorithm can be executed by the main function \texttt{hex\_binning()}, or can be run separately for more flexibility.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-2-1} \caption{Illustration of key steps of the hexagonal bining algorithm: (a) first, all hexagonal bin centroids are created; (b) second, full hexagon grid is created; and (c) finally, assign the data to hexagons.
+The centroids are used to generate hexagonal coordinates.
+Once the full hexagonal grid is generated, by computing the 2D euclidean distance from each NLDR points to hexagonal bin centroids, the nearest NLDR points are allocated to the nearest hexagoanl bin.}\label{fig:unnamed-chunk-2}
+\end{figure}
 
@@
 
+\hypertarget{prediction}{%
+\subsection{Prediction}\label{prediction}}
+
+\texttt{predict\_emb()} function is used to predict 2D embedding for a new \(p\)-D data point using the fitted model.
+This function is useful to predict 2D embedding irrespective of the NLDR technique.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-4-1} \caption{Predicted 2D embedding overlaid with S-curve training UMAP data: (a) Prediction of 2D embeddings for S-curve training data, and (b) Prediction of 2D embeddings for S-curve test data.
+Once the high-D model is generated, first, for the new data points the nearest high-D mappings of hexagonal bin centroids find by computing Euclidean distance.
+Then, mapping back the 2D hexagonal bin centroids gives the prediction for the new data points.}\label{fig:unnamed-chunk-4}
+\end{figure}
+
 \hypertarget{compute-residuals-and-mean-square-error-mse}{%
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1} \caption{Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges.
+The edges which has the length greater than $0.5$ are assigned as long edges.
+To obtain smooth representation in 2D space, the long edges are removed.}\label{fig:unnamed-chunk-5}
+\end{figure}
+
 \hypertarget{p-d-model-visualization}{%
@@
 
-\hypertarget{prediction}{%
-\subsection{Prediction}\label{prediction}}
-
-\texttt{predict\_emb()} function is used to predict 2D embedding for a new data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-2.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-2.pdf}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-2.pdf}
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-2.pdf}
 

### 2024-04-01 paper-quollr.R

@@ -16,6 +16,7 @@ library(kableExtra)
 library(ggplot2)
 library(dplyr)
 library(ggbeeswarm)
+library(patchwork)
 
 set.seed(20240110)
 
@@ -68,11 +69,23 @@ df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df,
 ## remove low-density hexagons
 hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
 
-p2_n <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+centroid_plot <- ggplot(data = all_centroids_df, aes(x = c_x, y = c_y)) +
+  geom_point(color = "#33a02c") +
+  coord_fixed() +
+  theme_void() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5,
+vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) 
+
+full_grid <- ggplot(data = hex_grid, aes(x = x, y = y)) +
+  geom_polygon(fill = "#ffffff", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_fixed() +
   theme_void() +
   theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
 vjust = -0.5),
@@ -81,15 +94,23 @@ vjust = -0.5),
         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
[…]

### 2024-04-01 paper-quollr.Rmd
Add vis ref

@@
 abstract: >
-  An abstract of less than 150 words.
+  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional ($p$-D) data.
+However, their complexity and parameter choices may lead to distrustful or misleading results.
+The R package \CRANpkg{quollr} is developed as a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.
+Clustering data from \CRANpkg{cardinalR} package, is used to illustrate the algorithm and its use within the paackage.
 draft: true
@@
 type: package
+creative_commons: CC BY
+date: "`r Sys.Date()`"
 preamble: >
@@
 output: 
-  rjtools::rjournal_article:
-    self_contained: yes
-    toc: no
+ rjtools::rjournal_web_article:
+    css: "style.css"
+    keep_md: true
 bibliography: RJreferences.bib
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE, 
-                      message = FALSE,
-                      echo = FALSE)
+knitr::opts_chunk$set(
+  echo = FALSE, 
+  cache=FALSE, 
+  message=FALSE, 
+  warning=FALSE)
 
@@
 library(patchwork)
+library(Rtsne)
 
@@
 
+<!-- research gap: add about hexbin pkg, and emphasize that in our package provide regular hexagons-->
+<!-- objective: introduce a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.--> 
 <!--intro with S-curve with 5 methods-->
@@
 <!--explain algorithm steps in summary-->
+Our algorithm mainly consists of two steps: (1) construct the model in 2D space, and (2) lift the model into high dimensions.
+Constructing the model in 2D space involves: (1) scaling NLDR data, (2) binning data, (3) obtain bin centroids/ means, and (4) triangulate bin centroids/ means.
+The implementation of these steps are described in detail in the following section.
 
@@
 
-```install.packages("quollr")```
+```r
+install.packages("quollr")
+```
 
@@
 
-```devtools::install_github("JayaniLakshika/quollr")```
+```r
+devtools::install_github("JayaniLakshika/quollr")
+```
 
@@
 When constructing the 2D model, the user can choose either to fir the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
+
+```{r, echo=TRUE, eval=FALSE}
+model_object <- fit_highd_model( training_data = s_curve_noise_training, 
+                                 nldr_df_with_id = s_curve_noise_umap_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_s_curve, 
+                                 num_bins_y = num_bins_y_s_curve, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_s_curve,
+                                 is_bin_centroid = TRUE,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+```
 
@@
 
+```{r, echo=TRUE, eval=FALSE}
+s_curve_noise_umap_scaled <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2")
+```
+
 <!-- add plot before and after scaling NLDR data-->
@@
 
+**Step 1: Compute hexagonal grid configurations**
+
+```{r, echo=TRUE, eval=FALSE}
+bin_list <- calc_bins(data = s_curve_noise_umap_scaled, x = "UMAP1", 
+                      y = "UMAP2", hex_size = NA, buffer_x = NA, buffer_y = NA)
+
+num_bins_x <- bin_list$num_x
+num_bins_y <- bin_list$num_y
+```
+
 The mains steps for the hexagonal binning algorithm can be executed by the main function `hex_binning()`, or can be run separately for more flexibility.
-<!-- add about main function for hexagonal bining-->
+<!-- add about main function for hexagonal binning-->
+
+```{r, echo=TRUE, eval=FALSE}
+hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x, 
+                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = NA, 
+                      col_start = "UMAP")
+
+```
 
@@
 
-**Step 1: Compute hexagonal grid configurations**
+**Step 2: Generate the hexagonal bin centroids**
 
-**Step 2: Generate the hexagonal bin centroids**
+```{r, echo=TRUE, eval=FALSE}
+centroid_list <- gen_centroids(data = s_curve_noise_umap_scaled, x = "UMAP1", 
+                               y = "UMAP2", num_bins_x = num_bins_x,
+                               num_bins_y = num_bins_y, x_start = NA,
+                               y_start = NA, buffer_x = NA, buffer_y = NA, 
+                               hex_size = NA)
+
+all_centroids_df <- as.data.frame(do.call(cbind, centroid_list))
+```
 
@@
 
+```{r, echo=TRUE, eval=FALSE}
+all_hex_coordinates_list <- gen_hex_coord(centroids_df = all_centroids_df,
+                                            hex_size = NA)
+```
+
 **Step 4: Assign NLDR data to hexagons**
+
+```{r, echo=TRUE, eval=FALSE}
+df_with_hb_id_list <- assign_data(data = s_curve_noise_umap_scaled |> 
+                                    dplyr::select(-ID), 
+                                  centroid_df = all_centroids_df, 
+                                  col_start = "UMAP")
+umap_data_with_hb_id <- as.data.frame(do.call(cbind, df_with_hb_id_list))
+```
+
+```{r, echo=TRUE, eval=FALSE}
+std_counts_list <- compute_std_counts(data_hex_id = df_with_hex_id)
+counts_df <- as.data.frame(do.call(cbind, std_counts_list))
+```
 
@@
 
+```{r, echo=TRUE, eval=FALSE}
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+```
+
+```{r, echo=TRUE, eval=FALSE}
+df_bin_means <- extract_hexbin_mean(nldr_df_with_hex_id = umap_with_hb_id,
+                                        counts_df = counts_df)
+```
+
 ### Triangulate bin centroids/ means
@@
 The `tri_bin_centroids()` is used to do triangulate bin centroids/ bin means. Then, `gen_edges()` compute the edges to obtain the triangular mesh which is the 2D model.
+
+```{r, echo=TRUE, eval=FALSE}
+tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
+tr_from_to_df <- gen_edges(tri_object = tr1_object)
+```
 
@@
 
-## Tuning model parameters
+```{r, echo=TRUE, eval=FALSE}
+df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID),
+                           umap_data_with_hb_id)
+
+df_bin <- avg_highd_data(data = df_all, col_start = "x")
+```
+
+## Model parameters
 
@@
 On the other hand, `find_lg_benchmark()` function is used to compute the threshold for removing long edges.
+
+```{r, echo=TRUE, eval=FALSE}
+df_bin_centroids_low <- df_bin_centroids |> dplyr::filter(std_counts <= 0.43)
+
+low_dens_hex <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids, 
+                  num_bins_x = num_bins_x, 
+                  df_bin_centroids_low = df_bin_centroids_low)
+```
+
+```{r, echo=TRUE, eval=FALSE}
+distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, start_x = "x_from", 
+                           start_y = "y_from", end_x = "x_to", end_y = "y_to",
+                           select_vars = c("from", "to", "distance"))
+
+benchmark_val <- find_lg_benchmark(distance_edges = distance_df, distance_col = "distance")
+```
 
@@
 
+```{r, echo=TRUE}
+gen_summary(test_data = s_curve_noise_training, prediction_df = pred_df_training,
+df_bin = df_bin_s_curve, col_start = "x")
+```
+
+```{r, echo=TRUE}
+gen_summary(test_data = s_curve_noise_test, prediction_df = pred_df_test,
+df_bin = df_bin_s_curve, col_start = "x")
+```
+
 <!--add a plot with total bins Vs MSE, non-empty bins Vs Error-->
@@
 To visualize the 2D model, mainly three functions are used.
-As shown in Figure 1, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`.
+As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`.
 After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges.
-As shown in Figure 1, the small and long edges are colored by black and red respectively.
-Following this, `vis_rmlg_mesh()` is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure 1).
+As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively.
+Following this, `vis_rmlg_mesh()` is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure \@ref(fig:mesh-plots)c).
 In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in 2D.
 
-```{r, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges.
+```{r mesh-plots, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges.
 The edges which has the length greater than $0.5$ are assigned as long edges.
@@
 
+```{r, echo=TRUE, eval=FALSE}
+tour <- show_langevitour(df = df_all, df_b = df_bin, 
+                 df_b_with_center_data = df_bin_centroids, benchmark_value = 0.5, 
+                 distance = distance_df, distance_col = "distance", 
+                 use_default_benchmark_val = FALSE, col_start = "x")
+```
+
 <!--need to add screenshots of the frames and add the link of the video-->
-```{r}
+```{r tour-s-curve}
 tour
@@
 # Application
+
+We illustrate the use of the algorithm by applying it to `clusters_different_shapes_diff_num_points` dataset in `cardinalR` package.
+The tSNE is used in the example for the demonstration (see Figure \@ref(fig:tsne-clust)).
 
@@
 
+<!--need to add screen shots of the original data-->
+
 ```{r}
@@
 ```{r}
-library(Rtsne)
-
 tSNE_fit <- clust_df |>
@@
         axis.title = element_text(size = 7))
+```
 
+```{r tsne-clust, out.width="100%", fig.cap="tSNE representation of data with perplexity: 39. Is this representation accurately capture the structure in high dimensions?"}
 tSNE_df_plot
 ```
+
+## Selecting parameter values for hexagonal binning
+
+We previously introduced three key parameters, *hex_size*, *x_start*, and *y_start*, in our explanation of the algorithm.
+While their best choice is unknown, a visualization tool enables their tuning such that user can select reasonable values.
+
+Consider the relationships between *total number of bins*, *starting point*, and MSE. Increase of *total number of bins* usually reduces the MSE.
+
+<!--hex_size, starting point (x_start, y_start)-->
+<!--talk about to use bin centroids or bin means-->
 
@@
         axis.text = element_text(size = 7)) +
-  ylab("MSE") +
+  ylab("log(MSE)") +
   xlab("total number of bins")
+```
 
+```{r, fig.cap="Parameter tuning plots, where the best choice of parameter at a large drop in MSE. Here, these are at *total number of bins* = 180."}
 mse_clust_plot
 ```
+
+## Selecting parameter values for the model
+
+<!--benchmark value to remove low density hexagons, long edges-->
+
+## Fit the model
+
+To perform the algorithm on the clustering data, we first scale the tSNE data.
+We then let *number of bins along the x-axis* be $21$, *number of bins along the y-axis* be $28$, and *hex_size* be 0.03; see [Selecting parameter values] for further discussion and guidance regarding selection of these choices.
 

### 2024-04-01 paper-quollr.tex
Add vis ref

@@
 \abstract{%
-An abstract of less than 150 words.
+Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional (\(p\)-D) data.
+However, their complexity and parameter choices may lead to distrustful or misleading results.
+The R package \CRANpkg{quollr} is developed as a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.
+Clustering data from \CRANpkg{cardinalR} package, is used to illustrate the algorithm and its use within the paackage.
 }
@@
 
-\hypertarget{tuning-model-parameters}{%
-\subsection{Tuning model parameters}\label{tuning-model-parameters}}
+\hypertarget{model-parameters}{%
+\subsection{Model parameters}\label{model-parameters}}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-4-1} \caption{Predicted 2D embedding overlaid with S-curve training UMAP data: (a) Prediction of 2D embeddings for S-curve training data, and (b) Prediction of 2D embeddings for S-curve test data.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-4-1} \caption{Predicted 2D embedding overlaid with UMAP embedding of S-curve training data: (a) prediction of 2D embeddings for S-curve training data, and (b) prediction of 2D embeddings for S-curve test data.
 Once the high-D model is generated, first, for the new data points the nearest high-D mappings of hexagonal bin centroids find by computing Euclidean distance.
@@
 As a Goodness of fit statistics for the model, \texttt{gen\_summary()} is used to compute residuals and MSE. For the fitted model, how residuals and MSE is changing is shown in Figure 2.
+
+\begin{verbatim}
+gen_summary(test_data = s_curve_noise_training, prediction_df = pred_df_training,
+df_bin = df_bin_s_curve, col_start = "x")
+\end{verbatim}
+
+\begin{verbatim}
+#> $error
+#> [1] 35.79497
+#> 
+#> $mse
+#> [1] 0.1114669
+#> 
+#> $aic
+#> [1] -815.8643
+\end{verbatim}
+
+\begin{verbatim}
+gen_summary(test_data = s_curve_noise_test, prediction_df = pred_df_test,
+df_bin = df_bin_s_curve, col_start = "x")
+\end{verbatim}
+
+\begin{verbatim}
+#> $error
+#> [1] 15.4298
+#> 
+#> $mse
+#> [1] 0.1546555
+#> 
+#> $aic
+#> [1] 9.352806
+\end{verbatim}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1} \caption{Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-7-1} \caption{Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges.
 The edges which has the length greater than $0.5$ are assigned as long edges.
-To obtain smooth representation in 2D space, the long edges are removed.}\label{fig:unnamed-chunk-5}
+To obtain smooth representation in 2D space, the long edges are removed.}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{paper-quollr_files/figure-latex/tour-s-curve-1.pdf}
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf}
+We illustrate the use of the algorithm by applying it to \texttt{clusters\_different\_shapes\_diff\_num\_points} dataset in \texttt{cardinalR} package.
+The tSNE is used in the example for the demonstration (see Figure \ref{fig:tsne-clust}).
 
@@
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-11-1.pdf}
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/tsne-clust-1} \caption{tSNE representation of data with perplexity: 39.
+Is this representation accurately capture the structure in high dimensions?}\label{fig:tsne-clust}
+\end{figure}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-12-2.pdf}
+\hypertarget{selecting-parameter-values-for-hexagonal-binning}{%
+\subsection{Selecting parameter values for hexagonal binning}\label{selecting-parameter-values-for-hexagonal-binning}}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-2.pdf}
+We previously introduced three key parameters, \emph{hex\_size}, \emph{x\_start}, and \emph{y\_start}, in our explanation of the algorithm.
+While their best choice is unknown, a visualization tool enables their tuning such that user can select reasonable values.
+
+Consider the relationships between \emph{total number of bins}, \emph{starting point}, and MSE. Increase of \emph{total number of bins} usually reduces the MSE.
+
+\begin{figure}
+\centering
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-13-1.pdf}
+\caption{\label{fig:unnamed-chunk-13}Parameter tuning plots, where the best choice of parameter at a large drop in MSE. Here, these are at \emph{total number of bins} = 180.}
+\end{figure}
+
+\hypertarget{selecting-parameter-values-for-the-model}{%
+\subsection{Selecting parameter values for the model}\label{selecting-parameter-values-for-the-model}}
+
+\hypertarget{fit-the-model}{%
+\subsection{Fit the model}\label{fit-the-model}}
+
+To perform the algorithm on the clustering data, we first scale the tSNE data.
+We then let \emph{number of bins along the x-axis} be \(21\), \emph{number of bins along the y-axis} be \(28\), and \emph{hex\_size} be 0.03; see {[}Selecting parameter values{]} for further discussion and guidance regarding selection of these choices.
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-14-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-14-2.pdf}
+
+\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-15-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-15-2.pdf}
 

### 2024-04-01 paper-quollr.Rmd
Add paper sections to intro and add bechnark argument to vis

@@
 
+The paper is organized as follows.
+Section **Methodology** details the main steps of the algorithm.
+In next section, introduces the implementation of `quollr` package on CRAN, including demonstration of the package's key functions and visualization capabilities.
+We illustrate the algorithm's functionality to study clustering data set in **Application** section, and describe a visual heuristic to describe parameter selection.
+Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our algorithm.
+
 # Methodology
@@
 
-trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value =
-benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
 trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
@@
   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+trimesh_removed_s_curve_umap_0.3 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.3, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_removed_s_curve_umap_0.3 <- trimesh_removed_s_curve_umap_0.3 +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+trimesh_removed_s_curve_umap_0.8 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.8, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+trimesh_removed_s_curve_umap_0.8 <- trimesh_removed_s_curve_umap_0.8 +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 
@@
 
+<!--discuss about default settings-->
+
 Mainly there are two model parameter need to consider: benchmark value to remove the low-density hexagons, and benchmark value to remove the long edges.
@@
 In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in 2D.
+Using small value of `benchmark_value`, will produce a triangular mesh with missing data structure; for example `benchmark_value = 0.3` shows two clusters rather than continuous structure, while `benchmark_value = 0.8` creates long edges and mislead the data structure in $p-D$ space.
 
@@
   plot_layout(ncol = 3) &
+  theme(plot.tag = element_text(size = 8))
+```
+
+```{r mesh-diffplots, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space with different benchmark value applied to remove long edges: (a) benchmark_value = 0.3, and (b) benchmark_value = 0.8.
+Small benchmark values will generate misleading data structures in 2D space, while; large bechmark value will generate more long edges in $p-D$ and miinterpret the $p-D$ data structure."}
+
+trimesh_removed_s_curve_umap_0.3 + trimesh_removed_s_curve_umap_0.8 +
+  plot_layout(ncol = 2) &
   theme(plot.tag = element_text(size = 8))

### 2024-04-02 paper-quollr.Rmd
Added dependency info

@@
 abstract: >
-  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional ($p$-D) data.
-However, their complexity and parameter choices may lead to distrustful or misleading results.
-The R package \CRANpkg{quollr} is developed as a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.
-Clustering data from \CRANpkg{cardinalR} package, is used to illustrate the algorithm and its use within the paackage.
+  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional ($p-D$) data in low-diemnsioanl space.
+However, their complexity and hyperparameter choices may lead to distrustful or misleading results.
+The R package \CRANpkg{quollr} is developed as a new tool to help to determine which method, which hyperparameter choice provide the most accurate representation of $p-D$ data.
+Clustering data from \CRANpkg{cardinalR} package, is used to illustrate the algorithm and its use within the package.
 draft: true
@@
 bibliography: RJreferences.bib
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 ```{r setup, include=FALSE}
+options(repos = "https://cloud.r-project.org") ## set up CRAN mirror
 knitr::opts_chunk$set(
@@
   message=FALSE, 
-  warning=FALSE)
+  warning=FALSE,
+  out.width = "100%")
 
@@
 library(Rtsne)
+library(readr)
+library(tools)
 
@@
 
+```{r}
+tSNE_data <- read_rds("data/s_curve/s_curve_tsne_27.rds")
+
+plot_list1 <- tSNE_data |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2)) +
+  geom_point(alpha=0.1, size=1, colour="#e41a1c") +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
+
+#(n-neighbors: 50)
+plot_list2 <- UMAP_data |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour="#377eb8") +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#(knn: 5)
+PHATE_data <- read_rds(file = "data/s_curve/s_curve_phate.rds")
+
+plot_list3 <- PHATE_data |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2)) +
+  geom_point(alpha=0.1, size=1, colour="#4daf4a") +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+TriMAP_data <- read_rds(file = "data/s_curve/s_curve_trimap.rds")
+
+#(n-inliers: 5, \n n-outliers: 4, n-random: 3)
+plot_list4 <- TriMAP_data |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2)) +
+  geom_point(alpha=0.1, size=1, colour="#984ea3") +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+PaCMAP_data <- read_rds(file = "data/s_curve/s_curve_pacmap.rds")
+
+#(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
+plot_list5 <- PaCMAP_data |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2)) +
+  geom_point(alpha=0.1, size=1, colour="#ff7f00") +
+  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r fig-nldervis, fig.cap="2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 27), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?"}
+
+plot_list1 + plot_list2 + plot_list3 + plot_list4 + plot_list5 +
+  plot_layout(ncol=5)
+```
+
 The paper is organized as follows.
@@
 ```
+
+## Package dependencies
+
+Understanding the dependencies of the "quollr" package is essential for smooth operation and error prevention.
+The following dependencies refer to the other R packages that `quollr` relies on to execute its functions effectively.
+
+```{r}
+package_dependencies("quollr")
+```
+
+## Usage
 

### 2024-04-02 paper-quollr.Rmd
Add intro to the paper

@@
 
-```{r}
-tSNE_data <- read_rds("data/s_curve/s_curve_tsne_27.rds")
-
-plot_list1 <- tSNE_data |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2)) +
-  geom_point(alpha=0.1, size=1, colour="#e41a1c") +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
-
-#(n-neighbors: 50)
-plot_list2 <- UMAP_data |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2)) +
-  geom_point(alpha=0.1, size=1, colour="#377eb8") +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#(knn: 5)
-PHATE_data <- read_rds(file = "data/s_curve/s_curve_phate.rds")
-
-plot_list3 <- PHATE_data |>
-  ggplot(aes(x = PHATE1,
-             y = PHATE2)) +
-  geom_point(alpha=0.1, size=1, colour="#4daf4a") +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-TriMAP_data <- read_rds(file = "data/s_curve/s_curve_trimap.rds")
-
-#(n-inliers: 5, \n n-outliers: 4, n-random: 3)
-plot_list4 <- TriMAP_data |>
-  ggplot(aes(x = TriMAP1,
-             y = TriMAP2)) +
-  geom_point(alpha=0.1, size=1, colour="#984ea3") +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-PaCMAP_data <- read_rds(file = "data/s_curve/s_curve_pacmap.rds")
-
-#(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
-plot_list5 <- PaCMAP_data |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2)) +
-  geom_point(alpha=0.1, size=1, colour="#ff7f00") +
-  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r fig-nldervis, fig.cap="2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 27), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?"}
-
-plot_list1 + plot_list2 + plot_list3 + plot_list4 + plot_list5 +
-  plot_layout(ncol=5)
-```
+This paper presents the R package, `quollr` which introduce a new tool in determining which NLDR technique and which hyperparameter choice gives most accurate representation of $p-D$ data.
+The methodology of the algorithm is explained in *cite the methodology paper*.
+Furthermore, the `quollr` package enables users to perform hexagonal binning [@dan2023], resulting in the generation of regular hexagons.
+The software is available from the Comphrehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).
 
 The paper is organized as follows.
-Section **Methodology** details the main steps of the algorithm.
 In next section, introduces the implementation of `quollr` package on CRAN, including demonstration of the package's key functions and visualization capabilities.
@@
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our algorithm.
-
-# Methodology
-
-<!--explain algorithm steps in summary-->
-Our algorithm mainly consists of two steps: (1) construct the model in 2D space, and (2) lift the model into high dimensions.
-Constructing the model in 2D space involves: (1) scaling NLDR data, (2) binning data, (3) obtain bin centroids/ means, and (4) triangulate bin centroids/ means.
-The implementation of these steps are described in detail in the following section.
 

### 2024-04-02 paper-quollr.Rmd
Added main conclusion context

@@
 
+This paper presents the R package `quollr` to develop a way to take the fitted model, as represented by the positions of points in 2D, and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.
+
+The paper includes a clustering example to illustrate how `quollr` is useful to assess which NLDR technique and which hyperparameter choice gives the most accurate representation.
+In addition, how to select parameters for hexagonal binning and fitting model are explained.
+
+Possible future improvements would be...<!--assess the preservation of local and glocal structure w.r.t 2D and high-D distance comparison-->
+
+This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of $p-D$ data.
+
 # Acknowledgements

### 2024-04-04 paper-quollr.Rmd
Added some changes to constructing 2D models

@@
 
-Understanding the dependencies of the "quollr" package is essential for smooth operation and error prevention.
+Understanding the dependencies of the `quollr` package is essential for smooth operation and error prevention.
 The following dependencies refer to the other R packages that `quollr` relies on to execute its functions effectively.
@@
 The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
-When constructing the 2D model, the user can choose either to fir the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
+When constructing the 2D model, the user can choose either to fit the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
 
@@
 
+Constructing the 2D model mainly contains (1) preprocessing, (2) binning NLD data, (3) obtain bin centroids/ means, and (4) triangulate bin centroids/ means.
+
 ### Preprocessing
@@
 The function `gen_scaled_data()` is used to prepare the NLDR data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
-
-```{r, echo=TRUE, eval=FALSE}
-s_curve_noise_umap_scaled <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", y = "UMAP2")
+The user can set up the ratio of the width and height of the hexagon using the `hex_ratio` argument.
+
+```{r, echo=TRUE, eval=FALSE}
+s_curve_noise_umap_scaled <- gen_scaled_data(data = s_curve_noise_umap, 
+                                             x = "UMAP1", y = "UMAP2", 
+                                             hex_ratio = NA)
+
 ```
@@
 <!-- add plot before and after scaling NLDR data-->
+<!-- also can discuss about the tuning hex_ratio-->
 
@@
 
+In binning data step, the objective is to perform hexagonal binning and assign NLDR data to regualr hexagons.
+
 **Step 1: Compute hexagonal grid configurations**
@@
 bin_list <- calc_bins(data = s_curve_noise_umap_scaled, x = "UMAP1", 
-                      y = "UMAP2", hex_size = NA, buffer_x = NA, buffer_y = NA)
+                      y = "UMAP2", hex_size = 0.1, buffer_x = NA, buffer_y = NA)
 
@@
 
-<!--
-## Usage
-
--   dependencies
-
--   basic example-->
-
 ## Tests
@@
 All functions have tests written and implemented using the \CRANpkg{testthat} [@testthat] in R.
+
+These tests illuminated the issues that allowed us to make meaningful changes and understand some pitfalls of the package.
+
+<!--in data processing, minus values, NA, Inf values for aspect_ratio or hex_ratio-->
+<!-- to check length of list objects-->
+<!-- add tests to find NAs, inf values are in NLDR data-->
 

### 2024-04-08 paper-quollr.Rmd
Find the best fit for the application dataset

@@
 ```{r, echo=FALSE}
-clusters_different_shapes_diff_num_points <- function(sample_size = 400, with_seed = NULL, cluster_size_vec = c(50, 50, 50, 50, 100, 100), num_gussian_clusters = 4, num_non_gaussian_clusters = 2,
-                                      cluster_sd_gau = 0.05, cluster_sd_non_gau = 0.1, num_dims = 7, a = 2, b = 4) {
+clusters_different_shapes_diff_num_points <- function(sample_size = 400, with_seed = NULL, cluster_size_vec = c(50, 50, 50, 50, 100, 100), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.05, cluster_sd_non_gau = 0.1, num_dims = 7, a = 2, b = 4) {
 
@@
 ```{r}
-clust_df <- clusters_different_shapes_diff_num_points(sample_size = 1500, with_seed = NULL, cluster_size_vec = c(250, 150, 150, 150, 350, 450), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.2, cluster_sd_non_gau = 0.3, num_dims = 7, a = 2, b = 4)
+clust_df <- clusters_different_shapes_diff_num_points(sample_size = 1500, with_seed = NULL, cluster_size_vec = c(250, 150, 150, 150, 350, 450), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.2, cluster_sd_non_gau = 0.3, num_dims = 5, a = 2, b = 4)
 
@@
   dplyr::mutate(ID = 1:NROW(tSNE_df))    
-## tSNE
-hex_size_vec <- seq(0.02, 2, by = 0.01)
+## Model with bin centroid
+hex_size_vec <- seq(0.03, 2, by = 0.01)
 vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
@@
 ### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_clust <- dplyr::bind_rows(not_dupli_df, duplicate_df) 
+mse_df_clust <- dplyr::bind_rows(not_dupli_df, duplicate_df) |>
+  dplyr::mutate(type = "bin centroids")
 ```
@@
 ```{r}
+
 mse_clust_plot <- ggplot(mse_df_clust, aes(x = num_bins,
@@
   geom_line() +
-   geom_vline(xintercept = 588, linetype="solid",
+   geom_vline(xintercept = 238, linetype="solid",
                 color = "red", size=0.8, alpha = 0.5) +
@@
 
-```{r, fig.cap="Parameter tuning plots, where the best choice of parameter at a large drop in MSE. Here, these are at *total number of bins* = 180."}
+```{r, fig.cap="Parameter tuning plots, where the best choice of parameter at a large drop in MSE. Here, these are at *total number of bins* = 238."}
 mse_clust_plot
@@
 To perform the algorithm on the clustering data, we first scale the tSNE data.
-We then let *number of bins along the x-axis* be $21$, *number of bins along the y-axis* be $28$, and *hex_size* be 0.03; see [Selecting parameter values] for further discussion and guidance regarding selection of these choices.
+We then let *number of bins along the x-axis* be $14$, *number of bins along the y-axis* be $27$, and *hex_size* be $0.05$; see [Selecting parameter values] for further discussion and guidance regarding selection of these choices.
 
@@
 ## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 21
-num_bins_y_tsne_gau <- 28
-hex_size_tsne_gau <- 0.03
-## non-empty:198
+num_bins_x_tsne_gau <- 14
+num_bins_y_tsne_gau <- 27
+hex_size_tsne_gau <- 0.05
+x_start_tsne_gau <- -0.0481
+y_start_tsne_gau <- 0.00833
+## non-empty:80
 hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
                       y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = x_start_tsne_gau, y_start = y_start_tsne_gau,
                       buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "tSNE")
@@
                                  num_bins_y = num_bins_y_tsne_gau, 
-                                 x_start = NA, y_start = NA, 
+                                 x_start = x_start_tsne_gau, y_start = y_start_tsne_gau, 
                                  buffer_x = NA, buffer_y = NA, 
@@
 benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+
 # ggplot() +
@@
                          lineTo = distance_df_small_edges$to,
-                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.7, NROW(df))),
+                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.5, NROW(df))),
                          levelColors = c("#6a3d9a", "#33a02c"))
-
-```
-
-<!-- model with bin means-->
-```{r}
-## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 21
-num_bins_y_tsne_gau <- 28
-hex_size_tsne_gau <- 0.03
-## non-empty:198
-hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
-                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start
-= "tSNE")
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" =
-"hb_id"))
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
-
-model_object_tsne_gau <- fit_highd_model( training_data = clust_df, 
-                                 nldr_df_with_id = tsne_gau_scaled, 
-                                 x = "tSNE1", y = "tSNE2", 
-                                 num_bins_x = num_bins_x_tsne_gau, 
-                                 num_bins_y = num_bins_y_tsne_gau, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_tsne_gau,
-                                 is_bin_centroid = FALSE,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "tSNE", 
-                                 col_start_highd = "x")
-df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
-df_bin_tsne_gau <- model_object_tsne_gau$df_bin
-## Triangulate bin centroids
-tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
-tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
-## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-# distance_plot <- plot_dist(distance_tsne_gau) +
-  # ylab(expression(d^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-        # axis.title = element_text(size = 12))
-
-# distance_plot
-## To find the benchmark value
-benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col =
-"distance")
-
-benchmark_tsne_gau <- 0.17
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
-trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, 
-                                          benchmark_value = benchmark_tsne_gau, 
-                                          tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size =
-3)
-trimesh_removed_tsne_gau
-df_all_tsne_gau <- dplyr::bind_cols(clust_df |> dplyr::select(-ID),
-tsne_data_with_hb_id_gau)
-### Define type column
-df <- df_all_tsne_gau |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
-df_b <- df_bin_tsne_gau |>
-  dplyr::filter(hb_id %in% df_bin_centroids_tsne_gau$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_tsne_gau$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-df_exe <- dplyr::bind_rows(df_b, df)
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_tsne_gau |>
-  dplyr::filter(distance < benchmark_tsne_gau)
-## Since erase brushing is considerd.
-langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = distance_df_small_edges$from,
-                         lineTo = distance_df_small_edges$to,
-                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)),
-rep(0.7, NROW(df))),
-                         levelColors = c("#6a3d9a", "#33a02c"))
-```
+```
+
+What can learn from the model:
+
+- There are five clusters. Four of them are Gaussian and two of them are long 2D clusters. Gaussian clusters are close to each other while others are far away.
+
+- Model preserved local and global structure.
+
+- Problem: Couldn't capture the variance within the clusters.
+
+- There is one edge that is not connected in 2D triangular mesh.
 

### 2024-04-09 paper-quollr.Rmd
Added more descriptions to the algorithm steps

@@
 
-The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `s_curve_noise_training` and `s_curve_noise_umap`.
+The following demonstration of the package's functionality assumes `quollr` has been loaded.
+We also want to load the built-in data sets `s_curve_noise_training`, `s_curve_noise_test` and `s_curve_noise_umap`.
+
+`s_curve_noise_training` is a $3-D$ S-curve dataset with additional four noise dimensions which is used to train the model.
+`s_curve_noise_test` is used for testing.
+`s_curve_noise_umap` is the UMAP $2-D$ embedding for `s_curve_noise_training` dataset.
+Each dataset contains a unique ID column.
 
@@
 When constructing the 2D model, the user can choose either to fit the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
+
+If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
+For example, if the user wishes to use already binning data, the `extract_hexbin_centroids()` function can be used directly.
+
+The number of bins along the x and y axes, hexagonal grid starting point, buffer size along the x and y axes, hexagon size, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model`, if they are not provided.
+They are created as they are needed throughout the following example.
 
@@
                                  x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_s_curve, 
-                                 num_bins_y = num_bins_y_s_curve, 
+                                 num_bins_x = 8, 
+                                 num_bins_y = 14, 
                                  x_start = NA, y_start = NA, 
                                  buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_s_curve,
+                                 hex_size = 0.1,
                                  is_bin_centroid = TRUE,
@@
 
-## Constructing the 2D model
-
-Constructing the 2D model mainly contains (1) preprocessing, (2) binning NLD data, (3) obtain bin centroids/ means, and (4) triangulate bin centroids/ means.
+## Constructing the 2-D model
+
+Constructing the $2-D$ model mainly contains (1) preprocessing, (2) binning NLD data, (3) obtain bin centroids/ means, and (4) triangulate bin centroids/ means.
 
@@
 The user can set up the ratio of the width and height of the hexagon using the `hex_ratio` argument.
+`hex_ratio` will be determined within `gen_scaled_data()`, if it is not provided.
 
@@
 
-In binning data step, the objective is to perform hexagonal binning and assign NLDR data to regualr hexagons.
+In binning data step, the objective is to perform hexagonal binning and assign NLDR data to regular hexagons.
 
 **Step 1: Compute hexagonal grid configurations**
+
+In the computation of hexagonal bin configurations, there are mainly two configurations to consider: the number of bins along the x and y axes.
+The `calc_bins()` function is used to compute these configurations.
+It uses the hexagonal size, and the buffer amount along the x and y axes.
+
+When computing the number of bins along the x and y axes, the process begins by calculating the range of the respective NLDR component.
+After that, the range is adjusted to accommodate the buffer amount.
+Then, the spacing between hexagons is determined based on hexagonal size.
+Finally, the number of bins along the axis is computed by dividing the adjusted range by the corresponding spacing.
 
@@
 The mains steps for the hexagonal binning algorithm can be executed by the main function `hex_binning()`, or can be run separately for more flexibility.
-<!-- add about main function for hexagonal binning-->
+The number of bins along the x and y axes, hexagonal grid starting point, buffer size along the x and y axes, and hexagon size are parameters that will be determined within `hex_binning()`, if they are not provided.
+The results from various steps in the process are illustrated in Figure \@ref(fig:hex-bin).
 
@@
 hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x, 
-                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      y = "UMAP2", num_bins_x = 8, 
+                      num_bins_y = 14, x_start = NA, y_start = NA, 
                       buffer_x = NA, buffer_y = NA, hex_size = NA, 
@@
 
-```{r, out.width="100%", fig.cap = "Illustration of key steps of the hexagonal bining algorithm: (a) first, all hexagonal bin centroids are created; (b) second, full hexagon grid is created; and (c) finally, assign the data to hexagons.
+```{r hex-bin, out.width="100%", fig.cap = "Illustration of key steps of the hexagonal bining algorithm: (a) first, all hexagonal bin centroids are created; (b) second, full hexagon grid is created; and (c) finally, assign the data to hexagons.
 The centroids are used to generate hexagonal coordinates.
@@
 
+The process of generating hexagonal bin centroids begins by defining the starting coordinates for the hexagons.
+With the number of bins computed along the x and y axes, along with the horizontal and vertical spacing, the centroids are computed by adding corresponding spacing to the hexagonal grid starting point along the axes.
+
+The `gen_centroids()` function is used to compute all the centroids within the hexagonal grid after providing the number of bins along the x and y axes, hexagonal grid starting point, buffer size along the x and y axes, and hexagon size.
+
 ```{r, echo=TRUE, eval=FALSE}
 centroid_list <- gen_centroids(data = s_curve_noise_umap_scaled, x = "UMAP1", 
-                               y = "UMAP2", num_bins_x = num_bins_x,
-                               num_bins_y = num_bins_y, x_start = NA,
+                               y = "UMAP2", num_bins_x = 8,
+                               num_bins_y = 14, x_start = NA,
                                y_start = NA, buffer_x = NA, buffer_y = NA, 
@@
 
+Once the centroids for the hexagonal grid are obtained, the next step is to compute the hexagonal coordinates.
+To obtain hexagonal coordinates by passing all the hexagonal centroids, `gen_hex_coord()` is used.
+
+<!--write more on how grid coordinates are computed-->
+
 ```{r, echo=TRUE, eval=FALSE}
@@
 **Step 4: Assign NLDR data to hexagons**
+
+After obtaining the centroids of the hexagonal bins for the entire grid, the process of assigning NLDR points to hexagons involves determining the nearest hexagonal bin for each NLDR point using the $2-D$ Euclidean distance.
+Then, the hexagonal ID of the nearest hexagon is assigned to the corresponding NLDR points.
+`assign_data()` used to do it after passing all the centroids.
 
@@
 
+Next, `compute_std_counts()` function is used to compute standardized number of points within each hexagon when provide the NLDR data along with their corresponding hexagonal IDs.
+
 ```{r, echo=TRUE, eval=FALSE}
@@
 
-To obtain hexagonal bin centroids, `extract_hexbin_centroids()`is used. The function `extract_hexbin_mean()` is used to obtain average within each hexagon.
+To obtain hexagonal bin centroid coordinates, `extract_hexbin_centroids()`is used. The dataframes of all the centroids and the standardized counts for hexagons need to be provided.
 
@@
 
+The function `extract_hexbin_mean()` is used to obtain mean value coordinates for each hexagon by passing the NLDR data along with their hexagonal IDs and standardized number of points within each hexagon.
+
 ```{r, echo=TRUE, eval=FALSE}
@@
 
-The `tri_bin_centroids()` is used to do triangulate bin centroids/ bin means. Then, `gen_edges()` compute the edges to obtain the triangular mesh which is the 2D model.
+The `tri_bin_centroids()` is used to triangulate bin centroids/ bin means. Then, `gen_edges()` compute the edges to obtain the triangular mesh after providing the triangulated data.
 
@@
 
-The $p$-D model is generated by `avg_highd_data()` by passing the $p$-D data.
+The `avg_highd_data()` is used to generate $p-D$ model by passing the $p-D$ and binned NLDR data. The $p-D$ model for NLDR data is the averaged $p-D$ data within hexagons.
 
@@
 
-`predict_emb()` function is used to predict 2D embedding for a new $p$-D data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
+The `predict_emb()` function is used to predict 2D embedding for a new $p-D$ data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
+
+In the prediction process, first, the nearest $p-D$ model point is identified for a given new $p-D$ point by computing $p-D$ Euclidean distance.
+Then, the corresponding $2-D$ bin centroid mapping for the identified $p-D$ model point is determined.
+Finally, the coordinates of the identified $2-D$ bin centroid is used as the predicted NLDR point for the new $p-D$ data point.
 
@@
 
-As a Goodness of fit statistics for the model, `gen_summary()` is used to compute residuals and MSE. For the fitted model, how residuals and MSE is changing is shown in Figure 2.
+As a Goodness of fit statistics for the model, `gen_summary()` is used to compute residuals and MSE.
+These metrics are used to assess how well the fitted model will capture the underlying structure of the $p-D$ data.
 

### 2024-06-04 paper-quollr.Rmd
Plan for paper

--- 2024-04-08 paper-quollr.Rmd
@@
 abstract: >
-  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional ($p-D$) data in low-diemnsioanl space.
-However, their complexity and hyperparameter choices may lead to distrustful or misleading results.
-The R package \CRANpkg{quollr} is developed as a new tool to help to determine which method, which hyperparameter choice provide the most accurate representation of $p-D$ data.
+  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional data in low-diemnsioanl space.
+However, their complexity and (hyper)parameter choices may lead to distrustful or misleading results.
+The R package \CRANpkg{quollr} is developed as a new tool to help to determine which method, which (hyper)parameter choice provide the most accurate representation of $p-D$ data.
 Clustering data from \CRANpkg{cardinalR} package, is used to illustrate the algorithm and its use within the package.
@@
 
-This paper presents the R package, `quollr` which introduce a new tool in determining which NLDR technique and which hyperparameter choice gives most accurate representation of $p-D$ data.
+This paper presents the R package, `quollr` which introduce a new visual tool in determining which NLDR technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
 The methodology of the algorithm is explained in *cite the methodology paper*.
 Furthermore, the `quollr` package enables users to perform hexagonal binning [@dan2023], resulting in the generation of regular hexagons.
-The software is available from the Comphrehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).
+The software is available from the Comprehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).
 
@@
 
-The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `s_curve_noise_training` and `s_curve_noise_umap`.
-
-The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
-When constructing the 2D model, the user can choose either to fit the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
-
-```{r, echo=TRUE, eval=FALSE}
-model_object <- fit_highd_model( training_data = s_curve_noise_training, 
-                                 nldr_df_with_id = s_curve_noise_umap_scaled, 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_s_curve, 
-                                 num_bins_y = num_bins_y_s_curve, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_s_curve,
-                                 is_bin_centroid = TRUE,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "x")
-```
-
 <!-- add about main function to gen 2D and highD models-->
@@
 
-```{r}
+## Constructing the 2-D model
 
-## Decide by looking at MSE plot
-num_bins_x_s_curve <- 8
-num_bins_y_s_curve <- 14
-hex_size_s_curve <- 0.1
-hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
-                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-## remove low-density hexagons
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+### Scaling the data
 
-centroid_plot <- ggplot(data = all_centroids_df, aes(x = c_x, y = c_y)) +
-  geom_point(color = "#33a02c") +
-  coord_fixed() +
-  theme_void() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5,
-vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6))
+### Computing hexagon grid configurations
 
-full_grid <- ggplot(data = hex_grid, aes(x = x, y = y)) +
-  geom_polygon(fill = "#ffffff", color = "black", aes(group = hex_poly_id)) +
-  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#33a02c") +
-  coord_fixed() +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
-vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6))
+### Binning the data
 
-std_pts_grid <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_point(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2), color = "black", alpha = 1) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_fixed() +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
-vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6))
-
-umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
-
-model_object <- fit_highd_model( training_data = s_curve_noise_training, 
-                                 nldr_df_with_id = s_curve_noise_umap_scaled, 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_s_curve, 
-                                 num_bins_y = num_bins_y_s_curve, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_s_curve,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "x")
-df_bin_centroids_s_curve <- model_object$df_bin_centroids
-df_bin_s_curve <- model_object$df_bin
-
-## Triangulate bin centroids
-tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
-tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
-## Compute 2D distances
-distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
-benchmark_s_curve <- 0.5
-
-trimesh_s_curve <- ggplot(df_bin_centroids_s_curve, aes(x = c_x, y = c_y)) +
-  geom_trimesh() +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
-  labs(colour = "")
-
-trimesh_colored_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve,
-tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-
-trimesh_colored_s_curve_umap <- trimesh_colored_s_curve_umap +
-  theme_linedraw() +
-  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
-  labs(colour = "")
-
-trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-trimesh_removed_s_curve_umap_0.3 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.3, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-trimesh_removed_s_curve_umap_0.3 <- trimesh_removed_s_curve_umap_0.3 +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-trimesh_removed_s_curve_umap_0.8 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.8, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-trimesh_removed_s_curve_umap_0.8 <- trimesh_removed_s_curve_umap_0.8 +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_s_curve <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), umap_data_with_hb_id)
-
-# show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
-
-### Define type column
-df <- df_all_s_curve |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin_s_curve |>
-  dplyr::filter(hb_id %in% df_bin_centroids_s_curve$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_s_curve$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-df_exe <- dplyr::bind_rows(df_b, df)
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_s_curve |>
-  dplyr::filter(distance < benchmark_s_curve)
-## Since erase brushing is considerd.
-tour <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = distance_df_small_edges$from,
-                         lineTo = distance_df_small_edges$to,
-                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(1, NROW(df))),
-                         levelColors = c("#6a3d9a", "#33a02c"))
-```
-
-## Constructing the 2D model
-
-Constructing the 2D model mainly contains (1) preprocessing, (2) binning NLD data, (3) obtain bin centroids/ means, and (4) triangulate bin centroids/ means.
-
-### Preprocessing
-<!--Discuss scaling the NLDR data-->
-
-The function `gen_scaled_data()` is used to prepare the NLDR data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
-The user can set up the ratio of the width and height of the hexagon using the `hex_ratio` argument.
-
-```{r, echo=TRUE, eval=FALSE}
-s_curve_noise_umap_scaled <- gen_scaled_data(data = s_curve_noise_umap, 
-                                             x = "UMAP1", y = "UMAP2", 
-                                             hex_ratio = NA)
-
-```
-
-<!-- add plot before and after scaling NLDR data-->
-<!-- also can discuss about the tuning hex_ratio-->
-
-### Binning data
-
-In binning data step, the objective is to perform hexagonal binning and assign NLDR data to regualr hexagons.
-
-**Step 1: Compute hexagonal grid configurations**
-
-```{r, echo=TRUE, eval=FALSE}
-bin_list <- calc_bins(data = s_curve_noise_umap_scaled, x = "UMAP1", 
-                      y = "UMAP2", hex_size = 0.1, buffer_x = NA, buffer_y = NA)
-
-num_bins_x <- bin_list$num_x
-num_bins_y <- bin_list$num_y
-```
-
-The mains steps for the hexagonal binning algorithm can be executed by the main function `hex_binning()`, or can be run separately for more flexibility.
-<!-- add about main function for hexagonal binning-->
-
-```{r, echo=TRUE, eval=FALSE}
-hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x, 
-                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = NA, 
-                      col_start = "UMAP")
-
-```
-
-```{r, out.width="100%", fig.cap = "Illustration of key steps of the hexagonal bining algorithm: (a) first, all hexagonal bin centroids are created; (b) second, full hexagon grid is created; and (c) finally, assign the data to hexagons.
-The centroids are used to generate hexagonal coordinates.
-Once the full hexagonal grid is generated, by computing the 2D euclidean distance from each NLDR points to hexagonal bin centroids, the nearest NLDR points are allocated to the nearest hexagoanl bin."}
-
-centroid_plot + full_grid + std_pts_grid +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 3) &
-  theme(plot.tag = element_text(size = 8))
-```
-
-**Step 2: Generate the hexagonal bin centroids**
-
-```{r, echo=TRUE, eval=FALSE}
-centroid_list <- gen_centroids(data = s_curve_noise_umap_scaled, x = "UMAP1", 
-                               y = "UMAP2", num_bins_x = num_bins_x,
-                               num_bins_y = num_bins_y, x_start = NA,
-                               y_start = NA, buffer_x = NA, buffer_y = NA, 
-                               hex_size = NA)
-
-all_centroids_df <- as.data.frame(do.call(cbind, centroid_list))
-```
-
-**Step 3: Create the hexagonal grid**
-
-```{r, echo=TRUE, eval=FALSE}
-all_hex_coordinates_list <- gen_hex_coord(centroids_df = all_centroids_df,
-                                            hex_size = NA)
-```
-
-**Step 4: Assign NLDR data to hexagons**
-
-```{r, echo=TRUE, eval=FALSE}
-df_with_hb_id_list <- assign_data(data = s_curve_noise_umap_scaled |> 
-                                    dplyr::select(-ID), 
-                                  centroid_df = all_centroids_df, 
-                                  col_start = "UMAP")
-umap_data_with_hb_id <- as.data.frame(do.call(cbind, df_with_hb_id_list))
-```
-
-```{r, echo=TRUE, eval=FALSE}
-std_counts_list <- compute_std_counts(data_hex_id = df_with_hex_id)
-counts_df <- as.data.frame(do.call(cbind, std_counts_list))
-```
-
-### Obtain bin centroids/ means
-
-To obtain hexagonal bin centroids, `extract_hexbin_centroids()`is used. The function `extract_hexbin_mean()` is used to obtain average within each hexagon.
-
-```{r, echo=TRUE, eval=FALSE}
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-```
-
-```{r, echo=TRUE, eval=FALSE}
-df_bin_means <- extract_hexbin_mean(nldr_df_with_hex_id = umap_with_hb_id,
-                                        counts_df = counts_df)
-```
-
-### Triangulate bin centroids/ means
-
-The `tri_bin_centroids()` is used to do triangulate bin centroids/ bin means. Then, `gen_edges()` compute the edges to obtain the triangular mesh which is the 2D model.
-
-```{r, echo=TRUE, eval=FALSE}
-tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
-tr_from_to_df <- gen_edges(tri_object = tr1_object)
-```
+### Indicating neighbors by line segments connecting centroids
 
 ## Lifting the model into high dimensions
-
-The $p$-D model is generated by `avg_highd_data()` by passing the $p$-D data.
-
-```{r, echo=TRUE, eval=FALSE}
-df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID),
-                           umap_data_with_hb_id)
-
-df_bin <- avg_highd_data(data = df_all, col_start = "x")
-```
 
@@
 
-Mainly there are two model parameter need to consider: benchmark value to remove the low-density hexagons, and benchmark value to remove the long edges.
-`find_low_dens_hex()` is used to find the hexagons which contains less number of points by considering the density of their neighboring points as well.
-In here, user can first decide which are the low-density hexagons and pass them to this function to check whether these removal of low-density hexagons can affect the model fit by looking at the the neighbors.
-On the other hand, `find_lg_benchmark()` function is used to compute the threshold for removing long edges.
-
-```{r, echo=TRUE, eval=FALSE}
-df_bin_centroids_low <- df_bin_centroids |> dplyr::filter(std_counts <= 0.43)
-
-low_dens_hex <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids, 
-                  num_bins_x = num_bins_x, 
-                  df_bin_centroids_low = df_bin_centroids_low)
-```
-
-```{r, echo=TRUE, eval=FALSE}
-distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, start_x = "x_from", 
-                           start_y = "y_from", end_x = "x_to", end_y = "y_to",
-                           select_vars = c("from", "to", "distance"))
-
-benchmark_val <- find_lg_benchmark(distance_edges = distance_df, distance_col = "distance")
-```
-
 ## Prediction
-
-`predict_emb()` function is used to predict 2D embedding for a new $p$-D data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
-
-```{r}
-## Prediction for training data
-pred_emb_list_training <- predict_emb(test_data = s_curve_noise_training, 
-                             df_bin_centroids = df_bin_centroids_s_curve, 
-                             df_bin = df_bin_s_curve, type_NLDR = "UMAP")
-pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list_training))
-
-pred_points_training <- ggplot(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2)) +
-  geom_point() +
-  geom_point(data = pred_df_training, aes(x = pred_UMAP_1, y = pred_UMAP_2), colour = "#de2d26") +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-## Prediction for test data
-pred_emb_list_test <- predict_emb(test_data = s_curve_noise_test, 
-                             df_bin_centroids = df_bin_centroids_s_curve, 
-                             df_bin = df_bin_s_curve, type_NLDR = "UMAP")
-pred_df_test <- as.data.frame(do.call(cbind, pred_emb_list_test))
-
-pred_points_test <- ggplot(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2)) +
-  geom_point() +
-  geom_point(data = pred_df_test, aes(x = pred_UMAP_1, y = pred_UMAP_2), colour = "#de2d26") +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r, out.width="100%", fig.cap = "Predicted 2D embedding overlaid with UMAP embedding of S-curve training data: (a) prediction of 2D embeddings for S-curve training data, and (b) prediction of 2D embeddings for S-curve test data.
-Once the high-D model is generated, first, for the new data points the nearest high-D mappings of hexagonal bin centroids find by computing Euclidean distance.
-Then, mapping back the 2D hexagonal bin centroids gives the prediction for the new data points."}
-
-pred_points_training + pred_points_test +
-  plot_layout(ncol = 2) &
-  theme(plot.tag = element_text(size = 8))
-```
 
@@
 
-As a Goodness of fit statistics for the model, `gen_summary()` is used to compute residuals and MSE. For the fitted model, how residuals and MSE is changing is shown in Figure 2.
-
-```{r, echo=TRUE}
-gen_summary(test_data = s_curve_noise_training, prediction_df = pred_df_training,
-df_bin = df_bin_s_curve, col_start = "x")
-```
-
-```{r, echo=TRUE}
-gen_summary(test_data = s_curve_noise_test, prediction_df = pred_df_test,
-df_bin = df_bin_s_curve, col_start = "x")
-```
-
-<!--add a plot with total bins Vs MSE, non-empty bins Vs Error-->
-
 ## Visualizations
 
-The package provides four basic visualizations which includes three visualizations regarding 2D model (static vis) and one is regrading $p$-D model (dynamic vis).
-Each visualization can be done their own functions which describes in this section.
+### $2\text{-}D$ model visualization
 
-<!--add trimesh, colored trimesh, removed edge trimesh, screenshots from langevitour with the model and data-->
-
-### 2D model visualization
-
-<!--2D model vis: triangular mesh, colored trimesh, removed edge trimesh-->
-To visualize the 2D model, mainly three functions are used.
-As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`.
-After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges.
-As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively.
-Following this, `vis_rmlg_mesh()` is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure \@ref(fig:mesh-plots)c).
-In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in 2D.
-Using small value of `benchmark_value`, will produce a triangular mesh with missing data structure; for example `benchmark_value = 0.3` shows two clusters rather than continuous structure, while `benchmark_value = 0.8` creates long edges and mislead the data structure in $p-D$ space.
-
-```{r mesh-plots, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges.
-The edges which has the length greater than $0.5$ are assigned as long edges.
-To obtain smooth representation in 2D space, the long edges are removed."}
-
-trimesh_s_curve + trimesh_colored_s_curve_umap + trimesh_removed_s_curve_umap +
-  plot_layout(ncol = 3) &
-  theme(plot.tag = element_text(size = 8))
-```
-
-```{r mesh-diffplots, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space with different benchmark value applied to remove long edges: (a) benchmark_value = 0.3, and (b) benchmark_value = 0.8.
-Small benchmark values will generate misleading data structures in 2D space, while; large bechmark value will generate more long edges in $p-D$ and miinterpret the $p-D$ data structure."}
-
-trimesh_removed_s_curve_umap_0.3 + trimesh_removed_s_curve_umap_0.8 +
-  plot_layout(ncol = 2) &
-  theme(plot.tag = element_text(size = 8))
-```
-
-### p-D model visualization
-
-<!--highD model vis: langevitour with the model-->
-
-The $p$-D model overlaid on data is visualized by the function `show_langevitour()`. This is a useful dynamic visualization result to validate visually whether the model fits the data well or not.
-
-```{r, echo=TRUE, eval=FALSE}
-tour <- show_langevitour(df = df_all, df_b = df_bin, 
-                 df_b_with_center_data = df_bin_centroids, benchmark_value = 0.5, 
-                 distance = distance_df, distance_col = "distance", 
-                 use_default_benchmark_val = FALSE, col_start = "x")
-```
-
-<!--need to add screenshots of the frames and add the link of the video-->
-```{r tour-s-curve}
-tour
-```
+### $p\text{-}D$ model visualization
 
@@
 
-<!--in data processing, minus values, NA, Inf values for aspect_ratio or hex_ratio-->
-<!-- to check length of list objects-->
-<!-- add tests to find NAs, inf values are in NLDR data-->
+<!--discuss a test with how a point in a intersection of two hexagonal going to evaluate-->
 
@@
 
-We illustrate the use of the algorithm by applying it to `clusters_different_shapes_diff_num_points` dataset in `cardinalR` package.
-The tSNE is used in the example for the demonstration (see Figure \@ref(fig:tsne-clust)).
+<!--with prism data-->
 
-<!--select a dataset from cardinalR-->
-<!-- diff shape clusters-->
-<!-- talk with best fit-->
-
-```{r, echo=FALSE}
-clusters_different_shapes_diff_num_points <- function(sample_size = 400, with_seed = NULL, cluster_size_vec = c(50, 50, 50, 50, 100, 100), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.05, cluster_sd_non_gau = 0.1, num_dims = 7, a = 2, b = 4) {
-
-  # To check the seed is not assigned
-  if (!is.null(with_seed)) {
-    set.seed(with_seed)
-  }
-
-  num_clusters <- num_gussian_clusters + num_non_gaussian_clusters
-
-  ## Generate Gaussian clusters
-
-  # Create a vector of possible values (0 and 1)
-  values <- c(0, 1)
-
-  # Create an expanded grid with 0's and 1's
-  mean_val_grid <- tidyr::expand_grid(!!!setNames(rep(list(values), num_dims),
-                                                  paste0("mean_dim", 1:num_dims)))
-
-  # To select combinations for assigned number of clusters
-
-  mean_val_grid_gau <- mean_val_grid |>
-    dplyr::slice_sample(n = num_gussian_clusters)
-
-  mean_val_grid_non_gau <- mean_val_grid |>
-    dplyr::slice_sample(n = num_non_gaussian_clusters)
-
-  # To generate empty tibble
-  column_names <- paste0(rep("x", num_dims), 1:num_dims)
-  df <- tibble(!!!setNames(rep(list(NULL), length(column_names)), column_names))
-
-  for (i in 1:num_gussian_clusters) {
-
-    # To filter the mean values for specific cluster
-    mean_val_for_cluster <- mean_val_grid_gau |>
-      dplyr::filter(dplyr::row_number() == i) |>
-      unlist(use.names = FALSE)
-
-    # Initialize an empty list to store the vectors with column
-    # values
-    dim_val_list <- list()
-
-    for (j in 1:num_dims) {
-
-      dim_val_list[[column_names[j]]] <- rnorm(cluster_size_vec[i], mean = mean_val_for_cluster[j],
-                                               sd = cluster_sd_gau)
-
-    }
-    # To generate a tibble for a cluster
-    df_gau_cluster <- tibble::as_tibble(dim_val_list)
-
-    df <- dplyr::bind_rows(df, df_gau_cluster)
-
-  }
-
-  for (i in 1:num_non_gaussian_clusters) {
-
-    phi <- runif(cluster_size_vec[(num_clusters - i)], max = 2*pi)
-    rho <- sqrt(runif(cluster_size_vec[(num_clusters - i)]))
-
-    # To filter the mean values for specific cluster
-    presence_of_elipse_cluster <- mean_val_grid_non_gau |>
-      dplyr::filter(dplyr::row_number() == i) |>
-      unlist(use.names = FALSE)
-
-    # Initialize an empty list to store the vectors with column
-    # values
-    dim_val_list_n <- list()
-
-    for (j in 1:num_dims) {
-      if(presence_of_elipse_cluster[j] == 1){
-        dim_val_list_n[[column_names[j]]] <- sqrt(a)*rho*cos(phi) + b
-        ## Surface of poolar coordinate
-      } else {
-        dim_val_list_n[[column_names[j]]] <- rnorm(cluster_size_vec[(num_clusters - i)], mean = 0,
-                                                   sd = cluster_sd_non_gau)
-
-      }
-
-    }
-    # To generate a tibble for a cluster
-    df_non_gau_cluster <- tibble::as_tibble(dim_val_list_n)
-
-    df <- dplyr::bind_rows(df, df_non_gau_cluster)
-
-  }
-
-  df
-
-}
-```
-
-<!--need to add screen shots of the original data-->
-
-```{r}
-clust_df <- clusters_different_shapes_diff_num_points(sample_size = 1500, with_seed = NULL, cluster_size_vec = c(250, 150, 150, 150, 350, 450), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.2, cluster_sd_non_gau = 0.3, num_dims = 5, a = 2, b = 4)
-
-langevitour::langevitour(clust_df)
-
-clust_df <- clust_df |>
-  dplyr::mutate(ID = row_number())
-```
-
-```{r}
-tSNE_fit <- clust_df |>
-  dplyr::select(-ID) |>
-  dplyr::select(where(is.numeric)) |>
-  Rtsne::Rtsne(perplexity = 39, pca = FALSE, pca_center = FALSE, normalize = FALSE, dims = 2)
-
-tSNE_df <- tSNE_fit$Y |> 
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::mutate(ID = clust_df$ID)
-
-names(tSNE_df) <- c("tSNE1", "tSNE2", "ID")
-
-tSNE_df_plot <- tSNE_df |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
-  geom_point(alpha=0.5, colour="#e41a1c", size = 0.5) +
-  coord_equal() +
-  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
-        axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7))
-```
-
-```{r tsne-clust, out.width="100%", fig.cap="tSNE representation of data with perplexity: 39. Is this representation accurately capture the structure in high dimensions?"}
-tSNE_df_plot
-```
-
-## Selecting parameter values for hexagonal binning
-
-We previously introduced three key parameters, *hex_size*, *x_start*, and *y_start*, in our explanation of the algorithm.
-While their best choice is unknown, a visualization tool enables their tuning such that user can select reasonable values.
-
-Consider the relationships between *total number of bins*, *starting point*, and MSE. Increase of *total number of bins* usually reduces the MSE.
-
-<!--hex_size, starting point (x_start, y_start)-->
-<!--talk about to use bin centroids or bin means-->
-
-<!-- MSE plot to find which is the effective number of bins to use-->
-```{r}
-tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_df, 
-                                    x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tSNE_df))    
-## Model with bin centroid
-hex_size_vec <- seq(0.03, 2, by = 0.01)
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
-mse_df_gau <- dplyr::bind_rows(vec)[0, ]
-mse_df_gau <- mse_df_gau |>
-  dplyr::mutate_if(is.character, as.numeric)
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = tsne_gau_scaled, 
-            x = "tSNE1", y = "tSNE2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = clust_df, 
-                                   nldr_df_with_id = tsne_gau_scaled, 
-                                   x = "tSNE1", y = "tSNE2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "tSNE", 
-                                   col_start_highd = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = clust_df, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = clust_df, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "x")
-
-  mse_df_gau <- mse_df_gau |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
-
-}
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_gau |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_gau |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_gau |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_clust <- dplyr::bind_rows(not_dupli_df, duplicate_df) |>
-  dplyr::mutate(type = "bin centroids")
-```
-
-```{r}
-
-mse_clust_plot <- ggplot(mse_df_clust, aes(x = num_bins,
-                                       y = log(mse)
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 238, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  ylab("log(MSE)") +
-  xlab("total number of bins")
-```
-
-```{r, fig.cap="Parameter tuning plots, where the best choice of parameter at a large drop in MSE. Here, these are at *total number of bins* = 238."}
-mse_clust_plot
-```
-
-## Selecting parameter values for the model
-
-<!--benchmark value to remove low density hexagons, long edges-->
-
-## Fit the model
-
-To perform the algorithm on the clustering data, we first scale the tSNE data.
-We then let *number of bins along the x-axis* be $14$, *number of bins along the y-axis* be $27$, and *hex_size* be $0.05$; see [Selecting parameter values] for further discussion and guidance regarding selection of these choices.
-
-<!-- model with bin centroids-->
-```{r}
-
-## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 14
-num_bins_y_tsne_gau <- 27
-hex_size_tsne_gau <- 0.05
-x_start_tsne_gau <- -0.0481
-y_start_tsne_gau <- 0.00833
-## non-empty:80
-hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
-                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = x_start_tsne_gau, y_start = y_start_tsne_gau,
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "tSNE")
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
-
-model_object_tsne_gau <- fit_highd_model( training_data = clust_df, 
-                                 nldr_df_with_id = tsne_gau_scaled, 
-                                 x = "tSNE1", y = "tSNE2", 
-                                 num_bins_x = num_bins_x_tsne_gau, 
-                                 num_bins_y = num_bins_y_tsne_gau, 
-                                 x_start = x_start_tsne_gau, y_start = y_start_tsne_gau, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_tsne_gau,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "tSNE", 
-                                 col_start_highd = "x")
-df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
-df_bin_tsne_gau <- model_object_tsne_gau$df_bin
-## Triangulate bin centroids
-tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
-tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
-## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-# distance_plot <- plot_dist(distance_tsne_gau) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-## To find the benchmark value
-benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
-trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value =
-benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-trimesh_removed_tsne_gau
-
-df_all_tsne_gau <- dplyr::bind_cols(clust_df |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
-
-### Define type column
-df <- df_all_tsne_gau |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin_tsne_gau |>
-  dplyr::filter(hb_id %in% df_bin_centroids_tsne_gau$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_tsne_gau$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-df_exe <- dplyr::bind_rows(df_b, df)
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_tsne_gau |>
-  dplyr::filter(distance < benchmark_tsne_gau)
-## Since erase brushing is considerd.
-langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = distance_df_small_edges$from,
-                         lineTo = distance_df_small_edges$to,
-                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.5, NROW(df))),
-                         levelColors = c("#6a3d9a", "#33a02c"))
-```
-
-What can learn from the model:
-
-- There are five clusters. Four of them are Gaussian and two of them are long 2D clusters. Gaussian clusters are close to each other while others are far away.
-
-- Model preserved local and global structure.
-
-- Problem: Couldn't capture the variance within the clusters.
-
-- There is one edge that is not connected in 2D triangular mesh.
-
-# Conclusion
+# Discussion
 
@@
 
-The paper includes a clustering example to illustrate how `quollr` is useful to assess which NLDR technique and which hyperparameter choice gives the most accurate representation.
+The paper includes a clustering example to illustrate how `quollr` is useful to assess which NLDR technique and which (hyper)parameter choice gives the most accurate representation.
 In addition, how to select parameters for hexagonal binning and fitting model are explained.

### 2024-06-04 paper_old_2.qmd
Plan for paper

---
title: "Untitled"
format: html
editor: visual
---

## Usage

The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `s_curve_noise_training`, `s_curve_noise_test` and `s_curve_noise_umap`.

`s_curve_noise_training` is a $3-D$ S-curve dataset with additional four noise dimensions which is used to train the model. `s_curve_noise_test` is used for testing. `s_curve_noise_umap` is the UMAP $2-D$ embedding for `s_curve_noise_training` dataset. Each dataset contains a unique ID column.

The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility. When constructing the 2D model, the user can choose either to fit the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.

If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step. For example, if the user wishes to use already binning data, the `extract_hexbin_centroids()` function can be used directly.

The number of bins along the x and y axes, hexagonal grid starting point, buffer size along the x and y axes, hexagon size, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model`, if they are not provided. They are created as they are needed throughout the following example.

```{r, echo=TRUE, eval=FALSE}
model_object <- fit_highd_model( training_data = s_curve_noise_training, 
                                 nldr_df_with_id = s_curve_noise_umap_scaled, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = 8, 
                                 num_bins_y = 14, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = 0.1,
                                 is_bin_centroid = TRUE,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "x")
```

<!-- add about main function to gen 2D and highD models-->
<!--Discuss the model can be generated with bin centroids or bin means-->

```{r}

## Decide by looking at MSE plot
num_bins_x_s_curve <- 8
num_bins_y_s_curve <- 14
hex_size_s_curve <- 0.1
hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
## remove low-density hexagons
hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

centroid_plot <- ggplot(data = all_centroids_df, aes(x = c_x, y = c_y)) +
  geom_point(color = "#33a02c") +
  coord_fixed() +
  theme_void() +
  theme(plot.title = element_text(size = 7, hjust = 0.5,
vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6))

full_grid <- ggplot(data = hex_grid, aes(x = x, y = y)) +
  geom_polygon(fill = "#ffffff", color = "black", aes(group = hex_poly_id)) +
  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#33a02c") +
  coord_fixed() +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6))

std_pts_grid <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_point(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2), color = "black", alpha = 1) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_fixed() +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6))

umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))

model_object <- fit_highd_model( training_data = s_curve_noise_training, 
                                 nldr_df_with_id = s_curve_noise_umap_scaled, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = num_bins_x_s_curve, 
                                 num_bins_y = num_bins_y_s_curve, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_s_curve,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "x")
df_bin_centroids_s_curve <- model_object$df_bin_centroids
df_bin_s_curve <- model_object$df_bin

## Triangulate bin centroids
tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
## Compute 2D distances
distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
benchmark_s_curve <- 0.5

trimesh_s_curve <- ggplot(df_bin_centroids_s_curve, aes(x = c_x, y = c_y)) +
  geom_trimesh() +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
  labs(colour = "")

trimesh_colored_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve,
tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")

trimesh_colored_s_curve_umap <- trimesh_colored_s_curve_umap +
  theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
  labs(colour = "")

trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

trimesh_removed_s_curve_umap_0.3 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.3, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
trimesh_removed_s_curve_umap_0.3 <- trimesh_removed_s_curve_umap_0.3 +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

trimesh_removed_s_curve_umap_0.8 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.8, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
trimesh_removed_s_curve_umap_0.8 <- trimesh_removed_s_curve_umap_0.8 +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

df_all_s_curve <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), umap_data_with_hb_id)

# show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")

### Define type column
df <- df_all_s_curve |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin_s_curve |>
  dplyr::filter(hb_id %in% df_bin_centroids_s_curve$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean
## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_s_curve$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)
df_exe <- dplyr::bind_rows(df_b, df)
## Set the maximum difference as the criteria
distance_df_small_edges <- distance_s_curve |>
  dplyr::filter(distance < benchmark_s_curve)
## Since erase brushing is considerd.
tour <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                         lineFrom = distance_df_small_edges$from,
                         lineTo = distance_df_small_edges$to,
                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(1, NROW(df))),
                         levelColors = c("#6a3d9a", "#33a02c"))
```

## Constructing the 2-D model

Constructing the $2-D$ model mainly contains (1) preprocessing, (2) binning NLD data, (3) obtain bin centroids/ means, and (4) triangulate bin centroids/ means.

### Preprocessing
<!--Discuss scaling the NLDR data-->

The function `gen_scaled_data()` is used to prepare the NLDR data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization. The user can set up the ratio of the width and height of the hexagon using the `hex_ratio` argument. `hex_ratio` will be determined within `gen_scaled_data()`, if it is not provided.

```{r, echo=TRUE, eval=FALSE}
s_curve_noise_umap_scaled <- gen_scaled_data(data = s_curve_noise_umap, 
                                             x = "UMAP1", y = "UMAP2", 
                                             hex_ratio = NA)

```

<!-- add plot before and after scaling NLDR data-->
<!-- also can discuss about the tuning hex_ratio-->

### Binning data

In binning data step, the objective is to perform hexagonal binning and assign NLDR data to regular hexagons.

**Step 1: Compute hexagonal grid configurations**

In the computation of hexagonal bin configurations, there are mainly two configurations to consider: the number of bins along the x and y axes. The `calc_bins()` function is used to compute these configurations. It uses the hexagonal size, and the buffer amount along the x and y axes.

When computing the number of bins along the x and y axes, the process begins by calculating the range of the respective NLDR component. After that, the range is adjusted to accommodate the buffer amount. Then, the spacing between hexagons is determined based on hexagonal size. Finally, the number of bins along the axis is computed by dividing the adjusted range by the corresponding spacing.

```{r, echo=TRUE, eval=FALSE}
bin_list <- calc_bins(data = s_curve_noise_umap_scaled, x = "UMAP1", 
                      y = "UMAP2", hex_size = 0.1, buffer_x = NA, buffer_y = NA)

num_bins_x <- bin_list$num_x
num_bins_y <- bin_list$num_y
```

The mains steps for the hexagonal binning algorithm can be executed by the main function `hex_binning()`, or can be run separately for more flexibility. The number of bins along the x and y axes, hexagonal grid starting point, buffer size along the x and y axes, and hexagon size are parameters that will be determined within `hex_binning()`, if they are not provided. The results from various steps in the process are illustrated in Figure \@ref(fig:hex-bin).

```{r, echo=TRUE, eval=FALSE}
hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = 8, 
                      num_bins_y = 14, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = NA, 
                      col_start = "UMAP")

```

```{r hex-bin, out.width="100%", fig.cap = "Illustration of key steps of the hexagonal bining algorithm: (a) first, all hexagonal bin centroids are created; (b) second, full hexagon grid is created; and (c) finally, assign the data to hexagons. The centroids are used to generate hexagonal coordinates. Once the full hexagonal grid is generated, by computing the 2D euclidean distance from each NLDR points to hexagonal bin centroids, the nearest NLDR points are allocated to the nearest hexagoanl bin."}

centroid_plot + full_grid + std_pts_grid +
  plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 3) &
  theme(plot.tag = element_text(size = 8))
```

**Step 2: Generate the hexagonal bin centroids**

The process of generating hexagonal bin centroids begins by defining the starting coordinates for the hexagons. With the number of bins computed along the x and y axes, along with the horizontal and vertical spacing, the centroids are computed by adding corresponding spacing to the hexagonal grid starting point along the axes.

The `gen_centroids()` function is used to compute all the centroids within the hexagonal grid after providing the number of bins along the x and y axes, hexagonal grid starting point, buffer size along the x and y axes, and hexagon size.

```{r, echo=TRUE, eval=FALSE}
centroid_list <- gen_centroids(data = s_curve_noise_umap_scaled, x = "UMAP1", 
                               y = "UMAP2", num_bins_x = 8,
                               num_bins_y = 14, x_start = NA,
                               y_start = NA, buffer_x = NA, buffer_y = NA, 
                               hex_size = NA)

all_centroids_df <- as.data.frame(do.call(cbind, centroid_list))
```

**Step 3: Create the hexagonal grid**

Once the centroids for the hexagonal grid are obtained, the next step is to compute the hexagonal coordinates. To obtain hexagonal coordinates by passing all the hexagonal centroids, `gen_hex_coord()` is used.

<!--write more on how grid coordinates are computed-->

```{r, echo=TRUE, eval=FALSE}
all_hex_coordinates_list <- gen_hex_coord(centroids_df = all_centroids_df,
                                            hex_size = NA)
```

**Step 4: Assign NLDR data to hexagons**

After obtaining the centroids of the hexagonal bins for the entire grid, the process of assigning NLDR points to hexagons involves determining the nearest hexagonal bin for each NLDR point using the $2-D$ Euclidean distance. Then, the hexagonal ID of the nearest hexagon is assigned to the corresponding NLDR points. `assign_data()` used to do it after passing all the centroids.

```{r, echo=TRUE, eval=FALSE}
df_with_hb_id_list <- assign_data(data = s_curve_noise_umap_scaled |> 
                                    dplyr::select(-ID), 
                                  centroid_df = all_centroids_df, 
                                  col_start = "UMAP")
umap_data_with_hb_id <- as.data.frame(do.call(cbind, df_with_hb_id_list))
```

Next, `compute_std_counts()` function is used to compute standardized number of points within each hexagon when provide the NLDR data along with their corresponding hexagonal IDs.

```{r, echo=TRUE, eval=FALSE}
std_counts_list <- compute_std_counts(data_hex_id = df_with_hex_id)
counts_df <- as.data.frame(do.call(cbind, std_counts_list))
```

### Obtain bin centroids/ means

To obtain hexagonal bin centroid coordinates, `extract_hexbin_centroids()`is used. The dataframes of all the centroids and the standardized counts for hexagons need to be provided.

```{r, echo=TRUE, eval=FALSE}
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
```

The function `extract_hexbin_mean()` is used to obtain mean value coordinates for each hexagon by passing the NLDR data along with their hexagonal IDs and standardized number of points within each hexagon.

```{r, echo=TRUE, eval=FALSE}
df_bin_means <- extract_hexbin_mean(nldr_df_with_hex_id = umap_with_hb_id,
                                        counts_df = counts_df)
```

### Triangulate bin centroids/ means

The `tri_bin_centroids()` is used to triangulate bin centroids/ bin means. Then, `gen_edges()` compute the edges to obtain the triangular mesh after providing the triangulated data.

```{r, echo=TRUE, eval=FALSE}
tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df <- gen_edges(tri_object = tr1_object)
```

## Lifting the model into high dimensions

The `avg_highd_data()` is used to generate $p-D$ model by passing the $p-D$ and binned NLDR data. The $p-D$ model for NLDR data is the averaged $p-D$ data within hexagons.

```{r, echo=TRUE, eval=FALSE}
df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID),
                           umap_data_with_hb_id)

df_bin <- avg_highd_data(data = df_all, col_start = "x")
```

## Model parameters

<!--discuss about default settings-->

Mainly there are two model parameter need to consider: benchmark value to remove the low-density hexagons, and benchmark value to remove the long edges. `find_low_dens_hex()` is used to find the hexagons which contains less number of points by considering the density of their neighboring points as well. In here, user can first decide which are the low-density hexagons and pass them to this function to check whether these removal of low-density hexagons can affect the model fit by looking at the the neighbors. On the other hand, `find_lg_benchmark()` function is used to compute the threshold for removing long edges.

```{r, echo=TRUE, eval=FALSE}
df_bin_centroids_low <- df_bin_centroids |> dplyr::filter(std_counts <= 0.43)

low_dens_hex <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids, 
                  num_bins_x = num_bins_x, 
                  df_bin_centroids_low = df_bin_centroids_low)
```

```{r, echo=TRUE, eval=FALSE}
distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, start_x = "x_from", 
                           start_y = "y_from", end_x = "x_to", end_y = "y_to",
                           select_vars = c("from", "to", "distance"))

benchmark_val <- find_lg_benchmark(distance_edges = distance_df, distance_col = "distance")
```

## Prediction

The `predict_emb()` function is used to predict 2D embedding for a new $p-D$ data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.

In the prediction process, first, the nearest $p-D$ model point is identified for a given new $p-D$ point by computing $p-D$ Euclidean distance. Then, the corresponding $2-D$ bin centroid mapping for the identified $p-D$ model point is determined. Finally, the coordinates of the identified $2-D$ bin centroid is used as the predicted NLDR point for the new $p-D$ data point.

```{r}
## Prediction for training data
pred_emb_list_training <- predict_emb(test_data = s_curve_noise_training, 
                             df_bin_centroids = df_bin_centroids_s_curve, 
                             df_bin = df_bin_s_curve, type_NLDR = "UMAP")
pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list_training))

pred_points_training <- ggplot(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2)) +
  geom_point() +
  geom_point(data = pred_df_training, aes(x = pred_UMAP_1, y = pred_UMAP_2), colour = "#de2d26") +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

## Prediction for test data
pred_emb_list_test <- predict_emb(test_data = s_curve_noise_test, 
                             df_bin_centroids = df_bin_centroids_s_curve, 
                             df_bin = df_bin_s_curve, type_NLDR = "UMAP")
pred_df_test <- as.data.frame(do.call(cbind, pred_emb_list_test))

pred_points_test <- ggplot(data = s_curve_noise_umap_scaled, aes(x = UMAP1, y = UMAP2)) +
  geom_point() +
  geom_point(data = pred_df_test, aes(x = pred_UMAP_1, y = pred_UMAP_2), colour = "#de2d26") +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r, out.width="100%", fig.cap = "Predicted 2D embedding overlaid with UMAP embedding of S-curve training data: (a) prediction of 2D embeddings for S-curve training data, and (b) prediction of 2D embeddings for S-curve test data. Once the high-D model is generated, first, for the new data points the nearest high-D mappings of hexagonal bin centroids find by computing Euclidean distance. Then, mapping back the 2D hexagonal bin centroids gives the prediction for the new data points."}

pred_points_training + pred_points_test +
  plot_layout(ncol = 2) &
  theme(plot.tag = element_text(size = 8))
```

## Compute residuals and Mean Square Error (MSE)

As a Goodness of fit statistics for the model, `gen_summary()` is used to compute residuals and MSE. These metrics are used to assess how well the fitted model will capture the underlying structure of the $p-D$ data.

```{r, echo=TRUE}
gen_summary(test_data = s_curve_noise_training, prediction_df = pred_df_training,
df_bin = df_bin_s_curve, col_start = "x")
```

```{r, echo=TRUE}
gen_summary(test_data = s_curve_noise_test, prediction_df = pred_df_test,
df_bin = df_bin_s_curve, col_start = "x")
```

<!--add a plot with total bins Vs MSE, non-empty bins Vs Error-->

## Visualizations

The package provides four basic visualizations which includes three visualizations regarding 2D model (static vis) and one is regrading $p$-D model (dynamic vis). Each visualization can be done their own functions which describes in this section.

<!--add trimesh, colored trimesh, removed edge trimesh, screenshots from langevitour with the model and data-->

### 2D model visualization

<!--2D model vis: triangular mesh, colored trimesh, removed edge trimesh-->
To visualize the 2D model, mainly three functions are used. As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`. After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges. As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively. Following this, `vis_rmlg_mesh()` is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure \@ref(fig:mesh-plots)c). In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in 2D. Using small value of `benchmark_value`, will produce a triangular mesh with missing data structure; for example `benchmark_value = 0.3` shows two clusters rather than continuous structure, while `benchmark_value = 0.8` creates long edges and mislead the data structure in $p-D$ space.

```{r mesh-plots, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space: (a) triangular mesh, (b), triangular mesh colored by edge type, and (c) triangular mesh after removing the long edges. The edges which has the length greater than $0.5$ are assigned as long edges. To obtain smooth representation in 2D space, the long edges are removed."}

trimesh_s_curve + trimesh_colored_s_curve_umap + trimesh_removed_s_curve_umap +
  plot_layout(ncol = 3) &
  theme(plot.tag = element_text(size = 8))
```

```{r mesh-diffplots, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space with different benchmark value applied to remove long edges: (a) benchmark_value = 0.3, and (b) benchmark_value = 0.8. Small benchmark values will generate misleading data structures in 2D space, while; large bechmark value will generate more long edges in $p-D$ and miinterpret the $p-D$ data structure."}

trimesh_removed_s_curve_umap_0.3 + trimesh_removed_s_curve_umap_0.8 +
  plot_layout(ncol = 2) &
  theme(plot.tag = element_text(size = 8))
```

### p-D model visualization

<!--highD model vis: langevitour with the model-->

The $p$-D model overlaid on data is visualized by the function `show_langevitour()`. This is a useful dynamic visualization result to validate visually whether the model fits the data well or not.

```{r, echo=TRUE, eval=FALSE}
tour <- show_langevitour(df = df_all, df_b = df_bin, 
                 df_b_with_center_data = df_bin_centroids, benchmark_value = 0.5, 
                 distance = distance_df, distance_col = "distance", 
                 use_default_benchmark_val = FALSE, col_start = "x")
```

<!--need to add screenshots of the frames and add the link of the video-->
```{r tour-s-curve}
tour
```

## Tests

All functions have tests written and implemented using the \CRANpkg{testthat} [@testthat] in R.

These tests illuminated the issues that allowed us to make meaningful changes and understand some pitfalls of the package.

<!--in data processing, minus values, NA, Inf values for aspect_ratio or hex_ratio-->
<!-- to check length of list objects-->
<!-- add tests to find NAs, inf values are in NLDR data-->

# Application

We illustrate the use of the algorithm by applying it to `clusters_different_shapes_diff_num_points` dataset in `cardinalR` package. The tSNE is used in the example for the demonstration (see Figure \@ref(fig:tsne-clust)).

<!--select a dataset from cardinalR-->
```{r}
library(cardinalR)

triangular_3d_data <- tri_3d(
  n = 1250, num_noise = 2, 
  min_n = -0.05, max_n = 0.05) |>
  tibble::as_tibble()

langevitour::langevitour(triangular_3d_data)
```

```{r}
library(uwot)
library(tibble)
library(dplyr)
library(ggplot2)

## Obtain UMAP
umap_pbmc <- umap(triangular_3d_data, n_neighbors = 35, n_components =  2,
                 metric = "euclidean", min_dist = 0.5, init = "spca")

umap_pbmc <- umap_pbmc |>
  as_tibble()

names(umap_pbmc) <- c("UMAP1", "UMAP2")

umap_pbmc <- umap_pbmc |>
  mutate(ID = 1:NROW(umap_pbmc))

nldr_prism <- umap_pbmc |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5)

nldr_prism
```

<!-- diff shape clusters-->
<!-- talk with best fit-->

```{r, echo=FALSE}
clusters_different_shapes_diff_num_points <- function(sample_size = 400, with_seed = NULL, cluster_size_vec = c(50, 50, 50, 50, 100, 100), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.05, cluster_sd_non_gau = 0.1, num_dims = 7, a = 2, b = 4) {

  # To check the seed is not assigned
  if (!is.null(with_seed)) {
    set.seed(with_seed)
  }

  num_clusters <- num_gussian_clusters + num_non_gaussian_clusters

  ## Generate Gaussian clusters

  # Create a vector of possible values (0 and 1)
  values <- c(0, 1)

  # Create an expanded grid with 0's and 1's
  mean_val_grid <- tidyr::expand_grid(!!!setNames(rep(list(values), num_dims),
                                                  paste0("mean_dim", 1:num_dims)))

  # To select combinations for assigned number of clusters

  mean_val_grid_gau <- mean_val_grid |>
    dplyr::slice_sample(n = num_gussian_clusters)

  mean_val_grid_non_gau <- mean_val_grid |>
    dplyr::slice_sample(n = num_non_gaussian_clusters)

  # To generate empty tibble
  column_names <- paste0(rep("x", num_dims), 1:num_dims)
  df <- tibble(!!!setNames(rep(list(NULL), length(column_names)), column_names))

  for (i in 1:num_gussian_clusters) {

    # To filter the mean values for specific cluster
    mean_val_for_cluster <- mean_val_grid_gau |>
      dplyr::filter(dplyr::row_number() == i) |>
      unlist(use.names = FALSE)

    # Initialize an empty list to store the vectors with column
    # values
    dim_val_list <- list()

    for (j in 1:num_dims) {

      dim_val_list[[column_names[j]]] <- rnorm(cluster_size_vec[i], mean = mean_val_for_cluster[j],
                                               sd = cluster_sd_gau)

    }
    # To generate a tibble for a cluster
    df_gau_cluster <- tibble::as_tibble(dim_val_list)

    df <- dplyr::bind_rows(df, df_gau_cluster)

  }

  for (i in 1:num_non_gaussian_clusters) {

    phi <- runif(cluster_size_vec[(num_clusters - i)], max = 2*pi)
    rho <- sqrt(runif(cluster_size_vec[(num_clusters - i)]))

    # To filter the mean values for specific cluster
    presence_of_elipse_cluster <- mean_val_grid_non_gau |>
      dplyr::filter(dplyr::row_number() == i) |>
      unlist(use.names = FALSE)

    # Initialize an empty list to store the vectors with column
    # values
    dim_val_list_n <- list()

    for (j in 1:num_dims) {
      if(presence_of_elipse_cluster[j] == 1){
        dim_val_list_n[[column_names[j]]] <- sqrt(a)*rho*cos(phi) + b
        ## Surface of poolar coordinate
      } else {
        dim_val_list_n[[column_names[j]]] <- rnorm(cluster_size_vec[(num_clusters - i)], mean = 0,
                                                   sd = cluster_sd_non_gau)

      }

    }
    # To generate a tibble for a cluster
    df_non_gau_cluster <- tibble::as_tibble(dim_val_list_n)

    df <- dplyr::bind_rows(df, df_non_gau_cluster)

  }

  df

}
```

<!--need to add screen shots of the original data-->

```{r}
clust_df <- clusters_different_shapes_diff_num_points(sample_size = 1500, with_seed = NULL, cluster_size_vec = c(250, 150, 150, 150, 350, 450), num_gussian_clusters = 4, num_non_gaussian_clusters = 2, cluster_sd_gau = 0.2, cluster_sd_non_gau = 0.3, num_dims = 5, a = 2, b = 4)

langevitour::langevitour(clust_df)

clust_df <- clust_df |>
  dplyr::mutate(ID = row_number())
```

```{r}
tSNE_fit <- clust_df |>
  dplyr::select(-ID) |>
  dplyr::select(where(is.numeric)) |>
  Rtsne::Rtsne(perplexity = 39, pca = FALSE, pca_center = FALSE, normalize = FALSE, dims = 2)

tSNE_df <- tSNE_fit$Y |> 
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::mutate(ID = clust_df$ID)

names(tSNE_df) <- c("tSNE1", "tSNE2", "ID")

tSNE_df_plot <- tSNE_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.5, colour="#e41a1c", size = 0.5) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.text = element_text(size = 5),
        axis.title = element_text(size = 7))
```

```{r tsne-clust, out.width="100%", fig.cap="tSNE representation of data with perplexity: 39. Is this representation accurately capture the structure in high dimensions?"}
tSNE_df_plot
```

## Selecting parameter values for hexagonal binning

We previously introduced three key parameters, *hex_size*, *x_start*, and *y_start*, in our explanation of the algorithm. While their best choice is unknown, a visualization tool enables their tuning such that user can select reasonable values.

Consider the relationships between *total number of bins*, *starting point*, and MSE. Increase of *total number of bins* usually reduces the MSE.

<!--hex_size, starting point (x_start, y_start)-->
<!--talk about to use bin centroids or bin means-->

<!-- MSE plot to find which is the effective number of bins to use-->
```{r}
tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_df, 
                                    x = "tSNE1", y = "tSNE2"))) |>
  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
                  "tSNE2" = "scaled_tSNE2")) |>
  dplyr::mutate(ID = 1:NROW(tSNE_df))    
## Model with bin centroid
hex_size_vec <- seq(0.03, 2, by = 0.01)
vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
mse_df_gau <- dplyr::bind_rows(vec)[0, ]
mse_df_gau <- mse_df_gau |>
  dplyr::mutate_if(is.character, as.numeric)
for (i in 1:length(hex_size_vec)) {

  num_bin_list <- calc_bins(data = tsne_gau_scaled, 
            x = "tSNE1", y = "tSNE2", 
            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)

  num_bins_x <- num_bin_list$num_x
  num_bins_y <- num_bin_list$num_y

  model_object <- fit_highd_model( training_data = clust_df, 
                                   nldr_df_with_id = tsne_gau_scaled, 
                                   x = "tSNE1", y = "tSNE2", 
                                   num_bins_x = num_bins_x, 
                                   num_bins_y = num_bins_y, 
                                   x_start = NA, y_start = NA, 
                                   buffer_x = NA, buffer_y = NA, 
                                   hex_size = hex_size_vec[i],
                                   is_rm_lwd_hex = FALSE, 
                                   benchmark_to_rm_lwd_hex = NA, 
                                   col_start_2d = "tSNE", 
                                   col_start_highd = "x")

  centroid_df_training <- model_object$df_bin_centroids
  avg_df_training <- model_object$df_bin

  pred_emb_list <- predict_emb(test_data = clust_df, 
                                  df_bin_centroids = centroid_df_training, 
                                  df_bin = avg_df_training, type_NLDR = "tSNE")

  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))

  eval_list <- gen_summary(test_data = clust_df, 
                                  prediction_df = pred_df_training, 
                                  df_bin = avg_df_training, col_start = "x")

  mse_df_gau <- mse_df_gau |>
    tibble::add_row(num_bins = num_bins_x * num_bins_y,
                    aic = eval_list$aic,
                    mse = eval_list$mse,
                    num_bins_x = num_bins_x,
                    num_bins_y = num_bins_y,
                    hex_size = hex_size_vec[i])

}
## If same total number of bins occurred only select ones with minimum error
### Obtain duplicate bins
dupli_bins <- mse_df_gau |> 
  dplyr::count(num_bins) |> 
  dplyr::filter(n > 1) |> 
  dplyr::pull(num_bins)
### Group split by duplicated bins
duplicate_df_list <- mse_df_gau |>
  dplyr::filter(num_bins %in% dupli_bins) |>
  dplyr::arrange(num_bins) |>
  dplyr::group_split(num_bins)
### Obtain one row from duplicates which have lowest error and hexsize
duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
for (i in 1:length(duplicate_df_list)) {

  dd <- duplicate_df_list[[i]] |>
    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))

  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)

}
### Obtain the mse_df with not duplicated bins
not_dupli_df <- mse_df_gau |>
  dplyr::filter(!(num_bins %in% dupli_bins))
### Combine duplicated and not duplicated(corrected) bins dfs
mse_df_clust <- dplyr::bind_rows(not_dupli_df, duplicate_df) |>
  dplyr::mutate(type = "bin centroids")
```

```{r}

mse_clust_plot <- ggplot(mse_df_clust, aes(x = num_bins,
                                       y = log(mse)
)) +
  geom_point() +
  geom_line() +
   geom_vline(xintercept = 238, linetype="solid",
                color = "red", size=0.8, alpha = 0.5) +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  ylab("log(MSE)") +
  xlab("total number of bins")
```

```{r, fig.cap="Parameter tuning plots, where the best choice of parameter at a large drop in MSE. Here, these are at *total number of bins* = 238."}
mse_clust_plot
```

## Selecting parameter values for the model

<!--benchmark value to remove low density hexagons, long edges-->

## Fit the model

To perform the algorithm on the clustering data, we first scale the tSNE data. We then let *number of bins along the x-axis* be $14$, *number of bins along the y-axis* be $27$, and *hex_size* be $0.05$; see [Selecting parameter values] for further discussion and guidance regarding selection of these choices.

<!-- model with bin centroids-->
```{r}

## Decide by looking at MSE plot
num_bins_x_tsne_gau <- 14
num_bins_y_tsne_gau <- 27
hex_size_tsne_gau <- 0.05
x_start_tsne_gau <- -0.0481
y_start_tsne_gau <- 0.00833
## non-empty:80
hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
                      num_bins_y = num_bins_y_tsne_gau, x_start = x_start_tsne_gau, y_start = y_start_tsne_gau,
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "tSNE")
## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()
hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()
tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))

model_object_tsne_gau <- fit_highd_model( training_data = clust_df, 
                                 nldr_df_with_id = tsne_gau_scaled, 
                                 x = "tSNE1", y = "tSNE2", 
                                 num_bins_x = num_bins_x_tsne_gau, 
                                 num_bins_y = num_bins_y_tsne_gau, 
                                 x_start = x_start_tsne_gau, y_start = y_start_tsne_gau, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_tsne_gau,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "tSNE", 
                                 col_start_highd = "x")
df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
df_bin_tsne_gau <- model_object_tsne_gau$df_bin
## Triangulate bin centroids
tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
## Compute 2D distances
distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))
# distance_plot <- plot_dist(distance_tsne_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot
## To find the benchmark value
benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value =
benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

trimesh_removed_tsne_gau

df_all_tsne_gau <- dplyr::bind_cols(clust_df |> dplyr::select(-ID), tsne_data_with_hb_id_gau)

### Define type column
df <- df_all_tsne_gau |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin_tsne_gau |>
  dplyr::filter(hb_id %in% df_bin_centroids_tsne_gau$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean
## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_tsne_gau$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)
df_exe <- dplyr::bind_rows(df_b, df)
## Set the maximum difference as the criteria
distance_df_small_edges <- distance_tsne_gau |>
  dplyr::filter(distance < benchmark_tsne_gau)
## Since erase brushing is considerd.
langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                         lineFrom = distance_df_small_edges$from,
                         lineTo = distance_df_small_edges$to,
                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.5, NROW(df))),
                         levelColors = c("#6a3d9a", "#33a02c"))
```

What can learn from the model:

- There are five clusters. Four of them are Gaussian and two of them are long 2D clusters. Gaussian clusters are close to each other while others are far away.

- Model preserved local and global structure.

- Problem: Couldn't capture the variance within the clusters.

- There is one edge that is not connected in 2D triangular mesh.

# Discussion

This paper presents the R package `quollr` to develop a way to take the fitted model, as represented by the positions of points in 2D, and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.

The paper includes a clustering example to illustrate how `quollr` is useful to assess which NLDR technique and which hyperparameter choice gives the most accurate representation. In addition, how to select parameters for hexagonal binning and fitting model are explained.

Possible future improvements would be...<!--assess the preservation of local and glocal structure w.r.t 2D and high-D distance comparison-->

This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of $p-D$ data.

# Acknowledgements

This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template. The source code for reproducing this paper can be found at: <https://github.com/JayaniLakshika/paper-quollr>.

### 2024-06-04 paper-quollr.Rmd
Select the dataset to evaluate from cardinalR

@@
 library(tools)
+library(cardinalR)
 
@@
 )
+
+interior_annotation <- function(label, position = c(0.92, 0.92)) {
+  annotation_custom(grid::textGrob(label = label,
+      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
+      gp = grid::gpar(cex = 1, col="grey70")))
+}
 ```
@@
 
-## Constructing the 2-D model
+## Constructing the $2\text{-}D$ model
+
+Constructing the $2\text{-}D$ model mainly contains (i) scaling, (ii) computing hexagon grid configurations, (iii) binning, and (iv) indicating neighbors by line segments connecting centroids.
 
 ### Scaling the data
+
+```{r}
+scaled_nldr_data <- gen_scaled_data(data = s_curve_noise_umap)
+
+scaled_nldr_data
+```
 
@@
 <!--with prism data-->
+<!-- my plan is to discuss which the appropriate representation from different parameter choices from tSNE-->
+<!--Why tSNE? thinking about the usability-->
+
+```{r}
+triangular_3d_data <- tri_3d(
+  n = 1250, num_noise = 2, 
+  min_n = -0.05, max_n = 0.05) |>
+  as_tibble()
+
+colnames(triangular_3d_data) <- paste0("x", 1:NCOL(triangular_3d_data))
+
+langevitour::langevitour(triangular_3d_data)
+```
+
+```{r}
+sc_ltr_pos <- c(0.08, 0.96)
+
+triangular_3d_data <- triangular_3d_data |>
+  mutate(ID = row_number())
+
+tSNE_fit <- triangular_3d_data |>
+  select(-ID) |>
+  Rtsne(perplexity = sqrt(NROW(triangular_3d_data)))
+
+tsne_prism <- tSNE_fit$Y |>
+  as_tibble() |>
+  mutate(ID = row_number())
+
+colnames(tsne_prism) <- c("tSNE1", "tSNE2", "ID")
+
+prism_scaled_obj <- gen_scaled_data(
+  data = tsne_prism)
+tsne_prism_scaled <- prism_scaled_obj$scaled_nldr
+
+nldr_prism <- tsne_prism_scaled |> 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5)
+
+nldr_prism
+```
+
+<!--To generate errors for different total number of bins-->
+```{r}
+## To initialize number of bins along the x-axis
+bin1_vec_prism <- 2:25
+
+lim1 <- prism_scaled_obj$lim1
+lim2 <- prism_scaled_obj$lim2
+r2_prism <- diff(lim2)/diff(lim1)
+
+error_prism <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_prism) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2
+
+  prism_model <- fit_highd_model(
+  training_data = triangular_3d_data,
+  emb_df = tsne_prism_scaled,
+  bin1 = xbins,
+  r2 = r2_prism,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = TRUE,
+  col_start_highd = "x"
+)
+
+  df_bin_centroids_prism <- prism_model$df_bin_centroids
+  df_bin_prism <- prism_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_prism,
+    df_bin = df_bin_prism,
+    training_data = triangular_3d_data,
+    newdata = NULL,
+    type_NLDR = "tSNE",
+    col_start = "x") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_prism))
+
+  error_prism <- bind_rows(error_prism, error_df)
+
+}
+
+mse_prism_b <- ggplot(error_prism, 
+                     aes(x = b, 
+                         y = log(MSE))) + 
+  geom_point(size = 1) +
+  geom_line() + 
+  geom_vline(xintercept = 420, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  ylab("log(MSE)") +
+  xlab("b") +
+  theme(axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7, angle = 90))
+
+mse_prism_b
+```
+
+```{r}
+## Compute hexbin parameters
+num_bins_x_prism <- 20
+lim1 <- prism_scaled_obj$lim1
+lim2 <- prism_scaled_obj$lim2
+r2_prism <- diff(lim2)/diff(lim1)
+
+prism_model <- fit_highd_model(
+  training_data = triangular_3d_data,
+  emb_df = tsne_prism_scaled,
+  bin1 = num_bins_x_prism,
+  r2 = r2_prism,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = TRUE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_prism <- prism_model$df_bin_centroids
+df_bin_prism <- prism_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_prism <- tri_bin_centroids(
+  df_bin_centroids_prism, x = "c_x", y = "c_y")
+tr_from_to_df_prism <- gen_edges(
+  tri_object = tr1_object_prism)
+
+## Compute 2D distances
+distance_prism <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_prism,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_prism <- find_lg_benchmark(
+  distance_edges = distance_prism,
+  distance_col = "distance")
+
+trimesh_removed_prism <- vis_rmlg_mesh(
+  distance_edges = distance_prism,
+  benchmark_value = benchmark_prism,
+  tr_coord_df = tr_from_to_df_prism,
+  distance_col = "distance") #+
+  #xlim(sc_xlims) + ylim(sc_ylims) +
+  #interior_annotation("a", sc_ltr_pos)
+
+trimesh_removed_prism
+```
+
+```{r}
+## Hexagonal binning to have regular hexagons
+hb_obj_prism <- hex_binning(
+  data = tsne_prism_scaled,
+  bin1 = num_bins_x_prism,
+  r2 = r2_prism)
+
+tsne_data_with_hb_id <- hb_obj_prism$data_hb_id
+
+df_all_prism <- dplyr::bind_cols(triangular_3d_data |> dplyr::select(-ID),
+                                tsne_data_with_hb_id)
+
+### Define type column
+df <- df_all_prism |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_prism |>
+  dplyr::filter(hb_id %in% df_bin_centroids_prism$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_prism$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+df_exe <- dplyr::bind_rows(df_b, df)
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_prism |>
+  dplyr::filter(distance < benchmark_prism)
+## Since erase brushing is considerd.
+
+langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = distance_df_small_edges$from,
+                         lineTo = distance_df_small_edges$to,
+                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.8, NROW(df))),
+                         levelColors = c("#6a3d9a", "#33a02c"))
+```
 

### 2024-06-04 paper-quollr.Rmd
Adding suggestion for appropriate model for prism
Added intro for prism data

@@
 
-<!--with prism data-->
-<!-- my plan is to discuss which the appropriate representation from different parameter choices from tSNE-->
-<!--Why tSNE? thinking about the usability-->
+<!--add intro to prism data-->
+
+To illustrate the method, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data".
+This dataset is generated using the `tri_3d` function from the `cardinalR` package.
+The dataset generation starts with the initialization of a starting point $p_0 = (x_0, y_0, z_0)$ randomly selected within the $3\text{-}D$ space using a uniform distribution.
+Four fixed corner points of a tetrahedron are defined as $c_1 = (0, 0, 0)$, $c_2 = (1, 0, 0)$, $c_3 = (0.5, 1, 0)$, and $c_4 = (0.5, 0.5, 1)$.
+For each point $p_i$ where $i = 1, 2, \ldots, n$, one of the corner points $c_j$ (where $j \in \{1, 2, 3, 4\}$) is selected randomly, and the new point $p_i$ is computed as the midpoint between the current point $p_{i-1}$ and the selected corner point $c_j$: $p_i = \frac{p_{i-1} + c_j}{2}$.
+This iterative process ensures that each new point moves closer to one of the corners, creating a fractal-like triangular distribution of points.
+The coordinates of each generated point $p_i$ form the variables $X_1, X_2, X_3$.
+The remaining variables $X_4, X_5$ are all uniform error, with small variance.
+We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+
+<!-- hoe the data are generated and from which package I got the data-->
 
@@
 ```{r}
-sc_ltr_pos <- c(0.08, 0.96)
-
 triangular_3d_data <- triangular_3d_data |>
   mutate(ID = row_number())
+```
+
+<!--added, I selected tSNE with perplexity = 35 to explore and -->
+
+```{r}
+sc_ltr_pos <- c(0.08, 0.96)
 
@@
   select(-ID) |>
-  Rtsne(perplexity = sqrt(NROW(triangular_3d_data)))
+  Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))
 
@@
 ```
+
+## Selecting parameter values for the model
+
+<!--with prism data-->
+<!-- my plan is to discuss which the appropriate representation from different parameter choices from tSNE-->
+<!--Why tSNE? thinking about the usability-->
 

### 2024-06-04 paper-quollr.Rmd
Added plan for the application

@@
 
-<!-- hoe the data are generated and from which package I got the data-->
+<!-- how the data are generated and from which package I got the data-->
 
@@
 
-nldr_prism <- tsne_prism_scaled |> 
+nldr_prism1 <- tsne_prism_scaled |> 
   ggplot(aes(x = tSNE1, y = tSNE2)) + 
@@
 
-nldr_prism
+nldr_prism1
 ```
 
-## Selecting parameter values for the model
+## Deciding an appropriate fit
+
+<!--select based on looking at b_non_empty vs log(MSE)-->
+<!--select a choice for b_non_empty-->
+
+## Fit the appropriate model
+
+### Selecting parameter values for the model
+
+#### Choice of bins
+
+#### Removal of low-density bins
+
+#### Removing long edges
 

### 2024-06-05 paper-quollr.Rmd
Add different representations for prism data

@@
 library(Rtsne)
+library(uwot)
 library(readr)
@@
 
-<!--add intro to prism data-->
-
 To illustrate the method, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data".
@@
 We would consider $T=(X_1, X_2, X_3)$ to be the true model.
-
-<!-- how the data are generated and from which package I got the data-->
 
@@
 
-<!--added, I selected tSNE with perplexity = 35 to explore and -->
-
 ```{r}
 sc_ltr_pos <- c(0.08, 0.96)
+```
 
+<!--tSNE with perplexity = 35-->
+```{r}
 tSNE_fit <- triangular_3d_data |>
@@
   geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
 
-nldr_prism1
+```{r}
+tSNE_fit2 <- triangular_3d_data |>
+  select(-ID) |>
+  Rtsne(perplexity = 6)
+
+tsne_prism2 <- tSNE_fit2$Y |>
+  as_tibble() |>
+  mutate(ID = row_number())
+
+colnames(tsne_prism2) <- c("tSNE1", "tSNE2", "ID")
+
+prism_scaled_obj2 <- gen_scaled_data(
+  data = tsne_prism2)
+tsne_prism_scaled2 <- prism_scaled_obj2$scaled_nldr
+
+nldr_prism2 <- tsne_prism_scaled2 |> 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+```{r}
+tSNE_fit3 <- triangular_3d_data |>
+  select(-ID) |>
+  Rtsne(perplexity = 58)
+
+tsne_prism3 <- tSNE_fit3$Y |>
+  as_tibble() |>
+  mutate(ID = row_number())
+
+colnames(tsne_prism3) <- c("tSNE1", "tSNE2", "ID")
+
+prism_scaled_obj3 <- gen_scaled_data(
+  data = tsne_prism3)
+tsne_prism_scaled3 <- prism_scaled_obj3$scaled_nldr
+
+nldr_prism3 <- tsne_prism_scaled3 |> 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+<!--UMAP with n_neighbors = 35-->
+```{r}
+umap_pbmc <- triangular_3d_data |>
+  select(-ID) |>
+  umap(n_neighbors = 35, n_components =  2, 
+       metric = "euclidean", min_dist = 0.5, init = "spca")
+
+umap_pbmc <- umap_pbmc |>
+  as_tibble()
+
+names(umap_pbmc) <- c("UMAP1", "UMAP2")
+
+umap_pbmc <- umap_pbmc |>
+  mutate(ID = 1:NROW(umap_pbmc))
+
+nldr_prism4 <- umap_pbmc |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+```{r}
+umap_pbmc2 <- triangular_3d_data |>
+  select(-ID) |>
+  umap(n_neighbors = 15, n_components =  2, 
+       metric = "euclidean", min_dist = 0.01, init = "spca")
+
+umap_pbmc2 <- umap_pbmc2 |>
+  as_tibble()
+
+names(umap_pbmc2) <- c("UMAP1", "UMAP2")
+
+umap_pbmc2 <- umap_pbmc2 |>
+  mutate(ID = 1:NROW(umap_pbmc))
+
+nldr_prism5 <- umap_pbmc2 |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+```{r}
+umap_pbmc3 <- triangular_3d_data |>
+  select(-ID) |>
+  umap(n_neighbors = 69, n_components =  2, 
+       metric = "euclidean", min_dist = 0.6, init = "spca")
+
+umap_pbmc3 <- umap_pbmc3 |>
+  as_tibble()
+
+names(umap_pbmc3) <- c("UMAP1", "UMAP2")
+
+umap_pbmc3 <- umap_pbmc3 |>
+  mutate(ID = 1:NROW(umap_pbmc))
+
+nldr_prism6 <- umap_pbmc3 |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+```{r}
+nldr_prism1 + nldr_prism2 + nldr_prism3 + 
+  nldr_prism4 + nldr_prism5 + nldr_prism6 +
+  plot_layout(ncol=3)
 ```

### 2024-06-05 paper-quollr.Rmd
Need to work on the best model

--- 2024-06-04 paper-quollr.Rmd
@@
 
-<!--add intro to prism data-->
-
-To illustrate the method, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data".
+To illustrate the algorithm, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data".
 This dataset is generated using the `tri_3d` function from the `cardinalR` package.
@@
 
-<!--added, I selected tSNE with perplexity = 35 to explore and -->
+```{r}
+sc_ltr_pos <- c(0.08, 0.96)
+```
+
+<!--tSNE with perplexity = 6-->
+```{r}
+tSNE_fit1 <- triangular_3d_data |>
+  select(-ID) |>
+  Rtsne(perplexity = 6)
+
+tsne_prism1 <- tSNE_fit1$Y |>
+  as_tibble() |>
+  mutate(ID = row_number())
+
+colnames(tsne_prism1) <- c("tSNE1", "tSNE2", "ID")
+
+prism_scaled_obj_tsne1 <- gen_scaled_data(
+  data = tsne_prism1)
+tsne_prism_scaled1 <- prism_scaled_obj_tsne1$scaled_nldr
+
+nldr_prism1 <- tsne_prism_scaled1 |> 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+<!--tSNE with perplexity = 35-->
+```{r}
+tSNE_fit2 <- triangular_3d_data |>
+  select(-ID) |>
+  Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))
+
+tsne_prism2 <- tSNE_fit2$Y |>
+  as_tibble() |>
+  mutate(ID = row_number())
+
+colnames(tsne_prism2) <- c("tSNE1", "tSNE2", "ID")
+
+prism_scaled_obj_tsne2 <- gen_scaled_data(
+  data = tsne_prism2)
+tsne_prism_scaled2 <- prism_scaled_obj_tsne2$scaled_nldr
+
+nldr_prism2 <- tsne_prism_scaled2 |> 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+<!--tSNE with perplexity = 58-->
+```{r}
+tSNE_fit3 <- triangular_3d_data |>
+  select(-ID) |>
+  Rtsne(perplexity = 58)
+
+tsne_prism3 <- tSNE_fit3$Y |>
+  as_tibble() |>
+  mutate(ID = row_number())
+
+colnames(tsne_prism3) <- c("tSNE1", "tSNE2", "ID")
+
+prism_scaled_obj_tsne3 <- gen_scaled_data(
+  data = tsne_prism3)
+tsne_prism_scaled3 <- prism_scaled_obj_tsne3$scaled_nldr
+
+nldr_prism3 <- tsne_prism_scaled3 |> 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+<!--UMAP with n_neighbors = 15 and min_dist = 0.3-->
+```{r}
+umap_prism1 <- triangular_3d_data |>
+  select(-ID) |>
+  umap(n_neighbors = 15, n_components =  2, 
+       metric = "euclidean", min_dist = 0.3, init = "spca")
+
+umap_prism1 <- umap_prism1 |>
+  as_tibble()
+
+names(umap_prism1) <- c("UMAP1", "UMAP2")
+
+umap_prism1 <- umap_prism1 |>
+  mutate(ID = 1:NROW(umap_prism1))
+
+prism_scaled_obj_umap1 <- gen_scaled_data(
+  data = umap_prism1)
+umap_prism_scaled1 <- prism_scaled_obj_umap1$scaled_nldr
+
+nldr_prism4 <- umap_prism_scaled1 |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+<!--UMAP with n_neighbors = 35 and min_dist = 0.5-->
+```{r}
+umap_prism2 <- triangular_3d_data |>
+  select(-ID) |>
+  umap(n_neighbors = 35, n_components =  2, 
+       metric = "euclidean", min_dist = 0.5, init = "spca")
+
+umap_prism2 <- umap_prism2 |>
+  as_tibble()
+
+names(umap_prism2) <- c("UMAP1", "UMAP2")
+
+umap_prism2 <- umap_prism2 |>
+  mutate(ID = 1:NROW(umap_prism2))
+
+prism_scaled_obj_umap2 <- gen_scaled_data(
+  data = umap_prism2)
+umap_prism_scaled2 <- prism_scaled_obj_umap2$scaled_nldr
+
+nldr_prism5 <- umap_prism_scaled2 |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
+
+<!--UMAP with n_neighbors = 69 and min_dist = 0.6-->
+```{r}
+umap_prism3 <- triangular_3d_data |>
+  select(-ID) |>
+  umap(n_neighbors = 69, n_components =  2, 
+       metric = "euclidean", min_dist = 0.6, init = "spca")
+
+umap_prism3 <- umap_prism3 |>
+  as_tibble()
+
+names(umap_prism3) <- c("UMAP1", "UMAP2")
+
+umap_prism3 <- umap_prism3 |>
+  mutate(ID = 1:NROW(umap_prism3))
+
+prism_scaled_obj_umap3 <- gen_scaled_data(
+  data = umap_prism3)
+umap_prism_scaled3 <- prism_scaled_obj_umap3$scaled_nldr
+
+nldr_prism6 <- umap_prism_scaled3 |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
+```
 
 ```{r}
-sc_ltr_pos <- c(0.08, 0.96)
+nldr_prism1 + nldr_prism2 + nldr_prism3 + 
+  nldr_prism4 + nldr_prism5 + nldr_prism6 +
+  plot_layout(ncol=3)
+```
+
+<!--need to add details regarding which differences can I see with different layouts-->
+
+## Deciding an appropriate fit
+
+<!--select based on looking at b_non_empty vs log(Error)-->
+<!--select a choice for b_non_empty-->
+
+<!--compute error with tSNE and different perplexity values-->
+```{r}
+error_prism_tsne <- data.frame(matrix(nrow = 0, ncol = 0))
+
+## To initialize number of bins along the x-axis
+bin1_vec_prism <- 2:25
+
+perplexity_vec <- c(6, 35, 58)
+
+for (perplexity_val in perplexity_vec) {
 
@@
     select(-ID) |>
-  Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))
+    Rtsne(perplexity = perplexity_val)
 
@@
 
-nldr_prism1 <- tsne_prism_scaled |> 
-  ggplot(aes(x = tSNE1, y = tSNE2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5)
-
-nldr_prism1
-```
-
-## Deciding an appropriate fit
-
-<!--select based on looking at b_non_empty vs log(MSE)-->
-<!--select a choice for b_non_empty-->
-
-## Fit the appropriate model
-
-### Selecting parameter values for the model
-
-#### Choice of bins
-
-#### Removal of low-density bins
-
-#### Removing long edges
-
-<!--with prism data-->
-<!-- my plan is to discuss which the appropriate representation from different parameter choices from tSNE-->
-<!--Why tSNE? thinking about the usability-->
-
-<!--To generate errors for different total number of bins-->
-```{r}
-## To initialize number of bins along the x-axis
-bin1_vec_prism <- 2:25
-
   lim1 <- prism_scaled_obj$lim1
@@
   r2_prism <- diff(lim2)/diff(lim1)
-
-error_prism <- data.frame(matrix(nrow = 0, ncol = 0))
 
@@
              b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_prism))
+             b_non_empty = NROW(df_bin_centroids_prism),
+             type = paste0("perplexity: ", perplexity_val))
 
-  error_prism <- bind_rows(error_prism, error_df)
+    error_prism_tsne <- bind_rows(error_prism_tsne, error_df)
 
@@
 
-mse_prism_b <- ggplot(error_prism, 
-                     aes(x = b, 
-                         y = log(MSE))) + 
-  geom_point(size = 1) +
-  geom_line() + 
-  geom_vline(xintercept = 420, linetype="solid", 
-             color = "black", linewidth=0.8, alpha = 0.5) +
-  ylab("log(MSE)") +
-  xlab("b") +
-  theme(axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, angle = 90))
-
-mse_prism_b
+}
 ```
 
+<!--compute error with UMAP and different parameter values-->
 ```{r}
-## Compute hexbin parameters
-num_bins_x_prism <- 20
+error_prism_umap <- data.frame(matrix(nrow = 0, ncol = 0))
+
+## To initialize number of bins along the x-axis
+bin1_vec_prism <- 2:25
+
+param_list <- list(param1 = c(15, 0.3), param2 = c(35, 0.5), param3 = c(69, 0.6))
+
+for (i in 1:length(param_list)) {
+
+  n_neighbors_val <- param_list[[i]][1]
+  min_dist_val <- param_list[[i]][2]
+
+  umap_prism <- triangular_3d_data |>
+    select(-ID) |>
+    umap(n_neighbors = n_neighbors_val, n_components =  2, 
+         metric = "euclidean", min_dist = min_dist_val, init = "spca")
+
+  umap_prism <- umap_prism |>
+    as_tibble()
+
+  names(umap_prism) <- c("UMAP1", "UMAP2")
+
+  umap_prism <- umap_prism |>
+    mutate(ID = 1:NROW(umap_prism))
+
+  prism_scaled_obj <- gen_scaled_data(
+    data = umap_prism)
+  umap_prism_scaled <- prism_scaled_obj$scaled_nldr
+
   lim1 <- prism_scaled_obj$lim1
@@
 
+  for (xbins in bin1_vec_prism) {
+
+    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2
+
     prism_model <- fit_highd_model(
       training_data = triangular_3d_data,
-  emb_df = tsne_prism_scaled,
+      emb_df = umap_prism_scaled,
+      bin1 = xbins,
+      r2 = r2_prism,
+      is_bin_centroid = TRUE,
+      is_rm_lwd_hex = TRUE,
+      col_start_highd = "x"
+    )
+
+    df_bin_centroids_prism <- prism_model$df_bin_centroids
+    df_bin_prism <- prism_model$df_bin
+
+    ## Compute error
+    error_df <- glance(
+      df_bin_centroids = df_bin_centroids_prism,
+      df_bin = df_bin_prism,
+      training_data = triangular_3d_data,
+      newdata = NULL,
+      type_NLDR = "UMAP",
+      col_start = "x") |>
+      mutate(bin1 = xbins,
+             bin2 = bin2,
+             b = bin1 * bin2,
+             b_non_empty = NROW(df_bin_centroids_prism),
+             type = paste0("n_neighbors: ", n_neighbors_val, 
+                           " min_dist: ", min_dist_val))
+
+    error_prism_umap <- bind_rows(error_prism_umap, error_df)
+
+  }
+
+}
+```
+
+```{r}
+error_prism <- bind_rows(error_prism_tsne, error_prism_umap) 
+error_prism <- error_prism |>
+  filter(b <= 625)
+```
+
+```{r}
+error_plot_prism <- ggplot(error_prism, 
+                           aes(x = b_non_empty, 
+                               y = log(Error), 
+                               group = type, 
+                               colour = type)) + 
+  geom_point(size = 1) +
+  geom_line() + 
+  geom_vline(xintercept = 112, linetype="solid",
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3',
+                              '#ff7f00','#a6cee3','#a65628','#f781bf')) +
+  ylab("log(Error)") +
+  xlab("number of non-empty bins") +
+  theme(axis.text.x = element_text(size = 5),
+        axis.text.y = element_text(size = 5),
+        axis.title.x = element_text(size = 5),
+        axis.title.y = element_text(size = 5, angle = 90),
+        legend.position = "bottom")
+
+error_plot_prism
+```
+
+<!--selected tSNE perplexity: 35-->
+
+## Fit the appropriate model
+
+```{r}
+## Compute hexbin parameters
+num_bins_x_prism <- 21
+lim1 <- prism_scaled_obj_tsne2$lim1
+lim2 <- prism_scaled_obj_tsne2$lim2
+r2_prism <- diff(lim2)/diff(lim1)
+
+prism_model <- fit_highd_model(
+  training_data = triangular_3d_data,
+  emb_df = tsne_prism_scaled2,
   bin1 = num_bins_x_prism,
@@
 hb_obj_prism <- hex_binning(
-  data = tsne_prism_scaled,
+  data = tsne_prism_scaled2,
   bin1 = num_bins_x_prism,
@@
 
+### Selecting parameter values for the model
+
+#### Choice of bins
+
+#### Removal of low-density bins
+
+#### Removing long edges
+
 # Discussion

### 2024-06-07 paper-quollr.Rmd
Fit an appropriate model for the prism data
Added scratch details about what I learnt from the model

@@
 
+### Construct model in $2\text{-}D$
+
+### Selecting parameter values for the model
+
+#### Choice of bins
+
+#### Removal of low-density bins
+
+#### Removing long edges
+
 ```{r}
@@
 
+### Lifting the model into $p\text{-}D$
+
 ```{r}
@@
 
-### Selecting parameter values for the model
+What can I see in high-D?
 
-#### Choice of bins
+- Four clusters that are really close
 
-#### Removal of low-density bins
+What can I see in 2D layout?
 
-#### Removing long edges
+- Several clusters (more that what see in high-D)
+
+What can learn from the model?
+
+- There are four clusters, but not as the same cluster what data shows.
+In 2D, the two of actual clusters (high-D clusters) are really close.
+Also, one of the big high-D cluster is separated to two sub clusters.
 

### 2024-06-07 paper-quollr.Rmd
Added scratch for the algorithm writing

@@
 
+The following demonstration of the package's functionality assumes `quollr` has been loaded.
+We also want to load the built-in data sets `s_curve_noise_training`, `s_curve_noise_test` and `s_curve_noise_umap`.
+
+`s_curve_noise_training` is a $3-D$ S-curve dataset with additional four noise dimensions which is used to train the model.
+`s_curve_noise_test` is used for prediction.
+`s_curve_noise_umap` is the UMAP $2-D$ embedding for `s_curve_noise_training` dataset.
+Each dataset contains a unique ID column.
+
+### Scaling the data
+
+```{r, echo=TRUE}
+scaled_nldr_obj_scurve <- gen_scaled_data(data = s_curve_noise_umap)
+
+s_curve_noise_umap_scaled <- scaled_nldr_obj_scurve$scaled_nldr
+
+lim1 <- scaled_nldr_obj_scurve$lim1
+lim2 <- scaled_nldr_obj_scurve$lim2
+r2 <- diff(lim2)/diff(lim1) 
+```
+
+The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
+When constructing the 2D model, the user can choose either to fit the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
+
+If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
+For example, if the user wishes to use already binning data, the `extract_hexbin_centroids()` function can be used directly.
+
+The number of bins along the x axis, the ratio of the ranges of the original embedding components, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model`, if they are not provided.
+They are created as they are needed throughout the following example.
+
+```{r, echo=TRUE}
+fit_highd_model(
+  training_data = s_curve_noise_training,
+  emb_df = s_curve_noise_umap_scaled, 
+  bin1 = 15, r2 = r2, 
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = TRUE,
+  benchmark_to_rm_lwd_hex = NULL,
+  col_start_highd = "x")
+```
+
 ## Constructing the $2\text{-}D$ model
@@
 Constructing the $2\text{-}D$ model mainly contains (i) scaling, (ii) computing hexagon grid configurations, (iii) binning, and (iv) indicating neighbors by line segments connecting centroids.
-
-### Scaling the data
-
-```{r}
-scaled_nldr_data <- gen_scaled_data(data = s_curve_noise_umap)
-
-scaled_nldr_data
-```
 

### 2024-06-07 paper-quollr.Rmd
Added scratch to find defaults

@@
   benchmark_to_rm_lwd_hex = NULL,
-  col_start_highd = "x")
+  col_start_highd = "x"
+  )
 ```
@@
 
+```{r, echo=TRUE}
+calc_bins_y(
+  bin1 = 10, 
+  r2 = r2
+  )
+```
+
 ### Binning the data
+
+```{r, echo=TRUE}
+hex_binning(
+  data = s_curve_noise_umap_scaled, 
+  bin1 = 10, 
+  r2 = r2
+  )
+```
 
@@
 
+```{r, echo=TRUE, eval=FALSE}
+tr1_object <- tri_bin_centroids(
+  hex_df = df_bin_centroids, 
+  x = "c_x", 
+  y = "c_y"
+  )
+
+tr_from_to_df <- gen_edges(
+  tri_object = tr1_object
+  )
+```
+
 ## Lifting the model into high dimensions
+
+```{r, echo=TRUE, eval=FALSE}
+umap_data_with_hb_id <- hb_obj$data_hb_id
+
+df_all <- bind_cols(
+  s_curve_noise_training |> dplyr::select(-ID), 
+  umap_data_with_hb_id
+  )
+
+df_bin <- avg_highd_data(
+  data = df_all, 
+  col_start = "x"
+  )
+```
 
@@
 <!--discuss about default settings-->
+
+### Choice of bins
+
+### Removal of low-density bins
+
+```{r, echo=TRUE, eval=FALSE}
+
+find_low_dens_hex(
+  df_bin_centroids_all = df_bin_centroids, 
+  bin1 = num_bins_x, 
+  df_bin_centroids_low = df_bin_centroids_low
+  )
+
+```
+
+### Removing long edges
+
+```{r, echo=TRUE, eval=FALSE}
+find_lg_benchmark(
+  distance_edges = distance_df, 
+  distance_col = "distance"
+  )
+```
 
@@
 
+```{r, echo=TRUE, eval=FALSE}
+vis_lg_mesh(
+  distance_edges = distance_df, 
+  benchmark_value = 0.75, 
+  tr_coord_df = tr_from_to_df, 
+  distance_col = "distance"
+  )
+```
+
+```{r, echo=TRUE, eval=FALSE}
+vis_rmlg_mesh(
+  distance_edges = distance_df, 
+  benchmark_value = 0.75, 
+  tr_coord_df = tr_from_to_df, 
+  distance_col = "distance"
+  )
+```
+
 ### $p\text{-}D$ model visualization
+
+```{r, echo=TRUE, eval=FALSE}
+show_langevitour(
+  df = df_all, 
+  df_b = df_bin, 
+  df_b_with_center_data = df_bin_centroids, 
+  benchmark_value = 0.75, 
+  distance = distance_df, 
+  distance_col = "distance", 
+  use_default_benchmark_val = FALSE, 
+  col_start = "x"
+  )
+```
 

### 2024-06-22 paper-quollr.Rmd
Update the abstract

@@
 ---
-title: "quollr: An R Package for Visualizing 2D Models from Nonlinear Dimension Reductions in High Dimensional Space"
+title: "quollr: An R Package for Visualizing $2-D$ Models from Non-linear Dimension Reductions in High Dimensional Space"
 abstract: >
-  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional data in low-diemnsioanl space.
-However, their complexity and (hyper)parameter choices may lead to distrustful or misleading results.
-The R package \CRANpkg{quollr} is developed as a new tool to help to determine which method, which (hyper)parameter choice provide the most accurate representation of $p-D$ data.
-Clustering data from \CRANpkg{cardinalR} package, is used to illustrate the algorithm and its use within the package.
+  Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (p\text{-}D) by applying a non-linear transformation.
+However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices.
+It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
+The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data.
+The `triangular_3d_data` data from the \CRANpkg{cardinalR} package is used to illustrate the algorithm and its application within the package.
 draft: true
@@
 In next section, introduces the implementation of `quollr` package on CRAN, including demonstration of the package's key functions and visualization capabilities.
-We illustrate the algorithm's functionality to study clustering data set in **Application** section, and describe a visual heuristic to describe parameter selection.
+We illustrate the algorithm's functionality to study about clustering data structure in **Application** section, and describe a visual heuristic to describe parameter selection.
 Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our algorithm.

### 2024-06-22 paper-quollr.Rmd
Update the main function

@@
 
-The following demonstration of the package's functionality assumes `quollr` has been loaded.
-We also want to load the built-in data sets `s_curve_noise_training`, `s_curve_noise_test` and `s_curve_noise_umap`.
+The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `s_curve_noise_training` and `s_curve_noise_umap`.
 
-`s_curve_noise_training` is a $3-D$ S-curve dataset with additional four noise dimensions which is used to train the model.
-`s_curve_noise_test` is used for prediction.
-`s_curve_noise_umap` is the UMAP $2-D$ embedding for `s_curve_noise_training` dataset.
-Each dataset contains a unique ID column.
+`s_curve_noise_training` is a $3\text{-}D$ S-curve data set with additional four noise dimensions which is used to train the model.
+`s_curve_noise_umap` is the UMAP $2\text{-}D$ embedding for `s_curve_noise_training` data set.
+Each data set contains a unique ID column.
 
 ### Scaling the data
+
+The algorithm begins by scaling NLDR data to a standard scale using the `gen_scaled_data()` function.
+This function standardizes the data and provides the original limits of embeddings and the aspect ratio.
 
@@
 The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
-When constructing the 2D model, the user can choose either to fit the 2D model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
+When constructing the $2\text{-}D$ model, the user can choose either to fit the $2\text{-}D$ model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
 
@@
 
-The number of bins along the x axis, the ratio of the ranges of the original embedding components, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model`, if they are not provided.
+The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model`, if they are not provided.
 They are created as they are needed throughout the following example.
@@
   emb_df = s_curve_noise_umap_scaled, 
-  bin1 = 15, r2 = r2, 
+  bin1 = 12, 
+  r2 = r2,
+  q = 0.07,
   is_bin_centroid = TRUE,

### 2024-06-22 paper-quollr.Rmd
Add details regarding the binning data

@@
 
-The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model`, if they are not provided.
+The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model()`, if they are not provided.
 They are created as they are needed throughout the following example.
+The function `fit_highd_model()` provides the fitted model in $2\text{-}D$, and $p\text{-}D$.
 
@@
 
-Constructing the $2\text{-}D$ model mainly contains (i) scaling, (ii) computing hexagon grid configurations, (iii) binning, and (iv) indicating neighbors by line segments connecting centroids.
+Constructing the $2\text{-}D$ model mainly contains (i) binning data, (ii) obtaining bin centroids, and (iii) indicating neighbors by line segments connecting centroids.
 
@@
 
+The configurations of the hexagonal grid is defined by the number of bins in each direction.
+To find the number of bins along the y-axis, `calc_bins_y()` is used.
+This function takes as input the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
+Additionally, this function provides the bin width.
+
 ```{r, echo=TRUE}
 calc_bins_y(
-  bin1 = 10, 
-  r2 = r2
+  bin1 = 12, 
+  r2 = r2,
+  q = 0.07
   )
@@
 
+Points are allocated to the bins they fall into based on the nearest centroid.
+The main steps of the hexagonal binning algorithm can be executed using the `hex_binning()` function, or they can be run separately for greater flexibility.
+The parameters used within `hex_binning()` include the scaled NLDR data, the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
+The output is an object of the `hex_bin_obj` class, which contains the number of bins in each direction, the coordinates of the hexagonal grid starting point, the details of bin centroids, the coordinates of bins, embedding components with their corresponding hexagon IDs, hex bins with their corresponding standardized counts, the total number of bins, the number of non-empty bins, and the points within each hexagon.
+
 ```{r, echo=TRUE}
@@
   data = s_curve_noise_umap_scaled, 
-  bin1 = 10, 
-  r2 = r2
+  bin1 = 12, 
+  r2 = r2,
+  q = 0.07
   )
 ```
+
+### Remove low-density hexagons
 
@@
   tri_object = tr1_object
+  )
+```
+
+```{r, echo=TRUE, eval=FALSE}
+find_lg_benchmark(
+  distance_edges = distance_df, 
+  distance_col = "distance"
   )
@@
   col_start = "x"
-  )
-```
-
-## Model parameters
-
-<!--discuss about default settings-->
-
-### Choice of bins
-
-### Removal of low-density bins
-
-```{r, echo=TRUE, eval=FALSE}
-
-find_low_dens_hex(
-  df_bin_centroids_all = df_bin_centroids, 
-  bin1 = num_bins_x, 
-  df_bin_centroids_low = df_bin_centroids_low
-  )
-
-```
-
-### Removing long edges
-
-```{r, echo=TRUE, eval=FALSE}
-find_lg_benchmark(
-  distance_edges = distance_df, 
-  distance_col = "distance"
   )
@@
 
-This paper presents the R package `quollr` to develop a way to take the fitted model, as represented by the positions of points in 2D, and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.
+This paper presents the R package `quollr` to develop a way to take the fitted model, as represented by the positions of points in $2\text{-}D$, and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.
 
@@
 
-This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of $p-D$ data.
+This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of $p\text{-}D$ data.
 

### 2024-06-22 paper-quollr.Rmd
Add indicating line segments section

@@
 
+<!--add each step separately-->
+
 ### Remove low-density hexagons
 
+```{r, echo=TRUE, eval=FALSE}
+find_low_dens_hex(
+  df_bin_centroids_all = df_bin_centroids, 
+  bin1 = 12, 
+  df_bin_centroids_low = df_bin_centroids_low
+  )
+
+```
+
 ### Indicating neighbors by line segments connecting centroids
+
+To indicate neighbors, the `tri_bin_centroids()` function is used to triangulate bin centroids.
+Following this, `gen_edges()` function computes the line segments that connect neighboring bins by providing the triangulated data.
+This results the coordinates that generate the connecting lines.
 

### 2024-06-22 paper-quollr.Rmd
Add removing long edge threshold explanation

@@
 ```
+
+In some cases, distant centroids may be connected, resulting in long line segments that can affect the smoothness of the $2\text{-}D$ representation.
+To address this issue, the `find_lg_benchmark()` function is used.
+This function computes a threshold based on the distances of line segments, determining when long edges should be removed.
 

### 2024-06-22 paper-quollr.Rmd
Add lift model to high-dimensions

@@
 ## Lifting the model into high dimensions
+
+The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each bin to represent bin centroids.
+This transformation is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data and their corresponding hexagonal bin IDs as inputs.
 

### 2024-06-22 paper-quollr.Rmd
Add remove low-density hexagons section

@@
 ### Remove low-density hexagons
+
+In certain scenarios, hexagonal bins may contain a few number of points.
+To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
+The `find_low_dens_hex()` function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
+Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
 

### 2024-06-22 paper-quollr.Rmd
Add section prediction

@@
 
+The `predict_emb()` function is used to predict $2\text{-}D$ embedding for a new $p\text{-}D$ data point using the fitted model.
+This function is useful to predict $2\text{-}D$ embedding irrespective of the NLDR technique.
+
+In the prediction process, first, the nearest $p\text{-}D$ model point is identified for a given new $p\text{-}D$ point by computing $p\text{-}D$ Euclidean distance.
+Then, the corresponding $2-D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
+Finally, the coordinates of the identified $2\text{-}D$ bin centroid is used as the predicted NLDR point for the new $p\text{-}D$ data point.
+
+```{r, echo=TRUE, eval=FALSE}
+predict_emb(
+  test_data = s_curve_noise_training, 
+  df_bin_centroids = df_bin_centroids, 
+  df_bin = df_bin, 
+  type_NLDR = "UMAP"
+  )
+```
+
 ## Compute residuals and Mean Square Error (MSE)
+
+As a Goodness of fit statistics for the model, `glance()` is used to compute residuals and MSE.
+These metrics are used to assess how well the fitted model will capture the underlying structure of the $p\text{-}D$ data.
+
+```{r, echo=TRUE, eval=FALSE}
+glance(
+  df_bin_centroids = df_bin_centroids, 
+  df_bin = df_bin, 
+  training_data = s_curve_noise_training, 
+  newdata = NULL, 
+  type_NLDR = "UMAP", 
+  col_start = "x"
+  )
+```
+
+```{r, echo=TRUE, eval=FALSE}
+augment(
+  df_bin_centroids = df_bin_centroids, 
+  df_bin = df_bin, 
+  training_data = s_curve_noise_training, 
+  newdata = NULL, 
+  type_NLDR = "UMAP", 
+  col_start = "x"
+  )
+```
 

### 2024-06-23 paper-quollr.Rmd
Added model summary section

@@
 
-In the prediction process, first, the nearest $p\text{-}D$ model point is identified for a given new $p\text{-}D$ point by computing $p\text{-}D$ Euclidean distance.
-Then, the corresponding $2-D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
-Finally, the coordinates of the identified $2\text{-}D$ bin centroid is used as the predicted NLDR point for the new $p\text{-}D$ data point.
+In the prediction process, first, the nearest $p\text{-}D$ model point is identified for a given new $p\text{-}D$ data point by computing $p\text{-}D$ Euclidean distance.
+Then, the corresponding $2\text{-}D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
+Finally, the coordinates of the identified $2\text{-}D$ bin centroid is used as the predicted NLDR embedding for the new $p\text{-}D$ data point.
 
@@
 ```
+
+Furthermore, `augment()` accepts $2\text{-}D$ and $p\text{-}D$ model points, and the $p\text{-}D$ data and adds information about each observation in the data set.
+Most commonly, this includes predicted values, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error.
+
+Users can pass data to `augment()` via either the `training_data` argument or the `newdata` argument.
+If data is passed to the `training_data` argument, it must be exactly the data that was used to fit the model.
+Alternatively, datasets can be passed to `newdata` to augment data that was not used during model fitting.
+This requires that at least all predictor variable columns used to fit the model are present.
+If the original outcome variable used to fit the model is not included in `newdata`, then no corresponding column will be included in the output.
+
+The augmented dataset is always returned as a `tibble::tibble` with the same number of rows as the passed dataset.
 

### 2024-06-23 paper-quollr.Rmd
Add visualization introduction paragraph

@@
 ## Visualizations
+
+The package provides five basic visualizations which includes one to visualize the full hexagonal grid in $2\text{-}D$, three visualizations related to the $2\text{-}D$ model (static visualizations), and one related to the $p\text{-}D$ model (dynamic visualization).
+Each visualization can be generated using its respective function, as described in this section.
 

### 2024-06-23 paper-quollr.Rmd
Add 2D vis functions
Add 2D vis description

@@
 <!--add each step separately-->
+<!--add hexbin notation image-->
 
@@
 
+The `geom_hexgrid()` function is used to plot the hexagonal grid from the provided centroid data set.
+
+```{r, echo=TRUE, eval=FALSE}
+ggplot() + 
+  geom_hexgrid(
+    data = all_centroids_df, 
+    aes(x = c_x, y = c_y)
+    ) +
+  coord_fixed()
+```
+
+To visualize the $2\text{-}D$ model, mainly three functions are used.
+As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`.
+After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges.
+As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively.
+Following this, `vis_rmlg_mesh()` is used to visualize smoothed $2\text{-}D$ model which is the $2\text{-}D$ model after removing the long edges (see Figure \@ref(fig:mesh-plots)c).
+In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in $2\text{-}D$.
+Using small value of `benchmark_value`, will produce a triangular mesh with missing data structure; for example `benchmark_value = 0.3` shows two clusters rather than continuous structure, while `benchmark_value = 0.8` creates long edges and mislead the data structure in $p\text{-}D$ space.
+
+```{r, echo=TRUE, eval=FALSE}
+ggplot() + 
+  geom_trimesh(
+    data = df_bin_centroids, 
+    aes(x = c_x, y = c_y)
+    ) +
+  coord_fixed()
+```
+
 ```{r, echo=TRUE, eval=FALSE}
@@
 ### $p\text{-}D$ model visualization
+
+Displaying the $p\text{-}D$ model overlaid on the data is done using the function `show_langevitour()`.
+This visualization is helpful for visually evaluating how well the model fits the data.
+The function requires several arguments: data along with their corresponding hexagonal bin ID, $2\text{-}D$ and $p\text{-}D$ model points, the threshold for removing long edges, and the distance data set.
 

### 2024-06-24 paper-quollr.Rmd
Update tests

@@
 
-All functions have tests written and implemented using the \CRANpkg{testthat} [@testthat] in R.
+All functions have tests written and implemented using the \CRANpkg{testthat} package in R.
+These tests illuminated issues, allowing meaningful changes and providing insights into some pitfalls of the package.
+For example, several tests check if there is an NLDR point at the intersection of two bins and determine which bin this point falls into.
 
-These tests illuminated the issues that allowed us to make meaningful changes and understand some pitfalls of the package.
+Figure xx shows an example of a simulated test set. When creating this test data set,
 
 <!--discuss a test with how a point in a intersection of two hexagonal going to evaluate-->
+<!--selecting the default values to q as 0.1-->
 

### 2024-06-24 paper-quollr.tex
Update tests

--- 2024-03-31 paper-quollr.tex
@@
 % !TeX root = RJwrapper.tex
-\title{quollr: An R Package for Visualizing 2D Models from Nonlinear Dimension Reductions in High Dimensional Space}
+\title{quollr: An R Package for Visualizing \(2-D\) Models from Non-linear Dimension Reductions in High Dimensional Space}
 
@@
 \abstract{%
-An abstract of less than 150 words.
+Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (p\text{-}D) by applying a non-linear transformation.
+However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices.
+It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
+The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data.
+The \texttt{triangular\_3d\_data} data from the \CRANpkg{cardinalR} package is used to illustrate the algorithm and its application within the package.
 }
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
-\hypertarget{methodology}{%
-\section{Methodology}\label{methodology}}
+This paper presents the R package, \texttt{quollr} which introduce a new visual tool in determining which NLDR technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
+The methodology of the algorithm is explained in \emph{cite the methodology paper}.
+Furthermore, the \texttt{quollr} package enables users to perform hexagonal binning (Carr et al.
+2023), resulting in the generation of regular hexagons.
+The software is available from the Comprehensive R Archive Network (CARN) at \url{https://CRAN.R-project.org/package=quollr}.
 
-\hypertarget{implementation}{%
-\section{Implementation}\label{implementation}}
+The paper is organized as follows.
+In next section, introduces the implementation of \texttt{quollr} package on CRAN, including demonstration of the package's key functions and visualization capabilities.
+We illustrate the algorithm's functionality to study about clustering data structure in \textbf{Application} section, and describe a visual heuristic to describe parameter selection.
+Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our algorithm.
+
+\section{Implementation}\label{implementation}
 
@@
 
-\texttt{install.packages("quollr")}
+\begin{verbatim}
+install.packages("quollr")
+\end{verbatim}
 
@@
 
-\texttt{devtools::install\_github("JayaniLakshika/quollr")}
+\begin{verbatim}
+devtools::install_github("JayaniLakshika/quollr")
+\end{verbatim}
+
+\subsection{Package dependencies}\label{package-dependencies}
+
+Understanding the dependencies of the \texttt{quollr} package is essential for smooth operation and error prevention.
+The following dependencies refer to the other R packages that \texttt{quollr} relies on to execute its functions effectively.
+
+\begin{verbatim}
+#> $quollr
+#>  [1] "dplyr"       "ggplot2"     "grid"        "interp"      "langevitour"
+#>  [6] "proxy"       "rlang"       "rsample"     "stats"       "tibble"     
+#> [11] "tidyselect"
+\end{verbatim}
+
+\subsection{Usage}\label{usage}
 
@@
 
+\texttt{s\_curve\_noise\_training} is a \(3\text{-}D\) S-curve data set with additional four noise dimensions which is used to train the model.
+\texttt{s\_curve\_noise\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{s\_curve\_noise\_training} data set.
+Each data set contains a unique ID column.
+
+\subsubsection{Scaling the data}\label{scaling-the-data}
+
+The algorithm begins by scaling NLDR data to a standard scale using the \texttt{gen\_scaled\_data()} function.
+This function standardizes the data and provides the original limits of embeddings and the aspect ratio.
+
+\begin{verbatim}
+scaled_nldr_obj_scurve <- gen_scaled_data(data = s_curve_noise_umap)
+
+s_curve_noise_umap_scaled <- scaled_nldr_obj_scurve$scaled_nldr
+
+lim1 <- scaled_nldr_obj_scurve$lim1
+lim2 <- scaled_nldr_obj_scurve$lim2
+r2 <- diff(lim2)/diff(lim1) 
+\end{verbatim}
+
 The mains steps for the algorithm can be executed by the main function \texttt{fit\_highd\_model()}, or can be run separately for more flexibility.
-When constructing the 2D model, the user can choose either to fir the 2D model with hexagonal bin centroids or bin means using \texttt{is\_bin\_centroid} argument.
+When constructing the \(2\text{-}D\) model, the user can choose either to fit the \(2\text{-}D\) model with hexagonal bin centroids or bin means using \texttt{is\_bin\_centroid} argument.
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-1-1.pdf}
+If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
+For example, if the user wishes to use already binning data, the \texttt{extract\_hexbin\_centroids()} function can be used directly.
 
-\hypertarget{constructing-the-2d-model}{%
-\subsection{Constructing the 2D model}\label{constructing-the-2d-model}}
+The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if \texttt{is\_rm\_lwd\_hex\ =\ TRUE}, benchmark value to remove low density hexagons are parameters that will be determined within \texttt{fit\_highd\_model()}, if they are not provided.
+They are created as they are needed throughout the following example.
+The function \texttt{fit\_highd\_model()} provides the fitted model in \(2\text{-}D\), and \(p\text{-}D\).
 
-\hypertarget{preprocessing}{%
-\subsubsection{Preprocessing}\label{preprocessing}}
+\begin{verbatim}
+fit_highd_model(
+  training_data = s_curve_noise_training,
+  emb_df = s_curve_noise_umap_scaled, 
+  bin1 = 12, 
+  r2 = r2,
+  q = 0.07,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = TRUE,
+  benchmark_to_rm_lwd_hex = NULL,
+  col_start_highd = "x"
+  )
+\end{verbatim}
 
-The function \texttt{gen\_scaled\_data()} is used to prepare the NLDR data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization.
+\begin{verbatim}
+#> $df_bin
+#> # A tibble: 37 x 8
+#>    hb_id       x1     x2     x3       x4       x5       x6        x7
+#>    <int>    <dbl>  <dbl>  <dbl>    <dbl>    <dbl>    <dbl>     <dbl>
+#>  1    27 -0.706   1.70   -1.65   0.00918  0.00261 -0.0236  -0.00199 
+#>  2    28 -0.214   1.45   -1.98  -0.00966 -0.00370  0.0573   0.000307
+#>  3    40  0.356   1.74   -1.89   0.0127  -0.00773 -0.0219  -0.000757
+#>  4    50 -0.209   0.453  -1.98   0.0185   0.0189  -0.0999  -0.00412 
+#>  5    52  0.00161 1.14   -2.00   0.00206 -0.00940 -0.0974  -0.00796 
+#>  6    61 -0.671   0.0922 -1.74  -0.00458  0.0127  -0.0613  -0.00643 
+#>  7    62 -0.375   0.328  -1.89  -0.00226 -0.00651 -0.0291   0.00120 
+#>  8    74 -0.175   0.0562 -1.98   0.00204 -0.00179  0.0112  -0.00111 
+#>  9    75  0.325   0.0821 -1.95   0.00292  0.0197  -0.0165  -0.000881
+#> 10   127  0.704   0.454  -0.290 -0.00394  0.0118  -0.00941 -0.00348 
+#> # i 27 more rows
+#> 
+#> $df_bin_centroids
+#> # A tibble: 37 x 5
+#>    hexID     c_x    c_y std_counts drop_empty
+#>    <int>   <dbl>  <dbl>      <dbl> <lgl>     
+#>  1    27  0.127  0.0295        1   FALSE     
+#>  2    28  0.226  0.0295        0.2 FALSE     
+#>  3    40  0.275  0.115         0.6 FALSE     
+#>  4    50  0.0287 0.200         0.2 FALSE     
+#>  5    52  0.226  0.200         0.2 FALSE     
+#>  6    61 -0.0207 0.286         0.2 FALSE     
+#>  7    62  0.0780 0.286         0.8 FALSE     
+#>  8    74  0.0287 0.371         0.4 FALSE     
+#>  9    75  0.127  0.371         0.2 FALSE     
+#> 10   127  0.522  0.713         0.4 FALSE     
+#> # i 27 more rows
+\end{verbatim}
 
-\hypertarget{binning-data}{%
-\subsubsection{Binning data}\label{binning-data}}
+\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) model}{Constructing the 2\textbackslash text\{-\}D model}}\label{constructing-the-2text-d-model}
 
-The mains steps for the hexagonal binning algorithm can be executed by the main function \texttt{hex\_binning()}, or can be run separately for more flexibility.
+Constructing the \(2\text{-}D\) model mainly contains (i) binning data, (ii) obtaining bin centroids, and (iii) indicating neighbors by line segments connecting centroids.
 
-\textbf{Step 1: Compute hexagonal grid configurations}
+\subsubsection{Computing hexagon grid configurations}\label{computing-hexagon-grid-configurations}
 
-\textbf{Step 2: Generate the hexagonal bin centroids}
+The configurations of the hexagonal grid is defined by the number of bins in each direction.
+To find the number of bins along the y-axis, \texttt{calc\_bins\_y()} is used.
+This function takes as input the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
+Additionally, this function provides the bin width.
 
-\textbf{Step 3: Create the hexagonal grid}
+\begin{verbatim}
+calc_bins_y(
+  bin1 = 12, 
+  r2 = r2,
+  q = 0.07
+  )
+\end{verbatim}
 
-\textbf{Step 4: Assign NLDR data to hexagons}
+\begin{verbatim}
+#> $bin2
+#> [1] 27
+#> 
+#> $a1
+#> [1] 0.09596332
+\end{verbatim}
 
-\hypertarget{obtain-bin-centroids-means}{%
-\subsubsection{Obtain bin centroids/ means}\label{obtain-bin-centroids-means}}
+\subsubsection{Binning the data}\label{binning-the-data}
 
-To obtain hexagonal bin centroids, \texttt{extract\_hexbin\_centroids()}is used. The function \texttt{extract\_hexbin\_mean()} is used to obtain average within each hexagon.
+Points are allocated to the bins they fall into based on the nearest centroid.
+The main steps of the hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or they can be run separately for greater flexibility.
+The parameters used within \texttt{hex\_binning()} include the scaled NLDR data, the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
+The output is an object of the \texttt{hex\_bin\_obj} class, which contains the number of bins in each direction, the coordinates of the hexagonal grid starting point, the details of bin centroids, the coordinates of bins, embedding components with their corresponding hexagon IDs, hex bins with their corresponding standardized counts, the total number of bins, the number of non-empty bins, and the points within each hexagon.
 
-\hypertarget{triangulate-bin-centroids-means}{%
-\subsubsection{Triangulate bin centroids/ means}\label{triangulate-bin-centroids-means}}
+\begin{verbatim}
+hex_binning(
+  data = s_curve_noise_umap_scaled, 
+  bin1 = 12, 
+  r2 = r2,
+  q = 0.07
+  )
+\end{verbatim}
 
-The \texttt{tri\_bin\_centroids()} is used to do triangulate bin centroids/ bin means. Then, \texttt{gen\_edges()} compute the edges to obtain the triangular mesh which is the 2D model.
+\begin{verbatim}
+#> $bins
+#> [1] 12 27
+#> 
+#> $start_point
+#> [1] -0.070000 -0.141359
+#> 
+#> $centroids
+#> # A tibble: 324 x 3
+#>    hexID     c_x    c_y
+#>    <int>   <dbl>  <dbl>
+#>  1     1 -0.07   -0.141
+#>  2     2  0.0287 -0.141
+#>  3     3  0.127  -0.141
+#>  4     4  0.226  -0.141
+#>  5     5  0.325  -0.141
+#>  6     6  0.423  -0.141
+#>  7     7  0.522  -0.141
+#>  8     8  0.621  -0.141
+#>  9     9  0.719  -0.141
+#> 10    10  0.818  -0.141
+#> # i 314 more rows
+#> 
+#> $hex_poly
+#> # A tibble: 1,944 x 3
+#>    hex_poly_id       x       y
+#>          <int>   <dbl>   <dbl>
+#>  1           1 -0.07   -0.0860
+#>  2           1 -0.118  -0.114 
+#>  3           1 -0.118  -0.169 
+#>  4           1 -0.07   -0.197 
+#>  5           1 -0.0220 -0.169 
+#>  6           1 -0.0220 -0.114 
+#>  7           2  0.0287 -0.0860
+#>  8           2 -0.0193 -0.114 
+#>  9           2 -0.0193 -0.169 
+#> 10           2  0.0287 -0.197 
+#> # i 1,934 more rows
+#> 
+#> $data_hb_id
+#> # A tibble: 75 x 4
+#>     UMAP1  UMAP2    ID hb_id
+#>     <dbl>  <dbl> <int> <int>
+#>  1 0.0804 0.320      1    62
+#>  2 0.739  1.00       2   165
+#>  3 0.840  1.08       3   178
+#>  4 0.167  0.0432     4    27
+#>  5 0.263  0.398      6    76
+#>  6 0.838  2.01       7   310
+#>  7 0.734  0.972      8   165
+#>  8 0.627  0.721      9   128
+#>  9 0.810  1.01      11   178
+#> 10 0.903  1.87      12   299
+#> # i 65 more rows
+#> 
+#> $std_cts
+#> # A tibble: 45 x 3
+#>    hb_id     n std_counts
+#>    <int> <int>      <dbl>
+#>  1    27     5        1  
+#>  2    28     1        0.2
+#>  3    38     1        0.2
+#>  4    40     3        0.6
+#>  5    50     1        0.2
+#>  6    52     1        0.2
+#>  7    61     1        0.2
+#>  8    62     4        0.8
+#>  9    64     1        0.2
+#> 10    74     2        0.4
+#> # i 35 more rows
+#> 
+#> $tot_bins
+#> [1] 324
+#> 
+#> $non_bins
+#> [1] 45
+#> 
+#> $pts_bins
+#> # A tibble: 45 x 2
+#>    hexID pts_list    
+#>    <int> <named list>
+#>  1    62 <int [75]>  
+#>  2   165 <int [75]>  
+#>  3   178 <int [75]>  
+#>  4    27 <int [75]>  
+#>  5    76 <int [75]>  
+#>  6   310 <int [75]>  
+#>  7   128 <int [75]>  
+#>  8   299 <int [75]>  
+#>  9    74 <int [75]>  
+#> 10   287 <int [75]>  
+#> # i 35 more rows
+#> 
+#> attr(,"class")
+#> [1] "hex_bin_obj"
+\end{verbatim}
 
-\hypertarget{lifting-the-model-into-high-dimensions}{%
-\subsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}}
+\subsubsection{Remove low-density hexagons}\label{remove-low-density-hexagons}
 
-The \(p\)-D model is generated by \texttt{avg\_highd\_data()} by passing the \(p\)-D data.
+In certain scenarios, hexagonal bins may contain a few number of points.
+To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
+The \texttt{find\_low\_dens\_hex()} function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
+Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
 
-\hypertarget{tuning-model-parameters}{%
-\subsection{Tuning model parameters}\label{tuning-model-parameters}}
+\begin{verbatim}
+find_low_dens_hex(
+  df_bin_centroids_all = df_bin_centroids, 
+  bin1 = 12, 
+  df_bin_centroids_low = df_bin_centroids_low
+  )
+\end{verbatim}
 
-Mainly there are two model parameter need to consider: benchmark value to remove the low-density hexagons, and benchmark value to remove the long edges.
-\texttt{find\_low\_dens\_hex()} is used to find the hexagons which contains less number of points by considering the density of their neighboring points as well.
-In here, user can first decide which are the low-density hexagons and pass them to this function to check whether these removal of low-density hexagons can affect the model fit by looking at the the neighbors.
-On the other hand, \texttt{find\_lg\_benchmark()} function is used to compute the threshold for removing long edges.
+\subsubsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
 
-\hypertarget{compute-residuals-and-mean-square-error-mse}{%
-\subsection{Compute residuals and Mean Square Error (MSE)}\label{compute-residuals-and-mean-square-error-mse}}
+To indicate neighbors, the \texttt{tri\_bin\_centroids()} function is used to triangulate bin centroids.
+Following this, \texttt{gen\_edges()} function computes the line segments that connect neighboring bins by providing the triangulated data.
+This results the coordinates that generate the connecting lines.
 
-As a Goodness of fit statistics for the model, \texttt{gen\_summary()} is used to compute residuals and MSE. For the fitted model, how residuals and MSE is changing is shown in Figure 2.
+\begin{verbatim}
+tr1_object <- tri_bin_centroids(
+  hex_df = df_bin_centroids, 
+  x = "c_x", 
+  y = "c_y"
+  )
 
-\hypertarget{visualizations}{%
-\subsection{Visualizations}\label{visualizations}}
+tr_from_to_df <- gen_edges(
+  tri_object = tr1_object
+  )
+\end{verbatim}
 
-The package provides four basic visualizations which includes three visualizations regarding 2D model (static vis) and one is regrading \(p\)-D model (dynamic vis).
-Each visualization can be done their own functions which describes in this section.
+In some cases, distant centroids may be connected, resulting in long line segments that can affect the smoothness of the \(2\text{-}D\) representation.
+To address this issue, the \texttt{find\_lg\_benchmark()} function is used.
+This function computes a threshold based on the distances of line segments, determining when long edges should be removed.
 
-\hypertarget{d-model-visualization}{%
-\subsubsection{2D model visualization}\label{d-model-visualization}}
+\begin{verbatim}
+find_lg_benchmark(
+  distance_edges = distance_df, 
+  distance_col = "distance"
+  )
+\end{verbatim}
 
-To visualize the 2D model, mainly three functions are used.
-As shown in Figure 1, \texttt{geom\_trimesh()} to visualize the triangular mesh by adding a new layer to \texttt{ggplot()}.
+\subsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}
+
+The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each bin to represent bin centroids.
+This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data and their corresponding hexagonal bin IDs as inputs.
+
+\begin{verbatim}
+umap_data_with_hb_id <- hb_obj$data_hb_id
+
+df_all <- bind_cols(
+  s_curve_noise_training |> dplyr::select(-ID), 
+  umap_data_with_hb_id
+  )
+
+df_bin <- avg_highd_data(
+  data = df_all, 
+  col_start = "x"
+  )
+\end{verbatim}
+
+\subsection{Prediction}\label{prediction}
+
+The \texttt{predict\_emb()} function is used to predict \(2\text{-}D\) embedding for a new \(p\text{-}D\) data point using the fitted model.
+This function is useful to predict \(2\text{-}D\) embedding irrespective of the NLDR technique.
+
+In the prediction process, first, the nearest \(p\text{-}D\) model point is identified for a given new \(p\text{-}D\) data point by computing \(p\text{-}D\) Euclidean distance.
+Then, the corresponding \(2\text{-}D\) bin centroid mapping for the identified \(p\text{-}D\) model point is determined.
+Finally, the coordinates of the identified \(2\text{-}D\) bin centroid is used as the predicted NLDR embedding for the new \(p\text{-}D\) data point.
+
+\begin{verbatim}
+predict_emb(
+  test_data = s_curve_noise_training, 
+  df_bin_centroids = df_bin_centroids, 
+  df_bin = df_bin, 
+  type_NLDR = "UMAP"
+  )
+\end{verbatim}
+
+\subsection{Compute residuals and Mean Square Error (MSE)}\label{compute-residuals-and-mean-square-error-mse}
+
+As a Goodness of fit statistics for the model, \texttt{glance()} is used to compute residuals and MSE.
+These metrics are used to assess how well the fitted model will capture the underlying structure of the \(p\text{-}D\) data.
+
+\begin{verbatim}
+glance(
+  df_bin_centroids = df_bin_centroids, 
+  df_bin = df_bin, 
+  training_data = s_curve_noise_training, 
+  newdata = NULL, 
+  type_NLDR = "UMAP", 
+  col_start = "x"
+  )
+\end{verbatim}
+
+Furthermore, \texttt{augment()} accepts \(2\text{-}D\) and \(p\text{-}D\) model points, and the \(p\text{-}D\) data and adds information about each observation in the data set.
+Most commonly, this includes predicted values, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error.
+
+Users can pass data to \texttt{augment()} via either the \texttt{training\_data} argument or the \texttt{newdata} argument.
+If data is passed to the \texttt{training\_data} argument, it must be exactly the data that was used to fit the model.
+Alternatively, datasets can be passed to \texttt{newdata} to augment data that was not used during model fitting.
+This requires that at least all predictor variable columns used to fit the model are present.
+If the original outcome variable used to fit the model is not included in \texttt{newdata}, then no corresponding column will be included in the output.
+
+The augmented dataset is always returned as a \texttt{tibble::tibble} with the same number of rows as the passed dataset.
+
+\begin{verbatim}
+augment(
+  df_bin_centroids = df_bin_centroids, 
+  df_bin = df_bin, 
+  training_data = s_curve_noise_training, 
+  newdata = NULL, 
+  type_NLDR = "UMAP", 
+  col_start = "x"
+  )
+\end{verbatim}
+
+\subsection{Visualizations}\label{visualizations}
+
+The package provides five basic visualizations which includes one to visualize the full hexagonal grid in \(2\text{-}D\), three visualizations related to the \(2\text{-}D\) model (static visualizations), and one related to the \(p\text{-}D\) model (dynamic visualization).
+Each visualization can be generated using its respective function, as described in this section.
+
+\subsubsection{\texorpdfstring{\(2\text{-}D\) model visualization}{2\textbackslash text\{-\}D model visualization}}\label{text-d-model-visualization}
+
+The \texttt{geom\_hexgrid()} function is used to plot the hexagonal grid from the provided centroid data set.
+
+\begin{verbatim}
+ggplot() + 
+  geom_hexgrid(
+    data = all_centroids_df, 
+    aes(x = c_x, y = c_y)
+    ) +
+  coord_fixed()
+\end{verbatim}
+
+To visualize the \(2\text{-}D\) model, mainly three functions are used.
+As shown in Figure \ref{fig:mesh-plots}a, \texttt{geom\_trimesh()} to visualize the triangular mesh by adding a new layer to \texttt{ggplot()}.
 After identifying benchmark value to remove long edge, \texttt{vis\_lg\_mesh()} is used to visualize the triangular mesh by coloring the small and long edges.
-As shown in Figure 1, the small and long edges are colored by black and red respectively.
-Following this, \texttt{vis\_rmlg\_mesh()} is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure 1).
-In \texttt{vis\_lg\_mesh()} and \texttt{vis\_rmlg\_mesh()}, \texttt{benchmark\_value} argument controls the edge removal in 2D.
+As shown in Figure \ref{fig:mesh-plots}b, the small and long edges are colored by black and red respectively.
+Following this, \texttt{vis\_rmlg\_mesh()} is used to visualize smoothed \(2\text{-}D\) model which is the \(2\text{-}D\) model after removing the long edges (see Figure \ref{fig:mesh-plots}c).
+In \texttt{vis\_lg\_mesh()} and \texttt{vis\_rmlg\_mesh()}, \texttt{benchmark\_value} argument controls the edge removal in \(2\text{-}D\).
+Using small value of \texttt{benchmark\_value}, will produce a triangular mesh with missing data structure; for example \texttt{benchmark\_value\ =\ 0.3} shows two clusters rather than continuous structure, while \texttt{benchmark\_value\ =\ 0.8} creates long edges and mislead the data structure in \(p\text{-}D\) space.
 
-\hypertarget{p-d-model-visualization}{%
-\subsubsection{p-D model visualization}\label{p-d-model-visualization}}
+\begin{verbatim}
+ggplot() + 
+  geom_trimesh(
+    data = df_bin_centroids, 
+    aes(x = c_x, y = c_y)
+    ) +
+  coord_fixed()
+\end{verbatim}
 
-The \(p\)-D model overlaid on data is visualized by the function \texttt{show\_langevitour()}.
-This is a useful dynamic visualization result to validate visually whether the model fits the data well or not.
+\begin{verbatim}
+vis_lg_mesh(
+  distance_edges = distance_df, 
+  benchmark_value = 0.75, 
+  tr_coord_df = tr_from_to_df, 
+  distance_col = "distance"
+  )
+\end{verbatim}
 
-\hypertarget{prediction}{%
-\subsection{Prediction}\label{prediction}}
+\begin{verbatim}
+vis_rmlg_mesh(
+  distance_edges = distance_df, 
+  benchmark_value = 0.75, 
+  tr_coord_df = tr_from_to_df, 
+  distance_col = "distance"
+  )
+\end{verbatim}
 
-\texttt{predict\_emb()} function is used to predict 2D embedding for a new data point using the fitted model. This function is useful to predict 2D embedding irrespective of the NLDR technique.
+\subsubsection{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
 
-\hypertarget{tests}{%
-\subsection{Tests}\label{tests}}
+Displaying the \(p\text{-}D\) model overlaid on the data is done using the function \texttt{show\_langevitour()}.
+This visualization is helpful for visually evaluating how well the model fits the data.
+The function requires several arguments: data along with their corresponding hexagonal bin ID, \(2\text{-}D\) and \(p\text{-}D\) model points, the threshold for removing long edges, and the distance data set.
+
+\begin{verbatim}
+show_langevitour(
+  df = df_all, 
+  df_b = df_bin, 
+  df_b_with_center_data = df_bin_centroids, 
+  benchmark_value = 0.75, 
+  distance = distance_df, 
+  distance_col = "distance", 
+  use_default_benchmark_val = FALSE, 
+  col_start = "x"
+  )
+\end{verbatim}
+
+\subsection{Tests}\label{tests}
 
@@
 
-\hypertarget{application}{%
-\section{Application}\label{application}}
+These tests illuminated the issues that allowed us to make meaningful changes and understand some pitfalls of the package.
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-4-1.pdf}
+\section{Application}\label{application}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}
+To illustrate the algorithm, we use \(5\text{-}D\) simulated data, which we call the ``triangular\_3d\_data''.
+This dataset is generated using the \texttt{tri\_3d} function from the \texttt{cardinalR} package.
+The dataset generation starts with the initialization of a starting point \(p_0 = (x_0, y_0, z_0)\) randomly selected within the \(3\text{-}D\) space using a uniform distribution.
+Four fixed corner points of a tetrahedron are defined as \(c_1 = (0, 0, 0)\), \(c_2 = (1, 0, 0)\), \(c_3 = (0.5, 1, 0)\), and \(c_4 = (0.5, 0.5, 1)\).
+For each point \(p_i\) where \(i = 1, 2, \ldots, n\), one of the corner points \(c_j\) (where \(j \in \{1, 2, 3, 4\}\)) is selected randomly, and the new point \(p_i\) is computed as the midpoint between the current point \(p_{i-1}\) and the selected corner point \(c_j\): \(p_i = \frac{p_{i-1} + c_j}{2}\).
+This iterative process ensures that each new point moves closer to one of the corners, creating a fractal-like triangular distribution of points.
+The coordinates of each generated point \(p_i\) form the variables \(X_1, X_2, X_3\).
+The remaining variables \(X_4, X_5\) are all uniform error, with small variance.
+We would consider \(T=(X_1, X_2, X_3)\) to be the true model.
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-2.pdf}
+\subsection{Deciding an appropriate fit}\label{deciding-an-appropriate-fit}
 
-\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-9-2.pdf}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\subsection{Fit the appropriate model}\label{fit-the-appropriate-model}
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\subsubsection{\texorpdfstring{Construct model in \(2\text{-}D\)}{Construct model in 2\textbackslash text\{-\}D}}\label{construct-model-in-2text-d}
+
+\subsubsection{Selecting parameter values for the model}\label{selecting-parameter-values-for-the-model}
+
+\paragraph{Choice of bins}\label{choice-of-bins}
+
+\paragraph{Removal of low-density bins}\label{removal-of-low-density-bins}
+
+\paragraph{Removing long edges}\label{removing-long-edges}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-32-1}
+
+\subsubsection{\texorpdfstring{Lifting the model into \(p\text{-}D\)}{Lifting the model into p\textbackslash text\{-\}D}}\label{lifting-the-model-into-ptext-d}
+
+What can I see in high-D?
+
+\begin{itemize}
+\tightlist
+\item
+  Four clusters that are really close
+\end{itemize}
+
+What can I see in 2D layout?
+
+\begin{itemize}
+\tightlist
+\item
+  Several clusters (more that what see in high-D)
+\end{itemize}
+
+What can learn from the model?
+
+\begin{itemize}
+\tightlist
+\item
+  There are four clusters, but not as the same cluster what data shows.
+In 2D, the two of actual clusters (high-D clusters) are really close.
+Also, one of the big high-D cluster is separated to two sub clusters.
+\end{itemize}
+
+\section{Discussion}\label{discussion}
+
+This paper presents the R package \texttt{quollr} to develop a way to take the fitted model, as represented by the positions of points in \(2\text{-}D\), and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.
+
+The paper includes a clustering example to illustrate how \texttt{quollr} is useful to assess which NLDR technique and which (hyper)parameter choice gives the most accurate representation.
+In addition, how to select parameters for hexagonal binning and fitting model are explained.
+
+Possible future improvements would be\ldots{}
+
+This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of \(p\text{-}D\) data.
+
+\section{Acknowledgements}\label{acknowledgements}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-testthat}{}}%
+\bibitem[\citeproctext]{ref-dan2023}
+Carr, Dan, ported by Nicholas Lewin-Koh, Martin Maechler, and contains copies of lattice functions written by Deepayan Sarkar.
+2023.
+\emph{Hexbin: Hexagonal Binning Routines}.
+\url{https://CRAN.R-project.org/package=hexbin}.
+
+\bibitem[\citeproctext]{ref-testthat}
 Wickham, Hadley. 2011. {``Testthat: Get Started with Testing.''} \emph{The R Journal} 3: 5--10. \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
+\bibitem[\citeproctext]{ref-knitr}
 Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
+\bibitem[\citeproctext]{ref-rmarkdown}
 Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{{R} Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.

### 2025-03-21 paper-quollr.Rmd
With new updates to the package
Changed the application to new one

@@
 
+### Interactivity
+
 ## Tests
@@
 
-# Application
+# Example
 
@@
 ## To initialize number of bins along the x-axis
-bin1_vec_prism <- 2:25
+bin1_vec_prism <- 4:25
 
@@
 
+# Application
+
+<!--https://arxiv.org/abs/2409.09227-->
+
+The data used in our study is from APOGEE Data Release 17 (DR17), which encompasses 19 chemical abundances
+and stellar parameters for 370,060 stars. This data was obtained through automated analysis of spectra using APOGEE
+Stellar Parameter and Chemical Abundances Pipeline (ASPCAP) software (Garc´ıa P´erez et al. 2016). We removed
+poor quality stars by ASPCAPFLAG and STARFLAG bitmasks to mitigate potential quality warnings and also
+removed outliers and applied standard scale, resulting in a final dataset comprising 133,891 stars.
+We focused on chemical abundances, which refer to the relative abundance of different chemical elements in the
+atmosphere of stars. To quantify the relative amounts of individual elements, it uses the abundance ratio as the
+logarithm of the ratio of two metallic elements in a star relative to their ratio in the sun. For instance, if a star has a
+[C/FE] value of -0.08, it means its carbon abundance is 83% that of the Sun. We are using 19 dimensions related to
+chemical abundances, including the metallicity [FE/H] and other 18 element abundances.
+
+Compare PCA, tSNE, UMAP, VAE, Autoencoders
+
 # Discussion

### 2025-03-30 paper-quollr.Rmd
Need to fix with new version of quollr

--- 2024-06-24 paper-quollr.Rmd
@@
 
-## Tests
+### Interactivity
 
-All functions have tests written and implemented using the \CRANpkg{testthat} package in R.
+<!-- ## Tests -->
+
+<!-- All functions have tests written and implemented using the \CRANpkg{testthat} package in R.
 These tests illuminated issues, allowing meaningful changes and providing insights into some pitfalls of the package.
 For example, several tests check if there is an NLDR point at the intersection of two bins and determine which bin this point falls into.
+-->
 
-Figure xx shows an example of a simulated test set. When creating this test data set,
+<!-- Figure xx shows an example of a simulated test set. When creating this test data set,  -->
 
-<!--discuss a test with how a point in a intersection of two hexagonal going to evaluate-->
-<!--selecting the default values to q as 0.1-->
+<!-- <!--discuss a test with how a point in a intersection of two hexagonal going to evaluate--> -->
+<!-- <!--selecting the default values to q as 0.1--> -->
 
-# Application
+# Example
 
@@
 
+# Application
+
+Need to find another application
+
 # Discussion

### 2025-03-30 paper-quollr.Rmd
Fixed the issue with new version of quollr

@@
 fit_highd_model(
-  training_data = s_curve_noise_training,
-  emb_df = s_curve_noise_umap_scaled, 
+  highd_data = s_curve_noise_training,
+  nldr_data = s_curve_noise_umap_scaled, 
   bin1 = 12, 
   r2 = r2,
-  q = 0.07,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = TRUE,
-  benchmark_to_rm_lwd_hex = NULL,
-  col_start_highd = "x"
+  q = 0.1,
+  is_bin_centroid = TRUE
   )
@@
   r2 = r2,
-  q = 0.07
+  q = 0.1
   )
@@
 df_bin <- avg_highd_data(
-  data = df_all, 
-  col_start = "x"
+  data = df_all
   )
@@
 predict_emb(
-  test_data = s_curve_noise_training, 
-  df_bin_centroids = df_bin_centroids, 
-  df_bin = df_bin, 
-  type_NLDR = "UMAP"
+  highd_data = s_curve_noise_training, 
+  model_2d = df_bin_centroids, 
+  model_highd = df_bin
   )
@@
 glance(
-  df_bin_centroids = df_bin_centroids, 
-  df_bin = df_bin, 
-  training_data = s_curve_noise_training, 
-  newdata = NULL, 
-  type_NLDR = "UMAP", 
-  col_start = "x"
+  highd_data = s_curve_noise_training, 
+  model_2d = df_bin_centroids, 
+  model_highd = df_bin
   )
@@
 augment(
-  df_bin_centroids = df_bin_centroids, 
-  df_bin = df_bin, 
-  training_data = s_curve_noise_training, 
-  newdata = NULL, 
-  type_NLDR = "UMAP", 
-  col_start = "x"
+  highd_data = s_curve_noise_training, 
+  model_2d = df_bin_centroids, 
+  model_highd = df_bin
   )
@@
 
+```{r, echo=TRUE, eval=FALSE}
+show_link_plots(
+  point_df = df_exe, 
+  edge_df = distance_small_df
+  )
+```
+
+```{r, echo=TRUE, eval=FALSE}
+show_error_link_plots(
+  point_df = df_exe, 
+  edge_df = distance_small_df
+  )
+```
+
 <!-- ## Tests -->
@@
 
-# Example
+<!-- # Example -->
 
-To illustrate the algorithm, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data".
+<!-- To illustrate the algorithm, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data".
 This dataset is generated using the `tri_3d` function from the `cardinalR` package.
@@
 We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+-->
 
-```{r}
-triangular_3d_data <- tri_3d(
-  n = 1250, num_noise = 2, 
-  min_n = -0.05, max_n = 0.05) |>
-  as_tibble()
+<!-- ```{r} -->
+<!-- triangular_3d_data <- tri_3d( -->
+<!--   n = 1250, num_noise = 2,  -->
+<!--   min_n = -0.05, max_n = 0.05) |> -->
+<!--   as_tibble()  -->
 
-colnames(triangular_3d_data) <- paste0("x", 1:NCOL(triangular_3d_data))
+<!-- colnames(triangular_3d_data) <- paste0("x", 1:NCOL(triangular_3d_data)) -->
 
-langevitour::langevitour(triangular_3d_data)
-```
+<!-- langevitour::langevitour(triangular_3d_data) -->
+<!-- ``` -->
 
-```{r}
-triangular_3d_data <- triangular_3d_data |>
-  mutate(ID = row_number())
-```
+<!-- ```{r} -->
+<!-- triangular_3d_data <- triangular_3d_data |> -->
+<!--   mutate(ID = row_number()) -->
+<!-- ``` -->
 
-```{r}
-sc_ltr_pos <- c(0.08, 0.96)
-```
+<!-- ```{r} -->
+<!-- sc_ltr_pos <- c(0.08, 0.96) -->
+<!-- ``` -->
 
-<!--tSNE with perplexity = 6-->
-```{r}
-tSNE_fit1 <- triangular_3d_data |>
-  select(-ID) |>
-  Rtsne(perplexity = 6)
+<!-- <!--tSNE with perplexity = 6--> -->
+<!-- ```{r} -->
+<!-- tSNE_fit1 <- triangular_3d_data |> -->
+<!--   select(-ID) |> -->
+<!--   Rtsne(perplexity = 6)  -->
 
-tsne_prism1 <- tSNE_fit1$Y |>
-  as_tibble() |>
-  mutate(ID = row_number())
+<!-- tsne_prism1 <- tSNE_fit1$Y |> -->
+<!--   as_tibble() |> -->
+<!--   mutate(ID = row_number()) -->
 
-colnames(tsne_prism1) <- c("tSNE1", "tSNE2", "ID")
+<!-- colnames(tsne_prism1) <- c("tSNE1", "tSNE2", "ID") -->
 
-prism_scaled_obj_tsne1 <- gen_scaled_data(
-  data = tsne_prism1)
-tsne_prism_scaled1 <- prism_scaled_obj_tsne1$scaled_nldr
+<!-- prism_scaled_obj_tsne1 <- gen_scaled_data( -->
+<!--   data = tsne_prism1) -->
+<!-- tsne_prism_scaled1 <- prism_scaled_obj_tsne1$scaled_nldr -->
 
-nldr_prism1 <- tsne_prism_scaled1 |> 
-  ggplot(aes(x = tSNE1, y = tSNE2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
-```
+<!-- nldr_prism1 <- tsne_prism_scaled1 |>  -->
+<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
+<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
+<!-- ``` -->
 
-<!--tSNE with perplexity = 35-->
-```{r}
-tSNE_fit2 <- triangular_3d_data |>
-  select(-ID) |>
-  Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))
+<!-- <!--tSNE with perplexity = 35--> -->
+<!-- ```{r} -->
+<!-- tSNE_fit2 <- triangular_3d_data |> -->
+<!--   select(-ID) |> -->
+<!--   Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))  -->
 
-tsne_prism2 <- tSNE_fit2$Y |>
-  as_tibble() |>
-  mutate(ID = row_number())
+<!-- tsne_prism2 <- tSNE_fit2$Y |> -->
+<!--   as_tibble() |> -->
+<!--   mutate(ID = row_number()) -->
 
-colnames(tsne_prism2) <- c("tSNE1", "tSNE2", "ID")
+<!-- colnames(tsne_prism2) <- c("tSNE1", "tSNE2", "ID") -->
 
-prism_scaled_obj_tsne2 <- gen_scaled_data(
-  data = tsne_prism2)
-tsne_prism_scaled2 <- prism_scaled_obj_tsne2$scaled_nldr
+<!-- prism_scaled_obj_tsne2 <- gen_scaled_data( -->
+<!--   data = tsne_prism2) -->
+<!-- tsne_prism_scaled2 <- prism_scaled_obj_tsne2$scaled_nldr -->
 
-nldr_prism2 <- tsne_prism_scaled2 |> 
-  ggplot(aes(x = tSNE1, y = tSNE2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
-```
+<!-- nldr_prism2 <- tsne_prism_scaled2 |>  -->
+<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
+<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
+<!-- ``` -->
 
-<!--tSNE with perplexity = 58-->
-```{r}
-tSNE_fit3 <- triangular_3d_data |>
-  select(-ID) |>
-  Rtsne(perplexity = 58)
+<!-- <!--tSNE with perplexity = 58--> -->
+<!-- ```{r} -->
+<!-- tSNE_fit3 <- triangular_3d_data |> -->
+<!--   select(-ID) |> -->
+<!--   Rtsne(perplexity = 58)  -->
 
-tsne_prism3 <- tSNE_fit3$Y |>
-  as_tibble() |>
-  mutate(ID = row_number())
+<!-- tsne_prism3 <- tSNE_fit3$Y |> -->
+<!--   as_tibble() |> -->
+<!--   mutate(ID = row_number()) -->
 
-colnames(tsne_prism3) <- c("tSNE1", "tSNE2", "ID")
+<!-- colnames(tsne_prism3) <- c("tSNE1", "tSNE2", "ID") -->
 
-prism_scaled_obj_tsne3 <- gen_scaled_data(
-  data = tsne_prism3)
-tsne_prism_scaled3 <- prism_scaled_obj_tsne3$scaled_nldr
+<!-- prism_scaled_obj_tsne3 <- gen_scaled_data( -->
+<!--   data = tsne_prism3) -->
+<!-- tsne_prism_scaled3 <- prism_scaled_obj_tsne3$scaled_nldr -->
 
-nldr_prism3 <- tsne_prism_scaled3 |> 
-  ggplot(aes(x = tSNE1, y = tSNE2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
-```
+<!-- nldr_prism3 <- tsne_prism_scaled3 |>  -->
+<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
+<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
+<!-- ``` -->
 
-<!--UMAP with n_neighbors = 15 and min_dist = 0.3-->
-```{r}
-umap_prism1 <- triangular_3d_data |>
-  select(-ID) |>
-  umap(n_neighbors = 15, n_components =  2, 
-       metric = "euclidean", min_dist = 0.3, init = "spca")
+<!-- <!--UMAP with n_neighbors = 15 and min_dist = 0.3--> -->
+<!-- ```{r} -->
+<!-- umap_prism1 <- triangular_3d_data |> -->
+<!--   select(-ID) |> -->
+<!--   umap(n_neighbors = 15, n_components =  2,  -->
+<!--        metric = "euclidean", min_dist = 0.3, init = "spca") -->
 
-umap_prism1 <- umap_prism1 |>
-  as_tibble()
+<!-- umap_prism1 <- umap_prism1 |> -->
+<!--   as_tibble() -->
 
-names(umap_prism1) <- c("UMAP1", "UMAP2")
+<!-- names(umap_prism1) <- c("UMAP1", "UMAP2") -->
 
-umap_prism1 <- umap_prism1 |>
-  mutate(ID = 1:NROW(umap_prism1))
+<!-- umap_prism1 <- umap_prism1 |> -->
+<!--   mutate(ID = 1:NROW(umap_prism1)) -->
 
-prism_scaled_obj_umap1 <- gen_scaled_data(
-  data = umap_prism1)
-umap_prism_scaled1 <- prism_scaled_obj_umap1$scaled_nldr
+<!-- prism_scaled_obj_umap1 <- gen_scaled_data( -->
+<!--   data = umap_prism1) -->
+<!-- umap_prism_scaled1 <- prism_scaled_obj_umap1$scaled_nldr -->
 
-nldr_prism4 <- umap_prism_scaled1 |> 
-  ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
-```
+<!-- nldr_prism4 <- umap_prism_scaled1 |>  -->
+<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
+<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
+<!-- ``` -->
 
-<!--UMAP with n_neighbors = 35 and min_dist = 0.5-->
-```{r}
-umap_prism2 <- triangular_3d_data |>
-  select(-ID) |>
-  umap(n_neighbors = 35, n_components =  2, 
-       metric = "euclidean", min_dist = 0.5, init = "spca")
+<!-- <!--UMAP with n_neighbors = 35 and min_dist = 0.5--> -->
+<!-- ```{r} -->
+<!-- umap_prism2 <- triangular_3d_data |> -->
+<!--   select(-ID) |> -->
+<!--   umap(n_neighbors = 35, n_components =  2,  -->
+<!--        metric = "euclidean", min_dist = 0.5, init = "spca") -->
 
-umap_prism2 <- umap_prism2 |>
-  as_tibble()
+<!-- umap_prism2 <- umap_prism2 |> -->
+<!--   as_tibble() -->
 
-names(umap_prism2) <- c("UMAP1", "UMAP2")
+<!-- names(umap_prism2) <- c("UMAP1", "UMAP2") -->
 
-umap_prism2 <- umap_prism2 |>
-  mutate(ID = 1:NROW(umap_prism2))
+<!-- umap_prism2 <- umap_prism2 |> -->
+<!--   mutate(ID = 1:NROW(umap_prism2)) -->
 
-prism_scaled_obj_umap2 <- gen_scaled_data(
-  data = umap_prism2)
-umap_prism_scaled2 <- prism_scaled_obj_umap2$scaled_nldr
+<!-- prism_scaled_obj_umap2 <- gen_scaled_data( -->
+<!--   data = umap_prism2) -->
+<!-- umap_prism_scaled2 <- prism_scaled_obj_umap2$scaled_nldr -->
 
-nldr_prism5 <- umap_prism_scaled2 |> 
-  ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
-```
+<!-- nldr_prism5 <- umap_prism_scaled2 |>  -->
+<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
+<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
+<!-- ``` -->
 
-<!--UMAP with n_neighbors = 69 and min_dist = 0.6-->
-```{r}
-umap_prism3 <- triangular_3d_data |>
-  select(-ID) |>
-  umap(n_neighbors = 69, n_components =  2, 
-       metric = "euclidean", min_dist = 0.6, init = "spca")
+<!-- <!--UMAP with n_neighbors = 69 and min_dist = 0.6--> -->
+<!-- ```{r} -->
+<!-- umap_prism3 <- triangular_3d_data |> -->
+<!--   select(-ID) |> -->
+<!--   umap(n_neighbors = 69, n_components =  2,  -->
+<!--        metric = "euclidean", min_dist = 0.6, init = "spca") -->
 
-umap_prism3 <- umap_prism3 |>
-  as_tibble()
+<!-- umap_prism3 <- umap_prism3 |> -->
+<!--   as_tibble() -->
 
-names(umap_prism3) <- c("UMAP1", "UMAP2")
+<!-- names(umap_prism3) <- c("UMAP1", "UMAP2") -->
 
-umap_prism3 <- umap_prism3 |>
-  mutate(ID = 1:NROW(umap_prism3))
+<!-- umap_prism3 <- umap_prism3 |> -->
+<!--   mutate(ID = 1:NROW(umap_prism3)) -->
 
-prism_scaled_obj_umap3 <- gen_scaled_data(
-  data = umap_prism3)
-umap_prism_scaled3 <- prism_scaled_obj_umap3$scaled_nldr
+<!-- prism_scaled_obj_umap3 <- gen_scaled_data( -->
+<!--   data = umap_prism3) -->
+<!-- umap_prism_scaled3 <- prism_scaled_obj_umap3$scaled_nldr -->
 
-nldr_prism6 <- umap_prism_scaled3 |> 
-  ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) 
-```
+<!-- nldr_prism6 <- umap_prism_scaled3 |>  -->
+<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
+<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
+<!-- ``` -->
 
-```{r}
-nldr_prism1 + nldr_prism2 + nldr_prism3 + 
-  nldr_prism4 + nldr_prism5 + nldr_prism6 +
-  plot_layout(ncol=3)
-```
+<!-- ```{r} -->
+<!-- nldr_prism1 + nldr_prism2 + nldr_prism3 +  -->
+<!--   nldr_prism4 + nldr_prism5 + nldr_prism6 + -->
+<!--   plot_layout(ncol=3) -->
+<!-- ``` -->
 
-<!--need to add details regarding which differences can I see with different layouts-->
+<!-- <!--need to add details regarding which differences can I see with different layouts--> -->
 
-## Deciding an appropriate fit
+<!-- ## Deciding an appropriate fit -->
 
-<!--select based on looking at b_non_empty vs log(Error)-->
-<!--select a choice for b_non_empty-->
+<!-- <!--select based on looking at b_non_empty vs log(Error)--> -->
+<!-- <!--select a choice for b_non_empty--> -->
 
-<!--compute error with tSNE and different perplexity values-->
-```{r}
-error_prism_tsne <- data.frame(matrix(nrow = 0, ncol = 0))
+<!-- <!--compute error with tSNE and different perplexity values--> -->
+<!-- ```{r} -->
+<!-- error_prism_tsne <- data.frame(matrix(nrow = 0, ncol = 0)) -->
 
-## To initialize number of bins along the x-axis
-bin1_vec_prism <- 4:25
+<!-- ## To initialize number of bins along the x-axis -->
+<!-- bin1_vec_prism <- 4:25  -->
 
-perplexity_vec <- c(6, 35, 58)
+<!-- perplexity_vec <- c(6, 35, 58) -->
 
-for (perplexity_val in perplexity_vec) {
+<!-- for (perplexity_val in perplexity_vec) { -->
 
-  tSNE_fit <- triangular_3d_data |>
-    select(-ID) |>
-    Rtsne(perplexity = perplexity_val)
+<!--   tSNE_fit <- triangular_3d_data |> -->
+<!--     select(-ID) |> -->
+<!--     Rtsne(perplexity = perplexity_val) -->
 
-  tsne_prism <- tSNE_fit$Y |>
-    as_tibble() |>
-    mutate(ID = row_number())
+<!--   tsne_prism <- tSNE_fit$Y |> -->
+<!--     as_tibble() |> -->
+<!--     mutate(ID = row_number()) -->
 
-  colnames(tsne_prism) <- c("tSNE1", "tSNE2", "ID")
+<!--   colnames(tsne_prism) <- c("tSNE1", "tSNE2", "ID") -->
 
-  prism_scaled_obj <- gen_scaled_data(
-    data = tsne_prism)
-  tsne_prism_scaled <- prism_scaled_obj$scaled_nldr
+<!--   prism_scaled_obj <- gen_scaled_data( -->
+<!--     data = tsne_prism) -->
+<!--   tsne_prism_scaled <- prism_scaled_obj$scaled_nldr -->
 
-  lim1 <- prism_scaled_obj$lim1
-  lim2 <- prism_scaled_obj$lim2
-  r2_prism <- diff(lim2)/diff(lim1)
+<!--   lim1 <- prism_scaled_obj$lim1 -->
+<!--   lim2 <- prism_scaled_obj$lim2 -->
+<!--   r2_prism <- diff(lim2)/diff(lim1)  -->
 
-  for (xbins in bin1_vec_prism) {
+<!--   for (xbins in bin1_vec_prism) { -->
 
-    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2
+<!--     bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2 -->
 
-    prism_model <- fit_highd_model(
-      training_data = triangular_3d_data,
-      emb_df = tsne_prism_scaled,
-      bin1 = xbins,
-      r2 = r2_prism,
-      is_bin_centroid = TRUE,
-      is_rm_lwd_hex = TRUE,
-      col_start_highd = "x"
-    )
+<!--     prism_model <- fit_highd_model( -->
+<!--       training_data = triangular_3d_data, -->
+<!--       emb_df = tsne_prism_scaled, -->
+<!--       bin1 = xbins, -->
+<!--       r2 = r2_prism, -->
+<!--       is_bin_centroid = TRUE, -->
+<!--       is_rm_lwd_hex = TRUE, -->
+<!--       col_start_highd = "x" -->
+<!--     ) -->
 
-    df_bin_centroids_prism <- prism_model$df_bin_centroids
-    df_bin_prism <- prism_model$df_bin
+<!--     df_bin_centroids_prism <- prism_model$df_bin_centroids -->
+<!--     df_bin_prism <- prism_model$df_bin -->
 
-    ## Compute error
-    error_df <- glance(
-      df_bin_centroids = df_bin_centroids_prism,
-      df_bin = df_bin_prism,
-      training_data = triangular_3d_data,
-      newdata = NULL,
-      type_NLDR = "tSNE",
-      col_start = "x") |>
-      mutate(bin1 = xbins,
-             bin2 = bin2,
-             b = bin1 * bin2,
-             b_non_empty = NROW(df_bin_centroids_prism),
-             type = paste0("perplexity: ", perplexity_val))
+<!--     ## Compute error -->
+<!--     error_df <- glance( -->
+<!--       df_bin_centroids = df_bin_centroids_prism, -->
+<!--       df_bin = df_bin_prism, -->
+<!--       training_data = triangular_3d_data, -->
+<!--       newdata = NULL, -->
+<!--       type_NLDR = "tSNE", -->
+<!--       col_start = "x") |> -->
+<!--       mutate(bin1 = xbins, -->
+<!--              bin2 = bin2, -->
+<!--              b = bin1 * bin2, -->
+<!--              b_non_empty = NROW(df_bin_centroids_prism), -->
+<!--              type = paste0("perplexity: ", perplexity_val)) -->
 
-    error_prism_tsne <- bind_rows(error_prism_tsne, error_df)
+<!--     error_prism_tsne <- bind_rows(error_prism_tsne, error_df) -->
 
-  }
+<!--   } -->
 
-}
-```
+<!-- } -->
+<!-- ``` -->
 
-<!--compute error with UMAP and different parameter values-->
-```{r}
-error_prism_umap <- data.frame(matrix(nrow = 0, ncol = 0))
+<!-- <!--compute error with UMAP and different parameter values--> -->
+<!-- ```{r} -->
+<!-- error_prism_umap <- data.frame(matrix(nrow = 0, ncol = 0)) -->
 
-## To initialize number of bins along the x-axis
-bin1_vec_prism <- 4:25
+<!-- ## To initialize number of bins along the x-axis -->
+<!-- bin1_vec_prism <- 4:25  -->
 
-param_list <- list(param1 = c(15, 0.3), param2 = c(35, 0.5), param3 = c(69, 0.6))
+<!-- param_list <- list(param1 = c(15, 0.3), param2 = c(35, 0.5), param3 = c(69, 0.6)) -->
 
-for (i in 1:length(param_list)) {
+<!-- for (i in 1:length(param_list)) { -->
 
-  n_neighbors_val <- param_list[[i]][1]
-  min_dist_val <- param_list[[i]][2]
+<!--   n_neighbors_val <- param_list[[i]][1] -->
+<!--   min_dist_val <- param_list[[i]][2] -->
 
-  umap_prism <- triangular_3d_data |>
-    select(-ID) |>
-    umap(n_neighbors = n_neighbors_val, n_components =  2, 
-         metric = "euclidean", min_dist = min_dist_val, init = "spca")
+<!--   umap_prism <- triangular_3d_data |> -->
+<!--     select(-ID) |> -->
+<!--     umap(n_neighbors = n_neighbors_val, n_components =  2,  -->
+<!--          metric = "euclidean", min_dist = min_dist_val, init = "spca") -->
 
-  umap_prism <- umap_prism |>
-    as_tibble()
+<!--   umap_prism <- umap_prism |> -->
+<!--     as_tibble() -->
 
-  names(umap_prism) <- c("UMAP1", "UMAP2")
+<!--   names(umap_prism) <- c("UMAP1", "UMAP2") -->
 
-  umap_prism <- umap_prism |>
-    mutate(ID = 1:NROW(umap_prism))
+<!--   umap_prism <- umap_prism |> -->
+<!--     mutate(ID = 1:NROW(umap_prism)) -->
 
-  prism_scaled_obj <- gen_scaled_data(
-    data = umap_prism)
-  umap_prism_scaled <- prism_scaled_obj$scaled_nldr
+<!--   prism_scaled_obj <- gen_scaled_data( -->
+<!--     data = umap_prism) -->
+<!--   umap_prism_scaled <- prism_scaled_obj$scaled_nldr -->
 
-  lim1 <- prism_scaled_obj$lim1
-  lim2 <- prism_scaled_obj$lim2
-  r2_prism <- diff(lim2)/diff(lim1)
+<!--   lim1 <- prism_scaled_obj$lim1 -->
+<!--   lim2 <- prism_scaled_obj$lim2 -->
+<!--   r2_prism <- diff(lim2)/diff(lim1)  -->
 
-  for (xbins in bin1_vec_prism) {
+<!--   for (xbins in bin1_vec_prism) { -->
 
-    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2
+<!--     bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2 -->
 
-    prism_model <- fit_highd_model(
-      training_data = triangular_3d_data,
-      emb_df = umap_prism_scaled,
-      bin1 = xbins,
-      r2 = r2_prism,
-      is_bin_centroid = TRUE,
-      is_rm_lwd_hex = TRUE,
-      col_start_highd = "x"
-    )
+<!--     prism_model <- fit_highd_model( -->
+<!--       training_data = triangular_3d_data, -->
+<!--       emb_df = umap_prism_scaled, -->
+<!--       bin1 = xbins, -->
+<!--       r2 = r2_prism, -->
+<!--       is_bin_centroid = TRUE, -->
+<!--       is_rm_lwd_hex = TRUE, -->
+<!--       col_start_highd = "x" -->
+<!--     ) -->
 
-    df_bin_centroids_prism <- prism_model$df_bin_centroids
-    df_bin_prism <- prism_model$df_bin
+<!--     df_bin_centroids_prism <- prism_model$df_bin_centroids -->
+<!--     df_bin_prism <- prism_model$df_bin -->
 
-    ## Compute error
-    error_df <- glance(
-      df_bin_centroids = df_bin_centroids_prism,
-      df_bin = df_bin_prism,
-      training_data = triangular_3d_data,
-      newdata = NULL,
-      type_NLDR = "UMAP",
-      col_start = "x") |>
-      mutate(bin1 = xbins,
-             bin2 = bin2,
-             b = bin1 * bin2,
-             b_non_empty = NROW(df_bin_centroids_prism),
-             type = paste0("n_neighbors: ", n_neighbors_val, 
-                           " min_dist: ", min_dist_val))
+<!--     ## Compute error -->
+<!--     error_df <- glance( -->
+<!--       df_bin_centroids = df_bin_centroids_prism, -->
+<!--       df_bin = df_bin_prism, -->
+<!--       training_data = triangular_3d_data, -->
+<!--       newdata = NULL, -->
+<!--       type_NLDR = "UMAP", -->
+<!--       col_start = "x") |> -->
+<!--       mutate(bin1 = xbins, -->
+<!--              bin2 = bin2, -->
+<!--              b = bin1 * bin2, -->
+<!--              b_non_empty = NROW(df_bin_centroids_prism), -->
+<!--              type = paste0("n_neighbors: ", n_neighbors_val,  -->
+<!--                            " min_dist: ", min_dist_val)) -->
 
-    error_prism_umap <- bind_rows(error_prism_umap, error_df)
+<!--     error_prism_umap <- bind_rows(error_prism_umap, error_df) -->
 
-  }
+<!--   } -->
 
-}
-```
+<!-- } -->
+<!-- ``` -->
 
-```{r}
-error_prism <- bind_rows(error_prism_tsne, error_prism_umap) 
-error_prism <- error_prism |>
-  filter(b <= 625)
-```
+<!-- ```{r} -->
+<!-- error_prism <- bind_rows(error_prism_tsne, error_prism_umap)  -->
+<!-- error_prism <- error_prism |> -->
+<!--   filter(b <= 625) -->
+<!-- ``` -->
 
-```{r}
-error_plot_prism <- ggplot(error_prism, 
-                           aes(x = b_non_empty, 
-                               y = log(Error), 
-                               group = type, 
-                               colour = type)) + 
-  geom_point(size = 1) +
-  geom_line() + 
-  geom_vline(xintercept = 112, linetype="solid",
-             color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3',
-                              '#ff7f00','#a6cee3','#a65628','#f781bf')) +
-  ylab("log(Error)") +
-  xlab("number of non-empty bins") +
-  theme(axis.text.x = element_text(size = 5),
-        axis.text.y = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, angle = 90),
-        legend.position = "bottom")
+<!-- ```{r} -->
+<!-- error_plot_prism <- ggplot(error_prism,  -->
+<!--                            aes(x = b_non_empty,  -->
+<!--                                y = log(Error),  -->
+<!--                                group = type,  -->
+<!--                                colour = type)) +  -->
+<!--   geom_point(size = 1) + -->
+<!--   geom_line() +  -->
+<!--   geom_vline(xintercept = 112, linetype="solid", -->
+<!--              color = "black", linewidth=0.8, alpha = 0.5) + -->
+<!--   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3', -->
+<!--                               '#ff7f00','#a6cee3','#a65628','#f781bf')) + -->
+<!--   ylab("log(Error)") + -->
+<!--   xlab("number of non-empty bins") + -->
+<!--   theme(axis.text.x = element_text(size = 5), -->
+<!--         axis.text.y = element_text(size = 5), -->
+<!--         axis.title.x = element_text(size = 5), -->
+<!--         axis.title.y = element_text(size = 5, angle = 90), -->
+<!--         legend.position = "bottom") -->
 
-error_plot_prism
-```
+<!-- error_plot_prism -->
+<!-- ``` -->
 
-<!--selected tSNE perplexity: 35-->
+<!-- <!--selected tSNE perplexity: 35--> -->
 
-## Fit the appropriate model
+<!-- ## Fit the appropriate model -->
 
-### Construct model in $2\text{-}D$
+<!-- ### Construct model in $2\text{-}D$ -->
 
-### Selecting parameter values for the model
+<!-- ### Selecting parameter values for the model -->
 
-#### Choice of bins
+<!-- #### Choice of bins -->
 
-#### Removal of low-density bins
+<!-- #### Removal of low-density bins -->
 
-#### Removing long edges
+<!-- #### Removing long edges -->
 
-```{r}
-## Compute hexbin parameters
-num_bins_x_prism <- 21
-lim1 <- prism_scaled_obj_tsne2$lim1
-lim2 <- prism_scaled_obj_tsne2$lim2
-r2_prism <- diff(lim2)/diff(lim1)
+<!-- ```{r} -->
+<!-- ## Compute hexbin parameters -->
+<!-- num_bins_x_prism <- 21 -->
+<!-- lim1 <- prism_scaled_obj_tsne2$lim1 -->
+<!-- lim2 <- prism_scaled_obj_tsne2$lim2 -->
+<!-- r2_prism <- diff(lim2)/diff(lim1)  -->
 
-prism_model <- fit_highd_model(
-  training_data = triangular_3d_data,
-  emb_df = tsne_prism_scaled2,
-  bin1 = num_bins_x_prism,
-  r2 = r2_prism,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = TRUE,
-  col_start_highd = "x"
-)
+<!-- prism_model <- fit_highd_model( -->
+<!--   training_data = triangular_3d_data, -->
+<!--   emb_df = tsne_prism_scaled2, -->
+<!--   bin1 = num_bins_x_prism, -->
+<!--   r2 = r2_prism, -->
+<!--   is_bin_centroid = TRUE, -->
+<!--   is_rm_lwd_hex = TRUE, -->
+<!--   col_start_highd = "x" -->
+<!-- ) -->
 
-df_bin_centroids_prism <- prism_model$df_bin_centroids
-df_bin_prism <- prism_model$df_bin
+<!-- df_bin_centroids_prism <- prism_model$df_bin_centroids -->
+<!-- df_bin_prism <- prism_model$df_bin -->
 
-## Triangulate bin centroids
-tr1_object_prism <- tri_bin_centroids(
-  df_bin_centroids_prism, x = "c_x", y = "c_y")
-tr_from_to_df_prism <- gen_edges(
-  tri_object = tr1_object_prism)
+<!-- ## Triangulate bin centroids -->
+<!-- tr1_object_prism <- tri_bin_centroids( -->
+<!--   df_bin_centroids_prism, x = "c_x", y = "c_y") -->
+<!-- tr_from_to_df_prism <- gen_edges( -->
+<!--   tri_object = tr1_object_prism) -->
 
-## Compute 2D distances
-distance_prism <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_prism,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
+<!-- ## Compute 2D distances -->
+<!-- distance_prism <- cal_2d_dist( -->
+<!--   tr_coord_df = tr_from_to_df_prism, -->
+<!--   start_x = "x_from", -->
+<!--   start_y = "y_from", -->
+<!--   end_x = "x_to", -->
+<!--   end_y = "y_to", -->
+<!--   select_vars = c("from", "to", "distance")) -->
 
-## To find the benchmark value
-benchmark_prism <- find_lg_benchmark(
-  distance_edges = distance_prism,
-  distance_col = "distance")
+<!-- ## To find the benchmark value -->
+<!-- benchmark_prism <- find_lg_benchmark( -->
+<!--   distance_edges = distance_prism, -->
+<!--   distance_col = "distance") -->
 
-trimesh_removed_prism <- vis_rmlg_mesh(
-  distance_edges = distance_prism,
-  benchmark_value = benchmark_prism,
-  tr_coord_df = tr_from_to_df_prism,
-  distance_col = "distance") #+
-#xlim(sc_xlims) + ylim(sc_ylims) +
-#interior_annotation("a", sc_ltr_pos)
+<!-- trimesh_removed_prism <- vis_rmlg_mesh( -->
+<!--   distance_edges = distance_prism, -->
+<!--   benchmark_value = benchmark_prism, -->
+<!--   tr_coord_df = tr_from_to_df_prism, -->
+<!--   distance_col = "distance") #+ -->
+<!-- #xlim(sc_xlims) + ylim(sc_ylims) + -->
+<!-- #interior_annotation("a", sc_ltr_pos) -->
 
-trimesh_removed_prism
-```
+<!-- trimesh_removed_prism -->
+<!-- ``` -->
 
-### Lifting the model into $p\text{-}D$
+<!-- ### Lifting the model into $p\text{-}D$ -->
 
-```{r}
-## Hexagonal binning to have regular hexagons
-hb_obj_prism <- hex_binning(
-  data = tsne_prism_scaled2,
-  bin1 = num_bins_x_prism,
-  r2 = r2_prism)
+<!-- ```{r} -->
+<!-- ## Hexagonal binning to have regular hexagons -->
+<!-- hb_obj_prism <- hex_binning( -->
+<!--   data = tsne_prism_scaled2, -->
+<!--   bin1 = num_bins_x_prism, -->
+<!--   r2 = r2_prism) -->
 
-tsne_data_with_hb_id <- hb_obj_prism$data_hb_id
+<!-- tsne_data_with_hb_id <- hb_obj_prism$data_hb_id -->
 
-df_all_prism <- dplyr::bind_cols(triangular_3d_data |> dplyr::select(-ID),
-                                 tsne_data_with_hb_id)
+<!-- df_all_prism <- dplyr::bind_cols(triangular_3d_data |> dplyr::select(-ID), -->
+<!--                                  tsne_data_with_hb_id) -->
 
-### Define type column
-df <- df_all_prism |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
+<!-- ### Define type column -->
+<!-- df <- df_all_prism |> -->
+<!--   dplyr::select(tidyselect::starts_with("x")) |> -->
+<!--   dplyr::mutate(type = "data") ## original dataset -->
 
-df_b <- df_bin_prism |>
-  dplyr::filter(hb_id %in% df_bin_centroids_prism$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
+<!-- df_b <- df_bin_prism |> -->
+<!--   dplyr::filter(hb_id %in% df_bin_centroids_prism$hexID) |> -->
+<!--   dplyr::mutate(type = "model") ## Data with summarized mean -->
 
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_prism$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
+<!-- ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data -->
+<!-- df_b <- df_b[match(df_bin_centroids_prism$hexID, df_b$hb_id),] |> -->
+<!--   dplyr::select(-hb_id) -->
 
-df_exe <- dplyr::bind_rows(df_b, df)
+<!-- df_exe <- dplyr::bind_rows(df_b, df) -->
 
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_prism |>
-  dplyr::filter(distance < benchmark_prism)
-## Since erase brushing is considerd.
+<!-- ## Set the maximum difference as the criteria -->
+<!-- distance_df_small_edges <- distance_prism |> -->
+<!--   dplyr::filter(distance < benchmark_prism) -->
+<!-- ## Since erase brushing is considerd. -->
 
-langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = distance_df_small_edges$from,
-                         lineTo = distance_df_small_edges$to,
-                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.8, NROW(df))),
-                         levelColors = c("#6a3d9a", "#33a02c"))
-```
+<!-- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], -->
+<!--                          lineFrom = distance_df_small_edges$from, -->
+<!--                          lineTo = distance_df_small_edges$to, -->
+<!--                          group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.8, NROW(df))), -->
+<!--                          levelColors = c("#6a3d9a", "#33a02c")) -->
+<!-- ``` -->
 
-What can I see in high-D?
+<!-- What can I see in high-D? -->
 
-- Four clusters that are really close
+<!-- - Four clusters that are really close  -->
 
-What can I see in 2D layout?
+<!-- What can I see in 2D layout? -->
 
-- Several clusters (more that what see in high-D)
+<!-- - Several clusters (more that what see in high-D) -->
 
-What can learn from the model?
+<!-- What can learn from the model? -->
 
-- There are four clusters, but not as the same cluster what data shows.
+<!-- - There are four clusters, but not as the same cluster what data shows.
 In 2D, the two of actual clusters (high-D clusters) are really close.
 Also, one of the big high-D cluster is separated to two sub clusters.
+-->
 

### 2025-03-30 paper-quollr.Rmd
Added initial link plots description

@@
 
-### Interactivity
+### Link plots
+
+There are mainly two interactive link plots can be generated.
+`show_link_plots()` helps to examine the fit.
+The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, and high-dimensional model data, and edge data where the from and to links of the edges.
+
+`show_error_link_plots()` helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
+The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
 

### 2025-04-04 paper-quollr.Rmd
Updated the intro

@@
 
-This paper presents the R package, `quollr` which introduce a new visual tool in determining which NLDR technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
+This paper presents the R package, `quollr` which introduce a new visual tool in determining which Non-linear dimension reduction (NLDR) technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
 The methodology of the algorithm is explained in *cite the methodology paper*.

### 2025-04-05 paper-quollr.Rmd
Found an application with FACS sells for Limb Muscles
Added MSE plot

@@
 }
+```
+
+```{r}
+# Plot MSE
+
+plot_mse <- function(error_df) {
+
+  ggplot(error_df,
+         aes(x = a1,
+             y = MSE,
+             colour = method)) +
+    geom_point(size = 0.8) +
+    geom_line(linewidth = 0.3) +
+    scale_y_log10() +
+    ylab("log(MSE)") +
+    xlab(expression(paste("binwidth (", a[1], ")"))) +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
+          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+          axis.ticks.x = element_line(),
+          axis.ticks.y = element_line(),
+          legend.position = "none",
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7),
+          axis.title.x = element_text(size = 7),
+          axis.title.y = element_text(size = 7))
+
+}
+
 ```
@@
 
+<https://www.nature.com/articles/s41586-018-0590-4#data-availability> <!--paper-->
+<https://figshare.com/articles/dataset/Robject_files_for_tissues_processed_by_Seurat/5821263/1> <!--data-->
+
+```{r}
+#| label: combine-error-data-muscles
+
+error_limb_umap <- read_rds("data/limb_muscles/error_limb_muscles_umap_n-neigbors_30_min-dist_0.1.rds")
+error_limb_tsne <- read_rds("data/limb_muscles/error_limb_muscles_tsne_perplexity_30.rds")
+error_limb_phate <- read_rds("data/limb_muscles/error_limb_muscles_phate_knn_5.rds")
+error_limb_trimap <- read_rds("data/limb_muscles/error_limb_muscles_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+error_limb_pacmap <- read_rds("data/limb_muscles/error_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+error_limb <- bind_rows(error_limb_umap, 
+                        error_limb_tsne,
+                        error_limb_phate,
+                        error_limb_trimap,
+                        error_limb_pacmap)
+
+error_limb <- error_limb |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+
+error_limb <- error_limb |>
+  mutate(method = factor(method,
+                         levels = c("UMAP_30_min_dist_0.1","tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2")))
+```
+
+```{r}
+#| label: error-comp-muscles
+
+error_plot_limb <- plot_mse(error_limb) +
+  scale_x_continuous(breaks = sort(unique(error_limb$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
+  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf'))
+
+```
+
 # Discussion

### 2025-04-05 paper-quollr.Rmd
Added NLDR plots

@@
 ```{r}
+#| label: read-limb-nldr
+# Read a variety of different NLDR representations of PBMC
+# and plot them on same aspect ratio
+clr_choice <- "#0077A3"
+umap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr1 <- umap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#999999') +
+  interior_annotation("a")
+
+tsne_limb <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_30.rds")
+
+nldr2 <- tsne_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("b")
+
+phate_limb <- read_rds("data/limb_muscles/facs_limb_muscles_phate_knn_5.rds")
+
+nldr3 <- phate_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("c")
+
+trimap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr4 <- trimap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("d")
+
+pacmap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr5 <- pacmap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  interior_annotation("e")
+```
+
+```{r}
 #| label: combine-error-data-muscles
@@
 
+<!-- Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
+Layout f is universally poor.
+Layouts a, c, g, h that show large separations between clusters are universally suboptimal.
+Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
+The choice of best is between layouts b and e, that have small separations between oddly shaped clusters.
+Layout e is the best choice.-->
+
+```{r}
+#| fig-cap: ""
+#| label: fig-limb-mse
+#| fig-pos: H
+#| out-width: 100%
+
+free(error_plot_limb) + wrap_plots(
+  nldr1, nldr2, nldr3, 
+  nldr4, nldr5, ncol = 2)
+```
+
 # Discussion

### 2025-04-05 paper-quollr.Rmd
Updated the layouts

@@
 
-error_limb_umap <- read_rds("data/limb_muscles/error_limb_muscles_umap_n-neigbors_30_min-dist_0.1.rds")
+error_limb_umap <- read_rds("data/limb_muscles/error_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
 error_limb_tsne <- read_rds("data/limb_muscles/error_limb_muscles_tsne_perplexity_30.rds")
@@
   mutate(method = factor(method,
-                         levels = c("UMAP_30_min_dist_0.1","tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2")))
+                         levels = c("UMAP_15_min_dist_0.1","tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2")))
 ```

### 2025-04-05 scripts/FACs/02_gen_embeddings.R

@@ -4,7 +4,7 @@ library(readr)
 library(conflicted)
 
 library(Rtsne)
-library(uwot)
+library(umap)
 library(phateR)
 library(reticulate)
 
@@ -18,7 +18,8 @@ use_python("~/miniforge3/envs/pcamp_env/bin/python")
 use_condaenv("pcamp_env")
 
 data <- read_rds(here::here("data/limb_muscles/facs_limb_muscles_pcs_10.rds"))
-
+data <- data |>
+  dplyr::select(-ID)
 
 ## tSNE (default)
 perplexity <- 30
@@ -56,7 +57,7 @@ UMAP_data <- UMAP_data |>
   mutate(ID = row_number())
 
 ## Run only once
-write_rds(UMAP_data, file = paste0("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_", n_neighbors, "_min-dist_", min_dist, ".rds"))
+write_rds(UMAP_data, file = paste0("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_", "15", "_min-dist_", "0.1", ".rds"))
 
 ## PHATE
 knn <- 5

### 2025-04-05 scripts/FACs/03_gen_mse_for_diff_methods.R

@@ -2,6 +2,8 @@ library(readr)
 library(quollr)
 library(dplyr)
 
+set.seed(20240110)
+
 data <- read_rds("data/limb_muscles/facs_limb_muscles_pcs_10.rds")
 
 ## For umap
@@ -48,14 +50,14 @@ for (xbins in bin1_vec_limb_muscles) {
            bin2 = bin2,
            b = bin1 * bin2,
            b_non_empty = NROW(df_bin_centroids_limb_muscles),
-           method = "UMAP_30_min_dist_0.1",
+           method = "UMAP_15_min_dist_0.1",
            a1 = a1)
 
   error_limb_muscles_umap <- bind_rows(error_limb_muscles_umap, error_df)
 
 }
 
-write_rds(error_limb_muscles_umap, "data/limb_muscles/error_limb_muscles_umap_n-neigbors_30_min-dist_0.1.rds")
+write_rds(error_limb_muscles_umap, "data/limb_muscles/error_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
 
 ###########
 

### 2025-04-05 paper-quollr.Rmd
Found the best layout

@@
   interior_annotation("e")
+
+tsne_limb2 <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_15.rds")
+
+nldr6 <- tsne_limb2 |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  interior_annotation("f")
 ```
@@
 
+error_limb_tsne2 <- read_rds("data/limb_muscles/error_limb_muscles_tsne_perplexity_15.rds")
+
 error_limb <- bind_rows(error_limb_umap, 
@@
                         error_limb_trimap,
-                        error_limb_pacmap)
+                        error_limb_pacmap,
+                        error_limb_tsne2)
 
@@
   filter(bin1 >= 5) |>
+  filter(a1 >= 0.03) |>
   group_by(method, a1) |>
@@
   mutate(method = factor(method,
-                         levels = c("UMAP_15_min_dist_0.1","tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2")))
+                         levels = c("UMAP_15_min_dist_0.1","tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2", "tsne_15")))
 ```
@@
   nldr1, nldr2, nldr3, 
-  nldr4, nldr5, ncol = 2)
+  nldr4, nldr5, nldr6, ncol = 2)
 ```

### 2025-04-05 scripts/FACs/03_gen_mse_for_diff_methods.R

@@ -279,4 +279,165 @@ for (xbins in bin1_vec_limb_muscles) {
 
 write_rds(error_limb_muscles_pacmap, "data/limb_muscles/error_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
 
+###########
 
+## For trimap
+trimap_limb_muscles <- read_rds("data/limb_muscles/facs_limb_muscles_trimap_n-inliers_5_n-outliers_4_n-random_3.rds")
+
+limb_scaled_obj_trimap <- gen_scaled_data(
+  data = trimap_limb_muscles)
+trimap_limb_muscles_scaled <- limb_scaled_obj_trimap$scaled_nldr
+
+lim1 <- limb_scaled_obj_trimap$lim1
+lim2 <- limb_scaled_obj_trimap$lim2
+r2_trimap <- diff(lim2)/diff(lim1)
+
+## To initialize number of bins along the x-axis
+bin1_vec_limb_muscles <- 2:32 #sqrt(NROW(data)/r2_trimap)
+
+error_limb_muscles_trimap <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_limb_muscles) {
+
+  hb_obj <- calc_bins_y(bin1 = xbins, r2 = r2_trimap, q = 0.1)
+
+  bin2 <- hb_obj$bin2
+  a1 <- hb_obj$a1
+
+  limb_model <- fit_highd_model(
+    highd_data = data,
+    nldr_data = trimap_limb_muscles_scaled,
+    bin1 = xbins,
+    r2 = r2_trimap,
+    q = 0.1,
+    is_bin_centroid = TRUE
+  )
+
+  df_bin_centroids_limb_muscles <- limb_model$df_bin_centroids
+  df_bin_limb_muscles <- limb_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    model_2d = df_bin_centroids_limb_muscles,
+    model_highd = df_bin_limb_muscles,
+    highd_data = data) |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_limb_muscles),
+           method = "trimap_n-inliers_5_n-outliers_4_n-random_3",
+           a1 = a1)
+
[…]

### 2025-04-05 paper-quollr.Rmd
Add script with additinal functions

@@
 ```{r}
-# Plot MSE
-
-plot_mse <- function(error_df) {
-
-  ggplot(error_df,
-         aes(x = a1,
-             y = MSE,
-             colour = method)) +
-    geom_point(size = 0.8) +
-    geom_line(linewidth = 0.3) +
-    scale_y_log10() +
-    ylab("log(MSE)") +
-    xlab(expression(paste("binwidth (", a[1], ")"))) +
-    theme_minimal() +
-    theme(panel.border = element_rect(fill = 'transparent'),
-          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-          axis.ticks.x = element_line(),
-          axis.ticks.y = element_line(),
-          legend.position = "none",
-          axis.text.x = element_text(size = 7),
-          axis.text.y = element_text(size = 7),
-          axis.title.x = element_text(size = 7),
-          axis.title.y = element_text(size = 7))
-
-}
-
+#| label: import-scripts
+source("scripts/additional_functions.R")
 ```

### 2025-04-05 paper-quollr.Rmd
Updated the trimesh

@@
 #| label: read-limb-nldr
-# Read a variety of different NLDR representations of PBMC
+# Read a variety of different NLDR representations of limb
 # and plot them on same aspect ratio
@@
 
+```{r}
+#| label: data-limb
+
+training_data_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pcs_10.rds")
+```
+
+```{r}
+#| label: tsne-num-bins-limb
+
+## Compute hexbin parameters
+num_bins_x_limb <- 16
+
+algo_obj_limb <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_limb, 
+  nldr_data = tsne_limb, 
+  num_x_bins = num_bins_x_limb)
+
+tsne_limb_scaled <- algo_obj_limb$nldr_scaled
+distance_limb <- algo_obj_limb$distance_df
+tr_from_to_df_limb <- algo_obj_limb$tr_from_to_df
+benchmark_limb <- algo_obj_limb$benchmark
+df_bin_centroids_limb <- algo_obj_limb$df_bin_centroids
+df_bin_limb <- algo_obj_limb$df_bin
+
+distance_df_small_edges_limb <- distance_limb |>
+  filter(distance < benchmark_limb) #benchmark_limb
+
+tr_from_to_df_limb <- right_join(
+  tr_from_to_df_limb, distance_df_small_edges_limb,
+  by = c("from", "to"))
+
+# umap_limb_scaled_with_cluster <- umap_limb_scaled |>
+#   mutate(cluster = umap_limb_scaled_with_cluster$cluster)
+
+trimesh_removed_limb <- ggplot() + 
+  geom_point(
+    data = tsne_limb_scaled,
+    aes(
+      x = emb1,
+      y = emb2
+    ),
+    alpha = 0.3,
+    color = "#a65628"
+  ) +
+    geom_segment(data = tr_from_to_df_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  #scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
+  interior_annotation("b1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+```
+
+```{r}
+#| label: tsne-best-num-bins-limb
+
+## Compute hexbin parameters
+num_bins_x_limb <- 13
+
+algo_obj_limb <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_limb, 
+  nldr_data = tsne_limb2, 
+  num_x_bins = num_bins_x_limb)
+
+tsne_limb_scaled_best <- algo_obj_limb$nldr_scaled
+distance_limb <- algo_obj_limb$distance_df
+tr_from_to_df_limb <- algo_obj_limb$tr_from_to_df
+benchmark_limb <- algo_obj_limb$benchmark
+df_bin_centroids_limb <- algo_obj_limb$df_bin_centroids
+df_bin_limb <- algo_obj_limb$df_bin
+
+distance_df_small_edges_limb <- distance_limb |>
+  filter(distance < benchmark_limb) #benchmark_limb
+
+tr_from_to_df_limb <- right_join(
+  tr_from_to_df_limb, distance_df_small_edges_limb,
+  by = c("from", "to"))
+
+# umap_limb_scaled_with_cluster <- umap_limb_scaled |>
+#   mutate(cluster = umap_limb_scaled_with_cluster$cluster)
+
+trimesh_removed_limb_best <- ggplot() + 
+  geom_point(
+    data = tsne_limb_scaled_best,
+    aes(
+      x = emb1,
+      y = emb2
+    ),
+    alpha = 0.3,
+    color = "#ff7f00"
+  ) +
+    geom_segment(data = tr_from_to_df_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  #scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
+  interior_annotation("b1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+```
+
+```{r}
+trimesh_removed_limb + trimesh_removed_limb_best +
+  plot_layout(ncol = 3)
+```
+
 # Discussion

### 2025-04-05 paper-quollr.Rmd
Try to optimize the model
Added langevitour output for both limb layouts

@@
 
-<!-- Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
+<!-- Assessing which of the 8 NLDR layouts on the limb3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
@@
 ## Compute hexbin parameters
-num_bins_x_limb <- 16
+num_bins_x_limb <- 25
 
@@
 ```{r}
+#| label: prep-limb-tsne-author-model-proj
+
+data_limb <- training_data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(hb_id %in% df_bin_centroids_limb$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb, df_b_limb)
+scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
+  as_tibble()
+
+scaled_limb_data <- scaled_limb |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_limb_data_model <- scaled_limb |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-limb-tsne-author-proj
+#| eval: false
+
+# data_limb_n <- data_limb |>
+#   select(-type) |>
+#   mutate(type = as.character(umap_limb_scaled_with_cluster$cluster))
+
+df_model_data_limb_n <- bind_rows(df_b_limb, data_limb)
+
+langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
+                         lineFrom = distance_df_small_edges_limb$from,
+                         lineTo = distance_df_small_edges_limb$to,
+                         group = factor(df_model_data_limb_n$type,
+                                        c("data", "model")),
+                         levelColors = c("#a65628", "#000000"))
+```
+
+```{r}
 #| label: tsne-best-num-bins-limb
@@
 ## Compute hexbin parameters
-num_bins_x_limb <- 13
+num_bins_x_limb <- 19
 
@@
 ```{r}
+#| label: prep-limb-tsne-best-model-proj
+
+data_limb <- training_data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(hb_id %in% df_bin_centroids_limb$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb, df_b_limb)
+scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
+  as_tibble()
+
+scaled_limb_data <- scaled_limb |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_limb_data_model <- scaled_limb |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-limb-tsne-best-proj
+#| eval: false
+
+# data_limb_n <- data_limb |>
+#   select(-type) |>
+#   mutate(type = as.character(umap_limb_scaled_with_cluster$cluster))
+
+df_model_data_limb_n <- bind_rows(df_b_limb, data_limb)
+
+langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
+                         lineFrom = distance_df_small_edges_limb$from,
+                         lineTo = distance_df_small_edges_limb$to,
+                         group = factor(df_model_data_limb_n$type,
+                                        c("data", "model")),
+                         levelColors = c("#ff7f00", "#000000"))
+```
+
+```{r}
 trimesh_removed_limb + trimesh_removed_limb_best +

### 2025-04-05 paper-quollr.Rmd
Deleted unnecessary scripts

@@
   #scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
-  interior_annotation("b1", cex = 2) +
+  interior_annotation("a1", cex = 2) +
   theme(

### 2025-04-05 scripts/function_scripts/functions_tri_hex_TriMAP.R

@@ -1,349 +1,0 @@
-Fit_TriMAP_data <- function(df_2_without_class, tem_dir){
-  write.csv(df_2_without_class, file.path(tem_dir, "df_2_without_class.csv"), row.names = FALSE,
-            quote = TRUE)
-}
-
-
-### Plot tSNE in 2D
-
-plot_TriMAP_2D <- function(tSNE_df){
-  tSNE_df_plot <- tSNE_df %>%
-    ggplot(aes(x = TriMAP1,
-               y = TriMAP2))+
-    geom_point() +
-    coord_equal()
-  return(tSNE_df_plot)
-}
-
-## To find optimal bin value
-
-find_opt_bin_val <- function (x)
-{
-  #h <- stats::IQR(x)
-  #if (h == 0)
-  #  h <- stats::mad(x, constant = 2)
-  #if (h > 0)
-  #  ceiling(diff(range(x))/(2 * h * length(x)^(-1/3)))
-  #else 1L
-  h <- 2 * IQR(x) / length(x)^(1/3) # bin width
-  return(h)
-}
-
-
-## To create hexbin object
-create_hex_bin <- function(tSNE_df, num_bins, shape){
-  hb <- hexbin(tSNE_df$TriMAP1, tSNE_df$TriMAP2, num_bins, IDs = TRUE, shape = shape)
-  return(hb)
-}
-
-### To Draw a hex bins with the points
-draw_hex_bin <- function(tSNE_df, num_bins){
-  hex_bin_plot <- ggplot(tSNE_df, aes(TriMAP1, TriMAP2)) +
-    geom_hex(bins = num_bins) + geom_point()+
-    scale_fill_viridis_c()
-
-  return(hex_bin_plot)
-}
-
-## To create the dataframe which contain coordinate values with the hexbin ID
-create_hexbin_ID_for_all <- function(df_2, tSNE_df, num_bins, shape){
-  ### Merge tSNE dataset in 2D with original dataset which contains 4D coordinates
-  df_2 <- df_2 %>%
-    mutate(ID=row_number())
-
-  df_new <- df_2 %>% inner_join(tSNE_df, by = "ID")
-
-  ### Fit hexbins and store hexbin IDs
-  hb <- create_hex_bin(tSNE_df, num_bins, shape)
-
-  ### Add hexbin Ids as a column to the original dataset
-  df_new$hb_id <- hb@cID
-
-  ## To change column names to lower case
-  names(df_new) <- tolower(names(df_new))
-
[…]

### 2025-04-05 scripts/gen_hex_centroid_from_scratch.R

@@ -1,231 +1,0 @@
-library(quollr)
-library(dplyr)
-library(ggplot2)
-
-## To compute the diameter of the hexagon
-cell_area <- 1
-cell_diameter <- sqrt(2 * cell_area / sqrt(3))
-
-# Define hexagonal grid parameters
-hex_size <- cell_diameter/2  # Size of the hexagons
-
-buffer_size <- 0.1  # Specify the buffer size here
-buffer_size <- hex_size/2  ## should less than hex_size
-
-
-num_bins_x <- 5 #Exact bynber of bins alng the x-axis
-
-shape_val <- calculate_effective_shape_value(.data = s_curve_noise_umap,
-                                             x = "UMAP1", y = "UMAP2")
-shape_val
-
-num_bins_y <- 2 * floor(((num_bins_x - 1) * shape_val)/sqrt(3) + 1.5001)
-num_bins_y
-
-x_bounds <- seq(min(s_curve_noise_umap$UMAP1) - buffer_size, max(s_curve_noise_umap$UMAP1) + buffer_size, length.out = 5)
-y_bounds <- seq(min(s_curve_noise_umap$UMAP2) - buffer_size, max(s_curve_noise_umap$UMAP2) + buffer_size, length.out = 12)
-
-
-# # Calculate middle points between consecutive values
-# middle_points_x <- (x_bounds[-length(x_bounds)] + x_bounds[-1]) / hex_size
-# middle_points_y <- (y_bounds[-length(y_bounds)] + y_bounds[-1]) / hex_size
-
-
-box_points <- expand.grid(x = x_bounds,
-                             y = y_bounds)
-
-## For each x-value find the even y value
-
-# Function to generate even y-values for a given x-value
-generate_even_y <- function(data) {
-  data |>
-    dplyr::arrange(y) |>
-    dplyr::mutate(is_even = ifelse(seq_along(y) %% 2 == 0, 1, 0))
-}
-
-# For each x-value, generate even y-values
-box_points <- box_points |>
-  dplyr::arrange(x) |>
-  dplyr::group_by(x) |>
-  dplyr::group_modify(~ generate_even_y(.x))
-
-# split_by_unique_x <- box_points |>
-#   dplyr::arrange(x) |>
-#   dplyr::group_by(x, .add = TRUE) |>
-#   dplyr::group_split()
-#
-# is_even_vec1 <- split_by_unique_x[[1]] |>
-#   dplyr::arrange(y) |>
[…]

### 2025-04-05 scripts/quollr_code_old.R

@@ -1,1428 +1,0 @@
-calculate_effective_x_bins <- function(.data, x = "UMAP1", cell_area = 1){
-
-  if (anyNA(.data[[rlang::as_string(rlang::ensym(x))]])) {
-    stop("NAs present")
-  }
-
-  if (any(is.infinite(.data |> dplyr::pull({{ x }})))) {
-    stop("Inf present")
-  }
-
-  if ((cell_area <= 0) || (is.infinite(cell_area))) {
-    stop("Invalid cell area value")
-
-  }
-
-  ## To compute the radius of the hexagon outer circle
-  cell_radius <- 2 * sqrt(2 * cell_area / sqrt(3))
-
-  ## To compute the range along x-axis
-  xwidth <- diff(range(.data |>
-                         dplyr::pull({{ x }})))
-
-  num_bins <- ceiling(xwidth/cell_diameter)
-  #num_bins <- floor(xwidth/cell_diameter + 1.5001)
-  num_bins
-
-}
-
-calculate_effective_y_bins <- function(.data, x = "UMAP2", y = "UMAP2", shape_val = NA, num_bins_x){
-
-  if (anyNA(.data[[rlang::as_string(rlang::ensym(y))]])) {
-    stop("NAs present")
-  }
-
-  if (any(is.infinite(.data |> dplyr::pull({{ y }})))) {
-    stop("Inf present")
-  }
-
-  if (is.na(shape_val)) {
-    shape_val <- calculate_effective_shape_value(.data = .data, x = x, y = y)
-
-  }
-
-  num_bins <- 2 * floor(((num_bins_x - 1) * shape_val)/sqrt(3) + 1.5001)
-  num_bins
-
-}
-
-calculate_effective_shape_value <- function(.data, x = UMAP1, y = UMAP2){
-
-  if (any(is.na(.data |> dplyr::pull({{ x }}))) || any(is.na(.data |> dplyr::pull({{ y }})))) {
-    stop("NAs present")
-  }
-
-  if (any(is.infinite(.data |> dplyr::pull({{ x }}))) || any(is.infinite(.data |> dplyr::pull({{ y }})))) {
-    stop("Inf present")
-  }
-
-  if ((length(.data |> dplyr::pull({{ x }})) == 1) || (length(.data |> dplyr::pull({{ y }})) == 1)) {
-    stop("Presence one observation only")
-
-  }
-
-  ## To compute the range along x-axis
[…]

### 2025-04-06 paper-quollr.Rmd
Added colors according to the clusters

@@
 training_data_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pcs_10.rds")
+
+cluster_df <- read_rds("data/limb_muscles/facs_limb_muscles_cluster_df.rds")
 ```
@@
 
-# umap_limb_scaled_with_cluster <- umap_limb_scaled |>
-#   mutate(cluster = umap_limb_scaled_with_cluster$cluster)
+tsne_limb_scaled_with_cluster <- inner_join(tsne_limb_scaled, cluster_df, by = "ID") |>
+  mutate(cluster.ids = as.character(cluster.ids))
 
@@
   geom_point(
-    data = tsne_limb_scaled,
+    data = tsne_limb_scaled_with_cluster,
     aes(
       x = emb1,
-      y = emb2
+      y = emb2,
+      color = cluster.ids
     ),
-    alpha = 0.3,
-    color = "#a65628"
+    alpha = 0.3#,
+    #color = "#a65628"
   ) +
@@
                linewidth = 1) +
-  #scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
   interior_annotation("a1", cex = 2) +
@@
 
-# data_limb_n <- data_limb |>
-#   select(-type) |>
-#   mutate(type = as.character(umap_limb_scaled_with_cluster$cluster))
-
-df_model_data_limb_n <- bind_rows(df_b_limb, data_limb)
+data_limb_n <- data_limb |>
+  select(-type) |>
+  mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
+
+df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
 
@@
                          group = factor(df_model_data_limb_n$type,
-                                        c("data", "model")),
-                         levelColors = c("#a65628", "#000000"))
+                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
+                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
 ```
@@
 
-# umap_limb_scaled_with_cluster <- umap_limb_scaled |>
-#   mutate(cluster = umap_limb_scaled_with_cluster$cluster)
+tsne_limb_scaled_best_with_cluster <- inner_join(tsne_limb_scaled_best, cluster_df, by = "ID") |>
+  mutate(cluster.ids = as.character(cluster.ids))
 
@@
   geom_point(
-    data = tsne_limb_scaled_best,
+    data = tsne_limb_scaled_best_with_cluster,
     aes(
       x = emb1,
-      y = emb2
+      y = emb2,
+      color = cluster.ids
     ),
-    alpha = 0.3,
-    color = "#ff7f00"
+    alpha = 0.3#,
+    #color = "#ff7f00"
   ) +
@@
                linewidth = 1) +
-  #scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
   interior_annotation("b1", cex = 2) +
@@
 
-# data_limb_n <- data_limb |>
-#   select(-type) |>
-#   mutate(type = as.character(umap_limb_scaled_with_cluster$cluster))
-
-df_model_data_limb_n <- bind_rows(df_b_limb, data_limb)
+df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
 
@@
                          group = factor(df_model_data_limb_n$type,
-                                        c("data", "model")),
-                         levelColors = c("#ff7f00", "#000000"))
+                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
+                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
 ```

### 2025-04-08 paper-quollr.Rmd
Updated layouts with same binwidths
paper renders in both html and pdf

@@
 output: 
- rjtools::rjournal_web_article:
-    css: "style.css"
+ rjtools::rjournal_article:
     keep_md: true
-bibliography: RJreferences.bib
+bibliography: paper-quollr.bib
 editor_options: 
@@
 #| label: tsne-num-bins-limb
+#| eval: false
 
@@
 #| label: prep-limb-tsne-author-model-proj
+#| eval: false
 
@@
 #| label: tsne-best-num-bins-limb
+#| eval: false
 
@@
 #| label: prep-limb-tsne-best-model-proj
+#| eval: false
 
@@
 ```{r}
+#| eval: false
 trimesh_removed_limb + trimesh_removed_limb_best +

### 2025-04-08 paper-quollr.tex
paper renders in both html and pdf

@@
 
-This paper presents the R package, \texttt{quollr} which introduce a new visual tool in determining which NLDR technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
+This paper presents the R package, \texttt{quollr} which introduce a new visual tool in determining which Non-linear dimension reduction (NLDR) technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
 The methodology of the algorithm is explained in \emph{cite the methodology paper}.
-Furthermore, the \texttt{quollr} package enables users to perform hexagonal binning (Carr et al.
-2023), resulting in the generation of regular hexagons.
+Furthermore, the \texttt{quollr} package enables users to perform hexagonal binning \citep{dan2023}, resulting in the generation of regular hexagons.
 The software is available from the Comprehensive R Archive Network (CARN) at \url{https://CRAN.R-project.org/package=quollr}.
@@
 #> $df_bin
-#> # A tibble: 37 x 8
+#> # A tibble: 72 x 8
 #>    hb_id     x1    x2    x3        x4        x5        x6         x7
 #>    <int>  <dbl> <dbl> <dbl>     <dbl>     <dbl>     <dbl>      <dbl>
-#>  1    27 -0.706   1.70   -1.65   0.00918  0.00261 -0.0236  -0.00199 
-#>  2    28 -0.214   1.45   -1.98  -0.00966 -0.00370  0.0573   0.000307
-#>  3    40  0.356   1.74   -1.89   0.0127  -0.00773 -0.0219  -0.000757
-#>  4    50 -0.209   0.453  -1.98   0.0185   0.0189  -0.0999  -0.00412 
-#>  5    52  0.00161 1.14   -2.00   0.00206 -0.00940 -0.0974  -0.00796 
-#>  6    61 -0.671   0.0922 -1.74  -0.00458  0.0127  -0.0613  -0.00643 
-#>  7    62 -0.375   0.328  -1.89  -0.00226 -0.00651 -0.0291   0.00120 
-#>  8    74 -0.175   0.0562 -1.98   0.00204 -0.00179  0.0112  -0.00111 
-#>  9    75  0.325   0.0821 -1.95   0.00292  0.0197  -0.0165  -0.000881
-#> 10   127  0.704   0.454  -0.290 -0.00394  0.0118  -0.00941 -0.00348 
-#> # i 27 more rows
+#>  1    13  0.996  1.77  1.07 -0.0117    0.00979  -0.0725   -0.00321  
+#>  2    14  0.829  1.79  1.50 -0.000753 -0.00194  -0.00936  -0.000454 
+#>  3    15  0.307  1.78  1.94 -0.00384   0.00212  -0.00117  -0.000159 
+#>  4    16 -0.175  1.74  1.98  0.00143   0.00320   0.00450  -0.0000763
+#>  5    18 -0.493  1.78  1.87 -0.0100    0.000680  0.0799   -0.000835 
+#>  6    26  0.962  1.22  1.22  0.00107  -0.00260  -0.000948  0.000234 
+#>  7    27  0.631  1.28  1.75  0.000266  0.00174  -0.00158   0.00114  
+#>  8    28  0.139  1.20  1.97 -0.00183   0.000279  0.00177  -0.000666 
+#>  9    29 -0.187  1.00  1.98 -0.00228   0.00461  -0.0506   -0.00472  
+#> 10    30 -0.562  1.23  1.82 -0.000563  0.00166  -0.00923  -0.000111 
+#> # i 62 more rows
 #> 
 #> $df_bin_centroids
-#> # A tibble: 37 x 5
-#>    hexID     c_x    c_y std_counts drop_empty
-#>    <int>   <dbl>  <dbl>      <dbl> <lgl>     
-#>  1    27  0.127  0.0295        1   FALSE     
-#>  2    28  0.226  0.0295        0.2 FALSE     
-#>  3    40  0.275  0.115         0.6 FALSE     
-#>  4    50  0.0287 0.200         0.2 FALSE     
-#>  5    52  0.226  0.200         0.2 FALSE     
-#>  6    61 -0.0207 0.286         0.2 FALSE     
-#>  7    62  0.0780 0.286         0.8 FALSE     
-#>  8    74  0.0287 0.371         0.4 FALSE     
-#>  9    75  0.127  0.371         0.2 FALSE     
-#> 10   127  0.522  0.713         0.4 FALSE     
-#> # i 27 more rows
+#> # A tibble: 72 x 6
+#>    hexID      c_x      c_y bin_counts std_counts drop_empty
+#>    <int>    <dbl>    <dbl>      <int>      <dbl> <lgl>     
+#>  1    13 -0.0455  0.000514          2     0.0194 FALSE     
+#>  2    14  0.0634  0.000514         81     0.786  FALSE     
+#>  3    15  0.172   0.000514         56     0.544  FALSE     
+#>  4    16  0.281   0.000514         55     0.534  FALSE     
+#>  5    18  0.499   0.000514          2     0.0194 FALSE     
+#>  6    26  0.00894 0.0949           70     0.680  FALSE     
+#>  7    27  0.118   0.0949           78     0.757  FALSE     
+#>  8    28  0.227   0.0949           86     0.835  FALSE     
+#>  9    29  0.336   0.0949            5     0.0485 FALSE     
+#> 10    30  0.445   0.0949           54     0.524  FALSE     
+#> # i 62 more rows
 \end{verbatim}
@@
 #> $bin2
-#> [1] 27
+#> [1] 13
 #> 
 #> $a1
-#> [1] 0.09596332
+#> [1] 0.1089367
+#> 
+#> $a2
+#> [1] 0.09434198
 \end{verbatim}
@@
 \begin{verbatim}
+#> $a1
+#> [1] 0.1089367
+#> 
+#> $a2
+#> [1] 0.09434198
+#> 
 #> $bins
-#> [1] 12 27
+#> [1] 12 13
 #> 
 #> $start_point
-#> [1] -0.070000 -0.141359
+#> [1] -0.10000000 -0.09382762
 #> 
 #> $centroids
-#> # A tibble: 324 x 3
+#> # A tibble: 156 x 3
 #>    hexID      c_x     c_y
 #>    <int>    <dbl>   <dbl>
-#>  1     1 -0.07   -0.141
-#>  2     2  0.0287 -0.141
-#>  3     3  0.127  -0.141
-#>  4     4  0.226  -0.141
-#>  5     5  0.325  -0.141
-#>  6     6  0.423  -0.141
-#>  7     7  0.522  -0.141
-#>  8     8  0.621  -0.141
-#>  9     9  0.719  -0.141
-#> 10    10  0.818  -0.141
-#> # i 314 more rows
+#>  1     1 -0.1     -0.0938
+#>  2     2  0.00894 -0.0938
+#>  3     3  0.118   -0.0938
+#>  4     4  0.227   -0.0938
+#>  5     5  0.336   -0.0938
+#>  6     6  0.445   -0.0938
+#>  7     7  0.554   -0.0938
+#>  8     8  0.663   -0.0938
+#>  9     9  0.771   -0.0938
+#> 10    10  0.880   -0.0938
+#> # i 146 more rows
 #> 
 #> $hex_poly
-#> # A tibble: 1,944 x 3
+#> # A tibble: 936 x 3
 #>    hex_poly_id        x       y
 #>          <int>    <dbl>   <dbl>
-#>  1           1 -0.07   -0.0860
-#>  2           1 -0.118  -0.114 
-#>  3           1 -0.118  -0.169 
-#>  4           1 -0.07   -0.197 
-#>  5           1 -0.0220 -0.169 
-#>  6           1 -0.0220 -0.114 
-#>  7           2  0.0287 -0.0860
-#>  8           2 -0.0193 -0.114 
-#>  9           2 -0.0193 -0.169 
-#> 10           2  0.0287 -0.197 
-#> # i 1,934 more rows
+#>  1           1 -0.1     -0.0309
+#>  2           1 -0.154   -0.0624
+#>  3           1 -0.154   -0.125 
+#>  4           1 -0.1     -0.157 
+#>  5           1 -0.0455  -0.125 
+#>  6           1 -0.0455  -0.0624
+#>  7           2  0.00894 -0.0309
+#>  8           2 -0.0455  -0.0624
+#>  9           2 -0.0455  -0.125 
+#> 10           2  0.00894 -0.157 
+#> # i 926 more rows
 #> 
 #> $data_hb_id
-#> # A tibble: 75 x 4
-#>     UMAP1  UMAP2    ID hb_id
+#> # A tibble: 3,750 x 4
+#>      emb1   emb2    ID hb_id
 #>     <dbl>  <dbl> <int> <int>
-#>  1 0.0804 0.320      1    62
-#>  2 0.739  1.00       2   165
-#>  3 0.840  1.08       3   178
-#>  4 0.167  0.0432     4    27
-#>  5 0.263  0.398      6    76
-#>  6 0.838  2.01       7   310
-#>  7 0.734  0.972      8   165
-#>  8 0.627  0.721      9   128
-#>  9 0.810  1.01      11   178
-#> 10 0.903  1.87      12   299
-#> # i 65 more rows
+#>  1 0.276  0.915      1   136
+#>  2 0.927  0.347      2    70
+#>  3 0.810  0.242      3    45
+#>  4 0.137  0.657      5    99
+#>  5 0.476  0.799      6   114
+#>  6 0.0485 0.0657     7    26
+#>  7 0.737  0.851      9   129
+#>  8 0.833  0.224     10    45
+#>  9 0.909  0.371     11    70
+#> 10 0.0890 0.194     12    38
+#> # i 3,740 more rows
 #> 
 #> $std_cts
-#> # A tibble: 45 x 3
+#> # A tibble: 72 x 3
 #>    hb_id     n std_counts
 #>    <int> <int>      <dbl>
-#>  1    27     5        1  
-#>  2    28     1        0.2
-#>  3    38     1        0.2
-#>  4    40     3        0.6
-#>  5    50     1        0.2
-#>  6    52     1        0.2
-#>  7    61     1        0.2
-#>  8    62     4        0.8
-#>  9    64     1        0.2
-#> 10    74     2        0.4
-#> # i 35 more rows
+#>  1    13     2     0.0194
+#>  2    14    81     0.786 
+#>  3    15    56     0.544 
+#>  4    16    55     0.534 
+#>  5    18     2     0.0194
+#>  6    26    70     0.680 
+#>  7    27    78     0.757 
+#>  8    28    86     0.835 
+#>  9    29     5     0.0485
+#> 10    30    54     0.524 
+#> # i 62 more rows
 #> 
 #> $tot_bins
-#> [1] 324
+#> [1] 156
 #> 
 #> $non_bins
-#> [1] 45
+#> [1] 72
 #> 
 #> $pts_bins
-#> # A tibble: 45 x 2
+#> # A tibble: 72 x 2
 #>    hexID pts_list     
 #>    <int> <named list> 
-#>  1    62 <int [75]>  
-#>  2   165 <int [75]>  
-#>  3   178 <int [75]>  
-#>  4    27 <int [75]>  
-#>  5    76 <int [75]>  
-#>  6   310 <int [75]>  
-#>  7   128 <int [75]>  
-#>  8   299 <int [75]>  
-#>  9    74 <int [75]>  
-#> 10   287 <int [75]>  
-#> # i 35 more rows
+#>  1   136 <int [3,750]>
+#>  2    70 <int [3,750]>
+#>  3    45 <int [3,750]>
+#>  4    99 <int [3,750]>
+#>  5   114 <int [3,750]>
+#>  6    26 <int [3,750]>
+#>  7   129 <int [3,750]>
+#>  8    38 <int [3,750]>
+#>  9   111 <int [3,750]>
+#> 10    14 <int [3,750]>
+#> # i 62 more rows
 #> 
@@
 
-\subsection{Tests}\label{tests}
+\subsubsection{Link plots}\label{link-plots}
 
-All functions have tests written and implemented using the \CRANpkg{testthat} (Wickham 2011) in R.
+There are mainly two interactive link plots can be generated.
+\texttt{show\_link\_plots()} helps to examine the fit.
+The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, and high-dimensional model data, and edge data where the from and to links of the edges.
 
-These tests illuminated the issues that allowed us to make meaningful changes and understand some pitfalls of the package.
+\texttt{show\_error\_link\_plots()} helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
+The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
+
+\begin{verbatim}
+show_link_plots(
+  point_df = df_exe, 
+  edge_df = distance_small_df
+  )
+\end{verbatim}
+
+\begin{verbatim}
+show_error_link_plots(
+  point_df = df_exe, 
+  edge_df = distance_small_df
+  )
+\end{verbatim}
+
+--\textgreater{}
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
+
+--\textgreater{}
 
@@
 
-To illustrate the algorithm, we use \(5\text{-}D\) simulated data, which we call the ``triangular\_3d\_data''.
-This dataset is generated using the \texttt{tri\_3d} function from the \texttt{cardinalR} package.
-The dataset generation starts with the initialization of a starting point \(p_0 = (x_0, y_0, z_0)\) randomly selected within the \(3\text{-}D\) space using a uniform distribution.
-Four fixed corner points of a tetrahedron are defined as \(c_1 = (0, 0, 0)\), \(c_2 = (1, 0, 0)\), \(c_3 = (0.5, 1, 0)\), and \(c_4 = (0.5, 0.5, 1)\).
-For each point \(p_i\) where \(i = 1, 2, \ldots, n\), one of the corner points \(c_j\) (where \(j \in \{1, 2, 3, 4\}\)) is selected randomly, and the new point \(p_i\) is computed as the midpoint between the current point \(p_{i-1}\) and the selected corner point \(c_j\): \(p_i = \frac{p_{i-1} + c_j}{2}\).
-This iterative process ensures that each new point moves closer to one of the corners, creating a fractal-like triangular distribution of points.
-The coordinates of each generated point \(p_i\) form the variables \(X_1, X_2, X_3\).
-The remaining variables \(X_4, X_5\) are all uniform error, with small variance.
-We would consider \(T=(X_1, X_2, X_3)\) to be the true model.
+Need to find another application
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
+\url{https://www.nature.com/articles/s41586-018-0590-4\#data-availability}
+\url{https://figshare.com/articles/dataset/Robject_files_for_tissues_processed_by_Seurat/5821263/1}
 
-\subsection{Deciding an appropriate fit}\label{deciding-an-appropriate-fit}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}
-
-\subsection{Fit the appropriate model}\label{fit-the-appropriate-model}
-
-\subsubsection{\texorpdfstring{Construct model in \(2\text{-}D\)}{Construct model in 2\textbackslash text\{-\}D}}\label{construct-model-in-2text-d}
-
-\subsubsection{Selecting parameter values for the model}\label{selecting-parameter-values-for-the-model}
-
-\paragraph{Choice of bins}\label{choice-of-bins}
-
-\paragraph{Removal of low-density bins}\label{removal-of-low-density-bins}
-
-\paragraph{Removing long edges}\label{removing-long-edges}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-32-1}
-
-\subsubsection{\texorpdfstring{Lifting the model into \(p\text{-}D\)}{Lifting the model into p\textbackslash text\{-\}D}}\label{lifting-the-model-into-ptext-d}
-
-What can I see in high-D?
-
-\begin{itemize}
-\tightlist
-\item
-  Four clusters that are really close
-\end{itemize}
-
-What can I see in 2D layout?
-
-\begin{itemize}
-\tightlist
-\item
-  Several clusters (more that what see in high-D)
-\end{itemize}
-
-What can learn from the model?
-
-\begin{itemize}
-\tightlist
-\item
-  There are four clusters, but not as the same cluster what data shows.
-In 2D, the two of actual clusters (high-D clusters) are really close.
-Also, one of the big high-D cluster is separated to two sub clusters.
-\end{itemize}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-limb-mse-1} \end{figure}
 
@@
 
-This article is created using \CRANpkg{knitr} (Xie 2015) and \CRANpkg{rmarkdown} (Xie, Allaire, and Grolemund 2018) in R with the \texttt{rjtools::rjournal\_article} template.
+This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
 The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-dan2023}
-Carr, Dan, ported by Nicholas Lewin-Koh, Martin Maechler, and contains copies of lattice functions written by Deepayan Sarkar.
-2023.
-\emph{Hexbin: Hexagonal Binning Routines}.
-\url{https://CRAN.R-project.org/package=hexbin}.
-
-\bibitem[\citeproctext]{ref-testthat}
-Wickham, Hadley. 2011. {``Testthat: Get Started with Testing.''} \emph{The R Journal} 3: 5--10. \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}.
-
-\bibitem[\citeproctext]{ref-knitr}
-Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
-
-\bibitem[\citeproctext]{ref-rmarkdown}
-Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{{R} Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
-
-\end{CSLReferences}
+\bibliography{paper-quollr.bib}
 

### 2025-04-08 paper-quollr.Rmd
Updated layout models

--- 2025-04-06 paper-quollr.Rmd
@@
 ## Compute hexbin parameters
-num_bins_x_limb <- 25
+num_bins_x_limb <- 19
 
@@
 
+tr_from_to_df_limb <- tr_from_to_df_limb |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(108, 99, 11, 231)))
+
 distance_df_small_edges_limb <- distance_limb |>
-  filter(distance < benchmark_limb) #benchmark_limb
+  filter(distance < 0.1) #benchmark_limb
 
@@
 ## Compute hexbin parameters
-num_bins_x_limb <- 19
+num_bins_x_limb <- 18
 
@@
 
+tr_from_to_df_limb <- tr_from_to_df_limb |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(298, 34, 48, 326)))
+
 distance_df_small_edges_limb <- distance_limb |>
-  filter(distance < benchmark_limb) #benchmark_limb
+  filter(distance < benchmark_limb) #|>
+  #mutate(ID = row_number()) |>
+  #filter(!(ID %in% c(298, 34, 48))) |>
+  #select(-ID) #benchmark_limb
 

### 2025-04-08 paper-quollr.Rmd

--- 2025-04-08 paper-quollr.Rmd
@@
 #| label: fig-limb-mse
-#| fig-pos: H
 #| out-width: 100%
@@
 distance_df_small_edges_limb <- distance_limb |>
-  filter(distance < benchmark_limb) #benchmark_limb
+  filter(distance < 0.1) #benchmark_limb
 

### 2025-04-08 paper-quollr.Rmd

@@
 
+tr_from_to_df_limb <- tr_from_to_df_limb |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(108, 99, 11, 231)))
+
 distance_df_small_edges_limb <- distance_limb |>

### 2025-04-08 paper-quollr.Rmd
Corrected the code

--- 2025-04-08 paper-quollr.Rmd
@@
 )
-
-interior_annotation <- function(label, position = c(0.92, 0.92)) {
-  annotation_custom(grid::textGrob(label = label,
-      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
-      gp = grid::gpar(cex = 1, col="grey70")))
-}
 ```
@@
 
-<!-- ## Tests -->
-
-<!-- All functions have tests written and implemented using the \CRANpkg{testthat} package in R.
-These tests illuminated issues, allowing meaningful changes and providing insights into some pitfalls of the package.
-For example, several tests check if there is an NLDR point at the intersection of two bins and determine which bin this point falls into.
--->
-
-<!-- Figure xx shows an example of a simulated test set. When creating this test data set,  -->
-
-<!-- <!--discuss a test with how a point in a intersection of two hexagonal going to evaluate--> -->
-<!-- <!--selecting the default values to q as 0.1--> -->
-
-<!-- # Example -->
-
-<!-- To illustrate the algorithm, we use $5\text{-}D$ simulated data, which we call the "triangular_3d_data".
-This dataset is generated using the `tri_3d` function from the `cardinalR` package.
-The dataset generation starts with the initialization of a starting point $p_0 = (x_0, y_0, z_0)$ randomly selected within the $3\text{-}D$ space using a uniform distribution.
-Four fixed corner points of a tetrahedron are defined as $c_1 = (0, 0, 0)$, $c_2 = (1, 0, 0)$, $c_3 = (0.5, 1, 0)$, and $c_4 = (0.5, 0.5, 1)$.
-For each point $p_i$ where $i = 1, 2, \ldots, n$, one of the corner points $c_j$ (where $j \in \{1, 2, 3, 4\}$) is selected randomly, and the new point $p_i$ is computed as the midpoint between the current point $p_{i-1}$ and the selected corner point $c_j$: $p_i = \frac{p_{i-1} + c_j}{2}$.
-This iterative process ensures that each new point moves closer to one of the corners, creating a fractal-like triangular distribution of points.
-The coordinates of each generated point $p_i$ form the variables $X_1, X_2, X_3$.
-The remaining variables $X_4, X_5$ are all uniform error, with small variance.
-We would consider $T=(X_1, X_2, X_3)$ to be the true model.
--->
-
-<!-- ```{r} -->
-<!-- triangular_3d_data <- tri_3d( -->
-<!--   n = 1250, num_noise = 2,  -->
-<!--   min_n = -0.05, max_n = 0.05) |> -->
-<!--   as_tibble()  -->
-
-<!-- colnames(triangular_3d_data) <- paste0("x", 1:NCOL(triangular_3d_data)) -->
-
-<!-- langevitour::langevitour(triangular_3d_data) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- triangular_3d_data <- triangular_3d_data |> -->
-<!--   mutate(ID = row_number()) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- sc_ltr_pos <- c(0.08, 0.96) -->
-<!-- ``` -->
-
-<!-- <!--tSNE with perplexity = 6--> -->
-<!-- ```{r} -->
-<!-- tSNE_fit1 <- triangular_3d_data |> -->
-<!--   select(-ID) |> -->
-<!--   Rtsne(perplexity = 6)  -->
-
-<!-- tsne_prism1 <- tSNE_fit1$Y |> -->
-<!--   as_tibble() |> -->
-<!--   mutate(ID = row_number()) -->
-
-<!-- colnames(tsne_prism1) <- c("tSNE1", "tSNE2", "ID") -->
-
-<!-- prism_scaled_obj_tsne1 <- gen_scaled_data( -->
-<!--   data = tsne_prism1) -->
-<!-- tsne_prism_scaled1 <- prism_scaled_obj_tsne1$scaled_nldr -->
-
-<!-- nldr_prism1 <- tsne_prism_scaled1 |>  -->
-<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
-<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
-<!-- ``` -->
-
-<!-- <!--tSNE with perplexity = 35--> -->
-<!-- ```{r} -->
-<!-- tSNE_fit2 <- triangular_3d_data |> -->
-<!--   select(-ID) |> -->
-<!--   Rtsne(perplexity = round(sqrt(NROW(triangular_3d_data))))  -->
-
-<!-- tsne_prism2 <- tSNE_fit2$Y |> -->
-<!--   as_tibble() |> -->
-<!--   mutate(ID = row_number()) -->
-
-<!-- colnames(tsne_prism2) <- c("tSNE1", "tSNE2", "ID") -->
-
-<!-- prism_scaled_obj_tsne2 <- gen_scaled_data( -->
-<!--   data = tsne_prism2) -->
-<!-- tsne_prism_scaled2 <- prism_scaled_obj_tsne2$scaled_nldr -->
-
-<!-- nldr_prism2 <- tsne_prism_scaled2 |>  -->
-<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
-<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
-<!-- ``` -->
-
-<!-- <!--tSNE with perplexity = 58--> -->
-<!-- ```{r} -->
-<!-- tSNE_fit3 <- triangular_3d_data |> -->
-<!--   select(-ID) |> -->
-<!--   Rtsne(perplexity = 58)  -->
-
-<!-- tsne_prism3 <- tSNE_fit3$Y |> -->
-<!--   as_tibble() |> -->
-<!--   mutate(ID = row_number()) -->
-
-<!-- colnames(tsne_prism3) <- c("tSNE1", "tSNE2", "ID") -->
-
-<!-- prism_scaled_obj_tsne3 <- gen_scaled_data( -->
-<!--   data = tsne_prism3) -->
-<!-- tsne_prism_scaled3 <- prism_scaled_obj_tsne3$scaled_nldr -->
-
-<!-- nldr_prism3 <- tsne_prism_scaled3 |>  -->
-<!--   ggplot(aes(x = tSNE1, y = tSNE2)) +  -->
-<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
-<!-- ``` -->
-
-<!-- <!--UMAP with n_neighbors = 15 and min_dist = 0.3--> -->
-<!-- ```{r} -->
-<!-- umap_prism1 <- triangular_3d_data |> -->
-<!--   select(-ID) |> -->
-<!--   umap(n_neighbors = 15, n_components =  2,  -->
-<!--        metric = "euclidean", min_dist = 0.3, init = "spca") -->
-
-<!-- umap_prism1 <- umap_prism1 |> -->
-<!--   as_tibble() -->
-
-<!-- names(umap_prism1) <- c("UMAP1", "UMAP2") -->
-
-<!-- umap_prism1 <- umap_prism1 |> -->
-<!--   mutate(ID = 1:NROW(umap_prism1)) -->
-
-<!-- prism_scaled_obj_umap1 <- gen_scaled_data( -->
-<!--   data = umap_prism1) -->
-<!-- umap_prism_scaled1 <- prism_scaled_obj_umap1$scaled_nldr -->
-
-<!-- nldr_prism4 <- umap_prism_scaled1 |>  -->
-<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
-<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
-<!-- ``` -->
-
-<!-- <!--UMAP with n_neighbors = 35 and min_dist = 0.5--> -->
-<!-- ```{r} -->
-<!-- umap_prism2 <- triangular_3d_data |> -->
-<!--   select(-ID) |> -->
-<!--   umap(n_neighbors = 35, n_components =  2,  -->
-<!--        metric = "euclidean", min_dist = 0.5, init = "spca") -->
-
-<!-- umap_prism2 <- umap_prism2 |> -->
-<!--   as_tibble() -->
-
-<!-- names(umap_prism2) <- c("UMAP1", "UMAP2") -->
-
-<!-- umap_prism2 <- umap_prism2 |> -->
-<!--   mutate(ID = 1:NROW(umap_prism2)) -->
-
-<!-- prism_scaled_obj_umap2 <- gen_scaled_data( -->
-<!--   data = umap_prism2) -->
-<!-- umap_prism_scaled2 <- prism_scaled_obj_umap2$scaled_nldr -->
-
-<!-- nldr_prism5 <- umap_prism_scaled2 |>  -->
-<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
-<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
-<!-- ``` -->
-
-<!-- <!--UMAP with n_neighbors = 69 and min_dist = 0.6--> -->
-<!-- ```{r} -->
-<!-- umap_prism3 <- triangular_3d_data |> -->
-<!--   select(-ID) |> -->
-<!--   umap(n_neighbors = 69, n_components =  2,  -->
-<!--        metric = "euclidean", min_dist = 0.6, init = "spca") -->
-
-<!-- umap_prism3 <- umap_prism3 |> -->
-<!--   as_tibble() -->
-
-<!-- names(umap_prism3) <- c("UMAP1", "UMAP2") -->
-
-<!-- umap_prism3 <- umap_prism3 |> -->
-<!--   mutate(ID = 1:NROW(umap_prism3)) -->
-
-<!-- prism_scaled_obj_umap3 <- gen_scaled_data( -->
-<!--   data = umap_prism3) -->
-<!-- umap_prism_scaled3 <- prism_scaled_obj_umap3$scaled_nldr -->
-
-<!-- nldr_prism6 <- umap_prism_scaled3 |>  -->
-<!--   ggplot(aes(x = UMAP1, y = UMAP2)) +  -->
-<!--   geom_point(alpha=0.5, colour="#000000", size = 0.5)  -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- nldr_prism1 + nldr_prism2 + nldr_prism3 +  -->
-<!--   nldr_prism4 + nldr_prism5 + nldr_prism6 + -->
-<!--   plot_layout(ncol=3) -->
-<!-- ``` -->
-
-<!-- <!--need to add details regarding which differences can I see with different layouts--> -->
-
-<!-- ## Deciding an appropriate fit -->
-
-<!-- <!--select based on looking at b_non_empty vs log(Error)--> -->
-<!-- <!--select a choice for b_non_empty--> -->
-
-<!-- <!--compute error with tSNE and different perplexity values--> -->
-<!-- ```{r} -->
-<!-- error_prism_tsne <- data.frame(matrix(nrow = 0, ncol = 0)) -->
-
-<!-- ## To initialize number of bins along the x-axis -->
-<!-- bin1_vec_prism <- 4:25  -->
-
-<!-- perplexity_vec <- c(6, 35, 58) -->
-
-<!-- for (perplexity_val in perplexity_vec) { -->
-
-<!--   tSNE_fit <- triangular_3d_data |> -->
-<!--     select(-ID) |> -->
-<!--     Rtsne(perplexity = perplexity_val) -->
-
-<!--   tsne_prism <- tSNE_fit$Y |> -->
-<!--     as_tibble() |> -->
-<!--     mutate(ID = row_number()) -->
-
-<!--   colnames(tsne_prism) <- c("tSNE1", "tSNE2", "ID") -->
-
-<!--   prism_scaled_obj <- gen_scaled_data( -->
-<!--     data = tsne_prism) -->
-<!--   tsne_prism_scaled <- prism_scaled_obj$scaled_nldr -->
-
-<!--   lim1 <- prism_scaled_obj$lim1 -->
-<!--   lim2 <- prism_scaled_obj$lim2 -->
-<!--   r2_prism <- diff(lim2)/diff(lim1)  -->
-
-<!--   for (xbins in bin1_vec_prism) { -->
-
-<!--     bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2 -->
-
-<!--     prism_model <- fit_highd_model( -->
-<!--       training_data = triangular_3d_data, -->
-<!--       emb_df = tsne_prism_scaled, -->
-<!--       bin1 = xbins, -->
-<!--       r2 = r2_prism, -->
-<!--       is_bin_centroid = TRUE, -->
-<!--       is_rm_lwd_hex = TRUE, -->
-<!--       col_start_highd = "x" -->
-<!--     ) -->
-
-<!--     df_bin_centroids_prism <- prism_model$df_bin_centroids -->
-<!--     df_bin_prism <- prism_model$df_bin -->
-
-<!--     ## Compute error -->
-<!--     error_df <- glance( -->
-<!--       df_bin_centroids = df_bin_centroids_prism, -->
-<!--       df_bin = df_bin_prism, -->
-<!--       training_data = triangular_3d_data, -->
-<!--       newdata = NULL, -->
-<!--       type_NLDR = "tSNE", -->
-<!--       col_start = "x") |> -->
-<!--       mutate(bin1 = xbins, -->
-<!--              bin2 = bin2, -->
-<!--              b = bin1 * bin2, -->
-<!--              b_non_empty = NROW(df_bin_centroids_prism), -->
-<!--              type = paste0("perplexity: ", perplexity_val)) -->
-
-<!--     error_prism_tsne <- bind_rows(error_prism_tsne, error_df) -->
-
-<!--   } -->
-
-<!-- } -->
-<!-- ``` -->
-
-<!-- <!--compute error with UMAP and different parameter values--> -->
-<!-- ```{r} -->
-<!-- error_prism_umap <- data.frame(matrix(nrow = 0, ncol = 0)) -->
-
-<!-- ## To initialize number of bins along the x-axis -->
-<!-- bin1_vec_prism <- 4:25  -->
-
-<!-- param_list <- list(param1 = c(15, 0.3), param2 = c(35, 0.5), param3 = c(69, 0.6)) -->
-
-<!-- for (i in 1:length(param_list)) { -->
-
-<!--   n_neighbors_val <- param_list[[i]][1] -->
-<!--   min_dist_val <- param_list[[i]][2] -->
-
-<!--   umap_prism <- triangular_3d_data |> -->
-<!--     select(-ID) |> -->
-<!--     umap(n_neighbors = n_neighbors_val, n_components =  2,  -->
-<!--          metric = "euclidean", min_dist = min_dist_val, init = "spca") -->
-
-<!--   umap_prism <- umap_prism |> -->
-<!--     as_tibble() -->
-
-<!--   names(umap_prism) <- c("UMAP1", "UMAP2") -->
-
-<!--   umap_prism <- umap_prism |> -->
-<!--     mutate(ID = 1:NROW(umap_prism)) -->
-
-<!--   prism_scaled_obj <- gen_scaled_data( -->
-<!--     data = umap_prism) -->
-<!--   umap_prism_scaled <- prism_scaled_obj$scaled_nldr -->
-
-<!--   lim1 <- prism_scaled_obj$lim1 -->
-<!--   lim2 <- prism_scaled_obj$lim2 -->
-<!--   r2_prism <- diff(lim2)/diff(lim1)  -->
-
-<!--   for (xbins in bin1_vec_prism) { -->
-
-<!--     bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_prism)$bin2 -->
-
-<!--     prism_model <- fit_highd_model( -->
-<!--       training_data = triangular_3d_data, -->
-<!--       emb_df = umap_prism_scaled, -->
-<!--       bin1 = xbins, -->
-<!--       r2 = r2_prism, -->
-<!--       is_bin_centroid = TRUE, -->
-<!--       is_rm_lwd_hex = TRUE, -->
-<!--       col_start_highd = "x" -->
-<!--     ) -->
-
-<!--     df_bin_centroids_prism <- prism_model$df_bin_centroids -->
-<!--     df_bin_prism <- prism_model$df_bin -->
-
-<!--     ## Compute error -->
-<!--     error_df <- glance( -->
-<!--       df_bin_centroids = df_bin_centroids_prism, -->
-<!--       df_bin = df_bin_prism, -->
-<!--       training_data = triangular_3d_data, -->
-<!--       newdata = NULL, -->
-<!--       type_NLDR = "UMAP", -->
-<!--       col_start = "x") |> -->
-<!--       mutate(bin1 = xbins, -->
-<!--              bin2 = bin2, -->
-<!--              b = bin1 * bin2, -->
-<!--              b_non_empty = NROW(df_bin_centroids_prism), -->
-<!--              type = paste0("n_neighbors: ", n_neighbors_val,  -->
-<!--                            " min_dist: ", min_dist_val)) -->
-
-<!--     error_prism_umap <- bind_rows(error_prism_umap, error_df) -->
-
-<!--   } -->
-
-<!-- } -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- error_prism <- bind_rows(error_prism_tsne, error_prism_umap)  -->
-<!-- error_prism <- error_prism |> -->
-<!--   filter(b <= 625) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- error_plot_prism <- ggplot(error_prism,  -->
-<!--                            aes(x = b_non_empty,  -->
-<!--                                y = log(Error),  -->
-<!--                                group = type,  -->
-<!--                                colour = type)) +  -->
-<!--   geom_point(size = 1) + -->
-<!--   geom_line() +  -->
-<!--   geom_vline(xintercept = 112, linetype="solid", -->
-<!--              color = "black", linewidth=0.8, alpha = 0.5) + -->
-<!--   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3', -->
-<!--                               '#ff7f00','#a6cee3','#a65628','#f781bf')) + -->
-<!--   ylab("log(Error)") + -->
-<!--   xlab("number of non-empty bins") + -->
-<!--   theme(axis.text.x = element_text(size = 5), -->
-<!--         axis.text.y = element_text(size = 5), -->
-<!--         axis.title.x = element_text(size = 5), -->
-<!--         axis.title.y = element_text(size = 5, angle = 90), -->
-<!--         legend.position = "bottom") -->
-
-<!-- error_plot_prism -->
-<!-- ``` -->
-
-<!-- <!--selected tSNE perplexity: 35--> -->
-
-<!-- ## Fit the appropriate model -->
-
-<!-- ### Construct model in $2\text{-}D$ -->
-
-<!-- ### Selecting parameter values for the model -->
-
-<!-- #### Choice of bins -->
-
-<!-- #### Removal of low-density bins -->
-
-<!-- #### Removing long edges -->
-
-<!-- ```{r} -->
-<!-- ## Compute hexbin parameters -->
-<!-- num_bins_x_prism <- 21 -->
-<!-- lim1 <- prism_scaled_obj_tsne2$lim1 -->
-<!-- lim2 <- prism_scaled_obj_tsne2$lim2 -->
-<!-- r2_prism <- diff(lim2)/diff(lim1)  -->
-
-<!-- prism_model <- fit_highd_model( -->
-<!--   training_data = triangular_3d_data, -->
-<!--   emb_df = tsne_prism_scaled2, -->
-<!--   bin1 = num_bins_x_prism, -->
-<!--   r2 = r2_prism, -->
-<!--   is_bin_centroid = TRUE, -->
-<!--   is_rm_lwd_hex = TRUE, -->
-<!--   col_start_highd = "x" -->
-<!-- ) -->
-
-<!-- df_bin_centroids_prism <- prism_model$df_bin_centroids -->
-<!-- df_bin_prism <- prism_model$df_bin -->
-
-<!-- ## Triangulate bin centroids -->
-<!-- tr1_object_prism <- tri_bin_centroids( -->
-<!--   df_bin_centroids_prism, x = "c_x", y = "c_y") -->
-<!-- tr_from_to_df_prism <- gen_edges( -->
-<!--   tri_object = tr1_object_prism) -->
-
-<!-- ## Compute 2D distances -->
-<!-- distance_prism <- cal_2d_dist( -->
-<!--   tr_coord_df = tr_from_to_df_prism, -->
-<!--   start_x = "x_from", -->
-<!--   start_y = "y_from", -->
-<!--   end_x = "x_to", -->
-<!--   end_y = "y_to", -->
-<!--   select_vars = c("from", "to", "distance")) -->
-
-<!-- ## To find the benchmark value -->
-<!-- benchmark_prism <- find_lg_benchmark( -->
-<!--   distance_edges = distance_prism, -->
-<!--   distance_col = "distance") -->
-
-<!-- trimesh_removed_prism <- vis_rmlg_mesh( -->
-<!--   distance_edges = distance_prism, -->
-<!--   benchmark_value = benchmark_prism, -->
-<!--   tr_coord_df = tr_from_to_df_prism, -->
-<!--   distance_col = "distance") #+ -->
-<!-- #xlim(sc_xlims) + ylim(sc_ylims) + -->
-<!-- #interior_annotation("a", sc_ltr_pos) -->
-
-<!-- trimesh_removed_prism -->
-<!-- ``` -->
-
-<!-- ### Lifting the model into $p\text{-}D$ -->
-
-<!-- ```{r} -->
-<!-- ## Hexagonal binning to have regular hexagons -->
-<!-- hb_obj_prism <- hex_binning( -->
-<!--   data = tsne_prism_scaled2, -->
-<!--   bin1 = num_bins_x_prism, -->
-<!--   r2 = r2_prism) -->
-
-<!-- tsne_data_with_hb_id <- hb_obj_prism$data_hb_id -->
-
-<!-- df_all_prism <- dplyr::bind_cols(triangular_3d_data |> dplyr::select(-ID), -->
-<!--                                  tsne_data_with_hb_id) -->
-
-<!-- ### Define type column -->
-<!-- df <- df_all_prism |> -->
-<!--   dplyr::select(tidyselect::starts_with("x")) |> -->
-<!--   dplyr::mutate(type = "data") ## original dataset -->
-
-<!-- df_b <- df_bin_prism |> -->
-<!--   dplyr::filter(hb_id %in% df_bin_centroids_prism$hexID) |> -->
-<!--   dplyr::mutate(type = "model") ## Data with summarized mean -->
-
-<!-- ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data -->
-<!-- df_b <- df_b[match(df_bin_centroids_prism$hexID, df_b$hb_id),] |> -->
-<!--   dplyr::select(-hb_id) -->
-
-<!-- df_exe <- dplyr::bind_rows(df_b, df) -->
-
-<!-- ## Set the maximum difference as the criteria -->
-<!-- distance_df_small_edges <- distance_prism |> -->
-<!--   dplyr::filter(distance < benchmark_prism) -->
-<!-- ## Since erase brushing is considerd. -->
-
-<!-- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], -->
-<!--                          lineFrom = distance_df_small_edges$from, -->
-<!--                          lineTo = distance_df_small_edges$to, -->
-<!--                          group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.8, NROW(df))), -->
-<!--                          levelColors = c("#6a3d9a", "#33a02c")) -->
-<!-- ``` -->
-
-<!-- What can I see in high-D? -->
-
-<!-- - Four clusters that are really close  -->
-
-<!-- What can I see in 2D layout? -->
-
-<!-- - Several clusters (more that what see in high-D) -->
-
-<!-- What can learn from the model? -->
-
-<!-- - There are four clusters, but not as the same cluster what data shows.
-In 2D, the two of actual clusters (high-D clusters) are really close.
-Also, one of the big high-D cluster is separated to two sub clusters.
--->
-
 # Application
@@
 #| label: tsne-num-bins-limb
-#| eval: false
 
@@
 #| label: prep-limb-tsne-author-model-proj
-#| eval: false
 
@@
 #| label: langevitour-limb-tsne-author-proj
-#| eval: false
 
@@
 #| label: tsne-best-num-bins-limb
-#| eval: false
 
@@
 #| label: prep-limb-tsne-best-model-proj
-#| eval: false
 
@@
 #| label: langevitour-limb-tsne-best-proj
-#| eval: false
 
@@
 ```{r}
-#| eval: false
 trimesh_removed_limb + trimesh_removed_limb_best +

### 2025-04-19 paper-quollr.tex
Corrected the code
Updated paper
New version

@@
 #> # A tibble: 72 x 2
-#>    hexID pts_list     
-#>    <int> <named list> 
-#>  1   136 <int [3,750]>
-#>  2    70 <int [3,750]>
-#>  3    45 <int [3,750]>
-#>  4    99 <int [3,750]>
-#>  5   114 <int [3,750]>
-#>  6    26 <int [3,750]>
-#>  7   129 <int [3,750]>
-#>  8    38 <int [3,750]>
-#>  9   111 <int [3,750]>
-#> 10    14 <int [3,750]>
+#>    hb_id pts_list  
+#>    <int> <list>    
+#>  1    13 <int [2]> 
+#>  2    14 <int [81]>
+#>  3    15 <int [56]>
+#>  4    16 <int [55]>
+#>  5    18 <int [2]> 
+#>  6    26 <int [70]>
+#>  7    27 <int [78]>
+#>  8    28 <int [86]>
+#>  9    29 <int [5]> 
+#> 10    30 <int [54]>
 #> # i 62 more rows
@@
 
---\textgreater{}
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
---\textgreater{}
-
 \section{Application}\label{application}
@@
 
-\begin{figure}[H]
+\begin{figure}
 \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-limb-mse-1} \end{figure}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-20-1}
 

### 2025-04-20 paper-quollr.Rmd
Updated paper
Updated my name

@@
 author:  
-  - name: Jayani P.G. Lakshika
+  - name: Jayani P. Gamage
     affiliation: Monash University
@@
 #| label: langevitour-limb-tsne-author-proj
+#| eval: false
 
@@
 #| label: langevitour-limb-tsne-best-proj
+#| eval: false
 

### 2025-04-20 paper-quollr.tex
Updated my name

@@
 
-\author{by Jayani P.G. Lakshika, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}
+\author{by Jayani P. Gamage, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}
 
@@
 \address{%
-Jayani P.G. Lakshika\\
+Jayani P. Gamage\\
 Monash University\\%

### 2025-05-07 paper-quollr.Rmd
Deleted unnecessary data files
Added S-curve with different UMAP parameters

@@
 The software is available from the Comprehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).
+
+Add the image generated with S-curve data
+
+```{r}
+#| label: scurve-diff-umap-layouts
+
+scurve_umap1 <- read_rds("data/scurve/s_curve_noise_umap.rds")
+
+nldr_scaled_obj <- gen_scaled_data(
+  data = scurve_umap1)
+umap_scaled1 <- nldr_scaled_obj$scaled_nldr
+
+nldr1 <- umap_scaled1 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  theme(aspect.ratio = 1)
+
+scurve_umap2 <- read_rds("data/scurve/s_curve_noise_umap2.rds")
+
+nldr_scaled_obj2 <- gen_scaled_data(
+  data = scurve_umap2)
+umap_scaled2 <- nldr_scaled_obj2$scaled_nldr
+
+nldr2 <- umap_scaled2 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  theme(aspect.ratio = 1)
+
+scurve_umap3 <- read_rds("data/scurve/s_curve_noise_umap3.rds")
+
+nldr_scaled_obj3 <- gen_scaled_data(
+  data = scurve_umap3)
+umap_scaled3 <- nldr_scaled_obj3$scaled_nldr
+
+nldr3 <- umap_scaled3 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  theme(aspect.ratio = 1)
+
+scurve_umap4 <- read_rds("data/scurve/s_curve_noise_umap4.rds")
+
+nldr_scaled_obj4 <- gen_scaled_data(
+  data = scurve_umap4)
+umap_scaled4 <- nldr_scaled_obj4$scaled_nldr
+
+nldr4 <- umap_scaled4 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  theme(aspect.ratio = 1)
+
+scurve_umap5 <- read_rds("data/scurve/s_curve_noise_umap5.rds")
+
+nldr_scaled_obj5 <- gen_scaled_data(
+  data = scurve_umap5)
+umap_scaled5 <- nldr_scaled_obj5$scaled_nldr
+
+nldr5 <- umap_scaled5 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#636363') +
+  theme(aspect.ratio = 1)
+
+scurve_umap6 <- read_rds("data/scurve/s_curve_noise_umap6.rds")
+
+nldr_scaled_obj6 <- gen_scaled_data(
+  data = scurve_umap6)
+umap_scaled6 <- nldr_scaled_obj6$scaled_nldr
+
+nldr6 <- umap_scaled6 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  theme(aspect.ratio = 1)
+```
+
+```{r}
+nldr1 + nldr2 + nldr3 + 
+  nldr4 + nldr5 + nldr6 +
+  plot_layout(ncol = 3)
+```
 

### 2025-05-07 paper-quollr.tex
Added S-curve with different UMAP parameters

@@
 The software is available from the Comprehensive R Archive Network (CARN) at \url{https://CRAN.R-project.org/package=quollr}.
+
+Add the image generated with S-curve data
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-1-1}
 
@@
 #> $df_bin
-#> # A tibble: 72 x 8
+#> # A tibble: 86 x 8
 #>    hb_id       x1    x2    x3        x4        x5        x6        x7
 #>    <int>    <dbl> <dbl> <dbl>     <dbl>     <dbl>     <dbl>     <dbl>
-#>  1    13  0.996  1.77  1.07 -0.0117    0.00979  -0.0725   -0.00321  
-#>  2    14  0.829  1.79  1.50 -0.000753 -0.00194  -0.00936  -0.000454 
-#>  3    15  0.307  1.78  1.94 -0.00384   0.00212  -0.00117  -0.000159 
-#>  4    16 -0.175  1.74  1.98  0.00143   0.00320   0.00450  -0.0000763
-#>  5    18 -0.493  1.78  1.87 -0.0100    0.000680  0.0799   -0.000835 
-#>  6    26  0.962  1.22  1.22  0.00107  -0.00260  -0.000948  0.000234 
-#>  7    27  0.631  1.28  1.75  0.000266  0.00174  -0.00158   0.00114  
-#>  8    28  0.139  1.20  1.97 -0.00183   0.000279  0.00177  -0.000666 
-#>  9    29 -0.187  1.00  1.98 -0.00228   0.00461  -0.0506   -0.00472  
-#> 10    30 -0.562  1.23  1.82 -0.000563  0.00166  -0.00923  -0.000111 
-#> # i 62 more rows
+#>  1    15  0.968   0.198  1.21  0.000518  0.00161   0.00240  -0.000547
+#>  2    16  0.730   0.142  1.67 -0.000689  0.000495  0.00168   0.000794
+#>  3    17  0.323   0.158  1.94 -0.00265  -0.00523  -0.00942  -0.00115 
+#>  4    18 -0.203   0.154  1.98 -0.00269  -0.00107  -0.0382   -0.00167 
+#>  5    19 -0.628   0.114  1.78 -0.000131 -0.00323   0.0347   -0.00288 
+#>  6    27  0.987   0.579  1.14 -0.00124  -0.00121   0.0199   -0.00363 
+#>  7    28  0.847   0.537  1.51  0.000374 -0.00215  -0.00901  -0.000548
+#>  8    29  0.469   0.476  1.87  0.00117  -0.000439 -0.0191    0.000251
+#>  9    30 -0.00747 0.491  1.99 -0.000570  0.00114  -0.000815  0.000786
+#> 10    31 -0.525   0.449  1.84  0.00141  -0.000575 -0.0112    0.000714
+#> # i 76 more rows
 #> 
 #> $df_bin_centroids
-#> # A tibble: 72 x 6
+#> # A tibble: 86 x 6
 #>    hexID   c_x      c_y bin_counts std_counts drop_empty
 #>    <int> <dbl>    <dbl>      <int>      <dbl> <lgl>     
-#>  1    13 -0.0455  0.000514          2     0.0194 FALSE     
-#>  2    14  0.0634  0.000514         81     0.786  FALSE     
-#>  3    15  0.172   0.000514         56     0.544  FALSE     
-#>  4    16  0.281   0.000514         55     0.534  FALSE     
-#>  5    18  0.499   0.000514          2     0.0194 FALSE     
-#>  6    26  0.00894 0.0949           70     0.680  FALSE     
-#>  7    27  0.118   0.0949           78     0.757  FALSE     
-#>  8    28  0.227   0.0949           86     0.835  FALSE     
-#>  9    29  0.336   0.0949            5     0.0485 FALSE     
-#> 10    30  0.445   0.0949           54     0.524  FALSE     
-#> # i 62 more rows
+#>  1    15 0.158 0.000959         43     0.518  FALSE     
+#>  2    16 0.262 0.000959         35     0.422  FALSE     
+#>  3    17 0.365 0.000959         21     0.253  FALSE     
+#>  4    18 0.468 0.000959         11     0.133  FALSE     
+#>  5    19 0.571 0.000959          8     0.0964 FALSE     
+#>  6    27 0.107 0.0904           10     0.120  FALSE     
+#>  7    28 0.210 0.0904           56     0.675  FALSE     
+#>  8    29 0.313 0.0904           43     0.518  FALSE     
+#>  9    30 0.416 0.0904           54     0.651  FALSE     
+#> 10    31 0.520 0.0904           42     0.506  FALSE     
+#> # i 76 more rows
 \end{verbatim}
@@
 #> $a1
-#> [1] 0.1089367
+#> [1] 0.1032943
 #> 
 #> $a2
-#> [1] 0.09434198
+#> [1] 0.08945548
 \end{verbatim}
@@
 #> $a1
-#> [1] 0.1089367
+#> [1] 0.1032943
 #> 
 #> $a2
-#> [1] 0.09434198
+#> [1] 0.08945548
 #> 
@@
 #> $start_point
-#> [1] -0.10000000 -0.09382762
+#> [1] -0.10000000 -0.08849688
 #> 
@@
 #>    <int>    <dbl>   <dbl>
-#>  1     1 -0.1     -0.0938
-#>  2     2  0.00894 -0.0938
-#>  3     3  0.118   -0.0938
-#>  4     4  0.227   -0.0938
-#>  5     5  0.336   -0.0938
-#>  6     6  0.445   -0.0938
-#>  7     7  0.554   -0.0938
-#>  8     8  0.663   -0.0938
-#>  9     9  0.771   -0.0938
-#> 10    10  0.880   -0.0938
+#>  1     1 -0.1     -0.0885
+#>  2     2  0.00329 -0.0885
+#>  3     3  0.107   -0.0885
+#>  4     4  0.210   -0.0885
+#>  5     5  0.313   -0.0885
+#>  6     6  0.416   -0.0885
+#>  7     7  0.520   -0.0885
+#>  8     8  0.623   -0.0885
+#>  9     9  0.726   -0.0885
+#> 10    10  0.830   -0.0885
 #> # i 146 more rows
@@
 #>          <int>    <dbl>   <dbl>
-#>  1           1 -0.1     -0.0309
-#>  2           1 -0.154   -0.0624
-#>  3           1 -0.154   -0.125 
-#>  4           1 -0.1     -0.157 
-#>  5           1 -0.0455  -0.125 
-#>  6           1 -0.0455  -0.0624
-#>  7           2  0.00894 -0.0309
-#>  8           2 -0.0455  -0.0624
-#>  9           2 -0.0455  -0.125 
-#> 10           2  0.00894 -0.157 
+#>  1           1 -0.1     -0.0289
+#>  2           1 -0.152   -0.0587
+#>  3           1 -0.152   -0.118 
+#>  4           1 -0.1     -0.148 
+#>  5           1 -0.0484  -0.118 
+#>  6           1 -0.0484  -0.0587
+#>  7           2  0.00329 -0.0289
+#>  8           2 -0.0484  -0.0587
+#>  9           2 -0.0484  -0.118 
+#> 10           2  0.00329 -0.148 
 #> # i 926 more rows
@@
 #>    <dbl>  <dbl> <int> <int>
-#>  1 0.276  0.915      1   136
-#>  2 0.927  0.347      2    70
-#>  3 0.810  0.242      3    45
-#>  4 0.137  0.657      5    99
-#>  5 0.476  0.799      6   114
-#>  6 0.0485 0.0657     7    26
-#>  7 0.737  0.851      9   129
-#>  8 0.833  0.224     10    45
-#>  9 0.909  0.371     11    70
-#> 10 0.0890 0.194     12    38
+#>  1 0.270 0.839      1   125
+#>  2 0.788 0.466      2    82
+#>  3 0.771 0.319      3    69
+#>  4 0.306 0.542      5    88
+#>  5 0.549 0.806      6   127
+#>  6 0.166 0.259      7    52
+#>  7 0.672 0.596      9   104
+#>  8 0.735 0.354     10    69
+#>  9 0.839 0.467     11    82
+#> 10 0.240 0.0903    12    28
 #> # i 3,740 more rows
@@
 #> $std_cts
-#> # A tibble: 72 x 3
+#> # A tibble: 86 x 3
 #>    hb_id     n std_counts
 #>    <int> <int>      <dbl>
-#>  1    13     2     0.0194
-#>  2    14    81     0.786 
-#>  3    15    56     0.544 
-#>  4    16    55     0.534 
-#>  5    18     2     0.0194
-#>  6    26    70     0.680 
-#>  7    27    78     0.757 
-#>  8    28    86     0.835 
-#>  9    29     5     0.0485
-#> 10    30    54     0.524 
-#> # i 62 more rows
+#>  1    15    43     0.518 
+#>  2    16    35     0.422 
+#>  3    17    21     0.253 
+#>  4    18    11     0.133 
+#>  5    19     8     0.0964
+#>  6    27    10     0.120 
+#>  7    28    56     0.675 
+#>  8    29    43     0.518 
+#>  9    30    54     0.651 
+#> 10    31    42     0.506 
+#> # i 76 more rows
 #> 
@@
 #> $non_bins
-#> [1] 72
+#> [1] 86
 #> 
 #> $pts_bins
-#> # A tibble: 72 x 2
+#> # A tibble: 86 x 2
 #>    hb_id pts_list  
 #>    <int> <list>    
-#>  1    13 <int [2]> 
-#>  2    14 <int [81]>
-#>  3    15 <int [56]>
-#>  4    16 <int [55]>
-#>  5    18 <int [2]> 
-#>  6    26 <int [70]>
-#>  7    27 <int [78]>
-#>  8    28 <int [86]>
-#>  9    29 <int [5]> 
-#> 10    30 <int [54]>
-#> # i 62 more rows
+#>  1    15 <int [43]>
+#>  2    16 <int [35]>
+#>  3    17 <int [21]>
+#>  4    18 <int [11]>
+#>  5    19 <int [8]> 
+#>  6    27 <int [10]>
+#>  7    28 <int [56]>
+#>  8    29 <int [43]>
+#>  9    30 <int [54]>
+#> 10    31 <int [42]>
+#> # i 76 more rows
 #> 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-20-1}
-
 \section{Discussion}\label{discussion}

### 2025-05-07 paper-quollr.Rmd
Added layouts for example
Started updating according to the package updates

@@
 
+## Installation
+
 The package can be installed from CRAN:
@@
 
-The development version can be installed from GitHub:
+and the development version can be installed from GitHub:
 
@@
 ```
+
+## Web site
+
+More documentation of the package can be found at the web site [https://jayanilakshika.github.io/quollr/](https://jayanilakshika.github.io/quollr/).
 
@@
 
+Add a table of data sets that quollr have
+
 ### Scaling the data
@@
 The algorithm begins by scaling NLDR data to a standard scale using the `gen_scaled_data()` function.
-This function standardizes the data and provides the original limits of embeddings and the aspect ratio.
+This function standardizes the data with first embedding scale to 0 and 1, The second embedding scales to 0 and the maximum of the second embedding.
+The resulting output provides the scaled NLDR data and the original limits of embeddings.
 
 ```{r, echo=TRUE}
-scaled_nldr_obj_scurve <- gen_scaled_data(data = s_curve_noise_umap)
+scurve_nldr_obj <- gen_scaled_data(nldr_data = scurve_umap)
 
-s_curve_noise_umap_scaled <- scaled_nldr_obj_scurve$scaled_nldr
+scurve_nldr_obj
+```
 
-lim1 <- scaled_nldr_obj_scurve$lim1
-lim2 <- scaled_nldr_obj_scurve$lim2
+```{r, echo=TRUE}
+scurve_umap_scaled <- scurve_nldr_obj$scaled_nldr
+
+lim1 <- scurve_nldr_obj$lim1
+lim2 <- scurve_nldr_obj$lim2
 r2 <- diff(lim2)/diff(lim1) 
 ```
+
+We can compute the `r2` which is the ratio of the ranges of the original embedding components.
 

### 2025-05-07 paper-quollr.Rmd
Updated sections

@@
 
+## Data sets
+
+Add a table of data sets that quollr have
+
 ## Usage
@@
 
-The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `s_curve_noise_training` and `s_curve_noise_umap`.
+The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `scurve` and `scurve_umap`.
 
-`s_curve_noise_training` is a $3\text{-}D$ S-curve data set with additional four noise dimensions which is used to train the model.
-`s_curve_noise_umap` is the UMAP $2\text{-}D$ embedding for `s_curve_noise_training` data set.
-Each data set contains a unique ID column.
+`scurve` is a $3\text{-}D$ S-curve data set with additional four noise dimensions which is used to build the model.
+`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data set with `n_neighbors` is $15$ and `min_dist` is $0.1$.
+Each data set contains a unique ID column that mapping `scurve` and `scurve_umap`.
 
-Add a table of data sets that quollr have
+### Scaling the Data
 
-### Scaling the data
-
-The algorithm begins by scaling NLDR data to a standard scale using the `gen_scaled_data()` function.
-This function standardizes the data with first embedding scale to 0 and 1, The second embedding scales to 0 and the maximum of the second embedding.
-The resulting output provides the scaled NLDR data and the original limits of embeddings.
+The algorithm starts by scaling the NLDR data to a standard range using the `gen_scaled_data()` function.
+This function standardizes the data so that the first embedding ranges from $0$ to $1$, while the second embedding scales from $0$ to the maximum value of the second embedding.
+The output includes the scaled NLDR data along with the original limits of the embeddings.
 
@@
 
-We can compute the `r2` which is the ratio of the ranges of the original embedding components.
+We can calculate the `r2`, which represents the ratio of the ranges of the original embedding components.
 
@@
 
-## Constructing the $2\text{-}D$ model
+## Constructing the $2\text{-}D$ Model
 
-Constructing the $2\text{-}D$ model mainly contains (i) binning data, (ii) obtaining bin centroids, and (iii) indicating neighbors by line segments connecting centroids.
+Constructing the $2\text{-}D$ model primarily involves (i) binning data, (ii) obtaining bin centroids, and (iii) connecting centroids with line segments to indicate neighbors.
 

### 2025-05-07 paper-quollr.Rmd
Added latex code for the topic

@@
 ---
-title: "quollr: An R Package for Visualizing $2-D$ Models from Non-linear Dimension Reductions in High Dimensional Space"
+title: "quollr: An R Package for Visualizing $2-\\text{D}$ Models from Non-linear Dimension Reductions in High Dimensional Space"
 abstract: >

### 2025-05-07 paper-quollr.Rmd
Updated abstract

@@
 The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data.
-The `triangular_3d_data` data from the \CRANpkg{cardinalR} package is used to illustrate the algorithm and its application within the package.
+The `scurve` data from the package is used to illustrate the algorithm.
+scRNA-seq data for Limb muscles of mice are used to discuss the usability of the package.
 draft: true

### 2025-05-08 paper-quollr.Rmd
Updated Usage intro

@@
 
-`scurve` is a $3\text{-}D$ S-curve data set with additional four noise dimensions which is used to build the model.
-`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data set with `n_neighbors` is $15$ and `min_dist` is $0.1$.
-Each data set contains a unique ID column that mapping `scurve` and `scurve_umap`.
+`scurve` is a $7\text{-}D$ simulated dataset.
+It is constructed by simulating $5000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
+The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
+`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $15$ and `min_dist` is $0.1$.
+Each data set contains a unique ID column that maps `scurve` and `scurve_umap`.
 

### 2025-05-08 paper-quollr.Rmd
Added main function intro

@@
 
+### Main function
+
+The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
+
+<!-- If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
+For example, if the user wishes to use already binning data, the `extract_hexbin_centroids()` function can be used directly.
+-->
+
+This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`bin1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`benchmark_highdens`).
+The function returns an object that includes the hexagonal object, the fitted model in both $2\text{-}D$, and $p\text{-}D$, and triangular mesh.
+
+```{r, echo=TRUE}
+fit_highd_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  bin1 = 15, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+```
+
 ### Scaling the Data
@@
 We can calculate the `r2`, which represents the ratio of the ranges of the original embedding components.
-
-The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
-When constructing the $2\text{-}D$ model, the user can choose either to fit the $2\text{-}D$ model with hexagonal bin centroids or bin means using `is_bin_centroid` argument.
-
-If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
-For example, if the user wishes to use already binning data, the `extract_hexbin_centroids()` function can be used directly.
-
-The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if `is_rm_lwd_hex = TRUE`, benchmark value to remove low density hexagons are parameters that will be determined within `fit_highd_model()`, if they are not provided.
-They are created as they are needed throughout the following example.
-The function `fit_highd_model()` provides the fitted model in $2\text{-}D$, and $p\text{-}D$.
-
-```{r, echo=TRUE}
-fit_highd_model(
-  highd_data = s_curve_noise_training,
-  nldr_data = s_curve_noise_umap_scaled, 
-  bin1 = 12, 
-  r2 = r2,
-  q = 0.1,
-  is_bin_centroid = TRUE
-  )
-```
 

### 2025-05-08 paper-quollr.Rmd
Updated the flow of generating 2D model

@@
 
-## Usage
+<!-- ## Usage -->
 
@@
 
-### Main function
+## Main function
 
@@
 
+## Constructing the $2\text{-}D$ Model
+
+Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) Remove low-density hexagons.
+
 ### Scaling the Data
@@
 ```
-
-```{r, echo=TRUE}
-scurve_umap_scaled <- scurve_nldr_obj$scaled_nldr
-
-lim1 <- scurve_nldr_obj$lim1
-lim2 <- scurve_nldr_obj$lim2
-r2 <- diff(lim2)/diff(lim1) 
-```
-
-We can calculate the `r2`, which represents the ratio of the ranges of the original embedding components.
-
-## Constructing the $2\text{-}D$ Model
-
-Constructing the $2\text{-}D$ model primarily involves (i) binning data, (ii) obtaining bin centroids, and (iii) connecting centroids with line segments to indicate neighbors.
 

### 2025-05-08 paper-quollr.Rmd
Added updated version

@@
 nldr_scaled_obj <- gen_scaled_data(
-  data = scurve_umap1)
+  nldr_data = scurve_umap)
 umap_scaled1 <- nldr_scaled_obj$scaled_nldr
@@
 nldr_scaled_obj2 <- gen_scaled_data(
-  data = scurve_umap2)
+  nldr_data = scurve_umap2)
 umap_scaled2 <- nldr_scaled_obj2$scaled_nldr
@@
 nldr_scaled_obj3 <- gen_scaled_data(
-  data = scurve_umap3)
+  nldr_data = scurve_umap3)
 umap_scaled3 <- nldr_scaled_obj3$scaled_nldr
@@
 nldr_scaled_obj4 <- gen_scaled_data(
-  data = scurve_umap4)
+  nldr_data = scurve_umap4)
 umap_scaled4 <- nldr_scaled_obj4$scaled_nldr
@@
 nldr_scaled_obj5 <- gen_scaled_data(
-  data = scurve_umap5)
+  nldr_data = scurve_umap5)
 umap_scaled5 <- nldr_scaled_obj5$scaled_nldr
@@
 nldr_scaled_obj6 <- gen_scaled_data(
-  data = scurve_umap6)
+  nldr_data = scurve_umap6)
 umap_scaled6 <- nldr_scaled_obj6$scaled_nldr
@@
 The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
-`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $15$ and `min_dist` is $0.1$.
+`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $46$ and `min_dist` is $0.9$.
 Each data set contains a unique ID column that maps `scurve` and `scurve_umap`.
@@
 calc_bins_y(
-  bin1 = 12, 
-  r2 = r2,
-  q = 0.1
-  )
+  nldr_obj = scurve_nldr_obj, 
+  bin1 = 4, 
+  q = 0.1)
 ```
@@
 hex_binning(
-  data = s_curve_noise_umap_scaled, 
-  bin1 = 12, 
-  r2 = r2,
-  q = 0.1
-  )
+  nldr_obj = scurve_nldr_obj, 
+  bin1 = 4, 
+  q = 0.1)
 ```
@@
 <!--add hexbin notation image-->
-
-### Remove low-density hexagons
-
-In certain scenarios, hexagonal bins may contain a few number of points.
-To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
-The `find_low_dens_hex()` function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
-
-```{r, echo=TRUE, eval=FALSE}
-find_low_dens_hex(
-  df_bin_centroids_all = df_bin_centroids, 
-  bin1 = 12, 
-  df_bin_centroids_low = df_bin_centroids_low
-  )
-
-```
 
@@
 
+### Remove low-density hexagons
+
+In certain scenarios, hexagonal bins may contain a few number of points.
+To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
+The `find_low_dens_hex()` function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
+Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
+
 ```{r, echo=TRUE, eval=FALSE}
-find_lg_benchmark(
-  distance_edges = distance_df, 
-  distance_col = "distance"
+find_low_dens_hex(
+  centroids_data, 
+  bin1 = 15, 
+  benchmark_mean_dens = 0.05
 )
+
 ```
@@
 
-The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each bin to represent bin centroids.
-This transformation is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data and their corresponding hexagonal bin IDs as inputs.
+The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each hexgaonal bin to represent bin centroids.
+This transformation is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).
 
 ```{r, echo=TRUE, eval=FALSE}
-umap_data_with_hb_id <- hb_obj$data_hb_id
-
-df_all <- bind_cols(
-  s_curve_noise_training |> dplyr::select(-ID), 
-  umap_data_with_hb_id
-  )
-
 df_bin <- avg_highd_data(
-  data = df_all
+  highd_data = scurve, 
+  scaled_nldr_hexid
 )
@@
 predict_emb(
-  highd_data = s_curve_noise_training, 
+  highd_data = scurve, 
   model_2d = df_bin_centroids, 
@@
 glance(
-  highd_data = s_curve_noise_training, 
+  highd_data = scurve, 
   model_2d = df_bin_centroids, 
@@
 augment(
-  highd_data = s_curve_noise_training, 
+  highd_data = scurve, 
   model_2d = df_bin_centroids, 
@@
 ```{r, echo=TRUE, eval=FALSE}
-vis_lg_mesh(
-  distance_edges = distance_df, 
-  benchmark_value = 0.75, 
-  tr_coord_df = tr_from_to_df, 
-  distance_col = "distance"
-  )
-```
-
-```{r, echo=TRUE, eval=FALSE}
-vis_rmlg_mesh(
+vis_mesh(
   distance_edges = distance_df, 
@@
 ```{r, echo=TRUE, eval=FALSE}
+df_exe <- comb_data_mode(
+  highd_data = scurve, 
+  model_highd = scurve_model_obj$model_highd, 
+  model_2d = scurve_model_obj$model_2d
+  )
+```
+
+```{r, echo=TRUE, eval=FALSE}
 show_langevitour(
-  df = df_all, 
-  df_b = df_bin, 
-  df_b_with_center_data = df_bin_centroids, 
-  benchmark_value = 0.75, 
-  distance = distance_df, 
-  distance_col = "distance", 
-  use_default_benchmark_val = FALSE, 
-  col_start = "x"
+  point_df, 
+  edge_df
   )

### 2025-05-08 paper-quollr.tex
Added layouts for example
Added updated version

@@
 % !TeX root = RJwrapper.tex
-\title{quollr: An R Package for Visualizing \(2-D\) Models from Non-linear Dimension Reductions in High Dimensional Space}
+\title{quollr: An R Package for Visualizing \(2-\text{D}\) Models from Non-linear Dimension Reductions in High Dimensional Space}
 
@@
 The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data.
-The \texttt{triangular\_3d\_data} data from the \CRANpkg{cardinalR} package is used to illustrate the algorithm and its application within the package.
+The \texttt{scurve} data from the package is used to illustrate the algorithm.
+scRNA-seq data for Limb muscles of mice are used to discuss the usability of the package.
 }
@@
 
+\subsection{Installation}\label{installation}
+
 The package can be installed from CRAN:
@@
 \end{verbatim}
+
+\subsection{Web site}\label{web-site}
+
+More documentation of the package can be found at the web site \url{https://jayanilakshika.github.io/quollr/}.
 
@@
 
-\subsection{Usage}\label{usage}
+\subsection{Data sets}\label{data-sets}
 
-The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded.
-We also want to load the built-in data sets \texttt{s\_curve\_noise\_training} and \texttt{s\_curve\_noise\_umap}.
+Add a table of data sets that quollr have
 
-\texttt{s\_curve\_noise\_training} is a \(3\text{-}D\) S-curve data set with additional four noise dimensions which is used to train the model.
-\texttt{s\_curve\_noise\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{s\_curve\_noise\_training} data set.
-Each data set contains a unique ID column.
+The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. We also want to load the built-in data sets \texttt{scurve} and \texttt{scurve\_umap}.
 
-\subsubsection{Scaling the data}\label{scaling-the-data}
+\texttt{scurve} is a \(7\text{-}D\) simulated dataset.
+It is constructed by simulating \(5000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\).
+The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error, with small variance.
+\texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors} is \(46\) and \texttt{min\_dist} is \(0.9\).
+Each data set contains a unique ID column that maps \texttt{scurve} and \texttt{scurve\_umap}.
 
-The algorithm begins by scaling NLDR data to a standard scale using the \texttt{gen\_scaled\_data()} function.
-This function standardizes the data and provides the original limits of embeddings and the aspect ratio.
-
-\begin{verbatim}
-scaled_nldr_obj_scurve <- gen_scaled_data(data = s_curve_noise_umap)
-
-s_curve_noise_umap_scaled <- scaled_nldr_obj_scurve$scaled_nldr
-
-lim1 <- scaled_nldr_obj_scurve$lim1
-lim2 <- scaled_nldr_obj_scurve$lim2
-r2 <- diff(lim2)/diff(lim1) 
-\end{verbatim}
+\subsection{Main function}\label{main-function}
 
 The mains steps for the algorithm can be executed by the main function \texttt{fit\_highd\_model()}, or can be run separately for more flexibility.
-When constructing the \(2\text{-}D\) model, the user can choose either to fit the \(2\text{-}D\) model with hexagonal bin centroids or bin means using \texttt{is\_bin\_centroid} argument.
 
-If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
-For example, if the user wishes to use already binning data, the \texttt{extract\_hexbin\_centroids()} function can be used directly.
-
-The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if \texttt{is\_rm\_lwd\_hex\ =\ TRUE}, benchmark value to remove low density hexagons are parameters that will be determined within \texttt{fit\_highd\_model()}, if they are not provided.
-They are created as they are needed throughout the following example.
-The function \texttt{fit\_highd\_model()} provides the fitted model in \(2\text{-}D\), and \(p\text{-}D\).
+This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{bin1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{benchmark\_highdens}).
+The function returns an object that includes the hexagonal object, the fitted model in both \(2\text{-}D\), and \(p\text{-}D\), and triangular mesh.
 
@@
 fit_highd_model(
-  highd_data = s_curve_noise_training,
-  nldr_data = s_curve_noise_umap_scaled, 
-  bin1 = 12, 
-  r2 = r2,
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  bin1 = 15, 
   q = 0.1, 
-  is_bin_centroid = TRUE
-  )
+  benchmark_highdens = 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> $df_bin
-#> # A tibble: 86 x 8
-#>    hb_id       x1    x2    x3        x4        x5        x6        x7
+#> $hb_obj
+#> $a1
+#> [1] 0.08326136
+#> 
+#> $a2
+#> [1] 0.07210645
+#> 
+#> $bins
+#> [1] 15 16
+#> 
+#> $start_point
+#> [1] -0.10000000 -0.08923607
+#> 
+#> $centroids
+#> # A tibble: 240 x 3
+#>    hexID     c_x     c_y
+#>    <int>   <dbl>   <dbl>
+#>  1     1 -0.1    -0.0892
+#>  2     2 -0.0167 -0.0892
+#>  3     3  0.0665 -0.0892
+#>  4     4  0.150  -0.0892
+#>  5     5  0.233  -0.0892
+#>  6     6  0.316  -0.0892
+#>  7     7  0.400  -0.0892
+#>  8     8  0.483  -0.0892
+#>  9     9  0.566  -0.0892
+#> 10    10  0.649  -0.0892
+#> # i 230 more rows
+#> 
+#> $hex_poly
+#> # A tibble: 1,440 x 3
+#>    hex_poly_id       x       y
+#>          <int>   <dbl>   <dbl>
+#>  1           1 -0.1    -0.0412
+#>  2           1 -0.142  -0.0652
+#>  3           1 -0.142  -0.113 
+#>  4           1 -0.1    -0.137 
+#>  5           1 -0.0584 -0.113 
+#>  6           1 -0.0584 -0.0652
+#>  7           2 -0.0167 -0.0412
+#>  8           2 -0.0584 -0.0652
+#>  9           2 -0.0584 -0.113 
+#> 10           2 -0.0167 -0.137 
+#> # i 1,430 more rows
+#> 
+#> $data_hb_id
+#> # A tibble: 5,000 x 4
+#>     emb1  emb2    ID hexID
+#>    <dbl> <dbl> <int> <int>
+#>  1 0.707 0.839     1   205
+#>  2 0.231 0.401     2   109
+#>  3 0.232 0.215     3    65
+#>  4 0.790 0.564     4   146
+#>  5 0.761 0.551     5   146
+#>  6 0.445 0.721     6   172
+#>  7 0.900 0.137     7    58
+#>  8 0.247 0.392     8   110
+#>  9 0.325 0.542     9   141
+#> 10 0.278 0.231    10    80
+#> # i 4,990 more rows
+#> 
+#> $std_cts
+#> # A tibble: 142 x 3
+#>    hexID bin_counts std_counts
+#>    <int>      <int>      <dbl>
+#>  1    21         12     0.16  
+#>  2    22         17     0.227 
+#>  3    23         22     0.293 
+#>  4    24         10     0.133 
+#>  5    25          7     0.0933
+#>  6    26         12     0.16  
+#>  7    27          1     0.0133
+#>  8    36         42     0.56  
+#>  9    37         42     0.56  
+#> 10    38         44     0.587 
+#> # i 132 more rows
+#> 
+#> $tot_bins
+#> [1] 240
+#> 
+#> $non_bins
+#> [1] 142
+#> 
+#> $pts_bins
+#> # A tibble: 142 x 2
+#>    hexID pts_list  
+#>    <int> <list>    
+#>  1    21 <int [12]>
+#>  2    22 <int [17]>
+#>  3    23 <int [22]>
+#>  4    24 <int [10]>
+#>  5    25 <int [7]> 
+#>  6    26 <int [12]>
+#>  7    27 <int [1]> 
+#>  8    36 <int [42]>
+#>  9    37 <int [42]>
+#> 10    38 <int [44]>
+#> # i 132 more rows
+#> 
+#> attr(,"class")
+#> [1] "hex_bin_obj"
+#> 
+#> $model_highd
+#> # A tibble: 130 x 8
+#>    hexID      x1    x2    x3         x4        x5       x6        x7
 #>    <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
-#>  1    15  0.968   0.198  1.21  0.000518  0.00161   0.00240  -0.000547
-#>  2    16  0.730   0.142  1.67 -0.000689  0.000495  0.00168   0.000794
-#>  3    17  0.323   0.158  1.94 -0.00265  -0.00523  -0.00942  -0.00115 
-#>  4    18 -0.203   0.154  1.98 -0.00269  -0.00107  -0.0382   -0.00167 
-#>  5    19 -0.628   0.114  1.78 -0.000131 -0.00323   0.0347   -0.00288 
-#>  6    27  0.987   0.579  1.14 -0.00124  -0.00121   0.0199   -0.00363 
-#>  7    28  0.847   0.537  1.51  0.000374 -0.00215  -0.00901  -0.000548
-#>  8    29  0.469   0.476  1.87  0.00117  -0.000439 -0.0191    0.000251
-#>  9    30 -0.00747 0.491  1.99 -0.000570  0.00114  -0.000815  0.000786
-#> 10    31 -0.525   0.449  1.84  0.00141  -0.000575 -0.0112    0.000714
-#> # i 76 more rows
+#>  1    21 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105 
+#>  2    22 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363
+#>  3    23 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153 
+#>  4    24 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128 
+#>  5    25  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195 
+#>  6    26  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172 
+#>  7    36 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338
+#>  8    37 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126
+#>  9    38 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780
+#> 10    39 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117 
+#> # i 120 more rows
 #> 
-#> $df_bin_centroids
-#> # A tibble: 86 x 6
-#>    hexID   c_x      c_y bin_counts std_counts drop_empty
-#>    <int> <dbl>    <dbl>      <int>      <dbl> <lgl>     
-#>  1    15 0.158 0.000959         43     0.518  FALSE     
-#>  2    16 0.262 0.000959         35     0.422  FALSE     
-#>  3    17 0.365 0.000959         21     0.253  FALSE     
-#>  4    18 0.468 0.000959         11     0.133  FALSE     
-#>  5    19 0.571 0.000959          8     0.0964 FALSE     
-#>  6    27 0.107 0.0904           10     0.120  FALSE     
-#>  7    28 0.210 0.0904           56     0.675  FALSE     
-#>  8    29 0.313 0.0904           43     0.518  FALSE     
-#>  9    30 0.416 0.0904           54     0.651  FALSE     
-#> 10    31 0.520 0.0904           42     0.506  FALSE     
-#> # i 76 more rows
+#> $model_2d
+#> # A tibble: 130 x 5
+#>    hexID   c_x     c_y bin_counts std_counts
+#>    <int> <dbl>   <dbl>      <dbl>      <dbl>
+#>  1    21 0.358 -0.0171         12     0.16  
+#>  2    22 0.441 -0.0171         17     0.227 
+#>  3    23 0.524 -0.0171         22     0.293 
+#>  4    24 0.608 -0.0171         10     0.133 
+#>  5    25 0.691 -0.0171          7     0.0933
+#>  6    26 0.774 -0.0171         12     0.16  
+#>  7    36 0.316  0.0550         42     0.56  
+#>  8    37 0.400  0.0550         42     0.56  
+#>  9    38 0.483  0.0550         44     0.587 
+#> 10    39 0.566  0.0550         39     0.52  
+#> # i 120 more rows
+#> 
+#> $trimesh_data
+#> # A tibble: 324 x 6
+#>     from    to x_from  y_from  x_to   y_to
+#>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>
+#>  1    16    26 0.275   0.127  0.233 0.199 
+#>  2    25    26 0.150   0.199  0.233 0.199 
+#>  3    25    37 0.150   0.199  0.191 0.271 
+#>  4    36    49 0.108   0.271  0.150 0.343 
+#>  5    36    37 0.108   0.271  0.191 0.271 
+#>  6     1     7 0.358  -0.0171 0.316 0.0550
+#>  7    48    49 0.0665  0.343  0.150 0.343 
+#>  8    37    38 0.191   0.271  0.275 0.271 
+#>  9    26    27 0.233   0.199  0.316 0.199 
+#> 10    27    38 0.316   0.199  0.275 0.271 
+#> # i 314 more rows
 \end{verbatim}
 
-\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) model}{Constructing the 2\textbackslash text\{-\}D model}}\label{constructing-the-2text-d-model}
+\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
 
-Constructing the \(2\text{-}D\) model mainly contains (i) binning data, (ii) obtaining bin centroids, and (iii) indicating neighbors by line segments connecting centroids.
+Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) Remove low-density hexagons.
+
+\subsubsection{Scaling the Data}\label{scaling-the-data}
+
+The algorithm starts by scaling the NLDR data to a standard range using the \texttt{gen\_scaled\_data()} function.
+This function standardizes the data so that the first embedding ranges from \(0\) to \(1\), while the second embedding scales from \(0\) to the maximum value of the second embedding.
+The output includes the scaled NLDR data along with the original limits of the embeddings.
+
+\begin{verbatim}
+scurve_nldr_obj <- gen_scaled_data(nldr_data = scurve_umap)
+
+scurve_nldr_obj
+\end{verbatim}
+
+\begin{verbatim}
+#> $scaled_nldr
+#> # A tibble: 5,000 x 3
+#>     emb1  emb2    ID
+#>    <dbl> <dbl> <int>
+#>  1 0.707 0.839     1
+#>  2 0.231 0.401     2
+#>  3 0.232 0.215     3
+#>  4 0.790 0.564     4
+#>  5 0.761 0.551     5
+#>  6 0.445 0.721     6
+#>  7 0.900 0.137     7
+#>  8 0.247 0.392     8
+#>  9 0.325 0.542     9
+#> 10 0.278 0.231    10
+#> # i 4,990 more rows
+#> 
+#> $lim1
+#> [1] -14.42166  13.32655
+#> 
+#> $lim2
+#> [1] -12.43687  12.32455
+\end{verbatim}
 
@@
 calc_bins_y(
-  bin1 = 12, 
-  r2 = r2,
-  q = 0.1
-  )
+  nldr_obj = scurve_nldr_obj, 
+  bin1 = 4, 
+  q = 0.1)
 \end{verbatim}
@@
 #> $bin2
-#> [1] 13
+#> [1] 5
 #> 
 #> $a1
-#> [1] 0.1032943
+#> [1] 0.3122301
 #> 
 #> $a2
-#> [1] 0.08945548
+#> [1] 0.2703992
 \end{verbatim}
@@
 hex_binning(
-  data = s_curve_noise_umap_scaled, 
-  bin1 = 12, 
-  r2 = r2,
-  q = 0.1
-  )
+  nldr_obj = scurve_nldr_obj, 
+  bin1 = 4, 
+  q = 0.1)
 \end{verbatim}
@@
 #> $a1
-#> [1] 0.1032943
+#> [1] 0.3122301
 #> 
 #> $a2
-#> [1] 0.08945548
+#> [1] 0.2703992
 #> 
 #> $bins
-#> [1] 12 13
+#> [1] 4 5
 #> 
 #> $start_point
-#> [1] -0.10000000 -0.08849688
+#> [1] -0.10000000 -0.08923607
 #> 
 #> $centroids
-#> # A tibble: 156 x 3
+#> # A tibble: 20 x 3
 #>    hexID     c_x     c_y
 #>    <int>   <dbl>   <dbl>
-#>  1     1 -0.1     -0.0885
-#>  2     2  0.00329 -0.0885
-#>  3     3  0.107   -0.0885
-#>  4     4  0.210   -0.0885
-#>  5     5  0.313   -0.0885
-#>  6     6  0.416   -0.0885
-#>  7     7  0.520   -0.0885
-#>  8     8  0.623   -0.0885
-#>  9     9  0.726   -0.0885
-#> 10    10  0.830   -0.0885
-#> # i 146 more rows
+#>  1     1 -0.1    -0.0892
+#>  2     2  0.212  -0.0892
+#>  3     3  0.524  -0.0892
+#>  4     4  0.837  -0.0892
+#>  5     5  0.0561  0.181 
+#>  6     6  0.368   0.181 
+#>  7     7  0.681   0.181 
+#>  8     8  0.993   0.181 
+#>  9     9 -0.1     0.452 
+#> 10    10  0.212   0.452 
+#> 11    11  0.524   0.452 
+#> 12    12  0.837   0.452 
+#> 13    13  0.0561  0.722 
+#> 14    14  0.368   0.722 
+#> 15    15  0.681   0.722 
+#> 16    16  0.993   0.722 
+#> 17    17 -0.1     0.992 
+#> 18    18  0.212   0.992 
+#> 19    19  0.524   0.992 
+#> 20    20  0.837   0.992 
 #> 
 #> $hex_poly
-#> # A tibble: 936 x 3
+#> # A tibble: 120 x 3
 #>    hex_poly_id       x         y
 #>          <int>   <dbl>     <dbl>
-#>  1           1 -0.1     -0.0289
-#>  2           1 -0.152   -0.0587
-#>  3           1 -0.152   -0.118 
-#>  4           1 -0.1     -0.148 
-#>  5           1 -0.0484  -0.118 
-#>  6           1 -0.0484  -0.0587
-#>  7           2  0.00329 -0.0289
-#>  8           2 -0.0484  -0.0587
-#>  9           2 -0.0484  -0.118 
-#> 10           2  0.00329 -0.148 
-#> # i 926 more rows
+#>  1           1 -0.1     0.0910  
+#>  2           1 -0.256   0.000897
+#>  3           1 -0.256  -0.179   
+#>  4           1 -0.1    -0.270   
+#>  5           1  0.0561 -0.179   
+#>  6           1  0.0561  0.000897
+#>  7           2  0.212   0.0910  
+#>  8           2  0.0561  0.000897
+#>  9           2  0.0561 -0.179   
+#> 10           2  0.212  -0.270   
+#> # i 110 more rows
 #> 
 #> $data_hb_id
-#> # A tibble: 3,750 x 4
-#>     emb1   emb2    ID hb_id
+#> # A tibble: 5,000 x 4
+#>     emb1  emb2    ID hexID
 #>    <dbl> <dbl> <int> <int>
-#>  1 0.270 0.839      1   125
-#>  2 0.788 0.466      2    82
-#>  3 0.771 0.319      3    69
-#>  4 0.306 0.542      5    88
-#>  5 0.549 0.806      6   127
-#>  6 0.166 0.259      7    52
-#>  7 0.672 0.596      9   104
-#>  8 0.735 0.354     10    69
-#>  9 0.839 0.467     11    82
-#> 10 0.240 0.0903    12    28
-#> # i 3,740 more rows
+#>  1 0.707 0.839     1    15
+#>  2 0.231 0.401     2    10
+#>  3 0.232 0.215     3     6
+#>  4 0.790 0.564     4    12
+#>  5 0.761 0.551     5    12
+#>  6 0.445 0.721     6    14
+#>  7 0.900 0.137     7     8
+#>  8 0.247 0.392     8    10
+#>  9 0.325 0.542     9    10
+#> 10 0.278 0.231    10     6
+#> # i 4,990 more rows
 #> 
 #> $std_cts
-#> # A tibble: 86 x 3
-#>    hb_id     n std_counts
+#> # A tibble: 16 x 3
+#>    hexID bin_counts std_counts
 #>    <int>      <int>      <dbl>
-#>  1    15    43     0.518 
-#>  2    16    35     0.422 
-#>  3    17    21     0.253 
-#>  4    18    11     0.133 
-#>  5    19     8     0.0964
-#>  6    27    10     0.120 
-#>  7    28    56     0.675 
-#>  8    29    43     0.518 
-#>  9    30    54     0.651 
-#> 10    31    42     0.506 
-#> # i 76 more rows
+#>  1     3        116    0.172  
+#>  2     4         82    0.122  
+#>  3     5        193    0.286  
+#>  4     6        674    1      
+#>  5     7        529    0.785  
+#>  6     8        243    0.361  
+#>  7     9         87    0.129  
+#>  8    10        601    0.892  
+#>  9    11        244    0.362  
+#> 10    12        355    0.527  
+#> 11    13        183    0.272  
+#> 12    14        627    0.930  
+#> 13    15        669    0.993  
+#> 14    16        318    0.472  
+#> 15    19          2    0.00297
+#> 16    20         77    0.114  
 #> 
 #> $tot_bins
-#> [1] 156
+#> [1] 20
 #> 
 #> $non_bins
-#> [1] 86
+#> [1] 16
 #> 
 #> $pts_bins
-#> # A tibble: 86 x 2
-#>    hb_id pts_list  
+#> # A tibble: 16 x 2
+#>    hexID pts_list   
 #>    <int> <list>     
-#>  1    15 <int [43]>
-#>  2    16 <int [35]>
-#>  3    17 <int [21]>
-#>  4    18 <int [11]>
-#>  5    19 <int [8]> 
-#>  6    27 <int [10]>
-#>  7    28 <int [56]>
-#>  8    29 <int [43]>
-#>  9    30 <int [54]>
-#> 10    31 <int [42]>
-#> # i 76 more rows
+#>  1     3 <int [116]>
+#>  2     4 <int [82]> 
+#>  3     5 <int [193]>
+#>  4     6 <int [674]>
+#>  5     7 <int [529]>
+#>  6     8 <int [243]>
+#>  7     9 <int [87]> 
+#>  8    10 <int [601]>
+#>  9    11 <int [244]>
+#> 10    12 <int [355]>
+#> 11    13 <int [183]>
+#> 12    14 <int [627]>
+#> 13    15 <int [669]>
+#> 14    16 <int [318]>
+#> 15    19 <int [2]>  
+#> 16    20 <int [77]> 
 #> 
@@
 #> [1] "hex_bin_obj"
-\end{verbatim}
-
-\subsubsection{Remove low-density hexagons}\label{remove-low-density-hexagons}
-
-In certain scenarios, hexagonal bins may contain a few number of points.
-To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
-The \texttt{find\_low\_dens\_hex()} function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
-
-\begin{verbatim}
-find_low_dens_hex(
-  df_bin_centroids_all = df_bin_centroids, 
-  bin1 = 12, 
-  df_bin_centroids_low = df_bin_centroids_low
-  )
 \end{verbatim}
@@
 
+\subsubsection{Remove low-density hexagons}\label{remove-low-density-hexagons}
+
+In certain scenarios, hexagonal bins may contain a few number of points.
+To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
+The \texttt{find\_low\_dens\_hex()} function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
+Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
+
 \begin{verbatim}
-find_lg_benchmark(
-  distance_edges = distance_df, 
-  distance_col = "distance"
+find_low_dens_hex(
+  centroids_data, 
+  bin1 = 15, 
+  benchmark_mean_dens = 0.05
 )
@@
 
-The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each bin to represent bin centroids.
-This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data and their corresponding hexagonal bin IDs as inputs.
+The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each hexgaonal bin to represent bin centroids.
+This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
 
 \begin{verbatim}
-umap_data_with_hb_id <- hb_obj$data_hb_id
-
-df_all <- bind_cols(
-  s_curve_noise_training |> dplyr::select(-ID), 
-  umap_data_with_hb_id
-  )
-
 df_bin <- avg_highd_data(
-  data = df_all
+  highd_data = scurve, 
+  scaled_nldr_hexid
 )
@@
 \begin{verbatim}
-vis_lg_mesh(
-  distance_edges = distance_df, 
-  benchmark_value = 0.75, 
-  tr_coord_df = tr_from_to_df, 
-  distance_col = "distance"
-  )
-\end{verbatim}
-
-\begin{verbatim}
-vis_rmlg_mesh(
+vis_mesh(
   distance_edges = distance_df, 
@@
 \begin{verbatim}
+df_exe <- comb_data_mode(
+  highd_data = scurve, 
+  model_highd = scurve_model_obj$model_highd, 
+  model_2d = scurve_model_obj$model_2d
+  )
+\end{verbatim}
+
+\begin{verbatim}
 show_langevitour(
-  df = df_all, 
-  df_b = df_bin, 
-  df_b_with_center_data = df_bin_centroids, 
-  benchmark_value = 0.75, 
-  distance = distance_df, 
-  distance_col = "distance", 
-  use_default_benchmark_val = FALSE, 
-  col_start = "x"
+  point_df, 
+  edge_df
   )

### 2025-05-08 paper-quollr.Rmd
Updated 2D model generation

@@
 ```{r, echo=TRUE}
-scurve_nldr_obj <- gen_scaled_data(nldr_data = scurve_umap)
+scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
 
-scurve_nldr_obj
+scurve_umap_obj
 ```
@@
 
-The configurations of the hexagonal grid is defined by the number of bins in each direction.
-To find the number of bins along the y-axis, `calc_bins_y()` is used.
-This function takes as input the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
-Additionally, this function provides the bin width.
+The configurations of a hexagonal grid are determined by the number of bins and the bin width in each direction.
+The function `calc_bins_y()` is used for this purpose.
+This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (`bin1`), and the buffer amount as a proportion.
 
@@
 calc_bins_y(
-  nldr_obj = scurve_nldr_obj, 
-  bin1 = 4, 
+  nldr_obj = scurve_umap_obj, 
+  bin1 = 15, 
   q = 0.1)
@@
 
-Points are allocated to the bins they fall into based on the nearest centroid.
-The main steps of the hexagonal binning algorithm can be executed using the `hex_binning()` function, or they can be run separately for greater flexibility.
-The parameters used within `hex_binning()` include the scaled NLDR data, the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
-The output is an object of the `hex_bin_obj` class, which contains the number of bins in each direction, the coordinates of the hexagonal grid starting point, the details of bin centroids, the coordinates of bins, embedding components with their corresponding hexagon IDs, hex bins with their corresponding standardized counts, the total number of bins, the number of non-empty bins, and the points within each hexagon.
+Points are allocated to bins based on the nearest centroid.
+The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its components can be run separately for added flexibility.
+The parameters used within `hex_binning()` include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (`nldr_obj`), the number of bins along the x-axis (`bin1`), and the buffer amount as a proportion of the data (`q`).
+The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagon IDs (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`tot_bins`), the number of non-empty bins (`non_bins`), and the points within each hexagon (`pts_bins`).
 
 ```{r, echo=TRUE}
-hex_binning(
-  nldr_obj = scurve_nldr_obj, 
-  bin1 = 4, 
+hb_obj <- hex_binning(
+  nldr_obj = scurve_umap_obj, 
+  bin1 = 15, 
   q = 0.1)
@@
 <!--add hexbin notation image-->
+
+### Obtaining bin centroids
+
+```{r, echo=TRUE}
+df_bin_centroids <- extract_hexbin_centroids(
+  centroids_data = hb_obj$centroids, 
+  counts_data = hb_obj$std_cts
+  )
+
+df_bin_centroids
+```
 
@@
 
-```{r, echo=TRUE, eval=FALSE}
-tr1_object <- tri_bin_centroids(
-  hex_df = df_bin_centroids, 
-  x = "c_x", 
-  y = "c_y"
+```{r, echo=TRUE}
+tr_object <- tri_bin_centroids(
+  centroids_data = df_bin_centroids
   )
 
-tr_from_to_df <- gen_edges(
-  tri_object = tr1_object
-  )
+trimesh <- gen_edges(tri_object = tr_object)
+trimesh
 ```
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 find_low_dens_hex(
-  centroids_data, 
+  centroids_data = df_bin_centroids, 
   bin1 = 15, 
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 df_bin <- avg_highd_data(
   highd_data = scurve, 
-  scaled_nldr_hexid
+  scaled_nldr_hexid = hb_obj$data_hb_id
 )
+
+df_bin
 ```
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 predict_emb(
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 glance(
@@
 ```{r, echo=TRUE, eval=FALSE}
-augment(
+model_error <- augment(
   highd_data = scurve, 
@@
   )
+
+model_error
 ```
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 ggplot() + 
   geom_hexgrid(
-    data = all_centroids_df, 
+    data = hb_obj$centroids, 
     aes(x = c_x, y = c_y)
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 ggplot() + 
   geom_trimesh(
-    data = df_bin_centroids, 
+    data = hb_obj$centroids, 
     aes(x = c_x, y = c_y)
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 vis_mesh(
-  distance_edges = distance_df, 
-  benchmark_value = 0.75, 
-  tr_coord_df = tr_from_to_df, 
-  distance_col = "distance"
+  trimesh_data = trimesh
   )
@@
 
-```{r, echo=TRUE, eval=FALSE}
-df_exe <- comb_data_mode(
+```{r, echo=TRUE}
+df_exe <- comb_data_model(
   highd_data = scurve, 
-  model_highd = scurve_model_obj$model_highd, 
-  model_2d = scurve_model_obj$model_2d
+  model_highd = df_bin, 
+  model_2d = df_bin_centroids
   )
+
+df_exe
 ```
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 show_langevitour(
-  point_df, 
-  edge_df
+  point_data = df_exe, 
+  edge_data = trimesh
   )
@@
 ```{r, echo=TRUE, eval=FALSE}
-show_link_plots(
-  point_df = df_exe, 
-  edge_df = distance_small_df
+df_exe <- comb_all_data_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  model_highd = scurve_model_obj$model_highd, 
+  model_2d = scurve_model_obj$model_2d
   )
@@
 ```{r, echo=TRUE, eval=FALSE}
+show_link_plots(
+  point_data = df_exe, 
+  edge_data = trimesh
+  )
+```
+
+```{r, echo=TRUE, eval=FALSE}
+df_exe <- comb_all_data_model_error(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  model_highd = scurve_model_obj$model_highd, 
+  model_2d = scurve_model_obj$model_2d, 
+  error_data = model_error)
+```
+
+```{r, echo=TRUE, eval=FALSE}
 show_error_link_plots(
-  point_df = df_exe, 
-  edge_df = distance_small_df
+  point_data = df_exe, 
+  edge_data = trimesh
   )

### 2025-05-08 paper-quollr.tex
Updated 2D model generation

@@
 \begin{verbatim}
-scurve_nldr_obj <- gen_scaled_data(nldr_data = scurve_umap)
+scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
 
-scurve_nldr_obj
+scurve_umap_obj
 \end{verbatim}
@@
 
-The configurations of the hexagonal grid is defined by the number of bins in each direction.
-To find the number of bins along the y-axis, \texttt{calc\_bins\_y()} is used.
-This function takes as input the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
-Additionally, this function provides the bin width.
+The configurations of a hexagonal grid are determined by the number of bins and the bin width in each direction.
+The function \texttt{calc\_bins\_y()} is used for this purpose.
+This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion.
 
@@
 #> $bin2
-#> [1] 5
+#> [1] 16
 #> 
 #> $a1
-#> [1] 0.3122301
+#> [1] 0.08326136
 #> 
 #> $a2
-#> [1] 0.2703992
+#> [1] 0.07210645
 \end{verbatim}
@@
 
-Points are allocated to the bins they fall into based on the nearest centroid.
-The main steps of the hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or they can be run separately for greater flexibility.
-The parameters used within \texttt{hex\_binning()} include the scaled NLDR data, the number of bins along the x-axis, the ratio of the ranges of the original embedding components, and the buffer amount as a proportion of the data.
-The output is an object of the \texttt{hex\_bin\_obj} class, which contains the number of bins in each direction, the coordinates of the hexagonal grid starting point, the details of bin centroids, the coordinates of bins, embedding components with their corresponding hexagon IDs, hex bins with their corresponding standardized counts, the total number of bins, the number of non-empty bins, and the points within each hexagon.
+Points are allocated to bins based on the nearest centroid.
+The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its components can be run separately for added flexibility.
+The parameters used within \texttt{hex\_binning()} include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (\texttt{nldr\_obj}), the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion of the data (\texttt{q}).
+The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{tot\_bins}), the number of non-empty bins (\texttt{non\_bins}), and the points within each hexagon (\texttt{pts\_bins}).
 
 \begin{verbatim}
-hex_binning(
-  nldr_obj = scurve_nldr_obj, 
-  bin1 = 4, 
+hb_obj <- hex_binning(
+  nldr_obj = scurve_umap_obj, 
+  bin1 = 15, 
   q = 0.1)
@@
 
+\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
+
 \begin{verbatim}
-#> $a1
-#> [1] 0.3122301
-#> 
-#> $a2
-#> [1] 0.2703992
-#> 
-#> $bins
-#> [1] 4 5
-#> 
-#> $start_point
-#> [1] -0.10000000 -0.08923607
-#> 
-#> $centroids
-#> # A tibble: 20 x 3
-#>    hexID     c_x     c_y
-#>    <int>   <dbl>   <dbl>
-#>  1     1 -0.1    -0.0892
-#>  2     2  0.212  -0.0892
-#>  3     3  0.524  -0.0892
-#>  4     4  0.837  -0.0892
-#>  5     5  0.0561  0.181 
-#>  6     6  0.368   0.181 
-#>  7     7  0.681   0.181 
-#>  8     8  0.993   0.181 
-#>  9     9 -0.1     0.452 
-#> 10    10  0.212   0.452 
-#> 11    11  0.524   0.452 
-#> 12    12  0.837   0.452 
-#> 13    13  0.0561  0.722 
-#> 14    14  0.368   0.722 
-#> 15    15  0.681   0.722 
-#> 16    16  0.993   0.722 
-#> 17    17 -0.1     0.992 
-#> 18    18  0.212   0.992 
-#> 19    19  0.524   0.992 
-#> 20    20  0.837   0.992 
-#> 
-#> $hex_poly
-#> # A tibble: 120 x 3
-#>    hex_poly_id       x         y
-#>          <int>   <dbl>     <dbl>
-#>  1           1 -0.1     0.0910  
-#>  2           1 -0.256   0.000897
-#>  3           1 -0.256  -0.179   
-#>  4           1 -0.1    -0.270   
-#>  5           1  0.0561 -0.179   
-#>  6           1  0.0561  0.000897
-#>  7           2  0.212   0.0910  
-#>  8           2  0.0561  0.000897
-#>  9           2  0.0561 -0.179   
-#> 10           2  0.212  -0.270   
-#> # i 110 more rows
-#> 
-#> $data_hb_id
-#> # A tibble: 5,000 x 4
-#>     emb1  emb2    ID hexID
-#>    <dbl> <dbl> <int> <int>
-#>  1 0.707 0.839     1    15
-#>  2 0.231 0.401     2    10
-#>  3 0.232 0.215     3     6
-#>  4 0.790 0.564     4    12
-#>  5 0.761 0.551     5    12
-#>  6 0.445 0.721     6    14
-#>  7 0.900 0.137     7     8
-#>  8 0.247 0.392     8    10
-#>  9 0.325 0.542     9    10
-#> 10 0.278 0.231    10     6
-#> # i 4,990 more rows
-#> 
-#> $std_cts
-#> # A tibble: 16 x 3
-#>    hexID bin_counts std_counts
-#>    <int>      <int>      <dbl>
-#>  1     3        116    0.172  
-#>  2     4         82    0.122  
-#>  3     5        193    0.286  
-#>  4     6        674    1      
-#>  5     7        529    0.785  
-#>  6     8        243    0.361  
-#>  7     9         87    0.129  
-#>  8    10        601    0.892  
-#>  9    11        244    0.362  
-#> 10    12        355    0.527  
-#> 11    13        183    0.272  
-#> 12    14        627    0.930  
-#> 13    15        669    0.993  
-#> 14    16        318    0.472  
-#> 15    19          2    0.00297
-#> 16    20         77    0.114  
-#> 
-#> $tot_bins
-#> [1] 20
-#> 
-#> $non_bins
-#> [1] 16
-#> 
-#> $pts_bins
-#> # A tibble: 16 x 2
-#>    hexID pts_list   
-#>    <int> <list>     
-#>  1     3 <int [116]>
-#>  2     4 <int [82]> 
-#>  3     5 <int [193]>
-#>  4     6 <int [674]>
-#>  5     7 <int [529]>
-#>  6     8 <int [243]>
-#>  7     9 <int [87]> 
-#>  8    10 <int [601]>
-#>  9    11 <int [244]>
-#> 10    12 <int [355]>
-#> 11    13 <int [183]>
-#> 12    14 <int [627]>
-#> 13    15 <int [669]>
-#> 14    16 <int [318]>
-#> 15    19 <int [2]>  
-#> 16    20 <int [77]> 
-#> 
-#> attr(,"class")
-#> [1] "hex_bin_obj"
+df_bin_centroids <- extract_hexbin_centroids(
+  centroids_data = hb_obj$centroids, 
+  counts_data = hb_obj$std_cts
+  )
+
+df_bin_centroids
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 240 x 5
+#>    hexID     c_x     c_y bin_counts std_counts
+#>    <int>   <dbl>   <dbl>      <dbl>      <dbl>
+#>  1     1 -0.1    -0.0892          0          0
+#>  2     2 -0.0167 -0.0892          0          0
+#>  3     3  0.0665 -0.0892          0          0
+#>  4     4  0.150  -0.0892          0          0
+#>  5     5  0.233  -0.0892          0          0
+#>  6     6  0.316  -0.0892          0          0
+#>  7     7  0.400  -0.0892          0          0
+#>  8     8  0.483  -0.0892          0          0
+#>  9     9  0.566  -0.0892          0          0
+#> 10    10  0.649  -0.0892          0          0
+#> # i 230 more rows
 \end{verbatim}
@@
 \begin{verbatim}
-tr1_object <- tri_bin_centroids(
-  hex_df = df_bin_centroids, 
-  x = "c_x", 
-  y = "c_y"
+tr_object <- tri_bin_centroids(
+  centroids_data = df_bin_centroids
   )
 
-tr_from_to_df <- gen_edges(
-  tri_object = tr1_object
-  )
+trimesh <- gen_edges(tri_object = tr_object)
+trimesh
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 324 x 6
+#>     from    to x_from  y_from  x_to   y_to
+#>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>
+#>  1    16    26 0.275   0.127  0.233 0.199 
+#>  2    25    26 0.150   0.199  0.233 0.199 
+#>  3    25    37 0.150   0.199  0.191 0.271 
+#>  4    36    49 0.108   0.271  0.150 0.343 
+#>  5    36    37 0.108   0.271  0.191 0.271 
+#>  6     1     7 0.358  -0.0171 0.316 0.0550
+#>  7    48    49 0.0665  0.343  0.150 0.343 
+#>  8    37    38 0.191   0.271  0.275 0.271 
+#>  9    26    27 0.233   0.199  0.316 0.199 
+#> 10    27    38 0.316   0.199  0.275 0.271 
+#> # i 314 more rows
 \end{verbatim}
@@
 find_low_dens_hex(
-  centroids_data, 
+  centroids_data = df_bin_centroids, 
   bin1 = 15, 
@@
 )
+\end{verbatim}
+
+\begin{verbatim}
+#>  [1]   1   2   3   4   5  11  12  13  14  15  16  17  18  19  30  31  32  33  46
+#> [20]  47  61  75  76  90 105 120 135 150 151 166 181 182 196 197 198 211 212 213
+#> [39] 214 215 225 226 227 228 229 230 231 232 233 234 235 236 237 238 239 240
 \end{verbatim}
@@
   highd_data = scurve, 
-  scaled_nldr_hexid
+  scaled_nldr_hexid = hb_obj$data_hb_id
 )
+
+df_bin
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 142 x 8
+#>    hexID      x1    x2    x3         x4        x5       x6        x7
+#>    <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
+#>  1    21 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105 
+#>  2    22 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363
+#>  3    23 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153 
+#>  4    24 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128 
+#>  5    25  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195 
+#>  6    26  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172 
+#>  7    27  0.719   1.99 1.70   0.0109    -0.00349  -0.0297  -0.00223 
+#>  8    36 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338
+#>  9    37 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126
+#> 10    38 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780
+#> # i 132 more rows
 \end{verbatim}
@@
 
+\begin{verbatim}
+#> # A tibble: 5,000 x 4
+#>    pred_emb_1 pred_emb_2    ID pred_hb_id
+#>         <dbl>      <dbl> <int>      <int>
+#>  1      0.691      0.848     1        205
+#>  2      0.191      0.416     2        109
+#>  3      0.316      0.199     3         66
+#>  4      0.774      0.560     4        146
+#>  5      0.774      0.560     5        146
+#>  6      0.441      0.704     6        172
+#>  7      0.941      0.127     7         58
+#>  8      0.275      0.416     8        110
+#>  9      0.358      0.560     9        141
+#> 10      0.316      0.343    10         96
+#> # i 4,990 more rows
+\end{verbatim}
+
 \subsection{Compute residuals and Mean Square Error (MSE)}\label{compute-residuals-and-mean-square-error-mse}
@@
   )
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 1 x 2
+#>   Error    MSE
+#>   <dbl>  <dbl>
+#> 1 1481. 0.0325
 \end{verbatim}
@@
   geom_hexgrid(
-    data = all_centroids_df, 
+    data = hb_obj$centroids, 
     aes(x = c_x, y = c_y)
@@
 \end{verbatim}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-14-1}
 
@@
   geom_trimesh(
-    data = df_bin_centroids, 
+    data = hb_obj$centroids, 
     aes(x = c_x, y = c_y)
@@
 
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-15-1}
+
 \begin{verbatim}
 vis_mesh(
-  distance_edges = distance_df, 
-  benchmark_value = 0.75, 
-  tr_coord_df = tr_from_to_df, 
-  distance_col = "distance"
+  trimesh_data = trimesh
   )
 \end{verbatim}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-16-1}
 

### 2025-05-08 paper-quollr.Rmd
Updated Link plots arguments

@@
   nldr_data = scurve_umap, 
-  model_highd = scurve_model_obj$model_highd, 
-  model_2d = scurve_model_obj$model_2d
+  model_highd = df_bin, 
+  model_2d = df_bin_centroids
   )
@@
   nldr_data = scurve_umap, 
-  model_highd = scurve_model_obj$model_highd, 
-  model_2d = scurve_model_obj$model_2d, 
-  error_data = model_error)
+  model_highd = df_bin, 
+  model_2d = df_bin_centroids, 
+  error_data = model_error
+  )
 ```

### 2025-05-08 paper-quollr.Rmd
Updated

@@
     aes(x = c_x, y = c_y)
-    ) +
-  coord_fixed()
+    ) 
 ```
@@
 
-```{r, echo=TRUE}
+```{r, echo=TRUE, eval=FALSE}
 df_exe <- comb_data_model(
@@
 
-```{r, echo=TRUE}
+```{r, echo=TRUE, eval=FALSE}
 show_langevitour(

### 2025-05-08 paper-quollr.tex
Updated

@@
 #> $trimesh_data
-#> # A tibble: 324 x 6
-#>     from    to x_from  y_from  x_to   y_to
-#>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>
-#>  1    16    26 0.275   0.127  0.233 0.199 
-#>  2    25    26 0.150   0.199  0.233 0.199 
-#>  3    25    37 0.150   0.199  0.191 0.271 
-#>  4    36    49 0.108   0.271  0.150 0.343 
-#>  5    36    37 0.108   0.271  0.191 0.271 
-#>  6     1     7 0.358  -0.0171 0.316 0.0550
-#>  7    48    49 0.0665  0.343  0.150 0.343 
-#>  8    37    38 0.191   0.271  0.275 0.271 
-#>  9    26    27 0.233   0.199  0.316 0.199 
-#> 10    27    38 0.316   0.199  0.275 0.271 
+#> # A tibble: 324 x 8
+#>     from    to x_from  y_from  x_to   y_to from_count to_count
+#>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>      <dbl>    <dbl>
+#>  1    50    65 0.275   0.127  0.233 0.199          65       67
+#>  2    64    65 0.150   0.199  0.233 0.199          39       67
+#>  3    64    79 0.150   0.199  0.191 0.271          39       62
+#>  4    78    94 0.108   0.271  0.150 0.343          62       34
+#>  5    78    79 0.108   0.271  0.191 0.271          62       62
+#>  6    21    36 0.358  -0.0171 0.316 0.0550         12       42
+#>  7    93    94 0.0665  0.343  0.150 0.343          45       34
+#>  8    79    80 0.191   0.271  0.275 0.271          62       68
+#>  9    65    66 0.233   0.199  0.316 0.199          67       56
+#> 10    66    80 0.316   0.199  0.275 0.271          56       68
 #> # i 314 more rows
@@
 \begin{verbatim}
-#> # A tibble: 324 x 6
-#>     from    to x_from  y_from  x_to   y_to
-#>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>
-#>  1    16    26 0.275   0.127  0.233 0.199 
-#>  2    25    26 0.150   0.199  0.233 0.199 
-#>  3    25    37 0.150   0.199  0.191 0.271 
-#>  4    36    49 0.108   0.271  0.150 0.343 
-#>  5    36    37 0.108   0.271  0.191 0.271 
-#>  6     1     7 0.358  -0.0171 0.316 0.0550
-#>  7    48    49 0.0665  0.343  0.150 0.343 
-#>  8    37    38 0.191   0.271  0.275 0.271 
-#>  9    26    27 0.233   0.199  0.316 0.199 
-#> 10    27    38 0.316   0.199  0.275 0.271 
-#> # i 314 more rows
+#> # A tibble: 670 x 8
+#>     from    to  x_from  y_from    x_to    y_to from_count to_count
+#>    <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
+#>  1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
+#>  2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
+#>  3     1    31 -0.1    -0.0892 -0.1     0.0550          0        0
+#>  4    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
+#>  5     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
+#>  6    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
+#>  7    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
+#>  8    31    46 -0.1     0.0550 -0.0584  0.127           0        0
+#>  9    32    47 -0.0167  0.0550  0.0249  0.127           0        0
+#> 10    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
+#> # i 660 more rows
 \end{verbatim}
@@
 \begin{verbatim}
-augment(
+model_error <- augment(
   highd_data = scurve, 
@@
   )
+
+model_error
 \end{verbatim}
@@
     aes(x = c_x, y = c_y)
-    ) +
-  coord_fixed()
+    ) 
 \end{verbatim}
@@
 \begin{verbatim}
-df_exe <- comb_data_mode(
+df_exe <- comb_data_model(
   highd_data = scurve, 
-  model_highd = scurve_model_obj$model_highd, 
-  model_2d = scurve_model_obj$model_2d
+  model_highd = df_bin, 
+  model_2d = df_bin_centroids
   )
+
+df_exe
 \end{verbatim}
@@
 show_langevitour(
-  point_df, 
-  edge_df
+  point_data = df_exe, 
+  edge_data = trimesh
   )
@@
 \begin{verbatim}
+df_exe <- comb_all_data_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  model_highd = df_bin, 
+  model_2d = df_bin_centroids
+  )
+\end{verbatim}
+
+\begin{verbatim}
 show_link_plots(
-  point_df = df_exe, 
-  edge_df = distance_small_df
+  point_data = df_exe, 
+  edge_data = trimesh
+  )
+\end{verbatim}
+
+\begin{verbatim}
+df_exe <- comb_all_data_model_error(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  model_highd = df_bin, 
+  model_2d = df_bin_centroids, 
+  error_data = model_error
   )
@@
 show_error_link_plots(
-  point_df = df_exe, 
-  edge_df = distance_small_df
+  point_data = df_exe, 
+  edge_data = trimesh
   )

### 2025-05-08 paper-quollr.tex
Got the trimesh with small edges

@@
 #> $trimesh_data
-#> # A tibble: 324 x 8
+#> # A tibble: 322 x 8
 #>     from    to x_from  y_from  x_to   y_to from_count to_count
@@
 #> 10    66    80 0.316   0.199  0.275 0.271          56       68
-#> # i 314 more rows
+#> # i 312 more rows
 \end{verbatim}
@@
 
-trimesh <- gen_edges(tri_object = tr_object)
+trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1, a2 = hb_obj$a2)
 trimesh
@@
 \begin{verbatim}
-#> # A tibble: 670 x 8
+#> # A tibble: 653 x 8
 #>     from    to  x_from  y_from    x_to    y_to from_count to_count
@@
 #>  2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
-#>  3     1    31 -0.1    -0.0892 -0.1     0.0550          0        0
-#>  4    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
-#>  5     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
-#>  6    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
-#>  7    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
-#>  8    31    46 -0.1     0.0550 -0.0584  0.127           0        0
-#>  9    32    47 -0.0167  0.0550  0.0249  0.127           0        0
-#> 10    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
-#> # i 660 more rows
+#>  3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
+#>  4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
+#>  5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
+#>  6    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
+#>  7    31    46 -0.1     0.0550 -0.0584  0.127           0        0
+#>  8    32    47 -0.0167  0.0550  0.0249  0.127           0        0
+#>  9    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
+#> 10     4    18  0.150  -0.0892  0.108  -0.0171          0        0
+#> # i 643 more rows
 \end{verbatim}

### 2025-05-08 paper-quollr.R

@@ -191,7 +191,7 @@ tr_object <- tri_bin_centroids(
   centroids_data = df_bin_centroids
   )
 
-trimesh <- gen_edges(tri_object = tr_object)
+trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1, a2 = hb_obj$a2)
 trimesh
 
 
@@ -252,8 +252,7 @@ ggplot() +
   geom_trimesh(
     data = hb_obj$centroids, 
     aes(x = c_x, y = c_y)
-    ) +
-  coord_fixed()
+    ) 
 
 
 ## ----echo=TRUE----------------------------------------------------------------

### 2025-05-08 paper-quollr.Rmd
Got the trimesh with small edges
Added new geom_trimesh

@@
 ---
-title: "quollr: An R Package for Visualizing $2-\\text{D}$ Models from Non-linear Dimension Reductions in High Dimensional Space"
+title: "quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space"
 abstract: >
@@
 
-trimesh <- gen_edges(tri_object = tr_object)
+trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1)
 trimesh

### 2025-05-08 paper-quollr.tex
Added new geom_trimesh

@@
 % !TeX root = RJwrapper.tex
-\title{quollr: An R Package for Visualizing \(2-\text{D}\) Models from Non-linear Dimension Reductions in High Dimensional Space}
+\title{quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space}
 
@@
 
-trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1, a2 = hb_obj$a2)
+trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1)
 trimesh

### 2025-05-09 paper-quollr.Rmd
Check the updates with show_langevitour

@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 model_error <- augment(
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
+# df_bin_centroids <- df_bin_centroids |>
+#   dplyr::filter(bin_counts > 10)
+
 df_exe <- comb_data_model(
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE, eval=knitr::is_html_output()}
+# trimesh <- trimesh |>
+#   dplyr::filter(bin_counts > 10)
+
 show_langevitour(
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 df_exe <- comb_all_data_model(
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE, eval=knitr::is_html_output()}
 show_link_plots(
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE}
 df_exe <- comb_all_data_model_error(
@@
 
-```{r, echo=TRUE, eval=FALSE}
+```{r, echo=TRUE, eval=knitr::is_html_output()}
 show_error_link_plots(

### 2025-05-09 paper-quollr.tex
Check the updates with show_langevitour

@@
 
+\begin{verbatim}
+#> # A tibble: 5,000 x 32
+#>       ID      x1     x2       x3       x4       x5       x6        x7 pred_hb_id
+#>    <int>   <dbl>  <dbl>    <dbl>    <dbl>    <dbl>    <dbl>     <dbl>      <int>
+#>  1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923  -0.00128         205
+#>  2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362  -0.00535         109
+#>  3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962   0.00335          66
+#>  4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716   0.00126         146
+#>  5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998   0.000677        146
+#>  6     6  0.818  0.927  -1.58    -0.00318 -0.00980  0.0989   0.00696         172
+#>  7     7  0.910  1.40    1.42     0.00699 -0.0182  -0.0710   0.00966          58
+#>  8     8 -0.0691 1.59    0.00239  0.0127  -0.0130   0.0396  -0.000185        110
+#>  9     9  0.859  1.59   -0.488   -0.0119   0.00421 -0.00440 -0.00595         141
+#> 10    10 -0.727  1.62    0.314    0.00251  0.0177  -0.0755  -0.00369          96
+#> # i 4,990 more rows
+#> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
+#> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
+#> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
+#> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
+#> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
+#> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>, ...
+\end{verbatim}
+
 \subsection{Visualizations}\label{visualizations}
@@
 \begin{verbatim}
+# df_bin_centroids <- df_bin_centroids |>
+#   dplyr::filter(bin_counts > 10)
+
 df_exe <- comb_data_model(
@@
 \begin{verbatim}
+#> # A tibble: 5,142 x 8
+#>         x1    x2    x3         x4        x5       x6        x7 type 
+#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr>
+#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model
+#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model
+#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model
+#>  4 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128  model
+#>  5  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195  model
+#>  6  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model
+#>  7  0.719   1.99 1.70   0.0109    -0.00349  -0.0297  -0.00223  model
+#>  8 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model
+#>  9 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model
+#> 10 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model
+#> # i 5,132 more rows
+\end{verbatim}
+
+\begin{verbatim}
+# trimesh <- trimesh |>
+#   dplyr::filter(bin_counts > 10)
+
 show_langevitour(

### 2025-05-09 paper-quollr.Rmd
Added hexagonal grid with the notifications

@@
 <!--add hexbin notation image-->
+
+```{r}
+#| label: code-illustration
+# Code to draw illustration for notation
+## hexagon binning to have regular hexagons
+hb_obj_notation <- hex_binning(
+  nldr_obj = scurve_umap_obj, 
+  bin1 = 9, 
+  q = 0.1)
+
+a1_temp <- hb_obj_notation$a1
+a2_temp <- hb_obj_notation$a2
+l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
+
+## Data set with all centroids
+all_centroids_df_temp <- hb_obj_notation$centroids
+hex_grid_temp <- hb_obj_notation$hex_poly
+
+hex_grid_temp45 <- hex_grid_temp |> 
+  filter(hex_poly_id == 45)
+
+start_pt <- all_centroids_df_temp |> 
+  filter(hexID == 1)
+d_rect <- tibble(x1min = 0, 
+                 x1max = 1,
+                 x2min = 0,
+                 x2max = diff(scurve_umap_obj$lim2)/diff(scurve_umap_obj$lim1)) # x2max = r2
+
+# To move the rectangle to ignore the overlap with the centroids
+# rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
+rect_adj <- tibble(x1 = -0.03, x2 = 0.03)
+
+a1 <- tibble(x = all_centroids_df_temp$c_x[4],
+             xend = all_centroids_df_temp$c_x[5],
+             y = all_centroids_df_temp$c_y[21],
+             yend = all_centroids_df_temp$c_y[21],
+             label = expression(a[1]))
+a2 <- tibble(x = all_centroids_df_temp$c_x[25],
+             xend = all_centroids_df_temp$c_x[25],
+             y = all_centroids_df_temp$c_y[25],
+             yend = all_centroids_df_temp$c_y[33],
+             label = expression(a[2]))
+l <- tibble(x = hex_grid_temp45$x[2],
+            xend = hex_grid_temp45$x[3],
+            y = hex_grid_temp45$y[2],
+            yend = hex_grid_temp45$y[3],
+            label = expression(l))
+
+hex_param_vis <- ggplot() + 
+    geom_polygon(data = hex_grid_temp, 
+                        aes(x = x, 
+                            y = y, 
+                            group = hex_poly_id),
+                 fill = "white", 
+                 color = "#bdbdbd") +
+    geom_point(data = all_centroids_df_temp, aes(
+      x = c_x, 
+      y = c_y), 
+      color = "#31a354", size = 0.9) +
+    geom_point(data = start_pt, aes(x = c_x, 
+                                    y = c_y), 
+               color = "black") + 
+    geom_rect(data=d_rect, 
+              aes(xmin = x1min - rect_adj$x1,# - rect_adj$s1, 
+                  xmax = x1max - rect_adj$x1,# - rect_adj$s1, 
+                  ymin = x2min - rect_adj$x2,# - rect_adj$s2, 
+                  ymax = x2max - rect_adj$x2),# - rect_adj$s2), 
+              fill = "white", 
+              color = "black", 
+              alpha = 0, 
+              linewidth = 0.7) +
+    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
+                                y=x2min - rect_adj$x2)) + 
+    geom_point(data=d_rect, aes(x=x1max - rect_adj$x1, 
+                                y=x2min - rect_adj$x2)) + 
+    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
+                                y=x2max - rect_adj$x2)) + 
+    annotate("text", x=d_rect$x1min - rect_adj$x1, 
+                     y=d_rect$x2min - rect_adj$x2,
+                     label = "(0,0)", 
+             hjust=-0.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1max - rect_adj$x1, 
+                     y=d_rect$x2min - rect_adj$x2,
+                     label = "(0,1)", 
+             hjust=1.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1min - rect_adj$x1, 
+                     y=d_rect$x2max - rect_adj$x2,
+                     label = expression(group("(", 
+                        list(0, y[2][max]),")")), 
+            hjust=-0.1, vjust=1.2) + 
+    geom_segment(data=d_rect, aes(
+      x = x1min  - rect_adj$x1, # 0 - 0.03, 
+      y = -0.31, 
+      xend = x1max - rect_adj$x1, #1 - 0.03, 
+      yend = -0.31), #-0.35),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+                 color = "black")+
+    annotate("text", x=0.5, y=-0.36, 
+             label = expression(r[1]), color = "black") +
+    geom_segment(data=d_rect, aes(
+      x = -0.25, 
+      y = x2min - rect_adj$x2, #0 - 0.05, 
+      xend = -0.25, 
+      yend = x2max - rect_adj$x2), #r2 - 0.05),
+      arrow = arrow(length = unit(0.03, "npc"),
+                       ends = "both"), 
+                 color = "black")+ 
+    annotate("text", x=-0.3, y=0.4, 
+             label = expression(r[2]), color = "black") +
+    geom_segment(data = a1, aes(
+      x = x, #-0.1 + 0.2087578, 
+      y = y, #-0.15, 
+      xend = xend, #-0.1 + 0.2087578*2, 
+      yend = yend), #-0.15),
+      arrow = arrow(length = unit(0.03, "npc"),
+        ends = "both"), 
+        color = "black")+ # a1 = 0.2087578
+    annotate("text", 
+             x=(a1$x+a1$xend)/2, 
+             y=a1$y, 
+             label = expression(a[1]), 
+             color = "black",
+             vjust = 1.2) +
+    geom_segment(data = a2, aes(
+      x = x, #-0.15, 
+      y = y, #-0.1*r2 + 0.1807896*2, 
+      xend = xend, #-0.15, 
+      yend = yend), #-0.1*r2 + 0.1807896*3),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+      color = "black") + # a2 = 0.1807896
+    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
+             label = expression(a[2]), 
+             color = "black", hjust=-0.2) +
+    annotate("text", x=-0.18, y=-0.24, 
+      label = expression(group("(", list(s[1], s[2]), ")")),
+      color = "black") +
+  geom_segment(data = l, aes(
+      x = x, #-0.15, 
+      y = y, #-0.1*r2 + 0.1807896*2, 
+      xend = xend, #-0.15, 
+      yend = yend), #-0.1*r2 + 0.1807896*3),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+      color = "black") + 
+    annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
+             label = expression(l), 
+             color = "black", hjust=-0.2) +
+  coord_equal()
+```
+
+```{r}
+#| label: fig-hex-param
+#| fig-cap: "The components of the hexagon grid illustrating notation."
+
+hex_param_vis
+```
 

### 2025-05-09 paper-quollr.Rmd
Completed the workflow

@@
 ```{r, echo=TRUE}
-# df_bin_centroids <- df_bin_centroids |>
-#   dplyr::filter(bin_counts > 10)
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::filter(bin_counts > 10)
 
@@
 ```{r, echo=TRUE, eval=knitr::is_html_output()}
-# trimesh <- trimesh |>
-#   dplyr::filter(bin_counts > 10)
+trimesh <- trimesh |>
+  dplyr::filter(from_count > 10,
+                to_count > 10)
+
+trimesh <- update_trimesh_index(trimesh)
 

### 2025-05-09 paper-quollr.tex
Added hexagonal grid with the notifications
Completed the workflow

@@
 #>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>      <dbl>    <dbl>
-#>  1    50    65 0.275   0.127  0.233 0.199          65       67
-#>  2    64    65 0.150   0.199  0.233 0.199          39       67
-#>  3    64    79 0.150   0.199  0.191 0.271          39       62
-#>  4    78    94 0.108   0.271  0.150 0.343          62       34
-#>  5    78    79 0.108   0.271  0.191 0.271          62       62
-#>  6    21    36 0.358  -0.0171 0.316 0.0550         12       42
-#>  7    93    94 0.0665  0.343  0.150 0.343          45       34
-#>  8    79    80 0.191   0.271  0.275 0.271          62       68
-#>  9    65    66 0.233   0.199  0.316 0.199          67       56
-#> 10    66    80 0.316   0.199  0.275 0.271          56       68
+#>  1    16    26 0.275   0.127  0.233 0.199          65       67
+#>  2    25    26 0.150   0.199  0.233 0.199          39       67
+#>  3    25    37 0.150   0.199  0.191 0.271          39       62
+#>  4    36    49 0.108   0.271  0.150 0.343          62       34
+#>  5    36    37 0.108   0.271  0.191 0.271          62       62
+#>  6     1     7 0.358  -0.0171 0.316 0.0550         12       42
+#>  7    48    49 0.0665  0.343  0.150 0.343          45       34
+#>  8    37    38 0.191   0.271  0.275 0.271          62       68
+#>  9    26    27 0.233   0.199  0.316 0.199          67       56
+#> 10    27    38 0.316   0.199  0.275 0.271          56       68
 #> # i 312 more rows
@@
 \end{verbatim}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-hex-param-1} \caption{The components of the hexagon grid illustrating notation.}\label{fig:fig-hex-param}
+\end{figure}
 
@@
 \begin{verbatim}
-# df_bin_centroids <- df_bin_centroids |>
-#   dplyr::filter(bin_counts > 10)
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::filter(bin_counts > 10)
 
@@
 \begin{verbatim}
-#> # A tibble: 5,142 x 8
+#> # A tibble: 5,122 x 8
 #>         x1    x2    x3         x4        x5       x6        x7 type 
@@
 #>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model
-#>  4 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128  model
-#>  5  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195  model
-#>  6  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model
-#>  7  0.719   1.99 1.70   0.0109    -0.00349  -0.0297  -0.00223  model
-#>  8 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model
-#>  9 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model
-#> 10 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model
-#> # i 5,132 more rows
+#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model
+#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model
+#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model
+#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model
+#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model
+#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model
+#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model
+#> # i 5,112 more rows
 \end{verbatim}
@@
 \begin{verbatim}
-# trimesh <- trimesh |>
-#   dplyr::filter(bin_counts > 10)
+trimesh <- trimesh |>
+  dplyr::filter(from_count > 10,
+                to_count > 10)
+
+trimesh <- update_trimesh_index(trimesh)
 

### 2025-05-09 paper-quollr.Rmd
Updated the image in the introduction, add section on finding the best fit

@@
 
-Add the image generated with S-curve data
-
-```{r}
-#| label: scurve-diff-umap-layouts
-
-scurve_umap1 <- read_rds("data/scurve/s_curve_noise_umap.rds")
-
-nldr_scaled_obj <- gen_scaled_data(
-  nldr_data = scurve_umap)
-umap_scaled1 <- nldr_scaled_obj$scaled_nldr
-
-nldr1 <- umap_scaled1 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+Add the image generated with the application
+
+```{r}
+#| label: read-limb-nldr
+# Read a variety of different NLDR representations of limb
+# and plot them on same aspect ratio
+clr_choice <- "#000000"
+umap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr1 <- umap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#999999') +
+  interior_annotation("a")
+
+nldr1_org <- umap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("a")
+
+tsne_limb <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_30.rds")
+
+nldr2 <- tsne_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("b")
+
+nldr2_org <- tsne_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("b")
+
+phate_limb <- read_rds("data/limb_muscles/facs_limb_muscles_phate_knn_5.rds")
+
+nldr3 <- phate_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  theme(aspect.ratio = 1)
-
-scurve_umap2 <- read_rds("data/scurve/s_curve_noise_umap2.rds")
-
-nldr_scaled_obj2 <- gen_scaled_data(
-  nldr_data = scurve_umap2)
-umap_scaled2 <- nldr_scaled_obj2$scaled_nldr
-
-nldr2 <- umap_scaled2 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  interior_annotation("c")
+
+nldr3_org <- phate_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("c")
+
+trimap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr4 <- trimap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("d")
+
+nldr4_org <- trimap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("d")
+
+pacmap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr5 <- pacmap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  interior_annotation("e")
+
+nldr5_org <- pacmap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("e")
+
+tsne_limb2 <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_15.rds")
+
+nldr6 <- tsne_limb2 |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  theme(aspect.ratio = 1)
-
-scurve_umap3 <- read_rds("data/scurve/s_curve_noise_umap3.rds")
-
-nldr_scaled_obj3 <- gen_scaled_data(
-  nldr_data = scurve_umap3)
-umap_scaled3 <- nldr_scaled_obj3$scaled_nldr
-
-nldr3 <- umap_scaled3 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  theme(aspect.ratio = 1)
-
-scurve_umap4 <- read_rds("data/scurve/s_curve_noise_umap4.rds")
-
-nldr_scaled_obj4 <- gen_scaled_data(
-  nldr_data = scurve_umap4)
-umap_scaled4 <- nldr_scaled_obj4$scaled_nldr
-
-nldr4 <- umap_scaled4 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
-  theme(aspect.ratio = 1)
-
-scurve_umap5 <- read_rds("data/scurve/s_curve_noise_umap5.rds")
-
-nldr_scaled_obj5 <- gen_scaled_data(
-  nldr_data = scurve_umap5)
-umap_scaled5 <- nldr_scaled_obj5$scaled_nldr
-
-nldr5 <- umap_scaled5 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#636363') +
-  theme(aspect.ratio = 1)
-
-scurve_umap6 <- read_rds("data/scurve/s_curve_noise_umap6.rds")
-
-nldr_scaled_obj6 <- gen_scaled_data(
-  nldr_data = scurve_umap6)
-umap_scaled6 <- nldr_scaled_obj6$scaled_nldr
-
-nldr6 <- umap_scaled6 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  theme(aspect.ratio = 1)
-```
-
-```{r}
-nldr1 + nldr2 + nldr3 + 
-  nldr4 + nldr5 + nldr6 +
+  interior_annotation("f")
+
+nldr6_org <- tsne_limb2 |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("f")
+```
+
+```{r}
+nldr1_org + nldr2_org + nldr3_org + 
+  nldr4_org + nldr5_org + nldr6_org +
   plot_layout(ncol = 3)
@@
 
+## Find the most appropriate fit
+
+```{r}
+#| label: scurve-diff-umap-layouts
+
+nldr_scaled_obj <- gen_scaled_data(
+  nldr_data = scurve_umap)
+umap_scaled1 <- nldr_scaled_obj$scaled_nldr
+
+nldr1 <- umap_scaled1 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  theme(aspect.ratio = 1)
+
+nldr_scaled_obj2 <- gen_scaled_data(
+  nldr_data = scurve_umap2)
+umap_scaled2 <- nldr_scaled_obj2$scaled_nldr
+
+nldr2 <- umap_scaled2 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  theme(aspect.ratio = 1)
+
+nldr_scaled_obj3 <- gen_scaled_data(
+  nldr_data = scurve_umap3)
+umap_scaled3 <- nldr_scaled_obj3$scaled_nldr
+
+nldr3 <- umap_scaled3 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  theme(aspect.ratio = 1)
+
+nldr_scaled_obj4 <- gen_scaled_data(
+  nldr_data = scurve_umap4)
+umap_scaled4 <- nldr_scaled_obj4$scaled_nldr
+
+nldr4 <- umap_scaled4 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  theme(aspect.ratio = 1)
+
+nldr_scaled_obj5 <- gen_scaled_data(
+  nldr_data = scurve_umap5)
+umap_scaled5 <- nldr_scaled_obj5$scaled_nldr
+
+nldr5 <- umap_scaled5 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#636363') +
+  theme(aspect.ratio = 1)
+
+nldr_scaled_obj6 <- gen_scaled_data(
+  nldr_data = scurve_umap6)
+umap_scaled6 <- nldr_scaled_obj6$scaled_nldr
+
+nldr6 <- umap_scaled6 |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  theme(aspect.ratio = 1)
+```
+
+```{r}
+nldr1 + nldr2 + nldr3 + 
+  nldr4 + nldr5 + nldr6 +
+  plot_layout(ncol = 3)
+```
+
 # Application
@@
 <https://figshare.com/articles/dataset/Robject_files_for_tissues_processed_by_Seurat/5821263/1> <!--data-->
-
-```{r}
-#| label: read-limb-nldr
-# Read a variety of different NLDR representations of limb
-# and plot them on same aspect ratio
-clr_choice <- "#0077A3"
-umap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr1 <- umap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#999999') +
-  interior_annotation("a")
-
-tsne_limb <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_30.rds")
-
-nldr2 <- tsne_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("b")
-
-phate_limb <- read_rds("data/limb_muscles/facs_limb_muscles_phate_knn_5.rds")
-
-nldr3 <- phate_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("c")
-
-trimap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr4 <- trimap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("d")
-
-pacmap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr5 <- pacmap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("e")
-
-tsne_limb2 <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_15.rds")
-
-nldr6 <- tsne_limb2 |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  interior_annotation("f")
-```
 

### 2025-05-09 paper-quollr.Rmd
Added error lot code for Scurve

@@
 library(dplyr)
-library(ggbeeswarm)
 library(patchwork)
-library(Rtsne)
-library(uwot)
 library(readr)
-library(tools)
-library(cardinalR)
 
@@
 
-```{r, echo=TRUE}
-vis_mesh(
-  trimesh_data = trimesh
-  )
-```
-
 ### $p\text{-}D$ model visualization
@@
 
-nldr1 <- umap_scaled1 |>
+scurve_nldr1 <- umap_scaled1 |>
   ggplot(aes(x = emb1,
@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1) +
+  interior_annotation("a", c(0.08, 0.9))
 
@@
 
-nldr2 <- umap_scaled2 |>
+scurve_nldr2 <- umap_scaled2 |>
   ggplot(aes(x = emb1,
@@
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1) +
+  interior_annotation("b")
 
@@
 
-nldr3 <- umap_scaled3 |>
+scurve_nldr3 <- umap_scaled3 |>
   ggplot(aes(x = emb1,
@@
   geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1) +
+  interior_annotation("c", c(0.08, 0.9))
 
@@
 
-nldr4 <- umap_scaled4 |>
+scurve_nldr4 <- umap_scaled4 |>
   ggplot(aes(x = emb1,
@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1) +
+  interior_annotation("d")
 
@@
 
-nldr5 <- umap_scaled5 |>
+scurve_nldr5 <- umap_scaled5 |>
   ggplot(aes(x = emb1,
@@
   geom_point(alpha=0.1, size=1, colour='#636363') +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1) +
+  interior_annotation("e")
 
@@
 
-nldr6 <- umap_scaled6 |>
+scurve_nldr6 <- umap_scaled6 |>
   ggplot(aes(x = emb1,
@@
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  theme(aspect.ratio = 1)
-```
-
-```{r}
-nldr1 + nldr2 + nldr3 + 
-  nldr4 + nldr5 + nldr6 +
-  plot_layout(ncol = 3)
+  theme(aspect.ratio = 1) +
+  interior_annotation("f")
+```
+
+```{r}
+scurve_error <- dplyr::bind_rows(scurve_umap_mse,
+                                 scurve_umap_mse2,
+                                 scurve_umap_mse3,
+                                 scurve_umap_mse4,
+                                 scurve_umap_mse5,
+                                 scurve_umap_mse6)
+
+scurve_error <- scurve_error |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+
+error_plot <- ggplot(scurve_error,
+         aes(x = a1,
+             y = sqrt(MSE),
+             color = method)) +
+    geom_point(size = 0.8) +
+    geom_line(linewidth = 0.3) +
+  scale_x_continuous(breaks =
+    sort(unique(scurve_error$a1))[
+      seq(1, length(
+        unique(scurve_error$a1)), 
+        by = 5)]) +
+  scale_color_manual(
+    values=c('#e41a1c','#ff7f00','#4daf4a', 
+             "#a65628",'#636363', '#984ea3'))+
+    ylab("RMSE") +
+    xlab(expression(paste("binwidth (", a[1], ")"))) +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
+          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+          axis.ticks.x = element_line(),
+          axis.ticks.y = element_line(),
+          legend.position = "none",
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7),
+          axis.title.x = element_text(size = 7),
+          axis.title.y = element_text(size = 7),
+          aspect.ratio = 1.5)
+```
+
+```{r}
+error_plot + wrap_plots(
+  scurve_nldr1, scurve_nldr2, 
+  scurve_nldr3, scurve_nldr4,
+  scurve_nldr5, scurve_nldr6, 
+  ncol = 2, widths = c(50, 50))
 ```

### 2025-05-09 paper-quollr.tex
Updated the image in the introduction, add section on finding the best fit
Added error lot code for Scurve

@@
 
-Add the image generated with S-curve data
+Add the image generated with the application
 
@@
 
-\begin{verbatim}
-vis_mesh(
-  trimesh_data = trimesh
-  )
-\end{verbatim}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-16-1}
-
 \subsubsection{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
@@
 
+\subsection{Find the most appropriate fit}\label{find-the-most-appropriate-fit}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
+
 \section{Application}\label{application}

### 2025-05-09 scripts/01_gen_mse_for_diff_methods.R

@@ -1,0 +1,4 @@
+library(quollr)
+library(dplyr)
+
+set.seed(20240110)

### 2025-05-09 paper-quollr.Rmd
Added the error plot

@@
 library(readr)
+library(tools)
 
@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  theme(aspect.ratio = 1) +
   interior_annotation("a", c(0.08, 0.9))
@@
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  theme(aspect.ratio = 1) +
   interior_annotation("b")
@@
   geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  theme(aspect.ratio = 1) +
   interior_annotation("c", c(0.08, 0.9))
@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  theme(aspect.ratio = 1) +
   interior_annotation("d")
@@
   geom_point(alpha=0.1, size=1, colour='#636363') +
-  theme(aspect.ratio = 1) +
   interior_annotation("e")
@@
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  theme(aspect.ratio = 1) +
   interior_annotation("f")
@@
 ```{r}
-error_plot + wrap_plots(
+free(error_plot) + wrap_plots(
   scurve_nldr1, scurve_nldr2, 
@@
   scurve_nldr5, scurve_nldr6, 
-  ncol = 2, widths = c(50, 50))
+  ncol = 2)
 ```

### 2025-05-09 paper-quollr.Rmd
Added main steps for hexbinning step wise

@@
 ```{r, echo=TRUE}
-calc_bins_y(
+bin_configs <- calc_bins_y(
   nldr_obj = scurve_umap_obj, 
@@
   q = 0.1)
+
+bin_configs
 ```
@@
 
+```{r, echo=TRUE}
+all_centroids_df <- gen_centroids(
+  nldr_obj = scurve_umap_obj, 
+  bin1 = 15, 
+  q = 0.1
+  )
+
+all_centroids_df
+```
+
+```{r, echo=TRUE}
+all_hex_coord <- gen_hex_coord(
+  centroids_data = all_centroids_df, 
+  a1 = bin_configs$a1
+  )
+
+all_hex_coord
+```
+
+```{r, echo=TRUE}
+umap_hex_id <- assign_data(
+  nldr_obj = scurve_umap_obj, 
+  centroids_data = all_centroids_df
+  )
+
+umap_hex_id
+```
+
+```{r, echo=TRUE}
+std_df <- compute_std_counts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+std_df
+```
+
+```{r, echo=TRUE}
+pts_df <- find_pts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+pts_df
+```
+
 ### Obtaining bin centroids
@@
 df_bin_centroids <- extract_hexbin_centroids(
-  centroids_data = hb_obj$centroids, 
+  centroids_data = all_centroids_df, 
   counts_data = hb_obj$std_cts

### 2025-05-09 paper-quollr.tex
Added the error plot
Added main steps for hexbinning step wise

@@
 \begin{verbatim}
-calc_bins_y(
+bin_configs <- calc_bins_y(
   nldr_obj = scurve_umap_obj, 
@@
   q = 0.1)
+
+bin_configs
 \end{verbatim}
@@
 
+\begin{verbatim}
+all_centroids_df <- gen_centroids(
+  nldr_obj = scurve_umap_obj, 
+  bin1 = 15, 
+  q = 0.1
+  )
+
+all_centroids_df
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 240 x 3
+#>    hexID     c_x     c_y
+#>    <int>   <dbl>   <dbl>
+#>  1     1 -0.1    -0.0892
+#>  2     2 -0.0167 -0.0892
+#>  3     3  0.0665 -0.0892
+#>  4     4  0.150  -0.0892
+#>  5     5  0.233  -0.0892
+#>  6     6  0.316  -0.0892
+#>  7     7  0.400  -0.0892
+#>  8     8  0.483  -0.0892
+#>  9     9  0.566  -0.0892
+#> 10    10  0.649  -0.0892
+#> # i 230 more rows
+\end{verbatim}
+
+\begin{verbatim}
+all_hex_coord <- gen_hex_coord(
+  centroids_data = all_centroids_df, 
+  a1 = bin_configs$a1
+  )
+
+all_hex_coord
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 1,440 x 3
+#>    hex_poly_id       x       y
+#>          <int>   <dbl>   <dbl>
+#>  1           1 -0.1    -0.0412
+#>  2           1 -0.142  -0.0652
+#>  3           1 -0.142  -0.113 
+#>  4           1 -0.1    -0.137 
+#>  5           1 -0.0584 -0.113 
+#>  6           1 -0.0584 -0.0652
+#>  7           2 -0.0167 -0.0412
+#>  8           2 -0.0584 -0.0652
+#>  9           2 -0.0584 -0.113 
+#> 10           2 -0.0167 -0.137 
+#> # i 1,430 more rows
+\end{verbatim}
+
+\begin{verbatim}
+umap_hex_id <- assign_data(
+  nldr_obj = scurve_umap_obj, 
+  centroids_data = all_centroids_df
+  )
+
+umap_hex_id
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 5,000 x 4
+#>     emb1  emb2    ID hexID
+#>    <dbl> <dbl> <int> <int>
+#>  1 0.707 0.839     1   205
+#>  2 0.231 0.401     2   109
+#>  3 0.232 0.215     3    65
+#>  4 0.790 0.564     4   146
+#>  5 0.761 0.551     5   146
+#>  6 0.445 0.721     6   172
+#>  7 0.900 0.137     7    58
+#>  8 0.247 0.392     8   110
+#>  9 0.325 0.542     9   141
+#> 10 0.278 0.231    10    80
+#> # i 4,990 more rows
+\end{verbatim}
+
+\begin{verbatim}
+std_df <- compute_std_counts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+std_df
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 142 x 3
+#>    hexID bin_counts std_counts
+#>    <int>      <int>      <dbl>
+#>  1    21         12     0.16  
+#>  2    22         17     0.227 
+#>  3    23         22     0.293 
+#>  4    24         10     0.133 
+#>  5    25          7     0.0933
+#>  6    26         12     0.16  
+#>  7    27          1     0.0133
+#>  8    36         42     0.56  
+#>  9    37         42     0.56  
+#> 10    38         44     0.587 
+#> # i 132 more rows
+\end{verbatim}
+
+\begin{verbatim}
+pts_df <- find_pts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+pts_df
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 142 x 2
+#>    hexID pts_list  
+#>    <int> <list>    
+#>  1    21 <int [12]>
+#>  2    22 <int [17]>
+#>  3    23 <int [22]>
+#>  4    24 <int [10]>
+#>  5    25 <int [7]> 
+#>  6    26 <int [12]>
+#>  7    27 <int [1]> 
+#>  8    36 <int [42]>
+#>  9    37 <int [42]>
+#> 10    38 <int [44]>
+#> # i 132 more rows
+\end{verbatim}
+
 \subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-14-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-19-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-15-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-20-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
 

### 2025-05-09 paper-quollr.R

@@ -174,10 +174,12 @@ scurve_umap_obj
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-calc_bins_y(
+bin_configs <- calc_bins_y(
   nldr_obj = scurve_umap_obj, 
   bin1 = 15, 
   q = 0.1)
+
+bin_configs
 
 
 ## ----echo=TRUE----------------------------------------------------------------
@@ -347,8 +349,52 @@ hex_param_vis
 
 
 ## ----echo=TRUE----------------------------------------------------------------
+all_centroids_df <- gen_centroids(
+  nldr_obj = scurve_umap_obj, 
+  bin1 = 15, 
+  q = 0.1
+  )
+
+all_centroids_df
+
+
+## ----echo=TRUE----------------------------------------------------------------
+all_hex_coord <- gen_hex_coord(
+  centroids_data = all_centroids_df, 
+  a1 = bin_configs$a1
+  )
+
+all_hex_coord
+
+
+## ----echo=TRUE----------------------------------------------------------------
+umap_hex_id <- assign_data(
+  nldr_obj = scurve_umap_obj, 
+  centroids_data = all_centroids_df
+  )
+
+umap_hex_id
+
+
+## ----echo=TRUE----------------------------------------------------------------
+std_df <- compute_std_counts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+std_df
+
+
+## ----echo=TRUE----------------------------------------------------------------
+pts_df <- find_pts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+pts_df
+
+
+## ----echo=TRUE----------------------------------------------------------------
 df_bin_centroids <- extract_hexbin_centroids(
-  centroids_data = hb_obj$centroids, 
+  centroids_data = all_centroids_df, 
   counts_data = hb_obj$std_cts
   )
 

### 2025-05-10 paper-quollr.Rmd
Start working on quollr introduction

@@
 The paper is organized as follows.
-In next section, introduces the implementation of `quollr` package on CRAN, including demonstration of the package's key functions and visualization capabilities.
-We illustrate the algorithm's functionality to study about clustering data structure in **Application** section, and describe a visual heuristic to describe parameter selection.
-Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our algorithm.
+The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
+In the application section, we illustrate the algorithm's functionality for studying clustering data structures.
+Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.
 

### 2025-05-12 paper-quollr.Rmd
Updated the abstract

@@
 abstract: >
-  Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (p\text{-}D) by applying a non-linear transformation.
+  Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation.
 However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices.
@@
 The `scurve` data from the package is used to illustrate the algorithm.
-scRNA-seq data for Limb muscles of mice are used to discuss the usability of the package.
+Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles are used to demonstrate the usability of the package.
 draft: true

### 2025-05-12 paper-quollr.tex
Start working on quollr introduction
Updated the abstract

@@
 \abstract{%
-Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (p\text{-}D) by applying a non-linear transformation.
+Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (\(p\text{-}D\)) by applying a non-linear transformation.
 However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices.
@@
 The \texttt{scurve} data from the package is used to illustrate the algorithm.
-scRNA-seq data for Limb muscles of mice are used to discuss the usability of the package.
+Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles are used to demonstrate the usability of the package.
 }
@@
 The paper is organized as follows.
-In next section, introduces the implementation of \texttt{quollr} package on CRAN, including demonstration of the package's key functions and visualization capabilities.
-We illustrate the algorithm's functionality to study about clustering data structure in \textbf{Application} section, and describe a visual heuristic to describe parameter selection.
-Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our algorithm.
+The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
+In the application section, we illustrate the algorithm's functionality for studying clustering data structures.
+Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.
 

### 2025-05-12 paper-quollr.Rmd
Updated the introduction

@@
 
-This paper presents the R package, `quollr` which introduce a new visual tool in determining which Non-linear dimension reduction (NLDR) technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
-The methodology of the algorithm is explained in *cite the methodology paper*.
-Furthermore, the `quollr` package enables users to perform hexagonal binning [@dan2023], resulting in the generation of regular hexagons.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \@ref(fig:example)).
+It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
+
+This paper presents the R package, `quollr` which introduce a new visual tool to evaluate which NLDR method and which (hyper)parameter choice gives the most accurate representation of high-dimensional data.
+The methodology of this algorithm is explained in *cite the methodology paper*.
 The software is available from the Comprehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).
-
-Add the image generated with the application
 
@@
 
-```{r}
+```{r fig-example, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "", out.width = "100%", fig.width= 6, fig.height= 4}
+
 nldr1_org + nldr2_org + nldr3_org + 

### 2025-05-12 paper-quollr.tex
Updated the introduction

@@
 
-This paper presents the R package, \texttt{quollr} which introduce a new visual tool in determining which Non-linear dimension reduction (NLDR) technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data.
-The methodology of the algorithm is explained in \emph{cite the methodology paper}.
-Furthermore, the \texttt{quollr} package enables users to perform hexagonal binning \citep{dan2023}, resulting in the generation of regular hexagons.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \ref{fig:example}).
+It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
+
+This paper presents the R package, \texttt{quollr} which introduce a new visual tool to evaluate which NLDR method and which (hyper)parameter choice gives the most accurate representation of high-dimensional data.
+The methodology of this algorithm is explained in \emph{cite the methodology paper}.
 The software is available from the Comprehensive R Archive Network (CARN) at \url{https://CRAN.R-project.org/package=quollr}.
 
-Add the image generated with the application
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-1-1}
+\begin{figure}
+\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-example-1} \end{figure}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-19-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-18-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-20-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-19-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
 

### 2025-05-12 paper-quollr.Rmd
Added intro figure title

@@
 
-```{r fig-example, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "", out.width = "100%", fig.width= 6, fig.height= 4}
+```{r fig-example, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
+Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
+Would their downstream analysis decisions vary based on the version they observe?
+Which representation most accurately represents the structure in high-dimensional space?", out.width = "100%", fig.width= 6, fig.height= 4}
 

### 2025-05-12 paper-quollr.tex
Added intro figure title

@@
 \begin{figure}
-\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-example-1} \end{figure}
+\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-example-1} \caption{Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
+Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
+Would their downstream analysis decisions vary based on the version they observe?
+Which representation most accurately represents the structure in high-dimensional space?}\label{fig:fig-example}
+\end{figure}
 

### 2025-05-12 paper-quollr.Rmd
Start working on the application

@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \@ref(fig:example)).
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure @fig-nldrlayouts).
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 
-```{r fig-example, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
+```{r fig-nldrlayouts, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
 Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
@@
 
-Need to find another application
+Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
+NLDR often used to summarise the discovered clusters, and help to understand the results.
+The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
+
+The question is whether this accurately represents the cluster structure in the data. Our method can help here, and also help to provide a better \gD{} layout, as needed.
 
@@
 
-training_data_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pcs_10.rds")
+data_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pcs_10.rds")
 
@@
 ```{r}
-#| label: tsne-num-bins-limb
-#| eval: false
-
-## Compute hexbin parameters
-num_bins_x_limb <- 19
-
-algo_obj_limb <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_limb, 
-  nldr_data = tsne_limb, 
-  num_x_bins = num_bins_x_limb)
-
-tsne_limb_scaled <- algo_obj_limb$nldr_scaled
-distance_limb <- algo_obj_limb$distance_df
-tr_from_to_df_limb <- algo_obj_limb$tr_from_to_df
-benchmark_limb <- algo_obj_limb$benchmark
-df_bin_centroids_limb <- algo_obj_limb$df_bin_centroids
-df_bin_limb <- algo_obj_limb$df_bin
-
-distance_df_small_edges_limb <- distance_limb |>
-  filter(distance < 0.1) #benchmark_limb
-
-tr_from_to_df_limb <- right_join(
-  tr_from_to_df_limb, distance_df_small_edges_limb,
-  by = c("from", "to"))
-
-tsne_limb_scaled_with_cluster <- inner_join(tsne_limb_scaled, cluster_df, by = "ID") |>
-  mutate(cluster.ids = as.character(cluster.ids))
-
-trimesh_removed_limb <- ggplot() + 
-  geom_point(
-    data = tsne_limb_scaled_with_cluster,
-    aes(
-      x = emb1,
-      y = emb2,
-      color = cluster.ids
-    ),
-    alpha = 0.3#,
-    #color = "#a65628"
-  ) +
-    geom_segment(data = tr_from_to_df_limb, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
-  interior_annotation("a1", cex = 2) +
-  theme(
-    aspect.ratio = 1
-  )
-```
-
-```{r}
-#| label: prep-limb-tsne-author-model-proj
-#| eval: false
-
-data_limb <- training_data_limb |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-df_b_limb <- df_bin_limb |>
-  dplyr::filter(hb_id %in% df_bin_centroids_limb$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hb_id),] |>
-  dplyr::select(-hb_id)
-
-# Apply the scaling
-df_model_data_limb <- bind_rows(data_limb, df_b_limb)
-scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
-  as_tibble()
-
-scaled_limb_data <- scaled_limb |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_limb_data_model <- scaled_limb |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-limb-tsne-author-proj
-#| eval: false
-
-data_limb_n <- data_limb |>
-  select(-type) |>
-  mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
-
-df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
-
-langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
-                         lineFrom = distance_df_small_edges_limb$from,
-                         lineTo = distance_df_small_edges_limb$to,
-                         group = factor(df_model_data_limb_n$type,
-                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
-                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
-```
-
-```{r}
-#| label: tsne-best-num-bins-limb
-#| eval: false
-
-## Compute hexbin parameters
-num_bins_x_limb <- 18
-
-algo_obj_limb <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_limb, 
-  nldr_data = tsne_limb2, 
-  num_x_bins = num_bins_x_limb)
-
-tsne_limb_scaled_best <- algo_obj_limb$nldr_scaled
-distance_limb <- algo_obj_limb$distance_df
-tr_from_to_df_limb <- algo_obj_limb$tr_from_to_df
-benchmark_limb <- algo_obj_limb$benchmark
-df_bin_centroids_limb <- algo_obj_limb$df_bin_centroids
-df_bin_limb <- algo_obj_limb$df_bin
-
-distance_df_small_edges_limb <- distance_limb |>
-  filter(distance < benchmark_limb) #benchmark_limb
-
-tr_from_to_df_limb <- right_join(
-  tr_from_to_df_limb, distance_df_small_edges_limb,
-  by = c("from", "to"))
-
-tsne_limb_scaled_best_with_cluster <- inner_join(tsne_limb_scaled_best, cluster_df, by = "ID") |>
-  mutate(cluster.ids = as.character(cluster.ids))
-
-trimesh_removed_limb_best <- ggplot() + 
-  geom_point(
-    data = tsne_limb_scaled_best_with_cluster,
-    aes(
-      x = emb1,
-      y = emb2,
-      color = cluster.ids
-    ),
-    alpha = 0.3#,
-    #color = "#ff7f00"
-  ) +
-    geom_segment(data = tr_from_to_df_limb, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
-  interior_annotation("b1", cex = 2) +
-  theme(
-    aspect.ratio = 1
-  )
-```
-
-```{r}
-#| label: prep-limb-tsne-best-model-proj
-#| eval: false
-
-data_limb <- training_data_limb |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-df_b_limb <- df_bin_limb |>
-  dplyr::filter(hb_id %in% df_bin_centroids_limb$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hb_id),] |>
-  dplyr::select(-hb_id)
-
-# Apply the scaling
-df_model_data_limb <- bind_rows(data_limb, df_b_limb)
-scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
-  as_tibble()
-
-scaled_limb_data <- scaled_limb |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_limb_data_model <- scaled_limb |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-limb-tsne-best-proj
-#| eval: false
-
-df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
-
-langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
-                         lineFrom = distance_df_small_edges_limb$from,
-                         lineTo = distance_df_small_edges_limb$to,
-                         group = factor(df_model_data_limb_n$type,
-                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
-                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
-```
-
-```{r}
-#| eval: false
-trimesh_removed_limb + trimesh_removed_limb_best +
-  plot_layout(ncol = 3)
+#| label: tsne-model-limb
+
 ```

### 2025-05-12 paper-quollr.tex
Start working on the application

--- 2025-05-12 paper-quollr.tex
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \ref{fig:example}).
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \citet{fig-nldrlayouts}).
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-example-1} \end{figure}
+\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-nldrlayouts-1} \caption{Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
+Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
+Would their downstream analysis decisions vary based on the version they observe?
+Which representation most accurately represents the structure in high-dimensional space?}\label{fig:fig-nldrlayouts}
+\end{figure}
 

### 2025-05-12 paper-quollr.R

@@ -148,7 +148,7 @@ nldr6_org <- tsne_limb2 |>
   interior_annotation("f")
 
 
-## ----fig-example, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice. Would their downstream analysis decisions vary based on the version they observe? Which representation most accurately represents the structure in high-dimensional space?", out.width = "100%", fig.width= 6, fig.height= 4----
+## ----fig-nldrlayouts, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice. Would their downstream analysis decisions vary based on the version they observe? Which representation most accurately represents the structure in high-dimensional space?", out.width = "100%", fig.width= 6, fig.height= 4----
 
 nldr1_org + nldr2_org + nldr3_org + 
   nldr4_org + nldr5_org + nldr6_org +

### 2025-05-12 paper-quollr.Rmd
Added intro to application

@@
 
+Single-cell RNA sequencing (scRNA-seq) is a popular and powerful technology that allows you to profile the whole transcriptome of a large number of individual cells [@andrews2021].
+
 Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
@@
 
-The question is whether this accurately represents the cluster structure in the data. Our method can help here, and also help to provide a better \gD{} layout, as needed.
+Limb muscle cells of mice in @tabula2018 are examined.
+There are $1067$ single cells, with $14997$ gene expressions.
+Following their pre-processing, tSNE was performed using ten principal components.
+@fig-NLDR-variety (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout.The question is whether this accurately represents the cluster structure in the data.
+Our method can help here, and also help to provide a better $2-\text{D}$ layout, as needed.
 
-<https://www.nature.com/articles/s41586-018-0590-4#data-availability> <!--paper-->
-<https://figshare.com/articles/dataset/Robject_files_for_tissues_processed_by_Seurat/5821263/1> <!--data-->
+<!-- https://www.nature.com/articles/s41586-018-0590-4#data-availability -->
+<!-- https://figshare.com/articles/dataset/Robject_files_for_tissues_processed_by_Seurat/5821263/1 -->
 

### 2025-05-12 paper-quollr.tex
Added intro to application

@@
 
-Need to find another application
+Single-cell RNA sequencing (scRNA-seq) is a popular and powerful technology that allows you to profile the whole transcriptome of a large number of individual cells \citep{andrews2021}.
 
-\url{https://www.nature.com/articles/s41586-018-0590-4\#data-availability}
-\url{https://figshare.com/articles/dataset/Robject_files_for_tissues_processed_by_Seurat/5821263/1}
+Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
+NLDR often used to summarise the discovered clusters, and help to understand the results.
+The purpose of this example is to \emph{illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data}.
+
+Limb muscle cells of mice in \citet{tabula2018} are examined.
+There are \(1067\) single cells, with \(14997\) gene expressions.
+Following their pre-processing, tSNE was performed using ten principal components.
+\citet{fig-NLDR-variety} (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout.The question is whether this accurately represents the cluster structure in the data.
+Our method can help here, and also help to provide a better \(2-\text{D}\) layout, as needed.
 

### 2025-05-12 paper-quollr.Rmd
Application figure title

@@
 ```{r}
-#| fig-cap: ""
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f)."
 #| label: fig-limb-mse

### 2025-05-12 paper-quollr.tex
Application figure title

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-limb-mse-1} \end{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-limb-mse-1} \caption{Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).}\label{fig:fig-limb-mse}
+\end{figure}
 

### 2025-05-12 paper-quollr.Rmd
Added langevitour of the model

@@
 
+tsne_limb_obj <- fit_highd_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb, 
+  bin1 = 19, 
+  q = 0.1, 
+  benchmark_highdens = 2)
+
+df_bin_centroids_limb <- tsne_limb_obj$model_2d
+df_bin_limb <- tsne_limb_obj$model_highd
+trimesh_data_limb <- tsne_limb_obj$trimesh_data
+tsne_limb_scaled <- tsne_limb_obj$nldr_obj$scaled_nldr
+
+tsne_limb_scaled_with_cluster <- inner_join(tsne_limb_scaled, cluster_df, by = "ID") |>
+  mutate(cluster = as.character(cluster.ids))
+
+trimesh_limb <- ggplot() + 
+  geom_point(
+    data = tsne_limb_scaled_with_cluster,
+    aes(
+      x = emb1,
+      y = emb2,
+      color = cluster
+    ),
+    alpha = 0.3
+  ) +
+  geom_segment(data = trimesh_data_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
+  interior_annotation("a1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+```
+
+```{r}
+#| label: prep-limb-tsne-author-model-proj
+
+data_limb_n <- data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(hexID %in% df_bin_centroids_limb$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
+scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
+  as_tibble()
+
+scaled_limb_data <- scaled_limb |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_limb_data_model <- scaled_limb |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r, eval=knitr::is_html_output()}
+#| label: langevitour-limb-tsne-author-proj
+
+data_limb_n <- data_limb_n |>
+  select(-type) |>
+  mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
+
+df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
+
+langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
+                         lineFrom = trimesh_data_limb$from,
+                         lineTo = trimesh_data_limb$to,
+                         group = factor(df_model_data_limb_n$type,
+                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
+                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
 ```

### 2025-05-12 paper-quollr.R

@@ -707,4 +707,88 @@ cluster_df <- read_rds("data/limb_muscles/facs_limb_muscles_cluster_df.rds")
 ## -----------------------------------------------------------------------------
 #| label: tsne-model-limb
 
+tsne_limb_obj <- fit_highd_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb, 
+  bin1 = 19, 
+  q = 0.1, 
+  benchmark_highdens = 2)
+
+df_bin_centroids_limb <- tsne_limb_obj$model_2d
+df_bin_limb <- tsne_limb_obj$model_highd
+trimesh_data_limb <- tsne_limb_obj$trimesh_data
+tsne_limb_scaled <- tsne_limb_obj$nldr_obj$scaled_nldr
+
+tsne_limb_scaled_with_cluster <- inner_join(tsne_limb_scaled, cluster_df, by = "ID") |>
+  mutate(cluster = as.character(cluster.ids))
+
+trimesh_limb <- ggplot() + 
+  geom_point(
+    data = tsne_limb_scaled_with_cluster,
+    aes(
+      x = emb1,
+      y = emb2,
+      color = cluster
+    ),
+    alpha = 0.3
+  ) +
+  geom_segment(data = trimesh_data_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
+  interior_annotation("a1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+
+
+## -----------------------------------------------------------------------------
+#| label: prep-limb-tsne-author-model-proj
+
+data_limb_n <- data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(hexID %in% df_bin_centroids_limb$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hexID),] |>
+  dplyr::select(-hexID) 
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
[…]

### 2025-05-12 paper-quollr.Rmd
Added tSNE best

@@
   q = 0.1, 
-  benchmark_highdens = 2)
+  benchmark_highdens = 5)
 
@@
   )
+
+# hex_grid <- tsne_limb_obj$hb_obj$hex_poly
+# counts_df <- tsne_limb_obj$hb_obj$std_cts
+# 
+# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hexID"))
+# 
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = bin_counts)) +
+#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed() +
+#   theme_minimal()
 ```
@@
 
+```{r}
+#| label: tsne-best-model-limb
+
+tsne_best_limb_obj <- fit_highd_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb2, 
+  bin1 = 19, 
+  q = 0.1, 
+  benchmark_highdens = 1)
+
+df_bin_centroids_limb <- tsne_best_limb_obj$model_2d
+df_bin_limb <- tsne_best_limb_obj$model_highd
+trimesh_data_limb <- tsne_best_limb_obj$trimesh_data
+tsne_limb_scaled <- tsne_best_limb_obj$nldr_obj$scaled_nldr
+
+tsne_limb_scaled_with_cluster <- inner_join(tsne_limb_scaled, cluster_df, by = "ID") |>
+  mutate(cluster = as.character(cluster.ids))
+
+trimesh_limb_best <- ggplot() + 
+  geom_point(
+    data = tsne_limb_scaled_with_cluster,
+    aes(
+      x = emb1,
+      y = emb2,
+      color = cluster
+    ),
+    alpha = 0.3
+  ) +
+  geom_segment(data = trimesh_data_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
+  interior_annotation("b1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+
+# hex_grid <- tsne_best_limb_obj$hb_obj$hex_poly
+# counts_df <- tsne_best_limb_obj$hb_obj$std_cts
+# 
+# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hexID"))
+# 
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = bin_counts)) +
+#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed() +
+#   theme_minimal()
+```
+
+```{r}
+#| label: prep-limb-tsne-best-model-proj
+
+data_limb_n <- data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(hexID %in% df_bin_centroids_limb$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
+scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
+  as_tibble()
+
+scaled_limb_data <- scaled_limb |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_limb_data_model <- scaled_limb |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r, eval=knitr::is_html_output()}
+#| label: langevitour-limb-tsne-best-proj
+
+data_limb_n <- data_limb_n |>
+  select(-type) |>
+  mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
+
+df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
+
+langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
+                         lineFrom = trimesh_data_limb$from,
+                         lineTo = trimesh_data_limb$to,
+                         group = factor(df_model_data_limb_n$type,
+                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
+                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
+```
+
+```{r}
+trimesh_limb + trimesh_limb_best +
+  plot_layout(ncol = 2)
+```
+
 # Discussion

### 2025-05-12 paper-quollr.Rmd
Find the cross reference

@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure @fig-nldrlayouts).
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \@ref(fig:nldr-layouts)).
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 
-```{r fig-nldrlayouts, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
+```{r nldr-layouts, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
 Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
@@
 
+\@ref(fig:limb-rmse)
+
 ```{r}
@@
 
-<!-- Assessing which of the 8 NLDR layouts on the limb3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
-Layout f is universally poor.
-Layouts a, c, g, h that show large separations between clusters are universally suboptimal.
-Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
-The choice of best is between layouts b and e, that have small separations between oddly shaped clusters.
-Layout e is the best choice.-->
-
-```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f)."
-#| label: fig-limb-mse
-#| out-width: 100%
+```{r limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
+Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
+Layout f is the best choice.", out.width="100%"}
 

### 2025-05-12 paper-quollr.tex
Added langevitour of the model
Added tSNE best
Find the cross reference

--- 2025-05-12 paper-quollr.tex
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \citet{fig-nldrlayouts}).
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \ref{fig:nldr-layouts}).
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-nldrlayouts-1} \caption{Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
+\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/nldr-layouts-1} \caption{Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
 Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
 Would their downstream analysis decisions vary based on the version they observe?
-Which representation most accurately represents the structure in high-dimensional space?}\label{fig:fig-nldrlayouts}
+Which representation most accurately represents the structure in high-dimensional space?}\label{fig:nldr-layouts}
 \end{figure}
@@
 \begin{verbatim}
+#> $nldr_obj
+#> $nldr_obj$scaled_nldr
+#> # A tibble: 5,000 x 3
+#>     emb1  emb2    ID
+#>    <dbl> <dbl> <int>
+#>  1 0.707 0.839     1
+#>  2 0.231 0.401     2
+#>  3 0.232 0.215     3
+#>  4 0.790 0.564     4
+#>  5 0.761 0.551     5
+#>  6 0.445 0.721     6
+#>  7 0.900 0.137     7
+#>  8 0.247 0.392     8
+#>  9 0.325 0.542     9
+#> 10 0.278 0.231    10
+#> # i 4,990 more rows
+#> 
+#> $nldr_obj$lim1
+#> [1] -14.42166  13.32655
+#> 
+#> $nldr_obj$lim2
+#> [1] -12.43687  12.32455
+#> 
+#> 
 #> $hb_obj
@@
 
+\ref{fig:limb-rmse}
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-limb-mse-1} \end{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
+Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
+Layout f is the best choice.}\label{fig:limb-rmse}
+\end{figure}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
 

### 2025-05-12 scripts/additional_functions.R

@@ -301,12 +301,12 @@ plot_mse <- function(error_df) {
 
   ggplot(error_df,
          aes(x = a1,
-             y = MSE,
+             y = sqrt(MSE),
              colour = method)) +
     geom_point(size = 0.8) +
     geom_line(linewidth = 0.3) +
-    scale_y_log10() +
-    ylab("log(MSE)") +
+    #scale_y_log10() +
+    ylab("RMSE") +
     xlab(expression(paste("binwidth (", a[1], ")"))) +
     theme_minimal() +
     theme(panel.border = element_rect(fill = 'transparent'),

### 2025-05-12 paper-quollr.Rmd
Updated the fig-label

@@
 Following their pre-processing, tSNE was performed using ten principal components.
-@fig-NLDR-variety (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout.The question is whether this accurately represents the cluster structure in the data.
+\@ref(fig:nldr-layouts) (b) is the reproduction of the published plot.
+The question is whether this accurately represents the cluster structure in the data.
 Our method can help here, and also help to provide a better $2-\text{D}$ layout, as needed.
@@
 <!-- https://figshare.com/articles/dataset/Robject_files_for_tissues_processed_by_Seurat/5821263/1 -->
-
-\@ref(fig:limb-rmse)
 
@@
   mutate(method = factor(method,
-                         levels = c("UMAP_15_min_dist_0.1","tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2", "tsne_15")))
+                         levels = c("UMAP_15_min_dist_0.1", "tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2", "tsne_15")))
 ```
@@
 
-```{r limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$).
+```{r limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data  (shown in \\@ref(fig:nldr-layouts)) is the better representation using RMSE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
@@
 
-```{r}
-#| label: data-limb
-
+```{r data-limb}
 data_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pcs_10.rds")
@@
 
-```{r}
+```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (\\@ref(fig:nldr-layouts) b) and the $2-\\text{D}$ layout selected (\\@ref(fig:nldr-layouts) f) by RMSE plot (\\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=10, fig.height=15}
+
+trimesh_limb + trimesh_limb_best +
+  trimesh_limb + trimesh_limb_best +
   trimesh_limb + trimesh_limb_best +

### 2025-05-12 paper-quollr.tex
Updated the fig-label

@@
 Following their pre-processing, tSNE was performed using ten principal components.
-\citet{fig-NLDR-variety} (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout.The question is whether this accurately represents the cluster structure in the data.
+\ref{fig:nldr-layouts} (b) is the reproduction of the published plot.
+The question is whether this accurately represents the cluster structure in the data.
 Our method can help here, and also help to provide a better \(2-\text{D}\) layout, as needed.
 
-\ref{fig:limb-rmse}
-
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data  (shown in \\ref{fig:nldr-layouts}) is the better representation using RMSE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (\\ref{fig:nldr-layouts} b) and the $2-\text{D}$ layout selected (\\ref{fig:nldr-layouts} f) by RMSE plot (\\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}
+\end{figure}
 

### 2025-05-12 paper-quollr.Rmd
Updated the MSE plot

@@
 ```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (\\@ref(fig:nldr-layouts) b) and the $2-\\text{D}$ layout selected (\\@ref(fig:nldr-layouts) f) by RMSE plot (\\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=10, fig.height=15}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=15, fig.height=10}
 
@@
   trimesh_limb + trimesh_limb_best +
-  plot_layout(ncol = 2)
+  plot_layout(nrow = 2)
 ```

### 2025-05-12 paper-quollr.R

@@ -901,10 +901,10 @@ scaled_limb_data_model <- scaled_limb |>
 #                          levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
 
 
-## ----model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (\\@ref(fig:nldr-layouts) b) and the $2-\\text{D}$ layout selected (\\@ref(fig:nldr-layouts) f) by RMSE plot (\\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters. The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=10, fig.height=15----
+## ----model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (\\@ref(fig:nldr-layouts) b) and the $2-\\text{D}$ layout selected (\\@ref(fig:nldr-layouts) f) by RMSE plot (\\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters. The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=15, fig.height=10----
 
 trimesh_limb + trimesh_limb_best +
   trimesh_limb + trimesh_limb_best +
   trimesh_limb + trimesh_limb_best +
-  plot_layout(ncol = 2)
+  plot_layout(nrow = 2)
 

### 2025-05-12 paper-quollr.Rmd
Started adding the data set table

@@
 
-Add a table of data sets that quollr have
+The `quollr` package comes with several data sets that load with the package.
+These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.
+
+```{r}
+datasets_tb <- tibble(dt = c("scurve",
+                             "scurve_umap",
+                             "scurve_umap_predict",
+                             "scurve_umap2",
+                             "scurve_umap3",
+                             "scurve_umap4",
+                             "scurve_umap5",
+                             "scurve_umap6",
+                             "scurve_model_obj",
+                             "scurve_umap_mse",
+                             "scurve_umap_mse2",
+                             "scurve_umap_mse3",
+                             "scurve_umap_mse4",
+                             "scurve_umap_mse5",
+                             "scurve_umap_mse6"), 
+                      text = c("scurve",
+                             "scurve_umap",
+                             "scurve_umap_predict",
+                             "scurve_umap2",
+                             "scurve_umap3",
+                             "scurve_umap4",
+                             "scurve_umap5",
+                             "scurve_umap6",
+                             "scurve_model_obj",
+                             "scurve_umap_mse",
+                             "scurve_umap_mse2",
+                             "scurve_umap_mse3",
+                             "scurve_umap_mse4",
+                             "scurve_umap_mse5",
+                             "scurve_umap_mse6"))
+```
+
+```{r datasets-tb-html, eval=knitr::is_html_output()}
+datasets_tb |> 
+  kable(caption = "cardinalR data sets", col.names = c("data", "explanation")) 
+```
+
+```{r datasets-tb-pdf, eval=knitr::is_latex_output()}
+datasets_tb |> 
+  kable(caption = "cquollr data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 

### 2025-05-12 paper-quollr.tex
Started adding the data set table

@@
 
-Add a table of data sets that quollr have
+The \texttt{quollr} package comes with several data sets that load with the package. These are described in Table \ref{tab:datasets-tb-pdf}.
+
+\begin{table}
+
+\caption{\label{tab:datasets-tb-pdf}cquollr data sets}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+data & explanation\\
+\midrule
+scurve & scurve\\
+scurve\_umap & scurve\_umap\\
+scurve\_umap\_predict & scurve\_umap\_predict\\
+scurve\_umap2 & scurve\_umap2\\
+scurve\_umap3 & scurve\_umap3\\
+\addlinespace
+scurve\_umap4 & scurve\_umap4\\
+scurve\_umap5 & scurve\_umap5\\
+scurve\_umap6 & scurve\_umap6\\
+scurve\_model\_obj & scurve\_model\_obj\\
+scurve\_umap\_mse & scurve\_umap\_mse\\
+\addlinespace
+scurve\_umap\_mse2 & scurve\_umap\_mse2\\
+scurve\_umap\_mse3 & scurve\_umap\_mse3\\
+scurve\_umap\_mse4 & scurve\_umap\_mse4\\
+scurve\_umap\_mse5 & scurve\_umap\_mse5\\
+scurve\_umap\_mse6 & scurve\_umap\_mse6\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-18-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-19-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-19-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-20-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
 

### 2025-05-12 paper-quollr.R

@@ -159,6 +159,51 @@ nldr1_org + nldr2_org + nldr3_org +
 package_dependencies("quollr")
 
 
+## -----------------------------------------------------------------------------
+datasets_tb <- tibble(dt = c("scurve",
+                             "scurve_umap",
+                             "scurve_umap_predict",
+                             "scurve_umap2",
+                             "scurve_umap3",
+                             "scurve_umap4",
+                             "scurve_umap5",
+                             "scurve_umap6",
+                             "scurve_model_obj",
+                             "scurve_umap_mse",
+                             "scurve_umap_mse2",
+                             "scurve_umap_mse3",
+                             "scurve_umap_mse4",
+                             "scurve_umap_mse5",
+                             "scurve_umap_mse6"), 
+                      text = c("scurve",
+                             "scurve_umap",
+                             "scurve_umap_predict",
+                             "scurve_umap2",
+                             "scurve_umap3",
+                             "scurve_umap4",
+                             "scurve_umap5",
+                             "scurve_umap6",
+                             "scurve_model_obj",
+                             "scurve_umap_mse",
+                             "scurve_umap_mse2",
+                             "scurve_umap_mse3",
+                             "scurve_umap_mse4",
+                             "scurve_umap_mse5",
+                             "scurve_umap_mse6"))
+
+
+## ----datasets-tb-html, eval=knitr::is_html_output()---------------------------
+# datasets_tb |>
+#   kable(caption = "cardinalR data sets", col.names = c("data", "explanation"))
+
+
+## ----datasets-tb-pdf, eval=knitr::is_latex_output()---------------------------
+datasets_tb |> 
+  kable(caption = "cquollr data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+
+
 ## ----echo=TRUE----------------------------------------------------------------
 fit_highd_model(
   highd_data = scurve, 

### 2025-05-12 scripts/additional_functions.R

@@ -123,177 +123,177 @@ scale_data_manual <- function(data, type_col) {
   return(data_scaled)
 }
 
-# Get projection
-
-get_projection <- function(projection, proj_scale, scaled_data,
-                           scaled_data_model, distance_df_small_edges,
-                           axis_param) {
-
-  projection_scaled <- projection * proj_scale
-  projected <- as.matrix(scaled_data) %*% projection_scaled
-
-  projected_df <- projected |>
-    tibble::as_tibble(.name_repair = "unique") |>
-    dplyr::rename(c("proj1" = "...1",
-                    "proj2" = "...2")) |>
-    dplyr::mutate(ID = dplyr::row_number())
-
-  projected_model <- as.matrix(scaled_data_model) %*% projection_scaled
-
-  projected_model_df <- projected_model |>
-    tibble::as_tibble(.name_repair = "unique") |>
-    dplyr::rename(c("proj1" = "...1",
-                    "proj2" = "...2")) |>
-    dplyr::mutate(ID = dplyr::row_number())
-
-  model_df <- dplyr::left_join(
-    distance_df_small_edges |> select(-distance),
-    projected_model_df,
-    by = c("from" = "ID"))
-
-  names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-  model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-  names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-  limits <- axis_param$limits
-  axis_scaled <- axis_param$axis_scaled
-  axis_pos_x <- axis_param$axis_pos_x
-  axis_pos_y <- axis_param$axis_pos_y
-  threshold <- axis_param$threshold
-
-  axes_obj <- gen_axes(
-    proj = projection * axis_scaled,
-    limits = limits,
-    axis_pos_x = axis_pos_x,
-    axis_pos_y = axis_pos_y,
-    axis_labels = names(scaled_data),
-    threshold = threshold)
-
-  axes <- axes_obj$axes
-  circle <- axes_obj$circle
-
-  return(list(projected_df = projected_df,
-              model_df = model_df,
-              axes = axes,
-              circle = circle))
-
-}
-
-# Plot projection
-
[…]

### 2025-05-12 paper-quollr.Rmd
Updated table

--- 2025-05-12 paper-quollr.Rmd
@@
 
-Add a table of data sets that quollr have
+The `quollr` package comes with several data sets that load with the package.
+These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.
+
+```{r}
+datasets_tb <- tibble(dt = c("scurve",
+                             "scurve_umap",
+                             "scurve_umap_predict",
+                             "scurve_umap2",
+                             "scurve_umap3",
+                             "scurve_umap4",
+                             "scurve_umap5",
+                             "scurve_umap6",
+                             "scurve_model_obj",
+                             "scurve_umap_mse",
+                             "scurve_umap_mse2",
+                             "scurve_umap_mse3",
+                             "scurve_umap_mse4",
+                             "scurve_umap_mse5",
+                             "scurve_umap_mse6"), 
+                      text = c("Simulated data to generate a S-curve with additional noise dimensions.",
+                             "The UMAP embedding with n_neighbors $46$ and min_dist $0.9$ for `S-curve`.",
+                             "scurve_umap_predict",
+                             "The UMAP embedding with n_neighbors $10$ and min_dist $0.4$ for `S-curve`.",
+                             "The UMAP embedding with n_neighbors $62$ and min_dist $0.1$ for `S-curve`.",
+                             "The UMAP embedding with n_neighbors $30$ and min_dist $0.5$ for `S-curve`.",
+                             "The UMAP embedding with n_neighbors $15$ and min_dist $0.5$ for `S-curve`.",
+                             "The UMAP embedding with n_neighbors $15$ and min_dist $0.1$ for `S-curve`.",
+                             "scurve_model_obj",
+                             "scurve_umap_mse",
+                             "scurve_umap_mse2",
+                             "scurve_umap_mse3",
+                             "scurve_umap_mse4",
+                             "scurve_umap_mse5",
+                             "scurve_umap_mse6"))
+```
+
+```{r datasets-tb-html, eval=knitr::is_html_output()}
+datasets_tb |> 
+  kable(caption = "quollr data sets", col.names = c("data", "explanation")) 
+```
+
+```{r datasets-tb-pdf, eval=knitr::is_latex_output()}
+datasets_tb |> 
+  kable(caption = "quollr data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
+  column_spec(1, width = "4cm") |>
+  column_spec(2, width = "8cm")
+```
 
@@
 This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`bin1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`benchmark_highdens`).
-The function returns an object that includes the hexagonal object, the fitted model in both $2\text{-}D$, and $p\text{-}D$, and triangular mesh.
+The function returns an object that includes the scaled NLDR object (`nldr_obj`), the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
 

### 2025-05-12 paper-quollr.tex
Updated table

@@
 
-\caption{\label{tab:datasets-tb-pdf}cquollr data sets}
+\caption{\label{tab:datasets-tb-pdf}quollr data sets}
 \centering
@@
 \midrule
-scurve & scurve\\
-scurve\_umap & scurve\_umap\\
+scurve & Simulated data to generate a S-curve with additional noise dimensions.\\
+scurve\_umap & The UMAP embedding with n\_neighbors \$46\$ and min\_dist \$0.9\$ for `S-curve`.\\
 scurve\_umap\_predict & scurve\_umap\_predict\\
-scurve\_umap2 & scurve\_umap2\\
-scurve\_umap3 & scurve\_umap3\\
+scurve\_umap2 & The UMAP embedding with n\_neighbors \$10\$ and min\_dist \$0.4\$ for `S-curve`.\\
+scurve\_umap3 & The UMAP embedding with n\_neighbors \$62\$ and min\_dist \$0.1\$ for `S-curve`.\\
 \addlinespace
-scurve\_umap4 & scurve\_umap4\\
-scurve\_umap5 & scurve\_umap5\\
-scurve\_umap6 & scurve\_umap6\\
+scurve\_umap4 & The UMAP embedding with n\_neighbors \$30\$ and min\_dist \$0.5\$ for `S-curve`.\\
+scurve\_umap5 & The UMAP embedding with n\_neighbors \$15\$ and min\_dist \$0.5\$ for `S-curve`.\\
+scurve\_umap6 & The UMAP embedding with n\_neighbors \$15\$ and min\_dist \$0.1\$ for `S-curve`.\\
 scurve\_model\_obj & scurve\_model\_obj\\
@@
 This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{bin1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{benchmark\_highdens}).
-The function returns an object that includes the hexagonal object, the fitted model in both \(2\text{-}D\), and \(p\text{-}D\), and triangular mesh.
+The function returns an object that includes the scaled NLDR object (\texttt{nldr\_obj}), the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
 

### 2025-05-12 paper-quollr.Rmd
Updated the dataset table

@@
                              "The UMAP embedding with n_neighbors $46$ and min_dist $0.9$ for `S-curve`.",
-                             "scurve_umap_predict",
+                             "The predicted UMAP embedding with n_neighbors $46$ and min_dist $0.9$ for `S-curve`.",
                              "The UMAP embedding with n_neighbors $10$ and min_dist $0.4$ for `S-curve`.",
@@
                              "The UMAP embedding with n_neighbors $15$ and min_dist $0.1$ for `S-curve`.",
-                             "scurve_model_obj",
-                             "scurve_umap_mse",
-                             "scurve_umap_mse2",
-                             "scurve_umap_mse3",
-                             "scurve_umap_mse4",
-                             "scurve_umap_mse5",
-                             "scurve_umap_mse6"))
+                             "The object contains the scaled NLDR object (`nldr_obj`), the hexagonal object (`hb_obj`), the fitted model in both $2\\text{-}D$ (`model_2d`), and $p\\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).",
+                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap`.",
+                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap2`.",
+                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap3`.",
+                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap4`.",
+                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap5`.",
+                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap6`."))
 ```

### 2025-05-12 paper-quollr.tex
Updated the dataset table

@@
 scurve\_umap & The UMAP embedding with n\_neighbors \$46\$ and min\_dist \$0.9\$ for `S-curve`.\\
-scurve\_umap\_predict & scurve\_umap\_predict\\
+scurve\_umap\_predict & The predicted UMAP embedding with n\_neighbors \$46\$ and min\_dist \$0.9\$ for `S-curve`.\\
 scurve\_umap2 & The UMAP embedding with n\_neighbors \$10\$ and min\_dist \$0.4\$ for `S-curve`.\\
@@
 scurve\_umap6 & The UMAP embedding with n\_neighbors \$15\$ and min\_dist \$0.1\$ for `S-curve`.\\
-scurve\_model\_obj & scurve\_model\_obj\\
-scurve\_umap\_mse & scurve\_umap\_mse\\
+scurve\_model\_obj & The object contains the scaled NLDR object (`nldr\_obj`), the hexagonal object (`hb\_obj`), the fitted model in both \$2\textbackslash{}text\{-\}D\$ (`model\_2d`), and \$p\textbackslash{}text\{-\}D\$ (`model\_highd`), and triangular mesh (`trimesh\_data`).\\
+scurve\_umap\_mse & The MSE corresponding to different binwidths for the model generated using `scurve\_umap`.\\
 \addlinespace
-scurve\_umap\_mse2 & scurve\_umap\_mse2\\
-scurve\_umap\_mse3 & scurve\_umap\_mse3\\
-scurve\_umap\_mse4 & scurve\_umap\_mse4\\
-scurve\_umap\_mse5 & scurve\_umap\_mse5\\
-scurve\_umap\_mse6 & scurve\_umap\_mse6\\
+scurve\_umap\_mse2 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap2`.\\
+scurve\_umap\_mse3 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap3`.\\
+scurve\_umap\_mse4 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap4`.\\
+scurve\_umap\_mse5 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap5`.\\
+scurve\_umap\_mse6 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap6`.\\
 \bottomrule

### 2025-05-12 paper-quollr.Rmd
Updated computing grid configurations

@@
 The function `calc_bins_y()` is used for this purpose.
-This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (`bin1`), and the buffer amount as a proportion.
+This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (`bin1`), and the buffer amount as a proportion (`q`).
 

### 2025-05-12 paper-quollr.tex
Updated computing grid configurations

@@
 The function \texttt{calc\_bins\_y()} is used for this purpose.
-This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion.
+This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion (\texttt{q}).
 

### 2025-05-12 paper-quollr.Rmd
Added steps for hexagonal binning

@@
 
+If the hexagonal binning process is run separately, it involves several steps: (i) generating all possible centroids in a hexagonal grid, (ii) creating the coordinates of the hexagons, (iii) assigning data points to their respective hexagons, (iv) computing the standardized number of points within each hexagon, and (v) mapping the points to their corresponding hexagonal bins.
+
+#### Generating all possible centroids in a hexagonal grid
+
 ```{r, echo=TRUE}
@@
 
+#### Creating the coordinates of the hexagons
+
 ```{r, echo=TRUE}
@@
 
+#### Assigning data points to their respective hexagons
+
 ```{r, echo=TRUE}
@@
 
+#### Computing the standardized number of points within each hexagon
+
 ```{r, echo=TRUE}
@@
 ```
+
+#### Mapping the points to their corresponding hexagonal bins
 

### 2025-05-12 paper-quollr.tex
Added steps for hexagonal binning

@@
 
+If the hexagonal binning process is run separately, it involves several steps: (i) generating all possible centroids in a hexagonal grid, (ii) creating the coordinates of the hexagons, (iii) assigning data points to their respective hexagons, (iv) computing the standardized number of points within each hexagon, and (v) mapping the points to their corresponding hexagonal bins.
+
+\paragraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
+
 \begin{verbatim}
@@
 
+\paragraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
+
 \begin{verbatim}
@@
 \end{verbatim}
+
+\paragraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
 
@@
 
+\paragraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
+
 \begin{verbatim}
@@
 \end{verbatim}
+
+\paragraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
 

### 2025-05-12 paper-quollr.Rmd
Added generate hexgonal centroids

@@
 
+The `gen_centroids()` function calculates the centroids of a hexagonal grid.
+
+The coordinate limits of the embedding (`lim1` and `lim2`) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
+The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagonal rows (`bin2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`bin1`) and buffer (`q`).
+
+Then, the centroids are computed iteratively.
+The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width.
+Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
+Vertical spacing (`a2`) is derived by $\sqrt{3}/2 \times a_1$.
+
+The y-coordinates for each row are similarly calculated, and paired with the x-coordinates based on whether the total number of rows is even or odd.
+In the case of an odd number of rows, the final row uses only the odd-row x-coordinates to maintain the alternating pattern.
+
+Finally, a tibble is returned containing a unique hexagon ID (`hexID`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2-\text{D}$ space.
+
 ```{r, echo=TRUE}

### 2025-05-12 paper-quollr.tex
Added generate hexgonal centroids

@@
 \paragraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
+
+The \texttt{gen\_centroids()} function calculates the centroids of a hexagonal grid.
+
+The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
+The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagonal rows (\texttt{bin2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{bin1}) and buffer (\texttt{q}).
+
+Then, the centroids are computed iteratively.
+The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width.
+Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
+Vertical spacing (\texttt{a2}) is derived by \(\sqrt{3}/2 \times a_1\).
+
+The y-coordinates for each row are similarly calculated, and paired with the x-coordinates based on whether the total number of rows is even or odd.
+In the case of an odd number of rows, the final row uses only the odd-row x-coordinates to maintain the alternating pattern.
+
+Finally, a tibble is returned containing a unique hexagon ID (\texttt{hexID}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2-\text{D}\) space.
 

### 2025-05-12 paper-quollr.Rmd
Add the section Creating the coordinates of the hexagons

@@
 
+Following the generation of hexagonal centroids, the `gen_hex_coord()` function constructs the coordinates of each hexagonal bin by defining its six polygonal vertices.
+These coordinates are used to visualize the hexagonal tessellation.
+
+Each hexagon is defined relative to its centroid $(C_x, C_y)$, with six vertices positioned equidistantly around the center.
+The function first verifies the presence of the required hexagon width parameter `a1`.
+This width determines the horizontal spacing.
+
+Two derived constants are calculated to define the relative distances to the vertices.
+The horizontal and vertical offset is defined as $dx = \frac{a_1}{2}$, and $dy = \frac{a_1}{\sqrt{3}}$ repectively.
+A vertical spacing factor $v = \frac{a_1}{2\sqrt{3}}$ refines vertical placement in staggered rows.
+
+With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid.
+These offsets form two vectors (`x_add_factor` and `y_add_factor`) corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
+
+For each centroid, six vertices are computed and assigned a polygon ID (`hex_poly_id`) corresponding to the centroid’s `hexID`.
+These points are aggregated into a tibble that includes the polygon ID (`hex_poly_id`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.
+
 ```{r, echo=TRUE}

### 2025-05-12 paper-quollr.tex
Add the section Creating the coordinates of the hexagons

@@
 \paragraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
+
+Following the generation of hexagonal centroids, the \texttt{gen\_hex\_coord()} function constructs the coordinates of each hexagonal bin by defining its six polygonal vertices.
+These coordinates are used to visualize the hexagonal tessellation.
+
+Each hexagon is defined relative to its centroid \((C_x, C_y)\), with six vertices positioned equidistantly around the center.
+The function first verifies the presence of the required hexagon width parameter \texttt{a1}.
+This width determines the horizontal spacing.
+
+Two derived constants are calculated to define the relative distances to the vertices.
+The horizontal and vertical offset is defined as \(dx = \frac{a_1}{2}\), and \(dy = \frac{a_1}{\sqrt{3}}\) repectively.
+A vertical spacing factor \(v = \frac{a_1}{2\sqrt{3}}\) refines vertical placement in staggered rows.
+
+With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid.
+These offsets form two vectors (\texttt{x\_add\_factor} and \texttt{y\_add\_factor}) corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
+
+For each centroid, six vertices are computed and assigned a polygon ID (\texttt{hex\_poly\_id}) corresponding to the centroid's \texttt{hexID}.
+These points are aggregated into a tibble that includes the polygon ID (\texttt{hex\_poly\_id}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
 

### 2025-05-12 paper-quollr.Rmd
Added Assigning data points to their respective hexagons

@@
 
+After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
+The `assign_data()` function performs this assignment by calculating the $2-\text{D}$ Euclidean distance between each point in the $2-\text{D}$ embedding and all hexagon centroids.
+
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2-\text{D}$ layout.
+It then selects the corresponding x and y coordinates of each hexagon’s centroid.
+
+Both the embedding coordinates and the centroid coordinates are converted to matrices to facilitate distance computations.
+The function uses the `proxy::dist()` method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
+For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon—and assigns the corresponding hexagon ID (`hexID`) to the point.
+
+The result is a data frame of the scaled $2-\text{D}$ embedding with an additional `hexID` column, indicating the hexagonal bin to which each point belongs.
+
 ```{r, echo=TRUE}

### 2025-05-12 paper-quollr.tex
Added Assigning data points to their respective hexagons

@@
 
+After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
+The \texttt{assign\_data()} function performs this assignment by calculating the \(2-\text{D}\) Euclidean distance between each point in the \(2-\text{D}\) embedding and all hexagon centroids.
+
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \(2-\text{D}\) layout.
+It then selects the corresponding x and y coordinates of each hexagon's centroid.
+
+Both the embedding coordinates and the centroid coordinates are converted to matrices to facilitate distance computations.
+The function uses the \texttt{proxy::dist()} method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
+For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon---and assigns the corresponding hexagon ID (\texttt{hexID}) to the point.
+
+The result is a data frame of the scaled \(2-\text{D}\) embedding with an additional \texttt{hexID} column, indicating the hexagonal bin to which each point belongs.
+
 \begin{verbatim}

### 2025-05-12 paper-quollr.Rmd
Added Mapping the points to their corresponding hexagonal bins

@@
 
+The `compute_std_counts()` function calculates both the raw and standardized counts for each hexagon.
+
+The function begins by grouping the data by `hexID` and counting the number of NLDR points falling within each bin.
+These raw counts are stored as `bin_counts`.
+To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin’s count by the maximum count across all bins.
+This yields a normalized metric, `std_counts`, ranging from 0 to 1.
+
 ```{r, echo=TRUE}
@@
 #### Mapping the points to their corresponding hexagonal bins
+
+The `find_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon in the NLDR space.
+
+The function first groups the input data by `hexID`, which represents the hexagon label associated with each point in the $2-\text{D}$ layout.
+Within each group, it collects the `ID`s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
+The resulting column, `pts_list`, contains all point identifiers associated with that hexagon.
 

### 2025-05-12 paper-quollr.tex
Added Mapping the points to their corresponding hexagonal bins

@@
 
+The \texttt{compute\_std\_counts()} function calculates both the raw and standardized counts for each hexagon.
+
+The function begins by grouping the data by \texttt{hexID} and counting the number of NLDR points falling within each bin.
+These raw counts are stored as \texttt{bin\_counts}.
+To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin's count by the maximum count across all bins.
+This yields a normalized metric, \texttt{std\_counts}, ranging from 0 to 1.
+
 \begin{verbatim}
@@
 \paragraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
+
+The \texttt{find\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon in the NLDR space.
+
+The function first groups the input data by \texttt{hexID}, which represents the hexagon label associated with each point in the \(2-\text{D}\) layout.
+Within each group, it collects the \texttt{ID}s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
+The resulting column, \texttt{pts\_list}, contains all point identifiers associated with that hexagon.
 

### 2025-05-12 paper-quollr.Rmd
Added Obtaining bin centroids

@@
 
+The `extract_hexbin_centroids()` function combines hexagonal bin coordinates, raw and standardized counts within each hexagons.
+
+This function begins by arranging the `counts_data` by `hexID` to ensure consistent ordering.
+It then performs a full join with `centroids_data`, aligning hexagon IDs between the two datasets to incorporate both hexagonal bin centroids and count metrics.
+After merging, the function handles missing values in the count columns: any `NA` values in `std_counts` or `bin_counts` are replaced with zeros.
+This ensures that hexagons with no assigned data points are retained in the output, with zero values for count-related fields.
+The resulting data frame contains the full set of hexagon centroids along with associated bin counts and standardized counts.
+
 ```{r, echo=TRUE}
@@
 ```
-
-In some cases, distant centroids may be connected, resulting in long line segments that can affect the smoothness of the $2\text{-}D$ representation.
-To address this issue, the `find_lg_benchmark()` function is used.
-This function computes a threshold based on the distances of line segments, determining when long edges should be removed.
 

### 2025-05-12 paper-quollr.tex
Added Obtaining bin centroids

@@
 
+The \texttt{extract\_hexbin\_centroids()} function combines hexagonal bin coordinates, raw and standardized counts within each hexagons.
+
+This function begins by arranging the \texttt{counts\_data} by \texttt{hexID} to ensure consistent ordering.
+It then performs a full join with \texttt{centroids\_data}, aligning hexagon IDs between the two datasets to incorporate both hexagonal bin centroids and count metrics.
+After merging, the function handles missing values in the count columns: any \texttt{NA} values in \texttt{std\_counts} or \texttt{bin\_counts} are replaced with zeros.
+This ensures that hexagons with no assigned data points are retained in the output, with zero values for count-related fields.
+The resulting data frame contains the full set of hexagon centroids along with associated bin counts and standardized counts.
+
 \begin{verbatim}
@@
 \end{verbatim}
-
-In some cases, distant centroids may be connected, resulting in long line segments that can affect the smoothness of the \(2\text{-}D\) representation.
-To address this issue, the \texttt{find\_lg\_benchmark()} function is used.
-This function computes a threshold based on the distances of line segments, determining when long edges should be removed.
 

### 2025-05-12 paper-quollr.Rmd
Added triangulation section

@@
 
-To indicate neighbors, the `tri_bin_centroids()` function is used to triangulate bin centroids.
-Following this, `gen_edges()` function computes the line segments that connect neighboring bins by providing the triangulated data.
-This results the coordinates that generate the connecting lines.
+To represent the neighborhood structure of hexagonal bins in a $2-\text{D}$ layout, we employ Delaunay triangulation on the centroids of hexagons.
+This geometric approach is used to infer which bins are considered neighbors.
+
+The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `tripack::tri.mesh()` function.
+This triangulation forms the structural basis for identifying adjacent bins.
 
@@
   )
-
+```
+
+The `gen_edges()` function uses this triangulation object to extract line segments between neighboring bins.
+It constructs a unique set of bin-to-bin connections by identifying the triangle edges and filtering duplicate or reversed links.
+Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function `calc_2d_dist()`.
+Only edges within a hexagon’s neighborhood radius (based on the hexagon side length `a1`) are retained.
+
+```{r, echo=TRUE}
 trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1)
+
+trimesh
+```
+
+The `update_trimesh_index()` function re-indexes the node IDs to ensure that edge identifiers are sequentially numbered and consistent with downstream analysis.
+
+```{r, echo=TRUE}
+trimesh <- update_trimesh_index(trimesh_data = trimesh)
+
 trimesh

### 2025-05-12 paper-quollr.tex
Added triangulation section

@@
 
-To indicate neighbors, the \texttt{tri\_bin\_centroids()} function is used to triangulate bin centroids.
-Following this, \texttt{gen\_edges()} function computes the line segments that connect neighboring bins by providing the triangulated data.
-This results the coordinates that generate the connecting lines.
+To represent the neighborhood structure of hexagonal bins in a \(2-\text{D}\) layout, we employ Delaunay triangulation on the centroids of hexagons.
+This geometric approach is used to infer which bins are considered neighbors.
+
+The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{tripack::tri.mesh()} function.
+This triangulation forms the structural basis for identifying adjacent bins.
 
@@
   )
-
+\end{verbatim}
+
+The \texttt{gen\_edges()} function uses this triangulation object to extract line segments between neighboring bins.
+It constructs a unique set of bin-to-bin connections by identifying the triangle edges and filtering duplicate or reversed links.
+Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function \texttt{calc\_2d\_dist()}.
+Only edges within a hexagon's neighborhood radius (based on the hexagon side length \texttt{a1}) are retained.
+
+\begin{verbatim}
 trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1)
+
+trimesh
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 653 x 8
+#>     from    to  x_from  y_from    x_to    y_to from_count to_count
+#>    <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
+#>  1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
+#>  2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
+#>  3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
+#>  4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
+#>  5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
+#>  6    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
+#>  7    31    46 -0.1     0.0550 -0.0584  0.127           0        0
+#>  8    32    47 -0.0167  0.0550  0.0249  0.127           0        0
+#>  9    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
+#> 10     4    18  0.150  -0.0892  0.108  -0.0171          0        0
+#> # i 643 more rows
+\end{verbatim}
+
+The \texttt{update\_trimesh\_index()} function re-indexes the node IDs to ensure that edge identifiers are sequentially numbered and consistent with downstream analysis.
+
+\begin{verbatim}
+trimesh <- update_trimesh_index(trimesh_data = trimesh)
+
 trimesh
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-19-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-21-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-20-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1}
 

### 2025-05-12 paper-quollr.R

@@ -452,7 +452,16 @@ tr_object <- tri_bin_centroids(
   centroids_data = df_bin_centroids
   )
 
+
+## ----echo=TRUE----------------------------------------------------------------
 trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1)
+
+trimesh
+
+
+## ----echo=TRUE----------------------------------------------------------------
+trimesh <- update_trimesh_index(trimesh_data = trimesh)
+
 trimesh
 
 

### 2025-05-12 paper-quollr.Rmd
Added remove low-density hexagon section

@@
 
-### Remove low-density hexagons
-
-In certain scenarios, hexagonal bins may contain a few number of points.
-To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
+### Identifying and removing low-density hexagons
+
+Not all hexagons contain meaningful information.
+Some may have very few or no data points due to the sparsity or shape of the underlying structure.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or "holes" in the $2-\text{D}$ structure, potentially disrupting the continuity of the representation.
+
+To address this, we propose a more nuanced method that evaluates each hexagon not only based on its own density, but also in the context of its immediate neighbors.
+The `find_low_dens_hex()` function identifies hexagonal bins with insufficient local support by calculating the average standardized count across their six neighboring bins.
+If this mean neighborhood density is below a user-defined threshold (e.g., 0.05), the hexagon is flagged for removal.
+
 The `find_low_dens_hex()` function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
+The `find_low_dens_hex()` function relies on a helper, `compute_mean_density_hex()`, which iterates over all hexagons and computes the average density across neighbors based on their `hexID` and a defined number of bins along the x-axis (`bin1`).
+The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.
 
@@
 
-The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each hexgaonal bin to represent bin centroids.
+The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each hexagonal bin to represent bin centroids.
 This transformation is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).

### 2025-05-12 paper-quollr.tex
Added remove low-density hexagon section

@@
 
-\subsubsection{Remove low-density hexagons}\label{remove-low-density-hexagons}
-
-In certain scenarios, hexagonal bins may contain a few number of points.
-To ensure comprehensive coverage of NLDR data, it is important to select hexagonal bins with a suitable number of data points.
+\subsubsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
+
+Not all hexagons contain meaningful information.
+Some may have very few or no data points due to the sparsity or shape of the underlying structure.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the \(2-\text{D}\) structure, potentially disrupting the continuity of the representation.
+
+To address this, we propose a more nuanced method that evaluates each hexagon not only based on its own density, but also in the context of its immediate neighbors.
+The \texttt{find\_low\_dens\_hex()} function identifies hexagonal bins with insufficient local support by calculating the average standardized count across their six neighboring bins.
+If this mean neighborhood density is below a user-defined threshold (e.g., 0.05), the hexagon is flagged for removal.
+
 The \texttt{find\_low\_dens\_hex()} function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-Users can initially identify low-density hexagons and then use this function to evaluate how removing them might affect the model fit by examining their neighbors.
+The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their \texttt{hexID} and a defined number of bins along the x-axis (\texttt{bin1}).
+The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.
 
@@
 
-The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each hexgaonal bin to represent bin centroids.
+The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each hexagonal bin to represent bin centroids.
 This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).

### 2025-05-13 paper-quollr.Rmd
Added descriptions for glance function

@@
 
+This function begins by renaming the columns of the input `model_highd` data frame to avoid naming conflicts during subsequent joins.
+It then uses the `predict_emb()` function to assign each point in the high-dimensional dataset to a corresponding bin in the 2D model, producing a prediction data frame that contains both the predicted bin assignment (`pred_hb_id`) and the original observation `ID`.
+
+The function joins this prediction output with both the high-dimensional model (to get mean bin coordinates in the original space) and the original high-dimensional data (to retrieve true coordinates).
+It then calculates squared differences between the true and predicted high-dimensional coordinates for each dimension, storing these as `error_square_x1`, `error_square_x2`, ..., up to the dimensionality of the data.
+
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the Mean squared error (MSE) which is the average of the total squared error per point.
+
+These metrics are returned in a tibble as `Error` (absolute error) and `MSE` (mean squared error).
+
 ```{r, echo=TRUE}
@@
 
-Users can pass data to `augment()` via either the `training_data` argument or the `newdata` argument.
-If data is passed to the `training_data` argument, it must be exactly the data that was used to fit the model.
-Alternatively, datasets can be passed to `newdata` to augment data that was not used during model fitting.
+Users can pass data to `augment()` via either the `highd_data` argument.
 This requires that at least all predictor variable columns used to fit the model are present.
@@
 ```
+
+### $2\text{-}D$ projection
 

### 2025-05-13 paper-quollr.tex
Added descriptions for glance function

@@
 
+This function begins by renaming the columns of the input \texttt{model\_highd} data frame to avoid naming conflicts during subsequent joins.
+It then uses the \texttt{predict\_emb()} function to assign each point in the high-dimensional dataset to a corresponding bin in the 2D model, producing a prediction data frame that contains both the predicted bin assignment (\texttt{pred\_hb\_id}) and the original observation \texttt{ID}.
+
+The function joins this prediction output with both the high-dimensional model (to get mean bin coordinates in the original space) and the original high-dimensional data (to retrieve true coordinates).
+It then calculates squared differences between the true and predicted high-dimensional coordinates for each dimension, storing these as \texttt{error\_square\_x1}, \texttt{error\_square\_x2}, \ldots, up to the dimensionality of the data.
+
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the Mean squared error (MSE) which is the average of the total squared error per point.
+
+These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{MSE} (mean squared error).
+
 \begin{verbatim}
@@
 
+\subsubsection{\texorpdfstring{\(2\text{-}D\) projection}{2\textbackslash text\{-\}D projection}}\label{text-d-projection}
+
 \subsubsection{Link plots}\label{link-plots}

### 2025-05-13 paper-quollr.Rmd
Added glance description

@@
 Furthermore, `augment()` accepts $2\text{-}D$ and $p\text{-}D$ model points, and the $p\text{-}D$ data and adds information about each observation in the data set.
-Most commonly, this includes predicted values, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error.
-
-Users can pass data to `augment()` via either the `highd_data` argument.
-This requires that at least all predictor variable columns used to fit the model are present.
-If the original outcome variable used to fit the model is not included in `newdata`, then no corresponding column will be included in the output.
-
+
+The function starts by renaming columns in the `model_highd` data frame to avoid conflicts.
+It then predicts the high-dimensional bin assignments for each point using the `predict_emb()` function, mapping each observation to its nearest $2-\text{D}$ bin (`pred_hb_id`).
+This prediction is joined with the mean high-dimensional coordinates of each bin from the model and with the original high-dimensional data.
+
+Next, the function computes residuals between each original coordinate (`x1`, `x2`, ..., `xp`) and the corresponding modeled coordinate (`model_high_d_x1`, ..., `model_high_d_xp`) across all dimensions.
+It calculates both squared errors and absolute errors per dimension.
+These are used to compute two aggregate diagnostic measures per point.
+First, the `row_wise_total_error` which is the total squared error across all dimensions, and the `row_wise_abs_error` which is the total absolute error across all dimensions.
+
+The final output is a data frame that combines the original IDs, high-dimensional coordinates, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
 The augmented dataset is always returned as a `tibble::tibble` with the same number of rows as the passed dataset.

### 2025-05-13 paper-quollr.tex
Added glance description

@@
 Furthermore, \texttt{augment()} accepts \(2\text{-}D\) and \(p\text{-}D\) model points, and the \(p\text{-}D\) data and adds information about each observation in the data set.
-Most commonly, this includes predicted values, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error.
-
-Users can pass data to \texttt{augment()} via either the \texttt{training\_data} argument or the \texttt{newdata} argument.
-If data is passed to the \texttt{training\_data} argument, it must be exactly the data that was used to fit the model.
-Alternatively, datasets can be passed to \texttt{newdata} to augment data that was not used during model fitting.
-This requires that at least all predictor variable columns used to fit the model are present.
-If the original outcome variable used to fit the model is not included in \texttt{newdata}, then no corresponding column will be included in the output.
-
+
+The function starts by renaming columns in the \texttt{model\_highd} data frame to avoid conflicts.
+It then predicts the high-dimensional bin assignments for each point using the \texttt{predict\_emb()} function, mapping each observation to its nearest \(2-\text{D}\) bin (\texttt{pred\_hb\_id}).
+This prediction is joined with the mean high-dimensional coordinates of each bin from the model and with the original high-dimensional data.
+
+Next, the function computes residuals between each original coordinate (\texttt{x1}, \texttt{x2}, \ldots, \texttt{xp}) and the corresponding modeled coordinate (\texttt{model\_high\_d\_x1}, \ldots, \texttt{model\_high\_d\_xp}) across all dimensions.
+It calculates both squared errors and absolute errors per dimension.
+These are used to compute two aggregate diagnostic measures per point.
+First, the \texttt{row\_wise\_total\_error} which is the total squared error across all dimensions, and the \texttt{row\_wise\_abs\_error} which is the total absolute error across all dimensions.
+
+The final output is a data frame that combines the original IDs, high-dimensional coordinates, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
 The augmented dataset is always returned as a \texttt{tibble::tibble} with the same number of rows as the passed dataset.

### 2025-05-13 paper-quollr.Rmd
Added intro for the visualization

@@
 
-The package provides five basic visualizations which includes one to visualize the full hexagonal grid in $2\text{-}D$, three visualizations related to the $2\text{-}D$ model (static visualizations), and one related to the $p\text{-}D$ model (dynamic visualization).
+The package offers several $2-\text{D}$ visualizations, including:
+
+- A full hexagonal grid,
+- A hexagonal grid that matches the data,
+- A full grid based on centroid triangulation,
+- A centroid triangulation grid that aligns with the data, 
+- A triangular mesh for any provided set of points.
+
+The generated p-D model, overlaid with the data, can also be visualized using `show_langevitour`.
+Additionally, it features a function for visualizing the $2-\text{D}$ projection of the fitted model overlaid on the data, called `plot_proj`.
+
+Furthermore, there are two interactive plots, `show_link_plots` and `show_error_link_plots`, which are designed to help diagnose the model.
+
 Each visualization can be generated using its respective function, as described in this section.
@@
 
-## Find the most appropriate fit
+# Find the most appropriate fit
 

### 2025-05-13 paper-quollr.tex
Added intro for the visualization

@@
 
-The package provides five basic visualizations which includes one to visualize the full hexagonal grid in \(2\text{-}D\), three visualizations related to the \(2\text{-}D\) model (static visualizations), and one related to the \(p\text{-}D\) model (dynamic visualization).
+The package offers several \(2-\text{D}\) visualizations, including:
+
+\begin{itemize}
+\tightlist
+\item
+  A full hexagonal grid,
+\item
+  A hexagonal grid that matches the data,
+\item
+  A full grid based on centroid triangulation,
+\item
+  A centroid triangulation grid that aligns with the data,
+\item
+  A triangular mesh for any provided set of points.
+\end{itemize}
+
+The generated p-D model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
+Additionally, it features a function for visualizing the \(2-\text{D}\) projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
+
+Furthermore, there are two interactive plots, \texttt{show\_link\_plots} and \texttt{show\_error\_link\_plots}, which are designed to help diagnose the model.
+
 Each visualization can be generated using its respective function, as described in this section.
@@
 
-\subsection{Find the most appropriate fit}\label{find-the-most-appropriate-fit}
+\section{Find the most appropriate fit}\label{find-the-most-appropriate-fit}
 

### 2025-05-13 paper-quollr.Rmd
Added hexgonal grid

@@
 
-### $2\text{-}D$ model visualization
-
-The `geom_hexgrid()` function is used to plot the hexagonal grid from the provided centroid data set.
+### Hexagonal grid
+
+The `geom_hexgrid()` function introduces a custom `ggplot2` layer designed for visualizing hexagonal grid on a provided set of bin centroids.
+
+To display the complete grid, users should supply all available bin centroids.
 
@@
 ```
+
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+
+```{r, echo=TRUE}
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::filter(bin_counts > 10)
+
+ggplot() + 
+  geom_hexgrid(
+    data = df_bin_centroids, 
+    aes(x = c_x, y = c_y)
+    ) 
+```
+
+### Triangular mesh
 

### 2025-05-13 paper-quollr.tex
Added hexgonal grid

@@
 
-\subsubsection{\texorpdfstring{\(2\text{-}D\) model visualization}{2\textbackslash text\{-\}D model visualization}}\label{text-d-model-visualization}
-
-The \texttt{geom\_hexgrid()} function is used to plot the hexagonal grid from the provided centroid data set.
+\subsubsection{Hexagonal grid}\label{hexagonal-grid}
+
+The \texttt{geom\_hexgrid()} function introduces a custom \texttt{ggplot2} layer designed for visualizing hexagonal grid on a provided set of bin centroids.
+
+To display the complete grid, users should supply all available bin centroids.
 
@@
 \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-21-1}
+
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+
+\begin{verbatim}
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::filter(bin_counts > 10)
+
+ggplot() + 
+  geom_hexgrid(
+    data = df_bin_centroids, 
+    aes(x = c_x, y = c_y)
+    ) 
+\end{verbatim}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
+
+\subsubsection{Triangular mesh}\label{triangular-mesh}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}
 

### 2025-05-13 paper-quollr.Rmd
Visualize the triangular mesh

@@
 
-To visualize the $2\text{-}D$ model, mainly three functions are used.
-As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`.
-After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges.
-As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively.
-Following this, `vis_rmlg_mesh()` is used to visualize smoothed $2\text{-}D$ model which is the $2\text{-}D$ model after removing the long edges (see Figure \@ref(fig:mesh-plots)c).
-In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in $2\text{-}D$.
-Using small value of `benchmark_value`, will produce a triangular mesh with missing data structure; for example `benchmark_value = 0.3` shows two clusters rather than continuous structure, while `benchmark_value = 0.8` creates long edges and mislead the data structure in $p\text{-}D$ space.
+The `geom_trimesh()` function introduces a custom `ggplot2` layer designed for visualizing $2\text{-}D$ wireframe on a provided set of bin centroids.
+
+To display the complete wireframe, users should supply all available bin centroids.
 
@@
     data = hb_obj$centroids, 
+    aes(x = c_x, y = c_y)
+    ) 
+```
+
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+
+```{r, echo=TRUE}
+ggplot() + 
+  geom_trimesh(
+    data = df_bin_centroids, 
     aes(x = c_x, y = c_y)

### 2025-05-13 paper-quollr.tex
Visualize the triangular mesh

@@
 
-To visualize the \(2\text{-}D\) model, mainly three functions are used.
-As shown in Figure \ref{fig:mesh-plots}a, \texttt{geom\_trimesh()} to visualize the triangular mesh by adding a new layer to \texttt{ggplot()}.
-After identifying benchmark value to remove long edge, \texttt{vis\_lg\_mesh()} is used to visualize the triangular mesh by coloring the small and long edges.
-As shown in Figure \ref{fig:mesh-plots}b, the small and long edges are colored by black and red respectively.
-Following this, \texttt{vis\_rmlg\_mesh()} is used to visualize smoothed \(2\text{-}D\) model which is the \(2\text{-}D\) model after removing the long edges (see Figure \ref{fig:mesh-plots}c).
-In \texttt{vis\_lg\_mesh()} and \texttt{vis\_rmlg\_mesh()}, \texttt{benchmark\_value} argument controls the edge removal in \(2\text{-}D\).
-Using small value of \texttt{benchmark\_value}, will produce a triangular mesh with missing data structure; for example \texttt{benchmark\_value\ =\ 0.3} shows two clusters rather than continuous structure, while \texttt{benchmark\_value\ =\ 0.8} creates long edges and mislead the data structure in \(p\text{-}D\) space.
+The \texttt{geom\_trimesh()} function introduces a custom \texttt{ggplot2} layer designed for visualizing \(2\text{-}D\) wireframe on a provided set of bin centroids.
+
+To display the complete wireframe, users should supply all available bin centroids.
 
@@
 \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
+
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+
+\begin{verbatim}
+ggplot() + 
+  geom_trimesh(
+    data = df_bin_centroids, 
+    aes(x = c_x, y = c_y)
+    ) 
+\end{verbatim}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-32-1}
 

### 2025-05-13 paper-quollr.Rmd
Added low-density removal to front

@@
 
+For simplicity, we remove low-density hexagons using a threshold of $10$.
+
+```{r, echo=TRUE}
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::filter(bin_counts > 10)
+
+trimesh <- trimesh |>
+  dplyr::filter(from_count > 10,
+                to_count > 10)
+
+trimesh <- update_trimesh_index(trimesh)
+
+```
+
 ## Lifting the model into high dimensions
@@
 ```{r, echo=TRUE}
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
-
 ggplot() + 
@@
 ```{r, echo=TRUE}
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
 
@@
 ```{r, echo=TRUE, eval=knitr::is_html_output()}
-trimesh <- trimesh |>
-  dplyr::filter(from_count > 10,
-                to_count > 10)
-
-trimesh <- update_trimesh_index(trimesh)
-
 show_langevitour(

### 2025-05-13 paper-quollr.tex
Added low-density removal to front

@@
 
+For simplicity, we remove low-density hexagons using a threshold of \(10\).
+
+\begin{verbatim}
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::filter(bin_counts > 10)
+
+trimesh <- trimesh |>
+  dplyr::filter(from_count > 10,
+                to_count > 10)
+
+trimesh <- update_trimesh_index(trimesh)
+\end{verbatim}
+
 \subsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-21-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
 
@@
 \begin{verbatim}
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
-
 ggplot() + 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
 
@@
 \begin{verbatim}
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
-
 df_exe <- comb_data_model(
@@
 \begin{verbatim}
-trimesh <- trimesh |>
-  dplyr::filter(from_count > 10,
-                to_count > 10)
-
-trimesh <- update_trimesh_index(trimesh)
-
 show_langevitour(
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-32-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-33-1}
 

### 2025-05-13 paper-quollr.R

@@ -475,6 +475,18 @@ find_low_dens_hex(
 
 
 ## ----echo=TRUE----------------------------------------------------------------
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::filter(bin_counts > 10)
+
+trimesh <- trimesh |>
+  dplyr::filter(from_count > 10,
+                to_count > 10)
+
+trimesh <- update_trimesh_index(trimesh)
+
+
+
+## ----echo=TRUE----------------------------------------------------------------
 df_bin <- avg_highd_data(
   highd_data = scurve, 
   scaled_nldr_hexid = hb_obj$data_hb_id
@@ -518,9 +530,6 @@ ggplot() +
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
-
 ggplot() + 
   geom_hexgrid(
     data = df_bin_centroids, 
@@ -545,8 +554,6 @@ ggplot() +
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
 
 df_exe <- comb_data_model(
   highd_data = scurve, 
@@ -558,12 +565,6 @@ df_exe
 
 
 ## ----echo=TRUE, eval=knitr::is_html_output()----------------------------------
-# trimesh <- trimesh |>
-#   dplyr::filter(from_count > 10,
-#                 to_count > 10)
-# 
-# trimesh <- update_trimesh_index(trimesh)
-# 
 # show_langevitour(
 #   point_data = df_exe,
 #   edge_data = trimesh

### 2025-05-13 paper-quollr.Rmd
Added high-d model vis

@@
 
-Displaying the $p\text{-}D$ model overlaid on the data is done using the function `show_langevitour()`.
-This visualization is helpful for visually evaluating how well the model fits the data.
-The function requires several arguments: data along with their corresponding hexagonal bin ID, $2\text{-}D$ and $p\text{-}D$ model points, the threshold for removing long edges, and the distance data set.
+To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualization using the `show_langevitour()` function.
+This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.
+
+Before plotting, the data needs to be organized into a combined format through the `comb_data_model()` function.
+This function takes three inputs: `highd_data` (the high-dimensional observations), `model_highd` (high-dimensional summaries for each bin), and `model_2d` (the hexagonal bin centroids of the model).
+It returns a tidy data frame combining both the data and the model.
+
+In this structure, the `type` variable distinguishes between original observations (`"data"`) and the bin-averaged model representation (`"model"`).
 
@@
 ```
+
+The `show_langevitour()` function then renders the visualization using the `langevitour` interface, displaying both types of points in a dynamic tour.
+The `edge_data` input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model’s structure.
+
+In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2-\text{D}$ wireframe structure mapped back to high-dimensional space.
 

### 2025-05-13 paper-quollr.tex
Added high-d model vis

@@
 
-Displaying the \(p\text{-}D\) model overlaid on the data is done using the function \texttt{show\_langevitour()}.
-This visualization is helpful for visually evaluating how well the model fits the data.
-The function requires several arguments: data along with their corresponding hexagonal bin ID, \(2\text{-}D\) and \(p\text{-}D\) model points, the threshold for removing long edges, and the distance data set.
+To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualization using the \texttt{show\_langevitour()} function.
+This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.
+
+Before plotting, the data needs to be organized into a combined format through the \texttt{comb\_data\_model()} function.
+This function takes three inputs: \texttt{highd\_data} (the high-dimensional observations), \texttt{model\_highd} (high-dimensional summaries for each bin), and \texttt{model\_2d} (the hexagonal bin centroids of the model).
+It returns a tidy data frame combining both the data and the model.
+
+In this structure, the \texttt{type} variable distinguishes between original observations (\texttt{"data"}) and the bin-averaged model representation (\texttt{"model"}).
 
@@
 
+The \texttt{show\_langevitour()} function then renders the visualization using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour.
+The \texttt{edge\_data} input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model's structure.
+
+In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2-\text{D}\) wireframe structure mapped back to high-dimensional space.
+
 \begin{verbatim}

### 2025-05-13 paper-quollr.Rmd
Initial update for link plot section
Added show_link_plots description

@@
 There are mainly two interactive link plots can be generated.
-`show_link_plots()` helps to examine the fit.
-The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, and high-dimensional model data, and edge data where the from and to links of the edges.
-
-`show_error_link_plots()` helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
-The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
+
+To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
+This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+Both views are interactively linked, enabling users to explore.
+
+Before visualization, the input data must be prepared using the `comb_all_data_model()` function.
+This function combines the high-dimensional data (`highd_data`), the NLDR output (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the 2D bin layout (`model_2d`):
+
+This combined dataset includes both the original observations and the bin-level model averages, labeled with a `type` variable for distinguishing between them.
 
@@
 
+The function `show_link_plots()` generates two side-by-side, interactively linked plots; a 2D scatterplot of the NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+
+These two views are linked using `crosstalk`, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
+This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
+
 ```{r, echo=TRUE, eval=knitr::is_html_output()}
@@
 ```
+
+`show_error_link_plots()` helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
+The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
 

### 2025-05-13 paper-quollr.tex
Initial update for link plot section
Added show_link_plots description

@@
 There are mainly two interactive link plots can be generated.
-\texttt{show\_link\_plots()} helps to examine the fit.
-The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, and high-dimensional model data, and edge data where the from and to links of the edges.
-
-\texttt{show\_error\_link\_plots()} helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
-The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
+
+To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
+This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+Both views are interactively linked, enabling users to explore.
+
+Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
+This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR output (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the 2D bin layout (\texttt{model\_2d}):
+
+This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.
 
@@
 
+The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a 2D scatterplot of the NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+
+These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
+This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
+
 \begin{verbatim}
@@
 \end{verbatim}
+
+\texttt{show\_error\_link\_plots()} helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
+The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
 

### 2025-05-15 paper-quollr.Rmd
Updated link plot texts

@@
 Before visualization, the input data must be prepared using the `comb_all_data_model()` function.
-This function combines the high-dimensional data (`highd_data`), the NLDR output (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the 2D bin layout (`model_2d`):
+This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2-\text{D}$ bin layout (`model_2d`):
 
@@
 
-The function `show_link_plots()` generates two side-by-side, interactively linked plots; a 2D scatterplot of the NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
-
-These two views are linked using `crosstalk`, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
-This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
+The function `show_link_plots()` generates two side-by-side, interactively linked plots; a $2-\text{D}$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function takes the output from `comb_all_data_model()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
+
+These two views are linked using `crosstalk`, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the `langevitour` output.
 
@@
 `show_error_link_plots()` helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
-The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
+
+Before visualization, the input data must be prepared using the `comb_all_data_model_error()` function.
+The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2-\text{D}$ model data (`model_2d`), and model error (`error_data`).
+
+This combined dataset includes both the original observations and the bin-level model averages, labeled with a `type` variable for distinguishing between them.
 
@@
   )
-```
+
+df_exe
+```
+
+The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2-\text{D}$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
+
+These two views are linked using `crosstalk`, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
+This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
+
+These three views are linked using `crosstalk`, allowing interactive selection of points in error plot and the NLDR plot to highlight corresponding structures in the `langevitour` output.
 
@@
 
-# Find the most appropriate fit
-
-```{r}
-#| label: scurve-diff-umap-layouts
-
-nldr_scaled_obj <- gen_scaled_data(
-  nldr_data = scurve_umap)
-umap_scaled1 <- nldr_scaled_obj$scaled_nldr
-
-scurve_nldr1 <- umap_scaled1 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("a", c(0.08, 0.9))
-
-nldr_scaled_obj2 <- gen_scaled_data(
-  nldr_data = scurve_umap2)
-umap_scaled2 <- nldr_scaled_obj2$scaled_nldr
-
-scurve_nldr2 <- umap_scaled2 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  interior_annotation("b")
-
-nldr_scaled_obj3 <- gen_scaled_data(
-  nldr_data = scurve_umap3)
-umap_scaled3 <- nldr_scaled_obj3$scaled_nldr
-
-scurve_nldr3 <- umap_scaled3 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("c", c(0.08, 0.9))
-
-nldr_scaled_obj4 <- gen_scaled_data(
-  nldr_data = scurve_umap4)
-umap_scaled4 <- nldr_scaled_obj4$scaled_nldr
-
-scurve_nldr4 <- umap_scaled4 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("d")
-
-nldr_scaled_obj5 <- gen_scaled_data(
-  nldr_data = scurve_umap5)
-umap_scaled5 <- nldr_scaled_obj5$scaled_nldr
-
-scurve_nldr5 <- umap_scaled5 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#636363') +
-  interior_annotation("e")
-
-nldr_scaled_obj6 <- gen_scaled_data(
-  nldr_data = scurve_umap6)
-umap_scaled6 <- nldr_scaled_obj6$scaled_nldr
-
-scurve_nldr6 <- umap_scaled6 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
-  geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  interior_annotation("f")
-```
-
-```{r}
-scurve_error <- dplyr::bind_rows(scurve_umap_mse,
-                                 scurve_umap_mse2,
-                                 scurve_umap_mse3,
-                                 scurve_umap_mse4,
-                                 scurve_umap_mse5,
-                                 scurve_umap_mse6)
-
-scurve_error <- scurve_error |>
-  mutate(a1 = round(a1, 2)) |>
-  filter(bin1 >= 5) |>
-  group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
-  ungroup()
-
-error_plot <- ggplot(scurve_error,
-         aes(x = a1,
-             y = sqrt(MSE),
-             color = method)) +
-    geom_point(size = 0.8) +
-    geom_line(linewidth = 0.3) +
-  scale_x_continuous(breaks =
-    sort(unique(scurve_error$a1))[
-      seq(1, length(
-        unique(scurve_error$a1)), 
-        by = 5)]) +
-  scale_color_manual(
-    values=c('#e41a1c','#ff7f00','#4daf4a', 
-             "#a65628",'#636363', '#984ea3'))+
-    ylab("RMSE") +
-    xlab(expression(paste("binwidth (", a[1], ")"))) +
-    theme_minimal() +
-    theme(panel.border = element_rect(fill = 'transparent'),
-          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-          axis.ticks.x = element_line(),
-          axis.ticks.y = element_line(),
-          legend.position = "none",
-          axis.text.x = element_text(size = 7),
-          axis.text.y = element_text(size = 7),
-          axis.title.x = element_text(size = 7),
-          axis.title.y = element_text(size = 7),
-          aspect.ratio = 1.5)
-```
-
-```{r}
-free(error_plot) + wrap_plots(
-  scurve_nldr1, scurve_nldr2, 
-  scurve_nldr3, scurve_nldr4,
-  scurve_nldr5, scurve_nldr6, 
-  ncol = 2)
-```
+<!-- # Find the most appropriate fit -->
+
+<!-- ```{r} -->
+<!-- #| label: scurve-diff-umap-layouts -->
+
+<!-- nldr_scaled_obj <- gen_scaled_data( -->
+<!--   nldr_data = scurve_umap) -->
+<!-- umap_scaled1 <- nldr_scaled_obj$scaled_nldr -->
+
+<!-- scurve_nldr1 <- umap_scaled1 |> -->
+<!--   ggplot(aes(x = emb1, -->
+<!--              y = emb2))+ -->
+<!--   geom_point(alpha=0.1, size=1, colour='#e41a1c') + -->
+<!--   interior_annotation("a", c(0.08, 0.9)) -->
+
+<!-- nldr_scaled_obj2 <- gen_scaled_data( -->
+<!--   nldr_data = scurve_umap2) -->
+<!-- umap_scaled2 <- nldr_scaled_obj2$scaled_nldr -->
+
+<!-- scurve_nldr2 <- umap_scaled2 |> -->
+<!--   ggplot(aes(x = emb1, -->
+<!--              y = emb2))+ -->
+<!--   geom_point(alpha=0.1, size=1, colour='#ff7f00') + -->
+<!--   interior_annotation("b") -->
+
+<!-- nldr_scaled_obj3 <- gen_scaled_data( -->
+<!--   nldr_data = scurve_umap3) -->
+<!-- umap_scaled3 <- nldr_scaled_obj3$scaled_nldr -->
+
+<!-- scurve_nldr3 <- umap_scaled3 |> -->
+<!--   ggplot(aes(x = emb1, -->
+<!--              y = emb2))+ -->
+<!--   geom_point(alpha=0.1, size=1, colour='#4daf4a') + -->
+<!--   interior_annotation("c", c(0.08, 0.9)) -->
+
+<!-- nldr_scaled_obj4 <- gen_scaled_data( -->
+<!--   nldr_data = scurve_umap4) -->
+<!-- umap_scaled4 <- nldr_scaled_obj4$scaled_nldr -->
+
+<!-- scurve_nldr4 <- umap_scaled4 |> -->
+<!--   ggplot(aes(x = emb1, -->
+<!--              y = emb2))+ -->
+<!--   geom_point(alpha=0.1, size=1, colour='#a65628') + -->
+<!--   interior_annotation("d") -->
+
+<!-- nldr_scaled_obj5 <- gen_scaled_data( -->
+<!--   nldr_data = scurve_umap5) -->
+<!-- umap_scaled5 <- nldr_scaled_obj5$scaled_nldr -->
+
+<!-- scurve_nldr5 <- umap_scaled5 |> -->
+<!--   ggplot(aes(x = emb1, -->
+<!--              y = emb2))+ -->
+<!--   geom_point(alpha=0.1, size=1, colour='#636363') + -->
+<!--   interior_annotation("e") -->
+
+<!-- nldr_scaled_obj6 <- gen_scaled_data( -->
+<!--   nldr_data = scurve_umap6) -->
+<!-- umap_scaled6 <- nldr_scaled_obj6$scaled_nldr -->
+
+<!-- scurve_nldr6 <- umap_scaled6 |> -->
+<!--   ggplot(aes(x = emb1, -->
+<!--              y = emb2))+ -->
+<!--   geom_point(alpha=0.1, size=1, colour='#984ea3') + -->
+<!--   interior_annotation("f") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- scurve_error <- dplyr::bind_rows(scurve_umap_mse, -->
+<!--                                  scurve_umap_mse2, -->
+<!--                                  scurve_umap_mse3, -->
+<!--                                  scurve_umap_mse4, -->
+<!--                                  scurve_umap_mse5, -->
+<!--                                  scurve_umap_mse6) -->
+
+<!-- scurve_error <- scurve_error |> -->
+<!--   mutate(a1 = round(a1, 2)) |> -->
+<!--   filter(bin1 >= 5) |> -->
+<!--   group_by(method, a1) |> -->
+<!--   filter(MSE == min(MSE)) |> -->
+<!--   ungroup() -->
+
+<!-- error_plot <- ggplot(scurve_error, -->
+<!--          aes(x = a1, -->
+<!--              y = sqrt(MSE), -->
+<!--              color = method)) + -->
+<!--     geom_point(size = 0.8) + -->
+<!--     geom_line(linewidth = 0.3) + -->
+<!--   scale_x_continuous(breaks = -->
+<!--     sort(unique(scurve_error$a1))[ -->
+<!--       seq(1, length( -->
+<!--         unique(scurve_error$a1)),  -->
+<!--         by = 5)]) + -->
+<!--   scale_color_manual( -->
+<!--     values=c('#e41a1c','#ff7f00','#4daf4a',  -->
+<!--              "#a65628",'#636363', '#984ea3'))+ -->
+<!--     ylab("RMSE") + -->
+<!--     xlab(expression(paste("binwidth (", a[1], ")"))) + -->
+<!--     theme_minimal() + -->
+<!--     theme(panel.border = element_rect(fill = 'transparent'), -->
+<!--           plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5), -->
+<!--           axis.ticks.x = element_line(), -->
+<!--           axis.ticks.y = element_line(), -->
+<!--           legend.position = "none", -->
+<!--           axis.text.x = element_text(size = 7), -->
+<!--           axis.text.y = element_text(size = 7), -->
+<!--           axis.title.x = element_text(size = 7), -->
+<!--           axis.title.y = element_text(size = 7), -->
+<!--           aspect.ratio = 1.5) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- free(error_plot) + wrap_plots( -->
+<!--   scurve_nldr1, scurve_nldr2,  -->
+<!--   scurve_nldr3, scurve_nldr4, -->
+<!--   scurve_nldr5, scurve_nldr6,  -->
+<!--   ncol = 2) -->
+<!-- ``` -->
 

### 2025-05-15 paper-quollr.tex
Updated link plot texts

@@
 #> # A tibble: 1 x 2
-#>   Error    MSE
+#>   Error  RMSE
 #>   <dbl> <dbl>
-#> 1 1481. 0.0325
+#> 1 1481. 0.180
 \end{verbatim}
@@
 Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
-This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR output (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the 2D bin layout (\texttt{model\_2d}):
+This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2-\text{D}\) bin layout (\texttt{model\_2d}):
 
@@
 
-The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a 2D scatterplot of the NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
-
-These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
-This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
+The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a \(2-\text{D}\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function takes the output from \texttt{comb\_all\_data\_model()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
+
+These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the \texttt{langevitour} output.
 
@@
 \texttt{show\_error\_link\_plots()} helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
-The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, high-dimensional model data, and model error, and edge data where the from and to links of the edges.
+
+Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
+The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2-\text{D}\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
+
+This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.
 
@@
   )
-\end{verbatim}
+
+df_exe
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 5,122 x 12
+#>         x1    x2    x3         x4        x5       x6        x7 type   emb1  emb2
+#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr> <dbl> <dbl>
+#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model    NA    NA
+#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model    NA    NA
+#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model    NA    NA
+#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model    NA    NA
+#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model    NA    NA
+#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model    NA    NA
+#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model    NA    NA
+#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model    NA    NA
+#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model    NA    NA
+#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model    NA    NA
+#> # i 5,112 more rows
+#> # i 2 more variables: sqrt_row_wise_total_error <dbl>, density <dbl>
+\end{verbatim}
+
+The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2-\text{D}\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
+
+These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
+This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
+
+These three views are linked using \texttt{crosstalk}, allowing interactive selection of points in error plot and the NLDR plot to highlight corresponding structures in the \texttt{langevitour} output.
 
@@
 \end{verbatim}
-
-\section{Find the most appropriate fit}\label{find-the-most-appropriate-fit}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-33-1}
 

### 2025-05-15 paper-quollr.Rmd
Removed the intro image

@@
 
-```{r}
-#| label: read-limb-nldr
-# Read a variety of different NLDR representations of limb
-# and plot them on same aspect ratio
-clr_choice <- "#000000"
-umap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr1 <- umap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#999999') +
-  interior_annotation("a")
-
-nldr1_org <- umap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("a")
-
-tsne_limb <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_30.rds")
-
-nldr2 <- tsne_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("b")
-
-nldr2_org <- tsne_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("b")
-
-phate_limb <- read_rds("data/limb_muscles/facs_limb_muscles_phate_knn_5.rds")
-
-nldr3 <- phate_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("c")
-
-nldr3_org <- phate_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("c")
-
-trimap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr4 <- trimap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("d")
-
-nldr4_org <- trimap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("d")
-
-pacmap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr5 <- pacmap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("e")
-
-nldr5_org <- pacmap_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("e")
-
-tsne_limb2 <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_15.rds")
-
-nldr6 <- tsne_limb2 |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  interior_annotation("f")
-
-nldr6_org <- tsne_limb2 |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("f")
-```
-
-```{r nldr-layouts, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
-Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
-Would their downstream analysis decisions vary based on the version they observe?
-Which representation most accurately represents the structure in high-dimensional space?", out.width = "100%", fig.width= 6, fig.height= 4}
-
-nldr1_org + nldr2_org + nldr3_org + 
-  nldr4_org + nldr5_org + nldr6_org +
-  plot_layout(ncol = 3)
-```
+Add software descriptive figure
 
@@
 
-### $2\text{-}D$ projection
-
 ### Link plots
@@
 ```{r}
+#| label: read-limb-nldr
+# Read a variety of different NLDR representations of limb
+# and plot them on same aspect ratio
+
+umap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr1 <- umap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#999999') +
+  interior_annotation("a")
+
+tsne_limb <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_30.rds")
+
+nldr2 <- tsne_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("b")
+
+phate_limb <- read_rds("data/limb_muscles/facs_limb_muscles_phate_knn_5.rds")
+
+nldr3 <- phate_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("c")
+
+trimap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr4 <- trimap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("d")
+
+pacmap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr5 <- pacmap_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  interior_annotation("e")
+
+tsne_limb2 <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_15.rds")
+
+nldr6 <- tsne_limb2 |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  interior_annotation("f")
+```
+
+```{r}
 #| label: combine-error-data-muscles

### 2025-05-15 paper-quollr.tex
Removed the intro image

@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/nldr-layouts-1} \caption{Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices.
-Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice.
-Would their downstream analysis decisions vary based on the version they observe?
-Which representation most accurately represents the structure in high-dimensional space?}\label{fig:nldr-layouts}
-\end{figure}
+Add software descriptive figure
 
@@
 
-\subsubsection{\texorpdfstring{\(2\text{-}D\) projection}{2\textbackslash text\{-\}D projection}}\label{text-d-projection}
-
 \subsubsection{Link plots}\label{link-plots}

### 2025-05-15 paper-quollr.Rmd
Recomputed the RMSE for limb data

@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \@ref(fig:nldr-layouts)).
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 
-```{r, echo=TRUE}
+```{r, echo=TRUE, eval=FALSE}
 fit_highd_model(
@@
 Following their pre-processing, tSNE was performed using ten principal components.
-\@ref(fig:nldr-layouts) (b) is the reproduction of the published plot.
+Figure \@ref(fig:limb-rmse) (b) is the reproduction of the published plot.
 The question is whether this accurately represents the cluster structure in the data.
@@
   group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
+  filter(RMSE == min(RMSE)) |>
   ungroup()
@@
 
-error_plot_limb <- plot_mse(error_limb) +
+error_plot_limb <- plot_rmse(error_limb) +
   scale_x_continuous(breaks = sort(unique(error_limb$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
@@
 
-```{r limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data  (shown in \\@ref(fig:nldr-layouts)) is the better representation using RMSE for varying binwidth ($a_1$).
+```{r limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
@@
 
-```{r, eval=knitr::is_html_output()}
-#| label: langevitour-limb-tsne-author-proj
+```{r langevitour-limb-tsne-author-proj, eval=knitr::is_html_output()}
 
@@
 
-```{r, eval=knitr::is_html_output()}
-#| label: langevitour-limb-tsne-best-proj
+```{r langevitour-limb-tsne-best-proj, eval=knitr::is_html_output()}
 
@@
 
-```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (\\@ref(fig:nldr-layouts) b) and the $2-\\text{D}$ layout selected (\\@ref(fig:nldr-layouts) f) by RMSE plot (\\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=15, fig.height=10}

### 2025-05-15 paper-quollr.tex
Recomputed the RMSE for limb data

@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \ref{fig:nldr-layouts}).
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 
-\begin{verbatim}
-#> $nldr_obj
-#> $nldr_obj$scaled_nldr
-#> # A tibble: 5,000 x 3
-#>     emb1  emb2    ID
-#>    <dbl> <dbl> <int>
-#>  1 0.707 0.839     1
-#>  2 0.231 0.401     2
-#>  3 0.232 0.215     3
-#>  4 0.790 0.564     4
-#>  5 0.761 0.551     5
-#>  6 0.445 0.721     6
-#>  7 0.900 0.137     7
-#>  8 0.247 0.392     8
-#>  9 0.325 0.542     9
-#> 10 0.278 0.231    10
-#> # i 4,990 more rows
-#> 
-#> $nldr_obj$lim1
-#> [1] -14.42166  13.32655
-#> 
-#> $nldr_obj$lim2
-#> [1] -12.43687  12.32455
-#> 
-#> 
-#> $hb_obj
-#> $a1
-#> [1] 0.08326136
-#> 
-#> $a2
-#> [1] 0.07210645
-#> 
-#> $bins
-#> [1] 15 16
-#> 
-#> $start_point
-#> [1] -0.10000000 -0.08923607
-#> 
-#> $centroids
-#> # A tibble: 240 x 3
-#>    hexID     c_x     c_y
-#>    <int>   <dbl>   <dbl>
-#>  1     1 -0.1    -0.0892
-#>  2     2 -0.0167 -0.0892
-#>  3     3  0.0665 -0.0892
-#>  4     4  0.150  -0.0892
-#>  5     5  0.233  -0.0892
-#>  6     6  0.316  -0.0892
-#>  7     7  0.400  -0.0892
-#>  8     8  0.483  -0.0892
-#>  9     9  0.566  -0.0892
-#> 10    10  0.649  -0.0892
-#> # i 230 more rows
-#> 
-#> $hex_poly
-#> # A tibble: 1,440 x 3
-#>    hex_poly_id       x       y
-#>          <int>   <dbl>   <dbl>
-#>  1           1 -0.1    -0.0412
-#>  2           1 -0.142  -0.0652
-#>  3           1 -0.142  -0.113 
-#>  4           1 -0.1    -0.137 
-#>  5           1 -0.0584 -0.113 
-#>  6           1 -0.0584 -0.0652
-#>  7           2 -0.0167 -0.0412
-#>  8           2 -0.0584 -0.0652
-#>  9           2 -0.0584 -0.113 
-#> 10           2 -0.0167 -0.137 
-#> # i 1,430 more rows
-#> 
-#> $data_hb_id
-#> # A tibble: 5,000 x 4
-#>     emb1  emb2    ID hexID
-#>    <dbl> <dbl> <int> <int>
-#>  1 0.707 0.839     1   205
-#>  2 0.231 0.401     2   109
-#>  3 0.232 0.215     3    65
-#>  4 0.790 0.564     4   146
-#>  5 0.761 0.551     5   146
-#>  6 0.445 0.721     6   172
-#>  7 0.900 0.137     7    58
-#>  8 0.247 0.392     8   110
-#>  9 0.325 0.542     9   141
-#> 10 0.278 0.231    10    80
-#> # i 4,990 more rows
-#> 
-#> $std_cts
-#> # A tibble: 142 x 3
-#>    hexID bin_counts std_counts
-#>    <int>      <int>      <dbl>
-#>  1    21         12     0.16  
-#>  2    22         17     0.227 
-#>  3    23         22     0.293 
-#>  4    24         10     0.133 
-#>  5    25          7     0.0933
-#>  6    26         12     0.16  
-#>  7    27          1     0.0133
-#>  8    36         42     0.56  
-#>  9    37         42     0.56  
-#> 10    38         44     0.587 
-#> # i 132 more rows
-#> 
-#> $tot_bins
-#> [1] 240
-#> 
-#> $non_bins
-#> [1] 142
-#> 
-#> $pts_bins
-#> # A tibble: 142 x 2
-#>    hexID pts_list  
-#>    <int> <list>    
-#>  1    21 <int [12]>
-#>  2    22 <int [17]>
-#>  3    23 <int [22]>
-#>  4    24 <int [10]>
-#>  5    25 <int [7]> 
-#>  6    26 <int [12]>
-#>  7    27 <int [1]> 
-#>  8    36 <int [42]>
-#>  9    37 <int [42]>
-#> 10    38 <int [44]>
-#> # i 132 more rows
-#> 
-#> attr(,"class")
-#> [1] "hex_bin_obj"
-#> 
-#> $model_highd
-#> # A tibble: 130 x 8
-#>    hexID      x1    x2    x3         x4        x5       x6        x7
-#>    <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
-#>  1    21 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105 
-#>  2    22 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363
-#>  3    23 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153 
-#>  4    24 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128 
-#>  5    25  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195 
-#>  6    26  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172 
-#>  7    36 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338
-#>  8    37 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126
-#>  9    38 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780
-#> 10    39 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117 
-#> # i 120 more rows
-#> 
-#> $model_2d
-#> # A tibble: 130 x 5
-#>    hexID   c_x     c_y bin_counts std_counts
-#>    <int> <dbl>   <dbl>      <dbl>      <dbl>
-#>  1    21 0.358 -0.0171         12     0.16  
-#>  2    22 0.441 -0.0171         17     0.227 
-#>  3    23 0.524 -0.0171         22     0.293 
-#>  4    24 0.608 -0.0171         10     0.133 
-#>  5    25 0.691 -0.0171          7     0.0933
-#>  6    26 0.774 -0.0171         12     0.16  
-#>  7    36 0.316  0.0550         42     0.56  
-#>  8    37 0.400  0.0550         42     0.56  
-#>  9    38 0.483  0.0550         44     0.587 
-#> 10    39 0.566  0.0550         39     0.52  
-#> # i 120 more rows
-#> 
-#> $trimesh_data
-#> # A tibble: 322 x 8
-#>     from    to x_from  y_from  x_to   y_to from_count to_count
-#>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>      <dbl>    <dbl>
-#>  1    16    26 0.275   0.127  0.233 0.199          65       67
-#>  2    25    26 0.150   0.199  0.233 0.199          39       67
-#>  3    25    37 0.150   0.199  0.191 0.271          39       62
-#>  4    36    49 0.108   0.271  0.150 0.343          62       34
-#>  5    36    37 0.108   0.271  0.191 0.271          62       62
-#>  6     1     7 0.358  -0.0171 0.316 0.0550         12       42
-#>  7    48    49 0.0665  0.343  0.150 0.343          45       34
-#>  8    37    38 0.191   0.271  0.275 0.271          62       68
-#>  9    26    27 0.233   0.199  0.316 0.199          67       56
-#> 10    27    38 0.316   0.199  0.275 0.271          56       68
-#> # i 312 more rows
-\end{verbatim}
-
 \subsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
@@
 Following their pre-processing, tSNE was performed using ten principal components.
-\ref{fig:nldr-layouts} (b) is the reproduction of the published plot.
+Figure \ref{fig:limb-rmse} (b) is the reproduction of the published plot.
 The question is whether this accurately represents the cluster structure in the data.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data  (shown in \\ref{fig:nldr-layouts}) is the better representation using RMSE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (\\ref{fig:nldr-layouts} b) and the $2-\text{D}$ layout selected (\\ref{fig:nldr-layouts} f) by RMSE plot (\\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \\ref{fig:limb-rmse} b) and the $2-\text{D}$ layout selected (Figure \\ref{fig:limb-rmse} f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-05-15 paper-quollr.R

@@ -107,13 +107,13 @@ datasets_tb |>
   column_spec(2, width = "8cm")
 
 
-## ----echo=TRUE----------------------------------------------------------------
-fit_highd_model(
-  highd_data = scurve, 
-  nldr_data = scurve_umap, 
-  bin1 = 15, 
-  q = 0.1, 
-  benchmark_highdens = 5)
+## ----echo=TRUE, eval=FALSE----------------------------------------------------
+# fit_highd_model(
+#   highd_data = scurve,
+#   nldr_data = scurve_umap,
+#   bin1 = 15,
+#   q = 0.1,
+#   benchmark_highdens = 5)
 
 
 ## ----echo=TRUE----------------------------------------------------------------
@@ -586,7 +586,7 @@ error_limb <- error_limb |>
   filter(bin1 >= 5) |>
   filter(a1 >= 0.03) |>
   group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
+  filter(RMSE == min(RMSE)) |>
   ungroup()
 
 error_limb <- error_limb |>
@@ -597,13 +597,13 @@ error_limb <- error_limb |>
 ## -----------------------------------------------------------------------------
 #| label: error-comp-muscles
 
-error_plot_limb <- plot_mse(error_limb) +
+error_plot_limb <- plot_rmse(error_limb) +
   scale_x_continuous(breaks = sort(unique(error_limb$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
   scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf')) 
 
 
 
-## ----limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data  (shown in \\@ref(fig:nldr-layouts)) is the better representation using RMSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases. Layout f is the best choice.", out.width="100%"----
[…]

### 2025-05-15 scripts/01_gen_mse_for_diff_methods.R

@@ -1,4 +1,0 @@
-library(quollr)
-library(dplyr)
-
-set.seed(20240110)

### 2025-05-15 paper-quollr.Rmd
Started adding the acknowlegment section

@@
 library(readr)
-library(tools)
+#library(tools)
 
@@
 More documentation of the package can be found at the web site [https://jayanilakshika.github.io/quollr/](https://jayanilakshika.github.io/quollr/).
-
-## Package dependencies
-
-Understanding the dependencies of the `quollr` package is essential for smooth operation and error prevention.
-The following dependencies refer to the other R packages that `quollr` relies on to execute its functions effectively.
-
-```{r}
-package_dependencies("quollr")
-```
 
@@
 This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template.
+These `R` packages were used for this work: `cli` [@gabor2025], `dplyr`, `ggplot2`, `interp` (>= 1.1-6), `langevitour`, `proxy`, `stats`, `tibble`, `tidyselect`, `crosstalk`, `plotly`, `kableExtra`, `patchwork`, and `readr`.
+
 The source code for reproducing this paper can be found at: <https://github.com/JayaniLakshika/paper-quollr>.

### 2025-05-15 paper-quollr.tex
Started adding the acknowlegment section

@@
 More documentation of the package can be found at the web site \url{https://jayanilakshika.github.io/quollr/}.
-
-\subsection{Package dependencies}\label{package-dependencies}
-
-Understanding the dependencies of the \texttt{quollr} package is essential for smooth operation and error prevention.
-The following dependencies refer to the other R packages that \texttt{quollr} relies on to execute its functions effectively.
-
-\begin{verbatim}
-#> $quollr
-#>  [1] "dplyr"       "ggplot2"     "grid"        "interp"      "langevitour"
-#>  [6] "proxy"       "rlang"       "rsample"     "stats"       "tibble"     
-#> [11] "tidyselect"
-\end{verbatim}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-21-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
 
@@
 This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
+These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr}, \texttt{ggplot2}, \texttt{interp} (\textgreater= 1.1-6), \texttt{langevitour}, \texttt{proxy}, \texttt{stats}, \texttt{tibble}, \texttt{tidyselect}, \texttt{crosstalk}, \texttt{plotly}, \texttt{kableExtra}, \texttt{patchwork}, and \texttt{readr}.
+
 The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.

### 2025-05-15 paper-quollr.R

@@ -21,7 +21,7 @@ library(ggplot2)
 library(dplyr)
 library(patchwork)
 library(readr)
-library(tools)
+#library(tools)
 
 set.seed(20240110)
 
@@ -56,10 +56,6 @@ theme_set(theme_linedraw() +
 ## -----------------------------------------------------------------------------
 #| label: import-scripts
 source("scripts/additional_functions.R")
-
-
-## -----------------------------------------------------------------------------
-package_dependencies("quollr")
 
 
 ## -----------------------------------------------------------------------------

### 2025-05-15 paper-quollr.Rmd
Acknowlege the packages

@@
 This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template.
-These `R` packages were used for this work: `cli` [@gabor2025], `dplyr`, `ggplot2`, `interp` (>= 1.1-6), `langevitour`, `proxy`, `stats`, `tibble`, `tidyselect`, `crosstalk`, `plotly`, `kableExtra`, `patchwork`, and `readr`.
+These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].
 

### 2025-05-15 paper-quollr.tex
Acknowlege the packages

@@
 This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
-These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr}, \texttt{ggplot2}, \texttt{interp} (\textgreater= 1.1-6), \texttt{langevitour}, \texttt{proxy}, \texttt{stats}, \texttt{tibble}, \texttt{tidyselect}, \texttt{crosstalk}, \texttt{plotly}, \texttt{kableExtra}, \texttt{patchwork}, and \texttt{readr}.
+These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.
 

### 2025-05-15 paper-quollr.Rmd
Removed website
Removed dataset section

@@
 
+# Algorithm
+
 # Implementation
@@
 
-The package can be installed from CRAN:
-
-```r
-install.packages("quollr")
-```
-
-and the development version can be installed from GitHub:
+The development version can be installed from GitHub:
 
@@
 devtools::install_github("JayaniLakshika/quollr")
-```
-
-## Web site
-
-More documentation of the package can be found at the web site [https://jayanilakshika.github.io/quollr/](https://jayanilakshika.github.io/quollr/).
-
-## Data sets
-
-The `quollr` package comes with several data sets that load with the package.
-These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:datasets-tb-html)', '\\@ref(tab:datasets-tb-pdf)'))`.
-
-```{r}
-datasets_tb <- tibble(dt = c("scurve",
-                             "scurve_umap",
-                             "scurve_umap_predict",
-                             "scurve_umap2",
-                             "scurve_umap3",
-                             "scurve_umap4",
-                             "scurve_umap5",
-                             "scurve_umap6",
-                             "scurve_model_obj",
-                             "scurve_umap_mse",
-                             "scurve_umap_mse2",
-                             "scurve_umap_mse3",
-                             "scurve_umap_mse4",
-                             "scurve_umap_mse5",
-                             "scurve_umap_mse6"), 
-                      text = c("Simulated data to generate a S-curve with additional noise dimensions.",
-                             "The UMAP embedding with n_neighbors $46$ and min_dist $0.9$ for `S-curve`.",
-                             "The predicted UMAP embedding with n_neighbors $46$ and min_dist $0.9$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $10$ and min_dist $0.4$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $62$ and min_dist $0.1$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $30$ and min_dist $0.5$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $15$ and min_dist $0.5$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $15$ and min_dist $0.1$ for `S-curve`.",
-                             "The object contains the scaled NLDR object (`nldr_obj`), the hexagonal object (`hb_obj`), the fitted model in both $2\\text{-}D$ (`model_2d`), and $p\\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap2`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap3`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap4`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap5`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap6`."))
-```
-
-```{r datasets-tb-html, eval=knitr::is_html_output()}
-datasets_tb |> 
-  kable(caption = "quollr data sets", col.names = c("data", "explanation")) 
-```
-
-```{r datasets-tb-pdf, eval=knitr::is_latex_output()}
-datasets_tb |> 
-  kable(caption = "quollr data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
-  column_spec(1, width = "4cm") |>
-  column_spec(2, width = "8cm")
 ```

### 2025-05-15 paper-quollr.tex
Removed website
Removed dataset section

@@
 
+\section{Algorithm}\label{algorithm}
+
 \section{Implementation}\label{implementation}
@@
 
-The package can be installed from CRAN:
-
-\begin{verbatim}
-install.packages("quollr")
-\end{verbatim}
-
-and the development version can be installed from GitHub:
+The development version can be installed from GitHub:
 
@@
 \end{verbatim}
-
-\subsection{Web site}\label{web-site}
-
-More documentation of the package can be found at the web site \url{https://jayanilakshika.github.io/quollr/}.
-
-\subsection{Data sets}\label{data-sets}
-
-The \texttt{quollr} package comes with several data sets that load with the package. These are described in Table \ref{tab:datasets-tb-pdf}.
-
-\begin{table}
-
-\caption{\label{tab:datasets-tb-pdf}quollr data sets}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
-\toprule
-data & explanation\\
-\midrule
-scurve & Simulated data to generate a S-curve with additional noise dimensions.\\
-scurve\_umap & The UMAP embedding with n\_neighbors \$46\$ and min\_dist \$0.9\$ for `S-curve`.\\
-scurve\_umap\_predict & The predicted UMAP embedding with n\_neighbors \$46\$ and min\_dist \$0.9\$ for `S-curve`.\\
-scurve\_umap2 & The UMAP embedding with n\_neighbors \$10\$ and min\_dist \$0.4\$ for `S-curve`.\\
-scurve\_umap3 & The UMAP embedding with n\_neighbors \$62\$ and min\_dist \$0.1\$ for `S-curve`.\\
-\addlinespace
-scurve\_umap4 & The UMAP embedding with n\_neighbors \$30\$ and min\_dist \$0.5\$ for `S-curve`.\\
-scurve\_umap5 & The UMAP embedding with n\_neighbors \$15\$ and min\_dist \$0.5\$ for `S-curve`.\\
-scurve\_umap6 & The UMAP embedding with n\_neighbors \$15\$ and min\_dist \$0.1\$ for `S-curve`.\\
-scurve\_model\_obj & The object contains the scaled NLDR object (`nldr\_obj`), the hexagonal object (`hb\_obj`), the fitted model in both \$2\textbackslash{}text\{-\}D\$ (`model\_2d`), and \$p\textbackslash{}text\{-\}D\$ (`model\_highd`), and triangular mesh (`trimesh\_data`).\\
-scurve\_umap\_mse & The MSE corresponding to different binwidths for the model generated using `scurve\_umap`.\\
-\addlinespace
-scurve\_umap\_mse2 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap2`.\\
-scurve\_umap\_mse3 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap3`.\\
-scurve\_umap\_mse4 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap4`.\\
-scurve\_umap\_mse5 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap5`.\\
-scurve\_umap\_mse6 & The MSE corresponding to different binwidths for the model generated using `scurve\_umap6`.\\
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-21-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-20-1}
 

### 2025-05-15 paper-quollr.R

@@ -58,51 +58,6 @@ theme_set(theme_linedraw() +
 source("scripts/additional_functions.R")
 
 
-## -----------------------------------------------------------------------------
-datasets_tb <- tibble(dt = c("scurve",
-                             "scurve_umap",
-                             "scurve_umap_predict",
-                             "scurve_umap2",
-                             "scurve_umap3",
-                             "scurve_umap4",
-                             "scurve_umap5",
-                             "scurve_umap6",
-                             "scurve_model_obj",
-                             "scurve_umap_mse",
-                             "scurve_umap_mse2",
-                             "scurve_umap_mse3",
-                             "scurve_umap_mse4",
-                             "scurve_umap_mse5",
-                             "scurve_umap_mse6"), 
-                      text = c("Simulated data to generate a S-curve with additional noise dimensions.",
-                             "The UMAP embedding with n_neighbors $46$ and min_dist $0.9$ for `S-curve`.",
-                             "The predicted UMAP embedding with n_neighbors $46$ and min_dist $0.9$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $10$ and min_dist $0.4$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $62$ and min_dist $0.1$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $30$ and min_dist $0.5$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $15$ and min_dist $0.5$ for `S-curve`.",
-                             "The UMAP embedding with n_neighbors $15$ and min_dist $0.1$ for `S-curve`.",
-                             "The object contains the scaled NLDR object (`nldr_obj`), the hexagonal object (`hb_obj`), the fitted model in both $2\\text{-}D$ (`model_2d`), and $p\\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap2`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap3`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap4`.",
-                             "The MSE corresponding to different binwidths for the model generated using `scurve_umap5`.",
[…]

### 2025-05-15 paper-quollr.Rmd
Added Usage section

@@
 
-<!-- ## Usage -->
-
-<!-- add about main function to gen 2D and highD models-->
-<!--Discuss the model can be generated with bin centroids or bin means-->
+## Usage
 
@@
 
-## Main function
+### Main function
 
@@
 
-## Constructing the $2\text{-}D$ Model
+### Constructing the $2\text{-}D$ Model
 
@@
 
-### Scaling the Data
+#### Scaling the Data
 
@@
 
-### Computing hexagon grid configurations
+#### Computing hexagon grid configurations
 
@@
 
-### Binning the data
+#### Binning the data
 
@@
 
-#### Generating all possible centroids in a hexagonal grid
+##### Generating all possible centroids in a hexagonal grid
 
@@
 
-#### Creating the coordinates of the hexagons
+##### Creating the coordinates of the hexagons
 
@@
 
-#### Assigning data points to their respective hexagons
+##### Assigning data points to their respective hexagons
 
@@
 
-#### Computing the standardized number of points within each hexagon
+##### Computing the standardized number of points within each hexagon
 
@@
 
-#### Mapping the points to their corresponding hexagonal bins
+##### Mapping the points to their corresponding hexagonal bins
 
@@
 
-### Obtaining bin centroids
+#### Obtaining bin centroids
 
@@
 
-### Indicating neighbors by line segments connecting centroids
+#### Indicating neighbors by line segments connecting centroids
 
@@
 
-### Identifying and removing low-density hexagons
+#### Identifying and removing low-density hexagons
 
@@
 
-## Lifting the model into high dimensions
+### Lifting the model into high dimensions
 
@@
 
-## Prediction
+### Prediction
 
@@
 
-## Compute residuals and Mean Square Error (MSE)
-
-As a Goodness of fit statistics for the model, `glance()` is used to compute residuals and MSE.
+### Compute residuals and Root Mean Square Error (RMSE)
+
+As a Goodness of fit statistics for the model, `glance()` is used to compute residuals and RMSE.
 These metrics are used to assess how well the fitted model will capture the underlying structure of the $p\text{-}D$ data.
@@
 
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the Mean squared error (MSE) which is the average of the total squared error per point.
-
-These metrics are returned in a tibble as `Error` (absolute error) and `MSE` (mean squared error).
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RMSE which is the average of the total squared error per point.
+
+These metrics are returned in a tibble as `Error` (absolute error) and `RMSE` (root mean squared error).
 
@@
 
-## Visualizations
+### Visualizations
 
@@
 
-### Hexagonal grid
+#### Hexagonal grid
 
@@
 
-### Triangular mesh
+#### Triangular mesh
 
@@
 
-### $p\text{-}D$ model visualization
+#### $p\text{-}D$ model visualization
 
@@
 
-### Link plots
+#### Link plots
 

### 2025-05-15 paper-quollr.tex
Added Usage section

@@
 
+\subsection{Usage}\label{usage}
+
 The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. We also want to load the built-in data sets \texttt{scurve} and \texttt{scurve\_umap}.
@@
 
-\subsection{Main function}\label{main-function}
+\subsubsection{Main function}\label{main-function}
 
@@
 
-\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
+\subsubsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
 
@@
 
-\subsubsection{Scaling the Data}\label{scaling-the-data}
+\paragraph{Scaling the Data}\label{scaling-the-data}
 
@@
 
-\subsubsection{Computing hexagon grid configurations}\label{computing-hexagon-grid-configurations}
+\paragraph{Computing hexagon grid configurations}\label{computing-hexagon-grid-configurations}
 
@@
 
-\subsubsection{Binning the data}\label{binning-the-data}
+\paragraph{Binning the data}\label{binning-the-data}
 
@@
 
-\paragraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
+\subparagraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
 
@@
 
-\paragraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
+\subparagraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
 
@@
 
-\paragraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
+\subparagraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
 
@@
 
-\paragraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
+\subparagraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
 
@@
 
-\paragraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
+\subparagraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
 
@@
 
-\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
+\paragraph{Obtaining bin centroids}\label{obtaining-bin-centroids}
 
@@
 
-\subsubsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
+\paragraph{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
 
@@
 
-\subsubsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
+\paragraph{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
 
@@
 
-\subsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}
+\subsubsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}
 
@@
 
-\subsection{Prediction}\label{prediction}
+\subsubsection{Prediction}\label{prediction}
 
@@
 
-\subsection{Compute residuals and Mean Square Error (MSE)}\label{compute-residuals-and-mean-square-error-mse}
+\subsubsection{Compute residuals and Root Mean Square Error (RMSE)}\label{compute-residuals-and-root-mean-square-error-rmse}
 
-As a Goodness of fit statistics for the model, \texttt{glance()} is used to compute residuals and MSE.
+As a Goodness of fit statistics for the model, \texttt{glance()} is used to compute residuals and RMSE.
 These metrics are used to assess how well the fitted model will capture the underlying structure of the \(p\text{-}D\) data.
@@
 
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the Mean squared error (MSE) which is the average of the total squared error per point.
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RMSE which is the average of the total squared error per point.
 
-These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{MSE} (mean squared error).
+These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{RMSE} (root mean squared error).
 
@@
 
-\subsection{Visualizations}\label{visualizations}
+\subsubsection{Visualizations}\label{visualizations}
 
@@
 
-\subsubsection{Hexagonal grid}\label{hexagonal-grid}
+\paragraph{Hexagonal grid}\label{hexagonal-grid}
 
@@
 
-\subsubsection{Triangular mesh}\label{triangular-mesh}
+\paragraph{Triangular mesh}\label{triangular-mesh}
 
@@
 
-\subsubsection{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
+\paragraph{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
 
@@
 
-\subsubsection{Link plots}\label{link-plots}
+\paragraph{Link plots}\label{link-plots}
 

### 2025-05-15 paper-quollr.Rmd
Add algorithm section

@@
 
+Objective: From this, we can understand how an NLDR warps the high-dimensional space and fits the data.
+
 Add software descriptive figure
@@
 The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
-In the application section, we illustrate the algorithm's functionality for studying clustering data structures.
+In the application section, we illustrate the algorithm's functionality for studying a clustering data structure.
 Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.
@@
 # Algorithm
+
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, (6) identifying and removing low-density hexagons, and (7) lifting the model into higher dimensions.
+A detailed description of the algorithm can be found in **cite the methodology paper**.
+
+Add a Figure explaining the steps
 
@@
 It is constructed by simulating $5000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
-The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
+The remaining variables $X_4$, $X_5$, $X_6$, $X_7$ are all uniform error, with small variance.
 `scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $46$ and `min_dist` is $0.9$.
@@
 
-#### Scaling the Data
+#### Scaling the data
 

### 2025-05-15 paper-quollr.tex
Add algorithm section

@@
 The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
-In the application section, we illustrate the algorithm's functionality for studying clustering data structures.
+In the application section, we illustrate the algorithm's functionality for studying a clustering data structure.
 Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.
@@
 \section{Algorithm}\label{algorithm}
+
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, (6) identifying and removing low-density hexagons, and (7) lifting the model into higher dimensions.
+A detailed description of the algorithm can be found in \textbf{cite the methodology paper}.
+
+Add a Figure explaining the steps
 
@@
 It is constructed by simulating \(5000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\).
-The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error, with small variance.
+The remaining variables \(X_4\), \(X_5\), \(X_6\), \(X_7\) are all uniform error, with small variance.
 \texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors} is \(46\) and \texttt{min\_dist} is \(0.9\).
@@
 
-\paragraph{Scaling the Data}\label{scaling-the-data}
+\paragraph{Scaling the data}\label{scaling-the-data}
 

### 2025-05-15 paper-quollr.Rmd
Start working on addimg the intro image

@@
   Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation.
-However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices.
+However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper-)parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
-The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data.
+The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper-)parameter choices provide the most accurate representation of high-dimensional data.
 The `scurve` data from the package is used to illustrate the algorithm.
@@
 library(readr)
+library(plotly)
+library(crosstalk)
 #library(tools)
@@
      legend.key.height = unit(0.25, 'cm'),
-     legend.key.width = unit(0.25, 'cm')
+     legend.key.width = unit(0.25, 'cm'),
+     plot.margin = margin(0, 0, 0, 0)
    )
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper-)parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
 
-This paper presents the R package, `quollr` which introduce a new visual tool to evaluate which NLDR method and which (hyper)parameter choice gives the most accurate representation of high-dimensional data.
-The methodology of this algorithm is explained in *cite the methodology paper*.
-The software is available from the Comprehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).
-
-Objective: From this, we can understand how an NLDR warps the high-dimensional space and fits the data.
-
-Add software descriptive figure
+This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
+
+```{r}
+clr_choice <- "#000000"
+
+scurve_model_obj <- fit_highd_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  bin1 = 15, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+scurve_umap_scaled <- scurve_model_obj$nldr_obj$scaled_nldr
+tr_from_to_df_scurve <- scurve_model_obj$trimesh_data
+df_bin_centroids_scurve <- scurve_model_obj$model_2d
+df_bin_scurve <- scurve_model_obj$model_highd
+hex_grid_scurve <- scurve_model_obj$hb_obj$hex_poly
+counts_df_scurve <- scurve_model_obj$hb_obj$std_cts
+
+hex_grid_with_counts_scurve <- left_join(
+  hex_grid_scurve, counts_df_scurve, 
+  by = c("hex_poly_id" = "hexID"))
+
+hex_grid_nonempty_scurve <- hex_grid_scurve |>
+  filter(hex_poly_id %in% df_bin_centroids_scurve$hexID)
+
+scurve_umap_plt <- ggplot(
+  scurve_umap_scaled, 
+  aes(x = emb1, y = emb2)) +
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
+
+hex_grid_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff") +
+  geom_point(data = scurve_umap_scaled, 
+             aes(x = emb1, y = emb2), 
+             alpha = 0.2, size = 0.5, color = clr_choice) +
+  geom_segment(data = tr_from_to_df_scurve,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#33a02c")
+
+hex_grid_scurve <- ggplotly(hex_grid_scurve, width = "600", 
+                     height = "600", tooltip = "none")
+
+df_exe <- comb_data_model(
+    highd_data = scurve, 
+    model_highd = df_bin_scurve, 
+    model_2d = df_bin_centroids_scurve
+)
+
+scurve_umap_model_vis <- show_langevitour(
+    point_data = df_exe, 
+    edge_data = tr_from_to_df_scurve
+)
+```
+
+```{r, eval=knitr::is_html_output()}
+
+crosstalk::bscols(
+    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+                   hex_grid_scurve,
+                   scurve_umap_model_vis),
+    device = "sm"
+  )
+```
 

### 2025-05-15 paper-quollr.tex
Start working on addimg the intro image

@@
 Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (\(p\text{-}D\)) by applying a non-linear transformation.
-However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices.
+However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper-)parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
-The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data.
+The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper-)parameter choices provide the most accurate representation of high-dimensional data.
 The \texttt{scurve} data from the package is used to illustrate the algorithm.
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper-)parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
 
-This paper presents the R package, \texttt{quollr} which introduce a new visual tool to evaluate which NLDR method and which (hyper)parameter choice gives the most accurate representation of high-dimensional data.
-The methodology of this algorithm is explained in \emph{cite the methodology paper}.
-The software is available from the Comprehensive R Archive Network (CARN) at \url{https://CRAN.R-project.org/package=quollr}.
-
-Objective: From this, we can understand how an NLDR warps the high-dimensional space and fits the data.
-
-Add software descriptive figure
+This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-21-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-22-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
 

### 2025-05-15 paper-quollr.R

@@ -21,6 +21,8 @@ library(ggplot2)
 library(dplyr)
 library(patchwork)
 library(readr)
+library(plotly)
+library(crosstalk)
 #library(tools)
 
 set.seed(20240110)
@@ -47,7 +49,8 @@ theme_set(theme_linedraw() +
      legend.title = element_text(size=5), 
      legend.text = element_text(size=4),
      legend.key.height = unit(0.25, 'cm'),
-     legend.key.width = unit(0.25, 'cm')
+     legend.key.width = unit(0.25, 'cm'),
+     plot.margin = margin(0, 0, 0, 0)
    )
 
 )
@@ -56,6 +59,80 @@ theme_set(theme_linedraw() +
 ## -----------------------------------------------------------------------------
 #| label: import-scripts
 source("scripts/additional_functions.R")
+
+
+## -----------------------------------------------------------------------------
+clr_choice <- "#000000"
+
+scurve_model_obj <- fit_highd_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  bin1 = 15, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+scurve_umap_scaled <- scurve_model_obj$nldr_obj$scaled_nldr
+tr_from_to_df_scurve <- scurve_model_obj$trimesh_data
+df_bin_centroids_scurve <- scurve_model_obj$model_2d
+df_bin_scurve <- scurve_model_obj$model_highd
+hex_grid_scurve <- scurve_model_obj$hb_obj$hex_poly
+counts_df_scurve <- scurve_model_obj$hb_obj$std_cts
+
+hex_grid_with_counts_scurve <- left_join(
+  hex_grid_scurve, counts_df_scurve, 
+  by = c("hex_poly_id" = "hexID"))
+
+hex_grid_nonempty_scurve <- hex_grid_scurve |>
+  filter(hex_poly_id %in% df_bin_centroids_scurve$hexID)
+
+scurve_umap_plt <- ggplot(
+  scurve_umap_scaled, 
+  aes(x = emb1, y = emb2)) +
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
+
+hex_grid_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff") +
[…]

### 2025-05-15 paper-quollr.Rmd
Added intro image

@@
 
+## 2-d vis model
 hex_grid_scurve <- ggplot(
@@
 
-hex_grid_scurve <- ggplotly(hex_grid_scurve, width = "600", 
-                     height = "600", tooltip = "none")
-
+hex_grid_scurve_int <- ggplotly(hex_grid_scurve, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
+## High-d vis model
 df_exe <- comb_data_model(
@@
 
-scurve_umap_model_vis <- show_langevitour(
-    point_data = df_exe, 
-    edge_data = tr_from_to_df_scurve
-)
+df <- df_exe |>
+    dplyr::filter(type == "data") ## original dataset
+
+df_b <- df_exe |>
+  dplyr::filter(type == "model") ## High-d model
+
+scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = tr_from_to_df_scurve$from,
+                         lineTo = tr_from_to_df_scurve$to,
+                         group = df_exe$type,
+                         pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
+                         levelColors = c("#000000", "#33a02c"),
+                         enableControls = FALSE,
+                         width = "421px", height = "421px")
+
+## Generate 2D projections
+
 ```
@@
     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
-                   hex_grid_scurve,
-                   scurve_umap_model_vis),
-    device = "sm"
-  )
+                   hex_grid_scurve_int,
+                  htmltools::div(
+                    style = "margin-top: 13px;",  # adjust px as needed
+                    scurve_umap_model_vis
+                  )),
+    device = "xs"
+  )
+```
+
+```{r, eval=knitr::is_latex_output()}
+
 ```
@@
 
+```{r, eval=knitr::is_latex_output()}
+
+```
+
 #### Link plots
@@
   )
+```
+
+```{r, eval=knitr::is_latex_output()}
+
 ```

### 2025-05-15 paper-quollr.R

@@ -90,6 +90,7 @@ scurve_umap_plt <- ggplot(
   aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
 
+## 2-d vis model
 hex_grid_scurve <- ggplot(
   data = hex_grid_with_counts_scurve, 
   aes(x = x, y = y)) +
@@ -110,29 +111,61 @@ hex_grid_scurve <- ggplot(
              aes(x = c_x, y = c_y), 
              size = 1, color = "#33a02c")
 
-hex_grid_scurve <- ggplotly(hex_grid_scurve, width = "600", 
-                     height = "600", tooltip = "none")
+hex_grid_scurve_int <- ggplotly(hex_grid_scurve, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
 
+
+## High-d vis model
 df_exe <- comb_data_model(
     highd_data = scurve, 
     model_highd = df_bin_scurve, 
     model_2d = df_bin_centroids_scurve
 )
 
-scurve_umap_model_vis <- show_langevitour(
-    point_data = df_exe, 
-    edge_data = tr_from_to_df_scurve
-)
+df <- df_exe |>
+    dplyr::filter(type == "data") ## original dataset
+
+df_b <- df_exe |>
+  dplyr::filter(type == "model") ## High-d model
+
+scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = tr_from_to_df_scurve$from,
+                         lineTo = tr_from_to_df_scurve$to,
+                         group = df_exe$type,
+                         pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
+                         levelColors = c("#000000", "#33a02c"),
+                         enableControls = FALSE,
+                         width = "421px", height = "421px")
+
+## Generate 2D projections
+
 
 
 ## ----eval=knitr::is_html_output()---------------------------------------------
 # 
 # crosstalk::bscols(
[…]

### 2025-05-15 paper-quollr.Rmd
Added one projection for S-curve data

@@
 
-## Generate 2D projections
-
 ```
@@
 
+```{r scurve-projections}
+
+df_b_scurve <- df_bin_scurve |>
+  dplyr::filter(hexID %in% df_bin_centroids_scurve$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_scurve <- df_b_scurve[match(df_bin_centroids_scurve$hexID, df_b_scurve$hexID),] |>
+  dplyr::select(-hexID)
+
+scurve_labeled <- scurve |>
+  select(-ID) |>
+  mutate(type = "data")
+
+# Apply the scaling
+df_model_data_scurve <- bind_rows(scurve_labeled, df_b_scurve)
+scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
+  as_tibble()
+
+scaled_scurve_data <- scaled_scurve |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_scurve_data_model <- scaled_scurve |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(0.10479,0.06673,0.19430,-0.14763,0.02861,-0.04302,0.06601),
+  c(0.11421,0.18556,-0.16860,-0.05033,-0.00219,0.04159,0.04285))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.0259))
+
+scurve_proj_umap_model1 <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 4,
+  is_category = FALSE) +
+  interior_annotation(label = "a1", cex = 2)
+
+```
+
 ```{r, eval=knitr::is_latex_output()}
-
+hex_grid_scurve + wrap_plots(
+  scurve_proj_umap_model1, scurve_proj_umap_model1,
+  scurve_proj_umap_model1, scurve_proj_umap_model1, 
+  ncol = 2)
 ```

### 2025-05-15 paper-quollr.R

@@ -147,8 +147,6 @@ scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          enableControls = FALSE,
                          width = "421px", height = "421px")
 
-## Generate 2D projections
-
 
 
 ## ----eval=knitr::is_html_output()---------------------------------------------
@@ -156,16 +154,73 @@ scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
 # crosstalk::bscols(
 #     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
 #                    hex_grid_scurve_int,
-#                   htmltools::div(
-#                     style = "margin-top: 13px;",  # adjust px as needed
-#                     scurve_umap_model_vis
-#                   )),
+#                     htmltools::div(
+#                       style = "margin-top: 13px;",  # adjust px as needed
+#                       scurve_umap_model_vis
+#                     )),
 #     device = "xs"
 #   )
 
 
-## ----eval=knitr::is_latex_output()--------------------------------------------
+## ----scurve-projections-------------------------------------------------------
 
+df_b_scurve <- df_bin_scurve |>
+  dplyr::filter(hexID %in% df_bin_centroids_scurve$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_scurve <- df_b_scurve[match(df_bin_centroids_scurve$hexID, df_b_scurve$hexID),] |>
+  dplyr::select(-hexID) 
+
+scurve_labeled <- scurve |>
+  select(-ID) |>
+  mutate(type = "data")
+
+# Apply the scaling
+df_model_data_scurve <- bind_rows(scurve_labeled, df_b_scurve)
+scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
+  as_tibble()
+
+scaled_scurve_data <- scaled_scurve |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_scurve_data_model <- scaled_scurve |>
+  filter(type == "model") |>
+  select(-type)
+
+
+## First projection
+projection <- cbind(
+  c(0.10479,0.06673,0.19430,-0.14763,0.02861,-0.04302,0.06601),
+  c(0.11421,0.18556,-0.16860,-0.05033,-0.00219,0.04159,0.04285))
+
+proj_obj1 <- get_projection(projection = projection, 
[…]

### 2025-05-15 paper-quollr.Rmd
Added projections

@@
   plot_limits = c(-0.35, 0.35), 
-  axis_text_size = 4,
+  axis_text_size = 3,
   is_category = FALSE) +
-  interior_annotation(label = "a1", cex = 2)
-
+  interior_annotation(label = "a1", cex = 1)
+
+## Second projection
+
+projection <- cbind(
+  c(0.01588,-0.13640,-0.17180,-0.15486,-0.09726,0.00858,0.01090),
+  c(0.21443,-0.00390,0.08182,-0.11746,0.06820,-0.05458,-0.08885))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.0259))
+
+scurve_proj_umap_model2 <- plot_proj(
+  proj_obj = proj_obj2,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 3,
+  is_category = FALSE) +
+  interior_annotation(label = "a2", cex = 1)
+
+## Third projection
+
+projection <- cbind(
+  c(0.02727,0.09154,-0.16383,-0.04931,-0.17013,0.11965,-0.01884),
+  c(-0.10510,-0.10971,-0.09084,0.19642,-0.07163,-0.02486,0.07965))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.031))
+
+scurve_proj_umap_model3 <- plot_proj(
+  proj_obj = proj_obj3,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 3,
+  is_category = FALSE) +
+  interior_annotation(label = "a3", cex = 1)
+
+## Fourth projection
+
+projection <- cbind(
+  c(0.17174,0.14255,0.11950,0.10344,-0.07148,-0.01656,0.04275),
+  c(0.09270,0.04486,-0.14397,0.08305,0.11783,0.16187,-0.06077))
+
+proj_obj4 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.031))
+
+scurve_proj_umap_model4 <- plot_proj(
+  proj_obj = proj_obj4,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 3,
+  is_category = FALSE) +
+  interior_annotation(label = "a4", cex = 1) 
 ```
@@
 hex_grid_scurve + wrap_plots(
-  scurve_proj_umap_model1, scurve_proj_umap_model1,
-  scurve_proj_umap_model1, scurve_proj_umap_model1, 
+  scurve_proj_umap_model1, scurve_proj_umap_model2,
+  scurve_proj_umap_model3, scurve_proj_umap_model4, 
   ncol = 2)

### 2025-05-15 paper-quollr.Rmd
Start add the algorith steps

@@
 
+hex_grid_poly_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff") +
+  geom_point(data = scurve_umap_scaled, 
+             aes(x = emb1, y = emb2), 
+             alpha = 0.2, size = 0.5, color = clr_choice)
+
+hex_centroids_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff")  +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#33a02c")
+
+wireframe_scurve <- ggplot() +
+  geom_segment(data = tr_from_to_df_scurve,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#33a02c")
+
 ## 2-d vis model
@@
 
-Add a Figure explaining the steps
+```{r, eval=knitr::is_latex_output()}
+scurve_umap_plt + hex_grid_poly_scurve + 
+  hex_centroids_scurve + wireframe_scurve + 
+  plot_layout(ncol = 5)
+```
 

### 2025-05-15 paper-quollr.tex
Added intro image
Added one projection for S-curve data
Start add the algorith steps

@@
 
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-3-1}
+
 The paper is organized as follows.
@@
 
-Add a Figure explaining the steps
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-4-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-26-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
 

### 2025-05-15 paper-quollr.R

@@ -90,6 +90,38 @@ scurve_umap_plt <- ggplot(
   aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
 
+hex_grid_poly_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff") +
+  geom_point(data = scurve_umap_scaled, 
+             aes(x = emb1, y = emb2), 
+             alpha = 0.2, size = 0.5, color = clr_choice)
+
+hex_centroids_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff")  +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#33a02c")
+
+wireframe_scurve <- ggplot() +
+  geom_segment(data = tr_from_to_df_scurve,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#33a02c")
+
 ## 2-d vis model
 hex_grid_scurve <- ggplot(
   data = hex_grid_with_counts_scurve, 
@@ -295,6 +327,12 @@ hex_grid_scurve + wrap_plots(
   scurve_proj_umap_model1, scurve_proj_umap_model2,
   scurve_proj_umap_model3, scurve_proj_umap_model4, 
   ncol = 2)
+
+
+## ----eval=knitr::is_latex_output()--------------------------------------------
+scurve_umap_plt + hex_grid_poly_scurve + 
+  hex_centroids_scurve + wireframe_scurve + 
+  plot_layout(ncol = 5)
 
 
 ## ----echo=TRUE, eval=FALSE----------------------------------------------------

### 2025-05-15 paper-quollr.Rmd
Add algorithm steps latex output

@@
 
+scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 hex_grid_poly_scurve <- ggplot(
@@
 
+hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 hex_centroids_scurve <- ggplot(
@@
              size = 1, color = "#33a02c")
+
+hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
 
@@
              size = 1, color = "#33a02c")
+
+wireframe_scurve_int <- ggplotly(wireframe_scurve, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
 
@@
 
-```{r, eval=knitr::is_latex_output()}
+```{r, eval=knitr::is_html_output()}
+
+crosstalk::bscols(
+    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+                   scurve_umap_plt_int, hex_grid_poly_scurve_int,
+                   hex_centroids_scurve_int, wireframe_scurve_int, 
+                    htmltools::div(
+                      style = "margin-top: 13px;",  # adjust px as needed
+                      scurve_umap_model_vis
+                    )),
+    device = "xs"
+  )
+```
+
+```{r, eval=knitr::is_latex_output(), out.width="100%"}
 scurve_umap_plt + hex_grid_poly_scurve +
+  scurve_proj_umap_model1 + scurve_proj_umap_model2 +
   hex_centroids_scurve + wireframe_scurve + 
-  plot_layout(ncol = 5)
+  scurve_proj_umap_model3 + scurve_proj_umap_model4 +
+  plot_layout(ncol = 4)
 ```

### 2025-05-15 paper-quollr.tex
Add algorithm steps latex output

@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-4-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-26-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-26-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
 

### 2025-05-15 paper-quollr.R

@@ -90,6 +90,19 @@ scurve_umap_plt <- ggplot(
   aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
 
+scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 hex_grid_poly_scurve <- ggplot(
   data = hex_grid_with_counts_scurve, 
   aes(x = x, y = y)) +
@@ -100,6 +113,19 @@ hex_grid_poly_scurve <- ggplot(
              aes(x = emb1, y = emb2), 
              alpha = 0.2, size = 0.5, color = clr_choice)
 
+hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 hex_centroids_scurve <- ggplot(
   data = hex_grid_with_counts_scurve, 
   aes(x = x, y = y)) +
@@ -109,6 +135,19 @@ hex_centroids_scurve <- ggplot(
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
              size = 1, color = "#33a02c")
+
+hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
[…]

### 2025-05-15 paper-quollr.Rmd
Updated the algorithm steps

--- 2025-05-15 paper-quollr.Rmd
@@
 
+scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
+                                width = "250", 
+                                height = "250", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 hex_grid_poly_scurve <- ggplot(
@@
 
+hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
+                                width = "250", 
+                                height = "250", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 hex_centroids_scurve <- ggplot(
@@
              size = 1, color = "#33a02c")
+
+hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
+                                width = "250", 
+                                height = "250", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
 
@@
              size = 1, color = "#33a02c")
+
+wireframe_scurve_int <- ggplotly(wireframe_scurve, 
+                                width = "250", 
+                                height = "250", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
 
@@
 
-```{r, eval=knitr::is_latex_output()}
+```{r, eval=knitr::is_html_output()}
+
+crosstalk::bscols(
+    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr;",
+                   scurve_umap_plt_int, hex_grid_poly_scurve_int,
+                   hex_centroids_scurve_int, wireframe_scurve_int,
+                    htmltools::div(
+                      style = "margin-top: 13px;",  # adjust px as needed
+                      scurve_umap_model_vis
+                    )),
+    device = "xs"
+  )
+```
+
+```{r, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25}
 scurve_umap_plt + hex_grid_poly_scurve +
   hex_centroids_scurve + wireframe_scurve + 
+  scurve_proj_umap_model1 +
   plot_layout(ncol = 5)

### 2025-05-15 paper-quollr.Rmd
Added show_langevitour projections

@@
 
-```{r, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25}
+```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25}
 scurve_umap_plt + hex_grid_poly_scurve +
@@
 ```{r, eval=knitr::is_latex_output()}
-
+scurve_proj_umap_model1 + scurve_proj_umap_model2 + scurve_proj_umap_model3 + scurve_proj_umap_model4 + 
+  plot_layout(ncol = 4)
 ```

### 2025-05-19 paper-quollr.tex
Updated the algorithm steps
HTML version is working

@@
 Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.
-
-\section{Algorithm}\label{algorithm}
-
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, (6) identifying and removing low-density hexagons, and (7) lifting the model into higher dimensions.
-A detailed description of the algorithm can be found in \textbf{cite the methodology paper}.
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1}
-
-\section{Implementation}\label{implementation}
-
-\subsection{Installation}\label{installation}
-
-The development version can be installed from GitHub:
-
-\begin{verbatim}
-devtools::install_github("JayaniLakshika/quollr")
-\end{verbatim}
-
-\subsection{Usage}\label{usage}
-
-The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. We also want to load the built-in data sets \texttt{scurve} and \texttt{scurve\_umap}.
-
-\texttt{scurve} is a \(7\text{-}D\) simulated dataset.
-It is constructed by simulating \(5000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\).
-The remaining variables \(X_4\), \(X_5\), \(X_6\), \(X_7\) are all uniform error, with small variance.
-\texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors} is \(46\) and \texttt{min\_dist} is \(0.9\).
-Each data set contains a unique ID column that maps \texttt{scurve} and \texttt{scurve\_umap}.
-
-\subsubsection{Main function}\label{main-function}
-
-The mains steps for the algorithm can be executed by the main function \texttt{fit\_highd\_model()}, or can be run separately for more flexibility.
-
-This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{bin1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{benchmark\_highdens}).
-The function returns an object that includes the scaled NLDR object (\texttt{nldr\_obj}), the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
-
-\begin{verbatim}
-fit_highd_model(
-  highd_data = scurve, 
-  nldr_data = scurve_umap, 
-  bin1 = 15, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-\end{verbatim}
-
-\subsubsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
-
-Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) Remove low-density hexagons.
-
-\paragraph{Scaling the data}\label{scaling-the-data}
-
-The algorithm starts by scaling the NLDR data to a standard range using the \texttt{gen\_scaled\_data()} function.
-This function standardizes the data so that the first embedding ranges from \(0\) to \(1\), while the second embedding scales from \(0\) to the maximum value of the second embedding.
-The output includes the scaled NLDR data along with the original limits of the embeddings.
-
-\begin{verbatim}
-scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
-
-scurve_umap_obj
-\end{verbatim}
-
-\begin{verbatim}
-#> $scaled_nldr
-#> # A tibble: 5,000 x 3
-#>     emb1  emb2    ID
-#>    <dbl> <dbl> <int>
-#>  1 0.707 0.839     1
-#>  2 0.231 0.401     2
-#>  3 0.232 0.215     3
-#>  4 0.790 0.564     4
-#>  5 0.761 0.551     5
-#>  6 0.445 0.721     6
-#>  7 0.900 0.137     7
-#>  8 0.247 0.392     8
-#>  9 0.325 0.542     9
-#> 10 0.278 0.231    10
-#> # i 4,990 more rows
-#> 
-#> $lim1
-#> [1] -14.42166  13.32655
-#> 
-#> $lim2
-#> [1] -12.43687  12.32455
-\end{verbatim}
-
-\paragraph{Computing hexagon grid configurations}\label{computing-hexagon-grid-configurations}
-
-The configurations of a hexagonal grid are determined by the number of bins and the bin width in each direction.
-The function \texttt{calc\_bins\_y()} is used for this purpose.
-This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion (\texttt{q}).
-
-\begin{verbatim}
-bin_configs <- calc_bins_y(
-  nldr_obj = scurve_umap_obj, 
-  bin1 = 15, 
-  q = 0.1)
-
-bin_configs
-\end{verbatim}
-
-\begin{verbatim}
-#> $bin2
-#> [1] 16
-#> 
-#> $a1
-#> [1] 0.08326136
-#> 
-#> $a2
-#> [1] 0.07210645
-\end{verbatim}
-
-\paragraph{Binning the data}\label{binning-the-data}
-
-Points are allocated to bins based on the nearest centroid.
-The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its components can be run separately for added flexibility.
-The parameters used within \texttt{hex\_binning()} include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (\texttt{nldr\_obj}), the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion of the data (\texttt{q}).
-The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{tot\_bins}), the number of non-empty bins (\texttt{non\_bins}), and the points within each hexagon (\texttt{pts\_bins}).
-
-\begin{verbatim}
-hb_obj <- hex_binning(
-  nldr_obj = scurve_umap_obj, 
-  bin1 = 15, 
-  q = 0.1)
-\end{verbatim}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-hex-param-1} \caption{The components of the hexagon grid illustrating notation.}\label{fig:fig-hex-param}
-\end{figure}
-
-If the hexagonal binning process is run separately, it involves several steps: (i) generating all possible centroids in a hexagonal grid, (ii) creating the coordinates of the hexagons, (iii) assigning data points to their respective hexagons, (iv) computing the standardized number of points within each hexagon, and (v) mapping the points to their corresponding hexagonal bins.
-
-\subparagraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
-
-The \texttt{gen\_centroids()} function calculates the centroids of a hexagonal grid.
-
-The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
-The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagonal rows (\texttt{bin2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{bin1}) and buffer (\texttt{q}).
-
-Then, the centroids are computed iteratively.
-The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width.
-Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
-Vertical spacing (\texttt{a2}) is derived by \(\sqrt{3}/2 \times a_1\).
-
-The y-coordinates for each row are similarly calculated, and paired with the x-coordinates based on whether the total number of rows is even or odd.
-In the case of an odd number of rows, the final row uses only the odd-row x-coordinates to maintain the alternating pattern.
-
-Finally, a tibble is returned containing a unique hexagon ID (\texttt{hexID}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2-\text{D}\) space.
-
-\begin{verbatim}
-all_centroids_df <- gen_centroids(
-  nldr_obj = scurve_umap_obj, 
-  bin1 = 15, 
-  q = 0.1
-  )
-
-all_centroids_df
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 240 x 3
-#>    hexID     c_x     c_y
-#>    <int>   <dbl>   <dbl>
-#>  1     1 -0.1    -0.0892
-#>  2     2 -0.0167 -0.0892
-#>  3     3  0.0665 -0.0892
-#>  4     4  0.150  -0.0892
-#>  5     5  0.233  -0.0892
-#>  6     6  0.316  -0.0892
-#>  7     7  0.400  -0.0892
-#>  8     8  0.483  -0.0892
-#>  9     9  0.566  -0.0892
-#> 10    10  0.649  -0.0892
-#> # i 230 more rows
-\end{verbatim}
-
-\subparagraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
-
-Following the generation of hexagonal centroids, the \texttt{gen\_hex\_coord()} function constructs the coordinates of each hexagonal bin by defining its six polygonal vertices.
-These coordinates are used to visualize the hexagonal tessellation.
-
-Each hexagon is defined relative to its centroid \((C_x, C_y)\), with six vertices positioned equidistantly around the center.
-The function first verifies the presence of the required hexagon width parameter \texttt{a1}.
-This width determines the horizontal spacing.
-
-Two derived constants are calculated to define the relative distances to the vertices.
-The horizontal and vertical offset is defined as \(dx = \frac{a_1}{2}\), and \(dy = \frac{a_1}{\sqrt{3}}\) repectively.
-A vertical spacing factor \(v = \frac{a_1}{2\sqrt{3}}\) refines vertical placement in staggered rows.
-
-With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid.
-These offsets form two vectors (\texttt{x\_add\_factor} and \texttt{y\_add\_factor}) corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
-
-For each centroid, six vertices are computed and assigned a polygon ID (\texttt{hex\_poly\_id}) corresponding to the centroid's \texttt{hexID}.
-These points are aggregated into a tibble that includes the polygon ID (\texttt{hex\_poly\_id}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
-
-\begin{verbatim}
-all_hex_coord <- gen_hex_coord(
-  centroids_data = all_centroids_df, 
-  a1 = bin_configs$a1
-  )
-
-all_hex_coord
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 1,440 x 3
-#>    hex_poly_id       x       y
-#>          <int>   <dbl>   <dbl>
-#>  1           1 -0.1    -0.0412
-#>  2           1 -0.142  -0.0652
-#>  3           1 -0.142  -0.113 
-#>  4           1 -0.1    -0.137 
-#>  5           1 -0.0584 -0.113 
-#>  6           1 -0.0584 -0.0652
-#>  7           2 -0.0167 -0.0412
-#>  8           2 -0.0584 -0.0652
-#>  9           2 -0.0584 -0.113 
-#> 10           2 -0.0167 -0.137 
-#> # i 1,430 more rows
-\end{verbatim}
-
-\subparagraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
-
-After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The \texttt{assign\_data()} function performs this assignment by calculating the \(2-\text{D}\) Euclidean distance between each point in the \(2-\text{D}\) embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \(2-\text{D}\) layout.
-It then selects the corresponding x and y coordinates of each hexagon's centroid.
-
-Both the embedding coordinates and the centroid coordinates are converted to matrices to facilitate distance computations.
-The function uses the \texttt{proxy::dist()} method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
-For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon---and assigns the corresponding hexagon ID (\texttt{hexID}) to the point.
-
-The result is a data frame of the scaled \(2-\text{D}\) embedding with an additional \texttt{hexID} column, indicating the hexagonal bin to which each point belongs.
-
-\begin{verbatim}
-umap_hex_id <- assign_data(
-  nldr_obj = scurve_umap_obj, 
-  centroids_data = all_centroids_df
-  )
-
-umap_hex_id
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 5,000 x 4
-#>     emb1  emb2    ID hexID
-#>    <dbl> <dbl> <int> <int>
-#>  1 0.707 0.839     1   205
-#>  2 0.231 0.401     2   109
-#>  3 0.232 0.215     3    65
-#>  4 0.790 0.564     4   146
-#>  5 0.761 0.551     5   146
-#>  6 0.445 0.721     6   172
-#>  7 0.900 0.137     7    58
-#>  8 0.247 0.392     8   110
-#>  9 0.325 0.542     9   141
-#> 10 0.278 0.231    10    80
-#> # i 4,990 more rows
-\end{verbatim}
-
-\subparagraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
-
-The \texttt{compute\_std\_counts()} function calculates both the raw and standardized counts for each hexagon.
-
-The function begins by grouping the data by \texttt{hexID} and counting the number of NLDR points falling within each bin.
-These raw counts are stored as \texttt{bin\_counts}.
-To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin's count by the maximum count across all bins.
-This yields a normalized metric, \texttt{std\_counts}, ranging from 0 to 1.
-
-\begin{verbatim}
-std_df <- compute_std_counts(
-  scaled_nldr_hexid = umap_hex_id
-  )
-
-std_df
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 142 x 3
-#>    hexID bin_counts std_counts
-#>    <int>      <int>      <dbl>
-#>  1    21         12     0.16  
-#>  2    22         17     0.227 
-#>  3    23         22     0.293 
-#>  4    24         10     0.133 
-#>  5    25          7     0.0933
-#>  6    26         12     0.16  
-#>  7    27          1     0.0133
-#>  8    36         42     0.56  
-#>  9    37         42     0.56  
-#> 10    38         44     0.587 
-#> # i 132 more rows
-\end{verbatim}
-
-\subparagraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
-
-The \texttt{find\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon in the NLDR space.
-
-The function first groups the input data by \texttt{hexID}, which represents the hexagon label associated with each point in the \(2-\text{D}\) layout.
-Within each group, it collects the \texttt{ID}s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
-The resulting column, \texttt{pts\_list}, contains all point identifiers associated with that hexagon.
-
-\begin{verbatim}
-pts_df <- find_pts(
-  scaled_nldr_hexid = umap_hex_id
-  )
-
-pts_df
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 142 x 2
-#>    hexID pts_list  
-#>    <int> <list>    
-#>  1    21 <int [12]>
-#>  2    22 <int [17]>
-#>  3    23 <int [22]>
-#>  4    24 <int [10]>
-#>  5    25 <int [7]> 
-#>  6    26 <int [12]>
-#>  7    27 <int [1]> 
-#>  8    36 <int [42]>
-#>  9    37 <int [42]>
-#> 10    38 <int [44]>
-#> # i 132 more rows
-\end{verbatim}
-
-\paragraph{Obtaining bin centroids}\label{obtaining-bin-centroids}
-
-The \texttt{extract\_hexbin\_centroids()} function combines hexagonal bin coordinates, raw and standardized counts within each hexagons.
-
-This function begins by arranging the \texttt{counts\_data} by \texttt{hexID} to ensure consistent ordering.
-It then performs a full join with \texttt{centroids\_data}, aligning hexagon IDs between the two datasets to incorporate both hexagonal bin centroids and count metrics.
-After merging, the function handles missing values in the count columns: any \texttt{NA} values in \texttt{std\_counts} or \texttt{bin\_counts} are replaced with zeros.
-This ensures that hexagons with no assigned data points are retained in the output, with zero values for count-related fields.
-The resulting data frame contains the full set of hexagon centroids along with associated bin counts and standardized counts.
-
-\begin{verbatim}
-df_bin_centroids <- extract_hexbin_centroids(
-  centroids_data = all_centroids_df, 
-  counts_data = hb_obj$std_cts
-  )
-
-df_bin_centroids
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 240 x 5
-#>    hexID     c_x     c_y bin_counts std_counts
-#>    <int>   <dbl>   <dbl>      <dbl>      <dbl>
-#>  1     1 -0.1    -0.0892          0          0
-#>  2     2 -0.0167 -0.0892          0          0
-#>  3     3  0.0665 -0.0892          0          0
-#>  4     4  0.150  -0.0892          0          0
-#>  5     5  0.233  -0.0892          0          0
-#>  6     6  0.316  -0.0892          0          0
-#>  7     7  0.400  -0.0892          0          0
-#>  8     8  0.483  -0.0892          0          0
-#>  9     9  0.566  -0.0892          0          0
-#> 10    10  0.649  -0.0892          0          0
-#> # i 230 more rows
-\end{verbatim}
-
-\paragraph{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
-
-To represent the neighborhood structure of hexagonal bins in a \(2-\text{D}\) layout, we employ Delaunay triangulation on the centroids of hexagons.
-This geometric approach is used to infer which bins are considered neighbors.
-
-The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{tripack::tri.mesh()} function.
-This triangulation forms the structural basis for identifying adjacent bins.
-
-\begin{verbatim}
-tr_object <- tri_bin_centroids(
-  centroids_data = df_bin_centroids
-  )
-\end{verbatim}
-
-The \texttt{gen\_edges()} function uses this triangulation object to extract line segments between neighboring bins.
-It constructs a unique set of bin-to-bin connections by identifying the triangle edges and filtering duplicate or reversed links.
-Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function \texttt{calc\_2d\_dist()}.
-Only edges within a hexagon's neighborhood radius (based on the hexagon side length \texttt{a1}) are retained.
-
-\begin{verbatim}
-trimesh <- gen_edges(tri_object = tr_object, a1 = hb_obj$a1)
-
-trimesh
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 653 x 8
-#>     from    to  x_from  y_from    x_to    y_to from_count to_count
-#>    <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-#>  1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
-#>  2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
-#>  3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
-#>  4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
-#>  5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
-#>  6    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
-#>  7    31    46 -0.1     0.0550 -0.0584  0.127           0        0
-#>  8    32    47 -0.0167  0.0550  0.0249  0.127           0        0
-#>  9    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
-#> 10     4    18  0.150  -0.0892  0.108  -0.0171          0        0
-#> # i 643 more rows
-\end{verbatim}
-
-The \texttt{update\_trimesh\_index()} function re-indexes the node IDs to ensure that edge identifiers are sequentially numbered and consistent with downstream analysis.
-
-\begin{verbatim}
-trimesh <- update_trimesh_index(trimesh_data = trimesh)
-
-trimesh
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 653 x 8
-#>     from    to  x_from  y_from    x_to    y_to from_count to_count
-#>    <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-#>  1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
-#>  2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
-#>  3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
-#>  4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
-#>  5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
-#>  6    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
-#>  7    31    46 -0.1     0.0550 -0.0584  0.127           0        0
-#>  8    32    47 -0.0167  0.0550  0.0249  0.127           0        0
-#>  9    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
-#> 10     4    18  0.150  -0.0892  0.108  -0.0171          0        0
-#> # i 643 more rows
-\end{verbatim}
-
-\paragraph{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
-
-Not all hexagons contain meaningful information.
-Some may have very few or no data points due to the sparsity or shape of the underlying structure.
-Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the \(2-\text{D}\) structure, potentially disrupting the continuity of the representation.
-
-To address this, we propose a more nuanced method that evaluates each hexagon not only based on its own density, but also in the context of its immediate neighbors.
-The \texttt{find\_low\_dens\_hex()} function identifies hexagonal bins with insufficient local support by calculating the average standardized count across their six neighboring bins.
-If this mean neighborhood density is below a user-defined threshold (e.g., 0.05), the hexagon is flagged for removal.
-
-The \texttt{find\_low\_dens\_hex()} function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their \texttt{hexID} and a defined number of bins along the x-axis (\texttt{bin1}).
-The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.
-
-\begin{verbatim}
-find_low_dens_hex(
-  centroids_data = df_bin_centroids, 
-  bin1 = 15, 
-  benchmark_mean_dens = 0.05
-)
-\end{verbatim}
-
-\begin{verbatim}
-#>  [1]   1   2   3   4   5  11  12  13  14  15  16  17  18  19  30  31  32  33  46
-#> [20]  47  61  75  76  90 105 120 135 150 151 166 181 182 196 197 198 211 212 213
-#> [39] 214 215 225 226 227 228 229 230 231 232 233 234 235 236 237 238 239 240
-\end{verbatim}
-
-For simplicity, we remove low-density hexagons using a threshold of \(10\).
-
-\begin{verbatim}
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
-
-trimesh <- trimesh |>
-  dplyr::filter(from_count > 10,
-                to_count > 10)
-
-trimesh <- update_trimesh_index(trimesh)
-\end{verbatim}
-
-\subsubsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}
-
-The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each hexagonal bin to represent bin centroids.
-This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
-
-\begin{verbatim}
-df_bin <- avg_highd_data(
-  highd_data = scurve, 
-  scaled_nldr_hexid = hb_obj$data_hb_id
-)
-
-df_bin
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 142 x 8
-#>    hexID      x1    x2    x3         x4        x5       x6        x7
-#>    <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
-#>  1    21 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105 
-#>  2    22 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363
-#>  3    23 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153 
-#>  4    24 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128 
-#>  5    25  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195 
-#>  6    26  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172 
-#>  7    27  0.719   1.99 1.70   0.0109    -0.00349  -0.0297  -0.00223 
-#>  8    36 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338
-#>  9    37 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126
-#> 10    38 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780
-#> # i 132 more rows
-\end{verbatim}
-
-\subsubsection{Prediction}\label{prediction}
-
-The \texttt{predict\_emb()} function is used to predict \(2\text{-}D\) embedding for a new \(p\text{-}D\) data point using the fitted model.
-This function is useful to predict \(2\text{-}D\) embedding irrespective of the NLDR technique.
-
-In the prediction process, first, the nearest \(p\text{-}D\) model point is identified for a given new \(p\text{-}D\) data point by computing \(p\text{-}D\) Euclidean distance.
-Then, the corresponding \(2\text{-}D\) bin centroid mapping for the identified \(p\text{-}D\) model point is determined.
-Finally, the coordinates of the identified \(2\text{-}D\) bin centroid is used as the predicted NLDR embedding for the new \(p\text{-}D\) data point.
-
-\begin{verbatim}
-predict_emb(
-  highd_data = scurve, 
-  model_2d = df_bin_centroids, 
-  model_highd = df_bin
-  )
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 5,000 x 4
-#>    pred_emb_1 pred_emb_2    ID pred_hb_id
-#>         <dbl>      <dbl> <int>      <int>
-#>  1      0.691      0.848     1        205
-#>  2      0.191      0.416     2        109
-#>  3      0.316      0.199     3         66
-#>  4      0.774      0.560     4        146
-#>  5      0.774      0.560     5        146
-#>  6      0.441      0.704     6        172
-#>  7      0.941      0.127     7         58
-#>  8      0.275      0.416     8        110
-#>  9      0.358      0.560     9        141
-#> 10      0.316      0.343    10         96
-#> # i 4,990 more rows
-\end{verbatim}
-
-\subsubsection{Compute residuals and Root Mean Square Error (RMSE)}\label{compute-residuals-and-root-mean-square-error-rmse}
-
-As a Goodness of fit statistics for the model, \texttt{glance()} is used to compute residuals and RMSE.
-These metrics are used to assess how well the fitted model will capture the underlying structure of the \(p\text{-}D\) data.
-
-This function begins by renaming the columns of the input \texttt{model\_highd} data frame to avoid naming conflicts during subsequent joins.
-It then uses the \texttt{predict\_emb()} function to assign each point in the high-dimensional dataset to a corresponding bin in the 2D model, producing a prediction data frame that contains both the predicted bin assignment (\texttt{pred\_hb\_id}) and the original observation \texttt{ID}.
-
-The function joins this prediction output with both the high-dimensional model (to get mean bin coordinates in the original space) and the original high-dimensional data (to retrieve true coordinates).
-It then calculates squared differences between the true and predicted high-dimensional coordinates for each dimension, storing these as \texttt{error\_square\_x1}, \texttt{error\_square\_x2}, \ldots, up to the dimensionality of the data.
-
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RMSE which is the average of the total squared error per point.
-
-These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{RMSE} (root mean squared error).
-
-\begin{verbatim}
-glance(
-  highd_data = scurve, 
-  model_2d = df_bin_centroids, 
-  model_highd = df_bin
-  )
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 1 x 2
-#>   Error  RMSE
-#>   <dbl> <dbl>
-#> 1 1481. 0.180
-\end{verbatim}
-
-Furthermore, \texttt{augment()} accepts \(2\text{-}D\) and \(p\text{-}D\) model points, and the \(p\text{-}D\) data and adds information about each observation in the data set.
-
-The function starts by renaming columns in the \texttt{model\_highd} data frame to avoid conflicts.
-It then predicts the high-dimensional bin assignments for each point using the \texttt{predict\_emb()} function, mapping each observation to its nearest \(2-\text{D}\) bin (\texttt{pred\_hb\_id}).
-This prediction is joined with the mean high-dimensional coordinates of each bin from the model and with the original high-dimensional data.
-
-Next, the function computes residuals between each original coordinate (\texttt{x1}, \texttt{x2}, \ldots, \texttt{xp}) and the corresponding modeled coordinate (\texttt{model\_high\_d\_x1}, \ldots, \texttt{model\_high\_d\_xp}) across all dimensions.
-It calculates both squared errors and absolute errors per dimension.
-These are used to compute two aggregate diagnostic measures per point.
-First, the \texttt{row\_wise\_total\_error} which is the total squared error across all dimensions, and the \texttt{row\_wise\_abs\_error} which is the total absolute error across all dimensions.
-
-The final output is a data frame that combines the original IDs, high-dimensional coordinates, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
-The augmented dataset is always returned as a \texttt{tibble::tibble} with the same number of rows as the passed dataset.
-
-\begin{verbatim}
-model_error <- augment(
-  highd_data = scurve, 
-  model_2d = df_bin_centroids, 
-  model_highd = df_bin
-  )
-
-model_error
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 5,000 x 32
-#>       ID      x1     x2       x3       x4       x5       x6        x7 pred_hb_id
-#>    <int>   <dbl>  <dbl>    <dbl>    <dbl>    <dbl>    <dbl>     <dbl>      <int>
-#>  1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923  -0.00128         205
-#>  2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362  -0.00535         109
-#>  3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962   0.00335          66
-#>  4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716   0.00126         146
-#>  5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998   0.000677        146
-#>  6     6  0.818  0.927  -1.58    -0.00318 -0.00980  0.0989   0.00696         172
-#>  7     7  0.910  1.40    1.42     0.00699 -0.0182  -0.0710   0.00966          58
-#>  8     8 -0.0691 1.59    0.00239  0.0127  -0.0130   0.0396  -0.000185        110
-#>  9     9  0.859  1.59   -0.488   -0.0119   0.00421 -0.00440 -0.00595         141
-#> 10    10 -0.727  1.62    0.314    0.00251  0.0177  -0.0755  -0.00369          96
-#> # i 4,990 more rows
-#> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
-#> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
-#> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
-#> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
-#> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
-#> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>, ...
-\end{verbatim}
-
-\subsubsection{Visualizations}\label{visualizations}
-
-The package offers several \(2-\text{D}\) visualizations, including:
-
-\begin{itemize}
-\tightlist
-\item
-  A full hexagonal grid,
-\item
-  A hexagonal grid that matches the data,
-\item
-  A full grid based on centroid triangulation,
-\item
-  A centroid triangulation grid that aligns with the data,
-\item
-  A triangular mesh for any provided set of points.
-\end{itemize}
-
-The generated p-D model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
-Additionally, it features a function for visualizing the \(2-\text{D}\) projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
-
-Furthermore, there are two interactive plots, \texttt{show\_link\_plots} and \texttt{show\_error\_link\_plots}, which are designed to help diagnose the model.
-
-Each visualization can be generated using its respective function, as described in this section.
-
-\paragraph{Hexagonal grid}\label{hexagonal-grid}
-
-The \texttt{geom\_hexgrid()} function introduces a custom \texttt{ggplot2} layer designed for visualizing hexagonal grid on a provided set of bin centroids.
-
-To display the complete grid, users should supply all available bin centroids.
-
-\begin{verbatim}
-ggplot() + 
-  geom_hexgrid(
-    data = hb_obj$centroids, 
-    aes(x = c_x, y = c_y)
-    ) 
-\end{verbatim}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
-
-If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
-
-\begin{verbatim}
-ggplot() + 
-  geom_hexgrid(
-    data = df_bin_centroids, 
-    aes(x = c_x, y = c_y)
-    ) 
-\end{verbatim}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-26-1}
-
-\paragraph{Triangular mesh}\label{triangular-mesh}
-
-The \texttt{geom\_trimesh()} function introduces a custom \texttt{ggplot2} layer designed for visualizing \(2\text{-}D\) wireframe on a provided set of bin centroids.
-
-To display the complete wireframe, users should supply all available bin centroids.
-
-\begin{verbatim}
-ggplot() + 
-  geom_trimesh(
-    data = hb_obj$centroids, 
-    aes(x = c_x, y = c_y)
-    ) 
-\end{verbatim}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
-
-If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
-
-\begin{verbatim}
-ggplot() + 
-  geom_trimesh(
-    data = df_bin_centroids, 
-    aes(x = c_x, y = c_y)
-    ) 
-\end{verbatim}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
-
-\paragraph{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
-
-To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualization using the \texttt{show\_langevitour()} function.
-This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.
-
-Before plotting, the data needs to be organized into a combined format through the \texttt{comb\_data\_model()} function.
-This function takes three inputs: \texttt{highd\_data} (the high-dimensional observations), \texttt{model\_highd} (high-dimensional summaries for each bin), and \texttt{model\_2d} (the hexagonal bin centroids of the model).
-It returns a tidy data frame combining both the data and the model.
-
-In this structure, the \texttt{type} variable distinguishes between original observations (\texttt{"data"}) and the bin-averaged model representation (\texttt{"model"}).
-
-\begin{verbatim}
-df_exe <- comb_data_model(
-  highd_data = scurve, 
-  model_highd = df_bin, 
-  model_2d = df_bin_centroids
-  )
-
-df_exe
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 5,122 x 8
-#>         x1    x2    x3         x4        x5       x6        x7 type 
-#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr>
-#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model
-#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model
-#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model
-#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model
-#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model
-#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model
-#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model
-#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model
-#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model
-#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model
-#> # i 5,112 more rows
-\end{verbatim}
-
-The \texttt{show\_langevitour()} function then renders the visualization using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour.
-The \texttt{edge\_data} input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model's structure.
-
-In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2-\text{D}\) wireframe structure mapped back to high-dimensional space.
-
-\begin{verbatim}
-show_langevitour(
-  point_data = df_exe, 
-  edge_data = trimesh
-  )
-\end{verbatim}
-
-\paragraph{Link plots}\label{link-plots}
-
-There are mainly two interactive link plots can be generated.
-
-To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
-This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
-Both views are interactively linked, enabling users to explore.
-
-Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
-This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2-\text{D}\) bin layout (\texttt{model\_2d}):
-
-This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.
-
-\begin{verbatim}
-df_exe <- comb_all_data_model(
-  highd_data = scurve, 
-  nldr_data = scurve_umap, 
-  model_highd = df_bin, 
-  model_2d = df_bin_centroids
-  )
-\end{verbatim}
-
-The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a \(2-\text{D}\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
-The function takes the output from \texttt{comb\_all\_data\_model()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
-
-These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the \texttt{langevitour} output.
-
-\begin{verbatim}
-show_link_plots(
-  point_data = df_exe, 
-  edge_data = trimesh
-  )
-\end{verbatim}
-
-\texttt{show\_error\_link\_plots()} helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
-
-Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
-The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2-\text{D}\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
-
-This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.
-
-\begin{verbatim}
-df_exe <- comb_all_data_model_error(
-  highd_data = scurve, 
-  nldr_data = scurve_umap, 
-  model_highd = df_bin, 
-  model_2d = df_bin_centroids, 
-  error_data = model_error
-  )
-
-df_exe
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 5,122 x 12
-#>         x1    x2    x3         x4        x5       x6        x7 type   emb1  emb2
-#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr> <dbl> <dbl>
-#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model    NA    NA
-#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model    NA    NA
-#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model    NA    NA
-#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model    NA    NA
-#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model    NA    NA
-#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model    NA    NA
-#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model    NA    NA
-#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model    NA    NA
-#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model    NA    NA
-#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model    NA    NA
-#> # i 5,112 more rows
-#> # i 2 more variables: sqrt_row_wise_total_error <dbl>, density <dbl>
-\end{verbatim}
-
-The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2-\text{D}\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
-The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
-
-These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
-This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
-
-These three views are linked using \texttt{crosstalk}, allowing interactive selection of points in error plot and the NLDR plot to highlight corresponding structures in the \texttt{langevitour} output.
-
-\begin{verbatim}
-show_error_link_plots(
-  point_data = df_exe, 
-  edge_data = trimesh
-  )
-\end{verbatim}
-
-\section{Application}\label{application}
-
-Single-cell RNA sequencing (scRNA-seq) is a popular and powerful technology that allows you to profile the whole transcriptome of a large number of individual cells \citep{andrews2021}.
-
-Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
-NLDR often used to summarise the discovered clusters, and help to understand the results.
-The purpose of this example is to \emph{illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data}.
-
-Limb muscle cells of mice in \citet{tabula2018} are examined.
-There are \(1067\) single cells, with \(14997\) gene expressions.
-Following their pre-processing, tSNE was performed using ten principal components.
-Figure \ref{fig:limb-rmse} (b) is the reproduction of the published plot.
-The question is whether this accurately represents the cluster structure in the data.
-Our method can help here, and also help to provide a better \(2-\text{D}\) layout, as needed.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
-Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.}\label{fig:limb-rmse}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \\ref{fig:limb-rmse} b) and the $2-\text{D}$ layout selected (Figure \\ref{fig:limb-rmse} f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}
-\end{figure}
-
-\section{Discussion}\label{discussion}
-
-This paper presents the R package \texttt{quollr} to develop a way to take the fitted model, as represented by the positions of points in \(2\text{-}D\), and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.
-
-The paper includes a clustering example to illustrate how \texttt{quollr} is useful to assess which NLDR technique and which (hyper)parameter choice gives the most accurate representation.
-In addition, how to select parameters for hexagonal binning and fitting model are explained.
-
-Possible future improvements would be\ldots{}
-
-This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of \(p\text{-}D\) data.
-
-\section{Acknowledgements}\label{acknowledgements}
-
-This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
-These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.
-
-The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.
 

### 2025-05-19 paper-quollr.R

@@ -380,753 +380,3 @@ hex_grid_scurve + wrap_plots(
   scurve_proj_umap_model3, scurve_proj_umap_model4, 
   ncol = 2)
 
-
-## ----eval=knitr::is_html_output()---------------------------------------------
-# 
-# crosstalk::bscols(
-#     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr;",
-#                    scurve_umap_plt_int, hex_grid_poly_scurve_int,
-#                    hex_centroids_scurve_int, wireframe_scurve_int,
-#                     htmltools::div(
-#                       style = "margin-top: 13px;",  # adjust px as needed
-#                       scurve_umap_model_vis
-#                     )),
-#     device = "xs"
-#   )
-
-
-## ----eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25----
-scurve_umap_plt + hex_grid_poly_scurve +
-  hex_centroids_scurve + wireframe_scurve + 
-  scurve_proj_umap_model1 +
-  plot_layout(ncol = 5)
-
-
-## ----echo=TRUE, eval=FALSE----------------------------------------------------
-# fit_highd_model(
-#   highd_data = scurve,
-#   nldr_data = scurve_umap,
-#   bin1 = 15,
-#   q = 0.1,
-#   benchmark_highdens = 5)
-
-
-## ----echo=TRUE----------------------------------------------------------------
-scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
-
-scurve_umap_obj
-
-
-## ----echo=TRUE----------------------------------------------------------------
-bin_configs <- calc_bins_y(
-  nldr_obj = scurve_umap_obj, 
-  bin1 = 15, 
-  q = 0.1)
-
-bin_configs
-
-
-## ----echo=TRUE----------------------------------------------------------------
-hb_obj <- hex_binning(
-  nldr_obj = scurve_umap_obj, 
-  bin1 = 15, 
-  q = 0.1)
-
-
-## -----------------------------------------------------------------------------
-#| label: code-illustration
-# Code to draw illustration for notation
-## hexagon binning to have regular hexagons
-hb_obj_notation <- hex_binning(
-  nldr_obj = scurve_umap_obj, 
-  bin1 = 9, 
-  q = 0.1)
-
-a1_temp <- hb_obj_notation$a1
-a2_temp <- hb_obj_notation$a2
[…]

### 2025-05-19 paper-quollr.Rmd
HTML version is working
Added figure position

@@
   \usepackage{array}
+  \usepackage{float}
 output: 
@@
 
-```{r, eval=knitr::is_html_output()}
+```{r, eval=knitr::is_html_output(), fig.cap="algorithm"}
 
@@
 
-```{r, eval=knitr::is_latex_output()}
+```{r, eval=knitr::is_latex_output(), fig.cap="algorithm", fig.pos='H'}
 hex_grid_scurve + wrap_plots(
@@
 
-```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25}
+```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H'}
 scurve_umap_plt + hex_grid_poly_scurve +

### 2025-05-19 paper-quollr.tex
Fixed the issue of rendering pdf (removed \maketitle)
Added figure position

--- 2025-05-15 paper-quollr.tex
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-3-1}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-3-1} \caption{algorithm}\label{fig:unnamed-chunk-3}
+\end{figure}
 
@@
 
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}
+
 \paragraph{Link plots}\label{link-plots}

### 2025-05-19 paper-quollr.R

@@ -233,7 +233,7 @@ scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
 
 
 
-## ----eval=knitr::is_html_output()---------------------------------------------
+## ----eval=knitr::is_html_output(), fig.cap="algorithm"------------------------
 # 
 # crosstalk::bscols(
 #     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
@@ -374,7 +374,7 @@ scurve_proj_umap_model4 <- plot_proj(
   interior_annotation(label = "a4", cex = 1) 
 
 
-## ----eval=knitr::is_latex_output()--------------------------------------------
+## ----eval=knitr::is_latex_output(), fig.cap="algorithm", fig.pos='H'----------
 hex_grid_scurve + wrap_plots(
   scurve_proj_umap_model1, scurve_proj_umap_model2,
   scurve_proj_umap_model3, scurve_proj_umap_model4, 
@@ -395,7 +395,7 @@ hex_grid_scurve + wrap_plots(
 #   )
 
 
-## ----eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25----
+## ----eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H'----
 scurve_umap_plt + hex_grid_poly_scurve +
   hex_centroids_scurve + wireframe_scurve + 
   scurve_proj_umap_model1 +

### 2025-05-19 paper-quollr.Rmd
Updated algorithm steps figure

@@
 
+sc_xlims <- c(-0.15, 1.15)
+sc_ylims <- c(-0.14, 1.05)
+
 scurve_umap_plt <- ggplot(
@@
   aes(x = emb1, y = emb2)) +
-  geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
 
 scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
-                                width = "250", 
-                                height = "250", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
@@
              aes(x = emb1, y = emb2), 
-             alpha = 0.2, size = 0.5, color = clr_choice)
+             alpha = 0.2, size = 0.5, color = clr_choice) + 
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
 
 hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
-                                width = "250", 
-                                height = "250", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
@@
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c")
+             size = 1, color = "#33a02c") +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
 
 hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
-                                width = "250", 
-                                height = "250", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
@@
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c")
+             size = 1, color = "#33a02c") + 
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
 
 wireframe_scurve_int <- ggplotly(wireframe_scurve, 
-                                width = "250", 
-                                height = "250", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>

### 2025-05-19 paper-quollr.R

@@ -85,14 +85,19 @@ hex_grid_with_counts_scurve <- left_join(
 hex_grid_nonempty_scurve <- hex_grid_scurve |>
   filter(hex_poly_id %in% df_bin_centroids_scurve$hexID)
 
+sc_xlims <- c(-0.15, 1.15)
+sc_ylims <- c(-0.14, 1.05)
+
 scurve_umap_plt <- ggplot(
   scurve_umap_scaled, 
   aes(x = emb1, y = emb2)) +
-  geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
 
 scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
-                                width = "250", 
-                                height = "250", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
   config(
     staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
@@ -111,11 +116,13 @@ hex_grid_poly_scurve <- ggplot(
                fill = "#ffffff") +
   geom_point(data = scurve_umap_scaled, 
              aes(x = emb1, y = emb2), 
-             alpha = 0.2, size = 0.5, color = clr_choice)
+             alpha = 0.2, size = 0.5, color = clr_choice) + 
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
 
 hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
-                                width = "250", 
-                                height = "250", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
   config(
     staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
@@ -134,11 +141,13 @@ hex_centroids_scurve <- ggplot(
                fill = "#ffffff")  +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c")
+             size = 1, color = "#33a02c") +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
 
 hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
-                                width = "250", 
-                                height = "250", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
   config(
[…]

### 2025-05-19 paper-quollr.Rmd
Updated algorithm step figures

@@
 scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
-                                width = "350", 
-                                height = "350", 
+                                width = "230", 
+                                height = "230", 
                                 tooltip = "none") |>
@@
 hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
-                                width = "350", 
-                                height = "350", 
+                                width = "230", 
+                                height = "230", 
                                 tooltip = "none") |>
@@
 hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
-                                width = "350", 
-                                height = "350", 
+                                width = "230", 
+                                height = "230", 
                                 tooltip = "none") |>
@@
 wireframe_scurve_int <- ggplotly(wireframe_scurve, 
-                                width = "350", 
-                                height = "350", 
+                                width = "230", 
+                                height = "230", 
                                 tooltip = "none") |>
@@
 
+scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = tr_from_to_df_scurve$from,
+                         lineTo = tr_from_to_df_scurve$to,
+                         group = df_exe$type,
+                         pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
+                         levelColors = c("#000000", "#33a02c"),
+                         enableControls = FALSE,
+                         width = "300px", height = "300px")
+
 ```
@@
                       style = "margin-top: 13px;",  # adjust px as needed
-                      scurve_umap_model_vis
+                      scurve_umap_model_vis_n
                     )),

### 2025-05-19 paper-quollr.Rmd
Need to find a way to remove the padding

@@
 crosstalk::bscols(
-    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr;",
-                   scurve_umap_plt_int, hex_grid_poly_scurve_int,
-                   hex_centroids_scurve_int, wireframe_scurve_int,
+    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr; gap: 0;",
+    htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_plt_int),
+    htmltools::div(style = "margin: 0; padding: 0;", hex_grid_poly_scurve_int),
+    htmltools::div(style = "margin: 0; padding: 0;", hex_centroids_scurve_int),
+    htmltools::div(style = "margin: 0; padding: 0;", wireframe_scurve_int),
     htmltools::div(
-                      style = "margin-top: 13px;",  # adjust px as needed
+      style = "margin-top: 13px;",
       scurve_umap_model_vis_n

### 2025-05-19 paper-quollr.R

@@ -404,13 +404,15 @@ hex_grid_scurve + wrap_plots(
 ## ----eval=knitr::is_html_output()---------------------------------------------
 # 
 # crosstalk::bscols(
-#     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr;",
-#                    scurve_umap_plt_int, hex_grid_poly_scurve_int,
-#                    hex_centroids_scurve_int, wireframe_scurve_int,
-#                     htmltools::div(
-#                       style = "margin-top: 13px;",  # adjust px as needed
-#                       scurve_umap_model_vis_n
-#                     )),
+#     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr; gap: 0;",
+#     htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_plt_int),
+#     htmltools::div(style = "margin: 0; padding: 0;", hex_grid_poly_scurve_int),
+#     htmltools::div(style = "margin: 0; padding: 0;", hex_centroids_scurve_int),
+#     htmltools::div(style = "margin: 0; padding: 0;", wireframe_scurve_int),
+#     htmltools::div(
+#       style = "margin-top: 13px;",
+#       scurve_umap_model_vis_n
+#     )),
 #     device = "xs"
 #   )
 

### 2025-05-19 paper-quollr.Rmd
Updated Figure 2 static image

--- 2025-05-19 paper-quollr.Rmd
@@
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
+  interior_annotation("a") +
   xlim(sc_xlims) +
@@
              alpha = 0.2, size = 0.5, color = clr_choice) + 
+  interior_annotation("b") +
   xlim(sc_xlims) +
@@
              size = 1, color = "#33a02c") +
+  interior_annotation("c") +
   xlim(sc_xlims) +
@@
              size = 1, color = "#33a02c") +
+  interior_annotation("d") +
   xlim(sc_xlims) +
@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, (6) identifying and removing low-density hexagons, and (7) lifting the model into higher dimensions.
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions.
 A detailed description of the algorithm can be found in **cite the methodology paper**.
@@
 crosstalk::bscols(
-    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr;",
-                   scurve_umap_plt_int, hex_grid_poly_scurve_int,
-                   hex_centroids_scurve_int, wireframe_scurve_int,
+    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr; gap: 0;",
+    htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_plt_int),
+    htmltools::div(style = "margin: 0; padding: 0;", hex_grid_poly_scurve_int),
+    htmltools::div(style = "margin: 0; padding: 0;", hex_centroids_scurve_int),
+    htmltools::div(style = "margin: 0; padding: 0;", wireframe_scurve_int),
     htmltools::div(
-                      style = "margin-top: 13px;",  # adjust px as needed
+      style = "margin-top: 13px;t",
       scurve_umap_model_vis_n
@@
 
-```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H'}
+```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout ($k=2$): (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+The `Scurve` data is shown."}
 scurve_umap_plt + hex_grid_poly_scurve +

### 2025-05-19 paper-quollr.tex
Updated Figure 2 static image

@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, (6) identifying and removing low-density hexagons, and (7) lifting the model into higher dimensions.
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions.
 A detailed description of the algorithm can be found in \textbf{cite the methodology paper}.
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1} \caption{Key steps for constructing the model on the UMAP layout ($k=2$): (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+The `Scurve` data is shown.}\label{fig:unnamed-chunk-5}
+\end{figure}
 

### 2025-05-19 paper-quollr.Rmd
Figure out how to handle reference with htmltool object

@@
 library(crosstalk)
+library(htmltools)
 #library(tools)
@@
 
-This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
+This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data ([Figure 1](#fig-algo)).
 
@@
 
-```{r, eval=knitr::is_html_output(), fig.cap="algorithm"}
-
+```{r, eval=knitr::is_html_output()}
+
+htmltools::browsable(
+  tagList(
+    tags$div(
+      id = "fig-algo",  # This is the anchor
       crosstalk::bscols(
-    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+        htmltools::div(
+          style = "display: grid; grid-template-columns: 1fr 1fr;",
           hex_grid_scurve_int,
-                    htmltools::div(
-                      style = "margin-top: 13px;",  # adjust px as needed
-                      scurve_umap_model_vis
-                    )),
+          htmltools::div(style = "margin-top: 13px;", scurve_umap_model_vis)
+        ),
         device = "xs"
-  )
+      ),
+      tags$p(tags$strong("Figure 1:"), " Algorithm to visualize the 2D model.")
+    )
+  )
+)
+
 ```

### 2025-05-19 paper-quollr.tex
Figure out how to handle reference with htmltool object

@@
 
-This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
+This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data (\hyperref[fig-algo]{Figure 1}).
 

### 2025-05-19 paper-quollr.Rmd
Added Figure title

@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions.
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions ([Figure 2](#fig-algo-steps)).
 A detailed description of the algorithm can be found in **cite the methodology paper**.
@@
 
+htmltools::browsable(
+  tagList(
+    tags$div(
+      id = "fig-algo-steps",  # This is the anchor
       crosstalk::bscols(
@@
         device = "xs"
-  )
-```
-
-```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout ($k=2$): (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+      ),
+      tags$p(tags$strong("Figure 2:"), "Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+The `Scurve` data is shown.")
+    )
+  )
+)
+```
+
+```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
 The `Scurve` data is shown."}

### 2025-05-19 paper-quollr.tex
Added Figure title

@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions.
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (\hyperref[fig-algo-steps]{Figure 2}).
 A detailed description of the algorithm can be found in \textbf{cite the methodology paper}.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1} \caption{Key steps for constructing the model on the UMAP layout ($k=2$): (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1} \caption{Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
 The `Scurve` data is shown.}\label{fig:unnamed-chunk-5}

### 2025-05-19 paper-quollr.Rmd
Fixed the references issue for HTML and PDF outputs

@@
 
-This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data ([Figure 1](#fig-algo)).
+This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data (`r knitr::asis_output(if (knitr::is_html_output()) { "[Figure 1](#fig-overview)"} else { "Figure \\@ref(fig:overview)"})`).
 
@@
     tags$div(
-      id = "fig-algo",  # This is the anchor
+      id = "fig-overview",  # This is the anchor
       crosstalk::bscols(
@@
 
-```{r, eval=knitr::is_latex_output(), fig.cap="algorithm", fig.pos='H'}
+```{r overview, eval=knitr::is_latex_output(), fig.cap="algorithm", fig.pos='H'}
 hex_grid_scurve + wrap_plots(
@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions ([Figure 2](#fig-algo-steps)).
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "[Figure 2](#fig-algo-steps)"} else { "Figure \\@ref(fig:algo-steps)"})`).
 A detailed description of the algorithm can be found in **cite the methodology paper**.
@@
 
-```{r, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
 The `Scurve` data is shown."}

### 2025-05-19 paper-quollr.tex
Fixed the references issue for HTML and PDF outputs

--- 2025-05-19 paper-quollr.tex
@@
 
-This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
+This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data (Figure \ref{fig:overview}).
 
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-3-1} \caption{algorithm}\label{fig:unnamed-chunk-3}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/overview-1} \caption{algorithm}\label{fig:overview}
 \end{figure}
@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions.
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (Figure \ref{fig:algo-steps}).
 A detailed description of the algorithm can be found in \textbf{cite the methodology paper}.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-5-1} \caption{Key steps for constructing the model on the UMAP layout ($k=2$): (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
-The `Scurve` data is shown.}\label{fig:unnamed-chunk-5}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+The `Scurve` data is shown.}\label{fig:algo-steps}
 \end{figure}
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-26-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-26-1}
 
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-29-1}
 

### 2025-05-19 paper-quollr.Rmd
Added sentence for main model approach

@@
 
-This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data (`r knitr::asis_output(if (knitr::is_html_output()) { "[Figure 1](#fig-overview)"} else { "Figure \\@ref(fig:overview)"})`).
+This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
+Starting with an NLDR layout, our approach is to create a wireframe model representation, that can be lifted and displayed in the high-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "[Figure 1](#fig-overview)"} else { "Figure \\@ref(fig:overview)"})`).
 

### 2025-05-19 paper-quollr.tex
Added sentence for main model approach

@@
 
-This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data (Figure \ref{fig:overview}).
+This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
+Starting with an NLDR layout, our approach is to create a wireframe model representation, that can be lifted and displayed in the high-dimensional space (Figure \ref{fig:overview}).
 

### 2025-05-19 paper-quollr.tex
Updated hex_param_vis configurations

@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-hex-param-1} \caption{The components of the hexagon grid illustrating notation.}\label{fig:fig-hex-param}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.5\linewidth]{paper-quollr_files/figure-latex/hex-param-1}
+
+}
+
+\caption{The components of the hexagon grid illustrating notation.}\label{fig:hex-param}
 \end{figure}

### 2025-05-19 paper-quollr.R

@@ -628,9 +628,7 @@ hex_param_vis <- ggplot() +
   coord_equal()
 
 
-## -----------------------------------------------------------------------------
-#| label: fig-hex-param
-#| fig-cap: "The components of the hexagon grid illustrating notation."
+## ----hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos='H'----
 
 hex_param_vis
 

### 2025-05-19 paper-quollr.Rmd
Updated hex_param_vis configurations
Updated Figure 4

@@
 
-```{r}
-#| label: fig-hex-param
-#| fig-cap: "The components of the hexagon grid illustrating notation."
+```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos='H'}
 
@@
 ```{r, echo=TRUE}
-ggplot() + 
+full_hexgrid <- ggplot() + 
   geom_hexgrid(
@@
 ```{r, echo=TRUE}
-ggplot() + 
+data_hexgrid <- ggplot() + 
   geom_hexgrid(
@@
 ```{r, echo=TRUE}
-ggplot() + 
+full_triangulation_grid <- ggplot() + 
   geom_trimesh(
@@
 ```{r, echo=TRUE}
-ggplot() + 
+data_triangulation_grid <- ggplot() + 
   geom_trimesh(
@@
     ) 
+```
+
+```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', eval=knitr::is_html_output()}
+
+full_hexgrid + data_hexgrid + 
+  full_triangulation_grid + data_triangulation_grid +
+  plot_layout(ncol = 4)
+```
+
+```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', eval=knitr::is_latex_output()}
+
+full_hexgrid + data_hexgrid + 
+  full_triangulation_grid + data_triangulation_grid +
+  plot_layout(ncol = 4)
 ```

### 2025-05-19 paper-quollr.tex
Updated Figure 4

@@
 \begin{verbatim}
-ggplot() + 
+full_hexgrid <- ggplot() + 
   geom_hexgrid(
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-23-1}
-
 If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
@@
 \begin{verbatim}
-ggplot() + 
+data_hexgrid <- ggplot() + 
   geom_hexgrid(
@@
 \end{verbatim}
-
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}
 
@@
 \begin{verbatim}
-ggplot() + 
+full_triangulation_grid <- ggplot() + 
   geom_trimesh(
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-25-1}
-
 If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
@@
 \begin{verbatim}
-ggplot() + 
+data_triangulation_grid <- ggplot() + 
   geom_trimesh(
@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-26-1}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/geom-outputs-pdf-1}
+
+}
+
+\caption{The outputs of `geom\_hexgrid` and `geom\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.}\label{fig:geom-outputs-pdf}
+\end{figure}
 

### 2025-05-19 paper-quollr.R

@@ -761,7 +761,7 @@ model_error
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-ggplot() + 
+full_hexgrid <- ggplot() + 
   geom_hexgrid(
     data = hb_obj$centroids, 
     aes(x = c_x, y = c_y)
@@ -769,7 +769,7 @@ ggplot() +
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-ggplot() + 
+data_hexgrid <- ggplot() + 
   geom_hexgrid(
     data = df_bin_centroids, 
     aes(x = c_x, y = c_y)
@@ -777,7 +777,7 @@ ggplot() +
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-ggplot() + 
+full_triangulation_grid <- ggplot() + 
   geom_trimesh(
     data = hb_obj$centroids, 
     aes(x = c_x, y = c_y)
@@ -785,11 +785,25 @@ ggplot() +
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-ggplot() + 
+data_triangulation_grid <- ggplot() + 
   geom_trimesh(
     data = df_bin_centroids, 
     aes(x = c_x, y = c_y)
     ) 
+
+
+## ----geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', eval=knitr::is_html_output()----
+# 
+# full_hexgrid + data_hexgrid +
+#   full_triangulation_grid + data_triangulation_grid +
+#   plot_layout(ncol = 4)
+
+
+## ----geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', eval=knitr::is_latex_output()----
+
+full_hexgrid + data_hexgrid + 
+  full_triangulation_grid + data_triangulation_grid +
+  plot_layout(ncol = 4)
 
 
 ## ----echo=TRUE----------------------------------------------------------------

### 2025-05-19 paper-quollr.Rmd
Updated trimesh figure configurations

@@
 
-```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', eval=knitr::is_html_output()}
+```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_html_output()}
 
@@
 
-```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', eval=knitr::is_latex_output()}
+```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_latex_output()}
 
@@
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.", out.width="100%"}
+Layout f is the best choice.", fig.pos='H', out.width="100%"}
 

### 2025-05-19 paper-quollr.Rmd
Updated Figure 6 HTML output

@@
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.", fig.pos='H', out.width="100%"}
-
-free(error_plot_limb) + wrap_plots(
-  nldr1, nldr2, nldr3, 
+Layout f is the best choice.", fig.pos='H'}
+
+error_plot_limb + 
+  wrap_plots(nldr1, nldr2, nldr3, 
   nldr4, nldr5, nldr6, ncol = 2) 
@@
 
+trimesh_limb_int <- ggplotly(trimesh_limb, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 # hex_grid <- tsne_limb_obj$hb_obj$hex_poly
@@
 
-langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
+limb_highd_vis <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
                          lineFrom = trimesh_data_limb$from,
@@
                                         c("0", "1", "2", "3", "4", "5", "6", "model")),
-                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"))
+                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
+                         enableControls = FALSE,
+                         width = "421px", height = "421px")
 ```
@@
 
+trimesh_limb_best_int <- ggplotly(trimesh_limb_best, 
+                                width = "350", 
+                                height = "350", 
+                                tooltip = "none") |>
+  config(
+    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
+    displayModeBar = FALSE,   # Hides the plotly toolbar
+    editable = FALSE,         # Disables annotations and editing
+    showTips = FALSE,         # Removes tooltip on hover
+    displaylogo = FALSE,      # Hides plotly logo
+    responsive = FALSE        # Disables responsive resizing
+  )
+
 # hex_grid <- tsne_best_limb_obj$hb_obj$hex_poly
@@
 
-langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
+limb_highd_vis_best <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
                          lineFrom = trimesh_data_limb$from,
@@
 ```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=15, fig.height=10}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 
@@
 
+```{r, eval=knitr::is_html_output()}
+
+htmltools::browsable(
+  tagList(
+    tags$div(
+      id = "fig-limb-model",  # This is the anchor
+      crosstalk::bscols(
+        htmltools::div(
+          style = "display: grid; grid-template-columns: 1fr 1fr;",
+          trimesh_limb_int,
+          htmltools::div(style = "margin-top: 13px;", limb_highd_vis),
+          trimesh_limb_best_int,
+          htmltools::div(style = "margin-top: 13px;", limb_highd_vis_best),
+        ),
+        device = "xs"
+      ),
+      tags$p(tags$strong("Figure 6:"), "Compare the published.")
+    )
+  )
+)
+
+```
+
 # Discussion

### 2025-05-20 paper-quollr.Rmd
Figure 6 need to find a way to remove row-wise gap

@@
 ```{r, eval=knitr::is_html_output()}
-
 htmltools::browsable(
-  tagList(
+  htmltools::tagList(
     tags$div(
-      id = "fig-limb-model",  # This is the anchor
-      crosstalk::bscols(
+      id = "fig-limb-model",  # Anchor for referencing
+      style = "text-align: center; margin-bottom: 1em;",
+
+      # 2x2 grid layout
       htmltools::div(
-          style = "display: grid; grid-template-columns: 1fr 1fr;",
+        style = "display: grid; grid-template-columns: 1fr 1fr; row-gap: 0px; justify-items: center;",
         trimesh_limb_int,
@@
         trimesh_limb_best_int,
-          htmltools::div(style = "margin-top: 13px;", limb_highd_vis_best),
+        htmltools::div(style = "margin-top: 13px;", limb_highd_vis_best)
       ),
-        device = "xs"
-      ),
-      tags$p(tags$strong("Figure 6:"), "Compare the published.")
+
+      # Caption
+      tags$p(
+        tags$strong("Figure 6:"), 
+        " Comparison of the published model with the best selected model."
+      )
     )

### 2025-05-20 paper-quollr.Rmd
Updated Figure 6 latex output

@@
 
-```{r langevitour-limb-tsne-author-proj, eval=knitr::is_html_output()}
+```{r langevitour-limb-tsne-author-proj}
 
@@
                          width = "421px", height = "421px")
+```
+
+```{r plot-proj}
+
+## First projection
+projection <- cbind(
+  c(-0.021068,0.007018,0.008770,0.016884,-0.008135,0.011458,-0.022323,0.007231,0.004976,-0.019298),
+  c(-0.017371,-0.008514,0.003231,-0.022945,-0.026264,-0.001571,0.008007,-0.001268,-0.016923,-0.006700))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 0.05,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -0.04, 
+                                              axis_pos_y = -0.04, 
+                                              threshold = 0.004))
+
+proj_obj1[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_model1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.05, 0.03), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "a2", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+
+## Second projection
+projection <- cbind(
+  c(-0.025678,-0.017472,0.015634,-0.014868,-0.000752,0.001050,-0.019524,-0.008076,-0.010808,-0.002319),
+  c(0.014526,-0.013526,-0.005147,-0.027612,-0.016919,0.005198,0.010710,0.004623,-0.007055,0.017842))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 0.05,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -0.03, 
+                                              axis_pos_y = -0.03, 
+                                              threshold = 0.004))
+
+proj_obj2[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_model2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.04, 0.04), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "a3", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+
 ```
@@
 
-```{r langevitour-limb-tsne-best-proj, eval=knitr::is_html_output()}
+```{r langevitour-limb-tsne-best-proj}
 
@@
 
+```{r plot-proj-best}
+
+## First projection
+projection <- cbind(
+  c(-0.021068,0.007018,0.008770,0.016884,-0.008135,0.011458,-0.022323,0.007231,0.004976,-0.019298),
+  c(-0.017371,-0.008514,0.003231,-0.022945,-0.026264,-0.001571,0.008007,-0.001268,-0.016923,-0.006700))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 0.05,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -0.04, 
+                                              axis_pos_y = -0.04, 
+                                              threshold = 0.004))
+
+proj_obj1[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_best_model1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.05, 0.03), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "b2", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+
+## Second projection
+projection <- cbind(
+  c(-0.025678,-0.017472,0.015634,-0.014868,-0.000752,0.001050,-0.019524,-0.008076,-0.010808,-0.002319),
+  c(0.014526,-0.013526,-0.005147,-0.027612,-0.016919,0.005198,0.010710,0.004623,-0.007055,0.017842))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 0.05,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -0.03, 
+                                              axis_pos_y = -0.03, 
+                                              threshold = 0.004))
+
+proj_obj2[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_best_model2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.04, 0.04), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "b3", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+
+```
+
 ```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
@@
 
-trimesh_limb + trimesh_limb_best +
-  trimesh_limb + trimesh_limb_best +
-  trimesh_limb + trimesh_limb_best +
+trimesh_limb + limb_proj_tsne_model1 +
+  limb_proj_tsne_model2 + trimesh_limb_best +
+  limb_proj_tsne_best_model1 + limb_proj_tsne_best_model2 +
   plot_layout(nrow = 2)

### 2025-05-20 paper-quollr.R

@@ -1076,22 +1076,81 @@ scaled_limb_data_model <- scaled_limb |>
   select(-type)
 
 
-## ----langevitour-limb-tsne-author-proj, eval=knitr::is_html_output()----------
-# 
-# data_limb_n <- data_limb_n |>
-#   select(-type) |>
-#   mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
-# 
-# df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
-# 
-# limb_highd_vis <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
-#                          lineFrom = trimesh_data_limb$from,
-#                          lineTo = trimesh_data_limb$to,
-#                          group = factor(df_model_data_limb_n$type,
-#                                         c("0", "1", "2", "3", "4", "5", "6", "model")),
-#                          levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
-#                          enableControls = FALSE,
-#                          width = "421px", height = "421px")
+## ----langevitour-limb-tsne-author-proj----------------------------------------
+
+data_limb_n <- data_limb_n |>
+  select(-type) |>
+  mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
+
+df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
+
+limb_highd_vis <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
+                         lineFrom = trimesh_data_limb$from,
+                         lineTo = trimesh_data_limb$to,
+                         group = factor(df_model_data_limb_n$type,
+                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
+                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
+                         enableControls = FALSE,
+                         width = "421px", height = "421px")
+
+
+## ----plot-proj----------------------------------------------------------------
+
+## First projection
+projection <- cbind(
+  c(-0.021068,0.007018,0.008770,0.016884,-0.008135,0.011458,-0.022323,0.007231,0.004976,-0.019298),
[…]

### 2025-05-20 paper-quollr.Rmd
Started working on Figure 6 title update

@@
         tags$strong("Figure 6:"), 
-        " Comparison of the published model with the best selected model."
+        "Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$."
       )

### 2025-05-20 paper-quollr.Rmd
Changed the position of the figure

@@
 ```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 

### 2025-05-20 paper-quollr.R

@@ -1327,7 +1327,7 @@ limb_proj_tsne_best_model2 <- plot_proj(
 
 
 
-## ----model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters. The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", out.width="100%", fig.width=15, fig.height=10, eval=knitr::is_latex_output()----
+## ----model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters. The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()----
 
 trimesh_limb + limb_proj_tsne_model1 +
   limb_proj_tsne_model2 + trimesh_limb_best +

### 2025-05-20 paper-quollr.Rmd
Found a way to refer the Figure inside figure (need to map with figure label)

@@
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.", fig.pos='H'}
+Layout f is the best choice.\\label{fig:limb-rmse}", fig.pos='H'}
 
@@
 
-```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\ref{fig:limb-rmse}b) and the $2-\\text{D}$ layout selected (Figure \\ref{fig:limb-rmse}f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}

### 2025-05-20 paper-quollr.tex
Updated trimesh figure configurations
Changed the position of the figure
Found a way to refer the Figure inside figure (need to map with figure label)

@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$).
@@
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.}\label{fig:limb-rmse}
+Layout f is the best choice.\label{fig:limb-rmse}}\label{fig:limb-rmse}
 \end{figure}
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \\ref{fig:limb-rmse} b) and the $2-\text{D}$ layout selected (Figure \\ref{fig:limb-rmse} f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \ref{fig:limb-rmse}b) and the $2-\text{D}$ layout selected (Figure \ref{fig:limb-rmse}f) by RMSE plot (Figure \ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-05-20 paper-quollr.Rmd
Removed hexgrid from figure 1

@@
   aes(x = x, y = y)) +
-  geom_polygon(color = "grey70", 
-               aes(group = hex_poly_id), 
-               fill = "#ffffff") +
   geom_point(data = scurve_umap_scaled, 

### 2025-05-20 paper-quollr.Rmd
Changed the color of the data points

@@
 ```{r}
-clr_choice <- "#000000"
+clr_choice <- "#bcbddc"
 
@@
              aes(x = emb1, y = emb2), 
-             alpha = 0.2, size = 0.5, color = clr_choice) +
+             size = 0.5, color = clr_choice) +
   geom_segment(data = tr_from_to_df_scurve,
@@
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
-                         levelColors = c("#000000", "#33a02c"),
+                         levelColors = c(clr_choice, "#33a02c"),
                          enableControls = FALSE,

### 2025-06-24 paper-quollr.Rmd
Added quollr motivation letter

@@
 
-all_hex_coord
+head(all_hex_coord, 5)
 ```
@@
 
-umap_hex_id
+head(umap_hex_id, 5)
 ```
@@
 
-std_df
+head(std_df, 5)
 ```
@@
 
-pts_df
+head(pts_df, 5)
 ```
@@
 
-df_bin_centroids
+head(df_bin_centroids, 5)
 ```
@@
 
-trimesh
+head(trimesh, 5)
 ```
@@
 
-df_bin
+head(df_bin, 5)
 ```
@@
 ```{r, echo=TRUE}
-predict_emb(
+predict_data <- predict_emb(
   highd_data = scurve, 
@@
   )
+
+head(predict_data, 5)
 ```
@@
 
-model_error
+head(model_error, 5)
 ```
@@
 ```{r, echo=TRUE}
-
 df_exe <- comb_data_model(
@@
   )
-
-df_exe
 ```
@@
 There are $1067$ single cells, with $14997$ gene expressions.
-Following their pre-processing, tSNE was performed using ten principal components.
+Following their pre-processing, different NLDR methods were performed using ten principal components.
 Figure \@ref(fig:limb-rmse) (b) is the reproduction of the published plot.

### 2025-06-24 paper-quollr.tex
Added quollr motivation letter

@@
 
-all_hex_coord
+head(all_hex_coord, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 1,440 x 3
 #>   hex_poly_id           x           y
-#>          <int>   <dbl>   <dbl>
-#>  1           1 -0.1    -0.0412
-#>  2           1 -0.142  -0.0652
-#>  3           1 -0.142  -0.113 
-#>  4           1 -0.1    -0.137 
-#>  5           1 -0.0584 -0.113 
-#>  6           1 -0.0584 -0.0652
-#>  7           2 -0.0167 -0.0412
-#>  8           2 -0.0584 -0.0652
-#>  9           2 -0.0584 -0.113 
-#> 10           2 -0.0167 -0.137 
-#> # i 1,430 more rows
+#> 1           1 -0.10000000 -0.04116510
+#> 2           1 -0.14163068 -0.06520059
+#> 3           1 -0.14163068 -0.11327155
+#> 4           1 -0.10000000 -0.13730704
+#> 5           1 -0.05836932 -0.11327155
 \end{verbatim}
@@
 
-umap_hex_id
+head(umap_hex_id, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5,000 x 4
+#> # A tibble: 5 x 4
 #>    emb1  emb2    ID hexID
@@
 #> 5 0.761 0.551     5   146
-#>  6 0.445 0.721     6   172
-#>  7 0.900 0.137     7    58
-#>  8 0.247 0.392     8   110
-#>  9 0.325 0.542     9   141
-#> 10 0.278 0.231    10    80
-#> # i 4,990 more rows
 \end{verbatim}
@@
 
-std_df
+head(std_df, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 142 x 3
+#> # A tibble: 5 x 3
 #>   hexID bin_counts std_counts
@@
 #> 5    25          7     0.0933
-#>  6    26         12     0.16  
-#>  7    27          1     0.0133
-#>  8    36         42     0.56  
-#>  9    37         42     0.56  
-#> 10    38         44     0.587 
-#> # i 132 more rows
 \end{verbatim}
@@
 
-pts_df
+head(pts_df, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 142 x 2
+#> # A tibble: 5 x 2
 #>   hexID pts_list  
@@
 #> 5    25 <int [7]>
-#>  6    26 <int [12]>
-#>  7    27 <int [1]> 
-#>  8    36 <int [42]>
-#>  9    37 <int [42]>
-#> 10    38 <int [44]>
-#> # i 132 more rows
 \end{verbatim}
@@
 
-df_bin_centroids
+head(df_bin_centroids, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 240 x 5
 #>   hexID         c_x         c_y bin_counts std_counts
-#>    <int>   <dbl>   <dbl>      <dbl>      <dbl>
-#>  1     1 -0.1    -0.0892          0          0
-#>  2     2 -0.0167 -0.0892          0          0
-#>  3     3  0.0665 -0.0892          0          0
-#>  4     4  0.150  -0.0892          0          0
-#>  5     5  0.233  -0.0892          0          0
-#>  6     6  0.316  -0.0892          0          0
-#>  7     7  0.400  -0.0892          0          0
-#>  8     8  0.483  -0.0892          0          0
-#>  9     9  0.566  -0.0892          0          0
-#> 10    10  0.649  -0.0892          0          0
-#> # i 230 more rows
+#> 1     1 -0.10000000 -0.08923607          0          0
+#> 2     2 -0.01673864 -0.08923607          0          0
+#> 3     3  0.06652271 -0.08923607          0          0
+#> 4     4  0.14978407 -0.08923607          0          0
+#> 5     5  0.23304543 -0.08923607          0          0
 \end{verbatim}
@@
 
-trimesh
+head(trimesh, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 653 x 8
+#> # A tibble: 5 x 8
 #>    from    to  x_from  y_from    x_to    y_to from_count to_count
@@
 #> 5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
-#>  6    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
-#>  7    31    46 -0.1     0.0550 -0.0584  0.127           0        0
-#>  8    32    47 -0.0167  0.0550  0.0249  0.127           0        0
-#>  9    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
-#> 10     4    18  0.150  -0.0892  0.108  -0.0171          0        0
-#> # i 643 more rows
 \end{verbatim}
@@
 
-df_bin
+head(df_bin, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 142 x 8
+#> # A tibble: 5 x 8
 #>   hexID      x1    x2    x3         x4        x5       x6        x7
@@
 #> 5    25  0.0760  1.93  2.00  0.00876    0.00447   0.00851 -0.00195
-#>  6    26  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172 
-#>  7    27  0.719   1.99 1.70   0.0109    -0.00349  -0.0297  -0.00223 
-#>  8    36 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338
-#>  9    37 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126
-#> 10    38 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780
-#> # i 132 more rows
 \end{verbatim}
@@
 \begin{verbatim}
-predict_emb(
+predict_data <- predict_emb(
   highd_data = scurve, 
@@
   )
+
+head(predict_data, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5,000 x 4
+#> # A tibble: 5 x 4
 #>   pred_emb_1 pred_emb_2    ID pred_hb_id
@@
 #> 5      0.774      0.560     5        146
-#>  6      0.441      0.704     6        172
-#>  7      0.941      0.127     7         58
-#>  8      0.275      0.416     8        110
-#>  9      0.358      0.560     9        141
-#> 10      0.316      0.343    10         96
-#> # i 4,990 more rows
 \end{verbatim}
@@
 
-model_error
+head(model_error, 5)
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5,000 x 32
+#> # A tibble: 5 x 32
 #>      ID      x1     x2       x3       x4       x5      x6        x7 pred_hb_id
@@
 #> 5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998  0.000677        146
-#>  6     6  0.818  0.927  -1.58    -0.00318 -0.00980  0.0989   0.00696         172
-#>  7     7  0.910  1.40    1.42     0.00699 -0.0182  -0.0710   0.00966          58
-#>  8     8 -0.0691 1.59    0.00239  0.0127  -0.0130   0.0396  -0.000185        110
-#>  9     9  0.859  1.59   -0.488   -0.0119   0.00421 -0.00440 -0.00595         141
-#> 10    10 -0.727  1.62    0.314    0.00251  0.0177  -0.0755  -0.00369          96
-#> # i 4,990 more rows
 #> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
@@
 #> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
-#> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>, ...
+#> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>,
+#> #   abs_error_x3 <dbl>, abs_error_x4 <dbl>, abs_error_x5 <dbl>, ...
 \end{verbatim}
@@
   )
-
-df_exe
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 5,122 x 8
-#>         x1    x2    x3         x4        x5       x6        x7 type 
-#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr>
-#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model
-#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model
-#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model
-#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model
-#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model
-#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model
-#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model
-#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model
-#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model
-#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model
-#> # i 5,112 more rows
 \end{verbatim}
@@
   )
-
-df_exe
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 5,122 x 12
-#>         x1    x2    x3         x4        x5       x6        x7 type   emb1  emb2
-#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr> <dbl> <dbl>
-#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model    NA    NA
-#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model    NA    NA
-#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model    NA    NA
-#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model    NA    NA
-#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model    NA    NA
-#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model    NA    NA
-#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model    NA    NA
-#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model    NA    NA
-#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model    NA    NA
-#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model    NA    NA
-#> # i 5,112 more rows
-#> # i 2 more variables: sqrt_row_wise_total_error <dbl>, density <dbl>
 \end{verbatim}

### 2025-06-24 paper-quollr.Rmd
Added source code link
Reword hyperparameter

@@
   Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation.
-However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper-)parameter choices.
+However, the complexity of the transformations and data structures can create wildly different representations depending on the method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
-The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper-)parameter choices provide the most accurate representation of high-dimensional data.
+The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which hyper-parameter choices provide the most accurate representation of high-dimensional data.
 The `scurve` data from the package is used to illustrate the algorithm.
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper-)parameter choices.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 
-This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of $p\text{-}D$ data.
+This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyper-parameter choice gives the most accurate representation of $p\text{-}D$ data.
 
 # Acknowledgements
+
+The source code for reproducing this paper can be found at: [https://github.com/JayaniLakshika/paper-quollr](https://github.com/JayaniLakshika/paper-quollr).
 
@@
 These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].
-
-The source code for reproducing this paper can be found at: <https://github.com/JayaniLakshika/paper-quollr>.

### 2025-06-24 paper-quollr.tex
Added source code link
Reword hyperparameter

@@
 Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (\(p\text{-}D\)) by applying a non-linear transformation.
-However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper-)parameter choices.
+However, the complexity of the transformations and data structures can create wildly different representations depending on the method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
-The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper-)parameter choices provide the most accurate representation of high-dimensional data.
+The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which hyper-parameter choices provide the most accurate representation of high-dimensional data.
 The \texttt{scurve} data from the package is used to illustrate the algorithm.
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper-)parameter choices.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 There are \(1067\) single cells, with \(14997\) gene expressions.
-Following their pre-processing, tSNE was performed using ten principal components.
+Following their pre-processing, different NLDR methods were performed using ten principal components.
 Figure \ref{fig:limb-rmse} (b) is the reproduction of the published plot.
@@
 
-This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyperparameter choice gives the most accurate representation of \(p\text{-}D\) data.
+This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyper-parameter choice gives the most accurate representation of \(p\text{-}D\) data.
 
@@
 
+The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.
+
 This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
 These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.
-
-The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.
 

### 2025-06-24 paper-quollr.R
Function with arrange RMSE and 2D layouts

@@ -962,9 +962,11 @@ error_plot_limb <- plot_rmse(error_limb) +
 
 ## ----limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases. Layout f is the best choice.\\label{fig:limb-rmse}", fig.pos='H'----
 
-error_plot_limb + 
-  wrap_plots(nldr1, nldr2, nldr3, 
-  nldr4, nldr5, nldr6, ncol = 2) 
+design <- gen_design(n_right = 6, ncol_right = 2)
+
+plot_rmse_layouts(plots = list(error_plot_limb, nldr1, 
+                               nldr2, nldr3, nldr4, 
+                               nldr5, nldr6), design = design)
 
 
 ## ----data-limb----------------------------------------------------------------

### 2025-06-24 paper-quollr.Rmd
Function with arrange RMSE and 2D layouts
Add the code

@@
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.\\label{fig:limb-rmse}", fig.pos='H'}
-
-error_plot_limb + 
-  wrap_plots(nldr1, nldr2, nldr3, 
-  nldr4, nldr5, nldr6, ncol = 2) 
+Layout f is the best choice.\\label{fig:limb-rmse}", fig.pos='H', echo=TRUE}
+
+design <- gen_design(n_right = 6, ncol_right = 2)
+
+plot_rmse_layouts(plots = list(error_plot_limb, nldr1, 
+                               nldr2, nldr3, nldr4, 
+                               nldr5, nldr6), design = design)
 ```

### 2025-07-01 paper-quollr.Rmd
Updated the notation parameters

@@
   nldr_data = scurve_umap, 
-  bin1 = 15, 
+  b1 = 15, 
   q = 0.1, 
@@
   hex_grid_scurve, counts_df_scurve, 
-  by = c("hex_poly_id" = "hexID"))
+  by = c("hex_poly_id" = "h"))
 
 hex_grid_nonempty_scurve <- hex_grid_scurve |>
-  filter(hex_poly_id %in% df_bin_centroids_scurve$hexID)
+  filter(hex_poly_id %in% df_bin_centroids_scurve$h)
 
@@
 df_b_scurve <- df_bin_scurve |>
-  dplyr::filter(hexID %in% df_bin_centroids_scurve$hexID) |>
+  dplyr::filter(h %in% df_bin_centroids_scurve$h) |>
   dplyr::mutate(type = "model") ## Data with summarized mean
 
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_scurve <- df_b_scurve[match(df_bin_centroids_scurve$hexID, df_b_scurve$hexID),] |>
-  dplyr::select(-hexID)
+## Reorder the rows of df_b according to the h order in df_b_with_center_data
+df_b_scurve <- df_b_scurve[match(df_bin_centroids_scurve$h, df_b_scurve$h),] |>
+  dplyr::select(-h)
 
@@
 
-This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`bin1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`benchmark_highdens`).
+This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`benchmark_highdens`).
 The function returns an object that includes the scaled NLDR object (`nldr_obj`), the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
@@
 The function `calc_bins_y()` is used for this purpose.
-This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (`bin1`), and the buffer amount as a proportion (`q`).
+This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (`b1`), and the buffer amount as a proportion (`q`).
 
@@
   nldr_obj = scurve_umap_obj, 
-  bin1 = 15, 
+  b1 = 15, 
   q = 0.1)
@@
 The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its components can be run separately for added flexibility.
-The parameters used within `hex_binning()` include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (`nldr_obj`), the number of bins along the x-axis (`bin1`), and the buffer amount as a proportion of the data (`q`).
+The parameters used within `hex_binning()` include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (`nldr_obj`), the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
 The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagon IDs (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`tot_bins`), the number of non-empty bins (`non_bins`), and the points within each hexagon (`pts_bins`).
@@
   nldr_obj = scurve_umap_obj, 
-  bin1 = 9, 
+  b1 = 9, 
   q = 0.1)
@@
 start_pt <- all_centroids_df_temp |> 
-  filter(hexID == 1)
+  filter(h == 1)
 d_rect <- tibble(x1min = 0, 
@@
 The coordinate limits of the embedding (`lim1` and `lim2`) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
-The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagonal rows (`bin2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`bin1`) and buffer (`q`).
+The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagonal rows (`bin2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`b1`) and buffer (`q`).
 
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (`hexID`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2-\text{D}$ space.
+Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2-\text{D}$ space.
 
@@
   nldr_obj = scurve_umap_obj, 
-  bin1 = 15, 
+  b1 = 15, 
   q = 0.1
@@
 
-For each centroid, six vertices are computed and assigned a polygon ID (`hex_poly_id`) corresponding to the centroid’s `hexID`.
+For each centroid, six vertices are computed and assigned a polygon ID (`hex_poly_id`) corresponding to the centroid’s `h`.
 These points are aggregated into a tibble that includes the polygon ID (`hex_poly_id`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.
@@
 The function uses the `proxy::dist()` method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
-For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon—and assigns the corresponding hexagon ID (`hexID`) to the point.
-
-The result is a data frame of the scaled $2-\text{D}$ embedding with an additional `hexID` column, indicating the hexagonal bin to which each point belongs.
+For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon—and assigns the corresponding hexagon ID (`h`) to the point.
+
+The result is a data frame of the scaled $2-\text{D}$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
 
@@
 
-The function begins by grouping the data by `hexID` and counting the number of NLDR points falling within each bin.
-These raw counts are stored as `bin_counts`.
+The function begins by grouping the data by `h` and counting the number of NLDR points falling within each bin.
+These raw counts are stored as `n_h`.
 To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin’s count by the maximum count across all bins.
-This yields a normalized metric, `std_counts`, ranging from 0 to 1.
+This yields a normalized metric, `w_h`, ranging from 0 to 1.
 
@@
 std_df <- compute_std_counts(
-  scaled_nldr_hexid = umap_hex_id
+  scaled_nldr_h = umap_hex_id
   )
@@
 
-The function first groups the input data by `hexID`, which represents the hexagon label associated with each point in the $2-\text{D}$ layout.
+The function first groups the input data by `h`, which represents the hexagon label associated with each point in the $2-\text{D}$ layout.
 Within each group, it collects the `ID`s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
@@
 
-This function begins by arranging the `counts_data` by `hexID` to ensure consistent ordering.
+This function begins by arranging the `counts_data` by `h` to ensure consistent ordering.
 It then performs a full join with `centroids_data`, aligning hexagon IDs between the two datasets to incorporate both hexagonal bin centroids and count metrics.
-After merging, the function handles missing values in the count columns: any `NA` values in `std_counts` or `bin_counts` are replaced with zeros.
+After merging, the function handles missing values in the count columns: any `NA` values in `w_h` or `n_h` are replaced with zeros.
 This ensures that hexagons with no assigned data points are retained in the output, with zero values for count-related fields.
@@
 The `find_low_dens_hex()` function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-The `find_low_dens_hex()` function relies on a helper, `compute_mean_density_hex()`, which iterates over all hexagons and computes the average density across neighbors based on their `hexID` and a defined number of bins along the x-axis (`bin1`).
+The `find_low_dens_hex()` function relies on a helper, `compute_mean_density_hex()`, which iterates over all hexagons and computes the average density across neighbors based on their `h` and a defined number of bins along the x-axis (`b1`).
 The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.
@@
   centroids_data = df_bin_centroids, 
-  bin1 = 15, 
+  b1 = 15, 
   benchmark_mean_dens = 0.05
@@
 df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(bin_counts > 10)
+  dplyr::filter(n_h > 10)
 
@@
 <!--   mutate(a1 = round(a1, 2)) |> -->
-<!--   filter(bin1 >= 5) |> -->
+<!--   filter(b1 >= 5) |> -->
 <!--   group_by(method, a1) |> -->
@@
   mutate(a1 = round(a1, 2)) |>
-  filter(bin1 >= 5) |>
+  filter(b1 >= 5) |>
   filter(a1 >= 0.03) |>
@@
   nldr_data = tsne_limb, 
-  bin1 = 19, 
+  b1 = 19, 
   q = 0.1, 
@@
 # 
-# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hexID"))
+# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "h"))
 # 
 # ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = bin_counts)) +
-#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = n_h)) +
+#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = h)) +
 #   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
@@
 df_b_limb <- df_bin_limb |>
-  dplyr::filter(hexID %in% df_bin_centroids_limb$hexID) |>
+  dplyr::filter(h %in% df_bin_centroids_limb$h) |>
   dplyr::mutate(type = "model") ## Data with summarized mean
 
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_limb <- df_b_limb[match(df_bin_centroids_limb$hexID, df_b_limb$hexID),] |>
-  dplyr::select(-hexID)
+## Reorder the rows of df_b according to the h order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
+  dplyr::select(-h)
 
@@
   nldr_data = tsne_limb2, 
-  bin1 = 19, 
+  b1 = 19, 
   q = 0.1, 

### 2025-07-01 paper-quollr.tex
Add the code
Updated the notation parameters

@@
 
-This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{bin1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{benchmark\_highdens}).
+This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{benchmark\_highdens}).
 The function returns an object that includes the scaled NLDR object (\texttt{nldr\_obj}), the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
@@
 The function \texttt{calc\_bins\_y()} is used for this purpose.
-This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion (\texttt{q}).
+This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion (\texttt{q}).
 
@@
 \begin{verbatim}
-#> $bin2
+#> $b2
 #> [1] 16
@@
 The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its components can be run separately for added flexibility.
-The parameters used within \texttt{hex\_binning()} include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (\texttt{nldr\_obj}), the number of bins along the x-axis (\texttt{bin1}), and the buffer amount as a proportion of the data (\texttt{q}).
+The parameters used within \texttt{hex\_binning()} include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (\texttt{nldr\_obj}), the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
 The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{tot\_bins}), the number of non-empty bins (\texttt{non\_bins}), and the points within each hexagon (\texttt{pts\_bins}).
@@
 The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
-The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagonal rows (\texttt{bin2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{bin1}) and buffer (\texttt{q}).
+The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagonal rows (\texttt{bin2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{b1}) and buffer (\texttt{q}).
 
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (\texttt{hexID}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2-\text{D}\) space.
+Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2-\text{D}\) space.
 
@@
 #> # A tibble: 240 x 3
-#>    hexID     c_x     c_y
+#>        h     c_x     c_y
 #>    <int>   <dbl>   <dbl>
@@
 
-For each centroid, six vertices are computed and assigned a polygon ID (\texttt{hex\_poly\_id}) corresponding to the centroid's \texttt{hexID}.
+For each centroid, six vertices are computed and assigned a polygon ID (\texttt{hex\_poly\_id}) corresponding to the centroid's \texttt{h}.
 These points are aggregated into a tibble that includes the polygon ID (\texttt{hex\_poly\_id}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
@@
 The function uses the \texttt{proxy::dist()} method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
-For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon---and assigns the corresponding hexagon ID (\texttt{hexID}) to the point.
+For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon---and assigns the corresponding hexagon ID (\texttt{h}) to the point.
 
-The result is a data frame of the scaled \(2-\text{D}\) embedding with an additional \texttt{hexID} column, indicating the hexagonal bin to which each point belongs.
+The result is a data frame of the scaled \(2-\text{D}\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
 
@@
 #> # A tibble: 5 x 4
-#>    emb1  emb2    ID hexID
+#>    emb1  emb2    ID     h
 #>   <dbl> <dbl> <int> <int>
@@
 
-The function begins by grouping the data by \texttt{hexID} and counting the number of NLDR points falling within each bin.
-These raw counts are stored as \texttt{bin\_counts}.
+The function begins by grouping the data by \texttt{h} and counting the number of NLDR points falling within each bin.
+These raw counts are stored as \texttt{n\_h}.
 To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin's count by the maximum count across all bins.
-This yields a normalized metric, \texttt{std\_counts}, ranging from 0 to 1.
+This yields a normalized metric, \texttt{w\_h}, ranging from 0 to 1.
 
@@
 #> # A tibble: 5 x 3
-#>   hexID bin_counts std_counts
+#>       h   n_h    w_h
 #>   <int> <int>  <dbl>
-#> 1    21         12     0.16  
-#> 2    22         17     0.227 
-#> 3    23         22     0.293 
-#> 4    24         10     0.133 
-#> 5    25          7     0.0933
+#> 1    21    12 0.0024
+#> 2    22    17 0.0034
+#> 3    23    22 0.0044
+#> 4    24    10 0.002 
+#> 5    25     7 0.0014
 \end{verbatim}
@@
 
-The function first groups the input data by \texttt{hexID}, which represents the hexagon label associated with each point in the \(2-\text{D}\) layout.
+The function first groups the input data by \texttt{h}, which represents the hexagon label associated with each point in the \(2-\text{D}\) layout.
 Within each group, it collects the \texttt{ID}s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
@@
 #> # A tibble: 5 x 2
-#>   hexID pts_list  
+#>       h pts_list  
 #>   <int> <list>    
@@
 
-This function begins by arranging the \texttt{counts\_data} by \texttt{hexID} to ensure consistent ordering.
+This function begins by arranging the \texttt{counts\_data} by \texttt{h} to ensure consistent ordering.
 It then performs a full join with \texttt{centroids\_data}, aligning hexagon IDs between the two datasets to incorporate both hexagonal bin centroids and count metrics.
-After merging, the function handles missing values in the count columns: any \texttt{NA} values in \texttt{std\_counts} or \texttt{bin\_counts} are replaced with zeros.
+After merging, the function handles missing values in the count columns: any \texttt{NA} values in \texttt{w\_h} or \texttt{n\_h} are replaced with zeros.
 This ensures that hexagons with no assigned data points are retained in the output, with zero values for count-related fields.
@@
 \begin{verbatim}
-#>   hexID         c_x         c_y bin_counts std_counts
+#>   h         c_x         c_y n_h w_h
 #> 1 1 -0.10000000 -0.08923607   0   0
@@
 The \texttt{find\_low\_dens\_hex()} function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their \texttt{hexID} and a defined number of bins along the x-axis (\texttt{bin1}).
+The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their \texttt{h} and a defined number of bins along the x-axis (\texttt{b1}).
 The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.
@@
 \begin{verbatim}
-#>  [1]   1   2   3   4   5  11  12  13  14  15  16  17  18  19  30  31  32  33  46
-#> [20]  47  61  75  76  90 105 120 135 150 151 166 181 182 196 197 198 211 212 213
-#> [39] 214 215 225 226 227 228 229 230 231 232 233 234 235 236 237 238 239 240
+#>   [1]   1   2   3   4   5   6   7   8   9  10  11  12  13  14  15  16  17  18
+#>  [19]  19  20  21  22  23  24  25  26  27  28  29  30  31  32  33  34  35  36
+#>  [37]  37  38  39  40  41  42  43  44  45  46  47  48  49  50  51  52  53  54
+#>  [55]  55  56  57  58  59  60  61  62  63  64  65  66  67  68  69  70  71  72
+#>  [73]  73  74  75  76  77  78  79  80  81  82  83  84  85  86  87  88  89  90
+#>  [91]  91  92  93  94  95  96  97  98  99 100 101 102 103 104 105 106 107 108
+#> [109] 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126
+#> [127] 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144
+#> [145] 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162
+#> [163] 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180
+#> [181] 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198
+#> [199] 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215 216
+#> [217] 217 218 219 220 221 222 223 224 225 226 227 228 229 230 231 232 233 234
+#> [235] 235 236 237 238 239 240
 \end{verbatim}
@@
 #> # A tibble: 5 x 8
-#>   hexID      x1    x2    x3         x4        x5       x6        x7
+#>       h      x1    x2    x3         x4        x5       x6        x7
 #>   <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
@@
 
+\begin{verbatim}
+design <- gen_design(n_right = 6, ncol_right = 2)
+
+plot_rmse_layouts(plots = list(error_plot_limb, nldr1, 
+                               nldr2, nldr3, nldr4, 
+                               nldr5, nldr6), design = design)
+\end{verbatim}
+
 \begin{figure}[H]

### 2025-07-01 paper-quollr.Rmd
Added a way to refer crosstalk output with fi-cap

@@
 This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
-Starting with an NLDR layout, our approach is to create a wireframe model representation, that can be lifted and displayed in the high-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "[Figure 1](#fig-overview)"} else { "Figure \\@ref(fig:overview)"})`).
+Starting with an NLDR layout, our approach is to create a wireframe model representation, that can be lifted and displayed in the high-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:overview)"})`).
 
@@
 
-```{r, eval=knitr::is_html_output()}
-
-htmltools::browsable(
-  tagList(
-    tags$div(
-      id = "fig-overview",  # This is the anchor
-      crosstalk::bscols(
+```{r overviewhtml, eval=knitr::is_html_output(), fig.cap="algorithm", fig.pos='H'}
+
+overviewfig <- bscols(
   htmltools::div(
@@
   device = "xs"
-      ),
-      tags$p(tags$strong("Figure 1:"), " Algorithm to visualize the 2D model.")
-    )
-  )
-)
-
+)
+
+class(overviewfig) <- c(class(overviewfig), "htmlwidget")
+
+overviewfig
 ```
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output()}
+```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="jkjio"}
 show_link_plots(
@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method can help here, and also help to provide a better $2-\text{D}$ layout, as needed.
+Our method help to provide a better $2-\text{D}$ layout.
+
+<!-- add function to generate 2D trimesh and the projections (PDF) and langevitour (HTML)-->
 

### 2025-07-01 paper-quollr.tex
Added a way to refer crosstalk output with fi-cap

@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-29-1}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
 
@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method can help here, and also help to provide a better \(2-\text{D}\) layout, as needed.
+Our method help to provide a better \(2-\text{D}\) layout.
 

### 2025-07-06 paper-quollr.Rmd
Added discussion section

@@
 
+To reduce computational overhead, the geometry calculations are implemented in C++ using `gen_hex_coord_cpp()`, which returns a `data.frame` of vertex coordinates.
+
 ```{r, echo=TRUE}
@@
 
+To accelerate this process, the nearest-neighbor search is implemented in C++ using `Rcpp` via the internal function `compute_highd_dist()`.
+
 ```{r, echo=TRUE}
@@
 These metrics are returned in a tibble as `Error` (absolute error) and `RMSE` (root mean squared error).
+The computation of total absolute error and RMSE is performed in C++ for efficiency using the internal `compute_errors()` function.
 
@@
 
+# Computational efficiency and optimization
+
+Several core computations within `quollr` are optimized using compiled C++ code via the `Rcpp` and `RcppArmadillo` packages.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), $2-\text{D}$ distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
+This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
+These C++ functions are not exported but are bundled within the package and fully accessible for inspection in the source code.
+
 # Application
@@
 
-This paper presents the R package `quollr` to develop a way to take the fitted model, as represented by the positions of points in $2\text{-}D$, and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.
-
-The paper includes a clustering example to illustrate how `quollr` is useful to assess which NLDR technique and which (hyper)parameter choice gives the most accurate representation.
-In addition, how to select parameters for hexagonal binning and fitting model are explained.
-
-Possible future improvements would be...<!--assess the preservation of local and glocal structure w.r.t 2D and high-D distance comparison-->
-
-This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyper-parameter choice gives the most accurate representation of $p\text{-}D$ data.
+The `quollr` package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in $2-\text{D}$ and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a $2-\text{D}$ layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate $2-\text{D}$ representations.
+
+In contrast to the common practice of visually inspecting scatterplots for clusters or patterns, `quollr` provides a quantitative route for evaluation.
+It enables the computation of RMSE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
+These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualisations using the `langevitour` package, allowing users to inspect where the model fits well and where it does not.
+
+To support efficient computation, particularly for large-scale datasets, several core functions in `quollr` are implemented in C++ using `Rcpp` and `RcppArmadillo`.
+These include functions for computing Euclidean distances in high-dimensional and $2-\text{D}$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RMSE and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in $2-\text{D}$.
+Additionally, `gen_hex_coord_cpp()` constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.
+These optimizations result in substantial performance gains compared to native R implementations, making the package responsive even when used in interactive contexts or on large datasets such as single-cell transcriptomic profiles.
+
+The modular structure of the package is designed to support both flexibility and reproducibility.
+Users can access individual functions to control each step of the pipeline such as scaling, binning, and triangulation or use the main function `fit_highd_model()` for end-to-end model construction.
+The diagnostics can be used not only to compare NLDR methods but also to tune binning parameters, assess layout stability, and detect local distortions in the embedding.
+
+There are several avenues for future development.
+While hexagonal binning provides a regular structure conducive to modeling, alternative spatial discretizations (e.g., adaptive binning or density-aware tessellations) could be explored to better capture varying data densities.
+Expanding support for additional distance metrics in the lifting and prediction steps may improve performance across different domains.
+Additionally, statistical inference tools could be introduced to assess the stability and robustness of the fitted model, which would enhance interpretability and confidence in the outcomes.
 

### 2025-07-06 paper-quollr.tex
Added discussion section

@@
 These points are aggregated into a tibble that includes the polygon ID (\texttt{hex\_poly\_id}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
+
+To reduce computational overhead, the geometry calculations are implemented in C++ using \texttt{gen\_hex\_coord\_cpp()}, which returns a \texttt{data.frame} of vertex coordinates.
 
@@
 
+To accelerate this process, the nearest-neighbor search is implemented in C++ using \texttt{Rcpp} via the internal function \texttt{compute\_highd\_dist()}.
+
 \begin{verbatim}
@@
 These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{RMSE} (root mean squared error).
+The computation of total absolute error and RMSE is performed in C++ for efficiency using the internal \texttt{compute\_errors()} function.
 
@@
 
+\section{Computational efficiency and optimization}\label{computational-efficiency-and-optimization}
+
+Several core computations within \texttt{quollr} are optimized using compiled C++ code via the \texttt{Rcpp} and \texttt{RcppArmadillo} packages.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), \(2-\text{D}\) distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
+This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
+These C++ functions are not exported but are bundled within the package and fully accessible for inspection in the source code.
+
 \section{Application}\label{application}
@@
 
-This paper presents the R package \texttt{quollr} to develop a way to take the fitted model, as represented by the positions of points in \(2\text{-}D\), and turn it into a high-dimensional wireframe to overlay on the data, viewing it with a tour.
+The \texttt{quollr} package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in \(2-\text{D}\) and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a \(2-\text{D}\) layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate \(2-\text{D}\) representations.
 
-The paper includes a clustering example to illustrate how \texttt{quollr} is useful to assess which NLDR technique and which (hyper)parameter choice gives the most accurate representation.
-In addition, how to select parameters for hexagonal binning and fitting model are explained.
+In contrast to the common practice of visually inspecting scatterplots for clusters or patterns, \texttt{quollr} provides a quantitative route for evaluation.
+It enables the computation of RMSE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
+These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualisations using the \texttt{langevitour} package, allowing users to inspect where the model fits well and where it does not.
 
-Possible future improvements would be\ldots{}
+To support efficient computation, particularly for large-scale datasets, several core functions in \texttt{quollr} are implemented in C++ using \texttt{Rcpp} and \texttt{RcppArmadillo}.
+These include functions for computing Euclidean distances in high-dimensional and \(2-\text{D}\) space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RMSE and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in \(2-\text{D}\).
+Additionally, \texttt{gen\_hex\_coord\_cpp()} constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.
+These optimizations result in substantial performance gains compared to native R implementations, making the package responsive even when used in interactive contexts or on large datasets such as single-cell transcriptomic profiles.
 
-This new tool provides an effective start point for automatically creating regular hexagons and help to evaluate which NLDR technique and which hyper-parameter choice gives the most accurate representation of \(p\text{-}D\) data.
+The modular structure of the package is designed to support both flexibility and reproducibility.
+Users can access individual functions to control each step of the pipeline such as scaling, binning, and triangulation or use the main function \texttt{fit\_highd\_model()} for end-to-end model construction.
+The diagnostics can be used not only to compare NLDR methods but also to tune binning parameters, assess layout stability, and detect local distortions in the embedding.
+
+There are several avenues for future development.
+While hexagonal binning provides a regular structure conducive to modeling, alternative spatial discretizations (e.g., adaptive binning or density-aware tessellations) could be explored to better capture varying data densities.
+Expanding support for additional distance metrics in the lifting and prediction steps may improve performance across different domains.
+Additionally, statistical inference tools could be introduced to assess the stability and robustness of the fitted model, which would enhance interpretability and confidence in the outcomes.
 

### 2025-07-06 paper-quollr.Rmd
Changed code apperance

@@
   warning=FALSE,
+  comment = ">",
   out.width = "100%")

### 2025-07-06 paper-quollr.tex
Changed code apperance

@@
 \begin{verbatim}
-#> $scaled_nldr
-#> # A tibble: 5,000 x 3
-#>     emb1  emb2    ID
-#>    <dbl> <dbl> <int>
-#>  1 0.707 0.839     1
-#>  2 0.231 0.401     2
-#>  3 0.232 0.215     3
-#>  4 0.790 0.564     4
-#>  5 0.761 0.551     5
-#>  6 0.445 0.721     6
-#>  7 0.900 0.137     7
-#>  8 0.247 0.392     8
-#>  9 0.325 0.542     9
-#> 10 0.278 0.231    10
-#> # i 4,990 more rows
-#> 
-#> $lim1
-#> [1] -14.42166  13.32655
-#> 
-#> $lim2
-#> [1] -12.43687  12.32455
+> $scaled_nldr
+> # A tibble: 5,000 x 3
+>     emb1  emb2    ID
+>    <dbl> <dbl> <int>
+>  1 0.707 0.839     1
+>  2 0.231 0.401     2
+>  3 0.232 0.215     3
+>  4 0.790 0.564     4
+>  5 0.761 0.551     5
+>  6 0.445 0.721     6
+>  7 0.900 0.137     7
+>  8 0.247 0.392     8
+>  9 0.325 0.542     9
+> 10 0.278 0.231    10
+> # i 4,990 more rows
+> 
+> $lim1
+> [1] -14.42166  13.32655
+> 
+> $lim2
+> [1] -12.43687  12.32455
 \end{verbatim}
@@
 \begin{verbatim}
-#> $b2
-#> [1] 16
-#> 
-#> $a1
-#> [1] 0.08326136
-#> 
-#> $a2
-#> [1] 0.07210645
+> $b2
+> [1] 16
+> 
+> $a1
+> [1] 0.08326136
+> 
+> $a2
+> [1] 0.07210645
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 240 x 3
-#>        h     c_x     c_y
-#>    <int>   <dbl>   <dbl>
-#>  1     1 -0.1    -0.0892
-#>  2     2 -0.0167 -0.0892
-#>  3     3  0.0665 -0.0892
-#>  4     4  0.150  -0.0892
-#>  5     5  0.233  -0.0892
-#>  6     6  0.316  -0.0892
-#>  7     7  0.400  -0.0892
-#>  8     8  0.483  -0.0892
-#>  9     9  0.566  -0.0892
-#> 10    10  0.649  -0.0892
-#> # i 230 more rows
+> # A tibble: 240 x 3
+>        h     c_x     c_y
+>    <int>   <dbl>   <dbl>
+>  1     1 -0.1    -0.0892
+>  2     2 -0.0167 -0.0892
+>  3     3  0.0665 -0.0892
+>  4     4  0.150  -0.0892
+>  5     5  0.233  -0.0892
+>  6     6  0.316  -0.0892
+>  7     7  0.400  -0.0892
+>  8     8  0.483  -0.0892
+>  9     9  0.566  -0.0892
+> 10    10  0.649  -0.0892
+> # i 230 more rows
 \end{verbatim}
@@
 \begin{verbatim}
-#>   hex_poly_id           x           y
-#> 1           1 -0.10000000 -0.04116510
-#> 2           1 -0.14163068 -0.06520059
-#> 3           1 -0.14163068 -0.11327155
-#> 4           1 -0.10000000 -0.13730704
-#> 5           1 -0.05836932 -0.11327155
+>   hex_poly_id           x           y
+> 1           1 -0.10000000 -0.04116510
+> 2           1 -0.14163068 -0.06520059
+> 3           1 -0.14163068 -0.11327155
+> 4           1 -0.10000000 -0.13730704
+> 5           1 -0.05836932 -0.11327155
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5 x 4
-#>    emb1  emb2    ID     h
-#>   <dbl> <dbl> <int> <int>
-#> 1 0.707 0.839     1   205
-#> 2 0.231 0.401     2   109
-#> 3 0.232 0.215     3    65
-#> 4 0.790 0.564     4   146
-#> 5 0.761 0.551     5   146
+> # A tibble: 5 x 4
+>    emb1  emb2    ID     h
+>   <dbl> <dbl> <int> <int>
+> 1 0.707 0.839     1   205
+> 2 0.231 0.401     2   109
+> 3 0.232 0.215     3    65
+> 4 0.790 0.564     4   146
+> 5 0.761 0.551     5   146
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5 x 3
-#>       h   n_h    w_h
-#>   <int> <int>  <dbl>
-#> 1    21    12 0.0024
-#> 2    22    17 0.0034
-#> 3    23    22 0.0044
-#> 4    24    10 0.002 
-#> 5    25     7 0.0014
+> # A tibble: 5 x 3
+>       h   n_h    w_h
+>   <int> <int>  <dbl>
+> 1    21    12 0.0024
+> 2    22    17 0.0034
+> 3    23    22 0.0044
+> 4    24    10 0.002 
+> 5    25     7 0.0014
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5 x 2
-#>       h pts_list  
-#>   <int> <list>    
-#> 1    21 <int [12]>
-#> 2    22 <int [17]>
-#> 3    23 <int [22]>
-#> 4    24 <int [10]>
-#> 5    25 <int [7]>
+> # A tibble: 5 x 2
+>       h pts_list  
+>   <int> <list>    
+> 1    21 <int [12]>
+> 2    22 <int [17]>
+> 3    23 <int [22]>
+> 4    24 <int [10]>
+> 5    25 <int [7]>
 \end{verbatim}
@@
 \begin{verbatim}
-#>   h         c_x         c_y n_h w_h
-#> 1 1 -0.10000000 -0.08923607   0   0
-#> 2 2 -0.01673864 -0.08923607   0   0
-#> 3 3  0.06652271 -0.08923607   0   0
-#> 4 4  0.14978407 -0.08923607   0   0
-#> 5 5  0.23304543 -0.08923607   0   0
+>   h         c_x         c_y n_h w_h
+> 1 1 -0.10000000 -0.08923607   0   0
+> 2 2 -0.01673864 -0.08923607   0   0
+> 3 3  0.06652271 -0.08923607   0   0
+> 4 4  0.14978407 -0.08923607   0   0
+> 5 5  0.23304543 -0.08923607   0   0
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5 x 8
-#>    from    to  x_from  y_from    x_to    y_to from_count to_count
-#>   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-#> 1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
-#> 2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
-#> 3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
-#> 4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
-#> 5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
+> # A tibble: 5 x 8
+>    from    to  x_from  y_from    x_to    y_to from_count to_count
+>   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
+> 1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
+> 2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
+> 3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
+> 4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
+> 5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
 \end{verbatim}
@@
 \begin{verbatim}
-#>   [1]   1   2   3   4   5   6   7   8   9  10  11  12  13  14  15  16  17  18
-#>  [19]  19  20  21  22  23  24  25  26  27  28  29  30  31  32  33  34  35  36
-#>  [37]  37  38  39  40  41  42  43  44  45  46  47  48  49  50  51  52  53  54
-#>  [55]  55  56  57  58  59  60  61  62  63  64  65  66  67  68  69  70  71  72
-#>  [73]  73  74  75  76  77  78  79  80  81  82  83  84  85  86  87  88  89  90
-#>  [91]  91  92  93  94  95  96  97  98  99 100 101 102 103 104 105 106 107 108
-#> [109] 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126
-#> [127] 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144
-#> [145] 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162
-#> [163] 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180
-#> [181] 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198
-#> [199] 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215 216
-#> [217] 217 218 219 220 221 222 223 224 225 226 227 228 229 230 231 232 233 234
-#> [235] 235 236 237 238 239 240
+>   [1]   1   2   3   4   5   6   7   8   9  10  11  12  13  14  15  16  17  18
+>  [19]  19  20  21  22  23  24  25  26  27  28  29  30  31  32  33  34  35  36
+>  [37]  37  38  39  40  41  42  43  44  45  46  47  48  49  50  51  52  53  54
+>  [55]  55  56  57  58  59  60  61  62  63  64  65  66  67  68  69  70  71  72
+>  [73]  73  74  75  76  77  78  79  80  81  82  83  84  85  86  87  88  89  90
+>  [91]  91  92  93  94  95  96  97  98  99 100 101 102 103 104 105 106 107 108
+> [109] 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126
+> [127] 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144
+> [145] 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162
+> [163] 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180
+> [181] 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198
+> [199] 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215 216
+> [217] 217 218 219 220 221 222 223 224 225 226 227 228 229 230 231 232 233 234
+> [235] 235 236 237 238 239 240
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5 x 8
-#>       h      x1    x2    x3         x4        x5       x6        x7
-#>   <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
-#> 1    21 -0.992   1.91  1.11 -0.000427   0.000624  0.00749  0.00105 
-#> 2    22 -0.906   1.93  1.41 -0.0000183  0.00331  -0.0204  -0.000363
-#> 3    23 -0.680   1.93  1.72 -0.000810  -0.00259  -0.00449  0.00153 
-#> 4    24 -0.272   1.93  1.96  0.00251    0.00668  -0.0460   0.00128 
-#> 5    25  0.0760  1.93  2.00  0.00876    0.00447   0.00851 -0.00195
+> # A tibble: 5 x 8
+>       h      x1    x2    x3         x4        x5       x6        x7
+>   <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
+> 1    21 -0.992   1.91  1.11 -0.000427   0.000624  0.00749  0.00105 
+> 2    22 -0.906   1.93  1.41 -0.0000183  0.00331  -0.0204  -0.000363
+> 3    23 -0.680   1.93  1.72 -0.000810  -0.00259  -0.00449  0.00153 
+> 4    24 -0.272   1.93  1.96  0.00251    0.00668  -0.0460   0.00128 
+> 5    25  0.0760  1.93  2.00  0.00876    0.00447   0.00851 -0.00195
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5 x 4
-#>   pred_emb_1 pred_emb_2    ID pred_hb_id
-#>        <dbl>      <dbl> <int>      <int>
-#> 1      0.691      0.848     1        205
-#> 2      0.191      0.416     2        109
-#> 3      0.316      0.199     3         66
-#> 4      0.774      0.560     4        146
-#> 5      0.774      0.560     5        146
+> # A tibble: 5 x 4
+>   pred_emb_1 pred_emb_2    ID pred_hb_id
+>        <dbl>      <dbl> <int>      <int>
+> 1      0.691      0.848     1        205
+> 2      0.191      0.416     2        109
+> 3      0.316      0.199     3         66
+> 4      0.774      0.560     4        146
+> 5      0.774      0.560     5        146
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 1 x 2
-#>   Error  RMSE
-#>   <dbl> <dbl>
-#> 1 1481. 0.180
+> # A tibble: 1 x 2
+>   Error  RMSE
+>   <dbl> <dbl>
+> 1 1481. 0.180
 \end{verbatim}
@@
 \begin{verbatim}
-#> # A tibble: 5 x 32
-#>      ID      x1     x2       x3       x4       x5      x6        x7 pred_hb_id
-#>   <int>   <dbl>  <dbl>    <dbl>    <dbl>    <dbl>   <dbl>     <dbl>      <int>
-#> 1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923 -0.00128         205
-#> 2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362 -0.00535         109
-#> 3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962  0.00335          66
-#> 4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716  0.00126         146
-#> 5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998  0.000677        146
-#> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
-#> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
-#> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
-#> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
-#> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
-#> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>,
-#> #   abs_error_x3 <dbl>, abs_error_x4 <dbl>, abs_error_x5 <dbl>, ...
+> # A tibble: 5 x 32
+>      ID      x1     x2       x3       x4       x5      x6        x7 pred_hb_id
+>   <int>   <dbl>  <dbl>    <dbl>    <dbl>    <dbl>   <dbl>     <dbl>      <int>
+> 1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923 -0.00128         205
+> 2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362 -0.00535         109
+> 3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962  0.00335          66
+> 4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716  0.00126         146
+> 5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998  0.000677        146
+> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
+> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
+> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
+> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
+> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
+> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>,
+> #   abs_error_x3 <dbl>, abs_error_x4 <dbl>, abs_error_x5 <dbl>, ...
 \end{verbatim}

### 2025-07-06 paper-quollr.Rmd
Added Figure title for algorithm plots

@@
 library(htmltools)
+library(crosstalk)
 #library(tools)
@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "[Figure 2](#fig-algo-steps)"} else { "Figure \\@ref(fig:algo-steps)"})`).
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`).
 A detailed description of the algorithm can be found in **cite the methodology paper**.
 
-```{r, eval=knitr::is_html_output()}
-
-htmltools::browsable(
-  tagList(
-    tags$div(
-      id = "fig-algo-steps",  # This is the anchor
-      crosstalk::bscols(
+```{r algo-step-html, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+The `Scurve` data is shown."}
+
+algofig <- crosstalk::bscols(
         htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr; gap: 0;",
@@
         device = "xs"
-      ),
-      tags$p(tags$strong("Figure 2:"), "Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
-The `Scurve` data is shown.")
-    )
-  )
-)
+      )
+
+class(algofig) <- c(class(algofig), "htmlwidget")
+
+algofig
 ```
@@
 Two derived constants are calculated to define the relative distances to the vertices.
-The horizontal and vertical offset is defined as $dx = \frac{a_1}{2}$, and $dy = \frac{a_1}{\sqrt{3}}$ repectively.
-A vertical spacing factor $v = \frac{a_1}{2\sqrt{3}}$ refines vertical placement in staggered rows.
+The horizontal and vertical offset is defined as $dx = a_1/2$, and $dy = a_1/\sqrt{3}$ repectively.
+A vertical spacing factor $v = a_1/2\sqrt{3}$ refines vertical placement in staggered rows.
 

### 2025-07-06 paper-quollr.tex
Added Figure title for algorithm plots

@@
 Two derived constants are calculated to define the relative distances to the vertices.
-The horizontal and vertical offset is defined as \(dx = \frac{a_1}{2}\), and \(dy = \frac{a_1}{\sqrt{3}}\) repectively.
-A vertical spacing factor \(v = \frac{a_1}{2\sqrt{3}}\) refines vertical placement in staggered rows.
+The horizontal and vertical offset is defined as \(dx = a_1/2\), and \(dy = a_1/\sqrt{3}\) repectively.
+A vertical spacing factor \(v = a_1/2\sqrt{3}\) refines vertical placement in staggered rows.
 

### 2025-07-06 paper-quollr.R

@@ -25,6 +25,7 @@ library(readr)
 library(plotly)
 library(crosstalk)
 library(htmltools)
+library(crosstalk)
 #library(tools)
 
 set.seed(20240110)
@@ -407,13 +408,9 @@ hex_grid_scurve + wrap_plots(
   ncol = 2)
 
 
-## ----eval=knitr::is_html_output()---------------------------------------------
+## ----algo-step-html, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions. The `Scurve` data is shown."----
 # 
-# htmltools::browsable(
-#   tagList(
-#     tags$div(
-#       id = "fig-algo-steps",  # This is the anchor
-#       crosstalk::bscols(
+# algofig <- crosstalk::bscols(
 #         htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr; gap: 0;",
 #                        htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_plt_int),
 #                        htmltools::div(style = "margin: 0; padding: 0;", hex_grid_poly_scurve_int),
@@ -424,11 +421,11 @@ hex_grid_scurve + wrap_plots(
 #                          scurve_umap_model_vis_n
 #                        )),
 #         device = "xs"
-#       ),
-#       tags$p(tags$strong("Figure 2:"), "Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions. The `Scurve` data is shown.")
-#     )
-#   )
-# )
+#       )
+# 
+# class(algofig) <- c(class(algofig), "htmlwidget")
+# 
+# algofig
 
 
 ## ----algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions. The `Scurve` data is shown."----

### 2025-07-06 paper-quollr.Rmd
Add figure title for Figure 5

@@
 The `Scurve` data is shown."}
+
 scurve_umap_plt + hex_grid_poly_scurve +
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output()}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.pos='H', fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
+The two twists visible in the UMAP layout can also be seen in the lifted model."}
+
 show_langevitour(
@@
 
-```{r, eval=knitr::is_latex_output()}
-scurve_proj_umap_model1 + scurve_proj_umap_model2 + scurve_proj_umap_model3 + scurve_proj_umap_model4 + 
+```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (in purple) in different projections.
+These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
+The two twists visible in the UMAP layout can also be seen in the lifted model."}
+
+scurve_proj_umap_model1 + scurve_proj_umap_model2 + 
+  scurve_proj_umap_model3 + scurve_proj_umap_model4 + 
   plot_layout(ncol = 4)

### 2025-07-06 paper-quollr.tex
Add figure title for Figure 5

@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1} \caption{$2 ext{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (in purple) in different projections.
+These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
+The two twists visible in the UMAP layout can also be seen in the lifted model.}\label{fig:unnamed-chunk-27}
+\end{figure}
 

### 2025-07-06 paper-quollr.Rmd
Updated link plot colors

@@
   point_data = df_exe, 
-  edge_data = trimesh
+  edge_data = trimesh, 
+  point_colour = clr_choice
   )

### 2025-07-06 paper-quollr.Rmd
Updated application views

@@
 
-```{r, eval=knitr::is_html_output()}
-htmltools::browsable(
-  htmltools::tagList(
-    tags$div(
-      id = "fig-limb-model",  # Anchor for referencing
-      style = "text-align: center; margin-bottom: 1em;",
-
-      # 2x2 grid layout
-      htmltools::div(
+```{r model-limb-html, eval=knitr::is_html_output(), fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$."}
+
+modellimbfig <- htmltools::div(
         style = "display: grid; grid-template-columns: 1fr 1fr; row-gap: 0px; justify-items: center;",
@@
         htmltools::div(style = "margin-top: 13px;", limb_highd_vis_best)
-      ),
-
-      # Caption
-      tags$p(
-        tags$strong("Figure 6:"), 
-        "Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$."
-      )
-    )
-  )
-)
-
+      )
+
+class(modellimbfig) <- c(class(modellimbfig), "htmlwidget")
+
+modellimbfig
 ```

### 2025-07-07 paper-quollr.Rmd
Replaced the word visualisation

@@
 The paper is organized as follows.
-The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
+The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualisation capabilities.
 In the application section, we illustrate the algorithm's functionality for studying a clustering data structure.
@@
 
-### Visualizations
-
-The package offers several $2-\text{D}$ visualizations, including:
+### visualisations
+
+The package offers several $2-\text{D}$ visualisations, including:
 
@@
 
-Each visualization can be generated using its respective function, as described in this section.
+Each visualisation can be generated using its respective function, as described in this section.
 
@@
 
-#### $p\text{-}D$ model visualization
-
-To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualization using the `show_langevitour()` function.
+#### $p\text{-}D$ model visualisation
+
+To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualisation using the `show_langevitour()` function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.
@@
 
-The `show_langevitour()` function then renders the visualization using the `langevitour` interface, displaying both types of points in a dynamic tour.
+The `show_langevitour()` function then renders the visualisation using the `langevitour` interface, displaying both types of points in a dynamic tour.
 The `edge_data` input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model’s structure.
 
-In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2-\text{D}$ wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2-\text{D}$ wireframe structure mapped back to high-dimensional space.
 
@@
 To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
-This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+This visualisation combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
 Both views are interactively linked, enabling users to explore.
 
-Before visualization, the input data must be prepared using the `comb_all_data_model()` function.
+Before visualisation, the input data must be prepared using the `comb_all_data_model()` function.
 This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2-\text{D}$ bin layout (`model_2d`):
@@
 
-Before visualization, the input data must be prepared using the `comb_all_data_model_error()` function.
+Before visualisation, the input data must be prepared using the `comb_all_data_model_error()` function.
 The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2-\text{D}$ model data (`model_2d`), and model error (`error_data`).

### 2025-07-07 paper-quollr.tex
Replaced the word visualisation

@@
 The paper is organized as follows.
-The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
+The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualisation capabilities.
 In the application section, we illustrate the algorithm's functionality for studying a clustering data structure.
@@
 
-\subsubsection{Visualizations}\label{visualizations}
+\subsubsection{visualisations}\label{visualisations}
 
-The package offers several \(2-\text{D}\) visualizations, including:
+The package offers several \(2-\text{D}\) visualisations, including:
 
@@
 
-\paragraph{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
+\paragraph{\texorpdfstring{\(p\text{-}D\) model visualisation}{p\textbackslash text\{-\}D model visualisation}}\label{ptext-d-model-visualisation}
 
-To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualization using the \texttt{show\_langevitour()} function.
+To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualisation using the \texttt{show\_langevitour()} function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.
@@
 
-The \texttt{show\_langevitour()} function then renders the visualization using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour.
+The \texttt{show\_langevitour()} function then renders the visualisation using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour.
 The \texttt{edge\_data} input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model's structure.
 
-In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2-\text{D}\) wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2-\text{D}\) wireframe structure mapped back to high-dimensional space.
 
@@
 To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
-This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+This visualisation combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
 Both views are interactively linked, enabling users to explore.
 
-Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
+Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
 This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2-\text{D}\) bin layout (\texttt{model\_2d}):
@@
 
-Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
+Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
 The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2-\text{D}\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).

### 2025-07-07 paper-quollr.Rmd
Added package installation

@@
 
+```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE}
+
+options(repos = c(CRAN = "https://cran.rstudio.com")) # Setup mirror
+
+packages_to_check <- c("remotes", "quollr", "tibble", "knitr", "kableExtra", "ggplot2", "dplyr", "patchwork", "readr", "plotly", "crosstalk", "htmltools")
+
+for (pkg in packages_to_check) {
+  if (!requireNamespace(pkg, quietly = TRUE)) {
+    message(paste("Installing package:", pkg))
+    install.packages(pkg)
+  } else {
+    installed_version <- packageVersion(pkg)
+    available_version <- tryCatch({
+      utils::packageDescription(pkg)$Version
+    }, error = function(e) NA) # Handle cases where package info isn't readily available
+
+    if (!is.na(available_version) && installed_version < package_version(available_version)) {
+      message(paste("A newer version of package", pkg, "is available. Updating..."))
+      install.packages(pkg)
+    } else {
+      message(paste("Package", pkg, "is up-to-date (version", installed_version, ")."))
+    }
+  }
+}
+
+```
+
 ```{r load-libraries}
@@
 library(htmltools)
-library(crosstalk)
-#library(tools)
 
@@
 
-### visualisations
+### Visualisations
 

### 2025-07-07 paper-quollr.R

@@ -13,6 +13,33 @@ knitr::opts_chunk$set(
 
 
 
+## ----install-libraries, include=FALSE, warning=FALSE, echo=FALSE--------------
+
+options(repos = c(CRAN = "https://cran.rstudio.com")) # Setup mirror
+
+packages_to_check <- c("remotes", "quollr", "tibble", "knitr", "kableExtra", "ggplot2", "dplyr", "patchwork", "readr", "plotly", "crosstalk", "htmltools")
+
+for (pkg in packages_to_check) {
+  if (!requireNamespace(pkg, quietly = TRUE)) {
+    message(paste("Installing package:", pkg))
+    install.packages(pkg)
+  } else {
+    installed_version <- packageVersion(pkg)
+    available_version <- tryCatch({
+      utils::packageDescription(pkg)$Version
+    }, error = function(e) NA) # Handle cases where package info isn't readily available
+
+    if (!is.na(available_version) && installed_version < package_version(available_version)) {
+      message(paste("A newer version of package", pkg, "is available. Updating..."))
+      install.packages(pkg)
+    } else {
+      message(paste("Package", pkg, "is up-to-date (version", installed_version, ")."))
+    }
+  }
+}
+
+
+
 ## ----load-libraries-----------------------------------------------------------
 library(quollr)
 library(tibble)
@@ -25,8 +52,6 @@ library(readr)
 library(plotly)
 library(crosstalk)
 library(htmltools)
-library(crosstalk)
-#library(tools)
 
 set.seed(20240110)
 

### 2025-07-08 paper-quollr.Rmd
converted the htmlwidget objects

@@
 ```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="jkjio"}
-show_link_plots(
+nldrdt_link <- show_link_plots(
   point_data = df_exe, 
@@
   )
+
+class(nldrdt_link) <- c(class(nldrdt_link), "htmlwidget")
+
+nldrdt_link
 ```
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output()}
-show_error_link_plots(
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap=""}
+errornldrdt_link <- show_error_link_plots(
   point_data = df_exe, 
@@
 )
+
+class(errornldrdt_link) <- c(class(errornldrdt_link), "htmlwidget")
+
+errornldrdt_link
 ```

### 2025-07-08 paper-quollr.tex
Added package installation
converted the htmlwidget objects

@@
 
-\subsubsection{visualisations}\label{visualisations}
+\subsubsection{Visualisations}\label{visualisations}
 
@@
 \begin{verbatim}
-show_link_plots(
+nldrdt_link <- show_link_plots(
   point_data = df_exe, 
@@
   )
+
+class(nldrdt_link) <- c(class(nldrdt_link), "htmlwidget")
+
+nldrdt_link
 \end{verbatim}
@@
 \begin{verbatim}
-show_error_link_plots(
+errornldrdt_link <- show_error_link_plots(
   point_data = df_exe, 
@@
 )
+
+class(errornldrdt_link) <- c(class(errornldrdt_link), "htmlwidget")
+
+errornldrdt_link
 \end{verbatim}

### 2025-08-04 paper-quollr.Rmd
Updated the paper with new version

@@
   hex_grid_scurve, counts_df_scurve, 
-  by = c("hex_poly_id" = "h"))
+  by = c("h" = "h"))
 
 hex_grid_nonempty_scurve <- hex_grid_scurve |>
-  filter(hex_poly_id %in% df_bin_centroids_scurve$h)
+  filter(h %in% df_bin_centroids_scurve$h)
 
@@
   geom_polygon(color = "grey70", 
-               aes(group = hex_poly_id), 
+               aes(group = h), 
                fill = "#ffffff") +
@@
   geom_polygon(color = "grey70", 
-               aes(group = hex_poly_id), 
+               aes(group = h), 
                fill = "#ffffff")  +
@@
 ```r
-devtools::install_github("JayaniLakshika/quollr")
+pak::pak("JayaniLakshika/quollr")
 ```
@@
 hex_grid_temp45 <- hex_grid_temp |> 
-  filter(hex_poly_id == 45)
+  filter(h == 45)
 
@@
                             y = y, 
-                            group = hex_poly_id),
+                            group = h),
                  fill = "white", 
@@
 
-For each centroid, six vertices are computed and assigned a polygon ID (`hex_poly_id`) corresponding to the centroid’s `h`.
-These points are aggregated into a tibble that includes the polygon ID (`hex_poly_id`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.
+For each centroid, six vertices are computed and assigned a polygon ID (`h`) corresponding to the centroid’s `h`.
+These points are aggregated into a tibble that includes the polygon ID (`h`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.
 
@@
 find_low_dens_hex(
-  centroids_data = df_bin_centroids, 
+  model_2d = df_bin_centroids, 
   b1 = 15, 
@@
 
-```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="jkjio"}
+```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+
 nldrdt_link <- show_link_plots(
@@
 # 
-# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "h"))
+# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("h" = "h"))
 # 

### 2025-08-04 paper-quollr.tex
Updated the paper with new version

@@
 \begin{verbatim}
-devtools::install_github("JayaniLakshika/quollr")
+pak::pak("JayaniLakshika/quollr")
 \end{verbatim}
@@
 > $scaled_nldr
-> # A tibble: 5,000 x 3
+> # A tibble: 1,000 x 3
 >     emb1  emb2    ID
 >    <dbl> <dbl> <int>
->  1 0.707 0.839     1
->  2 0.231 0.401     2
->  3 0.232 0.215     3
->  4 0.790 0.564     4
->  5 0.761 0.551     5
->  6 0.445 0.721     6
->  7 0.900 0.137     7
->  8 0.247 0.392     8
->  9 0.325 0.542     9
-> 10 0.278 0.231    10
-> # i 4,990 more rows
+>  1 0.277 0.913     1
+>  2 0.697 0.538     2
+>  3 0.779 0.399     3
+>  4 0.173 0.953     4
+>  5 0.218 0.983     5
+>  6 0.593 1.05      6
+>  7 0.180 0.210     7
+>  8 0.976 0.571     8
+>  9 0.803 0.829     9
+> 10 0.932 0.410    10
+> # i 990 more rows
 > 
 > $lim1
-> [1] -14.42166  13.32655
+> [1] -9.146398  8.552211
 > 
 > $lim2
-> [1] -12.43687  12.32455
+> [1] -10.36686  10.10691
 \end{verbatim}
@@
 > $b2
-> [1] 16
+> [1] 20
 > 
 > $a1
-> [1] 0.08326136
+> [1] 0.0834108
 > 
 > $a2
-> [1] 0.07210645
+> [1] 0.07223587
 \end{verbatim}
@@
 \begin{verbatim}
-> # A tibble: 240 x 3
+> # A tibble: 300 x 3
 >        h     c_x    c_y
 >    <int>   <dbl>  <dbl>
->  1     1 -0.1    -0.0892
->  2     2 -0.0167 -0.0892
->  3     3  0.0665 -0.0892
->  4     4  0.150  -0.0892
->  5     5  0.233  -0.0892
->  6     6  0.316  -0.0892
->  7     7  0.400  -0.0892
->  8     8  0.483  -0.0892
->  9     9  0.566  -0.0892
-> 10    10  0.649  -0.0892
-> # i 230 more rows
+>  1     1 -0.1    -0.116
+>  2     2 -0.0166 -0.116
+>  3     3  0.0668 -0.116
+>  4     4  0.150  -0.116
+>  5     5  0.234  -0.116
+>  6     6  0.317  -0.116
+>  7     7  0.400  -0.116
+>  8     8  0.484  -0.116
+>  9     9  0.567  -0.116
+> 10    10  0.651  -0.116
+> # i 290 more rows
 \end{verbatim}
@@
 
-For each centroid, six vertices are computed and assigned a polygon ID (\texttt{hex\_poly\_id}) corresponding to the centroid's \texttt{h}.
-These points are aggregated into a tibble that includes the polygon ID (\texttt{hex\_poly\_id}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
+For each centroid, six vertices are computed and assigned a polygon ID (\texttt{h}) corresponding to the centroid's \texttt{h}.
+These points are aggregated into a tibble that includes the polygon ID (\texttt{h}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
 
@@
 \begin{verbatim}
->   hex_poly_id           x           y
-> 1           1 -0.10000000 -0.04116510
-> 2           1 -0.14163068 -0.06520059
-> 3           1 -0.14163068 -0.11327155
-> 4           1 -0.10000000 -0.13730704
-> 5           1 -0.05836932 -0.11327155
+>   h          x           y
+> 1 1 -0.1000000 -0.06752289
+> 2 1 -0.1417054 -0.09160152
+> 3 1 -0.1417054 -0.13975877
+> 4 1 -0.1000000 -0.16383739
+> 5 1 -0.0582946 -0.13975877
 \end{verbatim}
@@
 >   <dbl> <dbl> <int> <int>
-> 1 0.707 0.839     1   205
-> 2 0.231 0.401     2   109
-> 3 0.232 0.215     3    65
-> 4 0.790 0.564     4   146
-> 5 0.761 0.551     5   146
+> 1 0.277 0.913     1   216
+> 2 0.697 0.538     2   145
+> 3 0.779 0.399     3   116
+> 4 0.173 0.953     4   229
+> 5 0.218 0.983     5   229
 \end{verbatim}
@@
 >   <int> <int> <dbl>
-> 1    21    12 0.0024
-> 2    22    17 0.0034
-> 3    23    22 0.0044
-> 4    24    10 0.002 
-> 5    25     7 0.0014
+> 1    34     1 0.001
+> 2    35    10 0.01 
+> 3    36     8 0.008
+> 4    37     8 0.008
+> 5    38     2 0.002
 \end{verbatim}
@@
 >   <int> <list>    
-> 1    21 <int [12]>
-> 2    22 <int [17]>
-> 3    23 <int [22]>
-> 4    24 <int [10]>
-> 5    25 <int [7]>
+> 1    34 <int [1]> 
+> 2    35 <int [10]>
+> 3    36 <int [8]> 
+> 4    37 <int [8]> 
+> 5    38 <int [2]>
 \end{verbatim}
@@
 >   h        c_x        c_y n_h w_h
-> 1 1 -0.10000000 -0.08923607   0   0
-> 2 2 -0.01673864 -0.08923607   0   0
-> 3 3  0.06652271 -0.08923607   0   0
-> 4 4  0.14978407 -0.08923607   0   0
-> 5 5  0.23304543 -0.08923607   0   0
+> 1 1 -0.1000000 -0.1156801   0   0
+> 2 2 -0.0165892 -0.1156801   0   0
+> 3 3  0.0668216 -0.1156801   0   0
+> 4 4  0.1502324 -0.1156801   0   0
+> 5 5  0.2336432 -0.1156801   0   0
 \end{verbatim}
@@
 >   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
-> 2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
-> 3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
-> 4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
-> 5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
+> 1     1     2 -0.1    -0.116  -0.0166 -0.116           0        0
+> 2    16    17 -0.0583 -0.0434  0.0251 -0.0434          0        0
+> 3    16    32 -0.0583 -0.0434 -0.0166  0.0288          0        0
+> 4     3    17  0.0668 -0.116   0.0251 -0.0434          0        0
+> 5    32    33 -0.0166  0.0288  0.0668  0.0288          0        0
 \end{verbatim}
@@
 > [217] 217 218 219 220 221 222 223 224 225 226 227 228 229 230 231 232 233 234
-> [235] 235 236 237 238 239 240
+> [235] 235 236 237 238 239 240 241 242 243 244 245 246 247 248 249 250 251 252
+> [253] 253 254 255 256 257 258 259 260 261 262 263 264 265 266 267 268 269 270
+> [271] 271 272 273 274 275 276 277 278 279 280 281 282 283 284 285 286 287 288
+> [289] 289 290 291 292 293 294 295 296 297 298 299 300
 \end{verbatim}
@@
 >   <int>  <dbl>  <dbl> <dbl>      <dbl>     <dbl>   <dbl>      <dbl>
-> 1    21 -0.992   1.91  1.11 -0.000427   0.000624  0.00749  0.00105 
-> 2    22 -0.906   1.93  1.41 -0.0000183  0.00331  -0.0204  -0.000363
-> 3    23 -0.680   1.93  1.72 -0.000810  -0.00259  -0.00449  0.00153 
-> 4    24 -0.272   1.93  1.96  0.00251    0.00668  -0.0460   0.00128 
-> 5    25  0.0760  1.93  2.00  0.00876    0.00447   0.00851 -0.00195
+> 1    34 0.958  0.0854  1.29  0.00265    0.0171    0.0876 -0.00249  
+> 2    35 0.818  0.116   1.56  0.00184    0.00361  -0.0318 -0.00377  
+> 3    36 0.544  0.111   1.83 -0.00341   -0.000303  0.0196  0.0000704
+> 4    37 0.279  0.128   1.95  0.0000880  0.00104  -0.0276 -0.000227 
+> 5    38 0.0567 0.119   2.00  0.00733    0.00238   0.0833 -0.00166
 \end{verbatim}
@@
 > # A tibble: 5 x 4
->   pred_emb_1 pred_emb_2    ID pred_hb_id
+>   pred_emb_1 pred_emb_2    ID pred_h
 >        <dbl>      <dbl> <int>  <int>
-> 1      0.691      0.848     1        205
-> 2      0.191      0.416     2        109
-> 3      0.316      0.199     3         66
-> 4      0.774      0.560     4        146
-> 5      0.774      0.560     5        146
+> 1         NA         NA     1    215
+> 2         NA         NA     2    145
+> 3         NA         NA     3    116
+> 4         NA         NA     4    229
+> 5         NA         NA     5    229
 \end{verbatim}
@@
 >   <dbl> <dbl>
-> 1 1481. 0.180
+> 1  265. 0.157
 \end{verbatim}
@@
 > # A tibble: 5 x 32
->      ID      x1     x2       x3       x4       x5      x6        x7 pred_hb_id
+>      ID      x1    x2       x3       x4       x5      x6       x7 pred_h
 >   <int>   <dbl> <dbl>    <dbl>    <dbl>    <dbl>   <dbl>    <dbl>  <int>
-> 1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923 -0.00128         205
-> 2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362 -0.00535         109
-> 3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962  0.00335          66
-> 4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716  0.00126         146
-> 5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998  0.000677        146
+> 1     1 -0.120  0.819 -1.99     0.0114   0.00351  0.0334  0.00638    215
+> 2     2 -0.0492 0.166  0.00121  0.0115  -0.0166  -0.0297  0.00509    145
+> 3     3 -0.774  0.651  0.367   -0.0172   0.00600  0.0211  0.00303    116
+> 4     4 -0.606  0.952 -1.80     0.0157  -0.00978 -0.0590 -0.00754    229
+> 5     5 -0.478  1.10  -1.88    -0.00423  0.00495 -0.0482 -0.00982    229
 > # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,

### 2025-08-04 paper-quollr.R

@@ -108,10 +108,10 @@ counts_df_scurve <- scurve_model_obj$hb_obj$std_cts
 
 hex_grid_with_counts_scurve <- left_join(
   hex_grid_scurve, counts_df_scurve, 
-  by = c("hex_poly_id" = "h"))
+  by = c("h" = "h"))
 
 hex_grid_nonempty_scurve <- hex_grid_scurve |>
-  filter(hex_poly_id %in% df_bin_centroids_scurve$h)
+  filter(h %in% df_bin_centroids_scurve$h)
 
 sc_xlims <- c(-0.15, 1.15)
 sc_ylims <- c(-0.14, 1.05)
@@ -141,7 +141,7 @@ hex_grid_poly_scurve <- ggplot(
   data = hex_grid_with_counts_scurve, 
   aes(x = x, y = y)) +
   geom_polygon(color = "grey70", 
-               aes(group = hex_poly_id), 
+               aes(group = h), 
                fill = "#ffffff") +
   geom_point(data = scurve_umap_scaled, 
              aes(x = emb1, y = emb2), 
@@ -167,7 +167,7 @@ hex_centroids_scurve <- ggplot(
   data = hex_grid_with_counts_scurve, 
   aes(x = x, y = y)) +
   geom_polygon(color = "grey70", 
-               aes(group = hex_poly_id), 
+               aes(group = h), 
                fill = "#ffffff")  +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
@@ -510,7 +510,7 @@ all_centroids_df_temp <- hb_obj_notation$centroids
 hex_grid_temp <- hb_obj_notation$hex_poly
 
 hex_grid_temp45 <- hex_grid_temp |> 
-  filter(hex_poly_id == 45)
+  filter(h == 45)
 
 start_pt <- all_centroids_df_temp |> 
   filter(h == 1)
@@ -544,7 +544,7 @@ hex_param_vis <- ggplot() +
     geom_polygon(data = hex_grid_temp, 
                         aes(x = x, 
                             y = y, 
-                            group = hex_poly_id),
+                            group = h),
                  fill = "white", 
                  color = "#bdbdbd") +
     geom_point(data = all_centroids_df_temp, aes(
@@ -722,7 +722,7 @@ head(trimesh, 5)
 
[…]

### 2025-08-04 paper-quollr.Rmd
Fixed some issue with new data

@@
   q = 0.1, 
-  benchmark_highdens = 5)
+  benchmark_highdens = 1)
 
@@
 The paper is organized as follows.
-The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualisation capabilities.
+The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
 In the application section, we illustrate the algorithm's functionality for studying a clustering data structure.
@@
 Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`).
-A detailed description of the algorithm can be found in **cite the methodology paper**.
+A detailed description of the algorithm can be found in @gamage2025c.
 
@@
 `scurve` is a $7\text{-}D$ simulated dataset.
-It is constructed by simulating $5000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
+It is constructed by simulating $1000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
 The remaining variables $X_4$, $X_5$, $X_6$, $X_7$ are all uniform error, with small variance.
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2-\text{D}$ space.
+Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2\text{-}D$ space.
 
@@
 After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The `assign_data()` function performs this assignment by calculating the $2-\text{D}$ Euclidean distance between each point in the $2-\text{D}$ embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2-\text{D}$ layout.
+The `assign_data()` function performs this assignment by calculating the $2\text{-}D$ Euclidean distance between each point in the $2\text{-}D$ embedding and all hexagon centroids.
+
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2\text{-}D$ layout.
 It then selects the corresponding x and y coordinates of each hexagon’s centroid.
@@
 
-The result is a data frame of the scaled $2-\text{D}$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
+The result is a data frame of the scaled $2\text{-}D$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
 
@@
 
-The function first groups the input data by `h`, which represents the hexagon label associated with each point in the $2-\text{D}$ layout.
+The function first groups the input data by `h`, which represents the hexagon label associated with each point in the $2\text{-}D$ layout.
 Within each group, it collects the `ID`s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
@@
 
-To represent the neighborhood structure of hexagonal bins in a $2-\text{D}$ layout, we employ Delaunay triangulation on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
@@
 Some may have very few or no data points due to the sparsity or shape of the underlying structure.
-Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or "holes" in the $2-\text{D}$ structure, potentially disrupting the continuity of the representation.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or "holes" in the $2\text{-}D$ structure, potentially disrupting the continuity of the representation.
 
@@
 The function starts by renaming columns in the `model_highd` data frame to avoid conflicts.
-It then predicts the high-dimensional bin assignments for each point using the `predict_emb()` function, mapping each observation to its nearest $2-\text{D}$ bin (`pred_hb_id`).
+It then predicts the high-dimensional bin assignments for each point using the `predict_emb()` function, mapping each observation to its nearest $2\text{-}D$ bin (`pred_hb_id`).
 This prediction is joined with the mean high-dimensional coordinates of each bin from the model and with the original high-dimensional data.
@@
 
-The package offers several $2-\text{D}$ visualisations, including:
+The package offers several $2\text{-}D$ visualisations, including:
 
@@
 The generated p-D model, overlaid with the data, can also be visualized using `show_langevitour`.
-Additionally, it features a function for visualizing the $2-\text{D}$ projection of the fitted model overlaid on the data, called `plot_proj`.
+Additionally, it features a function for visualizing the $2\text{-}D$ projection of the fitted model overlaid on the data, called `plot_proj`.
 
@@
 
-In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2-\text{D}$ wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped back to high-dimensional space.
 
@@
 Before visualisation, the input data must be prepared using the `comb_all_data_model()` function.
-This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2-\text{D}$ bin layout (`model_2d`):
+This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2\text{-}D$ bin layout (`model_2d`):
 
@@
 
-The function `show_link_plots()` generates two side-by-side, interactively linked plots; a $2-\text{D}$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function `show_link_plots()` generates two side-by-side, interactively linked plots; a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
 The function takes the output from `comb_all_data_model()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
@@
 Before visualisation, the input data must be prepared using the `comb_all_data_model_error()` function.
-The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2-\text{D}$ model data (`model_2d`), and model error (`error_data`).
+The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2\text{-}D$ model data (`model_2d`), and model error (`error_data`).
 
@@
 
-The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2-\text{D}$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
 The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
@@
 Several core computations within `quollr` are optimized using compiled C++ code via the `Rcpp` and `RcppArmadillo` packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), $2-\text{D}$ distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), $2\text{-}D$ distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
 This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better $2-\text{D}$ layout.
+Our method help to provide a better $2\text{-}D$ layout.
 
@@
 
-The `quollr` package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in $2-\text{D}$ and lifting it into high-dimensional space.
-This lifted model provides a direct way to assess how well a $2-\text{D}$ layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
-The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate $2-\text{D}$ representations.
+The `quollr` package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in $2\text{-}D$ and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a $2\text{-}D$ layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate $2\text{-}D$ representations.
 
@@
 To support efficient computation, particularly for large-scale datasets, several core functions in `quollr` are implemented in C++ using `Rcpp` and `RcppArmadillo`.
-These include functions for computing Euclidean distances in high-dimensional and $2-\text{D}$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RMSE and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in $2-\text{D}$.
+These include functions for computing Euclidean distances in high-dimensional and $2\text{-}D$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RMSE and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in $2\text{-}D$.
 Additionally, `gen_hex_coord_cpp()` constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-08-04 paper-quollr.tex
Fixed some issue with new data

@@
 The paper is organized as follows.
-The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualisation capabilities.
+The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
 In the application section, we illustrate the algorithm's functionality for studying a clustering data structure.
@@
 Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (Figure \ref{fig:algo-steps}).
-A detailed description of the algorithm can be found in \textbf{cite the methodology paper}.
+A detailed description of the algorithm can be found in \citet{gamage2025c}.
 
@@
 \texttt{scurve} is a \(7\text{-}D\) simulated dataset.
-It is constructed by simulating \(5000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\).
+It is constructed by simulating \(1000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\).
 The remaining variables \(X_4\), \(X_5\), \(X_6\), \(X_7\) are all uniform error, with small variance.
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2-\text{D}\) space.
+Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2\text{-}D\) space.
 
@@
 After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The \texttt{assign\_data()} function performs this assignment by calculating the \(2-\text{D}\) Euclidean distance between each point in the \(2-\text{D}\) embedding and all hexagon centroids.
+The \texttt{assign\_data()} function performs this assignment by calculating the \(2\text{-}D\) Euclidean distance between each point in the \(2\text{-}D\) embedding and all hexagon centroids.
 
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \(2-\text{D}\) layout.
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \(2\text{-}D\) layout.
 It then selects the corresponding x and y coordinates of each hexagon's centroid.
@@
 
-The result is a data frame of the scaled \(2-\text{D}\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
+The result is a data frame of the scaled \(2\text{-}D\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
 
@@
 
-The function first groups the input data by \texttt{h}, which represents the hexagon label associated with each point in the \(2-\text{D}\) layout.
+The function first groups the input data by \texttt{h}, which represents the hexagon label associated with each point in the \(2\text{-}D\) layout.
 Within each group, it collects the \texttt{ID}s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
@@
 
-To represent the neighborhood structure of hexagonal bins in a \(2-\text{D}\) layout, we employ Delaunay triangulation on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a \(2\text{-}D\) layout, we employ Delaunay triangulation on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
@@
 Some may have very few or no data points due to the sparsity or shape of the underlying structure.
-Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the \(2-\text{D}\) structure, potentially disrupting the continuity of the representation.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the \(2\text{-}D\) structure, potentially disrupting the continuity of the representation.
 
@@
 The function starts by renaming columns in the \texttt{model\_highd} data frame to avoid conflicts.
-It then predicts the high-dimensional bin assignments for each point using the \texttt{predict\_emb()} function, mapping each observation to its nearest \(2-\text{D}\) bin (\texttt{pred\_hb\_id}).
+It then predicts the high-dimensional bin assignments for each point using the \texttt{predict\_emb()} function, mapping each observation to its nearest \(2\text{-}D\) bin (\texttt{pred\_hb\_id}).
 This prediction is joined with the mean high-dimensional coordinates of each bin from the model and with the original high-dimensional data.
@@
 
-The package offers several \(2-\text{D}\) visualisations, including:
+The package offers several \(2\text{-}D\) visualisations, including:
 
@@
 The generated p-D model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
-Additionally, it features a function for visualizing the \(2-\text{D}\) projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
+Additionally, it features a function for visualizing the \(2\text{-}D\) projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
 
@@
 
-In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2-\text{D}\) wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2\text{-}D\) wireframe structure mapped back to high-dimensional space.
 
@@
 Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
-This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2-\text{D}\) bin layout (\texttt{model\_2d}):
+This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2\text{-}D\) bin layout (\texttt{model\_2d}):
 
@@
 
-The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a \(2-\text{D}\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
 The function takes the output from \texttt{comb\_all\_data\_model()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
@@
 Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
-The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2-\text{D}\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
+The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2\text{-}D\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
 
@@
 
-The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2-\text{D}\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
 The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
@@
 Several core computations within \texttt{quollr} are optimized using compiled C++ code via the \texttt{Rcpp} and \texttt{RcppArmadillo} packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), \(2-\text{D}\) distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), \(2\text{-}D\) distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
 This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better \(2-\text{D}\) layout.
+Our method help to provide a better \(2\text{-}D\) layout.
 
@@
 
-The \texttt{quollr} package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in \(2-\text{D}\) and lifting it into high-dimensional space.
-This lifted model provides a direct way to assess how well a \(2-\text{D}\) layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
-The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate \(2-\text{D}\) representations.
+The \texttt{quollr} package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in \(2\text{-}D\) and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a \(2\text{-}D\) layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate \(2\text{-}D\) representations.
 
@@
 To support efficient computation, particularly for large-scale datasets, several core functions in \texttt{quollr} are implemented in C++ using \texttt{Rcpp} and \texttt{RcppArmadillo}.
-These include functions for computing Euclidean distances in high-dimensional and \(2-\text{D}\) space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RMSE and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in \(2-\text{D}\).
+These include functions for computing Euclidean distances in high-dimensional and \(2\text{-}D\) space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RMSE and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in \(2\text{-}D\).
 Additionally, \texttt{gen\_hex\_coord\_cpp()} constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-08-04 paper-quollr.R

@@ -97,7 +97,7 @@ scurve_model_obj <- fit_highd_model(
   nldr_data = scurve_umap, 
   b1 = 15, 
   q = 0.1, 
-  benchmark_highdens = 5)
+  benchmark_highdens = 1)
 
 scurve_umap_scaled <- scurve_model_obj$nldr_obj$scaled_nldr
 tr_from_to_df_scurve <- scurve_model_obj$trimesh_data

### 2025-08-04 paper-quollr.Rmd
Fixed the error with low density bins

@@
   q = 0.1, 
-  benchmark_highdens = 5)
+  benchmark_highdens = 1)
 ```
@@
 df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(n_h > 10)
+  dplyr::filter(n_h > 1)
 
 trimesh <- trimesh |>
-  dplyr::filter(from_count > 10,
-                to_count > 10)
+  dplyr::filter(from_count > 1,
+                to_count > 1)
 

### 2025-08-04 paper-quollr.tex
Fixed the error with low density bins

@@
   q = 0.1, 
-  benchmark_highdens = 5)
+  benchmark_highdens = 1)
 \end{verbatim}
@@
 >        <dbl>      <dbl> <int>  <int>
-> 1         NA         NA     1    215
-> 2         NA         NA     2    145
-> 3         NA         NA     3    116
-> 4         NA         NA     4    229
-> 5         NA         NA     5    229
+> 1      0.234      0.896     1    215
+> 2      0.692      0.534     2    145
+> 3      0.776      0.390     3    116
+> 4      0.192      0.968     4    229
+> 5      0.192      0.968     5    229
 \end{verbatim}

### 2025-08-04 paper-quollr.R

@@ -467,7 +467,7 @@ scurve_umap_plt + hex_grid_poly_scurve +
 #   nldr_data = scurve_umap,
 #   b1 = 15,
 #   q = 0.1,
-#   benchmark_highdens = 5)
+#   benchmark_highdens = 1)
 
 
 ## ----echo=TRUE----------------------------------------------------------------
@@ -731,11 +731,11 @@ find_low_dens_hex(
 
 ## ----echo=TRUE----------------------------------------------------------------
 df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(n_h > 10)
+  dplyr::filter(n_h > 1)
 
 trimesh <- trimesh |>
-  dplyr::filter(from_count > 10,
-                to_count > 10)
+  dplyr::filter(from_count > 1,
+                to_count > 1)
 
 trimesh <- update_trimesh_index(trimesh)
 

### 2025-08-04 paper-quollr.Rmd
Updated color of data
Updated the limits of the plot

@@
 ```{r}
-clr_choice <- "#bcbddc"
+clr_choice <- "#66B2CC"
 
@@
 
-sc_xlims <- c(-0.15, 1.15)
-sc_ylims <- c(-0.14, 1.05)
+sc_xlims <- c(-0.25, 1.25)
+sc_ylims <- c(-0.25, 1.25)
 
@@
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c") +
+             size = 1, color = "#FF7755") +
   interior_annotation("c") +
@@
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c") +
+             size = 1, color = "#FF7755") +
   interior_annotation("d") +
@@
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c")
+             size = 1, color = "#FF7755")
 
@@
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
-                         levelColors = c(clr_choice, "#33a02c"),
+                         levelColors = c(clr_choice, "#FF7755"),
                          enableControls = FALSE,

### 2025-08-04 paper-quollr.R

@@ -113,8 +113,8 @@ hex_grid_with_counts_scurve <- left_join(
 hex_grid_nonempty_scurve <- hex_grid_scurve |>
   filter(h %in% df_bin_centroids_scurve$h)
 
-sc_xlims <- c(-0.15, 1.15)
-sc_ylims <- c(-0.14, 1.05)
+sc_xlims <- c(-0.25, 1.25)
+sc_ylims <- c(-0.25, 1.25)
 
 scurve_umap_plt <- ggplot(
   scurve_umap_scaled, 
@@ -171,7 +171,7 @@ hex_centroids_scurve <- ggplot(
                fill = "#ffffff")  +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c") +
+             size = 1, color = "#FF7755") +
   interior_annotation("c") +
   xlim(sc_xlims) +
   ylim(sc_ylims)
@@ -199,7 +199,7 @@ wireframe_scurve <- ggplot() +
                colour = "#000000") +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c") +
+             size = 1, color = "#FF7755") +
   interior_annotation("d") +
   xlim(sc_xlims) +
   ylim(sc_ylims)
@@ -233,7 +233,7 @@ hex_grid_scurve <- ggplot(
                colour = "#000000") +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
-             size = 1, color = "#33a02c")
+             size = 1, color = "#FF7755")
 
 hex_grid_scurve_int <- ggplotly(hex_grid_scurve, 
                                 width = "350", 
@@ -267,7 +267,7 @@ scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          lineTo = tr_from_to_df_scurve$to,
                          group = df_exe$type,
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
-                         levelColors = c(clr_choice, "#33a02c"),
+                         levelColors = c(clr_choice, "#FF7755"),
                          enableControls = FALSE,
                          width = "421px", height = "421px")
 
@@ -276,7 +276,7 @@ scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)]
[…]

### 2025-08-04 paper-quollr.Rmd
Updated the limit

@@
 
-The generated p-D model, overlaid with the data, can also be visualized using `show_langevitour`.
+The generated $p\text{-}D$ model, overlaid with the data, can also be visualized using `show_langevitour`.
 Additionally, it features a function for visualizing the $2\text{-}D$ projection of the fitted model overlaid on the data, called `plot_proj`.
@@
              y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
   interior_annotation("d")
@@
   scale_x_continuous(breaks = sort(unique(error_limb$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf'))
+  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#984ea3','#4daf4a','#ff7f00'))
 
@@
   q = 0.1, 
-  benchmark_highdens = 5)
+  benchmark_highdens = 0)
 
@@
   q = 0.1, 
-  benchmark_highdens = 1)
+  benchmark_highdens = 0)
 

### 2025-08-04 paper-quollr.R

@@ -933,7 +933,7 @@ trimap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_trimap_n-inliers_12
 nldr4 <- trimap_limb |>
   ggplot(aes(x = emb1,
              y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
   interior_annotation("d")
 
 pacmap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
@@ -989,7 +989,7 @@ error_limb <- error_limb |>
 
 error_plot_limb <- plot_rmse(error_limb) +
   scale_x_continuous(breaks = sort(unique(error_limb$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf')) 
+  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#984ea3','#4daf4a','#ff7f00')) 
 
 
 
@@ -1016,7 +1016,7 @@ tsne_limb_obj <- fit_highd_model(
   nldr_data = tsne_limb, 
   b1 = 19, 
   q = 0.1, 
-  benchmark_highdens = 5)
+  benchmark_highdens = 0)
 
 df_bin_centroids_limb <- tsne_limb_obj$model_2d
 df_bin_limb <- tsne_limb_obj$model_highd
@@ -1190,7 +1190,7 @@ tsne_best_limb_obj <- fit_highd_model(
   nldr_data = tsne_limb2, 
   b1 = 19, 
   q = 0.1, 
-  benchmark_highdens = 1)
+  benchmark_highdens = 0)
 
 df_bin_centroids_limb <- tsne_best_limb_obj$model_2d
 df_bin_limb <- tsne_best_limb_obj$model_highd

### 2025-08-17 paper-quollr.Rmd
Initial work adding NLDR link plots

@@
 
-```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (in purple) in different projections.
@@
 
-```{r, eval=knitr::is_latex_output()}
-
+```{r}
+## Create the first row of selection
+scurve_umap_scaled_select1 <- scurve_umap_scaled |>
+  mutate(select_area = if_else(emb1 <= 0.7 & emb1 >= 0.4 & emb2 <= 0.25 & emb2 >= 0, "selected", "deselected"))
+
+scurve_umap_scaled_select1_selected <- scurve_umap_scaled_select1 |>
+  filter(select_area == "selected")
+
+scurve_umap_scaled_select1_deselected <- scurve_umap_scaled_select1 |>
+  filter(select_area == "deselected")
+
+scurve_umap_plt_select1 <- ggplot(
+  scurve_umap_scaled_select1_deselected, 
+  aes(x = emb1, y = emb2)) +
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
+  geom_point(data = scurve_umap_scaled_select1_selected, aes(x = emb1, y = emb2), alpha = 0.5, color = "#756bb1", size = 0.5) +
+  interior_annotation("a1") +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
+
+## Create the second row of selection
+scurve_umap_scaled_select2 <- scurve_umap_scaled |>
+  mutate(select_area = if_else(emb1 <= 0.55 & emb1 >= 0.25 & emb2 <= 1.25 & emb2 >= 0.75, "selected", "deselected"))
+
+scurve_umap_scaled_select2_selected <- scurve_umap_scaled_select2 |>
+  filter(select_area == "selected")
+
+scurve_umap_scaled_select2_deselected <- scurve_umap_scaled_select2 |>
+  filter(select_area == "deselected")
+
+scurve_umap_plt_select2 <- ggplot(
+  scurve_umap_scaled_select2_deselected, 
+  aes(x = emb1, y = emb2)) +
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
+  geom_point(data = scurve_umap_scaled_select2_selected, aes(x = emb1, y = emb2), alpha = 0.5, color = "#756bb1", size = 0.5) +
+  interior_annotation("b1") +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
+```
+
+```{r, eval=knitr::is_latex_output(), fig.pos='H'}
+scurve_umap_plt_select1 + scurve_umap_plt_select1 +
+  scurve_umap_plt_select2 + scurve_umap_plt_select1 +
+  plot_layout(ncol = 2)
 ```

### 2025-08-17 paper-quollr.tex
Updated the limit
Initial work adding NLDR link plots

@@
 
-The generated p-D model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
+The generated \(p\text{-}D\) model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
 Additionally, it features a function for visualizing the \(2\text{-}D\) projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1} \caption{$2 ext{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (in purple) in different projections.
@@
 \end{verbatim}
+
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1}
 

### 2025-08-18 paper-quollr.Rmd
Added histograms for links plots

@@
 
+```{r}
+model_error <- model_error |>
+  bind_cols(scurve_umap_scaled |>
+              select(-ID))
+
+model_error <- model_error |>
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+
+# Compute density
+density_data <- density(model_error$sqrt_row_wise_total_error)
+density_df <- data.frame(x = density_data$x, y = density_data$y)
+
+# Add density values to the original dataset
+model_error <- model_error |>
+  mutate(density = approx(density_df$x, density_df$y, xout = sqrt_row_wise_total_error)$y)
+
+## Create the first row of selection
+model_error_select1 <- model_error |>
+  mutate(select_area = if_else(emb1 <= 0.7 & emb1 >= 0.4 & emb2 <= 0.25 & emb2 >= 0, "selected", "deselected"))
+
+model_error_select1_selected <- model_error_select1 |>
+  filter(select_area == "selected")
+
+model_error_select1_deselected <- model_error_select1 |>
+  filter(select_area == "deselected")
+
+error_plot_scurve_hist1 <- ggplot(model_error_select1_deselected, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density)) +
+  geom_point(alpha=0.7, colour = "#d9d9d9") +
+  geom_point(data = model_error_select1_selected, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density),
+      alpha=0.7, colour = "#756bb1", size = 3) +
+  xlab(paste("residual")) +
+  ylab("") +
+  interior_annotation("a1") +
+  theme_bw() +
+  theme(axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        aspect.ratio = 1)
+
+## Create the second row of selection
+model_error_select2 <- model_error |>
+  mutate(select_area = if_else(emb1 <= 0.55 & emb1 >= 0.25 & emb2 <= 1.25 & emb2 >= 0.75, "selected", "deselected"))
+
+model_error_select2_selected <- model_error_select2 |>
+  filter(select_area == "selected")
+
+model_error_select2_deselected <- model_error_select2 |>
+  filter(select_area == "deselected")
+
+error_plot_scurve_hist2 <- ggplot(model_error_select2_deselected, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density)) +
+  geom_point(alpha=0.7, colour = "#d9d9d9") +
+  geom_point(data = model_error_select2_selected, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density),
+      alpha=0.7, colour = "#756bb1", size = 3) +
+  xlab(paste("residual")) +
+  ylab("") +
+  interior_annotation("b1") +
+  theme_bw() +
+  theme(axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        aspect.ratio = 1)
+```
+
 ```{r, eval=knitr::is_latex_output()}
-
+error_plot_scurve_hist1 + scurve_umap_plt_select1 + scurve_umap_plt_select1 +
+  error_plot_scurve_hist2 + scurve_umap_plt_select2 + scurve_umap_plt_select1 +
+  plot_layout(ncol = 3)
 ```

### 2025-08-18 paper-quollr.R

@@ -937,8 +937,79 @@ df_exe <- comb_all_data_model_error(
 # errornldrdt_link
 
 
-## ----eval=knitr::is_latex_output()--------------------------------------------
+## -----------------------------------------------------------------------------
+model_error <- model_error |>
+  bind_cols(scurve_umap_scaled |>
+              select(-ID))
 
+model_error <- model_error |>
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+
+# Compute density
+density_data <- density(model_error$sqrt_row_wise_total_error)
+density_df <- data.frame(x = density_data$x, y = density_data$y)
+
+# Add density values to the original dataset
+model_error <- model_error |>
+  mutate(density = approx(density_df$x, density_df$y, xout = sqrt_row_wise_total_error)$y)
+
+## Create the first row of selection
+model_error_select1 <- model_error |>
+  mutate(select_area = if_else(emb1 <= 0.7 & emb1 >= 0.4 & emb2 <= 0.25 & emb2 >= 0, "selected", "deselected"))
+
+model_error_select1_selected <- model_error_select1 |>
+  filter(select_area == "selected")
+
+model_error_select1_deselected <- model_error_select1 |>
+  filter(select_area == "deselected")
+
+error_plot_scurve_hist1 <- ggplot(model_error_select1_deselected, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density)) +
+  geom_point(alpha=0.7, colour = "#d9d9d9") +
+  geom_point(data = model_error_select1_selected, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density),
+      alpha=0.7, colour = "#756bb1", size = 3) +
+  xlab(paste("residual")) +
+  ylab("") +
+  interior_annotation("a1") +
+  theme_bw() +
+  theme(axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        aspect.ratio = 1)
+
+## Create the second row of selection
+model_error_select2 <- model_error |>
+  mutate(select_area = if_else(emb1 <= 0.55 & emb1 >= 0.25 & emb2 <= 1.25 & emb2 >= 0.75, "selected", "deselected"))
+
+model_error_select2_selected <- model_error_select2 |>
+  filter(select_area == "selected")
+
+model_error_select2_deselected <- model_error_select2 |>
+  filter(select_area == "deselected")
+
[…]

### 2025-08-18 paper-quollr.Rmd
Updated the annotations

@@
   geom_point(data = scurve_umap_scaled_select1_selected, aes(x = emb1, y = emb2), alpha = 0.5, color = "#756bb1", size = 0.5) +
-  interior_annotation("a1") +
   xlim(sc_xlims) +
   ylim(sc_ylims)
+
+scurve_umap_plt_select1_lk <- scurve_umap_plt_select1 +
+  interior_annotation("a1")
+
+scurve_umap_plt_select1_lk2 <- scurve_umap_plt_select1 +
+  interior_annotation("a2")
 
@@
   geom_point(data = scurve_umap_scaled_select2_selected, aes(x = emb1, y = emb2), alpha = 0.5, color = "#756bb1", size = 0.5) +
-  interior_annotation("b1") +
   xlim(sc_xlims) +
   ylim(sc_ylims)
+
+scurve_umap_plt_select2_lk <- scurve_umap_plt_select2 +
+    interior_annotation("b1")
+
+scurve_umap_plt_select2_lk2 <- scurve_umap_plt_select2 +
+    interior_annotation("b2") 
 ```
@@
 ```{r, eval=knitr::is_latex_output(), fig.pos='H'}
-scurve_umap_plt_select1 + scurve_umap_plt_select1 +
-  scurve_umap_plt_select2 + scurve_umap_plt_select1 +
+scurve_umap_plt_select1_lk + scurve_umap_plt_select1 +
+  scurve_umap_plt_select2_lk + scurve_umap_plt_select1 +
   plot_layout(ncol = 2)
@@
 ```{r, eval=knitr::is_latex_output()}
-error_plot_scurve_hist1 + scurve_umap_plt_select1 + scurve_umap_plt_select1 +
-  error_plot_scurve_hist2 + scurve_umap_plt_select2 + scurve_umap_plt_select1 +
+error_plot_scurve_hist1 + scurve_umap_plt_select1_lk2 + scurve_umap_plt_select1 +
+  error_plot_scurve_hist2 + scurve_umap_plt_select2_lk2 + scurve_umap_plt_select1 +
   plot_layout(ncol = 3)

### 2025-08-18 paper-quollr.R

@@ -885,9 +885,14 @@ scurve_umap_plt_select1 <- ggplot(
   aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
   geom_point(data = scurve_umap_scaled_select1_selected, aes(x = emb1, y = emb2), alpha = 0.5, color = "#756bb1", size = 0.5) +
-  interior_annotation("a1") +
   xlim(sc_xlims) +
   ylim(sc_ylims)
+
+scurve_umap_plt_select1_lk <- scurve_umap_plt_select1 +
+  interior_annotation("a1")
+
+scurve_umap_plt_select1_lk2 <- scurve_umap_plt_select1 +
+  interior_annotation("a2") 
 
 ## Create the second row of selection
 scurve_umap_scaled_select2 <- scurve_umap_scaled |>
@@ -904,14 +909,19 @@ scurve_umap_plt_select2 <- ggplot(
   aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
   geom_point(data = scurve_umap_scaled_select2_selected, aes(x = emb1, y = emb2), alpha = 0.5, color = "#756bb1", size = 0.5) +
-  interior_annotation("b1") +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
+scurve_umap_plt_select2_lk <- scurve_umap_plt_select2 +
+    interior_annotation("b1") 
+
+scurve_umap_plt_select2_lk2 <- scurve_umap_plt_select2 +
+    interior_annotation("b2") 
+
 
 ## ----eval=knitr::is_latex_output(), fig.pos='H'-------------------------------
-scurve_umap_plt_select1 + scurve_umap_plt_select1 +
-  scurve_umap_plt_select2 + scurve_umap_plt_select1 +
+scurve_umap_plt_select1_lk + scurve_umap_plt_select1 +
+  scurve_umap_plt_select2_lk + scurve_umap_plt_select1 +
   plot_layout(ncol = 2)
 
 
@@ -1007,8 +1017,8 @@ error_plot_scurve_hist2 <- ggplot(model_error_select2_deselected,
 
 
 ## ----eval=knitr::is_latex_output()--------------------------------------------
-error_plot_scurve_hist1 + scurve_umap_plt_select1 + scurve_umap_plt_select1 +
[…]

### 2025-08-18 paper-quollr.Rmd
Added projection for link plot NLDR

@@
 
+proj_obj2[["cluster"]] <- factor(scurve_umap_scaled_select1$select_area,
+                                 levels=c("deselected", "selected"))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
+
+projected_df <- projected_df |>
+  dplyr::mutate(cluster = proj_obj2$cluster)
+
+scurve_proj_umap_model1_selected1 <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          colour = factor(cluster, levels = c("deselected", "selected")),
+          size = factor(cluster, levels = c("deselected", "selected"))),
+        alpha = 0.5) +
+      geom_segment(
+        data = model_df,
+        aes(
+          x = proj1_from,
+          y = proj2_from,
+          xend = proj1_to,
+          yend = proj2_to),
+        colour = "#000000",
+        linewidth = 0.8,
+        alpha = 0.4) +
+      geom_segment(
+          data=axes,
+          aes(x=x1, y=y1, xend=x2, yend=y2),
+          colour="grey70") +
+        geom_text(
+          data=axes,
+          aes(x=x2, y=y2),
+          label=rownames(axes),
+          colour="grey50",
+          size = 2.5) +
+        geom_path(
+          data=circle,
+          aes(x=c1, y=c2), colour="grey70") +
+        xlim(c(-0.35, 0.35)) +
+        ylim(c(-0.35, 0.35)) +
+  interior_annotation(label = "a2", cex = 1) +
+  scale_color_manual(values = c(clr_choice, '#756bb1')) +
+  scale_size_manual(values = c('deselected' = 1.0, 'selected' = 1.5)) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
+
 ## Create the second row of selection
@@
     interior_annotation("b2")
+
+proj_obj2[["cluster"]] <- factor(scurve_umap_scaled_select2$select_area,
+                                 levels=c("deselected", "selected"))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
+
+projected_df <- projected_df |>
+  dplyr::mutate(cluster = proj_obj2$cluster)
+
+scurve_proj_umap_model1_selected2 <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          colour = factor(cluster, levels = c("deselected", "selected")),
+          size = factor(cluster, levels = c("deselected", "selected"))),
+        alpha = 0.5) +
+      geom_segment(
+        data = model_df,
+        aes(
+          x = proj1_from,
+          y = proj2_from,
+          xend = proj1_to,
+          yend = proj2_to),
+        colour = "#000000",
+        linewidth = 0.8,
+        alpha = 0.4) +
+      geom_segment(
+          data=axes,
+          aes(x=x1, y=y1, xend=x2, yend=y2),
+          colour="grey70") +
+        geom_text(
+          data=axes,
+          aes(x=x2, y=y2),
+          label=rownames(axes),
+          colour="grey50",
+          size = 2.5) +
+        geom_path(
+          data=circle,
+          aes(x=c1, y=c2), colour="grey70") +
+        xlim(c(-0.35, 0.35)) +
+        ylim(c(-0.35, 0.35)) +
+  interior_annotation(label = "b2", cex = 1) +
+  scale_color_manual(values = c(clr_choice, '#756bb1')) +
+  scale_size_manual(values = c('deselected' = 1.0, 'selected' = 1.5)) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
 ```
@@
 ```{r, eval=knitr::is_latex_output(), fig.pos='H'}
-scurve_umap_plt_select1_lk + scurve_umap_plt_select1 +
-  scurve_umap_plt_select2_lk + scurve_umap_plt_select1 +
+scurve_umap_plt_select1_lk + scurve_proj_umap_model1_selected1 +
+  scurve_umap_plt_select2_lk + scurve_proj_umap_model1_selected2 +
   plot_layout(ncol = 2)

### 2025-08-18 paper-quollr.R

@@ -894,6 +894,57 @@ scurve_umap_plt_select1_lk <- scurve_umap_plt_select1 +
 scurve_umap_plt_select1_lk2 <- scurve_umap_plt_select1 +
   interior_annotation("a2") 
 
+proj_obj2[["cluster"]] <- factor(scurve_umap_scaled_select1$select_area,
+                                 levels=c("deselected", "selected"))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
+
+projected_df <- projected_df |>
+  dplyr::mutate(cluster = proj_obj2$cluster)
+
+scurve_proj_umap_model1_selected1 <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          colour = factor(cluster, levels = c("deselected", "selected")),
+          size = factor(cluster, levels = c("deselected", "selected"))),
+        alpha = 0.5) +
+      geom_segment(
+        data = model_df,
+        aes(
+          x = proj1_from,
+          y = proj2_from,
+          xend = proj1_to,
+          yend = proj2_to),
+        colour = "#000000",
+        linewidth = 0.8,
+        alpha = 0.4) +
+      geom_segment(
+          data=axes,
+          aes(x=x1, y=y1, xend=x2, yend=y2),
+          colour="grey70") +
+        geom_text(
+          data=axes,
+          aes(x=x2, y=y2),
+          label=rownames(axes),
+          colour="grey50",
+          size = 2.5) +
+        geom_path(
+          data=circle,
+          aes(x=c1, y=c2), colour="grey70") +
+        xlim(c(-0.35, 0.35)) +
+        ylim(c(-0.35, 0.35)) +
+  interior_annotation(label = "a2", cex = 1) +
+  scale_color_manual(values = c(clr_choice, '#756bb1')) +
+  scale_size_manual(values = c('deselected' = 1.0, 'selected' = 1.5)) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
+
 ## Create the second row of selection
 scurve_umap_scaled_select2 <- scurve_umap_scaled |>
   mutate(select_area = if_else(emb1 <= 0.55 & emb1 >= 0.25 & emb2 <= 1.25 & emb2 >= 0.75, "selected", "deselected"))
[…]

### 2025-08-18 paper-quollr.Rmd
Added the projections for linked plots

@@
 
+scurve_proj_umap_model1_selected1_dp <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          colour = factor(cluster, levels = c("deselected", "selected")),
+          size = factor(cluster, levels = c("deselected", "selected"))),
+        alpha = 0.5) +
+      geom_segment(
+        data = model_df,
+        aes(
+          x = proj1_from,
+          y = proj2_from,
+          xend = proj1_to,
+          yend = proj2_to),
+        colour = "#000000",
+        linewidth = 0.8,
+        alpha = 0.4) +
+      geom_segment(
+          data=axes,
+          aes(x=x1, y=y1, xend=x2, yend=y2),
+          colour="grey70") +
+        geom_text(
+          data=axes,
+          aes(x=x2, y=y2),
+          label=rownames(axes),
+          colour="grey50",
+          size = 2) +
+        geom_path(
+          data=circle,
+          aes(x=c1, y=c2), colour="grey70") +
+        xlim(c(-0.35, 0.35)) +
+        ylim(c(-0.35, 0.35)) +
+  interior_annotation(label = "a3", cex = 1) +
+  scale_color_manual(values = c(clr_choice, '#756bb1')) +
+  scale_size_manual(values = c('deselected' = 1.0, 'selected' = 1.5)) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
+
 ## Create the second row of selection
@@
         legend.position = "none")
+
+scurve_proj_umap_model1_selected2_dp <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          colour = factor(cluster, levels = c("deselected", "selected")),
+          size = factor(cluster, levels = c("deselected", "selected"))),
+        alpha = 0.5) +
+      geom_segment(
+        data = model_df,
+        aes(
+          x = proj1_from,
+          y = proj2_from,
+          xend = proj1_to,
+          yend = proj2_to),
+        colour = "#000000",
+        linewidth = 0.8,
+        alpha = 0.4) +
+      geom_segment(
+          data=axes,
+          aes(x=x1, y=y1, xend=x2, yend=y2),
+          colour="grey70") +
+        geom_text(
+          data=axes,
+          aes(x=x2, y=y2),
+          label=rownames(axes),
+          colour="grey50",
+          size = 2) +
+        geom_path(
+          data=circle,
+          aes(x=c1, y=c2), colour="grey70") +
+        xlim(c(-0.35, 0.35)) +
+        ylim(c(-0.35, 0.35)) +
+  interior_annotation(label = "b3", cex = 1) +
+  scale_color_manual(values = c(clr_choice, '#756bb1')) +
+  scale_size_manual(values = c('deselected' = 1.0, 'selected' = 1.5)) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
 ```
@@
 ```{r, eval=knitr::is_latex_output()}
-error_plot_scurve_hist1 + scurve_umap_plt_select1_lk2 + scurve_umap_plt_select1 +
-  error_plot_scurve_hist2 + scurve_umap_plt_select2_lk2 + scurve_umap_plt_select1 +
+error_plot_scurve_hist1 + scurve_umap_plt_select1_lk2 + scurve_proj_umap_model1_selected1_dp +
+  error_plot_scurve_hist2 + scurve_umap_plt_select2_lk2 + scurve_proj_umap_model1_selected2_dp +
   plot_layout(ncol = 3)

### 2025-08-18 paper-quollr.Rmd
Added Figure captions

@@
 
-```{r, eval=knitr::is_latex_output(), fig.pos='H'}
+```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
+Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
+The purple points indicate the selected subsets, which differ between rows.
+In (a1), the lower bridge of the `Scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
+In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection.
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP."}
+
 scurve_umap_plt_select1_lk + scurve_proj_umap_model1_selected1 +
@@
 
-```{r, eval=knitr::is_latex_output()}
+```{r, eval=knitr::is_latex_output(), fig.pos="H", fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `Scurve` model.
+Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\\text{-}D$ model with data (a3, b3).
+The purple points highlight selected subsets of the data, which differ across rows.
+In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `Scurve` in the high-dimensional projection (a3).
+In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure."}
+
 error_plot_scurve_hist1 + scurve_umap_plt_select1_lk2 + scurve_proj_umap_model1_selected1_dp +

### 2025-08-18 paper-quollr.tex
Added histograms for links plots
Added Figure captions

@@
 
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1}
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
+The purple points indicate the selected subsets, which differ between rows.
+In (a1), the lower bridge of the `Scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
+In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-30}
+\end{figure}
 
@@
 \end{verbatim}
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-34-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
+The purple points highlight selected subsets of the data, which differ across rows.
+In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `Scurve` in the high-dimensional projection (a3).
+In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-34}
+\end{figure}
 

### 2025-08-18 paper-quollr.R

@@ -1101,7 +1101,8 @@ scurve_proj_umap_model1_selected2_dp <- ggplot() +
         legend.position = "none")
 
 
-## ----eval=knitr::is_latex_output(), fig.pos='H'-------------------------------
+## ----eval=knitr::is_latex_output(), fig.pos='H', fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$. Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2). The purple points indicate the selected subsets, which differ between rows. In (a1), the lower bridge of the `Scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure. In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection. While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP."----
+
 scurve_umap_plt_select1_lk + scurve_proj_umap_model1_selected1 +
   scurve_umap_plt_select2_lk + scurve_proj_umap_model1_selected2 +
   plot_layout(ncol = 2)
@@ -1198,7 +1199,8 @@ error_plot_scurve_hist2 <- ggplot(model_error_select2_deselected,
         aspect.ratio = 1)
 
 
-## ----eval=knitr::is_latex_output()--------------------------------------------
[…]

### 2025-08-18 paper-quollr.Rmd
Added Figure caption in HTML version for linked plots

@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap=""}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+
 errornldrdt_link <- show_error_link_plots(
@@
 
-```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\ref{fig:limb-rmse}b) and the $2-\\text{D}$ layout selected (Figure \\ref{fig:limb-rmse}f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\ref{fig:limb-rmse}b) and the $2-\\text{D}$ layout selected (Figure \\ref{fig:limb-rmse}f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
@@
 
-```{r model-limb-html, eval=knitr::is_html_output(), fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+```{r model-limb-html, eval=knitr::is_html_output(), fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$."}

### 2025-08-18 paper-quollr.tex
Added Figure caption in HTML version for linked plots

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \ref{fig:limb-rmse}b) and the $2-\text{D}$ layout selected (Figure \ref{fig:limb-rmse}f) by RMSE plot (Figure \ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different (hyper)parameters.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \ref{fig:limb-rmse}b) and the $2-\text{D}$ layout selected (Figure \ref{fig:limb-rmse}f) by RMSE plot (Figure \ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-08-18 paper-quollr.Rmd
Updated the language

@@
 
-### Visualisations
-
-The package offers several $2\text{-}D$ visualisations, including:
+### Visualizations
+
+The package offers several $2\text{-}D$ visualizations, including:
 
@@
 
-Each visualisation can be generated using its respective function, as described in this section.
+Each visualization can be generated using its respective function, as described in this section.
 
@@
 
-#### $p\text{-}D$ model visualisation
-
-To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualisation using the `show_langevitour()` function.
+#### $p\text{-}D$ model visualization
+
+To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualization using the `show_langevitour()` function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.
@@
 
-The `show_langevitour()` function then renders the visualisation using the `langevitour` interface, displaying both types of points in a dynamic tour.
+The `show_langevitour()` function then renders the visualization using the `langevitour` interface, displaying both types of points in a dynamic tour.
 The `edge_data` input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model’s structure.
 
-In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped back to high-dimensional space.
 
@@
 To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
-This visualisation combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
 Both views are interactively linked, enabling users to explore.
 
-Before visualisation, the input data must be prepared using the `comb_all_data_model()` function.
+Before visualization, the input data must be prepared using the `comb_all_data_model()` function.
 This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2\text{-}D$ bin layout (`model_2d`):
@@
 
-Before visualisation, the input data must be prepared using the `comb_all_data_model_error()` function.
+Before visualization, the input data must be prepared using the `comb_all_data_model_error()` function.
 The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2\text{-}D$ model data (`model_2d`), and model error (`error_data`).
@@
 It enables the computation of RMSE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
-These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualisations using the `langevitour` package, allowing users to inspect where the model fits well and where it does not.
+These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualizations using the `langevitour` package, allowing users to inspect where the model fits well and where it does not.
 

### 2025-08-18 paper-quollr.tex
Updated the language

@@
 
-\subsubsection{Visualisations}\label{visualisations}
+\subsubsection{Visualizations}\label{visualizations}
 
-The package offers several \(2\text{-}D\) visualisations, including:
+The package offers several \(2\text{-}D\) visualizations, including:
 
@@
 
-\paragraph{\texorpdfstring{\(p\text{-}D\) model visualisation}{p\textbackslash text\{-\}D model visualisation}}\label{ptext-d-model-visualisation}
+\paragraph{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
 
-To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualisation using the \texttt{show\_langevitour()} function.
+To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualization using the \texttt{show\_langevitour()} function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.
@@
 
-The \texttt{show\_langevitour()} function then renders the visualisation using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour.
+The \texttt{show\_langevitour()} function then renders the visualization using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour.
 The \texttt{edge\_data} input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model's structure.
 
-In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2\text{-}D\) wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2\text{-}D\) wireframe structure mapped back to high-dimensional space.
 
@@
 To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
-This visualisation combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
 Both views are interactively linked, enabling users to explore.
 
-Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
+Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
 This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2\text{-}D\) bin layout (\texttt{model\_2d}):
@@
 
-Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
+Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
 The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2\text{-}D\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
@@
 It enables the computation of RMSE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
-These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualisations using the \texttt{langevitour} package, allowing users to inspect where the model fits well and where it does not.
+These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualizations using the \texttt{langevitour} package, allowing users to inspect where the model fits well and where it does not.
 

### 2025-08-18 paper-quollr.Rmd
Added Figure title for first figure

@@
 
-```{r overview, eval=knitr::is_latex_output(), fig.cap="algorithm", fig.pos='H'}
+```{r overview, eval=knitr::is_latex_output(), fig.cap="Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space.
+The left panel shows the NLDR layout with a triangular mesh overlay, forming the wireframe structure.
+This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
+Panels (a1–a4) display different $2\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H'}
 hex_grid_scurve + wrap_plots(

### 2025-08-18 paper-quollr.tex
Added Figure title for first figure

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/overview-1} \caption{algorithm}\label{fig:overview}
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/overview-1} \caption{Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space.
+The left panel shows the NLDR layout with a triangular mesh overlay, forming the wireframe structure.
+This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
+Panels (a1–a4) display different $2 ext{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.}\label{fig:overview}
 \end{figure}

### 2025-08-18 paper-quollr.Rmd
Updated Figure 2

@@
 
-sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.25, 1.25)
+sc_xlims <- c(-0.25, 1.35)
+sc_ylims <- c(-0.25, 1.35)
 
@@
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
-  interior_annotation("a") +
+  interior_annotation("a", cex = 3) +
   xlim(sc_xlims) +
@@
              alpha = 0.2, size = 0.5, color = clr_choice) + 
-  interior_annotation("b") +
+  interior_annotation("b", cex = 3) +
   xlim(sc_xlims) +
@@
              size = 1, color = "#FF7755") +
-  interior_annotation("c") +
+  interior_annotation("c", cex = 3) +
   xlim(sc_xlims) +
@@
              size = 1, color = "#FF7755") +
-  interior_annotation("d") +
+  interior_annotation("d", cex = 3) +
   xlim(sc_xlims) +
@@
 
+scurve_proj_umap_model1_dp <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(0.2, 0, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 7,
+  is_category = FALSE) +
+  interior_annotation(label = "e", cex = 3)
+
 ## Second projection
@@
   hex_centroids_scurve + wireframe_scurve + 
-  scurve_proj_umap_model1 +
+  scurve_proj_umap_model1_dp +
   plot_layout(ncol = 5)

### 2025-08-18 paper-quollr.R

@@ -113,14 +113,14 @@ hex_grid_with_counts_scurve <- left_join(
 hex_grid_nonempty_scurve <- hex_grid_scurve |>
   filter(h %in% df_bin_centroids_scurve$h)
 
-sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.25, 1.25)
+sc_xlims <- c(-0.25, 1.35)
+sc_ylims <- c(-0.25, 1.35)
 
 scurve_umap_plt <- ggplot(
   scurve_umap_scaled, 
   aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
-  interior_annotation("a") +
+  interior_annotation("a", cex = 3) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -146,7 +146,7 @@ hex_grid_poly_scurve <- ggplot(
   geom_point(data = scurve_umap_scaled, 
              aes(x = emb1, y = emb2), 
              alpha = 0.2, size = 0.5, color = clr_choice) + 
-  interior_annotation("b") +
+  interior_annotation("b", cex = 3) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -172,7 +172,7 @@ hex_centroids_scurve <- ggplot(
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
              size = 1, color = "#FF7755") +
-  interior_annotation("c") +
+  interior_annotation("c", cex = 3) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -200,7 +200,7 @@ wireframe_scurve <- ggplot() +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
              size = 1, color = "#FF7755") +
-  interior_annotation("d") +
+  interior_annotation("d", cex = 3) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -348,7 +348,15 @@ scurve_proj_umap_model1 <- plot_proj(
   plot_limits = c(-0.35, 0.35), 
   axis_text_size = 3,
   is_category = FALSE) +
-  interior_annotation(label = "a1", cex = 1) 
+  interior_annotation(label = "a1", cex = 1)
+
+scurve_proj_umap_model1_dp <- plot_proj(
+  proj_obj = proj_obj1,
[…]

### 2025-08-18 paper-quollr.Rmd
Updating the figure widths and heights

@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.pos='H', fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.pos='H', fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
@@
 
-```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
+```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
@@
 
-```{r, eval=knitr::is_latex_output(), fig.pos="H", fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `Scurve` model.
+```{r, eval=knitr::is_latex_output(), fig.pos="H", fig.width=15, fig.height=10, fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\\text{-}D$ model with data (a3, b3).

### 2025-08-18 paper-quollr.Rmd
Updated link plots

@@
 scurve_umap_plt_select1_lk <- scurve_umap_plt_select1 +
-  interior_annotation("a1")
+  interior_annotation("a1", cex = 3)
 
 scurve_umap_plt_select1_lk2 <- scurve_umap_plt_select1 +
-  interior_annotation("a2")
+  interior_annotation("a2", cex = 3)
 
@@
           colour="grey50",
-          size = 2.5) +
+          size = 5) +
         geom_path(
@@
         ylim(c(-0.35, 0.35)) +
-  interior_annotation(label = "a2", cex = 1) +
+  interior_annotation(label = "a2", cex = 3) +
   scale_color_manual(values = c(clr_choice, '#756bb1')) +
@@
           colour="grey50",
-          size = 2) +
+          size = 5) +
         geom_path(
@@
         ylim(c(-0.35, 0.35)) +
-  interior_annotation(label = "a3", cex = 1) +
+  interior_annotation(label = "a3", cex = 3) +
   scale_color_manual(values = c(clr_choice, '#756bb1')) +
@@
 scurve_umap_plt_select2_lk <- scurve_umap_plt_select2 +
-    interior_annotation("b1")
+    interior_annotation("b1", cex = 3)
 
 scurve_umap_plt_select2_lk2 <- scurve_umap_plt_select2 +
-    interior_annotation("b2")
+    interior_annotation("b2", cex = 3)
 
@@
         ylim(c(-0.35, 0.35)) +
-  interior_annotation(label = "b2", cex = 1) +
+  interior_annotation(label = "b2", cex = 3) +
   scale_color_manual(values = c(clr_choice, '#756bb1')) +
@@
         ylim(c(-0.35, 0.35)) +
-  interior_annotation(label = "b3", cex = 1) +
+  interior_annotation(label = "b3", cex = 3) +
   scale_color_manual(values = c(clr_choice, '#756bb1')) +
@@
   ylab("") +
-  interior_annotation("a1") +
+  interior_annotation("a1", cex = 3) +
   theme_bw() +
   theme(axis.ticks.y = element_blank(),
+        axis.title.x = element_text(size = 18),
+        axis.text.x = element_text(size = 15),
         axis.text.y = element_blank(),
@@
   ylab("") +
-  interior_annotation("b1") +
+  interior_annotation("b1", cex = 3) +
   theme_bw() +
   theme(axis.ticks.y = element_blank(),
+        axis.title.x = element_text(size = 18),
+        axis.text.x = element_text(size = 15),
         axis.text.y = element_blank(),

### 2025-08-18 paper-quollr.Rmd
Updated axis text size

@@
   plot_limits = c(-0.35, 0.35), 
-  axis_text_size = 3,
+  axis_text_size = 2,
   is_category = FALSE) +

### 2025-08-26 paper-quollr.Rmd
Paper changes up to the linked plots section

@@
 
+::: {.hidden}
+$$
+\newcommand{\pD}{$p\text{-}D$}
+\newcommand{\twoD}{$2\text{-}D$}
+$$
+:::
+
 ```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE}
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and hyper-parameter choices.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], can create hugely different representations depending on the selected method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 
-The following demonstration of the package's functionality assumes `quollr` has been loaded. We also want to load the built-in data sets `scurve` and `scurve_umap`.
-
-`scurve` is a $7\text{-}D$ simulated dataset.
-It is constructed by simulating $1000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
+The following demonstration of the package's functionality assumes `quollr` has been loaded. The built-in data sets `scurve` and `scurve_umap` are also used throughout.
+
+`scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $1000$ observations from
+
+$$
+\begin{aligned}
+\theta &\sim U(-3\pi/2, 3\pi/2), \\
+X_1 &= \sin(\theta), \\
+X_2 &\sim U(0, 2), \text{adding thickness to the S}, \\
+X_3 &= \text{sign}(\theta) \times (\cos(\theta) - 1).
+\end{aligned}
+$$
+
 The remaining variables $X_4$, $X_5$, $X_6$, $X_7$ are all uniform error, with small variance.
-`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $46$ and `min_dist` is $0.9$.
-Each data set contains a unique ID column that maps `scurve` and `scurve_umap`.
+`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors`$=46$ and `min_dist`$=0.9$.
+Each data set contains a unique ID column that maps the $7\text{-}D$ point in `scurve` to the corresponding $2\text{-}D$ `scurve_umap`.
 
@@
 
-Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) Remove low-density hexagons.
+Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
 
@@
 
-The algorithm starts by scaling the NLDR data to a standard range using the `gen_scaled_data()` function.
-This function standardizes the data so that the first embedding ranges from $0$ to $1$, while the second embedding scales from $0$ to the maximum value of the second embedding.
-The output includes the scaled NLDR data along with the original limits of the embeddings.
+The algorithm starts by scaling the NLDR data to to the range $[0, 1] \times [0, x_2]$, where $x_2$ is the maximum value of the second embedding, using the `gen_scaled_data()` function.
+The output includes the scaled NLDR data (`scaled_nldr`) along with the original limits of the embeddings (`lim1`, `lim2`).
 
@@
 The function `calc_bins_y()` is used for this purpose.
-This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (`b1`), and the buffer amount as a proportion (`q`).
+This function accepts (1) an object containing three elements: the first is the scaled NLDR data with two columns, and the second and third are the limits of the original NLDR data; (2) the number of bins along the x-axis (`b1`), and (3) the buffer amount as a proportion (`q`).
 
@@
 
-Points are allocated to bins based on the nearest centroid.
+Points are allocated to bins based on the nearest centroid of the hexagonal bins.
 The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its components can be run separately for added flexibility.
@@
 Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
-Vertical spacing (`a2`) is derived by $\sqrt{3}/2 \times a_1$.
+Vertical spacing (`a2`) is given by the height of a hexagon, $\sqrt{3}/2 \times a_1$.
 
@@
 With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid.
-These offsets form two vectors (`x_add_factor` and `y_add_factor`) corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
+These offsets form two vectors corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
 
@@
 
-The `compute_std_counts()` function calculates both the raw and standardized counts for each hexagon.
-
-The function begins by grouping the data by `h` and counting the number of NLDR points falling within each bin.
+The `compute_std_counts()` function calculates both the raw and standardized counts of points inside each hexagon.
+
+The function begins by grouping the data by hexagon ID (`h`) and counting the number of NLDR points falling within each bin.
 These raw counts are stored as `n_h`.
@@
 
-The `find_low_dens_hex()` function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-The `find_low_dens_hex()` function relies on a helper, `compute_mean_density_hex()`, which iterates over all hexagons and computes the average density across neighbors based on their `h` and a defined number of bins along the x-axis (`b1`).
+The `find_low_dens_hex()` function relies on a helper, `compute_mean_density_hex()`, which iterates over all hexagons and computes the average density across neighbors based on their hexagon ID (`h`) and a defined number of bins along the x-axis (`b1`).
 The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.
@@
 
-The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each hexagonal bin to represent bin centroids.
-This transformation is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).
+The final step involves lifting the fitted \twoD model into \pD.
+This is done by modelling a point in \pD as the  \pD mean of data points in the \twoD centroid.
+This is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).
 
@@
 
-The `predict_emb()` function is used to predict $2\text{-}D$ embedding for a new $p\text{-}D$ data point using the fitted model.
+The `predict_emb()` function is used to predict a point in a $2\text{-}D$ embedding for a new $p\text{-}D$ data point using the fitted model.
 This function is useful to predict $2\text{-}D$ embedding irrespective of the NLDR technique.
 
-In the prediction process, first, the nearest $p\text{-}D$ model point is identified for a given new $p\text{-}D$ data point by computing $p\text{-}D$ Euclidean distance.
+In the prediction process, first, the nearest $p\text{-}D$ model point is identified for the new $p\text{-}D$ data point by computing $p\text{-}D$ Euclidean distance.
 Then, the corresponding $2\text{-}D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
@@
 
-As a Goodness of fit statistics for the model, `glance()` is used to compute residuals and RMSE.
-These metrics are used to assess how well the fitted model will capture the underlying structure of the $p\text{-}D$ data.
+Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model. These metrics can be output using the  `glance()` function.
 
@@
 
-The function starts by renaming columns in the `model_highd` data frame to avoid conflicts.
-It then predicts the high-dimensional bin assignments for each point using the `predict_emb()` function, mapping each observation to its nearest $2\text{-}D$ bin (`pred_hb_id`).
-This prediction is joined with the mean high-dimensional coordinates of each bin from the model and with the original high-dimensional data.
+The function starts with the same process as is used in the `glance()` function to produce a predicted point in \pD for each point in the high-dimensional dataset.
 
@@
 
-To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualization using the `show_langevitour()` function.
+To visualise how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \pD using the `show_langevitour()` function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.

### 2025-08-26 paper-quollr.Rmd
Added first version rendered

@@
   \usepackage{float}
+  \newcommand{\pD}{$p\text{-}D$}
+  \newcommand{\twoD}{$2\text{-}D$}
 output: 
@@
 ```
-
-::: {.hidden}
-$$
-\newcommand{\pD}{$p\text{-}D$}
-\newcommand{\twoD}{$2\text{-}D$}
-$$
-:::
 

### 2025-08-26 paper-quollr.tex
Added first version rendered

@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and hyper-parameter choices.
+Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, can create hugely different representations depending on the selected method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 
-The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. We also want to load the built-in data sets \texttt{scurve} and \texttt{scurve\_umap}.
+The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. The built-in data sets \texttt{scurve} and \texttt{scurve\_umap} are also used throughout.
 
-\texttt{scurve} is a \(7\text{-}D\) simulated dataset.
-It is constructed by simulating \(1000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\).
+\texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(1000\) observations from
+
+\[
+\begin{aligned}
+\theta &\sim U(-3\pi/2, 3\pi/2), \\
+X_1 &= \sin(\theta), \\
+X_2 &\sim U(0, 2), \text{adding thickness to the S}, \\
+X_3 &= \text{sign}(\theta) \times (\cos(\theta) - 1).
+\end{aligned}
+\]
+
 The remaining variables \(X_4\), \(X_5\), \(X_6\), \(X_7\) are all uniform error, with small variance.
-\texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors} is \(46\) and \texttt{min\_dist} is \(0.9\).
-Each data set contains a unique ID column that maps \texttt{scurve} and \texttt{scurve\_umap}.
+\texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors}\(=46\) and \texttt{min\_dist}\(=0.9\).
+Each data set contains a unique ID column that maps the \(7\text{-}D\) point in \texttt{scurve} to the corresponding \(2\text{-}D\) \texttt{scurve\_umap}.
 
@@
 
-Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) Remove low-density hexagons.
+Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
 
@@
 
-The algorithm starts by scaling the NLDR data to a standard range using the \texttt{gen\_scaled\_data()} function.
-This function standardizes the data so that the first embedding ranges from \(0\) to \(1\), while the second embedding scales from \(0\) to the maximum value of the second embedding.
-The output includes the scaled NLDR data along with the original limits of the embeddings.
+The algorithm starts by scaling the NLDR data to to the range \([0, 1] \times [0, x_2]\), where \(x_2\) is the maximum value of the second embedding, using the \texttt{gen\_scaled\_data()} function.
+The output includes the scaled NLDR data (\texttt{scaled\_nldr}) along with the original limits of the embeddings (\texttt{lim1}, \texttt{lim2}).
 
@@
 The function \texttt{calc\_bins\_y()} is used for this purpose.
-This function accepts an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data, the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion (\texttt{q}).
+This function accepts (1) an object containing three elements: the first is the scaled NLDR data with two columns, and the second and third are the limits of the original NLDR data; (2) the number of bins along the x-axis (\texttt{b1}), and (3) the buffer amount as a proportion (\texttt{q}).
 
@@
 
-Points are allocated to bins based on the nearest centroid.
+Points are allocated to bins based on the nearest centroid of the hexagonal bins.
 The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its components can be run separately for added flexibility.
@@
 Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
-Vertical spacing (\texttt{a2}) is derived by \(\sqrt{3}/2 \times a_1\).
+Vertical spacing (\texttt{a2}) is given by the height of a hexagon, \(\sqrt{3}/2 \times a_1\).
 
@@
 With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid.
-These offsets form two vectors (\texttt{x\_add\_factor} and \texttt{y\_add\_factor}) corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
+These offsets form two vectors corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
 
@@
 
-The \texttt{compute\_std\_counts()} function calculates both the raw and standardized counts for each hexagon.
+The \texttt{compute\_std\_counts()} function calculates both the raw and standardized counts of points inside each hexagon.
 
-The function begins by grouping the data by \texttt{h} and counting the number of NLDR points falling within each bin.
+The function begins by grouping the data by hexagon ID (\texttt{h}) and counting the number of NLDR points falling within each bin.
 These raw counts are stored as \texttt{n\_h}.
@@
 
-The \texttt{find\_low\_dens\_hex()} function identifies hexagons with low point densities, considering the densities of their neighboring bins as well.
-The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their \texttt{h} and a defined number of bins along the x-axis (\texttt{b1}).
+The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their hexagon ID (\texttt{h}) and a defined number of bins along the x-axis (\texttt{b1}).
 The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.
@@
 
-The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each hexagonal bin to represent bin centroids.
-This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
+The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$.
+This is done by modelling a point in $p\text{-}D$ as the $p\text{-}D$ mean of data points in the $2\text{-}D$ centroid.
+This is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
 
@@
 
-The \texttt{predict\_emb()} function is used to predict \(2\text{-}D\) embedding for a new \(p\text{-}D\) data point using the fitted model.
+The \texttt{predict\_emb()} function is used to predict a point in a \(2\text{-}D\) embedding for a new \(p\text{-}D\) data point using the fitted model.
 This function is useful to predict \(2\text{-}D\) embedding irrespective of the NLDR technique.
 
-In the prediction process, first, the nearest \(p\text{-}D\) model point is identified for a given new \(p\text{-}D\) data point by computing \(p\text{-}D\) Euclidean distance.
+In the prediction process, first, the nearest \(p\text{-}D\) model point is identified for the new \(p\text{-}D\) data point by computing \(p\text{-}D\) Euclidean distance.
 Then, the corresponding \(2\text{-}D\) bin centroid mapping for the identified \(p\text{-}D\) model point is determined.
@@
 
-As a Goodness of fit statistics for the model, \texttt{glance()} is used to compute residuals and RMSE.
-These metrics are used to assess how well the fitted model will capture the underlying structure of the \(p\text{-}D\) data.
+Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model. These metrics can be output using the \texttt{glance()} function.
 
@@
 
-The function starts by renaming columns in the \texttt{model\_highd} data frame to avoid conflicts.
-It then predicts the high-dimensional bin assignments for each point using the \texttt{predict\_emb()} function, mapping each observation to its nearest \(2\text{-}D\) bin (\texttt{pred\_hb\_id}).
-This prediction is joined with the mean high-dimensional coordinates of each bin from the model and with the original high-dimensional data.
+The function starts with the same process as is used in the \texttt{glance()} function to produce a predicted point in $p\text{-}D$ for each point in the high-dimensional dataset.
 
@@
 
-To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualization using the \texttt{show\_langevitour()} function.
+To visualise how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the \texttt{show\_langevitour()} function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.

### 2025-08-31 paper-quollr.Rmd
Added citation for tripack

@@
 
-The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `tripack::tri.mesh()` function.
+The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `tripack::tri.mesh()` function (@albrecht2025).
 This triangulation forms the structural basis for identifying adjacent bins.

### 2025-08-31 paper-quollr.Rmd
Adde prediction with other methods
Added the sentence regarding small variance

@@
 
-The remaining variables $X_4$, $X_5$, $X_6$, $X_7$ are all uniform error, with small variance.
+The remaining variables $X_4$, $X_5$, $X_6$, and $X_7$ are included as low-magnitude noise dimensions, generated from uniform distributions with narrow ranges around zero: $X_4, X_5 \sim U(-0.02, 0.02)$, $X_6 \sim U(-0.1, 0.1)$, and $X_7 \sim U(-0.01, 0.01)$.
+These small variances ensure that the added dimensions contribute noise without obscuring the main geometric structure.
 `scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors`$=46$ and `min_dist`$=0.9$.
@@
 
+It is worth noting that while `predict_emb()` provides a general approach that works across methods, some NLDR techniques have their own built-in prediction mechanisms.
+For example, UMAP (@tomasz2023) supports direct prediction of embeddings for new data once a model is fitted.
+
 ### Compute residuals and Root Mean Square Error (RMSE)

### 2025-09-01 paper-quollr.Rmd
Updated threshold
Updated scaling data section

@@
 
-The algorithm starts by scaling the NLDR data to to the range $[0, 1] \times [0, x_2]$, where $x_2$ is the maximum value of the second embedding, using the `gen_scaled_data()` function.
+The algorithm starts by scaling the NLDR data to to the range $[0, 1] \times [0, y_{2,max}]$, where $y_{2,max} = r_2/r_1$ is the ratio of ranges of embedding components.
 The output includes the scaled NLDR data (`scaled_nldr`) along with the original limits of the embeddings (`lim1`, `lim2`).
@@
 
-For simplicity, we remove low-density hexagons using a threshold of $10$.
+For simplicity, we remove low-density hexagons using a threshold of $1$.
 

### 2025-09-01 paper-quollr.Rmd
Added the buffer amount definition and updated hex_binning runing individually

@@
 This function accepts (1) an object containing three elements: the first is the scaled NLDR data with two columns, and the second and third are the limits of the original NLDR data; (2) the number of bins along the x-axis (`b1`), and (3) the buffer amount as a proportion (`q`).
+The buffer ensures that the grid fully covers the data space by extending one hexagon width ($a_1$) and height ($a_2$) beyond the observed data in all directions.
+By default, $q = 0.1$, but it must be set to a value smaller than the minimum data value to avoid exceeding the data range.
 
@@
 Points are allocated to bins based on the nearest centroid of the hexagonal bins.
-The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its components can be run separately for added flexibility.
+The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its individual components can be run separately for added flexibility.
+While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the `hex_binning()` function automates this entire workflow.
 The parameters used within `hex_binning()` include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (`nldr_obj`), the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
@@
 
-If the hexagonal binning process is run separately, it involves several steps: (i) generating all possible centroids in a hexagonal grid, (ii) creating the coordinates of the hexagons, (iii) assigning data points to their respective hexagons, (iv) computing the standardized number of points within each hexagon, and (v) mapping the points to their corresponding hexagonal bins.
-
 ##### Generating all possible centroids in a hexagonal grid

### 2025-09-01 paper-quollr.Rmd
Updated polygon_ID description

@@
 
-For each centroid, six vertices are computed and assigned a polygon ID (`h`) corresponding to the centroid’s `h`.
-These points are aggregated into a tibble that includes the polygon ID (`h`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.
+For each centroid, six vertices are computed and assigned a polygon ID as the centroid.
+These vertices are then combined into a tibble that records the polygon ID (`h`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.
 

### 2025-09-01 paper-quollr.Rmd
Updated w_h definition
Updated elements in an object (`nldr_scaled_obj`)

@@
 This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`benchmark_highdens`).
-The function returns an object that includes the scaled NLDR object (`nldr_obj`), the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
+The function returns an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
 
@@
 The function `calc_bins_y()` is used for this purpose.
-This function accepts (1) an object containing three elements: the first is the scaled NLDR data with two columns, and the second and third are the limits of the original NLDR data; (2) the number of bins along the x-axis (`b1`), and (3) the buffer amount as a proportion (`q`).
+This function accepts (1) an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); (2) the number of bins along the x-axis (`b1`), and (3) the buffer amount as a proportion (`q`).
 The buffer ensures that the grid fully covers the data space by extending one hexagon width ($a_1$) and height ($a_2$) beyond the observed data in all directions.
@@
 While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the `hex_binning()` function automates this entire workflow.
-The parameters used within `hex_binning()` include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (`nldr_obj`), the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
+The parameters used within `hex_binning()` include an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
 The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagon IDs (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`tot_bins`), the number of non-empty bins (`non_bins`), and the points within each hexagon (`pts_bins`).
@@
 To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin’s count by the maximum count across all bins.
-This yields a normalized metric, `w_h`, ranging from 0 to 1.
+This yields a standardized bin counts, `w_h`, ranging from 0 to 1.
 

### 2025-09-08 paper-quollr.Rmd
Updated wordings

@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different $2\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H'}

### 2025-09-17 paper-quollr.Rmd
Updated wordings with hd_thresh

@@
   q = 0.1, 
-  benchmark_highdens = 1)
+  hd_thresh = 1)
 
@@
 
-This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`benchmark_highdens`).
+This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`hd_thresh`).
 The function returns an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
@@
   q = 0.1, 
-  benchmark_highdens = 1)
+  hd_thresh = 1)
 ```
@@
   q = 0.1, 
-  benchmark_highdens = 0)
+  hd_thresh = 0)
 

### 2025-09-17 paper-quollr.Rmd
Updated find_pts

@@
 
-scurve_umap_scaled <- scurve_model_obj$nldr_obj$scaled_nldr
+scurve_umap_scaled <- scurve_model_obj$nldr_scaled_obj$scaled_nldr
 tr_from_to_df_scurve <- scurve_model_obj$trimesh_data
@@
 bin_configs <- calc_bins_y(
-  nldr_obj = scurve_umap_obj, 
+  nldr_scaled_obj = scurve_umap_obj, 
   b1 = 15, 
@@
 hb_obj <- hex_binning(
-  nldr_obj = scurve_umap_obj, 
+  nldr_scaled_obj = scurve_umap_obj, 
   b1 = 15, 
@@
 
-```{r}
+```{r, eval=FALSE}
 #| label: code-illustration
@@
 hb_obj_notation <- hex_binning(
-  nldr_obj = scurve_umap_obj, 
+  nldr_scaled_obj = scurve_umap_obj, 
   b1 = 9, 
@@
 
-```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos='H'}
+```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos='H', eval=FALSE}
 
@@
 all_centroids_df <- gen_centroids(
-  nldr_obj = scurve_umap_obj, 
+  nldr_scaled_obj = scurve_umap_obj, 
   b1 = 15, 
@@
 umap_hex_id <- assign_data(
-  nldr_obj = scurve_umap_obj, 
+  nldr_scaled_obj = scurve_umap_obj, 
   centroids_data = all_centroids_df
@@
 
-The `find_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon in the NLDR space.
+The `group_hex_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon in the NLDR space.
 
@@
 ```{r, echo=TRUE}
-pts_df <- find_pts(
+pts_df <- group_hex_pts(
   scaled_nldr_hexid = umap_hex_id
@@
 trimesh_data_limb <- tsne_limb_obj$trimesh_data
-tsne_limb_scaled <- tsne_limb_obj$nldr_obj$scaled_nldr
+tsne_limb_scaled <- tsne_limb_obj$nldr_scaled_obj$scaled_nldr
 
@@
 trimesh_data_limb <- tsne_best_limb_obj$trimesh_data
-tsne_limb_scaled <- tsne_best_limb_obj$nldr_obj$scaled_nldr
+tsne_limb_scaled <- tsne_best_limb_obj$nldr_scaled_obj$scaled_nldr
 

### 2025-09-17 paper-quollr.Rmd
Updated glance and augment function implementations

@@
 <!-- If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
-For example, if the user wishes to use already binning data, the `extract_hexbin_centroids()` function can be used directly.
+For example, if the user wishes to use already binning data, the `merge_hexbin_centroids()` function can be used directly.
 -->
@@
 
-The `extract_hexbin_centroids()` function combines hexagonal bin coordinates, raw and standardized counts within each hexagons.
+The `merge_hexbin_centroids()` function combines hexagonal bin coordinates, raw and standardized counts within each hexagons.
 
@@
 ```{r, echo=TRUE}
-df_bin_centroids <- extract_hexbin_centroids(
+df_bin_centroids <- merge_hexbin_centroids(
   centroids_data = all_centroids_df, 
@@
   b1 = 15, 
-  benchmark_mean_dens = 0.05
+  md_thresh = 0.05
 )
@@
 glance(
-  highd_data = scurve, 
-  model_2d = df_bin_centroids, 
-  model_highd = df_bin
+  model_object = scurve_model_obj,
+  highd_data = scurve
   )
@@
 model_error <- augment(
-  highd_data = scurve, 
-  model_2d = df_bin_centroids, 
-  model_highd = df_bin
+  model_object = scurve_model_obj,
+  highd_data = scurve
   )

### 2025-09-17 paper-quollr.tex
Updated glance and augment function implementations

@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different $2 ext{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different $2\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.}\label{fig:overview}
@@
 
-The remaining variables \(X_4\), \(X_5\), \(X_6\), \(X_7\) are all uniform error, with small variance.
+The remaining variables \(X_4\), \(X_5\), \(X_6\), and \(X_7\) are included as low-magnitude noise dimensions, generated from uniform distributions with narrow ranges around zero: \(X_4, X_5 \sim U(-0.02, 0.02)\), \(X_6 \sim U(-0.1, 0.1)\), and \(X_7 \sim U(-0.01, 0.01)\).
+These small variances ensure that the added dimensions contribute noise without obscuring the main geometric structure.
 \texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors}\(=46\) and \texttt{min\_dist}\(=0.9\).
@@
 
-This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{benchmark\_highdens}).
-The function returns an object that includes the scaled NLDR object (\texttt{nldr\_obj}), the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
+This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{hd\_thresh}).
+The function returns an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
 
@@
   q = 0.1, 
-  benchmark_highdens = 1)
+  hd_thresh = 1)
 \end{verbatim}
@@
 
-The algorithm starts by scaling the NLDR data to to the range \([0, 1] \times [0, x_2]\), where \(x_2\) is the maximum value of the second embedding, using the \texttt{gen\_scaled\_data()} function.
+The algorithm starts by scaling the NLDR data to to the range \([0, 1] \times [0, y_{2,max}]\), where \(y_{2,max} = r_2/r_1\) is the ratio of ranges of embedding components.
 The output includes the scaled NLDR data (\texttt{scaled\_nldr}) along with the original limits of the embeddings (\texttt{lim1}, \texttt{lim2}).
@@
 The function \texttt{calc\_bins\_y()} is used for this purpose.
-This function accepts (1) an object containing three elements: the first is the scaled NLDR data with two columns, and the second and third are the limits of the original NLDR data; (2) the number of bins along the x-axis (\texttt{b1}), and (3) the buffer amount as a proportion (\texttt{q}).
+This function accepts (1) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (2) the number of bins along the x-axis (\texttt{b1}), and (3) the buffer amount as a proportion (\texttt{q}).
+The buffer ensures that the grid fully covers the data space by extending one hexagon width (\(a_1\)) and height (\(a_2\)) beyond the observed data in all directions.
+By default, \(q = 0.1\), but it must be set to a value smaller than the minimum data value to avoid exceeding the data range.
 
@@
 Points are allocated to bins based on the nearest centroid of the hexagonal bins.
-The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its components can be run separately for added flexibility.
-The parameters used within \texttt{hex\_binning()} include an object containing scaled NLDR data in the first and second columns, along with numeric vectors that represent the limits of the original NLDR data (\texttt{nldr\_obj}), the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
+The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its individual components can be run separately for added flexibility.
+While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the \texttt{hex\_binning()} function automates this entire workflow.
+The parameters used within \texttt{hex\_binning()} include an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
 The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{tot\_bins}), the number of non-empty bins (\texttt{non\_bins}), and the points within each hexagon (\texttt{pts\_bins}).
@@
 \end{verbatim}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.5\linewidth]{paper-quollr_files/figure-latex/hex-param-1}
-
-}
-
-\caption{The components of the hexagon grid illustrating notation.}\label{fig:hex-param}
-\end{figure}
-
-If the hexagonal binning process is run separately, it involves several steps: (i) generating all possible centroids in a hexagonal grid, (ii) creating the coordinates of the hexagons, (iii) assigning data points to their respective hexagons, (iv) computing the standardized number of points within each hexagon, and (v) mapping the points to their corresponding hexagonal bins.
 
@@
 
-For each centroid, six vertices are computed and assigned a polygon ID (\texttt{h}) corresponding to the centroid's \texttt{h}.
-These points are aggregated into a tibble that includes the polygon ID (\texttt{h}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
+For each centroid, six vertices are computed and assigned a polygon ID as the centroid.
+These vertices are then combined into a tibble that records the polygon ID (\texttt{h}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
 
@@
 To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin's count by the maximum count across all bins.
-This yields a normalized metric, \texttt{w\_h}, ranging from 0 to 1.
+This yields a standardized bin counts, \texttt{w\_h}, ranging from 0 to 1.
 
@@
 
-The \texttt{find\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon in the NLDR space.
+The \texttt{group\_hex\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon in the NLDR space.
 
@@
 \begin{verbatim}
-pts_df <- find_pts(
+pts_df <- group_hex_pts(
   scaled_nldr_hexid = umap_hex_id
@@
 
-The \texttt{extract\_hexbin\_centroids()} function combines hexagonal bin coordinates, raw and standardized counts within each hexagons.
+The \texttt{merge\_hexbin\_centroids()} function combines hexagonal bin coordinates, raw and standardized counts within each hexagons.
 
@@
 \begin{verbatim}
-df_bin_centroids <- extract_hexbin_centroids(
+df_bin_centroids <- merge_hexbin_centroids(
   centroids_data = all_centroids_df, 
@@
 
-The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{tripack::tri.mesh()} function.
+The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{tripack::tri.mesh()} function (\citet{albrecht2025}).
 This triangulation forms the structural basis for identifying adjacent bins.
@@
 \begin{verbatim}
-> # A tibble: 5 x 8
->    from    to  x_from  y_from    x_to    y_to from_count to_count
->   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1    -0.116  -0.0166 -0.116           0        0
-> 2    16    17 -0.0583 -0.0434  0.0251 -0.0434          0        0
-> 3    16    32 -0.0583 -0.0434 -0.0166  0.0288          0        0
-> 4     3    17  0.0668 -0.116   0.0251 -0.0434          0        0
-> 5    32    33 -0.0166  0.0288  0.0668  0.0288          0        0
+> # A tibble: 5 x 10
+>    from    to  x_from  y_from    x_to    y_to from_count to_count from_reindexed
+>   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>          <int>
+> 1     1     2 -0.1    -0.116  -0.0166 -0.116           0        0              1
+> 2    16    17 -0.0583 -0.0434  0.0251 -0.0434          0        0             16
+> 3    16    32 -0.0583 -0.0434 -0.0166  0.0288          0        0             16
+> 4     3    17  0.0668 -0.116   0.0251 -0.0434          0        0              3
+> 5    32    33 -0.0166  0.0288  0.0668  0.0288          0        0             32
+> # i 1 more variable: to_reindexed <int>
 \end{verbatim}
@@
 
-For simplicity, we remove low-density hexagons using a threshold of \(10\).
+For simplicity, we remove low-density hexagons using a threshold of \(1\).
 
@@
 
+It is worth noting that while \texttt{predict\_emb()} provides a general approach that works across methods, some NLDR techniques have their own built-in prediction mechanisms.
+For example, UMAP (\citet{tomasz2023}) supports direct prediction of embeddings for new data once a model is fitted.
+
 \subsubsection{Compute residuals and Root Mean Square Error (RMSE)}\label{compute-residuals-and-root-mean-square-error-rmse}
@@
 >   <dbl> <dbl>
-> 1  265. 0.157
+> 1  273. 0.163
 \end{verbatim}

### 2025-09-17 paper-quollr.Rmd
Updated with new implementation

@@
   q = 0.1, 
-  hd_thresh = 1)
+  hd_thresh = 0)
 
@@
 scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = tr_from_to_df_scurve$from,
-                         lineTo = tr_from_to_df_scurve$to,
+                         lineFrom = tr_from_to_df_scurve$from_reindexed,
+                         lineTo = tr_from_to_df_scurve$to_reindexed,
                          group = df_exe$type,
@@
 scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = tr_from_to_df_scurve$from,
-                         lineTo = tr_from_to_df_scurve$to,
+                         lineFrom = tr_from_to_df_scurve$from_reindexed,
+                         lineTo = tr_from_to_df_scurve$to_reindexed,
                          group = df_exe$type,
@@
   nldr_data = scurve_umap, 
-  b1 = 15, 
+  b1 = 25, 
   q = 0.1, 
@@
   nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 15, 
+  b1 = 25, 
   q = 0.1)
@@
   nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 15, 
+  b1 = 25, 
   q = 0.1
@@
   model_2d = df_bin_centroids, 
-  b1 = 15, 
+  b1 = 25, 
   md_thresh = 0.05
@@
 limb_highd_vis <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
-                         lineFrom = trimesh_data_limb$from,
-                         lineTo = trimesh_data_limb$to,
+                         lineFrom = trimesh_data_limb$from_reindexed,
+                         lineTo = trimesh_data_limb$to_reindexed,
                          group = factor(df_model_data_limb_n$type,
@@
 limb_highd_vis_best <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
-                         lineFrom = trimesh_data_limb$from,
-                         lineTo = trimesh_data_limb$to,
+                         lineFrom = trimesh_data_limb$from_reindexed,
+                         lineTo = trimesh_data_limb$to_reindexed,
                          group = factor(df_model_data_limb_n$type,

### 2025-09-17 paper-quollr.tex
Updated with new implementation

@@
 > $b2
-> [1] 20
+> [1] 33
 > 
 > $a1
-> [1] 0.0834108
+> [1] 0.04952516
 > 
 > $a2
-> [1] 0.07223587
+> [1] 0.04289005
 \end{verbatim}
@@
 \begin{verbatim}
-> # A tibble: 300 x 3
+> # A tibble: 825 x 3
 >        h       c_x    c_y
@@
 >  1     1 -0.1      -0.116
->  2     2 -0.0166 -0.116
->  3     3  0.0668 -0.116
->  4     4  0.150  -0.116
->  5     5  0.234  -0.116
->  6     6  0.317  -0.116
->  7     7  0.400  -0.116
->  8     8  0.484  -0.116
->  9     9  0.567  -0.116
-> 10    10  0.651  -0.116
-> # i 290 more rows
+>  2     2 -0.0505   -0.116
+>  3     3 -0.000950 -0.116
+>  4     4  0.0486   -0.116
+>  5     5  0.0981   -0.116
+>  6     6  0.148    -0.116
+>  7     7  0.197    -0.116
+>  8     8  0.247    -0.116
+>  9     9  0.296    -0.116
+> 10    10  0.346    -0.116
+> # i 815 more rows
 \end{verbatim}
@@
 >   h           x           y
-> 1 1 -0.1000000 -0.06752289
-> 2 1 -0.1417054 -0.09160152
-> 3 1 -0.1417054 -0.13975877
-> 4 1 -0.1000000 -0.16383739
-> 5 1 -0.0582946 -0.13975877
+> 1 1 -0.10000000 -0.08708678
+> 2 1 -0.12476258 -0.10138346
+> 3 1 -0.12476258 -0.12997683
+> 4 1 -0.10000000 -0.14427351
+> 5 1 -0.07523742 -0.12997683
 \end{verbatim}
@@
 >   <dbl> <dbl> <int> <int>
-> 1 0.277 0.913     1   216
-> 2 0.697 0.538     2   145
-> 3 0.779 0.399     3   116
-> 4 0.173 0.953     4   229
-> 5 0.218 0.983     5   229
+> 1 0.277 0.913     1   609
+> 2 0.697 0.538     2   392
+> 3 0.779 0.399     3   319
+> 4 0.173 0.953     4   631
+> 5 0.218 0.983     5   657
 \end{verbatim}
@@
 >   <int> <int> <dbl>
-> 1    34     1 0.001
-> 2    35    10 0.01 
-> 3    36     8 0.008
-> 4    37     8 0.008
-> 5    38     2 0.002
+> 1    82     1 0.001
+> 2    83     4 0.004
+> 3    84     1 0.001
+> 4    85     3 0.003
+> 5    86     1 0.001
 \end{verbatim}
@@
 >   <int> <list>   
-> 1    34 <int [1]> 
-> 2    35 <int [10]>
-> 3    36 <int [8]> 
-> 4    37 <int [8]> 
-> 5    38 <int [2]>
+> 1    82 <int [1]>
+> 2    83 <int [4]>
+> 3    84 <int [1]>
+> 4    85 <int [3]>
+> 5    86 <int [1]>
 \end{verbatim}
@@
 >   h           c_x        c_y n_h w_h
-> 1 1 -0.1000000 -0.1156801   0   0
-> 2 2 -0.0165892 -0.1156801   0   0
-> 3 3  0.0668216 -0.1156801   0   0
-> 4 4  0.1502324 -0.1156801   0   0
-> 5 5  0.2336432 -0.1156801   0   0
+> 1 1 -0.1000000000 -0.1156801   0   0
+> 2 2 -0.0504748374 -0.1156801   0   0
+> 3 3 -0.0009496749 -0.1156801   0   0
+> 4 4  0.0485754877 -0.1156801   0   0
+> 5 5  0.0981006503 -0.1156801   0   0
 \end{verbatim}
@@
 >   <int> <int>     <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1    -0.116  -0.0166 -0.116           0        0
-> 2    16    17 -0.0583 -0.0434  0.0251 -0.0434          0        0
-> 3    16    32 -0.0583 -0.0434 -0.0166  0.0288          0        0
-> 4     3    17  0.0668 -0.116   0.0251 -0.0434          0        0
-> 5    32    33 -0.0166  0.0288  0.0668  0.0288          0        0
+> 1     1     2 -0.1      -0.116  -0.0505 -0.116           0        0
+> 2    26    27 -0.0752   -0.0728 -0.0257 -0.0728          0        0
+> 3     3    27 -0.000950 -0.116  -0.0257 -0.0728          0        0
+> 4    26    52 -0.0752   -0.0728 -0.0505 -0.0299          0        0
+> 5    27    28 -0.0257   -0.0728  0.0238 -0.0728          0        0
 \end{verbatim}
@@
 > # A tibble: 5 x 10
->    from    to  x_from  y_from    x_to    y_to from_count to_count from_reindexed
->   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>          <int>
-> 1     1     2 -0.1    -0.116  -0.0166 -0.116           0        0              1
-> 2    16    17 -0.0583 -0.0434  0.0251 -0.0434          0        0             16
-> 3    16    32 -0.0583 -0.0434 -0.0166  0.0288          0        0             16
-> 4     3    17  0.0668 -0.116   0.0251 -0.0434          0        0              3
-> 5    32    33 -0.0166  0.0288  0.0668  0.0288          0        0             32
-> # i 1 more variable: to_reindexed <int>
+>    from    to    x_from  y_from    x_to    y_to from_count to_count
+>   <int> <int>     <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
+> 1     1     2 -0.1      -0.116  -0.0505 -0.116           0        0
+> 2    26    27 -0.0752   -0.0728 -0.0257 -0.0728          0        0
+> 3     3    27 -0.000950 -0.116  -0.0257 -0.0728          0        0
+> 4    26    52 -0.0752   -0.0728 -0.0505 -0.0299          0        0
+> 5    27    28 -0.0257   -0.0728  0.0238 -0.0728          0        0
+> # i 2 more variables: from_reindexed <int>, to_reindexed <int>
 \end{verbatim}
@@
 > [271] 271 272 273 274 275 276 277 278 279 280 281 282 283 284 285 286 287 288
-> [289] 289 290 291 292 293 294 295 296 297 298 299 300
+> [289] 289 290 291 292 293 294 295 296 297 298 299 300 301 302 303 304 305 306
+> [307] 307 308 309 310 311 312 313 314 315 316 317 318 319 320 321 322 323 324
+> [325] 325 326 327 328 329 330 331 332 333 334 335 336 337 338 339 340 341 342
+> [343] 343 344 345 346 347 348 349 350 351 352 353 354 355 356 357 358 359 360
+> [361] 361 362 363 364 365 366 367 368 369 370 371 372 373 374 375 376 377 378
+> [379] 379 380 381 382 383 384 385 386 387 388 389 390 391 392 393 394 395 396
+> [397] 397 398 399 400 401 402 403 404 405 406 407 408 409 410 411 412 413 414
+> [415] 415 416 417 418 419 420 421 422 423 424 425 426 427 428 429 430 431 432
+> [433] 433 434 435 436 437 438 439 440 441 442 443 444 445 446 447 448 449 450
+> [451] 451 452 453 454 455 456 457 458 459 460 461 462 463 464 465 466 467 468
+> [469] 469 470 471 472 473 474 475 476 477 478 479 480 481 482 483 484 485 486
+> [487] 487 488 489 490 491 492 493 494 495 496 497 498 499 500 501 502 503 504
+> [505] 505 506 507 508 509 510 511 512 513 514 515 516 517 518 519 520 521 522
+> [523] 523 524 525 526 527 528 529 530 531 532 533 534 535 536 537 538 539 540
+> [541] 541 542 543 544 545 546 547 548 549 550 551 552 553 554 555 556 557 558
+> [559] 559 560 561 562 563 564 565 566 567 568 569 570 571 572 573 574 575 576
+> [577] 577 578 579 580 581 582 583 584 585 586 587 588 589 590 591 592 593 594
+> [595] 595 596 597 598 599 600 601 602 603 604 605 606 607 608 609 610 611 612
+> [613] 613 614 615 616 617 618 619 620 621 622 623 624 625 626 627 628 629 630
+> [631] 631 632 633 634 635 636 637 638 639 640 641 642 643 644 645 646 647 648
+> [649] 649 650 651 652 653 654 655 656 657 658 659 660 661 662 663 664 665 666
+> [667] 667 668 669 670 671 672 673 674 675 676 677 678 679 680 681 682 683 684
+> [685] 685 686 687 688 689 690 691 692 693 694 695 696 697 698 699 700 701 702
+> [703] 703 704 705 706 707 708 709 710 711 712 713 714 715 716 717 718 719 720
+> [721] 721 722 723 724 725 726 727 728 729 730 731 732 733 734 735 736 737 738
+> [739] 739 740 741 742 743 744 745 746 747 748 749 750 751 752 753 754 755 756
+> [757] 757 758 759 760 761 762 763 764 765 766 767 768 769 770 771 772 773 774
+> [775] 775 776 777 778 779 780 781 782 783 784 785 786 787 788 789 790 791 792
+> [793] 793 794 795 796 797 798 799 800 801 802 803 804 805 806 807 808 809 810
+> [811] 811 812 813 814 815 816 817 818 819 820 821 822 823 824 825
 \end{verbatim}
@@
 >   <int> <dbl>  <dbl> <dbl>    <dbl>    <dbl>   <dbl>    <dbl>
-> 1    34 0.958  0.0854  1.29  0.00265    0.0171    0.0876 -0.00249  
-> 2    35 0.818  0.116   1.56  0.00184    0.00361  -0.0318 -0.00377  
-> 3    36 0.544  0.111   1.83 -0.00341   -0.000303  0.0196  0.0000704
-> 4    37 0.279  0.128   1.95  0.0000880  0.00104  -0.0276 -0.000227 
-> 5    38 0.0567 0.119   2.00  0.00733    0.00238   0.0833 -0.00166
+> 1    82 0.888 0.102   1.46  0.0145  -0.0118   0.0303 -0.00707
+> 2    83 0.744 0.123   1.66 -0.00950  0.00499 -0.0208 -0.00425
+> 3    84 0.591 0.0447  1.81 -0.0168   0.0173  -0.0485 -0.00112
+> 4    85 0.371 0.0937  1.93 -0.00109  0.00840  0.0268 -0.00286
+> 5    86 0.255 0.0447  1.97 -0.00862 -0.00733 -0.0826  0.00229
 \end{verbatim}
@@
 >        <dbl>      <dbl> <int>  <int>
-> 1      0.234      0.896     1    215
-> 2      0.692      0.534     2    145
-> 3      0.776      0.390     3    116
-> 4      0.192      0.968     4    229
-> 5      0.192      0.968     5    229
+> 1     NA         NA         1    609
+> 2      0.717      0.528     2    392
+> 3      0.791      0.399     3    319
+> 4      0.172      0.957     4    631
+> 5      0.197      0.999     5    657
 \end{verbatim}
@@
 >   <dbl> <dbl>
-> 1  273. 0.163
+> 1  265. 0.157
 \end{verbatim}

### 2025-09-17 paper-quollr.R

@@ -97,7 +97,7 @@ scurve_model_obj <- fit_highd_model(
   nldr_data = scurve_umap, 
   b1 = 15, 
   q = 0.1, 
-  hd_thresh = 1)
+  hd_thresh = 0)
 
 scurve_umap_scaled <- scurve_model_obj$nldr_scaled_obj$scaled_nldr
 tr_from_to_df_scurve <- scurve_model_obj$trimesh_data
@@ -263,8 +263,8 @@ df_b <- df_exe |>
   dplyr::filter(type == "model") ## High-d model
 
 scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = tr_from_to_df_scurve$from,
-                         lineTo = tr_from_to_df_scurve$to,
+                         lineFrom = tr_from_to_df_scurve$from_reindexed,
+                         lineTo = tr_from_to_df_scurve$to_reindexed,
                          group = df_exe$type,
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
                          levelColors = c(clr_choice, "#FF7755"),
@@ -272,8 +272,8 @@ scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          width = "421px", height = "421px")
 
 scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = tr_from_to_df_scurve$from,
-                         lineTo = tr_from_to_df_scurve$to,
+                         lineFrom = tr_from_to_df_scurve$from_reindexed,
+                         lineTo = tr_from_to_df_scurve$to_reindexed,
                          group = df_exe$type,
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
                          levelColors = c(clr_choice, "#FF7755"),
@@ -473,7 +473,7 @@ scurve_umap_plt + hex_grid_poly_scurve +
 # fit_highd_model(
 #   highd_data = scurve,
 #   nldr_data = scurve_umap,
-#   b1 = 15,
+#   b1 = 25,
 #   q = 0.1,
 #   hd_thresh = 1)
 
@@ -487,7 +487,7 @@ scurve_umap_obj
 ## ----echo=TRUE----------------------------------------------------------------
 bin_configs <- calc_bins_y(
   nldr_scaled_obj = scurve_umap_obj, 
[…]

### 2025-09-17 paper-quollr.Rmd
Initial work on adding notation table

@@
 These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].
+
+```{r notation-tb-html, eval=knitr::is_html_output()}
+# Notation used in the paper
+
+notation_df <- read_csv("misc/notation.csv")
+
+notation_df |> 
+  kable(caption = "Summary of notation for describing new methodology.") 
+```
+
+```{r notation-tb-pdf, eval=knitr::is_latex_output()}
+
+notation_df <- read_csv("misc/notation.csv")
+
+notation_df |> 
+  kable(caption = "Summary of notation for describing new methodology.", booktabs = T) |>
+  column_spec(1, width = "3cm") |>
+  column_spec(2, width = "12cm")
+```

### 2025-09-17 paper-quollr.tex
Initial work on adding notation table

@@
 
+\begin{table}
+
+\caption{\label{tab:notation-tb-pdf}Summary of notation for describing new methodology.}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
+\toprule
+Notation & Description\\
+\midrule
+\$n, p, k\$ & number of observations, variables, embedding dimension, respectively\\
+\$\textbackslash{}mathbfit\{X\}, \textbackslash{}mathbfit\{x\}\$ & \$p\$-dimensional data (population, sample)\\
+\$\textbackslash{}mathbfit\{y\}\$ & \$k\$-dimensional layout\\
+\$P\$ & orthonormal basis, generating a \$d\textbackslash{}text\{-\}dimensional\$ linear projection of \$p\$-dimensional data\\
+\$T\$ & true  model\\
+\addlinespace
+\$g\$ & functional mapping from \textbackslash{}pD\{\} to \textbackslash{}kD\{\}, especially as prescribed by NLDR\\
+\$\textbackslash{}mathbfit\{\textbackslash{}theta\}\$ & (Hyper-) parameters for NLDR method\\
+\$r\$ & ranges of the embedding components\\
+\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers\\
+\$(b\_1, b\_2)\$ & number of bins in each direction\\
+\addlinespace
+\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction\\
+\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid\\
+\$q\$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1\\
+\$m\$ & number of non-empty bins\\
+\$b\$ & number of  hexagons in the grid\\
+\addlinespace
+\$h\$ & hexagonal id\\
+\$l\$ & side length\\
+\$A\$ & area\\
+\$n\_h\$ & bin count\\
+\$w\_h\$ & standardised bin counts\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \bibliography{paper-quollr.bib}

### 2025-09-17 paper-quollr.R

@@ -1697,3 +1697,22 @@ trimesh_limb + limb_proj_tsne_model1 +
 # 
 # modellimbfig
 
+
+## ----notation-tb-html, eval=knitr::is_html_output()---------------------------
+# # Notation used in the paper
+# 
+# notation_df <- read_csv("misc/notation.csv")
+# 
+# notation_df |>
+#   kable(caption = "Summary of notation for describing new methodology.")
+
+
+## ----notation-tb-pdf, eval=knitr::is_latex_output()---------------------------
+
+notation_df <- read_csv("misc/notation.csv")
+
+notation_df |> 
+  kable(caption = "Summary of notation for describing new methodology.", booktabs = T) |>
+  column_spec(1, width = "3cm") |>
+  column_spec(2, width = "12cm")
+

### 2025-09-18 paper-quollr.Rmd
Fixed the counts to remove

@@
   nldr_data = scurve_umap, 
-  b1 = 15, 
+  b1 = 21, 
   q = 0.1, 
@@
 df_bin_centroids <- df_bin_centroids |>
-  dplyr::filter(n_h > 1)
+  dplyr::filter(n_h > 0)
 
 trimesh <- trimesh |>
-  dplyr::filter(from_count > 1,
-                to_count > 1)
+  dplyr::filter(from_count > 0,
+                to_count > 0)
 

### 2025-09-18 paper-quollr.tex
Fixed the issue
Fixed the counts to remove

@@
 >        <dbl>      <dbl> <int>  <int>
-> 1     NA         NA         1    609
+> 1      0.296      0.914     1    609
 > 2      0.717      0.528     2    392
@@
 >   <dbl> <dbl>
-> 1  265. 0.157
+> 1  196. 0.116
 \end{verbatim}
@@
 >   <int>   <dbl> <dbl>    <dbl>    <dbl>    <dbl>   <dbl>    <dbl>  <int>
-> 1     1 -0.120  0.819 -1.99     0.0114   0.00351  0.0334  0.00638    215
-> 2     2 -0.0492 0.166  0.00121  0.0115  -0.0166  -0.0297  0.00509    145
-> 3     3 -0.774  0.651  0.367   -0.0172   0.00600  0.0211  0.00303    116
-> 4     4 -0.606  0.952 -1.80     0.0157  -0.00978 -0.0590 -0.00754    229
-> 5     5 -0.478  1.10  -1.88    -0.00423  0.00495 -0.0482 -0.00982    229
+> 1     1 -0.120  0.819 -1.99     0.0114   0.00351  0.0334  0.00638    427
+> 2     2 -0.0492 0.166  0.00121  0.0115  -0.0166  -0.0297  0.00509    287
+> 3     3 -0.774  0.651  0.367   -0.0172   0.00600  0.0211  0.00303    226
+> 4     4 -0.606  0.952 -1.80     0.0157  -0.00978 -0.0590 -0.00754    446
+> 5     5 -0.478  1.10  -1.88    -0.00423  0.00495 -0.0482 -0.00982    468
 > # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
@@
 \midrule
-\$n, p, k\$ & number of observations, variables, embedding dimension, respectively\\
+\textbackslash{}\$n, p, k\textbackslash{}\$ & number of observations, variables, embedding dimension, respectively\\
 \$\textbackslash{}mathbfit\{X\}, \textbackslash{}mathbfit\{x\}\$ & \$p\$-dimensional data (population, sample)\\

### 2025-09-18 paper-quollr.Rmd
Used notation for 2-D

@@
 This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
-Starting with an NLDR layout, our approach is to create a wireframe model representation, that can be lifted and displayed in the high-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:overview)"})`).
+Starting with an NLDR layout, our approach is to create a \twoD wireframe model representation, that can be lifted and displayed in the high-dimensional (\pD) space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:overview)"})`).
 

### 2025-09-18 paper-quollr.tex
Used notation for 2-D

@@
 This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
-Starting with an NLDR layout, our approach is to create a wireframe model representation, that can be lifted and displayed in the high-dimensional space (Figure \ref{fig:overview}).
+Starting with an NLDR layout, our approach is to create a $2\text{-}D$ wireframe model representation, that can be lifted and displayed in the high-dimensional ($p\text{-}D$) space (Figure \ref{fig:overview}).
 

### 2025-09-18 paper-quollr.Rmd
Updated notation table

@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different \\twoD projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H'}
@@
   nldr_data = scurve_umap, 
-  b1 = 25, 
+  b1 = 21, 
   q = 0.1, 
-  hd_thresh = 1)
+  hd_thresh = 0)
 ```
@@
 
-Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
+Constructing the \twoD model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
 
@@
   column_spec(1, width = "3cm") |>
-  column_spec(2, width = "12cm")
-```
+  column_spec(2, width = "10cm") |>
+  column_spec(3, width = "3cm")
+```

### 2025-09-18 paper-quollr.tex
Updated notation table

@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different $2\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different $2\text{-}D$projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.}\label{fig:overview}
@@
   nldr_data = scurve_umap, 
-  b1 = 25, 
+  b1 = 21, 
   q = 0.1, 
-  hd_thresh = 1)
+  hd_thresh = 0)
 \end{verbatim}
@@
 
-Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
+Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
 
@@
 \centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{10cm}>{\raggedright\arraybackslash}p{3cm}}
 \toprule
-Notation & Description\\
+Notation & Description & Variable\\
 \midrule
-\textbackslash{}\$n, p, k\textbackslash{}\$ & number of observations, variables, embedding dimension, respectively\\
-\$\textbackslash{}mathbfit\{X\}, \textbackslash{}mathbfit\{x\}\$ & \$p\$-dimensional data (population, sample)\\
-\$\textbackslash{}mathbfit\{y\}\$ & \$k\$-dimensional layout\\
-\$P\$ & orthonormal basis, generating a \$d\textbackslash{}text\{-\}dimensional\$ linear projection of \$p\$-dimensional data\\
-\$T\$ & true  model\\
+\textbackslash{}\$n, p, k\textbackslash{}\$ & number of observations, variables, embedding dimension, respectively & NA\\
+\$\textbackslash{}mathbfit\{X\}, \textbackslash{}mathbfit\{x\}\$ & \$p\$-dimensional data (population, sample) & NA\\
+\$\textbackslash{}mathbfit\{y\}\$ & \$k\$-dimensional layout & NA\\
+\$P\$ & orthonormal basis, generating a \$d\textbackslash{}text\{-\}dimensional\$ linear projection of \$p\$-dimensional data & NA\\
+\$T\$ & true  model & NA\\
 \addlinespace
-\$g\$ & functional mapping from \textbackslash{}pD\{\} to \textbackslash{}kD\{\}, especially as prescribed by NLDR\\
-\$\textbackslash{}mathbfit\{\textbackslash{}theta\}\$ & (Hyper-) parameters for NLDR method\\
-\$r\$ & ranges of the embedding components\\
-\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers\\
-\$(b\_1, b\_2)\$ & number of bins in each direction\\
+\$g\$ & functional mapping from \textbackslash{}pD\{\} to \textbackslash{}kD\{\}, especially as prescribed by NLDR & NA\\
+\$\textbackslash{}mathbfit\{\textbackslash{}theta\}\$ & (Hyper-) parameters for NLDR method & NA\\
+\$r\$ & ranges of the embedding components & NA\\
+\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers & NA\\
+\$(b\_1, b\_2)\$ & number of bins in each direction & b1, b2\\
 \addlinespace
-\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction\\
-\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid\\
-\$q\$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1\\
-\$m\$ & number of non-empty bins\\
-\$b\$ & number of  hexagons in the grid\\
+\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction & NA\\
+\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid & NA\\
+\$q\$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1 & q\\
+\$m\$ & number of non-empty bins & NA\\
+\$b\$ & number of  hexagons in the grid & NA\\
 \addlinespace
-\$h\$ & hexagonal id\\
-\$l\$ & side length\\
-\$A\$ & area\\
-\$n\_h\$ & bin count\\
-\$w\_h\$ & standardised bin counts\\
+\$h\$ & hexagonal id & NA\\
+\$l\$ & side length & NA\\
+\$A\$ & area & NA\\
+\$n\_h\$ & bin count & NA\\
+\$w\_h\$ & standardised bin counts & NA\\
+\addlinespace
+NA & threshold to extract high density hexagons & hd\_thresh\\
 \bottomrule

### 2025-09-18 paper-quollr.Rmd
Updated section Computing hexagon grid configurations

@@
 
-#### Computing hexagon grid configurations
-
-The configurations of a hexagonal grid are determined by the number of bins and the bin width in each direction.
-The function `calc_bins_y()` is used for this purpose.
+#### Computing hexagon grid configuration
+
+The function `calc_bins_y()` determines the configuration of the hexagonal grid by computing the number of bins along the y-axis (`b2`), the hexagon width (`a1`), and height (`a2`).
 This function accepts (1) an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); (2) the number of bins along the x-axis (`b1`), and (3) the buffer amount as a proportion (`q`).

### 2025-09-18 paper-quollr.tex
Updated section Computing hexagon grid configurations

@@
 
-\paragraph{Computing hexagon grid configurations}\label{computing-hexagon-grid-configurations}
-
-The configurations of a hexagonal grid are determined by the number of bins and the bin width in each direction.
-The function \texttt{calc\_bins\_y()} is used for this purpose.
+\paragraph{Computing hexagon grid configuration}\label{computing-hexagon-grid-configuration}
+
+The function \texttt{calc\_bins\_y()} determines the configuration of the hexagonal grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon width (\texttt{a1}), and height (\texttt{a2}).
 This function accepts (1) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (2) the number of bins along the x-axis (\texttt{b1}), and (3) the buffer amount as a proportion (\texttt{q}).

### 2025-09-18 paper-quollr.Rmd
Updated binning data text

@@
 The parameters used within `hex_binning()` include an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
-The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagon IDs (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`tot_bins`), the number of non-empty bins (`non_bins`), and the points within each hexagon (`pts_bins`).
+The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagon IDs (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`b`), the number of non-empty bins (`m`), and the points within each hexagon (`pts_bins`).
 

### 2025-09-18 paper-quollr.tex
Updated binning data text

@@
 The parameters used within \texttt{hex\_binning()} include an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
-The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{tot\_bins}), the number of non-empty bins (\texttt{non\_bins}), and the points within each hexagon (\texttt{pts\_bins}).
+The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{b}), the number of non-empty bins (\texttt{m}), and the points within each hexagon (\texttt{pts\_bins}).
 
@@
 \$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers & NA\\
-\$(b\_1, b\_2)\$ & number of bins in each direction & b1, b2\\
+\$(b\_1, b\_2)\$ & number of bins in each direction & (b1, b2) = bins\\
 \addlinespace
-\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction & NA\\
-\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid & NA\\
+\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction & a1, a2\\
+\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid & start\_point\\
 \$q\$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1 & q\\
-\$m\$ & number of non-empty bins & NA\\
-\$b\$ & number of  hexagons in the grid & NA\\
+\$m\$ & number of non-empty bins & m\\
+\$b\$ & number of  hexagons in the grid & b\\
 \addlinespace

### 2025-09-18 paper-quollr.Rmd
Updated vertical spacing

@@
 The coordinate limits of the embedding (`lim1` and `lim2`) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
-The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagonal rows (`bin2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`b1`) and buffer (`q`).
+The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagons along y-axis (`b2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`b1`) and buffer (`q`).
 
@@
 Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
-Vertical spacing (`a2`) is given by the height of a hexagon, $\sqrt{3}/2 \times a_1$.
+Vertical spacing (`vs`) is given by $\sqrt{3}/2 \times a_1$.
 
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2\text{-}D$ space.
+Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the \twoD space.
 

### 2025-09-18 paper-quollr.tex
Updated vertical spacing

@@
 The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
-The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagonal rows (\texttt{bin2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{b1}) and buffer (\texttt{q}).
+The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagons along y-axis (\texttt{b2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{b1}) and buffer (\texttt{q}).
 
@@
 Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
-Vertical spacing (\texttt{a2}) is given by the height of a hexagon, \(\sqrt{3}/2 \times a_1\).
+Vertical spacing (\texttt{vs}) is given by \(\sqrt{3}/2 \times a_1\).
 
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2\text{-}D\) space.
+Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the $2\text{-}D$ space.
 
@@
 \$r\$ & ranges of the embedding components & NA\\
-\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers & NA\\
+\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers & (c\_x, c\_y)\\
 \$(b\_1, b\_2)\$ & number of bins in each direction & (b1, b2) = bins\\

### 2025-09-18 paper-quollr.Rmd
Updated notations

@@
 This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`hd_thresh`).
-The function returns an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
+The function returns an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both \twoD (`model_2d`), and \pD (`model_highd`), and triangular mesh (`trimesh_data`).
 
@@
 The function first verifies the presence of the required hexagon width parameter `a1`.
-This width determines the horizontal spacing.
+This width determines the horizontal spacing (`hs`).
 
@@
 The horizontal and vertical offset is defined as $dx = a_1/2$, and $dy = a_1/\sqrt{3}$ repectively.
-A vertical spacing factor $v = a_1/2\sqrt{3}$ refines vertical placement in staggered rows.
+A vertical spacing factor $vf = a_1/2\sqrt{3}$ refines vertical placement in staggered rows.
 
@@
 After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The `assign_data()` function performs this assignment by calculating the $2\text{-}D$ Euclidean distance between each point in the $2\text{-}D$ embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2\text{-}D$ layout.
-It then selects the corresponding x and y coordinates of each hexagon’s centroid.
+The `assign_data()` function performs this assignment by calculating the \twoD Euclidean distance between each point in the \twoD embedding and all hexagon centroids.
+
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \twoD layout. It then selects the corresponding x and y coordinates of each hexagon’s centroid.
 
@@
 
-The result is a data frame of the scaled $2\text{-}D$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
+The result is a data frame of the scaled \twoD embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
 
@@
 
-The function first groups the input data by `h`, which represents the hexagon label associated with each point in the $2\text{-}D$ layout.
+The function first groups the input data by `h`, which represents the hexagon label associated with each point in the \twoD layout.
 Within each group, it collects the `ID`s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
@@
 
-To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a \twoD layout, we employ Delaunay triangulation on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
@@
 Some may have very few or no data points due to the sparsity or shape of the underlying structure.
-Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or "holes" in the $2\text{-}D$ structure, potentially disrupting the continuity of the representation.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or "holes" in the \twoD structure, potentially disrupting the continuity of the representation.
 
@@
 This is done by modelling a point in \pD as the  \pD mean of data points in the \twoD centroid.
-This is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).
+This is performed using the `avg_highd_data()` function, which takes \pD data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).
 
@@
 
-The `predict_emb()` function is used to predict a point in a $2\text{-}D$ embedding for a new $p\text{-}D$ data point using the fitted model.
-This function is useful to predict $2\text{-}D$ embedding irrespective of the NLDR technique.
-
-In the prediction process, first, the nearest $p\text{-}D$ model point is identified for the new $p\text{-}D$ data point by computing $p\text{-}D$ Euclidean distance.
-Then, the corresponding $2\text{-}D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
-Finally, the coordinates of the identified $2\text{-}D$ bin centroid is used as the predicted NLDR embedding for the new $p\text{-}D$ data point.
+The `predict_emb()` function is used to predict a point in a \twoD embedding for a new \pD data point using the fitted model.
+This function is useful to predict \twoD embedding irrespective of the NLDR technique.
+
+In the prediction process, first, the nearest \pD model point is identified for the new \pD data point by computing \pD Euclidean distance.
+Then, the corresponding \twoD bin centroid mapping for the identified \pD model point is determined.
+Finally, the coordinates of the identified \twoD bin centroid is used as the predicted NLDR embedding for the new \pD data point.
 
@@
 
-Furthermore, `augment()` accepts $2\text{-}D$ and $p\text{-}D$ model points, and the $p\text{-}D$ data and adds information about each observation in the data set.
+Furthermore, `augment()` accepts \twoD and \pD model points, and the \pD data and adds information about each observation in the data set.
 
@@
 
-The package offers several $2\text{-}D$ visualizations, including:
+The package offers several \twoD visualizations, including:
 
@@
 
-The generated $p\text{-}D$ model, overlaid with the data, can also be visualized using `show_langevitour`.
-Additionally, it features a function for visualizing the $2\text{-}D$ projection of the fitted model overlaid on the data, called `plot_proj`.
+The generated \pD model, overlaid with the data, can also be visualized using `show_langevitour`.
+Additionally, it features a function for visualizing the \twoD projection of the fitted model overlaid on the data, called `plot_proj`.
 
@@
 
-The `geom_trimesh()` function introduces a custom `ggplot2` layer designed for visualizing $2\text{-}D$ wireframe on a provided set of bin centroids.
+The `geom_trimesh()` function introduces a custom `ggplot2` layer designed for visualizing \twoD wireframe on a provided set of bin centroids.
 
@@
 
-#### $p\text{-}D$ model visualization
-
-To visualise how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \pD using the `show_langevitour()` function.
+#### \pD model visualization
+
+To visualise how well the \pD model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \pD using the `show_langevitour()` function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.
@@
 
-In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \twoD wireframe structure mapped back to high-dimensional space.
 
@@
 
-To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
+To support interactive evaluation of how well the \pD model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
 This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
@@
 Before visualization, the input data must be prepared using the `comb_all_data_model()` function.
-This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2\text{-}D$ bin layout (`model_2d`):
+This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the \twoD bin layout (`model_2d`):
 
@@
 
-The function `show_link_plots()` generates two side-by-side, interactively linked plots; a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function `show_link_plots()` generates two side-by-side, interactively linked plots; a \twoD NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
 The function takes the output from `comb_all_data_model()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
@@
 Before visualization, the input data must be prepared using the `comb_all_data_model_error()` function.
-The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2\text{-}D$ model data (`model_2d`), and model error (`error_data`).
+The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), \twoD model data (`model_2d`), and model error (`error_data`).
 
@@
 
-The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a \twoD NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
 The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
@@
 Several core computations within `quollr` are optimized using compiled C++ code via the `Rcpp` and `RcppArmadillo` packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), $2\text{-}D$ distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), \twoD distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
 This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better $2\text{-}D$ layout.
+Our method help to provide a better \twoD layout.
 
@@
 
-The `quollr` package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in $2\text{-}D$ and lifting it into high-dimensional space.
-This lifted model provides a direct way to assess how well a $2\text{-}D$ layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
-The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate $2\text{-}D$ representations.
+The `quollr` package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in \twoD and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a \twoD layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate \twoD representations.
 
@@
 To support efficient computation, particularly for large-scale datasets, several core functions in `quollr` are implemented in C++ using `Rcpp` and `RcppArmadillo`.
-These include functions for computing Euclidean distances in high-dimensional and $2\text{-}D$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RMSE and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in $2\text{-}D$.
+These include functions for computing Euclidean distances in high-dimensional and \twoD space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RMSE and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in \twoD.
 Additionally, `gen_hex_coord_cpp()` constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-09-18 paper-quollr.tex
Updated notations

@@
 This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{hd\_thresh}).
-The function returns an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
+The function returns an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both $2\text{-}D$ (\texttt{model\_2d}), and $p\text{-}D$ (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
 
@@
 The function first verifies the presence of the required hexagon width parameter \texttt{a1}.
-This width determines the horizontal spacing.
+This width determines the horizontal spacing (\texttt{hs}).
 
@@
 The horizontal and vertical offset is defined as \(dx = a_1/2\), and \(dy = a_1/\sqrt{3}\) repectively.
-A vertical spacing factor \(v = a_1/2\sqrt{3}\) refines vertical placement in staggered rows.
+A vertical spacing factor \(vf = a_1/2\sqrt{3}\) refines vertical placement in staggered rows.
 
@@
 After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The \texttt{assign\_data()} function performs this assignment by calculating the \(2\text{-}D\) Euclidean distance between each point in the \(2\text{-}D\) embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \(2\text{-}D\) layout.
+The \texttt{assign\_data()} function performs this assignment by calculating the $2\text{-}D$ Euclidean distance between each point in the $2\text{-}D$ embedding and all hexagon centroids.
+
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2\text{-}D$ layout.
 It then selects the corresponding x and y coordinates of each hexagon's centroid.
@@
 
-The result is a data frame of the scaled \(2\text{-}D\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
+The result is a data frame of the scaled $2\text{-}D$ embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
 
@@
 
-The function first groups the input data by \texttt{h}, which represents the hexagon label associated with each point in the \(2\text{-}D\) layout.
+The function first groups the input data by \texttt{h}, which represents the hexagon label associated with each point in the $2\text{-}D$ layout.
 Within each group, it collects the \texttt{ID}s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
@@
 
-To represent the neighborhood structure of hexagonal bins in a \(2\text{-}D\) layout, we employ Delaunay triangulation on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
@@
 Some may have very few or no data points due to the sparsity or shape of the underlying structure.
-Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the \(2\text{-}D\) structure, potentially disrupting the continuity of the representation.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the $2\text{-}D$ structure, potentially disrupting the continuity of the representation.
 
@@
 This is done by modelling a point in $p\text{-}D$ as the $p\text{-}D$ mean of data points in the $2\text{-}D$ centroid.
-This is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
+This is performed using the \texttt{avg\_highd\_data()} function, which takes $p\text{-}D$ data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
 
@@
 
-The \texttt{predict\_emb()} function is used to predict a point in a \(2\text{-}D\) embedding for a new \(p\text{-}D\) data point using the fitted model.
-This function is useful to predict \(2\text{-}D\) embedding irrespective of the NLDR technique.
-
-In the prediction process, first, the nearest \(p\text{-}D\) model point is identified for the new \(p\text{-}D\) data point by computing \(p\text{-}D\) Euclidean distance.
-Then, the corresponding \(2\text{-}D\) bin centroid mapping for the identified \(p\text{-}D\) model point is determined.
-Finally, the coordinates of the identified \(2\text{-}D\) bin centroid is used as the predicted NLDR embedding for the new \(p\text{-}D\) data point.
+The \texttt{predict\_emb()} function is used to predict a point in a $2\text{-}D$ embedding for a new $p\text{-}D$ data point using the fitted model.
+This function is useful to predict $2\text{-}D$ embedding irrespective of the NLDR technique.
+
+In the prediction process, first, the nearest $p\text{-}D$ model point is identified for the new $p\text{-}D$ data point by computing $p\text{-}D$ Euclidean distance.
+Then, the corresponding $2\text{-}D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
+Finally, the coordinates of the identified $2\text{-}D$ bin centroid is used as the predicted NLDR embedding for the new $p\text{-}D$ data point.
 
@@
 
-Furthermore, \texttt{augment()} accepts \(2\text{-}D\) and \(p\text{-}D\) model points, and the \(p\text{-}D\) data and adds information about each observation in the data set.
+Furthermore, \texttt{augment()} accepts $2\text{-}D$ and $p\text{-}D$ model points, and the $p\text{-}D$ data and adds information about each observation in the data set.
 
@@
 
-The package offers several \(2\text{-}D\) visualizations, including:
+The package offers several $2\text{-}D$ visualizations, including:
 
@@
 
-The generated \(p\text{-}D\) model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
-Additionally, it features a function for visualizing the \(2\text{-}D\) projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
+The generated $p\text{-}D$ model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
+Additionally, it features a function for visualizing the $2\text{-}D$ projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
 
@@
 
-The \texttt{geom\_trimesh()} function introduces a custom \texttt{ggplot2} layer designed for visualizing \(2\text{-}D\) wireframe on a provided set of bin centroids.
+The \texttt{geom\_trimesh()} function introduces a custom \texttt{ggplot2} layer designed for visualizing $2\text{-}D$ wireframe on a provided set of bin centroids.
 
@@
 
-\paragraph{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
-
-To visualise how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the \texttt{show\_langevitour()} function.
+\paragraph{\texorpdfstring{$p\text{-}D$ model visualization}{model visualization}}\label{model-visualization}
+
+To visualise how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the \texttt{show\_langevitour()} function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.
@@
 
-In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2\text{-}D\) wireframe structure mapped back to high-dimensional space.
+In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped back to high-dimensional space.
 
@@
 
-To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
+To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
 This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
@@
 Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
-This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2\text{-}D\) bin layout (\texttt{model\_2d}):
+This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the $2\text{-}D$ bin layout (\texttt{model\_2d}):
 
@@
 
-The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
 The function takes the output from \texttt{comb\_all\_data\_model()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
@@
 Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
-The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2\text{-}D\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
+The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), $2\text{-}D$ model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
 
@@
 
-The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
 The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
@@
 Several core computations within \texttt{quollr} are optimized using compiled C++ code via the \texttt{Rcpp} and \texttt{RcppArmadillo} packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), \(2\text{-}D\) distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), $2\text{-}D$ distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
 This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better \(2\text{-}D\) layout.
+Our method help to provide a better $2\text{-}D$ layout.
 
@@
 
-The \texttt{quollr} package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in \(2\text{-}D\) and lifting it into high-dimensional space.
-This lifted model provides a direct way to assess how well a \(2\text{-}D\) layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
-The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate \(2\text{-}D\) representations.
+The \texttt{quollr} package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in $2\text{-}D$ and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a $2\text{-}D$ layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate $2\text{-}D$ representations.
 
@@
 To support efficient computation, particularly for large-scale datasets, several core functions in \texttt{quollr} are implemented in C++ using \texttt{Rcpp} and \texttt{RcppArmadillo}.
-These include functions for computing Euclidean distances in high-dimensional and \(2\text{-}D\) space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RMSE and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in \(2\text{-}D\).
+These include functions for computing Euclidean distances in high-dimensional and $2\text{-}D$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RMSE and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in $2\text{-}D$.
 Additionally, \texttt{gen\_hex\_coord\_cpp()} constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-09-18 paper-quollr.tex
Updated function outputs

@@
 
-The result is a data frame of the scaled $2\text{-}D$ embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
+The result is a \texttt{tibble} of the scaled $2\text{-}D$ embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
 

### 2025-09-18 paper-quollr.Rmd
Updated function outputs
Bib file has a problem..couldn't able to find

@@
     orcid: 0000-0002-6265-6481
-    email:  \email{jayani.piyadigamage@monash.edu}
+    email: jayani.piyadigamage@monash.edu
   - name: Dianne Cook
@@
 These small variances ensure that the added dimensions contribute noise without obscuring the main geometric structure.
-`scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors`$=46$ and `min_dist`$=0.9$.
-Each data set contains a unique ID column that maps the $7\text{-}D$ point in `scurve` to the corresponding $2\text{-}D$ `scurve_umap`.
+`scurve_umap` is the UMAP \twoD embedding for `scurve` data with `n_neighbors`$=46$ and `min_dist`$=0.9$.
+Each data set contains a unique ID column that maps the $7\text{-}D$ point in `scurve` to the corresponding \twoD `scurve_umap`.
 
@@
 
-<!-- If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step.
-For example, if the user wishes to use already binning data, the `merge_hexbin_centroids()` function can be used directly.
--->
-
 This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`hd_thresh`).
-The function returns an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both \twoD (`model_2d`), and \pD (`model_highd`), and triangular mesh (`trimesh_data`).
+The function returns an object of class `highd_vis_model` containing the scaled NLDR object (`nldr_scaled_obj`) with three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both \twoD (`model_2d`), and \pD (`model_highd`), and triangular mesh (`trimesh_data`).
 
@@
 
-To reduce computational overhead, the geometry calculations are implemented in C++ using `gen_hex_coord_cpp()`, which returns a `data.frame` of vertex coordinates.
+To reduce computational overhead, the geometry calculations are implemented in C++ using `gen_hex_coord_cpp()`, which returns a `tibble` of vertex coordinates.
 
@@
 
-The result is a data frame of the scaled \twoD embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
+The result is a `tibble` of the scaled \twoD embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
 
@@
 To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin’s count by the maximum count across all bins.
-This yields a standardized bin counts, `w_h`, ranging from 0 to 1.
+This yields a standardized bin counts, `w_h`, ranging from $0$ to $1$.
 
@@
 
-The function first groups the input data by `h`, which represents the hexagon label associated with each point in the \twoD layout.
-Within each group, it collects the `ID`s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
+The function first groups the input data by `h`, which represents the hexagon ID associated with each point in the \twoD layout.
+Within each group, it collects the `ID`s into a list, resulting in a summary where each row corresponds to a single hexagon.
 The resulting column, `pts_list`, contains all point identifiers associated with that hexagon.
@@
 This function begins by arranging the `counts_data` by `h` to ensure consistent ordering.
-It then performs a full join with `centroids_data`, aligning hexagon IDs between the two datasets to incorporate both hexagonal bin centroids and count metrics.
+It then performs a full join with `centroids_data`, aligning hexagon IDs (`h`) between the two datasets to incorporate both hexagonal bin centroids (`h`) and count metrics.
 After merging, the function handles missing values in the count columns: any `NA` values in `w_h` or `n_h` are replaced with zeros.
 This ensures that hexagons with no assigned data points are retained in the output, with zero values for count-related fields.
-The resulting data frame contains the full set of hexagon centroids along with associated bin counts and standardized counts.
+The resulting data contains the full set of hexagon centroids along with associated bin counts (`n_h`) and standardized counts (`w_h`).
 
@@
 
-To represent the neighborhood structure of hexagonal bins in a \twoD layout, we employ Delaunay triangulation on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a \twoD layout, we employ Delaunay triangulation [@lee1980;@alb2024] on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
 
-The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `tripack::tri.mesh()` function (@albrecht2025).
+The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `tripack::tri.mesh()` function [@albrecht2025].
 This triangulation forms the structural basis for identifying adjacent bins.
@@
 Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function `calc_2d_dist()`.
-Only edges within a hexagon’s neighborhood radius (based on the hexagon side length `a1`) are retained.
 
@@
 The `find_low_dens_hex()` function identifies hexagonal bins with insufficient local support by calculating the average standardized count across their six neighboring bins.
-If this mean neighborhood density is below a user-defined threshold (e.g., 0.05), the hexagon is flagged for removal.
+If this mean neighborhood density is below a user-defined threshold (e.g., $0.05$), the hexagon is flagged for removal.
 
@@
 )
-
-```
-
-For simplicity, we remove low-density hexagons using a threshold of $1$.
+```
+
+For simplicity, we remove low-density hexagons using a threshold of $0$.
 
@@
 trimesh <- update_trimesh_index(trimesh)
-
 ```
@@
 It is worth noting that while `predict_emb()` provides a general approach that works across methods, some NLDR techniques have their own built-in prediction mechanisms.
-For example, UMAP (@tomasz2023) supports direct prediction of embeddings for new data once a model is fitted.
+For example, UMAP [@tomasz2023] supports direct prediction of embeddings for new data once a model is fitted.
 
@@
 
-Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model. These metrics can be output using the  `glance()` function.
-
-This function begins by renaming the columns of the input `model_highd` data frame to avoid naming conflicts during subsequent joins.
-It then uses the `predict_emb()` function to assign each point in the high-dimensional dataset to a corresponding bin in the 2D model, producing a prediction data frame that contains both the predicted bin assignment (`pred_hb_id`) and the original observation `ID`.
-
-The function joins this prediction output with both the high-dimensional model (to get mean bin coordinates in the original space) and the original high-dimensional data (to retrieve true coordinates).
-It then calculates squared differences between the true and predicted high-dimensional coordinates for each dimension, storing these as `error_square_x1`, `error_square_x2`, ..., up to the dimensionality of the data.
+Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model.
+These metrics can be computed using the  `glance()` function, which provides a tidy output for evaluation.
+
+The function requires both the fitted model object returned by `fit_highd_model()` and \pD data to begin.
+The \pD model output (`model_highd`) is first renamed to avoid naming conflicts during subsequent data joins.
+It then uses the `predict_emb()` function to assign each point in the \pD dataset to a corresponding hexagon bin in the \twoD model, producing a prediction data frame that contains both the predicted bin assignment (`pred_h`) and the original observation `ID`.
+
+The function joins this prediction output with both the \pD model and the \pD data (to retrieve true coordinates).
+It then calculates squared differences between the original and predicted \pD coordinates for each dimension, storing these as `error_square_x1`, `error_square_x2`, ..., up to the dimensionality of the data.
 
@@
 
-Furthermore, `augment()` accepts \twoD and \pD model points, and the \pD data and adds information about each observation in the data set.
-
-The function starts with the same process as is used in the `glance()` function to produce a predicted point in \pD for each point in the high-dimensional dataset.
+Furthermore, `augment()` requires both the fitted model object returned by `fit_highd_model()` and \pD data to begin.
+It extends the fitted model by adding prediction results and error diagnostics to the original \pD data.
+
+The function starts with the same process as is used in the `glance()` function to produce a predicted point in \pD for each point in the \pD dataset.
 
@@
 
-The final output is a data frame that combines the original IDs, high-dimensional coordinates, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
+The final output is a data frame that combines the original IDs, high-dimensional data, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
 The augmented dataset is always returned as a `tibble::tibble` with the same number of rows as the passed dataset.
@@
 
-To visualise how well the \pD model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \pD using the `show_langevitour()` function.
-This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.
+To visualize how well the \pD model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \pD using the `show_langevitour()` function.
+This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` R package [@paul2023].
 

### 2025-09-18 paper-quollr.tex
Bib file has a problem..couldn't able to find

@@
 These small variances ensure that the added dimensions contribute noise without obscuring the main geometric structure.
-\texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors}\(=46\) and \texttt{min\_dist}\(=0.9\).
-Each data set contains a unique ID column that maps the \(7\text{-}D\) point in \texttt{scurve} to the corresponding \(2\text{-}D\) \texttt{scurve\_umap}.
+\texttt{scurve\_umap} is the UMAP $2\text{-}D$ embedding for \texttt{scurve} data with \texttt{n\_neighbors}\(=46\) and \texttt{min\_dist}\(=0.9\).
+Each data set contains a unique ID column that maps the \(7\text{-}D\) point in \texttt{scurve} to the corresponding $2\text{-}D$ \texttt{scurve\_umap}.
 
@@
 This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{hd\_thresh}).
-The function returns an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both $2\text{-}D$ (\texttt{model\_2d}), and $p\text{-}D$ (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
+The function returns an object of class \texttt{highd\_vis\_model} containing the scaled NLDR object (\texttt{nldr\_scaled\_obj}) with three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both $2\text{-}D$ (\texttt{model\_2d}), and $p\text{-}D$ (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
 
@@
 
-To reduce computational overhead, the geometry calculations are implemented in C++ using \texttt{gen\_hex\_coord\_cpp()}, which returns a \texttt{data.frame} of vertex coordinates.
+To reduce computational overhead, the geometry calculations are implemented in C++ using \texttt{gen\_hex\_coord\_cpp()}, which returns a \texttt{tibble} of vertex coordinates.
 
@@
 To enable comparisons across bins with varying densities, the function then standardizes these counts by dividing each bin's count by the maximum count across all bins.
-This yields a standardized bin counts, \texttt{w\_h}, ranging from 0 to 1.
+This yields a standardized bin counts, \texttt{w\_h}, ranging from \(0\) to \(1\).
 
@@
 
-The function first groups the input data by \texttt{h}, which represents the hexagon label associated with each point in the $2\text{-}D$ layout.
-Within each group, it collects the \texttt{ID}s into a list, resulting in a summary data frame where each row corresponds to a single hexagon.
+The function first groups the input data by \texttt{h}, which represents the hexagon ID associated with each point in the $2\text{-}D$ layout.
+Within each group, it collects the \texttt{ID}s into a list, resulting in a summary where each row corresponds to a single hexagon.
 The resulting column, \texttt{pts\_list}, contains all point identifiers associated with that hexagon.
@@
 This function begins by arranging the \texttt{counts\_data} by \texttt{h} to ensure consistent ordering.
-It then performs a full join with \texttt{centroids\_data}, aligning hexagon IDs between the two datasets to incorporate both hexagonal bin centroids and count metrics.
+It then performs a full join with \texttt{centroids\_data}, aligning hexagon IDs (\texttt{h}) between the two datasets to incorporate both hexagonal bin centroids (\texttt{h}) and count metrics.
 After merging, the function handles missing values in the count columns: any \texttt{NA} values in \texttt{w\_h} or \texttt{n\_h} are replaced with zeros.
 This ensures that hexagons with no assigned data points are retained in the output, with zero values for count-related fields.
-The resulting data frame contains the full set of hexagon centroids along with associated bin counts and standardized counts.
+The resulting data contains the full set of hexagon centroids along with associated bin counts (\texttt{n\_h}) and standardized counts (\texttt{w\_h}).
 
@@
 
-To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation \citep{lee1980, alb2024} on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
 
-The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{tripack::tri.mesh()} function (\citet{albrecht2025}).
+The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{tripack::tri.mesh()} function \citep{albrecht2025}.
 This triangulation forms the structural basis for identifying adjacent bins.
@@
 Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function \texttt{calc\_2d\_dist()}.
-Only edges within a hexagon's neighborhood radius (based on the hexagon side length \texttt{a1}) are retained.
 
@@
 The \texttt{find\_low\_dens\_hex()} function identifies hexagonal bins with insufficient local support by calculating the average standardized count across their six neighboring bins.
-If this mean neighborhood density is below a user-defined threshold (e.g., 0.05), the hexagon is flagged for removal.
+If this mean neighborhood density is below a user-defined threshold (e.g., \(0.05\)), the hexagon is flagged for removal.
 
@@
 \begin{verbatim}
->   [1]   1   2   3   4   5   6   7   8   9  10  11  12  13  14  15  16  17  18
->  [19]  19  20  21  22  23  24  25  26  27  28  29  30  31  32  33  34  35  36
->  [37]  37  38  39  40  41  42  43  44  45  46  47  48  49  50  51  52  53  54
->  [55]  55  56  57  58  59  60  61  62  63  64  65  66  67  68  69  70  71  72
->  [73]  73  74  75  76  77  78  79  80  81  82  83  84  85  86  87  88  89  90
->  [91]  91  92  93  94  95  96  97  98  99 100 101 102 103 104 105 106 107 108
-> [109] 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126
-> [127] 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144
-> [145] 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162
-> [163] 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180
-> [181] 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198
-> [199] 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215 216
-> [217] 217 218 219 220 221 222 223 224 225 226 227 228 229 230 231 232 233 234
-> [235] 235 236 237 238 239 240 241 242 243 244 245 246 247 248 249 250 251 252
-> [253] 253 254 255 256 257 258 259 260 261 262 263 264 265 266 267 268 269 270
-> [271] 271 272 273 274 275 276 277 278 279 280 281 282 283 284 285 286 287 288
-> [289] 289 290 291 292 293 294 295 296 297 298 299 300 301 302 303 304 305 306
-> [307] 307 308 309 310 311 312 313 314 315 316 317 318 319 320 321 322 323 324
-> [325] 325 326 327 328 329 330 331 332 333 334 335 336 337 338 339 340 341 342
-> [343] 343 344 345 346 347 348 349 350 351 352 353 354 355 356 357 358 359 360
-> [361] 361 362 363 364 365 366 367 368 369 370 371 372 373 374 375 376 377 378
-> [379] 379 380 381 382 383 384 385 386 387 388 389 390 391 392 393 394 395 396
-> [397] 397 398 399 400 401 402 403 404 405 406 407 408 409 410 411 412 413 414
-> [415] 415 416 417 418 419 420 421 422 423 424 425 426 427 428 429 430 431 432
-> [433] 433 434 435 436 437 438 439 440 441 442 443 444 445 446 447 448 449 450
-> [451] 451 452 453 454 455 456 457 458 459 460 461 462 463 464 465 466 467 468
-> [469] 469 470 471 472 473 474 475 476 477 478 479 480 481 482 483 484 485 486
-> [487] 487 488 489 490 491 492 493 494 495 496 497 498 499 500 501 502 503 504
-> [505] 505 506 507 508 509 510 511 512 513 514 515 516 517 518 519 520 521 522
-> [523] 523 524 525 526 527 528 529 530 531 532 533 534 535 536 537 538 539 540
-> [541] 541 542 543 544 545 546 547 548 549 550 551 552 553 554 555 556 557 558
-> [559] 559 560 561 562 563 564 565 566 567 568 569 570 571 572 573 574 575 576
-> [577] 577 578 579 580 581 582 583 584 585 586 587 588 589 590 591 592 593 594
-> [595] 595 596 597 598 599 600 601 602 603 604 605 606 607 608 609 610 611 612
-> [613] 613 614 615 616 617 618 619 620 621 622 623 624 625 626 627 628 629 630
-> [631] 631 632 633 634 635 636 637 638 639 640 641 642 643 644 645 646 647 648
-> [649] 649 650 651 652 653 654 655 656 657 658 659 660 661 662 663 664 665 666
-> [667] 667 668 669 670 671 672 673 674 675 676 677 678 679 680 681 682 683 684
-> [685] 685 686 687 688 689 690 691 692 693 694 695 696 697 698 699 700 701 702
-> [703] 703 704 705 706 707 708 709 710 711 712 713 714 715 716 717 718 719 720
-> [721] 721 722 723 724 725 726 727 728 729 730 731 732 733 734 735 736 737 738
-> [739] 739 740 741 742 743 744 745 746 747 748 749 750 751 752 753 754 755 756
-> [757] 757 758 759 760 761 762 763 764 765 766 767 768 769 770 771 772 773 774
-> [775] 775 776 777 778 779 780 781 782 783 784 785 786 787 788 789 790 791 792
-> [793] 793 794 795 796 797 798 799 800 801 802 803 804 805 806 807 808 809 810
-> [811] 811 812 813 814 815 816 817 818 819 820 821 822 823 824 825
-\end{verbatim}
-
-For simplicity, we remove low-density hexagons using a threshold of \(1\).
+>   [1]  82  83  84  85  86  93  94 107 108 109 110 111 112 113 118 119 120 131
+>  [19] 132 133 134 135 136 137 138 141 142 143 144 145 156 157 158 159 160 161
+>  [37] 162 163 164 165 166 167 168 169 170 171 181 182 183 184 185 186 187 188
+>  [55] 189 190 191 192 193 194 195 196 206 207 209 210 211 212 213 214 215 216
+>  [73] 217 218 219 220 221 222 231 232 233 234 235 236 237 239 240 241 242 243
+>  [91] 244 245 246 256 257 258 259 260 261 266 267 268 269 271 272 281 282 283
+> [109] 284 285 291 292 293 294 295 296 297 307 317 318 319 320 321 322 323 341
+> [127] 342 343 344 345 346 347 367 368 369 371 372 373 391 392 393 394 395 396
+> [145] 397 398 416 417 418 419 420 421 422 423 440 441 442 443 444 446 447 465
+> [163] 466 467 468 469 470 471 472 489 490 491 492 493 494 495 496 497 509 510
+> [181] 511 512 513 514 515 516 517 518 519 520 521 522 528 530 531 532 533 534
+> [199] 535 536 537 538 539 540 541 542 543 544 545 546 553 554 555 556 557 558
+> [217] 559 560 561 562 563 564 565 566 567 568 569 570 578 579 580 581 582 583
+> [235] 584 585 586 587 588 589 590 591 592 593 594 595 603 604 605 606 607 608
+> [253] 609 610 611 612 613 614 615 616 617 618 619 620 628 629 630 631 633 634
+> [271] 635 636 637 638 640 641 642 643 644 653 654 655 656 657 658 659 660 661
+> [289] 662 663 664 665 666 667 678 679 680 681 682 683 685 686 687 688 689 690
+> [307] 691 704 705 706 707 708 709 710 711 712 713 714 729 730 731 732 733 734
+> [325] 735 736 737 738 755 757
+\end{verbatim}
+
+For simplicity, we remove low-density hexagons using a threshold of \(0\).
 
@@
 It is worth noting that while \texttt{predict\_emb()} provides a general approach that works across methods, some NLDR techniques have their own built-in prediction mechanisms.
-For example, UMAP (\citet{tomasz2023}) supports direct prediction of embeddings for new data once a model is fitted.
+For example, UMAP \citep{tomasz2023} supports direct prediction of embeddings for new data once a model is fitted.
 
@@
 
-Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model. These metrics can be output using the \texttt{glance()} function.
-
-This function begins by renaming the columns of the input \texttt{model\_highd} data frame to avoid naming conflicts during subsequent joins.
-It then uses the \texttt{predict\_emb()} function to assign each point in the high-dimensional dataset to a corresponding bin in the 2D model, producing a prediction data frame that contains both the predicted bin assignment (\texttt{pred\_hb\_id}) and the original observation \texttt{ID}.
-
-The function joins this prediction output with both the high-dimensional model (to get mean bin coordinates in the original space) and the original high-dimensional data (to retrieve true coordinates).
-It then calculates squared differences between the true and predicted high-dimensional coordinates for each dimension, storing these as \texttt{error\_square\_x1}, \texttt{error\_square\_x2}, \ldots, up to the dimensionality of the data.
+Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model.
+These metrics can be computed using the \texttt{glance()} function, which provides a tidy output for evaluation.
+
+The function requires both the fitted model object returned by \texttt{fit\_highd\_model()} and $p\text{-}D$ data to begin.
+The $p\text{-}D$ model output (\texttt{model\_highd}) is first renamed to avoid naming conflicts during subsequent data joins.
+It then uses the \texttt{predict\_emb()} function to assign each point in the $p\text{-}D$ dataset to a corresponding hexagon bin in the $2\text{-}D$ model, producing a prediction data frame that contains both the predicted bin assignment (\texttt{pred\_h}) and the original observation \texttt{ID}.
+
+The function joins this prediction output with both the $p\text{-}D$ model and the $p\text{-}D$ data (to retrieve true coordinates).
+It then calculates squared differences between the original and predicted $p\text{-}D$ coordinates for each dimension, storing these as \texttt{error\_square\_x1}, \texttt{error\_square\_x2}, \ldots, up to the dimensionality of the data.
 
@@
 
-Furthermore, \texttt{augment()} accepts $2\text{-}D$ and $p\text{-}D$ model points, and the $p\text{-}D$ data and adds information about each observation in the data set.
-
-The function starts with the same process as is used in the \texttt{glance()} function to produce a predicted point in $p\text{-}D$ for each point in the high-dimensional dataset.
+Furthermore, \texttt{augment()} requires both the fitted model object returned by \texttt{fit\_highd\_model()} and $p\text{-}D$ data to begin.
+It extends the fitted model by adding prediction results and error diagnostics to the original $p\text{-}D$ data.
+
+The function starts with the same process as is used in the \texttt{glance()} function to produce a predicted point in $p\text{-}D$ for each point in the $p\text{-}D$ dataset.
 
@@
 
-The final output is a data frame that combines the original IDs, high-dimensional coordinates, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
+The final output is a data frame that combines the original IDs, high-dimensional data, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
 The augmented dataset is always returned as a \texttt{tibble::tibble} with the same number of rows as the passed dataset.
@@
 
-To visualise how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the \texttt{show\_langevitour()} function.
-This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.
+To visualize how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the \texttt{show\_langevitour()} function.
+This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} R package \citep{paul2023}.
 
@@
 \textit{ORCiD: \href{https://orcid.org/0000-0002-6265-6481}{0000-0002-6265-6481}}\\%
-\email{jayani.piyadigamage@monash.edu}%
+\href{mailto:jayani.piyadigamage@monash.edu}{\nolinkurl{jayani.piyadigamage@monash.edu}}%
 }

### 2025-09-18 paper-quollr.R

@@ -736,7 +736,6 @@ find_low_dens_hex(
 )
 
 
-
 ## ----echo=TRUE----------------------------------------------------------------
 df_bin_centroids <- df_bin_centroids |>
   dplyr::filter(n_h > 0)
@@ -746,7 +745,6 @@ trimesh <- trimesh |>
                 to_count > 0)
 
 trimesh <- update_trimesh_index(trimesh)
-
 
 
 ## ----echo=TRUE----------------------------------------------------------------

### 2025-09-21 paper-quollr.Rmd
Updated function definitions for glance and augment

@@
   \newcommand{\twoD}{$2\text{-}D$}
+  \PassOptionsToPackage{draft}{hyperref}
 output: 
@@
 glance(
-  model_object = scurve_model_obj,
+  x = scurve_model_obj,
   highd_data = scurve
@@
 model_error <- augment(
-  model_object = scurve_model_obj,
+  x = scurve_model_obj,
   highd_data = scurve

### 2025-10-03 paper-quollr.Rmd
Small language change on function arguments

@@
 While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the `hex_binning()` function automates this entire workflow.
-The parameters used within `hex_binning()` include an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
+The parameters used within `hex_binning()` are the object output from `gen_scaled_data` (`nldr_scaled_obj`); the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
 The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagon IDs (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`b`), the number of non-empty bins (`m`), and the points within each hexagon (`pts_bins`).

### 2025-10-03 paper-quollr.tex
rendered with Michael's changes

@@
 While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the \texttt{hex\_binning()} function automates this entire workflow.
-The parameters used within \texttt{hex\_binning()} include an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
+The parameters used within \texttt{hex\_binning()} are the object output from \texttt{gen\_scaled\_data} (\texttt{nldr\_scaled\_obj}); the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
 The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{b}), the number of non-empty bins (\texttt{m}), and the points within each hexagon (\texttt{pts\_bins}).

### 2025-10-03 paper-quollr.tex
Updated .bib

@@
 
-This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.
+This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.
 These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.

### 2025-10-03 paper-quollr.Rmd
Updated .bib
Updated references

@@
 
-To represent the neighborhood structure of hexagonal bins in a \twoD layout, we employ Delaunay triangulation [@lee1980;@alb2024] on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a \twoD layout, we employ Delaunay triangulation [@lee1980;@albrecht2024] on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
 
-The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `tripack::tri.mesh()` function [@albrecht2025].
+The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `interp::tri.mesh()` function [@albrecht2024].
 This triangulation forms the structural basis for identifying adjacent bins.
@@
 
-This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template.
+This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.
 These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].

### 2025-10-03 paper-quollr.tex
Updated references

@@
 
-To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation \citep{lee1980, alb2024} on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation \citep{lee1980, albrecht2024} on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
 
-The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{tripack::tri.mesh()} function \citep{albrecht2025}.
+The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{interp::tri.mesh()} function \citep{albrecht2024}.
 This triangulation forms the structural basis for identifying adjacent bins.

### 2025-10-03 paper-quollr.Rmd
Removed unncessary commented chunks

@@
   comment = ">",
-  out.width = "100%")
+  out.width = "100%",
+  htmlwidgets.sizing = TRUE)
 
@@
 
-<!-- # Find the most appropriate fit -->
-
-<!-- ```{r} -->
-<!-- #| label: scurve-diff-umap-layouts -->
-
-<!-- nldr_scaled_obj <- gen_scaled_data( -->
-<!--   nldr_data = scurve_umap) -->
-<!-- umap_scaled1 <- nldr_scaled_obj$scaled_nldr -->
-
-<!-- scurve_nldr1 <- umap_scaled1 |> -->
-<!--   ggplot(aes(x = emb1, -->
-<!--              y = emb2))+ -->
-<!--   geom_point(alpha=0.1, size=1, colour='#e41a1c') + -->
-<!--   interior_annotation("a", c(0.08, 0.9)) -->
-
-<!-- nldr_scaled_obj2 <- gen_scaled_data( -->
-<!--   nldr_data = scurve_umap2) -->
-<!-- umap_scaled2 <- nldr_scaled_obj2$scaled_nldr -->
-
-<!-- scurve_nldr2 <- umap_scaled2 |> -->
-<!--   ggplot(aes(x = emb1, -->
-<!--              y = emb2))+ -->
-<!--   geom_point(alpha=0.1, size=1, colour='#ff7f00') + -->
-<!--   interior_annotation("b") -->
-
-<!-- nldr_scaled_obj3 <- gen_scaled_data( -->
-<!--   nldr_data = scurve_umap3) -->
-<!-- umap_scaled3 <- nldr_scaled_obj3$scaled_nldr -->
-
-<!-- scurve_nldr3 <- umap_scaled3 |> -->
-<!--   ggplot(aes(x = emb1, -->
-<!--              y = emb2))+ -->
-<!--   geom_point(alpha=0.1, size=1, colour='#4daf4a') + -->
-<!--   interior_annotation("c", c(0.08, 0.9)) -->
-
-<!-- nldr_scaled_obj4 <- gen_scaled_data( -->
-<!--   nldr_data = scurve_umap4) -->
-<!-- umap_scaled4 <- nldr_scaled_obj4$scaled_nldr -->
-
-<!-- scurve_nldr4 <- umap_scaled4 |> -->
-<!--   ggplot(aes(x = emb1, -->
-<!--              y = emb2))+ -->
-<!--   geom_point(alpha=0.1, size=1, colour='#a65628') + -->
-<!--   interior_annotation("d") -->
-
-<!-- nldr_scaled_obj5 <- gen_scaled_data( -->
-<!--   nldr_data = scurve_umap5) -->
-<!-- umap_scaled5 <- nldr_scaled_obj5$scaled_nldr -->
-
-<!-- scurve_nldr5 <- umap_scaled5 |> -->
-<!--   ggplot(aes(x = emb1, -->
-<!--              y = emb2))+ -->
-<!--   geom_point(alpha=0.1, size=1, colour='#636363') + -->
-<!--   interior_annotation("e") -->
-
-<!-- nldr_scaled_obj6 <- gen_scaled_data( -->
-<!--   nldr_data = scurve_umap6) -->
-<!-- umap_scaled6 <- nldr_scaled_obj6$scaled_nldr -->
-
-<!-- scurve_nldr6 <- umap_scaled6 |> -->
-<!--   ggplot(aes(x = emb1, -->
-<!--              y = emb2))+ -->
-<!--   geom_point(alpha=0.1, size=1, colour='#984ea3') + -->
-<!--   interior_annotation("f") -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- scurve_error <- dplyr::bind_rows(scurve_umap_mse, -->
-<!--                                  scurve_umap_mse2, -->
-<!--                                  scurve_umap_mse3, -->
-<!--                                  scurve_umap_mse4, -->
-<!--                                  scurve_umap_mse5, -->
-<!--                                  scurve_umap_mse6) -->
-
-<!-- scurve_error <- scurve_error |> -->
-<!--   mutate(a1 = round(a1, 2)) |> -->
-<!--   filter(b1 >= 5) |> -->
-<!--   group_by(method, a1) |> -->
-<!--   filter(MSE == min(MSE)) |> -->
-<!--   ungroup() -->
-
-<!-- error_plot <- ggplot(scurve_error, -->
-<!--          aes(x = a1, -->
-<!--              y = sqrt(MSE), -->
-<!--              color = method)) + -->
-<!--     geom_point(size = 0.8) + -->
-<!--     geom_line(linewidth = 0.3) + -->
-<!--   scale_x_continuous(breaks = -->
-<!--     sort(unique(scurve_error$a1))[ -->
-<!--       seq(1, length( -->
-<!--         unique(scurve_error$a1)),  -->
-<!--         by = 5)]) + -->
-<!--   scale_color_manual( -->
-<!--     values=c('#e41a1c','#ff7f00','#4daf4a',  -->
-<!--              "#a65628",'#636363', '#984ea3'))+ -->
-<!--     ylab("RMSE") + -->
-<!--     xlab(expression(paste("binwidth (", a[1], ")"))) + -->
-<!--     theme_minimal() + -->
-<!--     theme(panel.border = element_rect(fill = 'transparent'), -->
-<!--           plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5), -->
-<!--           axis.ticks.x = element_line(), -->
-<!--           axis.ticks.y = element_line(), -->
-<!--           legend.position = "none", -->
-<!--           axis.text.x = element_text(size = 7), -->
-<!--           axis.text.y = element_text(size = 7), -->
-<!--           axis.title.x = element_text(size = 7), -->
-<!--           axis.title.y = element_text(size = 7), -->
-<!--           aspect.ratio = 1.5) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- free(error_plot) + wrap_plots( -->
-<!--   scurve_nldr1, scurve_nldr2,  -->
-<!--   scurve_nldr3, scurve_nldr4, -->
-<!--   scurve_nldr5, scurve_nldr6,  -->
-<!--   ncol = 2) -->
-<!-- ``` -->
-
 # Computational efficiency and optimization

### 2025-10-07 paper-quollr.Rmd
small change to installation section now being availability
Removed all the notations

@@
 
-## Installation
-
-The development version can be installed from GitHub:
-
-```r
-pak::pak("JayaniLakshika/quollr")
-```
+## Availability
+
+The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.
 
@@
 
-`scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $1000$ observations from
+The `scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $1000$ observations from
 
@@
 These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].
-
-```{r notation-tb-html, eval=knitr::is_html_output()}
-# Notation used in the paper
-
-notation_df <- read_csv("misc/notation.csv")
-
-notation_df |> 
-  kable(caption = "Summary of notation for describing new methodology.") 
-```
-
-```{r notation-tb-pdf, eval=knitr::is_latex_output()}
-
-notation_df <- read_csv("misc/notation.csv")
-
-notation_df |> 
-  kable(caption = "Summary of notation for describing new methodology.", booktabs = T) |>
-  column_spec(1, width = "3cm") |>
-  column_spec(2, width = "10cm") |>
-  column_spec(3, width = "3cm")
-```

### 2025-10-07 paper-quollr.tex
Rendered after Di's updates
Removed all the notations

@@
 
-\subsection{Installation}\label{installation}
+\subsection{Availability}\label{availability}
 
-The development version can be installed from GitHub:
-
-\begin{verbatim}
-pak::pak("JayaniLakshika/quollr")
-\end{verbatim}
+The package is available on CRAN, and the development version is available at \url{https://jayanilakshika.github.io/quollr/}.
 
@@
 
-\texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(1000\) observations from
+The \texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(1000\) observations from
 
@@
 
-\begin{table}
-
-\caption{\label{tab:notation-tb-pdf}Summary of notation for describing new methodology.}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{10cm}>{\raggedright\arraybackslash}p{3cm}}
-\toprule
-Notation & Description & Variable\\
-\midrule
-\textbackslash{}\$n, p, k\textbackslash{}\$ & number of observations, variables, embedding dimension, respectively & NA\\
-\$\textbackslash{}mathbfit\{X\}, \textbackslash{}mathbfit\{x\}\$ & \$p\$-dimensional data (population, sample) & NA\\
-\$\textbackslash{}mathbfit\{y\}\$ & \$k\$-dimensional layout & NA\\
-\$P\$ & orthonormal basis, generating a \$d\textbackslash{}text\{-\}dimensional\$ linear projection of \$p\$-dimensional data & NA\\
-\$T\$ & true  model & NA\\
-\addlinespace
-\$g\$ & functional mapping from \textbackslash{}pD\{\} to \textbackslash{}kD\{\}, especially as prescribed by NLDR & NA\\
-\$\textbackslash{}mathbfit\{\textbackslash{}theta\}\$ & (Hyper-) parameters for NLDR method & NA\\
-\$r\$ & ranges of the embedding components & NA\\
-\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers & (c\_x, c\_y)\\
-\$(b\_1, b\_2)\$ & number of bins in each direction & (b1, b2) = bins\\
-\addlinespace
-\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction & a1, a2\\
-\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid & start\_point\\
-\$q\$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1 & q\\
-\$m\$ & number of non-empty bins & m\\
-\$b\$ & number of  hexagons in the grid & b\\
-\addlinespace
-\$h\$ & hexagonal id & NA\\
-\$l\$ & side length & NA\\
-\$A\$ & area & NA\\
-\$n\_h\$ & bin count & NA\\
-\$w\_h\$ & standardised bin counts & NA\\
-\addlinespace
-NA & threshold to extract high density hexagons & hd\_thresh\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \bibliography{paper-quollr.bib}

### 2025-10-07 paper-quollr.Rmd
Updated some wordings RMSE to RWBSS

@@
 
-### Compute residuals and Root Mean Square Error (RMSE)
-
-Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model.
-These metrics can be computed using the  `glance()` function, which provides a tidy output for evaluation.
+### Compute residuals and Root Within Bin Sum of Square (RWBSS)
+
+Root Within Bin Sum of Square (RWBSS) are used as goodness of fit metrics for the model. These metrics can be computed using the  `glance()` function, which provides a tidy output for evaluation.
 
@@
 
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RMSE which is the average of the total squared error per point.
-
-These metrics are returned in a tibble as `Error` (absolute error) and `RMSE` (root mean squared error).
-The computation of total absolute error and RMSE is performed in C++ for efficiency using the internal `compute_errors()` function.
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RWBSS which is the average of the total squared error per point.
+
+These metrics are returned in a tibble as `Error` (absolute error) and `RWBSS` (root mean squared error).
+The computation of total absolute error and RWBSS is performed in C++ for efficiency using the internal `compute_errors()` function.
 
@@
 Following their pre-processing, different NLDR methods were performed using ten principal components.
-Figure \@ref(fig:limb-rmse) (b) is the reproduction of the published plot.
+Figure \@ref(fig:limb-rwbss) (b) is the reproduction of the published plot.
 The question is whether this accurately represents the cluster structure in the data.
@@
 
-```{r limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$).
+```{r limb-rwbss, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RWBSS for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.\\label{fig:limb-rmse}", fig.pos='H', echo=TRUE}
+Layout f is the best choice.\\label{fig:limb-rwbss}", fig.pos='H', echo=TRUE}
 
@@
 
-```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\ref{fig:limb-rmse}b) and the $2-\\text{D}$ layout selected (Figure \\ref{fig:limb-rmse}f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\ref{fig:limb-rwbss}b) and the $2-\\text{D}$ layout selected (Figure \\ref{fig:limb-rwbss}f) by RWBSS plot (Figure \\ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
@@
 
-```{r model-limb-html, eval=knitr::is_html_output(), fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rmse) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rmse) f) by RMSE plot (Figure \\@ref(fig:limb-rmse)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+```{r model-limb-html, eval=knitr::is_html_output(), fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rwbss) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rwbss) f) by RWBSS plot (Figure \\@ref(fig:limb-rwbss)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$."}
@@
 In contrast to the common practice of visually inspecting scatterplots for clusters or patterns, `quollr` provides a quantitative route for evaluation.
-It enables the computation of RMSE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
+It enables the computation of RWBSS and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
 These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualizations using the `langevitour` package, allowing users to inspect where the model fits well and where it does not.
@@
 These include functions for computing Euclidean distances in high-dimensional and \twoD space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RMSE and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in \twoD.
+For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RWBSS and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in \twoD.
 Additionally, `gen_hex_coord_cpp()` constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-10-07 paper-quollr.tex
Updated some wordings RMSE to RWBSS

@@
 
-\subsubsection{Compute residuals and Root Mean Square Error (RMSE)}\label{compute-residuals-and-root-mean-square-error-rmse}
+\subsubsection{Compute residuals and Root Within Bin Sum of Square (RWBSS)}\label{compute-residuals-and-root-within-bin-sum-of-square-rwbss}
 
-Residuals and root mean squared error (RMSE) are used as goodness of fit metrics for the model.
+Root Within Bin Sum of Square (RWBSS) are used as goodness of fit metrics for the model.
 These metrics can be computed using the \texttt{glance()} function, which provides a tidy output for evaluation.
@@
 
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RMSE which is the average of the total squared error per point.
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RWBSS which is the average of the total squared error per point.
 
-These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{RMSE} (root mean squared error).
-The computation of total absolute error and RMSE is performed in C++ for efficiency using the internal \texttt{compute\_errors()} function.
+These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{RWBSS} (root mean squared error).
+The computation of total absolute error and RWBSS is performed in C++ for efficiency using the internal \texttt{compute\_errors()} function.
 
@@
 Following their pre-processing, different NLDR methods were performed using ten principal components.
-Figure \ref{fig:limb-rmse} (b) is the reproduction of the published plot.
+Figure \ref{fig:limb-rwbss} (b) is the reproduction of the published plot.
 The question is whether this accurately represents the cluster structure in the data.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rwbss-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RWBSS for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.\label{fig:limb-rmse}}\label{fig:limb-rmse}
+Layout f is the best choice.\label{fig:limb-rwbss}}\label{fig:limb-rwbss}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \ref{fig:limb-rmse}b) and the $2-\text{D}$ layout selected (Figure \ref{fig:limb-rmse}f) by RMSE plot (Figure \ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \ref{fig:limb-rwbss}b) and the $2-\text{D}$ layout selected (Figure \ref{fig:limb-rwbss}f) by RWBSS plot (Figure \ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}
@@
 In contrast to the common practice of visually inspecting scatterplots for clusters or patterns, \texttt{quollr} provides a quantitative route for evaluation.
-It enables the computation of RMSE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
+It enables the computation of RWBSS and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
 These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualizations using the \texttt{langevitour} package, allowing users to inspect where the model fits well and where it does not.
@@
 These include functions for computing Euclidean distances in high-dimensional and $2\text{-}D$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RMSE and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in $2\text{-}D$.
+For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RWBSS and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in $2\text{-}D$.
 Additionally, \texttt{gen\_hex\_coord\_cpp()} constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-10-07 paper-quollr.R

@@ -1310,7 +1310,7 @@ error_plot_limb <- plot_rmse(error_limb) +
 
 
 
-## ----limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RMSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases. Layout f is the best choice.\\label{fig:limb-rmse}", fig.pos='H', echo=TRUE----
+## ----limb-rwbss, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RWBSS for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases. Layout f is the best choice.\\label{fig:limb-rwbss}", fig.pos='H', echo=TRUE----
 
 design <- gen_design(n_right = 6, ncol_right = 2)
 
@@ -1673,7 +1673,7 @@ limb_proj_tsne_best_model2 <- plot_proj(
 
 
 
-## ----model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\ref{fig:limb-rmse}b) and the $2-\\text{D}$ layout selected (Figure \\ref{fig:limb-rmse}f) by RMSE plot (Figure \\ref{fig:limb-rmse}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters. The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()----
[…]

### 2025-10-07 paper-quollr.Rmd
Updated the subsections

@@
 
-# Implementation
-
-## Availability
+# Usage
 
 The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.
-
-## Usage
 
@@
 
-### Constructing the $2\text{-}D$ Model
+## Constructing the $2\text{-}D$ Model
 
@@
 
-#### Scaling the data
+### Scaling the data
 
@@
 
-#### Computing hexagon grid configuration
+### Computing hexagon grid configuration
 
@@
 
-#### Binning the data
+### Binning the data
 
@@
 
-##### Generating all possible centroids in a hexagonal grid
+#### Generating all possible centroids in a hexagonal grid
 
@@
 
-##### Creating the coordinates of the hexagons
+#### Creating the coordinates of the hexagons
 
@@
 
-##### Assigning data points to their respective hexagons
+#### Assigning data points to their respective hexagons
 
@@
 
-##### Computing the standardized number of points within each hexagon
+#### Computing the standardized number of points within each hexagon
 
@@
 
-##### Mapping the points to their corresponding hexagonal bins
+#### Mapping the points to their corresponding hexagonal bins
 
@@
 
-#### Obtaining bin centroids
+### Obtaining bin centroids
 
@@
 
-#### Indicating neighbors by line segments connecting centroids
+### Indicating neighbors by line segments connecting centroids
 
@@
 
-#### Identifying and removing low-density hexagons
+### Identifying and removing low-density hexagons
 
@@
 
-### Lifting the model into high dimensions
+## Lifting the model into high dimensions
 
@@
 
-### Prediction
+## Prediction
 
@@
 
-### Compute residuals and Root Within Bin Sum of Square (RWBSS)
+## Compute residuals and Root Within Bin Sum of Square (RWBSS)
 
@@
 
-### Visualizations
+## Visualizations
 
@@
 
-#### Hexagonal grid
+### Hexagonal grid
 
@@
 
-#### Triangular mesh
+### Triangular mesh
 
@@
 
-#### \pD model visualization
+### \pD model visualization
 
@@
 
-#### Link plots
+### Link plots
 
@@
 
-# Computational efficiency and optimization
+# Implementation
+
+## Computational efficiency and optimization
 
@@
 
+```{r}
+df_exe_limb_int <- comb_all_data_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb, 
+  model_highd = df_bin_limb, 
+  model_2d = df_bin_centroids_limb
+)
+```
+
+```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+
+tsne_link_limb <- show_link_plots(
+  point_data = df_exe_limb_int, 
+  edge_data = trimesh_data_limb, 
+  point_colour = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')
+)
+
+class(tsne_link_limb) <- c(class(tsne_link_limb), "htmlwidget")
+
+tsne_link_limb
+```
+
+```{r}
+df_exe_limb_int_best <- comb_all_data_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb2, 
+  model_highd = df_bin_limb, 
+  model_2d = df_bin_centroids_limb
+)
+```
+
+```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+
+tsne_link_limb_best <- show_link_plots(
+  point_data = df_exe_limb_int_best, 
+  edge_data = trimesh_data_limb, 
+  point_colour = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')
+)
+
+class(tsne_link_limb_best) <- c(class(tsne_link_limb_best), "htmlwidget")
+
+tsne_link_limb_best
+```
+
 # Discussion

### 2025-10-07 paper-quollr.tex
Updated the subsections

@@
 
-\section{Implementation}\label{implementation}
-
-\subsection{Availability}\label{availability}
+\section{Usage}\label{usage}
 
 The package is available on CRAN, and the development version is available at \url{https://jayanilakshika.github.io/quollr/}.
-
-\subsection{Usage}\label{usage}
 
@@
 
-\subsubsection{Main function}\label{main-function}
+\subsection{Main function}\label{main-function}
 
@@
 
-\subsubsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
+\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
 
@@
 
-\paragraph{Scaling the data}\label{scaling-the-data}
+\subsubsection{Scaling the data}\label{scaling-the-data}
 
@@
 
-\paragraph{Computing hexagon grid configuration}\label{computing-hexagon-grid-configuration}
+\subsubsection{Computing hexagon grid configuration}\label{computing-hexagon-grid-configuration}
 
@@
 
-\paragraph{Binning the data}\label{binning-the-data}
+\subsubsection{Binning the data}\label{binning-the-data}
 
@@
 
-\subparagraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
+\paragraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
 
@@
 
-\subparagraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
+\paragraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
 
@@
 
-\subparagraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
+\paragraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
 
@@
 
-\subparagraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
+\paragraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
 
@@
 
-\subparagraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
+\paragraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
 
@@
 
-\paragraph{Obtaining bin centroids}\label{obtaining-bin-centroids}
+\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
 
@@
 
-\paragraph{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
+\subsubsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
 
@@
 
-\paragraph{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
+\subsubsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
 
@@
 
-\subsubsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}
+\subsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}
 
@@
 
-\subsubsection{Prediction}\label{prediction}
+\subsection{Prediction}\label{prediction}
 
@@
 
-\subsubsection{Compute residuals and Root Within Bin Sum of Square (RWBSS)}\label{compute-residuals-and-root-within-bin-sum-of-square-rwbss}
+\subsection{Compute residuals and Root Within Bin Sum of Square (RWBSS)}\label{compute-residuals-and-root-within-bin-sum-of-square-rwbss}
 
@@
 
-\subsubsection{Visualizations}\label{visualizations}
+\subsection{Visualizations}\label{visualizations}
 
@@
 
-\paragraph{Hexagonal grid}\label{hexagonal-grid}
+\subsubsection{Hexagonal grid}\label{hexagonal-grid}
 
@@
 
-\paragraph{Triangular mesh}\label{triangular-mesh}
+\subsubsection{Triangular mesh}\label{triangular-mesh}
 
@@
 
-\paragraph{\texorpdfstring{$p\text{-}D$ model visualization}{model visualization}}\label{model-visualization}
+\subsubsection{\texorpdfstring{$p\text{-}D$ model visualization}{model visualization}}\label{model-visualization}
 
@@
 
-\paragraph{Link plots}\label{link-plots}
+\subsubsection{Link plots}\label{link-plots}
 
@@
 
-\section{Computational efficiency and optimization}\label{computational-efficiency-and-optimization}
+\section{Implementation}\label{implementation}
+
+\subsection{Computational efficiency and optimization}\label{computational-efficiency-and-optimization}
 

### 2025-10-07 paper-quollr.R

@@ -1696,3 +1696,47 @@ trimesh_limb + limb_proj_tsne_model1 +
 # 
 # modellimbfig
 
+
+## -----------------------------------------------------------------------------
+df_exe_limb_int <- comb_all_data_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb, 
+  model_highd = df_bin_limb, 
+  model_2d = df_bin_centroids_limb
+)
+
+
+## ----tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."----
+# 
+# tsne_link_limb <- show_link_plots(
+#   point_data = df_exe_limb_int,
+#   edge_data = trimesh_data_limb,
+#   point_colour = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')
+# )
+# 
+# class(tsne_link_limb) <- c(class(tsne_link_limb), "htmlwidget")
+# 
+# tsne_link_limb
+
+
+## -----------------------------------------------------------------------------
+df_exe_limb_int_best <- comb_all_data_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb2, 
+  model_highd = df_bin_limb, 
+  model_2d = df_bin_centroids_limb
+)
+
+
+## ----tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."----
+# 
+# tsne_link_limb_best <- show_link_plots(
+#   point_data = df_exe_limb_int_best,
+#   edge_data = trimesh_data_limb,
+#   point_colour = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')
+# )
+# 
+# class(tsne_link_limb_best) <- c(class(tsne_link_limb_best), "htmlwidget")
+# 
+# tsne_link_limb_best
+

### 2025-10-07 paper-quollr.Rmd
Setup to page width

@@
 
-```{r overviewhtml, eval=knitr::is_html_output(), fig.cap="algorithm", fig.pos='H'}
+```{r overviewhtml, eval=knitr::is_html_output(), fig.cap="algorithm", fig.pos='H', layout = "l-page"}
 
@@
 ```{r algo-step-html, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
-The `Scurve` data is shown."}
+The `Scurve` data is shown.", layout = "l-page"}
 
@@
 
-```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
 
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
 
@@
 
-```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
 
@@
 
-```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}
+```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
 

### 2025-10-07 paper-quollr.R

@@ -283,7 +283,7 @@ scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)]
 
 
 
-## ----overviewhtml, eval=knitr::is_html_output(), fig.cap="algorithm", fig.pos='H'----
+## ----overviewhtml, eval=knitr::is_html_output(), fig.cap="algorithm", fig.pos='H', layout = "l-page"----
 # 
 # overviewfig <- bscols(
 #   htmltools::div(
@@ -442,7 +442,7 @@ hex_grid_scurve + wrap_plots(
   ncol = 2)
 
 
-## ----algo-step-html, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions. The `Scurve` data is shown."----
+## ----algo-step-html, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions. The `Scurve` data is shown.", layout = "l-page"----
 # 
 # algofig <- crosstalk::bscols(
 #         htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr; gap: 0;",
@@ -861,7 +861,7 @@ df_exe <- comb_all_data_model(
   )
 
 
-## ----linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."----
+## ----linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"----
 # 
 # nldrdt_link <- show_link_plots(
 #   point_data = df_exe,
@@ -1123,7 +1123,7 @@ df_exe <- comb_all_data_model_error(
   )
 
 
[…]

### 2025-10-15 paper-quollr.Rmd
Putting the implementation section earlier
Start updating the usage section

@@
 
-# Algorithm
+# Implementation
+
+## Computational efficiency and optimization
+
+Several core computations within `quollr` are optimized using compiled C++ code via the `Rcpp` and `RcppArmadillo` packages.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), \twoD distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
+This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
+These C++ functions are not exported but are bundled within the package and fully accessible for inspection in the source code.
+
+# Usage
+
+The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.
 
@@
 
-# Usage
-
-The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.
-
 The following demonstration of the package's functionality assumes `quollr` has been loaded. The built-in data sets `scurve` and `scurve_umap` are also used throughout.
 
-The `scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $1000$ observations from
-
-$$
-\begin{aligned}
-\theta &\sim U(-3\pi/2, 3\pi/2), \\
-X_1 &= \sin(\theta), \\
-X_2 &\sim U(0, 2), \text{adding thickness to the S}, \\
-X_3 &= \text{sign}(\theta) \times (\cos(\theta) - 1).
-\end{aligned}
-$$
-
-The remaining variables $X_4$, $X_5$, $X_6$, and $X_7$ are included as low-magnitude noise dimensions, generated from uniform distributions with narrow ranges around zero: $X_4, X_5 \sim U(-0.02, 0.02)$, $X_6 \sim U(-0.1, 0.1)$, and $X_7 \sim U(-0.01, 0.01)$.
-These small variances ensure that the added dimensions contribute noise without obscuring the main geometric structure.
+The `scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $1000$ observations.
+
 `scurve_umap` is the UMAP \twoD embedding for `scurve` data with `n_neighbors`$=46$ and `min_dist`$=0.9$.
@@
 
-# Implementation
-
-## Computational efficiency and optimization
-
-Several core computations within `quollr` are optimized using compiled C++ code via the `Rcpp` and `RcppArmadillo` packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), \twoD distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
-This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
-These C++ functions are not exported but are bundled within the package and fully accessible for inspection in the source code.
-
 # Application

### 2025-10-15 paper-quollr.tex
Putting the implementation section earlier
Start updating the usage section

@@
 
-\section{Algorithm}\label{algorithm}
+\section{Implementation}\label{implementation}
+
+\subsection{Computational efficiency and optimization}\label{computational-efficiency-and-optimization}
+
+Several core computations within \texttt{quollr} are optimized using compiled C++ code via the \texttt{Rcpp} and \texttt{RcppArmadillo} packages.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), $2\text{-}D$ distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
+This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
+These C++ functions are not exported but are bundled within the package and fully accessible for inspection in the source code.
+
+\section{Usage}\label{usage}
+
+The package is available on CRAN, and the development version is available at \url{https://jayanilakshika.github.io/quollr/}.
 
@@
 
-\section{Usage}\label{usage}
-
-The package is available on CRAN, and the development version is available at \url{https://jayanilakshika.github.io/quollr/}.
-
 The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. The built-in data sets \texttt{scurve} and \texttt{scurve\_umap} are also used throughout.
 
-The \texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(1000\) observations from
+The \texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(1000\) observations.
 
-\[
-\begin{aligned}
-\theta &\sim U(-3\pi/2, 3\pi/2), \\
-X_1 &= \sin(\theta), \\
-X_2 &\sim U(0, 2), \text{adding thickness to the S}, \\
-X_3 &= \text{sign}(\theta) \times (\cos(\theta) - 1).
-\end{aligned}
-\]
-
-The remaining variables \(X_4\), \(X_5\), \(X_6\), and \(X_7\) are included as low-magnitude noise dimensions, generated from uniform distributions with narrow ranges around zero: \(X_4, X_5 \sim U(-0.02, 0.02)\), \(X_6 \sim U(-0.1, 0.1)\), and \(X_7 \sim U(-0.01, 0.01)\).
-These small variances ensure that the added dimensions contribute noise without obscuring the main geometric structure.
 \texttt{scurve\_umap} is the UMAP $2\text{-}D$ embedding for \texttt{scurve} data with \texttt{n\_neighbors}\(=46\) and \texttt{min\_dist}\(=0.9\).
@@
 
-\section{Implementation}\label{implementation}
-
-\subsection{Computational efficiency and optimization}\label{computational-efficiency-and-optimization}
-
-Several core computations within \texttt{quollr} are optimized using compiled C++ code via the \texttt{Rcpp} and \texttt{RcppArmadillo} packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), $2\text{-}D$ distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
-This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
-These C++ functions are not exported but are bundled within the package and fully accessible for inspection in the source code.
-
 \section{Application}\label{application}

### 2025-10-15 paper-quollr.Rmd
Updated Usage section

@@
 
-The following demonstration of the package's functionality assumes `quollr` has been loaded. The built-in data sets `scurve` and `scurve_umap` are also used throughout.
-
-The `scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $1000$ observations.
-
-`scurve_umap` is the UMAP \twoD embedding for `scurve` data with `n_neighbors`$=46$ and `min_dist`$=0.9$.
-Each data set contains a unique ID column that maps the $7\text{-}D$ point in `scurve` to the corresponding \twoD `scurve_umap`.
+The following demonstration of the package's functionality assumes `quollr` has been loaded.
+To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
+The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
+
+To illustrate the workflow, we use the built-in example dataset `scurve`, a $7\text{-}D$ simulated dataset consisting of $1000$ observations.
+The first three variables define a $3\text{-}D$ S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.
+
+```{r, echo=TRUE}
+data("scurve")
+```
+
+For this example, we use the UMAP layout, which is produced using `n_neighbors =` $46$ and `min_dist =` $0.9$.
+
+```{r, echo=TRUE}
+library(umap)
+library(dplyr)
+
+scurve_umap <- umap(
+  scurve |> select(-ID),
+  config = modifyList(umap.defaults, list(
+    n_neighbors = 46,
+    n_components = 2,
+    min_dist = 0.9
+  ))
+)$layout |>
+  as_tibble(.name_repair = ~ paste0("emb", 1:2)) |>
+  mutate(ID = scurve$ID)
+```
 

### 2025-10-15 paper-quollr.tex
Updated Usage section

@@
 
-The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. The built-in data sets \texttt{scurve} and \texttt{scurve\_umap} are also used throughout.
+The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded.
+To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
+The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
 
-The \texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(1000\) observations.
+To illustrate the workflow, we use the built-in example dataset \texttt{scurve}, a \(7\text{-}D\) simulated dataset consisting of \(1000\) observations.
+The first three variables define a \(3\text{-}D\) S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.
 
-\texttt{scurve\_umap} is the UMAP $2\text{-}D$ embedding for \texttt{scurve} data with \texttt{n\_neighbors}\(=46\) and \texttt{min\_dist}\(=0.9\).
-Each data set contains a unique ID column that maps the \(7\text{-}D\) point in \texttt{scurve} to the corresponding $2\text{-}D$ \texttt{scurve\_umap}.
+\begin{verbatim}
+data("scurve")
+\end{verbatim}
+
+For this example, we use the UMAP layout, which is produced using \texttt{n\_neighbors\ =} \(46\) and \texttt{min\_dist\ =} \(0.9\).
+
+\begin{verbatim}
+library(umap)
+library(dplyr)
+
+scurve_umap <- umap(
+  scurve |> select(-ID),
+  config = modifyList(umap.defaults, list(
+    n_neighbors = 46,
+    n_components = 2,
+    min_dist = 0.9
+  ))
+)$layout |>
+  as_tibble(.name_repair = ~ paste0("emb", 1:2)) |>
+  mutate(ID = scurve$ID)
+\end{verbatim}
 
@@
 >    <dbl> <dbl> <int>
->  1 0.277 0.913     1
->  2 0.697 0.538     2
->  3 0.779 0.399     3
->  4 0.173 0.953     4
->  5 0.218 0.983     5
->  6 0.593 1.05      6
->  7 0.180 0.210     7
->  8 0.976 0.571     8
->  9 0.803 0.829     9
-> 10 0.932 0.410    10
+>  1 0.300 0.985     1
+>  2 0.721 0.688     2
+>  3 0.722 0.542     3
+>  4 0.196 1.01      4
+>  5 0.240 1.06      5
+>  6 0.621 1.19      6
+>  7 0.159 0.160     7
+>  8 0.979 0.661     8
+>  9 0.870 0.963     9
+> 10 0.885 0.501    10
 > # i 990 more rows
@@
 > $lim1
-> [1] -9.146398  8.552211
+> [1] -8.320406  8.522628
 > 
 > $lim2
-> [1] -10.36686  10.10691
+> [1] -11.732897   9.679631
 \end{verbatim}
@@
 > $b2
-> [1] 33
+> [1] 36
 > 
 > $a1
-> [1] 0.04952516
+> [1] 0.04943533
 > 
 > $a2
-> [1] 0.04289005
+> [1] 0.04281225
 \end{verbatim}
@@
 \begin{verbatim}
-> # A tibble: 825 x 3
+> # A tibble: 900 x 3
 >        h      c_x    c_y
 >    <int>    <dbl>  <dbl>
->  1     1 -0.1      -0.116
->  2     2 -0.0505   -0.116
->  3     3 -0.000950 -0.116
->  4     4  0.0486   -0.116
->  5     5  0.0981   -0.116
->  6     6  0.148    -0.116
->  7     7  0.197    -0.116
->  8     8  0.247    -0.116
->  9     9  0.296    -0.116
-> 10    10  0.346    -0.116
-> # i 815 more rows
+>  1     1 -0.1     -0.127
+>  2     2 -0.0506  -0.127
+>  3     3 -0.00113 -0.127
+>  4     4  0.0483  -0.127
+>  5     5  0.0977  -0.127
+>  6     6  0.147   -0.127
+>  7     7  0.197   -0.127
+>  8     8  0.246   -0.127
+>  9     9  0.295   -0.127
+> 10    10  0.345   -0.127
+> # i 890 more rows
 \end{verbatim}
@@
 >   h           x           y
-> 1 1 -0.10000000 -0.08708678
-> 2 1 -0.12476258 -0.10138346
-> 3 1 -0.12476258 -0.12997683
-> 4 1 -0.10000000 -0.14427351
-> 5 1 -0.07523742 -0.12997683
+> 1 1 -0.10000000 -0.09858838
+> 2 1 -0.12471766 -0.11285913
+> 3 1 -0.12471766 -0.14140063
+> 4 1 -0.10000000 -0.15567138
+> 5 1 -0.07528234 -0.14140063
 \end{verbatim}
@@
 >   <dbl> <dbl> <int> <int>
-> 1 0.277 0.913     1   609
-> 2 0.697 0.538     2   392
-> 3 0.779 0.399     3   319
-> 4 0.173 0.953     4   631
-> 5 0.218 0.983     5   657
+> 1 0.300 0.985     1   659
+> 2 0.721 0.688     2   492
+> 3 0.722 0.542     3   418
+> 4 0.196 1.01      4   657
+> 5 0.240 1.06      5   708
 \end{verbatim}
@@
 >   <int> <int> <dbl>
-> 1    82     1 0.001
-> 2    83     4 0.004
-> 3    84     1 0.001
-> 4    85     3 0.003
-> 5    86     1 0.001
+> 1    82     2 0.002
+> 2    83     1 0.001
+> 3   106     1 0.001
+> 4   107     5 0.005
+> 5   108     7 0.007
 \end{verbatim}
@@
 >   <int> <list>   
-> 1    82 <int [1]>
-> 2    83 <int [4]>
-> 3    84 <int [1]>
-> 4    85 <int [3]>
-> 5    86 <int [1]>
+> 1    82 <int [2]>
+> 2    83 <int [1]>
+> 3   106 <int [1]>
+> 4   107 <int [5]>
+> 5   108 <int [7]>
 \end{verbatim}
@@
 >   h          c_x        c_y n_h w_h
-> 1 1 -0.1000000000 -0.1156801   0   0
-> 2 2 -0.0504748374 -0.1156801   0   0
-> 3 3 -0.0009496749 -0.1156801   0   0
-> 4 4  0.0485754877 -0.1156801   0   0
-> 5 5  0.0981006503 -0.1156801   0   0
+> 1 1 -0.100000000 -0.1271299   0   0
+> 2 2 -0.050564674 -0.1271299   0   0
+> 3 3 -0.001129349 -0.1271299   0   0
+> 4 4  0.048305977 -0.1271299   0   0
+> 5 5  0.097741302 -0.1271299   0   0
 \end{verbatim}
@@
 >   <int> <int>    <dbl>   <dbl>    <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1      -0.116  -0.0505 -0.116           0        0
-> 2    26    27 -0.0752   -0.0728 -0.0257 -0.0728          0        0
-> 3     3    27 -0.000950 -0.116  -0.0257 -0.0728          0        0
-> 4    26    52 -0.0752   -0.0728 -0.0505 -0.0299          0        0
-> 5    27    28 -0.0257   -0.0728  0.0238 -0.0728          0        0
+> 1     1     2 -0.1     -0.127  -0.0506  -0.127           0        0
+> 2    26    27 -0.0753  -0.0843 -0.0258  -0.0843          0        0
+> 3     3    27 -0.00113 -0.127  -0.0258  -0.0843          0        0
+> 4    26    52 -0.0753  -0.0843 -0.0506  -0.0415          0        0
+> 5    52    53 -0.0506  -0.0415 -0.00113 -0.0415          0        0
 \end{verbatim}
@@
 >   <int> <int>    <dbl>   <dbl>    <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1      -0.116  -0.0505 -0.116           0        0
-> 2    26    27 -0.0752   -0.0728 -0.0257 -0.0728          0        0
-> 3     3    27 -0.000950 -0.116  -0.0257 -0.0728          0        0
-> 4    26    52 -0.0752   -0.0728 -0.0505 -0.0299          0        0
-> 5    27    28 -0.0257   -0.0728  0.0238 -0.0728          0        0
+> 1     1     2 -0.1     -0.127  -0.0506  -0.127           0        0
+> 2    26    27 -0.0753  -0.0843 -0.0258  -0.0843          0        0
+> 3     3    27 -0.00113 -0.127  -0.0258  -0.0843          0        0
+> 4    26    52 -0.0753  -0.0843 -0.0506  -0.0415          0        0
+> 5    52    53 -0.0506  -0.0415 -0.00113 -0.0415          0        0
 > # i 2 more variables: from_reindexed <int>, to_reindexed <int>
@@
 \begin{verbatim}
->   [1]  82  83  84  85  86  93  94 107 108 109 110 111 112 113 118 119 120 131
->  [19] 132 133 134 135 136 137 138 141 142 143 144 145 156 157 158 159 160 161
->  [37] 162 163 164 165 166 167 168 169 170 171 181 182 183 184 185 186 187 188
->  [55] 189 190 191 192 193 194 195 196 206 207 209 210 211 212 213 214 215 216
->  [73] 217 218 219 220 221 222 231 232 233 234 235 236 237 239 240 241 242 243
->  [91] 244 245 246 256 257 258 259 260 261 266 267 268 269 271 272 281 282 283
-> [109] 284 285 291 292 293 294 295 296 297 307 317 318 319 320 321 322 323 341
-> [127] 342 343 344 345 346 347 367 368 369 371 372 373 391 392 393 394 395 396
-> [145] 397 398 416 417 418 419 420 421 422 423 440 441 442 443 444 446 447 465
-> [163] 466 467 468 469 470 471 472 489 490 491 492 493 494 495 496 497 509 510
-> [181] 511 512 513 514 515 516 517 518 519 520 521 522 528 530 531 532 533 534
-> [199] 535 536 537 538 539 540 541 542 543 544 545 546 553 554 555 556 557 558
-> [217] 559 560 561 562 563 564 565 566 567 568 569 570 578 579 580 581 582 583
-> [235] 584 585 586 587 588 589 590 591 592 593 594 595 603 604 605 606 607 608
-> [253] 609 610 611 612 613 614 615 616 617 618 619 620 628 629 630 631 633 634
-> [271] 635 636 637 638 640 641 642 643 644 653 654 655 656 657 658 659 660 661
-> [289] 662 663 664 665 666 667 678 679 680 681 682 683 685 686 687 688 689 690
-> [307] 691 704 705 706 707 708 709 710 711 712 713 714 729 730 731 732 733 734
-> [325] 735 736 737 738 755 757
+>   [1]  82  83 106 107 108 109 131 132 133 134 135 156 157 158 159 160 161 162
+>  [19] 179 180 181 183 184 185 186 188 189 190 204 205 206 207 208 209 210 211
+>  [37] 214 215 216 217 228 229 230 231 232 233 234 235 236 238 239 240 241 242
+>  [55] 254 255 256 257 258 259 260 261 263 264 265 266 267 268 269 279 280 281
+>  [73] 283 284 285 286 287 288 289 290 293 294 305 306 307 308 309 310 311 312
+>  [91] 313 314 315 316 317 318 319 320 321 331 332 333 334 335 336 337 339 340
+> [109] 341 342 344 345 346 358 359 360 361 362 363 364 365 366 367 369 370 371
+> [127] 372 386 387 388 389 390 391 392 393 394 395 396 414 415 416 417 418 419
+> [145] 420 421 422 440 441 442 443 444 445 446 447 467 468 469 470 471 472 473
+> [163] 491 492 493 494 495 496 497 498 518 519 520 522 523 541 542 543 544 545
+> [181] 546 547 548 554 558 560 566 567 568 569 570 571 572 573 578 579 580 581
+> [199] 582 583 584 585 586 587 588 590 591 592 593 594 595 596 597 598 603 604
+> [217] 605 606 607 608 609 610 611 612 613 614 615 616 617 618 619 620 621 622
+> [235] 623 629 630 631 633 634 635 636 637 638 639 640 641 642 643 644 645 646
+> [253] 647 653 654 655 656 657 658 659 660 661 662 663 664 665 666 667 668 669
+> [271] 670 671 672 678 679 680 681 682 683 684 685 686 687 688 689 690 691 692
+> [289] 693 694 695 696 703 704 705 706 707 708 709 710 711 712 713 714 716 717
+> [307] 718 719 720 721 728 729 730 731 732 733 734 736 737 738 739 741 742 743
+> [325] 744 754 755 756 757 758 759 760 761 762 763 764 765 766 767 768 779 781
+> [343] 782 783 784 785 786 787 788 789 790 791 792 807 808 809 810 811 812 813
+> [361] 814 837
 \end{verbatim}
@@
 >   <int> <dbl> <dbl> <dbl>    <dbl>     <dbl>    <dbl>     <dbl>
-> 1    82 0.888 0.102   1.46  0.0145  -0.0118   0.0303 -0.00707
-> 2    83 0.744 0.123   1.66 -0.00950  0.00499 -0.0208 -0.00425
-> 3    84 0.591 0.0447  1.81 -0.0168   0.0173  -0.0485 -0.00112
-> 4    85 0.371 0.0937  1.93 -0.00109  0.00840  0.0268 -0.00286
-> 5    86 0.255 0.0447  1.97 -0.00862 -0.00733 -0.0826  0.00229
+> 1    82 0.989  1.78  1.13  0.00470  0.0142   -0.0547  -0.00428 
+> 2    83 0.960  1.93  1.28  0.00278 -0.0180   -0.0678  -0.00260 
+> 3   106 0.997  1.37  1.07  0.00154  0.00687   0.0289   0.00760 
+> 4   107 0.975  1.46  1.20 -0.00497 -0.000595  0.0139   0.000537
+> 5   108 0.931  1.64  1.36 -0.00706  0.00487  -0.00705  0.000383
 \end{verbatim}
@@
 >        <dbl>      <dbl> <int>  <int>
-> 1      0.296      0.914     1    609
-> 2      0.717      0.528     2    392
-> 3      0.791      0.399     3    319
-> 4      0.172      0.957     4    631
-> 5      0.197      0.999     5    657
+> 1      0.295      0.986     1    659
+> 2      0.716      0.686     2    492
+> 3      0.740      0.558     3    418
+> 4      0.197      0.986     4    657
+> 5      0.246      1.07      5    708
 \end{verbatim}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-29-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (in purple) in different projections.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-The two twists visible in the UMAP layout can also be seen in the lifted model.}\label{fig:unnamed-chunk-27}
+The two twists visible in the UMAP layout can also be seen in the lifted model.}\label{fig:unnamed-chunk-29}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-32-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
-While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-30}
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-32}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-34-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-36-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-34}
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-36}
 \end{figure}

### 2025-10-15 paper-quollr.R

@@ -470,6 +470,26 @@ scurve_umap_plt + hex_grid_poly_scurve +
   plot_layout(ncol = 5)
 
 
+## ----echo=TRUE----------------------------------------------------------------
+data("scurve")
+
+
+## ----echo=TRUE----------------------------------------------------------------
+library(umap)
+library(dplyr)
+
+scurve_umap <- umap(
+  scurve |> select(-ID),
+  config = modifyList(umap.defaults, list(
+    n_neighbors = 46,
+    n_components = 2,
+    min_dist = 0.9
+  ))
+)$layout |>
+  as_tibble(.name_repair = ~ paste0("emb", 1:2)) |>
+  mutate(ID = scurve$ID)
+
+
 ## ----echo=TRUE, eval=FALSE----------------------------------------------------
 # fit_highd_model(
 #   highd_data = scurve,

### 2025-10-15 paper-quollr.Rmd
Remove some outputs

@@
 scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
-
-scurve_umap_obj
 ```
@@
 ```{r, echo=TRUE}
-find_low_dens_hex(
+low_density_hex <- find_low_dens_hex(
   model_2d = df_bin_centroids, 
@@
   )
-
-head(model_error, 5)
 ```

### 2025-10-15 paper-quollr.tex
Remove some outputs

@@
 scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
-
-scurve_umap_obj
-\end{verbatim}
-
-\begin{verbatim}
-> $scaled_nldr
-> # A tibble: 1,000 x 3
->     emb1  emb2    ID
->    <dbl> <dbl> <int>
->  1 0.300 0.985     1
->  2 0.721 0.688     2
->  3 0.722 0.542     3
->  4 0.196 1.01      4
->  5 0.240 1.06      5
->  6 0.621 1.19      6
->  7 0.159 0.160     7
->  8 0.979 0.661     8
->  9 0.870 0.963     9
-> 10 0.885 0.501    10
-> # i 990 more rows
-> 
-> $lim1
-> [1] -8.320406  8.522628
-> 
-> $lim2
-> [1] -11.732897   9.679631
 \end{verbatim}
@@
 \begin{verbatim}
-find_low_dens_hex(
+low_density_hex <- find_low_dens_hex(
   model_2d = df_bin_centroids, 
@@
 )
-\end{verbatim}
-
-\begin{verbatim}
->   [1]  82  83 106 107 108 109 131 132 133 134 135 156 157 158 159 160 161 162
->  [19] 179 180 181 183 184 185 186 188 189 190 204 205 206 207 208 209 210 211
->  [37] 214 215 216 217 228 229 230 231 232 233 234 235 236 238 239 240 241 242
->  [55] 254 255 256 257 258 259 260 261 263 264 265 266 267 268 269 279 280 281
->  [73] 283 284 285 286 287 288 289 290 293 294 305 306 307 308 309 310 311 312
->  [91] 313 314 315 316 317 318 319 320 321 331 332 333 334 335 336 337 339 340
-> [109] 341 342 344 345 346 358 359 360 361 362 363 364 365 366 367 369 370 371
-> [127] 372 386 387 388 389 390 391 392 393 394 395 396 414 415 416 417 418 419
-> [145] 420 421 422 440 441 442 443 444 445 446 447 467 468 469 470 471 472 473
-> [163] 491 492 493 494 495 496 497 498 518 519 520 522 523 541 542 543 544 545
-> [181] 546 547 548 554 558 560 566 567 568 569 570 571 572 573 578 579 580 581
-> [199] 582 583 584 585 586 587 588 590 591 592 593 594 595 596 597 598 603 604
-> [217] 605 606 607 608 609 610 611 612 613 614 615 616 617 618 619 620 621 622
-> [235] 623 629 630 631 633 634 635 636 637 638 639 640 641 642 643 644 645 646
-> [253] 647 653 654 655 656 657 658 659 660 661 662 663 664 665 666 667 668 669
-> [271] 670 671 672 678 679 680 681 682 683 684 685 686 687 688 689 690 691 692
-> [289] 693 694 695 696 703 704 705 706 707 708 709 710 711 712 713 714 716 717
-> [307] 718 719 720 721 728 729 730 731 732 733 734 736 737 738 739 741 742 743
-> [325] 744 754 755 756 757 758 759 760 761 762 763 764 765 766 767 768 779 781
-> [343] 782 783 784 785 786 787 788 789 790 791 792 807 808 809 810 811 812 813
-> [361] 814 837
 \end{verbatim}
@@
   )
-
-head(model_error, 5)
-\end{verbatim}
-
-\begin{verbatim}
-> # A tibble: 5 x 32
->      ID      x1    x2       x3       x4       x5      x6       x7 pred_h
->   <int>   <dbl> <dbl>    <dbl>    <dbl>    <dbl>   <dbl>    <dbl>  <int>
-> 1     1 -0.120  0.819 -1.99     0.0114   0.00351  0.0334  0.00638    427
-> 2     2 -0.0492 0.166  0.00121  0.0115  -0.0166  -0.0297  0.00509    287
-> 3     3 -0.774  0.651  0.367   -0.0172   0.00600  0.0211  0.00303    226
-> 4     4 -0.606  0.952 -1.80     0.0157  -0.00978 -0.0590 -0.00754    446
-> 5     5 -0.478  1.10  -1.88    -0.00423  0.00495 -0.0482 -0.00982    468
-> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
-> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
-> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
-> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
-> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
-> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>,
-> #   abs_error_x3 <dbl>, abs_error_x4 <dbl>, abs_error_x5 <dbl>, ...
 \end{verbatim}

### 2025-10-15 paper-quollr.R

@@ -502,8 +502,6 @@ scurve_umap <- umap(
 ## ----echo=TRUE----------------------------------------------------------------
 scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
 
-scurve_umap_obj
-
 
 ## ----echo=TRUE----------------------------------------------------------------
 bin_configs <- calc_bins_y(
@@ -750,7 +748,7 @@ head(trimesh, 5)
 
 
 ## ----echo=TRUE----------------------------------------------------------------
-find_low_dens_hex(
+low_density_hex <- find_low_dens_hex(
   model_2d = df_bin_centroids, 
   b1 = 25, 
   md_thresh = 0.05
@@ -799,8 +797,6 @@ model_error <- augment(
   x = scurve_model_obj,
   highd_data = scurve
   )
-
-head(model_error, 5)
 
 
 ## ----echo=TRUE----------------------------------------------------------------

### 2025-10-15 paper-quollr.Rmd
Initial work adding paragraph for detourr implementation

@@
 
-In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \twoD wireframe structure mapped back to high-dimensional space.
-
 ```{r, echo=TRUE, eval=knitr::is_html_output(), fig.pos='H', fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
@@
 ```
+
+As an alternative to `langevitour`, users can explore the fitted \pD model using the `detourr` [@casper2025].
+The combined data object from `comb_data_model()` can be passed directly to the `detour()` function, where `tour_aes()` defines the projection variables and color mapping.
+The visualization is rendered using `show_scatter()`, which can display both data points and the model’s structural edges via the `edges` argument.
+
+```{r, echo=TRUE, eval=knitr::is_html_output()}
+library(detourr)
+
+detour(
+  df_exe,
+  tour_aes(
+    projection = starts_with("x"),
+    colour = type
+  )
+) |>
+  show_scatter(axes = TRUE, 
+               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
+               palette = c("#66B2CC", "#FF7755"))
+```
+
+In the resulting interactive visualization, blue points represent the high-dimensional data, orange points represent the model centroids from each bin, and the lines between model points reflect the \twoD wireframe structure mapped to high-dimensional space.
 

### 2025-10-15 paper-quollr.tex
Initial work adding paragraph for detourr implementation

@@
 
-In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped back to high-dimensional space.
-
 \begin{verbatim}
@@
 \end{figure}
+
+As an alternative to \texttt{langevitour}, users can explore the fitted $p\text{-}D$ model using the \texttt{detourr} \citep{casper2025}.
+The combined data object from \texttt{comb\_data\_model()} can be passed directly to the \texttt{detour()} function, where \texttt{tour\_aes()} defines the projection variables and color mapping.
+The visualization is rendered using \texttt{show\_scatter()}, which can display both data points and the model's structural edges via the \texttt{edges} argument.
+
+\begin{verbatim}
+library(detourr)
+
+detour(
+  df_exe,
+  tour_aes(
+    projection = starts_with("x"),
+    colour = type
+  )
+) |>
+  show_scatter(axes = TRUE, 
+               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
+               palette = c("#66B2CC", "#FF7755"))
+\end{verbatim}
+
+In the resulting interactive visualization, blue points represent the high-dimensional data, orange points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped to high-dimensional space.
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-32-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-33-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
-While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-32}
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-33}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-36-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-37-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-36}
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-37}
 \end{figure}

### 2025-10-15 paper-quollr.Rmd
Added the detour

@@
                palette = c("#66B2CC", "#FF7755"))
+
 ```

### 2025-10-16 paper-quollr.Rmd
planning implemetation section
Initial writing on adding detour as link plots

@@
 
+To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
+The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
+
 ## Computational efficiency and optimization
@@
 The following demonstration of the package's functionality assumes `quollr` has been loaded.
-To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
-The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
-The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
-
 To illustrate the workflow, we use the built-in example dataset `scurve`, a $7\text{-}D$ simulated dataset consisting of $1000$ observations.
@@
 
+As an alternative to generate tour, the link plots can be generate with `detourr` as well.
+To achieve this, users can construct the linked visualization manually using the `crosstalk` and `htmltools` packages.
+The two- and three-panel linked layouts are created by arranging the NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and a dynamic high-dimensional projection view (`langevitour_output` or `detourr_output`) side by side within a flexible grid layout.
+The `bscols()` function from `crosstalk` synchronizes point selection between these plots, ensuring linked brushing across visual components.
+
 # Application

### 2025-10-16 paper-quollr.tex
planning implemetation section
Initial writing on adding detour as link plots

@@
 
+To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
+The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
+
 \subsection{Computational efficiency and optimization}\label{computational-efficiency-and-optimization}
@@
 The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded.
-To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
-The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
-The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
-
 To illustrate the workflow, we use the built-in example dataset \texttt{scurve}, a \(7\text{-}D\) simulated dataset consisting of \(1000\) observations.
@@
 
+As an alternative to generate tour, the link plots can be generate with \texttt{detourr} as well.
+To achieve this, users can construct the linked visualization manually using the \texttt{crosstalk} and \texttt{htmltools} packages.
+The two- and three-panel linked layouts are created by arranging the NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and a dynamic high-dimensional projection view (\texttt{langevitour\_output} or \texttt{detourr\_output}) side by side within a flexible grid layout.
+The \texttt{bscols()} function from \texttt{crosstalk} synchronizes point selection between these plots, ensuring linked brushing across visual components.
+
 \section{Application}\label{application}

### 2025-10-16 paper-quollr.Rmd
Adde file structures into Usage
Added initial working for the implementation

@@
 
-To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
-The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
-The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
+The implementation of `quollr` emphasizes modularity, computational efficiency, and interoperability.
+Its design follows a layered architecture that separates **data preprocessing**, **model construction**, and **visualization**, ensuring clarity, extensibility, and compatibility with the wider R ecosystem.
+
+## Software architecture
+
+Internally, `quollr` is organized into five primary modules that correspond to key steps in the analysis pipeline:
+
+1.
+**Data preparation** — functions such as `scale_highd()` and `prepare_data()` standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+2. **Hexagonal binning** — the `hex_bin_data()` module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
+3. **Model fitting** — the `fit_highd_model()` function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
+4. **Error computation** — functions like `compute_errors()` and `aggregate_bin_error()` evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
+5.
+**Visualization and linkage** — plotting and tour functions (`show_hex_model()`, `show_link_plots()`, and `show_error_link_plots()`) integrate outputs into linked static and dynamic displays using `ggplot2`, `langevitour`, and `crosstalk`.
+
+Each module is internally independent but communicates through standardized data objects (see next section).
+This modular design simplifies maintenance and allows developers to extend individual components—such as substituting different binning strategies, embedding methods, or visualization tools—without altering the overall workflow.
+
+## Data objects
+
+The internal data objects in `quollr` adhere to the **tidy data** principle, ensuring that every variable forms a column, every observation forms a row, and every type of observational unit forms a table.
+This structure supports compatibility with the `tidyverse` and other R visualization frameworks.
+
+The main object classes include:
+
+* **`highd_data`** – a tibble containing the original high-dimensional observations with a unique identifier (`ID`) and variable columns prefixed with `"x"` (e.g., `x1`, `x2`, …).
+* **`nldr_data`** – a tibble containing two-dimensional embeddings, labeled as `emb1` and `emb2`, matched to the same `ID`s.
+* **`highd_vis_model`** – the primary model object returned by `fit_highd_model()`, containing two key components: the high-dimensional summaries for each bin and their corresponding two-dimensional positions.
+* **`hex_bin_obj`** – the internal representation of the hexagonal grid, including bin coordinates (`a1`, `a2`), centroids, and normalized weights (`w_h`).
+
+These objects provide a coherent foundation for data transfer across `quollr`’s modules—from binning to model fitting and visualization—supporting reproducibility and transparency in analyses.
 
@@
 The following demonstration of the package's functionality assumes `quollr` has been loaded.
+To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
+The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
+
 To illustrate the workflow, we use the built-in example dataset `scurve`, a $7\text{-}D$ simulated dataset consisting of $1000$ observations.

### 2025-10-16 paper-quollr.tex
Adde file structures into Usage
Added initial working for the implementation

@@
 
-To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
-The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
-The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
+The implementation of \texttt{quollr} emphasizes modularity, computational efficiency, and interoperability.
+Its design follows a layered architecture that separates \textbf{data preprocessing}, \textbf{model construction}, and \textbf{visualization}, ensuring clarity, extensibility, and compatibility with the wider R ecosystem.
+
+\subsection{Software architecture}\label{software-architecture}
+
+Internally, \texttt{quollr} is organized into five primary modules that correspond to key steps in the analysis pipeline:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \textbf{Data preparation} --- functions such as \texttt{scale\_highd()} and \texttt{prepare\_data()} standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+\item
+  \textbf{Hexagonal binning} --- the \texttt{hex\_bin\_data()} module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
+\item
+  \textbf{Model fitting} --- the \texttt{fit\_highd\_model()} function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
+\item
+  \textbf{Error computation} --- functions like \texttt{compute\_errors()} and \texttt{aggregate\_bin\_error()} evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
+\item
+  \textbf{Visualization and linkage} --- plotting and tour functions (\texttt{show\_hex\_model()}, \texttt{show\_link\_plots()}, and \texttt{show\_error\_link\_plots()}) integrate outputs into linked static and dynamic displays using \texttt{ggplot2}, \texttt{langevitour}, and \texttt{crosstalk}.
+\end{enumerate}
+
+Each module is internally independent but communicates through standardized data objects (see next section).
+This modular design simplifies maintenance and allows developers to extend individual components---such as substituting different binning strategies, embedding methods, or visualization tools---without altering the overall workflow.
+
+\subsection{Data objects}\label{data-objects}
+
+The internal data objects in \texttt{quollr} adhere to the \textbf{tidy data} principle, ensuring that every variable forms a column, every observation forms a row, and every type of observational unit forms a table.
+This structure supports compatibility with the \texttt{tidyverse} and other R visualization frameworks.
+
+The main object classes include:
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{\texttt{highd\_data}} -- a tibble containing the original high-dimensional observations with a unique identifier (\texttt{ID}) and variable columns prefixed with \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, \ldots).
+\item
+  \textbf{\texttt{nldr\_data}} -- a tibble containing two-dimensional embeddings, labeled as \texttt{emb1} and \texttt{emb2}, matched to the same \texttt{ID}s.
+\item
+  \textbf{\texttt{highd\_vis\_model}} -- the primary model object returned by \texttt{fit\_highd\_model()}, containing two key components: the high-dimensional summaries for each bin and their corresponding two-dimensional positions.
+\item
+  \textbf{\texttt{hex\_bin\_obj}} -- the internal representation of the hexagonal grid, including bin coordinates (\texttt{a1}, \texttt{a2}), centroids, and normalized weights (\texttt{w\_h}).
+\end{itemize}
+
+These objects provide a coherent foundation for data transfer across \texttt{quollr}'s modules---from binning to model fitting and visualization---supporting reproducibility and transparency in analyses.
 
@@
 The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded.
+To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
+The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
+
 To illustrate the workflow, we use the built-in example dataset \texttt{scurve}, a \(7\text{-}D\) simulated dataset consisting of \(1000\) observations.

### 2025-10-18 paper-quollr.Rmd
Initial update for implementation section
Initial work on the quollr implementation

@@
 
-The implementation of `quollr` emphasizes modularity, computational efficiency, and interoperability.
-Its design follows a layered architecture that separates **data preprocessing**, **model construction**, and **visualization**, ensuring clarity, extensibility, and compatibility with the wider R ecosystem.
+The implementation of `quollr` is designed to be modular, efficient, and easy to extend.
+The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration.
+This modular structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
 

### 2025-10-18 paper-quollr.tex
Initial update for implementation section
Initial work on the quollr implementation

@@
 
-The implementation of \texttt{quollr} emphasizes modularity, computational efficiency, and interoperability.
-Its design follows a layered architecture that separates \textbf{data preprocessing}, \textbf{model construction}, and \textbf{visualization}, ensuring clarity, extensibility, and compatibility with the wider R ecosystem.
+The implementation of \texttt{quollr} is designed to be modular, efficient, and easy to extend.
+The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration.
+This modular structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
 

### 2025-10-18 paper-quollr.Rmd
Updated implementation intro

@@
 
-The implementation of `quollr` is designed to be modular, efficient, and easy to extend.
+The implementation of `quollr` is designed to be efficient, and easy to extend.
 The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration.
-This modular structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
+This package structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
 

### 2025-10-18 paper-quollr.tex
Updated implementation intro

@@
 
-The implementation of \texttt{quollr} is designed to be modular, efficient, and easy to extend.
+The implementation of \texttt{quollr} is designed to be efficient, and easy to extend.
 The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration.
-This modular structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
+This package structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
 

### 2025-10-18 paper-quollr.Rmd
Developed the outputs with detour (links)

@@
 
-1.
-**Data preparation** — functions such as `scale_highd()` and `prepare_data()` standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
-2. **Hexagonal binning** — the `hex_bin_data()` module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
-3. **Model fitting** — the `fit_highd_model()` function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
-4. **Error computation** — functions like `compute_errors()` and `aggregate_bin_error()` evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
+1. Data preparation — functions such as `gen_scaled_data()` standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+2. Hexagonal binning — the `hex_bin_data()` module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
+3. Model fitting — the `fit_highd_model()` function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
+4. Error computation — functions like `compute_errors()` and `aggregate_bin_error()` evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
 5.
-**Visualization and linkage** — plotting and tour functions (`show_hex_model()`, `show_link_plots()`, and `show_error_link_plots()`) integrate outputs into linked static and dynamic displays using `ggplot2`, `langevitour`, and `crosstalk`.
+Visualization and linkage — plotting and tour functions (`show_hex_model()`, `show_link_plots()`, and `show_error_link_plots()`) integrate outputs into linked static and dynamic displays using `ggplot2`, `langevitour`, and `crosstalk`.
 
 Each module is internally independent but communicates through standardized data objects (see next section).
-This modular design simplifies maintenance and allows developers to extend individual components—such as substituting different binning strategies, embedding methods, or visualization tools—without altering the overall workflow.
+This modular design simplifies maintenance and allows developers to extend individual components such as substituting different binning strategies, embedding methods, or visualization tools without altering the overall workflow.
 
@@
 ) |>
-  show_scatter(axes = TRUE, 
+  tour_path(grand_tour(2), 
+                    max_bases=50, fps = 60) |>
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
-               palette = c("#66B2CC", "#FF7755"))
-
+               palette = c("#66B2CC", "#FF7755"),
+               width = "600px", height = "600px")
 ```
@@
 
-As an alternative to generate tour, the link plots can be generate with `detourr` as well.
-To achieve this, users can construct the linked visualization manually using the `crosstalk` and `htmltools` packages.
-The two- and three-panel linked layouts are created by arranging the NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and a dynamic high-dimensional projection view (`langevitour_output` or `detourr_output`) side by side within a flexible grid layout.
-The `bscols()` function from `crosstalk` synchronizes point selection between these plots, ensuring linked brushing across visual components.
+As an alternative to using `langevitour`, link plots can also be generated with the `detourr` package.
+In this setup, users can manually construct the linked visualization using the `crosstalk` [@joe2023] and `htmltools` [@joe2024] packages.
+The interactive layout is created by arranging the \twoD NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid.
+The NLDR and error plots are rendered with `ggplotly()` [@chapman2020] to enable interactive linking.
+The `bscols()` function from `crosstalk` manages synchronization across these panels, allowing for linked brushing and coordinated selection between these interactive plots.
+
+A two-panel linked plot combining the NLDR view and the tour from `detourr` can be created as follows:
+
+```{r}
+point_colours <- c("#66B2CC", "#FF7755")
+point_sizes <- c(0, 1)
+
+num_highd_col <- df_exe |>
+  dplyr::select(starts_with("x")) |>
+  NCOL()
+
+df_all <- df_exe |>
+  dplyr::filter(type == "data") ## original dataset
+
+df_b <- df_exe |>
+  dplyr::filter(type == "model") ## High-d model
+
+shared_df <- crosstalk::SharedData$new(df_exe)
+
+nldr_plt <- shared_df |>
+  ggplot(aes(x = emb1, y = emb2)) +
+  geom_point(alpha=0.5, colour=point_colours[1], size = 0.5) +
+  theme_linedraw() +
+  theme(
+    #aspect.ratio = 1,
+    plot.background = element_rect(fill = 'transparent', colour = NA),
+    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+    panel.background = element_rect(fill = 'transparent',
+                                    colour = NA),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    axis.title.x = element_blank(), axis.title.y = element_blank(),
+    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(), axis.ticks.y = element_blank()
+  )
+
+nldr_plt <- ggplotly(nldr_plt, width = 600,
+                     height = 600, tooltip = "none") |>
+  style(unselected=list(marker=list(opacity=1))) |>
+  highlight(on="plotly_selected", off="plotly_deselect") |>
+  config(displayModeBar = FALSE)
+
+```
+
+```{r, echo=TRUE, eval=knitr::is_html_output(), layout = "l-page"}
+
+detourr_output <- detour(
+  shared_df,
+  tour_aes(
+    projection = starts_with("x"),
+    colour = type
+  )
+) |>
+  tour_path(grand_tour(2), 
+                    max_bases=50, fps = 60) |>
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
+               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
+               palette = c("#66B2CC", "#FF7755"),
+                width = "600px", height = "600px")
+
+crosstalk::bscols(
+    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+                   nldr_plt,
+                   htmltools::div(style = "margin-top: 20px;", detourr_output)
+    ),
+    device = "xs"
+  )
+```
+
+For analyses that include model error visualization, a three-panel view can be constructed by adding the error distribution plot (`error_plt`):
+
+```{r}
+
+num_highd_col <- df_exe |>
+  dplyr::select(starts_with("x")) |>
+  NCOL()
+
+df_all <- df_exe |>
+  dplyr::filter(type == "data") ## original dataset
+
+df_b <- df_exe |>
+  dplyr::filter(type == "model") ## High-d model
+
+shared_df <- crosstalk::SharedData$new(df_exe)
+
+error_plt <- shared_df |>
+  ggplot(aes(x=sqrt_row_wise_total_error, y = density)) +
+  geom_point(colour = point_colours[1]) +
+  xlab(expression(e[hj])) +
+  ylab("") +
+  theme_bw() +
+  theme(
+    #aspect.ratio = 1,
+    plot.background = element_rect(fill = 'transparent', colour = NA),
+    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+    panel.background = element_rect(fill = 'transparent',
+                                    colour = NA),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    axis.title.x = element_blank(), axis.title.y = element_blank(),
+    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(), axis.ticks.y = element_blank()
+  )
+
+error_plt <- ggplotly(error_plt, width = 400,
+                      height = 400, tooltip = "none") |>
+  style(unselected=list(marker=list(opacity=1))) |>
+  highlight(on="plotly_selected", off="plotly_deselect") |>
+  config(displayModeBar = FALSE)
+
+nldr_plt <- shared_df |>
+  ggplot(aes(x = emb1, y = emb2)) +
+  geom_point(alpha=0.5, colour=point_colours[1], size = 0.5) +
+  theme_linedraw() +
+  theme(
+    #aspect.ratio = 1,
+    plot.background = element_rect(fill = 'transparent', colour = NA),
+    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+    panel.background = element_rect(fill = 'transparent',
+                                    colour = NA),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    axis.title.x = element_blank(), axis.title.y = element_blank(),
+    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(), axis.ticks.y = element_blank()
+  )
+
+nldr_plt <- ggplotly(nldr_plt, width = 400,
+                     height = 400, tooltip = "none") |>
+  style(unselected=list(marker=list(opacity=1))) |>
+  highlight(on="plotly_selected", off="plotly_deselect") |>
+  config(displayModeBar = FALSE)
+```
+
+```{r, echo=TRUE, eval=knitr::is_html_output(), layout = "l-page"}
+
+detourr_output <- detour(
+  shared_df,
+  tour_aes(
+    projection = starts_with("x"),
+    colour = type
+  )
+) |>
+  tour_path(grand_tour(2), 
+                    max_bases=50, fps = 60) |>
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
+               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
+               palette = c("#66B2CC", "#FF7755"),
+                width = "600px", height = "600px")
+
+crosstalk::bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr 1fr;",
+    error_plt, 
+    nldr_plt,
+    htmltools::div(style = "margin-top: 20px;", detourr_output)
+  ),
+  device = "xs"
+)
+```
 
@@
 This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.
-These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].
+These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `htmltools` [@joe2024], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].

### 2025-10-18 paper-quollr.tex
Developed the outputs with detour (links)

@@
 \item
-  \textbf{Data preparation} --- functions such as \texttt{scale\_highd()} and \texttt{prepare\_data()} standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+  Data preparation --- functions such as \texttt{gen\_scaled\_data()} standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
 \item
-  \textbf{Hexagonal binning} --- the \texttt{hex\_bin\_data()} module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
+  Hexagonal binning --- the \texttt{hex\_bin\_data()} module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
 \item
-  \textbf{Model fitting} --- the \texttt{fit\_highd\_model()} function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
+  Model fitting --- the \texttt{fit\_highd\_model()} function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
 \item
-  \textbf{Error computation} --- functions like \texttt{compute\_errors()} and \texttt{aggregate\_bin\_error()} evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
+  Error computation --- functions like \texttt{compute\_errors()} and \texttt{aggregate\_bin\_error()} evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
 \item
-  \textbf{Visualization and linkage} --- plotting and tour functions (\texttt{show\_hex\_model()}, \texttt{show\_link\_plots()}, and \texttt{show\_error\_link\_plots()}) integrate outputs into linked static and dynamic displays using \texttt{ggplot2}, \texttt{langevitour}, and \texttt{crosstalk}.
+  Visualization and linkage --- plotting and tour functions (\texttt{show\_hex\_model()}, \texttt{show\_link\_plots()}, and \texttt{show\_error\_link\_plots()}) integrate outputs into linked static and dynamic displays using \texttt{ggplot2}, \texttt{langevitour}, and \texttt{crosstalk}.
 \end{enumerate}
@@
 Each module is internally independent but communicates through standardized data objects (see next section).
-This modular design simplifies maintenance and allows developers to extend individual components---such as substituting different binning strategies, embedding methods, or visualization tools---without altering the overall workflow.
+This modular design simplifies maintenance and allows developers to extend individual components such as substituting different binning strategies, embedding methods, or visualization tools without altering the overall workflow.
 
@@
 ) |>
-  show_scatter(axes = TRUE, 
+  tour_path(grand_tour(2), 
+                    max_bases=50, fps = 60) |>
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
-               palette = c("#66B2CC", "#FF7755"))
+               palette = c("#66B2CC", "#FF7755"),
+               width = "600px", height = "600px")
 \end{verbatim}
@@
 
-As an alternative to generate tour, the link plots can be generate with \texttt{detourr} as well.
-To achieve this, users can construct the linked visualization manually using the \texttt{crosstalk} and \texttt{htmltools} packages.
-The two- and three-panel linked layouts are created by arranging the NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and a dynamic high-dimensional projection view (\texttt{langevitour\_output} or \texttt{detourr\_output}) side by side within a flexible grid layout.
-The \texttt{bscols()} function from \texttt{crosstalk} synchronizes point selection between these plots, ensuring linked brushing across visual components.
+As an alternative to using \texttt{langevitour}, link plots can also be generated with the \texttt{detourr} package.
+In this setup, users can manually construct the linked visualization using the \texttt{crosstalk} \citep{joe2023} and \texttt{htmltools} \citep{joe2024} packages.
+The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid.
+The NLDR and error plots are rendered with \texttt{ggplotly()} \citep{chapman2020} to enable interactive linking.
+The \texttt{bscols()} function from \texttt{crosstalk} manages synchronization across these panels, allowing for linked brushing and coordinated selection between these interactive plots.
+
+A two-panel linked plot combining the NLDR view and the tour from \texttt{detourr} can be created as follows:
+
+\begin{verbatim}
+detourr_output <- detour(
+  shared_df,
+  tour_aes(
+    projection = starts_with("x"),
+    colour = type
+  )
+) |>
+  tour_path(grand_tour(2), 
+                    max_bases=50, fps = 60) |>
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
+               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
+               palette = c("#66B2CC", "#FF7755"),
+                width = "600px", height = "600px")
+
+crosstalk::bscols(
+    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+                   nldr_plt,
+                   htmltools::div(style = "margin-top: 20px;", detourr_output)
+    ),
+    device = "xs"
+  )
+\end{verbatim}
+
+For analyses that include model error visualization, a three-panel view can be constructed by adding the error distribution plot (\texttt{error\_plt}):
+
+\begin{verbatim}
+detourr_output <- detour(
+  shared_df,
+  tour_aes(
+    projection = starts_with("x"),
+    colour = type
+  )
+) |>
+  tour_path(grand_tour(2), 
+                    max_bases=50, fps = 60) |>
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
+               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
+               palette = c("#66B2CC", "#FF7755"),
+                width = "600px", height = "600px")
+
+crosstalk::bscols(
+  htmltools::div(
+    style = "display: grid; grid-template-columns: 1fr 1fr 1fr;",
+    error_plt, 
+    nldr_plt,
+    htmltools::div(style = "margin-top: 20px;", detourr_output)
+  ),
+  device = "xs"
+)
+\end{verbatim}
 
@@
 This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.
-These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.
+These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{htmltools} \citep{joe2024}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.
 

### 2025-10-19 paper-quollr.Rmd
Added implementation workflow

@@
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H'}
+
 hex_grid_scurve + wrap_plots(
@@
 The implementation of `quollr` is designed to be efficient, and easy to extend.
-The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration.
-This package structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
+The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration (Figure `r '\\@ref(fig:workflow)'`).
+This package structure makes the code easier to maintain and allows new features to be added without changing the existing functionality.
+
+```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow for "}
+
+knitr::include_graphics("figures/quollr_workflow.png")
+```
 
@@
 
-Internally, `quollr` is organized into five primary modules that correspond to key steps in the analysis pipeline:
-
-1. Data preparation — functions such as `gen_scaled_data()` standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
-2. Hexagonal binning — the `hex_bin_data()` module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
-3. Model fitting — the `fit_highd_model()` function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
-4. Error computation — functions like `compute_errors()` and `aggregate_bin_error()` evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
-5.
-Visualization and linkage — plotting and tour functions (`show_hex_model()`, `show_link_plots()`, and `show_error_link_plots()`) integrate outputs into linked static and dynamic displays using `ggplot2`, `langevitour`, and `crosstalk`.
+The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, \twoD model construction, lifting into \pD, prediction, error computation, visualization, and interactivity.
+Each module performs a distinct task and communicates through data objects.
+
+1. Data preprocessing — The function `gen_scaled_data()` standardizes the embedding data, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+
+2. Construct \twoD model — `hex_binning()`, `merge_hexbin_centroids()`, `tri_bin_centroids()`, `gen_edges()`, `update_trimesh_index()`
+
+3. Lift the model into \pD{} — the `avg_highd_data()`
+
+4. Prediction - `predict_emb()`
+
+5. Error computation — `glance()`
+
+6. Visualization — `geom_hexgrid()`, `geom_trimesh()`, `show_langevitour()`
+
+7. Interactivity - `show_link_plots()`, and `show_error_link_plots()`
 
@@
 
-The internal data objects in `quollr` adhere to the **tidy data** principle, ensuring that every variable forms a column, every observation forms a row, and every type of observational unit forms a table.
-This structure supports compatibility with the `tidyverse` and other R visualization frameworks.
-
-The main object classes include:
-
-* **`highd_data`** – a tibble containing the original high-dimensional observations with a unique identifier (`ID`) and variable columns prefixed with `"x"` (e.g., `x1`, `x2`, …).
-* **`nldr_data`** – a tibble containing two-dimensional embeddings, labeled as `emb1` and `emb2`, matched to the same `ID`s.
-* **`highd_vis_model`** – the primary model object returned by `fit_highd_model()`, containing two key components: the high-dimensional summaries for each bin and their corresponding two-dimensional positions.
-* **`hex_bin_obj`** – the internal representation of the hexagonal grid, including bin coordinates (`a1`, `a2`), centroids, and normalized weights (`w_h`).
-
-These objects provide a coherent foundation for data transfer across `quollr`’s modules—from binning to model fitting and visualization—supporting reproducibility and transparency in analyses.
+The internal data objects follow the tidy data principle: each variable is stored in a column, each observation in a row, and each type of information in its own table.
+This structure makes the package easy to use with the `tidyverse` and other R visualization tools.
+
+### Input objects
+
+- `highd_data` – a tibble containing the original high-dimensional observations with a unique identifier (`ID`) and variable columns prefixed with `"x"` (e.g., `x1`, `x2`, …).
+
+- `nldr_data` – a tibble containing two-dimensional embeddings, labeled as `emb1` and `emb2`, matched to the same `ID`s.
+
+### Generated objects
+
+- `scaled_nldr_obj` — the output of `gen_scaled_data()`, which rescales the embedding to the range $[0, 1] \times [0, y_{2,\text{max}}]$, where $y_{2,\text{max}} = r_2 / r_1$ is the ratio of the embedding ranges.
+It includes the scaled coordinates (`scaled_nldr`) and the original limits (`lim1`, `lim2`).
+
+- `hex_bin_obj` – the object created by `hex_binning()`, which defines the structure of the two-dimensional hexagonal grid used in modeling.
+It includes the grid spacing (`a1`, `a2`), the number of bins along each axis, the centroids of all hexagons, polygon coordinates for plotting, and the mapping of each data point to its assigned hexagon.
+
+- `highd_vis_model` — the main model object returned by `fit_highd_model()`.
+It stores all components of the fitted visualization model, including the scaled NLDR data (`nldr_scaled_obj`), the hexagonal bin structure (`hb_obj`), the averaged \pD summaries for each bin (`model_highd`), the corresponding \twoD bin centroids (`model_2d`), and the triangulated mesh connecting neighboring bins (`trimesh_data`).
 
@@
   nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 25, 
+  b1 = 21, 
   q = 0.1)
@@
   nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 25, 
+  b1 = 21, 
   q = 0.1
@@
 
-all_centroids_df
+head(all_centroids_df, 5)
 ```
@@
   model_2d = df_bin_centroids, 
-  b1 = 25, 
+  b1 = 21, 
   md_thresh = 0.05
@@
   )
-
-trimesh_limb_int <- ggplotly(trimesh_limb, 
-                                width = "350", 
-                                height = "350", 
-                                tooltip = "none") |>
-  config(
-    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
-    displayModeBar = FALSE,   # Hides the plotly toolbar
-    editable = FALSE,         # Disables annotations and editing
-    showTips = FALSE,         # Removes tooltip on hover
-    displaylogo = FALSE,      # Hides plotly logo
-    responsive = FALSE        # Disables responsive resizing
-  )
-
-# hex_grid <- tsne_limb_obj$hb_obj$hex_poly
-# counts_df <- tsne_limb_obj$hb_obj$std_cts
-# 
-# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("h" = "h"))
-# 
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = n_h)) +
-#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = h)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed() +
-#   theme_minimal()
 ```
@@
 
-```{r langevitour-limb-tsne-author-proj}
-
-data_limb_n <- data_limb_n |>
-  select(-type) |>
-  mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
-
-df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
-
-limb_highd_vis <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
-                         lineFrom = trimesh_data_limb$from_reindexed,
-                         lineTo = trimesh_data_limb$to_reindexed,
-                         group = factor(df_model_data_limb_n$type,
+```{r}
+
+df <- inner_join(data_limb, tsne_limb_scaled_with_cluster, by = "ID")
+
+### Define type column
+df <- df |>
+  select(starts_with("x"), starts_with("emb"), "cluster") |>
+  mutate(type = tsne_limb_scaled_with_cluster$cluster) ## original dataset
+
+df_b <- df_bin_limb |>
+  filter(h %in% df_bin_centroids_limb$h) |>
+  mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in model_2d
+df_b <- df_b[match(df_bin_centroids_limb$h, df_b$h),] |>
+  tidyr::drop_na() |>
+  select(-h)
+
+df_exe_limb_int <- bind_rows(df_b, df)
+
+point_data <- df_exe_limb_int 
+edge_data <- trimesh_data_limb
+
+num_highd_col <- point_data |>
+  dplyr::select(starts_with("x")) |>
+  NCOL()
+
+shared_df <- crosstalk::SharedData$new(point_data)
+
+nldr_plt <- shared_df |>
+  ggplot(aes(x = emb1, y = emb2, colour = cluster)) +
+  geom_point(alpha=0.5, size = 0.8) +
+  geom_segment(data = trimesh_data_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 0.5) +
+  scale_color_manual(values = c("0" = '#66c2a5',"1" = '#fc8d62',"2" = '#8da0cb',"3" = '#e78ac3',"4" = '#a6d854',"5" = '#ffd92f',"6" = '#e5c494')) +
+  theme_linedraw() +
+  theme(
+    #aspect.ratio = 1,
+    plot.background = element_rect(fill = 'transparent', colour = NA),
+    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+    panel.background = element_rect(fill = 'transparent',
+                                    colour = NA),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    axis.title.x = element_blank(), axis.title.y = element_blank(),
+    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+    legend.position = "none"
+  )
+
+nldr_plt <- ggplotly(nldr_plt, width = 600,
+                     height = 600, tooltip = "none") |>
+  style(unselected=list(marker=list(opacity=1))) |>
+  highlight(on="plotly_selected", off="plotly_deselect") |>
+  config(displayModeBar = FALSE)
+
+langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
+                                               lineFrom = edge_data$from_reindexed,
+                                               lineTo = edge_data$to_reindexed,
+                                               group = factor(point_data$type,
                                                               c("0", "1", "2", "3", "4", "5", "6", "model")),
+                                               pointSize = append(rep(point_sizes[1], NROW(df_b)),
+                                                                  rep(point_sizes[2], NROW(df))),
                                                levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
-                         enableControls = FALSE,
-                         width = "421px", height = "421px")
+                                               link=shared_df,
+                                               linkFilter=FALSE)
+
+linked_plt_int <- crosstalk::bscols(
+  htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+                 nldr_plt,
+                 htmltools::div(style = "margin-top: 20px;", langevitour_output)
+  ),
+  device = "xs"
+)
 ```
@@
 
-trimesh_limb_best_int <- ggplotly(trimesh_limb_best, 
-                                width = "350", 
-                                height = "350", 
-                                tooltip = "none") |>
-  config(
-    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
-    displayModeBar = FALSE,   # Hides the plotly toolbar
-    editable = FALSE,         # Disables annotations and editing
-    showTips = FALSE,         # Removes tooltip on hover
-    displaylogo = FALSE,      # Hides plotly logo
-    responsive = FALSE        # Disables responsive resizing
-  )
-
-# hex_grid <- tsne_best_limb_obj$hb_obj$hex_poly
-# counts_df <- tsne_best_limb_obj$hb_obj$std_cts
-# 
-# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "h"))
-# 
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = n_h)) +
-#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = h)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed() +
-#   theme_minimal()
+```
+
+```{r}
+df <- inner_join(data_limb, tsne_limb_scaled_with_cluster, by = "ID")
+
+### Define type column
+df <- df |>
+  select(starts_with("x"), starts_with("emb"), "cluster") |>
+  mutate(type = tsne_limb_scaled_with_cluster$cluster) ## original dataset
+
+df_b <- df_bin_limb |>
+  filter(h %in% df_bin_centroids_limb$h) |>
+  mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in model_2d
+df_b <- df_b[match(df_bin_centroids_limb$h, df_b$h),] |>
+  tidyr::drop_na() |>
+  select(-h)
+
+df_exe_limb_int <- bind_rows(df_b, df)
+
+point_data <- df_exe_limb_int 
+edge_data <- trimesh_data_limb
+
+num_highd_col <- point_data |>
+  dplyr::select(starts_with("x")) |>
+  NCOL()
+
+shared_df <- crosstalk::SharedData$new(point_data)
+
+nldr_plt <- shared_df |>
+  ggplot(aes(x = emb1, y = emb2, colour = cluster)) +
+  geom_point(alpha=0.5, size = 0.8) +
+  geom_segment(data = trimesh_data_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 0.5) +
+  scale_color_manual(values = c("0" = '#66c2a5',"1" = '#fc8d62',"2" = '#8da0cb',"3" = '#e78ac3',"4" = '#a6d854',"5" = '#ffd92f',"6" = '#e5c494')) +
+  theme_linedraw() +
+  theme(
+    #aspect.ratio = 1,
+    plot.background = element_rect(fill = 'transparent', colour = NA),
+    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+    panel.background = element_rect(fill = 'transparent',
+                                    colour = NA),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    axis.title.x = element_blank(), axis.title.y = element_blank(),
+    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+    legend.position = "none"
+  )
+
+nldr_plt <- ggplotly(nldr_plt, width = 600,
+                     height = 600, tooltip = "none") |>
+  style(unselected=list(marker=list(opacity=1))) |>
+  highlight(on="plotly_selected", off="plotly_deselect") |>
+  config(displayModeBar = FALSE)
+
+langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
+                                               lineFrom = edge_data$from_reindexed,
+                                               lineTo = edge_data$to_reindexed,
+                                               group = factor(point_data$type,
+                                                              c("0", "1", "2", "3", "4", "5", "6", "model")),
+                                               pointSize = append(rep(point_sizes[1], NROW(df_b)),
+                                                                  rep(point_sizes[2], NROW(df))),
+                                               levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
+                                               link=shared_df,
+                                               linkFilter=FALSE)
+
+linked_plt_best <- crosstalk::bscols(
+  htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+                 nldr_plt,
+                 htmltools::div(style = "margin-top: 20px;", langevitour_output)
+  ),
+  device = "xs"
+)
 ```
@@
   select(-type)
-```
-
-```{r langevitour-limb-tsne-best-proj}
-
-data_limb_n <- data_limb_n |>
-  select(-type) |>
-  mutate(type = as.character(tsne_limb_scaled_with_cluster$cluster.ids))
-
-df_model_data_limb_n <- bind_rows(df_b_limb, data_limb_n)
-
-limb_highd_vis_best <- langevitour::langevitour(df_model_data_limb_n[1:(length(df_model_data_limb_n)-1)],
-                         lineFrom = trimesh_data_limb$from_reindexed,
-                         lineTo = trimesh_data_limb$to_reindexed,
-                         group = factor(df_model_data_limb_n$type,
-                                        c("0", "1", "2", "3", "4", "5", "6", "model")),
-                         levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
-                         enableControls = FALSE,
-                         width = "421px", height = "421px")
 ```
@@
 
-```{r model-limb-html, eval=knitr::is_html_output(), fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\@ref(fig:limb-rwbss) b) and the $2-\\text{D}$ layout selected (Figure \\@ref(fig:limb-rwbss) f) by RWBSS plot (Figure \\@ref(fig:limb-rwbss)) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$."}
-
-modellimbfig <- htmltools::div(
-        style = "display: grid; grid-template-columns: 1fr 1fr; row-gap: 0px; justify-items: center;",
-        trimesh_limb_int,
-        htmltools::div(style = "margin-top: 13px;", limb_highd_vis),
-        trimesh_limb_best_int,
-        htmltools::div(style = "margin-top: 13px;", limb_highd_vis_best)
-      )
-
-class(modellimbfig) <- c(class(modellimbfig), "htmlwidget")
-
-modellimbfig
-```
-
-```{r}
-df_exe_limb_int <- comb_all_data_model(
-  highd_data = data_limb, 
-  nldr_data = tsne_limb, 
-  model_highd = df_bin_limb, 
-  model_2d = df_bin_centroids_limb
-)
-```
-
 ```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
 
-tsne_link_limb <- show_link_plots(
-  point_data = df_exe_limb_int, 
-  edge_data = trimesh_data_limb, 
-  point_colour = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')
-)
-
-class(tsne_link_limb) <- c(class(tsne_link_limb), "htmlwidget")
-
-tsne_link_limb
-```
-
-```{r}
-df_exe_limb_int_best <- comb_all_data_model(
-  highd_data = data_limb, 
-  nldr_data = tsne_limb2, 
-  model_highd = df_bin_limb, 
-  model_2d = df_bin_centroids_limb
-)
+class(linked_plt_int) <- c(class(linked_plt_int), "htmlwidget")
+
+linked_plt_int
 ```
@@
 
-tsne_link_limb_best <- show_link_plots(
-  point_data = df_exe_limb_int_best, 
-  edge_data = trimesh_data_limb, 
-  point_colour = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')
-)
-
-class(tsne_link_limb_best) <- c(class(tsne_link_limb_best), "htmlwidget")
-
-tsne_link_limb_best
+class(linked_plt_best) <- c(class(linked_plt_best), "htmlwidget")
+
+linked_plt_best
 ```

### 2025-10-19 paper-quollr.tex
Added implementation workflow

@@
 The implementation of \texttt{quollr} is designed to be efficient, and easy to extend.
-The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration.
-This package structure makes the code easier to maintain and allows new features to be added without disrupting the existing functionality.
+The package is organized into a series of logical components that reflect the main stages of the workflow: data preprocessing, model fitting, low-density bin removal, prediction, visualization, and interactive exploration (Figure \ref{fig:workflow}).
+This package structure makes the code easier to maintain and allows new features to be added without changing the existing functionality.
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{figures/quollr_workflow} \caption{Workflow for }\label{fig:workflow}
+\end{figure}
 
@@
 
-Internally, \texttt{quollr} is organized into five primary modules that correspond to key steps in the analysis pipeline:
+The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, $2\text{-}D$ model construction, lifting into $p\text{-}D$, prediction, error computation, visualization, and interactivity.
+Each module performs a distinct task and communicates through data objects.
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Data preparation --- functions such as \texttt{gen\_scaled\_data()} standardize the input, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
-\item
-  Hexagonal binning --- the \texttt{hex\_bin\_data()} module constructs a two-dimensional hexagonal grid and aggregates local summaries of the high-dimensional data for each bin.
-\item
-  Model fitting --- the \texttt{fit\_highd\_model()} function computes high-dimensional centroids and pairwise bin connections to build a structured manifold representation.
-\item
-  Error computation --- functions like \texttt{compute\_errors()} and \texttt{aggregate\_bin\_error()} evaluate local and global discrepancies between the model and data, enabling diagnostic visualization.
-\item
-  Visualization and linkage --- plotting and tour functions (\texttt{show\_hex\_model()}, \texttt{show\_link\_plots()}, and \texttt{show\_error\_link\_plots()}) integrate outputs into linked static and dynamic displays using \texttt{ggplot2}, \texttt{langevitour}, and \texttt{crosstalk}.
+\item
+  Data preprocessing --- The function \texttt{gen\_scaled\_data()} standardizes the embedding data, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+\item
+  Construct $2\text{-}D$ model --- \texttt{hex\_binning()}, \texttt{merge\_hexbin\_centroids()}, \texttt{tri\_bin\_centroids()}, \texttt{gen\_edges()}, \texttt{update\_trimesh\_index()}
+\item
+  Lift the model into $p\text{-}D${} --- the \texttt{avg\_highd\_data()}
+\item
+  Prediction - \texttt{predict\_emb()}
+\item
+  Error computation --- \texttt{glance()}
+\item
+  Visualization --- \texttt{geom\_hexgrid()}, \texttt{geom\_trimesh()}, \texttt{show\_langevitour()}
+\item
+  Interactivity - \texttt{show\_link\_plots()}, and \texttt{show\_error\_link\_plots()}
 \end{enumerate}
@@
 
-The internal data objects in \texttt{quollr} adhere to the \textbf{tidy data} principle, ensuring that every variable forms a column, every observation forms a row, and every type of observational unit forms a table.
-This structure supports compatibility with the \texttt{tidyverse} and other R visualization frameworks.
-
-The main object classes include:
+The internal data objects follow the tidy data principle: each variable is stored in a column, each observation in a row, and each type of information in its own table.
+This structure makes the package easy to use with the \texttt{tidyverse} and other R visualization tools.
+
+\subsubsection{Input objects}\label{input-objects}
 
 \begin{itemize}
-\tightlist
-\item
-  \textbf{\texttt{highd\_data}} -- a tibble containing the original high-dimensional observations with a unique identifier (\texttt{ID}) and variable columns prefixed with \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, \ldots).
-\item
-  \textbf{\texttt{nldr\_data}} -- a tibble containing two-dimensional embeddings, labeled as \texttt{emb1} and \texttt{emb2}, matched to the same \texttt{ID}s.
-\item
-  \textbf{\texttt{highd\_vis\_model}} -- the primary model object returned by \texttt{fit\_highd\_model()}, containing two key components: the high-dimensional summaries for each bin and their corresponding two-dimensional positions.
-\item
-  \textbf{\texttt{hex\_bin\_obj}} -- the internal representation of the hexagonal grid, including bin coordinates (\texttt{a1}, \texttt{a2}), centroids, and normalized weights (\texttt{w\_h}).
+\item
+  \texttt{highd\_data} -- a tibble containing the original high-dimensional observations with a unique identifier (\texttt{ID}) and variable columns prefixed with \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, \ldots).
+\item
+  \texttt{nldr\_data} -- a tibble containing two-dimensional embeddings, labeled as \texttt{emb1} and \texttt{emb2}, matched to the same \texttt{ID}s.
 \end{itemize}
 
-These objects provide a coherent foundation for data transfer across \texttt{quollr}'s modules---from binning to model fitting and visualization---supporting reproducibility and transparency in analyses.
+\subsubsection{Generated objects}\label{generated-objects}
+
+\begin{itemize}
+\item
+  \texttt{scaled\_nldr\_obj} --- the output of \texttt{gen\_scaled\_data()}, which rescales the embedding to the range \([0, 1] \times [0, y_{2,\text{max}}]\), where \(y_{2,\text{max}} = r_2 / r_1\) is the ratio of the embedding ranges.
+It includes the scaled coordinates (\texttt{scaled\_nldr}) and the original limits (\texttt{lim1}, \texttt{lim2}).
+\item
+  \texttt{hex\_bin\_obj} -- the object created by \texttt{hex\_binning()}, which defines the structure of the two-dimensional hexagonal grid used in modeling.
+It includes the grid spacing (\texttt{a1}, \texttt{a2}), the number of bins along each axis, the centroids of all hexagons, polygon coordinates for plotting, and the mapping of each data point to its assigned hexagon.
+\item
+  \texttt{highd\_vis\_model} --- the main model object returned by \texttt{fit\_highd\_model()}.
+It stores all components of the fitted visualization model, including the scaled NLDR data (\texttt{nldr\_scaled\_obj}), the hexagonal bin structure (\texttt{hb\_obj}), the averaged $p\text{-}D$ summaries for each bin (\texttt{model\_highd}), the corresponding $2\text{-}D$ bin centroids (\texttt{model\_2d}), and the triangulated mesh connecting neighboring bins (\texttt{trimesh\_data}).
+\end{itemize}
 
@@
 > $b2
-> [1] 36
+> [1] 30
 > 
 > $a1
-> [1] 0.04943533
+> [1] 0.05966332
 > 
 > $a2
-> [1] 0.04281225
+> [1] 0.05166995
 \end{verbatim}
@@
 
-all_centroids_df
-\end{verbatim}
-
-\begin{verbatim}
-> # A tibble: 900 x 3
+head(all_centroids_df, 5)
+\end{verbatim}
+
+\begin{verbatim}
+> # A tibble: 5 x 3
 >       h     c_x    c_y
@@
 > 1     1 -0.1    -0.127
->  2     2 -0.0506  -0.127
->  3     3 -0.00113 -0.127
->  4     4  0.0483  -0.127
->  5     5  0.0977  -0.127
->  6     6  0.147   -0.127
->  7     7  0.197   -0.127
->  8     8  0.246   -0.127
->  9     9  0.295   -0.127
-> 10    10  0.345   -0.127
-> # i 890 more rows
+> 2     2 -0.0403 -0.127
+> 3     3  0.0193 -0.127
+> 4     4  0.0790 -0.127
+> 5     5  0.139  -0.127
 \end{verbatim}
@@
 >   h           x           y
-> 1 1 -0.10000000 -0.09858838
-> 2 1 -0.12471766 -0.11285913
-> 3 1 -0.12471766 -0.14140063
-> 4 1 -0.10000000 -0.15567138
-> 5 1 -0.07528234 -0.14140063
+> 1 1 -0.10000000 -0.09268324
+> 2 1 -0.12983166 -0.10990656
+> 3 1 -0.12983166 -0.14435320
+> 4 1 -0.10000000 -0.16157652
+> 5 1 -0.07016834 -0.14435320
 \end{verbatim}
@@
 >   <dbl> <dbl> <int> <int>
-> 1 0.300 0.985     1   659
-> 2 0.721 0.688     2   492
-> 3 0.722 0.542     3   418
-> 4 0.196 1.01      4   657
-> 5 0.240 1.06      5   708
+> 1 0.300 0.985     1   448
+> 2 0.721 0.688     2   351
+> 3 0.722 0.542     3   287
+> 4 0.196 1.01      4   468
+> 5 0.240 1.06      5   489
 \end{verbatim}
@@
 >   <int> <int> <dbl>
-> 1    82     2 0.002
-> 2    83     1 0.001
-> 3   106     1 0.001
-> 4   107     5 0.005
-> 5   108     7 0.007
+> 1    49     1 0.001
+> 2    68     1 0.001
+> 3    69     8 0.008
+> 4    70     5 0.005
+> 5    89     3 0.003
 \end{verbatim}
@@
 >   <int> <list>   
-> 1    82 <int [2]>
-> 2    83 <int [1]>
-> 3   106 <int [1]>
-> 4   107 <int [5]>
-> 5   108 <int [7]>
+> 1    49 <int [1]>
+> 2    68 <int [1]>
+> 3    69 <int [8]>
+> 4    70 <int [5]>
+> 5    89 <int [3]>
 \end{verbatim}
@@
 >   h         c_x        c_y n_h w_h
-> 1 1 -0.100000000 -0.1271299   0   0
-> 2 2 -0.050564674 -0.1271299   0   0
-> 3 3 -0.001129349 -0.1271299   0   0
-> 4 4  0.048305977 -0.1271299   0   0
-> 5 5  0.097741302 -0.1271299   0   0
+> 1 1 -0.10000000 -0.1271299   0   0
+> 2 2 -0.04033668 -0.1271299   0   0
+> 3 3  0.01932665 -0.1271299   0   0
+> 4 4  0.07898997 -0.1271299   0   0
+> 5 5  0.13865330 -0.1271299   0   0
 \end{verbatim}
@@
 >   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1     -0.127  -0.0506  -0.127           0        0
-> 2    26    27 -0.0753  -0.0843 -0.0258  -0.0843          0        0
-> 3     3    27 -0.00113 -0.127  -0.0258  -0.0843          0        0
-> 4    26    52 -0.0753  -0.0843 -0.0506  -0.0415          0        0
-> 5    52    53 -0.0506  -0.0415 -0.00113 -0.0415          0        0
+> 1     1     2 -0.1    -0.127  -0.0403 -0.127           0        0
+> 2    22    23 -0.0702 -0.0755 -0.0105 -0.0755          0        0
+> 3    22    44 -0.0702 -0.0755 -0.0403 -0.0238          0        0
+> 4     3    23  0.0193 -0.127  -0.0105 -0.0755          0        0
+> 5    44    45 -0.0403 -0.0238  0.0193 -0.0238          0        0
 \end{verbatim}
@@
 > # A tibble: 5 x 10
->    from    to   x_from  y_from     x_to    y_to from_count to_count
->   <int> <int>    <dbl>   <dbl>    <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1     -0.127  -0.0506  -0.127           0        0
-> 2    26    27 -0.0753  -0.0843 -0.0258  -0.0843          0        0
-> 3     3    27 -0.00113 -0.127  -0.0258  -0.0843          0        0
-> 4    26    52 -0.0753  -0.0843 -0.0506  -0.0415          0        0
-> 5    52    53 -0.0506  -0.0415 -0.00113 -0.0415          0        0
-> # i 2 more variables: from_reindexed <int>, to_reindexed <int>
+>    from    to  x_from  y_from    x_to    y_to from_count to_count from_reindexed
+>   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>          <int>
+> 1     1     2 -0.1    -0.127  -0.0403 -0.127           0        0              1
+> 2    22    23 -0.0702 -0.0755 -0.0105 -0.0755          0        0             22
+> 3    22    44 -0.0702 -0.0755 -0.0403 -0.0238          0        0             22
+> 4     3    23  0.0193 -0.127  -0.0105 -0.0755          0        0              3
+> 5    44    45 -0.0403 -0.0238  0.0193 -0.0238          0        0             44
+> # i 1 more variable: to_reindexed <int>
 \end{verbatim}
@@
 >   <int> <dbl> <dbl> <dbl>     <dbl>    <dbl>     <dbl>     <dbl>
-> 1    82 0.989  1.78  1.13  0.00470  0.0142   -0.0547  -0.00428 
-> 2    83 0.960  1.93  1.28  0.00278 -0.0180   -0.0678  -0.00260 
-> 3   106 0.997  1.37  1.07  0.00154  0.00687   0.0289   0.00760 
-> 4   107 0.975  1.46  1.20 -0.00497 -0.000595  0.0139   0.000537
-> 5   108 0.931  1.64  1.36 -0.00706  0.00487  -0.00705  0.000383
+> 1    49 0.999  1.93  1.05  0.00276   0.0133  -0.0530   -0.00270 
+> 2    68 0.997  1.37  1.07  0.00154   0.00687  0.0289    0.00760 
+> 3    69 0.960  1.58  1.27 -0.00709   0.00759 -0.000368 -0.000292
+> 4    70 0.935  1.81  1.35 -0.000338 -0.00713 -0.0278   -0.00474 
+> 5    89 0.985  1.13  1.14 -0.00597  -0.00830  0.0196   -0.00140
 \end{verbatim}
@@
 >        <dbl>      <dbl> <int>  <int>
-> 1      0.295      0.986     1    659
-> 2      0.716      0.686     2    492
-> 3      0.740      0.558     3    418
-> 4      0.197      0.986     4    657
-> 5      0.246      1.07      5    708
+> 1      0.318      1.01      1    470
+> 2      0.735      0.700     2    351
+> 3      0.705      0.545     3    287
+> 4      0.198      1.01      4    468
+> 5      0.228      1.06      5    489
 \end{verbatim}

### 2025-10-19 paper-quollr.R

@@ -436,10 +436,16 @@ scurve_proj_umap_model4 <- plot_proj(
 
 
 ## ----overview, eval=knitr::is_latex_output(), fig.cap="Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space. The left panel shows the NLDR layout with a triangular mesh overlay, forming the wireframe structure. This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved. Panels (a1–a4) display different \\twoD projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible. The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H'----
+
 hex_grid_scurve + wrap_plots(
   scurve_proj_umap_model1, scurve_proj_umap_model2,
   scurve_proj_umap_model3, scurve_proj_umap_model4, 
   ncol = 2)
+
+
+## ----workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow for "-------
+
+knitr::include_graphics("figures/quollr_workflow.png")
 
 
 ## ----algo-step-html, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions. The `Scurve` data is shown.", layout = "l-page"----
@@ -506,7 +512,7 @@ scurve_umap_obj <- gen_scaled_data(nldr_data = scurve_umap)
 ## ----echo=TRUE----------------------------------------------------------------
 bin_configs <- calc_bins_y(
   nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 25, 
+  b1 = 21, 
   q = 0.1)
 
 bin_configs
@@ -515,7 +521,7 @@ bin_configs
 ## ----echo=TRUE----------------------------------------------------------------
 hb_obj <- hex_binning(
   nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 25, 
+  b1 = 21, 
   q = 0.1)
 
 
@@ -679,11 +685,11 @@ hb_obj <- hex_binning(
 ## ----echo=TRUE----------------------------------------------------------------
[…]

### 2025-10-19 paper-quollr.Rmd
Added screeshots for high-de model

@@
 library(htmltools)
+library(detourr)
 
@@
 
-```{r, echo=TRUE}
+```{r, echo=TRUE, eval=FALSE}
 library(umap)
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output()}
-library(detourr)
+```{r, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+
+knitr::include_graphics(c("figures/model_proj1_detourr.png",
+                          "figures/model_proj2_detourr.png"))
+```
+
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout."}
 
@@
                     max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
+  show_scatter(axes = TRUE, size = 1, alpha = 0.8, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),

### 2025-10-19 paper-quollr.tex
Added screeshots for high-de model

@@
 > $b2
-> [1] 30
+> [1] 28
 > 
 > $a1
-> [1] 0.05966332
+> [1] 0.05869649
 > 
 > $a2
-> [1] 0.05166995
+> [1] 0.05083265
 \end{verbatim}
@@
 >   <int>   <dbl>  <dbl>
-> 1     1 -0.1    -0.127
-> 2     2 -0.0403 -0.127
-> 3     3  0.0193 -0.127
-> 4     4  0.0790 -0.127
-> 5     5  0.139  -0.127
+> 1     1 -0.1    -0.116
+> 2     2 -0.0413 -0.116
+> 3     3  0.0174 -0.116
+> 4     4  0.0761 -0.116
+> 5     5  0.135  -0.116
 \end{verbatim}
@@
 >   h           x           y
-> 1 1 -0.10000000 -0.09268324
-> 2 1 -0.12983166 -0.10990656
-> 3 1 -0.12983166 -0.14435320
-> 4 1 -0.10000000 -0.16157652
-> 5 1 -0.07016834 -0.14435320
+> 1 1 -0.10000000 -0.08179171
+> 2 1 -0.12934824 -0.09873593
+> 3 1 -0.12934824 -0.13262436
+> 4 1 -0.10000000 -0.14956858
+> 5 1 -0.07065176 -0.13262436
 \end{verbatim}
@@
 >   <dbl> <dbl> <int> <int>
-> 1 0.300 0.985     1   448
-> 2 0.721 0.688     2   351
-> 3 0.722 0.542     3   287
-> 4 0.196 1.01      4   468
-> 5 0.240 1.06      5   489
+> 1 0.277 0.913     1   427
+> 2 0.697 0.538     2   287
+> 3 0.779 0.399     3   226
+> 4 0.173 0.953     4   446
+> 5 0.218 0.983     5   468
 \end{verbatim}
@@
 >   <int> <int> <dbl>
-> 1    49     1 0.001
+> 1    58     4 0.004
 > 2    68     1 0.001
-> 3    69     8 0.008
-> 4    70     5 0.005
-> 5    89     3 0.003
+> 3    69     5 0.005
+> 4    70     6 0.006
+> 5    71     9 0.009
 \end{verbatim}
@@
 >   <int> <list>   
-> 1    49 <int [1]>
+> 1    58 <int [4]>
 > 2    68 <int [1]>
-> 3    69 <int [8]>
-> 4    70 <int [5]>
-> 5    89 <int [3]>
+> 3    69 <int [5]>
+> 4    70 <int [6]>
+> 5    71 <int [9]>
 \end{verbatim}
@@
 >   h         c_x        c_y n_h w_h
-> 1 1 -0.10000000 -0.1271299   0   0
-> 2 2 -0.04033668 -0.1271299   0   0
-> 3 3  0.01932665 -0.1271299   0   0
-> 4 4  0.07898997 -0.1271299   0   0
-> 5 5  0.13865330 -0.1271299   0   0
+> 1 1 -0.10000000 -0.1156801   0   0
+> 2 2 -0.04130351 -0.1156801   0   0
+> 3 3  0.01739298 -0.1156801   0   0
+> 4 4  0.07608947 -0.1156801   0   0
+> 5 5  0.13478596 -0.1156801   0   0
 \end{verbatim}
@@
 >   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
-> 1     1     2 -0.1    -0.127  -0.0403 -0.127           0        0
-> 2    22    23 -0.0702 -0.0755 -0.0105 -0.0755          0        0
-> 3    22    44 -0.0702 -0.0755 -0.0403 -0.0238          0        0
-> 4     3    23  0.0193 -0.127  -0.0105 -0.0755          0        0
-> 5    44    45 -0.0403 -0.0238  0.0193 -0.0238          0        0
+> 1     1     2 -0.1    -0.116  -0.0413 -0.116           0        0
+> 2    22    23 -0.0707 -0.0648 -0.0120 -0.0648          0        0
+> 3    22    44 -0.0707 -0.0648 -0.0413 -0.0140          0        0
+> 4     3    23  0.0174 -0.116  -0.0120 -0.0648          0        0
+> 5    44    45 -0.0413 -0.0140  0.0174 -0.0140          0        0
 \end{verbatim}
@@
 >   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>          <int>
-> 1     1     2 -0.1    -0.127  -0.0403 -0.127           0        0              1
-> 2    22    23 -0.0702 -0.0755 -0.0105 -0.0755          0        0             22
-> 3    22    44 -0.0702 -0.0755 -0.0403 -0.0238          0        0             22
-> 4     3    23  0.0193 -0.127  -0.0105 -0.0755          0        0              3
-> 5    44    45 -0.0403 -0.0238  0.0193 -0.0238          0        0             44
+> 1     1     2 -0.1    -0.116  -0.0413 -0.116           0        0              1
+> 2    22    23 -0.0707 -0.0648 -0.0120 -0.0648          0        0             22
+> 3    22    44 -0.0707 -0.0648 -0.0413 -0.0140          0        0             22
+> 4     3    23  0.0174 -0.116  -0.0120 -0.0648          0        0              3
+> 5    44    45 -0.0413 -0.0140  0.0174 -0.0140          0        0             44
 > # i 1 more variable: to_reindexed <int>
@@
 >   <int>  <dbl>  <dbl> <dbl>    <dbl>    <dbl>    <dbl>    <dbl>
-> 1    49 0.999  1.93  1.05  0.00276   0.0133  -0.0530   -0.00270 
-> 2    68 0.997  1.37  1.07  0.00154   0.00687  0.0289    0.00760 
-> 3    69 0.960  1.58  1.27 -0.00709   0.00759 -0.000368 -0.000292
-> 4    70 0.935  1.81  1.35 -0.000338 -0.00713 -0.0278   -0.00474 
-> 5    89 0.985  1.13  1.14 -0.00597  -0.00830  0.0196   -0.00140
+> 1    58 -0.371 1.91    1.92 -0.00827 0.00189   0.0170   0.00281
+> 2    68  0.958 0.0854  1.29  0.00265 0.0171    0.0876  -0.00249
+> 3    69  0.855 0.0917  1.51  0.00512 0.000325 -0.0130  -0.00395
+> 4    70  0.731 0.129   1.68 -0.00433 0.00211  -0.0356  -0.00240
+> 5    71  0.474 0.108   1.88 -0.00260 0.000128  0.00785  0.00170
 \end{verbatim}
@@
 >        <dbl>      <dbl> <int>  <int>
-> 1      0.318      1.01      1    470
-> 2      0.735      0.700     2    351
-> 3      0.705      0.545     3    287
-> 4      0.198      1.01      4    468
-> 5      0.228      1.06      5    489
+> 1      0.252      0.901     1    427
+> 2      0.692      0.545     2    287
+> 3      0.780      0.393     3    226
+> 4      0.164      0.952     4    446
+> 5      0.193      1.00      5    468
 \end{verbatim}
@@
 
-\begin{verbatim}
-library(detourr)
-
+\begin{figure}[H]
+\includegraphics[width=0.5\linewidth]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.}\label{fig:unnamed-chunk-30}
+\end{figure}
+
+\begin{verbatim}
 detour(
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-33-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-34-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
-While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-33}
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-34}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-37-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-38-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-37}
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-38}
 \end{figure}

### 2025-10-19 paper-quollr.R

@@ -53,6 +53,7 @@ library(readr)
 library(plotly)
 library(crosstalk)
 library(htmltools)
+library(detourr)
 
 set.seed(20240110)
 
@@ -480,20 +481,20 @@ scurve_umap_plt + hex_grid_poly_scurve +
 data("scurve")
 
 
-## ----echo=TRUE----------------------------------------------------------------
-library(umap)
-library(dplyr)
-
-scurve_umap <- umap(
-  scurve |> select(-ID),
-  config = modifyList(umap.defaults, list(
-    n_neighbors = 46,
-    n_components = 2,
-    min_dist = 0.9
-  ))
-)$layout |>
-  as_tibble(.name_repair = ~ paste0("emb", 1:2)) |>
-  mutate(ID = scurve$ID)
+## ----echo=TRUE, eval=FALSE----------------------------------------------------
+# library(umap)
+# library(dplyr)
+# 
+# scurve_umap <- umap(
+#   scurve |> select(-ID),
+#   config = modifyList(umap.defaults, list(
+#     n_neighbors = 46,
+#     n_components = 2,
+#     min_dist = 0.9
+#   ))
+# )$layout |>
+#   as_tibble(.name_repair = ~ paste0("emb", 1:2)) |>
+#   mutate(ID = scurve$ID)
 
 
 ## ----echo=TRUE, eval=FALSE----------------------------------------------------
@@ -874,8 +875,13 @@ scurve_proj_umap_model1 + scurve_proj_umap_model2 +
   plot_layout(ncol = 4)
 
 
-## ----echo=TRUE, eval=knitr::is_html_output()----------------------------------
-# library(detourr)
+## ----out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()----
+
+knitr::include_graphics(c("figures/model_proj1_detourr.png",
+                          "figures/model_proj2_detourr.png"))
+
+
+## ----echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout."----
 # 
 # detour(
 #   df_exe,
@@ -1308,7 +1314,7 @@ nldr_plt <- ggplotly(nldr_plt, width = 600,
 # ) |>
 #   tour_path(grand_tour(2),
[…]

### 2025-10-19 paper-quollr.Rmd
Added screenshots for detourr outputs

@@
 
-```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
+```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.", layout = "l-page"}
 
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.", layout = "l-page"}
 
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), layout = "l-page"}
+```{r, echo=TRUE, fig.cap="Link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", eval=knitr::is_html_output(), layout = "l-page"}
 
@@
   )
+```
+
+```{r, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+
+knitr::include_graphics(c("figures/model_link_proj1_detourr.png",
+                          "figures/model_link_proj2_detourr.png"))
 ```
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), layout = "l-page"}
+```{r, echo=TRUE, fig.cap="Link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", eval=knitr::is_html_output(), layout = "l-page"}
 
@@
                     max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
+  show_scatter(axes = TRUE, size = 1, alpha = 0.8, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
                palette = c("#66B2CC", "#FF7755"),
-                width = "600px", height = "600px")
+                width = "500px", height = "500px")
 
@@
 )
+```
+
+```{r, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+
+knitr::include_graphics(c("figures/model_link_error_proj1_detourr.png",
+                          "figures/model_link_error_proj2_detourr.png"))
 ```

### 2025-10-19 paper-quollr.tex
Added screenshots for detourr outputs

@@
 
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-41}
+\end{figure}
+
 For analyses that include model error visualization, a three-panel view can be constructed by adding the error distribution plot (\texttt{error\_plt}):
@@
 \end{verbatim}
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-44}
+\end{figure}
 

### 2025-10-19 paper-quollr.Rmd
Updated software architeture section

@@
 
-1. Data preprocessing — The function `gen_scaled_data()` standardizes the embedding data, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
-
-2. Construct \twoD model — `hex_binning()`, `merge_hexbin_centroids()`, `tri_bin_centroids()`, `gen_edges()`, `update_trimesh_index()`
-
-3. Lift the model into \pD{} — the `avg_highd_data()`
-
-4. Prediction - `predict_emb()`
-
-5. Error computation — `glance()`
-
-6. Visualization — `geom_hexgrid()`, `geom_trimesh()`, `show_langevitour()`
-
-7. Interactivity - `show_link_plots()`, and `show_error_link_plots()`
-
-Each module is internally independent but communicates through standardized data objects (see next section).
-This modular design simplifies maintenance and allows developers to extend individual components such as substituting different binning strategies, embedding methods, or visualization tools without altering the overall workflow.
+1. Data preprocessing: The function `gen_scaled_data()` standardizes the embedding data, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+
+2.
+Construct \twoD model: A series of functions `hex_binning()`, `merge_hexbin_centroids()`, `tri_bin_centroids()`, `gen_edges()`, and `update_trimesh_index()` generate the hexagonal grid, compute bin centroids, and connect the triangular mesh that defines local neighborhoods in the \twoD space.
+
+3. Lift the model into \pD: The function `avg_highd_data()` computes the average of the high-dimensional variables for each bin, linking the \twoD representation back to the original data space.
+
+4. Prediction: The function `predict_emb()` estimates the embedding of new high-dimensional observations based on the fitted model.
+
+5. Error computation: The `glance()` and `augment()` function summarizes model performance by comparing the predicted and original embeddings.
+
+6. Visualization: Functions such as `geom_hexgrid()`, `geom_trimesh()`, and `show_langevitour()` provide tools for exploring the fitted models through static and dynamic visualizations.
+
+7.
+Interactivity: The functions `comb_all_data_model()` and `show_link_plots()` generate interactive linked visualizations that connect the \twoD NLDR layout, the corresponding tour view, and the fitted model.
+Similarly, `comb_all_data_model_error()` and `show_error_link_plots()` integrate the error distribution with the \twoD embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
+
+Each module is internally independent but connected through data objects (see next section).
+This modular design simplifies maintenance and allows developers to extend individual components such as substituting different binning approaches, extracting centroids, or visualization tools without altering the overall workflow.
 
@@
 
-- `highd_data` – a tibble containing the original high-dimensional observations with a unique identifier (`ID`) and variable columns prefixed with `"x"` (e.g., `x1`, `x2`, …).
-
-- `nldr_data` – a tibble containing two-dimensional embeddings, labeled as `emb1` and `emb2`, matched to the same `ID`s.
+- `highd_data`: a tibble containing the original high-dimensional observations with a unique identifier (`ID`) and variable columns prefixed with `"x"` (e.g., `x1`, `x2`, …).
+
+- `nldr_data`: a tibble containing two-dimensional embeddings, labeled as `emb1` and `emb2`, matched to the same `ID`s.
 
@@
 
-- `scaled_nldr_obj` — the output of `gen_scaled_data()`, which rescales the embedding to the range $[0, 1] \times [0, y_{2,\text{max}}]$, where $y_{2,\text{max}} = r_2 / r_1$ is the ratio of the embedding ranges.
+- `scaled_nldr_obj`: the output of `gen_scaled_data()`, which rescales the embedding to the range $[0, 1] \times [0, y_{2,\text{max}}]$, where $y_{2,\text{max}} = r_2 / r_1$ is the ratio of the embedding ranges.
 It includes the scaled coordinates (`scaled_nldr`) and the original limits (`lim1`, `lim2`).
 
-- `hex_bin_obj` – the object created by `hex_binning()`, which defines the structure of the two-dimensional hexagonal grid used in modeling.
+- `hex_bin_obj`: the object created by `hex_binning()`, which defines the structure of the two-dimensional hexagonal grid used in modeling.
 It includes the grid spacing (`a1`, `a2`), the number of bins along each axis, the centroids of all hexagons, polygon coordinates for plotting, and the mapping of each data point to its assigned hexagon.
 
-- `highd_vis_model` — the main model object returned by `fit_highd_model()`.
+- `highd_vis_model`: the main model object returned by `fit_highd_model()`.
 It stores all components of the fitted visualization model, including the scaled NLDR data (`nldr_scaled_obj`), the hexagonal bin structure (`hb_obj`), the averaged \pD summaries for each bin (`model_highd`), the corresponding \twoD bin centroids (`model_2d`), and the triangulated mesh connecting neighboring bins (`trimesh_data`).

### 2025-10-19 paper-quollr.tex
Updated software architeture section

@@
 \item
-  Data preprocessing --- The function \texttt{gen\_scaled\_data()} standardizes the embedding data, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
-\item
-  Construct $2\text{-}D$ model --- \texttt{hex\_binning()}, \texttt{merge\_hexbin\_centroids()}, \texttt{tri\_bin\_centroids()}, \texttt{gen\_edges()}, \texttt{update\_trimesh\_index()}
-\item
-  Lift the model into $p\text{-}D${} --- the \texttt{avg\_highd\_data()}
-\item
-  Prediction - \texttt{predict\_emb()}
-\item
-  Error computation --- \texttt{glance()}
-\item
-  Visualization --- \texttt{geom\_hexgrid()}, \texttt{geom\_trimesh()}, \texttt{show\_langevitour()}
-\item
-  Interactivity - \texttt{show\_link\_plots()}, and \texttt{show\_error\_link\_plots()}
+  Data preprocessing: The function \texttt{gen\_scaled\_data()} standardizes the embedding data, manage variable naming, and ensure consistent identifiers across high-dimensional and embedded datasets.
+\item
+  Construct $2\text{-}D$ model: A series of functions \texttt{hex\_binning()}, \texttt{merge\_hexbin\_centroids()}, \texttt{tri\_bin\_centroids()}, \texttt{gen\_edges()}, and \texttt{update\_trimesh\_index()} generate the hexagonal grid, compute bin centroids, and connect the triangular mesh that defines local neighborhoods in the $2\text{-}D$ space.
+\item
+  Lift the model into $p\text{-}D$: The function \texttt{avg\_highd\_data()} computes the average of the high-dimensional variables for each bin, linking the $2\text{-}D$ representation back to the original data space.
+\item
+  Prediction: The function \texttt{predict\_emb()} estimates the embedding of new high-dimensional observations based on the fitted model.
+\item
+  Error computation: The \texttt{glance()} and \texttt{augment()} function summarizes model performance by comparing the predicted and original embeddings.
+\item
+  Visualization: Functions such as \texttt{geom\_hexgrid()}, \texttt{geom\_trimesh()}, and \texttt{show\_langevitour()} provide tools for exploring the fitted models through static and dynamic visualizations.
+\item
+  Interactivity: The functions \texttt{comb\_all\_data\_model()} and \texttt{show\_link\_plots()} generate interactive linked visualizations that connect the $2\text{-}D$ NLDR layout, the corresponding tour view, and the fitted model.
+Similarly, \texttt{comb\_all\_data\_model\_error()} and \texttt{show\_error\_link\_plots()} integrate the error distribution with the $2\text{-}D$ embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
 \end{enumerate}
 
-Each module is internally independent but communicates through standardized data objects (see next section).
-This modular design simplifies maintenance and allows developers to extend individual components such as substituting different binning strategies, embedding methods, or visualization tools without altering the overall workflow.
+Each module is internally independent but connected through data objects (see next section).
+This modular design simplifies maintenance and allows developers to extend individual components such as substituting different binning approaches, extracting centroids, or visualization tools without altering the overall workflow.
 
@@
 \item
-  \texttt{highd\_data} -- a tibble containing the original high-dimensional observations with a unique identifier (\texttt{ID}) and variable columns prefixed with \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, \ldots).
-\item
-  \texttt{nldr\_data} -- a tibble containing two-dimensional embeddings, labeled as \texttt{emb1} and \texttt{emb2}, matched to the same \texttt{ID}s.
+  \texttt{highd\_data}: a tibble containing the original high-dimensional observations with a unique identifier (\texttt{ID}) and variable columns prefixed with \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, \ldots).
+\item
+  \texttt{nldr\_data}: a tibble containing two-dimensional embeddings, labeled as \texttt{emb1} and \texttt{emb2}, matched to the same \texttt{ID}s.
 \end{itemize}
@@
 \item
-  \texttt{scaled\_nldr\_obj} --- the output of \texttt{gen\_scaled\_data()}, which rescales the embedding to the range \([0, 1] \times [0, y_{2,\text{max}}]\), where \(y_{2,\text{max}} = r_2 / r_1\) is the ratio of the embedding ranges.
+  \texttt{scaled\_nldr\_obj}: the output of \texttt{gen\_scaled\_data()}, which rescales the embedding to the range \([0, 1] \times [0, y_{2,\text{max}}]\), where \(y_{2,\text{max}} = r_2 / r_1\) is the ratio of the embedding ranges.
 It includes the scaled coordinates (\texttt{scaled\_nldr}) and the original limits (\texttt{lim1}, \texttt{lim2}).
 \item
-  \texttt{hex\_bin\_obj} -- the object created by \texttt{hex\_binning()}, which defines the structure of the two-dimensional hexagonal grid used in modeling.
+  \texttt{hex\_bin\_obj}: the object created by \texttt{hex\_binning()}, which defines the structure of the two-dimensional hexagonal grid used in modeling.
 It includes the grid spacing (\texttt{a1}, \texttt{a2}), the number of bins along each axis, the centroids of all hexagons, polygon coordinates for plotting, and the mapping of each data point to its assigned hexagon.
 \item
-  \texttt{highd\_vis\_model} --- the main model object returned by \texttt{fit\_highd\_model()}.
+  \texttt{highd\_vis\_model}: the main model object returned by \texttt{fit\_highd\_model()}.
 It stores all components of the fitted visualization model, including the scaled NLDR data (\texttt{nldr\_scaled\_obj}), the hexagonal bin structure (\texttt{hb\_obj}), the averaged $p\text{-}D$ summaries for each bin (\texttt{model\_highd}), the corresponding $2\text{-}D$ bin centroids (\texttt{model\_2d}), and the triangulated mesh connecting neighboring bins (\texttt{trimesh\_data}).

### 2025-10-21 paper-quollr.Rmd
Started editing HTML page

@@
 This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
-Starting with an NLDR layout, our approach is to create a \twoD wireframe model representation, that can be lifted and displayed in the high-dimensional (\pD) space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:overview)"})`).
+Starting with an NLDR layout, our approach is to create a \twoD wireframe model representation, that can be lifted and displayed in the high-dimensional (\pD{}) space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:overview)"})`).
 

### 2025-10-21 paper-quollr.tex
Started editing HTML page

@@
 This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
-Starting with an NLDR layout, our approach is to create a $2\text{-}D$ wireframe model representation, that can be lifted and displayed in the high-dimensional ($p\text{-}D$) space (Figure \ref{fig:overview}).
+Starting with an NLDR layout, our approach is to create a $2\text{-}D$ wireframe model representation, that can be lifted and displayed in the high-dimensional ($p\text{-}D${}) space (Figure \ref{fig:overview}).
 

### 2025-10-21 paper-quollr.Rmd
Updated Figure 1

@@
 scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
@@
 hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
@@
 hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
@@
 wireframe_scurve_int <- ggplotly(wireframe_scurve, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
@@
                          enableControls = FALSE,
-                         width = "421px", height = "421px")
+                         width = "400px", height = "350px")
 
@@
                          enableControls = FALSE,
-                         width = "300px", height = "300px")
+                         width = "300px", height = "350px")
 
@@
   htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr;",
-    hex_grid_scurve_int,
-    htmltools::div(style = "margin-top: 13px;", scurve_umap_model_vis)
+    style = "
+      display: flex;
+      justify-content: center;  /* centers the whole grid */
+      margin: 0 auto;
+      padding: 0;
+    ",
+    htmltools::div(
+      style = "
+        display: grid;
+        grid-template-columns: 1fr 1fr;
+        gap: 0px;
+        align-items: start;
+        justify-items: center;
+        margin: 0;
+        padding: 0;
+      ",
+      htmltools::div(style = "margin: 0; padding: 0;", hex_grid_scurve_int),
+      htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_model_vis)
+    )
   ),
@@
 
+# Ensure it's treated as a widget
 class(overviewfig) <- c(class(overviewfig), "htmlwidget")
@@
 
-algofig <- crosstalk::bscols(
-        htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr 1fr 1fr 1fr; gap: 0;",
-                       htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_plt_int),
-                       htmltools::div(style = "margin: 0; padding: 0;", hex_grid_poly_scurve_int),
-                       htmltools::div(style = "margin: 0; padding: 0;", hex_centroids_scurve_int),
-                       htmltools::div(style = "margin: 0; padding: 0;", wireframe_scurve_int),
+algofig <- bscols(
   htmltools::div(
-                         style = "margin-top: 13px;t",
-                         scurve_umap_model_vis_n
-                       )),
+    style = "
+      display: flex;
+      justify-content: center;     /* centers the grid horizontally */
+      align-items: center;          /* optional: centers vertically if space allows */
+      margin: 0 auto;
+      padding: 0;
+    ",
+    htmltools::div(
+      style = "
+        display: grid;
+        grid-template-columns: 1fr 1fr;
+        gap: 0px;
+        align-items: start;
+        justify-items: center;
+        margin: 0;
+        padding: 0;
+      ",
+      htmltools::div(style = 'margin: 0; padding: 0;', hex_grid_poly_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0;', hex_centroids_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0;', wireframe_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0;', scurve_umap_model_vis_n)
+    )
+  ),
   device = "xs"
@@
 
+# Treat as HTML widget
 class(algofig) <- c(class(algofig), "htmlwidget")
@@
 
-scurve_umap_plt + hex_grid_poly_scurve +
+hex_grid_poly_scurve +
   hex_centroids_scurve + wireframe_scurve + 
   scurve_proj_umap_model1_dp +
-  plot_layout(ncol = 5)
+  plot_layout(ncol = 4)
 ```

### 2025-10-21 paper-quollr.R

@@ -127,8 +127,8 @@ scurve_umap_plt <- ggplot(
   ylim(sc_ylims)
 
 scurve_umap_plt_int <- ggplotly(scurve_umap_plt, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
   config(
     staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
@@ -153,8 +153,8 @@ hex_grid_poly_scurve <- ggplot(
   ylim(sc_ylims)
 
 hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
   config(
     staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
@@ -179,8 +179,8 @@ hex_centroids_scurve <- ggplot(
   ylim(sc_ylims)
 
 hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
   config(
     staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
@@ -207,8 +207,8 @@ wireframe_scurve <- ggplot() +
   ylim(sc_ylims)
 
 wireframe_scurve_int <- ggplotly(wireframe_scurve, 
-                                width = "230", 
-                                height = "230", 
+                                width = "350", 
+                                height = "350", 
                                 tooltip = "none") |>
   config(
     staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
@@ -271,7 +271,7 @@ scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
                          levelColors = c(clr_choice, "#FF7755"),
                          enableControls = FALSE,
-                         width = "421px", height = "421px")
[…]

### 2025-10-21 paper-quollr.Rmd
Updated Figure 1

@@
                          enableControls = FALSE,
-                         width = "400px", height = "350px")
+                         width = "400px", height = "380px")
 
@@
       htmltools::div(style = "margin: 0; padding: 0;", hex_grid_scurve_int),
-      htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_model_vis)
+      htmltools::div(style = "margin: 0; margin-top: 13px; padding: 0;", scurve_umap_model_vis)
     )

### 2025-10-21 paper-quollr.R

@@ -271,7 +271,7 @@ scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
                          levelColors = c(clr_choice, "#FF7755"),
                          enableControls = FALSE,
-                         width = "400px", height = "350px")
+                         width = "400px", height = "380px")
 
 scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          lineFrom = tr_from_to_df_scurve$from_reindexed,
@@ -305,7 +305,7 @@ scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)]
 #         padding: 0;
 #       ",
 #       htmltools::div(style = "margin: 0; padding: 0;", hex_grid_scurve_int),
-#       htmltools::div(style = "margin: 0; padding: 0;", scurve_umap_model_vis)
+#       htmltools::div(style = "margin: 0; margin-top: 13px; padding: 0;", scurve_umap_model_vis)
 #     )
 #   ),
 #   device = "xs"

### 2025-10-21 paper-quollr.Rmd
Updated Figure 3

@@
 hex_grid_poly_scurve_int <- ggplotly(hex_grid_poly_scurve, 
-                                width = "350", 
-                                height = "350", 
+                                width = "400", 
+                                height = "380", 
                                 tooltip = "none") |>
@@
 hex_centroids_scurve_int <- ggplotly(hex_centroids_scurve, 
-                                width = "350", 
-                                height = "350", 
+                                width = "400", 
+                                height = "380", 
                                 tooltip = "none") |>
@@
                                 width = "350", 
-                                height = "350", 
+                                height = "320", 
                                 tooltip = "none") |>
@@
                                 width = "350", 
-                                height = "350", 
+                                height = "380", 
                                 tooltip = "none") |>
@@
                          enableControls = FALSE,
-                         width = "300px", height = "350px")
+                         width = "400px", height = "380px")
 
@@
 
-```{r algo-step-html, eval=knitr::is_html_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+```{r algo-step-html, eval=knitr::is_html_output(), fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
 The `Scurve` data is shown.", layout = "l-page"}
@@
       ",
-      htmltools::div(style = 'margin: 0; padding: 0;', hex_grid_poly_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0;', hex_centroids_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0;', wireframe_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0;', scurve_umap_model_vis_n)
+      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 225px;', hex_grid_poly_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px;', hex_centroids_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 280px; width: 225px;', wireframe_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; margin-left: 60px; height: 300px;', scurve_umap_model_vis_n)
     )

### 2025-10-22 paper-quollr.Rmd
Added Figure 1 title

@@
 
-```{r overviewhtml, eval=knitr::is_html_output(), fig.cap="algorithm", fig.pos='H', layout = "l-page"}
+```{r overviewhtml, eval=knitr::is_html_output(), fig.cap="Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space.
+The left panel shows the NLDR layout with a triangular mesh overlay, forming the wireframe structure.
+This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
+Panels (a1–a4) display different \\twoD projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', layout = "l-page"}
 
@@
 The source code for reproducing this paper can be found at: [https://github.com/JayaniLakshika/paper-quollr](https://github.com/JayaniLakshika/paper-quollr).
-
 This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.

### 2025-10-22 paper-quollr.Rmd
Updated Figure 6 title

@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout."}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+These views illustrate how the lifted wireframe model captures the structure of the Scurve data.
+The two twists visible in the UMAP layout can also be seen in the lifted model."}
 

### 2025-10-22 paper-quollr.R

@@ -914,7 +914,7 @@ knitr::include_graphics(c("figures/model_proj1_detourr.png",
                           "figures/model_proj2_detourr.png"))
 
 
-## ----echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout."----
+## ----echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout. These views illustrate how the lifted wireframe model captures the structure of the Scurve data. The two twists visible in the UMAP layout can also be seen in the lifted model."----
 # 
 # detour(
 #   df_exe,

### 2025-10-23 paper-quollr.Rmd
Updated Figure 2 title

@@
 
-```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow for "}
+```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Overview of the `quollr` workflow and software architecture.
+The process begins with NLDR and $p\\text{-}D$ data inputs, followed by data preprocessing and hexagonal binning.
+Centroids are computed and triangulated to form the $2\\text{-}D$ mesh, which is then lifted into the $p\\text{-}D$ space.
+Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\\text{-}D$ and $2\\text{-}D$ representations."}
 

### 2025-10-23 paper-quollr.tex
Added Figure 1 title
Updated Figure 2 title

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/quollr_workflow} \caption{Workflow for }\label{fig:workflow}
+\includegraphics[width=1\linewidth]{figures/quollr_workflow} \caption{Overview of the `quollr` workflow and software architecture.
+The process begins with NLDR and $p\text{-}D$ data inputs, followed by data preprocessing and hexagonal binning.
+Centroids are computed and triangulated to form the $2\text{-}D$ mesh, which is then lifted into the $p\text{-}D$ space.
+Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\text{-}D$ and $2\text{-}D$ representations.}\label{fig:workflow}
 \end{figure}
@@
 The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.
-
 This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.

### 2025-10-24 paper-quollr.Rmd
Updated Figure 7 title.

@@
 ```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
-Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (in purple) in different projections.
+Each panel (a1–a4) shows the model (blue) overlaid on `Scurve` data (black) in different projections.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
@@
 
-```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.", layout = "l-page"}
+```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the NLDR layout preserves the underlying `Scurve` structure.
+Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"}
 
@@
 
-```{r, echo=TRUE, fig.cap="Link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", eval=knitr::is_html_output(), layout = "l-page"}
+```{r, echo=TRUE}
 
@@
 
+```
+
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
+Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the NLDR layout preserves the underlying `Scurve` structure.
+To highlight points in `detourr`, use the Select button.
+Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"}
+
 crosstalk::bscols(
@@
 
-```{r, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold'}
 
@@
 
-```{r, echo=TRUE, fig.cap="Link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", eval=knitr::is_html_output(), layout = "l-page"}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", layout = "l-page"}
 

### 2025-10-24 paper-quollr.tex
Updated Figure 7 title.

@@
 \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-29-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
-Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (in purple) in different projections.
+Each panel (a1–a4) shows the model (blue) overlaid on `Scurve` data (black) in different projections.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
@@
                 width = "600px", height = "600px")
-
+\end{verbatim}
+
+\begin{verbatim}
 crosstalk::bscols(
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-41}
+\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-42}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-44}
+\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-45}
 \end{figure}

### 2025-10-24 paper-quollr.Rmd
Updated Figure 8 caption

@@
 
-```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
-Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the NLDR layout preserves the underlying `Scurve` structure.
+```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure.
 Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"}
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.", layout = "l-page"}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
+Points with higher residuals typically correspond to sparse or bridging regions in the UMAP layout and the less dense ends of the Scurve, whereas low-residual points tend to cluster within denser, well-preserved regions.
+Double-clicking clears the selection and resets all highlights.", layout = "l-page"}
 
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
-Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the NLDR layout preserves the underlying `Scurve` structure.
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
+Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure.
 To highlight points in `detourr`, use the Select button.
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", layout = "l-page"}
+```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
+Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
+Points with higher residuals typically correspond to sparse or bridging regions in the UMAP layout and the less dense ends of the Scurve, whereas low-residual points tend to cluster within denser, well-preserved regions.
+To highlight points in `detourr`, use the Select button.
+Double-clicking clears the selection and resets all highlights.", layout = "l-page"}
 

### 2025-10-24 paper-quollr.Rmd
Start updating figure captions

@@
 
-```{r model-limb, fig.cap="Compare the published $2-\\text{D}$ layout (Figure \\ref{fig:limb-rwbss}b) and the $2-\\text{D}$ layout selected (Figure \\ref{fig:limb-rwbss}f) by RWBSS plot (Figure \\ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-rwbss}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-rwbss}f) by RWBSS plot (Figure \\ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
@@
 
-```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
+```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout b (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
+Each color represents a different muscle group, with triangulated hexagon centroids on the NLDR layout connected to their corresponding regions in the high-dimensional projection using `langevitour`.
+Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
+Double-clicking clears the selection and removes highlights.", layout = "l-page"}
 
@@
 
-```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right).", layout = "l-page"}
+```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout f (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
+Each color represents a different muscle group, with triangulated hexagon centroids on the NLDR layout connected to their corresponding regions in the high-dimensional projection using `langevitour`.
+Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
+Double-clicking clears the selection and removes highlights.", layout = "l-page"}
 

### 2025-10-24 paper-quollr.tex
Start updating figure captions

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2-\text{D}$ layout (Figure \ref{fig:limb-rwbss}b) and the $2-\text{D}$ layout selected (Figure \ref{fig:limb-rwbss}f) by RWBSS plot (Figure \ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-rwbss}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-rwbss}f) by RWBSS plot (Figure \ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-10-24 paper-quollr.Rmd
Fixed the detourr caption problem

@@
 
-crosstalk::bscols(
+lndet_link <- crosstalk::bscols(
     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
@@
   )
+
+class(lndet_link) <- c(class(lndet_link), "htmlwidget")
+
+lndet_link
 ```
@@
 
-crosstalk::bscols(
+erlndet_link <- crosstalk::bscols(
   htmltools::div(
@@
 )
+
+class(erlndet_link) <- c(class(erlndet_link), "htmlwidget")
+
+erlndet_link
 ```

### 2025-10-24 paper-quollr.tex
Fixed the detourr caption problem

@@
 \begin{verbatim}
-crosstalk::bscols(
+lndet_link <- crosstalk::bscols(
     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
@@
   )
+
+class(lndet_link) <- c(class(lndet_link), "htmlwidget")
+
+lndet_link
 \end{verbatim}
@@
 )
+
+class(erlndet_link) <- c(class(erlndet_link), "htmlwidget")
+
+erlndet_link
 \end{verbatim}

### 2025-10-24 paper-quollr.R

@@ -1356,13 +1356,18 @@ detourr_output <- detour(
 
 ## ----echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`. Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure. To highlight points in `detourr`, use the Select button. Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"----
 # 
-# crosstalk::bscols(
+# lndet_link <- crosstalk::bscols(
 #     htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
 #                    nldr_plt,
 #                    htmltools::div(style = "margin-top: 20px;", detourr_output)
 #     ),
 #     device = "xs"
 #   )
+# 
+# 
+# class(lndet_link) <- c(class(lndet_link), "htmlwidget")
+# 
+# lndet_link
 
 
 ## ----eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold'----
@@ -1450,7 +1455,7 @@ nldr_plt <- ggplotly(nldr_plt, width = 400,
 #                palette = c("#66B2CC", "#FF7755"),
 #                 width = "500px", height = "500px")
 # 
-# crosstalk::bscols(
+# erlndet_link <- crosstalk::bscols(
 #   htmltools::div(
 #     style = "display: grid; grid-template-columns: 1fr 1fr 1fr;",
 #     error_plt,
@@ -1459,6 +1464,10 @@ nldr_plt <- ggplotly(nldr_plt, width = 400,
 #   ),
 #   device = "xs"
 # )
+# 
+# class(erlndet_link) <- c(class(erlndet_link), "htmlwidget")
+# 
+# erlndet_link
 
 
 ## ----out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()----

### 2025-10-24 paper-quollr.Rmd
Updated Figure 3

@@
              alpha = 0.2, size = 0.5, color = clr_choice) + 
-  interior_annotation("b", cex = 3) +
+  interior_annotation("a", cex = 3) +
   xlim(sc_xlims) +
@@
              size = 1, color = "#FF7755") +
-  interior_annotation("c", cex = 3) +
+  interior_annotation("b", cex = 3) +
   xlim(sc_xlims) +
@@
              size = 1, color = "#FF7755") +
-  interior_annotation("d", cex = 3) +
+  interior_annotation("c", cex = 3) +
   xlim(sc_xlims) +
@@
   is_category = FALSE) +
-  interior_annotation(label = "e", cex = 3)
+  interior_annotation(label = "d", cex = 3)
 
@@
 
-```{r algo-step-html, eval=knitr::is_html_output(), fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+```{r algo-step-html, eval=knitr::is_html_output(), fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `Scurve` data is shown.", layout = "l-page"}
@@
 
-```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `Scurve` data is shown."}

### 2025-10-24 paper-quollr.tex
Updated Figure 3

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `Scurve` data is shown.}\label{fig:algo-steps}

### 2025-10-24 paper-quollr.R

@@ -148,7 +148,7 @@ hex_grid_poly_scurve <- ggplot(
   geom_point(data = scurve_umap_scaled, 
              aes(x = emb1, y = emb2), 
              alpha = 0.2, size = 0.5, color = clr_choice) + 
-  interior_annotation("b", cex = 3) +
+  interior_annotation("a", cex = 3) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -174,7 +174,7 @@ hex_centroids_scurve <- ggplot(
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
              size = 1, color = "#FF7755") +
-  interior_annotation("c", cex = 3) +
+  interior_annotation("b", cex = 3) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -202,7 +202,7 @@ wireframe_scurve <- ggplot() +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
              size = 1, color = "#FF7755") +
-  interior_annotation("d", cex = 3) +
+  interior_annotation("c", cex = 3) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -375,7 +375,7 @@ scurve_proj_umap_model1_dp <- plot_proj(
   plot_limits = c(-0.35, 0.35), 
   axis_text_size = 7,
   is_category = FALSE) +
-  interior_annotation(label = "e", cex = 3)
+  interior_annotation(label = "d", cex = 3)
 
 ## Second projection
 
@@ -466,7 +466,7 @@ hex_grid_scurve + wrap_plots(
 knitr::include_graphics("figures/quollr_workflow.png")
 
 
-## ----algo-step-html, eval=knitr::is_html_output(), fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) NLDR data, (b) hexagon bins, (c) bin centroids, (d) triangulated centroids, and (e) lifting the model into high dimensions. The `Scurve` data is shown.", layout = "l-page"----
+## ----algo-step-html, eval=knitr::is_html_output(), fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions. The `Scurve` data is shown.", layout = "l-page"----
 # 
 # algofig <- bscols(
 #   htmltools::div(
[…]

### 2025-10-24 paper-quollr.Rmd
Updated Figure 5, 6 captions

@@
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-The two twists visible in the UMAP layout can also be seen in the lifted model."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model."}
 
@@
 ```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
-Each panel (a1–a4) shows the model (blue) overlaid on `Scurve` data (black) in different projections.
+Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-The two twists visible in the UMAP layout can also be seen in the lifted model."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model."}
 
@@
 
-```{r, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 

### 2025-10-24 paper-quollr.tex
Updated Figure 5, 6 captions

@@
 \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-29-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
-Each panel (a1–a4) shows the model (blue) overlaid on `Scurve` data (black) in different projections.
+Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-The two twists visible in the UMAP layout can also be seen in the lifted model.}\label{fig:unnamed-chunk-29}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:unnamed-chunk-29}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=0.5\linewidth]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.}\label{fig:unnamed-chunk-30}
+\includegraphics[width=0.5\linewidth]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:unnamed-chunk-30}
 \end{figure}

### 2025-10-25 paper-quollr.Rmd
Updated Figure 6 title
Updated some figure captions

@@
 These views illustrate how the lifted wireframe model captures the structure of the Scurve data.
-The two twists visible in the UMAP layout can also be seen in the lifted model."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model."}
 
@@
 
-```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout b (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
+```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the layout b (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
 Each color represents a different muscle group, with triangulated hexagon centroids on the NLDR layout connected to their corresponding regions in the high-dimensional projection using `langevitour`.
@@
 
-```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the NLDR layout f (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
+```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the layout f (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
 Each color represents a different muscle group, with triangulated hexagon centroids on the NLDR layout connected to their corresponding regions in the high-dimensional projection using `langevitour`.

### 2025-11-08 paper-quollr.Rmd
Changed the notations to view in HTML as well

@@
   \usepackage{float}
-  \newcommand{\pD}{$p\text{-}D$}
-  \newcommand{\twoD}{$2\text{-}D$}
 output: 
@@
 This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
-Starting with an NLDR layout, our approach is to create a \twoD wireframe model representation, that can be lifted and displayed in the high-dimensional (\pD{}) space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:overview)"})`).
+Starting with an NLDR layout, our approach is to create a $2\text{-}D$ wireframe model representation, that can be lifted and displayed in the high-dimensional ($p\text{-}D$) space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:overviewhtml)"} else { "Figure \\@ref(fig:overview)"})`).
 
@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different \\twoD projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different $2\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', layout = "l-page"}
@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different \\twoD projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different $2\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H'}
@@
 
-The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, \twoD model construction, lifting into \pD, prediction, error computation, visualization, and interactivity.
+The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, $2\text{-}D$ model construction, lifting into $p\text{-}D$, prediction, error computation, visualization, and interactivity.
 Each module performs a distinct task and communicates through data objects.
@@
 2.
-Construct \twoD model: A series of functions `hex_binning()`, `merge_hexbin_centroids()`, `tri_bin_centroids()`, `gen_edges()`, and `update_trimesh_index()` generate the hexagonal grid, compute bin centroids, and connect the triangular mesh that defines local neighborhoods in the \twoD space.
-
-3. Lift the model into \pD: The function `avg_highd_data()` computes the average of the high-dimensional variables for each bin, linking the \twoD representation back to the original data space.
+Construct $2\text{-}D$ model: A series of functions `hex_binning()`, `merge_hexbin_centroids()`, `tri_bin_centroids()`, `gen_edges()`, and `update_trimesh_index()` generate the hexagonal grid, compute bin centroids, and connect the triangular mesh that defines local neighborhoods in the $2\text{-}D$ space.
+
+3.
+Lift the model into $p\text{-}D$: The function `avg_highd_data()` computes the average of the high-dimensional variables for each bin, linking the $2\text{-}D$ representation back to the original data space.
 
@@
 7.
-Interactivity: The functions `comb_all_data_model()` and `show_link_plots()` generate interactive linked visualizations that connect the \twoD NLDR layout, the corresponding tour view, and the fitted model.
-Similarly, `comb_all_data_model_error()` and `show_error_link_plots()` integrate the error distribution with the \twoD embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
+Interactivity: The functions `comb_all_data_model()` and `show_link_plots()` generate interactive linked visualizations that connect the $2\text{-}D$ NLDR layout, the corresponding tour view, and the fitted model.
+Similarly, `comb_all_data_model_error()` and `show_error_link_plots()` integrate the error distribution with the $2\text{-}D$ embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
 
@@
 - `highd_vis_model`: the main model object returned by `fit_highd_model()`.
-It stores all components of the fitted visualization model, including the scaled NLDR data (`nldr_scaled_obj`), the hexagonal bin structure (`hb_obj`), the averaged \pD summaries for each bin (`model_highd`), the corresponding \twoD bin centroids (`model_2d`), and the triangulated mesh connecting neighboring bins (`trimesh_data`).
+It stores all components of the fitted visualization model, including the scaled NLDR data (`nldr_scaled_obj`), the hexagonal bin structure (`hb_obj`), the averaged $p\text{-}D$ summaries for each bin (`model_highd`), the corresponding $2\text{-}D$ bin centroids (`model_2d`), and the triangulated mesh connecting neighboring bins (`trimesh_data`).
 
@@
 Several core computations within `quollr` are optimized using compiled C++ code via the `Rcpp` and `RcppArmadillo` packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), \twoD distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (`compute_highd_dist()`), error metrics (`compute_errors()`), $2\text{-}D$ distance calculations (`calc_2d_dist_cpp()`), and generation of hexagon coordinates (`gen_hex_coord_cpp()`) are handled internally in C++.
 This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
@@
 This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`hd_thresh`).
-The function returns an object of class `highd_vis_model` containing the scaled NLDR object (`nldr_scaled_obj`) with three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both \twoD (`model_2d`), and \pD (`model_highd`), and triangular mesh (`trimesh_data`).
+The function returns an object of class `highd_vis_model` containing the scaled NLDR object (`nldr_scaled_obj`) with three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
 
@@
 
-Constructing the \twoD model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
+Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
 
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the \twoD space.
+Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2\text{-}D$ space.
 
@@
 After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The `assign_data()` function performs this assignment by calculating the \twoD Euclidean distance between each point in the \twoD embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \twoD layout. It then selects the corresponding x and y coordinates of each hexagon’s centroid.
+The `assign_data()` function performs this assignment by calculating the $2\text{-}D$ Euclidean distance between each point in the $2\text{-}D$ embedding and all hexagon centroids.
+
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2\text{-}D$ layout.
+It then selects the corresponding x and y coordinates of each hexagon’s centroid.
 
@@
 
-The result is a `tibble` of the scaled \twoD embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
+The result is a `tibble` of the scaled $2\text{-}D$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
 
@@
 
-The function first groups the input data by `h`, which represents the hexagon ID associated with each point in the \twoD layout.
+The function first groups the input data by `h`, which represents the hexagon ID associated with each point in the $2\text{-}D$ layout.
 Within each group, it collects the `ID`s into a list, resulting in a summary where each row corresponds to a single hexagon.
@@
 
-To represent the neighborhood structure of hexagonal bins in a \twoD layout, we employ Delaunay triangulation [@lee1980;@albrecht2024] on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation [@lee1980;@albrecht2024] on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
@@
 Some may have very few or no data points due to the sparsity or shape of the underlying structure.
-Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or "holes" in the \twoD structure, potentially disrupting the continuity of the representation.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or "holes" in the $2\text{-}D$ structure, potentially disrupting the continuity of the representation.
 
@@
 
-The final step involves lifting the fitted \twoD model into \pD.
-This is done by modelling a point in \pD as the  \pD mean of data points in the \twoD centroid.
-This is performed using the `avg_highd_data()` function, which takes \pD data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).
+The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$.
+This is done by modelling a point in $p\text{-}D$ as the  $p\text{-}D$ mean of data points in the $2\text{-}D$ centroid.
+This is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).
 
@@
 
-The `predict_emb()` function is used to predict a point in a \twoD embedding for a new \pD data point using the fitted model.
-This function is useful to predict \twoD embedding irrespective of the NLDR technique.
-
-In the prediction process, first, the nearest \pD model point is identified for the new \pD data point by computing \pD Euclidean distance.
-Then, the corresponding \twoD bin centroid mapping for the identified \pD model point is determined.
-Finally, the coordinates of the identified \twoD bin centroid is used as the predicted NLDR embedding for the new \pD data point.
+The `predict_emb()` function is used to predict a point in a $2\text{-}D$ embedding for a new $p\text{-}D$ data point using the fitted model.
+This function is useful to predict $2\text{-}D$ embedding irrespective of the NLDR technique.
+
+In the prediction process, first, the nearest $p\text{-}D$ model point is identified for the new $p\text{-}D$ data point by computing $p\text{-}D$ Euclidean distance.
+Then, the corresponding $2\text{-}D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
+Finally, the coordinates of the identified $2\text{-}D$ bin centroid is used as the predicted NLDR embedding for the new $p\text{-}D$ data point.
 
@@
 
-## Compute residuals and Root Within Bin Sum of Square (RWBSS)
-
-Root Within Bin Sum of Square (RWBSS) are used as goodness of fit metrics for the model. These metrics can be computed using the  `glance()` function, which provides a tidy output for evaluation.
-
-The function requires both the fitted model object returned by `fit_highd_model()` and \pD data to begin.
-The \pD model output (`model_highd`) is first renamed to avoid naming conflicts during subsequent data joins.
-It then uses the `predict_emb()` function to assign each point in the \pD dataset to a corresponding hexagon bin in the \twoD model, producing a prediction data frame that contains both the predicted bin assignment (`pred_h`) and the original observation `ID`.
-
-The function joins this prediction output with both the \pD model and the \pD data (to retrieve true coordinates).
-It then calculates squared differences between the original and predicted \pD coordinates for each dimension, storing these as `error_square_x1`, `error_square_x2`, ..., up to the dimensionality of the data.
-
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RWBSS which is the average of the total squared error per point.
-
-These metrics are returned in a tibble as `Error` (absolute error) and `RWBSS` (root mean squared error).
-The computation of total absolute error and RWBSS is performed in C++ for efficiency using the internal `compute_errors()` function.
+## Compute residuals and hexbin error (HBE)
+
+Hexbin error (HBE) are used as goodness of fit metrics for the model. These metrics can be computed using the  `glance()` function, which provides a tidy output for evaluation.
+
+The function requires both the fitted model object returned by `fit_highd_model()` and $p\text{-}D$ data to begin.
+The $p\text{-}D$ model output (`model_highd`) is first renamed to avoid naming conflicts during subsequent data joins.
+It then uses the `predict_emb()` function to assign each point in the $p\text{-}D$ dataset to a corresponding hexagon bin in the $2\text{-}D$ model, producing a prediction data frame that contains both the predicted bin assignment (`pred_h`) and the original observation `ID`.
+
+The function joins this prediction output with both the $p\text{-}D$ model and the $p\text{-}D$ data (to retrieve true coordinates).
+It then calculates squared differences between the original and predicted $p\text{-}D$ coordinates for each dimension, storing these as `error_square_x1`, `error_square_x2`, ..., up to the dimensionality of the data.
+
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the HBE which is the average of the total squared error per point.
+
+These metrics are returned in a tibble as `Error` (absolute error) and `HBE` (root mean squared error).
+The computation of total absolute error and HBE is performed in C++ for efficiency using the internal `compute_errors()` function.
 
@@
 
-Furthermore, `augment()` requires both the fitted model object returned by `fit_highd_model()` and \pD data to begin.
-It extends the fitted model by adding prediction results and error diagnostics to the original \pD data.
-
-The function starts with the same process as is used in the `glance()` function to produce a predicted point in \pD for each point in the \pD dataset.
+Furthermore, `augment()` requires both the fitted model object returned by `fit_highd_model()` and $p\text{-}D$ data to begin.
+It extends the fitted model by adding prediction results and error diagnostics to the original $p\text{-}D$ data.
+
+The function starts with the same process as is used in the `glance()` function to produce a predicted point in $p\text{-}D$ for each point in the $p\text{-}D$ dataset.
 
@@
 
-The package offers several \twoD visualizations, including:
+The package offers several $2\text{-}D$ visualizations, including:
 
@@
 
-The generated \pD model, overlaid with the data, can also be visualized using `show_langevitour`.
-Additionally, it features a function for visualizing the \twoD projection of the fitted model overlaid on the data, called `plot_proj`.
+The generated $p\text{-}D$ model, overlaid with the data, can also be visualized using `show_langevitour`.
+Additionally, it features a function for visualizing the $2\text{-}D$ projection of the fitted model overlaid on the data, called `plot_proj`.
 
@@
 
-The `geom_trimesh()` function introduces a custom `ggplot2` layer designed for visualizing \twoD wireframe on a provided set of bin centroids.
+The `geom_trimesh()` function introduces a custom `ggplot2` layer designed for visualizing $2\text{-}D$ wireframe on a provided set of bin centroids.
 
@@
 
-### \pD model visualization
-
-To visualize how well the \pD model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \pD using the `show_langevitour()` function.
+### $p\text{-}D$ model visualization
+
+To visualize how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the `show_langevitour()` function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` R package [@paul2023].
@@
 
-As an alternative to `langevitour`, users can explore the fitted \pD model using the `detourr` [@casper2025].
+As an alternative to `langevitour`, users can explore the fitted $p\text{-}D$ model using the `detourr` [@casper2025].
 The combined data object from `comb_data_model()` can be passed directly to the `detour()` function, where `tour_aes()` defines the projection variables and color mapping.
@@
 
-In the resulting interactive visualization, blue points represent the high-dimensional data, orange points represent the model centroids from each bin, and the lines between model points reflect the \twoD wireframe structure mapped to high-dimensional space.
+In the resulting interactive visualization, blue points represent the high-dimensional data, orange points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped to high-dimensional space.
 
@@
 
-To support interactive evaluation of how well the \pD model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
+To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
 This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
@@
 Before visualization, the input data must be prepared using the `comb_all_data_model()` function.
-This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the \twoD bin layout (`model_2d`):
+This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2\text{-}D$ bin layout (`model_2d`):
 
@@
 
-The function `show_link_plots()` generates two side-by-side, interactively linked plots; a \twoD NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function `show_link_plots()` generates two side-by-side, interactively linked plots; a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
 The function takes the output from `comb_all_data_model()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
@@
 Before visualization, the input data must be prepared using the `comb_all_data_model_error()` function.
-The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), \twoD model data (`model_2d`), and model error (`error_data`).
+The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2\text{-}D$ model data (`model_2d`), and model error (`error_data`).
 
@@
 
-The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a \twoD NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
 The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
@@
 In this setup, users can manually construct the linked visualization using the `crosstalk` [@joe2023] and `htmltools` [@joe2024] packages.
-The interactive layout is created by arranging the \twoD NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid.
+The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid.
 The NLDR and error plots are rendered with `ggplotly()` [@chapman2020] to enable interactive linking.
@@
 Following their pre-processing, different NLDR methods were performed using ten principal components.
-Figure \@ref(fig:limb-rwbss) (b) is the reproduction of the published plot.
+Figure \@ref(fig:limb-hbe) (b) is the reproduction of the published plot.
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better \twoD layout.
+Our method help to provide a better $2\text{-}D$ layout.
 
@@
   group_by(method, a1) |>
-  filter(RMSE == min(RMSE)) |>
+  filter(HBE == min(HBE)) |>
   ungroup()
@@
 
-error_plot_limb <- plot_rmse(error_limb) +
+error_plot_limb <- plot_hbe(error_limb) +
   scale_x_continuous(breaks = sort(unique(error_limb$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
@@
 
-```{r limb-rwbss, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RWBSS for varying binwidth ($a_1$).
+```{r limb-hbe, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.\\label{fig:limb-rwbss}", fig.pos='H', echo=TRUE}
+Layout f is the best choice.\\label{fig:limb-hbe}", fig.pos='H', echo=TRUE}
 
@@
 
-plot_rmse_layouts(plots = list(error_plot_limb, nldr1, 
+plot_hbe_layouts(plots = list(error_plot_limb, nldr1, 
                              nldr2, nldr3, nldr4, 
@@
 
-```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-rwbss}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-rwbss}f) by RWBSS plot (Figure \\ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
@@
 
-The `quollr` package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in \twoD and lifting it into high-dimensional space.
-This lifted model provides a direct way to assess how well a \twoD layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
-The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate \twoD representations.
+The `quollr` package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in $2\text{-}D$ and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a $2\text{-}D$ layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate $2\text{-}D$ representations.
 
 In contrast to the common practice of visually inspecting scatterplots for clusters or patterns, `quollr` provides a quantitative route for evaluation.
-It enables the computation of RWBSS and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
+It enables the computation of HBE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
 These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualizations using the `langevitour` package, allowing users to inspect where the model fits well and where it does not.
@@
 To support efficient computation, particularly for large-scale datasets, several core functions in `quollr` are implemented in C++ using `Rcpp` and `RcppArmadillo`.
-These include functions for computing Euclidean distances in high-dimensional and \twoD space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates RWBSS and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in \twoD.
+These include functions for computing Euclidean distances in high-dimensional and $2\text{-}D$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, `compute_highd_dist()` accelerates nearest neighbor lookup in high-dimensional space, `compute_errors()` calculates HBE and total absolute error efficiently, and `calc_2d_dist_cpp()` speeds up distance calculations in $2\text{-}D$.
 Additionally, `gen_hex_coord_cpp()` constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-11-08 paper-quollr.tex
Changed the notations to view in HTML as well

@@
 This paper presents the R package, \texttt{quollr}, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
-Starting with an NLDR layout, our approach is to create a $2\text{-}D$ wireframe model representation, that can be lifted and displayed in the high-dimensional ($p\text{-}D${}) space (Figure \ref{fig:overview}).
+Starting with an NLDR layout, our approach is to create a \(2\text{-}D\) wireframe model representation, that can be lifted and displayed in the high-dimensional (\(p\text{-}D\)) space (Figure \ref{fig:overview}).
 
@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different $2\text{-}D$projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different $2 ext{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.}\label{fig:overview}
@@
 
-The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, $2\text{-}D$ model construction, lifting into $p\text{-}D$, prediction, error computation, visualization, and interactivity.
+The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, \(2\text{-}D\) model construction, lifting into \(p\text{-}D\), prediction, error computation, visualization, and interactivity.
 Each module performs a distinct task and communicates through data objects.
@@
 \item
-  Construct $2\text{-}D$ model: A series of functions \texttt{hex\_binning()}, \texttt{merge\_hexbin\_centroids()}, \texttt{tri\_bin\_centroids()}, \texttt{gen\_edges()}, and \texttt{update\_trimesh\_index()} generate the hexagonal grid, compute bin centroids, and connect the triangular mesh that defines local neighborhoods in the $2\text{-}D$ space.
-\item
-  Lift the model into $p\text{-}D$: The function \texttt{avg\_highd\_data()} computes the average of the high-dimensional variables for each bin, linking the $2\text{-}D$ representation back to the original data space.
+  Construct \(2\text{-}D\) model: A series of functions \texttt{hex\_binning()}, \texttt{merge\_hexbin\_centroids()}, \texttt{tri\_bin\_centroids()}, \texttt{gen\_edges()}, and \texttt{update\_trimesh\_index()} generate the hexagonal grid, compute bin centroids, and connect the triangular mesh that defines local neighborhoods in the \(2\text{-}D\) space.
+\item
+  Lift the model into \(p\text{-}D\): The function \texttt{avg\_highd\_data()} computes the average of the high-dimensional variables for each bin, linking the \(2\text{-}D\) representation back to the original data space.
 \item
@@
 \item
-  Interactivity: The functions \texttt{comb\_all\_data\_model()} and \texttt{show\_link\_plots()} generate interactive linked visualizations that connect the $2\text{-}D$ NLDR layout, the corresponding tour view, and the fitted model.
-Similarly, \texttt{comb\_all\_data\_model\_error()} and \texttt{show\_error\_link\_plots()} integrate the error distribution with the $2\text{-}D$ embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
+  Interactivity: The functions \texttt{comb\_all\_data\_model()} and \texttt{show\_link\_plots()} generate interactive linked visualizations that connect the \(2\text{-}D\) NLDR layout, the corresponding tour view, and the fitted model.
+Similarly, \texttt{comb\_all\_data\_model\_error()} and \texttt{show\_error\_link\_plots()} integrate the error distribution with the \(2\text{-}D\) embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
 \end{enumerate}
@@
   \texttt{highd\_vis\_model}: the main model object returned by \texttt{fit\_highd\_model()}.
-It stores all components of the fitted visualization model, including the scaled NLDR data (\texttt{nldr\_scaled\_obj}), the hexagonal bin structure (\texttt{hb\_obj}), the averaged $p\text{-}D$ summaries for each bin (\texttt{model\_highd}), the corresponding $2\text{-}D$ bin centroids (\texttt{model\_2d}), and the triangulated mesh connecting neighboring bins (\texttt{trimesh\_data}).
+It stores all components of the fitted visualization model, including the scaled NLDR data (\texttt{nldr\_scaled\_obj}), the hexagonal bin structure (\texttt{hb\_obj}), the averaged \(p\text{-}D\) summaries for each bin (\texttt{model\_highd}), the corresponding \(2\text{-}D\) bin centroids (\texttt{model\_2d}), and the triangulated mesh connecting neighboring bins (\texttt{trimesh\_data}).
 \end{itemize}
@@
 Several core computations within \texttt{quollr} are optimized using compiled C++ code via the \texttt{Rcpp} and \texttt{RcppArmadillo} packages.
-While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), $2\text{-}D$ distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
+While the user interacts with high-level R functions, performance-critical steps such as nearest-neighbor searches (\texttt{compute\_highd\_dist()}), error metrics (\texttt{compute\_errors()}), \(2\text{-}D\) distance calculations (\texttt{calc\_2d\_dist\_cpp()}), and generation of hexagon coordinates (\texttt{gen\_hex\_coord\_cpp()}) are handled internally in C++.
 This design provides significant speedups when analyzing large datasets while maintaining a user-friendly R interface.
@@
 This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{hd\_thresh}).
-The function returns an object of class \texttt{highd\_vis\_model} containing the scaled NLDR object (\texttt{nldr\_scaled\_obj}) with three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both $2\text{-}D$ (\texttt{model\_2d}), and $p\text{-}D$ (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
+The function returns an object of class \texttt{highd\_vis\_model} containing the scaled NLDR object (\texttt{nldr\_scaled\_obj}) with three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
 
@@
 
-Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
+Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
 
@@
 
-Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the $2\text{-}D$ space.
+Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2\text{-}D\) space.
 
@@
 After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The \texttt{assign\_data()} function performs this assignment by calculating the $2\text{-}D$ Euclidean distance between each point in the $2\text{-}D$ embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2\text{-}D$ layout.
+The \texttt{assign\_data()} function performs this assignment by calculating the \(2\text{-}D\) Euclidean distance between each point in the \(2\text{-}D\) embedding and all hexagon centroids.
+
+First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \(2\text{-}D\) layout.
 It then selects the corresponding x and y coordinates of each hexagon's centroid.
@@
 
-The result is a \texttt{tibble} of the scaled $2\text{-}D$ embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
+The result is a \texttt{tibble} of the scaled \(2\text{-}D\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
 
@@
 
-The function first groups the input data by \texttt{h}, which represents the hexagon ID associated with each point in the $2\text{-}D$ layout.
+The function first groups the input data by \texttt{h}, which represents the hexagon ID associated with each point in the \(2\text{-}D\) layout.
 Within each group, it collects the \texttt{ID}s into a list, resulting in a summary where each row corresponds to a single hexagon.
@@
 
-To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation \citep{lee1980, albrecht2024} on the centroids of hexagons.
+To represent the neighborhood structure of hexagonal bins in a \(2\text{-}D\) layout, we employ Delaunay triangulation \citep{lee1980, albrecht2024} on the centroids of hexagons.
 This geometric approach is used to infer which bins are considered neighbors.
@@
 Some may have very few or no data points due to the sparsity or shape of the underlying structure.
-Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the $2\text{-}D$ structure, potentially disrupting the continuity of the representation.
+Simply removing hexagons with low counts (e.g., fewer than a fixed threshold) can lead to gaps or ``holes'' in the \(2\text{-}D\) structure, potentially disrupting the continuity of the representation.
 
@@
 
-The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$.
-This is done by modelling a point in $p\text{-}D$ as the $p\text{-}D$ mean of data points in the $2\text{-}D$ centroid.
-This is performed using the \texttt{avg\_highd\_data()} function, which takes $p\text{-}D$ data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
+The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\).
+This is done by modelling a point in \(p\text{-}D\) as the \(p\text{-}D\) mean of data points in the \(2\text{-}D\) centroid.
+This is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).
 
@@
 
-The \texttt{predict\_emb()} function is used to predict a point in a $2\text{-}D$ embedding for a new $p\text{-}D$ data point using the fitted model.
-This function is useful to predict $2\text{-}D$ embedding irrespective of the NLDR technique.
-
-In the prediction process, first, the nearest $p\text{-}D$ model point is identified for the new $p\text{-}D$ data point by computing $p\text{-}D$ Euclidean distance.
-Then, the corresponding $2\text{-}D$ bin centroid mapping for the identified $p\text{-}D$ model point is determined.
-Finally, the coordinates of the identified $2\text{-}D$ bin centroid is used as the predicted NLDR embedding for the new $p\text{-}D$ data point.
+The \texttt{predict\_emb()} function is used to predict a point in a \(2\text{-}D\) embedding for a new \(p\text{-}D\) data point using the fitted model.
+This function is useful to predict \(2\text{-}D\) embedding irrespective of the NLDR technique.
+
+In the prediction process, first, the nearest \(p\text{-}D\) model point is identified for the new \(p\text{-}D\) data point by computing \(p\text{-}D\) Euclidean distance.
+Then, the corresponding \(2\text{-}D\) bin centroid mapping for the identified \(p\text{-}D\) model point is determined.
+Finally, the coordinates of the identified \(2\text{-}D\) bin centroid is used as the predicted NLDR embedding for the new \(p\text{-}D\) data point.
 
@@
 
-\subsection{Compute residuals and Root Within Bin Sum of Square (RWBSS)}\label{compute-residuals-and-root-within-bin-sum-of-square-rwbss}
-
-Root Within Bin Sum of Square (RWBSS) are used as goodness of fit metrics for the model.
-These metrics can be computed using the \texttt{glance()} function, which provides a tidy output for evaluation.
-
-The function requires both the fitted model object returned by \texttt{fit\_highd\_model()} and $p\text{-}D$ data to begin.
-The $p\text{-}D$ model output (\texttt{model\_highd}) is first renamed to avoid naming conflicts during subsequent data joins.
-It then uses the \texttt{predict\_emb()} function to assign each point in the $p\text{-}D$ dataset to a corresponding hexagon bin in the $2\text{-}D$ model, producing a prediction data frame that contains both the predicted bin assignment (\texttt{pred\_h}) and the original observation \texttt{ID}.
-
-The function joins this prediction output with both the $p\text{-}D$ model and the $p\text{-}D$ data (to retrieve true coordinates).
-It then calculates squared differences between the original and predicted $p\text{-}D$ coordinates for each dimension, storing these as \texttt{error\_square\_x1}, \texttt{error\_square\_x2}, \ldots, up to the dimensionality of the data.
-
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RWBSS which is the average of the total squared error per point.
-
-These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{RWBSS} (root mean squared error).
-The computation of total absolute error and RWBSS is performed in C++ for efficiency using the internal \texttt{compute\_errors()} function.
+\subsection{Compute residuals and hexbin error (HBE)}\label{compute-residuals-and-hexbin-error-hbe}
+
+Hexbin error (HBE) are used as goodness of fit metrics for the model. These metrics can be computed using the \texttt{glance()} function, which provides a tidy output for evaluation.
+
+The function requires both the fitted model object returned by \texttt{fit\_highd\_model()} and \(p\text{-}D\) data to begin.
+The \(p\text{-}D\) model output (\texttt{model\_highd}) is first renamed to avoid naming conflicts during subsequent data joins.
+It then uses the \texttt{predict\_emb()} function to assign each point in the \(p\text{-}D\) dataset to a corresponding hexagon bin in the \(2\text{-}D\) model, producing a prediction data frame that contains both the predicted bin assignment (\texttt{pred\_h}) and the original observation \texttt{ID}.
+
+The function joins this prediction output with both the \(p\text{-}D\) model and the \(p\text{-}D\) data (to retrieve true coordinates).
+It then calculates squared differences between the original and predicted \(p\text{-}D\) coordinates for each dimension, storing these as \texttt{error\_square\_x1}, \texttt{error\_square\_x2}, \ldots, up to the dimensionality of the data.
+
+From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the HBE which is the average of the total squared error per point.
+
+These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{HBE} (root mean squared error).
+The computation of total absolute error and HBE is performed in C++ for efficiency using the internal \texttt{compute\_errors()} function.
 
@@
 > # A tibble: 1 x 2
->   Error  RMSE
+>   Error   HBE
 >   <dbl> <dbl>
@@
 
-Furthermore, \texttt{augment()} requires both the fitted model object returned by \texttt{fit\_highd\_model()} and $p\text{-}D$ data to begin.
-It extends the fitted model by adding prediction results and error diagnostics to the original $p\text{-}D$ data.
-
-The function starts with the same process as is used in the \texttt{glance()} function to produce a predicted point in $p\text{-}D$ for each point in the $p\text{-}D$ dataset.
+Furthermore, \texttt{augment()} requires both the fitted model object returned by \texttt{fit\_highd\_model()} and \(p\text{-}D\) data to begin.
+It extends the fitted model by adding prediction results and error diagnostics to the original \(p\text{-}D\) data.
+
+The function starts with the same process as is used in the \texttt{glance()} function to produce a predicted point in \(p\text{-}D\) for each point in the \(p\text{-}D\) dataset.
 
@@
 
-The package offers several $2\text{-}D$ visualizations, including:
+The package offers several \(2\text{-}D\) visualizations, including:
 
@@
 
-The generated $p\text{-}D$ model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
-Additionally, it features a function for visualizing the $2\text{-}D$ projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
+The generated \(p\text{-}D\) model, overlaid with the data, can also be visualized using \texttt{show\_langevitour}.
+Additionally, it features a function for visualizing the \(2\text{-}D\) projection of the fitted model overlaid on the data, called \texttt{plot\_proj}.
 
@@
 
-The \texttt{geom\_trimesh()} function introduces a custom \texttt{ggplot2} layer designed for visualizing $2\text{-}D$ wireframe on a provided set of bin centroids.
+The \texttt{geom\_trimesh()} function introduces a custom \texttt{ggplot2} layer designed for visualizing \(2\text{-}D\) wireframe on a provided set of bin centroids.
 
@@
 
-\subsubsection{\texorpdfstring{$p\text{-}D$ model visualization}{model visualization}}\label{model-visualization}
-
-To visualize how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the \texttt{show\_langevitour()} function.
+\subsubsection{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}
+
+To visualize how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \(p\text{-}D\) using the \texttt{show\_langevitour()} function.
 This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} R package \citep{paul2023}.
@@
 
-As an alternative to \texttt{langevitour}, users can explore the fitted $p\text{-}D$ model using the \texttt{detourr} \citep{casper2025}.
+As an alternative to \texttt{langevitour}, users can explore the fitted \(p\text{-}D\) model using the \texttt{detourr} \citep{casper2025}.
 The combined data object from \texttt{comb\_data\_model()} can be passed directly to the \texttt{detour()} function, where \texttt{tour\_aes()} defines the projection variables and color mapping.
@@
 
-In the resulting interactive visualization, blue points represent the high-dimensional data, orange points represent the model centroids from each bin, and the lines between model points reflect the $2\text{-}D$ wireframe structure mapped to high-dimensional space.
+In the resulting interactive visualization, blue points represent the high-dimensional data, orange points represent the model centroids from each bin, and the lines between model points reflect the \(2\text{-}D\) wireframe structure mapped to high-dimensional space.
 
@@
 
-To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
+To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
 This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
@@
 Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
-This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the $2\text{-}D$ bin layout (\texttt{model\_2d}):
+This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2\text{-}D\) bin layout (\texttt{model\_2d}):
 
@@
 
-The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
 The function takes the output from \texttt{comb\_all\_data\_model()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
@@
 Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
-The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), $2\text{-}D$ model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
+The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2\text{-}D\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
 
@@
 
-The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
 The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
@@
 In this setup, users can manually construct the linked visualization using the \texttt{crosstalk} \citep{joe2023} and \texttt{htmltools} \citep{joe2024} packages.
-The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid.
+The interactive layout is created by arranging the \(2\text{-}D\) NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid.
 The NLDR and error plots are rendered with \texttt{ggplotly()} \citep{chapman2020} to enable interactive linking.
@@
 Following their pre-processing, different NLDR methods were performed using ten principal components.
-Figure \ref{fig:limb-rwbss} (b) is the reproduction of the published plot.
+Figure \ref{fig:limb-hbe} (b) is the reproduction of the published plot.
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better $2\text{-}D$ layout.
+Our method help to provide a better \(2\text{-}D\) layout.
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rwbss-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using RWBSS for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-hbe-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.\label{fig:limb-rwbss}}\label{fig:limb-rwbss}
-\end{figure}
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-rwbss}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-rwbss}f) by RWBSS plot (Figure \ref{fig:limb-rwbss}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+Layout f is the best choice.\label{fig:limb-hbe}}\label{fig:limb-hbe}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}
@@
 
-The \texttt{quollr} package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in $2\text{-}D$ and lifting it into high-dimensional space.
-This lifted model provides a direct way to assess how well a $2\text{-}D$ layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
-The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate $2\text{-}D$ representations.
+The \texttt{quollr} package introduces a new framework for interpreting NLDR outputs by fitting a geometric wireframe model in \(2\text{-}D\) and lifting it into high-dimensional space.
+This lifted model provides a direct way to assess how well a \(2\text{-}D\) layout, produced by methods such as tSNE, UMAP, PHATE, TriMAP, or PaCMAP, preserves the structure of the original high-dimensional data.
+The approach offers both numerical and visual diagnostics to support the selection of NLDR methods and tuning hyper-parameters that produce the most accurate \(2\text{-}D\) representations.
 
 In contrast to the common practice of visually inspecting scatterplots for clusters or patterns, \texttt{quollr} provides a quantitative route for evaluation.
-It enables the computation of RWBSS and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
+It enables the computation of HBE and residuals between the original high-dimensional data and the lifted model, offering interpretable diagnostics.
 These diagnostics are complemented by interactive linked plots and high-dimensional dynamic visualizations using the \texttt{langevitour} package, allowing users to inspect where the model fits well and where it does not.
@@
 To support efficient computation, particularly for large-scale datasets, several core functions in \texttt{quollr} are implemented in C++ using \texttt{Rcpp} and \texttt{RcppArmadillo}.
-These include functions for computing Euclidean distances in high-dimensional and $2\text{-}D$ space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
-For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates RWBSS and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in $2\text{-}D$.
+These include functions for computing Euclidean distances in high-dimensional and \(2\text{-}D\) space, identifying nearest centroids, calculating residual errors, and generating polygonal coordinates of hexagons.
+For instance, \texttt{compute\_highd\_dist()} accelerates nearest neighbor lookup in high-dimensional space, \texttt{compute\_errors()} calculates HBE and total absolute error efficiently, and \texttt{calc\_2d\_dist\_cpp()} speeds up distance calculations in \(2\text{-}D\).
 Additionally, \texttt{gen\_hex\_coord\_cpp()} constructs the coordinates for hexagonal bins based on their centroids with minimal overhead.

### 2025-11-08 paper-quollr.Rmd
Citec 2D vis

@@
 
-The package offers several $2\text{-}D$ visualizations, including:
+The package offers several $2\text{-}D$ visualizations (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:geom-outputs)"} else { "Figure \\@ref(fig:geom-outputs-pdf)"})`), including:
 
@@
 Furthermore, there are two interactive plots, `show_link_plots` and `show_error_link_plots`, which are designed to help diagnose the model.
-
 Each visualization can be generated using its respective function, as described in this section.
@@
 
-To display the complete grid, users should supply all available bin centroids.
+To display the complete grid, users should supply all available bin centroids (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:geom-outputs)a"} else { "Figure \\@ref(fig:geom-outputs-pdf)a"})`).
 
@@
 
-If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:geom-outputs)b"} else { "Figure \\@ref(fig:geom-outputs-pdf)b"})`).
 
@@
 
-To display the complete wireframe, users should supply all available bin centroids.
+To display the complete wireframe, users should supply all available bin centroids (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:geom-outputs)c"} else { "Figure \\@ref(fig:geom-outputs-pdf)c"})`).
 
@@
 
-If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:geom-outputs)d"} else { "Figure \\@ref(fig:geom-outputs-pdf)d"})`).
 

### 2025-11-08 paper-quollr.tex
Citec 2D vis

@@
 
-The package offers several \(2\text{-}D\) visualizations, including:
+The package offers several \(2\text{-}D\) visualizations (Figure \ref{fig:geom-outputs-pdf}), including:
 
@@
 Furthermore, there are two interactive plots, \texttt{show\_link\_plots} and \texttt{show\_error\_link\_plots}, which are designed to help diagnose the model.
-
 Each visualization can be generated using its respective function, as described in this section.
@@
 
-To display the complete grid, users should supply all available bin centroids.
+To display the complete grid, users should supply all available bin centroids (Figure \ref{fig:geom-outputs-pdf}a).
 
@@
 
-If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed (Figure \ref{fig:geom-outputs-pdf}b).
 
@@
 
-To display the complete wireframe, users should supply all available bin centroids.
+To display the complete wireframe, users should supply all available bin centroids (Figure \ref{fig:geom-outputs-pdf}c).
 
@@
 
-If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed.
+If the goal is to plot only the subset of hexagons that correspond to bins containing data points, then only the centroids associated with those bins should be passed (Figure \ref{fig:geom-outputs-pdf}d).
 

### 2025-11-08 paper-quollr.Rmd
Added detourr figure titles

@@
 
-To visualize how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the `show_langevitour()` function.
+To visualize how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a tour of the model in $p\text{-}D$ using the `show_langevitour()` function (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-highd-model-html)"} else { "Figure \\@ref(fig:scurve-highd-model-pdf)"})`).
 This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` R package [@paul2023].
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.pos='H', fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r scurve-highd-model-html, echo=TRUE, eval=knitr::is_html_output(), fig.pos='H', fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
@@
 
-```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
@@
 
-As an alternative to `langevitour`, users can explore the fitted $p\text{-}D$ model using the `detourr` [@casper2025].
+As an alternative to `langevitour`, users can explore the fitted $p\text{-}D$ model using the `detourr` [@casper2025] (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-highd-model-detourrhtml)"} else { "Figure \\@ref(fig:scurve-highd-model-detourrpdf)"})`).
 The combined data object from `comb_data_model()` can be passed directly to the `detour()` function, where `tour_aes()` defines the projection variables and color mapping.
@@
 
-```{r, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
+```{r scurve-highd-model-detourrpdf, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r scurve-highd-model-detourrhtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 These views illustrate how the lifted wireframe model captures the structure of the Scurve data.

### 2025-11-08 paper-quollr.tex
Added detourr figure titles

@@
 
-To visualize how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \(p\text{-}D\) using the \texttt{show\_langevitour()} function.
+To visualize how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a tour of the model in \(p\text{-}D\) using the \texttt{show\_langevitour()} function (Figure \ref{fig:scurve-highd-model-pdf}).
 This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} R package \citep{paul2023}.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-29-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:unnamed-chunk-29}
-\end{figure}
-
-As an alternative to \texttt{langevitour}, users can explore the fitted \(p\text{-}D\) model using the \texttt{detourr} \citep{casper2025}.
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-pdf}
+\end{figure}
+
+As an alternative to \texttt{langevitour}, users can explore the fitted \(p\text{-}D\) model using the \texttt{detourr} \citep{casper2025} (Figure \ref{fig:scurve-highd-model-detourrpdf}).
 The combined data object from \texttt{comb\_data\_model()} can be passed directly to the \texttt{detour()} function, where \texttt{tour\_aes()} defines the projection variables and color mapping.
@@
 \includegraphics[width=0.5\linewidth]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:unnamed-chunk-30}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-detourrpdf}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-34-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
-While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-34}
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-30}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-38-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-34-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-38}
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-34}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-42}
+\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-38}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-45}
+\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-41}
 \end{figure}

### 2025-11-08 paper-quollr.Rmd
Adde link error plots vis references

@@
 To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
-This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+This visualization combines two complementary views: the NLDR layout and a dynamic tour of the model ovelaid the data in the high-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-nldrlink-html)"} else { "Figure \\@ref(fig:scurve-nldrlink-pdf)"})`).
 Both views are interactively linked, enabling users to explore.
@@
 
-```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+```{r scurve-nldrlink-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
 Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure.
@@
 
-```{r, eval=knitr::is_latex_output(), fig.pos='H', fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
+```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos='H', fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
@@
 
-The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
+The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-html)"} else { "Figure \\@ref(fig:scurve-linkerror-pdf)"})`).
 The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+```{r scurve-linkerror-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
 Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
@@
 
-```{r, eval=knitr::is_latex_output(), fig.pos="H", fig.width=15, fig.height=10, fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `Scurve` model.
+```{r scurve-linkerror-pdf, eval=knitr::is_latex_output(), fig.pos="H", fig.width=15, fig.height=10, fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\\text{-}D$ model with data (a3, b3).
@@
 In this setup, users can manually construct the linked visualization using the `crosstalk` [@joe2023] and `htmltools` [@joe2024] packages.
-The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid.
+The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-nldrlink-dthtml)"} else { "Figure \\@ref(fig:scurve-nldrlink-dtpdf)"})`).
 The NLDR and error plots are rendered with `ggplotly()` [@chapman2020] to enable interactive linking.
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
+```{r scurve-nldrlink-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
 Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure.
@@
 
-```{r, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold'}
+```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold'}
 
@@
 
-For analyses that include model error visualization, a three-panel view can be constructed by adding the error distribution plot (`error_plt`):
+For analyses that include model error visualization (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-dthtml)"} else { "Figure \\@ref(fig:scurve-linkerror-dtpdf)"})`), a three-panel view can be constructed by adding the error distribution plot (`error_plt`):
 
@@
 
-```{r, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
+```{r scurve-linkerror-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
 Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
@@
 
-```{r, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 

### 2025-11-08 paper-quollr.tex
Adde link error plots vis references

@@
 To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
-This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space.
+This visualization combines two complementary views: the NLDR layout and a dynamic tour of the model ovelaid the data in the high-dimensional space (Figure \ref{fig:scurve-nldrlink-pdf}).
 Both views are interactively linked, enabling users to explore.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-30-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
-While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:unnamed-chunk-30}
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:scurve-nldrlink-pdf}
 \end{figure}
@@
 
-The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
+The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model (Figure \ref{fig:scurve-linkerror-pdf}).
 The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-34-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:unnamed-chunk-34}
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:scurve-linkerror-pdf}
 \end{figure}
@@
 In this setup, users can manually construct the linked visualization using the \texttt{crosstalk} \citep{joe2023} and \texttt{htmltools} \citep{joe2024} packages.
-The interactive layout is created by arranging the \(2\text{-}D\) NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid.
+The interactive layout is created by arranging the \(2\text{-}D\) NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid (Figure \ref{fig:scurve-nldrlink-dtpdf}).
 The NLDR and error plots are rendered with \texttt{ggplotly()} \citep{chapman2020} to enable interactive linking.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-38}
-\end{figure}
-
-For analyses that include model error visualization, a three-panel view can be constructed by adding the error distribution plot (\texttt{error\_plt}):
+\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-nldrlink-dtpdf}
+\end{figure}
+
+For analyses that include model error visualization (Figure \ref{fig:scurve-linkerror-dtpdf}), a three-panel view can be constructed by adding the error distribution plot (\texttt{error\_plt}):
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:unnamed-chunk-41}
+\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-linkerror-dtpdf}
 \end{figure}

### 2025-11-08 paper-quollr.R

@@ -940,7 +940,7 @@ df_exe <- comb_all_data_model(
   )
 
 
-## ----linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`. Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure. Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"----
+## ----scurve-nldrlink-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`. Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure. Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"----
 # 
 # nldrdt_link <- show_link_plots(
 #   point_data = df_exe,
@@ -1185,7 +1185,7 @@ scurve_proj_umap_model1_selected2_dp <- ggplot() +
         legend.position = "none")
 
 
-## ----eval=knitr::is_latex_output(), fig.pos='H', fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$. Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2). The purple points indicate the selected subsets, which differ between rows. In (a1), the lower bridge of the `Scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure. In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection. While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP."----
[…]

### 2025-11-08 paper-quollr.Rmd
Added reference to application section
Corrected figure caption maths

@@
 This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
-Panels (a1–a4) display different $2\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
 The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', layout = "l-page"}
@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better $2\text{-}D$ layout.
+Our method help to provide a better $2\text{-}D$ layout (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:tsne-link-limb), \\@ref(fig:tsne-link-limb-best)"} else { "Figure \\@ref(fig:model-limb)"})`).
 

### 2025-11-08 paper-quollr.tex
Added reference to application section
Corrected figure caption maths

@@
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better \(2\text{-}D\) layout.
+Our method help to provide a better \(2\text{-}D\) layout (Figure \ref{fig:model-limb}).
 

### 2025-11-09 paper-quollr.Rmd
Start updating Figure 4
Updated Figure 4

@@
 
+```{r}
+
+full_hexgrid <- full_hexgrid + 
+  xlim(c(-0.2, 1.32)) +
+  ylim(c(-0.2, 1.32)) + interior_annotation("a", cex = 3)
+
+data_hexgrid <- data_hexgrid + 
+  interior_annotation("b", cex = 3)
+
+full_triangulation_grid <- full_triangulation_grid + 
+  xlim(c(-0.2, 1.32)) +
+  ylim(c(-0.2, 1.32)) + interior_annotation("c", cex = 3)
+
+data_triangulation_grid <- data_triangulation_grid + 
+  interior_annotation("d", cex = 3)
+
+```
+
 ```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_html_output()}
@@
 
-full_hexgrid + data_hexgrid + 
-  full_triangulation_grid + data_triangulation_grid +
+ full_hexgrid + 
+  data_hexgrid +  
+  full_triangulation_grid + 
+  data_triangulation_grid +
   plot_layout(ncol = 4)
@@
 The NLDR and error plots are rendered with `ggplotly()` [@chapman2020] to enable interactive linking.
-The `bscols()` function from `crosstalk` manages synchronization across these panels, allowing for linked brushing and coordinated selection between these interactive plots.
+The `bscols()` function from `crosstalk` manages synchronization across these panels, allowing for linked brushing between these interactive plots.
 

### 2025-11-09 paper-quollr.tex
Updated Figure 4

@@
 The NLDR and error plots are rendered with \texttt{ggplotly()} \citep{chapman2020} to enable interactive linking.
-The \texttt{bscols()} function from \texttt{crosstalk} manages synchronization across these panels, allowing for linked brushing and coordinated selection between these interactive plots.
+The \texttt{bscols()} function from \texttt{crosstalk} manages synchronization across these panels, allowing for linked brushing between these interactive plots.
 

### 2025-11-09 paper-quollr.Rmd
Started adding fig.alt

@@
 Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
-The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', layout = "l-page"}
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', layout = "l-page", fig.alt="The UMAP layout with a triangular mesh on the left and 2-D projections of the lifted wireframe overlaying the data on the right.
+The mesh illustrates how local neighborhoods in the 2-D embedding correspond to curved sheet structures in the original high-dimensional S-curve data."}
 
@@
 Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
-The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H'}
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', fig.alt="The UMAP layout with a triangular mesh on the left and the tour view of the lifted wireframe overlaying the data on the right.
+The mesh illustrates how local neighborhoods in the 2-D embedding correspond to curved sheet structures in the original high-dimensional S-curve data."}
 

### 2025-11-09 paper-quollr.tex
Started adding fig.alt

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/overview-1} \caption{Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space.
+\includegraphics[width=1\linewidth,alt={The UMAP layout with a triangular mesh on the left and the tour view of the lifted wireframe overlaying the data on the right.
+The mesh illustrates how local neighborhoods in the 2-D embedding correspond to curved sheet structures in the original high-dimensional S-curve data.}]{paper-quollr_files/figure-latex/overview-1} \caption{Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space.
 The left panel shows the NLDR layout with a triangular mesh overlay, forming the wireframe structure.

### 2025-11-09 paper-quollr.Rmd
Added fig.alt

@@
 Centroids are computed and triangulated to form the $2\\text{-}D$ mesh, which is then lifted into the $p\\text{-}D$ space.
-Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\\text{-}D$ and $2\\text{-}D$ representations."}
+Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\\text{-}D$ and $2\\text{-}D$ representations.", fig.alt = "Flowchart diagram illustrating the quollr workflow.
+It shows inputs of NLDR and high-dimensional data progressing through preprocessing, hexagonal binning, centroid computation, and triangulation to form a 2-D mesh.
+The mesh is lifted into the high-dimensional space for prediction and error analysis, with interactive tools linking the high- and low-dimensional representations."}
 
@@
 ```{r algo-step-html, eval=knitr::is_html_output(), fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `Scurve` data is shown.", layout = "l-page"}
+The `Scurve` data is shown.", layout = "l-page", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
 
@@
 ```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `Scurve` data is shown."}
+The `Scurve` data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
 
@@
 
-```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos='H', eval=FALSE}
+```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos='H', fig.alt="The components of the hexagon grid illustrating notation.", eval=FALSE}
 
@@
 
-```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_html_output()}
+```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_html_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
 
@@
 
-```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_latex_output()}
+```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_latex_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
 
@@
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Interactive langevitour visualization showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
+The views reveal how the wireframe follows the underlying S-curve structure and highlight regions with sparse or missing data in the original layout."}
 
@@
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
+Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data."}
 
@@
 ```{r scurve-highd-model-detourrpdf, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
+The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 
@@
 These views illustrate how the lifted wireframe model captures the structure of the Scurve data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Multiple views from the detourr output showing the lifted high-dimensional wireframe model based on the S-curve UMAP layout.
+The wireframe reflects the curved structure of the S-curve data and reveals areas with sparse or no data in the original layout."}
 
@@
 Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure.
-Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"}
+Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right, displayed using langevitour.
+Brushing highlights matching points across both views, enabling comparison of structure preservation between the 2D layout and the high-dimensional data.", layout = "l-page"}
 
@@
 In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection.
-While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP."}
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP.", fig.alt = "Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
+Panels (a1, b1) show selected regions in the UMAP layout, and panels (a2, b2) display the corresponding points in the high-dimensional tour view.
+Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
+The figure illustrates how continuous structures in the 7D S-curve appear as separate clusters in the 2D UMAP layout, revealing distortion effects."}
 
@@
 Points with higher residuals typically correspond to sparse or bridging regions in the UMAP layout and the less dense ends of the Scurve, whereas low-residual points tend to cluster within denser, well-preserved regions.
-Double-clicking clears the selection and resets all highlights.", layout = "l-page"}
+Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel link plot showing the residual distribution (left), UMAP layout (middle), and fitted model overlaid with S-curve data in seven dimensions (right) using langevitour.
+Brushing highlights corresponding points across all panels, illustrating how high-residual points align with sparse or bridging regions in the UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-page"}
 
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure."}
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.", fig.alt = "Six-panel figure showing linked views of residuals and UMAP layouts for the 7D S-curve model.
+Panels (a1–a3) highlight high-residual points in purple, corresponding to sparse bridging regions in the UMAP layout and less dense ends of the high-dimensional S-curve.
+Panels (b1–b3) highlight low-residual points, mapping to denser regions in the UMAP and thicker bands in the high-dimensional projection.
+The figure illustrates how residuals reveal distortions in UMAP embeddings."}
 
@@
 To highlight points in `detourr`, use the Select button.
-Double-clicking on either view clears the selection and dehighlights all points.", layout = "l-page"}
+Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right using detourr.
+Brushing or selecting points in either view highlights the corresponding points in the other panel, allowing assessment of how well the UMAP layout preserves the high-dimensional S-curve structure.", layout = "l-page"}
 
@@
 
-```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold'}
+```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2D layout corresponds to the 7D data structure.", fig.pos="H", fig.show='hold'}
 
@@
 To highlight points in `detourr`, use the Select button.
-Double-clicking clears the selection and resets all highlights.", layout = "l-page"}
+Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel link plot showing the residual distribution (left), UMAP layout (middle), and fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr.
+Brushing or selecting points in any panel highlights corresponding points across all views, illustrating how high-residual points correspond to sparse or bridging regions in UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-page"}
 
@@
 
-```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 
@@
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
-Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
-Layout f is the best choice.\\label{fig:limb-hbe}", fig.pos='H', echo=TRUE}
+Layout d performs well at large binwidth (where the binwidth is not enough to capture the data structure) and poorly as the bin width decreases.
+Layout f is the best choice.\\label{fig:limb-hbe}", fig.alt = "Evaluation of six NLDR layouts (a–f) on limb muscle data using HBE across different binwidths (a1).
+Line and point colors correspond to layouts.
+Layout d performs well at large binwidths but poorly as binwidth decreases, while layout f consistently provides the best representation.", fig.pos='H', echo=TRUE}
 
@@
 ```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published 2D layout (b) and the HBE-selected 2D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10 dimensions, using t-SNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", layout = "l-page"}
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the published 2D layout (b) of limb muscle data on the left and the fitted high-dimensional model in 10D on the right using langevitour.
+Colors indicate different muscle groups.
+Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the 2D embedding correspond to structures in the original 10D space.", layout = "l-page"}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", layout = "l-page"}
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the HBE-selected 2D layout (f) of limb muscle data on the left and the fitted high-dimensional model in 10D on the right using langevitour.
+Colors indicate different muscle groups.
+Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the 2D embedding correspond to structures in the original 10D space.", layout = "l-page"}
 

### 2025-11-09 paper-quollr.tex
Added fig.alt

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/quollr_workflow} \caption{Overview of the `quollr` workflow and software architecture.
+\includegraphics[width=1\linewidth,alt={Flowchart diagram illustrating the quollr workflow.
+It shows inputs of NLDR and high-dimensional data progressing through preprocessing, hexagonal binning, centroid computation, and triangulation to form a 2-D mesh.
+The mesh is lifted into the high-dimensional space for prediction and error analysis, with interactive tools linking the high- and low-dimensional representations.}]{figures/quollr_workflow} \caption{Overview of the `quollr` workflow and software architecture.
 The process begins with NLDR and $p\text{-}D$ data inputs, followed by data preprocessing and hexagonal binning.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
+\includegraphics[width=1\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `Scurve` data is shown.}\label{fig:algo-steps}
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/geom-outputs-pdf-1}
+{\centering \includegraphics[width=1\linewidth,alt={Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region.}]{paper-quollr_files/figure-latex/geom-outputs-pdf-1}
 
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+\includegraphics[width=1\linewidth,alt={Four-panel figure showing 2D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
+Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
@@
 \begin{figure}[H]
-\includegraphics[width=0.5\linewidth]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
+\includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
+The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
+The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-detourrpdf}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth,alt={Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
+Panels (a1, b1) show selected regions in the UMAP layout, and panels (a2, b2) display the corresponding points in the high-dimensional tour view.
+Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
+The figure illustrates how continuous structures in the 7D S-curve appear as separate clusters in the 2D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+\includegraphics[width=1\linewidth,alt={Six-panel figure showing linked views of residuals and UMAP layouts for the 7D S-curve model.
+Panels (a1–a3) highlight high-residual points in purple, corresponding to sparse bridging regions in the UMAP layout and less dense ends of the high-dimensional S-curve.
+Panels (b1–b3) highlight low-residual points, mapping to denser regions in the UMAP and thicker bands in the high-dimensional projection.
+The figure illustrates how residuals reveal distortions in UMAP embeddings.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-nldrlink-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2D layout corresponds to the 7D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2D layout corresponds to the 7D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-nldrlink-dtpdf}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-linkerror-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-linkerror-dtpdf}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-hbe-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth,alt={Evaluation of six NLDR layouts (a–f) on limb muscle data using HBE across different binwidths (a1).
+Line and point colors correspond to layouts.
+Layout d performs well at large binwidths but poorly as binwidth decreases, while layout f consistently provides the best representation.}]{paper-quollr_files/figure-latex/limb-hbe-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
-Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases.
+Layout d performs well at large binwidth (where the binwidth is not enough to capture the data structure) and poorly as the bin width decreases.
 Layout f is the best choice.\label{fig:limb-hbe}}\label{fig:limb-hbe}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+\includegraphics[width=1\linewidth,alt={Comparison of published 2D layout (b) and the HBE-selected 2D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10 dimensions, using t-SNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-11-09 paper-quollr.Rmd
Started updating fig.alt

@@
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
 Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data."}
@@
 Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right, displayed using langevitour.
-Brushing highlights matching points across both views, enabling comparison of structure preservation between the 2D layout and the high-dimensional data.", layout = "l-page"}
+Brushing highlights matching points across both views, enabling comparison of structure preservation between the 2-D layout and the high-dimensional data.", layout = "l-page"}
 
@@
 Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
-The figure illustrates how continuous structures in the 7D S-curve appear as separate clusters in the 2D UMAP layout, revealing distortion effects."}
+The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects."}
 
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.", fig.alt = "Six-panel figure showing linked views of residuals and UMAP layouts for the 7D S-curve model.
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.", fig.alt = "Six-panel figure showing linked views of residuals and UMAP layouts for the 7-D S-curve model.
 Panels (a1–a3) highlight high-residual points in purple, corresponding to sparse bridging regions in the UMAP layout and less dense ends of the high-dimensional S-curve.
@@
 
-```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2D layout corresponds to the 7D data structure.", fig.pos="H", fig.show='hold'}
+```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.", fig.pos="H", fig.show='hold'}
 
@@
 
-```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 
@@
 ```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published 2D layout (b) and the HBE-selected 2D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10 dimensions, using t-SNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the published 2D layout (b) of limb muscle data on the left and the fitted high-dimensional model in 10D on the right using langevitour.
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the published 2-D layout (b) of limb muscle data on the left and the fitted high-dimensional model in 10-D on the right using langevitour.
 Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the 2D embedding correspond to structures in the original 10D space.", layout = "l-page"}
+Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the 2-D embedding correspond to structures in the original 10-D space.", layout = "l-page"}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the HBE-selected 2D layout (f) of limb muscle data on the left and the fitted high-dimensional model in 10D on the right using langevitour.
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the HBE-selected 2-D layout (f) of limb muscle data on the left and the fitted high-dimensional model in 10-D on the right using langevitour.
 Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the 2D embedding correspond to structures in the original 10D space.", layout = "l-page"}
+Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the 2-D embedding correspond to structures in the original 10-D space.", layout = "l-page"}
 

### 2025-11-09 paper-quollr.tex
Started updating fig.alt

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Four-panel figure showing 2D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
+\includegraphics[width=1\linewidth,alt={Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
 Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
@@
 Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
-The figure illustrates how continuous structures in the 7D S-curve appear as separate clusters in the 2D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Six-panel figure showing linked views of residuals and UMAP layouts for the 7D S-curve model.
+\includegraphics[width=1\linewidth,alt={Six-panel figure showing linked views of residuals and UMAP layouts for the 7-D S-curve model.
 Panels (a1–a3) highlight high-residual points in purple, corresponding to sparse bridging regions in the UMAP layout and less dense ends of the high-dimensional S-curve.
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2D layout corresponds to the 7D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2D layout corresponds to the 7D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-nldrlink-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-nldrlink-dtpdf}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-linkerror-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-linkerror-dtpdf}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Comparison of published 2D layout (b) and the HBE-selected 2D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10 dimensions, using t-SNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+\includegraphics[width=1\linewidth,alt={Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-11-09 paper-quollr.R

@@ -921,7 +921,7 @@ df_exe <- comb_data_model(
 #   )
 
 
-## ----scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout. Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections. These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data. Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout. Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data."----
+## ----scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout. Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections. These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data. Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout. Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data."----
 
 scurve_proj_umap_model1 + scurve_proj_umap_model2 + 
   scurve_proj_umap_model3 + scurve_proj_umap_model4 + 
@@ -960,7 +960,7 @@ df_exe <- comb_all_data_model(
   )
 
 
[…]

### 2025-11-09 paper-quollr.Rmd
Started upadting link plot section

@@
 
-There are mainly two interactive link plots can be generated.
-
-To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`.
-This visualization combines two complementary views: the NLDR layout and a dynamic tour of the model ovelaid the data in the high-dimensional space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-nldrlink-html)"} else { "Figure \\@ref(fig:scurve-nldrlink-pdf)"})`).
-Both views are interactively linked, enabling users to explore.
-
-Before visualization, the input data must be prepared using the `comb_all_data_model()` function.
-This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2\text{-}D$ bin layout (`model_2d`):
-
+Two types of interactive link plots can be generated to assess the model fits everywhere, or better in some subspaces, or completely mismatch the data.
+The plots are linked using `crosstalk`, which allows interactive brushing: selecting or brushing points in one plot automatically highlights the corresponding points in the other linked views.
+
+First, the function `show_link_plots()` provides linking a $2\text{-}D$ NLDR layout and a tour (via `langevitour`) of the model overlaid the data (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-nldrlink-html)"} else { "Figure \\@ref(fig:scurve-nldrlink-pdf)"})`).
+
+The `point_data` for `show_link_plots()` must be prepared using the `comb_all_data_model()` function.
+This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2\text{-}D$ bin layout (`model_2d`).
 This combined dataset includes both the original observations and the bin-level model averages, labeled with a `type` variable for distinguishing between them.
+
+The `show_link_plots()` function also takes `edge_data` which defines connections between neighboring bins.
 
@@
 ```
-
-The function `show_link_plots()` generates two side-by-side, interactively linked plots; a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.
-The function takes the output from `comb_all_data_model()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
-
-These two views are linked using `crosstalk`, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the `langevitour` output.
 
@@
 
-`show_error_link_plots()` helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
+The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-html)"} else { "Figure \\@ref(fig:scurve-linkerror-pdf)"})`).
+The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
 
@@
 The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2\text{-}D$ model data (`model_2d`), and model error (`error_data`).
-
 This combined dataset includes both the original observations and the bin-level model averages, labeled with a `type` variable for distinguishing between them.
@@
 ```
-
-The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-html)"} else { "Figure \\@ref(fig:scurve-linkerror-pdf)"})`).
-The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
-
-These two views are linked using `crosstalk`, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
-This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
-
-These three views are linked using `crosstalk`, allowing interactive selection of points in error plot and the NLDR plot to highlight corresponding structures in the `langevitour` output.
 

### 2025-11-09 paper-quollr.tex
Started upadting link plot section

@@
 
-There are mainly two interactive link plots can be generated.
-
-To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}.
-This visualization combines two complementary views: the NLDR layout and a dynamic tour of the model ovelaid the data in the high-dimensional space (Figure \ref{fig:scurve-nldrlink-pdf}).
-Both views are interactively linked, enabling users to explore.
-
-Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function.
-This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2\text{-}D\) bin layout (\texttt{model\_2d}):
-
+Two types of interactive link plots can be generated to assess the model fits everywhere, or better in some subspaces, or completely mismatch the data.
+The plots are linked using \texttt{crosstalk}, which allows interactive brushing: selecting or brushing points in one plot automatically highlights the corresponding points in the other linked views.
+
+First, the function \texttt{show\_link\_plots()} provides linking a \(2\text{-}D\) NLDR layout and a tour (via \texttt{langevitour}) of the model overlaid the data (Figure \ref{fig:scurve-nldrlink-pdf}).
+
+The \texttt{point\_data} for \texttt{show\_link\_plots()} must be prepared using the \texttt{comb\_all\_data\_model()} function.
+This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2\text{-}D\) bin layout (\texttt{model\_2d}).
 This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.
+
+The \texttt{show\_link\_plots()} function also takes \texttt{edge\_data} which defines connections between neighboring bins.
 
@@
 \end{verbatim}
-
-The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.
-The function takes the output from \texttt{comb\_all\_data\_model()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
-
-These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the \texttt{langevitour} output.
 
@@
 
-\texttt{show\_error\_link\_plots()} helps to see investigate whether the model fits the points everywhere or fits better in some places, or simply mismatches the pattern.
+The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model (Figure \ref{fig:scurve-linkerror-pdf}).
+The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
 
@@
 The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2\text{-}D\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
-
 This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.
@@
 \end{verbatim}
-
-The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model (Figure \ref{fig:scurve-linkerror-pdf}).
-The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
-
-These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection.
-This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
-
-These three views are linked using \texttt{crosstalk}, allowing interactive selection of points in error plot and the NLDR plot to highlight corresponding structures in the \texttt{langevitour} output.
 

### 2025-11-09 paper-quollr.Rmd
Shorten the link plot section

@@
 
-The `point_data` for `show_link_plots()` must be prepared using the `comb_all_data_model()` function.
+The `point_data` for `show_link_plots()` can be prepared using the `comb_all_data_model()` function.
 This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2\text{-}D$ bin layout (`model_2d`).
 This combined dataset includes both the original observations and the bin-level model averages, labeled with a `type` variable for distinguishing between them.
-
-The `show_link_plots()` function also takes `edge_data` which defines connections between neighboring bins.
+Also, the `show_link_plots()` function takes `edge_data`, which defines connections between neighboring bins.
 
@@
 
-The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-html)"} else { "Figure \\@ref(fig:scurve-linkerror-pdf)"})`).
+The function `show_error_link_plots()` generates three side-by-side, interactively linked plots; a error distribution, a $2\text{-}D$ NLDR layout, and a tour (via `langevitour`) of the model overlaid the data (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-html)"} else { "Figure \\@ref(fig:scurve-linkerror-pdf)"})`).
 The function takes the output from `comb_all_data_model_error()` (`point_data`) and `edge_data` which defines connections between neighboring bins.
 
-Before visualization, the input data must be prepared using the `comb_all_data_model_error()` function.
+The `point_data` can be generated using the `comb_all_data_model_error()` function.
 The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2\text{-}D$ model data (`model_2d`), and model error (`error_data`).
@@
 
-As an alternative to using `langevitour`, link plots can also be generated with the `detourr` package.
-In this setup, users can manually construct the linked visualization using the `crosstalk` [@joe2023] and `htmltools` [@joe2024] packages.
-The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-nldrlink-dthtml)"} else { "Figure \\@ref(fig:scurve-nldrlink-dtpdf)"})`).
-The NLDR and error plots are rendered with `ggplotly()` [@chapman2020] to enable interactive linking.
+As an alternative to using `langevitour`, the tour can also be generated with the `detourr`.
+In this setup, users can manually construct the linked visualization using the `crosstalk` [@joe2023] and `htmltools` [@joe2024].
+The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid.
+The NLDR and error plots are made with `ggplotly()` [@chapman2020] to enable interactive linking.
 The `bscols()` function from `crosstalk` manages synchronization across these panels, allowing for linked brushing between these interactive plots.
 
-A two-panel linked plot combining the NLDR view and the tour from `detourr` can be created as follows:
+A two-panel linked plot (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-nldrlink-dthtml)"} else { "Figure \\@ref(fig:scurve-nldrlink-dtpdf)"})`) combining the NLDR view and the tour from `detourr` can be created as follows:
 
@@
 
-For analyses that include model error visualization (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-dthtml)"} else { "Figure \\@ref(fig:scurve-linkerror-dtpdf)"})`), a three-panel view can be constructed by adding the error distribution plot (`error_plt`):
+Also, a three-panel view (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-dthtml)"} else { "Figure \\@ref(fig:scurve-linkerror-dtpdf)"})`) can be constructed by adding the error distribution plot (`error_plt`):
 

### 2025-11-09 paper-quollr.tex
Shorten the link plot section

@@
 
-The \texttt{point\_data} for \texttt{show\_link\_plots()} must be prepared using the \texttt{comb\_all\_data\_model()} function.
+The \texttt{point\_data} for \texttt{show\_link\_plots()} can be prepared using the \texttt{comb\_all\_data\_model()} function.
 This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2\text{-}D\) bin layout (\texttt{model\_2d}).
 This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.
-
-The \texttt{show\_link\_plots()} function also takes \texttt{edge\_data} which defines connections between neighboring bins.
+Also, the \texttt{show\_link\_plots()} function takes \texttt{edge\_data}, which defines connections between neighboring bins.
 
@@
 
-The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model (Figure \ref{fig:scurve-linkerror-pdf}).
+The function \texttt{show\_error\_link\_plots()} generates three side-by-side, interactively linked plots; a error distribution, a \(2\text{-}D\) NLDR layout, and a tour (via \texttt{langevitour}) of the model overlaid the data (Figure \ref{fig:scurve-linkerror-pdf}).
 The function takes the output from \texttt{comb\_all\_data\_model\_error()} (\texttt{point\_data}) and \texttt{edge\_data} which defines connections between neighboring bins.
 
-Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function.
+The \texttt{point\_data} can be generated using the \texttt{comb\_all\_data\_model\_error()} function.
 The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2\text{-}D\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
@@
 
-As an alternative to using \texttt{langevitour}, link plots can also be generated with the \texttt{detourr} package.
-In this setup, users can manually construct the linked visualization using the \texttt{crosstalk} \citep{joe2023} and \texttt{htmltools} \citep{joe2024} packages.
-The interactive layout is created by arranging the \(2\text{-}D\) NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid (Figure \ref{fig:scurve-nldrlink-dtpdf}).
-The NLDR and error plots are rendered with \texttt{ggplotly()} \citep{chapman2020} to enable interactive linking.
+As an alternative to using \texttt{langevitour}, the tour can also be generated with the \texttt{detourr}.
+In this setup, users can manually construct the linked visualization using the \texttt{crosstalk} \citep{joe2023} and \texttt{htmltools} \citep{joe2024}.
+The interactive layout is created by arranging the \(2\text{-}D\) NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid.
+The NLDR and error plots are made with \texttt{ggplotly()} \citep{chapman2020} to enable interactive linking.
 The \texttt{bscols()} function from \texttt{crosstalk} manages synchronization across these panels, allowing for linked brushing between these interactive plots.
 
-A two-panel linked plot combining the NLDR view and the tour from \texttt{detourr} can be created as follows:
+A two-panel linked plot (Figure \ref{fig:scurve-nldrlink-dtpdf}) combining the NLDR view and the tour from \texttt{detourr} can be created as follows:
 
@@
 
-For analyses that include model error visualization (Figure \ref{fig:scurve-linkerror-dtpdf}), a three-panel view can be constructed by adding the error distribution plot (\texttt{error\_plt}):
+Also, a three-panel view (Figure \ref{fig:scurve-linkerror-dtpdf}) can be constructed by adding the error distribution plot (\texttt{error\_plt}):
 

### 2025-11-09 paper-quollr.Rmd
Fixed the figure positions

@@
 Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
-The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', layout = "l-page", fig.alt="The UMAP layout with a triangular mesh on the left and 2-D projections of the lifted wireframe overlaying the data on the right.
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos="!ht", layout = "l-page", fig.alt="The UMAP layout with a triangular mesh on the left and 2-D projections of the lifted wireframe overlaying the data on the right.
 The mesh illustrates how local neighborhoods in the 2-D embedding correspond to curved sheet structures in the original high-dimensional S-curve data."}
@@
 Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
-The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos='H', fig.alt="The UMAP layout with a triangular mesh on the left and the tour view of the lifted wireframe overlaying the data on the right.
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos="!ht", fig.alt="The UMAP layout with a triangular mesh on the left and the tour view of the lifted wireframe overlaying the data on the right.
 The mesh illustrates how local neighborhoods in the 2-D embedding correspond to curved sheet structures in the original high-dimensional S-curve data."}
@@
 
-```{r algo-step-html, eval=knitr::is_html_output(), fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
+```{r algo-step-html, eval=knitr::is_html_output(), fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `Scurve` data is shown.", layout = "l-page", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
@@
 
-```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos='H', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
+```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `Scurve` data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
@@
 
-## Constructing the $2\text{-}D$ Model
-
-Constructing the $2\text{-}D$ model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
-
-### Scaling the data
+The model generation involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, (v) removing low-density hexagons, and (vi) lifting the model to high-dimensions.
+
+## Scaling the data
 
@@
 
-### Computing hexagon grid configuration
+## Computing hexagon grid configuration
 
@@
 
-### Binning the data
+## Binning the data
 
@@
 
-```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos='H', fig.alt="The components of the hexagon grid illustrating notation.", eval=FALSE}
+```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos="!ht", fig.alt="The components of the hexagon grid illustrating notation.", eval=FALSE}
 
@@
 
-#### Generating all possible centroids in a hexagonal grid
+### Generating all possible centroids in a hexagonal grid
 
@@
 
-#### Creating the coordinates of the hexagons
+### Creating the coordinates of the hexagons
 
@@
 
-#### Assigning data points to their respective hexagons
+### Assigning data points to their respective hexagons
 
@@
 
-#### Computing the standardized number of points within each hexagon
+### Computing the standardized number of points within each hexagon
 
@@
 
-#### Mapping the points to their corresponding hexagonal bins
+### Mapping the points to their corresponding hexagonal bins
 
@@
 
-### Obtaining bin centroids
+## Obtaining bin centroids
 
@@
 
-### Indicating neighbors by line segments connecting centroids
+## Indicating neighbors by line segments connecting centroids
 
@@
 
-### Identifying and removing low-density hexagons
+## Identifying and removing low-density hexagons
 
@@
 
-```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_html_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
+```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_html_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
 
@@
 
-```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos='H', fig.height=5, fig.width=20, eval=knitr::is_latex_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
+```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_latex_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
 
@@
 
-```{r scurve-highd-model-html, echo=TRUE, eval=knitr::is_html_output(), fig.pos='H', fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r scurve-highd-model-html, echo=TRUE, eval=knitr::is_html_output(), fig.pos="!ht", fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
@@
 
-```{r scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos='H', fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
@@
 
-```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos='H', fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
+```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
@@
 Line and point colors correspond to layouts.
-Layout d performs well at large binwidths but poorly as binwidth decreases, while layout f consistently provides the best representation.", fig.pos='H', echo=TRUE}
+Layout d performs well at large binwidths but poorly as binwidth decreases, while layout f consistently provides the best representation.", fig.pos="!ht", echo=TRUE}
 
@@
 ```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos='H', fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 

### 2025-11-09 paper-quollr.tex
Fixed the figure positions

@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={The UMAP layout with a triangular mesh on the left and the tour view of the lifted wireframe overlaying the data on the right.
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
@@
 
-\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}
-
-Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) removing low-density hexagons.
-
-\subsubsection{Scaling the data}\label{scaling-the-data}
+The model generation involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, (v) removing low-density hexagons, and (vi) lifting the model to high-dimensions.
+
+\subsection{Scaling the data}\label{scaling-the-data}
 
@@
 
-\subsubsection{Computing hexagon grid configuration}\label{computing-hexagon-grid-configuration}
+\subsection{Computing hexagon grid configuration}\label{computing-hexagon-grid-configuration}
 
@@
 
-\subsubsection{Binning the data}\label{binning-the-data}
+\subsection{Binning the data}\label{binning-the-data}
 
@@
 
-\paragraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
+\subsubsection{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
 
@@
 
-\paragraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
+\subsubsection{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
 
@@
 
-\paragraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
+\subsubsection{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
 
@@
 
-\paragraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
+\subsubsection{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
 
@@
 
-\paragraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
+\subsubsection{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
 
@@
 
-\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
+\subsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
 
@@
 
-\subsubsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
+\subsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
 
@@
 
-\subsubsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
+\subsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Evaluation of six NLDR layouts (a–f) on limb muscle data using HBE across different binwidths (a1).
@@
 
-\begin{figure}[H]
+\begin{figure}[!ht]
 \includegraphics[width=1\linewidth,alt={Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.

### 2025-11-09 paper-quollr.Rmd
Changed the scurve

@@
 ```{r algo-step-html, eval=knitr::is_html_output(), fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `Scurve` data is shown.", layout = "l-page", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
+The `scurve` data is shown.", layout = "l-page", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
 
@@
 ```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `Scurve` data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
+The `scurve` data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
 
@@
 
-```{r scurve-highd-model-html, echo=TRUE, eval=knitr::is_html_output(), fig.pos="!ht", fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
-These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
+```{r scurve-highd-model-html, echo=TRUE, eval=knitr::is_html_output(), fig.pos="!ht", fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
+These views illustrate how the lifted wireframe model captures the structure of the `scurve` data.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Interactive langevitour visualization showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
@@
 
-```{r scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
-Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
-These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
+```{r scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
+Each panel (a1–a4) shows the model (black) overlaid on `scurve` data (blue) in different projections.
+These views illustrate how the lifted wireframe model captures the structure of the `scurve` data.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
@@
 
-```{r scurve-highd-model-detourrpdf, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
+```{r scurve-highd-model-detourrpdf, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `scurve` UMAP layout using `detourr`.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
@@
 
-```{r scurve-highd-model-detourrhtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
+```{r scurve-highd-model-detourrhtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
 These views illustrate how the lifted wireframe model captures the structure of the Scurve data.
@@
 ```{r scurve-nldrlink-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
-Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure.
+Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `scurve` structure.
 Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right, displayed using langevitour.
@@
 
-```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `Scurve` structure in $7\\text{-}D$.
+```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `scurve` structure in $7\\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
 The purple points indicate the selected subsets, which differ between rows.
-In (a1), the lower bridge of the `Scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
-In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection.
+In (a1), the lower bridge of the `scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
+In (b1), a different region near the upper arm of the `scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection.
 While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP.", fig.alt = "Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
@@
 
-```{r scurve-linkerror-pdf, eval=knitr::is_latex_output(), fig.pos="H", fig.width=15, fig.height=10, fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `Scurve` model.
+```{r scurve-linkerror-pdf, eval=knitr::is_latex_output(), fig.pos="H", fig.width=15, fig.height=10, fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\\text{-}D$ model with data (a3, b3).
 The purple points highlight selected subsets of the data, which differ across rows.
-In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `Scurve` in the high-dimensional projection (a3).
-In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
+In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `scurve` in the high-dimensional projection (a3).
+In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `scurve` in the projection (b3).
 This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.", fig.alt = "Six-panel figure showing linked views of residuals and UMAP layouts for the 7-D S-curve model.
@@
 ```{r scurve-nldrlink-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
-Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `Scurve` structure.
+Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `scurve` structure.
 To highlight points in `detourr`, use the Select button.

### 2025-11-09 paper-quollr.tex
Changed the scurve

@@
 \includegraphics[width=1\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `Scurve` data is shown.}\label{fig:algo-steps}
+The `scurve` data is shown.}\label{fig:algo-steps}
 \end{figure}
@@
 \includegraphics[width=1\linewidth,alt={Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
-Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout.
-Each panel (a1–a4) shows the model (black) overlaid on `Scurve` data (blue) in different projections.
-These views illustrate how the lifted wireframe model captures the structure of the `Scurve` data.
+Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
+Each panel (a1–a4) shows the model (black) overlaid on `scurve` data (blue) in different projections.
+These views illustrate how the lifted wireframe model captures the structure of the `scurve` data.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-pdf}
@@
 The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `Scurve` UMAP layout using `detourr`.
+The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `scurve` UMAP layout using `detourr`.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-detourrpdf}
@@
 Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
-The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `Scurve` structure in $7\text{-}D$.
+The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `scurve` structure in $7\text{-}D$.
 Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
 The purple points indicate the selected subsets, which differ between rows.
-In (a1), the lower bridge of the `Scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
-In (b1), a different region near the upper arm of the `Scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
+In (a1), the lower bridge of the `scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
+In (b1), a different region near the upper arm of the `scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
 While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:scurve-nldrlink-pdf}
@@
 Panels (b1–b3) highlight low-residual points, mapping to denser regions in the UMAP and thicker bands in the high-dimensional projection.
-The figure illustrates how residuals reveal distortions in UMAP embeddings.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `Scurve` model.
+The figure illustrates how residuals reveal distortions in UMAP embeddings.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `scurve` model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
 The purple points highlight selected subsets of the data, which differ across rows.
-In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `Scurve` in the high-dimensional projection (a3).
-In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `Scurve` in the projection (b3).
+In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `scurve` in the high-dimensional projection (a3).
+In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `scurve` in the projection (b3).
 This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:scurve-linkerror-pdf}

### 2025-11-10 paper-quollr.Rmd
did some rewording in the implementation section

@@
 
-The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, $2\text{-}D$ model construction, lifting into $p\text{-}D$, prediction, error computation, visualization, and interactivity.
+The package is organized into seven core modules corresponding to stages of the workflow: preprocessing, $2\text{-}D$ model construction, lifting the model into $p\text{-}D$, prediction, error computation, visualizations, and interactivity.
 Each module performs a distinct task and communicates through data objects.
@@
 
-6. Visualization: Functions such as `geom_hexgrid()`, `geom_trimesh()`, and `show_langevitour()` provide tools for exploring the fitted models through static and dynamic visualizations.
+6. Visualizations: Functions such as `geom_hexgrid()`, `geom_trimesh()`, and `show_langevitour()` provide tools for exploring the fitted models through static and dynamic visualizations.
 
@@
 Interactivity: The functions `comb_all_data_model()` and `show_link_plots()` generate interactive linked visualizations that connect the $2\text{-}D$ NLDR layout, the corresponding tour view, and the fitted model.
-Similarly, `comb_all_data_model_error()` and `show_error_link_plots()` integrate the error distribution with the $2\text{-}D$ embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
+Similarly, `comb_all_data_model_error()` and `show_error_link_plots()` integrate the error distribution with the $2\text{-}D$ embedding and tour view, enabling interactivity across multiple plots.
 
@@
 The internal data objects follow the tidy data principle: each variable is stored in a column, each observation in a row, and each type of information in its own table.
-This structure makes the package easy to use with the `tidyverse` and other R visualization tools.
+This structure makes the package easy to use with the `tidyverse` and other visualization tools.
 
@@
 
-- `hex_bin_obj`: the object created by `hex_binning()`, which defines the structure of the two-dimensional hexagonal grid used in modeling.
-It includes the grid spacing (`a1`, `a2`), the number of bins along each axis, the centroids of all hexagons, polygon coordinates for plotting, and the mapping of each data point to its assigned hexagon.
+- `hex_bin_obj`: the object created by `hex_binning()`, which inlcudes hex binning configurations.
+It includes the binwidth (`a1`), binheight (`a2`), the number of bins along each axis (`b1`, `b2`), the centroids of all hexagons, polygon coordinates, and the assignment of each data point to the hexagon.
 
 - `highd_vis_model`: the main model object returned by `fit_highd_model()`.
-It stores all components of the fitted visualization model, including the scaled NLDR data (`nldr_scaled_obj`), the hexagonal bin structure (`hb_obj`), the averaged $p\text{-}D$ summaries for each bin (`model_highd`), the corresponding $2\text{-}D$ bin centroids (`model_2d`), and the triangulated mesh connecting neighboring bins (`trimesh_data`).
+It stores all components of the fitted model, including the scaled NLDR data (`nldr_scaled_obj`), the hexagonal bin configurations (`hb_obj`), the averaged $p\text{-}D$ summaries for each bin (`model_highd`), the corresponding $2\text{-}D$ bin centroids (`model_2d`), and the triangulated mesh connecting neighboring bins (`trimesh_data`).
 

### 2025-11-10 paper-quollr.tex
did some rewording in the implementation section

@@
 
-The package is organized into seven core modules corresponding to stages of the analysis workflow: preprocessing, \(2\text{-}D\) model construction, lifting into \(p\text{-}D\), prediction, error computation, visualization, and interactivity.
+The package is organized into seven core modules corresponding to stages of the workflow: preprocessing, \(2\text{-}D\) model construction, lifting the model into \(p\text{-}D\), prediction, error computation, visualizations, and interactivity.
 Each module performs a distinct task and communicates through data objects.
@@
 \item
-  Visualization: Functions such as \texttt{geom\_hexgrid()}, \texttt{geom\_trimesh()}, and \texttt{show\_langevitour()} provide tools for exploring the fitted models through static and dynamic visualizations.
+  Visualizations: Functions such as \texttt{geom\_hexgrid()}, \texttt{geom\_trimesh()}, and \texttt{show\_langevitour()} provide tools for exploring the fitted models through static and dynamic visualizations.
 \item
   Interactivity: The functions \texttt{comb\_all\_data\_model()} and \texttt{show\_link\_plots()} generate interactive linked visualizations that connect the \(2\text{-}D\) NLDR layout, the corresponding tour view, and the fitted model.
-Similarly, \texttt{comb\_all\_data\_model\_error()} and \texttt{show\_error\_link\_plots()} integrate the error distribution with the \(2\text{-}D\) embedding and dynamic high-dimensional view, enabling interactivity across multiple visual components.
+Similarly, \texttt{comb\_all\_data\_model\_error()} and \texttt{show\_error\_link\_plots()} integrate the error distribution with the \(2\text{-}D\) embedding and tour view, enabling interactivity across multiple plots.
 \end{enumerate}
@@
 The internal data objects follow the tidy data principle: each variable is stored in a column, each observation in a row, and each type of information in its own table.
-This structure makes the package easy to use with the \texttt{tidyverse} and other R visualization tools.
+This structure makes the package easy to use with the \texttt{tidyverse} and other visualization tools.
 
@@
 \item
-  \texttt{hex\_bin\_obj}: the object created by \texttt{hex\_binning()}, which defines the structure of the two-dimensional hexagonal grid used in modeling.
-It includes the grid spacing (\texttt{a1}, \texttt{a2}), the number of bins along each axis, the centroids of all hexagons, polygon coordinates for plotting, and the mapping of each data point to its assigned hexagon.
+  \texttt{hex\_bin\_obj}: the object created by \texttt{hex\_binning()}, which inlcudes hex binning configurations.
+It includes the binwidth (\texttt{a1}), binheight (\texttt{a2}), the number of bins along each axis (\texttt{b1}, \texttt{b2}), the centroids of all hexagons, polygon coordinates, and the assignment of each data point to the hexagon.
 \item
   \texttt{highd\_vis\_model}: the main model object returned by \texttt{fit\_highd\_model()}.
-It stores all components of the fitted visualization model, including the scaled NLDR data (\texttt{nldr\_scaled\_obj}), the hexagonal bin structure (\texttt{hb\_obj}), the averaged \(p\text{-}D\) summaries for each bin (\texttt{model\_highd}), the corresponding \(2\text{-}D\) bin centroids (\texttt{model\_2d}), and the triangulated mesh connecting neighboring bins (\texttt{trimesh\_data}).
+It stores all components of the fitted model, including the scaled NLDR data (\texttt{nldr\_scaled\_obj}), the hexagonal bin configurations (\texttt{hb\_obj}), the averaged \(p\text{-}D\) summaries for each bin (\texttt{model\_highd}), the corresponding \(2\text{-}D\) bin centroids (\texttt{model\_2d}), and the triangulated mesh connecting neighboring bins (\texttt{trimesh\_data}).
 \end{itemize}

### 2025-11-10 paper-quollr.Rmd
Checking wordings

@@
 
-- `hex_bin_obj`: the object created by `hex_binning()`, which inlcudes hex binning configurations.
+- `hex_bin_obj`: the object created by `hex_binning()`, which includes hexagon grid configurations.
 It includes the binwidth (`a1`), binheight (`a2`), the number of bins along each axis (`b1`, `b2`), the centroids of all hexagons, polygon coordinates, and the assignment of each data point to the hexagon.
@@
 The following demonstration of the package's functionality assumes `quollr` has been loaded.
-To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+To begin the workflow, user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
 The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
@@
 
-For this example, we use the UMAP layout, which is produced using `n_neighbors =` $46$ and `min_dist =` $0.9$.
+For this example, we use the UMAP layout, which is used `n_neighbors =` $46$ and `min_dist =` $0.9$ as hyper-parameters.
 
@@
 
-The model generation involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, (v) removing low-density hexagons, and (vi) lifting the model to high-dimensions.
-
 ## Scaling the data
@@
 
-The function `calc_bins_y()` determines the configuration of the hexagonal grid by computing the number of bins along the y-axis (`b2`), the hexagon width (`a1`), and height (`a2`).
-This function accepts (1) an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); (2) the number of bins along the x-axis (`b1`), and (3) the buffer amount as a proportion (`q`).
+The function `calc_bins_y()` determines the configuration of the hexagon grid by computing the number of bins along the y-axis (`b2`), the hexagon binwidth (`a1`), and binheight (`a2`).
+This function accepts (i) an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); (ii) the number of bins along the x-axis (`b1`), and (iii) the buffer amount as a proportion (`q`).
 The buffer ensures that the grid fully covers the data space by extending one hexagon width ($a_1$) and height ($a_2$) beyond the observed data in all directions.

### 2025-11-10 paper-quollr.tex
Checking wordings

@@
 \item
-  \texttt{hex\_bin\_obj}: the object created by \texttt{hex\_binning()}, which inlcudes hex binning configurations.
+  \texttt{hex\_bin\_obj}: the object created by \texttt{hex\_binning()}, which includes hexagon grid configurations.
 It includes the binwidth (\texttt{a1}), binheight (\texttt{a2}), the number of bins along each axis (\texttt{b1}, \texttt{b2}), the centroids of all hexagons, polygon coordinates, and the assignment of each data point to the hexagon.
@@
 The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded.
-To begin the workflow, users need two inputs: the high-dimensional dataset and the corresponding nonlinear dimensionality reduction (NLDR) layout.
+To begin the workflow, user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
 The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
@@
 
-For this example, we use the UMAP layout, which is produced using \texttt{n\_neighbors\ =} \(46\) and \texttt{min\_dist\ =} \(0.9\).
+For this example, we use the UMAP layout, which is used \texttt{n\_neighbors\ =} \(46\) and \texttt{min\_dist\ =} \(0.9\) as hyper-parameters.
 
@@
 
-The model generation involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, (v) removing low-density hexagons, and (vi) lifting the model to high-dimensions.
-
 \subsection{Scaling the data}\label{scaling-the-data}
@@
 
-The function \texttt{calc\_bins\_y()} determines the configuration of the hexagonal grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon width (\texttt{a1}), and height (\texttt{a2}).
-This function accepts (1) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (2) the number of bins along the x-axis (\texttt{b1}), and (3) the buffer amount as a proportion (\texttt{q}).
+The function \texttt{calc\_bins\_y()} determines the configuration of the hexagon grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon binwidth (\texttt{a1}), and binheight (\texttt{a2}).
+This function accepts (i) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (ii) the number of bins along the x-axis (\texttt{b1}), and (iii) the buffer amount as a proportion (\texttt{q}).
 The buffer ensures that the grid fully covers the data space by extending one hexagon width (\(a_1\)) and height (\(a_2\)) beyond the observed data in all directions.

### 2025-11-11 paper-quollr.Rmd
Will change after comments later

@@
 
-The `gen_centroids()` function calculates the centroids of a hexagonal grid.
-
-The coordinate limits of the embedding (`lim1` and `lim2`) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
-The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagons along y-axis (`b2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`b1`) and buffer (`q`).
-
-Then, the centroids are computed iteratively.
-The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width.
-Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
-Vertical spacing (`vs`) is given by $\sqrt{3}/2 \times a_1$.
-
-The y-coordinates for each row are similarly calculated, and paired with the x-coordinates based on whether the total number of rows is even or odd.
-In the case of an odd number of rows, the final row uses only the odd-row x-coordinates to maintain the alternating pattern.
-
-Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2\text{-}D$ space.
+The `gen_centroids()` function computes the centroids of a hexagonal grid.
+The function first validates the input parameters, ensuring that the number of bins along the x-axis (`b1`) is at least two and that the buffer width (`q`) lies within the recommended range of 0.05–0.2.
+
+The coordinate limits of the embedding (`lim1` and `lim2`) are used to calculate the aspect ratio between the two axes, which determines the vertical scaling of the grid.
+The helper function `calc_bins_y()` is then called to compute the number of bins along the y-axis (`b2`) and the hexagon binwidth (`a1`) given `b1` and `q`.
+The hexagon grid starting points are defined by `s1` and `s2`.
+These are set to $-q$ and $-q \times r2$, where `r2` is the is the aspect ratio between the two axes.
+
+Centroids are generated by alternating the x-coordinates of adjacent rows to create the hexagonal tiling.
+Odd-numbered rows use x-coordinates spaced by the hexagon width (`a1`), while even-numbered rows are staggered by half this distance (`a1/2`).
+The vertical spacing between rows is $v_s = (\sqrt{3}/2) a_1$.
+
+The y-coordinates are constructed using this spacing, and paired with x-coordinates depending on whether the total number of rows (`b2`) is even or odd.
+For an odd number of rows, the final row includes only the odd-row x-coordinates to preserve the alternating pattern.
+
+The output is a tibble containing a unique hexagon ID (`h`) and centroid coordinates (`c_x`, `c_y`), which together define the layout of the complete hexagonal grid over the $2\text{-}D$ space.
 
@@
 
-Each hexagon is defined relative to its centroid $(C_x, C_y)$, with six vertices positioned equidistantly around the center.
-The function first verifies the presence of the required hexagon width parameter `a1`.
-This width determines the horizontal spacing (`hs`).
-
-Two derived constants are calculated to define the relative distances to the vertices.
-The horizontal and vertical offset is defined as $dx = a_1/2$, and $dy = a_1/\sqrt{3}$ repectively.
-A vertical spacing factor $vf = a_1/2\sqrt{3}$ refines vertical placement in staggered rows.
-
-With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid.
-These offsets form two vectors corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
-
-For each centroid, six vertices are computed and assigned a polygon ID as the centroid.
-These vertices are then combined into a tibble that records the polygon ID (`h`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.
-
-To reduce computational overhead, the geometry calculations are implemented in C++ using `gen_hex_coord_cpp()`, which returns a `tibble` of vertex coordinates.
+The function first verifies that the hexagon width (`a1`) is provided, as it determines the overall scale of the grid.
+Each hexagon is defined relative to its centroid $(C_x, C_y)$, with six vertices positioned symmetrically around the center.
+
+Three constants are derived to define the relative distances to the vertices:
+$dx = a_1/2, \quad dy = a_1/\sqrt{3}, \quad vf = a_1/(2\sqrt{3})$.
+
+Here, `dx` and `dy` define the horizontal and vertical offsets, while `vf` refines vertical positioning to maintain geometric alignment across rows.
+
+Using these constants, two offset vectors are defined for the six vertices relative to the centroid: $x_{\text{add}} = (0, -dx, -dx, 0, dx, dx), \quad y_{\text{add}} = (dy, vf, -vf, -dy, -vf, vf)$.
+
+These offsets correspond to the six directions around each centroid: top, top-left, bottom-left, bottom, bottom-right, and top-right.
+
+For each centroid, the coordinates of its six vertices are computed as $x_i = C_x + x_{\text{add},i}, \quad y_i = C_y + y_{\text{add},i}$, and assigned a polygon ID equal to the centroid identifier (`h`).
+The resulting coordinates are combined into a tibble containing the polygon ID (`h`) and vertex coordinates (`x`, `y`).
+
+To improve computational efficiency, the vertex generation is implemented in C++ via the helper function `gen_hex_coord_cpp()`.
+This routine iterates over all centroids, applies the vertex offsets, and returns a tibble that defines the complete set of hexagon geometries in $2\text{-}D$ space.
 

### 2025-11-11 paper-quollr.tex
Will change after comments later

@@
 
-The \texttt{gen\_centroids()} function calculates the centroids of a hexagonal grid.
-
-The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing.
-The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagons along y-axis (\texttt{b2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{b1}) and buffer (\texttt{q}).
-
-Then, the centroids are computed iteratively.
-The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width.
-Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect.
-Vertical spacing (\texttt{vs}) is given by \(\sqrt{3}/2 \times a_1\).
-
-The y-coordinates for each row are similarly calculated, and paired with the x-coordinates based on whether the total number of rows is even or odd.
-In the case of an odd number of rows, the final row uses only the odd-row x-coordinates to maintain the alternating pattern.
-
-Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2\text{-}D\) space.
+The \texttt{gen\_centroids()} function computes the centroids of a hexagonal grid.
+The function first validates the input parameters, ensuring that the number of bins along the x-axis (\texttt{b1}) is at least two and that the buffer width (\texttt{q}) lies within the recommended range of 0.05--0.2.
+
+The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to calculate the aspect ratio between the two axes, which determines the vertical scaling of the grid.
+The helper function \texttt{calc\_bins\_y()} is then called to compute the number of bins along the y-axis (\texttt{b2}) and the hexagon binwidth (\texttt{a1}) given \texttt{b1} and \texttt{q}.
+The hexagon grid starting points are defined by \texttt{s1} and \texttt{s2}.
+These are set to \(-q\) and \(-q \times r2\), where \texttt{r2} is the is the aspect ratio between the two axes.
+
+Centroids are generated by alternating the x-coordinates of adjacent rows to create the hexagonal tiling.
+Odd-numbered rows use x-coordinates spaced by the hexagon width (\texttt{a1}), while even-numbered rows are staggered by half this distance (\texttt{a1/2}).
+The vertical spacing between rows is \(v_s = (\sqrt{3}/2) a_1\).
+
+The y-coordinates are constructed using this spacing, and paired with x-coordinates depending on whether the total number of rows (\texttt{b2}) is even or odd.
+For an odd number of rows, the final row includes only the odd-row x-coordinates to preserve the alternating pattern.
+
+The output is a tibble containing a unique hexagon ID (\texttt{h}) and centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which together define the layout of the complete hexagonal grid over the \(2\text{-}D\) space.
 
@@
 
-Each hexagon is defined relative to its centroid \((C_x, C_y)\), with six vertices positioned equidistantly around the center.
-The function first verifies the presence of the required hexagon width parameter \texttt{a1}.
-This width determines the horizontal spacing (\texttt{hs}).
-
-Two derived constants are calculated to define the relative distances to the vertices.
-The horizontal and vertical offset is defined as \(dx = a_1/2\), and \(dy = a_1/\sqrt{3}\) repectively.
-A vertical spacing factor \(vf = a_1/2\sqrt{3}\) refines vertical placement in staggered rows.
-
-With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid.
-These offsets form two vectors corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.
-
-For each centroid, six vertices are computed and assigned a polygon ID as the centroid.
-These vertices are then combined into a tibble that records the polygon ID (\texttt{h}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.
-
-To reduce computational overhead, the geometry calculations are implemented in C++ using \texttt{gen\_hex\_coord\_cpp()}, which returns a \texttt{tibble} of vertex coordinates.
+The function first verifies that the hexagon width (\texttt{a1}) is provided, as it determines the overall scale of the grid.
+Each hexagon is defined relative to its centroid \((C_x, C_y)\), with six vertices positioned symmetrically around the center.
+
+Three constants are derived to define the relative distances to the vertices:
+\(dx = a_1/2, \quad dy = a_1/\sqrt{3}, \quad vf = a_1/(2\sqrt{3})\).
+
+Here, \texttt{dx} and \texttt{dy} define the horizontal and vertical offsets, while \texttt{vf} refines vertical positioning to maintain geometric alignment across rows.
+
+Using these constants, two offset vectors are defined for the six vertices relative to the centroid: \(x_{\text{add}} = (0, -dx, -dx, 0, dx, dx), \quad y_{\text{add}} = (dy, vf, -vf, -dy, -vf, vf)\).
+
+These offsets correspond to the six directions around each centroid: top, top-left, bottom-left, bottom, bottom-right, and top-right.
+
+For each centroid, the coordinates of its six vertices are computed as \(x_i = C_x + x_{\text{add},i}, \quad y_i = C_y + y_{\text{add},i}\), and assigned a polygon ID equal to the centroid identifier (\texttt{h}).
+The resulting coordinates are combined into a tibble containing the polygon ID (\texttt{h}) and vertex coordinates (\texttt{x}, \texttt{y}).
+
+To improve computational efficiency, the vertex generation is implemented in C++ via the helper function \texttt{gen\_hex\_coord\_cpp()}.
+This routine iterates over all centroids, applies the vertex offsets, and returns a tibble that defines the complete set of hexagon geometries in \(2\text{-}D\) space.
 

### 2025-11-30 paper-quollr.Rmd
Move usage to before implementation and re-order many things in implementation

@@
 
+# Usage
+
+The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.
+
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`).
+A detailed description of the algorithm can be found in @gamage2025c.
+
+The user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
+The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
+The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
+The built-in example datasets, `scurve` and `scurve_umap` demonstrates these structures.
+
+To run the entire model pipeline, we can use the `fit_high_model()` function.
+This function requires: the high-dimensional data (`highd_data`), the embedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of the data (`q`), and a benchmark value to identify high-density hexagons (`hd_thresh`).
+
+The function returns an object of class `highd_vis_model` containing the scaled NLDR object (`nldr_scaled_obj`) with three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
+
+```{r, echo=TRUE, eval=TRUE}
+model_highd <- fit_highd_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  b1 = 21, 
+  q = 0.1, 
+  hd_thresh = 0)
+```
+
+The resulting model can then be shown in a tour using a two-step process:
+
+```{r, echo=TRUE, eval=TRUE}
+combined_data <- comb_data_model(
+  highd_data = scurve,
+  model_highd = model_highd$model_highd,
+  model_2d = model_highd$model_2d
+)
+
+tour_view <- show_langevitour(
+  point_data = combined_data,
+  edge_data = model_highd$trimesh_data
+)
+```
+
 # Implementation
@@
 
-# Usage
-
-The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.
-
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`).
-A detailed description of the algorithm can be found in @gamage2025c.
+## Pipeline implementation
 
@@
 
-The following demonstration of the package's functionality assumes `quollr` has been loaded.
-To begin the workflow, user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
-The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
-The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
-
-To illustrate the workflow, we use the built-in example dataset `scurve`, a $7\text{-}D$ simulated dataset consisting of $1000$ observations.
-The first three variables define a $3\text{-}D$ S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.
-
-```{r, echo=TRUE}
-data("scurve")
-```
-
-For this example, we use the UMAP layout, which is used `n_neighbors =` $46$ and `min_dist =` $0.9$ as hyper-parameters.
-
-```{r, echo=TRUE, eval=FALSE}
-library(umap)
-library(dplyr)
-
-scurve_umap <- umap(
-  scurve |> select(-ID),
-  config = modifyList(umap.defaults, list(
-    n_neighbors = 46,
-    n_components = 2,
-    min_dist = 0.9
-  ))
-)$layout |>
-  as_tibble(.name_repair = ~ paste0("emb", 1:2)) |>
-  mutate(ID = scurve$ID)
-```
-
-## Main function
-
-The mains steps for the algorithm can be executed by the main function `fit_highd_model()`, or can be run separately for more flexibility.
-
-This function requires several parameters: the high-dimensional data (`highd_data`), the emdedding data (`nldr_data`), the number of bins along the x-axis (`b1`), the buffer amount as a proportion of data (`q`), and benchmark value to extract high density hexagons (`hd_thresh`).
-The function returns an object of class `highd_vis_model` containing the scaled NLDR object (`nldr_scaled_obj`) with three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); the hexagonal object (`hb_obj`), the fitted model in both $2\text{-}D$ (`model_2d`), and $p\text{-}D$ (`model_highd`), and triangular mesh (`trimesh_data`).
-
-```{r, echo=TRUE, eval=FALSE}
-fit_highd_model(
-  highd_data = scurve, 
-  nldr_data = scurve_umap, 
-  b1 = 21, 
-  q = 0.1, 
-  hd_thresh = 0)
-```
-
-## Scaling the data
+In this section, we demonstrate the implementation of each step of the pipeline discussed in the Usage section (`fit_highd_model` and `comb_data_model`).
+Each step can be run independently to ensure flexibility in the modelling aproach.
 
@@
 
-## Computing hexagon grid configuration
-
-The function `calc_bins_y()` determines the configuration of the hexagon grid by computing the number of bins along the y-axis (`b2`), the hexagon binwidth (`a1`), and binheight (`a2`).
+### Hexagon binning
+
+The function `calc_bins_y()` determines the configuration of the hexagon grid by computing the number of bins along the y-axis (`b2`), the hexagon binwidth (`a1`), and hexagon binheight (`a2`).
 This function accepts (i) an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); (ii) the number of bins along the x-axis (`b1`), and (iii) the buffer amount as a proportion (`q`).
@@
 ```
-
-## Binning the data
 
@@
 
+The `group_hex_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon in the NLDR space.
+
+The function first groups the input data by `h`, which represents the hexagon ID associated with each point in the $2\text{-}D$ layout.
+Within each group, it collects the `ID`s into a list, resulting in a summary where each row corresponds to a single hexagon.
+The resulting column, `pts_list`, contains all point identifiers associated with that hexagon.
+
+```{r, echo=TRUE}
+pts_df <- group_hex_pts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+head(pts_df, 5)
+```
+
 ### Computing the standardized number of points within each hexagon
@@
 
-### Mapping the points to their corresponding hexagonal bins
-
-The `group_hex_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon in the NLDR space.
-
-The function first groups the input data by `h`, which represents the hexagon ID associated with each point in the $2\text{-}D$ layout.
-Within each group, it collects the `ID`s into a list, resulting in a summary where each row corresponds to a single hexagon.
-The resulting column, `pts_list`, contains all point identifiers associated with that hexagon.
-
-```{r, echo=TRUE}
-pts_df <- group_hex_pts(
-  scaled_nldr_hexid = umap_hex_id
-  )
-
-head(pts_df, 5)
-```
-
-## Obtaining bin centroids
+### Obtaining bin centroids
 
@@
 
-## Indicating neighbors by line segments connecting centroids
+### Indicating neighbors by line segments connecting centroids
 
 To represent the neighborhood structure of hexagonal bins in a $2\text{-}D$ layout, we employ Delaunay triangulation [@lee1980;@albrecht2024] on the centroids of hexagons.
-This geometric approach is used to infer which bins are considered neighbors.
 
 The `tri_bin_centroids()` function generates a triangulation object from the x and y coordinates of hexagon centroids using the `interp::tri.mesh()` function [@albrecht2024].
-This triangulation forms the structural basis for identifying adjacent bins.
 
@@
 It constructs a unique set of bin-to-bin connections by identifying the triangle edges and filtering duplicate or reversed links.
-Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function `calc_2d_dist()`.
+Each edge is then annotated with its start and end coordinates, and a Euclidean distance between the coordinates.
 

### 2025-11-30 paper-quollr.tex
Move usage to before implementation and re-order many things in implementation

@@
 
+\section{Usage}\label{usage}
+
+The package is available on CRAN, and the development version is available at \url{https://jayanilakshika.github.io/quollr/}.
+
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (Figure \ref{fig:algo-steps}).
+A detailed description of the algorithm can be found in \citet{gamage2025c}.
+
+The user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
+The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
+The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
+The built-in example datasets, \texttt{scurve} and \texttt{scurve\_umap} demonstrates these structures.
+
+To run the entire model pipeline, we can use the \texttt{fit\_high\_model()} function.
+This function requires: the high-dimensional data (\texttt{highd\_data}), the embedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of the data (\texttt{q}), and a benchmark value to identify high-density hexagons (\texttt{hd\_thresh}).
+
+The function returns an object of class \texttt{highd\_vis\_model} containing the scaled NLDR object (\texttt{nldr\_scaled\_obj}) with three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
+
+\begin{verbatim}
+model_highd <- fit_highd_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  b1 = 21, 
+  q = 0.1, 
+  hd_thresh = 0)
+\end{verbatim}
+
+The resulting model can then be shown in a tour using a two-step process:
+
+\begin{verbatim}
+combined_data <- comb_data_model(
+  highd_data = scurve,
+  model_highd = model_highd$model_highd,
+  model_2d = model_highd$model_2d
+)
+
+tour_view <- show_langevitour(
+  point_data = combined_data,
+  edge_data = model_highd$trimesh_data
+)
+\end{verbatim}
+
 \section{Implementation}\label{implementation}
@@
 
-\section{Usage}\label{usage}
-
-The package is available on CRAN, and the development version is available at \url{https://jayanilakshika.github.io/quollr/}.
-
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (Figure \ref{fig:algo-steps}).
-A detailed description of the algorithm can be found in \citet{gamage2025c}.
+\subsection{Pipeline implementation}\label{pipeline-implementation}
 
@@
 
-The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded.
-To begin the workflow, user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
-The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
-The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
-
-To illustrate the workflow, we use the built-in example dataset \texttt{scurve}, a \(7\text{-}D\) simulated dataset consisting of \(1000\) observations.
-The first three variables define a \(3\text{-}D\) S-shaped manifold, while the remaining four variables introduce low-magnitude uniform noise, yielding a structured yet noisy high-dimensional dataset.
-
-\begin{verbatim}
-data("scurve")
-\end{verbatim}
-
-For this example, we use the UMAP layout, which is used \texttt{n\_neighbors\ =} \(46\) and \texttt{min\_dist\ =} \(0.9\) as hyper-parameters.
-
-\begin{verbatim}
-library(umap)
-library(dplyr)
-
-scurve_umap <- umap(
-  scurve |> select(-ID),
-  config = modifyList(umap.defaults, list(
-    n_neighbors = 46,
-    n_components = 2,
-    min_dist = 0.9
-  ))
-)$layout |>
-  as_tibble(.name_repair = ~ paste0("emb", 1:2)) |>
-  mutate(ID = scurve$ID)
-\end{verbatim}
-
-\subsection{Main function}\label{main-function}
-
-The mains steps for the algorithm can be executed by the main function \texttt{fit\_highd\_model()}, or can be run separately for more flexibility.
-
-This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{b1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{hd\_thresh}).
-The function returns an object of class \texttt{highd\_vis\_model} containing the scaled NLDR object (\texttt{nldr\_scaled\_obj}) with three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); the hexagonal object (\texttt{hb\_obj}), the fitted model in both \(2\text{-}D\) (\texttt{model\_2d}), and \(p\text{-}D\) (\texttt{model\_highd}), and triangular mesh (\texttt{trimesh\_data}).
-
-\begin{verbatim}
-fit_highd_model(
-  highd_data = scurve, 
-  nldr_data = scurve_umap, 
-  b1 = 21, 
-  q = 0.1, 
-  hd_thresh = 0)
-\end{verbatim}
+In this section, we demonstrate the implementation of each step of the pipeline discussed in the Usage section (\texttt{fit\_highd\_model} and \texttt{comb\_data\_model}).
+Each step can be run independently to ensure flexibility in the modelling aproach.
 

### 2025-11-30 paper-quollr.tex
Added the current version

@@
 
-\subsection{Scaling the data}\label{scaling-the-data}
-
 The algorithm starts by scaling the NLDR data to to the range \([0, 1] \times [0, y_{2,max}]\), where \(y_{2,max} = r_2/r_1\) is the ratio of ranges of embedding components.
@@
 
-\subsection{Computing hexagon grid configuration}\label{computing-hexagon-grid-configuration}
-
-The function \texttt{calc\_bins\_y()} determines the configuration of the hexagon grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon binwidth (\texttt{a1}), and binheight (\texttt{a2}).
+\subsubsection{Hexagon binning}\label{hexagon-binning}
+
+The function \texttt{calc\_bins\_y()} determines the configuration of the hexagon grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon binwidth (\texttt{a1}), and hexagon binheight (\texttt{a2}).
 This function accepts (i) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (ii) the number of bins along the x-axis (\texttt{b1}), and (iii) the buffer amount as a proportion (\texttt{q}).
@@
 
-\subsection{Binning the data}\label{binning-the-data}
-
 Points are allocated to bins based on the nearest centroid of the hexagonal bins.
@@
 
+The \texttt{group\_hex\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon in the NLDR space.
+
+The function first groups the input data by \texttt{h}, which represents the hexagon ID associated with each point in the \(2\text{-}D\) layout.
+Within each group, it collects the \texttt{ID}s into a list, resulting in a summary where each row corresponds to a single hexagon.
+The resulting column, \texttt{pts\_list}, contains all point identifiers associated with that hexagon.
+
+\begin{verbatim}
+pts_df <- group_hex_pts(
+  scaled_nldr_hexid = umap_hex_id
+  )
+
+head(pts_df, 5)
+\end{verbatim}
+
+\begin{verbatim}
+> # A tibble: 5 x 2
+>       h pts_list 
+>   <int> <list>   
+> 1    58 <int [4]>
+> 2    68 <int [1]>
+> 3    69 <int [5]>
+> 4    70 <int [6]>
+> 5    71 <int [9]>
+\end{verbatim}
+
 \subsubsection{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
@@
 
-\subsubsection{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}
-
-The \texttt{group\_hex\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon in the NLDR space.
-
-The function first groups the input data by \texttt{h}, which represents the hexagon ID associated with each point in the \(2\text{-}D\) layout.
-Within each group, it collects the \texttt{ID}s into a list, resulting in a summary where each row corresponds to a single hexagon.
-The resulting column, \texttt{pts\_list}, contains all point identifiers associated with that hexagon.
-
-\begin{verbatim}
-pts_df <- group_hex_pts(
-  scaled_nldr_hexid = umap_hex_id
-  )
-
-head(pts_df, 5)
-\end{verbatim}
-
-\begin{verbatim}
-> # A tibble: 5 x 2
->       h pts_list 
->   <int> <list>   
-> 1    58 <int [4]>
-> 2    68 <int [1]>
-> 3    69 <int [5]>
-> 4    70 <int [6]>
-> 5    71 <int [9]>
-\end{verbatim}
-
-\subsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
+\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
 
@@
 
-\subsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
+\subsubsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
 
 To represent the neighborhood structure of hexagonal bins in a \(2\text{-}D\) layout, we employ Delaunay triangulation \citep{lee1980, albrecht2024} on the centroids of hexagons.
-This geometric approach is used to infer which bins are considered neighbors.
 
 The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{interp::tri.mesh()} function \citep{albrecht2024}.
-This triangulation forms the structural basis for identifying adjacent bins.
 
@@
 It constructs a unique set of bin-to-bin connections by identifying the triangle edges and filtering duplicate or reversed links.
-Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function \texttt{calc\_2d\_dist()}.
+Each edge is then annotated with its start and end coordinates, and a Euclidean distance between the coordinates.
 

### 2025-12-01 paper-quollr.Rmd
Added package installation code chunk

@@
 ---
-title: "quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space"
+title: "quollr: An R Package for Visualizing 2-D Models from Nonlinear Dimension Reductions in High-Dimensional Space"
 abstract: >
-  Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation.
+  Nonlinear dimension reduction methods provide a low-dimensional representation of high-dimensional data by applying a Nonlinear transformation.
 However, the complexity of the transformations and data structures can create wildly different representations depending on the method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
-The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which hyper-parameter choices provide the most accurate representation of high-dimensional data.
-The `scurve` data from the package is used to illustrate the algorithm.
+The R package quollr has been developed as a new visual tool to determine which method and which hyper-parameter choices provide the most accurate representation of high-dimensional data.
+The scurve data from the package is used to illustrate the algorithm.
 Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles are used to demonstrate the usability of the package.
@@
 
-options(repos = c(CRAN = "https://cran.rstudio.com")) # Setup mirror
-
-packages_to_check <- c("remotes", "quollr", "tibble", "knitr", "kableExtra", "ggplot2", "dplyr", "patchwork", "readr", "plotly", "crosstalk", "htmltools")
-
-for (pkg in packages_to_check) {
-  if (!requireNamespace(pkg, quietly = TRUE)) {
-    message(paste("Installing package:", pkg))
-    install.packages(pkg)
-  } else {
-    installed_version <- packageVersion(pkg)
-    available_version <- tryCatch({
-      utils::packageDescription(pkg)$Version
-    }, error = function(e) NA) # Handle cases where package info isn't readily available
-
-    if (!is.na(available_version) && installed_version < package_version(available_version)) {
-      message(paste("A newer version of package", pkg, "is available. Updating..."))
-      install.packages(pkg)
-    } else {
-      message(paste("Package", pkg, "is up-to-date (version", installed_version, ")."))
+# Ensure remotes is available
+if (!requireNamespace("remotes", quietly = TRUE)) {
+  install.packages("remotes")
 }
+
+# Read the package list
+pkgs_raw <- readLines("_Rpackages.txt")
+pkgs_raw <- pkgs_raw[nzchar(pkgs_raw)]  # remove empty lines
+
+# Split into package and version
+parts <- strsplit(pkgs_raw, "==")
+pkgs <- vapply(parts, `[[`, "", 1)
+versions <- vapply(parts, `[[`, "", 2)
+
+# Install packages with exact versions
+for (i in seq_along(pkgs)) {
+  pkg <- pkgs[i]
+  ver <- versions[i]
+
+  if (!requireNamespace(pkg, quietly = TRUE) ||
+      as.character(packageVersion(pkg)) != ver) {
+
+    message(sprintf("Installing %s (%s)", pkg, ver))
+    remotes::install_version(pkg, version = ver, upgrade = "never")
   }
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], can create hugely different representations depending on the selected method and hyper-parameter choices.
+Nonlinear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], can create hugely different representations depending on the selected method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 ```{r, echo=TRUE, eval=TRUE}
-model_highd <- fit_highd_model(
+model_obj <- fit_highd_model(
   highd_data = scurve, 
@@
   highd_data = scurve,
-  model_highd = model_highd$model_highd,
-  model_2d = model_highd$model_2d
+  model_highd = model_obj$model_highd,
+  model_2d = model_obj$model_2d
 )
@@
   point_data = combined_data,
-  edge_data = model_highd$trimesh_data
+  edge_data = model_obj$trimesh_data
 )
@@
 The result is a `tibble` of the scaled $2\text{-}D$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs. 
+  ?
 

### 2025-12-01 paper-quollr.tex
Added package installation code chunk

@@
 % !TeX root = RJwrapper.tex
-\title{quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space}
+\title{quollr: An R Package for Visualizing 2-D Models from Nonlinear Dimension Reductions in High-Dimensional Space}
 
@@
 \abstract{%
-Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (\(p\text{-}D\)) by applying a non-linear transformation.
+Nonlinear dimension reduction methods provide a low-dimensional representation of high-dimensional data by applying a Nonlinear transformation.
 However, the complexity of the transformations and data structures can create wildly different representations depending on the method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
-The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which hyper-parameter choices provide the most accurate representation of high-dimensional data.
-The \texttt{scurve} data from the package is used to illustrate the algorithm.
+The R package quollr has been developed as a new visual tool to determine which method and which hyper-parameter choices provide the most accurate representation of high-dimensional data.
+The scurve data from the package is used to illustrate the algorithm.
 Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles are used to demonstrate the usability of the package.
@@
 
-Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, can create hugely different representations depending on the selected method and hyper-parameter choices.
+Nonlinear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, can create hugely different representations depending on the selected method and hyper-parameter choices.
 It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.
@@
 \begin{verbatim}
-model_highd <- fit_highd_model(
+model_obj <- fit_highd_model(
   highd_data = scurve, 
@@
 The result is a \texttt{tibble} of the scaled \(2\text{-}D\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
+?
 

### 2025-12-02 paper-quollr.Rmd
Updated merge_hexbin_cetroid description

@@
 
-This function begins by arranging the `counts_data` by `h` to ensure consistent ordering.
-It then performs a full join with `centroids_data`, aligning hexagon IDs (`h`) between the two datasets to incorporate both hexagonal bin centroids (`h`) and count metrics.
+This function performs a full join with `centroids_data`, aligning hexagon IDs (`h`) between the two datasets to incorporate both hexagonal bin centroids (`h`) and count metrics.
 After merging, the function handles missing values in the count columns: any `NA` values in `w_h` or `n_h` are replaced with zeros.

### 2025-12-02 paper-quollr.tex
Updated merge_hexbin_cetroid description

@@
 
-This function begins by arranging the \texttt{counts\_data} by \texttt{h} to ensure consistent ordering.
-It then performs a full join with \texttt{centroids\_data}, aligning hexagon IDs (\texttt{h}) between the two datasets to incorporate both hexagonal bin centroids (\texttt{h}) and count metrics.
+This function performs a full join with \texttt{centroids\_data}, aligning hexagon IDs (\texttt{h}) between the two datasets to incorporate both hexagonal bin centroids (\texttt{h}) and count metrics.
 After merging, the function handles missing values in the count columns: any \texttt{NA} values in \texttt{w\_h} or \texttt{n\_h} are replaced with zeros.

### 2025-12-02 paper-quollr.Rmd
Updated hexagon ID to hexagon

@@
 
-The output is a tibble containing a unique hexagon ID (`h`) and centroid coordinates (`c_x`, `c_y`), which together define the layout of the complete hexagonal grid over the $2\text{-}D$ space.
+The output is a tibble containing a unique hexagon (`h`) and centroid coordinates (`c_x`, `c_y`), which together define the layout of the complete hexagonal grid over the $2\text{-}D$ space.
 
@@
 The function uses the `proxy::dist()` method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
-For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon—and assigns the corresponding hexagon ID (`h`) to the point.
+For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon—and assigns the corresponding hexagon (`h`) to the point.
 
@@
 
-The function first groups the input data by `h`, which represents the hexagon ID associated with each point in the $2\text{-}D$ layout.
+The function first groups the input data by `h`, which represents the hexagon associated with each point in the $2\text{-}D$ layout.
 Within each group, it collects the `ID`s into a list, resulting in a summary where each row corresponds to a single hexagon.
@@
 
-The function begins by grouping the data by hexagon ID (`h`) and counting the number of NLDR points falling within each bin.
+The function begins by grouping the data by hexagon (`h`) and counting the number of NLDR points falling within each bin.
 These raw counts are stored as `n_h`.
@@
 
-The `find_low_dens_hex()` function relies on a helper, `compute_mean_density_hex()`, which iterates over all hexagons and computes the average density across neighbors based on their hexagon ID (`h`) and a defined number of bins along the x-axis (`b1`).
+The `find_low_dens_hex()` function relies on a helper, `compute_mean_density_hex()`, which iterates over all hexagons and computes the average density across neighbors based on their hexagon (`h`) and a defined number of bins along the x-axis (`b1`).
 The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.

### 2025-12-02 paper-quollr.tex
Updated hexagon ID to hexagon

@@
 
-The output is a tibble containing a unique hexagon ID (\texttt{h}) and centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which together define the layout of the complete hexagonal grid over the \(2\text{-}D\) space.
+The output is a tibble containing a unique hexagon (\texttt{h}) and centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which together define the layout of the complete hexagonal grid over the \(2\text{-}D\) space.
 
@@
 The function uses the \texttt{proxy::dist()} method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
-For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon---and assigns the corresponding hexagon ID (\texttt{h}) to the point.
+For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon---and assigns the corresponding hexagon (\texttt{h}) to the point.
 
@@
 
-The function first groups the input data by \texttt{h}, which represents the hexagon ID associated with each point in the \(2\text{-}D\) layout.
+The function first groups the input data by \texttt{h}, which represents the hexagon associated with each point in the \(2\text{-}D\) layout.
 Within each group, it collects the \texttt{ID}s into a list, resulting in a summary where each row corresponds to a single hexagon.
@@
 
-The function begins by grouping the data by hexagon ID (\texttt{h}) and counting the number of NLDR points falling within each bin.
+The function begins by grouping the data by hexagon (\texttt{h}) and counting the number of NLDR points falling within each bin.
 These raw counts are stored as \texttt{n\_h}.
@@
 
-The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their hexagon ID (\texttt{h}) and a defined number of bins along the x-axis (\texttt{b1}).
+The \texttt{find\_low\_dens\_hex()} function relies on a helper, \texttt{compute\_mean\_density\_hex()}, which iterates over all hexagons and computes the average density across neighbors based on their hexagon (\texttt{h}) and a defined number of bins along the x-axis (\texttt{b1}).
 The hexagonal layout assumes a fixed grid structure, so neighbor IDs are computed by positional offsets.

### 2025-12-02 paper-quollr.Rmd
Updated hexgon IDs to hexagons

@@
 The parameters used within `hex_binning()` are the object output from `gen_scaled_data` (`nldr_scaled_obj`); the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
-The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagon IDs (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`b`), the number of non-empty bins (`m`), and the points within each hexagon (`pts_bins`).
+The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagons (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`b`), the number of non-empty bins (`m`), and the points within each hexagon (`pts_bins`).
 
@@
 
-The `group_hex_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon in the NLDR space.
+The `group_hex_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon.
 
@@
 
-This function performs a full join with `centroids_data`, aligning hexagon IDs (`h`) between the two datasets to incorporate both hexagonal bin centroids (`h`) and count metrics.
+This function performs a full join with `centroids_data`, aligning hexagons (`h`) between the two datasets to incorporate both hexagonal bin centroids (`h`) and count metrics.
 After merging, the function handles missing values in the count columns: any `NA` values in `w_h` or `n_h` are replaced with zeros.

### 2025-12-02 paper-quollr.tex
Updated hexgon IDs to hexagons

@@
 The parameters used within \texttt{hex\_binning()} are the object output from \texttt{gen\_scaled\_data} (\texttt{nldr\_scaled\_obj}); the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
-The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagon IDs (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{b}), the number of non-empty bins (\texttt{m}), and the points within each hexagon (\texttt{pts\_bins}).
+The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagons (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{b}), the number of non-empty bins (\texttt{m}), and the points within each hexagon (\texttt{pts\_bins}).
 
@@
 
-The \texttt{group\_hex\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon in the NLDR space.
+The \texttt{group\_hex\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon.
 
@@
 
-This function performs a full join with \texttt{centroids\_data}, aligning hexagon IDs (\texttt{h}) between the two datasets to incorporate both hexagonal bin centroids (\texttt{h}) and count metrics.
+This function performs a full join with \texttt{centroids\_data}, aligning hexagons (\texttt{h}) between the two datasets to incorporate both hexagonal bin centroids (\texttt{h}) and count metrics.
 After merging, the function handles missing values in the count columns: any \texttt{NA} values in \texttt{w\_h} or \texttt{n\_h} are replaced with zeros.

### 2025-12-02 paper-quollr.Rmd
Updated Figure 3

@@
              aes(x = emb1, y = emb2), 
-             alpha = 0.2, size = 0.5, color = clr_choice) + 
+             alpha = 0.8, size = 0.5, color = clr_choice) + 
   interior_annotation("a", cex = 3) +
@@
 ## Pipeline implementation
+
+In this section, we demonstrate the implementation of each step of the pipeline discussed in the Usage section (`fit_highd_model` and `comb_data_model`).
+Each step can be run independently to ensure flexibility in the modelling aproach.
 
@@
 ```
-
-In this section, we demonstrate the implementation of each step of the pipeline discussed in the Usage section (`fit_highd_model` and `comb_data_model`).
-Each step can be run independently to ensure flexibility in the modelling aproach.
 
@@
 The result is a `tibble` of the scaled $2\text{-}D$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
-  ?
 
@@
 ### Computing the standardized number of points within each hexagon
+
+After assiging data to the hexagons,
 

### 2025-12-04 paper-quollr.Rmd
usage section and installing packages chunk modified

@@
 
-```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE}
+```{r install-libraries, include=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
 
@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`).
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions.
 A detailed description of the algorithm can be found in @gamage2025c.
@@
 ```
+
+which produces the model and data plot(s) shown in `r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`.
 

### 2025-12-04 paper-quollr.tex
Updated Figure 3
usage section and installing packages chunk modified

@@
 
-Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions (Figure \ref{fig:algo-steps}).
+Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, and (6) lifting the model into high dimensions.
 A detailed description of the algorithm can be found in \citet{gamage2025c}.
@@
 \end{verbatim}
+
+which produces the model and data plot(s) shown in Figure \ref{fig:algo-steps}.
 
@@
 
+In this section, we demonstrate the implementation of each step of the pipeline discussed in the Usage section (\texttt{fit\_highd\_model} and \texttt{comb\_data\_model}).
+Each step can be run independently to ensure flexibility in the modelling aproach.
+
 \begin{figure}[!ht]
@@
 \end{figure}
-
-In this section, we demonstrate the implementation of each step of the pipeline discussed in the Usage section (\texttt{fit\_highd\_model} and \texttt{comb\_data\_model}).
-Each step can be run independently to ensure flexibility in the modelling aproach.
 
@@
 The result is a \texttt{tibble} of the scaled \(2\text{-}D\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
-?
 
@@
 \subsubsection{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
+
+After assiging data to the hexagons,
 

### 2025-12-04 paper-quollr.R

@@ -14,35 +14,35 @@ knitr::opts_chunk$set(
 
 
 
-## ----install-libraries, include=FALSE, warning=FALSE, echo=FALSE--------------
-
-# Ensure remotes is available
-if (!requireNamespace("remotes", quietly = TRUE)) {
-  install.packages("remotes")
-}
-
-# Read the package list
-pkgs_raw <- readLines("_Rpackages.txt")
-pkgs_raw <- pkgs_raw[nzchar(pkgs_raw)]  # remove empty lines
-
-# Split into package and version
-parts <- strsplit(pkgs_raw, "==")
-pkgs <- vapply(parts, `[[`, "", 1)
-versions <- vapply(parts, `[[`, "", 2)
-
-# Install packages with exact versions
-for (i in seq_along(pkgs)) {
-  pkg <- pkgs[i]
-  ver <- versions[i]
-
-  if (!requireNamespace(pkg, quietly = TRUE) ||
-      as.character(packageVersion(pkg)) != ver) {
-
-    message(sprintf("Installing %s (%s)", pkg, ver))
-    remotes::install_version(pkg, version = ver, upgrade = "never")
-  }
-}
-
+## ----install-libraries, include=FALSE, warning=FALSE, echo=FALSE, eval=FALSE----
+# 
+# # Ensure remotes is available
+# if (!requireNamespace("remotes", quietly = TRUE)) {
+#   install.packages("remotes")
+# }
+# 
+# # Read the package list
+# pkgs_raw <- readLines("_Rpackages.txt")
+# pkgs_raw <- pkgs_raw[nzchar(pkgs_raw)]  # remove empty lines
+# 
+# # Split into package and version
+# parts <- strsplit(pkgs_raw, "==")
+# pkgs <- vapply(parts, `[[`, "", 1)
+# versions <- vapply(parts, `[[`, "", 2)
+# 
+# # Install packages with exact versions
+# for (i in seq_along(pkgs)) {
+#   pkg <- pkgs[i]
+#   ver <- versions[i]
+# 
+#   if (!requireNamespace(pkg, quietly = TRUE) ||
+#       as.character(packageVersion(pkg)) != ver) {
+# 
+#     message(sprintf("Installing %s (%s)", pkg, ver))
+#     remotes::install_version(pkg, version = ver, upgrade = "never")
+#   }
+# }
+# 
 
 
 ## ----load-libraries-----------------------------------------------------------

### 2025-12-04 paper-quollr.Rmd
layout of fig 2 in pdf, and it needs to be linked plots not link plots

@@
   scurve_proj_umap_model1_dp +
-  plot_layout(ncol = 4)
+  plot_layout(ncol = 2)
 ```
@@
 
-### Link plots
-
-Two types of interactive link plots can be generated to assess the model fits everywhere, or better in some subspaces, or completely mismatch the data.
+### Linked plots
+
+Two types of interactively linked plots can be generated to assess the model fits everywhere, or better in some subspaces, or completely mismatch the data.
 The plots are linked using `crosstalk`, which allows interactive brushing: selecting or brushing points in one plot automatically highlights the corresponding points in the other linked views.

### 2025-12-04 paper-quollr.tex
layout of fig 2 in pdf, and it needs to be linked plots not link plots

@@
 
-\subsubsection{Link plots}\label{link-plots}
-
-Two types of interactive link plots can be generated to assess the model fits everywhere, or better in some subspaces, or completely mismatch the data.
+\subsubsection{Linked plots}\label{linked-plots}
+
+Two types of interactively linked plots can be generated to assess the model fits everywhere, or better in some subspaces, or completely mismatch the data.
 The plots are linked using \texttt{crosstalk}, which allows interactive brushing: selecting or brushing points in one plot automatically highlights the corresponding points in the other linked views.

### 2025-12-04 paper-quollr.R

@@ -533,7 +533,7 @@ knitr::include_graphics("figures/quollr_workflow.png")
 hex_grid_poly_scurve +
   hex_centroids_scurve + wireframe_scurve + 
   scurve_proj_umap_model1_dp +
-  plot_layout(ncol = 4)
+  plot_layout(ncol = 2)
 
 
 ## ----echo=TRUE----------------------------------------------------------------

### 2025-12-04 paper-quollr.Rmd
sizing figures

@@
 Panels (a1–a4) display different $2\\text{-}D$ projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
-The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos="!ht", layout = "l-page", fig.alt="The UMAP layout with a triangular mesh on the left and 2-D projections of the lifted wireframe overlaying the data on the right.
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions.", fig.pos="!ht", layout = "l-body", fig.alt="The UMAP layout with a triangular mesh on the left and 2-D projections of the lifted wireframe overlaying the data on the right.
 The mesh illustrates how local neighborhoods in the 2-D embedding correspond to curved sheet structures in the original high-dimensional S-curve data."}
@@
 ```{r algo-step-html, eval=knitr::is_html_output(), fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `scurve` data is shown.", layout = "l-page", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
+The `scurve` data is shown.", layout = "l-body", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
 
@@
 
-```{r algo-steps, eval=knitr::is_latex_output(), out.width="100%", fig.height=5, fig.width=25, fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
+```{r algo-steps, eval=knitr::is_latex_output(), out.width="80%", fig.height=5, fig.width=25, fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `scurve` data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
@@
                     max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 0.5, alpha = 0.5, 
+  show_scatter(axes = TRUE, size = 2, alpha = 0.5, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
@@
 
-```{r scurve-nldrlink-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+```{r scurve-nldrlink-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
 Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `scurve` structure.
@@
 ```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `scurve` structure in $7\\text{-}D$.
-Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
+Two sets of plots are interactively linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
 The purple points indicate the selected subsets, which differ between rows.

### 2025-12-04 paper-quollr.tex
sizing figures

@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
+\includegraphics[width=0.8\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `scurve` data is shown.}\label{fig:algo-steps}
@@
 The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `scurve` structure in $7\text{-}D$.
-Two sets of plots are linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
+Two sets of plots are interactively linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
 The purple points indicate the selected subsets, which differ between rows.

### 2025-12-04 paper-quollr.Rmd
more text sizing

@@
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
-  interior_annotation("a", cex = 3) +
+  interior_annotation("a", cex = 1) +
   xlim(sc_xlims) +
@@
              alpha = 0.8, size = 0.5, color = clr_choice) + 
-  interior_annotation("a", cex = 3) +
+  interior_annotation("a", cex = 1) +
   xlim(sc_xlims) +
@@
              size = 1, color = "#FF7755") +
-  interior_annotation("b", cex = 3) +
+  interior_annotation("b", cex = 1) +
   xlim(sc_xlims) +
@@
              size = 1, color = "#FF7755") +
-  interior_annotation("c", cex = 3) +
+  interior_annotation("c", cex = 1) +
   xlim(sc_xlims) +
@@
   plot_limits = c(-0.35, 0.35), 
-  axis_text_size = 7,
+  axis_text_size = 3,
   is_category = FALSE) +
-  interior_annotation(label = "d", cex = 3)
+  interior_annotation(label = "d", cex = 1)
 
@@
 
-```{r algo-steps, eval=knitr::is_latex_output(), out.width="80%", fig.height=5, fig.width=25, fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
+```{r algo-steps, eval=knitr::is_latex_output(), out.width="80%", fig.height=5, fig.width=25, fig.pos="!ht", fig.align='center', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `scurve` data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
@@
                     max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 2, alpha = 0.5, 
+  show_scatter(axes = TRUE, size = 1.5, alpha = 0.5, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),

### 2025-12-04 paper-quollr.tex
more text sizing

@@
 \begin{figure}[!ht]
-\includegraphics[width=0.8\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1} \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
+
+{\centering \includegraphics[width=0.8\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1}
+
+}
+
+\caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
 The `scurve` data is shown.}\label{fig:algo-steps}

### 2025-12-04 paper-quollr.R

@@ -126,7 +126,7 @@ scurve_umap_plt <- ggplot(
   scurve_umap_scaled, 
   aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
-  interior_annotation("a", cex = 3) +
+  interior_annotation("a", cex = 1) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -152,7 +152,7 @@ hex_grid_poly_scurve <- ggplot(
   geom_point(data = scurve_umap_scaled, 
              aes(x = emb1, y = emb2), 
              alpha = 0.8, size = 0.5, color = clr_choice) + 
-  interior_annotation("a", cex = 3) +
+  interior_annotation("a", cex = 1) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -178,7 +178,7 @@ hex_centroids_scurve <- ggplot(
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
              size = 1, color = "#FF7755") +
-  interior_annotation("b", cex = 3) +
+  interior_annotation("b", cex = 1) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -206,7 +206,7 @@ wireframe_scurve <- ggplot() +
   geom_point(data = df_bin_centroids_scurve, 
              aes(x = c_x, y = c_y), 
              size = 1, color = "#FF7755") +
-  interior_annotation("c", cex = 3) +
+  interior_annotation("c", cex = 1) +
   xlim(sc_xlims) +
   ylim(sc_ylims)
 
@@ -377,9 +377,9 @@ scurve_proj_umap_model1_dp <- plot_proj(
   proj_obj = proj_obj1,
   point_param = c(0.2, 0, clr_choice), # size, alpha, color
   plot_limits = c(-0.35, 0.35), 
-  axis_text_size = 7,
+  axis_text_size = 3,
   is_category = FALSE) +
-  interior_annotation(label = "d", cex = 3)
+  interior_annotation(label = "d", cex = 1)
 
 ## Second projection
 
@@ -528,7 +528,7 @@ knitr::include_graphics("figures/quollr_workflow.png")
 # algofig
 
 
[…]

### 2025-12-04 paper-quollr.Rmd
New PDF

@@
 ```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published $2\\text{-}D$ layout (b) and the HBE-selected $2\\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the published 2-D layout (b) of limb muscle data on the left and the fitted high-dimensional model in 10-D on the right using langevitour.
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the published $2\\text{-}D$ layout (b) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
 Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the 2-D embedding correspond to structures in the original 10-D space.", layout = "l-page"}
+Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-page"}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the HBE-selected 2-D layout (f) of limb muscle data on the left and the fitted high-dimensional model in 10-D on the right using langevitour.
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the HBE-selected $2\\text{-}D$ layout (f) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
 Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the 2-D embedding correspond to structures in the original 10-D space.", layout = "l-page"}
+Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-page"}
 

### 2025-12-04 paper-quollr.tex
New PDF

@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+\includegraphics[width=1\linewidth,alt={Comparison of published $2\text{-}D$ layout (b) and the HBE-selected $2\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-12-04 paper-quollr.R

@@ -2030,7 +2030,7 @@ limb_proj_tsne_best_model2 <- plot_proj(
 
 
 
-## ----model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters. The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published 2-D layout (b) and the HBE-selected 2-D layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in 10-D, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()----
+## ----model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters. The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published $2\\text{-}D$ layout (b) and the HBE-selected $2\\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()----
 
 trimesh_limb + limb_proj_tsne_model1 +
   limb_proj_tsne_model2 + trimesh_limb_best +
@@ -2038,14 +2038,14 @@ trimesh_limb + limb_proj_tsne_model1 +
   plot_layout(nrow = 2)
 
 
[…]

### 2025-12-08 paper-quollr.Rmd
Fixed the Figure 3

@@
       ",
-      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 225px;', hex_grid_poly_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 300px;', hex_grid_poly_scurve_int),
       htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px;', hex_centroids_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 280px; width: 225px;', wireframe_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 300px;', wireframe_scurve_int),
       htmltools::div(style = 'margin: 0; padding: 0; margin-left: 60px; height: 300px;', scurve_umap_model_vis_n)
@@
 
-nldr_plt <- ggplotly(nldr_plt, width = 600,
-                     height = 600, tooltip = "none") |>
+nldr_plt <- ggplotly(nldr_plt, width = 300,
+                     height = 300, tooltip = "none") |>
   style(unselected=list(marker=list(opacity=1))) |>
@@
                                                link=shared_df,
-                                               linkFilter=FALSE)
+                                               linkFilter=FALSE,
+                                               width = "200px", height = "100px")
 
 linked_plt_best <- crosstalk::bscols(
-  htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
-                 nldr_plt,
-                 htmltools::div(style = "margin-top: 20px;", langevitour_output)
+    htmltools::div(
+        style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: center;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
+        nldr_plt, langevitour_output
     ),

### 2025-12-08 paper-quollr.R

@@ -513,9 +513,9 @@ knitr::include_graphics("figures/quollr_workflow.png")
 #         margin: 0;
 #         padding: 0;
 #       ",
-#       htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 225px;', hex_grid_poly_scurve_int),
+#       htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 300px;', hex_grid_poly_scurve_int),
 #       htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px;', hex_centroids_scurve_int),
-#       htmltools::div(style = 'margin: 0; padding: 0; height: 280px; width: 225px;', wireframe_scurve_int),
+#       htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 300px;', wireframe_scurve_int),
 #       htmltools::div(style = 'margin: 0; padding: 0; margin-left: 60px; height: 300px;', scurve_umap_model_vis_n)
 #     )
 #   ),
@@ -1915,8 +1915,8 @@ nldr_plt <- shared_df |>
     legend.position = "none"
   )
 
-nldr_plt <- ggplotly(nldr_plt, width = 600,
-                     height = 600, tooltip = "none") |>
+nldr_plt <- ggplotly(nldr_plt, width = 300,
+                     height = 300, tooltip = "none") |>
   style(unselected=list(marker=list(opacity=1))) |>
   highlight(on="plotly_selected", off="plotly_deselect") |>
   config(displayModeBar = FALSE)
@@ -1931,14 +1931,20 @@ langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
                                                                   rep(point_sizes[2], NROW(df))),
                                                levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
                                                link=shared_df,
-                                               linkFilter=FALSE)
+                                               linkFilter=FALSE,
+                                               width = "200px", height = "100px")
 
 linked_plt_best <- crosstalk::bscols(
-  htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
-                 nldr_plt,
-                 htmltools::div(style = "margin-top: 20px;", langevitour_output)
-  ),
-  device = "xs"
+    htmltools::div(
+        style = "display: grid; grid-template-columns: 1fr 1fr;
[…]

### 2025-12-08 paper-quollr.Rmd
Fixed the caption width problem

@@
 Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right, displayed using langevitour.
-Brushing highlights matching points across both views, enabling comparison of structure preservation between the 2-D layout and the high-dimensional data.", layout = "l-page"}
+Brushing highlights matching points across both views, enabling comparison of structure preservation between the 2-D layout and the high-dimensional data.", layout = "l-body"}
 
@@
 Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel link plot showing the residual distribution (left), UMAP layout (middle), and fitted model overlaid with S-curve data in seven dimensions (right) using langevitour.
-Brushing highlights corresponding points across all panels, illustrating how high-residual points align with sparse or bridging regions in the UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-page"}
+Brushing highlights corresponding points across all panels, illustrating how high-residual points align with sparse or bridging regions in the UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-body"}
 
@@
                palette = c("#66B2CC", "#FF7755"),
-                width = "600px", height = "600px")
+                width = "300px", height = "300px")
 
@@
 Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right using detourr.
-Brushing or selecting points in either view highlights the corresponding points in the other panel, allowing assessment of how well the UMAP layout preserves the high-dimensional S-curve structure.", layout = "l-page"}
+Brushing or selecting points in either view highlights the corresponding points in the other panel, allowing assessment of how well the UMAP layout preserves the high-dimensional S-curve structure.", layout = "l-body"}
 
 lndet_link <- crosstalk::bscols(
-    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
-                   nldr_plt,
-                   htmltools::div(style = "margin-top: 20px;", detourr_output)
+    htmltools::div(
+        style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
+        nldr_plt, htmltools::div(style = "margin-top: 20px;", detourr_output)
     ),
@@
 Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel link plot showing the residual distribution (left), UMAP layout (middle), and fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr.
-Brushing or selecting points in any panel highlights corresponding points across all views, illustrating how high-residual points correspond to sparse or bridging regions in UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-page"}
+Brushing or selecting points in any panel highlights corresponding points across all views, illustrating how high-residual points correspond to sparse or bridging regions in UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-body"}
 
@@
 Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-page"}
+Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-body"}
 
@@
 Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-page"}
+Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-body"}
 

### 2025-12-08 paper-quollr.tex
Fixed the caption width problem

@@
                palette = c("#66B2CC", "#FF7755"),
-                width = "600px", height = "600px")
+                width = "300px", height = "300px")
 \end{verbatim}
@@
 lndet_link <- crosstalk::bscols(
-    htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
-                   nldr_plt,
-                   htmltools::div(style = "margin-top: 20px;", detourr_output)
+    htmltools::div(
+        style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
+        nldr_plt, htmltools::div(style = "margin-top: 20px;", detourr_output)
     ),

### 2025-12-08 paper-quollr.Rmd
Updated detourr link plots

@@
 
-error_plt <- ggplotly(error_plt, width = 400,
-                      height = 400, tooltip = "none") |>
+error_plt <- ggplotly(error_plt, width = 200,
+                      height = 200, tooltip = "none") |>
   style(unselected=list(marker=list(opacity=1))) |>
@@
 
-nldr_plt <- ggplotly(nldr_plt, width = 400,
-                     height = 400, tooltip = "none") |>
+nldr_plt <- ggplotly(nldr_plt, width = 200,
+                     height = 200, tooltip = "none") |>
   style(unselected=list(marker=list(opacity=1))) |>
@@
                palette = c("#66B2CC", "#FF7755"),
-                width = "500px", height = "500px")
+                width = "200px", height = "200px")
 
@@
     htmltools::div(
-    style = "display: grid; grid-template-columns: 1fr 1fr 1fr;",
+        style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
         error_plt, 

### 2025-12-08 paper-quollr.Rmd
Fixed applicarion figures and detourr figures

@@
 The user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
-The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `"x"` (e.g., `x1`, `x2`, etc.).
+The high-dimensional data must contain a unique `ID` column, with data columns prefixed by the letter `x` (e.g., `x1`, `x2`, etc.).
 The NLDR dataset should include embedding coordinates labeled as `emb1` and `emb2`, ensuring one-to-one correspondence with the high-dimensional data through the shared `ID`.
@@
 
-- `highd_data`: a tibble containing the original high-dimensional observations with a unique identifier (`ID`) and variable columns prefixed with `"x"` (e.g., `x1`, `x2`, …).
+- `highd_data`: a tibble containing the original high-dimensional observations with a unique identifier (`ID`) and variable columns prefixed with `x` (e.g., `x1`, `x2`, …).
 
@@
 Points are allocated to bins based on the nearest centroid of the hexagonal bins.
-The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its individual components can be run separately for added flexibility.
+The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its individual steps can be run separately.
 While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the `hex_binning()` function automates this entire workflow.
@@
 
-After assiging data to the hexagons,
-
 The `compute_std_counts()` function calculates both the raw and standardized counts of points inside each hexagon.
@@
 
-```{r geom-outputs-pdf, fig.cap="The outputs of `geom\\_hexgrid` and `geom\\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_latex_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
+```{r geom-outputs-pdf, fig.cap="The outputs of \\texttt{geom\\_hexgrid} and \\texttt{geom\\_trimesh} include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_latex_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
 
@@
 
-In this structure, the `type` variable distinguishes between original observations (`"data"`) and the bin-averaged model representation (`"model"`).
+In this structure, the `type` variable distinguishes between original observations (`data`) and the bin-averaged model representation (`model`).
 
@@
 
-```{r scurve-highd-model-html, echo=TRUE, eval=knitr::is_html_output(), fig.pos="!ht", fig.width=20, fig.height=5, fig.cap="`langevitour` output of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
-These views illustrate how the lifted wireframe model captures the structure of the `scurve` data.
+```{r scurve-highd-model-html, echo=TRUE, eval=knitr::is_html_output(), fig.pos="!ht", fig.width=20, fig.height=5, fig.cap="\\texttt{langevitour} output of the lifted high-dimensional wireframe model from the \\texttt{scurve} UMAP layout.
+These views illustrate how the lifted wireframe model captures the structure of the \\texttt{scurve} data.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Interactive langevitour visualization showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
@@
 
-```{r scurve-highd-model-detourrpdf, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the `scurve` UMAP layout using `detourr`.
+```{r scurve-highd-model-detourrpdf, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the \\texttt{scurve} UMAP layout using \\texttt{detourr}.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
@@
 
-```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.", fig.pos="H", fig.show='hold'}
+```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.", fig.pos="H", fig.show='hold'}
 
@@
 
-```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 
@@
 
-nldr_plt <- ggplotly(nldr_plt, width = 600,
-                     height = 600, tooltip = "none") |>
+nldr_plt <- ggplotly(nldr_plt, width = 250,
+                     height = 250, tooltip = "none") |>
   style(unselected=list(marker=list(opacity=1))) |>
@@
                                                link=shared_df,
-                                               linkFilter=FALSE)
+                                               linkFilter=FALSE,
+                                               width = "350px", height = "250px")
 
 linked_plt_int <- crosstalk::bscols(
-  htmltools::div(style="display: grid; grid-template-columns: 1fr 1fr;",
+    htmltools::div(
+        style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: center;
+    justify-items: center;
+    margin: 0;
+    padding: 0;",
         nldr_plt, 
-                 htmltools::div(style = "margin-top: 20px;", langevitour_output)
+        htmltools::div(style = "margin-top: 30px;", langevitour_output)
     ),
@@
 )
+
 ```
@@
 
-nldr_plt <- ggplotly(nldr_plt, width = 300,
-                     height = 300, tooltip = "none") |>
+nldr_plt <- ggplotly(nldr_plt, width = 250,
+                     height = 250, tooltip = "none") |>
   style(unselected=list(marker=list(opacity=1))) |>
@@
                                                linkFilter=FALSE,
-                                               width = "200px", height = "100px")
+                                               width = "350px", height = "250px")
 
@@
     padding: 0;",
-        nldr_plt, langevitour_output
+        nldr_plt, 
+        htmltools::div(style = "margin-top: 20px;", langevitour_output)
     ),

### 2025-12-08 paper-quollr.tex
Fixed applicarion figures and detourr figures

@@
 The user needs two inputs: the high-dimensional dataset and the corresponding NLDR data.
-The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, etc.).
+The high-dimensional data must contain a unique \texttt{ID} column, with data columns prefixed by the letter \texttt{x} (e.g., \texttt{x1}, \texttt{x2}, etc.).
 The NLDR dataset should include embedding coordinates labeled as \texttt{emb1} and \texttt{emb2}, ensuring one-to-one correspondence with the high-dimensional data through the shared \texttt{ID}.
@@
 \item
-  \texttt{highd\_data}: a tibble containing the original high-dimensional observations with a unique identifier (\texttt{ID}) and variable columns prefixed with \texttt{"x"} (e.g., \texttt{x1}, \texttt{x2}, \ldots).
+  \texttt{highd\_data}: a tibble containing the original high-dimensional observations with a unique identifier (\texttt{ID}) and variable columns prefixed with \texttt{x} (e.g., \texttt{x1}, \texttt{x2}, \ldots).
 \item
@@
 Points are allocated to bins based on the nearest centroid of the hexagonal bins.
-The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its individual components can be run separately for added flexibility.
+The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its individual steps can be run separately.
 While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the \texttt{hex\_binning()} function automates this entire workflow.
@@
 \subsubsection{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}
-
-After assiging data to the hexagons,
 
@@
 
-\caption{The outputs of `geom\_hexgrid` and `geom\_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.}\label{fig:geom-outputs-pdf}
+\caption{The outputs of \texttt{geom\_hexgrid} and \texttt{geom\_trimesh} include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.}\label{fig:geom-outputs-pdf}
 \end{figure}
@@
 
-In this structure, the \texttt{type} variable distinguishes between original observations (\texttt{"data"}) and the bin-averaged model representation (\texttt{"model"}).
+In this structure, the \texttt{type} variable distinguishes between original observations (\texttt{data}) and the bin-averaged model representation (\texttt{model}).
 
@@
 The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the `scurve` UMAP layout using `detourr`.
+The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout using \texttt{detourr}.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-detourrpdf}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-nldrlink-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-nldrlink-dtpdf}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using `detourr`.}\label{fig:scurve-linkerror-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-linkerror-dtpdf}
 \end{figure}

### 2025-12-08 paper-quollr.Rmd
Started adding labels for figure 3

@@
       ",
-      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 300px;', hex_grid_poly_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px;', hex_centroids_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 300px;', wireframe_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; margin-left: 60px; height: 300px;', scurve_umap_model_vis_n)
+      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 300px; text-align: left;', htmltools::h4("a"), hex_grid_poly_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px; text-align: left;', htmltools::h4("b"), hex_centroids_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 300px;  text-align: left;', htmltools::h4("c"), wireframe_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; margin-left: 60px; height: 300px;  text-align: left;', htmltools::h4("d"), scurve_umap_model_vis_n)
     )

### 2025-12-08 paper-quollr.R

@@ -513,10 +513,10 @@ knitr::include_graphics("figures/quollr_workflow.png")
 #         margin: 0;
 #         padding: 0;
 #       ",
-#       htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 300px;', hex_grid_poly_scurve_int),
-#       htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px;', hex_centroids_scurve_int),
-#       htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 300px;', wireframe_scurve_int),
-#       htmltools::div(style = 'margin: 0; padding: 0; margin-left: 60px; height: 300px;', scurve_umap_model_vis_n)
+#       htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 300px; text-align: left;', htmltools::h4("a"), hex_grid_poly_scurve_int),
+#       htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px; text-align: left;', htmltools::h4("b"), hex_centroids_scurve_int),
+#       htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 300px;  text-align: left;', htmltools::h4("c"), wireframe_scurve_int),
+#       htmltools::div(style = 'margin: 0; padding: 0; height: 300px;  text-align: left;', htmltools::h4("d"), scurve_umap_model_vis_n)
 #     )
 #   ),
 #   device = "xs"

### 2025-12-08 paper-quollr.Rmd
Found the way to fix the issue of selecting, deselecting and add the box

--- 2025-12-08 paper-quollr.Rmd
@@
       ",
-      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 300px;', hex_grid_poly_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 400px;', hex_centroids_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 300px;', wireframe_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; margin-left: 60px; height: 300px;', scurve_umap_model_vis_n)
+      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 370px; text-align: left;', htmltools::h4("a"), hex_grid_poly_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 300px; text-align: left;', htmltools::h4("b"), hex_centroids_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 370px;  text-align: left;', htmltools::h4("c"), wireframe_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 300px; text-align: left;', htmltools::h4("d"), htmltools::div(
+    style = "margin-left: 90px;",
+    scurve_umap_model_vis_n
+  ))
     )
@@
 
-nldr_plt <- shared_df |>
-  ggplot(aes(x = emb1, y = emb2)) +
-  geom_point(alpha=0.5, colour=point_colours[1], size = 0.5) +
-  theme_linedraw() +
-  theme(
-    #aspect.ratio = 1,
-    plot.background = element_rect(fill = 'transparent', colour = NA),
-    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-    panel.background = element_rect(fill = 'transparent',
-                                    colour = NA),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.title.x = element_blank(), axis.title.y = element_blank(),
-    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-    axis.text.y = element_blank(), axis.ticks.y = element_blank()
-  )
-
-nldr_plt <- ggplotly(nldr_plt, width = 300,
-                     height = 300, tooltip = "none") |>
-  style(unselected=list(marker=list(opacity=1))) |>
+nldr_plt <- plot_ly(
+    shared_df,
+    x = ~emb1, y = ~emb2,
+    type = "scatter",
+    mode = "markers",
+    marker = list(
+        color = point_colours[1],
+        size = 3,
+        opacity = 0.5
+    ),
+    hoverinfo = "none"
+) |>
+    layout(
+        width = 300,
+        height = 300,
+        xaxis = list(
+            title = "",
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            linecolor = "black",
+            mirror = TRUE
+        ),
+        yaxis = list(
+            title = "",
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            linecolor = "black",
+            mirror = TRUE
+        ),
+        margin = list(l = 20, r = 20, t = 20, b = 20),
+        dragmode = "select"
+    ) |> 
+    style(selected   = list(marker = list(opacity = 1)), unselected=list(marker=list(opacity=1))) |>
     highlight(on="plotly_selected", off="plotly_deselect") |>
     config(displayModeBar = FALSE)
-
 ```

### 2025-12-08 paper-quollr.Rmd
Fixed Figure 10

@@
 
-error_plt <- shared_df |>
-  ggplot(aes(x=sqrt_row_wise_total_error, y = density)) +
-  geom_point(colour = point_colours[1]) +
-  xlab(expression(e[hj])) +
-  ylab("") +
-  theme_bw() +
-  theme(
-    #aspect.ratio = 1,
-    plot.background = element_rect(fill = 'transparent', colour = NA),
-    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-    panel.background = element_rect(fill = 'transparent',
-                                    colour = NA),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.title.x = element_blank(), axis.title.y = element_blank(),
-    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-    axis.text.y = element_blank(), axis.ticks.y = element_blank()
-  )
-
-error_plt <- ggplotly(error_plt, width = 200,
-                      height = 200, tooltip = "none") |>
-  style(unselected=list(marker=list(opacity=1))) |>
+error_plt <- plot_ly(
+    shared_df,
+    x = ~sqrt_row_wise_total_error, y = ~density,
+    type = "scatter",
+    mode = "markers",
+    marker = list(
+        color = point_colours[1],
+        size = 3,
+        opacity = 0.5
+    ),
+    hoverinfo = "none"
+) |>
+    layout(
+        width = 250,
+        height = 250,
+        xaxis = list(
+            title = "",
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            linecolor = "black",
+            mirror = TRUE
+        ),
+        yaxis = list(
+            title = "",
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            linecolor = "black",
+            mirror = TRUE
+        ),
+        margin = list(l = 20, r = 20, t = 20, b = 20),
+        dragmode = "select"
+    ) |> 
+    style(selected   = list(marker = list(opacity = 1)), unselected=list(marker=list(opacity=1))) |>
     highlight(on="plotly_selected", off="plotly_deselect") |>
@@
 
-nldr_plt <- shared_df |>
-  ggplot(aes(x = emb1, y = emb2)) +
-  geom_point(alpha=0.5, colour=point_colours[1], size = 0.5) +
-  theme_linedraw() +
-  theme(
-    #aspect.ratio = 1,
-    plot.background = element_rect(fill = 'transparent', colour = NA),
-    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-    panel.background = element_rect(fill = 'transparent',
-                                    colour = NA),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.title.x = element_blank(), axis.title.y = element_blank(),
-    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-    axis.text.y = element_blank(), axis.ticks.y = element_blank()
-  )
-
-nldr_plt <- ggplotly(nldr_plt, width = 200,
-                     height = 200, tooltip = "none") |>
-  style(unselected=list(marker=list(opacity=1))) |>
+nldr_plt <- plot_ly(
+    shared_df,
+    x = ~emb1, y = ~emb2,
+    type = "scatter",
+    mode = "markers",
+    marker = list(
+        color = point_colours[1],
+        size = 3,
+        opacity = 0.5
+    ),
+    hoverinfo = "none"
+) |>
+    layout(
+        width = 250,
+        height = 250,
+        xaxis = list(
+            title = "",
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            linecolor = "black",
+            mirror = TRUE
+        ),
+        yaxis = list(
+            title = "",
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            linecolor = "black",
+            mirror = TRUE
+        ),
+        margin = list(l = 20, r = 20, t = 20, b = 20),
+        dragmode = "select"
+    ) |> 
+    style(selected   = list(marker = list(opacity = 1)), unselected=list(marker=list(opacity=1))) |>
     highlight(on="plotly_selected", off="plotly_deselect") |>
@@
                     max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 1, alpha = 0.8, 
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.8, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
                palette = c("#66B2CC", "#FF7755"),
-                width = "200px", height = "200px")
+                width = "250px", height = "250px")
 
@@
     margin: 0;
-    padding: 0;",
+    padding: 0;
+    height: 300px;
+    width: 400px;",
         error_plt, 

### 2025-12-09 paper-quollr.Rmd
Fixed Figure 12 and 13

@@
 
-nldr_plt <- shared_df |>
-  ggplot(aes(x = emb1, y = emb2, colour = cluster)) +
-  geom_point(alpha=0.5, size = 0.8) +
-  geom_segment(data = trimesh_data_limb, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 0.5) +
-  scale_color_manual(values = c("0" = '#66c2a5',"1" = '#fc8d62',"2" = '#8da0cb',"3" = '#e78ac3',"4" = '#a6d854',"5" = '#ffd92f',"6" = '#e5c494')) +
-  theme_linedraw() +
-  theme(
-    #aspect.ratio = 1,
-    plot.background = element_rect(fill = 'transparent', colour = NA),
-    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-    panel.background = element_rect(fill = 'transparent',
-                                    colour = NA),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.title.x = element_blank(), axis.title.y = element_blank(),
-    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-    legend.position = "none"
-  )
-
-nldr_plt <- ggplotly(nldr_plt, width = 250,
-                     height = 250, tooltip = "none") |>
-  style(unselected=list(marker=list(opacity=1))) |>
-  highlight(on="plotly_selected", off="plotly_deselect") |>
+cluster_colors <- c(
+    "0" = '#66c2a5', "1" = '#fc8d62', "2" = '#8da0cb',
+    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#e5c494'
+)
+
+# Scatter points
+nldr_plt <- plot_ly(
+    data = shared_df,
+    x = ~emb1, y = ~emb2,
+    type = "scatter", mode = "markers",
+    color = ~cluster, colors = cluster_colors,
+    marker = list(size = 3, opacity = 0.5),
+    hoverinfo = "skip",  # <-- skip hover completely
+    showlegend = FALSE
+)
+
+# Add line segments
+if (nrow(trimesh_data_limb) > 0) {
+    nldr_plt <- nldr_plt |>
+        add_segments(
+            data = trimesh_data_limb,
+            x = ~x_from, y = ~y_from,
+            xend = ~x_to, yend = ~y_to,
+            line = list(color = "#000000", width = 0.5),
+            inherit = FALSE,
+            showlegend = FALSE,
+            hoverinfo = "skip"  # <-- skip hover
+        )
+}
+
+# Layout: axes, border, no tooltip
+nldr_plt <- nldr_plt |>
+    layout(
+        width = 310, height = 310,
+        margin = list(l = 20, r = 20, t = 20, b = 20),
+        xaxis = list(title = "", showgrid = FALSE, zeroline = FALSE,
+                     showticklabels = FALSE, ticks = "", linecolor = "black", mirror = TRUE),
+        yaxis = list(title = "", showgrid = FALSE, zeroline = FALSE,
+                     showticklabels = FALSE, ticks = "", linecolor = "black", mirror = TRUE)
+    ) |>
+    style(
+        selected = list(marker = list(opacity = 1)),
+        unselected = list(marker = list(opacity = 1))
+    ) |>
+    highlight(
+        on = "plotly_selected",
+        off = "plotly_deselect"
+    ) |>
     config(displayModeBar = FALSE)
@@
                                                linkFilter=FALSE,
-                                               width = "350px", height = "250px")
+                                               width = "370px", height = "370px")
 
@@
     gap: 0px;
-    align-items: center;
+    align-items: start;
     justify-items: center;
     margin: 0;
-    padding: 0;",
+    padding: 0;
+    height: 380px;
+    width: 500px",
         nldr_plt, 
-        htmltools::div(style = "margin-top: 30px;", langevitour_output)
+        htmltools::div(style = "margin-top: 20px;", langevitour_output)
     ),
@@
     gap: 0px;
-    align-items: center;
+    align-items: start;
     justify-items: center;
     margin: 0;
-    padding: 0;",
+    padding: 0;
+    height: 380px;
+    width: 500px",
         nldr_plt, 

### 2025-12-09 paper-quollr.R

@@ -1765,38 +1765,55 @@ num_highd_col <- point_data |>
 
 shared_df <- crosstalk::SharedData$new(point_data)
 
-nldr_plt <- shared_df |>
-  ggplot(aes(x = emb1, y = emb2, colour = cluster)) +
-  geom_point(alpha=0.5, size = 0.8) +
-  geom_segment(data = trimesh_data_limb, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 0.5) +
-  scale_color_manual(values = c("0" = '#66c2a5',"1" = '#fc8d62',"2" = '#8da0cb',"3" = '#e78ac3',"4" = '#a6d854',"5" = '#ffd92f',"6" = '#e5c494')) +
-  theme_linedraw() +
-  theme(
-    #aspect.ratio = 1,
-    plot.background = element_rect(fill = 'transparent', colour = NA),
-    plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-    panel.background = element_rect(fill = 'transparent',
-                                    colour = NA),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.title.x = element_blank(), axis.title.y = element_blank(),
-    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-    legend.position = "none"
-  )
+cluster_colors <- c(
+    "0" = '#66c2a5', "1" = '#fc8d62', "2" = '#8da0cb',
+    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#e5c494'
+)
 
-nldr_plt <- ggplotly(nldr_plt, width = 250,
-                     height = 250, tooltip = "none") |>
-  style(unselected=list(marker=list(opacity=1))) |>
-  highlight(on="plotly_selected", off="plotly_deselect") |>
-  config(displayModeBar = FALSE)
+# Scatter points
+nldr_plt <- plot_ly(
+    data = shared_df,
+    x = ~emb1, y = ~emb2,
+    type = "scatter", mode = "markers",
+    color = ~cluster, colors = cluster_colors,
+    marker = list(size = 3, opacity = 0.5),
+    hoverinfo = "skip",  # <-- skip hover completely
[…]

### 2025-12-09 paper-quollr.Rmd
Updated Figure 9

@@
                     max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 1, alpha = 0.8, 
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.8, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
@@
     margin: 0;
-    padding: 0;",
+    padding: 0;
+    height: 350px;
+    width: 300px;",
         nldr_plt, htmltools::div(style = "margin-top: 20px;", detourr_output)

### 2025-12-09 paper-quollr.R

@@ -1383,7 +1383,7 @@ detourr_output <- detour(
 ) |>
   tour_path(grand_tour(2), 
                     max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 1, alpha = 0.8, 
+  show_scatter(axes = TRUE, size = 0.5, alpha = 0.8, 
                edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
                palette = c("#66B2CC", "#FF7755"),
                 width = "300px", height = "300px")
@@ -1399,7 +1399,9 @@ detourr_output <- detour(
 #     align-items: start;
 #     justify-items: center;
 #     margin: 0;
-#     padding: 0;",
+#     padding: 0;
+#     height: 350px;
+#     width: 300px;",
 #         nldr_plt, htmltools::div(style = "margin-top: 20px;", detourr_output)
 #     ),
 #     device = "xs"

### 2025-12-09 paper-quollr.Rmd
Added script paths

@@
 
+<!--scripts/FACs/02_gen_embeddings.R-->
 ```{r}
@@
 
+<!--scripts/FACs/03_gen_mse_for_diff_methods.R-->
 ```{r}
@@
 
+<!--scripts/FACs/01_data_preprocessing.R-->
 ```{r data-limb}

### 2025-12-10 paper-quollr.Rmd
Updated link plots

@@
 scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = tr_from_to_df_scurve$from_reindexed,
-                         lineTo = tr_from_to_df_scurve$to_reindexed,
+                         lineFrom = tr_from_to_df_scurve$from,
+                         lineTo = tr_from_to_df_scurve$to,
                          group = df_exe$type,
@@
 scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = tr_from_to_df_scurve$from_reindexed,
-                         lineTo = tr_from_to_df_scurve$to_reindexed,
+                         lineFrom = tr_from_to_df_scurve$from,
+                         lineTo = tr_from_to_df_scurve$to,
                          group = df_exe$type,
@@
 
+crosstalk::bscols(
+    htmltools::div(
+        style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;
+    height: 380px;
+    width: 500px",
         nldrdt_link
+    ),
+    device = "xs"
+)
 ```
@@
 langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
-                                               lineFrom = edge_data$from_reindexed,
-                                               lineTo = edge_data$to_reindexed,
+                                               lineFrom = edge_data$from,
+                                               lineTo = edge_data$to,
                                                group = factor(point_data$type,

### 2025-12-10 paper-quollr.Rmd
Updated figure

--- 2025-12-09 paper-quollr.Rmd
@@
       ",
-      htmltools::div(style = 'margin: 0; padding: 0;  height: 300px; width: 370px; text-align: left;', htmltools::h4("a"), hex_grid_poly_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 300px; text-align: left;', htmltools::h4("b"), hex_centroids_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 300px; width: 370px;  text-align: left;', htmltools::h4("c"), wireframe_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 350px; width: 300px; text-align: left;', htmltools::h4("d"), htmltools::div(
-    style = "margin-left: 90px;",
+      htmltools::div(style = 'margin: 0; padding: 0;  height: 450px; width: 200px; text-align: left;', htmltools::h5("a"), hex_grid_poly_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 450px; width: 200px; text-align: left;', htmltools::h5("b"), hex_centroids_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 200px;  text-align: left;', htmltools::h5("c"), wireframe_scurve_int),
+      htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 450px; text-align: left;', htmltools::div(
+    style = "margin-left: 120px;",
+    htmltools::h5("d"),
     scurve_umap_model_vis_n

### 2025-12-10 paper-quollr.Rmd
Corrected application figures

@@
 langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
-                                               lineFrom = edge_data$from,
-                                               lineTo = edge_data$to,
+                                               lineFrom = edge_data$from_reindexed,
+                                               lineTo = edge_data$to_reindexed,
                                                group = factor(point_data$type,

### 2025-12-10 paper-quollr.R

@@ -1833,8 +1833,8 @@ nldr_plt <- nldr_plt |>
 
 
 langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
-                                               lineFrom = edge_data$from,
-                                               lineTo = edge_data$to,
+                                               lineFrom = edge_data$from_reindexed,
+                                               lineTo = edge_data$to_reindexed,
                                                group = factor(point_data$type,
                                                               c("0", "1", "2", "3", "4", "5", "6", "model")),
                                                pointSize = append(rep(point_sizes[1], NROW(df_b)),
@@ -2045,8 +2045,8 @@ nldr_plt <- nldr_plt |>
 
 
 langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
-                                               lineFrom = edge_data$from,
-                                               lineTo = edge_data$to,
+                                               lineFrom = edge_data$from_reindexed,
+                                               lineTo = edge_data$to_reindexed,
                                                group = factor(point_data$type,
                                                               c("0", "1", "2", "3", "4", "5", "6", "model")),
                                                pointSize = append(rep(point_sizes[1], NROW(df_b)),

### 2025-12-10 paper-quollr.Rmd
Updated scurve codeblock in pdf figure titles

@@
 ```{r algo-steps, eval=knitr::is_latex_output(), out.width="80%", fig.height=5, fig.width=25, fig.pos="!ht", fig.align='center', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `scurve` data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
+The \\texttt{scurve} data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
 
@@
 
-```{r scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
-Each panel (a1–a4) shows the model (black) overlaid on `scurve` data (blue) in different projections.
-These views illustrate how the lifted wireframe model captures the structure of the `scurve` data.
+```{r scurve-highd-model-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.cap="$2\\text{-}D$ projections of the lifted high-dimensional wireframe model from the \\texttt{scurve} UMAP layout.
+Each panel (a1–a4) shows the model (black) overlaid on \\texttt{scurve} data (blue) in different projections.
+These views illustrate how the lifted wireframe model captures the structure of the \\texttt{scurve} data.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
@@
 
-```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and `scurve` structure in $7\\text{-}D$.
+```{r scurve-nldrlink-pdf, eval=knitr::is_latex_output(), fig.pos="!ht", fig.width=10, fig.height=10, fig.cap="Exploring the correspondence between UMAP layout and \\texttt{scurve} structure in $7\\text{-}D$.
 Two sets of plots are interactively linked: UMAP layout (a1, b1) and projection of $7\\text{-}D$ model and data (a2, b2).
 The purple points indicate the selected subsets, which differ between rows.
-In (a1), the lower bridge of the `scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
-In (b1), a different region near the upper arm of the `scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection.
+In (a1), the lower bridge of the \\texttt{scurve} is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
+In (b1), a different region near the upper arm of the \\texttt{scurve} is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection.
 While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP.", fig.alt = "Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
@@
 
-```{r scurve-linkerror-pdf, eval=knitr::is_latex_output(), fig.pos="H", fig.width=15, fig.height=10, fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ `scurve` model.
+```{r scurve-linkerror-pdf, eval=knitr::is_latex_output(), fig.pos="H", fig.width=15, fig.height=10, fig.cap="Exploring residuals in relation to UMAP layouts using a $7\\text{-}D$ \\texttt{scurve} model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\\text{-}D$ model with data (a3, b3).
 The purple points highlight selected subsets of the data, which differ across rows.
-In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `scurve` in the high-dimensional projection (a3).
-In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `scurve` in the projection (b3).
+In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the \\texttt{scurve} in the high-dimensional projection (a3).
+In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the \\texttt{scurve} in the projection (b3).
 This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.", fig.alt = "Six-panel figure showing linked views of residuals and UMAP layouts for the 7-D S-curve model.

### 2025-12-10 paper-quollr.tex
Updated link plots
Updated scurve codeblock in pdf figure titles

@@
 \caption{Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `scurve` data is shown.}\label{fig:algo-steps}
+The \texttt{scurve} data is shown.}\label{fig:algo-steps}
 \end{figure}
@@
 \includegraphics[width=1\linewidth,alt={Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
-Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
-Each panel (a1–a4) shows the model (black) overlaid on `scurve` data (blue) in different projections.
-These views illustrate how the lifted wireframe model captures the structure of the `scurve` data.
+Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout.
+Each panel (a1–a4) shows the model (black) overlaid on \texttt{scurve} data (blue) in different projections.
+These views illustrate how the lifted wireframe model captures the structure of the \texttt{scurve} data.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-pdf}
@@
 
+crosstalk::bscols(
+    htmltools::div(
+        style = "display: grid; grid-template-columns: 1fr 1fr;
+    gap: 0px;
+    align-items: start;
+    justify-items: center;
+    margin: 0;
+    padding: 0;
+    height: 380px;
+    width: 500px",
         nldrdt_link
+    ),
+    device = "xs"
+)
 \end{verbatim}
@@
 Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
-The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and `scurve` structure in $7\text{-}D$.
+The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and \texttt{scurve} structure in $7\text{-}D$.
 Two sets of plots are interactively linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
 The purple points indicate the selected subsets, which differ between rows.
-In (a1), the lower bridge of the `scurve` is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
-In (b1), a different region near the upper arm of the `scurve` is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
+In (a1), the lower bridge of the \texttt{scurve} is highlighted, which corresponds in (a2) to points spanning across both arms of the high-dimensional structure.
+In (b1), a different region near the upper arm of the \texttt{scurve} is selected, and in (b2) these points map onto one side of the curved manifold in $7\text{-}D$ projection.
 While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\text{-}D$ space, highlighting distortions introduced by UMAP.}\label{fig:scurve-nldrlink-pdf}
@@
 Panels (b1–b3) highlight low-residual points, mapping to denser regions in the UMAP and thicker bands in the high-dimensional projection.
-The figure illustrates how residuals reveal distortions in UMAP embeddings.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ `scurve` model.
+The figure illustrates how residuals reveal distortions in UMAP embeddings.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ \texttt{scurve} model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
 The purple points highlight selected subsets of the data, which differ across rows.
-In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the `scurve` in the high-dimensional projection (a3).
-In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the `scurve` in the projection (b3).
+In the top row (a1–a3), points with higher residuals (a1) are selected, corresponding to the sparse bridging region in the UMAP layout (a2) and the less dense end of the \texttt{scurve} in the high-dimensional projection (a3).
+In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the \texttt{scurve} in the projection (b3).
 This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.}\label{fig:scurve-linkerror-pdf}

### 2025-12-10 paper-quollr.Rmd
Added detourr to acknowledgment

@@
 
-The source code for reproducing this paper can be found at: [https://github.com/JayaniLakshika/paper-quollr](https://github.com/JayaniLakshika/paper-quollr).
+The source code for reproducing this paper can be found at: [github.com/JayaniLakshika/paper-quollr](https://github.com/JayaniLakshika/paper-quollr).
 This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template.
-These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `htmltools` [@joe2024], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].
+These `R` packages were used for this work: `cli` [@gabor2025], `dplyr` [@hadley2023], `ggplot2` [@hadley2016], `interp` (>= 1.1-6) [@albrecht2024], `langevitour` [@paul2023], `detourr` [@casper2025], `proxy`[@david2022], `stats` [@core2025], `tibble` [@kirill2023], `tidyselect` [@lionel2024], `crosstalk` [@joe2023], `plotly` [@chapman2020], `htmltools` [@joe2024], `kableExtra` [@hao2024], `patchwork` [@thomas2024], and `readr` [@hadley2024].

### 2025-12-10 paper-quollr.tex
Added detourr to acknowledgment

@@
 
-The source code for reproducing this paper can be found at: \url{https://github.com/JayaniLakshika/paper-quollr}.
+The source code for reproducing this paper can be found at: \href{https://github.com/JayaniLakshika/paper-quollr}{github.com/JayaniLakshika/paper-quollr}.
 This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.
-These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{htmltools} \citep{joe2024}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.
+These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{dplyr} \citep{hadley2023}, \texttt{ggplot2} \citep{hadley2016}, \texttt{interp} (\textgreater= 1.1-6) \citep{albrecht2024}, \texttt{langevitour} \citep{paul2023}, \texttt{detourr} \citep{casper2025}, \texttt{proxy}\citep{david2022}, \texttt{stats} \citep{core2025}, \texttt{tibble} \citep{kirill2023}, \texttt{tidyselect} \citep{lionel2024}, \texttt{crosstalk} \citep{joe2023}, \texttt{plotly} \citep{chapman2020}, \texttt{htmltools} \citep{joe2024}, \texttt{kableExtra} \citep{hao2024}, \texttt{patchwork} \citep{thomas2024}, and \texttt{readr} \citep{hadley2024}.
 

### 2025-12-10 paper-quollr.Rmd
Corrected html reference issue

@@
 
-The package offers several $2\text{-}D$ visualizations (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:geom-outputs)"} else { "Figure \\@ref(fig:geom-outputs-pdf)"})`), including:
-
-- A full hexagonal grid,
-- A hexagonal grid that matches the data,
-- A full grid based on centroid triangulation,
-- A centroid triangulation grid that aligns with the data, 
-- A triangular mesh for any provided set of points.
+The package offers several $2\text{-}D$ visualizations (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:geom-outputs)"} else { "Figure \\@ref(fig:geom-outputs-pdf)"})`), including a full hexagonal grid, a hexagonal grid that matches the data, a full grid based on centroid triangulation, a centroid triangulation grid that aligns with the data, and a triangular mesh for any provided set of points.
 
@@
 
-class(nldrdt_link) <- c(class(nldrdt_link), "htmlwidget")
-
-crosstalk::bscols(
+nldrdt_link <- crosstalk::bscols(
     htmltools::div(
@@
 )
+
+class(nldrdt_link) <- c(class(nldrdt_link), "htmlwidget")
+
+nldrdt_link
 ```

### 2025-12-10 paper-quollr.tex
Corrected html reference issue

@@
 
-The package offers several \(2\text{-}D\) visualizations (Figure \ref{fig:geom-outputs-pdf}), including:
-
-\begin{itemize}
-\tightlist
-\item
-  A full hexagonal grid,
-\item
-  A hexagonal grid that matches the data,
-\item
-  A full grid based on centroid triangulation,
-\item
-  A centroid triangulation grid that aligns with the data,
-\item
-  A triangular mesh for any provided set of points.
-\end{itemize}
+The package offers several \(2\text{-}D\) visualizations (Figure \ref{fig:geom-outputs-pdf}), including a full hexagonal grid, a hexagonal grid that matches the data, a full grid based on centroid triangulation, a centroid triangulation grid that aligns with the data, and a triangular mesh for any provided set of points.
 
@@
 )
+
+class(nldrdt_link) <- c(class(nldrdt_link), "htmlwidget")
+
+nldrdt_link
 \end{verbatim}

### 2025-12-10 paper-quollr.Rmd
Updated quollr github link
Updated link plot section to reduce the number of pages
Removed unnecessary code

@@
 
-The package is available on CRAN, and the development version is available at https://jayanilakshika.github.io/quollr/.
+The package is available on CRAN, and the development version is available at [github.com/JayaniLakshika/quollr](https://github.com/JayaniLakshika/quollr).
 
@@
 
-<!--add each step separately-->
-<!--add hexbin notation image-->
-
-```{r, eval=FALSE}
-#| label: code-illustration
-# Code to draw illustration for notation
-## hexagon binning to have regular hexagons
-hb_obj_notation <- hex_binning(
-  nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 9, 
-  q = 0.1)
-
-a1_temp <- hb_obj_notation$a1
-a2_temp <- hb_obj_notation$a2
-l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
-
-## Data set with all centroids
-all_centroids_df_temp <- hb_obj_notation$centroids
-hex_grid_temp <- hb_obj_notation$hex_poly
-
-hex_grid_temp45 <- hex_grid_temp |> 
-  filter(h == 45)
-
-start_pt <- all_centroids_df_temp |> 
-  filter(h == 1)
-d_rect <- tibble(x1min = 0, 
-                 x1max = 1,
-                 x2min = 0,
-                 x2max = diff(scurve_umap_obj$lim2)/diff(scurve_umap_obj$lim1)) # x2max = r2
-
-# To move the rectangle to ignore the overlap with the centroids
-# rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
-rect_adj <- tibble(x1 = -0.03, x2 = 0.03)
-
-a1 <- tibble(x = all_centroids_df_temp$c_x[4],
-             xend = all_centroids_df_temp$c_x[5],
-             y = all_centroids_df_temp$c_y[21],
-             yend = all_centroids_df_temp$c_y[21],
-             label = expression(a[1]))
-a2 <- tibble(x = all_centroids_df_temp$c_x[25],
-             xend = all_centroids_df_temp$c_x[25],
-             y = all_centroids_df_temp$c_y[25],
-             yend = all_centroids_df_temp$c_y[33],
-             label = expression(a[2]))
-l <- tibble(x = hex_grid_temp45$x[2],
-            xend = hex_grid_temp45$x[3],
-            y = hex_grid_temp45$y[2],
-            yend = hex_grid_temp45$y[3],
-            label = expression(l))
-
-hex_param_vis <- ggplot() + 
-    geom_polygon(data = hex_grid_temp, 
-                        aes(x = x, 
-                            y = y, 
-                            group = h),
-                 fill = "white", 
-                 color = "#bdbdbd") +
-    geom_point(data = all_centroids_df_temp, aes(
-      x = c_x, 
-      y = c_y), 
-      color = "#31a354", size = 0.9) +
-    geom_point(data = start_pt, aes(x = c_x, 
-                                    y = c_y), 
-               color = "black") + 
-    geom_rect(data=d_rect, 
-              aes(xmin = x1min - rect_adj$x1,# - rect_adj$s1, 
-                  xmax = x1max - rect_adj$x1,# - rect_adj$s1, 
-                  ymin = x2min - rect_adj$x2,# - rect_adj$s2, 
-                  ymax = x2max - rect_adj$x2),# - rect_adj$s2), 
-              fill = "white", 
-              color = "black", 
-              alpha = 0, 
-              linewidth = 0.7) +
-    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
-                                y=x2min - rect_adj$x2)) + 
-    geom_point(data=d_rect, aes(x=x1max - rect_adj$x1, 
-                                y=x2min - rect_adj$x2)) + 
-    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
-                                y=x2max - rect_adj$x2)) + 
-    annotate("text", x=d_rect$x1min - rect_adj$x1, 
-                     y=d_rect$x2min - rect_adj$x2,
-                     label = "(0,0)", 
-             hjust=-0.1, vjust=-0.3) + 
-    annotate("text", x=d_rect$x1max - rect_adj$x1, 
-                     y=d_rect$x2min - rect_adj$x2,
-                     label = "(0,1)", 
-             hjust=1.1, vjust=-0.3) + 
-    annotate("text", x=d_rect$x1min - rect_adj$x1,
-                     y=d_rect$x2max - rect_adj$x2,
-                     label = expression(group("(",
-                        list(0, y[2][max]),")")),
-            hjust=-0.1, vjust=1.2) +
-    geom_segment(data=d_rect, aes(
-      x = x1min  - rect_adj$x1, # 0 - 0.03, 
-      y = -0.31, 
-      xend = x1max - rect_adj$x1, #1 - 0.03, 
-      yend = -0.31), #-0.35),
-      arrow = arrow(length = unit(0.03, "npc"),
-                               ends = "both"), 
-                 color = "black")+
-    annotate("text", x=0.5, y=-0.36, 
-             label = expression(r[1]), color = "black") +
-    geom_segment(data=d_rect, aes(
-      x = -0.25, 
-      y = x2min - rect_adj$x2, #0 - 0.05, 
-      xend = -0.25, 
-      yend = x2max - rect_adj$x2), #r2 - 0.05),
-      arrow = arrow(length = unit(0.03, "npc"),
-                       ends = "both"), 
-                 color = "black")+ 
-    annotate("text", x=-0.3, y=0.4, 
-             label = expression(r[2]), color = "black") +
-    geom_segment(data = a1, aes(
-      x = x, #-0.1 + 0.2087578, 
-      y = y, #-0.15, 
-      xend = xend, #-0.1 + 0.2087578*2, 
-      yend = yend), #-0.15),
-      arrow = arrow(length = unit(0.03, "npc"),
-        ends = "both"), 
-        color = "black")+ # a1 = 0.2087578
-    annotate("text", 
-             x=(a1$x+a1$xend)/2, 
-             y=a1$y, 
-             label = expression(a[1]), 
-             color = "black",
-             vjust = 1.2) +
-    geom_segment(data = a2, aes(
-      x = x, #-0.15, 
-      y = y, #-0.1*r2 + 0.1807896*2, 
-      xend = xend, #-0.15, 
-      yend = yend), #-0.1*r2 + 0.1807896*3),
-      arrow = arrow(length = unit(0.03, "npc"),
-                               ends = "both"), 
-      color = "black") + # a2 = 0.1807896
-    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
-             label = expression(a[2]), 
-             color = "black", hjust=-0.2) +
-    annotate("text", x=-0.18, y=-0.24, 
-      label = expression(group("(", list(s[1], s[2]), ")")),
-      color = "black") +
-  geom_segment(data = l, aes(
-      x = x, #-0.15, 
-      y = y, #-0.1*r2 + 0.1807896*2, 
-      xend = xend, #-0.15, 
-      yend = yend), #-0.1*r2 + 0.1807896*3),
-      arrow = arrow(length = unit(0.03, "npc"),
-                               ends = "both"), 
-      color = "black") + 
-    annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
-             label = expression(l), 
-             color = "black", hjust=-0.2) +
-  coord_equal()
-```
-
-```{r hex-param, fig.cap="The components of the hexagon grid illustrating notation.", out.width="50%", fig.align='center', fig.pos="!ht", fig.alt="The components of the hexagon grid illustrating notation.", eval=FALSE}
-
-hex_param_vis
-```
-
 ### Generating all possible centroids in a hexagonal grid
@@
         style = "display: grid; grid-template-columns: 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;
-    height: 380px;
-    width: 500px",
+    gap: 0px; align-items: start; justify-items: center; margin: 0; padding: 0;
+    height: 380px; width: 500px",
         nldrdt_link
@@
         style = "display: grid; grid-template-columns: 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;
-    height: 350px;
-    width: 300px;",
+    gap: 0px; align-items: start; justify-items: center; margin: 0; padding: 0;
+    height: 350px; width: 300px;",
         nldr_plt, htmltools::div(style = "margin-top: 20px;", detourr_output)
@@
         style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
-    gap: 0px;
-    align-items: start;
-    justify-items: center;
-    margin: 0;
-    padding: 0;
-    height: 300px;
-    width: 400px;",
-        error_plt, 
-        nldr_plt, 
+    gap: 0px; align-items: start; justify-items: center; margin: 0; padding: 0;
+    height: 300px; width: 400px;",
+        error_plt, nldr_plt, 
     htmltools::div(style = "margin-top: 20px;", detourr_output)

### 2025-12-10 paper-quollr.R

@@ -560,163 +560,6 @@ hb_obj <- hex_binning(
   q = 0.1)
 
 
-## ----eval=FALSE---------------------------------------------------------------
-#| label: code-illustration
-# # Code to draw illustration for notation
-# ## hexagon binning to have regular hexagons
-# hb_obj_notation <- hex_binning(
-#   nldr_scaled_obj = scurve_umap_obj,
-#   b1 = 9,
-#   q = 0.1)
-# 
-# a1_temp <- hb_obj_notation$a1
-# a2_temp <- hb_obj_notation$a2
-# l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
-# 
-# ## Data set with all centroids
-# all_centroids_df_temp <- hb_obj_notation$centroids
-# hex_grid_temp <- hb_obj_notation$hex_poly
-# 
-# hex_grid_temp45 <- hex_grid_temp |>
-#   filter(h == 45)
-# 
-# start_pt <- all_centroids_df_temp |>
-#   filter(h == 1)
-# d_rect <- tibble(x1min = 0,
-#                  x1max = 1,
-#                  x2min = 0,
-#                  x2max = diff(scurve_umap_obj$lim2)/diff(scurve_umap_obj$lim1)) # x2max = r2
-# 
-# # To move the rectangle to ignore the overlap with the centroids
-# # rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
-# rect_adj <- tibble(x1 = -0.03, x2 = 0.03)
-# 
-# 
-# a1 <- tibble(x = all_centroids_df_temp$c_x[4],
-#              xend = all_centroids_df_temp$c_x[5],
-#              y = all_centroids_df_temp$c_y[21],
-#              yend = all_centroids_df_temp$c_y[21],
-#              label = expression(a[1]))
-# a2 <- tibble(x = all_centroids_df_temp$c_x[25],
-#              xend = all_centroids_df_temp$c_x[25],
-#              y = all_centroids_df_temp$c_y[25],
-#              yend = all_centroids_df_temp$c_y[33],
-#              label = expression(a[2]))
-# l <- tibble(x = hex_grid_temp45$x[2],
-#             xend = hex_grid_temp45$x[3],
-#             y = hex_grid_temp45$y[2],
-#             yend = hex_grid_temp45$y[3],
-#             label = expression(l))
-# 
-# hex_param_vis <- ggplot() +
[…]

### 2025-12-10 paper-quollr.Rmd
Updated hexbinning section

@@
 
-The function `calc_bins_y()` determines the configuration of the hexagon grid by computing the number of bins along the y-axis (`b2`), the hexagon binwidth (`a1`), and hexagon binheight (`a2`).
-This function accepts (i) an object (`nldr_scaled_obj`) containing three elements: the first is the scaled NLDR data (`scaled_nldr`), and the second and third are the limits of the original NLDR data (`lim1` and `lim2`); (ii) the number of bins along the x-axis (`b1`), and (iii) the buffer amount as a proportion (`q`).
-The buffer ensures that the grid fully covers the data space by extending one hexagon width ($a_1$) and height ($a_2$) beyond the observed data in all directions.
-By default, $q = 0.1$, but it must be set to a value smaller than the minimum data value to avoid exceeding the data range.
+The hexagon binning process builds a complete hexagonal tessellation over the $2\text{-}D$ NLDR embedding and assigns points to bins for downstream modeling.
+The workflow proceeds through several steps, beginning with selecting an appropriate grid configurations and ending with assigning NLDR points to their corresponding hexagons.
+
+The first step is to determine the configuration of the hexagonal grid.
+The function `calc_bins_y()` computes the number of bins along the y-axis (`b2`), the hexagon width (`a1`), and the hexagon height (`a2`), based on three inputs: (i) the scaled NLDR object (`nldr_scaled_obj`, containing the scaled embedding and its $2\text{-}D$ limits), (ii) the chosen number of bins along the x-axis (`b1`), and (iii) a buffer proportion (`q`) that expands the grid slightly beyond the observed data range.
+This buffer ensures that the tessellation fully covers the embedding.
+By default, `q = 0.1`, but in practice it must be chosen to avoid exceeding the domain of the scaled NLDR data.
 
@@
 
-Points are allocated to bins based on the nearest centroid of the hexagonal bins.
-The hexagonal binning algorithm can be executed using the `hex_binning()` function, or its individual steps can be run separately.
-While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the `hex_binning()` function automates this entire workflow.
-The parameters used within `hex_binning()` are the object output from `gen_scaled_data` (`nldr_scaled_obj`); the number of bins along the x-axis (`b1`), and the buffer amount as a proportion of the data (`q`).
-The output is an object of the `hex_bin_obj` class, which contains the bin widths in each direction (`a1`, `a2`), the number of bins in each direction (`bins`), the coordinates of the hexagonal grid starting point (`start_point`), the details of bin centroids (`centroids`), the coordinates of bins (`hex_poly`), NLDR components with their corresponding hexagons (`data_hb_id`), hex bins with their corresponding standardized counts (`std_cts`), the total number of bins (`b`), the number of non-empty bins (`m`), and the points within each hexagon (`pts_bins`).
+Once the grid configurations are known, the next step is to construct the full set of hexagon centroids.
+The function `gen_centroids()` uses the bin configuration and the embedding limits to determine the horizontal and vertical spacing of rows, including the staggered row offsets required for hexagonal tiling.
+Odd rows place centroids directly above one another, while even rows shift by half the hexagon width to create the interlocking structure.
+Vertical spacing is determined by $v_s = (\sqrt{3}/2) a_1$, with starting positions defined by the buffer-adjusted limits.
+This produces a complete grid of potential hexagons covering the $2\text{-}D$ space.
+
+```{r, echo=TRUE}
+all_centroids_df <- gen_centroids(
+  nldr_scaled_obj = scurve_umap_obj, 
+  b1 = 21, 
+  q = 0.1
+)
+
+head(all_centroids_df, 5)
+```
+
+After determining the centroid locations, each hexagon’s polygonal boundary is generated using the `gen_hex_coord()` function, which constructs the six vertices of every hexagon by applying fixed geometric offsets derived from the hexagon width, $a_1$.
+Specifically, the horizontal and vertical offsets are defined as $dx = a_1/2$, $dy = a_1/\sqrt{3}$, and $vf = a_1/(2\sqrt{3})$, and these constants are used to position each vertex relative to the centroid, producing the complete set of hexagon boundaries.
+
+These constants define how far each vertex lies from the centroid in the six characteristic directions.
+For efficiency, the vertex generation is implemented in C++ and returns a tibble listing all polygon vertices, uniquely indexed by hexagons.
+
+```{r, echo=TRUE}
+all_hex_coord <- gen_hex_coord(
+  centroids_data = all_centroids_df, 
+  a1 = bin_configs$a1
+)
+
+head(all_hex_coord, 5)
+```
+
+The next step is to assign each NLDR point to its closest hexagon.
+The `assign_data()` function computes all pairwise Euclidean distances between the $2\text{-}D$ NLDR embedding and the centroid coordinates, identifying the nearest centroid for every observation.
+Each point is assigned a hexagon index (`h`), producing a version of the embedding annotated with bin membership.
+
+```{r, echo=TRUE}
+umap_hex_id <- assign_data(
+  nldr_scaled_obj = scurve_umap_obj, 
+  centroids_data = all_centroids_df
+)
+
+head(umap_hex_id, 5)
+```
+
+Finally, the list of points within each hexagon can be extracted using `group_hex_pts()`.
+This step collapses the point-level assignments into a hexagon-level summary by grouping by `h` and collecting the IDs of all points belonging to that polygon.
+The result is a tidy representation of the binning structure.
+
+```{r, echo=TRUE}
+pts_df <- group_hex_pts(
+  scaled_nldr_hexid = umap_hex_id
+)
+
+head(pts_df, 5)
+```
+
+Although each component of the workflow can be run independently, the `hex_binning()` function automates the entire sequence—from grid construction to point assignment—and returns a `hex_bin_obj` containing the bin dimensions (`a1`, `a2`), the full grid geometry (`centroids`, `hex_poly`), bin membership (`data_hb_id`), standardized counts (`std_cts`), the total number of bins (`b`), non-empty bins (`m`), and the list of points per bin (`pts_bins`).
 
@@
   q = 0.1)
-```
-
-### Generating all possible centroids in a hexagonal grid
-
-The `gen_centroids()` function computes the centroids of a hexagonal grid.
-The function first validates the input parameters, ensuring that the number of bins along the x-axis (`b1`) is at least two and that the buffer width (`q`) lies within the recommended range of 0.05–0.2.
-
-The coordinate limits of the embedding (`lim1` and `lim2`) are used to calculate the aspect ratio between the two axes, which determines the vertical scaling of the grid.
-The helper function `calc_bins_y()` is then called to compute the number of bins along the y-axis (`b2`) and the hexagon binwidth (`a1`) given `b1` and `q`.
-The hexagon grid starting points are defined by `s1` and `s2`.
-These are set to $-q$ and $-q \times r2$, where `r2` is the is the aspect ratio between the two axes.
-
-Centroids are generated by alternating the x-coordinates of adjacent rows to create the hexagonal tiling.
-Odd-numbered rows use x-coordinates spaced by the hexagon width (`a1`), while even-numbered rows are staggered by half this distance (`a1/2`).
-The vertical spacing between rows is $v_s = (\sqrt{3}/2) a_1$.
-
-The y-coordinates are constructed using this spacing, and paired with x-coordinates depending on whether the total number of rows (`b2`) is even or odd.
-For an odd number of rows, the final row includes only the odd-row x-coordinates to preserve the alternating pattern.
-
-The output is a tibble containing a unique hexagon (`h`) and centroid coordinates (`c_x`, `c_y`), which together define the layout of the complete hexagonal grid over the $2\text{-}D$ space.
-
-```{r, echo=TRUE}
-all_centroids_df <- gen_centroids(
-  nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 21, 
-  q = 0.1
-  )
-
-head(all_centroids_df, 5)
-```
-
-### Creating the coordinates of the hexagons
-
-Following the generation of hexagonal centroids, the `gen_hex_coord()` function constructs the coordinates of each hexagonal bin by defining its six polygonal vertices.
-These coordinates are used to visualize the hexagonal tessellation.
-
-The function first verifies that the hexagon width (`a1`) is provided, as it determines the overall scale of the grid.
-Each hexagon is defined relative to its centroid $(C_x, C_y)$, with six vertices positioned symmetrically around the center.
-
-Three constants are derived to define the relative distances to the vertices:
-$dx = a_1/2, \quad dy = a_1/\sqrt{3}, \quad vf = a_1/(2\sqrt{3})$.
-
-Here, `dx` and `dy` define the horizontal and vertical offsets, while `vf` refines vertical positioning to maintain geometric alignment across rows.
-
-Using these constants, two offset vectors are defined for the six vertices relative to the centroid: $x_{\text{add}} = (0, -dx, -dx, 0, dx, dx), \quad y_{\text{add}} = (dy, vf, -vf, -dy, -vf, vf)$.
-
-These offsets correspond to the six directions around each centroid: top, top-left, bottom-left, bottom, bottom-right, and top-right.
-
-For each centroid, the coordinates of its six vertices are computed as $x_i = C_x + x_{\text{add},i}, \quad y_i = C_y + y_{\text{add},i}$, and assigned a polygon ID equal to the centroid identifier (`h`).
-The resulting coordinates are combined into a tibble containing the polygon ID (`h`) and vertex coordinates (`x`, `y`).
-
-To improve computational efficiency, the vertex generation is implemented in C++ via the helper function `gen_hex_coord_cpp()`.
-This routine iterates over all centroids, applies the vertex offsets, and returns a tibble that defines the complete set of hexagon geometries in $2\text{-}D$ space.
-
-```{r, echo=TRUE}
-all_hex_coord <- gen_hex_coord(
-  centroids_data = all_centroids_df, 
-  a1 = bin_configs$a1
-  )
-
-head(all_hex_coord, 5)
-```
-
-### Assigning data points to their respective hexagons
-
-After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The `assign_data()` function performs this assignment by calculating the $2\text{-}D$ Euclidean distance between each point in the $2\text{-}D$ embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the $2\text{-}D$ layout.
-It then selects the corresponding x and y coordinates of each hexagon’s centroid.
-
-Both the embedding coordinates and the centroid coordinates are converted to matrices to facilitate distance computations.
-The function uses the `proxy::dist()` method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
-For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon—and assigns the corresponding hexagon (`h`) to the point.
-
-The result is a `tibble` of the scaled $2\text{-}D$ embedding with an additional `h` column, indicating the hexagonal bin to which each point belongs.
-
-```{r, echo=TRUE}
-umap_hex_id <- assign_data(
-  nldr_scaled_obj = scurve_umap_obj, 
-  centroids_data = all_centroids_df
-  )
-
-head(umap_hex_id, 5)
-```
-
-The `group_hex_pts()` function extracts the list of data point identifiers (`ID`) assigned to each hexagon.
-
-The function first groups the input data by `h`, which represents the hexagon associated with each point in the $2\text{-}D$ layout.
-Within each group, it collects the `ID`s into a list, resulting in a summary where each row corresponds to a single hexagon.
-The resulting column, `pts_list`, contains all point identifiers associated with that hexagon.
-
-```{r, echo=TRUE}
-pts_df <- group_hex_pts(
-  scaled_nldr_hexid = umap_hex_id
-  )
-
-head(pts_df, 5)
 ```
@@
 
-```{r scurve-linkerror-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
+```{r scurve-linkerror-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the `scurve` data in $7\\text{-}D$ (right) using `detourr`.
 Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
@@
 
-```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \\texttt{scurve} data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 

### 2025-12-10 paper-quollr.tex
Updated quollr github link
Updated link plot section to reduce the number of pages
Updated hexbinning section

@@
 
-The package is available on CRAN, and the development version is available at \url{https://jayanilakshika.github.io/quollr/}.
+The package is available on CRAN, and the development version is available at \href{https://github.com/JayaniLakshika/quollr}{github.com/JayaniLakshika/quollr}.
 
@@
 
-The function \texttt{calc\_bins\_y()} determines the configuration of the hexagon grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon binwidth (\texttt{a1}), and hexagon binheight (\texttt{a2}).
-This function accepts (i) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (ii) the number of bins along the x-axis (\texttt{b1}), and (iii) the buffer amount as a proportion (\texttt{q}).
-The buffer ensures that the grid fully covers the data space by extending one hexagon width (\(a_1\)) and height (\(a_2\)) beyond the observed data in all directions.
-By default, \(q = 0.1\), but it must be set to a value smaller than the minimum data value to avoid exceeding the data range.
+The hexagon binning process builds a complete hexagonal tessellation over the \(2\text{-}D\) NLDR embedding and assigns points to bins for downstream modeling.
+The workflow proceeds through several steps, beginning with selecting an appropriate grid configurations and ending with assigning NLDR points to their corresponding hexagons.
+
+The first step is to determine the configuration of the hexagonal grid.
+The function \texttt{calc\_bins\_y()} computes the number of bins along the y-axis (\texttt{b2}), the hexagon width (\texttt{a1}), and the hexagon height (\texttt{a2}), based on three inputs: (i) the scaled NLDR object (\texttt{nldr\_scaled\_obj}, containing the scaled embedding and its \(2\text{-}D\) limits), (ii) the chosen number of bins along the x-axis (\texttt{b1}), and (iii) a buffer proportion (\texttt{q}) that expands the grid slightly beyond the observed data range.
+This buffer ensures that the tessellation fully covers the embedding.
+By default, \texttt{q\ =\ 0.1}, but in practice it must be chosen to avoid exceeding the domain of the scaled NLDR data.
 
@@
 
-Points are allocated to bins based on the nearest centroid of the hexagonal bins.
-The hexagonal binning algorithm can be executed using the \texttt{hex\_binning()} function, or its individual steps can be run separately.
-While running the process step by step would involve generating centroids, constructing hexagon coordinates, assigning points to bins, standardizing counts, and mapping the data back to hexagons, the \texttt{hex\_binning()} function automates this entire workflow.
-The parameters used within \texttt{hex\_binning()} are the object output from \texttt{gen\_scaled\_data} (\texttt{nldr\_scaled\_obj}); the number of bins along the x-axis (\texttt{b1}), and the buffer amount as a proportion of the data (\texttt{q}).
-The output is an object of the \texttt{hex\_bin\_obj} class, which contains the bin widths in each direction (\texttt{a1}, \texttt{a2}), the number of bins in each direction (\texttt{bins}), the coordinates of the hexagonal grid starting point (\texttt{start\_point}), the details of bin centroids (\texttt{centroids}), the coordinates of bins (\texttt{hex\_poly}), NLDR components with their corresponding hexagons (\texttt{data\_hb\_id}), hex bins with their corresponding standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{b}), the number of non-empty bins (\texttt{m}), and the points within each hexagon (\texttt{pts\_bins}).
-
-\begin{verbatim}
-hb_obj <- hex_binning(
-  nldr_scaled_obj = scurve_umap_obj, 
-  b1 = 21, 
-  q = 0.1)
-\end{verbatim}
-
-\subsubsection{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}
-
-The \texttt{gen\_centroids()} function computes the centroids of a hexagonal grid.
-The function first validates the input parameters, ensuring that the number of bins along the x-axis (\texttt{b1}) is at least two and that the buffer width (\texttt{q}) lies within the recommended range of 0.05--0.2.
-
-The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to calculate the aspect ratio between the two axes, which determines the vertical scaling of the grid.
-The helper function \texttt{calc\_bins\_y()} is then called to compute the number of bins along the y-axis (\texttt{b2}) and the hexagon binwidth (\texttt{a1}) given \texttt{b1} and \texttt{q}.
-The hexagon grid starting points are defined by \texttt{s1} and \texttt{s2}.
-These are set to \(-q\) and \(-q \times r2\), where \texttt{r2} is the is the aspect ratio between the two axes.
-
-Centroids are generated by alternating the x-coordinates of adjacent rows to create the hexagonal tiling.
-Odd-numbered rows use x-coordinates spaced by the hexagon width (\texttt{a1}), while even-numbered rows are staggered by half this distance (\texttt{a1/2}).
-The vertical spacing between rows is \(v_s = (\sqrt{3}/2) a_1\).
-
-The y-coordinates are constructed using this spacing, and paired with x-coordinates depending on whether the total number of rows (\texttt{b2}) is even or odd.
-For an odd number of rows, the final row includes only the odd-row x-coordinates to preserve the alternating pattern.
-
-The output is a tibble containing a unique hexagon (\texttt{h}) and centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which together define the layout of the complete hexagonal grid over the \(2\text{-}D\) space.
+Once the grid configurations are known, the next step is to construct the full set of hexagon centroids.
+The function \texttt{gen\_centroids()} uses the bin configuration and the embedding limits to determine the horizontal and vertical spacing of rows, including the staggered row offsets required for hexagonal tiling.
+Odd rows place centroids directly above one another, while even rows shift by half the hexagon width to create the interlocking structure.
+Vertical spacing is determined by \(v_s = (\sqrt{3}/2) a_1\), with starting positions defined by the buffer-adjusted limits.
+This produces a complete grid of potential hexagons covering the \(2\text{-}D\) space.
 
@@
 
-\subsubsection{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}
-
-Following the generation of hexagonal centroids, the \texttt{gen\_hex\_coord()} function constructs the coordinates of each hexagonal bin by defining its six polygonal vertices.
-These coordinates are used to visualize the hexagonal tessellation.
-
-The function first verifies that the hexagon width (\texttt{a1}) is provided, as it determines the overall scale of the grid.
-Each hexagon is defined relative to its centroid \((C_x, C_y)\), with six vertices positioned symmetrically around the center.
-
-Three constants are derived to define the relative distances to the vertices:
-\(dx = a_1/2, \quad dy = a_1/\sqrt{3}, \quad vf = a_1/(2\sqrt{3})\).
-
-Here, \texttt{dx} and \texttt{dy} define the horizontal and vertical offsets, while \texttt{vf} refines vertical positioning to maintain geometric alignment across rows.
-
-Using these constants, two offset vectors are defined for the six vertices relative to the centroid: \(x_{\text{add}} = (0, -dx, -dx, 0, dx, dx), \quad y_{\text{add}} = (dy, vf, -vf, -dy, -vf, vf)\).
-
-These offsets correspond to the six directions around each centroid: top, top-left, bottom-left, bottom, bottom-right, and top-right.
-
-For each centroid, the coordinates of its six vertices are computed as \(x_i = C_x + x_{\text{add},i}, \quad y_i = C_y + y_{\text{add},i}\), and assigned a polygon ID equal to the centroid identifier (\texttt{h}).
-The resulting coordinates are combined into a tibble containing the polygon ID (\texttt{h}) and vertex coordinates (\texttt{x}, \texttt{y}).
-
-To improve computational efficiency, the vertex generation is implemented in C++ via the helper function \texttt{gen\_hex\_coord\_cpp()}.
-This routine iterates over all centroids, applies the vertex offsets, and returns a tibble that defines the complete set of hexagon geometries in \(2\text{-}D\) space.
+After determining the centroid locations, each hexagon's polygonal boundary is generated using the \texttt{gen\_hex\_coord()} function, which constructs the six vertices of every hexagon by applying fixed geometric offsets derived from the hexagon width, \(a_1\).
+Specifically, the horizontal and vertical offsets are defined as \(dx = a_1/2\), \(dy = a_1/\sqrt{3}\), and \(vf = a_1/(2\sqrt{3})\), and these constants are used to position each vertex relative to the centroid, producing the complete set of hexagon boundaries.
+
+These constants define how far each vertex lies from the centroid in the six characteristic directions.
+For efficiency, the vertex generation is implemented in C++ and returns a tibble listing all polygon vertices, uniquely indexed by hexagons.
 
@@
 
-\subsubsection{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}
-
-After generating the centroids that define the hexagonal grid, the next step is to assign each point in the NLDR embedding to its nearest hexagonal bin.
-The \texttt{assign\_data()} function performs this assignment by calculating the \(2\text{-}D\) Euclidean distance between each point in the \(2\text{-}D\) embedding and all hexagon centroids.
-
-First, the function extracts the first two dimensions of the scaled NLDR embedding, which represent the \(2\text{-}D\) layout.
-It then selects the corresponding x and y coordinates of each hexagon's centroid.
-
-Both the embedding coordinates and the centroid coordinates are converted to matrices to facilitate distance computations.
-The function uses the \texttt{proxy::dist()} method to compute a pairwise Euclidean distance matrix between all NLDR points and all centroids.
-For each NLDR point, the function identifies the index of the centroid with the smallest distance representing the closest hexagon---and assigns the corresponding hexagon (\texttt{h}) to the point.
-
-The result is a \texttt{tibble} of the scaled \(2\text{-}D\) embedding with an additional \texttt{h} column, indicating the hexagonal bin to which each point belongs.
+The next step is to assign each NLDR point to its closest hexagon.
+The \texttt{assign\_data()} function computes all pairwise Euclidean distances between the \(2\text{-}D\) NLDR embedding and the centroid coordinates, identifying the nearest centroid for every observation.
+Each point is assigned a hexagon index (\texttt{h}), producing a version of the embedding annotated with bin membership.
 
@@
 
-The \texttt{group\_hex\_pts()} function extracts the list of data point identifiers (\texttt{ID}) assigned to each hexagon.
-
-The function first groups the input data by \texttt{h}, which represents the hexagon associated with each point in the \(2\text{-}D\) layout.
-Within each group, it collects the \texttt{ID}s into a list, resulting in a summary where each row corresponds to a single hexagon.
-The resulting column, \texttt{pts\_list}, contains all point identifiers associated with that hexagon.
+Finally, the list of points within each hexagon can be extracted using \texttt{group\_hex\_pts()}.
+This step collapses the point-level assignments into a hexagon-level summary by grouping by \texttt{h} and collecting the IDs of all points belonging to that polygon.
+The result is a tidy representation of the binning structure.
 
@@
 > 5    71 <int [9]>
+\end{verbatim}
+
+Although each component of the workflow can be run independently, the \texttt{hex\_binning()} function automates the entire sequence---from grid construction to point assignment---and returns a \texttt{hex\_bin\_obj} containing the bin dimensions (\texttt{a1}, \texttt{a2}), the full grid geometry (\texttt{centroids}, \texttt{hex\_poly}), bin membership (\texttt{data\_hb\_id}), standardized counts (\texttt{std\_cts}), the total number of bins (\texttt{b}), non-empty bins (\texttt{m}), and the list of points per bin (\texttt{pts\_bins}).
+
+\begin{verbatim}
+hb_obj <- hex_binning(
+  nldr_scaled_obj = scurve_umap_obj, 
+  b1 = 21, 
+  q = 0.1)
 \end{verbatim}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-linkerror-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \texttt{scurve} data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-linkerror-dtpdf}
 \end{figure}

### 2025-12-10 paper-quollr.Rmd
Added information regarding why trimesh index should be updated

@@
 The `update_trimesh_index()` function re-indexes the node IDs to ensure that edge identifiers are sequentially numbered and consistent with downstream analysis.
+This sequential ordering of edges is essential because software such as `langevitour` and `detourr` rely on one-to-one mapping between edges and their corresponding vertices to visualize the mesh.
 
@@
 
-## Identifying and removing low-density hexagons
+### Identifying and removing low-density hexagons
 

### 2025-12-10 paper-quollr.tex
Added information regarding why trimesh index should be updated

@@
 The \texttt{update\_trimesh\_index()} function re-indexes the node IDs to ensure that edge identifiers are sequentially numbered and consistent with downstream analysis.
+This sequential ordering of edges is essential because software such as \texttt{langevitour} and \texttt{detourr} rely on one-to-one mapping between edges and their corresponding vertices to visualize the mesh.
 
@@
 
-\subsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
+\subsubsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}
 

### 2025-12-10 paper-quollr.Rmd
Bit summarise the residual section

@@
 
-Hexbin error (HBE) are used as goodness of fit metrics for the model. These metrics can be computed using the  `glance()` function, which provides a tidy output for evaluation.
-
-The function requires both the fitted model object returned by `fit_highd_model()` and $p\text{-}D$ data to begin.
-The $p\text{-}D$ model output (`model_highd`) is first renamed to avoid naming conflicts during subsequent data joins.
-It then uses the `predict_emb()` function to assign each point in the $p\text{-}D$ dataset to a corresponding hexagon bin in the $2\text{-}D$ model, producing a prediction data frame that contains both the predicted bin assignment (`pred_h`) and the original observation `ID`.
-
-The function joins this prediction output with both the $p\text{-}D$ model and the $p\text{-}D$ data (to retrieve true coordinates).
-It then calculates squared differences between the original and predicted $p\text{-}D$ coordinates for each dimension, storing these as `error_square_x1`, `error_square_x2`, ..., up to the dimensionality of the data.
-
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the HBE which is the average of the total squared error per point.
-
-These metrics are returned in a tibble as `Error` (absolute error) and `HBE` (root mean squared error).
-The computation of total absolute error and HBE is performed in C++ for efficiency using the internal `compute_errors()` function.
+Hexbin error (HBE) serves as a goodness-of-fit metric for evaluating the high-dimensional to $2\text{-}D$ mapping.
+The `glance()` function computes these summary diagnostics by combining the fitted model returned by `fit_highd_model()` with the $p\text{-}D$ data.
+After renaming the model output to avoid join conflicts, `predict_emb()` is used to assign each observation to a hexagon in the $2\text{-}D$ embedding.
+The resulting bin assignments are joined with both the model output and the $p\text{-}D$ data, allowing squared and absolute differences between the true and modeled $p\text{-}D$ coordinates to be computed for every dimension.
+Total absolute error (`Error`) and hexbin error (`HBE`, the root mean squared error) are then obtained using an efficient C++ implementation (`compute_errors()`), and returned in a tidy tibble.
 
@@
 
-Furthermore, `augment()` requires both the fitted model object returned by `fit_highd_model()` and $p\text{-}D$ data to begin.
-It extends the fitted model by adding prediction results and error diagnostics to the original $p\text{-}D$ data.
-
-The function starts with the same process as is used in the `glance()` function to produce a predicted point in $p\text{-}D$ for each point in the $p\text{-}D$ dataset.
-
-Next, the function computes residuals between each original coordinate (`x1`, `x2`, ..., `xp`) and the corresponding modeled coordinate (`model_high_d_x1`, ..., `model_high_d_xp`) across all dimensions.
-It calculates both squared errors and absolute errors per dimension.
-These are used to compute two aggregate diagnostic measures per point.
-First, the `row_wise_total_error` which is the total squared error across all dimensions, and the `row_wise_abs_error` which is the total absolute error across all dimensions.
-
-The final output is a data frame that combines the original IDs, high-dimensional data, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
-The augmented dataset is always returned as a `tibble::tibble` with the same number of rows as the passed dataset.
+The `augment()` function provides point-level diagnostics by appending predictions and residuals to the original $p\text{-}D$ data.
+Using the same prediction process as `glance()`, it computes dimension-wise residuals, squared errors, and absolute errors, along with two aggregate measures per observation: the total squared error (`row_wise_total_error`) and the total absolute error (`row_wise_abs_error`).
+The final output is a tibble containing IDs (`ID`), $p\text{-}D$ data, predicted hexagons (`h`), modeled coordinates, and all residual diagnostics, with one row per observation.
 

### 2025-12-10 paper-quollr.tex
Bit summarise the residual section

@@
 
-Hexbin error (HBE) are used as goodness of fit metrics for the model. These metrics can be computed using the \texttt{glance()} function, which provides a tidy output for evaluation.
-
-The function requires both the fitted model object returned by \texttt{fit\_highd\_model()} and \(p\text{-}D\) data to begin.
-The \(p\text{-}D\) model output (\texttt{model\_highd}) is first renamed to avoid naming conflicts during subsequent data joins.
-It then uses the \texttt{predict\_emb()} function to assign each point in the \(p\text{-}D\) dataset to a corresponding hexagon bin in the \(2\text{-}D\) model, producing a prediction data frame that contains both the predicted bin assignment (\texttt{pred\_h}) and the original observation \texttt{ID}.
-
-The function joins this prediction output with both the \(p\text{-}D\) model and the \(p\text{-}D\) data (to retrieve true coordinates).
-It then calculates squared differences between the original and predicted \(p\text{-}D\) coordinates for each dimension, storing these as \texttt{error\_square\_x1}, \texttt{error\_square\_x2}, \ldots, up to the dimensionality of the data.
-
-From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the HBE which is the average of the total squared error per point.
-
-These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{HBE} (root mean squared error).
-The computation of total absolute error and HBE is performed in C++ for efficiency using the internal \texttt{compute\_errors()} function.
+Hexbin error (HBE) serves as a goodness-of-fit metric for evaluating the high-dimensional to \(2\text{-}D\) mapping.
+The \texttt{glance()} function computes these summary diagnostics by combining the fitted model returned by \texttt{fit\_highd\_model()} with the \(p\text{-}D\) data.
+After renaming the model output to avoid join conflicts, \texttt{predict\_emb()} is used to assign each observation to a hexagon in the \(2\text{-}D\) embedding.
+The resulting bin assignments are joined with both the model output and the \(p\text{-}D\) data, allowing squared and absolute differences between the true and modeled \(p\text{-}D\) coordinates to be computed for every dimension.
+Total absolute error (\texttt{Error}) and hexbin error (\texttt{HBE}, the root mean squared error) are then obtained using an efficient C++ implementation (\texttt{compute\_errors()}), and returned in a tidy tibble.
 
@@
 
-Furthermore, \texttt{augment()} requires both the fitted model object returned by \texttt{fit\_highd\_model()} and \(p\text{-}D\) data to begin.
-It extends the fitted model by adding prediction results and error diagnostics to the original \(p\text{-}D\) data.
-
-The function starts with the same process as is used in the \texttt{glance()} function to produce a predicted point in \(p\text{-}D\) for each point in the \(p\text{-}D\) dataset.
-
-Next, the function computes residuals between each original coordinate (\texttt{x1}, \texttt{x2}, \ldots, \texttt{xp}) and the corresponding modeled coordinate (\texttt{model\_high\_d\_x1}, \ldots, \texttt{model\_high\_d\_xp}) across all dimensions.
-It calculates both squared errors and absolute errors per dimension.
-These are used to compute two aggregate diagnostic measures per point.
-First, the \texttt{row\_wise\_total\_error} which is the total squared error across all dimensions, and the \texttt{row\_wise\_abs\_error} which is the total absolute error across all dimensions.
-
-The final output is a data frame that combines the original IDs, high-dimensional data, predicted bin IDs, modeled coordinates, residuals, row wise total error, absolute error for the fitted values, and row wise total absolute error for each observation.
-The augmented dataset is always returned as a \texttt{tibble::tibble} with the same number of rows as the passed dataset.
+The \texttt{augment()} function provides point-level diagnostics by appending predictions and residuals to the original \(p\text{-}D\) data.
+Using the same prediction process as \texttt{glance()}, it computes dimension-wise residuals, squared errors, and absolute errors, along with two aggregate measures per observation: the total squared error (\texttt{row\_wise\_total\_error}) and the total absolute error (\texttt{row\_wise\_abs\_error}).
+The final output is a tibble containing IDs (\texttt{ID}), \(p\text{-}D\) data, predicted hexagons (\texttt{h}), modeled coordinates, and all residual diagnostics, with one row per observation.
 

### 2025-12-10 paper-quollr.R

@@ -673,14 +673,14 @@ head(predict_data, 5)
 glance(
   x = scurve_model_obj,
   highd_data = scurve
-  )
+)
 
 
 ## ----echo=TRUE----------------------------------------------------------------
 model_error <- augment(
   x = scurve_model_obj,
   highd_data = scurve
-  )
+)
 
 
 ## ----echo=TRUE----------------------------------------------------------------

### 2025-12-10 paper-quollr.Rmd
Updated Figure 3 adding the box

@@
 
-wireframe_scurve_int <- ggplotly(wireframe_scurve, 
-                                width = "350", 
-                                height = "320", 
-                                tooltip = "none") |>
+wireframe_scurve_int <- plot_ly() |>
+    # Wireframe segments
+    add_segments(
+        data = tr_from_to_df_scurve,
+        x = ~x_from, y = ~y_from,
+        xend = ~x_to, yend = ~y_to,
+        line = list(color = "#000000", width = 0.5),
+        hoverinfo = "none",
+        showlegend = FALSE
+    ) |>
+    # Centroid points
+    add_markers(
+        data = df_bin_centroids_scurve,
+        x = ~c_x, y = ~c_y,
+        marker = list(size = 4, color = "#FF7755"),
+        hoverinfo = "none",
+        showlegend = FALSE
+    ) |>
+    # Layout with border
+    layout(
+        width = 350,
+        height = 320,
+        xaxis = list(
+            range = sc_xlims,
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            title = ""
+        ),
+        yaxis = list(
+            range = sc_ylims,
+            showgrid = FALSE,
+            zeroline = FALSE,
+            showticklabels = FALSE,
+            ticks = "",
+            title = ""
+        ),
+
+        # ADD A BOX AROUND THE ENTIRE PLOT PANEL
+        shapes = list(
+            list(
+                type = "rect",
+                xref = "paper", yref = "paper",
+                x0 = 0, y0 = 0,
+                x1 = 1, y1 = 1,
+                line = list(color = "black", width = 1)
+            )
+        )
+    ) |>
     config(
-    staticPlot = TRUE,        # Disables all interactivity (no hover, zoom, pan)
-    displayModeBar = FALSE,   # Hides the plotly toolbar
-    editable = FALSE,         # Disables annotations and editing
-    showTips = FALSE,         # Removes tooltip on hover
-    displaylogo = FALSE,      # Hides plotly logo
-    responsive = FALSE        # Disables responsive resizing
+        staticPlot = TRUE,
+        displayModeBar = FALSE,
+        editable = FALSE,
+        showTips = FALSE,
+        displaylogo = FALSE,
+        responsive = FALSE
     )

### 2025-12-11 paper-quollr.Rmd
Trying to chnage the width

@@
                          enableControls = FALSE,
-                         width = "400px", height = "380px")
+                         width = "400px", height = "450px")
 
@@
       htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 200px;  text-align: left;', htmltools::h5("c"), wireframe_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 450px; text-align: left;', htmltools::div(
-    style = "margin-left: 120px;",
+      htmltools::div(style = 'margin: 0; padding: 0; height: 600px; width: 450px; text-align: left;', htmltools::div(
+    style = "margin-left: 120px; margin-top: 50px;",
     htmltools::h5("d"),

### 2025-12-11 paper-quollr.Rmd
Fixed Figure 3

--- 2025-12-10 paper-quollr.Rmd
@@
                          enableControls = FALSE,
-                         width = "400px", height = "380px")
+                         width = "400px", height = "550px")
 
@@
       htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 200px;  text-align: left;', htmltools::h5("c"), wireframe_scurve_int),
-      htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 450px; text-align: left;', htmltools::div(
-    style = "margin-left: 120px;",
+      htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 485px; text-align: left;', htmltools::div(
+    style = "margin-left: 140px;",
     htmltools::h5("d"),
-    scurve_umap_model_vis_n
+    htmltools::div(style = "margin-left: 30px; margin-top: 47px; height: 360px; width: 360px;", scurve_umap_model_vis_n)
   ))

### 2025-12-11 paper-quollr.R

@@ -330,7 +330,7 @@ scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)]
                          pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
                          levelColors = c(clr_choice, "#FF7755"),
                          enableControls = FALSE,
-                         width = "400px", height = "450px")
+                         width = "400px", height = "550px")
 
 
 
@@ -562,10 +562,10 @@ knitr::include_graphics("figures/quollr_workflow.png")
 #       htmltools::div(style = 'margin: 0; padding: 0;  height: 450px; width: 200px; text-align: left;', htmltools::h5("a"), hex_grid_poly_scurve_int),
 #       htmltools::div(style = 'margin: 0; padding: 0; height: 450px; width: 200px; text-align: left;', htmltools::h5("b"), hex_centroids_scurve_int),
 #       htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 200px;  text-align: left;', htmltools::h5("c"), wireframe_scurve_int),
-#       htmltools::div(style = 'margin: 0; padding: 0; height: 600px; width: 450px; text-align: left;', htmltools::div(
-#     style = "margin-left: 120px; margin-top: 50px;",
+#       htmltools::div(style = 'margin: 0; padding: 0; height: 400px; width: 485px; text-align: left;', htmltools::div(
+#     style = "margin-left: 140px;",
 #     htmltools::h5("d"),
-#     scurve_umap_model_vis_n
+#     htmltools::div(style = "margin-left: 30px; margin-top: 47px; height: 360px; width: 360px;", scurve_umap_model_vis_n)
 #   ))
 #     )
 #   ),

### 2025-12-12 paper-quollr.Rmd
link changed to linked and re-worded captions

@@
 
-```{r scurve-nldrlink-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+```{r scurve-nldrlink-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
 Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `scurve` structure.
-Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right, displayed using langevitour.
+Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel linked plots showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right, displayed using langevitour.
 Brushing highlights matching points across both views, enabling comparison of structure preservation between the 2-D layout and the high-dimensional data.", layout = "l-body"}
@@
 
-```{r scurve-linkerror-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
+```{r scurve-linkerror-html, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `langevitour`.
 Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
 Points with higher residuals typically correspond to sparse or bridging regions in the UMAP layout and the less dense ends of the Scurve, whereas low-residual points tend to cluster within denser, well-preserved regions.
-Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel link plot showing the residual distribution (left), UMAP layout (middle), and fitted model overlaid with S-curve data in seven dimensions (right) using langevitour.
+Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel linked plots showing the residual distribution (left), UMAP layout (middle), and fitted model overlaid with S-curve data in seven dimensions (right) using langevitour.
 Brushing highlights corresponding points across all panels, illustrating how high-residual points align with sparse or bridging regions in the UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-body"}
@@
 
-```{r scurve-nldrlink-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
+```{r scurve-nldrlink-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
 Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `scurve` structure.
 To highlight points in `detourr`, use the Select button.
-Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel link plot showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right using detourr.
+Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactively linked plots showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right using detourr.
 Brushing or selecting points in either view highlights the corresponding points in the other panel, allowing assessment of how well the UMAP layout preserves the high-dimensional S-curve structure.", layout = "l-body"}
@@
 
-```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.", fig.pos="H", fig.show='hold'}
+```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the linked plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the linked between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.", fig.pos="H", fig.show='hold'}
 
@@
 
-```{r scurve-linkerror-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the `scurve` data in $7\\text{-}D$ (right) using `detourr`.
+```{r scurve-linkerror-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots conecting the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the `scurve` data in $7\\text{-}D$ (right) using `detourr`.
 Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
@@
 To highlight points in `detourr`, use the Select button.
-Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel link plot showing the residual distribution (left), UMAP layout (middle), and fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr.
+Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel linked plots showing the residual distribution (left), UMAP layout (middle), and fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr.
 Brushing or selecting points in any panel highlights corresponding points across all views, illustrating how high-residual points correspond to sparse or bridging regions in UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-body"}
@@
 
-```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \\texttt{scurve} data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the linked plots allowing exploration of the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \\texttt{scurve} data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing linked bushing between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 
@@
 
-```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the layout b (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
+```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting layout b (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
 Each color represents a different muscle group, with triangulated hexagon centroids on the NLDR layout connected to their corresponding regions in the high-dimensional projection using `langevitour`.
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the published $2\\text{-}D$ layout (b) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel linked plots showing the published $2\\text{-}D$ layout (b) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
 Colors indicate different muscle groups.
@@
 
-```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Interactive link plots showing the link between the layout f (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
+```{r tsne-link-limb-best, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting layout f (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
 Each color represents a different muscle group, with triangulated hexagon centroids on the NLDR layout connected to their corresponding regions in the high-dimensional projection using `langevitour`.
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel link plot showing the HBE-selected $2\\text{-}D$ layout (f) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
+Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel linked plots showing the HBE-selected $2\\text{-}D$ layout (f) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
 Colors indicate different muscle groups.

### 2025-12-12 paper-quollr.tex
link changed to linked and re-worded captions

@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-nldrlink-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing the linked between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the linked between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the linked plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-nldrlink-dtpdf}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the link between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the link plots showing the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \texttt{scurve} data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-linkerror-dtpdf}
+\includegraphics[width=1\linewidth,alt={Screenshots showing linked bushing between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing linked bushing between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the linked plots allowing exploration of the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \texttt{scurve} data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-linkerror-dtpdf}
 \end{figure}

### 2025-12-12 paper-quollr.Rmd
working on the application section - needs discussion - binwidth is too big i think

@@
 Figure \@ref(fig:limb-hbe) (b) is the reproduction of the published plot.
+XXX This was generated using t-NSNE with perplexity=30, the default hyperparameters.
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better $2\text{-}D$ layout (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:tsne-link-limb), \\@ref(fig:tsne-link-limb-best)"} else { "Figure \\@ref(fig:model-limb)"})`).
-
+Note that the cluster variable is not used to produce the $2\text{-}D$ layout.
+
+We illustrate how to use `quollr` to assess whether this is a reasonable layout.
+Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
+Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init_random_MN-ratio=0.5, FP-ratio=2); layout f was produced by t-SNE (perplexity=15).
+
+The HBE plot indicates that the two t-SNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
+There are small visual difference between the two layouts.
+Both support that there are 5-6 clusters.
+Layout b has slightly more space between clusters.
+Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom.
 <!-- add function to generate 2D trimesh and the projections (PDF) and langevitour (HTML)-->
@@
 
+`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:tsne-link-limb), \\@ref(fig:tsne-link-limb-best)"} else { "Figure \\@ref(fig:model-limb)"})` show how to examine the resulting models of the layouts overlaid on the data in high dimensions.
+Point color represents the cluster reported in the published paper.
+In each case the best model is produced using binwidth=XXX (and why).
+The plots in these figures are  best understand using small steps.
+
+1.
+Examine the model in the data space for each, by looking at the tour views.
+In each case, the clustering doesn't quite match the separations in the data, and both models help see this.
+For example, the orange cluster (1) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
+2.
+Because 6 clusters are hard to examine together, use the menu to select just one cluster to view at at time.
+Selecting just cluster 1 might help you see the explanation above, that a small group is quite separate from the main group.
+This suggests both the layouts and the clustering that groups these together might be wrong.
+Now change to focus on cluster 5 (yellow).
+This group is a fairly large sparse cluster but it is separated from the other points.
+Both layouts are right in separating these points from the others.
+3.
+To examine where the layouts differ, examine clusters 4 (green) and 6 (cream), by selecting just these two.
+Layout b has them close, but layout f has them far apart.
+(It might also help to include cluster 5 here because layout b has this group close to the cluster 4 also).
+In the tour view, we can see that they three clusters are separated clusters in different directions away from the most of the other points.
+They are both correct in this.
+It may have been better to place them all in different corners of the layout, but they have preserved the most important aspect that they are separated clusters.
+That they are all close together in layout b could be incorrectly interpreted as close in high-dimensions though.
+
 ```{r}

### 2025-12-12 paper-quollr.tex
working on the application section - needs discussion - binwidth is too big i think

@@
 Figure \ref{fig:limb-hbe} (b) is the reproduction of the published plot.
+XXX This was generated using t-NSNE with perplexity=30, the default hyperparameters.
 The question is whether this accurately represents the cluster structure in the data.
-Our method help to provide a better \(2\text{-}D\) layout (Figure \ref{fig:model-limb}).
+Note that the cluster variable is not used to produce the \(2\text{-}D\) layout.
+
+We illustrate how to use \texttt{quollr} to assess whether this is a reasonable layout.
+Figure \ref{fig:limb-hbe} shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
+Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init\_random\_MN-ratio=0.5, FP-ratio=2); layout f was produced by t-SNE (perplexity=15).
+
+The HBE plot indicates that the two t-SNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
+There are small visual difference between the two layouts.
+Both support that there are 5-6 clusters.
+Layout b has slightly more space between clusters.
+Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom.
 
@@
 \end{figure}
+
+Figure \ref{fig:model-limb} show how to examine the resulting models of the layouts overlaid on the data in high dimensions.
+Point color represents the cluster reported in the published paper.
+In each case the best model is produced using binwidth=XXX (and why).
+The plots in these figures are best understand using small steps.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Examine the model in the data space for each, by looking at the tour views.
+In each case, the clustering doesn't quite match the separations in the data, and both models help see this.
+For example, the orange cluster (1) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
+\item
+  Because 6 clusters are hard to examine together, use the menu to select just one cluster to view at at time.
+Selecting just cluster 1 might help you see the explanation above, that a small group is quite separate from the main group.
+This suggests both the layouts and the clustering that groups these together might be wrong.
+Now change to focus on cluster 5 (yellow).
+This group is a fairly large sparse cluster but it is separated from the other points.
+Both layouts are right in separating these points from the others.
+3.
+To examine where the layouts differ, examine clusters 4 (green) and 6 (cream), by selecting just these two.
+Layout b has them close, but layout f has them far apart.
+(It might also help to include cluster 5 here because layout b has this group close to the cluster 4 also).
+In the tour view, we can see that they three clusters are separated clusters in different directions away from the most of the other points.
+They are both correct in this.
+It may have been better to place them all in different corners of the layout, but they have preserved the most important aspect that they are separated clusters.
+That they are all close together in layout b could be incorrectly interpreted as close in high-dimensions though.
+\end{enumerate}
 

### 2025-12-14 extra/paper-quollr-code.qmd
Added another future work

---
title: "paper-quollr-code"
format: html
---

```{r, fig.cap="Examining the differences in bill sizes of three species of penguins. Colour indicates species. Gentoo bills are smaller in depth but longer. Chinstrap and Adelie bills have similarly large depth but Chinstrap bills are longer."}

library(ggplot2)
library(palmerpenguins)
p <- penguins |>  
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(13, 22, 2)) +
  scale_y_continuous(breaks = seq(30, 60, 10)) +
  theme(aspect.ratio = 1)
p
```

### 2025-12-14 paper-quollr.Rmd
Added another future work

@@
   nldr_data = tsne_limb, 
-  b1 = 19, 
+  b1 = 21, 
   q = 0.1, 
@@
   nldr_data = tsne_limb2, 
-  b1 = 19, 
+  b1 = 21, 
   q = 0.1, 
@@
 Additionally, statistical inference tools could be introduced to assess the stability and robustness of the fitted model, which would enhance interpretability and confidence in the outcomes.
+Currently, linked plot functionality is limited: selecting a cluster in the `langevitour` controls does not yet highlight the corresponding cluster in the $2\text{-}D$ layout.
+Implementing this feature in the future would allow users to clearly explore cluster-specific relationships between the `langevitour` and $2\text{-}D$ layout.
 

### 2025-12-14 paper-quollr.tex
Added another future work

@@
 Additionally, statistical inference tools could be introduced to assess the stability and robustness of the fitted model, which would enhance interpretability and confidence in the outcomes.
+Currently, linked plot functionality is limited: selecting a cluster in the \texttt{langevitour} controls does not yet highlight the corresponding cluster in the \(2\text{-}D\) layout.
+Implementing this feature in the future would allow users to clearly explore cluster-specific relationships between the \texttt{langevitour} and \(2\text{-}D\) layout.
 

### 2025-12-14 paper-quollr.R

@@ -1561,7 +1561,7 @@ cluster_df <- read_rds("data/limb_muscles/facs_limb_muscles_cluster_df.rds")
 tsne_limb_obj <- fit_highd_model(
   highd_data = data_limb, 
   nldr_data = tsne_limb, 
-  b1 = 19, 
+  b1 = 21, 
   q = 0.1, 
   hd_thresh = 0)
 
@@ -1802,7 +1802,7 @@ limb_proj_tsne_model2 <- plot_proj(
 tsne_best_limb_obj <- fit_highd_model(
   highd_data = data_limb, 
   nldr_data = tsne_limb2, 
-  b1 = 19, 
+  b1 = 21, 
   q = 0.1, 
   hd_thresh = 0)
 

### 2025-12-14 paper-quollr.Rmd
Chose another binwidth

@@
 Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init_random_MN-ratio=0.5, FP-ratio=2); layout f was produced by t-SNE (perplexity=15).
-
-The HBE plot indicates that the two t-SNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
+Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init_random_MN-ratio=0.5, FP-ratio=2); layout f was produced by tSNE (perplexity=15).
+
+The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
 There are small visual difference between the two layouts.
@@
   nldr_data = tsne_limb, 
-  b1 = 21, 
+  b1 = 28, 
   q = 0.1, 
@@
 Point color represents the cluster reported in the published paper.
-In each case the best model is produced using binwidth=XXX (and why).
+In each case the best model is produced using binwidth=0.06 (and why).
 The plots in these figures are  best understand using small steps.
@@
   nldr_data = tsne_limb2, 
-  b1 = 21, 
+  b1 = 28, 
   q = 0.1, 

### 2025-12-14 paper-quollr.tex
Chose another binwidth

@@
 Figure \ref{fig:limb-hbe} shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init\_random\_MN-ratio=0.5, FP-ratio=2); layout f was produced by t-SNE (perplexity=15).
-
-The HBE plot indicates that the two t-SNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
+Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init\_random\_MN-ratio=0.5, FP-ratio=2); layout f was produced by tSNE (perplexity=15).
+
+The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
 There are small visual difference between the two layouts.
@@
 Point color represents the cluster reported in the published paper.
-In each case the best model is produced using binwidth=XXX (and why).
+In each case the best model is produced using binwidth=0.06 (and why).
 The plots in these figures are best understand using small steps.

### 2025-12-14 paper-quollr.Rmd
Updated the application section

@@
 Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
-NLDR often used to summarise the discovered clusters, and help to understand the results.
+NLDR often used to summarize the discovered clusters, and help to understand the results.
 The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
@@
 Figure \@ref(fig:limb-hbe) (b) is the reproduction of the published plot.
-XXX This was generated using t-NSNE with perplexity=30, the default hyperparameters.
+This was generated using tSNE with $\text{perplexity}=30$, the default hyper-parameters.
 The question is whether this accurately represents the cluster structure in the data.
@@
 Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init_random_MN-ratio=0.5, FP-ratio=2); layout f was produced by tSNE (perplexity=15).
-
-The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
+Layout a is produce by UMAP ($\text{neighbors}=5, ~\text{minimum distance}=0.1$); layout c was produced by PHATE ($\text{knn}=5$); layout d was produced by TriMAP ($\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3$); layout e was produced by by PaCMAP ($\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2$); layout f was produced by tSNE ($\text{perplexity}=15$).
+
+The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of $15$ outperforms the other.
 There are small visual difference between the two layouts.
-Both support that there are 5-6 clusters.
+Both support that there are $5-6$ clusters.
 Layout b has slightly more space between clusters.
@@
 For example, the orange cluster (1) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
+
 2.
@@
 Both layouts are right in separating these points from the others.
+
 3.

### 2025-12-14 paper-quollr.tex
Updated the application section

@@
 Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
-NLDR often used to summarise the discovered clusters, and help to understand the results.
+NLDR often used to summarize the discovered clusters, and help to understand the results.
 The purpose of this example is to \emph{illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data}.
@@
 Figure \ref{fig:limb-hbe} (b) is the reproduction of the published plot.
-XXX This was generated using t-NSNE with perplexity=30, the default hyperparameters.
+This was generated using tSNE with \(\text{perplexity}=30\), the default hyper-parameters.
 The question is whether this accurately represents the cluster structure in the data.
@@
 Figure \ref{fig:limb-hbe} shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init\_random\_MN-ratio=0.5, FP-ratio=2); layout f was produced by tSNE (perplexity=15).
-
-The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other.
+Layout a is produce by UMAP (\(\text{neighbors}=5, ~\text{minimum distance}=0.1\)); layout c was produced by PHATE (\(\text{knn}=5\)); layout d was produced by TriMAP (\(\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3\)); layout e was produced by by PaCMAP (\(\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2\)); layout f was produced by tSNE (\(\text{perplexity}=15\)).
+
+The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of \(15\) outperforms the other.
 There are small visual difference between the two layouts.
-Both support that there are 5-6 clusters.
+Both support that there are \(5-6\) clusters.
 Layout b has slightly more space between clusters.
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
 \item
@@
 Both layouts are right in separating these points from the others.
-3.
+\item
   To examine where the layouts differ, examine clusters 4 (green) and 6 (cream), by selecting just these two.

### 2025-12-14 paper-quollr.Rmd
Put the already generated layout first

@@
 Following their pre-processing, different NLDR methods were performed using ten principal components.
-Figure \@ref(fig:limb-hbe) (b) is the reproduction of the published plot.
+Figure \@ref(fig:limb-hbe) (a) is the reproduction of the published plot.
 This was generated using tSNE with $\text{perplexity}=30$, the default hyper-parameters.
@@
 Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout a is produce by UMAP ($\text{neighbors}=5, ~\text{minimum distance}=0.1$); layout c was produced by PHATE ($\text{knn}=5$); layout d was produced by TriMAP ($\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3$); layout e was produced by by PaCMAP ($\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2$); layout f was produced by tSNE ($\text{perplexity}=15$).
+Layout b is produce by UMAP ($\text{neighbors}=5, ~\text{minimum distance}=0.1$); layout c was produced by PHATE ($\text{knn}=5$); layout d was produced by TriMAP ($\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3$); layout e was produced by by PaCMAP ($\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2$); layout f was produced by tSNE ($\text{perplexity}=15$).
 
@@
 Both support that there are $5-6$ clusters.
-Layout b has slightly more space between clusters.
+Layout a has slightly more space between clusters.
 Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom.
@@
 
+tsne_limb <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_30.rds")
+
+nldr1 <- tsne_limb |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("a")
+
 umap_limb <- read_rds("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_15_min-dist_0.1.rds")
 
-nldr1 <- umap_limb |>
+nldr2 <- umap_limb |>
   ggplot(aes(x = emb1,
@@
   geom_point(alpha=0.1, size=1, colour='#999999') +
-  interior_annotation("a")
-
-tsne_limb <- read_rds("data/limb_muscles/facs_limb_muscles_tsne_perplexity_30.rds")
-
-nldr2 <- tsne_limb |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
   interior_annotation("b")
@@
   mutate(method = factor(method,
-                         levels = c("UMAP_15_min_dist_0.1", "tsne_30", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2", "tsne_15")))
+                         levels = c("tsne_30", "UMAP_15_min_dist_0.1", "phate_5", "trimap_n-inliers_12_n-outliers_4_n-random_3", "pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2", "tsne_15")))
 ```
@@
   scale_x_continuous(breaks = sort(unique(error_limb$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#984ea3','#4daf4a','#ff7f00'))
+  scale_color_manual(values=c('#a65628','#999999','#e41a1c','#984ea3','#4daf4a','#ff7f00'))
 
@@
 To examine where the layouts differ, examine clusters 4 (green) and 6 (cream), by selecting just these two.
-Layout b has them close, but layout f has them far apart.
+Layout a has them close, but layout f has them far apart.
 (It might also help to include cluster 5 here because layout b has this group close to the cluster 4 also).
@@
 It may have been better to place them all in different corners of the layout, but they have preserved the most important aspect that they are separated clusters.
-That they are all close together in layout b could be incorrectly interpreted as close in high-dimensions though.
+That they are all close together in layout a could be incorrectly interpreted as close in high-dimensions though.
 
@@
   scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
-  interior_annotation("b1", cex = 2) +
+  interior_annotation("f1", cex = 2) +
   theme(
@@
   is_category = TRUE) +
-  interior_annotation(label = "b2", cex = 2) + 
+  interior_annotation(label = "f2", cex = 2) + 
   scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
@@
   is_category = TRUE) +
-  interior_annotation(label = "b3", cex = 2) + 
+  interior_annotation(label = "f3", cex = 2) + 
   scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
@@
 
-```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}b) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}a) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published $2\\text{-}D$ layout (b) and the HBE-selected $2\\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
@@
 
-```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting layout b (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
+```{r tsne-link-limb, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting layout a (left) and the fitted model overlaid with the limb muscle data in $10\\text{-}D$ (right).
 Each color represents a different muscle group, with triangulated hexagon centroids on the NLDR layout connected to their corresponding regions in the high-dimensional projection using `langevitour`.

### 2025-12-14 paper-quollr.tex
Put the already generated layout first

@@
 Following their pre-processing, different NLDR methods were performed using ten principal components.
-Figure \ref{fig:limb-hbe} (b) is the reproduction of the published plot.
+Figure \ref{fig:limb-hbe} (a) is the reproduction of the published plot.
 This was generated using tSNE with \(\text{perplexity}=30\), the default hyper-parameters.
@@
 Figure \ref{fig:limb-hbe} shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout a is produce by UMAP (\(\text{neighbors}=5, ~\text{minimum distance}=0.1\)); layout c was produced by PHATE (\(\text{knn}=5\)); layout d was produced by TriMAP (\(\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3\)); layout e was produced by by PaCMAP (\(\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2\)); layout f was produced by tSNE (\(\text{perplexity}=15\)).
+Layout b is produce by UMAP (\(\text{neighbors}=5, ~\text{minimum distance}=0.1\)); layout c was produced by PHATE (\(\text{knn}=5\)); layout d was produced by TriMAP (\(\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3\)); layout e was produced by by PaCMAP (\(\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2\)); layout f was produced by tSNE (\(\text{perplexity}=15\)).
 
@@
 Both support that there are \(5-6\) clusters.
-Layout b has slightly more space between clusters.
+Layout a has slightly more space between clusters.
 Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom.
@@
   To examine where the layouts differ, examine clusters 4 (green) and 6 (cream), by selecting just these two.
-Layout b has them close, but layout f has them far apart.
+Layout a has them close, but layout f has them far apart.
 (It might also help to include cluster 5 here because layout b has this group close to the cluster 4 also).
@@
 It may have been better to place them all in different corners of the layout, but they have preserved the most important aspect that they are separated clusters.
-That they are all close together in layout b could be incorrectly interpreted as close in high-dimensions though.
+That they are all close together in layout a could be incorrectly interpreted as close in high-dimensions though.
 \end{enumerate}
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Comparison of published $2\text{-}D$ layout (b) and the HBE-selected $2\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}b) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
+\includegraphics[width=1\linewidth,alt={Comparison of published $2\text{-}D$ layout (b) and the HBE-selected $2\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}a) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
 The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}

### 2025-12-14 paper-quollr.Rmd
Updated paper organization in the introduction
Updating teh application section

@@
 The paper is organized as follows.
+The usage section explains how users can fit a complete model pipeline from hexagonal binning of a $2\text{-}D$ layout to lifting the representation back into $p\text{-}D$ and how the resulting objects can be explored interactively using tour.
 The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
@@
 Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout b is produce by UMAP ($\text{neighbors}=5, ~\text{minimum distance}=0.1$); layout c was produced by PHATE ($\text{knn}=5$); layout d was produced by TriMAP ($\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3$); layout e was produced by by PaCMAP ($\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2$); layout f was produced by tSNE ($\text{perplexity}=15$).
+Layout b is produce by UMAP ($\text{neighbors}=5, ~\text{minimum distance}=0.1$); layout c was produced by PHATE ($\text{knn}=5$); layout d was produced by TriMAP ($\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3$); layout e was produced by by PaCMAP ($\text{neighbors}=10,~\text{init = "random"}, ~\text{MN-ratio}=0.5,~\text{FP-ratio}=2$); layout f was produced by tSNE ($\text{perplexity}=15$).
 
@@
 Point color represents the cluster reported in the published paper.
-In each case the best model is produced using binwidth=0.06 (and why).
+In each case the best model is produced using $\text{binwidth}=0.06$ (and why).
 The plots in these figures are best understand using small steps.
@@
 In each case, the clustering doesn't quite match the separations in the data, and both models help see this.
-For example, the orange cluster (1) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
+For example, the orange cluster ($1$) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
 
 2.
-Because 6 clusters are hard to examine together, use the menu to select just one cluster to view at at time.
-Selecting just cluster 1 might help you see the explanation above, that a small group is quite separate from the main group.
+Because $6$ clusters are hard to examine together, use the menu to select just one cluster to view at at time.
+Selecting just cluster $1$ might help you see the explanation above, that a small group is quite separate from the main group.
 This suggests both the layouts and the clustering that groups these together might be wrong.
-Now change to focus on cluster 5 (yellow).
+Now change to focus on cluster $5$ (yellow).
 This group is a fairly large sparse cluster but it is separated from the other points.
@@
 3.
-To examine where the layouts differ, examine clusters 4 (green) and 6 (cream), by selecting just these two.
+To examine where the layouts differ, examine clusters $4$ (green) and $6$ (cream), by selecting just these two.
 Layout a has them close, but layout f has them far apart.
-(It might also help to include cluster 5 here because layout b has this group close to the cluster 4 also).
+(It might also help to include cluster $5$ here because layout a has this group close to the cluster $4$ also).
 In the tour view, we can see that they three clusters are separated clusters in different directions away from the most of the other points.

### 2025-12-14 paper-quollr.tex
Updated paper organization in the introduction
Updating teh application section

@@
 The paper is organized as follows.
+The usage section explains how users can fit a complete model pipeline from hexagonal binning of a \(2\text{-}D\) layout to lifting the representation back into \(p\text{-}D\) and how the resulting objects can be explored interactively using tour.
 The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities.
@@
 Figure \ref{fig:limb-hbe} shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
-Layout b is produce by UMAP (\(\text{neighbors}=5, ~\text{minimum distance}=0.1\)); layout c was produced by PHATE (\(\text{knn}=5\)); layout d was produced by TriMAP (\(\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3\)); layout e was produced by by PaCMAP (\(\text{neighbors}=10,~\text{init = "random"}~\text{MN-ratio}=0.5,~\text{FP-ratio}=2\)); layout f was produced by tSNE (\(\text{perplexity}=15\)).
+Layout b is produce by UMAP (\(\text{neighbors}=5, ~\text{minimum distance}=0.1\)); layout c was produced by PHATE (\(\text{knn}=5\)); layout d was produced by TriMAP (\(\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3\)); layout e was produced by by PaCMAP (\(\text{neighbors}=10,~\text{init = "random"}, ~\text{MN-ratio}=0.5,~\text{FP-ratio}=2\)); layout f was produced by tSNE (\(\text{perplexity}=15\)).
 
@@
 Point color represents the cluster reported in the published paper.
-In each case the best model is produced using binwidth=0.06 (and why).
+In each case the best model is produced using \(\text{binwidth}=0.06\) (and why).
 The plots in these figures are best understand using small steps.
@@
 In each case, the clustering doesn't quite match the separations in the data, and both models help see this.
-For example, the orange cluster (1) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
-\item
-  Because 6 clusters are hard to examine together, use the menu to select just one cluster to view at at time.
-Selecting just cluster 1 might help you see the explanation above, that a small group is quite separate from the main group.
+For example, the orange cluster (\(1\)) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
+\item
+  Because \(6\) clusters are hard to examine together, use the menu to select just one cluster to view at at time.
+Selecting just cluster \(1\) might help you see the explanation above, that a small group is quite separate from the main group.
 This suggests both the layouts and the clustering that groups these together might be wrong.
-Now change to focus on cluster 5 (yellow).
+Now change to focus on cluster \(5\) (yellow).
 This group is a fairly large sparse cluster but it is separated from the other points.
@@
 \item
-  To examine where the layouts differ, examine clusters 4 (green) and 6 (cream), by selecting just these two.
+  To examine where the layouts differ, examine clusters \(4\) (green) and \(6\) (cream), by selecting just these two.
 Layout a has them close, but layout f has them far apart.
-(It might also help to include cluster 5 here because layout b has this group close to the cluster 4 also).
+(It might also help to include cluster \(5\) here because layout a has this group close to the cluster \(4\) also).
 In the tour view, we can see that they three clusters are separated clusters in different directions away from the most of the other points.

### 2025-12-14 paper-quollr.Rmd
Added the reason on why we chose this binwidth

@@
 Point color represents the cluster reported in the published paper.
-In each case the best model is produced using $\text{binwidth}=0.06$ (and why).
+In each case the best model is produced using $\text{binwidth}=0.06$.
+A binwidth of $0.06$ is used because it is small enough to show local structure and differences between clusters, but large enough to avoid breaking the layout into too many small pieces that would make the plots hard to interpret.
 The plots in these figures are best understand using small steps.

### 2025-12-14 paper-quollr.tex
Added the reason on why we chose this binwidth

@@
 Point color represents the cluster reported in the published paper.
-In each case the best model is produced using \(\text{binwidth}=0.06\) (and why).
+In each case the best model is produced using \(\text{binwidth}=0.06\).
+A binwidth of \(0.06\) is used because it is small enough to show local structure and differences between clusters, but large enough to avoid breaking the layout into too many small pieces that would make the plots hard to interpret.
 The plots in these figures are best understand using small steps.

### 2025-12-14 paper-quollr.Rmd
Removed detourr linked plot section

@@
 
-As an alternative to using `langevitour`, the tour can also be generated with the `detourr`.
-In this setup, users can manually construct the linked visualization using the `crosstalk` [@joe2023] and `htmltools` [@joe2024].
-The interactive layout is created by arranging the $2\text{-}D$ NLDR plot (`nldr_plt`), the optional error distribution plot (`error_plt`), and the tour view produced with `detourr`, side by side within a flexible grid.
-The NLDR and error plots are made with `ggplotly()` [@chapman2020] to enable interactive linking.
-The `bscols()` function from `crosstalk` manages synchronization across these panels, allowing for linked brushing between these interactive plots.
-
-A two-panel linked plot (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-nldrlink-dthtml)"} else { "Figure \\@ref(fig:scurve-nldrlink-dtpdf)"})`) combining the NLDR view and the tour from `detourr` can be created as follows:
-
-```{r}
-point_colours <- c("#66B2CC", "#FF7755")
-point_sizes <- c(0, 1)
-
-num_highd_col <- df_exe |>
-  dplyr::select(starts_with("x")) |>
-  NCOL()
-
-df_all <- df_exe |>
-  dplyr::filter(type == "data") ## original dataset
-
-df_b <- df_exe |>
-  dplyr::filter(type == "model") ## High-d model
-
-shared_df <- crosstalk::SharedData$new(df_exe)
-
-nldr_plt <- plot_ly(
-    shared_df,
-    x = ~emb1, y = ~emb2,
-    type = "scatter",
-    mode = "markers",
-    marker = list(
-        color = point_colours[1],
-        size = 3,
-        opacity = 0.5
-    ),
-    hoverinfo = "none"
-) |>
-    layout(
-        width = 300,
-        height = 300,
-        xaxis = list(
-            title = "",
-            showgrid = FALSE,
-            zeroline = FALSE,
-            showticklabels = FALSE,
-            ticks = "",
-            linecolor = "black",
-            mirror = TRUE
-        ),
-        yaxis = list(
-            title = "",
-            showgrid = FALSE,
-            zeroline = FALSE,
-            showticklabels = FALSE,
-            ticks = "",
-            linecolor = "black",
-            mirror = TRUE
-        ),
-        margin = list(l = 20, r = 20, t = 20, b = 20),
-        dragmode = "select"
-    ) |> 
-    style(selected   = list(marker = list(opacity = 1)), unselected=list(marker=list(opacity=1))) |>
-    highlight(on="plotly_selected", off="plotly_deselect") |>
-    config(displayModeBar = FALSE)
-```
-
-```{r, echo=TRUE}
-
-detourr_output <- detour(
-  shared_df,
-  tour_aes(
-    projection = starts_with("x"),
-    colour = type
-  )
-) |>
-  tour_path(grand_tour(2), 
-                    max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 0.5, alpha = 0.8, 
-               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
-               palette = c("#66B2CC", "#FF7755"),
-                width = "300px", height = "300px")
-
-```
-
-```{r scurve-nldrlink-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots connecting the UMAP layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using `detourr`.
-Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `scurve` structure.
-To highlight points in `detourr`, use the Select button.
-Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactively linked plots showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right using detourr.
-Brushing or selecting points in either view highlights the corresponding points in the other panel, allowing assessment of how well the UMAP layout preserves the high-dimensional S-curve structure.", layout = "l-body"}
-
-lndet_link <- crosstalk::bscols(
-    htmltools::div(
-        style = "display: grid; grid-template-columns: 1fr 1fr;
-    gap: 0px; align-items: start; justify-items: center; margin: 0; padding: 0;
-    height: 350px; width: 300px;",
-        nldr_plt, htmltools::div(style = "margin-top: 20px;", detourr_output)
-    ),
-    device = "xs"
-)
-
-class(lndet_link) <- c(class(lndet_link), "htmlwidget")
-
-lndet_link
-```
-
-```{r scurve-nldrlink-dtpdf, eval=knitr::is_latex_output(), out.width=c("100%", "100%"), fig.cap="Screenshots of the linked plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing the linked between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.", fig.pos="H", fig.show='hold'}
-
-knitr::include_graphics(c("figures/model_link_proj1_detourr.png",
-                          "figures/model_link_proj2_detourr.png"))
-```
-
-Also, a three-panel view (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:scurve-linkerror-dthtml)"} else { "Figure \\@ref(fig:scurve-linkerror-dtpdf)"})`) can be constructed by adding the error distribution plot (`error_plt`):
-
-```{r}
-
-num_highd_col <- df_exe |>
-  dplyr::select(starts_with("x")) |>
-  NCOL()
-
-df_all <- df_exe |>
-  dplyr::filter(type == "data") ## original dataset
-
-df_b <- df_exe |>
-  dplyr::filter(type == "model") ## High-d model
-
-shared_df <- crosstalk::SharedData$new(df_exe)
-
-error_plt <- plot_ly(
-    shared_df,
-    x = ~sqrt_row_wise_total_error, y = ~density,
-    type = "scatter",
-    mode = "markers",
-    marker = list(
-        color = point_colours[1],
-        size = 3,
-        opacity = 0.5
-    ),
-    hoverinfo = "none"
-) |>
-    layout(
-        width = 250,
-        height = 250,
-        xaxis = list(
-            title = "",
-            showgrid = FALSE,
-            zeroline = FALSE,
-            showticklabels = FALSE,
-            ticks = "",
-            linecolor = "black",
-            mirror = TRUE
-        ),
-        yaxis = list(
-            title = "",
-            showgrid = FALSE,
-            zeroline = FALSE,
-            showticklabels = FALSE,
-            ticks = "",
-            linecolor = "black",
-            mirror = TRUE
-        ),
-        margin = list(l = 20, r = 20, t = 20, b = 20),
-        dragmode = "select"
-    ) |> 
-    style(selected   = list(marker = list(opacity = 1)), unselected=list(marker=list(opacity=1))) |>
-    highlight(on="plotly_selected", off="plotly_deselect") |>
-    config(displayModeBar = FALSE)
-
-nldr_plt <- plot_ly(
-    shared_df,
-    x = ~emb1, y = ~emb2,
-    type = "scatter",
-    mode = "markers",
-    marker = list(
-        color = point_colours[1],
-        size = 3,
-        opacity = 0.5
-    ),
-    hoverinfo = "none"
-) |>
-    layout(
-        width = 250,
-        height = 250,
-        xaxis = list(
-            title = "",
-            showgrid = FALSE,
-            zeroline = FALSE,
-            showticklabels = FALSE,
-            ticks = "",
-            linecolor = "black",
-            mirror = TRUE
-        ),
-        yaxis = list(
-            title = "",
-            showgrid = FALSE,
-            zeroline = FALSE,
-            showticklabels = FALSE,
-            ticks = "",
-            linecolor = "black",
-            mirror = TRUE
-        ),
-        margin = list(l = 20, r = 20, t = 20, b = 20),
-        dragmode = "select"
-    ) |> 
-    style(selected   = list(marker = list(opacity = 1)), unselected=list(marker=list(opacity=1))) |>
-    highlight(on="plotly_selected", off="plotly_deselect") |>
-    config(displayModeBar = FALSE)
-```
-
-```{r scurve-linkerror-dthtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Interactively linked plots conecting the residual distribution (left), the UMAP layout (middle) and the fitted model overlaid with the `scurve` data in $7\\text{-}D$ (right) using `detourr`.
-Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
-Points with higher residuals typically correspond to sparse or bridging regions in the UMAP layout and the less dense ends of the Scurve, whereas low-residual points tend to cluster within denser, well-preserved regions.
-To highlight points in `detourr`, use the Select button.
-Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel linked plots showing the residual distribution (left), UMAP layout (middle), and fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr.
-Brushing or selecting points in any panel highlights corresponding points across all views, illustrating how high-residual points correspond to sparse or bridging regions in UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-body"}
-
-detourr_output <- detour(
-  shared_df,
-  tour_aes(
-    projection = starts_with("x"),
-    colour = type
-  )
-) |>
-  tour_path(grand_tour(2), 
-                    max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 0.5, alpha = 0.8, 
-               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
-               palette = c("#66B2CC", "#FF7755"),
-                width = "250px", height = "250px")
-
-erlndet_link <- crosstalk::bscols(
-    htmltools::div(
-        style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
-    gap: 0px; align-items: start; justify-items: center; margin: 0; padding: 0;
-    height: 300px; width: 400px;",
-        error_plt, nldr_plt, 
-    htmltools::div(style = "margin-top: 20px;", detourr_output)
-    ),
-    device = "xs"
-)
-
-class(erlndet_link) <- c(class(erlndet_link), "htmlwidget")
-
-erlndet_link
-```
-
-```{r scurve-linkerror-dtpdf, out.width=c("100%", "100%"), fig.cap="Screenshots of the linked plots allowing exploration of the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \\texttt{scurve} data in $7\\text{-}D$ (right) using \\texttt{detourr}.", fig.alt = "Screenshots showing linked bushing between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
-
-knitr::include_graphics(c("figures/model_link_error_proj1_detourr.png",
-                          "figures/model_link_error_proj2_detourr.png"))
-```
-
 # Application
@@
 
+point_sizes <- c(0, 1)
+
 # Scatter points

### 2025-12-14 paper-quollr.tex
Removed detourr linked plot section

@@
 
-As an alternative to using \texttt{langevitour}, the tour can also be generated with the \texttt{detourr}.
-In this setup, users can manually construct the linked visualization using the \texttt{crosstalk} \citep{joe2023} and \texttt{htmltools} \citep{joe2024}.
-The interactive layout is created by arranging the \(2\text{-}D\) NLDR plot (\texttt{nldr\_plt}), the optional error distribution plot (\texttt{error\_plt}), and the tour view produced with \texttt{detourr}, side by side within a flexible grid.
-The NLDR and error plots are made with \texttt{ggplotly()} \citep{chapman2020} to enable interactive linking.
-The \texttt{bscols()} function from \texttt{crosstalk} manages synchronization across these panels, allowing for linked brushing between these interactive plots.
-
-A two-panel linked plot (Figure \ref{fig:scurve-nldrlink-dtpdf}) combining the NLDR view and the tour from \texttt{detourr} can be created as follows:
-
-\begin{verbatim}
-detourr_output <- detour(
-  shared_df,
-  tour_aes(
-    projection = starts_with("x"),
-    colour = type
-  )
-) |>
-  tour_path(grand_tour(2), 
-                    max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 0.5, alpha = 0.8, 
-               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
-               palette = c("#66B2CC", "#FF7755"),
-                width = "300px", height = "300px")
-\end{verbatim}
-
-\begin{verbatim}
-lndet_link <- crosstalk::bscols(
-    htmltools::div(
-        style = "display: grid; grid-template-columns: 1fr 1fr;
-    gap: 0px; align-items: start; justify-items: center; margin: 0; padding: 0;
-    height: 350px; width: 300px;",
-        nldr_plt, htmltools::div(style = "margin-top: 20px;", detourr_output)
-    ),
-    device = "xs"
-)
-
-class(lndet_link) <- c(class(lndet_link), "htmlwidget")
-
-lndet_link
-\end{verbatim}
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing the linked between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing the linked between the NLDR layout on the left and the fitted high-dimensional model with S-curve data on the right using detourr, illustrating how the 2-D layout corresponds to the 7-D data structure.}]{figures/model_link_proj2_detourr} \caption{Screenshots of the linked plots showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-nldrlink-dtpdf}
-\end{figure}
-
-Also, a three-panel view (Figure \ref{fig:scurve-linkerror-dtpdf}) can be constructed by adding the error distribution plot (\texttt{error\_plt}):
-
-\begin{verbatim}
-detourr_output <- detour(
-  shared_df,
-  tour_aes(
-    projection = starts_with("x"),
-    colour = type
-  )
-) |>
-  tour_path(grand_tour(2), 
-                    max_bases=50, fps = 60) |>
-  show_scatter(axes = TRUE, size = 0.5, alpha = 0.8, 
-               edges = as.matrix(trimesh[, c("from_reindexed", "to_reindexed")]),
-               palette = c("#66B2CC", "#FF7755"),
-                width = "250px", height = "250px")
-
-erlndet_link <- crosstalk::bscols(
-    htmltools::div(
-        style = "display: grid; grid-template-columns: 1fr 1fr 1fr;
-    gap: 0px; align-items: start; justify-items: center; margin: 0; padding: 0;
-    height: 300px; width: 400px;",
-        error_plt, nldr_plt, 
-    htmltools::div(style = "margin-top: 20px;", detourr_output)
-    ),
-    device = "xs"
-)
-
-class(erlndet_link) <- c(class(erlndet_link), "htmlwidget")
-
-erlndet_link
-\end{verbatim}
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Screenshots showing linked bushing between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj1_detourr} \includegraphics[width=1\linewidth,alt={Screenshots showing linked bushing between the residual distribution (left), NLDR layout (middle), and the fitted high-dimensional model with S-curve data in seven dimensions (right) using detourr, illustrating how residuals correspond to structures in both the 2-D layout and the high-dimensional data.}]{figures/model_link_error_proj2_detourr} \caption{Screenshots of the linked plots allowing exploration of the relationship between the distribution of residuals (left), NLDR layout (middle) and the fitted model overlaid with the \texttt{scurve} data in $7\text{-}D$ (right) using \texttt{detourr}.}\label{fig:scurve-linkerror-dtpdf}
-\end{figure}
-
 \section{Application}\label{application}

### 2025-12-14 paper-quollr.Rmd
Added setence with linked plots
slight addition to application sentence

@@
 
+In addition to `langevitour`, linked plots can also be constructed using tour views generated with the `detourr` package.
+(see [vignette: 8.
+Linked plots with `detourr`](https://jayanilakshika.github.io/quollr/articles/quollr8linkeddetourr.html) for details.)
+
 # Application
@@
 This group is a fairly large sparse cluster but it is separated from the other points.
-Both layouts are right in separating these points from the others.
+Both layouts are right in separating these points from the others but the fit for layout f is slightly better.
 

### 2025-12-14 paper-quollr.tex
Added setence with linked plots
slight addition to application sentence

@@
 
+In addition to \texttt{langevitour}, linked plots can also be constructed using tour views generated with the \texttt{detourr} package.
+(see \href{https://jayanilakshika.github.io/quollr/articles/quollr8linkeddetourr.html}{vignette: 8.
+Linked plots with \texttt{detourr}} for details.)
+
 \section{Application}\label{application}
@@
 This group is a fairly large sparse cluster but it is separated from the other points.
-Both layouts are right in separating these points from the others.
+Both layouts are right in separating these points from the others but the fit for layout f is slightly better.
 \item

### 2025-12-14 extra/paper-quollr-code.Rmd
Rendered new paper quollr version

---
title: "paper-quollr-code"
format: pdf
---

```{r, fig.cap="Line plot."}
#| label: fig-unemployment

library(autoalt)
library(ggplot2)
library(dplyr)

# Built-in datasets from ggplot2 and base R
data("economics", package = "ggplot2")
data("airquality")

economics |>
  ggplot(aes(x = date, y = unemploy)) +
  geom_line(linewidth = 0.6) +
  labs(
    title = "Total Number of Unemployed Persons in the US Over Time",
    x = "Year",
    y = "Number of Unemployed (thousands)"
  ) +
  theme_minimal()

```

### 2025-12-14 extra/paper-quollr-code.Rmd
Fixed the issues

@@
 ---
-title: "paper-quollr-code"
+title: "Sample quollr"
+author: "Chat-GPT"
+date: "2025-09-10"
 format: pdf
@@
 
-```{r, fig.cap="Line plot."}
-#| label: fig-unemployment
-
-library(autoalt)
+```{r}
+#| label: setup
+#| echo: false
+#| message: false
+#| warning: false
+library(quollr)
+library(tibble)
+library(knitr)
+library(kableExtra)
 library(ggplot2)
 library(dplyr)
+library(patchwork)
+library(readr)
 
-# Built-in datasets from ggplot2 and base R
-data("economics", package = "ggplot2")
-data("airquality")
+# Scale high-d data
 
-economics |>
-  ggplot(aes(x = date, y = unemploy)) +
-  geom_line(linewidth = 0.6) +
-  labs(
-    title = "Total Number of Unemployed Persons in the US Over Time",
-    x = "Year",
-    y = "Number of Unemployed (thousands)"
-  ) +
-  theme_minimal()
+# Center the data by subtracting the mean of each column
+center_data <- function(data) {
+  apply(data, 2, function(col) col - mean(col))
+}
 
+# Function to scale data manually
+scale_data_manual <- function(data, type_col) {
+  # Step 1: Center the data (mean 0)
+  data_centered <- center_data(data |> select(-all_of(type_col)))
+
+  # Step 2: Calculate the standard deviation of each dimension
+  sds <- apply(data_centered, 2, sd)
+
+  # Step 3: Scale each dimension to have the range [0, 1]
+  data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
+
+  # Step 4: Scale dimensions according to their variation
+  # The dimension with the highest standard deviation is scaled to [-1, 1]
+  # Other dimensions are scaled to smaller ranges based on their standard deviations
+  max_sd <- max(sds)
+
+  # Normalize the standard deviations to get scaling factors
+  scaling_factors <- sds / max_sd
+
+  for (i in seq_along(scaling_factors)) {
+    data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
+  }
+
+  # Combine the scaled data with the 'type' column and return as a tibble
+  data_scaled <- as_tibble(data_scaled) %>%
+    mutate(!!type_col := data[[type_col]])
+
+  return(data_scaled)
+}
+
+# creating Standardization function
+standardize = function(x){
+  z <- (x - mean(x)) / sd(x)
+  return( z)
+}
+
+# Solve quadratic function
+
+quad <- function(a = 3, b = 2 * a2, c = -(a2^2 + a1^2))
+{
+  a <- as.complex(a)
+  answer <- c((-b + sqrt(b^2 - 4 * a * c)) / (2 * a),
+              (-b - sqrt(b^2 - 4 * a * c)) / (2 * a))
+  if(all(Im(answer) == 0)) answer <- Re(answer)
+  if(answer[1] == answer[2]) return(answer[1])
+  answer[answer>0] ## only positive
+}
 ```
+
+```{r}
+#| label: fig-overview
+#| fig-cap: "Wireframe model representation of the NLDR layout, lifted and displayed in high-dimensional space.
+The left panel shows the NLDR layout with a triangular mesh overlay, forming the wireframe structure.
+This mesh can be lifted into higher dimensions and projected to examine how the geometric structure of the data is preserved.
+Panels (a1–a4) display different 2D projections of the lifted wireframe, where the underlying curved sheet structure of the data is more clearly visible.
+The triangulated mesh highlights how local neighborhoods in the layout correspond to relationships in the high-dimensional space, enabling diagnostics of distortion and preservation across dimensions."
+#| echo: false
+# Plot MSE
+
+plot_hbe <- function(error_df) {
+
+  ggplot(error_df,
+         aes(x = a1,
+             y = HBE,
+             colour = method)) +
+    geom_point(size = 0.8) +
+    geom_line(linewidth = 0.3) +
+    ylab("HBE") +
+    xlab(expression(paste("binwidth (", a[1], ")"))) +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
+          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+          axis.ticks.x = element_line(),
+          axis.ticks.y = element_line(),
+          legend.position = "none",
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7),
+          axis.title.x = element_text(size = 7),
+          axis.title.y = element_text(size = 7),
+          plot.margin = margin(0, 0, 0, 0))
+
+}
+
+set.seed(20240110)
+
+clr_choice <- "#66B2CC"
+
+scurve_model_obj <- fit_highd_model(
+  highd_data = scurve, 
+  nldr_data = scurve_umap, 
+  b1 = 21, 
+  q = 0.1, 
+  hd_thresh = 0)
+
+scurve_umap_scaled <- scurve_model_obj$nldr_scaled_obj$scaled_nldr
+tr_from_to_df_scurve <- scurve_model_obj$trimesh_data
+df_bin_centroids_scurve <- scurve_model_obj$model_2d
+df_bin_scurve <- scurve_model_obj$model_highd
+hex_grid_scurve <- scurve_model_obj$hb_obj$hex_poly
+counts_df_scurve <- scurve_model_obj$hb_obj$std_cts
+
+hex_grid_with_counts_scurve <- left_join(
+  hex_grid_scurve, counts_df_scurve, 
+  by = c("h" = "h"))
+
+hex_grid_nonempty_scurve <- hex_grid_scurve |>
+  filter(h %in% df_bin_centroids_scurve$h)
+
+sc_xlims <- c(-0.25, 1.35)
+sc_ylims <- c(-0.25, 1.35)
+
+scurve_umap_plt <- ggplot(
+  scurve_umap_scaled, 
+  aes(x = emb1, y = emb2)) +
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5) +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
+
+hex_grid_poly_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = h), 
+               fill = "#ffffff") +
+  geom_point(data = scurve_umap_scaled, 
+             aes(x = emb1, y = emb2), 
+             alpha = 0.8, size = 0.5, color = clr_choice) + 
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
+
+hex_centroids_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "grey70", 
+               aes(group = h), 
+               fill = "#ffffff")  +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#FF7755") +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
+
+wireframe_scurve <- ggplot() +
+  geom_segment(data = tr_from_to_df_scurve,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#FF7755") +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
+
+## 2-d vis model
+hex_grid_scurve <- ggplot(
+  data = hex_grid_with_counts_scurve, 
+  aes(x = x, y = y)) +
+  geom_point(data = scurve_umap_scaled, 
+             aes(x = emb1, y = emb2), 
+             size = 0.5, color = clr_choice) +
+  geom_segment(data = tr_from_to_df_scurve,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = df_bin_centroids_scurve, 
+             aes(x = c_x, y = c_y), 
+             size = 1, color = "#FF7755")
+
+## High-d vis model
+df_exe <- comb_data_model(
+    highd_data = scurve, 
+    model_highd = df_bin_scurve, 
+    model_2d = df_bin_centroids_scurve
+)
+
+df <- df_exe |>
+    dplyr::filter(type == "data") ## original dataset
+
+df_b <- df_exe |>
+  dplyr::filter(type == "model") ## High-d model
+
+scurve_umap_model_vis <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = tr_from_to_df_scurve$from_reindexed,
+                         lineTo = tr_from_to_df_scurve$to_reindexed,
+                         group = df_exe$type,
+                         pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
+                         levelColors = c(clr_choice, "#FF7755"),
+                         enableControls = FALSE,
+                         width = "400px", height = "380px")
+
+scurve_umap_model_vis_n <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+                         lineFrom = tr_from_to_df_scurve$from_reindexed,
+                         lineTo = tr_from_to_df_scurve$to_reindexed,
+                         group = df_exe$type,
+                         pointSize = append(rep(2, NROW(df_b)), rep(1, NROW(df))),
+                         levelColors = c(clr_choice, "#FF7755"),
+                         enableControls = FALSE,
+                         width = "400px", height = "550px")
+
+df_b_scurve <- df_bin_scurve |>
+  dplyr::filter(h %in% df_bin_centroids_scurve$h) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in df_b_with_center_data
+df_b_scurve <- df_b_scurve[match(df_bin_centroids_scurve$h, df_b_scurve$h),] |>
+  dplyr::select(-h)
+
+scurve_labeled <- scurve |>
+  select(-ID) |>
+  mutate(type = "data")
+
+# Apply the scaling
+df_model_data_scurve <- bind_rows(scurve_labeled, df_b_scurve)
+scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
+  as_tibble()
+
+scaled_scurve_data <- scaled_scurve |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_scurve_data_model <- scaled_scurve |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(0.10479,0.06673,0.19430,-0.14763,0.02861,-0.04302,0.06601),
+  c(0.11421,0.18556,-0.16860,-0.05033,-0.00219,0.04159,0.04285))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.0259))
+
+scurve_proj_umap_model1 <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 2,
+  is_category = FALSE)
+
+scurve_proj_umap_model1_dp <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(0.2, 0, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 3,
+  is_category = FALSE)
+
+## Second projection
+
+projection <- cbind(
+  c(0.01588,-0.13640,-0.17180,-0.15486,-0.09726,0.00858,0.01090),
+  c(0.21443,-0.00390,0.08182,-0.11746,0.06820,-0.05458,-0.08885))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.0259))
+
+scurve_proj_umap_model2 <- plot_proj(
+  proj_obj = proj_obj2,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 2,
+  is_category = FALSE)
+
+## Third projection
+
+projection <- cbind(
+  c(0.02727,0.09154,-0.16383,-0.04931,-0.17013,0.11965,-0.01884),
+  c(-0.10510,-0.10971,-0.09084,0.19642,-0.07163,-0.02486,0.07965))
+
+proj_obj3 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.031))
+
+scurve_proj_umap_model3 <- plot_proj(
+  proj_obj = proj_obj3,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 2,
+  is_category = FALSE)
+
+## Fourth projection
+
+projection <- cbind(
+  c(0.17174,0.14255,0.11950,0.10344,-0.07148,-0.01656,0.04275),
+  c(0.09270,0.04486,-0.14397,0.08305,0.11783,0.16187,-0.06077))
+
+proj_obj4 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.25, 
+                                              axis_pos_y = -0.25, 
+                                              threshold = 0.031))
+
+scurve_proj_umap_model4 <- plot_proj(
+  proj_obj = proj_obj4,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  plot_limits = c(-0.35, 0.35), 
+  axis_text_size = 2,
+  is_category = FALSE)
+
+hex_grid_scurve + wrap_plots(
+  scurve_proj_umap_model1, scurve_proj_umap_model2,
+  scurve_proj_umap_model3, scurve_proj_umap_model4, 
+  ncol = 2)
+```
+
+# Introduction
+
+This report explores two publicly available datasets: economics from the ggplot2 package and the classic airquality dataset from base R.
+The economics dataset provides monthly time-series data on the US economy, including variables such as unemployment and personal consumption expenditure.
+The airquality dataset records daily air quality measurements in New York, including ozone concentration, solar radiation, and temperature.
+
+We use a selection of visualisations to highlight long-term economic trends and seasonal patterns in air quality.
+
+This paper presents the R package, `quollr`, which is useful for understanding how NLDR warps high-dimensional space and fits the data.
+Starting with an NLDR layout, our approach is to create a $2\text{-}D$ wireframe model representation, that can be lifted and displayed in the high-dimensional @fig-overview.

### 2025-12-14 paper-quollr.Rmd
Fixed the issues

@@
 
-which produces the model and data plot(s) shown in `r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`.
+which produces the model and data plot shown in `r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:algo-step-html)"} else { "Figure \\@ref(fig:algo-steps)"})`.
 
@@
 
+In the implementation examples, points are shown without cluster-based colouring unless explicitly stated.
+When points are coloured by cluster, linked plot functionality is currently only partially supported: selections made in the `langevitour` controls are not reflected in the corresponding highlights in the interactive $2\text{-}D$ layout.
+As a result, cluster-specific exploration must be carried out separately in the tour view and the interactive $2\text{-}D$ layout.
 In addition to `langevitour`, linked plots can also be constructed using tour views generated with the `detourr` package.
@@
 Additionally, statistical inference tools could be introduced to assess the stability and robustness of the fitted model, which would enhance interpretability and confidence in the outcomes.
-Currently, linked plot functionality is limited: selecting a cluster in the `langevitour` controls does not yet highlight the corresponding cluster in the $2\text{-}D$ layout.
-Implementing this feature in the future would allow users to clearly explore cluster-specific relationships between the `langevitour` and $2\text{-}D$ layout.
+A potential extension of the current implementation would be to synchronize cluster selections between the tour view and the linked $2\text{-}D$ layout, enabling more direct cluster-specific comparisons across views.
 

### 2025-12-14 paper-quollr.tex
Fixed the issues

@@
 
-which produces the model and data plot(s) shown in Figure \ref{fig:algo-steps}.
+which produces the model and data plot shown in Figure \ref{fig:algo-steps}.
 
@@
 
+In the implementation examples, points are shown without cluster-based colouring unless explicitly stated.
+When points are coloured by cluster, linked plot functionality is currently only partially supported: selections made in the \texttt{langevitour} controls are not reflected in the corresponding highlights in the interactive \(2\text{-}D\) layout.
+As a result, cluster-specific exploration must be carried out separately in the tour view and the interactive \(2\text{-}D\) layout.
 In addition to \texttt{langevitour}, linked plots can also be constructed using tour views generated with the \texttt{detourr} package.
@@
 Additionally, statistical inference tools could be introduced to assess the stability and robustness of the fitted model, which would enhance interpretability and confidence in the outcomes.
-Currently, linked plot functionality is limited: selecting a cluster in the \texttt{langevitour} controls does not yet highlight the corresponding cluster in the \(2\text{-}D\) layout.
-Implementing this feature in the future would allow users to clearly explore cluster-specific relationships between the \texttt{langevitour} and \(2\text{-}D\) layout.
+A potential extension of the current implementation would be to synchronize cluster selections between the tour view and the linked \(2\text{-}D\) layout, enabling more direct cluster-specific comparisons across views.
 

### 2025-12-15 paper-quollr.Rmd
started fixing the projections

@@
 df_model_data_scurve <- bind_rows(scurve_labeled, df_b_scurve)
-scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
-  as_tibble()
+scaled_scurve <- center_data(df_model_data_scurve[,-8]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_scurve$type)
 
@@
 df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
-scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
-  as_tibble()
+scaled_limb <- center_data(df_model_data_limb[,-11]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_limb$type)
 

### 2025-12-15 paper-quollr.R

@@ -383,8 +383,9 @@ scurve_labeled <- scurve |>
 
 # Apply the scaling
 df_model_data_scurve <- bind_rows(scurve_labeled, df_b_scurve)
-scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
-  as_tibble()
+scaled_scurve <- center_data(df_model_data_scurve[,-8]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_scurve$type)
 
 scaled_scurve_data <- scaled_scurve |>
   filter(type == "data") |>
@@ -1384,8 +1385,9 @@ df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
 
 # Apply the scaling
 df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
-scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
-  as_tibble()
+scaled_limb <- center_data(df_model_data_limb[,-11]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_limb$type)
 
 scaled_limb_data <- scaled_limb |>
   filter(type == "data") |>
@@ -1736,8 +1738,9 @@ df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
 
 # Apply the scaling
 df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
-scaled_limb <- scale_data_manual(df_model_data_limb, "type") |>
-  as_tibble()
+scaled_limb <- center_data(df_model_data_limb[,-11]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_limb$type)
 
 scaled_limb_data <- scaled_limb |>
   filter(type == "data") |>

### 2025-12-17 paper-quollr.Rmd
Updated Figure 1 and 3

@@
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.8, 
+                                              axis_scaled = 3,
+                                              axis_pos_x = -0.35, 
+                                              axis_pos_y = -0.35, 
+                                              threshold = 0.0623))
+
+scurve_proj_umap_model1 <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-0.5, 0.55), 
+  axis_text_size = 2,
+  is_category = FALSE) +
+  interior_annotation(label = "a1", cex = 1)
+
+scurve_proj_umap_model1_dp <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(0.2, 0, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-0.5, 0.55), 
+  axis_text_size = 3,
+  is_category = FALSE) +
+  interior_annotation(label = "d", cex = 1)
+
+## Second projection
+
+projection <- cbind(
+  c(0.01588,-0.13640,-0.17180,-0.15486,-0.09726,0.00858,0.01090),
+  c(0.21443,-0.00390,0.08182,-0.11746,0.06820,-0.05458,-0.08885))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.6, 
+                                              axis_scaled = 3,
+                                              axis_pos_x = -0.35, 
+                                              axis_pos_y = -0.35, 
+                                              threshold = 0.05))
+
+scurve_proj_umap_model2 <- plot_proj(
+  proj_obj = proj_obj2,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-0.45, 0.4), 
+  axis_text_size = 2,
+  is_category = FALSE) +
+  interior_annotation(label = "a2", cex = 1)
+
+## Third projection
+
+projection <- cbind(
+  c(0.02727,0.09154,-0.16383,-0.04931,-0.17013,0.11965,-0.01884),
+  c(-0.10510,-0.10971,-0.09084,0.19642,-0.07163,-0.02486,0.07965))
+
+proj_obj3 <- get_projection(projection = projection, 
                             highd_data = scaled_scurve_data, 
@@
                             axis_param = list(limits = 0.5, 
-                                              axis_scaled = 2,
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.0259))
-
-scurve_proj_umap_model1 <- plot_proj(
-  proj_obj = proj_obj1,
+                                              axis_scaled = 3,
+                                              axis_pos_x = -0.34, 
+                                              axis_pos_y = -0.34, 
+                                              threshold = 0.0462))
+
+scurve_proj_umap_model3 <- plot_proj(
+  proj_obj = proj_obj3,
   point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
-  plot_limits = c(-0.35, 0.35), 
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-0.43, 0.35), 
   axis_text_size = 2,
   is_category = FALSE) +
-  interior_annotation(label = "a1", cex = 1)
-
-scurve_proj_umap_model1_dp <- plot_proj(
-  proj_obj = proj_obj1,
-  point_param = c(0.2, 0, clr_choice), # size, alpha, color
-  plot_limits = c(-0.35, 0.35), 
-  axis_text_size = 3,
-  is_category = FALSE) +
-  interior_annotation(label = "d", cex = 1)
-
-## Second projection
+  interior_annotation(label = "a3", cex = 1)
+
+## Fourth projection
 
 projection <- cbind(
-  c(0.01588,-0.13640,-0.17180,-0.15486,-0.09726,0.00858,0.01090),
-  c(0.21443,-0.00390,0.08182,-0.11746,0.06820,-0.05458,-0.08885))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
+  c(0.17174,0.14255,0.11950,0.10344,-0.07148,-0.01656,0.04275),
+  c(0.09270,0.04486,-0.14397,0.08305,0.11783,0.16187,-0.06077))
+
+proj_obj4 <- get_projection(projection = projection, 
                             highd_data = scaled_scurve_data, 
@@
                             trimesh_data = tr_from_to_df_scurve, 
-                            axis_param = list(limits = 0.5, 
-                                              axis_scaled = 2,
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.0259))
-
-scurve_proj_umap_model2 <- plot_proj(
-  proj_obj = proj_obj2,
-  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
-  plot_limits = c(-0.35, 0.35), 
-  axis_text_size = 2,
-  is_category = FALSE) +
-  interior_annotation(label = "a2", cex = 1)
-
-## Third projection
-
-projection <- cbind(
-  c(0.02727,0.09154,-0.16383,-0.04931,-0.17013,0.11965,-0.01884),
-  c(-0.10510,-0.10971,-0.09084,0.19642,-0.07163,-0.02486,0.07965))
-
-proj_obj3 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
-                            highd_data = scaled_scurve_data, 
-                            model_highd = scaled_scurve_data_model, 
-                            trimesh_data = tr_from_to_df_scurve, 
-                            axis_param = list(limits = 0.5, 
-                                              axis_scaled = 2,
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.031))
-
-scurve_proj_umap_model3 <- plot_proj(
-  proj_obj = proj_obj3,
-  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
-  plot_limits = c(-0.35, 0.35), 
-  axis_text_size = 2,
-  is_category = FALSE) +
-  interior_annotation(label = "a3", cex = 1)
-
-## Fourth projection
-
-projection <- cbind(
-  c(0.17174,0.14255,0.11950,0.10344,-0.07148,-0.01656,0.04275),
-  c(0.09270,0.04486,-0.14397,0.08305,0.11783,0.16187,-0.06077))
-
-proj_obj4 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
-                            highd_data = scaled_scurve_data, 
-                            model_highd = scaled_scurve_data_model, 
-                            trimesh_data = tr_from_to_df_scurve, 
-                            axis_param = list(limits = 0.5, 
-                                              axis_scaled = 2,
-                                              axis_pos_x = -0.25, 
-                                              axis_pos_y = -0.25, 
-                                              threshold = 0.031))
+                            axis_param = list(limits = 0.6, 
+                                              axis_scaled = 3,
+                                              axis_pos_x = -0.35, 
+                                              axis_pos_y = -0.35, 
+                                              threshold = 0.048))
 
@@
   point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
-  plot_limits = c(-0.35, 0.35), 
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-0.45, 0.46), 
   axis_text_size = 2,
@@
 proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
                             highd_data = scaled_limb_data, 
@@
 proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
                             highd_data = scaled_limb_data, 

### 2025-12-17 paper-quollr.Rmd
Updated Linked plot projections

@@
 
+proj_obj2_cp <- get_projection(projection = projection, 
+                            highd_data = scaled_scurve_data, 
+                            model_highd = scaled_scurve_data_model, 
+                            trimesh_data = tr_from_to_df_scurve, 
+                            axis_param = list(limits = 0.6, 
+                                              axis_scaled = 3,
+                                              axis_pos_x = -0.4, 
+                                              axis_pos_y = -0.4, 
+                                              threshold = 0.05))
+
 scurve_proj_umap_model2 <- plot_proj(
@@
 
-proj_obj2[["cluster"]] <- factor(scurve_umap_scaled_select1$select_area,
+proj_obj2_cp[["cluster"]] <- factor(scurve_umap_scaled_select1$select_area,
                                  levels=c("deselected", "selected"))
 
-projected_df <- proj_obj2$projected_df
-model_df <- proj_obj2$model_df
-axes <- proj_obj2$axes
-circle <- proj_obj2$circle
+projected_df <- proj_obj2_cp$projected_df
+model_df <- proj_obj2_cp$model_df
+axes <- proj_obj2_cp$axes
+circle <- proj_obj2_cp$circle
 
 projected_df <- projected_df |>
-  dplyr::mutate(cluster = proj_obj2$cluster)
+  dplyr::mutate(cluster = proj_obj2_cp$cluster)
 
@@
         colour = "#000000",
-        linewidth = 0.8,
-        alpha = 0.4) +
+        linewidth = 0.5,
+        alpha = 0.3) +
       geom_segment(
@@
           aes(x=c1, y=c2), colour="grey70") +
-        xlim(c(-0.35, 0.35)) +
-        ylim(c(-0.35, 0.35)) +
+        xlim(c(-0.5, 0.47)) +
+        ylim(c(-0.5, 0.47)) +
   interior_annotation(label = "a2", cex = 3) +
@@
           aes(x=c1, y=c2), colour="grey70") +
-        xlim(c(-0.35, 0.35)) +
-        ylim(c(-0.35, 0.35)) +
+        xlim(c(-0.5, 0.47)) +
+        ylim(c(-0.5, 0.47)) +
   interior_annotation(label = "a3", cex = 3) +
@@
 
-proj_obj2[["cluster"]] <- factor(scurve_umap_scaled_select2$select_area,
+proj_obj2_cp[["cluster"]] <- factor(scurve_umap_scaled_select2$select_area,
                                  levels=c("deselected", "selected"))
 
-projected_df <- proj_obj2$projected_df
-model_df <- proj_obj2$model_df
-axes <- proj_obj2$axes
-circle <- proj_obj2$circle
+projected_df <- proj_obj2_cp$projected_df
+model_df <- proj_obj2_cp$model_df
+axes <- proj_obj2_cp$axes
+circle <- proj_obj2_cp$circle
 
 projected_df <- projected_df |>
-  dplyr::mutate(cluster = proj_obj2$cluster)
+  dplyr::mutate(cluster = proj_obj2_cp$cluster)
 
@@
           aes(x=c1, y=c2), colour="grey70") +
-        xlim(c(-0.35, 0.35)) +
-        ylim(c(-0.35, 0.35)) +
+        xlim(c(-0.5, 0.47)) +
+        ylim(c(-0.5, 0.47)) +
   interior_annotation(label = "b2", cex = 3) +
@@
           aes(x=c1, y=c2), colour="grey70") +
-        xlim(c(-0.35, 0.35)) +
-        ylim(c(-0.35, 0.35)) +
+        xlim(c(-0.5, 0.47)) +
+        ylim(c(-0.5, 0.47)) +
   interior_annotation(label = "b3", cex = 3) +

### 2025-12-17 paper-quollr.Rmd
Updated application projections

@@
                             trimesh_data = trimesh_data_limb, 
-                            axis_param = list(limits = 0.05,
+                            axis_param = list(limits = 1.5,
                                               axis_scaled = 20, 
-                                              axis_pos_x = -0.04, 
-                                              axis_pos_y = -0.04, 
-                                              threshold = 0.004))
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.13))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.05, 0.03), 
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-1.3, 1.4), 
   axis_text_size = 4,
@@
                             trimesh_data = trimesh_data_limb, 
-                            axis_param = list(limits = 0.05,
+                            axis_param = list(limits = 1.5,
                                               axis_scaled = 20, 
-                                              axis_pos_x = -0.03, 
-                                              axis_pos_y = -0.03, 
-                                              threshold = 0.004))
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.14))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.04, 0.04), 
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-1.3, 1.3), 
   axis_text_size = 4,
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.05, 0.03), 
+  plot_limits = c(-1.3, 1.4), 
   axis_text_size = 4,
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.04, 0.04), 
+  plot_limits = c(-1.3, 1.3), 
   axis_text_size = 4,

### 2025-12-17 paper-quollr.R

@@ -1532,18 +1532,19 @@ proj_obj1 <- get_projection(projection = projection,
                             highd_data = scaled_limb_data, 
                             model_highd = scaled_limb_data_model, 
                             trimesh_data = trimesh_data_limb, 
-                            axis_param = list(limits = 0.05,
+                            axis_param = list(limits = 1.5,
                                               axis_scaled = 20, 
-                                              axis_pos_x = -0.04, 
-                                              axis_pos_y = -0.04, 
-                                              threshold = 0.004))
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.13))
 
 proj_obj1[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
 
 limb_proj_tsne_model1 <- plot_proj(
   proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.05, 0.03), 
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-1.3, 1.4), 
   axis_text_size = 4,
   is_category = TRUE) +
   interior_annotation(label = "a2", cex = 2) + 
@@ -1558,18 +1559,19 @@ proj_obj2 <- get_projection(projection = projection,
                             highd_data = scaled_limb_data, 
                             model_highd = scaled_limb_data_model, 
                             trimesh_data = trimesh_data_limb, 
-                            axis_param = list(limits = 0.05,
+                            axis_param = list(limits = 1.5,
                                               axis_scaled = 20, 
-                                              axis_pos_x = -0.03, 
-                                              axis_pos_y = -0.03, 
-                                              threshold = 0.004))
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.14))
 
 proj_obj2[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
 
 limb_proj_tsne_model2 <- plot_proj(
   proj_obj = proj_obj2, 
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.04, 0.04), 
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-1.3, 1.3), 
   axis_text_size = 4,
   is_category = TRUE) +
   interior_annotation(label = "a3", cex = 2) + 
[…]

### 2025-12-17 paper-quollr.Rmd
Updated application cluster color

@@
                linewidth = 1) +
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3')) +
   interior_annotation("a1", cex = 2) +
@@
     "0" = '#66c2a5', "1" = '#fc8d62', "2" = '#8da0cb',
-    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#e5c494'
+    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#a6cee3'
 )
@@
 # Add line segments
-if (nrow(trimesh_data_limb) > 0) {
 nldr_plt <- nldr_plt |>
@@
     )
-}
 
@@
                                                                   rep(point_sizes[2], NROW(df))),
-                                               levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494', "#000000"),
+                                               levelColors = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3', "#000000"),
                                                link=shared_df,
@@
   interior_annotation(label = "a2", cex = 2) + 
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
 
@@
   interior_annotation(label = "a3", cex = 2) + 
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
 
@@
                linewidth = 1) +
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3')) +
   interior_annotation("f1", cex = 2) +
@@
   interior_annotation(label = "f2", cex = 2) + 
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
 
@@
   interior_annotation(label = "f3", cex = 2) + 
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494'))
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
 

### 2025-12-17 paper-quollr.R

@@ -1372,7 +1372,7 @@ trimesh_limb <- ggplot() +
                  yend = y_to),
                colour = "#000000",
                linewidth = 1) +
-  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#e5c494')) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3')) +
   interior_annotation("a1", cex = 2) +
   theme(
     aspect.ratio = 1
@@ -1440,7 +1440,7 @@ shared_df <- crosstalk::SharedData$new(point_data)
 
 cluster_colors <- c(
     "0" = '#66c2a5', "1" = '#fc8d62', "2" = '#8da0cb',
-    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#e5c494'
+    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#a6cee3'
 )
 
 point_sizes <- c(0, 1)
@@ -1457,18 +1457,16 @@ nldr_plt <- plot_ly(
 )
 
 # Add line segments
-if (nrow(trimesh_data_limb) > 0) {
-    nldr_plt <- nldr_plt |>
-        add_segments(
-            data = trimesh_data_limb,
-            x = ~x_from, y = ~y_from,
-            xend = ~x_to, yend = ~y_to,
-            line = list(color = "#000000", width = 0.5),
-            inherit = FALSE,
-            showlegend = FALSE,
-            hoverinfo = "skip"  # <-- skip hover
-        )
-}
+nldr_plt <- nldr_plt |>
+    add_segments(
+        data = trimesh_data_limb,
+        x = ~x_from, y = ~y_from,
+        xend = ~x_to, yend = ~y_to,
+        line = list(color = "#000000", width = 0.5),
+        inherit = FALSE,
+        showlegend = FALSE,
+        hoverinfo = "skip"  # <-- skip hover
+    )
 
 # Layout: axes, border, no tooltip
 nldr_plt <- nldr_plt |>
@@ -1498,7 +1496,7 @@ langevitour_output <- langevitour::langevitour(point_data[1:num_highd_col],
                                                               c("0", "1", "2", "3", "4", "5", "6", "model")),
                                                pointSize = append(rep(point_sizes[1], NROW(df_b)),
[…]

### 2025-12-17 paper-quollr.Rmd
Updated trimesh color in application

@@
                colour = "#000000",
-               linewidth = 1) +
+               linewidth = 1,
+               alpha = 0.3) +
   scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3')) +
@@
   )
-```
-
-```{r}
-#| label: prep-limb-tsne-author-model-proj
-
-data_limb_n <- data_limb |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-df_b_limb <- df_bin_limb |>
-  dplyr::filter(h %in% df_bin_centroids_limb$h) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the h order in df_b_with_center_data
-df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
-  dplyr::select(-h)
-
-# Apply the scaling
-df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
-scaled_limb <- center_data(df_model_data_limb[,-11]) |>
-  as_tibble() |>
-  mutate(type = df_model_data_limb$type)
-
-scaled_limb_data <- scaled_limb |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_limb_data_model <- scaled_limb |>
-  filter(type == "model") |>
-  select(-type)
 ```
@@
 
+```{r}
+#| label: prep-limb-tsne-author-model-proj
+
+data_limb_n <- data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(h %in% df_bin_centroids_limb$h) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
+  dplyr::select(-h)
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
+scaled_limb <- center_data(df_model_data_limb[,-11]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_limb$type)
+
+scaled_limb_data <- scaled_limb |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_limb_data_model <- scaled_limb |>
+  filter(type == "model") |>
+  select(-type)
+```
+
 ```{r plot-proj}
@@
 3.
-To examine where the layouts differ, examine clusters $4$ (green) and $6$ (cream), by selecting just these two.
+To examine where the layouts differ, examine clusters $4$ (green) and $6$ (blue), by selecting just these two.
 Layout a has them close, but layout f has them far apart.

### 2025-12-17 paper-quollr.R

@@ -1371,42 +1371,13 @@ trimesh_limb <- ggplot() +
                  xend = x_to, 
                  yend = y_to),
                colour = "#000000",
-               linewidth = 1) +
+               linewidth = 1,
+               alpha = 0.3) +
   scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3')) +
   interior_annotation("a1", cex = 2) +
   theme(
     aspect.ratio = 1
   )
-
-
-## -----------------------------------------------------------------------------
-#| label: prep-limb-tsne-author-model-proj
-
-data_limb_n <- data_limb |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-df_b_limb <- df_bin_limb |>
-  dplyr::filter(h %in% df_bin_centroids_limb$h) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the h order in df_b_with_center_data
-df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
-  dplyr::select(-h) 
-
-# Apply the scaling
-df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
-scaled_limb <- center_data(df_model_data_limb[,-11]) |>
-  as_tibble() |>
-  mutate(type = df_model_data_limb$type)
-
-scaled_limb_data <- scaled_limb |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_limb_data_model <- scaled_limb |>
-  filter(type == "model") |>
-  select(-type)
 
 
 ## -----------------------------------------------------------------------------
@@ -1519,6 +1490,36 @@ linked_plt_int <- crosstalk::bscols(
 
 
 
+## -----------------------------------------------------------------------------
+#| label: prep-limb-tsne-author-model-proj
+
+data_limb_n <- data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(h %in% df_bin_centroids_limb$h) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
+  dplyr::select(-h) 
+
[…]

### 2025-12-17 paper-quollr.Rmd
Started adding alt text

@@
 Layout d performs well at large binwidth (where the binwidth is not enough to capture the data structure) and poorly as the bin width decreases.
-Layout f is the best choice.\\label{fig:limb-hbe}", fig.alt = "Evaluation of six NLDR layouts (a–f) on limb muscle data using HBE across different binwidths (a1).
-Line and point colors correspond to layouts.
-Layout d performs well at large binwidths but poorly as binwidth decreases, while layout f consistently provides the best representation.", fig.pos="!ht", echo=TRUE}
+Layout f is the best choice.\\label{fig:limb-hbe}", fig.alt = "A multi-panel figure shows five alternative 2D embeddings of the same high-dimensional data, arranged side by side (panels b–f).
+Each panel is a scatterplot with two unlabeled embedding axes (dimension 1 on the horizontal axis and dimension 2 on the vertical axis, each spanning a similar, roughly symmetric range around zero).
+Points represent the same set of observations in every panel, plotted with identical colour or marker scheme so that corresponding observations can be visually compared across layouts.
+Panel b shows a UMAP layout (neighbors = 5, minimum distance = 0.1), panel c a PHATE layout (knn = 5), panel d a TriMAP layout (inliers = 12, outliers = 4, random = 3), panel e a PaCMAP layout (neighbors = 10, init = random, MN-ratio = 0.5, FP-ratio = 2), and panel f a tSNE layout (perplexity = 15).
+Across panels, similar clusters of points appear but are arranged differently in space: some layouts show more clearly separated compact clusters, while others show overlapping or elongated groups, reflecting how each method balances local versus global structure.
+The accompanying HBE plot (typically a separate panel) summarizes model fit quality for each layout, with one value per method indicating that some embeddings provide a more faithful representation of the original data structure than others.", fig.pos="!ht", echo=TRUE}
 

### 2025-12-17 paper-quollr.tex
Updated trimesh color in application
Started adding alt text

@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Evaluation of six NLDR layouts (a–f) on limb muscle data using HBE across different binwidths (a1).
-Line and point colors correspond to layouts.
-Layout d performs well at large binwidths but poorly as binwidth decreases, while layout f consistently provides the best representation.}]{paper-quollr_files/figure-latex/limb-hbe-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
+\includegraphics[width=1\linewidth,alt={A multi-panel figure shows five alternative 2D embeddings of the same high-dimensional data, arranged side by side (panels b–f).
+Each panel is a scatterplot with two unlabeled embedding axes (dimension 1 on the horizontal axis and dimension 2 on the vertical axis, each spanning a similar, roughly symmetric range around zero).
+Points represent the same set of observations in every panel, plotted with identical colour or marker scheme so that corresponding observations can be visually compared across layouts.
+Panel b shows a UMAP layout (neighbors = 5, minimum distance = 0.1), panel c a PHATE layout (knn = 5), panel d a TriMAP layout (inliers = 12, outliers = 4, random = 3), panel e a PaCMAP layout (neighbors = 10, init = random, MN-ratio = 0.5, FP-ratio = 2), and panel f a tSNE layout (perplexity = 15).
+Across panels, similar clusters of points appear but are arranged differently in space: some layouts show more clearly separated compact clusters, while others show overlapping or elongated groups, reflecting how each method balances local versus global structure.
+The accompanying HBE plot (typically a separate panel) summarizes model fit quality for each layout, with one value per method indicating that some embeddings provide a more faithful representation of the original data structure than others.}]{paper-quollr_files/figure-latex/limb-hbe-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
@@
 \item
-  To examine where the layouts differ, examine clusters \(4\) (green) and \(6\) (cream), by selecting just these two.
+  To examine where the layouts differ, examine clusters \(4\) (green) and \(6\) (blue), by selecting just these two.
 Layout a has them close, but layout f has them far apart.

### 2025-12-19 paper-quollr.Rmd
Added some alt-text

@@
 Centroids are computed and triangulated to form the $2\\text{-}D$ mesh, which is then lifted into the $p\\text{-}D$ space.
-Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\\text{-}D$ and $2\\text{-}D$ representations.", fig.alt = "Flowchart diagram illustrating the quollr workflow.
-It shows inputs of NLDR and high-dimensional data progressing through preprocessing, hexagonal binning, centroid computation, and triangulation to form a 2-D mesh.
-The mesh is lifted into the high-dimensional space for prediction and error analysis, with interactive tools linking the high- and low-dimensional representations."}
+Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\\text{-}D$ and $2\\text{-}D$ representations.", fig.alt = "The figure is a schematic diagram illustrating the quollr workflow and software architecture.
+Rectangular boxes represent processing steps and data objects, including NLDR and p-D data inputs, data preprocessing, hexagonal binning, centroid computation, triangulation, and construction of a 2-D mesh.
+Arrows indicate the flow of information between steps, from the input data through the 2-D mesh, and then from the mesh being lifted into the p-D space.
+Additional boxes show prediction and error computation for new data, as well as interactive components that link the p-D and 2-D representations.
+The diagram is laid out sequentially, showing how data move through the system from input to interactive output."}
+
+# # GitHub raw URL of the figure
+# fig_url <- "https://raw.githubusercontent.com/JayaniLakshika/paper-quollr/main/figures/quollr_workflow.png"
+# 
+# # Determine figure path depending on output
+# if (knitr::is_html_output()) {
+#   # HTML can use the URL directly
+#   fig_path <- fig_url
+# } else {
+#   # PDF needs a local copy
+#   temp_file <- tempfile(fileext = ".png")
+#   download.file(fig_url, destfile = temp_file, mode = "wb")
+#   fig_path <- temp_file
+# }
+# 
+# knitr::include_graphics(fig_path)
 
@@
 ```{r algo-step-html, eval=knitr::is_html_output(), fig.pos="!ht", fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The `scurve` data is shown.", layout = "l-body", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
+The `scurve` data is shown.", layout = "l-body", fig.alt = "The figure consists of four panels illustrating steps in constructing a model on a UMAP layout using the scurve data.
+Panel (a) is a static 2-D scatter plot of the UMAP embedding with points grouped into hexagonal bins.
+Panel (b) shows the same 2-D layout with a single centroid point displayed for each hexagonal bin.
+Panel (c) shows the centroids connected by straight line segments to form a triangulated mesh across the 2-D layout.
+Panel (d) is atour view in which the triangulated 2-D mesh is lifted into the original high-dimensional space and displayed as a continuously changing 2-D projection; the axes represent linear combinations of the original dimensions, and the mesh rotates and changes orientation as the view updates.
+The 2-D panels use bounded numeric axes with roughly square aspect ratios, while the tour view shows the lifted structure from multiple viewing angles."}
 
@@
 ```{r algo-steps, eval=knitr::is_latex_output(), out.width="80%", fig.height=5, fig.width=25, fig.pos="!ht", fig.align='center', fig.cap="Key steps for constructing the model on the UMAP layout: (a) hexagon bins, (b) bin centroids, (c) triangulated centroids, and (d) lifting the model into high dimensions.
-The \\texttt{scurve} data is shown.", fig.alt = "Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space."}
+The \\texttt{scurve} data is shown.", fig.alt = "The figure is a multi-panel diagram illustrating four steps in constructing a model on a UMAP layout using the scurve data.
+Panel (a) shows a 2-D UMAP scatter plot with points grouped into hexagonal bins covering the layout.
+Panel (b) shows the same layout with a centroid point displayed for each hexagon bin.
+Panel (c) shows the centroids connected by straight line segments to form a triangulated mesh across the 2-D layout.
+Panel (d) shows this 2-D triangulated mesh lifted into the original high-dimensional space, where the mesh appears as a curved surface.
+The horizontal and vertical axes in the 2-D panels represent the two UMAP coordinates with bounded numeric ranges, while the lifted view represents coordinates in the higher-dimensional space without explicit axis scales."}
 
@@
 
-```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_html_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
+```{r geom-outputs, fig.cap="The outputs of `geom_hexgrid` and `geom_trimesh` include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_html_output(), fig.alt = "The figure consists of four 2-D panels illustrating hexagonal grids and triangulated meshes.
+Panel (a) shows a complete hexagonal grid covering a rectangular 2-D region, with all hexagons drawn regardless of whether data are present.
+Panel (b) shows a hexagonal grid restricted to bins that contain data points, with only those hexagons displayed and gaps where bins are empty.
+Panel (c) shows a full triangulated mesh constructed from centroids across the entire grid region, with centroids connected by straight line segments to form triangles.
+Panel (d) shows a triangulated mesh constructed only from centroids corresponding to non-empty bins, aligned with the data layout and leaving open areas where no centroids are present.
+In all panels, the horizontal and vertical axes represent continuous 2-D coordinates with bounded numeric ranges, and the grids and meshes are drawn with uniform line styles and colours."}
 
@@
 
-```{r geom-outputs-pdf, fig.cap="The outputs of \\texttt{geom\\_hexgrid} and \\texttt{geom\\_trimesh} include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_latex_output(), fig.alt = "Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region."}
+```{r geom-outputs-pdf, fig.cap="The outputs of \\texttt{geom\\_hexgrid} and \\texttt{geom\\_trimesh} include: (a) a complete hexagonal grid, (b) a hexagonal grid that corresponds with the data, (c) a full grid based on centroid triangulation, and (d) a centroid triangulation grid that aligns with the data.", fig.align='center', fig.pos="!ht", fig.height=5, fig.width=20, eval=knitr::is_latex_output(), fig.alt = "The figure consists of four 2-D panels illustrating hexagonal grids and triangulated meshes.
+Panel (a) shows a complete hexagonal grid covering a rectangular 2-D region, with all hexagons drawn regardless of whether data are present.
+Panel (b) shows a hexagonal grid restricted to bins that contain data points, with only those hexagons displayed and gaps where bins are empty.
+Panel (c) shows a full triangulated mesh constructed from centroids across the entire grid region, with centroids connected by straight line segments to form triangles.
+Panel (d) shows a triangulated mesh constructed only from centroids corresponding to non-empty bins, aligned with the data layout and leaving open areas where no centroids are present.
+In all panels, the horizontal and vertical axes represent continuous 2-D coordinates with bounded numeric ranges, and the grids and meshes are drawn with uniform line styles and colours."}
 
@@
 These views illustrate how the lifted wireframe model captures the structure of the \\texttt{scurve} data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Interactive langevitour visualization showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The views reveal how the wireframe follows the underlying S-curve structure and highlight regions with sparse or missing data in the original layout."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "The figure shows an animated langevitour of the lifted high-dimensional wireframe model derived from the scurve UMAP layout.
+At each moment, a 2-D projection is displayed in which the horizontal and vertical axes represent linear combinations of the original high-dimensional coordinates, with continuous numeric ranges.
+The wireframe model appears as a network of connected line segments drawn in black, while the scurve data points are overlaid in blue as individual markers.
+As the projection changes smoothly over time, the wireframe and data rotate and change orientation together, revealing different alignments between the model and the data.
+In some views, parts of the wireframe extend into areas with few or no data points, leaving visible gaps.
+The display maintains a roughly square aspect ratio throughout the tour."}
 
@@
 These views illustrate how the lifted wireframe model captures the structure of the \\texttt{scurve} data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
-Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data."}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "The figure shows four panels of 2-D scatter plots (a1–a4), each displaying a different projection of the lifted high-dimensional wireframe model derived from the scurve UMAP layout.
+In each panel, the horizontal and vertical axes represent two projected coordinates with continuous numeric ranges.
+The wireframe model is drawn in black as a wireframe of connected line segments, and the scurve data points are overlaid in blue as individual markers.
+The relative positions and orientations of the wireframe and data points vary across panels as different projections are shown.
+In some regions, the wireframe extends into areas with few or no data points, leaving visible gaps between the model and the data.
+All panels have similar aspect ratios and use consistent colour and symbol encodings to distinguish the model from the data."}
 
@@
 ```{r scurve-highd-model-detourrpdf, out.width=c("50%", "50%"), fig.cap="Screenshots of the lifted high-dimensional wireframe model from the \\texttt{scurve} UMAP layout using \\texttt{detourr}.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "The figure shows several screenshots of 2-D projections of the lifted high-dimensional wireframe model derived from the scurve UMAP layout using detourr.
+In each screenshot, the horizontal and vertical axes represent projected coordinates with continuous numeric ranges.
+The wireframe model is displayed as a set of connected line segments, and the scurve data points are shown as individual markers overlaid on the wireframe.
+The orientation and shape of the wireframe and point cloud differ across screenshots as different projections are shown.
+In some views, parts of the wireframe extend into areas with few or no data points, leaving visible gaps between the model and the data.
+All screenshots use consistent colour and line styles and have similar aspect ratios.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 
@@
 ```{r scurve-highd-model-detourrhtml, echo=TRUE, eval=knitr::is_html_output(), fig.cap="`detourr` output of the lifted high-dimensional wireframe model from the `scurve` UMAP layout.
-These views illustrate how the lifted wireframe model captures the structure of the Scurve data.
-Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "Multiple views from the detourr output showing the lifted high-dimensional wireframe model based on the S-curve UMAP layout.
-The wireframe reflects the curved structure of the S-curve data and reveals areas with sparse or no data in the original layout."}
+These views illustrate how the lifted wireframe model captures the structure of the `scurve` data.
+Regions with sparse or no data in the UMAP layout are also visible in the lifted model.", fig.alt = "The figure shows an animated detourr tour of the lifted high-dimensional wireframe model derived from the scurve UMAP layout.
+The display presents a sequence of changing 2-D projections, where the horizontal and vertical axes represent linear combinations of the original high-dimensional coordinates with continuous numeric ranges.
+The wireframe model appears as a network of connected line segments, and the scurve data points are overlaid as individual markers.
+As the tour progresses, the orientation and shape of the wireframe and point cloud change smoothly, revealing different alignments between the model and the data.
+In some views, sections of the wireframe extend into regions with few or no data points, leaving visible gaps.
+The visualization maintains a roughly square aspect ratio and consistent visual encodings throughout the tour."}
 
@@
 Brushing in either view highlights the corresponding subset of points across both representations, allowing examination of how well the UMAP layout preserves the underlying `scurve` structure.
-Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "Interactive two-panel linked plots showing the connection between the UMAP layout on the left and the fitted model with S-curve data in seven dimensions on the right, displayed using langevitour.
-Brushing highlights matching points across both views, enabling comparison of structure preservation between the 2-D layout and the high-dimensional data.", layout = "l-body"}
+Double-clicking on either view clears the selection and dehighlights all points.", fig.alt = "The figure shows two interactively linked panels.
+The left panel displays a 2-D UMAP layout of the scurve data, with horizontal and vertical axes representing the UMAP coordinates on continuous numeric scales.
+The right panel shows a 2-D projection of the 7-D data and fitted model using langevitour, where each axis represents a linear combination of the original seven dimensions.
+Data points are plotted as individual markers, and brushing a subset of points in either panel highlights the corresponding points in the other panel using a distinct colour.
+The highlighted points update dynamically as selections change, and double-clicking on either panel clears all highlights.
+The display maintains roughly square aspect ratios in both panels and uses consistent visual encodings for data and selected subsets throughout the interactive view.", layout = "l-body"}
 
@@
 In (b1), a different region near the upper arm of the \\texttt{scurve} is selected, and in (b2) these points map onto one side of the curved manifold in $7\\text{-}D$ projection.
-While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP.", fig.alt = "Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
-Panels (a1, b1) show selected regions in the UMAP layout, and panels (a2, b2) display the corresponding points in the high-dimensional tour view.
-Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
-The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects."}
+While the UMAP layout suggests distinct local clusters, the linked tour views reveal how these selections trace continuous structures in the $7\\text{-}D$ space, highlighting distortions introduced by UMAP.", fig.alt = "The figure consists of four linked panels arranged in two rows and two columns.
+Panels (a1) and (b1) show 2-D scatter plots of a UMAP layout, with the horizontal and vertical axes representing the two UMAP coordinates on continuous numeric scales.
+Panels (a2) and (b2) show corresponding 2-D projections from a tour of the 7-D data and model, where each axis represents a linear combination of the original seven dimensions.
+Data points are plotted as individual markers, and selected subsets are highlighted in purple while unselected points appear in a contrasting colour.
+In the top row, a set of points highlighted in the lower region of the UMAP layout in (a1) is linked to highlighted points in the tour view in (a2).
+In the bottom row, a different highlighted region in the UMAP layout in (b1) is linked to a different set of highlighted points in the tour view in (b2).
+The linked selections update consistently across panels, showing how the same subsets of observations appear in the UMAP layout and in different projections of the high-dimensional space."}
 
@@
 Brushing in any view highlights the corresponding subset across all panels, allowing detailed inspection of how residuals relate to both the UMAP embedding and the $p\\text{-}D$ structure.
-Points with higher residuals typically correspond to sparse or bridging regions in the UMAP layout and the less dense ends of the Scurve, whereas low-residual points tend to cluster within denser, well-preserved regions.
-Double-clicking clears the selection and resets all highlights.", fig.alt = "Interactive three-panel linked plots showing the residual distribution (left), UMAP layout (middle), and fitted model overlaid with S-curve data in seven dimensions (right) using langevitour.
-Brushing highlights corresponding points across all panels, illustrating how high-residual points align with sparse or bridging regions in the UMAP and less dense ends of the S-curve, while low-residual points cluster in denser regions.", layout = "l-body"}
+Points with higher residuals typically correspond to sparse or bridging regions in the UMAP layout and the less dense ends of the `scurve`, whereas low-residual points tend to cluster within denser, well-preserved regions.
+Double-clicking clears the selection and resets all highlights.", fig.alt = "The figure shows three interactively linked panels.
+The left panel displays a univariate plot of model residuals, with the horizontal axis representing residual values and the vertical axis representing density.
+The middle panel shows a 2-D UMAP layout of the scurve data, with horizontal and vertical axes representing the two UMAP coordinates.
+The right panel shows a 2-D projection of the 7-D data and fitted model using langevitour, where each axis represents a linear combination of the original dimensions.
+Points are plotted as individual markers in all panels, and brushing a subset of points in any panel highlights the corresponding points across the other two panels using a distinct colour.
+Selections update dynamically as the user interacts with any panel, and double-clicking clears all highlights.
+All panels maintain roughly square aspect ratios and consistent visual encodings for selected and unselected points throughout the interactive view.", layout = "l-body"}
 
@@
 In the bottom row (b1–b3), points with lower residuals (b1) are highlighted, which map to one side of the dense region in the NLDR layout (b2) and to a thicker band of the \\texttt{scurve} in the projection (b3).
-This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.", fig.alt = "Six-panel figure showing linked views of residuals and UMAP layouts for the 7-D S-curve model.
-Panels (a1–a3) highlight high-residual points in purple, corresponding to sparse bridging regions in the UMAP layout and less dense ends of the high-dimensional S-curve.
-Panels (b1–b3) highlight low-residual points, mapping to denser regions in the UMAP and thicker bands in the high-dimensional projection.
-The figure illustrates how residuals reveal distortions in UMAP embeddings."}
+This comparison illustrates how residuals can help diagnose distortions in UMAP, with high-residual points often concentrated in sparse or stretched regions of the structure.", fig.alt = "The figure consists of six panels arranged in two rows and three columns, showing residuals and their correspondence between UMAP and high-dimensional projections of the scurve data.
+In each row, the left panel (a1, b1) is a univariate plot of model residuals, with the horizontal axis representing residual values and the vertical axis representing frequency or density; points selected for highlighting appear in purple.
+The middle panel (a2, b2) shows a 2-D UMAP layout, with horizontal and vertical axes representing the two UMAP coordinates; the selected points from the left panel are highlighted in purple, while unselected points appear in a contrasting colour.
+The right panel (a3, b3) shows a 2-D projection of the 7-D data and model using a tour view, where axes represent linear combinations of the original dimensions; highlighted points correspond to those selected in the residuals and UMAP panels, while unselected points are shown in a contrasting colour.
+Across the panels, the purple highlights differ between the top and bottom rows, showing different subsets of points.
+All panels maintain roughly square aspect ratios and use consistent visual encodings for selection and non-selection."}
 
@@
 ```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}a) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "Comparison of published $2\\text{-}D$ layout (b) and the HBE-selected $2\\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "The figure shows a 2-D scatter plot of a tSNE layout for the Limb muscle dataset (n=1067) with seven clusters in 10 dimensions.
+Each point represents one observation, with point colour indicating the cluster label reported in the original paper.
+The horizontal and vertical axes represent the two tSNE coordinates, both on roughly symmetric numeric scales around zero.
+Points form several visually distinct clusters, with some local variation visible within clusters.
+A model overlay with a binwidth of 0.06 shows the local density structure, providing a background grid that highlights differences between clusters without fragmenting the plot.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel linked plots showing the published $2\\text{-}D$ layout (b) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
-Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, illustrating how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-body"}
+Double-clicking clears the selection and removes highlights.", fig.alt = "The figure shows two interactively linked panels.
+The left panel displays a 2-D NLDR layout of the Limb muscle dataset (n=1067), with each point representing an observation and coloured according to its muscle group.
+The right panel shows a 2-D projection of the 10-D data and fitted model using langevitour, where each axis represents a linear combination of the original dimensions.
+Triangulated hexagon centroids on the NLDR layout are connected to their corresponding regions in the high-dimensional projection.
+Brushing a subset of points in either panel highlights the corresponding points in both panels using a distinct colour, and double-clicking clears all selections.
+Both panels maintain roughly square aspect ratios and consistent visual encodings for selected and unselected points throughout the interactive view.", layout = "l-body"}
 
@@
 Brushing in either view highlights the same subset of points across both panels, enabling exploration of how clusters in the low-dimensional embedding align with structures in the original high-dimensional space.
-Double-clicking clears the selection and removes highlights.", fig.alt = "Interactive two-panel linked plots showing the HBE-selected $2\\text{-}D$ layout (f) of limb muscle data on the left and the fitted high-dimensional model in $10\\text{-}D$ on the right using langevitour.
-Colors indicate different muscle groups.
-Brushing or selecting points in either panel highlights the corresponding points in the other, showing how clusters in the $2\\text{-}D$ embedding correspond to structures in the original $10\\text{-}D$ space.", layout = "l-body"}
+Double-clicking clears the selection and removes highlights.", fig.alt = "The figure shows two interactively linked panels.
+The left panel displays a 2-D NLDR layout (layout f) of the Limb muscle dataset (n=1067), with each point representing an observation and coloured by muscle group.
+The right panel shows a 2-D projection of the 10-D data and fitted model using langevitour, where each axis represents a linear combination of the original dimensions.
+Triangulated hexagon centroids on the NLDR layout are connected to their corresponding regions in the high-dimensional projection.
+Brushing a subset of points in either panel highlights the same points in both panels using a distinct colour, and double-clicking clears all selections.
+Both panels maintain roughly square aspect ratios and consistent visual encodings for selected and unselected points throughout the interactive view.", layout = "l-body"}
 

### 2025-12-19 Archive/paper-quollr.Rmd
Updated spellings and try to make submit .zip file

--- 2025-12-19 paper-quollr.Rmd
@@
 
-```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}a) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "The figure shows a 2-D scatter plot of a tSNE layout for the Limb muscle dataset (n=1067) with seven clusters in 10 dimensions.
-Each point represents one observation, with point colour indicating the cluster label reported in the original paper.
-The horizontal and vertical axes represent the two tSNE coordinates, both on roughly symmetric numeric scales around zero.
-Points form several visually distinct clusters, with some local variation visible within clusters.
-A model overlay with a binwidth of 0.06 shows the local density structure, providing a background grid that highlights differences between clusters without fragmenting the plot.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+```{r model-limb, fig.cap="Representative views of two selected NLDR layouts for the Limb muscle dataset ($n=1067$), shown row-wise.
+The top row (a1–a3) corresponds to the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}a), and the bottom row (f1–f3) corresponds to the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) selected using the HBE plot.
+In each row, the left panel (a1, f1) shows the NLDR embedding with points coloured by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2, f2) and right (a3, f3) panels show two different $2\\text{-}D$ projections of the fitted model and data in the original $10\\text{-}D$ space, with the same triangular mesh displayed.
+Together, these panels summarize how the low-dimensional layouts relate to the underlying high-dimensional structure across different viewing directions.", fig.alt = "The figure consists of six panels arranged in two rows and three columns, showing representative views of two NLDR layouts for the Limb muscle dataset (n=1067).
+In the top row (a1–a3), the left panel (a1) shows the published 2-D NLDR embedding with points coloured by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2) and right (a3) panels show two different 2-D projections of the corresponding 10-D data and fitted model, with points plotted as individual markers and the same triangular mesh connecting centroids.
+In the bottom row (f1–f3), the left panel (f1) shows the 2-D NLDR layout selected using the HBE plot, again with coloured points and triangulated centroids, while panels (f2) and (f3) show two different 2-D projections of the associated 10-D structure with the same mesh displayed.
+Across all panels, axes represent continuous numeric coordinates, colour encodes muscle group, and the triangular mesh is used consistently to show correspondence between the low-dimensional layouts and the high-dimensional projections.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 

### 2025-12-19 Archive/paper-quollr.tex
Updated spellings and try to make submit .zip file

--- 2025-12-17 paper-quollr.tex
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Flowchart diagram illustrating the quollr workflow.
-It shows inputs of NLDR and high-dimensional data progressing through preprocessing, hexagonal binning, centroid computation, and triangulation to form a 2-D mesh.
-The mesh is lifted into the high-dimensional space for prediction and error analysis, with interactive tools linking the high- and low-dimensional representations.}]{figures/quollr_workflow} \caption{Overview of the `quollr` workflow and software architecture.
+\includegraphics[width=1\linewidth,alt={The figure is a schematic diagram illustrating the quollr workflow and software architecture.
+Rectangular boxes represent processing steps and data objects, including NLDR and p-D data inputs, data preprocessing, hexagonal binning, centroid computation, triangulation, and construction of a 2-D mesh.
+Arrows indicate the flow of information between steps, from the input data through the 2-D mesh, and then from the mesh being lifted into the p-D space.
+Additional boxes show prediction and error computation for new data, as well as interactive components that link the p-D and 2-D representations.
+The diagram is laid out sequentially, showing how data move through the system from input to interactive output.}]{figures/quollr_workflow} \caption{Overview of the `quollr` workflow and software architecture.
 The process begins with NLDR and $p\text{-}D$ data inputs, followed by data preprocessing and hexagonal binning.
@@
 
-{\centering \includegraphics[width=0.8\linewidth,alt={Four-panel figure showing the main steps in building the model on a UMAP layout using S-curve data: (a) hexagonal binning of points, (b) calculation of bin centroids, (c) triangulation of centroids to form a mesh, and (d) lifting the triangulated model into the high-dimensional space.}]{paper-quollr_files/figure-latex/algo-steps-1}
+{\centering \includegraphics[width=0.8\linewidth,alt={The figure is a multi-panel diagram illustrating four steps in constructing a model on a UMAP layout using the scurve data.
+Panel (a) shows a 2-D UMAP scatter plot with points grouped into hexagonal bins covering the layout.
+Panel (b) shows the same layout with a centroid point displayed for each hexagon bin.
+Panel (c) shows the centroids connected by straight line segments to form a triangulated mesh across the 2-D layout.
+Panel (d) shows this 2-D triangulated mesh lifted into the original high-dimensional space, where the mesh appears as a curved surface.
+The horizontal and vertical axes in the 2-D panels represent the two UMAP coordinates with bounded numeric ranges, while the lifted view represents coordinates in the higher-dimensional space without explicit axis scales.}]{paper-quollr_files/figure-latex/algo-steps-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,alt={Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region.}]{paper-quollr_files/figure-latex/geom-outputs-pdf-1}
+{\centering \includegraphics[width=1\linewidth,alt={The figure consists of four 2-D panels illustrating hexagonal grids and triangulated meshes.
+Panel (a) shows a complete hexagonal grid covering a rectangular 2-D region, with all hexagons drawn regardless of whether data are present.
+Panel (b) shows a hexagonal grid restricted to bins that contain data points, with only those hexagons displayed and gaps where bins are empty.
+Panel (c) shows a full triangulated mesh constructed from centroids across the entire grid region, with centroids connected by straight line segments to form triangles.
+Panel (d) shows a triangulated mesh constructed only from centroids corresponding to non-empty bins, aligned with the data layout and leaving open areas where no centroids are present.
+In all panels, the horizontal and vertical axes represent continuous 2-D coordinates with bounded numeric ranges, and the grids and meshes are drawn with uniform line styles and colours.}]{paper-quollr_files/figure-latex/geom-outputs-pdf-1}
 
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
-Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout.
+\includegraphics[width=1\linewidth,alt={The figure shows four panels of 2-D scatter plots (a1–a4), each displaying a different projection of the lifted high-dimensional wireframe model derived from the scurve UMAP layout.
+In each panel, the horizontal and vertical axes represent two projected coordinates with continuous numeric ranges.
+The wireframe model is drawn in black as a wireframe of connected line segments, and the scurve data points are overlaid in blue as individual markers.
+The relative positions and orientations of the wireframe and data points vary across panels as different projections are shown.
+In some regions, the wireframe extends into areas with few or no data points, leaving visible gaps between the model and the data.
+All panels have similar aspect ratios and use consistent colour and symbol encodings to distinguish the model from the data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on \texttt{scurve} data (blue) in different projections.
@@
 \begin{figure}[H]
-\includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout using \texttt{detourr}.
+\includegraphics[width=0.5\linewidth,alt={The figure shows several screenshots of 2-D projections of the lifted high-dimensional wireframe model derived from the scurve UMAP layout using detourr.
+In each screenshot, the horizontal and vertical axes represent projected coordinates with continuous numeric ranges.
+The wireframe model is displayed as a set of connected line segments, and the scurve data points are shown as individual markers overlaid on the wireframe.
+The orientation and shape of the wireframe and point cloud differ across screenshots as different projections are shown.
+In some views, parts of the wireframe extend into areas with few or no data points, leaving visible gaps between the model and the data.
+All screenshots use consistent colour and line styles and have similar aspect ratios.}]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth,alt={The figure shows several screenshots of 2-D projections of the lifted high-dimensional wireframe model derived from the scurve UMAP layout using detourr.
+In each screenshot, the horizontal and vertical axes represent projected coordinates with continuous numeric ranges.
+The wireframe model is displayed as a set of connected line segments, and the scurve data points are shown as individual markers overlaid on the wireframe.
+The orientation and shape of the wireframe and point cloud differ across screenshots as different projections are shown.
+In some views, parts of the wireframe extend into areas with few or no data points, leaving visible gaps between the model and the data.
+All screenshots use consistent colour and line styles and have similar aspect ratios.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout using \texttt{detourr}.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-detourrpdf}
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
-Panels (a1, b1) show selected regions in the UMAP layout, and panels (a2, b2) display the corresponding points in the high-dimensional tour view.
-Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
-The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and \texttt{scurve} structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth,alt={The figure consists of four linked panels arranged in two rows and two columns.
+Panels (a1) and (b1) show 2-D scatter plots of a UMAP layout, with the horizontal and vertical axes representing the two UMAP coordinates on continuous numeric scales.
+Panels (a2) and (b2) show corresponding 2-D projections from a tour of the 7-D data and model, where each axis represents a linear combination of the original seven dimensions.
+Data points are plotted as individual markers, and selected subsets are highlighted in purple while unselected points appear in a contrasting colour.
+In the top row, a set of points highlighted in the lower region of the UMAP layout in (a1) is linked to highlighted points in the tour view in (a2).
+In the bottom row, a different highlighted region in the UMAP layout in (b1) is linked to a different set of highlighted points in the tour view in (b2).
+The linked selections update consistently across panels, showing how the same subsets of observations appear in the UMAP layout and in different projections of the high-dimensional space.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and \texttt{scurve} structure in $7\text{-}D$.
 Two sets of plots are interactively linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Six-panel figure showing linked views of residuals and UMAP layouts for the 7-D S-curve model.
-Panels (a1–a3) highlight high-residual points in purple, corresponding to sparse bridging regions in the UMAP layout and less dense ends of the high-dimensional S-curve.
-Panels (b1–b3) highlight low-residual points, mapping to denser regions in the UMAP and thicker bands in the high-dimensional projection.
-The figure illustrates how residuals reveal distortions in UMAP embeddings.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ \texttt{scurve} model.
+\includegraphics[width=1\linewidth,alt={The figure consists of six panels arranged in two rows and three columns, showing residuals and their correspondence between UMAP and high-dimensional projections of the scurve data.
+In each row, the left panel (a1, b1) is a univariate plot of model residuals, with the horizontal axis representing residual values and the vertical axis representing frequency or density; points selected for highlighting appear in purple.
+The middle panel (a2, b2) shows a 2-D UMAP layout, with horizontal and vertical axes representing the two UMAP coordinates; the selected points from the left panel are highlighted in purple, while unselected points appear in a contrasting colour.
+The right panel (a3, b3) shows a 2-D projection of the 7-D data and model using a tour view, where axes represent linear combinations of the original dimensions; highlighted points correspond to those selected in the residuals and UMAP panels, while unselected points are shown in a contrasting colour.
+Across the panels, the purple highlights differ between the top and bottom rows, showing different subsets of points.
+All panels maintain roughly square aspect ratios and use consistent visual encodings for selection and non-selection.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ \texttt{scurve} model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Comparison of published $2\text{-}D$ layout (b) and the HBE-selected $2\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}a) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}
+\includegraphics[width=1\linewidth,alt={The figure consists of six panels arranged in two rows and three columns, showing representative views of two NLDR layouts for the Limb muscle dataset (n=1067).
+In the top row (a1–a3), the left panel (a1) shows the published 2-D NLDR embedding with points coloured by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2) and right (a3) panels show two different 2-D projections of the corresponding 10-D data and fitted model, with points plotted as individual markers and the same triangular mesh connecting centroids.
+In the bottom row (f1–f3), the left panel (f1) shows the 2-D NLDR layout selected using the HBE plot, again with coloured points and triangulated centroids, while panels (f2) and (f3) show two different 2-D projections of the associated 10-D structure with the same mesh displayed.
+Across all panels, axes represent continuous numeric coordinates, colour encodes muscle group, and the triangular mesh is used consistently to show correspondence between the low-dimensional layouts and the high-dimensional projections.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Representative views of two selected NLDR layouts for the Limb muscle dataset ($n=1067$), shown row-wise.
+The top row (a1–a3) corresponds to the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}a), and the bottom row (f1–f3) corresponds to the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) selected using the HBE plot.
+In each row, the left panel (a1, f1) shows the NLDR embedding with points coloured by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2, f2) and right (a3, f3) panels show two different $2\text{-}D$ projections of the fitted model and data in the original $10\text{-}D$ space, with the same triangular mesh displayed.
+Together, these panels summarize how the low-dimensional layouts relate to the underlying high-dimensional structure across different viewing directions.}\label{fig:model-limb}
 \end{figure}

### 2025-12-19 extra/paper-quollr-code.Rmd
Updated spellings and try to make submit .zip file

--- 2025-12-14 extra/paper-quollr-code.Rmd
@@
 
-```{r, fig.cap="Line plot."}
-#| label: fig-unemployment
+<!-- 20 pages-->
 
-library(autoalt)
-library(ggplot2)
-library(dplyr)
+# Application
 
-# Built-in datasets from ggplot2 and base R
-data("economics", package = "ggplot2")
-data("airquality")
+Single-cell RNA sequencing (scRNA-seq) is a popular and powerful technology that allows you to profile the whole transcriptome of a large number of individual cells [@andrews2021].
 
-economics |>
-  ggplot(aes(x = date, y = unemploy)) +
-  geom_line(linewidth = 0.6) +
-  labs(
-    title = "Total Number of Unemployed Persons in the US Over Time",
-    x = "Year",
-    y = "Number of Unemployed (thousands)"
+Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
+NLDR often used to summarize the discovered clusters, and help to understand the results.
+The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
+
+Limb muscle cells of mice in @tabula2018 are examined.
+There are $1067$ single cells, with $14997$ gene expressions.
+Following their pre-processing, different NLDR methods were performed using ten principal components.
+Figure \@ref(fig:limb-hbe) (a) is the reproduction of the published plot.
+This was generated using tSNE with $\text{perplexity}=30$, the default hyper-parameters.
+The question is whether this accurately represents the cluster structure in the data.
+Note that the cluster variable is not used to produce the $2\text{-}D$ layout.
+
+We illustrate how to use `quollr` to assess whether this is a reasonable layout.
+Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits.
+Layout b is produce by UMAP ($\text{neighbors}=5, ~\text{minimum distance}=0.1$); layout c was produced by PHATE ($\text{knn}=5$); layout d was produced by TriMAP ($\text{number of inliers}=12,~\text{outliers}=4,~\text{random}=3$); layout e was produced by by PaCMAP ($\text{neighbors}=10,~\text{init = "random"}, ~\text{MN-ratio}=0.5,~\text{FP-ratio}=2$); layout f was produced by tSNE ($\text{perplexity}=15$).
+
+The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of $15$ outperforms the other.
+There are small visual difference between the two layouts.
+Both support that there are $5-6$ clusters.
+Layout a has slightly more space between clusters.
+Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom.
+
+`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:tsne-link-limb), \\@ref(fig:tsne-link-limb-best)"} else { "Figure \\@ref(fig:model-limb)"})` show how to examine the resulting models of the layouts overlaid on the data in high dimensions.
+Point color represents the cluster reported in the published paper.
+In each case the best model is produced using $\text{binwidth}=0.06$.
+A binwidth of $0.06$ is used because it is small enough to show local structure and differences between clusters, but large enough to avoid breaking the layout into too many small pieces that would make the plots hard to interpret.
+The plots in these figures are best understand using small steps.
+
+1.
+Examine the model in the data space for each, by looking at the tour views.
+In each case, the clustering doesn't quite match the separations in the data, and both models help see this.
+For example, the orange cluster ($1$) should probably be split into more than one cluster because both models show large stretched lines connecting a small group far from the remaining orange points.
+
+2.
+Because $6$ clusters are hard to examine together, use the menu to select just one cluster to view at at time.
+Selecting just cluster $1$ might help you see the explanation above, that a small group is quite separate from the main group.
+This suggests both the layouts and the clustering that groups these together might be wrong.
+Now change to focus on cluster $5$ (yellow).
+This group is a fairly large sparse cluster but it is separated from the other points.
+Both layouts are right in separating these points from the others but the fit for layout f is slightly better.
+
+3.
+To examine where the layouts differ, examine clusters $4$ (green) and $6$ (blue), by selecting just these two.
+Layout a has them close, but layout f has them far apart.
+(It might also help to include cluster $5$ here because layout a has this group close to the cluster $4$ also).
+In the tour view, we can see that they three clusters are separated clusters in different directions away from the most of the other points.
+They are both correct in this.
+It may have been better to place them all in different corners of the layout, but they have preserved the most important aspect that they are separated clusters.
+That they are all close together in layout a could be incorrectly interpreted as close in high-dimensions though.
+
+```{r model-limb, fig.cap="Representative views of two selected NLDR layouts for the Limb muscle dataset ($n=1067$), shown row-wise.
+The top row (a1–a3) corresponds to the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}a), and the bottom row (f1–f3) corresponds to the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) selected using the HBE plot.
+In each row, the left panel (a1, f1) shows the NLDR embedding with points coloured by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2, f2) and right (a3, f3) panels show two different $2\\text{-}D$ projections of the fitted model and data in the original $10\\text{-}D$ space, with the same triangular mesh displayed.
+Together, these panels summarize how the low-dimensional layouts relate to the underlying high-dimensional structure across different viewing directions.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+
+set.seed(20240110)
+
+theme_set(theme_linedraw() +
+   theme(
+     aspect.ratio = 1,
+     plot.background = element_rect(fill = 'transparent', colour = NA),
+     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+     panel.background = element_rect(fill = 'transparent', 
+                                     colour = NA),
+     panel.grid.major = element_blank(), 
+     panel.grid.minor = element_blank(), 
+     axis.title.x = element_blank(), axis.title.y = element_blank(),
+     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+     legend.background = element_rect(fill = 'transparent', 
+                                      colour = NA),
+     legend.key = element_rect(fill = 'transparent', 
+                               colour = NA),
+     legend.position = "none", 
+     legend.title = element_text(size=5), 
+     legend.text = element_text(size=4),
+     legend.key.height = unit(0.25, 'cm'),
+     legend.key.width = unit(0.25, 'cm'),
+     plot.margin = margin(0, 0, 0, 0)
+   )
+
+)
+
+# Add plot label
+
+interior_annotation <- function(label, position = c(0.9, 0.9), cex = 1, col="grey70") {
+  annotation_custom(grid::textGrob(label = label,
+                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
+                                   gp = grid::gpar(cex = cex, col=col)))
+}
+
+# Scale high-d data
+
+# Center the data by subtracting the mean of each column
+center_data <- function(data) {
+  center_values <- colMeans(data)
+  data_centered <- sweep(data, 2, center_values, FUN = "-")  # subtract means
+  data_centered
+}
+
+# Function to scale data manually
+# scale_data_manual <- function(data, type_col) {
+#   # Step 1: Center the data (mean 0)
+#   data_centered <- center_data(data |> select(-all_of(type_col)))
+#
+#   # Step 2: Calculate the standard deviation of each dimension
+#   sds <- apply(data_centered, 2, sd)
+#
+#   # Step 3: Scale each dimension to have the range [0, 1]
+#   data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
+#
+#   # Step 4: Scale dimensions according to their variation
+#   # The dimension with the highest standard deviation is scaled to [-1, 1]
+#   # Other dimensions are scaled to smaller ranges based on their standard deviations
+#   max_sd <- max(sds)
+#
+#   # Normalize the standard deviations to get scaling factors
+#   scaling_factors <- sds / max_sd
+#
+#   for (i in seq_along(scaling_factors)) {
+#     data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
+#   }
+#
+#   # Combine the scaled data with the 'type' column and return as a tibble
+#   data_scaled <- as_tibble(data_scaled) %>%
+#     mutate(!!type_col := data[[type_col]])
+#
+#   return(data_scaled)
+# }
+
+# Plot MSE
+
+plot_hbe <- function(error_df) {
+
+  ggplot(error_df,
+         aes(x = a1,
+             y = HBE,
+             colour = method)) +
+    geom_point(size = 0.8) +
+    geom_line(linewidth = 0.3) +
+    ylab("HBE") +
+    xlab(expression(paste("binwidth (", a[1], ")"))) +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
+          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+          axis.ticks.x = element_line(),
+          axis.ticks.y = element_line(),
+          legend.position = "none",
+          axis.text.x = element_text(size = 7),
+          axis.text.y = element_text(size = 7),
+          axis.title.x = element_text(size = 7),
+          axis.title.y = element_text(size = 7),
+          plot.margin = margin(0, 0, 0, 0))
+
+}
+
+# creating Standardization function
+standardize = function(x){
+  z <- (x - mean(x)) / sd(x)
+  return( z)
+}
+
+# Solve quadratic function
+
+quad <- function(a = 3, b = 2 * a2, c = -(a2^2 + a1^2))
+{
+  a <- as.complex(a)
+  answer <- c((-b + sqrt(b^2 - 4 * a * c)) / (2 * a),
+              (-b - sqrt(b^2 - 4 * a * c)) / (2 * a))
+  if(all(Im(answer) == 0)) answer <- Re(answer)
+  if(answer[1] == answer[2]) return(answer[1])
+  answer[answer>0] ## only positive
+}
+
+data_limb <- read_rds("data/limb_muscles/facs_limb_muscles_pcs_10.rds")
+
+cluster_df <- read_rds("data/limb_muscles/facs_limb_muscles_cluster_df.rds")
+
+tsne_limb_obj <- fit_highd_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb, 
+  b1 = 28, 
+  q = 0.1, 
+  hd_thresh = 0)
+
+df_bin_centroids_limb <- tsne_limb_obj$model_2d
+df_bin_limb <- tsne_limb_obj$model_highd
+trimesh_data_limb <- tsne_limb_obj$trimesh_data
+tsne_limb_scaled <- tsne_limb_obj$nldr_scaled_obj$scaled_nldr
+
+tsne_limb_scaled_with_cluster <- inner_join(tsne_limb_scaled, cluster_df, by = "ID") |>
+  mutate(cluster = as.character(cluster.ids))
+
+trimesh_limb <- ggplot() + 
+  geom_point(
+    data = tsne_limb_scaled_with_cluster,
+    aes(
+      x = emb1,
+      y = emb2,
+      color = cluster
+    ),
+    alpha = 0.3
   ) +
-  theme_minimal()
+  geom_segment(data = trimesh_data_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1,
+               alpha = 0.3) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3')) +
+  interior_annotation("a1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
 
+df <- inner_join(data_limb, tsne_limb_scaled_with_cluster, by = "ID")
+
+### Define type column
+df <- df |>
+  select(starts_with("x"), starts_with("emb"), "cluster") |>
+  mutate(type = tsne_limb_scaled_with_cluster$cluster) ## original dataset
+
+df_b <- df_bin_limb |>
+  filter(h %in% df_bin_centroids_limb$h) |>
+  mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in model_2d
+df_b <- df_b[match(df_bin_centroids_limb$h, df_b$h),] |>
+  tidyr::drop_na() |>
+  select(-h)
+
+df_exe_limb_int <- bind_rows(df_b, df)
+
+point_data <- df_exe_limb_int 
+edge_data <- trimesh_data_limb
+
+num_highd_col <- point_data |>
+  dplyr::select(starts_with("x")) |>
+  NCOL()
+
+shared_df <- crosstalk::SharedData$new(point_data)
+
+cluster_colors <- c(
+    "0" = '#66c2a5', "1" = '#fc8d62', "2" = '#8da0cb',
+    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#a6cee3'
+)
+
+point_sizes <- c(0, 1)
+
+data_limb_n <- data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(h %in% df_bin_centroids_limb$h) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
+  dplyr::select(-h)
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
+scaled_limb <- center_data(df_model_data_limb[,-11]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_limb$type)
+
+scaled_limb_data <- scaled_limb |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_limb_data_model <- scaled_limb |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(-0.021068,0.007018,0.008770,0.016884,-0.008135,0.011458,-0.022323,0.007231,0.004976,-0.019298),
+  c(-0.017371,-0.008514,0.003231,-0.022945,-0.026264,-0.001571,0.008007,-0.001268,-0.016923,-0.006700))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.13))
+
+proj_obj1[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_model1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-1.3, 1.4), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "a2", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
+
+## Second projection
+projection <- cbind(
+  c(-0.025678,-0.017472,0.015634,-0.014868,-0.000752,0.001050,-0.019524,-0.008076,-0.010808,-0.002319),
+  c(0.014526,-0.013526,-0.005147,-0.027612,-0.016919,0.005198,0.010710,0.004623,-0.007055,0.017842))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.14))
+
+proj_obj2[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_model2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  line_param = c(0.5, 0.3, "#000000"),
+  plot_limits = c(-1.3, 1.3), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "a3", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
+
+tsne_best_limb_obj <- fit_highd_model(
+  highd_data = data_limb, 
+  nldr_data = tsne_limb2, 
+  b1 = 28, 
+  q = 0.1, 
+  hd_thresh = 0)
+
+df_bin_centroids_limb <- tsne_best_limb_obj$model_2d
+df_bin_limb <- tsne_best_limb_obj$model_highd
+trimesh_data_limb <- tsne_best_limb_obj$trimesh_data
+tsne_limb_scaled <- tsne_best_limb_obj$nldr_scaled_obj$scaled_nldr
+
+tsne_limb_scaled_with_cluster <- inner_join(tsne_limb_scaled, cluster_df, by = "ID") |>
+  mutate(cluster = as.character(cluster.ids))
+
+trimesh_limb_best <- ggplot() + 
+  geom_point(
+    data = tsne_limb_scaled_with_cluster,
+    aes(
+      x = emb1,
+      y = emb2,
+      color = cluster
+    ),
+    alpha = 0.3
+  ) +
+  geom_segment(data = trimesh_data_limb, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1,
+               alpha = 0.3) +
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3')) +
+  interior_annotation("f1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+
+df <- inner_join(data_limb, tsne_limb_scaled_with_cluster, by = "ID")
+
+### Define type column
+df <- df |>
+  select(starts_with("x"), starts_with("emb"), "cluster") |>
+  mutate(type = tsne_limb_scaled_with_cluster$cluster) ## original dataset
+
+df_b <- df_bin_limb |>
+  filter(h %in% df_bin_centroids_limb$h) |>
+  mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in model_2d
+df_b <- df_b[match(df_bin_centroids_limb$h, df_b$h),] |>
+  tidyr::drop_na() |>
+  select(-h)
+
+df_exe_limb_int <- bind_rows(df_b, df)
+
+point_data <- df_exe_limb_int 
+edge_data <- trimesh_data_limb
+
+num_highd_col <- point_data |>
+  dplyr::select(starts_with("x")) |>
+  NCOL()
+
+shared_df <- crosstalk::SharedData$new(point_data)
+
+cluster_colors <- c(
+    "0" = '#66c2a5', "1" = '#fc8d62', "2" = '#8da0cb',
+    "3" = '#e78ac3', "4" = '#a6d854', "5" = '#ffd92f', "6" = '#a6cee3'
+)
+
+data_limb_n <- data_limb |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_limb <- df_bin_limb |>
+  dplyr::filter(h %in% df_bin_centroids_limb$h) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the h order in df_b_with_center_data
+df_b_limb <- df_b_limb[match(df_bin_centroids_limb$h, df_b_limb$h),] |>
+  dplyr::select(-h)
+
+# Apply the scaling
+df_model_data_limb <- bind_rows(data_limb_n, df_b_limb)
+scaled_limb <- center_data(df_model_data_limb[,-11]) |>
+  as_tibble() |>
+  mutate(type = df_model_data_limb$type)
+
+scaled_limb_data <- scaled_limb |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_limb_data_model <- scaled_limb |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(-0.021068,0.007018,0.008770,0.016884,-0.008135,0.011458,-0.022323,0.007231,0.004976,-0.019298),
+  c(-0.017371,-0.008514,0.003231,-0.022945,-0.026264,-0.001571,0.008007,-0.001268,-0.016923,-0.006700))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.13))
+
+proj_obj1[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_best_model1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.3, 1.4), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "f2", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
+
+## Second projection
+projection <- cbind(
+  c(-0.025678,-0.017472,0.015634,-0.014868,-0.000752,0.001050,-0.019524,-0.008076,-0.010808,-0.002319),
+  c(0.014526,-0.013526,-0.005147,-0.027612,-0.016919,0.005198,0.010710,0.004623,-0.007055,0.017842))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            highd_data = scaled_limb_data, 
+                            model_highd = scaled_limb_data_model, 
+                            trimesh_data = trimesh_data_limb, 
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 20, 
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.14))
+
+proj_obj2[["cluster"]] <- as.character(tsne_limb_scaled_with_cluster$cluster)
+
+limb_proj_tsne_best_model2 <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1.3, 1.3), 
+  axis_text_size = 4,
+  is_category = TRUE) +
+  interior_annotation(label = "f3", cex = 2) + 
+  scale_color_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f','#a6cee3'))
+
+trimesh_limb + limb_proj_tsne_model1 +
+  limb_proj_tsne_model2 + trimesh_limb_best +
+  limb_proj_tsne_best_model1 + limb_proj_tsne_best_model2 +
+  plot_layout(nrow = 2)
 ```

### 2025-12-19 paper-quollr.Rmd
Updated spellings and try to make submit .zip file

@@
 
-# # GitHub raw URL of the figure
-# fig_url <- "https://raw.githubusercontent.com/JayaniLakshika/paper-quollr/main/figures/quollr_workflow.png"
-# 
-# # Determine figure path depending on output
-# if (knitr::is_html_output()) {
-#   # HTML can use the URL directly
-#   fig_path <- fig_url
-# } else {
-#   # PDF needs a local copy
-#   temp_file <- tempfile(fileext = ".png")
-#   download.file(fig_url, destfile = temp_file, mode = "wb")
-#   fig_path <- temp_file
-# }
-# 
-# knitr::include_graphics(fig_path)
-
 knitr::include_graphics("figures/quollr_workflow.png")
@@
 Panel (d) shows a triangulated mesh constructed only from centroids corresponding to non-empty bins, aligned with the data layout and leaving open areas where no centroids are present.
-In all panels, the horizontal and vertical axes represent continuous 2-D coordinates with bounded numeric ranges, and the grids and meshes are drawn with uniform line styles and colours."}
+In all panels, the horizontal and vertical axes represent continuous 2-D coordinates with bounded numeric ranges, and the grids and meshes are drawn with uniform line styles and colors."}
 
@@
 In some regions, the wireframe extends into areas with few or no data points, leaving visible gaps between the model and the data.
-All panels have similar aspect ratios and use consistent colour and symbol encodings to distinguish the model from the data."}
+All panels have similar aspect ratios and use consistent color and symbol encodings to distinguish the model from the data."}
 
@@
 In some views, parts of the wireframe extend into areas with few or no data points, leaving visible gaps between the model and the data.
-All screenshots use consistent colour and line styles and have similar aspect ratios.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
+All screenshots use consistent color and line styles and have similar aspect ratios.", fig.pos="H", fig.show='hold', eval=knitr::is_latex_output()}
 
@@
 The right panel shows a 2-D projection of the 7-D data and fitted model using langevitour, where each axis represents a linear combination of the original seven dimensions.
-Data points are plotted as individual markers, and brushing a subset of points in either panel highlights the corresponding points in the other panel using a distinct colour.
+Data points are plotted as individual markers, and brushing a subset of points in either panel highlights the corresponding points in the other panel using a distinct color.
 The highlighted points update dynamically as selections change, and double-clicking on either panel clears all highlights.
@@
 Panels (a2) and (b2) show corresponding 2-D projections from a tour of the 7-D data and model, where each axis represents a linear combination of the original seven dimensions.
-Data points are plotted as individual markers, and selected subsets are highlighted in purple while unselected points appear in a contrasting colour.
+Data points are plotted as individual markers, and selected subsets are highlighted in purple while unselected points appear in a contrasting color.
 In the top row, a set of points highlighted in the lower region of the UMAP layout in (a1) is linked to highlighted points in the tour view in (a2).
@@
 The right panel shows a 2-D projection of the 7-D data and fitted model using langevitour, where each axis represents a linear combination of the original dimensions.
-Points are plotted as individual markers in all panels, and brushing a subset of points in any panel highlights the corresponding points across the other two panels using a distinct colour.
+Points are plotted as individual markers in all panels, and brushing a subset of points in any panel highlights the corresponding points across the other two panels using a distinct color.
 Selections update dynamically as the user interacts with any panel, and double-clicking clears all highlights.
@@
 In each row, the left panel (a1, b1) is a univariate plot of model residuals, with the horizontal axis representing residual values and the vertical axis representing frequency or density; points selected for highlighting appear in purple.
-The middle panel (a2, b2) shows a 2-D UMAP layout, with horizontal and vertical axes representing the two UMAP coordinates; the selected points from the left panel are highlighted in purple, while unselected points appear in a contrasting colour.
-The right panel (a3, b3) shows a 2-D projection of the 7-D data and model using a tour view, where axes represent linear combinations of the original dimensions; highlighted points correspond to those selected in the residuals and UMAP panels, while unselected points are shown in a contrasting colour.
+The middle panel (a2, b2) shows a 2-D UMAP layout, with horizontal and vertical axes representing the two UMAP coordinates; the selected points from the left panel are highlighted in purple, while unselected points appear in a contrasting color.
+The right panel (a3, b3) shows a 2-D projection of the 7-D data and model using a tour view, where axes represent linear combinations of the original dimensions; highlighted points correspond to those selected in the residuals and UMAP panels, while unselected points are shown in a contrasting color.
 Across the panels, the purple highlights differ between the top and bottom rows, showing different subsets of points.
@@
 
-In the implementation examples, points are shown without cluster-based colouring unless explicitly stated.
-When points are coloured by cluster, linked plot functionality is currently only partially supported: selections made in the `langevitour` controls are not reflected in the corresponding highlights in the interactive $2\text{-}D$ layout.
+In the implementation examples, points are shown without cluster-based coloring unless explicitly stated.
+When points are colored by cluster, linked plot functionality is currently only partially supported: selections made in the `langevitour` controls are not reflected in the corresponding highlights in the interactive $2\text{-}D$ layout.
 As a result, cluster-specific exploration must be carried out separately in the tour view and the interactive $2\text{-}D$ layout.
@@
 ```{r limb-hbe, fig.cap="Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
+Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d performs well at large binwidth (where the binwidth is not enough to capture the data structure) and poorly as the bin width decreases.
@@
 Each panel is a scatterplot with two unlabeled embedding axes (dimension 1 on the horizontal axis and dimension 2 on the vertical axis, each spanning a similar, roughly symmetric range around zero).
-Points represent the same set of observations in every panel, plotted with identical colour or marker scheme so that corresponding observations can be visually compared across layouts.
+Points represent the same set of observations in every panel, plotted with identical color or marker scheme so that corresponding observations can be visually compared across layouts.
 Panel b shows a UMAP layout (neighbors = 5, minimum distance = 0.1), panel c a PHATE layout (knn = 5), panel d a TriMAP layout (inliers = 12, outliers = 4, random = 3), panel e a PaCMAP layout (neighbors = 10, init = random, MN-ratio = 0.5, FP-ratio = 2), and panel f a tSNE layout (perplexity = 15).
@@
 
-```{r model-limb, fig.cap="Compare the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}a) and the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) by HBE plot (Figure \\ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\\text{-}D$.", fig.alt = "The figure shows a 2-D scatter plot of a tSNE layout for the Limb muscle dataset (n=1067) with seven clusters in 10 dimensions.
-Each point represents one observation, with point colour indicating the cluster label reported in the original paper.
-The horizontal and vertical axes represent the two tSNE coordinates, both on roughly symmetric numeric scales around zero.
-Points form several visually distinct clusters, with some local variation visible within clusters.
-A model overlay with a binwidth of 0.06 shows the local density structure, providing a background grid that highlights differences between clusters without fragmenting the plot.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
+```{r model-limb, fig.cap="Representative views of two selected NLDR layouts for the Limb muscle dataset ($n=1067$), shown row-wise.
+The top row (a1–a3) corresponds to the published $2\\text{-}D$ layout (Figure \\ref{fig:limb-hbe}a), and the bottom row (f1–f3) corresponds to the $2\\text{-}D$ layout selected (Figure \\ref{fig:limb-hbe}f) selected using the HBE plot.
+In each row, the left panel (a1, f1) shows the NLDR embedding with points colored by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2, f2) and right (a3, f3) panels show two different $2\\text{-}D$ projections of the fitted model and data in the original $10\\text{-}D$ space, with the same triangular mesh displayed.
+Together, these panels summarize how the low-dimensional layouts relate to the underlying high-dimensional structure across different viewing directions.", fig.alt = "The figure consists of six panels arranged in two rows and three columns, showing representative views of two NLDR layouts for the Limb muscle dataset (n=1067).
+In the top row (a1–a3), the left panel (a1) shows the published 2-D NLDR embedding with points colored by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2) and right (a3) panels show two different 2-D projections of the corresponding 10-D data and fitted model, with points plotted as individual markers and the same triangular mesh connecting centroids.
+In the bottom row (f1–f3), the left panel (f1) shows the 2-D NLDR layout selected using the HBE plot, again with colored points and triangulated centroids, while panels (f2) and (f3) show two different 2-D projections of the associated 10-D structure with the same mesh displayed.
+Across all panels, axes represent continuous numeric coordinates, color encodes muscle group, and the triangular mesh is used consistently to show correspondence between the low-dimensional layouts and the high-dimensional projections.", fig.pos="!ht", fig.width=15, fig.height=10, eval=knitr::is_latex_output()}
 
@@
 Double-clicking clears the selection and removes highlights.", fig.alt = "The figure shows two interactively linked panels.
-The left panel displays a 2-D NLDR layout of the Limb muscle dataset (n=1067), with each point representing an observation and coloured according to its muscle group.
+The left panel displays a 2-D NLDR layout of the Limb muscle dataset (n=1067), with each point representing an observation and colored according to its muscle group.
 The right panel shows a 2-D projection of the 10-D data and fitted model using langevitour, where each axis represents a linear combination of the original dimensions.
 Triangulated hexagon centroids on the NLDR layout are connected to their corresponding regions in the high-dimensional projection.
-Brushing a subset of points in either panel highlights the corresponding points in both panels using a distinct colour, and double-clicking clears all selections.
+Brushing a subset of points in either panel highlights the corresponding points in both panels using a distinct color, and double-clicking clears all selections.
 Both panels maintain roughly square aspect ratios and consistent visual encodings for selected and unselected points throughout the interactive view.", layout = "l-body"}
@@
 Double-clicking clears the selection and removes highlights.", fig.alt = "The figure shows two interactively linked panels.
-The left panel displays a 2-D NLDR layout (layout f) of the Limb muscle dataset (n=1067), with each point representing an observation and coloured by muscle group.
+The left panel displays a 2-D NLDR layout (layout f) of the Limb muscle dataset (n=1067), with each point representing an observation and colored by muscle group.
 The right panel shows a 2-D projection of the 10-D data and fitted model using langevitour, where each axis represents a linear combination of the original dimensions.
 Triangulated hexagon centroids on the NLDR layout are connected to their corresponding regions in the high-dimensional projection.
-Brushing a subset of points in either panel highlights the same points in both panels using a distinct colour, and double-clicking clears all selections.
+Brushing a subset of points in either panel highlights the same points in both panels using a distinct color, and double-clicking clears all selections.
 Both panels maintain roughly square aspect ratios and consistent visual encodings for selected and unselected points throughout the interactive view.", layout = "l-body"}

### 2025-12-19 paper-quollr.tex
Updated spellings and try to make submit .zip file

@@
 
-{\centering \includegraphics[width=1\linewidth,alt={Four-panel figure showing outputs from geom_hexgrid and geom_trimesh functions: (a) the full hexagonal grid, (b) the subset of the grid that overlaps with the data, (c) a complete triangulated grid based on centroids, and (d) the triangulated centroid grid corresponding to the data region.}]{paper-quollr_files/figure-latex/geom-outputs-pdf-1}
+{\centering \includegraphics[width=1\linewidth,alt={The figure consists of four 2-D panels illustrating hexagonal grids and triangulated meshes.
+Panel (a) shows a complete hexagonal grid covering a rectangular 2-D region, with all hexagons drawn regardless of whether data are present.
+Panel (b) shows a hexagonal grid restricted to bins that contain data points, with only those hexagons displayed and gaps where bins are empty.
+Panel (c) shows a full triangulated mesh constructed from centroids across the entire grid region, with centroids connected by straight line segments to form triangles.
+Panel (d) shows a triangulated mesh constructed only from centroids corresponding to non-empty bins, aligned with the data layout and leaving open areas where no centroids are present.
+In all panels, the horizontal and vertical axes represent continuous 2-D coordinates with bounded numeric ranges, and the grids and meshes are drawn with uniform line styles and colors.}]{paper-quollr_files/figure-latex/geom-outputs-pdf-1}
 
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Four-panel figure showing 2-D projections of the lifted high-dimensional wireframe model from the S-curve UMAP layout.
-Each panel displays the black wireframe model overlaid on blue S-curve data points, illustrating how the model captures the curved structure and revealing regions with sparse or no data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout.
+\includegraphics[width=1\linewidth,alt={The figure shows four panels of 2-D scatter plots (a1–a4), each displaying a different projection of the lifted high-dimensional wireframe model derived from the scurve UMAP layout.
+In each panel, the horizontal and vertical axes represent two projected coordinates with continuous numeric ranges.
+The wireframe model is drawn in black as a wireframe of connected line segments, and the scurve data points are overlaid in blue as individual markers.
+The relative positions and orientations of the wireframe and data points vary across panels as different projections are shown.
+In some regions, the wireframe extends into areas with few or no data points, leaving visible gaps between the model and the data.
+All panels have similar aspect ratios and use consistent color and symbol encodings to distinguish the model from the data.}]{paper-quollr_files/figure-latex/scurve-highd-model-pdf-1} \caption{$2\text{-}D$ projections of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout.
 Each panel (a1–a4) shows the model (black) overlaid on \texttt{scurve} data (blue) in different projections.
@@
 \begin{figure}[H]
-\includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth,alt={Screenshots from the detourr interface showing the lifted high-dimensional wireframe model derived from the S-curve UMAP layout.
-The images highlight how the wireframe represents the data structure and expose regions with sparse or missing data.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout using \texttt{detourr}.
+\includegraphics[width=0.5\linewidth,alt={The figure shows several screenshots of 2-D projections of the lifted high-dimensional wireframe model derived from the scurve UMAP layout using detourr.
+In each screenshot, the horizontal and vertical axes represent projected coordinates with continuous numeric ranges.
+The wireframe model is displayed as a set of connected line segments, and the scurve data points are shown as individual markers overlaid on the wireframe.
+The orientation and shape of the wireframe and point cloud differ across screenshots as different projections are shown.
+In some views, parts of the wireframe extend into areas with few or no data points, leaving visible gaps between the model and the data.
+All screenshots use consistent color and line styles and have similar aspect ratios.}]{figures/model_proj1_detourr} \includegraphics[width=0.5\linewidth,alt={The figure shows several screenshots of 2-D projections of the lifted high-dimensional wireframe model derived from the scurve UMAP layout using detourr.
+In each screenshot, the horizontal and vertical axes represent projected coordinates with continuous numeric ranges.
+The wireframe model is displayed as a set of connected line segments, and the scurve data points are shown as individual markers overlaid on the wireframe.
+The orientation and shape of the wireframe and point cloud differ across screenshots as different projections are shown.
+In some views, parts of the wireframe extend into areas with few or no data points, leaving visible gaps between the model and the data.
+All screenshots use consistent color and line styles and have similar aspect ratios.}]{figures/model_proj2_detourr} \caption{Screenshots of the lifted high-dimensional wireframe model from the \texttt{scurve} UMAP layout using \texttt{detourr}.
 Regions with sparse or no data in the UMAP layout are also visible in the lifted model.}\label{fig:scurve-highd-model-detourrpdf}
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Four-panel figure with linked views showing correspondence between the UMAP layout and the S-curve structure in seven dimensions.
-Panels (a1, b1) show selected regions in the UMAP layout, and panels (a2, b2) display the corresponding points in the high-dimensional tour view.
-Purple points indicate selected subsets: the lower bridge of the S-curve in (a1, a2) and the upper arm region in (b1, b2).
-The figure illustrates how continuous structures in the 7-D S-curve appear as separate clusters in the 2-D UMAP layout, revealing distortion effects.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and \texttt{scurve} structure in $7\text{-}D$.
+\includegraphics[width=1\linewidth,alt={The figure consists of four linked panels arranged in two rows and two columns.
+Panels (a1) and (b1) show 2-D scatter plots of a UMAP layout, with the horizontal and vertical axes representing the two UMAP coordinates on continuous numeric scales.
+Panels (a2) and (b2) show corresponding 2-D projections from a tour of the 7-D data and model, where each axis represents a linear combination of the original seven dimensions.
+Data points are plotted as individual markers, and selected subsets are highlighted in purple while unselected points appear in a contrasting color.
+In the top row, a set of points highlighted in the lower region of the UMAP layout in (a1) is linked to highlighted points in the tour view in (a2).
+In the bottom row, a different highlighted region in the UMAP layout in (b1) is linked to a different set of highlighted points in the tour view in (b2).
+The linked selections update consistently across panels, showing how the same subsets of observations appear in the UMAP layout and in different projections of the high-dimensional space.}]{paper-quollr_files/figure-latex/scurve-nldrlink-pdf-1} \caption{Exploring the correspondence between UMAP layout and \texttt{scurve} structure in $7\text{-}D$.
 Two sets of plots are interactively linked: UMAP layout (a1, b1) and projection of $7\text{-}D$ model and data (a2, b2).
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth,alt={Six-panel figure showing linked views of residuals and UMAP layouts for the 7-D S-curve model.
-Panels (a1–a3) highlight high-residual points in purple, corresponding to sparse bridging regions in the UMAP layout and less dense ends of the high-dimensional S-curve.
-Panels (b1–b3) highlight low-residual points, mapping to denser regions in the UMAP and thicker bands in the high-dimensional projection.
-The figure illustrates how residuals reveal distortions in UMAP embeddings.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ \texttt{scurve} model.
+\includegraphics[width=1\linewidth,alt={The figure consists of six panels arranged in two rows and three columns, showing residuals and their correspondence between UMAP and high-dimensional projections of the scurve data.
+In each row, the left panel (a1, b1) is a univariate plot of model residuals, with the horizontal axis representing residual values and the vertical axis representing frequency or density; points selected for highlighting appear in purple.
+The middle panel (a2, b2) shows a 2-D UMAP layout, with horizontal and vertical axes representing the two UMAP coordinates; the selected points from the left panel are highlighted in purple, while unselected points appear in a contrasting color.
+The right panel (a3, b3) shows a 2-D projection of the 7-D data and model using a tour view, where axes represent linear combinations of the original dimensions; highlighted points correspond to those selected in the residuals and UMAP panels, while unselected points are shown in a contrasting color.
+Across the panels, the purple highlights differ between the top and bottom rows, showing different subsets of points.
+All panels maintain roughly square aspect ratios and use consistent visual encodings for selection and non-selection.}]{paper-quollr_files/figure-latex/scurve-linkerror-pdf-1} \caption{Exploring residuals in relation to UMAP layouts using a $7\text{-}D$ \texttt{scurve} model.
 Three views are linked: distribution of residuals (a1, b1), UMAP layout (a2, b2), and projection of the $7\text{-}D$ model with data (a3, b3).
@@
 
-In the implementation examples, points are shown without cluster-based colouring unless explicitly stated.
-When points are coloured by cluster, linked plot functionality is currently only partially supported: selections made in the \texttt{langevitour} controls are not reflected in the corresponding highlights in the interactive \(2\text{-}D\) layout.
+In the implementation examples, points are shown without cluster-based coloring unless explicitly stated.
+When points are colored by cluster, linked plot functionality is currently only partially supported: selections made in the \texttt{langevitour} controls are not reflected in the corresponding highlights in the interactive \(2\text{-}D\) layout.
 As a result, cluster-specific exploration must be carried out separately in the tour view and the interactive \(2\text{-}D\) layout.
@@
 The accompanying HBE plot (typically a separate panel) summarizes model fit quality for each layout, with one value per method indicating that some embeddings provide a more faithful representation of the original data structure than others.}]{paper-quollr_files/figure-latex/limb-hbe-1} \caption{Assessing which of the 6 NLDR layouts on the limb muscle data is the better representation using HBE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
+Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d performs well at large binwidth (where the binwidth is not enough to capture the data structure) and poorly as the bin width decreases.
@@
 \begin{figure}[!ht]
-\includegraphics[width=1\linewidth,alt={Comparison of published $2\text{-}D$ layout (b) and the HBE-selected $2\text{-}D$ layout (f) for the limb muscle data (n = 1067) with seven differently shaped clusters in $10\text{-}D$, using tSNE, UMAP, PHATE, TriMAP, and PaCMAP with various hyper-parameters.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Compare the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}a) and the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) by HBE plot (Figure \ref{fig:limb-hbe}) from the tSNE, UMAP, PHATE, TriMAP, and PaCMAP with different hyper-parameters.
-The Limb muscle data ($n =  1067$) has seven close different shaped clusters in $10\text{-}D$.}\label{fig:model-limb}
+\includegraphics[width=1\linewidth,alt={The figure consists of six panels arranged in two rows and three columns, showing representative views of two NLDR layouts for the Limb muscle dataset (n=1067).
+In the top row (a1–a3), the left panel (a1) shows the published 2-D NLDR embedding with points colored by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2) and right (a3) panels show two different 2-D projections of the corresponding 10-D data and fitted model, with points plotted as individual markers and the same triangular mesh connecting centroids.
+In the bottom row (f1–f3), the left panel (f1) shows the 2-D NLDR layout selected using the HBE plot, again with colored points and triangulated centroids, while panels (f2) and (f3) show two different 2-D projections of the associated 10-D structure with the same mesh displayed.
+Across all panels, axes represent continuous numeric coordinates, color encodes muscle group, and the triangular mesh is used consistently to show correspondence between the low-dimensional layouts and the high-dimensional projections.}]{paper-quollr_files/figure-latex/model-limb-1} \caption{Representative views of two selected NLDR layouts for the Limb muscle dataset ($n=1067$), shown row-wise.
+The top row (a1–a3) corresponds to the published $2\text{-}D$ layout (Figure \ref{fig:limb-hbe}a), and the bottom row (f1–f3) corresponds to the $2\text{-}D$ layout selected (Figure \ref{fig:limb-hbe}f) selected using the HBE plot.
+In each row, the left panel (a1, f1) shows the NLDR embedding with points colored by muscle group and overlaid with triangulated hexagon centroids.
+The middle (a2, f2) and right (a3, f3) panels show two different $2\text{-}D$ projections of the fitted model and data in the original $10\text{-}D$ space, with the same triangular mesh displayed.
+Together, these panels summarize how the low-dimensional layouts relate to the underlying high-dimensional structure across different viewing directions.}\label{fig:model-limb}
 \end{figure}

### 2025-12-20 paper-quollr-submission-v1/paper-quollr.tex
Prepare submission for R Journal

--- 2025-12-19 paper-quollr.tex
@@
 Additional boxes show prediction and error computation for new data, as well as interactive components that link the p-D and 2-D representations.
-The diagram is laid out sequentially, showing how data move through the system from input to interactive output.}]{figures/quollr_workflow} \caption{Overview of the `quollr` workflow and software architecture.
+The diagram is laid out sequentially, showing how data move through the system from input to interactive output.}]{figures/quollr_workflow} \caption{Overview of the \texttt{quollr} workflow and software architecture.
 The process begins with NLDR and $p\text{-}D$ data inputs, followed by data preprocessing and hexagonal binning.

### 2025-12-20 paper-quollr-submission-v1/scripts/FACs/01_data_preprocessing.R

@@ -1,0 +1,88 @@
+## Tabula Muris: single organ analysis (Limb Muscles)
+## Paper: https://www.nature.com/articles/s41586-018-0590-4#data-availability
+## https://github.com/czbiohub-sf/tabula-muris/blob/master/00_data_ingest/FACS_Notebook.Rmd
+## Data: https://figshare.com/articles/dataset/Single-cell_RNA-seq_data_from_Smart-seq2_sequencing_of_FACS_sorted_cells_v2_/5829687/7
+
+library(Seurat)
+library(dplyr)
+library(Matrix)
+library(stringr)
+library(here)
+library(tidyverse)
+
+organ = "Limb_Muscle"
+
+raw.data <- read.csv(here("data", "limb_muscles",paste0(organ,"-counts.csv")),row.names = 1)
+meta.data <- read.csv(here("data", "limb_muscles", "metadata_FACS.csv"))
+
+plates <- str_split(colnames(raw.data),"[.]", simplify = TRUE)[,2]
+
+rownames(meta.data) <- meta.data$plate.barcode
+cell.meta.data <- meta.data[plates,]
+rownames(cell.meta.data) <- colnames(raw.data)
+
+# Find ERCC's, compute the percent ERCC, and drop them from the raw data.
+erccs <- grep(pattern = "^ERCC-", x = rownames(x = raw.data), value = TRUE)
+percent.ercc <- Matrix::colSums(raw.data[erccs, ])/Matrix::colSums(raw.data)
+ercc.index <- grep(pattern = "^ERCC-", x = rownames(x = raw.data), value = FALSE)
+raw.data <- raw.data[-ercc.index,]
+
+# Create the Seurat object with all the data
+tiss <- CreateSeuratObject(counts = raw.data, min.cells = 5, min.features = 5) # Changed raw.data to counts
+
+tiss <- AddMetaData(object = tiss, cell.meta.data)
+tiss[["percent.ercc"]] <- percent.ercc # Updated AddMetaData for v5
+# Change default name for sums of counts from nUMI to nReads
+tiss@meta.data <- tiss@meta.data %>% rename(nReads = nCount_RNA, nGene = nFeature_RNA) # Updated rename for v5
+
+ribo.genes <- grep(pattern = "^Rp[sl][[:digit:]]", x = rownames(x = tiss), value = TRUE) # Changed tiss@data to tiss
[…]

### 2025-12-20 paper-quollr-submission-v1/scripts/FACs/02_gen_embeddings.R

@@ -1,0 +1,151 @@
+library(dplyr)
+library(tibble)
+library(readr)
+library(conflicted)
+
+library(Rtsne)
+library(umap)
+library(phateR)
+library(reticulate)
+
+set.seed(20240110)
+
+conflicts_prefer(umap::umap)
+conflicts_prefer(dplyr::select)
+conflicts_prefer(dplyr::filter)
+
+use_python("~/miniforge3/envs/pcamp_env/bin/python")
+use_condaenv("pcamp_env")
+
+data <- read_rds(here::here("data/limb_muscles/facs_limb_muscles_pcs_10.rds"))
+data <- data |>
+  dplyr::select(-ID)
+
+## tSNE (default)
+perplexity <- 30
+
+tSNE_fit <- data |>
+  dplyr::select(where(is.numeric)) |>
+  Rtsne::Rtsne(perplexity = perplexity,
+               pca = FALSE)
+
+tSNE_data <- tSNE_fit$Y |>
+  tibble::as_tibble(.name_repair = "unique")
+names(tSNE_data) <- c("emb1", "emb2")
+
+tSNE_data <- tSNE_data |>
+  mutate(ID = row_number())
+
+write_rds(tSNE_data, file = paste0("data/limb_muscles/facs_limb_muscles_tsne_perplexity_", perplexity, ".rds"))
+
+## UMAP
+
+# Create a config list with the desired parameters
+umap_config <- umap.defaults
+umap_config$n_neighbors <- 15      # Set the number of neighbors
+umap_config$n_components <- 2    # Set the number of output dimensions (typically 2 or 3)
+umap_config$min_dist <- 0.1
+
+UMAP_fit <- umap(data, config = umap_config)
+
+UMAP_data <- UMAP_fit$layout |>
+  as_tibble()
+
+names(UMAP_data) <- c("emb1", "emb2")
+
+UMAP_data <- UMAP_data |>
+  mutate(ID = row_number())
+
+## Run only once
+write_rds(UMAP_data, file = paste0("data/limb_muscles/facs_limb_muscles_umap_n-neigbors_", "15", "_min-dist_", "0.1", ".rds"))
+
+## PHATE
+knn <- 5
+
+PHATE_data <- phate(data, knn = knn)
+PHATE_data <- as_tibble(PHATE_data$embedding)
+
+names(PHATE_data) <- c("emb1", "emb2")
+
+PHATE_data <- PHATE_data |>
+  mutate(ID = row_number())
+
[…]

### 2025-12-20 paper-quollr-submission-v1/scripts/FACs/06_comp_distances_clust_wise.R

@@ -1,0 +1,180 @@
+library(tidyverse)
+
+dist_df <- read_rds("data/limb_muscles/facs_limb_dist_df_with clusters.rds")
+
+## Within clusters
+
+### For cluster 0
+
+dist_df_clust0 <- dist_df |>
+  filter(from_cluster == 0) |>
+  filter(to_cluster == 0)
+
+dist_df_clust0 |>
+  ggplot(aes(x = dist_highd,
+             y = dist_2d_tsne_30)) +
+  geom_point(alpha=0.5) +
+  theme(aspect.ratio = 1)
+
+dist_df_clust0 |>
+  ggplot(aes(x = dist_highd,
+             y = dist_2d_tsne_15)) +
+  geom_point(alpha=0.5) +
+  theme(aspect.ratio = 1)
+
+### For cluster 1
+
+dist_df_clust1 <- dist_df |>
+  filter(from_cluster == 1) |>
+  filter(to_cluster == 1)
+
+dist_df_clust1 |>
+  ggplot(aes(x = dist_highd,
+             y = dist_2d_tsne_30)) +
+  geom_point(alpha=0.5) +
+  theme(aspect.ratio = 1)
+
+dist_df_clust1 |>
+  ggplot(aes(x = dist_highd,
+             y = dist_2d_tsne_15)) +
+  geom_point(alpha=0.5) +
+  theme(aspect.ratio = 1)
+
+### For cluster 2
+
+dist_df_clust2 <- dist_df |>
+  filter(from_cluster == 2) |>
+  filter(to_cluster == 2)
+
+dist_df_clust2 |>
+  ggplot(aes(x = dist_highd,
+             y = dist_2d_tsne_30)) +
+  geom_point(alpha=0.5) +
+  theme(aspect.ratio = 1)
+
+dist_df_clust2 |>
+  ggplot(aes(x = dist_highd,
+             y = dist_2d_tsne_15)) +
+  geom_point(alpha=0.5) +
+  theme(aspect.ratio = 1)
+
+### For cluster 3
+
+dist_df_clust3 <- dist_df |>
+  filter(from_cluster == 3) |>
+  filter(to_cluster == 3)
+
+dist_df_clust3 |>
+  ggplot(aes(x = dist_highd,
+             y = dist_2d_tsne_30)) +
+  geom_point(alpha=0.5) +
+  theme(aspect.ratio = 1)
+
+dist_df_clust3 |>
+  ggplot(aes(x = dist_highd,
[…]

### 2025-12-20 paper-quollr.R

@@ -545,7 +545,7 @@ tour_view <- show_langevitour(
 )
 
 
-## ----workflow, out.width = "100%", fig.pos="H", fig.cap="Overview of the `quollr` workflow and software architecture. The process begins with NLDR and $p\\text{-}D$ data inputs, followed by data preprocessing and hexagonal binning. Centroids are computed and triangulated to form the $2\\text{-}D$ mesh, which is then lifted into the $p\\text{-}D$ space. Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\\text{-}D$ and $2\\text{-}D$ representations.", fig.alt = "The figure is a schematic diagram illustrating the quollr workflow and software architecture. Rectangular boxes represent processing steps and data objects, including NLDR and p-D data inputs, data preprocessing, hexagonal binning, centroid computation, triangulation, and construction of a 2-D mesh. Arrows indicate the flow of information between steps, from the input data through the 2-D mesh, and then from the mesh being lifted into the p-D space. Additional boxes show prediction and error computation for new data, as well as interactive components that link the p-D and 2-D representations. The diagram is laid out sequentially, showing how data move through the system from input to interactive output."----
+## ----workflow, out.width = "100%", fig.pos="H", fig.cap="Overview of the \\texttt{quollr} workflow and software architecture. The process begins with NLDR and $p\\text{-}D$ data inputs, followed by data preprocessing and hexagonal binning. Centroids are computed and triangulated to form the $2\\text{-}D$ mesh, which is then lifted into the $p\\text{-}D$ space. Predictions and error computations are performed on new data, while interactive functions enable dynamic linking between the $p\\text{-}D$ and $2\\text{-}D$ representations.", fig.alt = "The figure is a schematic diagram illustrating the quollr workflow and software architecture. Rectangular boxes represent processing steps and data objects, including NLDR and p-D data inputs, data preprocessing, hexagonal binning, centroid computation, triangulation, and construction of a 2-D mesh. Arrows indicate the flow of information between steps, from the input data through the 2-D mesh, and then from the mesh being lifted into the p-D space. Additional boxes show prediction and error computation for new data, as well as interactive components that link the p-D and 2-D representations. The diagram is laid out sequentially, showing how data move through the system from input to interactive output."----
 
 knitr::include_graphics("figures/quollr_workflow.png")
 
